### 2016-03-23 paper.tex
basic files added

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage{bbm}
\usepackage{algorithm,algorithmic}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}
\usepackage{wrapfig}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{ulem}
\usepackage{multirow}

\def\B#1{\bm{#1}}
%\def\B#1{\mathbf{#1}}
\def\trans{\mathsf{T}}

%\renewcommand{\labelitemi}{--}

\newtheorem{theorem}{Theorem} \newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}{Remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{The Default Mode of Brain Function\\ Viewed By Control Theory}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\DeclareMathOperator{\proj}{proj}
\DeclareMathOperator{\softmax}{softmax}
\DeclareMathOperator{\prox}{prox}
\DeclareMathOperator{\Prox}{Prox}
\DeclareMathOperator{\im}{im}

% macros from michael's .tex
\DeclareMathOperator{\dist}{dist} % The distance.
\DeclareMathOperator{\argmin}{argmin}
\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\Id}{Id}
\DeclareMathOperator{\abs}{abs}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}

\newcommand{\suggestadd}[1]{{\color{blue} #1}}
\newcommand{\suggestremove}[1]{{\color{red} \sout{#1}}}

% \nipsfinalcopy % Uncomment for camera-ready version
\nipsfinaltrue
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\author{Elvis Dohmatob, (Bert Kappen), Danilo Bzdok\\
  INRIA, Parietal team, Saclay, France\\
  CEA, Neurospin, Gif-sur-Yvette, France\\
  firstname.lastname@inria.fr}

\maketitle

\begin{abstract}
ABC
%

% OUR keywords
%\textbf{\\keywords}: curse of dimensionality; semi-supervised learning;
%fMRI; systems neuroscience

% official NIPS keywords
\textbf{\\keywords}: Systems Biology, Cognitive Science, Autonomous Learning Systems

\end{abstract}

\section{Introduction}
%
abc

\section{Methods}
%
\paragraph{Data.}

\section{Experimental Results}
\paragraph{Serial versus parallel structure discovery and classification.}

\section{Discussion and Conclusion}

\paragraph{Acknowledgment}
{\small The research leading to these results has received funding from the
European Union Seventh Framework Programme (FP7/2007-2013)
under grant agreement no. 604102 (Human Brain Project).
Data were provided by the Human Connectome Project.
Further support was received from
the German National Academic Foundation (D.B.).
}

\small
\bibliographystyle{splncs03}
\bibliography{nips_refs}

\end{document}

### 2016-03-23 paper.tex
some intro notes

@@
 %
-abc
+\paragraph{The Human Default Mode Network.}
+\begin{itemize}
+  \item includes brain regions with highest baseline energy consumption in humans
+  \item most active in non-disturbed mind-wandering without external influence
+  \item tends to deactivate when humans engage in goal-direct tasks
+  \item fluctuations in the DMN are related to lapses during and performance in
+  such externally structured/focused tasks
+  \item includes brain regions that come online first when waking up from
+  anesthesia
+  \item its network nodes are late to myelinate, an indicator of cognitive sophistication
+  \item includes the most advanced processes hierarchies and has
+  no direct connections with sensory regions processing external input
+  \item the DMN's conceivable key role in the continuous environmental
+  tracking in a generative, integrative process might explain both its highest energy consumption in the brain and its intimate coupling with conscious awareness
+  \item Indeed, patients with right IPL (=DMN) damage have particular difficulties
+  with multistep actions (Hartmann et al. 2005)
+  and with imagined action (Sirigu et al. 1996).
+  \item Lesion of the hippocampus, feeding memory and spatial information, to the
+  DMN impairs future and hypothetical thinking (cf. Hassabis PNAS)
+  \item a lesion study on disturbed sleep (i.e., a state of mind
+  independent of sensory stimulation but dependent on
+  internally generated information) exclusively identified
+  the dmPFC (Koenigs et al., 2010). Third, another VLSM study exclusively
+  related the IFG and TPJ, both more strongly connected to the
+  dmPFC in our study, to inner speech
+  \item Bálint's syndrome: neurological disorder of conscious
+  awareness resulting from damage in bilateral parietal cortex;
+  "psychic paralysis of gaze", that is, the inability for internally
+  motivated, targeted saccades hindering fixation of currently not
+  attended features in the field of view (i.e., oculomotor apraxia).
+  unable to bind various individual features of the visual
+  environment into an integrated whole (i.e., simultanagnosia)
+  as well as unable to navigate hand movement to a targeted
+  object helped by vision (i.e., optic ataxia). 
+  difficulty avoiding running into objects placed ahead;
+  challenged in estimating distances as well as perceiving
+  spatial depth, object sizes and object orientation
+\end{itemize}
+
+\paragraph{The Predictive Coding Framework.}
+\begin{itemize}
+  \item It is one of the most frequently cited hypotheses of default mode function
+  \item Within this model the brain is conceptualized as a Bayesian machine
+  \item Predictive coding is a framework that, in a hierarchical setting,
+  is equivalent to empirical Bayesian inference
+  \item External sensory input is feed-forward/bottom-up processed in the brain
+  \item Their processing is compared against expected input
+  \item The prediction error, i.e. the difference between sensory input
+  and internal prediction, is computed at each level and passed to
+  higher levels via forward connections
+  \item in case of a mismatch, back-projections top-down modulate
+  input processing
+  \item ‘contextual integration’ for top-down modulation of sensorimotor
+  processing by context-specific a-priori information.
+  \item Closely related to Friston's free-energy principle: brain as inference engine
+  biological systems, including brains, must minimize the long-term average of surprise;
+  formulate perception as a constructive process based on internal or generative models
+  \item A generative model of the world
+\end{itemize}
+
+\paragraph{Control Theory.}
+\begin{itemize}
+\item mathematically formalized dynamics systems with input, output and
+feedback loops that produce error signals for adaptation
+\item typically realized by (non-linear) differential equations
+\item has been used to describe neural systems in the past
+\item gradient update = learn/adapt system dynamics by error occurrence
+\item related to notions of game theory
+\end{itemize}
 
@@
 \paragraph{Acknowledgment}
-{\small The research leading to these results has received funding from the
-European Union Seventh Framework Programme (FP7/2007-2013)
-under grant agreement no. 604102 (Human Brain Project).
-Data were provided by the Human Connectome Project.
-Further support was received from
-the German National Academic Foundation (D.B.).
-}
+% {\small The research leading to these results has received funding from the
+% European Union Seventh Framework Programme (FP7/2007-2013)
+% under grant agreement no. 604102 (Human Brain Project).
+% Further support was received from
+% the German National Academic Foundation (D.B.).
+% }
 

### 2016-03-23 build_paper.sh

@@ -25,3 +25,5 @@ pdflatex -interaction=nonstopmode -file-line-error-style -synctex=1 paper
 # makeindex -s paper.idx -t paper.glsdef -o paper.acr paper.acn
 makeglossaries paper
 pdflatex -interaction=nonstopmodxe -file-line-error-style -synctex=1 paper
+
+open paper.pdf

### 2016-07-03 paper.tex
methods: linear dynamical system + lqr feedback controller as a model for functional organization and operator

@@
 
+\def\Id{\mathbf{I}}
+\def\1{\mathbf{1}}
+\def\X{\mathbf{X}}
+\def\U{\mathbf{U}}
+\def\V{\mathbf{V}}
+\def\v{\mathbf{v}}
+\def\u{\mathbf{u}}
+\def\z{\mathbf{z}}
+\def\Y{\mathbf{Y}}
+\def\A{\mathbf{A}}
+\def\B{\mathbf{B}}
+\def\C{\mathbf{C}}
+\def\N{\mathbf{N}}
+\def\R{\mathbf{R}}
+\def\Q{\mathbf{Q}}
+\def\P{\mathbf{P}}
+\def\K{\mathbf{K}}
+\def\a{\mathbf{a}}
+\def\b{\mathbf{b}}
+\def\s{\mathbf{s}}
+\def\x{\mathbf{x}}
+
 \newcommand{\suggestadd}[1]{{\color{blue} #1}}
@@
 \begin{abstract}
-ABC
+We use linear time-varying dynamic control theory to study the infamous (human)  Default Mode Network.
 %
@@
 \section{Methods}
-%
+\subsection{Linear control network model for brain organization}
+Consider the following linear time-invariant (LTI) dynamical system as a (toy) model for high-level brain function
+\begin{equation}
+  \dot{\x}(t) = \A\x(t) + \B\u(t)
+\end{equation}
+Here, the $n$-by-$n$ matrix $\A$ denotes a model of the brain's wiring (for example a resting state connectome computed from an anatomical atlas), while the $n$-by-$k$ ``input matrix'' $\B$ describes which nodes can be controlled by us, an external \textit{controller}, via medical intervention or a careful choice of stimulus presentation, for example.
+At time $t \ge 0$, let  the $n$-vector  $\x(t) := (\x_1(t),\ldots,\x_n(t))$ describe the state of the network (one value for each node).
+The aim is to supply values of $\u_1(t),\ldots,\u_k(t)$ for $k \le n$ controls as a function of the time $t$,  take the system from any prescribed initial state $\x(0) = \x^{\text{init}} \in \mathbb R^n$ to any prescribed final state $\x(\tau) = \x^{\text{fin}} \in \mathbb R^n$ in finite time $\tau < \infty$.
+When such a controlling is possible, we say that the system $(\A,\B)$ is \textit{controllable}.
+A precise sufficient and necessary condition for such controllability is the Kalman condition: $(\A,\B)$ is controllable iff the $n \times nk$ \textit{controllability matrix}
+\begin{equation}
+  \C := (\B|\A\B|\ldots|\A^{n-1}\B)
+\end{equation}
+has full rank, i.e
+\begin{equation}
+  rank(\C) = n.
+\end{equation}
+
+\paragraph{How many controls do we need at best ?}
+In the thermodynamic limit ($n \rightarrow \infty$), statistical physics \textcolor{blue}{[Liu et al. 2011]} gives the extremely good estimate
+\begin{equation}
+  n_0(\langle k\rangle, \gamma) \approx e^{-\frac{1}{2}\left(\frac{\gamma-2}{\gamma - 1}\right)\langle k \rangle},
+\end{equation}
+where $\gamma \ge 1$ is the \textit{scale-free} parameter for the node-degree distribution of the network described by $\A$, and $\langle k \rangle$ is the mean degree.
+Letting $\gamma \rightarrow \infty$, one recovers the Erdos-Renyi value
+\begin{equation}
+  n_0(\langle k\rangle) = e^{-\frac{1}{2}\langle k\rangle},
+\end{equation}
+which agrees perfectly with its known analytic value. One identifies two radically different regimes:
+\begin{itemize}
+  \item the manageable regime $\gamma > \gamma_c := 2$, in which $n_0 < 1$, and so we only need fewer controls than total number of nodes, and
+\item the unmanageable regime ``$\gamma \le \gamma_c$'', in which  $n_0 = 1$, where each node must be controlled explicitly.
+\end{itemize}
+
+\paragraph{Why linear dynamics ?}
+As explained in \textcolor{blue}{[Liu et al. 2011]}, the choice of linear dynamics over a more general nonlinear dynamics can be justified as follows:
+\begin{itemize}
+  \item  Conclusions drawn from linear dynamics can be
+extended to nonlinear systems
+\item If the controllability matrix of the linearized system
+has full rank at all points, then it is sufficient for most
+systems to say that the actual nonlinear system is
+controllable (i.e. small signal model)
+\end{itemize}
+
+\subsection{Optimal control: an LQR feedback controller}
+We propose to use a linear quadratic regulator (LQR) for the feedback controller.
+Thus consider the time-varying \text{value function} $V: [0, \tau] \times \mathbb R^n \rightarrow \mathbb R$, defined by
+\begin{equation}
+  \begin{split}
+    &V(t, \z) := \min_{\u} \frac{1}{2}\x_{\text{fin}}^T\Q_{\text{fin}}\x_{\text{fin}} + \int_{0}^\tau\left(\frac{1}{2}\x(t)^T\Q\x(t) + \frac{1}{2}\u(t)^T\R\u(t)\right)dt,\\
+    &\text{subject to } \x(t) = \z, \dot{\x}(t) = \A\x(t) + \B\u(t),
+  \end{split}
+  \label{eq:hj}
+\end{equation}
+where the matrices $\Q$ and $\R$, specified with design considerations, are restricted to be positive semi-definite and positive definite, respectively.
+A classical calculation (remnicent of the \textit{Pontryagin minimization principle}) reveals that the optimal feedback control in \eqref{eq:hj} is given by
+\begin{equation}
+  \u(t) = -\K(t)\x(t),
+\end{equation}
+where
+\begin{equation}
+  \K(t) := \R^{-1}\B^T\P(t)
+\end{equation}
+and the time-varying matrix $\P(t)$ solves following matricial Riccati equation
+\begin{equation}
+\begin{split}
+&-\dot{\P}(t) = \A^T\P(t) + \P(t)\A - \P(t)\B\R^{-1}\B^T\P(t) + \Q\\
+&\text{subject to }\P(\tau) = \Q_{\text{fin}}.
+\end{split}
+\end{equation}
+
+\paragraph{Meta brain.}
+
 \paragraph{Data.}

### 2016-07-12 paper.tex
methods: linear dynamical system + lqr feedback controller as a model for functional organization and operator
added some refs

@@
   \item Indeed, patients with right IPL (=DMN) damage have particular difficulties
-  with multistep actions (Hartmann et al. 2005)
+  with multistory actions (Hartmann et al. 2005)
   and with imagined action (Sirigu et al. 1996).
@@
   \item ‘contextual integration’ for top-down modulation of sensorimotor
-  processing by context-specific a-priori information.
+  processing by context-specific a priori information.
   \item Closely related to Friston's free-energy principle: brain as inference engine
@@
   \dot{\x}(t) = \A\x(t) + \B\u(t)
+  \label{eq:lti}
 \end{equation}
@@
 \paragraph{How many controls do we need at best ?}
-In the thermodynamic limit ($n \rightarrow \infty$), statistical physics \textcolor{blue}{[Liu et al. 2011]} gives the extremely good estimate
+In the thermodynamic limit ($n \rightarrow \infty$), statistical physics \cite{Liu2011} gives the extremely good estimate
 \begin{equation}
@@
 \paragraph{Why linear dynamics ?}
-As explained in \textcolor{blue}{[Liu et al. 2011]}, the choice of linear dynamics over a more general nonlinear dynamics can be justified as follows:
+As explained in \cite{liu2011}, the choice of linear dynamics over a more general nonlinear dynamics can be justified as follows:
 \begin{itemize}
   \item  Conclusions drawn from linear dynamics can be
-extended to nonlinear systems
+extended to nonlinear systems.
 \item If the controllability matrix of the linearized system
@@
 systems to say that the actual nonlinear system is
-controllable (i.e. small signal model)
+controllable (i.e. small signal model).
 \end{itemize}
+
+\paragraph{A note on stability.}
+For stability in the LTI model \eqref{eq:lti} and hence in the constrained path integral \eqref{eq:hj} below, a sufficient (and necessary ?) condition is that all eigenvalues of the $A$ have negative real-part.
+One way to impose this is to add self-loops with a small negative weight.
+\paragraph{Meta brain.}
 
@@
 We propose to use a linear quadratic regulator (LQR) for the feedback controller.
-Thus consider the time-varying \text{value function} $V: [0, \tau] \times \mathbb R^n \rightarrow \mathbb R$, defined by
+Thus, consider the time-varying \text{value function} $V: [0, \tau] \times \mathbb R^n \rightarrow \mathbb R$, defined by the following Hamilton-Jacobi cost-to-go
 \begin{equation}
@@
 \end{equation}
-where the matrices $\Q$ and $\R$, specified with design considerations, are restricted to be positive semi-definite and positive definite, respectively.
-A classical calculation (remnicent of the \textit{Pontryagin minimization principle}) reveals that the optimal feedback control in \eqref{eq:hj} is given by
+where the matrices $\Q$ and $\R$, precised by design considerations and subject to meta-optimization, are restricted to be positive semi-definite and positive definite, respectively.
+A classical calculation (reminiscent of the \textit{Pontryagin minimization principle}) reveals that the optimal feedback control in \eqref{eq:hj} is given by
 \begin{equation}
@@
 \begin{equation}
-  \K(t) := \R^{-1}\B^T\P(t)
+  \K(t) := \R^{-1}\B^T\P(t)\text{ (Kalman gain matrix)}
 \end{equation}
-and the time-varying matrix $\P(t)$ solves following matricial Riccati equation
+and the time-varying positive semi-definite matrix $\P(t)$ solves following differential Riccati equation (DRE)
 \begin{equation}
@@
 \end{equation}
-
-\paragraph{Meta brain.}
+\subsection{Minimal control energy controller}
+\subsection{Related works}
+The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the choice $\Q_{\text{fin}} = \R = \textbf{I}$ and $\Q = \textbf{0}$, but with rather ad-hoc handle on stability issues and driver nodes...
 

### 2016-07-12 paper.tex
typo

@@
 \paragraph{Why linear dynamics ?}
-As explained in \cite{liu2011}, the choice of linear dynamics over a more general nonlinear dynamics can be justified as follows:
+As explained in \cite{Liu2011}, the choice of linear dynamics over a more general nonlinear dynamics can be justified as follows:
 \begin{itemize}

### 2016-07-12 paper.tex
more typos

@@
 \subsection{Related works}
-The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the choice $\Q_{\text{fin}} = \R = \textbf{I}$ and $\Q = \textbf{0}$, but with rather ad-hoc handle on stability issues and driver nodes...
+The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and driver nodes...
 

### 2016-07-12 paper.tex
fixups

--- 2016-07-12 paper.tex
@@
 \subsection{Related works}
-The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the choice $\Q_{\text{fin}} = \R = \textbf{I}$ and $\Q = \textbf{0}$, but with rather ad-hoc handle on stability issues and driver nodes...
+The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
 

### 2016-07-12 paper.tex
grammar

@@
 \begin{equation}
-  \dot{\x}(t) = \A\x(t) + \B\u(t)
+  \dot{\x}(t) = \A\x(t) + \B\u(t).
   \label{eq:lti}
@@
 Here, the $n$-by-$n$ matrix $\A$ denotes a model of the brain's wiring (for example a resting state connectome computed from an anatomical atlas), while the $n$-by-$k$ ``input matrix'' $\B$ describes which nodes can be controlled by us, an external \textit{controller}, via medical intervention or a careful choice of stimulus presentation, for example.
-At time $t \ge 0$, let  the $n$-vector  $\x(t) := (\x_1(t),\ldots,\x_n(t))$ describe the state of the network (one value for each node).
-The aim is to supply values of $\u_1(t),\ldots,\u_k(t)$ for $k \le n$ controls as a function of the time $t$,  take the system from any prescribed initial state $\x(0) = \x^{\text{init}} \in \mathbb R^n$ to any prescribed final state $\x(\tau) = \x^{\text{fin}} \in \mathbb R^n$ in finite time $\tau < \infty$.
+At time $t \ge 0$, let  the $n$-vector  $\x(t) := (\x_1(t),\ldots,\x_n(t))$ encodes the state of the network (one value for each node).
+The aim is to supply values of $\u_1(t),\ldots,\u_k(t)$ for $k \le n$ controls as a function of the time $t$,  to take the system from any prescribed initial state $\x(0) = \x^{\text{init}} \in \mathbb R^n$ to any prescribed final state $\x(\tau) = \x^{\text{fin}} \in \mathbb R^n$ in finite time $\tau < \infty$.
 When such a controlling is possible, we say that the system $(\A,\B)$ is \textit{controllable}.

### 2016-07-15 paper.tex
first anatomy

--- 2016-07-12 paper.tex
@@
 
-\title{The Default Mode of Brain Function\\ Viewed By Control Theory}
+\title{A Computational Account of Default-Mode Function\\by Control Theory and Reinforcement Learning}
 
@@
 
-\def\Id{\mathbf{I}}
-\def\1{\mathbf{1}}
-\def\X{\mathbf{X}}
-\def\U{\mathbf{U}}
-\def\V{\mathbf{V}}
-\def\v{\mathbf{v}}
-\def\u{\mathbf{u}}
-\def\z{\mathbf{z}}
-\def\Y{\mathbf{Y}}
-\def\A{\mathbf{A}}
-\def\B{\mathbf{B}}
-\def\C{\mathbf{C}}
-\def\N{\mathbf{N}}
-\def\R{\mathbf{R}}
-\def\Q{\mathbf{Q}}
-\def\P{\mathbf{P}}
-\def\K{\mathbf{K}}
-\def\a{\mathbf{a}}
-\def\b{\mathbf{b}}
-\def\s{\mathbf{s}}
-\def\x{\mathbf{x}}
-
 \newcommand{\suggestadd}[1]{{\color{blue} #1}}
@@
 \begin{abstract}
-We use linear time-varying dynamic control theory to study the infamous (human)  Default Mode Network.
+ABC
 %
@@
 \end{abstract}
+
+the DMN is often proposed to be dedicated to continuous auto- biographical memory retrieval, environmental assessment, emotion ap- praisal, and reward contingency evaluation when letting the mind go.
+
+Constantly tracking and predicting changes in a variable environment is believed to then shape the internal milieu and upcoming external behav- ior adaptively
+
+If this computational models hold some truth in them, the DMNs conceivable key role in the continuous environmental tracking in a generative, integrative process might explain both its highest energy consumption in the brain and its intimate coupling with conscious awareness.
 
@@
   \item Indeed, patients with right IPL (=DMN) damage have particular difficulties
-  with multistory actions (Hartmann et al. 2005)
+  with multistep actions (Hartmann et al. 2005)
   and with imagined action (Sirigu et al. 1996).
@@
   \item ‘contextual integration’ for top-down modulation of sensorimotor
-  processing by context-specific a priori information.
+  processing by context-specific a-priori information.
   \item Closely related to Friston's free-energy principle: brain as inference engine
@@
 \item related to notions of game theory
+\item switching between states -> typically reduces to a sequential processing model
+\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
+\item policy: action to take given a state
+\item value: prediction of future reward
+\item model: predict what the environment will do next
+\item model-baseed RI learning: we try to understand the environment; it is initially unknown
+\item special to RL: exploitation vs. exploration
+\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
+\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
+\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
 \end{itemize}
 
+\paragtaph{States and Actions in Neuroscience}
+
+state-dependent variables such as hunger, thirst and
+even wealth
+
+\paragraph{Anatomy}
+ involvement of the PMC across the boundaries of classical cognitive domains
+mPFC and PCC = highest metabolic consumption in the brain/highest level of basal glucose energy consumption in humans
+
+vmPFC:
+the vmPFC sub- serves predominantly non-ambiguous subjective-value-related evaluative processes driven by bottom-up pathways
+memory-informed reward and risk estimation of self-relevant environmental stimuli.
+the vmPFC might be more closely associated with orches- trating adapted behavior by bottom-up-driven processing of “what matters now”
+
+Most importantly, independent whole-brain analyses from structural neuroimaging studies related the gray-matter volume (GMV) of the vmPFC to indices of social competence and social network complex- ity in both humans and monkeys (Lebreton et al., 2009; Powell et al., 2010; Lewis et al., 2011; Sallet et al., 2011).
+To our knowl- edge, none of these four correlations have been found yet for the dmPFC.
+Consequently, vmPFC, rather than dmPFC, anatomy appears to predict an individual’s social behavioral dispositions and social network properties,
+
+With respect to our seeds, inter-individual differences in social skills or social networks were most often related to morphological differences in the human and monkey vmPFC, in stark contrast to the dmPFC
+
+self-related behavior guided by stimulus evalua- tion and reward-learning, a voxel-based lesion-symptom map- ping (VLSM) study in 344 neurological patients demonstrated functional-anatomical specificity of the vmPFC for value-based decision-making (Gläscher et al., 2012).
+
+As to the vmPFC subnetwork, the GMV of the vmPFC and VS correlated with indices of social reward attitudes and behavior (Lebreton et al., 2009), concur- ring with vmPFC’s relation to the NAc and reward-related tasks.
+
+In both MACM and RS analy- ses, the vmPFC was more strongly connected with the nucleus accumbens (NAc), hippocampus (HC), posterior cingulate cor- tex (PCC), and retrosplenial cortex (RSC)
+In both functional decoding analyses, the vmPFC was selectively associated with reward related tasks
+
+processing approach- and avoidance-relevant stimuli
+
+Glaescher: lesions -> value-based decision making  in 344 individuals 
+based decision-making, which included the orbitofrontal, ventro- medial, and frontopolar cortex.
+
+Indeed, the vmPFC, but not dmPFC, has been observed to have monosynap- tical connections with the ventral striatum (VS, which anatom- ically includes the NAc) in axonal tracing studies in monkeys (Haber et al., 1995; Ferry et al., 2000).
+This is consistent with our results and probabilistic diffusion tensor imaging (DTI) trac- tography in humans and monkeys (Croxson et al., 2005) that quantified the VS to be substantially more likely connected to the vmPFC than dmPFC in both species.
+
+the vmPFC is preferentially connected with limbic and reward-related medial brain areas.
+The NAc is thought to be linked to reward mechanisms that may not only modulate motivated behavior towards basic survival needs
+
+Such real or imagined bodily states, believed to be represented in the vmPFC, proba- bly operate as a bioregulatory disposition governing cognition and decision making (Damasio, 1996; Nauta, 1971),
+
+dmPFC:
+dmPFC subserves predominantly ambiguous amodal metacogni- tive processes driven by top-down pathways
+
+largely sensory-independent, highly abstract (hence, less tangible) processes across time, space, and content domains.
+might be top-down modulated by more dmPFC subserved higher reflective and hypothetical processing.
+consistent with present functional decoding, neural activity in the dmPFC, rather than vmPFC, has been con- sistently interpreted to underlie inference, representation, and assessment of one’s own and others’ mental states in functional neuroimaging research
+
+ dmPFC (but not vmPFC) activity was related to the proficiency decline of mental state inference in elderly (Moran et al., 2012), cognitive regulation of one’s own emotional states (Ochsner et al., 2004b) and inference of another person’s emotional states (Ochsner et al., 2004a), as well as self-reported (Wagner et al., 2011) and experimentally measured (Zaki et al., 2009) proficiency in emotional state inference.
+
+Thus, mental state inference necessarily relies on the generation of probabilistic internal information.
+Supported by dmPFC’s functional association with episodic memory retrieval, such prima vista non-mnemonic construction processes are likely to be subserved by the neural network underlying retrieval of past and imagination of future scenes as indicated by recent neu- roimaging experiments and meta-analyses
+
+congruently characterizes the dmPFC as a “mental sketchpad” (Goldman-Rakic, 1996) potentially impli- cated in modeling and binding plausible self- and other-related scenarios instructed by semantic concepts
+
+ sensory-independent de novo generation of mean- ing representations can only be expected from highly associative, integrative brain areas such as those of the dmPFC subnetwork
+
+the dmPFC subserves a domain-independent neural process
+
+the present results support the dmPFC’s possible involvement in domain-overarching computational mechanisms given its connections to highly associative brain areas and function- ally relation to different complex psychological processes.
+
+the dmPFC was more strongly connected with the inferior frontal gyrus (IFG), temporo-parietal junction (TPJ), and middle temporal gyrus (MTG)
+
+Concluding from previous and present connectivity findings, the dmPFC is preferentially connected with high association and heteromodal cortical areas of the lateral frontal, temporal, and parietal lobe.
+
+it is interesting to note that the vmPFC is more strongly con- nected to medial components of the default mode network (i.e., HC, PCC, RSC), whereas the dmPFC is more strongly con- nected to its lateral components (i.e., IFG, TPJ, and MTG).
+This dmPFC subnetwork was repeatedly related to self-focused reflec- tion (Andrews-Hanna et al., 2010), contemplation of others’ (Mar, 2011) and one’s own (Lombardo et al., 2009) mental states, mental navigation of the body in space (Maguire et al., 1997), semantic processing (Binder et al., 2009), as well as scene con- struction processes when envisioning past, fictitious, and future events (Hassabis et al., 2007; Spreng et al., 2009; Bzdok et al., 2013).
+
+a VLSM study on disturbed sleep (i.e., a state of mind independent of sensory stimulation but dependent on internally generated informa- tion) exclusively identified the dmPFC (Koenigs et al., 2010).
+-> experience replay during sleep
+
+PCC (gradients): rewards (exploitation) and gathering information about alter- native options (exploration)
+Such strategic decisions should incorporate not only recent reward history, but also opportunity costs and environmental statistics
+
+single-cell recordings in the monkey PCC demonstrated this brain region's sensitivity to subjective target utility (McCoy and Platt, 2005) and integration across individual outcomes in decision making (Pearson et al., 2009).
+ng attraction or aversion to options with uncertain or risky rewards
+. CGp activation was better predicted by the
+subjective salience of a chosen target than by its actual value.
+
+experimental psychologists and behavioral ecologists
+have long argued that decision making depends on the conversion
+of external variables into a common internal currency of value.
+
+electrophys- iological research in animals implicated the PCC in strategic selection (Pearson et al., 2009), risk assessment (McCoy and Platt, 2005), and outcome-contingent behavioral modulation (Hayden et al., 2008), while the RSC was implicated in nav- igation and approach-avoidance behavior (Vann et al., 2009).
+
+(McCoy and Platt, 2005):
+t internal representations of value lie at the very core of
+decision making,
+visual gambling task used in the current study affords a unique
+opportunity to examine neural activity under conditions in which
+subjective value, but not objective value, varied, as indicated by
+subjects’ choices
+-> value matrix is not the same in each agent
+h the direction of impending
+Our data extend those
+findings, demonstrating a direct relationship between subjective preferences
+for uncertain rewards, or the opportunity to harvest a
+relatively large reward, and the activity of neurons thought to participate
+in the allocation of attention
+movement, as shown previously24,30, and the uncertainty of rewards
+associated with this movement
+CGp appears to carry spatial information that is scaled by
+subjective preferences for particular patterns of reward outcome.
+Our data indicate that CGp
+neurons signal subjective biases for uncertain rewards (or, perhaps, the
+potential to receive a large reward) rather than objective target value.
+the spatial sensitivity of CGp neurons was enhanced under
+conditions of high risk.
+-> delay discounting / non-immediate reward -> reinforcement learning
+
+Pearson2009:
+balance. Here we show that CGp neurons distinguish be- tween exploratory and exploitative decisions made by mon- keys in a dynamic foraging task.
+ option switching [11] and integrate this information across multiple trials [11], but the present
+
+ Together, these results invite the hypothesis that CGp is part of a network that monitors the outcomes of individual deci- sions and integrates that information into higher-level strate- gies spanning multiple choices
+
+keys in a dynamic foraging task.
+Moreover, firing rates of these neurons predict in graded fashion the strategy most likely to be selected on upcoming trials.
+This encoding is distinct from switching between targets and is independent of the absolute magnitudes of rewards.
+These observations implicate CGp in the integration of individual outcomes across decision making and the modification of strategy in dynamic environments.
+recorded the activity of single cingulate cortex (CGp) neurons
+
+choosing among multiple tar-
+gets whose relative values changed dynamically,
+neurons in posterior cingulate cortex signaled the distinction between trials on which monkeys pursued an exploratory rather than an exploitative strategy.
+**learn its current value and integrate this information with their statistical knowledge of the environment to predict its relative value on upcoming trials.**
+
+Hayden2008: adapativity of subjective reward schedules tht impact the action choices
+ lesion (Gabriel, 1990) studies strongly implicate CGp in working mem- ory, attention, and general arousal (Raichle and Mintun, 2006)— all processes that contribute to adaptive decision making.
+-> gradient defect
+
+corded activity of single neurons in monkeys per- forming a gambling task in which the reward outcome of each choice strongly influenced subsequent choices.
+choices.
+We found that CGp neurons signaled reward outcomes in a nonlinear fashion and that outcome- contingent modulations in firing rate persisted into subsequent trials.
+Moreover, firing rate on any one trial predicted switching to the alternative option on the next trial.
+Finally, microstimulation in CGp follow- ing risky choices promoted a preference reversal for the safe option on the following trial.
+Collectively, these results demonstrate that CGp directly contrib- utes to the evaluative processes that support dynamic changes in decision making in volatile environments.
+
+ these observations suggest the hypothesis that CGp contributes to the integration of actions and their out- comes and thereby influences subsequent changes in behavior.
+ We confirmed a causal role for CGp in outcome evaluation and behavioral adjustment by showing that microstimulation following the most desired reward out- come increased the likelihood of exploring the alternative on the next trial.
+
+signal decision outcomes in a nonlinear fashion, maintain this information in a buffer between trials, and predict future changes in behavior.
+Moreover, microstimulation in CGp promotes explo- ration of the previously antipreferred option.
+
+highlight the dynamic nature of information processing in CGp, which is not restricted to any single epoch or aspect of task performance but instead continuously adapts to changes in both the external environment and internal milieu.
+
+(Patho-)Physiologically, meta- bolic fluctuations in the human PMC have been closely related to vari- ous instances of altered conscious awareness, including anesthesia (Fiset et al., 1999), sleep (Maquet, 2000), and restoration from vegeta- tive states (Laureys et al., 1999).
+
+Electrophysiologically, gamma band re- cordings in humans (Dastjerdi et al., 2011) and single-cell recordings in monkeys (Hayden et al., 2009) revealed activity reductions in the PMC during attentionally demanding tasks compared to rest.
+
+ have long been speculated to reflect constant contemplation of (external) environment and (internal) memory
+
+xons in parts of the PMC myelinate comparatively late during postnatal development in monkeys (Goldman- Rakic, 1987).
+Such late postnatal myelination is generally believed to occur in the phylogenetically most developed associations regions (Flechsig, 1920), t
+
+only cluster 2 was congruently coupled (across MACM and RSFC) with the amygdala (involved in significance evaluation) and the nucleus accumbens (involved in reward evaluation) 
+s well as exclusively functionally asso- ciated with facial appraisal.
+Similarly, the human vPCC was the only PMC region connected to the laterobasal (rather than centromedial or superficial) nuclei group of the amygdala (Bzdok et al., 2013b), which is an amygdalar subregion probably devoted to continuously scanning environmental input for biological significance (Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011; Ghods-Sharifi et al., 2009; LeDoux, 2007).
+
+vPCC:
+the ventral visual stream—the “what” system for object process- ing (Ungerleider and Haxby, 1994; Ungerleider and Mishkin, 1982; Vogt et al., 2006), contrary to the dPCC's conceivable relation to the dorsal “where” system
+
+dPCC:
+This view is in line with the well-known integration of the dPCC in the dorsal visual stream—the “where” system for spatial processing (Ungerleider and Haxby, 1994; Ungerleider and Mishkin, 1982; Vogt et al., 2006), in contrast to vPCC's relation to the ventral “what” system
+
+RSC:
+Regarding memory processing, lesions in the RSC, as compared to other PMC regions, is most consistently associated with anterograde and retrograde memory impairments of different types of sensory infor- mation in rabbits (Gabriel and Talk, 2001) and humans (Rudge and Warrington, 1991; Valenstein et al., 1987).
+current findings and earlier lesion reports can be parsimoniously reconciled by the previously proposed notion that the RSC mediates between the organism's egocentric (i.e., focused on sensory input) and allocentric (i.e., focused on world knowledge) perspective frames (Burgess, 2008; Epstein, 2008; Valiquette and McNamara, 2007).
+
+BOTH TPJ:
+
+RTPJ:
+
+LTPJ:
+
+PMC:
+covert reallocation of spatial attention (Gitelman et al., 1999), mediation between internal and external focus (Leech and Sharp, 2014), computation of environmental statistics (Pearson et al., 2009), and self-referential visuospatial imagery (Cavanna and Trimble, 2006)
+
+HC:
+
+The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
+
+the GMV of the entorhinal cortex (connectionally and functionally closely coupled with the HC) correlated with social network size (Kanai et al., 2012)
+
 \section{Methods}
-\subsection{Linear control network model for brain organization}
-Consider the following linear time-invariant (LTI) dynamical system as a (toy) model for high-level brain function
-\begin{equation}
-  \dot{\x}(t) = \A\x(t) + \B\u(t)
-  \label{eq:lti}
-\end{equation}
-Here, the $n$-by-$n$ matrix $\A$ denotes a model of the brain's wiring (for example a resting state connectome computed from an anatomical atlas), while the $n$-by-$k$ ``input matrix'' $\B$ describes which nodes can be controlled by us, an external \textit{controller}, via medical intervention or a careful choice of stimulus presentation, for example.
-At time $t \ge 0$, let  the $n$-vector  $\x(t) := (\x_1(t),\ldots,\x_n(t))$ describe the state of the network (one value for each node).
-The aim is to supply values of $\u_1(t),\ldots,\u_k(t)$ for $k \le n$ controls as a function of the time $t$,  take the system from any prescribed initial state $\x(0) = \x^{\text{init}} \in \mathbb R^n$ to any prescribed final state $\x(\tau) = \x^{\text{fin}} \in \mathbb R^n$ in finite time $\tau < \infty$.
-When such a controlling is possible, we say that the system $(\A,\B)$ is \textit{controllable}.
-A precise sufficient and necessary condition for such controllability is the Kalman condition: $(\A,\B)$ is controllable iff the $n \times nk$ \textit{controllability matrix}
-\begin{equation}
-  \C := (\B|\A\B|\ldots|\A^{n-1}\B)
-\end{equation}
-has full rank, i.e
-\begin{equation}
-  rank(\C) = n.
-\end{equation}
+%
+\paragraph{Reinforcement Learning.}
+\begin{itemize}
+\item optimal way to make decision
+\item closely related to reward system in the brain
+\item no supervisor, behavior is selected by trial-and-error approach
+\item agent moves around and thus influences the input data that is used for optimal decisions
+\item success and failure can only be evaluted post-hoc
+\item reward is a scalar feedback signal, nothing more
+\item the same formalism can frame a whole range of diverse real-world problems
+\item operation on the world are direct, the feed by reward is indirect
 
-\paragraph{How many controls do we need at best ?}
-In the thermodynamic limit ($n \rightarrow \infty$), statistical physics \cite{Liu2011} gives the extremely good estimate
-\begin{equation}
-  n_0(\langle k\rangle, \gamma) \approx e^{-\frac{1}{2}\left(\frac{\gamma-2}{\gamma - 1}\right)\langle k \rangle},
-\end{equation}
-where $\gamma \ge 1$ is the \textit{scale-free} parameter for the node-degree distribution of the network described by $\A$, and $\langle k \rangle$ is the mean degree.
-Letting $\gamma \rightarrow \infty$, one recovers the Erdos-Renyi value
-\begin{equation}
-  n_0(\langle k\rangle) = e^{-\frac{1}{2}\langle k\rangle},
-\end{equation}
-which agrees perfectly with its known analytic value. One identifies two radically different regimes:
-\begin{itemize}
-  \item the manageable regime $\gamma > \gamma_c := 2$, in which $n_0 < 1$, and so we only need fewer controls than total number of nodes, and
-\item the unmanageable regime ``$\gamma \le \gamma_c$'', in which  $n_0 = 1$, where each node must be controlled explicitly.
 \end{itemize}
-
-\paragraph{Why linear dynamics ?}
-As explained in \cite{Liu2011}, the choice of linear dynamics over a more general nonlinear dynamics can be justified as follows:
-\begin{itemize}
-  \item  Conclusions drawn from linear dynamics can be
-extended to nonlinear systems.
-\item If the controllability matrix of the linearized system
-has full rank at all points, then it is sufficient for most
-systems to say that the actual nonlinear system is
-controllable (i.e. small signal model).
-\end{itemize}
-
-\paragraph{A note on stability.}
-For stability in the LTI model \eqref{eq:lti} and hence in the constrained path integral \eqref{eq:hj} below, a sufficient (and necessary ?) condition is that all eigenvalues of the $A$ have negative real-part.
-One way to impose this is to add self-loops with a small negative weight.
-\paragraph{Meta brain.}
-
-\subsection{Optimal control: an LQR feedback controller}
-We propose to use a linear quadratic regulator (LQR) for the feedback controller.
-Thus, consider the time-varying \text{value function} $V: [0, \tau] \times \mathbb R^n \rightarrow \mathbb R$, defined by the following Hamilton-Jacobi cost-to-go
-\begin{equation}
-  \begin{split}
-    &V(t, \z) := \min_{\u} \frac{1}{2}\x_{\text{fin}}^T\Q_{\text{fin}}\x_{\text{fin}} + \int_{0}^\tau\left(\frac{1}{2}\x(t)^T\Q\x(t) + \frac{1}{2}\u(t)^T\R\u(t)\right)dt,\\
-    &\text{subject to } \x(t) = \z, \dot{\x}(t) = \A\x(t) + \B\u(t),
-  \end{split}
-  \label{eq:hj}
-\end{equation}
-where the matrices $\Q$ and $\R$, precised by design considerations and subject to meta-optimization, are restricted to be positive semi-definite and positive definite, respectively.
-A classical calculation (reminiscent of the \textit{Pontryagin minimization principle}) reveals that the optimal feedback control in \eqref{eq:hj} is given by
-\begin{equation}
-  \u(t) = -\K(t)\x(t),
-\end{equation}
-where
-\begin{equation}
-  \K(t) := \R^{-1}\B^T\P(t)\text{ (Kalman gain matrix)}
-\end{equation}
-and the time-varying positive semi-definite matrix $\P(t)$ solves following differential Riccati equation (DRE)
-\begin{equation}
-\begin{split}
-&-\dot{\P}(t) = \A^T\P(t) + \P(t)\A - \P(t)\B\R^{-1}\B^T\P(t) + \Q\\
-&\text{subject to }\P(\tau) = \Q_{\text{fin}}.
-\end{split}
-\end{equation}
-\subsection{Minimal control energy controller}
-\subsection{Related works}
-The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and driver nodes...
-
-\paragraph{Data.}
 

### 2016-07-15 paper.tex
restored lost formula + itemized anatomy stuff

--- 2016-07-12 paper.tex
@@
 \item related to notions of game theory
+\item switching between states -> typically reduces to a sequential processing model
+\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
+\item policy: action to take given a state
+\item value: prediction of future reward
+\item model: predict what the environment will do next
+\item model-baseed RI learning: we try to understand the environment; it is initially unknown
+\item special to RL: exploitation vs. exploration
+\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
+\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
+\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
+\end{itemize}
+
+\paragraph{States and Actions in Neuroscience}
+
+state-dependent variables such as hunger, thirst and
+even wealth
+
+\paragraph{Anatomy}
+ involvement of the PMC across the boundaries of classical cognitive domains
+mPFC and PCC = highest metabolic consumption in the brain/highest level of basal glucose energy consumption in humans
+
+\begin{itemize}
+  \item
+    vmPFC:
+    \begin{itemize}
+    \item
+the vmPFC sub- serves predominantly non-ambiguous subjective-value-related evaluative processes driven by bottom-up pathways
+memory-informed reward and risk estimation of self-relevant environmental stimuli.
+\item
+the vmPFC might be more closely associated with orches- trating adapted behavior by bottom-up-driven processing of “what matters now”
+
+\item
+Most importantly, independent whole-brain analyses from structural neuroimaging studies related the gray-matter volume (GMV) of the vmPFC to indices of social competence and social network complex- ity in both humans and monkeys (Lebreton et al., 2009; Powell et al., 2010; Lewis et al., 2011; Sallet et al., 2011).
+To our knowl- edge, none of these four correlations have been found yet for the dmPFC.
+Consequently, vmPFC, rather than dmPFC, anatomy appears to predict an individual’s social behavioral dispositions and social network properties,
+
+\item
+With respect to our seeds, inter-individual differences in social skills or social networks were most often related to morphological differences in the human and monkey vmPFC, in stark contrast to the dmPFC
+
+\item
+self-related behavior guided by stimulus evalua- tion and reward-learning, a voxel-based lesion-symptom map- ping (VLSM) study in 344 neurological patients demonstrated functional-anatomical specificity of the vmPFC for value-based decision-making (Gläscher et al., 2012).
+
+\item
+As to the vmPFC subnetwork, the GMV of the vmPFC and VS correlated with indices of social reward attitudes and behavior (Lebreton et al., 2009), concur- ring with vmPFC’s relation to the NAc and reward-related tasks.
+
+\item
+In both MACM and RS analy- ses, the vmPFC was more strongly connected with the nucleus accumbens (NAc), hippocampus (HC), posterior cingulate cor- tex (PCC), and retrosplenial cortex (RSC)
+In both functional decoding analyses, the vmPFC was selectively associated with reward related tasks
+
+\item
+processing approach- and avoidance-relevant stimuli
+
+\item
+Glaescher: lesions -> value-based decision making  in 344 individuals
+based decision-making, which included the orbitofrontal, ventro- medial, and frontopolar cortex.
+
+\item
+Indeed, the vmPFC, but not dmPFC, has been observed to have monosynap- tical connections with the ventral striatum (VS, which anatom- ically includes the NAc) in axonal tracing studies in monkeys (Haber et al., 1995; Ferry et al., 2000).
+This is consistent with our results and probabilistic diffusion tensor imaging (DTI) trac- tography in humans and monkeys (Croxson et al., 2005) that quantified the VS to be substantially more likely connected to the vmPFC than dmPFC in both species.
+
+\item
+the vmPFC is preferentially connected with limbic and reward-related medial brain areas.
+The NAc is thought to be linked to reward mechanisms that may not only modulate motivated behavior towards basic survival needs
+
+\item
+Such real or imagined bodily states, believed to be represented in the vmPFC, proba- bly operate as a bioregulatory disposition governing cognition and decision making (Damasio, 1996; Nauta, 1971),
+
+\item
+\end{itemize}
+
+\item
+  dmPFC:
+  \begin{itemize}
+    \item
+dmPFC subserves predominantly ambiguous amodal metacogni- tive processes driven by top-down pathways
+\item
+largely sensory-independent, highly abstract (hence, less tangible) processes across time, space, and content domains.
+might be top-down modulated by more dmPFC subserved higher reflective and hypothetical processing.
+consistent with present functional decoding, neural activity in the dmPFC, rather than vmPFC, has been con- sistently interpreted to underlie inference, representation, and assessment of one’s own and others’ mental states in functional neuroimaging research
+
+\item
+ dmPFC (but not vmPFC) activity was related to the proficiency decline of mental state inference in elderly (Moran et al., 2012), cognitive regulation of one’s own emotional states (Ochsner et al., 2004b) and inference of another person’s emotional states (Ochsner et al., 2004a), as well as self-reported (Wagner et al., 2011) and experimentally measured (Zaki et al., 2009) proficiency in emotional state inference.
+
+ \item
+Thus, mental state inference necessarily relies on the generation of probabilistic internal information.
+Supported by dmPFC’s functional association with episodic memory retrieval, such prima vista non-mnemonic construction processes are likely to be subserved by the neural network underlying retrieval of past and imagination of future scenes as indicated by recent neu- roimaging experiments and meta-analyses
+
+\item
+congruently characterizes the dmPFC as a “mental sketchpad” (Goldman-Rakic, 1996) potentially impli- cated in modeling and binding plausible self- and other-related scenarios instructed by semantic concepts
+
+\item
+ sensory-independent de novo generation of mean- ing representations can only be expected from highly associative, integrative brain areas such as those of the dmPFC subnetwork
+
+\item
+the dmPFC subserves a domain-independent neural process
+
+\item
+the present results support the dmPFC’s possible involvement in domain-overarching computational mechanisms given its connections to highly associative brain areas and function- ally relation to different complex psychological processes.
+
+\item
+the dmPFC was more strongly connected with the inferior frontal gyrus (IFG), temporo-parietal junction (TPJ), and middle temporal gyrus (MTG)
+
+\item
+Concluding from previous and present connectivity findings, the dmPFC is preferentially connected with high association and heteromodal cortical areas of the lateral frontal, temporal, and parietal lobe.
+
+\item
+it is interesting to note that the vmPFC is more strongly con- nected to medial components of the default mode network (i.e., HC, PCC, RSC), whereas the dmPFC is more strongly con- nected to its lateral components (i.e., IFG, TPJ, and MTG).
+This dmPFC subnetwork was repeatedly related to self-focused reflec- tion (Andrews-Hanna et al., 2010), contemplation of others’ (Mar, 2011) and one’s own (Lombardo et al., 2009) mental states, mental navigation of the body in space (Maguire et al., 1997), semantic processing (Binder et al., 2009), as well as scene con- struction processes when envisioning past, fictitious, and future events (Hassabis et al., 2007; Spreng et al., 2009; Bzdok et al., 2013).
+
+\item
+a VLSM study on disturbed sleep (i.e., a state of mind independent of sensory stimulation but dependent on internally generated informa- tion) exclusively identified the dmPFC (Koenigs et al., 2010).
+-> experience replay during sleep
+
+\end{itemize}
+
+  \item
+    PCC (gradients):
+    \begin{itemize}
+      \item rewards (exploitation) and gathering information about alter- native options (exploration)
+Such strategic decisions should incorporate not only recent reward history, but also opportunity costs and environmental statistics
+
+\item
+single-cell recordings in the monkey PCC demonstrated this brain region's sensitivity to subjective target utility (McCoy and Platt, 2005) and integration across individual outcomes in decision making (Pearson et al., 2009).
+ng attraction or aversion to options with uncertain or risky rewards
+. CGp activation was better predicted by the
+subjective salience of a chosen target than by its actual value.
+
+\item
+experimental psychologists and behavioral ecologists
+have long argued that decision making depends on the conversion
+of external variables into a common internal currency of value.
+
+\item
+electrophys- iological research in animals implicated the PCC in strategic selection (Pearson et al., 2009), risk assessment (McCoy and Platt, 2005), and outcome-contingent behavioral modulation (Hayden et al., 2008), while the RSC was implicated in nav- igation and approach-avoidance behavior (Vann et al., 2009).
+
+\item
+(McCoy and Platt, 2005):
+t internal representations of value lie at the very core of
+decision making,
+visual gambling task used in the current study affords a unique
+opportunity to examine neural activity under conditions in which
+subjective value, but not objective value, varied, as indicated by
+subjects’ choices
+-> value matrix is not the same in each agent
+h the direction of impending
+Our data extend those
+findings, demonstrating a direct relationship between subjective preferences
+for uncertain rewards, or the opportunity to harvest a
+relatively large reward, and the activity of neurons thought to participate
+in the allocation of attention
+movement, as shown previously24,30, and the uncertainty of rewards
+associated with this movement
+CGp appears to carry spatial information that is scaled by
+subjective preferences for particular patterns of reward outcome.
+Our data indicate that CGp
+neurons signal subjective biases for uncertain rewards (or, perhaps, the
+potential to receive a large reward) rather than objective target value.
+the spatial sensitivity of CGp neurons was enhanced under
+conditions of high risk.
+-> delay discounting / non-immediate reward -> reinforcement learning
+
+\item
+Pearson2009:
+balance. Here we show that CGp neurons distinguish be- tween exploratory and exploitative decisions made by mon- keys in a dynamic foraging task.
+ option switching [11] and integrate this information across multiple trials [11], but the present
+
+ \item
+ Together, these results invite the hypothesis that CGp is part of a network that monitors the outcomes of individual deci- sions and integrates that information into higher-level strate- gies spanning multiple choices
+
+ \item
+keys in a dynamic foraging task.
+Moreover, firing rates of these neurons predict in graded fashion the strategy most likely to be selected on upcoming trials.
+This encoding is distinct from switching between targets and is independent of the absolute magnitudes of rewards.
+These observations implicate CGp in the integration of individual outcomes across decision making and the modification of strategy in dynamic environments.
+recorded the activity of single cingulate cortex (CGp) neurons
+
+\item
+choosing among multiple tar-
+gets whose relative values changed dynamically,
+neurons in posterior cingulate cortex signaled the distinction between trials on which monkeys pursued an exploratory rather than an exploitative strategy.
+**learn its current value and integrate this information with their statistical knowledge of the environment to predict its relative value on upcoming trials.**
+
+\item
+Hayden2008: adapativity of subjective reward schedules tht impact the action choices
+ lesion (Gabriel, 1990) studies strongly implicate CGp in working mem- ory, attention, and general arousal (Raichle and Mintun, 2006)— all processes that contribute to adaptive decision making.
+-> gradient defect
+
+\item
+corded activity of single neurons in monkeys per- forming a gambling task in which the reward outcome of each choice strongly influenced subsequent choices.
+choices.
+We found that CGp neurons signaled reward outcomes in a nonlinear fashion and that outcome- contingent modulations in firing rate persisted into subsequent trials.
+Moreover, firing rate on any one trial predicted switching to the alternative option on the next trial.
+Finally, microstimulation in CGp follow- ing risky choices promoted a preference reversal for the safe option on the following trial.
+Collectively, these results demonstrate that CGp directly contrib- utes to the evaluative processes that support dynamic changes in decision making in volatile environments.
+
+\item
+ these observations suggest the hypothesis that CGp contributes to the integration of actions and their out- comes and thereby influences subsequent changes in behavior.
+ We confirmed a causal role for CGp in outcome evaluation and behavioral adjustment by showing that microstimulation following the most desired reward out- come increased the likelihood of exploring the alternative on the next trial.
+
+ \item
+signal decision outcomes in a nonlinear fashion, maintain this information in a buffer between trials, and predict future changes in behavior.
+Moreover, microstimulation in CGp promotes explo- ration of the previously antipreferred option.
+
+\item
+highlight the dynamic nature of information processing in CGp, which is not restricted to any single epoch or aspect of task performance but instead continuously adapts to changes in both the external environment and internal milieu.
+
+\item
+(Patho-)Physiologically, meta- bolic fluctuations in the human PMC have been closely related to vari- ous instances of altered conscious awareness, including anesthesia (Fiset et al., 1999), sleep (Maquet, 2000), and restoration from vegeta- tive states (Laureys et al., 1999).
+
+  \item
+Electrophysiologically, gamma band re- cordings in humans (Dastjerdi et al., 2011) and single-cell recordings in monkeys (Hayden et al., 2009) revealed activity reductions in the PMC during attentionally demanding tasks compared to rest.
+
+\item
+ have long been speculated to reflect constant contemplation of (external) environment and (internal) memory
+
+ \item
+   xons in parts of the PMC myelinate comparatively late during postnatal development in monkeys (Goldman- Rakic, 1987).
+Such late postnatal myelination is generally believed to occur in the phylogenetically most developed associations regions (Flechsig, 1920), t
+
+\item
+only cluster 2 was congruently coupled (across MACM and RSFC) with the amygdala (involved in significance evaluation) and the nucleus accumbens (involved in reward evaluation)
+s well as exclusively functionally asso- ciated with facial appraisal.
+Similarly, the human vPCC was the only PMC region connected to the laterobasal (rather than centromedial or superficial) nuclei group of the amygdala (Bzdok et al., 2013b), which is an amygdalar subregion probably devoted to continuously scanning environmental input for biological significance (Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011; Ghods-Sharifi et al., 2009; LeDoux, 2007).
+\end{itemize}
+
+  \item
+    vPCC:
+the ventral visual stream—the “what” system for object process- ing (Ungerleider and Haxby, 1994; Ungerleider and Mishkin, 1982; Vogt et al., 2006), contrary to the dPCC's conceivable relation to the dorsal “where” system
+\item
+  dPCC:
+This view is in line with the well-known integration of the dPCC in the dorsal visual stream—the “where” system for spatial processing (Ungerleider and Haxby, 1994; Ungerleider and Mishkin, 1982; Vogt et al., 2006), in contrast to vPCC's relation to the ventral “what” system
+
+\item
+RSC:
+Regarding memory processing, lesions in the RSC, as compared to other PMC regions, is most consistently associated with anterograde and retrograde memory impairments of different types of sensory infor- mation in rabbits (Gabriel and Talk, 2001) and humans (Rudge and Warrington, 1991; Valenstein et al., 1987).
+current findings and earlier lesion reports can be parsimoniously reconciled by the previously proposed notion that the RSC mediates between the organism's egocentric (i.e., focused on sensory input) and allocentric (i.e., focused on world knowledge) perspective frames (Burgess, 2008; Epstein, 2008; Valiquette and McNamara, 2007).
+
+\item
+BOTH TPJ:
+
+\item
+RTPJ:
+
+\item
+
+LTPJ:
+
+\item
+  PMC:
+  \begin{itemize}
+    \item
+covert reallocation of spatial attention (Gitelman et al., 1999), mediation between internal and external focus (Leech and Sharp, 2014), computation of environmental statistics (Pearson et al., 2009), and self-referential visuospatial imagery (Cavanna and Trimble, 2006)
+  \end{itemize}
+
+\item
+HC:
+\begin{itemize}
+  \item
+The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
+
+\item
+the GMV of the entorhinal cortex (connectionally and functionally closely coupled with the HC) correlated with social network size (Kanai et al., 2012)
+\end{itemize}
 \end{itemize}
@@
 
-\paragraph{Data.}
-
 \section{Experimental Results}

### 2016-07-15 paper.tex
hippocamp notes

--- 2016-07-15 paper.tex
@@
 
+"statistical control of the default brain function"
+
 If this computational models hold some truth in them, the DMNs conceivable key role in the continuous environmental tracking in a generative, integrative process might explain both its highest energy consumption in the brain and its intimate coupling with conscious awareness.
@@
 %
+
+Global objectives:
+a) frequentist alternative to Fristons paper -> „de-Friston-ize"
+b) emphasize more reinforcement learning aspects
+c) direct mapping to parts of the default mode network -> be much more specific
+
 \paragraph{The Human Default Mode Network.}
@@
 
+- In fact, phase-locked hippocampal–cortical loops have been proposed to correlate with the acquisition and retrieval of memories
+- formation of mnemonic neuronal representations
+- **a large body of animal and human literature relates
+hippocampal-prefrontal theta to memory and memory to the DMN**
+-
+
 \section{Methods}

### 2016-07-15 paper.tex
hipp comments
new anat notes

--- 2016-07-15 paper.tex
@@
 
+RTPJ:
 \item
-RTPJ:
+Corbetta2008:
+Although both attentional networks have been most exten- sively investigated in vision, the available evidence indicates a supramodal function (Driver and Spence, 1998; Macaluso et al., 2002).
+The ventral network (right TPJ, right IFG) registers salient events in the environment not only in the visual but also in the auditory and tactile modalities (Downar et al., 2000)
+Several lines of evidence indicate that two cortico-cortical neural systems are involved in attending to environmental stimuli (Corbetta and Shulman, 2002).
+A dorsal frontoparietal network.
+A second system, the ventral frontoparietal network, is not ac- tivated by expectations or task preparation but responds along with the dorsal network when behaviorally relevant objects (or targets) are detected (Corbetta et al., 2000).
+Both dorsal and ventral networks are also activated during reorienting, with en- hanced responses during the detection of targets that appear at unattended locations.
+For example, enhanced responses are observed when subjects are cued to expect a target at one location but it unexpectedly appears at another (i.e., ‘‘invalid’’ targets in the Posner spatial cueing paradigm)
+or when a target appears infre- quently, as in ‘‘oddball’’ paradigms.
+
+Intimately related to prediction error signaling:
+An early theory of how the two networks interact (Corbetta and Shulman, 2002) proposed that when attention is reoriented to a new source 
+of information (stimulus-driven reorienting), output from the ven- tral network interrupts (as a ‘‘circuit breaker’’) ongoing selection in the dorsal network, which in turn shifts attention toward the novel object of interest.
+
+relevance is the critical factor that determines whether an object activates the ventral network (Downar et al., 2001).
+The ventral network might be considered a prime candi-date for mediating orienting to salient but unimportant stimuli, i.e., exogenous attention (Posner and Cohen, 1984), be- cause under passive conditions it is highly responsive to distinctive sensory
+events in all modalities (Downar et al., 2000). The ventral network is well activated by stimuli that are important, even if they are not very distinctive. 
+CURRENT OPINION: VA network does underlie  stimulus-driven reorienting to environmental stimuli based on their task relevance.
+An important conclusion from these neuroimaging studies is that the psychological dis- tinction between exogenous and endogenous orienting (Jo- nides, 1981) may not map onto different neural systems.
+
+\item
+
+\item
+
+\item
 
@@
 the GMV of the entorhinal cortex (connectionally and functionally closely coupled with the HC) correlated with social network size (Kanai et al., 2012)
-\end{itemize}
+
+\item 
+In fact, phase-locked hippocampal?<80><93>cortical loops have been proposed t
+formation of mnemonic neuronal representations
+**a large body of animal and human literature relates
+hippocampal-prefrontal theta to memory and memory to the DMN**
+
+\item 
+A special case is the hippocampus whose highly recursive connection matrix is thought to function as a large ‘autoassocia- tor’, allowing the reconstruction of entire episodes from remembered fragments.
+
+\item
+hippocampal pyramidal cells during rest and sleep produce strongly coherent ensemble bursts believed to be critical in transferring information to the neocortex.
+
+\item
+sequential single- cell recording could reveal such cooperative patterns
+
+\item
+buszaki2004 theta oscillations:
+Using large-scale recordings, researchers can follow complex patterns for extended time periods and determine whether their modification by experience will influence self-generated patterns in the absence of envi- ronmental inputs.
+Spike sequences, imposed upon the network by behavioral manipulations, recur spontaneously during subsequent sleep
+indicating that neurons organize themselves into preferred cell assemblies, and the seeds of emergence stem from experi- ence-related activity.
+ Such internally driven, self-organized assemblies may reflect mechanisms that give rise to cognitive phenomena
+ Sharp-wave ripples are believed to be the transfer mechanism of information from the hippocampus to neocortex during off-line (eg, sleep) states.
+Neurons, mostly studied in the hippocampus, which fire action potentials only when the animal is in a specific place in a given environment (that place where a cell fires is called the “place field” of that cell).
+-> self-play in deepmind Go
+
+\item
+statistical structure of hippocampal activity ->
+spatial maps of surroundings are encoded
+
+\item
+Numerous experiments have demonstrated that hippocampal neurons show place-related firing while the rat explores or traverses its environment so that each assembly of hippocampal principal cells defines a particular position of space
+
+\item
+Diba + Buzsaki 2007
+We report that temporal spike sequences from hippocampal place neurons of rats on an elevated track recurred in reverse order at the end of a run, but in forward order in anticipation of the run, coinciding with sharp waves.
+Vector distances between the place fields were reflected in the temporal structure of these sequences.
+This bidirectional re-enactment of temporal sequences may contribute to the establishment of higher-order associations in episodic memory.
+
+\item
+coding, transfer, and consolidation of
+information
+This emerging picture of hippocampal dy- namics suggests that neurons at the opposite ends of the distributions may convey differ- ent but complementary types of information.
+The ever-active minority of place cells may be responsible for generalizing across envi- ronments and affords the brain the capacity to regard no situation as completely un- known because every alley, mountain, river, or room has elements of previously experi- enced similar situations.
+In many situations,
+this minority provides the “best guess” of the hippocampus and offers “good enough” solutions to get by.
+On the other hand, the majority of less active neurons constitute a large reservoir that can be mobilized (7) to precisely distinguish one situation from an- other and incorporate novel ones as distinct.
+
 \end{itemize}

### 2016-07-15 paper.tex
free-energy principle is here

--- 2016-07-15 paper.tex
@@
 the GMV of the entorhinal cortex (connectionally and functionally closely coupled with the HC) correlated with social network size (Kanai et al., 2012)
+
+\item 
+ In fact, phase-locked hippocampal?<80><93>cortical loops have been proposed t
+ formation of mnemonic neuronal representations
+ **a large body of animal and human literature relates
+hippocampal-prefrontal theta to memory and memory to the DMN**
 \end{itemize}
@@
 
+\subsection{A thermodynamic model for bounded rationality, aka robust optimality}
+\begin{itemize}
+  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
+  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
+  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
+    \end{itemize}
+
+\paragraph{Utility functions and conjugate pairs.}
+  Let $(\Omega, \mathcal F, \P)$ be a probablity space.
+A function $\U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $\U(A|B) := \U(A \cap B) - \U(B)$ has the following propertites:
+  \begin{itemize}
+  \item additivity: $\U(A_1 \cap A_2 | B) = \U(A_1|B) + \U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
+  \item statistic: there exists a function $f_{\U} :\mathbb R_+ \rightarrow \mathbb R$ such that $\U(A|B) = f(\P(A|B))$, for all events
+    $A,B \in \mathcal F$.
+  \item monotonicity: $f_{\U}$ is strictly increasing.
+    \end{itemize}
+
+\begin{theorem}
+  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $\U(A|B) \equiv f(\P(A|B))$ any probability space $(\Omega, \mathcal F, \P)$ and utility function $\U$ thereupon are of the form
+  \begin{equation}
+    f = \alpha \log(.),
+    \label{eq:boltzmann}
+  \end{equation}
+  where $\alpha > 0$.
+\end{theorem}
+
+\textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
+
+Such $\U$ and $\P$ are said to form a \textit{conjugate pair} at temperator $\alpha$.
+
+\paragraph{Example.}
+Given utility function $U$ a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-\U(\omega))_{\omega \in \Omega}$ is defined to be probability measure (on thesame measurable space)
+  \begin{equation}
+    \P_{\U}(A) = \sum_{\omega \in A}\frac{1}{Z(\alpha)}\exp\left(\frac{1}{\alpha}\U(\omega)\right), \; \forall A \in \mathcal F.
+  \end{equation}
+$Z(\alpha)$ is a normalization constant such that $\P_{\U}(\Omega) = 1$.  It's not hard to see that $\U$ and $\P_{\U}$  form a conjugate pair.
+
+  \paragraph{Free-utility functional.}
+    Let $(\U, \P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
+Given another probability measure $\P'$ on the same space, define it's \textit{free utility} as
+    \begin{equation}
+      J(\P'|\U, \P) = \mathbb E_{\P'}(\U) + \alpha \mathbb H(\P'),
+    \label{eq:free_u}
+    \end{equation}
+    where $\mathbb H(\P') := -\sum_{\omega \in \Omega}\P'(\omega)\log(\P'(\omega))$ is the entropy of $\P'$.
+
+  It's not difficult to show that
+  \begin{equation}
+    J(\P'|\U,\P) \le J(\P|\U,\P) = \sum_{\omega \in \Omega}\U(\omega) =: \U(\Omega).
+  \end{equation}
+  In particular, if $\P = \P_{\U}$ the Gibbs measure, then the upper bound above reduces to the \textit{log-partition function}
+  \begin{equation}
+    J(\P'|\U,\P) \le \U(\Omega) = -\alpha \log\left(\sum_{\omega \in \Omega}e^{\frac{1}{\alpha}\U(\omega)}\right).
+    \end{equation}
+
+  \paragraph{The free-energy / utility principle.}
+  We are now in shape to introduce the notion of free-energy for transition of priors, and a variational principle for optimizing it.
+Consider thus an initial system described by a conjugate pair $(\U_{\text{ini}}, \P_{\text{ini}}) \rightarrow \P_{\text{fin}}$ at temperature $\alpha > 0$.
+We want to transform this to a new one by adding constraints represented by the utility function $\delta \U$.
+The resulting system has final utility $\U_{\text{fin}} = \U_{\text{ini}} + \delta\U$.
+The difference in free-utility is then
+  \begin{equation}
+    \delta J = J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{\P_{\text{fin}}}(\delta \U)}_{\textbf{expected gain in utility}} + \alpha \underbrace{D_{\text{KL}}(\P_{\text{ini}}\|\P_{\text{fin}})}_{\textbf{complexity of change}},
+    \label{eq:free}
+  \end{equation}
+  where $D_{\text{KL}}(\P_{\text{ini}}\|\P_{\text{fin}})$ is the Kullback-Leibler divergence.
+  In the above formula, we've extensively used the fact that $(\U_{\text{ini}}, \P_{\text{ini}})$ is a congugate pair and so $\U_{\text{ini}}(\omega) \equiv \alpha \log(\P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
+  The two terms in \eqref{eq:free} (expected gain in utility or negative workdone, and the complexity of the transition)
+
 \section{Experimental Results}

### 2016-07-15 paper.tex
anat rtpj

--- 2016-07-15 paper.tex
@@
 An important conclusion from these neuroimaging studies is that the psychological dis- tinction between exogenous and endogenous orienting (Jo- nides, 1981) may not map onto different neural systems.
+"changing the control of behavior from one environ- mental input to another".
+Similar reorienting mechanisms may also be involved in shifting from a broad range of ‘‘internally directed’’ processes in order to deal with environmental events, as when interrupting memory retrieval (‘‘did I lock the car door?’’) to respond to a sudden stimulus (‘‘is that my cell phone ring- ing?’’).
+-> medites between the Default-Mode function and other more environmentally directed processing goals.
 

### 2016-07-15 paper.tex
thing
more free energy and cold/hot thermodynamic limits

@@
 
-\item
-
-\item
-
-\item
-
 LTPJ:
@@
 On the other hand, the majority of less active neurons constitute a large reservoir that can be mobilized (7) to precisely distinguish one situation from an- other and incorporate novel ones as distinct.
-
+\end{itemize}
 \end{itemize}
@@
 \begin{equation}
-  n_0(\langle k\rangle, \gamma) \approx e^{-\frac{1}{2}\left(\frac{\gamma-2}{\gamma - 1}\right)\langle k \rangle},
+  n_0(\langle k\rangle, \gamma) \approx \exp\left(-\frac{1}{2}\left(\frac{\gamma-2}{\gamma - 1}\right)\langle k \rangle\right),
 \end{equation}
@@
 \begin{equation}
-  n_0(\langle k\rangle) = e^{-\frac{1}{2}\langle k\rangle},
+  n_0(\langle k\rangle) = \exp\left(-\frac{1}{2}\langle k\rangle\right),
 \end{equation}
@@
 
+\subsection{A thermodynamic model for bounded rationality, aka robust optimality}
+\begin{itemize}
+  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
+  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
+  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
+    \end{itemize}
+The material presented is a revisit of \textcolor{blue}{[Braun 2011]}.
+\paragraph{Utility functions and conjugate pairs.}
+  Let $(\Omega, \mathcal F, \P)$ be a probablity space.
+A function $\U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $\U(A|B) := \U(A \cap B) - \U(B)$ has the following propertites:
+  \begin{itemize}
+  \item additivity: $\U(A_1 \cap A_2 | B) = \U(A_1|B) + \U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
+  \item statistic: there exists a function $f_{\U} :\mathbb R_+ \rightarrow \mathbb R$ such that $\U(A|B) = f(\P(A|B))$, for all events
+    $A,B \in \mathcal F$.
+  \item monotonicity: $f_{\U}$ is strictly increasing.
+    \end{itemize}
+
+\begin{theorem}
+  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $\U(A|B) \equiv f(\P(A|B))$ any probability space $(\Omega, \mathcal F, \P)$ and utility function $\U$ thereupon are of the form
+  \begin{equation}
+    f = \alpha \log(.),
+    \label{eq:boltzmann}
+  \end{equation}
+  where $\alpha > 0$.
+\end{theorem}
+
+\textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
+
+Such $\U$ and $\P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
+
+\paragraph{Example.}
+Given a utility function $\U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-\U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
+  \begin{equation}
+    \P(A) = \sum_{\omega \in A}\frac{1}{Z(\alpha)}\exp\left(\frac{1}{\alpha}\U(\omega)\right), \; \forall A \in \mathcal F.
+  \end{equation}
+$Z(\alpha)$ is a normalization constant such that $\P(\Omega) = 1$.  It's not hard to see that $\U$ and the $\P$ above form a conjugate pair.
+
+  \paragraph{Free-utility functional.}
+    Let $(\U, \P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
+Given another probability measure $\P'$ on the same space, define it's \textit{free utility} as
+    \begin{equation}
+      J(\P'|\U, \P) = \mathbb E_{\P'}(\U) + \alpha \mathbb H(\P'),
+    \label{eq:free_u}
+    \end{equation}
+    where $\mathbb H(\P') := -\sum_{\omega \in \Omega}\P'(\omega)\log(\P'(\omega))$ is the entropy of $\P'$.
+
+  It's not difficult to show that
+  \begin{equation}
+    J(\P'|\U,\P) \le J(\P|\U,\P) = \sum_{\omega \in \Omega}\U(\omega) =: \U(\Omega).
+  \end{equation}
+  In particular, if $\P$ is the Gibbs measure corresponding to $\U$, then the upper bound above reduces to the \textit{log-partition function}
+  \begin{equation}
+    J(\P'|\U,\P) \le \U(\Omega) = -\alpha \log\left(\sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}\U(\omega)\right)\right).
+    \end{equation}
+
+  \paragraph{The free-energy / utility principle.}
+  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
+Consider thus an initial system described by a conjugate pair $(\U_{\text{ini}}, \P_{\text{ini}})$ at temperature $\alpha > 0$.
+We want to transform this to a new model by adding constraints represented by the utility function $\delta \U$.
+The resulting system has final utility $\U_{\text{fin}} = \U_{\text{ini}} + \delta\U$.
+The difference in free-utility is then
+  \begin{equation}
+    \delta J_{(\U_{\text{ini}}, \P_{\text{ini}}) \rightarrow \P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{\P_{\text{fin}}}(\delta \U)}_{\textbf{expected gain in utility}} + \alpha \underbrace{D_{\text{KL}}(\P_{\text{ini}}\|\P_{\text{fin}})}_{\textbf{complexity of change}},
+    \label{eq:free}
+  \end{equation}
+  where $D_{\text{KL}}(\P_{\text{ini}}\|\P_{\text{fin}})$ is the Kullback-Leibler divergence.
+  In the above formula, we've extensively used the fact that $(\U_{\text{ini}}, \P_{\text{ini}})$ is a congugate pair and so $\U_{\text{ini}}(\omega) \equiv \alpha \log(\P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
+  The two terms in \eqref{eq:free} (expected gain in utility or negative workdone, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making.
+They formalize
+a trade-off between an expected utility $\delta \U$ (first term) and
+the information cost of transforming Pi
+into $\P_{\text{fin}}$ (second
+term). In this interpretation $\P_{\text{ini}}$ represents an initial choice
+probability or policy, which includes the special case of the
+uniform distribution where the decision-maker has initially no
+preferences between the different choices. The probability measure $\P_{\text{fin}}$ is the final choice probability that we are looking for since it
+considers the utility constraint U∗
+that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $\P_{\text{fin}}(\omega)$
+\begin{equation}
+\P^*_{\text{fin}} := \underset{\P_{\text{fin}}}{\text{argmax }} \delta J_{(\U_{\text{ini}}, \P_{\text{ini}}) \rightarrow \P_{\text{fin}}}
+  \end{equation}
+
+By differentiating the RHS of \eqref{eq:free} w.r.t $\P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
+
+\begin{equation}
+  \P^*_{\text{fin}}(\omega) \propto \P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\U(\omega)\right).
+\end{equation}
+
+Two limit cases are worth considering.
+\begin{itemize}
+\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\delta J_{(\U_{\text{ini}}, \P_{\text{ini}}) \rightarrow \P_{\text{fin}}} \approx \mathbb E_{\P'_{\text{fin}}}(\delta \U)$, and so it's optimal to take $\P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*)$, where $\omega^* := \argmax_{\omega \in \Omega}\U(\omega)$.
+This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
+  \item \textbf{High-temperature $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $\P^*_{\text{fin}}(\omega) \equiv \P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
+  \end{itemize}
 \section{Experimental Results}

### 2016-07-15 paper.tex
more

--- 2016-07-15 paper.tex
@@
 
-\item
+\item LTPJ:
 
-\item
-
-\item
-
-\item
-
-LTPJ:
-
-\item
-  PMC:
+\item PMC:
   \begin{itemize}
@@
 
-\item
-HC:
+\item HC:
 \begin{itemize}
@@
 
+\subsection{A thermodynamic model for bounded rationality, aka robust optimality}
+\begin{itemize}
+  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
+  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
+  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
+    \end{itemize}
+The material presented is a revisit of \textcolor{blue}{[Braun 2011]}.
+\paragraph{Utility functions and conjugate pairs.}
+  Let $(\Omega, \mathcal F, P)$ be a probablity space.
+A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
+  \begin{itemize}
+  \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
+  \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
+    $A,B \in \mathcal F$.
+  \item monotonicity: $f_{U}$ is strictly increasing.
+    \end{itemize}
+
+\begin{theorem}
+  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
+  \begin{equation}
+    f = \alpha \log(.),
+    \label{eq:boltzmann}
+  \end{equation}
+  where $\alpha > 0$.
+\end{theorem}
+
+\textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
+
+Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
+
+\paragraph{Example.}
+Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
+  \begin{equation}
+    P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
+  \end{equation}
+  where
+  $Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)$ is a normalization constant called the \textit{partition function} of $U$.
+  It's not hard to see that $U$ and the $P$ above form a conjugate pair.
+
+  \paragraph{Free-utility functional.}
+    Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
+Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
+    \begin{equation}
+      J(P'|U, P) = \mathbb E_{P'}(U) + \alpha \mathbb H(P'),
+    \label{eq:free_u}
+    \end{equation}
+    where $\mathbb H(P') := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))$ is the entropy of $P'$.
+
+  It's not difficult to show that
+  \begin{equation}
+    J(P'|U,P) \le J(P|U,P) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
+  \end{equation}
+  In particular, if $P$ is the Gibbs measure corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
+  \begin{equation}
+    J(P'|U,P) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
+    \end{equation}
+
+  \paragraph{The free-energy / utility principle.}
+  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
+Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
+We want to transform this to a new model by adding constraints represented by the utility function $\delta U$.
+The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \delta U$.
+The difference in free-utility is then
+  \begin{equation}
+    \delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\delta U)}_{\textbf{expected gain in utility}} + \alpha \underbrace{D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity of change}},
+    \label{eq:free}
+  \end{equation}
+  where $D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})$ is the Kullback-Leibler divergence.
+  In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
+  The two terms in \eqref{eq:free} (expected gain in utility or negative workdone, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making.
+They formalize
+a trade-off between an expected utility $\delta U$ (first term) and
+the information cost of transforming Pi
+into $P_{\text{fin}}$ (second
+term). In this interpretation $P_{\text{ini}}$ represents an initial choice
+probability or policy, which includes the special case of the
+uniform distribution where the decision-maker has initially no
+preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
+considers the utility constraint U∗
+that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
+\begin{equation}
+P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
+  \end{equation}
+
+By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
+
+\begin{equation}
+  P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}U(\omega)\right).
+\end{equation}
+
+Two limit cases are worth considering.
+\begin{itemize}
+\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \mathbb E_{P'_{\text{fin}}}(\delta U)$, and so it's optimal to take $P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*)$, where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$.
+This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
+  \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
+  \end{itemize}
 \section{Experimental Results}

### 2016-07-15 paper.tex
fixups

@@
 
-  \paragraph{The free-energy / utility principle.}
+  \paragraph{The free-energy / utility principle (of Friston ?).}
   We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
 Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
-We want to transform this to a new model by adding constraints represented by the utility function $\delta U$.
-The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \delta U$.
+We want to transform this to a new model by adding constraints represented by the utility function $\Delta U$.
+The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \Delta U$.
 The difference in free-utility is then
   \begin{equation}
-    \delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\delta U)}_{\textbf{expected gain in utility}} + \alpha \underbrace{D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity of change}},
+    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\Delta U)}_{\textbf{expected gain in utility}} + \alpha \underbrace{D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity of change}},
     \label{eq:free}
@@
 They formalize
-a trade-off between an expected utility $\delta U$ (first term) and
+a trade-off between an expected utility $\Delta U$ (first term) and
 the information cost of transforming Pi
@@
 \begin{equation}
-P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
+P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
   \end{equation}
@@
 \begin{equation}
-  P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}U(\omega)\right).
+  P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
 \end{equation}
@@
 \begin{itemize}
-\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \mathbb E_{P'_{\text{fin}}}(\delta U)$, and so it's optimal to take $P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*)$, where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$.
+\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \mathbb E_{P'_{\text{fin}}}(\Delta U)$, and so it's optimal to take $P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*)$, where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$.
 This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.

### 2016-07-15 paper.tex
fixups

@@
     \end{equation}
-    where $\mathbb H(P') := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))$ is the entropy of $P'$.
-
-  It's not difficult to show that
+    where $\mathbb H(P') := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))$ is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$).
+It's not difficult to establish the upper bound
   \begin{equation}

### 2016-07-15 paper.tex
phrasing

@@
   \begin{equation}
-    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\Delta U)}_{\textbf{expected gain in utility}} + \alpha \underbrace{D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity of change}},
+    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\Delta U)}_{\textbf{accuracy}} + \alpha \underbrace{D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
     \label{eq:free}
@@
   In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
-  The two terms in \eqref{eq:free} (expected gain in utility or negative workdone, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making.
-They formalize
+  The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
 a trade-off between an expected utility $\Delta U$ (first term) and

### 2016-07-16 paper.tex
more fixups

@@
   \begin{equation}
-    J(P'|U,P) \le J(P|U,P) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
+    J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
   \end{equation}
@@
   \begin{equation}
-    J(P'|U,P) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
+    J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
     \end{equation}
@@
   \begin{equation}
-    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\Delta U)}_{\textbf{accuracy}} + \alpha \underbrace{D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
+    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\Delta U)}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
     \label{eq:free}
   \end{equation}
-  where $D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})$ is the Kullback-Leibler divergence.
+  where
+  \begin{equation}
+    D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
+  \end{equation}
+  is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
 In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
@@
 \begin{itemize}
-\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \mathbb E_{P'_{\text{fin}}}(\Delta U)$, and so it's optimal to take $P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*)$, where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$.
-This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
+\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \mathbb E_{P'_{\text{fin}}}(\Delta U)$, and so it's optimal to take
+  $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
+  where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
   \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.

### 2016-07-16 paper.tex
more

@@
 \end{itemize}
+
+To conclude this section, let's note that \textcolor{blue}{[Braun 2011]} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
+For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...
+
 \section{Experimental Results}

### 2016-07-16 paper.tex
more TPJ anat notes

--- 2016-07-15 paper.tex
@@
 \begin{abstract}
-ABC
-%
-
-% OUR keywords
-%\textbf{\\keywords}: curse of dimensionality; semi-supervised learning;
-%fMRI; systems neuroscience
-
-% official NIPS keywords
-\textbf{\\keywords}: Systems Biology, Cognitive Science, Autonomous Learning Systems
-
-\end{abstract}
 
@@
 
-\section{Introduction}
+network that might be involved in the automated prediction of social events that modulate behavior
+
+\end{abstract}
+
+% official NIPS keywords
+\textbf{\\keywords}: Systems Biology, Cognitive Science, Autonomous Learning Systems
+
+\section{Introduction to the Default-Mode Network}
 %
 \paragraph{The Human Default Mode Network.}
+scene construction processes when envisioning past, ficti- tious, and future events (Hassabis et al., 2007; Spreng et al., 2009).
+
+It was observed that the more external stimuli are predictable, the more reflection processes become detached from the actual sensory environment and the more stimulus-independent, self-focused thoughts occur (Mason et al., 2007).
+
 \begin{itemize}
@@
   DMN impairs future and hypothetical thinking (cf. Hassabis PNAS)
-  \item a lesion study on disturbed sleep (i.e., a state of mind
-  independent of sensory stimulation but dependent on
-  internally generated information) exclusively identified
-  the dmPFC (Koenigs et al., 2010). Third, another VLSM study exclusively
-  related the IFG and TPJ, both more strongly connected to the
-  dmPFC in our study, to inner speech
-  \item Bálint's syndrome: neurological disorder of conscious
-  awareness resulting from damage in bilateral parietal cortex;
-  "psychic paralysis of gaze", that is, the inability for internally
-  motivated, targeted saccades hindering fixation of currently not
-  attended features in the field of view (i.e., oculomotor apraxia).
-  unable to bind various individual features of the visual
-  environment into an integrated whole (i.e., simultanagnosia)
-  as well as unable to navigate hand movement to a targeted
-  object helped by vision (i.e., optic ataxia).
-  difficulty avoiding running into objects placed ahead;
-  challenged in estimating distances as well as perceiving
-  spatial depth, object sizes and object orientation
+
+  \item
+  autobiographical memory sup- plies numerous building blocks of social semantic knowl- edge (Bar 2007; Binder et al.
+2009).
+These isolated conceptual scripts may be reassembled to enable the fore- casting of future events (Tulving 1983, 1985; Schacter et al.
+2007).
+Similar brain mechanisms in remembering past episodes and envisioning future circumstances is supported by their engagement of identical brain areas, as evidenced by a quantitative meta-analysis (Spreng et al.
+2009).
+Moreover, retrograde amnesic patients were repor- ted to be impaired not only in prospection but also in imagining novel experience (Hassabis et al.
+2007).
+These findings suggest a single neural network for mentally constructing plausible semantic scenarios of detached sit- uations regardless of temporal orientation (Buckner and Carroll 2007; Hassabis and Maguire 2007).
+Indeed, con- struction of detached probabilistic scenes has been argued to influence ongoing decision making by estimating out- comes of behavioral choices (Boyer 2008; Suddendorf and Corballis 2007; Schilbach et al.
+2008b).
 \end{itemize}
@@
 mPFC and PCC = highest metabolic consumption in the brain/highest level of basal glucose energy consumption in humans
+
+PFC
+Patients with prefrontal lesions have problems in adapting to novel situations and stimuli
+Stuss, D. T. & Benson, D. F. The Frontal Lobes (Raven, New York, 1986).
 
@@
 \item
+\item Bálint's syndrome: neurological disorder of conscious
+awareness resulting from damage in bilateral parietal cortex;
+"psychic paralysis of gaze", that is, the inability for internally
+motivated, targeted saccades hindering fixation of currently not
+attended features in the field of view (i.e., oculomotor apraxia).
+unable to bind various individual features of the visual
+environment into an integrated whole (i.e., simultanagnosia)
+as well as unable to navigate hand movement to a targeted
+object helped by vision (i.e., optic ataxia).
+difficulty avoiding running into objects placed ahead;
+challenged in estimating distances as well as perceiving
+spatial depth, object sizes and object orientation
+
+  \item
     vPCC:
@@
 -> medites between the Default-Mode function and other more environmentally directed processing goals.
+inducing network reconfiguration
 
-\item LTPJ:
+\item
+blanke:
+the core region of the human vestibular cortex is situated close to the angular gyrus10.
+It is possible that the experience of dissociation of self from the body is a result of failure to integrate complex somatosensory and vestibular information.
+Initial stimulations (n 3; 2.0–3.0 mA) induced vestibular responses, in which the patient reported that she was “sinking into the bed” or “falling from a height”.
 
-\item PMC:
+\item
+high-level social cognition also engenerded expectation violations
+-> that's why coincides with "agency" meta-analytic location
+(Decety 2007):
+suggest that activation in the TPJ during social cognition may therefore rely on a lower-level computational mech- anism involved in generating, testing, and correcting internal predictions about external sensory events.
+
+\item
+Sridharan: RTPJ involved in bottom-up saliency detection
+
+\item
+Corbetta2002:
+ We suggest that the TPJ is also strongly engaged by stimuli that are behaviourally relevant but require a change in the current task set. 
+ -> exogeneously induced state transition?
+ <=> MCMC in HC / PCC -> slow? endogeneous changes in state transitions by the policy and value matrices
+ For example, the alarm at the museum is not part of the task set of listening to the guide’s discussion of Hieronymous Bosch, but it is clearly a behaviourally relevant stimulus.
+
+\item
+visuo- proprioceptive conflict (Balslev et al., 2005), and multi-modal detection of sensory changes (Downar et al., 2000)
+brain areas connected to aRTPJ further responded to all changes of visual, auditory, or tactile stimulation in a multi-modal fMRI study (Downar et al., 2000).
+
+\item
+direct electrical stimulation of the RTPJ during neurosurgery was associated with altered perception and stimulus awareness (Blanke et al., 2002).
+
+\item
+the aRTPJ network is central for sensorimotor control by integrating supramodal stimulus-guided attention and action initia- tion during externally structured tasks.
+
+\item
+Bzdok2013
+
+pRTPJ cluster was also functionally associated with deception tasks
+
+qualify RTPJ as a po- tential switch between exteroceptive and interoceptive mind sets implemented by their networks.
+The reciprocal nature of those mind sets is well illustrated by consistent reports of continuous shifting be- tween externally (i.e., sensory) and internally (i.e., self- and presumably social/memory-) oriented processing in controlled laboratory settings and daily routine (Smallwood et al., 2007).
+
+\item
+the difference between the two task groups could be parsimoniously explained by the required atten- tion to either the external world or self.
+world -> policy matrix driven computation
+self -> value matrix driven computation
+
+\item
+significantly reduced impact of intentions after transient TPJ disruption (Young et al. 2010b)
+
+LTPJ:
+
+\item
+only the left TPJ was associated with language processes, whereas only the right TPJ was functionally associated with attention processes
+
+\item 
+Wernicke's area:
+involved in accessing and integrating world knowlege.
+involved in the comprehension or understanding of written and spoken language Damage caused to Wernicke's area results in receptive, fluent aphasia.
+This means that the person with aphasia will be able to fluently connect words, but the phrases will lack meaning.
+A major impairment of language comprehension, while speech retains a natural-sounding rhythm and a relatively normal syntax.
+Language as a result is largely meaningless.
+difficulty finding words, putting words in the wrong order, substituting sounds and/or words, difficulty understanding spoken utterances, difficulty recognizing some words by sight, difficulty writting letters
+
+\item a lesion study on disturbed sleep (i.e., a state of mind
+independent of sensory stimulation but dependent on
+internally generated information) exclusively identified
+the dmPFC (Koenigs et al., 2010). Third, another VLSM study exclusively
+related the IFG and TPJ, both more strongly connected to the
+dmPFC in our study, to inner speech
+
+\item
+the left and right TPJ further differ according to their cytoarchitectonic borders and gyral pattern (Caspers et al., 2008; Caspers et al., 2006)
+
+\item
+  PMC:
   \begin{itemize}
@@
 
-\item HC:
+\item
+HC:
 \begin{itemize}
@@
 On the other hand, the majority of less active neurons constitute a large reservoir that can be mobilized (7) to precisely distinguish one situation from an- other and incorporate novel ones as distinct.
-\end{itemize}
+
+\item
+damage to even the hippocampus alone is sufficient to cause amnesia +
+envisoning themselves in future and even hypothetical scenarios/mentally constructing new experiences
+
+The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
+
+\item
+Hassabias2007:
+Recollection of these rich autobiographical or episodic memories has been likened to mentally traveling back in time and re-experiencing one’s past (1).
+It has long been known that the hippocampus and related medial temporal lobe structures play a critical role in supporting episodic memory
+
+In fact, episodic memory and imagining or constructing events share striking similarities in terms of the psychological processes engaged (19–21).
+These include imagery (22), sense of presence (1), retrieval of semantic information and multimodal details (23), and narrative structure (22).
+
+visual imagery itself is not believed to be realized in the HC
+
 \end{itemize}
@@
 \begin{equation}
-  n_0(\langle k\rangle, \gamma) \approx \exp\left(-\frac{1}{2}\left(\frac{\gamma-2}{\gamma - 1}\right)\langle k \rangle\right),
+  n_0(\langle k\rangle, \gamma) \approx e^{-\frac{1}{2}\left(\frac{\gamma-2}{\gamma - 1}\right)\langle k \rangle},
 \end{equation}
@@
 \begin{equation}
-  n_0(\langle k\rangle) = \exp\left(-\frac{1}{2}\langle k\rangle\right),
+  n_0(\langle k\rangle) = e^{-\frac{1}{2}\langle k\rangle},
 \end{equation}
@@
     \end{itemize}
-The material presented is a revisit of \textcolor{blue}{[Braun 2011]}.
+
 \paragraph{Utility functions and conjugate pairs.}
-  Let $(\Omega, \mathcal F, P)$ be a probablity space.
-A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
+  Let $(\Omega, \mathcal F, \P)$ be a probablity space.
+A function $\U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $\U(A|B) := \U(A \cap B) - \U(B)$ has the following propertites:
   \begin{itemize}
-  \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
-  \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
+  \item additivity: $\U(A_1 \cap A_2 | B) = \U(A_1|B) + \U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
+  \item statistic: there exists a function $f_{\U} :\mathbb R_+ \rightarrow \mathbb R$ such that $\U(A|B) = f(\P(A|B))$, for all events
     $A,B \in \mathcal F$.
-  \item monotonicity: $f_{U}$ is strictly increasing.
+  \item monotonicity: $f_{\U}$ is strictly increasing.
     \end{itemize}
@@
 \begin{theorem}
-  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
+  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $\U(A|B) \equiv f(\P(A|B))$ any probability space $(\Omega, \mathcal F, \P)$ and utility function $\U$ thereupon are of the form
   \begin{equation}
@@
 
-Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
+Such $\U$ and $\P$ are said to form a \textit{conjugate pair} at temperator $\alpha$.
 
 \paragraph{Example.}
-Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
+Given utility function $U$ a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-\U(\omega))_{\omega \in \Omega}$ is defined to be probability measure (on thesame measurable space)
   \begin{equation}
-    P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
+    \P_{\U}(A) = \sum_{\omega \in A}\frac{1}{Z(\alpha)}\exp\left(\frac{1}{\alpha}\U(\omega)\right), \; \forall A \in \mathcal F.
   \end{equation}
-  where
-  $Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)$ is a normalization constant called the \textit{partition function} of $U$.
-  It's not hard to see that $U$ and the $P$ above form a conjugate pair.
+$Z(\alpha)$ is a normalization constant such that $\P_{\U}(\Omega) = 1$.  It's not hard to see that $\U$ and $\P_{\U}$  form a conjugate pair.
 
   \paragraph{Free-utility functional.}
-    Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
-Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
+    Let $(\U, \P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
+Given another probability measure $\P'$ on the same space, define it's \textit{free utility} as
     \begin{equation}
-      J(P'|U, P) = \mathbb E_{P'}(U) + \alpha \mathbb H(P'),
+      J(\P'|\U, \P) = \mathbb E_{\P'}(\U) + \alpha \mathbb H(\P'),
     \label{eq:free_u}
     \end{equation}
-    where $\mathbb H(P') := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))$ is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$).
-It's not difficult to establish the upper bound
+    where $\mathbb H(\P') := -\sum_{\omega \in \Omega}\P'(\omega)\log(\P'(\omega))$ is the entropy of $\P'$.
+
+  It's not difficult to show that
   \begin{equation}
-    J(P'|U,P) \le J(P|U,P) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
+    J(\P'|\U,\P) \le J(\P|\U,\P) = \sum_{\omega \in \Omega}\U(\omega) =: \U(\Omega).
   \end{equation}
-  In particular, if $P$ is the Gibbs measure corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
+  In particular, if $\P = \P_{\U}$ the Gibbs measure, then the upper bound above reduces to the \textit{log-partition function}
   \begin{equation}
-    J(P'|U,P) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
+    J(\P'|\U,\P) \le \U(\Omega) = -\alpha \log\left(\sum_{\omega \in \Omega}e^{\frac{1}{\alpha}\U(\omega)}\right).
     \end{equation}
 
-  \paragraph{The free-energy / utility principle (of Friston ?).}
-  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
-Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
-We want to transform this to a new model by adding constraints represented by the utility function $\Delta U$.
-The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \Delta U$.
+  \paragraph{The free-energy / utility principle.}
+  We are now in shape to introduce the notion of free-energy for transition of priors, and a variational principle for optimizing it.
+Consider thus an initial system described by a conjugate pair $(\U_{\text{ini}}, \P_{\text{ini}}) \rightarrow \P_{\text{fin}}$ at temperature $\alpha > 0$.
+We want to transform this to a new one by adding constraints represented by the utility function $\delta \U$.
+The resulting system has final utility $\U_{\text{fin}} = \U_{\text{ini}} + \delta\U$.
 The difference in free-utility is then
   \begin{equation}
-    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\Delta U)}_{\textbf{accuracy}} + \alpha \underbrace{D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
+    \delta J = J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{\P_{\text{fin}}}(\delta \U)}_{\textbf{expected gain in utility}} + \alpha \underbrace{D_{\text{KL}}(\P_{\text{ini}}\|\P_{\text{fin}})}_{\textbf{complexity of change}},
     \label{eq:free}
   \end{equation}
-  where $D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})$ is the Kullback-Leibler divergence.
-  In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
-  The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
-a trade-off between an expected utility $\Delta U$ (first term) and
-the information cost of transforming Pi
-into $P_{\text{fin}}$ (second
-term). In this interpretation $P_{\text{ini}}$ represents an initial choice
-probability or policy, which includes the special case of the
-uniform distribution where the decision-maker has initially no
-preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
-considers the utility constraint U∗
-that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
-\begin{equation}
-P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
-  \end{equation}
+  where $D_{\text{KL}}(\P_{\text{ini}}\|\P_{\text{fin}})$ is the Kullback-Leibler divergence.
+  In the above formula, we've extensively used the fact that $(\U_{\text{ini}}, \P_{\text{ini}})$ is a congugate pair and so $\U_{\text{ini}}(\omega) \equiv \alpha \log(\P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
+  The two terms in \eqref{eq:free} (expected gain in utility or negative workdone, and the complexity of the transition)
 
-By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
-
-\begin{equation}
-  P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
-\end{equation}
-
-Two limit cases are worth considering.
-\begin{itemize}
-\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \mathbb E_{P'_{\text{fin}}}(\Delta U)$, and so it's optimal to take $P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*)$, where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$.
-This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
-  \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
-  \end{itemize}
 \section{Experimental Results}
@@
 \section{Discussion and Conclusion}
+
+formalize and predict
 

### 2016-07-16 paper.tex

--- 2016-07-16 paper.tex
@@
 -> medites between the Default-Mode function and other more environmentally directed processing goals.
+inducing network reconfiguration
+
+\item
+blanke:
+the core region of the human vestibular cortex is situated close to the angular gyrus10.
+It is possible that the experience of dissociation of self from the body is a result of failure to integrate complex somatosensory and vestibular information.
+Initial stimulations (n 3; 2.0–3.0 mA) induced vestibular responses, in which the patient reported that she was “sinking into the bed” or “falling from a height”.
+
+\item
+high-level social cognition also engenerded expectation violations
+-> that's why coincides with "agency" meta-analytic location
+(Decety 2007):
+suggest that activation in the TPJ during social cognition may therefore rely on a lower-level computational mech- anism involved in generating, testing, and correcting internal predictions about external sensory events.
+
+\item
+Sridharan: RTPJ involved in bottom-up saliency detection
+
+\item
+Corbetta2002:
+ We suggest that the TPJ is also strongly engaged by stimuli that are behaviourally relevant but require a change in the current task set. 
+ -> exogeneously induced state transition?
+ <=> MCMC in HC / PCC -> slow? endogeneous changes in state transitions by the policy and value matrices
+ For example, the alarm at the museum is not part of the task set of listening to the guide’s discussion of Hieronymous Bosch, but it is clearly a behaviourally relevant stimulus.
+
+\item
+visuo- proprioceptive conflict (Balslev et al., 2005), and multi-modal detection of sensory changes (Downar et al., 2000)
+brain areas connected to aRTPJ further responded to all changes of visual, auditory, or tactile stimulation in a multi-modal fMRI study (Downar et al., 2000).
+
+\item
+direct electrical stimulation of the RTPJ during neurosurgery was associated with altered perception and stimulus awareness (Blanke et al., 2002).
+
+\item
+the aRTPJ network is central for sensorimotor control by integrating supramodal stimulus-guided attention and action initia- tion during externally structured tasks.
+
+\item
+Bzdok2013
+
+pRTPJ cluster was also functionally associated with deception tasks
+
+qualify RTPJ as a po- tential switch between exteroceptive and interoceptive mind sets implemented by their networks.
+The reciprocal nature of those mind sets is well illustrated by consistent reports of continuous shifting be- tween externally (i.e., sensory) and internally (i.e., self- and presumably social/memory-) oriented processing in controlled laboratory settings and daily routine (Smallwood et al., 2007).
 
 \item LTPJ:
+
+\item
+the difference between the two task groups could be parsimoniously explained by the required atten- tion to either the external world or self.
+world -> policy matrix driven computation
+self -> value matrix driven computation
+
+\item
+significantly reduced impact of intentions after transient TPJ disruption (Young et al. 2010b)
+
+LTPJ:
+
+\item
+only the left TPJ was associated with language processes, whereas only the right TPJ was functionally associated with attention processes
+
+\item 
+Wernicke's area:
+involved in accessing and integrating world knowlege.
+involved in the comprehension or understanding of written and spoken language Damage caused to Wernicke's area results in receptive, fluent aphasia.
+This means that the person with aphasia will be able to fluently connect words, but the phrases will lack meaning.
+A major impairment of language comprehension, while speech retains a natural-sounding rhythm and a relatively normal syntax.
+Language as a result is largely meaningless.
+difficulty finding words, putting words in the wrong order, substituting sounds and/or words, difficulty understanding spoken utterances, difficulty recognizing some words by sight, difficulty writting letters
+
+\item a lesion study on disturbed sleep (i.e., a state of mind
+independent of sensory stimulation but dependent on
+internally generated information) exclusively identified
+the dmPFC (Koenigs et al., 2010). Third, another VLSM study exclusively
+related the IFG and TPJ, both more strongly connected to the
+dmPFC in our study, to inner speech
+
+\item
+the left and right TPJ further differ according to their cytoarchitectonic borders and gyral pattern (Caspers et al., 2008; Caspers et al., 2006)
 
@@
 
+formalize and predict
+
 \paragraph{Acknowledgment}

### 2016-07-16 paper.tex
impose structure

@@
 %
-\paragraph{The Human Default Mode Network.}
 scene construction processes when envisioning past, ficti- tious, and future events (Hassabis et al., 2007; Spreng et al., 2009).
@@
 
-\paragraph{The Predictive Coding Framework.}
+The Predictive Coding Framework.
 \begin{itemize}
@@
 
+\section{Introduction to Control Theory and Reinforcement Learning}
 \paragraph{Control Theory.}
@@
 
-\paragraph{States and Actions in Neuroscience}
-
-state-dependent variables such as hunger, thirst and
-even wealth
-
-\paragraph{Anatomy}
+\section{The Neurobiological Properties of the Default-Mode Network}
  involvement of the PMC across the boundaries of classical cognitive domains
@@
 \item
+covert reallocation of spatial attention (Gitelman et al., 1999), mediation between internal and external focus (Leech and Sharp, 2014), computation of environmental statistics (Pearson et al., 2009), and self-referential visuospatial imagery (Cavanna and Trimble, 2006)
+
+\item
 single-cell recordings in the monkey PCC demonstrated this brain region's sensitivity to subjective target utility (McCoy and Platt, 2005) and integration across individual outcomes in decision making (Pearson et al., 2009).
@@
 Similarly, the human vPCC was the only PMC region connected to the laterobasal (rather than centromedial or superficial) nuclei group of the amygdala (Bzdok et al., 2013b), which is an amygdalar subregion probably devoted to continuously scanning environmental input for biological significance (Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011; Ghods-Sharifi et al., 2009; LeDoux, 2007).
-\end{itemize}
 
-\item
 \item Bálint's syndrome: neurological disorder of conscious
@@
 This view is in line with the well-known integration of the dPCC in the dorsal visual stream—the “where” system for spatial processing (Ungerleider and Haxby, 1994; Ungerleider and Mishkin, 1982; Vogt et al., 2006), in contrast to vPCC's relation to the ventral “what” system
+\end{itemize}
 
+\item RSC:
+  \begin{itemize}
 \item
-RSC:
 Regarding memory processing, lesions in the RSC, as compared to other PMC regions, is most consistently associated with anterograde and retrograde memory impairments of different types of sensory infor- mation in rabbits (Gabriel and Talk, 2001) and humans (Rudge and Warrington, 1991; Valenstein et al., 1987).
@@
 \item
-BOTH TPJ:
+the left and right TPJ further differ according to their cytoarchitectonic borders and gyral pattern (Caspers et al., 2008; Caspers et al., 2006)
 
+\item
+Neuroscientific investigations on hemispheric specialization have converged to the left cerebral hemisphere as dominant for language functions (Broca 1865, Wernicke 1874, Lichtheim 1885) and the right hemisphere as dominant for attention functions (Gazzaniga, Bogen et al.
+1965, Sperry 1982, Stephan, Marshall et al.
+2003).
+  \end{itemize}
+
+  \item
     RTPJ:
+    \begin{itemize}
 \item
@@
 The reciprocal nature of those mind sets is well illustrated by consistent reports of continuous shifting be- tween externally (i.e., sensory) and internally (i.e., self- and presumably social/memory-) oriented processing in controlled laboratory settings and daily routine (Smallwood et al., 2007).
+\end{itemize}
 
-\item LTPJ:
-
+\item
+  LTPJ
+  \begin{itemize}
 \item
@@
 
-LTPJ:
+\item
+complex conceptual linguistic functions are proposed to be more strongly left-lateralized -> in line with overall left-lateralization of human language
+
+\item
+Episodic and explicit memory retrieval strongly draws on complex semantic processing and contributes to social cognitive processes (see section 4.4.).
+Syntactic processing, on the other hand, is a core language process that is closely intermingled with semantic processing.
+It refers to the hierarchical sequencing of words and their meanings (Price, 2010) and is mandatory for sentence processing in both social cognitive and language tasks.
 
@@
 dmPFC in our study, to inner speech
-
-\item
-the left and right TPJ further differ according to their cytoarchitectonic borders and gyral pattern (Caspers et al., 2008; Caspers et al., 2006)
-
-\item PMC:
-  \begin{itemize}
-    \item
-covert reallocation of spatial attention (Gitelman et al., 1999), mediation between internal and external focus (Leech and Sharp, 2014), computation of environmental statistics (Pearson et al., 2009), and self-referential visuospatial imagery (Cavanna and Trimble, 2006)
 \end{itemize}
@@
 
+  \end{itemize}
 \end{itemize}
@@
 
-\section{Experimental Results}
-\paragraph{Serial versus parallel structure discovery and classification.}
+\section{Relation to Other Cognitive Models (Danilo)}
+\paragraph{Predictive Coding & Bayesian Brain Hypothesis}
 
-\section{Discussion and Conclusion}
+\paragraph{Semantic Hypothesis}
+
+\paragraph{Sentinel Hypothesis}
+
+\section{Relation to Other Statistical Models (Elvis)}
+\paragraph{Control Theory}
+
+\paragraph{Q-Learning}
+
+\paragraph{Free Energy Principle}
+
+\section{Conclusion}
 

### 2016-07-16 paper.tex
fixed
added ToC
fixup
small stuff
new stuff

@@
 
-\section{Introduction to the Default-Mode Network}
+\tableofcontents
+
+\section{The Default-Mode Network}
 %
-scene construction processes when envisioning past, ficti- tious, and future events (Hassabis et al., 2007; Spreng et al., 2009).
+When left unperturbed, the human brain is not at rest.
+Rather, the brain continues to metabolize large quantities of oxygen and glucose energy to maintain inter-neuronal information transfer in the absence of an externally structured task (Bzdok and Eickhoff, 2016a).
+This baseline energy demand is subject to surprisingly little modulations due to the cognitive load posed by environmental challenges.
+What has early been described as the "stream of consciousness" in psychology (James, 1890).
+%
+A few years ago, the so-called "default mode network" was discovered entirely by accident (Shulman et al., 1997).
+A coherent set of brain regions consistently increased in neural activity during stimulus-independent thought, which was proposed to reflect the neural correlates underlying unfocused everyday mind wandering (Raichle et al., 2001).
+That is, in the beginning of the 21st century, brain imaging was the first technology to allow for the discovery of a coherent brain network that subserves a set of baseline mental activity (Bzdok et al., 2012; IBzdok et al., 2015).
+%
+15 years ago neuroscientists have discovered an energy-demanding brain network with unique properties.
+The "default-mode network" (DMN) appeared exclusive in task-induced deactivation during various psychological experiments.
+It was later argued to be systematically anti-correlated with brain regions that subserve task performance.
+Today, many authors speculate that the DMN implements continuous thinking about others’ and one’s own mind states, as well as adaptive envisioning of past, hypothetical, and future events.
+This network might have emerged to continuously predict environmental events using mental imagery as an evolutionary advantage.
+%
+Since 2001, the neurobiological properties of the default mode network have been investigated and reported in more than 3,000 neuroimaging publications 
+The DMN was initially believed to represent the neural correlates of unconstrained mind-wandering because of its pervasively observed activity decrease during a large array of tasks.
+This processing of unknown information categories in the DMN has been argued to mediate an evolutionarily conserved function for the individual.
+This is all the more likely because the DMN contains the two biggest hotspots of energy consumption in the entire central nervous system.
+DMN activity also persists to a substantial degree during the early stages of sleep (Horovitz et al., 2008) and under anesthesia (Greicius et al., 2008).
+However, information processing in the DMN has also repeatedly been shown to impact human behavior.
+Goal-directed task performance improves with decreased activity in default-mode areas (Weissman et al., 2006) and increased DMN activity is linked to more task-independent, potentially disturbing thoughts (Mason et al., 2007).
+%
+the DMN is likely to control the interplay between perception–action cycles and mental imagery (Bzdok et al., 2013a; Bzdok et al., 2013b).
+In particular, the DMN is likely to subserve a computational mechanism by its connections to highly associative brain areas that is key for human thought across time, splace, and content domains at the interface of external world and self (IBzdok et al., 2015).
+%
+
+scene construction processes when envisioning past, fictitious, and future events (Hassabis et al., 2007; Spreng et al., 2009).
 
@@
 2008b).
+  \item
+  DMN exists in monkeys [73] and rats [74]. 
+  \item clinical research has corroborated default-mode network dysfunction in various psychiatric and neurological disorders
 \end{itemize}
 
-The Predictive Coding Framework.
-\begin{itemize}
-  \item It is one of the most frequently cited hypotheses of default mode function
-  \item Within this model the brain is conceptualized as a Bayesian machine
-  \item Predictive coding is a framework that, in a hierarchical setting,
-  is equivalent to empirical Bayesian inference
-  \item External sensory input is feed-forward/bottom-up processed in the brain
-  \item Their processing is compared against expected input
-  \item The prediction error, i.e. the difference between sensory input
-  and internal prediction, is computed at each level and passed to
-  higher levels via forward connections
-  \item in case of a mismatch, back-projections top-down modulate
-  input processing
-  \item ‘contextual integration’ for top-down modulation of sensorimotor
-  processing by context-specific a-priori information.
-  \item Closely related to Friston's free-energy principle: brain as inference engine
-  biological systems, including brains, must minimize the long-term average of surprise;
-  formulate perception as a constructive process based on internal or generative models
-  \item A generative model of the world
-\end{itemize}
+The DMN poses an unprecedented challenge to our neuroscientific understanding and methodological inventory.
 
-\section{Introduction to Control Theory and Reinforcement Learning}
+Despite ramifications in health and disease, the functional role of the DMN poses an unprecedented challenge to the conventional interpretation strategies and existing methodological arsenal in the neurosciences.
+
+Represents the physiological and pathological instantiation of a human beings' default mental repertoire, the nature of which remains largely obscure.
+
+The DMN has now urged major attention in the neuroscientific community (Raichle et al., 2001, >6600 citations).
+
+\section{Control Theory and Reinforcement Learning}
 \paragraph{Control Theory.}
@@
 Patients with prefrontal lesions have problems in adapting to novel situations and stimuli
-Stuss, D. T. & Benson, D. F. The Frontal Lobes (Raven, New York, 1986).
+Stuss, D. T. \& Benson, D. F. The Frontal Lobes (Raven, New York, 1986).
+
+mechanistic relevance of each node -> we deconstruct
 
@@
 \section{Relation to Other Cognitive Models (Danilo)}
-\paragraph{Predictive Coding & Bayesian Brain Hypothesis}
+\paragraph{Predictive Coding and Bayesian Brain Hypothesis}
+abc
+The Predictive Coding Framework.
+construction of detached probabilistic scenes has been argued to influence ongoing decision-making by estimating outcomes of behavioral choices.
+
+\begin{itemize}
+  \item It is one of the most frequently cited hypotheses of default mode function
+  \item Within this model the brain is conceptualized as a Bayesian machine
+  \item Predictive coding is a framework that, in a hierarchical setting,
+  is equivalent to empirical Bayesian inference
+  \item External sensory input is feed-forward/bottom-up processed in the brain
+  \item Their processing is compared against expected input
+  \item The prediction error, i.e. the difference between sensory input
+  and internal prediction, is computed at each level and passed to
+  higher levels via forward connections
+  \item in case of a mismatch, back-projections top-down modulate
+  input processing
+  \item ‘contextual integration’ for top-down modulation of sensorimotor
+  processing by context-specific a-priori information.
+  \item Closely related to Friston's free-energy principle: brain as inference engine
+  biological systems, including brains, must minimize the long-term average of surprise;
+  formulate perception as a constructive process based on internal or generative models
+  \item A generative model of the world
+\end{itemize}
 
 \paragraph{Semantic Hypothesis}
+abc
 
 \paragraph{Sentinel Hypothesis}
+many investigators speculate that this central neural network reflects the brain’s relentless tracking and prediction of environmental events to adaptively optimize the organism's future action (Schacter et al., 2007).
+
+Cognitive neuroimaging studies linked the DMN to the contemplation of others’ (Mar, 2011) and one’s thoughts (Lombardo et al., 2009) mental states,
+
+DMN does social cognition -> but it exists in rats and monkeys with much more poorly developped social capacities
+
+How can a same neurobiological circuit be equally important for baseline house-keeping functions and specific task performance?
+
+Our computational account can explain why the DMN is implicated in both a goal-directed task and an idlying rest cognitive set?
+-> task is policy/value updates to optmize short-term action / rest updates for mid- and long-term action
 
@@
 \paragraph{Control Theory}
+abc
 
 \paragraph{Q-Learning}
+abc
 
 \paragraph{Free Energy Principle}
+abc
 
 \section{Conclusion}
+What single brain function could be most important for existence and survival of the species?
+From the perspective of the human condition, it remains unclear what the nature and purpose of a domain-overarching computational mechanism implemented in the high association cortex could be to warriant the high energetic costs.
 
-formalize and predict
+; formalize and predict; offer statistical control on the brain's default function/operations
 

### 2016-07-30 paper.tex
more
finished: notes for each DMN node to a model sub-task

@@
 
+The model reinterprets many known effects
+
 If this computational models hold some truth in them, the DMNs conceivable key role in the continuous environmental tracking in a generative, integrative process might explain both its highest energy consumption in the brain and its intimate coupling with conscious awareness.
@@
 network that might be involved in the automated prediction of social events that modulate behavior
+
+We focus not on normative models (that describe what the brain does), but on process models (that describe how the brain does it)
 
@@
 %
+
+hierarchically deep models of the world
+
+sensory exchanges with the world
+
+show how a formal approach can provide generic explanations for psycho- pathology that are physiologically grounded
+
+quantitative (and parametric) characterisations offered by computational
+
+d Biological systems are homeostatic (or allostatic), which
+means that they minimize the dispersion (entropy) of their
+interoceptive and exteroceptive states.
+d Entropy is the average of surprise over time, which means
+that biological systems minimize the surprise associated
+with their sensory states at each point in time.
+d In statistics, surprise is the negative logarithm of B
+
+perception as hypothesis testing
+
+Consider an organism that, in their present situation, is confronted by several choices of what to do next.
+Being able to accurately and richly mentally enact possible future states before making a decision would help to evaluate the desirability of different outcomes and also the planning processes needed to make them happen.
+
+ the use of this constructive process goes far beyond simply predicting the future, to the general evaluation of fitness for purpose
 
@@
 
+vmPFC regions respond more to the value difference that will be acted on in the current trial (executed in red), and dmPFC to the currently irrelevant value difference (modeled in blue) {Nicolle, 2012 #3750} individual.
+Instead activity in vmPFC reflects a choice preference that is executed and rostral dmPFC a choice preference that is modeled.
+Thus, dorsal regions of rostral mPFC contain value representations in the frame of reference of a modeled (as opposed to an actuated) preference irrespective of whether this applies to another’s or to one’s own likely actions and goals.
+Likewise, ventral mPFC contains a representation of value in the frame of reference of an executed choice, even if this executed choice reflects one’s own or another’s preferences.
+dmPFC that reflects the values and preferences of another individual (here temporally discounted at a rate specific to the individual), even when they are not directly relevant to the task at hand.
+Critically, we also show this activity.
+While the simplest interpretation of this effect is that the region is simulating one’s own, currently irrelevant, preferences an alternative possibility is that the activity is projec- ting one’s own values into the mind of the partner, while simu- lating the partner’s choices.
+In essence, estimating the extent to which my own values would influence my partner, if they were making the choice.
+This iterated reasoning would be.
+
 mechanistic relevance of each node -> we deconstruct
+
+DMN function has been linked to the recollection of past episodes and the simulation of future episodes
+
+Simony 2016
+This implies that the processing timescale of the DMN is longer than the length of single paragraphs (B40s) and that information accumulated from prior paragraphs can affect ongoing processing in this system
 
@@
 the vmPFC might be more closely associated with orches- trating adapted behavior by bottom-up-driven processing of “what matters now”
+
+\item
+the vmPFC to monitoring others’ (Lotze et al., 2007) and one’s own (Lane et al., 1997; Phan et al., 2004) emotional responses, that is, other’s (external) emotional reac- tions and one’s own (visceral) arousal.
+Such real or imagined bodily states, believed to be represented in the vmPFC, proba- bly operate as a bioregulatory disposition governing cognition and decision making (Damasio, 1996; Nauta, 1971), in line with the vmPFC’s functional association with general cognition and reward processing.
 
@@
 As to the vmPFC subnetwork, the GMV of the vmPFC and VS correlated with indices of social reward attitudes and behavior (Lebreton et al., 2009), concur- ring with vmPFC’s relation to the NAc and reward-related tasks.
+
+\item
+In studies of decision making and neuroeconomics, ventromedial prefrontal cortex (vmPFC) con- sistently reflects an individual’s subjective valuation (Behrens et al., 2008; Boorman et al., 2009; Knutson et al., 2005; Plass- mann et al., 2007).
+By contrast, its dorsomedial neighbor (rostral
+
+\item
+NAC was proposed to subserve model-free reinforcement learning in humans
+(O'Doherty 2015)
 
@@
 dmPFC subserves predominantly ambiguous amodal metacogni- tive processes driven by top-down pathways
+
+\item
+A class of neurons in macaque dorsomedial frontal cortex, including the presupplementary area, responds preferentially to the actions that another macaque makes rather than their own actions (Yoshida et al., 2011)
+Many of the same neurons differentiate between another macaque’s actions that were unsuccessful, because the circumstances changed, and actions that were unsuccessful because the other macaque did not make the best choice given the circum- stances.
+The macaques adjusted their own subsequent actions in different ways after these two types of obser- vations.
+There is also activity in human dorsomedial prefrontal cortex that reflects expectations about what an observed person will do and errors in such predictions (Suzuki et al., 2012; van Schie et al., 2004)
+
 \item
@@
 \item
+hippocampal damage [21,38–40] is associated with an impairment in thinking about one’s personal future
+
+\item
+They found that amnesic patients with bilateral hippocampal damage were markedly impaired relative to matched control subjects at imagining new fictitious experiences (Figure 1).
+Moreover, Hassabis et al.
+[21] identified a possible source for this deficit.
+Although patients were able to produce considerable detail when asked to imagine fictitious new experiences, their descriptions lacked spatial coherence and were instead fragmented and considerably less rich [21].
+
+\item
+ empirical evidence hinting at a two-process function of the hippocampus comes from structural MRI studies of expert navigators (London taxi drivers) who show increased grey matter volume in posterior hippocampus seemingly at the expense of reduced grey matter volume in anterior hippocampus (Maguire et al.
+2006b).
+Moreover, their increased spatial knowledge appears to come at a cost to the acquisition of new visual associative information (Maguire et al.
+2006b; Woollett & Maguire 2009).
+
+\item
+Schachter/Buckner:
+The network described here is tied to the medial temporal lobe system, which is traditionally considered almost exclusively in the context of remembering the past. We suspect the adaptive value of
+episodic memory is not solely in its ability to afford mental reconstruction of the past but rather in its contribution to building mental models – simulations – of what might happen next or other perspectives on the immediate envir- onment, such as what others are thinking (see Ref.
+[63]).
+As
+
+\item
 the GMV of the entorhinal cortex (connectionally and functionally closely coupled with the HC) correlated with social network size (Kanai et al., 2012)
@@
   where
-  $Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)$ is a normalization constant called the \textit{partition function} of $U$.
+
+  \begin{equation}
+    Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
+  \end{equation}
+  is a normalization constant called the \textit{partition function} of $U$.
   It's not hard to see that $U$ and the $P$ above form a conjugate pair.
@@
 
+Friston2014:
+
+This perspective shifts away from the brain as a passive filter of sensations (or an elaborate stimulus– response link) towards a view of the brain as a statistical organ that generates hypotheses or fantasies that are tested against sensory evidence.
+In short, the brain is now considered a phantastic organ (from Greek phantastikos, the ability to create mental images).
+For many people, this perspective can be traced back to Hermann von Helmholtz and the notion of unconscious inference: that is, a pre-rational mechanism by which visual impressions are formed (eg, the seemingly automatic but erroneous belief that the sun rises and sets in the sky, as opposed to the truth that the Earth rotates around it).
+
+Predictive coding is not a normative or descriptive scheme, it is a process theory with a biologically plausible basis—there is now much circumstantial anatomical and physiological evidence for predictive coding in the brain.12–16 In this scheme, neuronal representations in higher levels of cortical hierarchies generate predictions of representations in lower levels.
+These top-down predictions are compared with representations at the lower level to form a prediction
+error (associated with the activity of superficial pyramidal cells).
+This mismatch signal is passed back up the hierarchy, to update higher representations (associated with the activity of deep pyramidal cells).
+
+A description of biological processes as minimising (variational) free energy—a statistical measure of the surprise or improbability of sensory data, under a model of how those data were generated.
+
+predictive coding represents a biologically plausible scheme that enables the brain to update beliefs about the world with sensory samples
+
+ individuals spend most of their time predicting the internal (proprioceptive) and external (exteroceptive) consequences of behaviour (both their own and that of others).
+To fully appreciate the bilateral nature of these predictions, inference can be considered in an embodied context.
+In this setting, perception can be understood as resolving exteroceptive prediction errors by selecting the predictions that best explain sensations.
+Conversely, behaviour suppresses pro- prioceptive prediction error by changing proprioceptive sensations.
+This suppression relies on classic reflexes, in which equilibrium points are set by descending proprioceptive predictions (figure 1).15 This process is called active inference
+
+providing a rich dynamical repertoire that enables the brain to respond quickly to changing inputs
+
+the formal constraints implicit in predictive coding mandate modulatory gain control for ascending prediction errors
+
 \paragraph{Semantic Hypothesis}

### 2016-08-01 paper.tex
fixed broken latex...

@@
 
-vmPFC regions respond more to the value difference that will be acted on in the current trial (executed in red), and dmPFC to the currently irrelevant value difference (modeled in blue) {Nicolle, 2012 #3750} individual.
+vmPFC regions respond more to the value difference that will be acted on in the current trial (executed in red), and dmPFC to the currently irrelevant value difference (modeled in blue) {Nicolle, 2012 \#3750} individual.
 Instead activity in vmPFC reflects a choice preference that is executed and rostral dmPFC a choice preference that is modeled.
@@
 Moreover, their increased spatial knowledge appears to come at a cost to the acquisition of new visual associative information (Maguire et al.
-2006b; Woollett & Maguire 2009).
+2006b; Woollett \& Maguire 2009).
 

### 2016-08-04 paper.tex
added a unified free-energy principle based on helmholtz machines (hinto et al. 1995, etc.)

@@
 \usepackage{amsmath,amsfonts,amsthm}
+\usepackage{mdframed}
 \usepackage{bbm}
@@
 \section{Methods}
+
 \subsection{Linear control network model for brain organization}
@@
 \end{equation}
-\subsection{Minimal control energy controller}
-\subsection{Related works}
+%% \subsection{Minimal control energy controller}
+\subsubsection{Related works}
 The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
 
-\subsection{A thermodynamic model for bounded rationality, aka robust optimality}
-\begin{itemize}
-  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
-  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
-  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
-    \end{itemize}
-The material presented is a revisit of \textcolor{blue}{[Braun 2011]}.
-\paragraph{Utility functions and conjugate pairs.}
-  Let $(\Omega, \mathcal F, P)$ be a probablity space.
+\subsection{Demystifying the free-energy principle}
+In this section we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision making.
+This ideas were first developed by Hinton et al.
+in the early 90s in the form of the so-called  \textit{Helmholtz machine}.
+Theories like Friston's free-energy principle will then emerge as particular instanstances of this general framework, with particular design choises.
+For example the Friston theory assumes the brain uses a (problematic) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
+\begin{table}[H]
+  \begin{tabular}{p{2cm}|p{11cm}}
+         \hline
+         \textbf{variable}    & \textbf{short description}  \\ \hline
+             $\langle E\rangle_p$ & Expectation (a.k.a average) of the random quantity $E$
+         .r.t to the probability density $p$.\\ \hline
+         $\mathcal H$ & The usual information-theoretic entropy operator.\\ \hline
+         $D_{KL}(p||q)$ & The Kullback-Leibler divergence between the probability densities $p$ and $q$. \\ \hline
+             $\zeta$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's hidden state $b$ and sensory inputs $s$, i.e $\zeta = (s, b).$ \\ \hline
+             $\psi$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
+             $p_G(.|\zeta)$ & Generative density for ...\\ \hline
+         $p_R(.|\zeta)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
+         $F_G(\zeta)$ & Helmholtz free-energy for the model $p_G$ of generating the observation $\zeta$.\\ \hline
+         $F^R_G(\zeta)$ & Variational Helmholtz free-energy from $R$ to $G$.  Note that $F^G_G = F_G$.\\ \hline
+  \end{tabular}
+  \caption{Table of notations.}
+\end{table}
+\subsubsection{Helmholtz free-energy and the generative model}
+Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
+So,
+\begin{eqnarray}
+  \begin{split}
+    \text{generative surprise } &= -\log(p_G(\zeta)) = -\log(p_G(\zeta)) \times 1 = -\log(P_G(\zeta))\sum_{\psi}p_G(\psi |\zeta)\\
+    &= -\sum_{\psi}p_G(\psi, \zeta)\log(p_G(\zeta))
+    =-\sum_{\psi}p_G(\psi |\zeta)\log(p_G(\psi, \zeta)/p_G(\psi|\zeta))\\
+    &= \sum_{\psi}p_G(\psi |\zeta)\log(p_G(\psi|\zeta))-\sum_{\psi}p_G(\psi |\zeta)\log  (p_G(\psi, \zeta))\\
+    &= -\langle \log  (p_G(., \zeta)) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta))\\
+    &= \langle E_G(., \zeta) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta)) := F_G(\zeta)
+  \end{split}
+  \label{eq:helm}
+\end{eqnarray}
+where $E_G(\psi, \zeta)$ is the energy of an fictive thermodynamic system defined by setting
+\begin{equation}
+  E_G(\psi, \zeta) := -\log(p_G(\psi, \zeta)).
+  \label{eq:gibbs}
+\end{equation}
+%% In the above, the variable $\psi$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
+The last quantity $F_G(\zeta) := \langle E_G(., \zeta) \rangle_{p_G(.
+|\zeta)} - \mathcal H(p_G(.
+|\zeta))$ in \eqref{eq:helm} can be is nothing other than the Helmholtz free-energy for the physical system correspond to the external world, prescribed by the generative density $p_G(\psi|\zeta)$ for the occupation of \textit{macrostates} $\psi$.
+Thus generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
+
+The goal of the agent is then to optimize over the generative model $G$ (i.e over the generative density $p_G(.|s)$, so as to minimize their surprise.
+It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(s)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
+A workaround is then to introduce a second density $p_R(.|\zeta)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\zeta)$: the former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
+This primal-dual idea is at the heart ot the (variational) free-energy principle that we'll introduce shortly.
+
+\subsubsection{Variational Helmholtz free-energy and the recognition density}
+In this subsection we'll develop a very insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
+As an avant-gout of what is to come shorty, let's just that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound of it instead.
+
+Now, invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
+distribution
+\begin{equation}
+  p_G(\psi|\zeta) = \frac{p_G(\psi|\zeta)}{p_G(\zeta)} = \frac{\exp(-E_G(\psi, \zeta))}{Z_G(\zeta)} = \frac{\exp(-E_G(\psi, \zeta))}{Z_G(\zeta)},
+\end{equation}
+where $Z_G(\zeta) := \log(p_G(\zeta)) = \sum_{\psi'}\exp(-E_G(\psi',\zeta))$, the normalizing partition function for the model \ref{eq:gibbs}.
+Whence $\forall \psi$, $Z_G(\zeta) = p_G(\zeta) = \exp(-E_G(\psi, \zeta)) / p_G(\psi|\zeta)$, and so we have the invariance relation
+\begin{equation}
+  -\log(Z_G(\zeta)) = -\log(p_G(\zeta)) = F_G(\zeta) = E_G(\psi, \zeta) + \log(p_G(\psi|\zeta)).
+  \end{equation}
+Now, in the equation above, the LHS only depends on the generative model $G$ and the data point $s$: it doesn't depend on the hidden variable $\psi$, etc.
+So, taking expectations w.r.t an arbitrary density\footnote{The conditioning in $P_R(.|\zeta)$ is because this density is selected from a world in which the sensory inputs and internal brain state vector $\zeta$ is assumed already observed.} $p_R(.|\zeta)$ (the subscrit $R$ stands for ``recognition'', and the terminology will become clear in a moment) yields
+\begin{equation}
+  \begin{split}
+    F_G(\zeta) &= -\log(Z_G(\zeta)) = \langle E_G(., \zeta)\rangle_{P_R(.|\zeta)} + \sum_{\psi}p_R(\psi|\zeta)\log(p_G(\psi|\zeta))\\
+    &= \langle E_G(., \zeta)\rangle_{P_R(.|\zeta)} - \mathcal H(p_R(.|\zeta)) - \sum_{\psi}p_R(\psi|\zeta)\log(p_R(\psi|\zeta)/p_G(\psi|\zeta))\\
+    &= F^R_G(\zeta) - D_{KL}(P_R(.|\zeta) || P_G(.|\zeta)),
+  \end{split}
+  \label{eq:fe}
+\end{equation}
+where $F^R_G(\zeta)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
+\begin{equation}
+  F^R_G(\zeta) := \langle E_G(., \zeta)\rangle_{P_R(.|\zeta)} - \mathcal H(p_R(.|\zeta))
+\end{equation}
+and $D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))$ is the Kullback-Leibler divergence between the $p_R(.|\zeta)$ and the generative density $p_G(.|\zeta)$. Note that $F^G_G = F_G$.
+
+\subsubsection{A general free-energy principle}
+We can resume the situation as follows:
+\begin{mdframed}
+\begin{equation}
+  \begin{split}
+    \text{generative surprise } &:= -\log(p_G(\zeta)) = \underbrace{F_G(\zeta)}_{\text{generative free-energy}} \\
+    &=\underbrace{F^R_G(\zeta)}_{\text{variational free-energy}} - \underbrace{D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))}_{\text{complexity}} \\
+    &\le F^R_G(\zeta),
+    \text{ with equalitity if }p_R(.|\zeta) = p_G(.|\zeta),
+    \end{split}
+\end{equation}
+\end{mdframed}
+where we have used the fact that KL divergence is always nonnegative. In fact, we have the following theorem
+\begin{theorem}
+  It holds that
+  $$
+  \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
+  i.e
+  \begin{equation}
+    \min_{G}F_G(\zeta) = \min_{G,R}F^R_G(\zeta)
+    \end{equation}
+\end{theorem}
+
+\subsubsection{Helmholtz machines}
+...
+\subsection{The wake-sleep algorithm}
+...
+\subsection{Friston's active-inference: an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model for the brain}
+...
+
+%% \subsection{A thermodynamic model for bounded rationality, aka robust optimality}
+%% \begin{itemize}
+%%   \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
+%%   \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
+%%   \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
+%%     \end{itemize}
+%% The material presented is a revisit of \cite{braun2011path}. See also \cite{ortega2013thermodynamics}
+%% \paragraph{Utility functions and conjugate pairs.}
+%%   Let $(\Omega, \mathcal F, P)$ be a probablity space.
 A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
-  \begin{itemize}
-  \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
-  \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
-    $A,B \in \mathcal F$.
-  \item monotonicity: $f_{U}$ is strictly increasing.
-    \end{itemize}
-
-\begin{theorem}
-  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
-  \begin{equation}
-    f = \alpha \log(.),
-    \label{eq:boltzmann}
-  \end{equation}
-  where $\alpha > 0$.
-\end{theorem}
-
-\textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
-
-Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
-
-\paragraph{Example.}
-Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
-  \begin{equation}
-    P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
-  \end{equation}
-  where
-
-  \begin{equation}
-    Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
-  \end{equation}
-  is a normalization constant called the \textit{partition function} of $U$.
-  It's not hard to see that $U$ and the $P$ above form a conjugate pair.
-
-  \paragraph{Free-utility functional.}
-    Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
+%%   \begin{itemize}
+%%   \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
+%%   \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
+%%     $A,B \in \mathcal F$.
+%%   \item monotonicity: $f_{U}$ is strictly increasing.
+%%     \end{itemize}
+
+%% \begin{theorem}
+%%   The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
+%%   \begin{equation}
+%%     f = \alpha \log(.),
+%%     \label{eq:boltzmann}
+%%   \end{equation}
+%%   where $\alpha > 0$.
+%% \end{theorem}
+
+%% \textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
+
+%% Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
+
+%% \paragraph{Example.}
+%% Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
+%%   \begin{equation}
+%%     P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
+%%   \end{equation}
+%%   where
+
+%%   \begin{equation}
+%%     Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
+%%   \end{equation}
+%%   is a normalization constant called the \textit{partition function} of $U$.
+%%   It's not hard to see that $U$ and the $P$ above form a conjugate pair.
+
+%%   \paragraph{Free-utility functional.}
+%%     Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
 Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
-    \begin{equation}
-      J(P'|U, P) = \mathbb E_{P'}(U) + \alpha \mathbb H(P'),
-    \label{eq:free_u}
-    \end{equation}
-    where $\mathbb H(P') := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))$ is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$).
-It's not difficult to establish the upper bound
-  \begin{equation}
-    J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
-  \end{equation}
-  In particular, if $P$ is the Gibbs measure corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
-  \begin{equation}
-    J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
-    \end{equation}
-
-  \paragraph{The free-energy / utility principle (of Friston ?).}
-  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
+%%     \begin{equation}
+%%       J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
+%%     \label{eq:free_u}
+%%     \end{equation}
+%%     where
+%%     \begin{equation}
+%%       \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
+%%     \end{equation}
+%%     is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
+%%   \begin{equation}
+%%     J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
+%%   \end{equation}
+%%   In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
+%%   \begin{equation}
+%%     J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
+%%     \end{equation}
+
+%%   \paragraph{The free-energy / utility principle (of Friston ?).}
+%%   We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
 Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
@@
 The difference in free-utility is then
-  \begin{equation}
-    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\Delta U)}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
-    \label{eq:free}
-  \end{equation}
-  where
-  \begin{equation}
-    D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
-  \end{equation}
-  is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
-In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
-  The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
-a trade-off between an expected utility $\Delta U$ (first term) and
-the information cost of transforming Pi
-into $P_{\text{fin}}$ (second
-term). In this interpretation $P_{\text{ini}}$ represents an initial choice
-probability or policy, which includes the special case of the
-uniform distribution where the decision-maker has initially no
-preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
-considers the utility constraint U∗
-that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
-\begin{equation}
-P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
-  \end{equation}
-
-By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
-
-\begin{equation}
-  P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
-\end{equation}
-
-Two limit cases are worth considering.
-\begin{itemize}
-\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \mathbb E_{P'_{\text{fin}}}(\Delta U)$, and so it's optimal to take
-  $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
-  where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
-  \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
-\end{itemize}
-
-To conclude this section, let's note that \textcolor{blue}{[Braun 2011]} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
+%%   \begin{equation}
+%%     \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
+%%     \label{eq:free}
+%%   \end{equation}
+%%   where
+%%   \begin{equation}
+%%     D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
+%%   \end{equation}
+%%   is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
+%% In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
+%%   The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
+%% a trade-off between an expected utility $\Delta U$ (first term) and
+%% the information cost of transforming Pi
+%% into $P_{\text{fin}}$ (second
+%% term). In this interpretation $P_{\text{ini}}$ represents an initial choice
+%% probability or policy, which includes the special case of the
+%% uniform distribution where the decision-maker has initially no
+%% preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
+%% considers the utility constraint U∗
+%% that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
+%% \begin{equation}
+%% P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
+%%   \end{equation}
+
+%% By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
+
+%% \begin{equation}
+%%   P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
+%% \end{equation}
+
+%% Two limit cases are worth considering.
+%% \begin{itemize}
+%% \item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
+%%   $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
+%%   where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
+%%   \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
+%% \end{itemize}
+
+%% To conclude this section, let's note that \cite{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
 For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...
@@
 \section{Relation to Other Statistical Models (Elvis)}
-\paragraph{Control Theory}
+\paragraph{Control Theory.}
 abc
 
-\paragraph{Q-Learning}
+\paragraph{Reinforcement Learning.}
 abc
 
-\paragraph{Free Energy Principle}
-abc
+\paragraph{Free Energy Principle.}
+The free-energy principle in it's present form (including notions like ``generative density'', ``recognition density'', etc.) can be traceback to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which the introduced the so-called Helmholtz machine...
+
+\begin{itemize}
+  \item Ortega et al. 2013 \cite{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
+\item Friston's critic of RL \cite{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
+pioneered in the 1950s by Richard Bellman and colleagues [2]. To
+optimise control a~pð Þ x~ under this formulation, we have to: (i)
+assume the hidden states are available to the agent and (ii) solve
+Equation 18 for the value-function. Solving for the value-function
+is a non-trivial problem and usually involves backwards induction
+or some approximation scheme like reinforcement-learning [4–6].
+The free-energy formulation circumvents these problems by
+prescribing the policy in terms of free-energy, which encodes
+optimal control (Equation 6)"
+  \end{itemize}
 

### 2016-08-04 paper.tex
fixups

@@
 \subsection{Demystifying the free-energy principle}
-In this section we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision making.
+In this section we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
 This ideas were first developed by Hinton et al.
 in the early 90s in the form of the so-called  \textit{Helmholtz machine}.
-Theories like Friston's free-energy principle will then emerge as particular instanstances of this general framework, with particular design choises.
+Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises.
 For example the Friston theory assumes the brain uses a (problematic) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
@@
          \textbf{variable}    & \textbf{short description}  \\ \hline
-             $\langle E\rangle_p$ & Expectation (a.k.a average) of the random quantity $E$
-         .r.t to the probability density $p$.\\ \hline
+         $\langle E\rangle_p$ & Expectation (a.k.a average) of the
+         random quantity $E$ w.r.t to the probability density $p$.\\ \hline
          $\mathcal H$ & The usual information-theoretic entropy operator.\\ \hline
@@
          $p_R(.|\zeta)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
-         $F_G(\zeta)$ & Helmholtz free-energy for the model $p_G$ of generating the observation $\zeta$.\\ \hline
-         $F^R_G(\zeta)$ & Variational Helmholtz free-energy from $R$ to $G$.  Note that $F^G_G = F_G$.\\ \hline
+         $F_G(\zeta)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\zeta$. This measures the surprise incured upon observing $\zeta$ generated by the model $G$.\\ \hline
+         $F^R_G(\zeta)$ & Variational Helmholtz free-energy from model $R$
+         to $G$.  Note that $F^G_G = F_G$.\\ \hline
   \end{tabular}
@@
   \end{equation}
-Now, in the equation above, the LHS only depends on the generative model $G$ and the data point $s$: it doesn't depend on the hidden variable $\psi$, etc.
-So, taking expectations w.r.t an arbitrary density\footnote{The conditioning in $P_R(.|\zeta)$ is because this density is selected from a world in which the sensory inputs and internal brain state vector $\zeta$ is assumed already observed.} $p_R(.|\zeta)$ (the subscrit $R$ stands for ``recognition'', and the terminology will become clear in a moment) yields
+Now, in the equation above, the LHS only depends on the generative model $G$ and the data point $\zeta$: it doesn't depend on the hidden variable $\psi$, etc.
+So, taking expectations w.r.t an arbitrary density\footnote{The conditioning in $P_R(.|\zeta)$ is because this density is selected from a world in which the sensory inputs and internal brain state vector $\zeta$ is assumed already observed.} $p_R(.|\zeta)$ yields
 \begin{equation}

### 2016-08-04 paper.tex
fixups

@@
 
-\subsection{Demystifying the free-energy principle}
+\subsection{Demystifying the free-energy principle and active-inference}
 In this section we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.

### 2016-08-04 paper.tex
more

@@
 Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises.
-For example the Friston theory assumes the brain uses a (problematic) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
+For example the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial, e.g) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
 \begin{table}[H]
@@
 |\zeta)} - \mathcal H(p_G(.
-|\zeta))$ in \eqref{eq:helm} can be is nothing other than the Helmholtz free-energy for the physical system correspond to the external world, prescribed by the generative density $p_G(\psi|\zeta)$ for the occupation of \textit{macrostates} $\psi$.
+|\zeta))$ in \eqref{eq:helm} is nothing other than the Helmholtz free-energy for the physical system correspond to the external world, prescribed by the generative density $p_G(\psi|\zeta)$ for the occupation of \textit{macrostates} $\psi$.
 Thus generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
@@
 A workaround is then to introduce a second density $p_R(.|\zeta)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\zeta)$: the former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
-This primal-dual idea is at the heart ot the (variational) free-energy principle that we'll introduce shortly.
+This primal-dual idea, first proposed in Hinton et al.
+1995, is at the heart ot the general free-energy principle that we'll introduce shortly.
 
@@
 In this subsection we'll develop a very insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
-As an avant-gout of what is to come shorty, let's just that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound of it instead.
+As an avant-gout of what is to come shorty, let's just that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
 

### 2016-08-04 paper.tex
more typos

@@
          $D_{KL}(p||q)$ & The Kullback-Leibler divergence between the probability densities $p$ and $q$. \\ \hline
-             $\zeta$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's hidden state $b$ and sensory inputs $s$, i.e $\zeta = (s, b).$ \\ \hline
+             $\zeta$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e $\zeta = (s, b).$ \\ \hline
              $\psi$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline

### 2016-08-04 paper.tex
some friston critics

@@
 
-The goal of the agent is then to optimize over the generative model $G$ (i.e over the generative density $p_G(.|s)$, so as to minimize their surprise.
+The goal of the brain is then to optimize over the generative model $G$ (to iteratively or analytically modify the generative density $p_G(.|s)$, so as to minimize their surprise.
 It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(s)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
@@
 \subsubsection{Helmholtz machines}
-...
+% \begin{mdframed}
+  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data. Precisely
+$$
+p_G(\psi^{(l)}|\zeta) = \Pi_{k=1}^{h_l}p_G(\psi_k^{(l)} | \zeta),\hspace{1cm}
+p_R(\psi^{(l)}|\zeta) = \Pi_{k=1}^{h_l}p_R(\psi_k^{(l)} |\zeta)
+$$
+% \end{mdframed}
 \subsection{The wake-sleep algorithm}
 ...
-\subsection{Friston's active-inference: an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model for the brain}
-...
-
-%% \subsection{A thermodynamic model for bounded rationality, aka robust optimality}
-%% \begin{itemize}
-%%   \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
-%%   \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
-%%   \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
-%%     \end{itemize}
-%% The material presented is a revisit of \cite{braun2011path}. See also \cite{ortega2013thermodynamics}
-%% \paragraph{Utility functions and conjugate pairs.}
-%%   Let $(\Omega, \mathcal F, P)$ be a probablity space.
+\subsection{Friston's active-inference}
+This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
+
+The following critics can be made:
+\begin{itemize}
+  \item 
+    As noted by Dayan et al.
+(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
+    \end{itemize}
+
+\subsection{A thermodynamic model for bounded rationality, aka robust optimality}
+\begin{itemize}
+  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
+  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
+  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
+    \end{itemize}
+The material presented is a revisit of \cite{braun2011path}. See also \cite{ortega2013thermodynamics}
+\paragraph{Utility functions and conjugate pairs.}
+  Let $(\Omega, \mathcal F, P)$ be a probablity space.
 A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
-%%   \begin{itemize}
-%%   \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
-%%   \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
-%%     $A,B \in \mathcal F$.
-%%   \item monotonicity: $f_{U}$ is strictly increasing.
-%%     \end{itemize}
-
-%% \begin{theorem}
-%%   The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
-%%   \begin{equation}
-%%     f = \alpha \log(.),
-%%     \label{eq:boltzmann}
-%%   \end{equation}
-%%   where $\alpha > 0$.
-%% \end{theorem}
-
-%% \textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
-
-%% Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
-
-%% \paragraph{Example.}
-%% Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
-%%   \begin{equation}
-%%     P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
-%%   \end{equation}
-%%   where
-
-%%   \begin{equation}
-%%     Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
-%%   \end{equation}
-%%   is a normalization constant called the \textit{partition function} of $U$.
-%%   It's not hard to see that $U$ and the $P$ above form a conjugate pair.
-
-%%   \paragraph{Free-utility functional.}
-%%     Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
+  \begin{itemize}
+  \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
+  \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
+    $A,B \in \mathcal F$.
+  \item monotonicity: $f_{U}$ is strictly increasing.
+    \end{itemize}
+
+\begin{theorem}
+  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
+  \begin{equation}
+    f = \alpha \log(.),
+    \label{eq:boltzmann}
+  \end{equation}
+  where $\alpha > 0$.
+\end{theorem}
+
+\textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
+
+Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
+
+\paragraph{Example.}
+Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
+  \begin{equation}
+    P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
+  \end{equation}
+  where
+
+  \begin{equation}
+    Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
+  \end{equation}
+  is a normalization constant called the \textit{partition function} of $U$.
+  It's not hard to see that $U$ and the $P$ above form a conjugate pair.
+
+  \paragraph{Free-utility functional.}
+    Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
 Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
-%%     \begin{equation}
-%%       J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
-%%     \label{eq:free_u}
-%%     \end{equation}
-%%     where
-%%     \begin{equation}
-%%       \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
-%%     \end{equation}
-%%     is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
-%%   \begin{equation}
-%%     J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
-%%   \end{equation}
-%%   In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
-%%   \begin{equation}
-%%     J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
-%%     \end{equation}
-
-%%   \paragraph{The free-energy / utility principle (of Friston ?).}
-%%   We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
+    \begin{equation}
+      J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
+    \label{eq:free_u}
+    \end{equation}
+    where
+    \begin{equation}
+      \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
+    \end{equation}
+    is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
+  \begin{equation}
+    J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
+  \end{equation}
+  In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
+  \begin{equation}
+    J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
+    \end{equation}
+
+  \paragraph{The free-energy / utility principle (of Friston ?).}
+  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
 Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
@@
 The difference in free-utility is then
-%%   \begin{equation}
-%%     \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
-%%     \label{eq:free}
-%%   \end{equation}
-%%   where
-%%   \begin{equation}
-%%     D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
-%%   \end{equation}
-%%   is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
-%% In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
-%%   The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
-%% a trade-off between an expected utility $\Delta U$ (first term) and
-%% the information cost of transforming Pi
-%% into $P_{\text{fin}}$ (second
-%% term). In this interpretation $P_{\text{ini}}$ represents an initial choice
-%% probability or policy, which includes the special case of the
-%% uniform distribution where the decision-maker has initially no
-%% preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
-%% considers the utility constraint U∗
-%% that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
-%% \begin{equation}
-%% P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
-%%   \end{equation}
-
-%% By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
-
-%% \begin{equation}
-%%   P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
-%% \end{equation}
-
-%% Two limit cases are worth considering.
-%% \begin{itemize}
-%% \item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
-%%   $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
-%%   where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
-%%   \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
-%% \end{itemize}
-
-%% To conclude this section, let's note that \cite{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
+  \begin{equation}
+    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
+    \label{eq:free}
+  \end{equation}
+  where
+  \begin{equation}
+    D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
+  \end{equation}
+  is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
+In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
+  The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
+a trade-off between an expected utility $\Delta U$ (first term) and
+the information cost of transforming Pi
+into $P_{\text{fin}}$ (second
+term). In this interpretation $P_{\text{ini}}$ represents an initial choice
+probability or policy, which includes the special case of the
+uniform distribution where the decision-maker has initially no
+preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
+considers the utility constraint U∗
+that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
+\begin{equation}
+P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
+  \end{equation}
+
+By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
+
+\begin{equation}
+  P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
+\end{equation}
+
+Two limit cases are worth considering.
+\begin{itemize}
+\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
+  $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
+  where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
+  \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
+\end{itemize}
+
+To conclude this section, let's note that \cite{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
 For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...

### 2016-08-05 paper.tex
more

--- 2016-08-04 paper.tex
@@
 \subsection{Demystifying the free-energy principle and active-inference}
+The free-energy principle in it's present form (including notions like ``generative density'', ``recognition density'', etc.) can be traceback to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which the introduced the so-called Helmholtz machine...
+
 In this section we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
 This ideas were first developed by Hinton et al.
-in the early 90s in the form of the so-called  \textit{Helmholtz machine}.
+in the early 90s in the form of the so-called  \textit{Helmholtz machine}, a hierarchical factorial directional deep-belief net.
 Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises.
@@
 \end{eqnarray}
-where $E_G(\psi, \zeta)$ is the energy of an fictive thermodynamic system defined by setting
+where $E_G(\psi, \zeta)$ is the energy of a fictive thermodynamic system defined by setting
 \begin{equation}
@@
 %% In the above, the variable $\psi$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
-The last quantity $F_G(\zeta) := \langle E_G(., \zeta) \rangle_{p_G(.
-|\zeta)} - \mathcal H(p_G(.
-|\zeta))$ in \eqref{eq:helm} is nothing other than the Helmholtz free-energy for the physical system correspond to the external world, prescribed by the generative density $p_G(\psi|\zeta)$ for the occupation of \textit{macrostates} $\psi$.
+The last quantity in \eqref{eq:helm} defined formally as
+\begin{equation}
+  F_G(\zeta) := \langle E_G(., \zeta) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta))
+\end{equation}
+is the \textit{Helmholtz free-energy} of the external world, modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of \textit{macrostates} $\psi$.
 Thus generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
 
-The goal of the agent is then to optimize over the generative model $G$ (i.e over the generative density $p_G(.|s)$, so as to minimize their surprise.
+The goal of the brain is then to optimize over the generative model $G$ (to iteratively or analytically modify the generative density $p_G(.|s)$, so as to minimize their surprise.
 It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(s)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
@@
 where $Z_G(\zeta) := \log(p_G(\zeta)) = \sum_{\psi'}\exp(-E_G(\psi',\zeta))$, the normalizing partition function for the model \ref{eq:gibbs}.
-Whence $\forall \psi$, $Z_G(\zeta) = p_G(\zeta) = \exp(-E_G(\psi, \zeta)) / p_G(\psi|\zeta)$, and so we have the invariance relation
-\begin{equation}
-  -\log(Z_G(\zeta)) = -\log(p_G(\zeta)) = F_G(\zeta) = E_G(\psi, \zeta) + \log(p_G(\psi|\zeta)).
+Whence $\forall \psi$, $p_G(\zeta) = Z_G(\zeta) = \exp(-E_G(\psi, \zeta)) / p_G(\psi|\zeta)$, and so we have the invariance relation
+\begin{equation}
+F_G(\zeta)\overset{\eqref{eq:helm}}{=} -\log(p_G(\zeta)) = -\log(Z_G(\zeta)) = E_G(\psi, \zeta) + \log(p_G(\psi|\zeta)).
   \end{equation}
@@
 \end{mdframed}
-where we have used the fact that KL divergence is always nonnegative. In fact, we have the following theorem
-\begin{theorem}
-  It holds that
+where we have used the fact that KL divergence is always nonnegative.
+%% In fact, we have the following theorem
+%% \begin{theorem}
+%%   It holds that
+%%   $$
+%%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
+%%   i.e
+%%   \begin{equation}
+%%     \min_{G}F_G(\zeta) = \min_{G,R}F^R_G(\zeta)
+%%     \end{equation}
+%% \end{theorem}
+
+\subsubsection{Helmholtz machines}
+% \begin{mdframed}
+  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data. Precisely
 $$
-  \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
-  i.e
-  \begin{equation}
-    \min_{G}F_G(\zeta) = \min_{G,R}F^R_G(\zeta)
-    \end{equation}
-\end{theorem}
-
-\subsubsection{Helmholtz machines}
-...
+p_G(\psi^{(l)}|\zeta) = \Pi_{k=1}^{h_l}p_G(\psi_k^{(l)} | \zeta),\hspace{1cm}
+p_R(\psi^{(l)}|\zeta) = \Pi_{k=1}^{h_l}p_R(\psi_k^{(l)} |\zeta)
+$$
+% \end{mdframed}
 \subsection{The wake-sleep algorithm}
 ...
-\subsection{Friston's active-inference: an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model for the brain}
-...
-
-%% \subsection{A thermodynamic model for bounded rationality, aka robust optimality}
+\subsection{Friston's active-inference}
+This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
+
+The following critics can be made:
+\begin{itemize}
+  \item 
+    As noted by Dayan et al.
+(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
+    \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
+    \end{itemize}
+
+\subsubsection{Free-energy minimization via a stochastic gradient variation bayes estimator}
+\textbf{XXX see Kingma 2014 -- 2014: the poor boy is minimizing FE without knowing!}
+\begin{equation}
+  \begin{split}
+    F_G^R(\zeta) &:= \langle E_G(., \zeta) \rangle_{p_R(. |\zeta)} - \mathcal H(p_R(. |\zeta)) = \langle E_G(., \zeta) - \log(p_R(.|\zeta))\rangle_{P_R(.|\zeta)}\\
+    &=\langle \Delta E_{G \rightarrow R}(., \zeta)\rangle_{P_R(.|\zeta)}
+    \end{split}
+\end{equation}
+
+For fixed $p_G(.|\zeta)$, the optimal choice for $p_R(.|\zeta)$ is given analytically by
+\begin{equation}
+  p_R(\psi|\zeta) \propto p_G(\psi|\zeta)\exp(-\Delta E_{G \rightarrow R}(\psi, \zeta)).
+\end{equation}
+
+%% \subsection{A thermodynamic model for bounded rationality, aka robust
+%%   optimality}
 %% \begin{itemize}
@@
 \paragraph{Free Energy Principle.}
-The free-energy principle in it's present form (including notions like ``generative density'', ``recognition density'', etc.) can be traceback to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which the introduced the so-called Helmholtz machine...
-
 \begin{itemize}

### 2016-08-05 paper.tex
stuf

--- 2016-08-04 paper.tex
@@
 \usepackage{amsmath,amsfonts,amsthm}
-\usepackage{mdframed}
 \usepackage{bbm}
@@
 
-vmPFC regions respond more to the value difference that will be acted on in the current trial (executed in red), and dmPFC to the currently irrelevant value difference (modeled in blue) {Nicolle, 2012 \#3750} individual.
+vmPFC regions respond more to the value difference that will be acted on in the current trial (executed in red), and dmPFC to the currently irrelevant value difference (modeled in blue) {Nicolle, 2012 #3750} individual.
 Instead activity in vmPFC reflects a choice preference that is executed and rostral dmPFC a choice preference that is modeled.
@@
 This implies that the processing timescale of the DMN is longer than the length of single paragraphs (B40s) and that information accumulated from prior paragraphs can affect ongoing processing in this system
+Our results suggest that the DMN accumulates and integrates information over minutes.
+Stimulus-induced correlation patterns in the DMN were largest for the intact story, greatly decreased in the paragraph scramble condition, and were almost entirely abolished in the word scramble condition.
 
@@
 \item
+Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later
+
+\item
 hippocampal damage [21,38–40] is associated with an impairment in thinking about one’s personal future
@@
 Moreover, their increased spatial knowledge appears to come at a cost to the acquisition of new visual associative information (Maguire et al.
-2006b; Woollett \& Maguire 2009).
+2006b; Woollett & Maguire 2009).
 
@@
 \section{Methods}
-
 \subsection{Linear control network model for brain organization}
@@
 \end{equation}
-%% \subsection{Minimal control energy controller}
-\subsubsection{Related works}
+\subsection{Minimal control energy controller}
+\subsection{Related works}
 The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
-
-\subsection{Demystifying the free-energy principle and active-inference}
-In this section we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
-This ideas were first developed by Hinton et al.
-in the early 90s in the form of the so-called  \textit{Helmholtz machine}.
-Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises.
-For example the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial, e.g) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
-\begin{table}[H]
-  \begin{tabular}{p{2cm}|p{11cm}}
-         \hline
-         \textbf{variable}    & \textbf{short description}  \\ \hline
-         $\langle E\rangle_p$ & Expectation (a.k.a average) of the
-         random quantity $E$ w.r.t to the probability density $p$.\\ \hline
-         $\mathcal H$ & The usual information-theoretic entropy operator.\\ \hline
-         $D_{KL}(p||q)$ & The Kullback-Leibler divergence between the probability densities $p$ and $q$. \\ \hline
-             $\zeta$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e $\zeta = (s, b).$ \\ \hline
-             $\psi$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
-             $p_G(.|\zeta)$ & Generative density for ...\\ \hline
-         $p_R(.|\zeta)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
-         $F_G(\zeta)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\zeta$. This measures the surprise incured upon observing $\zeta$ generated by the model $G$.\\ \hline
-         $F^R_G(\zeta)$ & Variational Helmholtz free-energy from model $R$
-         to $G$.  Note that $F^G_G = F_G$.\\ \hline
-  \end{tabular}
-  \caption{Table of notations.}
-\end{table}
-\subsubsection{Helmholtz free-energy and the generative model}
-Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
-So,
-\begin{eqnarray}
-  \begin{split}
-    \text{generative surprise } &= -\log(p_G(\zeta)) = -\log(p_G(\zeta)) \times 1 = -\log(P_G(\zeta))\sum_{\psi}p_G(\psi |\zeta)\\
-    &= -\sum_{\psi}p_G(\psi, \zeta)\log(p_G(\zeta))
-    =-\sum_{\psi}p_G(\psi |\zeta)\log(p_G(\psi, \zeta)/p_G(\psi|\zeta))\\
-    &= \sum_{\psi}p_G(\psi |\zeta)\log(p_G(\psi|\zeta))-\sum_{\psi}p_G(\psi |\zeta)\log  (p_G(\psi, \zeta))\\
-    &= -\langle \log  (p_G(., \zeta)) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta))\\
-    &= \langle E_G(., \zeta) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta)) := F_G(\zeta)
-  \end{split}
-  \label{eq:helm}
-\end{eqnarray}
-where $E_G(\psi, \zeta)$ is the energy of an fictive thermodynamic system defined by setting
-\begin{equation}
-  E_G(\psi, \zeta) := -\log(p_G(\psi, \zeta)).
-  \label{eq:gibbs}
-\end{equation}
-%% In the above, the variable $\psi$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
-The last quantity $F_G(\zeta) := \langle E_G(., \zeta) \rangle_{p_G(.
-|\zeta)} - \mathcal H(p_G(.
-|\zeta))$ in \eqref{eq:helm} is nothing other than the Helmholtz free-energy for the physical system correspond to the external world, prescribed by the generative density $p_G(\psi|\zeta)$ for the occupation of \textit{macrostates} $\psi$.
-Thus generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
-
-The goal of the brain is then to optimize over the generative model $G$ (to iteratively or analytically modify the generative density $p_G(.|s)$, so as to minimize their surprise.
-It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(s)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
-A workaround is then to introduce a second density $p_R(.|\zeta)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\zeta)$: the former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
-This primal-dual idea, first proposed in Hinton et al.
-1995, is at the heart ot the general free-energy principle that we'll introduce shortly.
-
-\subsubsection{Variational Helmholtz free-energy and the recognition density}
-In this subsection we'll develop a very insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
-As an avant-gout of what is to come shorty, let's just that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
-
-Now, invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
-distribution
-\begin{equation}
-  p_G(\psi|\zeta) = \frac{p_G(\psi|\zeta)}{p_G(\zeta)} = \frac{\exp(-E_G(\psi, \zeta))}{Z_G(\zeta)} = \frac{\exp(-E_G(\psi, \zeta))}{Z_G(\zeta)},
-\end{equation}
-where $Z_G(\zeta) := \log(p_G(\zeta)) = \sum_{\psi'}\exp(-E_G(\psi',\zeta))$, the normalizing partition function for the model \ref{eq:gibbs}.
-Whence $\forall \psi$, $Z_G(\zeta) = p_G(\zeta) = \exp(-E_G(\psi, \zeta)) / p_G(\psi|\zeta)$, and so we have the invariance relation
-\begin{equation}
-  -\log(Z_G(\zeta)) = -\log(p_G(\zeta)) = F_G(\zeta) = E_G(\psi, \zeta) + \log(p_G(\psi|\zeta)).
-  \end{equation}
-Now, in the equation above, the LHS only depends on the generative model $G$ and the data point $\zeta$: it doesn't depend on the hidden variable $\psi$, etc.
-So, taking expectations w.r.t an arbitrary density\footnote{The conditioning in $P_R(.|\zeta)$ is because this density is selected from a world in which the sensory inputs and internal brain state vector $\zeta$ is assumed already observed.} $p_R(.|\zeta)$ yields
-\begin{equation}
-  \begin{split}
-    F_G(\zeta) &= -\log(Z_G(\zeta)) = \langle E_G(., \zeta)\rangle_{P_R(.|\zeta)} + \sum_{\psi}p_R(\psi|\zeta)\log(p_G(\psi|\zeta))\\
-    &= \langle E_G(., \zeta)\rangle_{P_R(.|\zeta)} - \mathcal H(p_R(.|\zeta)) - \sum_{\psi}p_R(\psi|\zeta)\log(p_R(\psi|\zeta)/p_G(\psi|\zeta))\\
-    &= F^R_G(\zeta) - D_{KL}(P_R(.|\zeta) || P_G(.|\zeta)),
-  \end{split}
-  \label{eq:fe}
-\end{equation}
-where $F^R_G(\zeta)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
-\begin{equation}
-  F^R_G(\zeta) := \langle E_G(., \zeta)\rangle_{P_R(.|\zeta)} - \mathcal H(p_R(.|\zeta))
-\end{equation}
-and $D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))$ is the Kullback-Leibler divergence between the $p_R(.|\zeta)$ and the generative density $p_G(.|\zeta)$. Note that $F^G_G = F_G$.
-
-\subsubsection{A general free-energy principle}
-We can resume the situation as follows:
-\begin{mdframed}
-\begin{equation}
-  \begin{split}
-    \text{generative surprise } &:= -\log(p_G(\zeta)) = \underbrace{F_G(\zeta)}_{\text{generative free-energy}} \\
-    &=\underbrace{F^R_G(\zeta)}_{\text{variational free-energy}} - \underbrace{D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))}_{\text{complexity}} \\
-    &\le F^R_G(\zeta),
-    \text{ with equalitity if }p_R(.|\zeta) = p_G(.|\zeta),
-    \end{split}
-\end{equation}
-\end{mdframed}
-where we have used the fact that KL divergence is always nonnegative. In fact, we have the following theorem
-\begin{theorem}
-  It holds that
-  $$
-  \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
-  i.e
-  \begin{equation}
-    \min_{G}F_G(\zeta) = \min_{G,R}F^R_G(\zeta)
-    \end{equation}
-\end{theorem}
-
-\subsubsection{Helmholtz machines}
-% \begin{mdframed}
-  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data. Precisely
-$$
-p_G(\psi^{(l)}|\zeta) = \Pi_{k=1}^{h_l}p_G(\psi_k^{(l)} | \zeta),\hspace{1cm}
-p_R(\psi^{(l)}|\zeta) = \Pi_{k=1}^{h_l}p_R(\psi_k^{(l)} |\zeta)
-$$
-% \end{mdframed}
-\subsection{The wake-sleep algorithm}
-...
-\subsection{Friston's active-inference}
-This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
-
-The following critics can be made:
-\begin{itemize}
-  \item 
-    As noted by Dayan et al.
-(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
-    \end{itemize}
 
@@
     \end{itemize}
-The material presented is a revisit of \cite{braun2011path}. See also \cite{ortega2013thermodynamics}
+The material presented is a revisit of \textcolor{blue}{[Braun 2011]}.
 \paragraph{Utility functions and conjugate pairs.}
@@
     \begin{equation}
-      J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
+      J(P'|U, P) = \mathbb E_{P'}(U) + \alpha \mathbb H(P'),
     \label{eq:free_u}
     \end{equation}
-    where
-    \begin{equation}
-      \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
-    \end{equation}
-    is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
+    where $\mathbb H(P') := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))$ is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$).
+It's not difficult to establish the upper bound
   \begin{equation}
@@
   \end{equation}
-  In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
+  In particular, if $P$ is the Gibbs measure corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
   \begin{equation}
@@
   \begin{equation}
-    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
+    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\Delta U)}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
     \label{eq:free}
@@
   \begin{equation}
-    D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
+    D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
   \end{equation}
@@
 \begin{itemize}
-\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
+\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \mathbb E_{P'_{\text{fin}}}(\Delta U)$, and so it's optimal to take
   $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
@@
 
-To conclude this section, let's note that \cite{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
+To conclude this section, let's note that \textcolor{blue}{[Braun 2011]} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
 For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...
@@
 \section{Relation to Other Statistical Models (Elvis)}
-\paragraph{Control Theory.}
+\paragraph{Control Theory}
 abc
 
-\paragraph{Reinforcement Learning.}
+\paragraph{Q-Learning}
 abc
 
-\paragraph{Free Energy Principle.}
-The free-energy principle in it's present form (including notions like ``generative density'', ``recognition density'', etc.) can be traceback to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which the introduced the so-called Helmholtz machine...
-
-\begin{itemize}
-  \item Ortega et al. 2013 \cite{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
-\item Friston's critic of RL \cite{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
-pioneered in the 1950s by Richard Bellman and colleagues [2]. To
-optimise control a~pð Þ x~ under this formulation, we have to: (i)
-assume the hidden states are available to the agent and (ii) solve
-Equation 18 for the value-function. Solving for the value-function
-is a non-trivial problem and usually involves backwards induction
-or some approximation scheme like reinforcement-learning [4–6].
-The free-energy formulation circumvents these problems by
-prescribing the policy in terms of free-energy, which encodes
-optimal control (Equation 6)"
-  \end{itemize}
+\paragraph{Free Energy Principle}
+abc
 

### 2016-08-05 paper.tex

--- 2016-08-05 paper.tex
@@
 The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
+
+\item
+Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later
 

### 2016-08-05 paper.tex
added kingma stuf

--- 2016-08-05 paper.tex
@@
          \hline
-         \textbf{variable}    & \textbf{short description}  \\ \hline
+         \textbf{symbol}    & \textbf{description}  \\ \hline
          $\langle E\rangle_p$ & Expectation (a.k.a average) of the
-         random quantity $E$ w.r.t to the probability density $p$.\\ \hline
-         $\mathcal H$ & The usual information-theoretic entropy operator.\\ \hline
-         $D_{KL}(p||q)$ & The Kullback-Leibler divergence between the probability densities $p$ and $q$. \\ \hline
+         random quantity $E$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)E(z)$.\\ \hline
+         $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
+          with the usual convention $0 \log(0) := 0$.\\ \hline
+         $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
              $\zeta$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e $\zeta = (s, b).$ \\ \hline
@@
 
-\subsubsection{Helmholtz machines}
+\subsubsection{Helmholtz machines and the wake-sleep algorithm}
 % \begin{mdframed}
@@
 % \end{mdframed}
-\subsection{The wake-sleep algorithm}
-...
-\subsection{Friston's active-inference}
+
+\subsubsection{Friston's active-inference}
 This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
@@
 
-\subsubsection{Free-energy minimization via a stochastic gradient variation bayes estimator}
-\textbf{XXX see Kingma 2014 -- 2014: the poor boy is minimizing FE without knowing!}
-\begin{equation}
+\subsection{SGVB: an algorithm which truely minimizes the free-energy bound}
+Following Kingma et al. 2014 \cite{kingma2013auto}, define the data-dependent auxiliary random function
+\begin{eqnarray}
+  f_{G,R}(., \zeta) :\psi \mapsto \log(p_G(\psi,\zeta)) - \log(p_R(\psi|\zeta)).
+\end{eqnarray}
+Then we can rewrite the
+variational free-energy as
+\begin{eqnarray*}
   \begin{split}
-    F_G^R(\zeta) &:= \langle E_G(., \zeta) \rangle_{p_R(. |\zeta)} - \mathcal H(p_R(. |\zeta)) = \langle E_G(., \zeta) - \log(p_R(.|\zeta))\rangle_{P_R(.|\zeta)}\\
-    &=\langle \Delta E_{G \rightarrow R}(., \zeta)\rangle_{P_R(.|\zeta)}
+    F_G^R(\zeta) &:= \langle E_G(., \zeta) \rangle_{p_R(. |\zeta)} - \mathcal H(p_R(. |\zeta)) = \langle E_G(., \zeta) + \log(p_R(.|\zeta))\rangle_{P_R(.|\zeta)}\\
+    &=\langle -\log(p_G(.,\zeta)) + \log(p_R(.|\zeta))\rangle_{P_R(.|\zeta)}\\
+    &= -\langle f_{G,R}\rangle_{P_R(.|\zeta)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\psi^{(m)}), \text{ with }\psi^{(1)},\ldots,\psi^{(M)} \sim p_R(.|\zeta), \text{ and }M \rightarrow \infty.
     \end{split}
-\end{equation}
-
-For fixed $p_G(.|\zeta)$, the optimal choice for $p_R(.|\zeta)$ is given analytically by
-\begin{equation}
-  p_R(\psi|\zeta) \propto p_G(\psi|\zeta)\exp(-\Delta E_{G \rightarrow R}(\psi, \zeta)).
-\end{equation}
+\end{eqnarray*}
+
+\begin{mdframed}
+  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\zeta)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\textbf{W}^R$ ?
+\end{mdframed}
+\paragraph{Solution: The reparametrization trick.}
+
+%% For fixed $p_G(.|\zeta)$, the optimal choice for $p_R(.|\zeta)$ is given analytically by
+%% \begin{equation}
+%%   p_R(\psi|\zeta) \propto p_G(\psi|\zeta)\exp(-\Delta E_{G \rightarrow R}(\psi, \zeta)).
+%% \end{equation}
 

### 2016-08-05 paper/paper.tex
REFACTOR: moved things around to clean up the codebase

--- 2016-08-05 paper.tex
@@
 \subsection{Demystifying the free-energy principle and active-inference}
-The free-energy principle in it's present form (including notions like ``generative density'', ``recognition density'', etc.) can be traceback to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which the introduced the so-called Helmholtz machine...
+The free-energy principle in it's present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net.
 
@@
 This ideas were first developed by Hinton et al.
-in the early 90s in the form of the so-called  \textit{Helmholtz machine}, a hierarchical factorial directional deep-belief net.
+in the early 90s in building their Helmholtz machine.
 Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises.
@@
 \end{eqnarray}
-where $E_G(\psi, \zeta)$ is the energy of a fictive thermodynamic system defined by setting
+where $E_G(\psi, \zeta)$ is the energy at \textit{macrostate} $\psi$ of a fictive thermodynamic system defined by setting
 \begin{equation}
@@
 \end{equation}
-is the \textit{Helmholtz free-energy} of the external world, modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of \textit{macrostates} $\psi$.
+is the \textit{Helmholtz free-energy} of the external world, modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of macrostates $\psi$.
 Thus generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
@@
 This primal-dual idea, first proposed in Hinton et al.
-1995, is at the heart ot the general free-energy principle that we'll introduce shortly.
-
-\subsubsection{Variational Helmholtz free-energy and the recognition density}
+1995, is at the heart of the general free-energy principle that we'll introduce shortly.
+
+\subsubsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
 In this subsection we'll develop a very insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
@@
 
-\subsection{SGVB: an algorithm which truely minimizes the free-energy bound}
+\subsection{Free-energy minimization via bayesian auto-encoders}
 Following Kingma et al. 2014 \cite{kingma2013auto}, define the data-dependent auxiliary random function
@@
 \bibliographystyle{splncs03}
-\bibliography{nips_refs}
+\bibliography{refs}
 

### 2016-08-08 paper/paper.tex
more stuff

@@
 %% \subsection{Minimal control energy controller}
-\subsubsection{Related works}
-The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
 
@@
 For example the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial, e.g) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
+
+\subsubsection{Helmholtz free-energy and the generative model}
 \begin{table}[H]
@@
          \textbf{symbol}    & \textbf{description}  \\ \hline
-         $\langle E\rangle_p$ & Expectation (a.k.a average) of the
-         random quantity $E$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)E(z)$.\\ \hline
+         $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
+         random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
          $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
@@
 \end{table}
-\subsubsection{Helmholtz free-energy and the generative model}
 Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
@@
 \end{equation}
-is the \textit{Helmholtz free-energy} of the external world, modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of macrostates $\psi$.
+is the \textit{Helmholtz free-energy}. %% , modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of macrostates $\psi$
 Thus generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
@@
 It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(s)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
-A workaround is then to introduce a second density $p_R(.|\zeta)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\zeta)$: the former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
+A workaround is then to introduce a second density $p_R(.|\zeta)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\zeta)$, as a trick for doing approximate inference.
+The former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
 This primal-dual idea, first proposed in Hinton et al.
@@
 In this subsection we'll develop a very insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
-As an avant-gout of what is to come shorty, let's just that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
+As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
 
@@
 \end{equation}
-where $Z_G(\zeta) := \log(p_G(\zeta)) = \sum_{\psi'}\exp(-E_G(\psi',\zeta))$, the normalizing partition function for the model \ref{eq:gibbs}.
+where $Z_G(\zeta) := \log(p_G(\zeta)) = \sum_{\psi'}\exp(-E_G(\psi',\zeta))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
 Whence $\forall \psi$, $p_G(\zeta) = Z_G(\zeta) = \exp(-E_G(\psi, \zeta)) / p_G(\psi|\zeta)$, and so we have the invariance relation
@@
   \begin{split}
-    \text{generative surprise } &:= -\log(p_G(\zeta)) = \underbrace{F_G(\zeta)}_{\text{generative free-energy}} \\
-    &=\underbrace{F^R_G(\zeta)}_{\text{variational free-energy}} - \underbrace{D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))}_{\text{complexity}} \\
+    \text{generative surprise } &:= -\log(p_G(\zeta)) = F_G(\zeta) \\
+    &=\underbrace{F^R_G(\zeta)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))}_{\text{complexity}} \\
     &\le F^R_G(\zeta),
@@
 % \begin{mdframed}
-  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data. Precisely
+  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e input from lower more primitive layers).
+Precisely
 $$
@@
     As noted by Dayan et al.
-(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
+(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
     \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
@@
 \paragraph{Control Theory.}
-abc
+\begin{itemize}
+  \item The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
+\end{itemize}
 

### 2016-08-09 paper/paper.tex
more

@@
 \def\P{\mathbf{P}}
+\def\W{\mathbf{W}}
 \def\K{\mathbf{K}}
@@
 \subsection{Demystifying the free-energy principle and active-inference}
-The free-energy principle in it's present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net.
+The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net.
 
@@
 
-Now, invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
+Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
 distribution
@@
 \begin{mdframed}
-  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\zeta)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\textbf{W}^R$ ?
+  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\zeta)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
 \end{mdframed}
 \paragraph{Solution: The reparametrization trick.}
+\begin{itemize}
+\item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
+\item Set $\phi = g(\W^{R}, \zeta, \epsilon)$, where $g$ is an appropriate smooth function
+  \begin{itemize}
+  \item results in a sample $\phi \sim p_R(.|\zeta)$, from the correct
+    posterior
+    \end{itemize}
+\end{itemize}
+
+\paragraph{Examples.}
+\begin{table}[H]
+  \begin{tabular}{p{2cm}|p{1.5cm}|p{1.5cm}|p{2.2cm}|p{5cm}}
+         \hline
+         Example & $p_R(.|\zeta)$ & $p_{\text{noise}}$ & $g(\W^R,\zeta,\epsilon)$ & Also \\ \hline
+         Normal dist. & $\mathcal N(\mu,\sigma)$ & $\mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
+         Student’s t, Logistic, Uniform, Triangular, ... \\ \hline
+         Exponential & $\mathcal \exp(\lambda)$ & $\mathcal U(0, 1)$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
+\\ \hline
+         Other & $\log\mathcal N(\mu,\sigma)$ & $\mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
+  \end{tabular}
+  \caption{Reparametrization trick for a variety of models. Kingma et al. 2014 \cite{kingma2013auto}.}
+\end{table}
 

### 2016-08-15 paper/paper.tex
some new stuff

@@
 
-% macros from michael's .tex
+% some useful macros
 \DeclareMathOperator{\dist}{dist} % The distance.
@@
          $F_G(\zeta)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\zeta$. This measures the surprise incured upon observing $\zeta$ generated by the model $G$.\\ \hline
-         $F^R_G(\zeta)$ & Variational Helmholtz free-energy from model $R$
-         to $G$.  Note that $F^G_G = F_G$.\\ \hline
+         $F^R_G(\zeta)$ & Variational Helmholtz free-energy from model $G$
+          to $R$.  Note that $F^G_G = F_G$.\\ \hline
   \end{tabular}
@@
   \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
-    \end{itemize}
-
-\subsection{Free-energy minimization via bayesian auto-encoders}
+  \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
+    echnical difficulty in getting derivatives of loss function w.r.t
+    recognition wieghts $\W^R$).
+  \item This difficulty was removed in the 2010s by Kingma et al., an other groups, via some kind of ``reparametrization trick''.
+
+    \end{itemize}
+
+\subsection{Free-energy minimization via variational auto-encoders}
 Following Kingma et al. 2014 \cite{kingma2013auto}, define the data-dependent auxiliary random function
@@
 \item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
-\item Set $\phi = g(\W^{R}, \zeta, \epsilon)$, where $g$ is an appropriate smooth function
+\item Set $\phi = g(\W^{R}, \zeta, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
   \begin{itemize}
@@
 
-\paragraph{Examples.}
 \begin{table}[H]
-  \begin{tabular}{p{2cm}|p{1.5cm}|p{1.5cm}|p{2.2cm}|p{5cm}}
+  \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
          \hline
-         Example & $p_R(.|\zeta)$ & $p_{\text{noise}}$ & $g(\W^R,\zeta,\epsilon)$ & Also \\ \hline
-         Normal dist. & $\mathcal N(\mu,\sigma)$ & $\mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
+         Posterior & $p_R(.|\zeta)$ & noise & $g(\W^R,\zeta,\epsilon)$ & Also \\ \hline
+         Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
          Student’s t, Logistic, Uniform, Triangular, ... \\ \hline
-         Exponential & $\mathcal \exp(\lambda)$ & $\mathcal U(0, 1)$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
+         Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
 \\ \hline
-         Other & $\log\mathcal N(\mu,\sigma)$ & $\mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
+         Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
   \end{tabular}
@@
 \end{table}
+
+The result is a scheme for training DBNs via good-old backprop!
 

### 2016-09-01 paper/paper.tex
typos
more

@@
 
-\title{A Computational Account of Default-Mode Function\\by Control Theory and Reinforcement Learning}
+\title{A toy computational account of default-mode function by control theory and reinforcement learning}
 
@@
 \subsection{Demystifying the free-energy principle and active-inference}
-The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net.
-
-In this section we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
+The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
+
+In this section, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
 This ideas were first developed by Hinton et al.
@@
 \end{equation}
-is the \textit{Helmholtz free-energy}. %% , modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of macrostates $\psi$
+is the \textit{Helmholtz free-energy} (at unit temperature!). %% , modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of macrostates $\psi$
 Thus generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
 
-The goal of the brain is then to optimize over the generative model $G$ (to iteratively or analytically modify the generative density $p_G(.|s)$, so as to minimize their surprise.
+The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|s)$, so as to minimize surprise.
 It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(s)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
@@
 \subsubsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
-In this subsection we'll develop a very insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
+In this subsection, we'll present an insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
 As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
@@
     &\le F^R_G(\zeta),
-    \text{ with equalitity if }p_R(.|\zeta) = p_G(.|\zeta),
+    \text{ with equality if }p_R(.|\zeta) = p_G(.|\zeta),
     \end{split}
@@
   \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
-    echnical difficulty in getting derivatives of loss function w.r.t
-    recognition wieghts $\W^R$).
-  \item This difficulty was removed in the 2010s by Kingma et al., an other groups, via some kind of ``reparametrization trick''.
-
+    technical difficulty in getting derivatives of loss function w.r.t
+    recognition weights $\W^R$).
+  \item This difficulty was removed in the 2010s by Kingma et al., an other groups, via a ``reparametrization trick''.
     \end{itemize}
@@
 
-\section{Relation to Other Cognitive Models (Danilo)}
+\section{Relation to other cognitive models (Danilo)}
 \paragraph{Predictive Coding and Bayesian Brain Hypothesis}
@@
 
-\section{Relation to Other Statistical Models (Elvis)}
-\paragraph{Control Theory.}
+\section{Relation to other statistical models (Elvis)}
+\paragraph{Control theory.}
 \begin{itemize}

### 2016-09-22 paper/paper.tex
arranged passages

--- 2016-08-09 paper/paper.tex
@@
 
-\title{A Computational Account of Default-Mode Function\\by Control Theory and Reinforcement Learning}
+\title{A Computational Account of Default-Mode Function\\
+by Control Theory and Reinforcement Learning}
 
@@
 \def\P{\mathbf{P}}
-\def\W{\mathbf{W}}
 \def\K{\mathbf{K}}
@@
 \begin{abstract}
-
-the DMN is often proposed to be dedicated to continuous auto- biographical memory retrieval, environmental assessment, emotion ap- praisal, and reward contingency evaluation when letting the mind go.
-
-Constantly tracking and predicting changes in a variable environment is believed to then shape the internal milieu and upcoming external behav- ior adaptively
-
-The model reinterprets many known effects
-
-If this computational models hold some truth in them, the DMNs conceivable key role in the continuous environmental tracking in a generative, integrative process might explain both its highest energy consumption in the brain and its intimate coupling with conscious awareness.
-
-network that might be involved in the automated prediction of social events that modulate behavior
-
-We focus not on normative models (that describe what the brain does), but on process models (that describe how the brain does it)
+The default mode network (DMN) is the neural representation of the human
+baseline mental activity.
+%
+Many research streams agree on its likely role in evolutionarily adaptive envisioning of past and hypothetical scenarios to predict the environment.
+The DMN would hence be dedicated to continuous autobiographical memory retrieval, generate hypothetical outcomes,
+and reward contingency evaluation when letting the mind go.
+It explains its highest energy consumption in the brain and its intimate coupling with conscious awareness.
+%
+This concept paper proposes a process model that describes
+\textit{how} the DMN may actually implement continuous
+environmental assessment and prediction.
+DMN function is recast from the mathematical perspective of control theory and
+reinforcement learning.
+Neurobiological evidence on the human and animal DMN
+is linked to recent progress in autonomous systems research
+in the machine-learning domain.
+We argue that this engineering account can parsimoneously
+explain a variety of existing neuroscientific hypotheses of DMN function.
+%
+Formal models of the computational processes subserved by the DMN
+could offer statistical control of predictive human behavior.
 
@@
 % official NIPS keywords
-\textbf{\\keywords}: Systems Biology, Cognitive Science, Autonomous Learning Systems
+\textbf{\\keywords}: Systems Biology, mind wandering, cognitive science,
+autonomous learning systems
+\textbf{\\Get opinion from}: Danielle Bassett, Kai Vogeley, Mohammad Shakir
 
@@
 
+Its societal and economic relevance is certified by involvement in various psychiatric disorders, including schizophrenia, autism, and depression.
+
 sensory exchanges with the world
@@
 \section{The Neurobiological Properties of the Default-Mode Network}
+mPFC and PCC = highest metabolic consumption in the brain/highest level of basal glucose energy consumption in humans
+
 involvement of the PMC across the boundaries of classical cognitive domains
-mPFC and PCC = highest metabolic consumption in the brain/highest level of basal glucose energy consumption in humans
+
+mechanistic relevance of each node -> we deconstruct
 
@@
 dmPFC that reflects the values and preferences of another individual (here temporally discounted at a rate specific to the individual), even when they are not directly relevant to the task at hand.
-Critically, we also show this activity.
-While the simplest interpretation of this effect is that the region is simulating one’s own, currently irrelevant, preferences an alternative possibility is that the activity is projec- ting one’s own values into the mind of the partner, while simu- lating the partner’s choices.
-In essence, estimating the extent to which my own values would influence my partner, if they were making the choice.
-This iterated reasoning would be.
-
-mechanistic relevance of each node -> we deconstruct
-
-DMN function has been linked to the recollection of past episodes and the simulation of future episodes
 
@@
     \item
-the vmPFC sub- serves predominantly non-ambiguous subjective-value-related evaluative processes driven by bottom-up pathways
-memory-informed reward and risk estimation of self-relevant environmental stimuli.
+the vmPFC sub- serves predominantly non-ambiguous subjective-value-related evaluative processes
+reward-informed and risk estimation of self-relevant environmental stimuli.
 \item
@@
 the vmPFC to monitoring others’ (Lotze et al., 2007) and one’s own (Lane et al., 1997; Phan et al., 2004) emotional responses, that is, other’s (external) emotional reac- tions and one’s own (visceral) arousal.
-Such real or imagined bodily states, believed to be represented in the vmPFC, proba- bly operate as a bioregulatory disposition governing cognition and decision making (Damasio, 1996; Nauta, 1971), in line with the vmPFC’s functional association with general cognition and reward processing.
+Such real or imagined bodily states, believed to be represented in the vmPFC, proba- bly operate as a bioregulatory disposition governing cognition and decision making (Damasio, 1996; Nauta, 1971)
 
@@
 \item
+Glaescher: lesions -> value-based decision making  in 344 individuals
+based decision-making, which included the orbitofrontal, ventro- medial, and frontopolar cortex.
+
+\item
 As to the vmPFC subnetwork, the GMV of the vmPFC and VS correlated with indices of social reward attitudes and behavior (Lebreton et al., 2009), concur- ring with vmPFC’s relation to the NAc and reward-related tasks.
@@
 \item
-In studies of decision making and neuroeconomics, ventromedial prefrontal cortex (vmPFC) con- sistently reflects an individual’s subjective valuation (Behrens et al., 2008; Boorman et al., 2009; Knutson et al., 2005; Plass- mann et al., 2007).
-By contrast, its dorsomedial neighbor (rostral
+In studies of decision making and neuroeconomics, ventromedial prefrontal cortex (vmPFC) con- sistently reflects an individual’s subjective valuation (Behrens et al., 2008; Boorman et al., 2009; Knutson et al., 2005; Plass- mann et al., 2007)
 
@@
 \item
+Schuck 2016: Performance within and across participants was related to state encoding in OFC;
+“cognitive map” of task space in which the current state of the task is represented, and this representation is especially critical for behavior when states are unobservable from sensory input
+
+\item
 In both MACM and RS analy- ses, the vmPFC was more strongly connected with the nucleus accumbens (NAc), hippocampus (HC), posterior cingulate cor- tex (PCC), and retrosplenial cortex (RSC)
@@
 \item
-processing approach- and avoidance-relevant stimuli
-
-\item
-Glaescher: lesions -> value-based decision making  in 344 individuals
-based decision-making, which included the orbitofrontal, ventro- medial, and frontopolar cortex.
-
-\item
 Indeed, the vmPFC, but not dmPFC, has been observed to have monosynap- tical connections with the ventral striatum (VS, which anatom- ically includes the NAc) in axonal tracing studies in monkeys (Haber et al., 1995; Ferry et al., 2000).
@@
 dmPFC subserves predominantly ambiguous amodal metacogni- tive processes driven by top-down pathways
+
+\item
+sensory-independent de novo generation of meaning representations
+
+\item
+underlie inference, representation, and assessment
+such as in one’s own and others’ mental states in functional neuroimaging research
+
+\item
+There is also activity in human dorsomedial prefrontal cortex that reflects expectations about what an observed person will do and errors in such predictions (Suzuki et al., 2012; van Schie et al., 2004)
+largely sensory-independent, highly abstract (hence, less tangible) processes across time, space, and content domains.
+hypothetical processing.
 
@@
 The macaques adjusted their own subsequent actions in different ways after these two types of obser- vations.
-There is also activity in human dorsomedial prefrontal cortex that reflects expectations about what an observed person will do and errors in such predictions (Suzuki et al., 2012; van Schie et al., 2004)
-
-\item
-largely sensory-independent, highly abstract (hence, less tangible) processes across time, space, and content domains.
-might be top-down modulated by more dmPFC subserved higher reflective and hypothetical processing.
-consistent with present functional decoding, neural activity in the dmPFC, rather than vmPFC, has been con- sistently interpreted to underlie inference, representation, and assessment of one’s own and others’ mental states in functional neuroimaging research
-
-\item
- dmPFC (but not vmPFC) activity was related to the proficiency decline of mental state inference in elderly (Moran et al., 2012), cognitive regulation of one’s own emotional states (Ochsner et al., 2004b) and inference of another person’s emotional states (Ochsner et al., 2004a), as well as self-reported (Wagner et al., 2011) and experimentally measured (Zaki et al., 2009) proficiency in emotional state inference.
+
+\item
+ dmPFC (but not vmPFC) activity was related to the proficiency decline of mental state inference in elderly (Moran et al., 2012)
+
+ \item
+ congruently characterizes the dmPFC as a “mental sketchpad” (Goldman-Rakic, 1996) potentially impli- cated in modeling and binding plausible self- and other-related scenarios instructed by semantic concepts
 
@@
 
-\item
-congruently characterizes the dmPFC as a “mental sketchpad” (Goldman-Rakic, 1996) potentially impli- cated in modeling and binding plausible self- and other-related scenarios instructed by semantic concepts
-
-\item
- sensory-independent de novo generation of mean- ing representations can only be expected from highly associative, integrative brain areas such as those of the dmPFC subnetwork
-
-\item
-the dmPFC subserves a domain-independent neural process
-
-\item
-the present results support the dmPFC’s possible involvement in domain-overarching computational mechanisms given its connections to highly associative brain areas and function- ally relation to different complex psychological processes.
-
-\item
-the dmPFC was more strongly connected with the inferior frontal gyrus (IFG), temporo-parietal junction (TPJ), and middle temporal gyrus (MTG)
-
-\item
-Concluding from previous and present connectivity findings, the dmPFC is preferentially connected with high association and heteromodal cortical areas of the lateral frontal, temporal, and parietal lobe.
-
-\item
-it is interesting to note that the vmPFC is more strongly con- nected to medial components of the default mode network (i.e., HC, PCC, RSC), whereas the dmPFC is more strongly con- nected to its lateral components (i.e., IFG, TPJ, and MTG).
-This dmPFC subnetwork was repeatedly related to self-focused reflec- tion (Andrews-Hanna et al., 2010), contemplation of others’ (Mar, 2011) and one’s own (Lombardo et al., 2009) mental states, mental navigation of the body in space (Maguire et al., 1997), semantic processing (Binder et al., 2009), as well as scene con- struction processes when envisioning past, fictitious, and future events (Hassabis et al., 2007; Spreng et al., 2009; Bzdok et al., 2013).
-
-\item
-a VLSM study on disturbed sleep (i.e., a state of mind independent of sensory stimulation but dependent on internally generated informa- tion) exclusively identified the dmPFC (Koenigs et al., 2010).
--> experience replay during sleep
-
-\end{itemize}
-
-  \item
-    PCC (gradients):
-    \begin{itemize}
-      \item rewards (exploitation) and gathering information about alter- native options (exploration)
-Such strategic decisions should incorporate not only recent reward history, but also opportunity costs and environmental statistics
-
-\item
-covert reallocation of spatial attention (Gitelman et al., 1999), mediation between internal and external focus (Leech and Sharp, 2014), computation of environmental statistics (Pearson et al., 2009), and self-referential visuospatial imagery (Cavanna and Trimble, 2006)
-
-\item
+scene con- struction processes when envisioning past, fictitious, and future events (Hassabis et al., 2007; Spreng et al., 2009; Bzdok et al., 2013).
+
+\end{itemize}
+
+  \item
+    PMC (gradients):
+    \begin{itemize}
+      \item
+      PMC myelinate comparatively late during postnatal development in monkeys (Goldman- Rakic, 1987). generally a sign of phylogenetic sophistication (Flechsig, 1920)
+      \item
+      (Patho-)Physiologically, meta- bolic fluctuations in the human PMC have been closely related to vari- ous instances of altered conscious awareness, including anesthesia (Fiset et al., 1999), sleep (Maquet, 2000), and restoration from vegeta- tive states (Laureys et al., 1999).
+      \item
+      electrophys- iological research in animals implicated the PCC in strategic selection (Pearson et al., 2009), risk assessment (McCoy and Platt, 2005), and outcome-contingent behavioral modulation (Hayden et al., 2008), while the RSC was implicated in nav- igation and approach-avoidance behavior (Vann et al., 2009).
+
+      \item
+      Electrophysiologically, gamma band re- cordings in humans (Dastjerdi et al., 2011) and single-cell recordings in monkeys (Hayden et al., 2009) revealed
+      \textit{activity reductions} in the PMC
+      during task performance compared to rest.
+
+      \item
+      have long been speculated to reflect constant contemplation of (external) environment and (internal) memory, as an "inner minds eye"
+
+      \item Bálint's syndrome: neurological disorder of conscious
+      awareness resulting from damage in bilateral parietal cortex.
+      inability to bind various individual features of the visual
+      environment into an integrated whole (i.e., simultanagnosia)
+      as well as to direct action towards 
+      current unattended environmental objects
+      (i.e., optic ataxia).
+
+      This can viewed as an impairment in the gathering
+      of information about alternative options (i.e., exploration) as well as
+      leveraging these environmental opportunities
+      (i.e., exploitation).
+
+      \item RSC:
+      \begin{itemize}
+        \item
+        Regarding memory processing, lesions in the RSC, as compared to other PMC regions, is most consistently associated with anterograde and retrograde memory impairments of different types of sensory infor- mation in rabbits (Gabriel and Talk, 2001) and humans (Rudge and Warrington, 1991; Valenstein et al., 1987).
+        current findings and earlier lesion reports can be parsimoniously reconciled by the previously proposed notion that the RSC mediates between the organism's egocentric (i.e., focused on sensory input) and allocentric (i.e., focused on world knowledge) perspective frames (Burgess, 2008; Epstein, 2008; Valiquette and McNamara, 2007).
+
+      \item
+      neurons in posterior cingulate cortex signaled the distinction between trials on which monkeys pursued an exploratory rather than an exploitative strategy.
+
+      \item
+      Pearson2009:
+      balance. Here we show that CGp neurons distinguish be- tween exploratory and exploitative decisions made by mon- keys in a dynamic foraging task.
+
+      \item
+      computation of environmental statistics (Pearson et al., 2009):
+
       single-cell recordings in the monkey PCC demonstrated this brain region's sensitivity to subjective target utility (McCoy and Platt, 2005) and integration across individual outcomes in decision making (Pearson et al., 2009).
@@
 
-\item
-experimental psychologists and behavioral ecologists
-have long argued that decision making depends on the conversion
-of external variables into a common internal currency of value.
-
-\item
-electrophys- iological research in animals implicated the PCC in strategic selection (Pearson et al., 2009), risk assessment (McCoy and Platt, 2005), and outcome-contingent behavioral modulation (Hayden et al., 2008), while the RSC was implicated in nav- igation and approach-avoidance behavior (Vann et al., 2009).
-
+      -> internal currency of value
+
+      motivated, targeted saccades hindering fixation of currently not
+      attended features in the field of view (i.e., oculomotor apraxia).
+
+      \item
+      only cluster 2 was congruently coupled (across MACM and RSFC) with the amygdala (involved in significance evaluation) and the nucleus accumbens (involved in reward evaluation)
+      s well as exclusively functionally asso- ciated with facial appraisal.
+Similarly, the human vPCC was the only PMC region connected to the laterobasal (rather than centromedial or superficial) nuclei group of the amygdala (Bzdok et al., 2013b), which is an amygdalar subregion probably devoted to continuously scanning environmental input for biological significance (Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011; Ghods-Sharifi et al., 2009; LeDoux, 2007).
       \item
       (McCoy and Platt, 2005):
-t internal representations of value lie at the very core of
+      internal representations of subjective value lie at the very core of
       decision making,
-visual gambling task used in the current study affords a unique
-opportunity to examine neural activity under conditions in which
-subjective value, but not objective value, varied, as indicated by
+      task where subjective value, but not objective value, varied, as indicated by
       subjects’ choices
@@
 
-\item
-Pearson2009:
-balance. Here we show that CGp neurons distinguish be- tween exploratory and exploitative decisions made by mon- keys in a dynamic foraging task.
- option switching [11] and integrate this information across multiple trials [11], but the present
-
- \item
- Together, these results invite the hypothesis that CGp is part of a network that monitors the outcomes of individual deci- sions and integrates that information into higher-level strate- gies spanning multiple choices
+Such strategic decisions should incorporate not only recent reward history, but also opportunity costs and environmental statistics
+
+\item
+microstimulation in CGp promotes explo- ration of the previously antipreferred option.
 
@@
 \item
-choosing among multiple tar-
-gets whose relative values changed dynamically,
-neurons in posterior cingulate cortex signaled the distinction between trials on which monkeys pursued an exploratory rather than an exploitative strategy.
+outcome- contingent modulations in firing rate persisted into subsequent trials.
+Moreover, firing rate on any one trial predicted switching to the alternative option on the next trial.
+Finally, microstimulation in CGp follow- ing risky choices promoted a preference reversal for the safe option on the following trial.
+Collectively, these results demonstrate that CGp directly contrib- utes to the evaluative processes that support dynamic changes in decision making in volatile environments.
+
+\item
+IN SUM, these results suggest that the PCC is part of a network that monitors the subjective outcomes of possible decisions and integrates that information into higher-level strategies
 **learn its current value and integrate this information with their statistical knowledge of the environment to predict its relative value on upcoming trials.**
-
+\item
+information processing in the CGp continuously adapts to changes in both the external environment and internal milieu.
 \item
@@
 -> gradient defect
-
-\item
-corded activity of single neurons in monkeys per- forming a gambling task in which the reward outcome of each choice strongly influenced subsequent choices.
-choices.
-We found that CGp neurons signaled reward outcomes in a nonlinear fashion and that outcome- contingent modulations in firing rate persisted into subsequent trials.
-Moreover, firing rate on any one trial predicted switching to the alternative option on the next trial.
-Finally, microstimulation in CGp follow- ing risky choices promoted a preference reversal for the safe option on the following trial.
-Collectively, these results demonstrate that CGp directly contrib- utes to the evaluative processes that support dynamic changes in decision making in volatile environments.
-
-\item
- these observations suggest the hypothesis that CGp contributes to the integration of actions and their out- comes and thereby influences subsequent changes in behavior.
- We confirmed a causal role for CGp in outcome evaluation and behavioral adjustment by showing that microstimulation following the most desired reward out- come increased the likelihood of exploring the alternative on the next trial.
-
- \item
-signal decision outcomes in a nonlinear fashion, maintain this information in a buffer between trials, and predict future changes in behavior.
-Moreover, microstimulation in CGp promotes explo- ration of the previously antipreferred option.
-
-\item
-highlight the dynamic nature of information processing in CGp, which is not restricted to any single epoch or aspect of task performance but instead continuously adapts to changes in both the external environment and internal milieu.
-
-\item
-(Patho-)Physiologically, meta- bolic fluctuations in the human PMC have been closely related to vari- ous instances of altered conscious awareness, including anesthesia (Fiset et al., 1999), sleep (Maquet, 2000), and restoration from vegeta- tive states (Laureys et al., 1999).
-
-  \item
-Electrophysiologically, gamma band re- cordings in humans (Dastjerdi et al., 2011) and single-cell recordings in monkeys (Hayden et al., 2009) revealed activity reductions in the PMC during attentionally demanding tasks compared to rest.
-
-\item
- have long been speculated to reflect constant contemplation of (external) environment and (internal) memory
-
- \item
-   xons in parts of the PMC myelinate comparatively late during postnatal development in monkeys (Goldman- Rakic, 1987).
-Such late postnatal myelination is generally believed to occur in the phylogenetically most developed associations regions (Flechsig, 1920), t
-
-\item
-only cluster 2 was congruently coupled (across MACM and RSFC) with the amygdala (involved in significance evaluation) and the nucleus accumbens (involved in reward evaluation)
-s well as exclusively functionally asso- ciated with facial appraisal.
-Similarly, the human vPCC was the only PMC region connected to the laterobasal (rather than centromedial or superficial) nuclei group of the amygdala (Bzdok et al., 2013b), which is an amygdalar subregion probably devoted to continuously scanning environmental input for biological significance (Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011; Ghods-Sharifi et al., 2009; LeDoux, 2007).
-
-\item Bálint's syndrome: neurological disorder of conscious
-awareness resulting from damage in bilateral parietal cortex;
-"psychic paralysis of gaze", that is, the inability for internally
-motivated, targeted saccades hindering fixation of currently not
-attended features in the field of view (i.e., oculomotor apraxia).
-unable to bind various individual features of the visual
-environment into an integrated whole (i.e., simultanagnosia)
-as well as unable to navigate hand movement to a targeted
-object helped by vision (i.e., optic ataxia).
-difficulty avoiding running into objects placed ahead;
-challenged in estimating distances as well as perceiving
-spatial depth, object sizes and object orientation
-
-  \item
-    vPCC:
-the ventral visual stream—the “what” system for object process- ing (Ungerleider and Haxby, 1994; Ungerleider and Mishkin, 1982; Vogt et al., 2006), contrary to the dPCC's conceivable relation to the dorsal “where” system
-\item
-  dPCC:
-This view is in line with the well-known integration of the dPCC in the dorsal visual stream—the “where” system for spatial processing (Ungerleider and Haxby, 1994; Ungerleider and Mishkin, 1982; Vogt et al., 2006), in contrast to vPCC's relation to the ventral “what” system
-\end{itemize}
-
-\item RSC:
-  \begin{itemize}
-\item
-Regarding memory processing, lesions in the RSC, as compared to other PMC regions, is most consistently associated with anterograde and retrograde memory impairments of different types of sensory infor- mation in rabbits (Gabriel and Talk, 2001) and humans (Rudge and Warrington, 1991; Valenstein et al., 1987).
-current findings and earlier lesion reports can be parsimoniously reconciled by the previously proposed notion that the RSC mediates between the organism's egocentric (i.e., focused on sensory input) and allocentric (i.e., focused on world knowledge) perspective frames (Burgess, 2008; Epstein, 2008; Valiquette and McNamara, 2007).
+\item
+integration of actions and their out- comes and thereby influences subsequent changes in behavior.
+We confirmed a causal role for CGp in outcome evaluation and behavioral adjustment
+
+\item HC:
+\begin{itemize}
+  \item
+  Schachter/Buckner:
+  The network described here is tied to the medial temporal lobe system, which is traditionally considered almost exclusively in the context of remembering the past. We suspect the adaptive value of
+  episodic memory is not solely in its ability to afford mental reconstruction of the past but rather in its contribution to building mental models – simulations – of what might happen next or other perspectives on the immediate envir- onment, such as what others are thinking
+
+  \item
+The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
+
+\item
+hippocampal damage [21,38–40] is associated with an impairment in thinking about one’s personal future
+
+\item
+bilateral hippocampal damage were markedly impaired relative to matched control subjects at imagining new fictitious experiences (Hassabis), their descriptions lacked spatial coherence and were instead fragmented and considerably less rich
+
+\item 
+A special case is the hippocampus whose highly recursive connection matrix is thought to function as a large ‘autoassocia- tor’, allowing the reconstruction of entire episodes from remembered fragments.
+
+\item
+ structural MRI studies of expert navigators (London taxi drivers) who show increased grey matter volume in posterior hippocampus seemingly at the expense of reduced grey matter volume in anterior hippocampus (Maguire et al.
+2006b)
+
+\item
+hippocampal pyramidal cells during rest and sleep produce strongly coherent ensemble bursts believed to be critical in transferring information to the neocortex.
+
+\item
+statistical structure of hippocampal activity ->
+spatial maps of surroundings are encoded
+buszaki2004 theta oscillations:
+Using large-scale recordings, researchers can follow complex patterns for extended time periods and determine whether their modification by experience will influence self-generated patterns in the absence of envi- ronmental inputs.
+Spike sequences, imposed upon the network by behavioral manipulations, recur spontaneously during subsequent sleep
+indicating that neurons organize themselves into preferred cell assemblies, and the seeds of emergence stem from experi- ence-related activity.
+
+Neurons, mostly studied in the hippocampus, which fire action potentials only when the animal is in a specific place in a given environment (that place where a cell fires is called the “place field” of that cell).
+
+Numerous experiments have demonstrated that hippocampal neurons show place-related firing while the rat explores or traverses its environment so that each assembly of hippocampal principal cells defines a particular position of space
+
+\item
+Diba + Buzsaki 2007
+We report that temporal spike sequences from hippocampal place neurons of rats on an elevated track recurred in reverse order at the end of a run, but in forward order in anticipation of the run, coinciding with sharp waves.
+Vector distances between the place fields were reflected in the temporal structure of these sequences.
+This bidirectional re-enactment of temporal sequences may contribute to the establishment of higher-order associations in episodic memory.
+
+-> self-play in deepmind Go
+
+\item
+coding, transfer, and consolidation of information:
+This emerging picture of hippocampal dy- namics suggests that neurons at the opposite ends of the distributions may convey differ- ent but complementary types of information.
+The ever-active minority of place cells may be responsible for generalizing across envi- ronments and affords the brain the capacity to regard no situation as completely un- known because every alley, mountain, river, or room has elements of previously experi- enced similar situations.
+In many situations,
+this minority provides the “best guess” of the hippocampus and offers “good enough” solutions to get by.
+On the other hand, the majority of less active neurons constitute a large reservoir that can be mobilized (7) to precisely distinguish one situation from an- other and incorporate novel ones as distinct.
+
+\item
+damage to even the hippocampus alone is sufficient to cause amnesia +
+envisoning themselves in future and even hypothetical scenarios/mentally constructing new experiences
+
+The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
+
+\item
+Hassabias2007:
+Recollection of these rich autobiographical or episodic memories has been likened to mentally traveling back in time and re-experiencing one’s past (1).
+It has long been known that the hippocampus and related medial temporal lobe structures play a critical role in supporting episodic memory
+
+In fact, episodic memory and imagining or constructing events share striking similarities in terms of the psychological processes engaged (19–21).
+These include imagery (22), sense of presence (1), retrieval of semantic information and multimodal details (23), and narrative structure (22).
+
+visual imagery itself is not believed to be realized in the HC
+
+\item
+Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later
+
+\end{itemize}
+
+\item
+the aRTPJ network is central for sensorimotor control by integrating supramodal stimulus-guided attention and action initia- tion during externally structured tasks.
+
+\item
+visuo- proprioceptive conflict (Balslev et al., 2005), and multi-modal detection of sensory changes (Downar et al., 2000)
+brain areas connected to aRTPJ further responded to all changes of visual, auditory, or tactile stimulation in a multi-modal fMRI study (Downar et al., 2000).
+
+\item
+direct electrical stimulation of the RTPJ during neurosurgery was associated with altered perception and stimulus awareness (Blanke et al., 2002).
 
@@
     \begin{itemize}
+\item
+Sridharan: RTPJ involved in bottom-up saliency detection
+
 \item
@@
 \item
-blanke:
-the core region of the human vestibular cortex is situated close to the angular gyrus10.
-It is possible that the experience of dissociation of self from the body is a result of failure to integrate complex somatosensory and vestibular information.
-Initial stimulations (n 3; 2.0–3.0 mA) induced vestibular responses, in which the patient reported that she was “sinking into the bed” or “falling from a height”.
-
-\item
-high-level social cognition also engenerded expectation violations
--> that's why coincides with "agency" meta-analytic location
-(Decety 2007):
-suggest that activation in the TPJ during social cognition may therefore rely on a lower-level computational mech- anism involved in generating, testing, and correcting internal predictions about external sensory events.
-
-\item
-Sridharan: RTPJ involved in bottom-up saliency detection
+Bzdok2013
+
+qualify RTPJ as a po- tential switch between exteroceptive and interoceptive mind sets implemented by their networks.
+The reciprocal nature of those mind sets is well illustrated by consistent reports of continuous shifting be- tween externally (i.e., sensory) and internally (i.e., self- and presumably social/memory-) oriented processing in controlled laboratory settings and daily routine (Smallwood et al., 2007).
+\end{itemize}
 
@@
 \item
-visuo- proprioceptive conflict (Balslev et al., 2005), and multi-modal detection of sensory changes (Downar et al., 2000)
-brain areas connected to aRTPJ further responded to all changes of visual, auditory, or tactile stimulation in a multi-modal fMRI study (Downar et al., 2000).
-
-\item
-direct electrical stimulation of the RTPJ during neurosurgery was associated with altered perception and stimulus awareness (Blanke et al., 2002).
-
-\item
-the aRTPJ network is central for sensorimotor control by integrating supramodal stimulus-guided attention and action initia- tion during externally structured tasks.
-
-\item
-Bzdok2013
-
-pRTPJ cluster was also functionally associated with deception tasks
-
-qualify RTPJ as a po- tential switch between exteroceptive and interoceptive mind sets implemented by their networks.
-The reciprocal nature of those mind sets is well illustrated by consistent reports of continuous shifting be- tween externally (i.e., sensory) and internally (i.e., self- and presumably social/memory-) oriented processing in controlled laboratory settings and daily routine (Smallwood et al., 2007).
-\end{itemize}
-
-\item
   LTPJ
   \begin{itemize}
-\item
-the difference between the two task groups could be parsimoniously explained by the required atten- tion to either the external world or self.
-world -> policy matrix driven computation
-self -> value matrix driven computation
-
-\item
-significantly reduced impact of intentions after transient TPJ disruption (Young et al. 2010b)
-
-\item
-complex conceptual linguistic functions are proposed to be more strongly left-lateralized -> in line with overall left-lateralization of human language
-
-\item
-Episodic and explicit memory retrieval strongly draws on complex semantic processing and contributes to social cognitive processes (see section 4.4.).
-Syntactic processing, on the other hand, is a core language process that is closely intermingled with semantic processing.
-It refers to the hierarchical sequencing of words and their meanings (Price, 2010) and is mandatory for sentence processing in both social cognitive and language tasks.
-
-\item
-only the left TPJ was associated with language processes, whereas only the right TPJ was functionally associated with attention processes
-
     \item 
@@
 
+\item
+significantly reduced impact of intentions after transient TPJ disruption (Young et al. 2010b)
+
+\item
+complex conceptual linguistic functions are proposed to be more strongly left-lateralized -> in line with overall left-lateralization of human language
+
+\item
+Episodic and explicit memory retrieval strongly draws on complex semantic processing and contributes to social cognitive processes (see section 4.4.).
+Syntactic processing, on the other hand, is a core language process that is closely intermingled with semantic processing.
+It refers to the hierarchical sequencing of words and their meanings (Price, 2010) and is mandatory for sentence processing in both social cognitive and language tasks.
+
 \item a lesion study on disturbed sleep (i.e., a state of mind
@@
 
-\item HC:
-\begin{itemize}
-  \item
-The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
-
-\item
-Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later
-
-\item
-hippocampal damage [21,38–40] is associated with an impairment in thinking about one’s personal future
-
-\item
-They found that amnesic patients with bilateral hippocampal damage were markedly impaired relative to matched control subjects at imagining new fictitious experiences (Figure 1).
-Moreover, Hassabis et al.
-[21] identified a possible source for this deficit.
-Although patients were able to produce considerable detail when asked to imagine fictitious new experiences, their descriptions lacked spatial coherence and were instead fragmented and considerably less rich [21].
-
-\item
- empirical evidence hinting at a two-process function of the hippocampus comes from structural MRI studies of expert navigators (London taxi drivers) who show increased grey matter volume in posterior hippocampus seemingly at the expense of reduced grey matter volume in anterior hippocampus (Maguire et al.
-2006b).
-Moreover, their increased spatial knowledge appears to come at a cost to the acquisition of new visual associative information (Maguire et al.
-2006b; Woollett \& Maguire 2009).
-
-\item
-Schachter/Buckner:
-The network described here is tied to the medial temporal lobe system, which is traditionally considered almost exclusively in the context of remembering the past. We suspect the adaptive value of
-episodic memory is not solely in its ability to afford mental reconstruction of the past but rather in its contribution to building mental models – simulations – of what might happen next or other perspectives on the immediate envir- onment, such as what others are thinking (see Ref.
-[63]).
-As
-
-\item
-the GMV of the entorhinal cortex (connectionally and functionally closely coupled with the HC) correlated with social network size (Kanai et al., 2012)
-
-\item 
-In fact, phase-locked hippocampal?<80><93>cortical loops have been proposed t
-formation of mnemonic neuronal representations
-**a large body of animal and human literature relates
-hippocampal-prefrontal theta to memory and memory to the DMN**
-
-\item 
-A special case is the hippocampus whose highly recursive connection matrix is thought to function as a large ‘autoassocia- tor’, allowing the reconstruction of entire episodes from remembered fragments.
-
-\item
-hippocampal pyramidal cells during rest and sleep produce strongly coherent ensemble bursts believed to be critical in transferring information to the neocortex.
-
-\item
-sequential single- cell recording could reveal such cooperative patterns
-
-\item
-buszaki2004 theta oscillations:
-Using large-scale recordings, researchers can follow complex patterns for extended time periods and determine whether their modification by experience will influence self-generated patterns in the absence of envi- ronmental inputs.
-Spike sequences, imposed upon the network by behavioral manipulations, recur spontaneously during subsequent sleep
-indicating that neurons organize themselves into preferred cell assemblies, and the seeds of emergence stem from experi- ence-related activity.
- Such internally driven, self-organized assemblies may reflect mechanisms that give rise to cognitive phenomena
- Sharp-wave ripples are believed to be the transfer mechanism of information from the hippocampus to neocortex during off-line (eg, sleep) states.
-Neurons, mostly studied in the hippocampus, which fire action potentials only when the animal is in a specific place in a given environment (that place where a cell fires is called the “place field” of that cell).
--> self-play in deepmind Go
-
-\item
-statistical structure of hippocampal activity ->
-spatial maps of surroundings are encoded
-
-\item
-Numerous experiments have demonstrated that hippocampal neurons show place-related firing while the rat explores or traverses its environment so that each assembly of hippocampal principal cells defines a particular position of space
-
-\item
-Diba + Buzsaki 2007
-We report that temporal spike sequences from hippocampal place neurons of rats on an elevated track recurred in reverse order at the end of a run, but in forward order in anticipation of the run, coinciding with sharp waves.
-Vector distances between the place fields were reflected in the temporal structure of these sequences.
-This bidirectional re-enactment of temporal sequences may contribute to the establishment of higher-order associations in episodic memory.
-
-\item
-coding, transfer, and consolidation of
-information
-This emerging picture of hippocampal dy- namics suggests that neurons at the opposite ends of the distributions may convey differ- ent but complementary types of information.
-The ever-active minority of place cells may be responsible for generalizing across envi- ronments and affords the brain the capacity to regard no situation as completely un- known because every alley, mountain, river, or room has elements of previously experi- enced similar situations.
-In many situations,
-this minority provides the “best guess” of the hippocampus and offers “good enough” solutions to get by.
-On the other hand, the majority of less active neurons constitute a large reservoir that can be mobilized (7) to precisely distinguish one situation from an- other and incorporate novel ones as distinct.
-
-\item
-damage to even the hippocampus alone is sufficient to cause amnesia +
-envisoning themselves in future and even hypothetical scenarios/mentally constructing new experiences
-
-The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
-
-\item
-Hassabias2007:
-Recollection of these rich autobiographical or episodic memories has been likened to mentally traveling back in time and re-experiencing one’s past (1).
-It has long been known that the hippocampus and related medial temporal lobe structures play a critical role in supporting episodic memory
-
-In fact, episodic memory and imagining or constructing events share striking similarities in terms of the psychological processes engaged (19–21).
-These include imagery (22), sense of presence (1), retrieval of semantic information and multimodal details (23), and narrative structure (22).
-
-visual imagery itself is not believed to be realized in the HC
-
-  \end{itemize}
+\item
+the difference between the two task groups could be parsimoniously explained by the required atten- tion to either the external world or self.
+world -> policy matrix driven computation
+self -> value matrix driven computation
+
 \end{itemize}
@@
 %% \subsection{Minimal control energy controller}
+\subsubsection{Related works}
+The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
 
 \subsection{Demystifying the free-energy principle and active-inference}
-The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net.
+The free-energy principle in it's present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net.
 
@@
 For example the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial, e.g) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
-
-\subsubsection{Helmholtz free-energy and the generative model}
 \begin{table}[H]
@@
          \textbf{symbol}    & \textbf{description}  \\ \hline
-         $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
-         random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
+         $\langle E\rangle_p$ & Expectation (a.k.a average) of the
+         random quantity $E$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)E(z)$.\\ \hline
          $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
@@
 \end{table}
+\subsubsection{Helmholtz free-energy and the generative model}
 Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
@@
 \end{equation}
-is the \textit{Helmholtz free-energy}. %% , modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of macrostates $\psi$
+is the \textit{Helmholtz free-energy} of the external world, modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of macrostates $\psi$.
 Thus generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
@@
 It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(s)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
-A workaround is then to introduce a second density $p_R(.|\zeta)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\zeta)$, as a trick for doing approximate inference.
-The former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
+A workaround is then to introduce a second density $p_R(.|\zeta)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\zeta)$: the former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
 This primal-dual idea, first proposed in Hinton et al.
@@
 In this subsection we'll develop a very insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
-As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
-
-Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
+As an avant-gout of what is to come shorty, let's just that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
+
+Now, invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
 distribution
@@
 \end{equation}
-where $Z_G(\zeta) := \log(p_G(\zeta)) = \sum_{\psi'}\exp(-E_G(\psi',\zeta))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
+where $Z_G(\zeta) := \log(p_G(\zeta)) = \sum_{\psi'}\exp(-E_G(\psi',\zeta))$, the normalizing partition function for the model \ref{eq:gibbs}.
 Whence $\forall \psi$, $p_G(\zeta) = Z_G(\zeta) = \exp(-E_G(\psi, \zeta)) / p_G(\psi|\zeta)$, and so we have the invariance relation
@@
   \begin{split}
-    \text{generative surprise } &:= -\log(p_G(\zeta)) = F_G(\zeta) \\
-    &=\underbrace{F^R_G(\zeta)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))}_{\text{complexity}} \\
+    \text{generative surprise } &:= -\log(p_G(\zeta)) = \underbrace{F_G(\zeta)}_{\text{generative free-energy}} \\
+    &=\underbrace{F^R_G(\zeta)}_{\text{variational free-energy}} - \underbrace{D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))}_{\text{complexity}} \\
     &\le F^R_G(\zeta),
@@
 % \begin{mdframed}
-  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e input from lower more primitive layers).
-Precisely
+  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data. Precisely
 $$
@@
     As noted by Dayan et al.
-(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
+(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
     \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
@@
 \begin{mdframed}
-  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\zeta)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
+  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\zeta)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\textbf{W}^R$ ?
 \end{mdframed}
 \paragraph{Solution: The reparametrization trick.}
-\begin{itemize}
-\item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
-\item Set $\phi = g(\W^{R}, \zeta, \epsilon)$, where $g$ is an appropriate smooth function
-  \begin{itemize}
-  \item results in a sample $\phi \sim p_R(.|\zeta)$, from the correct
-    posterior
-    \end{itemize}
-\end{itemize}
-
-\paragraph{Examples.}
-\begin{table}[H]
-  \begin{tabular}{p{2cm}|p{1.5cm}|p{1.5cm}|p{2.2cm}|p{5cm}}
-         \hline
-         Example & $p_R(.|\zeta)$ & $p_{\text{noise}}$ & $g(\W^R,\zeta,\epsilon)$ & Also \\ \hline
-         Normal dist. & $\mathcal N(\mu,\sigma)$ & $\mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
-         Student’s t, Logistic, Uniform, Triangular, ... \\ \hline
-         Exponential & $\mathcal \exp(\lambda)$ & $\mathcal U(0, 1)$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
-\\ \hline
-         Other & $\log\mathcal N(\mu,\sigma)$ & $\mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
-  \end{tabular}
-  \caption{Reparametrization trick for a variety of models. Kingma et al. 2014 \cite{kingma2013auto}.}
-\end{table}
 
@@
 \paragraph{Control Theory.}
-\begin{itemize}
-  \item The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
-\end{itemize}
+abc
 

### 2016-09-25 paper/paper.tex
interm

@@
 Many research streams agree on its likely role in evolutionarily adaptive envisioning of past and hypothetical scenarios to predict the environment.
-The DMN would hence be dedicated to continuous autobiographical memory retrieval, generate hypothetical outcomes,
+The DMN would hence be dedicated to continuous autobiographical memory retrieval,
+generation of hypothetical outcomes,
 and reward contingency evaluation when letting the mind go.
@@
 
-\section{The Neurobiological Properties of the Default-Mode Network}
+\section{Known Neurobiological Properties of the Default-Mode Network}
 mPFC and PCC = highest metabolic consumption in the brain/highest level of basal glucose energy consumption in humans
 
+scene con- struction processes when envisioning past, fictitious, and future events (Hassabis et al., 2007; Spreng et al., 2009; Bzdok et al., 2013).
+it is not random but purposeful
+draws on conceptual knowledge acquired through experience
+operating on internal sources of information
+Computational processes with a focus on the internal world,
+when the environment is highly predictable and
+external stimuli are irrelevant.
 involvement of the PMC across the boundaries of classical cognitive domains
@@
 
-\begin{itemize}
-  \item
-    vmPFC:
-    \begin{itemize}
-    \item
-the vmPFC sub- serves predominantly non-ambiguous subjective-value-related evaluative processes
-reward-informed and risk estimation of self-relevant environmental stimuli.
-\item
-the vmPFC might be more closely associated with orches- trating adapted behavior by bottom-up-driven processing of “what matters now”
-
-\item
-the vmPFC to monitoring others’ (Lotze et al., 2007) and one’s own (Lane et al., 1997; Phan et al., 2004) emotional responses, that is, other’s (external) emotional reac- tions and one’s own (visceral) arousal.
-Such real or imagined bodily states, believed to be represented in the vmPFC, proba- bly operate as a bioregulatory disposition governing cognition and decision making (Damasio, 1996; Nauta, 1971)
-
-\item
-Most importantly, independent whole-brain analyses from structural neuroimaging studies related the gray-matter volume (GMV) of the vmPFC to indices of social competence and social network complex- ity in both humans and monkeys (Lebreton et al., 2009; Powell et al., 2010; Lewis et al., 2011; Sallet et al., 2011).
-To our knowl- edge, none of these four correlations have been found yet for the dmPFC.
-Consequently, vmPFC, rather than dmPFC, anatomy appears to predict an individual’s social behavioral dispositions and social network properties,
-
-\item
-With respect to our seeds, inter-individual differences in social skills or social networks were most often related to morphological differences in the human and monkey vmPFC, in stark contrast to the dmPFC
-
-\item
-self-related behavior guided by stimulus evalua- tion and reward-learning, a voxel-based lesion-symptom map- ping (VLSM) study in 344 neurological patients demonstrated functional-anatomical specificity of the vmPFC for value-based decision-making (Gläscher et al., 2012).
-
-\item
-Glaescher: lesions -> value-based decision making  in 344 individuals
-based decision-making, which included the orbitofrontal, ventro- medial, and frontopolar cortex.
-
-\item
-As to the vmPFC subnetwork, the GMV of the vmPFC and VS correlated with indices of social reward attitudes and behavior (Lebreton et al., 2009), concur- ring with vmPFC’s relation to the NAc and reward-related tasks.
-
-\item
-In studies of decision making and neuroeconomics, ventromedial prefrontal cortex (vmPFC) con- sistently reflects an individual’s subjective valuation (Behrens et al., 2008; Boorman et al., 2009; Knutson et al., 2005; Plass- mann et al., 2007)
-
-\item
-NAC was proposed to subserve model-free reinforcement learning in humans
-(O'Doherty 2015)
-
-\item
-Schuck 2016: Performance within and across participants was related to state encoding in OFC;
-“cognitive map” of task space in which the current state of the task is represented, and this representation is especially critical for behavior when states are unobservable from sensory input
-
-\item
-In both MACM and RS analy- ses, the vmPFC was more strongly connected with the nucleus accumbens (NAc), hippocampus (HC), posterior cingulate cor- tex (PCC), and retrosplenial cortex (RSC)
-In both functional decoding analyses, the vmPFC was selectively associated with reward related tasks
-
-\item
-Indeed, the vmPFC, but not dmPFC, has been observed to have monosynap- tical connections with the ventral striatum (VS, which anatom- ically includes the NAc) in axonal tracing studies in monkeys (Haber et al., 1995; Ferry et al., 2000).
-This is consistent with our results and probabilistic diffusion tensor imaging (DTI) trac- tography in humans and monkeys (Croxson et al., 2005) that quantified the VS to be substantially more likely connected to the vmPFC than dmPFC in both species.
-
-\item
-the vmPFC is preferentially connected with limbic and reward-related medial brain areas.
-The NAc is thought to be linked to reward mechanisms that may not only modulate motivated behavior towards basic survival needs
-
-\item
-Such real or imagined bodily states, believed to be represented in the vmPFC, proba- bly operate as a bioregulatory disposition governing cognition and decision making (Damasio, 1996; Nauta, 1971),
-
-\item
-\end{itemize}
-
-\item
-  dmPFC:
-  \begin{itemize}
-    \item
-dmPFC subserves predominantly ambiguous amodal metacogni- tive processes driven by top-down pathways
-
-\item
-sensory-independent de novo generation of meaning representations
-
-\item
-underlie inference, representation, and assessment
-such as in one’s own and others’ mental states in functional neuroimaging research
-
-\item
-There is also activity in human dorsomedial prefrontal cortex that reflects expectations about what an observed person will do and errors in such predictions (Suzuki et al., 2012; van Schie et al., 2004)
-largely sensory-independent, highly abstract (hence, less tangible) processes across time, space, and content domains.
-hypothetical processing.
-
-\item
-A class of neurons in macaque dorsomedial frontal cortex, including the presupplementary area, responds preferentially to the actions that another macaque makes rather than their own actions (Yoshida et al., 2011)
-Many of the same neurons differentiate between another macaque’s actions that were unsuccessful, because the circumstances changed, and actions that were unsuccessful because the other macaque did not make the best choice given the circum- stances.
-The macaques adjusted their own subsequent actions in different ways after these two types of obser- vations.
-
-\item
- dmPFC (but not vmPFC) activity was related to the proficiency decline of mental state inference in elderly (Moran et al., 2012)
-
- \item
- congruently characterizes the dmPFC as a “mental sketchpad” (Goldman-Rakic, 1996) potentially impli- cated in modeling and binding plausible self- and other-related scenarios instructed by semantic concepts
-
- \item
-Thus, mental state inference necessarily relies on the generation of probabilistic internal information.
-Supported by dmPFC’s functional association with episodic memory retrieval, such prima vista non-mnemonic construction processes are likely to be subserved by the neural network underlying retrieval of past and imagination of future scenes as indicated by recent neu- roimaging experiments and meta-analyses
-
-scene con- struction processes when envisioning past, fictitious, and future events (Hassabis et al., 2007; Spreng et al., 2009; Bzdok et al., 2013).
-
-\end{itemize}
-
-  \item
-    PMC (gradients):
-    \begin{itemize}
-      \item
-      PMC myelinate comparatively late during postnatal development in monkeys (Goldman- Rakic, 1987). generally a sign of phylogenetic sophistication (Flechsig, 1920)
-      \item
-      (Patho-)Physiologically, meta- bolic fluctuations in the human PMC have been closely related to vari- ous instances of altered conscious awareness, including anesthesia (Fiset et al., 1999), sleep (Maquet, 2000), and restoration from vegeta- tive states (Laureys et al., 1999).
-      \item
-      electrophys- iological research in animals implicated the PCC in strategic selection (Pearson et al., 2009), risk assessment (McCoy and Platt, 2005), and outcome-contingent behavioral modulation (Hayden et al., 2008), while the RSC was implicated in nav- igation and approach-avoidance behavior (Vann et al., 2009).
-
-      \item
-      Electrophysiologically, gamma band re- cordings in humans (Dastjerdi et al., 2011) and single-cell recordings in monkeys (Hayden et al., 2009) revealed
-      \textit{activity reductions} in the PMC
-      during task performance compared to rest.
-
-      \item
-      have long been speculated to reflect constant contemplation of (external) environment and (internal) memory, as an "inner minds eye"
-
-      \item Bálint's syndrome: neurological disorder of conscious
-      awareness resulting from damage in bilateral parietal cortex.
+The PMC myelinates relatively late during postnatal development in monkeys
+(Goldman-Rakic, 1987), generally considered to
+be a sign of phylogenetic sophistication (Flechsig, 1920).
+Physiological and disturbed metabolic fluctuations in the
+human PMC have been repeated related to
+phenomena of changed conscious awareness,
+including anesthesia (Fiset et al., 1999),
+sleep (Maquet, 2000), and forms of coma (Laureys et al., 1999).      
+%
+The PMC has long been speculated to reflect constant computation of
+the (external) environmental statistics and its (internal) representation
+as an "inner minds eye" (Cavanna et al., 2006).
+For instance, Bálint's syndrome is a neurological disorder of conscious
+awareness that resulting from damage in the bilateral parietal cortex.
+Patients are plaigued by an
 inability to bind various individual features of the visual
 environment into an integrated whole (i.e., simultanagnosia)
-      as well as to direct action towards 
+as well as inability to direct action towards 
 current unattended environmental objects
 (i.e., optic ataxia).
-
-      This can viewed as an impairment in the gathering
-      of information about alternative options (i.e., exploration) as well as
-      leveraging these environmental opportunities
-      (i.e., exploitation).
-
-      \item RSC:
-      \begin{itemize}
-        \item
-        Regarding memory processing, lesions in the RSC, as compared to other PMC regions, is most consistently associated with anterograde and retrograde memory impairments of different types of sensory infor- mation in rabbits (Gabriel and Talk, 2001) and humans (Rudge and Warrington, 1991; Valenstein et al., 1987).
-        current findings and earlier lesion reports can be parsimoniously reconciled by the previously proposed notion that the RSC mediates between the organism's egocentric (i.e., focused on sensory input) and allocentric (i.e., focused on world knowledge) perspective frames (Burgess, 2008; Epstein, 2008; Valiquette and McNamara, 2007).
-
-      \item
-      neurons in posterior cingulate cortex signaled the distinction between trials on which monkeys pursued an exploratory rather than an exploitative strategy.
-
-      \item
-      Pearson2009:
-      balance. Here we show that CGp neurons distinguish be- tween exploratory and exploitative decisions made by mon- keys in a dynamic foraging task.
-
-      \item
-      computation of environmental statistics (Pearson et al., 2009):
-
-      single-cell recordings in the monkey PCC demonstrated this brain region's sensitivity to subjective target utility (McCoy and Platt, 2005) and integration across individual outcomes in decision making (Pearson et al., 2009).
-      ng attraction or aversion to options with uncertain or risky rewards
-      . CGp activation was better predicted by the
-      subjective salience of a chosen target than by its actual value.
-
-      -> internal currency of value
-
-      motivated, targeted saccades hindering fixation of currently not
-      attended features in the field of view (i.e., oculomotor apraxia).
-
-      \item
-      only cluster 2 was congruently coupled (across MACM and RSFC) with the amygdala (involved in significance evaluation) and the nucleus accumbens (involved in reward evaluation)
-      s well as exclusively functionally asso- ciated with facial appraisal.
-Similarly, the human vPCC was the only PMC region connected to the laterobasal (rather than centromedial or superficial) nuclei group of the amygdala (Bzdok et al., 2013b), which is an amygdalar subregion probably devoted to continuously scanning environmental input for biological significance (Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011; Ghods-Sharifi et al., 2009; LeDoux, 2007).
-      \item
-      (McCoy and Platt, 2005):
-      internal representations of subjective value lie at the very core of
-      decision making,
-      task where subjective value, but not objective value, varied, as indicated by
-      subjects’ choices
-      -> value matrix is not the same in each agent
-      h the direction of impending
-      Our data extend those
-      findings, demonstrating a direct relationship between subjective preferences
-      for uncertain rewards, or the opportunity to harvest a
-      relatively large reward, and the activity of neurons thought to participate
-      in the allocation of attention
-      movement, as shown previously24,30, and the uncertainty of rewards
-      associated with this movement
-      CGp appears to carry spatial information that is scaled by
-      subjective preferences for particular patterns of reward outcome.
-      Our data indicate that CGp
-      neurons signal subjective biases for uncertain rewards (or, perhaps, the
-      potential to receive a large reward) rather than objective target value.
-      the spatial sensitivity of CGp neurons was enhanced under
-      conditions of high risk.
-      -> delay discounting / non-immediate reward -> reinforcement learning
-
-Such strategic decisions should incorporate not only recent reward history, but also opportunity costs and environmental statistics
-
-\item
-microstimulation in CGp promotes explo- ration of the previously antipreferred option.
-
-\item
-keys in a dynamic foraging task.
-Moreover, firing rates of these neurons predict in graded fashion the strategy most likely to be selected on upcoming trials.
-This encoding is distinct from switching between targets and is independent of the absolute magnitudes of rewards.
-These observations implicate CGp in the integration of individual outcomes across decision making and the modification of strategy in dynamic environments.
-recorded the activity of single cingulate cortex (CGp) neurons
-
-\item
-outcome- contingent modulations in firing rate persisted into subsequent trials.
-Moreover, firing rate on any one trial predicted switching to the alternative option on the next trial.
-Finally, microstimulation in CGp follow- ing risky choices promoted a preference reversal for the safe option on the following trial.
-Collectively, these results demonstrate that CGp directly contrib- utes to the evaluative processes that support dynamic changes in decision making in volatile environments.
-
-\item
-IN SUM, these results suggest that the PCC is part of a network that monitors the subjective outcomes of possible decisions and integrates that information into higher-level strategies
-**learn its current value and integrate this information with their statistical knowledge of the environment to predict its relative value on upcoming trials.**
-\item
-information processing in the CGp continuously adapts to changes in both the external environment and internal milieu.
-\item
-Hayden2008: adapativity of subjective reward schedules tht impact the action choices
-lesion (Gabriel, 1990) studies strongly implicate CGp in working mem- ory, attention, and general arousal (Raichle and Mintun, 2006)— all processes that contribute to adaptive decision making.
--> gradient defect
-\item
-integration of actions and their out- comes and thereby influences subsequent changes in behavior.
-We confirmed a causal role for CGp in outcome evaluation and behavioral adjustment
-
-\item HC:
-\begin{itemize}
-  \item
-  Schachter/Buckner:
-  The network described here is tied to the medial temporal lobe system, which is traditionally considered almost exclusively in the context of remembering the past. We suspect the adaptive value of
-  episodic memory is not solely in its ability to afford mental reconstruction of the past but rather in its contribution to building mental models – simulations – of what might happen next or other perspectives on the immediate envir- onment, such as what others are thinking
-
-  \item
-The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
-
-\item
-hippocampal damage [21,38–40] is associated with an impairment in thinking about one’s personal future
-
-\item
-bilateral hippocampal damage were markedly impaired relative to matched control subjects at imagining new fictitious experiences (Hassabis), their descriptions lacked spatial coherence and were instead fragmented and considerably less rich
-
-\item 
-A special case is the hippocampus whose highly recursive connection matrix is thought to function as a large ‘autoassocia- tor’, allowing the reconstruction of entire episodes from remembered fragments.
-
-\item
- structural MRI studies of expert navigators (London taxi drivers) who show increased grey matter volume in posterior hippocampus seemingly at the expense of reduced grey matter volume in anterior hippocampus (Maguire et al.
-2006b)
-
-\item
-hippocampal pyramidal cells during rest and sleep produce strongly coherent ensemble bursts believed to be critical in transferring information to the neocortex.
-
-\item
-statistical structure of hippocampal activity ->
-spatial maps of surroundings are encoded
-buszaki2004 theta oscillations:
-Using large-scale recordings, researchers can follow complex patterns for extended time periods and determine whether their modification by experience will influence self-generated patterns in the absence of envi- ronmental inputs.
-Spike sequences, imposed upon the network by behavioral manipulations, recur spontaneously during subsequent sleep
-indicating that neurons organize themselves into preferred cell assemblies, and the seeds of emergence stem from experi- ence-related activity.
-
-Neurons, mostly studied in the hippocampus, which fire action potentials only when the animal is in a specific place in a given environment (that place where a cell fires is called the “place field” of that cell).
-
-Numerous experiments have demonstrated that hippocampal neurons show place-related firing while the rat explores or traverses its environment so that each assembly of hippocampal principal cells defines a particular position of space
-
-\item
-Diba + Buzsaki 2007
-We report that temporal spike sequences from hippocampal place neurons of rats on an elevated track recurred in reverse order at the end of a run, but in forward order in anticipation of the run, coinciding with sharp waves.
-Vector distances between the place fields were reflected in the temporal structure of these sequences.
-This bidirectional re-enactment of temporal sequences may contribute to the establishment of higher-order associations in episodic memory.
-
+This can be viewed as a high-level impairment in the gathering
+of information about alternative objects (i.e., exploration) as well as
+leveraging these environmental opportunities (i.e., exploitation).
+The human PMC was congruently coupled in two functional connectivity modalities
+with the amygdala
+(involved in significance evaluation) and
+the nucleus accumbens (involved in reward evaluation).
+Specifically, the human ventral posterior cingulate cortex was
+the PMC region most connected to the laterobasal
+(rather than centromedial or superficial) nuclei group
+of the amygdala (Bzdok et al., 2013b),
+which is an amygdalar subregion probably devoted to
+continuously scanning environmental input
+for biological significance assessment
+(Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011; Ghods-Sharifi et al., 2009; LeDoux, 2007).
+%
+Indeed,
+electrophysiological recordings in animals implicated the PMC in
+strategic selection (Pearson et al., 2009), risk assessment (McCoy and Platt, 2005), and outcome-contingent behavioral modulation (Hayden et al., 2008),
+while its retrosplenial portion was
+more specifically implicated in approach-avoidance behavior (Vann et al., 2009).
+Neuron spiking activity in the PMC allowed distinguishing
+whether a monkey would persue an exploratory or exploitative
+behavioral strategy in a complex food foraging task (Pearson2009).
+Single-cell recordings in the monkey PMC
+demonstrated this brain region's sensitivity to
+subjective target utility (McCoy and Platt, 2005) and integration
+across individual decision-making instances (Pearson et al., 2009).
+This DMN node encoded the
+preference or aversion to options with uncertain reward outcomes
+and its spiking activity was more associated with
+subjectively perceived relevance of a chosen object
+than by its factual value, based on an internal currency of value
+In fact, direct stimulation of PMC neurons
+promoted exploratory action towards options
+with unsafe reward outcomes that were previously shunned.
+Graded changes in firing rates of PMC neurons
+that indicated changes in upcoming trials were
+distinct from neuronal firing indicating switching between choices.
+%
+The retrosplenial portion of the PMC can subserve evaluation of reward
+contingencies and action possibilities by integrating these with
+information from memory and altered perspective frames.
+Regarding memory retrieval, retrosplenial lesions have been
+consistently associated with anterograde and retrograde memory impairments
+of various kinds of sensory information
+in rabbits (Gabriel and Talk, 2001) and humans
+(Rudge and Warrington, 1991; Valenstein et al., 1987).
+Regarding perspective frames, this PMC subregion has been
+proposed to mediates between the organism's egocentric
+(i.e., focused on sensory input) and
+allocentric (i.e., focused on world knowledge) viewpoints
+in animals and humans
+(Burgess, 2008; Epstein, 2008; Valiquette and McNamara, 2007).
+%
+The PCC may consequently monitor the subjective outcomes
+of possible decisions and integrates that information
+with memory, perspective frames, and
+reward schedules into higher-level strategies. 
+Perceived value that differs across individuals updates
+statistical knowledge of the environment
+to predict delayed reward opportunities in the future.
+In doing so, the PMC continuously adapts to changes
+in both the external environment and internal milieu
+that modulate strategic behavioral adjustment in volatile environments
+to map possiblity spaces.
+
+The dmPFC subserves predominantly ambiguous amodal processes
+across time, space, and content domains in
+sensory-independent top-down pathways.
+This DMN node has been described as a “mental sketchpad” (Goldman-Rakic, 1996),
+potentially implicated in de-novo generation and binding
+of meaning representations instructed by stored semantics and memories.
+%
+The dmPFC may thus enable inference, representation, and assessment
+of one's own and other individuals' action and thoughts.
+For instance, neural activity in human dmPFC
+reflected expectations about other peoples actions and errors thereof
+(Suzuki et al., 2012; van Schie et al., 2004).
+dmPFC activity indeed explained the proficiency decline
+of inferring other peoples thoughts in the elderly (Moran et al., 2012).
+Some dmPFC neurons in macaque monkeys exhibited a preference
+for processing others', rather than own, behavior
+with fine-grained adjustment of contextual circumstances (Yoshida et al., 2011).
+%
+Such highly abstract neural computations necessarily rely on the
+generation of probabilistic internal information drawing from
+episodic memory retrieval, non-mnemonic scene construction processes,
+and explicit knowledge of the external world.
+
+The vmPFC subserves less ambiguous subjective-value-related evaluative processes
+reward-informed and risk estimation of self-relevant environmental stimuli.
+This DMN node is more closely associated with
+orchestrating adapted behavior by bottom-up-driven
+processing of “what matters now”.
+Quantative lesion findings across  344 human individuals confirmed
+a gross impairment in value-based decision making
+(Glaescher et al., 2016).
+The vmPFC is preferentially connected with limbic and reward-related areas.
+The vmPFC has been observed to have monosynap- tical connections with the ventral striatum (VS, which anatom- ically includes the NAc) in axonal tracing studies in monkeys (Haber et al., 1995; Ferry et al., 2000).
+The NAc is thought to be linked to reward mechanisms that may not only modulate motivated behavior towards basic survival needs, but also
+subserve model-free reinforcement learning in humans more broadly
+(O'Doherty 2015).
+This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys (Croxson et al., 2005) that quantified the VS to be substantially more likely connected to the vmPFC than dmPFC in both species.
+Two functional connectivity modalities in humans strongly connected
+the vmPFC with the nucleus accumbens (NAc), hippocampus (HC),
+posterior cingulate cortex (PCC), and retrosplenial cortex (RSC).
+%
+The vmPFC is often proposed to be involved in (external) emotional
+reactions and own (visceral) arousal.
+Real or imaginged bodily states could be mapped in the vmPFC
+as a bioregulatory disposition governing cognition
+and decision making (Damasio, 1996; Nauta, 1971)
+In studies of decision making and neuroeconomics, ventromedial prefrontal cortex (vmPFC) con- sistently reflects an individual’s subjective valuation (Behrens et al., 2008; Boorman et al., 2009; Knutson et al., 2005; Plass- mann et al., 2007)
+This may be why performance within and across participants
+was related to state encoding in OFC.
+Such a “cognitive map” of action space was argued to encode
+the current task state even when states are unobservable from sensory input,
+which was shown to be critical for behavior.
+The GMV of the vmPFC and VS correlated with indices of social reward attitudes and value-guided behavior (Lebreton et al., 2009).
+Additionally,
+independent whole-brain analyses from structural
+neuroimaging studies related the gray-matter volume (GMV) of the vmPFC
+(more consistently than any other
+brain region) to indices of
+social competence and social network complexity,
+among the most complicated decision that humans and monkeys take
+(Lebreton et al., 2009; Powell et al., 2010; Lewis et al., 2011; Sallet et al., 2011).
+%
+
+The HC is well known to be involved in memory and
+spatial navigation in animals and humans
+(von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
+Its highly recursive anatomical architecture
+may be specifically designed to allowing reconstruction
+whole episode of experience from memory fragments.
+%
+While the HC in the medial temporal lobe system
+is traditionally believed to allow remembering the past,
+there is now increasing evidence for a role
+in constructing mental models in general (Schachter/Buckner).
+Indeed,
+hippocampal damage [21,38–40] is
+not only associated with an impairment in reexperiencing the past (i.e., amnesia),
+but also thinking about one’s own future and
+imagining new fictitious experiences more broadly (Hassabis/PNAS).
+Mental scenes created by hippocampus-lesioned patients exposed a lack of
+spatial integrity, richness in detail, and overall coherence.
+%
+Single-cell recordings in the animal hippocampus revealed
+some constantly active neuronal ensembles whose firing coincided with
+distinct locations in space while the animal navigated through its surroundings.
+London taxi drivers, individuals with high performance in spatial navigation,
+were shown to exhibit increased grey matter volume in the
+posterior hippocampus (Maguire et al. 2006b).
+But encoding and generative reconstruction in the hippocampus extends
+beyond mere spatial knowledge of the environment.
+Based on large-scale recordings of hippocampal neuronal populations,
+complex spiking patterns can be followed across extended periods including
+their modification of input-free self-generated patterns
+after environmental events (buszaki2004 theta oscillations).
+Specific spiking sequences, that were elicited in experimental task conditions,
+have been shown to be reenacted spontaneously during sleep.
+Moreover, spike sequences measured in hippocampal place cells of rats
+featured reoccurred directly after experimental trials
+as well as directly before upcoming experimental trials (Diba + Buzsaki 2007).
+Such hippocampal ensemble burts during rest and sleep
+have been proposed to be critical in communicating local information
+to the neocortex for long-term storage.
+These mechanisms of the hippocampus probably make important contributions to the
+recollection of autobiographical memory episodes and other
+reexperienced or newly generated mental scenarios (Hassabias2007).
+%
+The HC thus orchestrates elements of experienced environmental aspects for
+consolidations based on reenactment and for integration into 
+rich mental scene construction. In this view, the HC my even influence
+ongoing perception in the current environment (Maguire 2016).
 -> self-play in deepmind Go
-
-\item
-coding, transfer, and consolidation of information:
-This emerging picture of hippocampal dy- namics suggests that neurons at the opposite ends of the distributions may convey differ- ent but complementary types of information.
-The ever-active minority of place cells may be responsible for generalizing across envi- ronments and affords the brain the capacity to regard no situation as completely un- known because every alley, mountain, river, or room has elements of previously experi- enced similar situations.
-In many situations,
-this minority provides the “best guess” of the hippocampus and offers “good enough” solutions to get by.
-On the other hand, the majority of less active neurons constitute a large reservoir that can be mobilized (7) to precisely distinguish one situation from an- other and incorporate novel ones as distinct.
-
-\item
-damage to even the hippocampus alone is sufficient to cause amnesia +
-envisoning themselves in future and even hypothetical scenarios/mentally constructing new experiences
-
-The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
-
-\item
-Hassabias2007:
-Recollection of these rich autobiographical or episodic memories has been likened to mentally traveling back in time and re-experiencing one’s past (1).
-It has long been known that the hippocampus and related medial temporal lobe structures play a critical role in supporting episodic memory
-
-In fact, episodic memory and imagining or constructing events share striking similarities in terms of the psychological processes engaged (19–21).
-These include imagery (22), sense of presence (1), retrieval of semantic information and multimodal details (23), and narrative structure (22).
-
-visual imagery itself is not believed to be realized in the HC
-
-\item
-Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later
-
-\end{itemize}
-
-\item
-the aRTPJ network is central for sensorimotor control by integrating supramodal stimulus-guided attention and action initia- tion during externally structured tasks.
-
-\item
-visuo- proprioceptive conflict (Balslev et al., 2005), and multi-modal detection of sensory changes (Downar et al., 2000)
-brain areas connected to aRTPJ further responded to all changes of visual, auditory, or tactile stimulation in a multi-modal fMRI study (Downar et al., 2000).
-
-\item
-direct electrical stimulation of the RTPJ during neurosurgery was associated with altered perception and stimulus awareness (Blanke et al., 2002).
-
-\item
-the left and right TPJ further differ according to their cytoarchitectonic borders and gyral pattern (Caspers et al., 2008; Caspers et al., 2006)
-
-\item
-Neuroscientific investigations on hemispheric specialization have converged to the left cerebral hemisphere as dominant for language functions (Broca 1865, Wernicke 1874, Lichtheim 1885) and the right hemisphere as dominant for attention functions (Gazzaniga, Bogen et al.
-1965, Sperry 1982, Stephan, Marshall et al.
-2003).
-  \end{itemize}
-
-  \item
-    RTPJ:
-    \begin{itemize}
-\item
-Sridharan: RTPJ involved in bottom-up saliency detection
-
-\item
-Corbetta2008:
-Although both attentional networks have been most exten- sively investigated in vision, the available evidence indicates a supramodal function (Driver and Spence, 1998; Macaluso et al., 2002).
-The ventral network (right TPJ, right IFG) registers salient events in the environment not only in the visual but also in the auditory and tactile modalities (Downar et al., 2000)
-Several lines of evidence indicate that two cortico-cortical neural systems are involved in attending to environmental stimuli (Corbetta and Shulman, 2002).
-A dorsal frontoparietal network.
-A second system, the ventral frontoparietal network, is not ac- tivated by expectations or task preparation but responds along with the dorsal network when behaviorally relevant objects (or targets) are detected (Corbetta et al., 2000).
-Both dorsal and ventral networks are also activated during reorienting, with en- hanced responses during the detection of targets that appear at unattended locations.
-For example, enhanced responses are observed when subjects are cued to expect a target at one location but it unexpectedly appears at another (i.e., ‘‘invalid’’ targets in the Posner spatial cueing paradigm)
-or when a target appears infre- quently, as in ‘‘oddball’’ paradigms.
-
-Intimately related to prediction error signaling:
-An early theory of how the two networks interact (Corbetta and Shulman, 2002) proposed that when attention is reoriented to a new source 
-of information (stimulus-driven reorienting), output from the ven- tral network interrupts (as a ‘‘circuit breaker’’) ongoing selection in the dorsal network, which in turn shifts attention toward the novel object of interest.
-
-relevance is the critical factor that determines whether an object activates the ventral network (Downar et al., 2001).
-The ventral network might be considered a prime candi-date for mediating orienting to salient but unimportant stimuli, i.e., exogenous attention (Posner and Cohen, 1984), be- cause under passive conditions it is highly responsive to distinctive sensory
-events in all modalities (Downar et al., 2000). The ventral network is well activated by stimuli that are important, even if they are not very distinctive. 
-CURRENT OPINION: VA network does underlie  stimulus-driven reorienting to environmental stimuli based on their task relevance.
-An important conclusion from these neuroimaging studies is that the psychological dis- tinction between exogenous and endogenous orienting (Jo- nides, 1981) may not map onto different neural systems.
-"changing the control of behavior from one environ- mental input to another".
-Similar reorienting mechanisms may also be involved in shifting from a broad range of ‘‘internally directed’’ processes in order to deal with environmental events, as when interrupting memory retrieval (‘‘did I lock the car door?’’) to respond to a sudden stimulus (‘‘is that my cell phone ring- ing?’’).
--> medites between the Default-Mode function and other more environmentally directed processing goals.
-inducing network reconfiguration
-
-\item
-Bzdok2013
-
-qualify RTPJ as a po- tential switch between exteroceptive and interoceptive mind sets implemented by their networks.
-The reciprocal nature of those mind sets is well illustrated by consistent reports of continuous shifting be- tween externally (i.e., sensory) and internally (i.e., self- and presumably social/memory-) oriented processing in controlled laboratory settings and daily routine (Smallwood et al., 2007).
-\end{itemize}
-
-\item
-Corbetta2002:
- We suggest that the TPJ is also strongly engaged by stimuli that are behaviourally relevant but require a change in the current task set. 
- -> exogeneously induced state transition?
- <=> MCMC in HC / PCC -> slow? endogeneous changes in state transitions by the policy and value matrices
- For example, the alarm at the museum is not part of the task set of listening to the guide’s discussion of Hieronymous Bosch, but it is clearly a behaviourally relevant stimulus.
-
-\item
-  LTPJ
-  \begin{itemize}
-    \item 
-    Wernicke's area:
-    involved in accessing and integrating world knowlege.
-    involved in the comprehension or understanding of written and spoken language Damage caused to Wernicke's area results in receptive, fluent aphasia.
-This means that the person with aphasia will be able to fluently connect words, but the phrases will lack meaning.
-    A major impairment of language comprehension, while speech retains a natural-sounding rhythm and a relatively normal syntax.
-Language as a result is largely meaningless.
-difficulty finding words, putting words in the wrong order, substituting sounds and/or words, difficulty understanding spoken utterances, difficulty recognizing some words by sight, difficulty writting letters
-
-\item
-significantly reduced impact of intentions after transient TPJ disruption (Young et al. 2010b)
-
-\item
-complex conceptual linguistic functions are proposed to be more strongly left-lateralized -> in line with overall left-lateralization of human language
-
-\item
-Episodic and explicit memory retrieval strongly draws on complex semantic processing and contributes to social cognitive processes (see section 4.4.).
-Syntactic processing, on the other hand, is a core language process that is closely intermingled with semantic processing.
-It refers to the hierarchical sequencing of words and their meanings (Price, 2010) and is mandatory for sentence processing in both social cognitive and language tasks.
-
-\item a lesion study on disturbed sleep (i.e., a state of mind
-independent of sensory stimulation but dependent on
-internally generated information) exclusively identified
-the dmPFC (Koenigs et al., 2010). Third, another VLSM study exclusively
-related the IFG and TPJ, both more strongly connected to the
-dmPFC in our study, to inner speech
-\end{itemize}
-
-\item
-the difference between the two task groups could be parsimoniously explained by the required atten- tion to either the external world or self.
+-> Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later
+
+Finally,
+the right and left TPJ are known to have hemispheric differences
+according to their cytoarchitectonic borders and gyrification pattern
+(Caspers et al., 2008; Caspers et al., 2006).
+Neuroscientific investigations on hemispheric functional specialization
+have converged to the right versus left cerebral hemisphere as dominant for
+attention versus language functions (Broca 1865, Wernicke 1874, Lichtheim 1885;
+Gazzaniga, Bogen et al. 1965, Sperry 1982, Stephan, Marshall et al. 2003).
+%
+The aRTPJ is a key node network is central for
+action initiation during externally structured tasks and
+sensorimotor control by integrating supramodal stimulus-guided attention
+(Corbetta et al., 2002).
+Involvement of this DMN node was repeatedly reported in
+visuo-proprioceptive conflict (Balslev et al., 2005) and
+multi-modal detection of sensory changes across
+visual, auditory, or tactile stimulation in a multi-modal fMRI study
+(Downar et al., 2000).
+In humans, direct electrical stimulation of the
+RTPJ during neurosurgery was associated with altered perception
+and stimulus awareness (Blanke et al., 2002).
+%
+Importantly, the RTPJ has been shown to be intimately related to
+supramodal prediction error signaling.
+Neural activity in the RTPJ has been argued to be responsible 
+for stimulus-driven attentional reallocation to 
+salient and surprising sources of information
+as a ‘‘circuit breaker’’ that recalibrates control and maintenance systems
+(Bzdok et al., 2013; Corbetta et al., 2008).
+In the face of large discrepancies between actual and previously predicted
+environmental events the RTPJ acts a potential switch between
+externally-oriented mind sets focussed on the
+sensory world and internally-oriented mind sets focussed
+on self-relevant mind-wandering of possible mental scenarios.
+Transient RTPJ in humans for instance disrupted the
+impact of predicted intentions of other individuals (Young et al. 2010b),
+a capacity believed to be subserved by the DMN.
+The RTPJ might hence be an important player that shifts away
+from the default active ‘‘internally directed’’ processes
+to deal with immediate environmental objects and contexts.
+-> (slow) endogeneous changes in state transitions by the policy and value matrices
+For example, the alarm at the museum is not part of the task set of listening to the guide’s discussion of Hieronymous Bosch, but it is clearly a behaviourally relevant stimulus. <=> MCMC in HC / PCC
+
+The left TPJ in turn exhibits a close topographical relationship to
+Wernicke's area
+involved in the comprehension or understanding of written and spoken language.
+Neurological patients with damage caused to Wernicke's area
+have a major impairment of language comprehension
+when listening to others or reading a book.
+Their speech
+preserves natural rhythm and about normal syntax, yet the
+voiced sentences are devoid of meaning (i.e., aphasia).
+Abstracting from the typical semantic interpretations in linguistics
+and neuropsychology,
+the LTPJ probably mediates access to and integrating of world knowlege.
+For instance, LTPJ lesions also entail problems in recognizing
+others' pantomimed action towards objects
+without obvious relation to processing any language content
+(Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
+%
+Also inner speech hinges on knowledge retrieval of statistical structure
+about the physical and inter-personal world.
+Multivariate volumetric analyses showed that the internal production of
+formulated thought ("language of the mind") is closely related to the left TPJ
+(Geva et al., 2011).
+A number of neuroimaging studies demonstrated LTPJ activity responses
+to sementically aberrant words in sentences, which suggests a role
+in binding a given semantic concept into an integrated whole.
+Further,
+episodic memory recall and imagination strongly draws on
+complex world knowledge.
+Isolated building blocks of world statistics probably get reassembled
+in internally generated visual scenarios that
+navigate present action, weigh hypothetical possibilities, and forcast the future.
+%
+The LTPJ may hence facilitate the automated prediction of events
+by incorporating experience-derived models of the world
+into ongoing action, planning, and problem solving.
 world -> policy matrix driven computation
 self -> value matrix driven computation
-
-\end{itemize}
 

### 2016-09-25 paper/paper.tex
minor fixes

@@
 have been proposed to be critical in communicating local information
-to the neocortex for long-term storage.
+to the neocortex for long-term storage, which includes the DMN.
 These mechanisms of the hippocampus probably make important contributions to the
@@
 from the default active ‘‘internally directed’’ processes
-to deal with immediate environmental objects and contexts.
+to deal instead with immediate environmental objects and contexts.
 -> (slow) endogeneous changes in state transitions by the policy and value matrices

### 2016-09-26 paper/paper.tex
minor things

@@
 autonomous learning systems
-\textbf{\\Get opinion from}: Danielle Bassett, Kai Vogeley, Mohammad Shakir
+\textbf{\\Get opinion from}: Danielle Bassett, Kai Vogeley, Mohammad Shakir,
+Guillaume Dumas, Smallwood
 
@@
 
+The DMN can be viewed as a smoothing kernel extending
+from the relatively recent past to the relatively close future
+that constantly estimates complex non-linear gain functions.
+-> This is why neuroimaging research revealed activity increases in the DMN
+when 
+-> This may be why perfect memory has been evolutionarily de-selected,
+although it is provably feasible by the neurobiological hardware in humans
+
 \paragraph{Acknowledgment}

### 2016-09-26 paper/paper.tex
first intro draft

@@
 %
-Many research streams agree on its likely role in evolutionarily adaptive envisioning of past and hypothetical scenarios to predict the environment.
+Many research streams agree on its likely role in evolutionarily adaptive
+envisioning of past and hypothetical scenarios to predict the environment.
 The DMN would hence be dedicated to continuous autobiographical memory retrieval,
@@
 and reward contingency evaluation when letting the mind go.
-It explains its highest energy consumption in the brain and its intimate coupling with conscious awareness.
+It explains its highest energy consumption in the brain and
+its intimate coupling with conscious awareness.
 %
@@
 
-\section{The Default-Mode Network}
+\section{Introduction}
 %
 When left unperturbed, the human brain is not at rest.
-Rather, the brain continues to metabolize large quantities of oxygen and glucose energy to maintain inter-neuronal information transfer in the absence of an externally structured task (Bzdok and Eickhoff, 2016a).
-This baseline energy demand is subject to surprisingly little modulations due to the cognitive load posed by environmental challenges.
-What has early been described as the "stream of consciousness" in psychology (James, 1890).
-%
-A few years ago, the so-called "default mode network" was discovered entirely by accident (Shulman et al., 1997).
-A coherent set of brain regions consistently increased in neural activity during stimulus-independent thought, which was proposed to reflect the neural correlates underlying unfocused everyday mind wandering (Raichle et al., 2001).
-That is, in the beginning of the 21st century, brain imaging was the first technology to allow for the discovery of a coherent brain network that subserves a set of baseline mental activity (Bzdok et al., 2012; IBzdok et al., 2015).
-%
-15 years ago neuroscientists have discovered an energy-demanding brain network with unique properties.
-The "default-mode network" (DMN) appeared exclusive in task-induced deactivation during various psychological experiments.
-It was later argued to be systematically anti-correlated with brain regions that subserve task performance.
-Today, many authors speculate that the DMN implements continuous thinking about others’ and one’s own mind states, as well as adaptive envisioning of past, hypothetical, and future events.
-This network might have emerged to continuously predict environmental events using mental imagery as an evolutionary advantage.
-%
-Since 2001, the neurobiological properties of the default mode network have been investigated and reported in more than 3,000 neuroimaging publications 
-The DMN was initially believed to represent the neural correlates of unconstrained mind-wandering because of its pervasively observed activity decrease during a large array of tasks.
-This processing of unknown information categories in the DMN has been argued to mediate an evolutionarily conserved function for the individual.
-This is all the more likely because the DMN contains the two biggest hotspots of energy consumption in the entire central nervous system.
-DMN activity also persists to a substantial degree during the early stages of sleep (Horovitz et al., 2008) and under anesthesia (Greicius et al., 2008).
-However, information processing in the DMN has also repeatedly been shown to impact human behavior.
-Goal-directed task performance improves with decreased activity in default-mode areas (Weissman et al., 2006) and increased DMN activity is linked to more task-independent, potentially disturbing thoughts (Mason et al., 2007).
-%
-the DMN is likely to control the interplay between perception–action cycles and mental imagery (Bzdok et al., 2013a; Bzdok et al., 2013b).
-In particular, the DMN is likely to subserve a computational mechanism by its connections to highly associative brain areas that is key for human thought across time, splace, and content domains at the interface of external world and self (IBzdok et al., 2015).
-%
-
-hierarchically deep models of the world
-
-Its societal and economic relevance is certified by involvement in various psychiatric disorders, including schizophrenia, autism, and depression.
-
-sensory exchanges with the world
-
-show how a formal approach can provide generic explanations for psycho- pathology that are physiologically grounded
-
-quantitative (and parametric) characterisations offered by computational
-
-d Biological systems are homeostatic (or allostatic), which
-means that they minimize the dispersion (entropy) of their
-interoceptive and exteroceptive states.
-d Entropy is the average of surprise over time, which means
-that biological systems minimize the surprise associated
-with their sensory states at each point in time.
-d In statistics, surprise is the negative logarithm of B
-
-perception as hypothesis testing
-
-Consider an organism that, in their present situation, is confronted by several choices of what to do next.
-Being able to accurately and richly mentally enact possible future states before making a decision would help to evaluate the desirability of different outcomes and also the planning processes needed to make them happen.
-
- the use of this constructive process goes far beyond simply predicting the future, to the general evaluation of fitness for purpose
-
-scene construction processes when envisioning past, fictitious, and future events (Hassabis et al., 2007; Spreng et al., 2009).
-
-It was observed that the more external stimuli are predictable, the more reflection processes become detached from the actual sensory environment and the more stimulus-independent, self-focused thoughts occur (Mason et al., 2007).
-
-\begin{itemize}
-  \item includes brain regions with highest baseline energy consumption in humans
-  \item most active in non-disturbed mind-wandering without external influence
-  \item tends to deactivate when humans engage in goal-direct tasks
-  \item fluctuations in the DMN are related to lapses during and performance in
-  such externally structured/focused tasks
-  \item includes brain regions that come online first when waking up from
-  anesthesia
-  \item its network nodes are late to myelinate, an indicator of cognitive sophistication
-  \item includes the most advanced processes hierarchies and has
-  no direct connections with sensory regions processing external input
-  \item the DMN's conceivable key role in the continuous environmental
-  tracking in a generative, integrative process might explain both its highest energy consumption in the brain and its intimate coupling with conscious awareness
-  \item Indeed, patients with right IPL (=DMN) damage have particular difficulties
-  with multistep actions (Hartmann et al. 2005)
-  and with imagined action (Sirigu et al. 1996).
-  \item Lesion of the hippocampus, feeding memory and spatial information, to the
-  DMN impairs future and hypothetical thinking (cf. Hassabis PNAS)
-
-  \item
-  autobiographical memory sup- plies numerous building blocks of social semantic knowl- edge (Bar 2007; Binder et al.
-2009).
-These isolated conceptual scripts may be reassembled to enable the fore- casting of future events (Tulving 1983, 1985; Schacter et al.
-2007).
-Similar brain mechanisms in remembering past episodes and envisioning future circumstances is supported by their engagement of identical brain areas, as evidenced by a quantitative meta-analysis (Spreng et al.
-2009).
-Moreover, retrograde amnesic patients were repor- ted to be impaired not only in prospection but also in imagining novel experience (Hassabis et al.
-2007).
-These findings suggest a single neural network for mentally constructing plausible semantic scenarios of detached sit- uations regardless of temporal orientation (Buckner and Carroll 2007; Hassabis and Maguire 2007).
-Indeed, con- struction of detached probabilistic scenes has been argued to influence ongoing decision making by estimating out- comes of behavioral choices (Boyer 2008; Suddendorf and Corballis 2007; Schilbach et al.
-2008b).
-  \item
-  DMN exists in monkeys [73] and rats [74]. 
-  \item clinical research has corroborated default-mode network dysfunction in various psychiatric and neurological disorders
-\end{itemize}
-
-The DMN poses an unprecedented challenge to our neuroscientific understanding and methodological inventory.
-
-Despite ramifications in health and disease, the functional role of the DMN poses an unprecedented challenge to the conventional interpretation strategies and existing methodological arsenal in the neurosciences.
-
-Represents the physiological and pathological instantiation of a human beings' default mental repertoire, the nature of which remains largely obscure.
-
-The DMN has now urged major attention in the neuroscientific community (Raichle et al., 2001, >6600 citations).
-
-\section{Control Theory and Reinforcement Learning}
-\paragraph{Control Theory.}
-\begin{itemize}
-\item mathematically formalized dynamics systems with input, output and
-feedback loops that produce error signals for adaptation
-\item typically realized by (non-linear) differential equations
-\item has been used to describe neural systems in the past
-\item gradient update = learn/adapt system dynamics by error occurrence
-\item related to notions of game theory
-\item switching between states -> typically reduces to a sequential processing model
-\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
-\item policy: action to take given a state
-\item value: prediction of future reward
-\item model: predict what the environment will do next
-\item model-baseed RI learning: we try to understand the environment; it is initially unknown
-\item special to RL: exploitation vs. exploration
-\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
-\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
-\item Tenenbaum2011Science: Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines (70–74).
-\end{itemize}
+Rather, the brain continues to metabolize large quantities of
+oxygen and glucose energy to maintain inter-neuronal information
+transfer in the absence of concrete behavioral goal (Bzdok and Eickhoff, 2016a).
+This baseline energy demand is subject to surprisingly little modulations
+when mentally processing environmental challenges.
+What has early been described as the "stream of consciousness"
+in psychology (James, 1890), found its neurobiological manifestation 
+in the so-called "default mode network" (DMN) was discovered entirely by accident
+(Shulman et al., 1997). This regrouping of some of the highest regions in
+the human association cortex (Mesulam et al., 1998) consistently
+increased in neural activity during
+unfocused everyday mind wandering (Raichle et al., 2001). That is, in
+the beginning of the 21st century, brain imaging was the first technology
+to allow for the discovery of a unique brain network that subserves a
+set of baseline mental activity (Bzdok et al., 2012; Bzdok et al., 2015).
+
+The "default-mode network" (DMN) appeared to be exclusive in task-induced
+deactivation during various psychological experiments.
+This processing of unknown information categories in the DMN has
+been argued to mediate an evolutionarily conserved function for the individual.
+This is all the more likely because the DMN contains the two
+biggest hotspots of energy consumption in the entire central nervous system.
+DMN activity also persists to a substantial degree during the early stages of
+sleep (Horovitz et al., 2008) and under anesthesia (Greicius et al., 2008).
+Today, many authors believe that the DMN implements
+probabilistic estimation of past, hypothetical, and future events. This
+brain network
+might have emerged to continuously predict environmental events using
+mental imagery as an evolutionary advantage.
+%
+However, information processing in the DMN has also repeatedly
+been shown to directly impact human behavior. Goal-directed task performance
+improved with decreased activity in default-mode areas (Weissman et al., 2006)
+and increased DMN activity was linked to more task-independent,
+potentially disturbing thoughts (Mason et al., 2007). 
+%
+The overarching function of the DMN is mystified by
+controlling the interplay between perception–action cycles and
+maintaining a baseline contemplations
+across time, splace, and content domains at the interface of external world
+and self (Bzdok et al., 2015).
+
+Consider an organism faced by choosing the next action
+and being informed by rich
+reenactment of really happened, hypothetically conceived, and
+predicted future events to perfect ongoing decision making.
+Control theory is a particularly attractive computational framework
+to describe, quantify, and predict the autonomously acting systems,
+perhaps including the human brain.
+It is known that, the more external world is predictable,
+the more mental activity becomes detached from the actual sensory environment
+and the more stimulus-independent thoughts occur
+(Mason et al., 2007)
+that may contribute to optimized control of the organism.
+A policy matrix captures the possible action than can be performed
+on the environment given a current state. A value matrix 
+maps maps expected rewards to environmental objects and events.
+Switching between states reduces to a sequential processing model.
+In a model-free reinforcement learning approach,
+we try to understand the environment always better
+starting, at birth, with an environment that initially unknown.
+Informed by gradient updates,
+the adaptive DMN dynamics learned by feed-back loops
+that are calibrated by error occurrence.
+Reinforcement learning frameworks can naturally embed DMN function
+into the tension between exploitative action with immediate reward and
+explorative action with longer-term reward schedules.
+The diversity of DMN implications in human cognitive processes
+can be parsimoneously recast as minimizing prediction error
+and maximizing reward outcome at various time scales.
+Such a biologically meaningful optimization objective
+may be solved by a stochastic approximation
+based on a brain implementation of Markov Chain Monte Carlo sampling
+(Tenenbaum2011Science). Even (necessarily imperfect) memories
+of past experience, random mind-wandering and dreams during sleep
+may be purposeful as building blocks to iteratively improve
+the predictive DMN machinery to optimize behavior of the organism.
 
 \section{Known Neurobiological Properties of the Default-Mode Network}
-mPFC and PCC = highest metabolic consumption in the brain/highest level of basal glucose energy consumption in humans
-
-scene con- struction processes when envisioning past, fictitious, and future events (Hassabis et al., 2007; Spreng et al., 2009; Bzdok et al., 2013).
-it is not random but purposeful
-draws on conceptual knowledge acquired through experience
-operating on internal sources of information
-Computational processes with a focus on the internal world,
-when the environment is highly predictable and
-external stimuli are irrelevant.
-involvement of the PMC across the boundaries of classical cognitive domains
-
-mechanistic relevance of each node -> we deconstruct
-
-PFC
-Patients with prefrontal lesions have problems in adapting to novel situations and stimuli
-Stuss, D. T. \& Benson, D. F. The Frontal Lobes (Raven, New York, 1986).
-
-vmPFC regions respond more to the value difference that will be acted on in the current trial (executed in red), and dmPFC to the currently irrelevant value difference (modeled in blue) {Nicolle, 2012 \#3750} individual.
-Instead activity in vmPFC reflects a choice preference that is executed and rostral dmPFC a choice preference that is modeled.
-Thus, dorsal regions of rostral mPFC contain value representations in the frame of reference of a modeled (as opposed to an actuated) preference irrespective of whether this applies to another’s or to one’s own likely actions and goals.
-Likewise, ventral mPFC contains a representation of value in the frame of reference of an executed choice, even if this executed choice reflects one’s own or another’s preferences.
-dmPFC that reflects the values and preferences of another individual (here temporally discounted at a rate specific to the individual), even when they are not directly relevant to the task at hand.
-
-Simony 2016
-This implies that the processing timescale of the DMN is longer than the length of single paragraphs (B40s) and that information accumulated from prior paragraphs can affect ongoing processing in this system
-Our results suggest that the DMN accumulates and integrates information over minutes.
-Stimulus-induced correlation patterns in the DMN were largest for the intact story, greatly decreased in the paragraph scramble condition, and were almost entirely abolished in the word scramble condition.
-
+We will attempt to deconstruct the overall DMN function
+into the mechanistic relevance of each individual network node
+by reviewing experimental neuroscience evidence.
+%
+The PMC and the dmPFC within the DMN
+are responsible for the highest metabolic turn-over
+of glucose energy consumption in humans (Raichle).
 The PMC myelinates relatively late during postnatal development in monkeys
@@
 electrophysiological recordings in animals implicated the PMC in
-strategic selection (Pearson et al., 2009), risk assessment (McCoy and Platt, 2005), and outcome-contingent behavioral modulation (Hayden et al., 2008),
+strategic selection (Pearson et al., 2009), risk assessment (McCoy and Platt, 2005),
+and outcome-contingent behavioral modulation (Hayden et al., 2008),
 while its retrosplenial portion was
@@
 distinct from neuronal firing indicating switching between choices.
+Also in humans, 
+the DMN has been shown to gather and integrate information
+over different paragraphs in an fMRI study on auditory narratives (Simony 2016).
 %
@@
 of one's own and other individuals' action and thoughts.
-For instance, neural activity in human dmPFC
+Patients with neurological lesions in the prefrontal cortex
+are now to adapt to novel situations and stimuli with difficulty
+(Stuss, D. T. \& Benson, D. F. The Frontal Lobes, Raven, New York, 1986).
+Specifically, neural activity in human dmPFC
 reflected expectations about other peoples actions and errors thereof
@@
 and explicit knowledge of the external world.
+%
+According to computational neuroimaging experiments,
+the dmPFC activity preferentially models stimulus-value associations of
+possible actions that are not actually executed,
+whereas the vmPFC activity models value in the environment that leads performed actions
+{Nicolle, 2012 \#3750}.
 
@@
 (Bzdok et al., 2013; Corbetta et al., 2008).
+Indeed, patients with right TPJ damage have particular difficulties
+with multistep actions (Hartmann et al. 2005).
 In the face of large discrepancies between actual and previously predicted
@@
 
+At the least, we propose an alternative vocabulary to
+describe and discuss neuroscientific studies on the DMN.
+
 The DMN can be viewed as a smoothing kernel extending

### 2016-09-27 paper/paper.tex
updates in a train

@@
 environmental assessment and prediction.
-DMN function is recast from the mathematical perspective of control theory and
+DMN function is recast in mathematical terms
+from the perspective of control theory and
 reinforcement learning.
@@
 in the machine-learning domain.
-We argue that this engineering account can parsimoneously
-explain a variety of existing neuroscientific hypotheses of DMN function.
+We argue that this engineering approach to DMN function can parsimoneously
+explain a variety of existing neuroscientific hypotheses.
 %
-Formal models of the computational processes subserved by the DMN
+Formal models of the computational processes subserved in the DMN
 could offer statistical control of predictive human behavior.
@@
 \textbf{\\Get opinion from}: Danielle Bassett, Kai Vogeley, Mohammad Shakir,
-Guillaume Dumas, Smallwood
+Guillaume Dumas, Smallwood, Tenenbaum
 
@@
 oxygen and glucose energy to maintain inter-neuronal information
-transfer in the absence of concrete behavioral goal (Bzdok and Eickhoff, 2016a).
+transfer in the absence of a focused behavioral goal (Raichle 2001;
+Shulman, 1999).
 This baseline energy demand is subject to surprisingly little modulations
-when mentally processing environmental challenges.
+when mentally processing environmental challenges
+(Bzdok and Eickhoff, 2016a).
 What has early been described as the "stream of consciousness"
 in psychology (James, 1890), found its neurobiological manifestation 
-in the so-called "default mode network" (DMN) was discovered entirely by accident
-(Shulman et al., 1997). This regrouping of some of the highest regions in
+in the so-called "default mode network" (DMN)
+(Shulman et al., 1997). This networked collection of some of the highest regions in
 the human association cortex (Mesulam et al., 1998) consistently
@@
 the beginning of the 21st century, brain imaging was the first technology
-to allow for the discovery of a unique brain network that subserves a
-set of baseline mental activity (Bzdok et al., 2012; Bzdok et al., 2015).
+to allow for the discovery of a unique brain network that subserves
+baseline mental activity (Bzdok et al., 2012; Bzdok et al., 2015).
 
-The "default-mode network" (DMN) appeared to be exclusive in task-induced
+The DMN appeared to be exclusive in task-induced
 deactivation during various psychological experiments.
@@
 biggest hotspots of energy consumption in the entire central nervous system.
-DMN activity also persists to a substantial degree during the early stages of
+DMN activity also persists to a substantial degree during
 sleep (Horovitz et al., 2008) and under anesthesia (Greicius et al., 2008).
@@
 and increased DMN activity was linked to more task-independent,
-potentially disturbing thoughts (Mason et al., 2007). 
+yet sometimes useful thoughts (Mason et al., 2007; Seli et al., 2016). 
 %
-The overarching function of the DMN is mystified by
-controlling the interplay between perception–action cycles and
-maintaining a baseline contemplations
-across time, splace, and content domains at the interface of external world
-and self (Bzdok et al., 2015).
+Understanding of the overarching DMN function is complicated by
+simultaneously controlling perception–action cycles and
+maintaining baseline contemplations
+across time, splace, and content domains
+interfacing the external world and the self.
 
-Consider an organism faced by choosing the next action
-and being informed by rich
+Consider an agent faced by choosing the next action
+and being guided by rich
 reenactment of really happened, hypothetically conceived, and
-predicted future events to perfect ongoing decision making.
-Control theory is a particularly attractive computational framework
-to describe, quantify, and predict the autonomously acting systems,
-perhaps including the human brain.
-It is known that, the more external world is predictable,
+predicted future events to perfect ongoing behavioral performance.
+A particularly attractive computational framework
+to describe, quantify, and predict autonomously acting systems like the brain
+can be the combination of control theory and reinforcement learning.
+It is known that, the more the external world is predictable,
 the more mental activity becomes detached from the actual sensory environment
 and the more stimulus-independent thoughts occur
-(Mason et al., 2007)
-that may contribute to optimized control of the organism.
-A policy matrix captures the possible action than can be performed
-on the environment given a current state. A value matrix 
+(Mason et al., 2007).
+These "offline" processes may however contribute to optimized control of the organism.
+We formalize
+a policy matrix to capture the space of possible actions that can be performed
+on the environment given the current state. A value matrix 
 maps maps expected rewards to environmental objects and events.
 Switching between states reduces to a sequential processing model.
-In a model-free reinforcement learning approach,
+In a model-based reinforcement learning approach,
 we try to understand the environment always better
-starting, at birth, with an environment that initially unknown.
-Informed by gradient updates,
-the adaptive DMN dynamics learned by feed-back loops
+starting, at birth, with an environment that is completely unknown.
+Informed by outcomes of performed action,
+the DMN dynamics are constantly adapted in feed-back loops
 that are calibrated by error occurrence.
-Reinforcement learning frameworks can naturally embed DMN function
-into the tension between exploitative action with immediate reward and
+A DMN framework incorporating reinforcement learning
+can naturally embed human behavior
+into the tension between exploitative action with immediate gains and
 explorative action with longer-term reward schedules.
-The diversity of DMN implications in human cognitive processes
+The implication of the DMN in a diversity of human cognitive processes
 can be parsimoneously recast as minimizing prediction error
-and maximizing reward outcome at various time scales.
-Such a biologically meaningful optimization objective
+based on pervasive probabilistic simulations,
+thus maximizing reward outcome at various time scales.
+Such a purposeful optimization objective
 may be solved by a stochastic approximation
@@
 of past experience, random mind-wandering and dreams during sleep
-may be purposeful as building blocks to iteratively improve
-the predictive DMN machinery to optimize behavior of the organism.
+may provide meanginful building blocks to iteratively improve
+the predictive DMN machinery to optimize the behavior of the organism.
 
@@
 We will attempt to deconstruct the overall DMN function
-into the mechanistic relevance of each individual network node
-by reviewing experimental neuroscience evidence.
+by the mechanistic relevance of each individual network node
+based on published evidence from neuroscience experiments.
 %
@@
 Physiological and disturbed metabolic fluctuations in the
-human PMC have been repeated related to
+human PMC have been repeatedly related to
 phenomena of changed conscious awareness,
@@
 The PMC has long been speculated to reflect constant computation of
-the (external) environmental statistics and its (internal) representation
+the environmental statistics and its internal representation
 as an "inner minds eye" (Cavanna et al., 2006).
-For instance, Bálint's syndrome is a neurological disorder of conscious
-awareness that resulting from damage in the bilateral parietal cortex.
-Patients are plaigued by an
+For instance, B\'alint's syndrome is a neurological disorder of conscious
+awareness that results from damage in the bilateral parietal cortex.
+Patients are plagued by an
 inability to bind various individual features of the visual
@@
 leveraging these environmental opportunities (i.e., exploitation).
-The human PMC was congruently coupled in two functional connectivity modalities
+Congruently,
+the human PMC was coupled in across two functional connectivity modalities
 with the amygdala
@@
 (rather than centromedial or superficial) nuclei group
-of the amygdala (Bzdok et al., 2013b),
-which is an amygdalar subregion probably devoted to
-continuously scanning environmental input
+of the amygdala (Bzdok et al., 2013b).
+This amygdalar subregion has been proposed to
+continuously scan environmental input
 for biological significance assessment
-(Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011; Ghods-Sharifi et al., 2009; LeDoux, 2007).
+(Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011;
+Ghods-Sharifi et al., 2009; LeDoux, 2007).
 %
@@
 subjectively perceived relevance of a chosen object
-than by its factual value, based on an internal currency of value
+than by its factual value, based on an internal currency of value.
 In fact, direct stimulation of PMC neurons
@@
 Graded changes in firing rates of PMC neurons
-that indicated changes in upcoming trials were
-distinct from neuronal firing indicating switching between choices.
+indicated changing choices in upcoming trials and their neural patterns were
+distinct from spike firings that indicated choosing either option.
 Also in humans, 
@@
 %
-The retrosplenial portion of the PMC can subserve evaluation of reward
-contingencies and action possibilities by integrating these with
+The retrosplenial portion of the PMC can subserve action possibilities
+and evaluation of reward contingencies by integrating these with
 information from memory and altered perspective frames.
@@
 Regarding perspective frames, this PMC subregion has been
-proposed to mediates between the organism's egocentric
+proposed to mediate between the organism's egocentric
 (i.e., focused on sensory input) and
@@
 Perceived value that differs across individuals updates
-statistical knowledge of the environment
+statistical assessment of the environment
 to predict delayed reward opportunities in the future.
@@
 in both the external environment and internal milieu
-that modulate strategic behavioral adjustment in volatile environments
-to map possiblity spaces.
+that modulate strategic behavioral adjustment in volatile environments.
 
@@
 Patients with neurological lesions in the prefrontal cortex
-are now to adapt to novel situations and stimuli with difficulty
+are known to adapt to novel situations and stimuli with difficulty
 (Stuss, D. T. \& Benson, D. F. The Frontal Lobes, Raven, New York, 1986).
@@
 generation of probabilistic internal information drawing from
-episodic memory retrieval, non-mnemonic scene construction processes,
+episodic memory retrieval, generative construction processes,
 and explicit knowledge of the external world.
@@
 The vmPFC is preferentially connected with limbic and reward-related areas.
-The vmPFC has been observed to have monosynap- tical connections with the ventral striatum (VS, which anatom- ically includes the NAc) in axonal tracing studies in monkeys (Haber et al., 1995; Ferry et al., 2000).
-The NAc is thought to be linked to reward mechanisms that may not only modulate motivated behavior towards basic survival needs, but also
-subserve model-free reinforcement learning in humans more broadly
+The vmPFC has been observed to have monosynaptical connections
+with the nucleus accumbens
+in axonal tracing studies in monkeys (Haber et al., 1995; Ferry et al., 2000).
+The NAc is thought to be linked to reward mechanisms that not
+only modulate motivated behavior towards basic survival needs but also
+subserve reinforcement learning in humans more broadly
 (O'Doherty 2015).
-This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys (Croxson et al., 2005) that quantified the VS to be substantially more likely connected to the vmPFC than dmPFC in both species.
+This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys
+(Croxson et al., 2005) that
+quantified this brain regions to
+be substantially more likely connected to the vmPFC than dmPFC in both species.
 Two functional connectivity modalities in humans strongly connected
-the vmPFC with the nucleus accumbens (NAc), hippocampus (HC),
-posterior cingulate cortex (PCC), and retrosplenial cortex (RSC).
+the vmPFC with the nucleus accumbens, hippocampus (HC),
+and posteriomedial cortex.
 %
@@
 and decision making (Damasio, 1996; Nauta, 1971)
-In studies of decision making and neuroeconomics, ventromedial prefrontal cortex (vmPFC) con- sistently reflects an individual’s subjective valuation (Behrens et al., 2008; Boorman et al., 2009; Knutson et al., 2005; Plass- mann et al., 2007)
+In neuroeconomic studies of human decision making,
+the vmPFC consistently reflects an individual’s subjective
+valuation (Behrens et al., 2008; Boorman et al., 2009; Knutson et al., 2005; Plassmann et al., 2007)
 This may be why performance within and across participants
@@
 which was shown to be critical for behavior.
-The GMV of the vmPFC and VS correlated with indices of social reward attitudes and value-guided behavior (Lebreton et al., 2009).
+The GMV of the vmPFC and nucleus accumbens
+correlated with indices of social reward attitudes and
+value-guided behavior (Lebreton et al., 2009).
 Additionally,
 independent whole-brain analyses from structural
-neuroimaging studies related the gray-matter volume (GMV) of the vmPFC
-(more consistently than any other
-brain region) to indices of
+neuroimaging studies related the gray-matter volume GMV of the vmPFC,
+more consistently than any other
+brain region, to indices of
 social competence and social network complexity,
@@
 Its highly recursive anatomical architecture
-may be specifically designed to allowing reconstruction
-whole episode of experience from memory fragments.
+may be specifically designed to allow reconstructing
+entire episodes of experience from memory fragments.
 %
@@
 there is now increasing evidence for a role
-in constructing mental models in general (Schachter/Buckner).
+in constructing mental models in general (Schachter/Buckner;
+Maguire 2016).
 Indeed,
-hippocampal damage [21,38–40] is
+hippocampal damage is
 not only associated with an impairment in reexperiencing the past (i.e., amnesia),
@@
 Moreover, spike sequences measured in hippocampal place cells of rats
-featured reoccurred directly after experimental trials
+featured reoccurrence directly after experimental trials
 as well as directly before upcoming experimental trials (Diba + Buzsaki 2007).
@@
 have been proposed to be critical in communicating local information
-to the neocortex for long-term storage, which includes the DMN.
+to the neocortex for long-term storage, potentially also in the DMN.
 These mechanisms of the hippocampus probably make important contributions to the
@@
 Importantly, the RTPJ has been shown to be intimately related to
-supramodal prediction error signaling.
+supramodal prediction and error signaling.
 Neural activity in the RTPJ has been argued to be responsible 

### 2016-09-29 paper/paper.tex
blah
RL jewel

--- 2016-09-26 paper/paper.tex
@@
 autonomous learning systems
-\textbf{\\Get opinion from}: Danielle Bassett, Kai Vogeley, Mohammad Shakir,
-Guillaume Dumas, Smallwood
+\textbf{\\Get opinion from}: Danielle Bassett, Kai Vogeley, Mohammad Shakir
 
@@
 %% \subsection{Minimal control energy controller}
-\subsubsection{Related works}
-The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
 
 \subsection{Demystifying the free-energy principle and active-inference}
-The free-energy principle in it's present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net.
-
-In this section we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
+The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
+
+In this section, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
 This ideas were first developed by Hinton et al.
@@
 \end{equation}
-is the \textit{Helmholtz free-energy} of the external world, modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of macrostates $\psi$.
+is the \textit{Helmholtz free-energy} (at unit temperature!). %% , modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of macrostates $\psi$
 Thus generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
 
-The goal of the brain is then to optimize over the generative model $G$ (to iteratively or analytically modify the generative density $p_G(.|s)$, so as to minimize their surprise.
+The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|s)$, so as to minimize surprise.
 It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(s)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
-A workaround is then to introduce a second density $p_R(.|\zeta)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\zeta)$: the former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
+A workaround is then to introduce a second density $p_R(.|\zeta)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\zeta)$, as a trick for doing approximate inference.
+The former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
 This primal-dual idea, first proposed in Hinton et al.
@@
 \subsubsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
-In this subsection we'll develop a very insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
-As an avant-gout of what is to come shorty, let's just that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
-
-Now, invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
+In this subsection, we'll present an insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
+As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
+
+Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
 distribution
@@
   \begin{split}
-    \text{generative surprise } &:= -\log(p_G(\zeta)) = \underbrace{F_G(\zeta)}_{\text{generative free-energy}} \\
-    &=\underbrace{F^R_G(\zeta)}_{\text{variational free-energy}} - \underbrace{D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))}_{\text{complexity}} \\
+    \text{generative surprise } &:= -\log(p_G(\zeta)) = F_G(\zeta) \\
+    &=\underbrace{F^R_G(\zeta)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))}_{\text{complexity}} \\
     &\le F^R_G(\zeta),
-    \text{ with equalitity if }p_R(.|\zeta) = p_G(.|\zeta),
+    \text{ with equality if }p_R(.|\zeta) = p_G(.|\zeta),
     \end{split}
@@
     As noted by Dayan et al.
-(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
+(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
   \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
+  \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
+    technical difficulty in getting derivatives of loss function w.r.t
+    recognition weights $\W^R$).
+  \item This difficulty was removed in the 2010s by Kingma et al., an other groups, via a ``reparametrization trick''.
     \end{itemize}
 
-\subsection{Free-energy minimization via bayesian auto-encoders}
+\subsection{Free-energy minimization via variational auto-encoders}
 Following Kingma et al. 2014 \cite{kingma2013auto}, define the data-dependent auxiliary random function
@@
 \begin{mdframed}
-  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\zeta)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\textbf{W}^R$ ?
+  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\zeta)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
 \end{mdframed}
 \paragraph{Solution: The reparametrization trick.}
+\begin{itemize}
+\item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
+\item Set $\phi = g(\W^{R}, \zeta, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
+  \begin{itemize}
+  \item results in a sample $\phi \sim p_R(.|\zeta)$, from the correct
+    posterior
+    \end{itemize}
+\end{itemize}
+
+\begin{table}[H]
+  \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
+         \hline
+         Posterior & $p_R(.|\zeta)$ & noise & $g(\W^R,\zeta,\epsilon)$ & Also \\ \hline
+         Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
+         Student’s t, Logistic, Uniform, Triangular, ... \\ \hline
+         Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
+\\ \hline
+         Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
+  \end{tabular}
+  \caption{Reparametrization trick for a variety of models. Kingma et al. 2014 \cite{kingma2013auto}.}
+\end{table}
+
+The result is a scheme for training DBNs via good-old backprop!
 
@@
 
-\section{Relation to Other Cognitive Models (Danilo)}
+\section{Reinforcement-Learning (RL) as a computational model for DMN function}
+\subsection{General RL theory}
+In a general reinforcement-learning (RL) framework, an \textit{agent} interacts with an unknown
+environment by taking \textit{actions / decisions} $a_t$ which trigger changes in the
+\textit{state} of the environment
+$s_t \rightarrow s_{t+1}$, accompanied with a real-valued \textit{reward} $r_t$ (or regret!,
+since it can be
+negative). This should be thought of as an instantaneous satisfaction accompanying the execution of
+an action. Here the subscript $t$ denotes time, and can be discrete or continuous.
+%% The unknown environment $\mathcal E$ is generally taken as stochastic, but having some kind of
+%% memory.
+A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
+This is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
+\begin{itemize}
+\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
+\item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
+  \text{laugh}, \text{empathize}, \text{empathize}\}.$
+\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$, so that $r(s, a)$ is the instant reward for taken action $a$ at state $s$.
+\item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
+  $\mu(s)$ is the probability that the environment starts-off in state $s$.
+  \item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1]$, so that
+    $\rho(s,a,s') = \mathbb P(s'|s, a)$ is the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such transitions be Markovian.
+\end{itemize}
+
+\begin{remark}
+  Given a system $(\mathcal S, \mathcal A, r, \rho, \mu)$ satisfying all the axioms for an MDP
+  except the Markov property, we can always transform it into an MDP by considering a compounded
+  version of the states
+  $$s_t \leftarrow (s_0,a_0,r_0,s_1,a_1,s_2\ldots,s_t,a_{t-1},r_{t-1},s_{t})$$ made of the systems
+  trajectory up to an including the last time window.
+\end{remark}
+
+The behavior of the agent is governed by some kind of policy, which maps states to sampling probabilities over candidate actions to perform from that state.
+The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
+
+Since an action taken now might have repacautions long into the feature, it turns out that the
+quantity to optimize is not the instantaneous rewards $r(s_t, a_t)$, a cumulative version which
+takes into account the future. A standard for modeling this accumulation is the so-called
+time-discounted \textit{cumulative reward} function
+\begin{equation}
+  \label{eq:cumr}
+  R(s_t, a_t) = r(s_t, s_t) + \gamma r(s_{t+1}, s_{t+1}) +\ldots + \gamma^Tr(s_{t+T}, s_{t+T})
+  = r(s_t, s_t) + \gamma \sum_{\tau=1}^{T}\gamma^{\tau-1}r(s_{t+\tau},a_{t+\tau}),
+  \end{equation}
+for some \textit{discount factor} $\gamma \in [0, 1]$, and horizon $T$.
+The choice $\gamma = 0$ (or $T=0$)  corresponds to a perfectly myopic agent who is solely concerned about their immediate rewards and have no regards for the future.
+For there to be any learning at all, a technical requirement is that $\gamma < 1$ \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
+Thus the agent is positively biased towards considering rewards which are will
+come soon. Note that $R(s_t, a_t)$ defined above is a stochastic quantity which depends on the
+trajectory of a policy $\pi$,
+\begin{eqnarray*}
+  \begin{split}
+    s_0 &\sim \mu\\
+    a_0 &\sim \pi(.|s_0)\\
+    s_1 &\sim \rho(.|s_0) \text{ and collect rewards }r(s_0, a_0)\\
+    a_1 &\sim \pi(.|s_1)\\
+    s_2 &\sim \rho(.|s_1) \text{ and collect rewards }r(s_1, a_1)\\
+    \ldots\\
+    a_t &\sim \pi(.|s_t)\\
+    s_{t+1} &\sim \rho(.|s_t) \text{ and collect rewards }r(s_t, a_t)\\
+    \ldots
+  \end{split}
+\end{eqnarray*}
+
+The goal is then to sample from a policy $a_t \sim \pi$ so that on average, $R(s_t, a_t)$ is maximized. Formally, the agent's goal is to maximize the $Q$-value at $(s_t, a_t)$, defined by
+\begin{equation}
+  \label{eq:qval}
+  Q^{\pi}(s_t,a_t) = \mathbb E [R(s_t, a_t)]
+\end{equation}
+
+where the expectation is over all possible trajectories from $(s_t, a_t)$, under the policy $\pi$.
+
+By construction, in general, maximizing the expectation in  \eqref{eq:qval} is intractable due to
+the enormous number of possible trajectories to consider.
+One popular simplification that removes this issue is to only consider the sub-family of \textit{deterministic policies}, which map each state to the best single best action which should be taken at that state, and therefore define \textbf{functions from states to actions}.
+
+Q-learning (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
+\textit{greedy policy}
+\begin{equation}
+  \pi(s) = \pi_{\tilde{Q}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a),
+\end{equation}
+where $\tilde{Q}(s, a)$ is the current estimate of the state-action value table. In the sequel, except otherwise stated, the policies discussed will be of this form.
+The hallmark of this trick wherein one only considers deterministic policies is the so-called
+\textit{Bellman's equation}
+\begin{equation}
+  Q^\pi(s_t, a_t) = \mathbb E_{s_{t+1} \sim \rho(.|s_t)} [r(s_t,a_t) + \gamma Q^\pi(s_{t+1}, \pi(s_{t+1}))]
+  \label{eq:bellman}
+\end{equation}
+which is a fixed-point equation for the $Q$-value operator of a deterministic policy $\pi$.
+Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which
+depend on the environment, and so can be learned off-policy by observing state-transitions
+triggered by another behavioral policy, which can be stochastic.
+
+\section{Q-learning with value approximation}
+At state $s \in \mathcal S$, the DMN agent takes an action $a \in \mathcal A$ (for example,
+$a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.) by sampling
+from its current policy matrix, and collects a reward $r$ (measured on their internal currency of
+worldly values ?), and the environment transitions to a new $s' \in \mathcal S$. At the end of
+this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced: this represents an exemplar
+behavior of the agent in such a situation, and is recorded in replay memory:
+$\mathcal D \leftarrow \mathcal D \cup \{e\}$.
+Now, replay consists in sampling (uniform or importance-weighted ?)  mini-batches of experiences
+$e = (s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$ and trying to approximate
+the would-be $Q$-value at $(s, a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
+$r + \gamma \max_{a'} Q(s', a'|\theta^Q)$, with the output of a parametrized regression model $(s,a,r,s') \mapsto {Q}(s, a|\theta^Q)$, where $Q(.|\theta^Q)$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q$.
+Computing an optimal policy then corresponds to finding the parameters $\theta^Q$ which minimize the following loss function
+\begin{equation}
+  \mathcal L(\theta^Q)
+  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q))^2]
+  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - y)^2],
+  \label{eq:oracle}
+  \end{equation}
+where
+$ y = r + \gamma \max_{a'} Q(s', a'|\theta^Q) = r + \gamma Q(s', \pi(s')|\theta^Q)$ is the prediction target.
+% While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
+
+For example, a general regression model with a kernel $(s, a) \mapsto \phi(s, a)$ would be of the
+form
+
+$${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
+$\phi$ would represent a high-level hand-crafted representation of the state-action pairs into an appropriate feature space.
+Such are the techniques that have been proposed by [sutton et al.
+xyzw, Song et al.
+2016 (NIPS), and Parr et al.
+2003 (JMLR)].
+A breakthrough alternative proposed by Google DeepMind [Minh et al.
+2014, Silver et al 2015] is to learn this representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which are by far the current state-of-the-art in RL.
+
+The entire model can then be effectively trained via \textit{backprop} wherein online stochastic
+gradient descent is used to minimize the variational objective \eqref{eq:oracle}.
+
+\begin{remark}
+Note that most RL algorithms, including \textit{Temporal Difference (TD)} learning [Sutton et al.,
+  Bertsekas] can be cast in this general variational framework.
+\end{remark}
+
+\subsection{A Deep Q-learning for DMN function}
+\textbf{XXX Put the diagrams and process charts here.}
+
+\paragraph*{Some notes:}
+\begin{itemize}
+\item The brain does backprop (see Hinton's recent talk at Stanford)
+\item ...
+\end{itemize}
+
+\section{Relation to other cognitive models (Danilo)}
 \paragraph{Predictive Coding and Bayesian Brain Hypothesis}
@@
 
-\section{Relation to Other Statistical Models (Elvis)}
-\paragraph{Control Theory.}
-abc
+\section{Relation to other statistical models (Elvis)}
+\paragraph{Control theory.}
+\begin{itemize}
+  \item The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
+\end{itemize}
 
@@
 
-The DMN can be viewed as a smoothing kernel extending
-from the relatively recent past to the relatively close future
-that constantly estimates complex non-linear gain functions.
--> This is why neuroimaging research revealed activity increases in the DMN
-when 
--> This may be why perfect memory has been evolutionarily de-selected,
-although it is provably feasible by the neurobiological hardware in humans
-
 \paragraph{Acknowledgment}

### 2016-09-29 paper/paper.tex

--- 2016-09-27 paper/paper.tex
@@
 Further,
-episodic memory recall and imagination strongly draws on
+episodic memory recall and imagination strongly draw on
 complex world knowledge.

### 2016-09-29 paper/paper.tex
more

@@
 
-\section{Reinforcement-Learning (RL) as a computational model for DMN function}
-\subsection{General RL theory}
+\section{Reinforcement-Learning as a computational model for DMN function}
+\subsection{General Reinforcment-Learning (RL) theory}
 In a general reinforcement-learning (RL) framework, an \textit{agent} interacts with an unknown

### 2016-09-29 paper/paper.tex
more

@@
 
-Since an action taken now might have repacautions long into the feature, it turns out that the
+Since an action taken now might have repercussion long into the feature, it turns out that the
 quantity to optimize is not the instantaneous rewards $r(s_t, a_t)$, a cumulative version which

### 2016-09-29 paper/paper.tex
more

@@
 
-Q-learning (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
+\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
 \textit{greedy policy}
@@
   \pi(s) = \pi_{\tilde{Q}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a),
-\end{equation}
-where $\tilde{Q}(s, a)$ is the current estimate of the state-action value table. In the sequel, except otherwise stated, the policies discussed will be of this form.
+  \label{eq:qlearning}
+\end{equation}
+where $\tilde{Q}(s, a)$ is the current estimate of the state-action value table.
 The hallmark of this trick wherein one only considers deterministic policies is the so-called
@@
 
-\section{Q-learning with value approximation}
-At state $s \in \mathcal S$, the DMN agent takes an action $a \in \mathcal A$ (for example,
+\subsection{Q-learning with value approximation}
+As already mentioned in the previous section, Q-learning optimizes over the class of
+deterministic policies of the form \eqref{eq:qlearning}.
+At state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$ (for example,
 $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.) by sampling
-from its current policy matrix, and collects a reward $r$ (measured on their internal currency of
-worldly values ?), and the environment transitions to a new $s' \in \mathcal S$. At the end of
+from its current policy matrix, and collects a reward $r$,
+% (measured on their internal currency of worldly values ?),
+and the environment transitions to a new $s' \in \mathcal S$. At the end of
 this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced: this represents an exemplar
@@
 Now, replay consists in sampling (uniform or importance-weighted ?)  mini-batches of experiences
-$e = (s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$ and trying to approximate
+$e = (s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ and trying to
+approximate
 the would-be $Q$-value at $(s, a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
@@
 2014, Silver et al 2015] is to learn this representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which are by far the current state-of-the-art in RL.
+In such a case, $\theta^Q$ would correspond to ..
 

### 2016-09-29 paper/paper.tex
more stuff

@@
   \text{laugh}, \text{empathize}, \text{empathize}\}.$
-\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$, so that $r(s, a)$ is the instant reward for taken action $a$ at state $s$.
+\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$, so that $r(s, a)$ is the instant reward for taking action $a$ at state $s$.
 \item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
@@
 \end{equation}
-which is a fixed-point equation for the $Q$-value operator of a deterministic policy $\pi$.
+which is a fixed-point equation for the deterministic policy $\pi$.
 Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which

### 2016-09-29 paper/paper.tex
more stuff

@@
 
-Since an action taken now might have repercussion long into the feature, it turns out that the
+Since an action taken now might have repercussions long into the feature, it turns out that the
 quantity to optimize is not the instantaneous rewards $r(s_t, a_t)$, a cumulative version which

### 2016-09-29 paper/paper.tex
some comments for later

@@
   version of the states
-  $$s_t \leftarrow (s_0,a_0,r_0,s_1,a_1,s_2\ldots,s_t,a_{t-1},r_{t-1},s_{t})$$ made of the systems
-  trajectory up to an including the last time window.
+  $$s_t \leftarrow (s_0,a_0,r_0,s_1,a_1,s_2\ldots,s_t,a_{t-1},r_{t-1},s_{t})$$ made of the system's
+  history up to an including time $t$.
 \end{remark}
 
-The behavior of the agent is governed by some kind of policy, which maps states to sampling probabilities over candidate actions to perform from that state.
-The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
+The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
+sampling probabilities over candidate actions to perform from that state. The  goal of RL is then
+adapt this policy so as to  maximize the cumulative rewards.
 
@@
     a_0 &\sim \pi(.|s_0)\\
-    s_1 &\sim \rho(.|s_0) \text{ and collect rewards }r(s_0, a_0)\\
+    s_1 &\sim \rho(.|s_0,a_0) \text{ and collect rewards }r(s_0, a_0)\\
     a_1 &\sim \pi(.|s_1)\\
-    s_2 &\sim \rho(.|s_1) \text{ and collect rewards }r(s_1, a_1)\\
+    s_2 &\sim \rho(.|s_1,a_1) \text{ and collect rewards }r(s_1, a_1)\\
     \ldots\\
     a_t &\sim \pi(.|s_t)\\
-    s_{t+1} &\sim \rho(.|s_t) \text{ and collect rewards }r(s_t, a_t)\\
+    s_{t+1} &\sim \rho(.|s_t,a_t) \text{ and collect rewards }r(s_t, a_t)\\
     \ldots
@@
 \item The brain does backprop (see Hinton's recent talk at Stanford)
-\item ...
+\item model weights $\theta^Q$ should correspond to connections between neurons
+\item humm, looks like such a knowledge would simply translate into constraints on the DqN weights $\theta^Q$ (... at least, part of such knowledge)
+  ok, gotta think about the model-free / model-based reconsiliation
 \end{itemize}

### 2016-09-29 paper/paper.tex
a pass on RL section

@@
 \subsection{General Reinforcment-Learning (RL) theory}
-In a general reinforcement-learning (RL) framework, an \textit{agent} interacts with an unknown
-environment by taking \textit{actions / decisions} $a_t$ which trigger changes in the
+In a general reinforcement-learning (RL) framework, an \textit{agent}
+interacts with an unknown environment in trial-and-error fashion by taking
+\textit{actions / decisions} $a_t$ which trigger changes in the
 \textit{state} of the environment
-$s_t \rightarrow s_{t+1}$, accompanied with a real-valued \textit{reward} $r_t$ (or regret!,
-since it can be
-negative). This should be thought of as an instantaneous satisfaction accompanying the execution of
-an action. Here the subscript $t$ denotes time, and can be discrete or continuous.
+$s_t \rightarrow s_{t+1}$, accompanied with a real-valued \textit{reward}
+$r_t$ (or regret!, since it can be negative). This reward should be thought
+of as an instantaneous satisfaction accompanying the execution of
+an action. Here the subscript $t$ denotes time, and can be discrete or
+continuous.
 %% The unknown environment $\mathcal E$ is generally taken as stochastic, but having some kind of
@@
   \label{eq:cumr}
-  R(s_t, a_t) = r(s_t, s_t) + \gamma r(s_{t+1}, s_{t+1}) +\ldots + \gamma^Tr(s_{t+T}, s_{t+T})
+  R(s_t, a_t) = r(s_t, s_t) + \ldots + \gamma^Tr(s_{t+T}, s_{t+T})
   = r(s_t, s_t) + \gamma \sum_{\tau=1}^{T}\gamma^{\tau-1}r(s_{t+\tau},a_{t+\tau}),
@@
 Thus the agent is positively biased towards considering rewards which are will
-come soon. Note that $R(s_t, a_t)$ defined above is a stochastic quantity which depends on the
-trajectory of a policy $\pi$,
+come soon. Note that $R(s_t, a_t)$ defined above is a stochastic quantity which depends on the trajectory generated by the policy $\pi$, upto but not including time $t$
 \begin{eqnarray*}
@@
     \ldots\\
-    a_t &\sim \pi(.|s_t)\\
-    s_{t+1} &\sim \rho(.|s_t,a_t) \text{ and collect rewards }r(s_t, a_t)\\
-    \ldots
+    a_{t-1} &\sim \pi(.|s_{t-1})\\
+    s_{t} &\sim \rho(.|s_{t-1},a_{t-1}) \text{ and collect rewards }r(s_{t-1}, a_{t-1})
   \end{split}
@@
 
-The goal is then to sample from a policy $a_t \sim \pi$ so that on average, $R(s_t, a_t)$ is maximized. Formally, the agent's goal is to maximize the $Q$-value at $(s_t, a_t)$, defined by
+%% The goal is then to sample from a policy $a_t \sim \pi$ so that the expected
+%% value of $R(s_t, a_t)$ is maximized.
+Formally, the agent's goal is then to maximize the state-action value for
+taking action $a_t$ and state $s_t$ defined by
 \begin{equation}
@@
 
-where the expectation is over all possible trajectories from $(s_t, a_t)$, under the policy $\pi$.
+where the expectation is over all possible trajectories successor
+trajectories of state $s_t$.
 
@@
 the enormous number of possible trajectories to consider.
-One popular simplification that removes this issue is to only consider the sub-family of \textit{deterministic policies}, which map each state to the best single best action which should be taken at that state, and therefore define \textbf{functions from states to actions}.
+One popular simplification that removes this issue is to only consider the sub-family of \textit{deterministic policies}, which map each state to the best single best action which should be taken at that state, and therefore define functions from states to actions.
 
@@
 The hallmark of this trick wherein one only considers deterministic policies is the so-called
-\textit{Bellman's equation}
+\textit{Bellman's equation} [Sutton and Barto 1998]
 \begin{equation}
@@
 \begin{itemize}
+  \item Actions are continuous
   \item The brain does backprop (see Hinton's recent talk at Stanford)

### 2016-09-29 paper/paper.tex
more stuff

@@
 \section{Reinforcement-Learning as a computational model for DMN function}
+The main idea is to show that the DMN has biological substrates (gradient
+computation, backprop, experience replay, etc.) for all the
+major components of an efficient RL algorithm, and all these components
+integrate in a plausible way.
 \subsection{General Reinforcment-Learning (RL) theory}
@@
 
-\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
+\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy \textit{critic-only} algorithm, uses the
 \textit{greedy policy}
@@
 
-\subsection{Q-learning with value approximation}
+\subsection{Q-learning with value approximation and replay}
 As already mentioned in the previous section, Q-learning optimizes over the class of
 deterministic policies of the form \eqref{eq:qlearning}.
-At state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$ (for example,
-$a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.) by sampling
-from its current policy matrix, and collects a reward $r$,
+At state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
+% (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
+by sampling from its current policy matrix, and collects a reward $r$,
 % (measured on their internal currency of worldly values ?),
 and the environment transitions to a new $s' \in \mathcal S$. At the end of
-this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced: this represents an exemplar
-behavior of the agent in such a situation, and is recorded in replay memory:
+this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar behavior of the agent and is recorded in replay memory:
 $\mathcal D \leftarrow \mathcal D \cup \{e\}$.
 Now, replay consists in sampling (uniform or importance-weighted ?)  mini-batches of experiences
-$e = (s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ and trying to
+$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ and trying to
 approximate
@@
 where
-$ y = r + \gamma \max_{a'} Q(s', a'|\theta^Q) = r + \gamma Q(s', \pi(s')|\theta^Q)$ is the prediction target.
+$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q) = r + \gamma Q(s', \pi(s')|\theta^Q)$ is the prediction target.
 % While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
@@
 Note that most RL algorithms, including \textit{Temporal Difference (TD)} learning [Sutton et al.,
-  Bertsekas] can be cast in this general variational framework.
+  Bertsekas] and SARSA can be cast in this general variational framework.
 \end{remark}
@@
 \begin{itemize}
-  \item Actions are continuous
+  \item Actions are continuous, whilst states are continuous.
   \item The brain does backprop (see Hinton's recent talk at Stanford)

### 2016-09-29 paper/paper.tex
grammar

@@
   \text{laugh}, \text{empathize}, \text{empathize}\}.$
-\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$, so that $r(s, a)$ is the instant reward for taking action $a$ at state $s$.
+\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$, so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
 \item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
@@
 The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
-sampling probabilities over candidate actions to perform from that state. The  goal of RL is then
+sampling probabilities over candidate actions to perform from thin state. The  goal of RL is then
 adapt this policy so as to  maximize the cumulative rewards.
@@
 the enormous number of possible trajectories to consider.
-One popular simplification that removes this issue is to only consider the sub-family of \textit{deterministic policies}, which map each state to the best single best action which should be taken at that state, and therefore define functions from states to actions.
+One popular simplification that removes this issue is to only consider the sub-family of \textit{deterministic policies}, which map each state to the best single best action which should be taken at thin state, and therefore define functions from states to actions.
 
@@
 deterministic policies of the form \eqref{eq:qlearning}.
-At state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
+in state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
 % (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)

### 2016-09-29 paper/paper.tex
glitches

@@
 \item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
-  \text{laugh}, \text{empathize}, \text{empathize}\}.$
+  \text{laugh}, \text{sympathize}, \text{empathize}\}.$
 \item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$, so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
@@
 approximate
-the would-be $Q$-value at $(s, a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
-$r + \gamma \max_{a'} Q(s', a'|\theta^Q)$, with the output of a parametrized regression model $(s,a,r,s') \mapsto {Q}(s, a|\theta^Q)$, where $Q(.|\theta^Q)$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q$.
+the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
+$r + \gamma \max_{a'} Q(s', a'|\theta^Q)$, with the output of a parametrized regression model $(s,a,r,s') \mapsto {Q}(s, a|\theta^Q)$, where $Q(.,.|\theta^Q)$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q$.
 Computing an optimal policy then corresponds to finding the parameters $\theta^Q$ which minimize the following loss function

### 2016-09-29 paper/paper.tex
a note on model-based RL + more

@@
 integrate in a plausible way.
-\subsection{General Reinforcment-Learning (RL) theory}
+\subsection{Markov Decision Processes}
 In a general reinforcement-learning (RL) framework, an \textit{agent}
-interacts with an unknown environment in trial-and-error fashion by taking
-\textit{actions / decisions} $a_t$ which trigger changes in the
+interacts with an  environment in a trial-by-error fashion by taking
+actions / decisions $a$, which trigger changes in the
 \textit{state} of the environment
-$s_t \rightarrow s_{t+1}$, accompanied with a real-valued \textit{reward}
-$r_t$ (or regret!, since it can be negative). This reward should be thought
+$s \rightarrow s'$, accompanied with a real-valued \textit{reward}
+$r = r(s, a)$ (or regret!, since it can be negative) collected by the
+agent. This reward should be thought
 of as an instantaneous satisfaction accompanying the execution of
@@
 continuous.
-%% The unknown environment $\mathcal E$ is generally taken as stochastic, but having some kind of
+%% The environment $\mathcal E$ is generally taken as stochastic, but having some kind of
 %% memory.
@@
   $\mu(s)$ is the probability that the environment starts-off in state $s$.
-  \item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1]$, so that
-    $\rho(s,a,s') = \mathbb P(s'|s, a)$ is the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such transitions be Markovian.
+  \item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto \rho(s'|s,a)$, the probability of moving to state $s'$ if action $a$ is taken from state $s$.
+In addition, one requires that such transitions be Markovian.
 \end{itemize}
@@
 
+\paragraph{A note on model-based RL.}
+Knowledge the agent has about the environment can be encoded in the
+functions $r$, $\mu$, $\rho$, and / or $\pi$, as is done in
+\textit{model-based} RL. For example if a rat in a maze knows, for sure,
+that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food, they can assume:
+\begin{itemize}
+\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
+\item $\rho(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
+\item etc.
+\end{itemize}
+
+\subsection{Value functions and policies}
 The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
@@
 
-\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy \textit{critic-only} algorithm, uses the
+\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
 \textit{greedy policy}
 \begin{equation}
-  \pi(s) = \pi_{\tilde{Q}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a),
+  \pi(s) = \argmax_{a \in \mathcal A}{Q}(s, a).
   \label{eq:qlearning}
 \end{equation}
-where $\tilde{Q}(s, a)$ is the current estimate of the state-action value table.
-The hallmark of this trick wherein one only considers deterministic policies is the so-called
-\textit{Bellman's equation} [Sutton and Barto 1998]
+% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
+% table.
+The celebrated \textit{Bellman's equation} [Sutton and Barto 1998] then takes the simple form
 \begin{equation}
@@
 
-\subsection{Q-learning with value approximation and replay}
+\subsection{Efficient RL with value approximation and replay}
 As already mentioned in the previous section, Q-learning optimizes over the class of
@@
 and the environment transitions to a new $s' \in \mathcal S$. At the end of
-this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar behavior of the agent and is recorded in replay memory:
+this step, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar behavior of the agent and is recorded in replay memory:
 $\mathcal D \leftarrow \mathcal D \cup \{e\}$.
-Now, replay consists in sampling (uniform or importance-weighted ?)  mini-batches of experiences
+Now, replay consists in sampling (uniform or importance-weighted\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
 $(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ and trying to

### 2016-09-30 paper/paper.tex
link to TD learning

@@
 the enormous number of possible trajectories to consider.
-One popular simplification that removes this issue is to only consider the sub-family of \textit{deterministic policies}, which map each state to the best single best action which should be taken at thin state, and therefore define functions from states to actions.
+One popular simplification that removes this issue is to only consider the sub-family of
+\textit{deterministic policies}, which map each state to the best single best action which
+should be taken at thin state, and therefore define functions from states to actions.
 
@@
 \begin{equation}
-  Q^\pi(s_t, a_t) = \mathbb E_{s_{t+1} \sim \rho(.|s_t)} [r(s_t,a_t) + \gamma Q^\pi(s_{t+1}, \pi(s_{t+1}))]
+  Q^\pi(s_t, a_t) = \mathbb E_{s_{t+1} \sim \rho(.|s_t,a_t)} [r(s_t,a_t) + \gamma Q^\pi(s_{t+1}, \pi(s_{t+1}))]
   \label{eq:bellman}
@@
 \subsection{Efficient RL with value approximation and replay}
-As already mentioned in the previous section, Q-learning optimizes over the class of
-deterministic policies of the form \eqref{eq:qlearning}.
-in state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
+%As already mentioned in the previous section, Q-learning optimizes over the class of
+% deterministic policies of the form \eqref{eq:qlearning}.
+We need ways to approximate the state-action value function
+which scale up to large state spaces and / or action spaces.
+In state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
 % (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
@@
 % (measured on their internal currency of worldly values ?),
-and the environment transitions to a new $s' \in \mathcal S$. At the end of
-this step, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar behavior of the agent and is recorded in replay memory:
-$\mathcal D \leftarrow \mathcal D \cup \{e\}$.
-Now, replay consists in sampling (uniform or importance-weighted\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
+and the environment transitions to a new state $s' \in \mathcal S$. At the end of
+this step, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar
+behavior of the agent and is recorded in replay memory buffer:
+$\mathcal D \leftarrow \mathcal D \cup \{e\}$ (possibly discarding the oldest entries to make space).
+Now, at iteration $k+1$, replay consists in sampling (uniform or importance-weighted
+\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
 $(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ and trying to
@@
 the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
-$r + \gamma \max_{a'} Q(s', a'|\theta^Q)$, with the output of a parametrized regression model $(s,a,r,s') \mapsto {Q}(s, a|\theta^Q)$, where $Q(.,.|\theta^Q)$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q$.
-Computing an optimal policy then corresponds to finding the parameters $\theta^Q$ which minimize the following loss function
-\begin{equation}
-  \mathcal L(\theta^Q)
-  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q))^2]
-  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - y)^2],
+$r + \gamma \max_{a'} Q(s', a'|\theta^Q_k)$, with the output of a parametrized regression model $(s,a,r,s')
+\mapsto {Q}(s, a|\theta^Q_{k})$, where $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q_{k+1}$.
+Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function
+\begin{equation}
+  \mathcal L(\theta^Q_{k+1})
+  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
+  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
   \label{eq:oracle}
@@
 where
-$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q) = r + \gamma Q(s', \pi(s')|\theta^Q)$ is the prediction target.
+$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
 % While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
-
-For example, a general regression model with a kernel $(s, a) \mapsto \phi(s, a)$ would be of the
+For example, a general linear model with a kernel $\phi$ would be of the
 form
-
 $${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
-$\phi$ would represent a high-level hand-crafted representation of the state-action pairs into an appropriate feature space.
-Such are the techniques that have been proposed by [sutton et al.
-xyzw, Song et al.
-2016 (NIPS), and Parr et al.
-2003 (JMLR)].
-A breakthrough alternative proposed by Google DeepMind [Minh et al.
-2014, Silver et al 2015] is to learn this representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which are by far the current state-of-the-art in RL.
-In such a case, $\theta^Q$ would correspond to ..
+$\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action $(s,a)$ into an appropriate
+feature space. Such are the techniques that have been proposed by [sutton et al. xyzw, Song et al. 2016 (NIPS),
+  and Parr et al. 2003 (JMLR)].
+A breakthrough alternative proposed by Google DeepMind [Minh et al. 2014, Silver et al 2015] is to learn this
+representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which
+are by far the current state-of-the-art in RL. In such a case, $\theta^Q$ would correspond to ..
 
@@
 
-\begin{remark}
+\paragraph{Parameter update.}
+\begin{equation}
+  \nabla_{\theta^Q_{k+1}}\mathcal L(\theta^Q_{k+1})
+  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
+  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
+    \underbrace{\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1})}_{\text{???}}\right],
+  \label{eq:oracle}
+\end{equation}
+
+\paragraph{Link to classical RL algorithms.}
 Note that most RL algorithms, including \textit{Temporal Difference (TD)} learning [Sutton et al.,
-  Bertsekas] and SARSA can be cast in this general variational framework.
-\end{remark}
+  Bertsekas] and SARSA can be cast in this general variational framework. For example TD can corresponds
+to the above framework with a linear value approximator with feature encoding $\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$,
+and so $$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}$$
+...
 

### 2016-10-01 paper/paper.tex
more

@@
   Bertsekas] and SARSA can be cast in this general variational framework. For example TD can corresponds
-to the above framework with a linear value approximator with feature encoding $\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$,
+to the above framework with a linear value approximator with feature encoding $\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$
+on the grid $\mathcal S \times \mathcal A$,
 and so $$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}$$

### 2016-10-01 paper/paper.tex
more

--- 2016-09-30 paper/paper.tex
@@
 Note that most RL algorithms, including \textit{Temporal Difference (TD)} learning [Sutton et al.,
-  Bertsekas] and SARSA can be cast in this general variational framework. For example TD can corresponds
-to the above framework with a linear value approximator with feature encoding $\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$,
-and so $$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}$$
-...
+  Bertsekas] and SARSA can be cast in this general variational framework. For example TD corresponds
+to the above framework using a linear value approximator with feature encoding
+$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid $\mathcal S \times \mathcal A$,
+and so $$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
+%% Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
+%% $$\theta^Q(s,a) \leftarrow \theta^Q(s,a) + (Q(s, a|\theta^Q_{k+1}) - y)$$ ??
 

### 2016-10-01 paper/paper.tex
new comments from Silver video 1+2

--- 2016-09-30 paper/paper.tex
@@
 continuous.
-%% The environment $\mathcal E$ is generally taken as stochastic, but having some kind of
-%% memory.
+The environment $\mathcal E$ is generally taken as stochastic,
+is partially observed and 
+has some kind of memory.
 A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
+In RL, MDPs are the mathematical instantiation of the environment model,
+which generally can incorporate action probabilities (policy matrix)
+and/or value function given a state.
+(Silver: The environment can be fully defined by state space and
+transition probability matrix.)
 This is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
 \begin{itemize}
-\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
+\item $\mathcal S$ is any function of the history.
+The state exhibits the Markov property
+$\mathbb{P}(S_{t+1}|S_{t})=\mathbb{P}(S_{t+1}|S_{1}, \dots, S_{t})$.
+Consequently, the future states are independent of past states and only depend on the present.
+% \item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
 \item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
@@
 Thus the agent is positively biased towards considering rewards which are will
-come soon. Note that $R(s_t, a_t)$ defined above is a stochastic quantity which depends on the trajectory generated by the policy $\pi$, upto but not including time $t$
+come soon. The reward importance hence decays with distance in the future,
+given that there is more uncertainty in the farsighted future.
+The reward discount $\gamma$ calibrates the delay discounting
+behavior of the intelligent agent.
+
+Note that $R(s_t, a_t)$ defined above is a stochastic quantity which depends on the trajectory generated by the policy $\pi$, upto but not including time $t$
 \begin{eqnarray*}
@@
 the enormous number of possible trajectories to consider.
-One popular simplification that removes this issue is to only consider the sub-family of
-\textit{deterministic policies}, which map each state to the best single best action which
-should be taken at thin state, and therefore define functions from states to actions.
+One popular simplification that removes this issue is to only consider the sub-family of \textit{deterministic policies}, which map each state to the best single best action which should be taken at thin state, and therefore define functions from states to actions.
 
@@
 \begin{equation}
-  Q^\pi(s_t, a_t) = \mathbb E_{s_{t+1} \sim \rho(.|s_t,a_t)} [r(s_t,a_t) + \gamma Q^\pi(s_{t+1}, \pi(s_{t+1}))]
+  Q^\pi(s_t, a_t) = \mathbb E_{s_{t+1} \sim \rho(.|s_t)} [r(s_t,a_t) + \gamma Q^\pi(s_{t+1}, \pi(s_{t+1}))]
   \label{eq:bellman}
@@
 \subsection{Efficient RL with value approximation and replay}
-%As already mentioned in the previous section, Q-learning optimizes over the class of
-% deterministic policies of the form \eqref{eq:qlearning}.
-We need ways to approximate the state-action value function
-which scale up to large state spaces and / or action spaces.
-In state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
+As already mentioned in the previous section, Q-learning optimizes over the class of
+deterministic policies of the form \eqref{eq:qlearning}.
+in state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
 % (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
@@
 % (measured on their internal currency of worldly values ?),
-and the environment transitions to a new state $s' \in \mathcal S$. At the end of
-this step, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar
-behavior of the agent and is recorded in replay memory buffer:
-$\mathcal D \leftarrow \mathcal D \cup \{e\}$ (possibly discarding the oldest entries to make space).
-Now, at iteration $k+1$, replay consists in sampling (uniform or importance-weighted
-\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
+and the environment transitions to a new $s' \in \mathcal S$. At the end of
+this step, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar behavior of the agent and is recorded in replay memory:
+$\mathcal D \leftarrow \mathcal D \cup \{e\}$.
+Now, replay consists in sampling (uniform or importance-weighted\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
 $(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ and trying to
@@
 the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
-$r + \gamma \max_{a'} Q(s', a'|\theta^Q_k)$, with the output of a parametrized regression model $(s,a,r,s')
-\mapsto {Q}(s, a|\theta^Q_{k})$, where $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q_{k+1}$.
-Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function
-\begin{equation}
-  \mathcal L(\theta^Q_{k+1})
-  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
-  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
+$r + \gamma \max_{a'} Q(s', a'|\theta^Q)$, with the output of a parametrized regression model $(s,a,r,s') \mapsto {Q}(s, a|\theta^Q)$, where $Q(.,.|\theta^Q)$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q$.
+Computing an optimal policy then corresponds to finding the parameters $\theta^Q$ which minimize the following loss function
+\begin{equation}
+  \mathcal L(\theta^Q)
+  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q))^2]
+  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - y)^2],
   \label{eq:oracle}
@@
 where
-$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
+$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q) = r + \gamma Q(s', \pi(s')|\theta^Q)$ is the prediction target.
 % While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
-For example, a general linear model with a kernel $\phi$ would be of the
+
+For example, a general regression model with a kernel $(s, a) \mapsto \phi(s, a)$ would be of the
 form
+
 $${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
-$\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action $(s,a)$ into an appropriate
-feature space. Such are the techniques that have been proposed by [sutton et al. xyzw, Song et al. 2016 (NIPS),
-  and Parr et al. 2003 (JMLR)].
-A breakthrough alternative proposed by Google DeepMind [Minh et al. 2014, Silver et al 2015] is to learn this
-representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which
-are by far the current state-of-the-art in RL. In such a case, $\theta^Q$ would correspond to ..
+$\phi$ would represent a high-level hand-crafted representation of the state-action pairs into an appropriate feature space.
+Such are the techniques that have been proposed by [sutton et al.
+xyzw, Song et al.
+2016 (NIPS), and Parr et al.
+2003 (JMLR)].
+A breakthrough alternative proposed by Google DeepMind [Minh et al.
+2014, Silver et al 2015] is to learn this representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which are by far the current state-of-the-art in RL.
+In such a case, $\theta^Q$ would correspond to ..
 
@@
 
-\paragraph{Parameter update.}
-\begin{equation}
-  \nabla_{\theta^Q_{k+1}}\mathcal L(\theta^Q_{k+1})
-  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
-  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
-    \underbrace{\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1})}_{\text{???}}\right],
-  \label{eq:oracle}
-\end{equation}
-
-\paragraph{Link to classical RL algorithms.}
+\begin{remark}
 Note that most RL algorithms, including \textit{Temporal Difference (TD)} learning [Sutton et al.,
-  Bertsekas] and SARSA can be cast in this general variational framework. For example TD can corresponds
-to the above framework with a linear value approximator with feature encoding $\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$,
-and so $$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}$$
-...
+  Bertsekas] and SARSA can be cast in this general variational framework.
+\end{remark}
 

### 2016-10-02 paper/paper.tex
rushworth2016
truc

@@
 with fine-grained adjustment of contextual circumstances (Yoshida et al., 2011).
+Also the topographically neighboring dorsal anterior cingulate cortex
+has been linked to computing values and efforts of
+persisting a behavioral plan versus switching the
+environmental context in several lesion studies (Rushworth NN 2016).
 %
@@
 trajectories of state $s_t$.
+Given that each $\pi$ correspondence to an exhaustive description of
+behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
+the set of actions given states in finite time.
 
@@
 the enormous number of possible trajectories to consider.
+(If the Bellman-based computational ideas have a corresponding
+neurobiological implementation, this provides a framework of how
+optimal behavior could be computable in the human brain.)
 One popular simplification that removes this issue is to only consider the sub-family of
@@
 optimal control (Equation 6)"
+\item It is difficult to compute the free energy principle
+in practice, which makes it physiologically less plausible.
   \end{itemize}

### 2016-10-02 paper/paper.tex
synthesis I

@@
 
-\subsection{A Deep Q-learning for DMN function}
+\subsection{Correspondence between DMN function and RL framework}
 \textbf{XXX Put the diagrams and process charts here.}
+\begin{itemize}
+  \item The policy matrix represent the repertoire of possible actions
+  on the world given a current state. It represents the probabilites of
+  choosing actions to be executed in a certain situation.
+  \item Given a state, the value function represents the estimated utility of
+  being in that state, which is subjective to the agent. 
+  For state s there is set of actions a.
+  It provides a quantitative answer to the question:
+  \"What state is desirable for a specific agent.\"
+  For all actions being in one of the states s,
+  the Q value function provides the subjective utility
+  of executing a specific action.
+  It provides a quantitative answer to the question:
+  \"Which is the best action to take in the current situation.\"
+  \item The DMN may subserve
+  constant exploration of possible future actions and their
+  cumulative outcomes. Implicit computation of future choices 
+  provides an explanation for purposeful mind-wandering.
+  \item hippocampus -> provides perturbed action-transition-state-reward samples
+  as a block of "imagined", "hypothesized", "recalled" experience;
+  the small variations in these experience blocks allow searching
+  a larger space of parameters and possible experiences.
+  In the absence of environmental input and feedback
+  (e.g., mind-wandering or sleep) pseudo-experiencing possible
+  future scenarios and action outcomes.
+
+\end{itemize}
 

### 2016-10-02 paper/paper.tex
fixing yo sh*t

@@
 ongoing perception in the current environment (Maguire 2016).
--> self-play in deepmind Go
--> Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later
+$\implies$ self-play in deepmind Go
+$\implies$ Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later
 
@@
   It provides a quantitative answer to the question:
-  \"What state is desirable for a specific agent.\"
+  "What state is desirable for a specific agent."
   For all actions being in one of the states s,
@@
   It provides a quantitative answer to the question:
-  \"Which is the best action to take in the current situation.\"
+  "Which is the best action to take in the current situation."
   \item The DMN may subserve

### 2016-10-02 paper/paper.tex
more musings

--- 2016-10-02 paper/paper.tex
@@
 (Silver: The environment can be fully defined by state space and
-transition probability matrix.)
+transition probability matrix.
+Across lifetime, the MDP naturally incorporates always better
+models of the environment due to \points.)
 This is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
@@
 
+($\mu$ gives the initially random distributions over the states,
+the initially chaotic model of the agent at birth.)
 %% The goal is then to sample from a policy $a_t \sim \pi$ so that the expected
@@
 \begin{itemize}
-  \item The policy matrix represent the repertoire of possible actions
-  on the world given a current state. It represents the probabilites of
+  \item The policy matrix represents the repertoire of possible actions
+  on the world given a current state. It encodes the probabilites of
   choosing actions to be executed in a certain situation.
@@
   future scenarios and action outcomes.
+  \item
+  \textit{Inference} in the human brain reduces to generalization of
+  policy and value adaptations from sampled experiences to
+  successful action choices and reward predictions in future states.
 
@@
 \begin{itemize}
+  \item A common point between the Free Energy principle and our framework
+  is placing the minimization of suprise at the core of brain function.
+  In RL, surprise minimization reduces to accurate prediction of 
+  rewarding outcomes in the future.
   \item Ortega et al. 2013 \cite{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''

### 2016-10-02 paper/paper.tex
wip

--- 2016-10-02 paper/paper.tex
@@
 Since an action taken now might have repercussions long into the feature, it turns out that the
-quantity to optimize is not the instantaneous rewards $r(s_t, a_t)$, a cumulative version which
-takes into account the future. A standard for modeling this accumulation is the so-called
+quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
+cumulative version which takes into account the future. A standard for
+modeling this accumulation is the so-called
 time-discounted \textit{cumulative reward} function
@@
   \label{eq:cumr}
-  R(s_t, a_t) = r(s_t, s_t) + \ldots + \gamma^Tr(s_{t+T}, s_{t+T})
-  = r(s_t, s_t) + \gamma \sum_{\tau=1}^{T}\gamma^{\tau-1}r(s_{t+\tau},a_{t+\tau}),
-  \end{equation}
-for some \textit{discount factor} $\gamma \in [0, 1]$, and horizon $T$.
-The choice $\gamma = 0$ (or $T=0$)  corresponds to a perfectly myopic agent who is solely concerned about their immediate rewards and have no regards for the future.
-For there to be any learning at all, a technical requirement is that $\gamma < 1$ \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
-Thus the agent is positively biased towards considering rewards which are will
-come soon. The reward importance hence decays with distance in the future,
-given that there is more uncertainty in the farsighted future.
-The reward discount $\gamma$ calibrates the delay discounting
-behavior of the intelligent agent.
-
-Note that $R(s_t, a_t)$ defined above is a stochastic quantity which depends on the trajectory generated by the policy $\pi$, upto but not including time $t$
+  R^\pi(s_0, a_0) = r(s_0, s_0) + \gamma r(s_1, a_1) + \ldots + 
+  = \sum_{t=0}^{\infty}\gamma^{t}r(s_{t},a_{t}),
+\end{equation}
+which measures the cumulative reward of starting off at state $s_0$ taking action $a_0$, then following a policy $\pi$ for choosing actions henceforth,
+i.e
+
 \begin{eqnarray*}
   \begin{split}
-    s_0 &\sim \mu\\
-    a_0 &\sim \pi(.|s_0)\\
+    s_0, a_0 &\leftarrow s, a\\
     s_1 &\sim \rho(.|s_0,a_0) \text{ and collect rewards }r(s_0, a_0)\\
@@
     a_{t-1} &\sim \pi(.|s_{t-1})\\
-    s_{t} &\sim \rho(.|s_{t-1},a_{t-1}) \text{ and collect rewards }r(s_{t-1}, a_{t-1})
+    s_{t} &\sim \rho(.|s_{t-1},a_{t-1}) \text{ and collect rewards }r(s_{t-1}, a_{t-1})\\
+    \ldots
   \end{split}
@@
 
-%% The goal is then to sample from a policy $a_t \sim \pi$ so that the expected
-%% value of $R(s_t, a_t)$ is maximized.
-Formally, the agent's goal is then to maximize the state-action value for
-taking action $a_t$ and state $s_t$ defined by
+In \eqref{eq:cumr}, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}.
+The choice $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their immediate rewards and have no regards for the future.
+For there to be any learning at all, a technical requirement is that $\gamma < 1$.
+% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
+Thus the agent is positively biased towards considering rewards which are will
+come soon. The reward importance hence decays with distance in the future,
+given that there is more uncertainty in the farsighted future.
+The reward discount factor $\gamma$ can thus be seen as calibrating
+the delay discounting behavior of the intelligent agent.
+
+Note that $R^\pi(s_0, a_0)$ defined above is a stochastic quantity which depends on the trajectory generated by the policy $\pi$ from state $s$ onwards.
+The agent's goal is find a polyciy $\pi$ which maximizes the expected cumulative value of state-stat-pairs $(a_0,s_0)$, namely
 \begin{equation}
   \label{eq:qval}
-  Q^{\pi}(s_t,a_t) = \mathbb E [R(s_t, a_t)]
-\end{equation}
-
-where the expectation is over all possible trajectories successor
-trajectories of state $s_t$.
-Given that each $\pi$ correspondence to an exhaustive description of
-behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
-the set of actions given states in finite time.
+  Q^{\pi}(s_0,a_0) = \mathbb E [R(s_0, a_0)]
+\end{equation}
+
+where the expectation is over all possible trajectories generate by the
+policy $\pi$, starting from state $s_0$ and action $a_0$ taken at this
+state.
+%% Given that each $\pi$ correspondence to an exhaustive description of
+%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
+%% the set of actions given states in finite time.
 
 By construction, in general, maximizing the expectation in  \eqref{eq:qval} is intractable due to
-the enormous number of possible trajectories to consider.
+the enormous number of possible trajectories the agent might need to
+consider.
 (If the Bellman-based computational ideas have a corresponding
@@
 \begin{equation}
-  Q^\pi(s_t, a_t) = \mathbb E_{s_{t+1} \sim \rho(.|s_t,a_t)} [r(s_t,a_t) + \gamma Q^\pi(s_{t+1}, \pi(s_{t+1}))]
+  Q^\pi(s, a) = \mathbb E_{s' \sim \rho(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))]
   \label{eq:bellman}
@@
   being in that state, which is subjective to the agent. 
-  For state s there is set of actions a.
-  It provides a quantitative answer to the question:
-  "What state is desirable for a specific agent."
-  For all actions being in one of the states s,
-  the Q value function provides the subjective utility
-  of executing a specific action.
-  It provides a quantitative answer to the question:
-  "Which is the best action to take in the current situation."
+  For a state $s$, the value function at $s$ namely $V(s)$ define provides a quantitative answer to the question:
+  "How desirable is it to be in state $s$ ?"
+  For each action $a$ that can be taken from state $s$,
+  the Q-value  $Q(s, a)$ defined in \eqref{eq:qval} provides the subjective
+  utility of executing a specific action; it answers the question
+  "What is the expected utiltiy of taking action $a$ in this situation ?"
   \item The DMN may subserve

### 2016-10-02 paper/paper.tex
thoughts in a train

--- 2016-10-02 paper/paper.tex
@@
 $r = r(s, a)$ (or regret!, since it can be negative) collected by the
-agent. This reward should be thought
+agent. In this view, the environmental is partially controlled by
+the action of the agent.
+This reward should be thought
 of as an instantaneous satisfaction accompanying the execution of
@@
   successful action choices and reward predictions in future states.
+  \item In the RL view, plasticity in the DMN arises naturally.
+  If an agent behaving optimally in a given environment moves
+  to novel, yet unexperienced environments, reward prediction errors will
+  massively increase. This will lead to adaptation of the policy matrix
+  and value function until the system converges to a
+  new steady-state of optimal action decisions given certain states.
 

### 2016-10-02 paper/paper.tex

--- 2016-10-02 paper/paper.tex
@@
 ongoing perception in the current environment (Maguire 2016).
-$\implies$ self-play in deepmind Go
-$\implies$ Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later
+-> self-play in deepmind Go
+-> Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later
 
@@
 
+($\mu$ gives the initially random distributions over the states,
+the initially chaotic model of the agent at birth.)
+%% The goal is then to sample from a policy $a_t \sim \pi$ so that the expected
+%% value of $R(s_t, a_t)$ is maximized.
+Formally, the agent's goal is then to maximize the state-action value for
+taking action $a_t$ and state $s_t$ defined by
 In \eqref{eq:cumr}, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}.
@@
 The agent's goal is find a polyciy $\pi$ which maximizes the expected cumulative value of state-stat-pairs $(a_0,s_0)$, namely
+
 \begin{equation}
@@
   future scenarios and action outcomes.
+  \item
+  \textit{Inference} in the human brain reduces to generalization of
+  policy and value adaptations from sampled experiences to
+  successful action choices and reward predictions in future states.
+  \item In the RL view, plasticity in the DMN arises naturally.
+  If an agent behaving optimally in a given environment moves
+  to novel, yet unexperienced environments, reward prediction errors will
+  massively increase. This will lead to adaptation of the policy matrix
+  and value function until the system converges to a
+  new steady-state of optimal action decisions given certain states.
 

### 2016-10-02 paper/paper.tex
notes on a train II

@@
 describe and discuss neuroscientific studies on the DMN.
-
+%
+In principle,
+neuroscientific experiments can be designed that operationalize
+the set of action, value, and state variables that determine
+the behavior of RL systems. This makes proposed machine-learning
+approach to DMN function not only practically computable but
+also neuroscientifically falsifiable.
+%
 The DMN can be viewed as a smoothing kernel extending

### 2016-10-03 paper/paper.tex
RL-PC number I

@@
 \paragraph{Predictive Coding and Bayesian Brain Hypothesis}
-abc
-The Predictive Coding Framework.
-construction of detached probabilistic scenes has been argued to influence ongoing decision-making by estimating outcomes of behavioral choices.
+The predictive coding framework 
+is one of the most frequently evoked hypotheses of default mode function.
+Cortical responses emerge from a functional network of continuously interacting high and low hierarchical levels,
+with a dynamic intercourse that performs inference on the world by reconciling gaps between sensory input and prior information. 
+Feed-forward sensory experience is constantly modulated by
+modulatory top-down modulation at various hierarchical processing levels.
+At each stage of processing abstraction level,
+an internally generated prediction of the sensory input is
+directly compared against actual external input.
+A prediction error, i.e. the difference between sensory input
+and internal prediction,
+occurring at one of the processing levels
+incurs plasticity changes of neuronal back-projections (i.e., model parameters)
+for gradually improved future prediction of the environment.
+Indeed,
+back projection outnumber by far the input projection in the brain
+Salin, P.A. and Bullier, J. (1995) Corticocortical connections in the visual system: structure and function. Physiol. Rev. 75, 107–154.
+This process permits updates of beliefs about the world
+to best fit the incoming sensory examples (Friston2014).
+The generative model of how data are generated in the
+world would hence be incorporated by
+the current neuronal wiring.
+Contextual integration is maintained by top-down modulation of sensorimotor
+processing by context-specific a-priori information.
+formulate perception as a constructive process
+based on internal or generative models.
 
 \begin{itemize}
-  \item It is one of the most frequently cited hypotheses of default mode function
-  \item Within this model the brain is conceptualized as a Bayesian machine
-  \item Predictive coding is a framework that, in a hierarchical setting,
-  is equivalent to empirical Bayesian inference
-  \item External sensory input is feed-forward/bottom-up processed in the brain
-  \item Their processing is compared against expected input
-  \item The prediction error, i.e. the difference between sensory input
-  and internal prediction, is computed at each level and passed to
-  higher levels via forward connections
-  \item in case of a mismatch, back-projections top-down modulate
-  input processing
-  \item ‘contextual integration’ for top-down modulation of sensorimotor
-  processing by context-specific a-priori information.
-  \item Closely related to Friston's free-energy principle: brain as inference engine
-  biological systems, including brains, must minimize the long-term average of surprise;
-  formulate perception as a constructive process based on internal or generative models
-  \item A generative model of the world
+  \item Both accounts are process theories, supported by
+  neurobiological evidence, that
+  view the brain as a \"statistical organ\" (Friston 2014).
+  \item Both have a notion of brain plasticity with increased
+  adaption of neuronal wiring when faced by unprecedented environmental
+  challenges.
+  \item hierarchical aspects from predictive coding in RL lies in the
+  nested prediction of probable upcoming actions and rewards
+  \item Both model consequences of action. In RL, the horizon of that
+  look into the future is manifest in the $\gamma$ parameter, but not
+  explicitly modeled in the PC account.
+  \item mismatch negativity of PC is immanent in reward prediction error
+  by the difference between actually predicted reward of an action given
+  a state and the reward predicted by non-linear regression with graded
+  discounting of future rewards
+  \item Sensory experience is a generative process from both views.
+  In predictive coding sensory perception of the external world
+  is a generative experience due to modulatory top down experience at
+  various hierarchical levels of sensory input processing.
+  In RL, the environmental model is incorporated in the Markov decision
+  process, \textit{which can be fully recovered based on the last
+  state alone}.
+  \item Both provide a parsimoneous explanation why
+  humans seem to minimize processing of incoming information
+  when the environment is predictable and allocate attention when
+  novel stimuli are encountered.
+  Similarly, the both propose explanations why
+  predicting the future is inherently
+  linked to information from the past; similarly for independence across
+  geographical spaces.
 \end{itemize}
 
-Friston2014:
-
-This perspective shifts away from the brain as a passive filter of sensations (or an elaborate stimulus– response link) towards a view of the brain as a statistical organ that generates hypotheses or fantasies that are tested against sensory evidence.
-In short, the brain is now considered a phantastic organ (from Greek phantastikos, the ability to create mental images).
-For many people, this perspective can be traced back to Hermann von Helmholtz and the notion of unconscious inference: that is, a pre-rational mechanism by which visual impressions are formed (eg, the seemingly automatic but erroneous belief that the sun rises and sets in the sky, as opposed to the truth that the Earth rotates around it).
-
-Predictive coding is not a normative or descriptive scheme, it is a process theory with a biologically plausible basis—there is now much circumstantial anatomical and physiological evidence for predictive coding in the brain.12–16 In this scheme, neuronal representations in higher levels of cortical hierarchies generate predictions of representations in lower levels.
-These top-down predictions are compared with representations at the lower level to form a prediction
-error (associated with the activity of superficial pyramidal cells).
-This mismatch signal is passed back up the hierarchy, to update higher representations (associated with the activity of deep pyramidal cells).
-
-A description of biological processes as minimising (variational) free energy—a statistical measure of the surprise or improbability of sensory data, under a model of how those data were generated.
-
-predictive coding represents a biologically plausible scheme that enables the brain to update beliefs about the world with sensory samples
-
- individuals spend most of their time predicting the internal (proprioceptive) and external (exteroceptive) consequences of behaviour (both their own and that of others).
-To fully appreciate the bilateral nature of these predictions, inference can be considered in an embodied context.
-In this setting, perception can be understood as resolving exteroceptive prediction errors by selecting the predictions that best explain sensations.
-Conversely, behaviour suppresses pro- prioceptive prediction error by changing proprioceptive sensations.
-This suppression relies on classic reflexes, in which equilibrium points are set by descending proprioceptive predictions (figure 1).15 This process is called active inference
-
-providing a rich dynamical repertoire that enables the brain to respond quickly to changing inputs
-
-the formal constraints implicit in predictive coding mandate modulatory gain control for ascending prediction errors
-
 \paragraph{Semantic Hypothesis}
-abc
+
+The proposal promoted here is that the brain is continually engaged in generating predictions, and that these predictions rely on associative activation. 
+what people do when their mental capacity is not completely consumed by a specific task is to generate associations
+Traditionally this process has been considered as recognition, classification, or even a type of memory retrie- val, but in the present context I treat this process as analogy- making instead. 
+in analogy, the emphasis is on ‘what is it like?
+Analogy- based mappings of properties manifest themselves in processes ranging from perception and memory [17] to stereotypic judgments and prejudice
+While our existing memories are used to derive analo- gies and activate predictions, they are constantly being updated.
+The analogical process, in addition to affording the interpretation of our environment, subsequently aug- ments previous representations in a way that fosters increasingly flexible future analogies.
+(Bar)
+
+phantastic organ (from Greek phantastikos, the ability to create mental images)
+Friston2014
+
+this perspective can be traced back to Hermann von Helmholtz and the notion of unconscious inference: that is, a pre-rational mechanism by which visual impressions are formed (eg, the seemingly automatic but erroneous belief that the sun rises and sets in the sky, as opposed to the truth that the Earth rotates around it).
+
+mem- ory-based predictions, and these predictions are generated continually either based on gist information gleaned from the senses or driven by thought. The emphasis in this.
+The proposed account integrates three primary components.
+The first is associations, which are formed by a lifetime of extracting repeating patterns and statisti- cal regularities from our environment, and storing them in memory.
+The second is the concept of analogies, whereby
+Corresponding author: Bar, M. (bar@nmr.mgh.harvard.edu). Available online 4 June 2007.
+we seek correspondence between a novel input and existing representations in memory (e.g.
+‘what does this look like?’).
+Finally, these analogies activate associated representa- tions that translate into predictions
+(Bar)
+
+A second important functional account of the DMN is the 'semantic hypothesis'.
+It holds that retrieval of building blocks of world knowledge enables language and other high-level faculties.
+The DMN central involvement in semantic processing results in its implication in many a-priori unrelated aspects of human behavior, such as reasoning, planning, problem solving and decision making (Binder, Desai, Graves, & Conant, 2009; Schilbach et al., 2008; Suddendorf & Corballis, 2007).
+This is also illustrated by the substantial relationship between semantic knowledge and words (Binder et al., 2009).
+The engagement of DMN in language processing matches the observation that the large majority of the DMN regions such as TPJ receive no direct input from primary sensory areas acting more as a “hermeneutic hub” (Binder et al., 2009; Seghier, 2013).
+The DMN activation rather enables the organization and processing of internally generated information in absence of stimulus presentation (Binder et al., 1999).
+The proposed role of the DMN in semantic processing is further highlighted by the convergence of these regions in a reciprocal relationship as an interface between the allocentric (bird’s eye view) and egocentric (first person view) perspective; in this way self-reflection, social knowledge, and autobiographical memories are integrated.
+Moreover, the semantic account exquisitely explains the engagement of DMN in both recalling past and envisioning future (Demblon, Bahri, & D'Argembeau, 2016; Hassabis, Kumaran, Vann, & Maguire, 2007; Schacter et al., 2007; Spreng, Mar, & Kim, 2009), in that the internally stored in language world knowledge may lay the contextual foundations for envisioning future events as well as memory processes (Bzdok et al., 2012).
+Hence, the reported retrograde amnesia and the simultaneous impairment of imagining novel experience across patients with hippocampus lesions ties in well with this model (Hassabis & Maguire, 2007).
+In other words, DMN activation might underlie a time dimension independent cognitive process of constructing contextual scenarios out of detached memories (Buckner & Carroll, 2007; Gaesser, Spreng, McLelland, Addis, & Schacter, 2013; Hassabis & Maguire, 2007).
+Summing up the 'semantic account', DMN offers the conceptual framework that underlies not only language but also other major high-level functions of human behavior.
+
+\begin{itemize}
+  \item The notion of semantic or knowledge association is
+  incorporated into the MDP as a function of the state and action
+  history.
+  \item Both: there is no evidence to indicate that predictions of various levels of complexity, abstraction, timescale and purpose use mech- anisms that are qualitatively different
+\end{itemize}
 
@@
 -> task is policy/value updates to optmize short-term action / rest updates for mid- and long-term action
+
+ The 'sentinel hypothesis' is a major functional account to explain DMN activity that this canonical network acts as a cohesive unit.
+This coherent neurobiological network would be prominently engaged as a “radar” as well as in self-referential information processing including mind wandering, autobiographical recall, mentalizing, reflection about past and future (Buckner, Andrews-Hanna, & Schacter, 2008; Schilbach, Eickhoff, Rotarska-Jagiela, Fink, & Vogeley, 2008).
+Acting as a sentinel, DMN functions as a broad focus manager that performs continuous environmental tracking for salient and unexpected events (Buckner et al., 2008).
+This default activity may allow the detection of possible predators and potentially serves an evolutionarily conserved purpose by means of attention allocation (Raichle et al., 2001).
+The “radar” function may be efficiently supported by the activation patterns of DMN in diffuse attention condition tasks with targets appearing in various locations (Hahn, Ross, & Stein, 2007).
+On the other hand, the salient information perceived from the external environment activates self-referential information processing (Krienen, Tu, & Buckner, 2010).
+A conceptualization of these functions suggests two highly adaptive interacting DMN subsystems (Andrews-Hanna, Saxe, & Yarkoni, 2014; Schacter, Addis, & Buckner, 2007).
+However, the sentinel hypothesis may not exhaustively explain the variety of findings on the DMN connectivity with memory-related structures such as hippocampus (Giugni, Vadala, De Vincentiis, Colica, & Bastianello, 2010; McCormick et al., 2014).
+Summing up this first candidate account for DMN function, the sentinel hypothesis heavily emphasizes on the DMN as a single structure with a continuous sentinel role increasing during self-referential tasks.
 
@@
 \begin{itemize}
+  \item Closely related to Friston's free-energy principle: brain as inference engine
+  biological systems, including brains, must minimize the long-term average of surprise;
   \item A common point between the Free Energy principle and our framework

### 2016-10-04 paper/paper.tex
better .tex on RL

--- 2016-10-02 paper/paper.tex
@@
 
-\section{Reinforcement-Learning as a computational model for DMN function}
-The main idea is to show that the DMN has biological substrates (gradient
+\section{Reinforcement Learning as a computational model for DMN function}
+The main idea of this section is to argue that the DMN has biological substrates (gradient
 computation, backprop, experience replay, etc.) for all the
-major components of an efficient RL algorithm, and all these components
-integrate in a plausible way.
-\subsection{Markov Decision Processes}
-In a general reinforcement-learning (RL) framework, an \textit{agent}
+major components of an efficient \textit{Reinforcement Learning} algorithm, and all these
+components integrate in a plausible way.
+Reinforcement Learning (RL) is learning to achieve a goal by interacting with an external
+environment. In a general RL framework, an \textit{agent}
 interacts with an  environment in a trial-by-error fashion by taking
@@
 \textit{state} of the environment
-$s \rightarrow s'$, accompanied with a real-valued \textit{reward}
+$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
 $r = r(s, a)$ (or regret!, since it can be negative) collected by the
-agent. In this view, the environmental is partially controlled by
-the action of the agent.
-This reward should be thought
+agent. This reward should be thought
 of as an instantaneous satisfaction accompanying the execution of
-an action. Here the subscript $t$ denotes time, and can be discrete or
-continuous.
+an action.
+%% Here the subscript $t$ denotes time, and can be discrete or
+%% continuous.
 The environment $\mathcal E$ is generally taken as stochastic,
-is partially observed and 
-has some kind of memory.
+is partially observed, in the sense that only part of the current state is observed by
+the agent.
+\subsection{Markov Decision Processes}
 A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
-In RL, MDPs are the mathematical instantiation of the environment model,
-which generally can incorporate action probabilities (policy matrix)
-and/or value function given a state.
-(Silver: The environment can be fully defined by state space and
-transition probability matrix.
-Across lifetime, the MDP naturally incorporates always better
-models of the environment due to \points.)
-This is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
+An MDP is
+%% In RL, MDPs are the mathematical instantiation of the environment model,
+%% which generally can incorporate action probabilities (policy matrix)
+%% and/or value function given a state.
+%% (Silver: The environment can be fully defined by state space and
+%% transition probability matrix.)
+is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
 \begin{itemize}
-\item $\mathcal S$ is any function of the history.
-The state exhibits the Markov property
-$\mathbb{P}(S_{t+1}|S_{t})=\mathbb{P}(S_{t+1}|S_{1}, \dots, S_{t})$.
-Consequently, the future states are independent of past states and only depend on the present.
-% \item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
+\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
+\item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto \rho(s'|s,a)$,
+  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
+  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present.  
 \item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
@@
   $\mu(s)$ is the probability that the environment starts-off in state $s$.
-  \item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto \rho(s'|s,a)$, the probability of moving to state $s'$ if action $a$ is taken from state $s$.
-In addition, one requires that such transitions be Markovian.
 \end{itemize}
@@
   version of the states
-  $$s_t \leftarrow (s_0,a_0,r_0,s_1,a_1,s_2\ldots,s_t,a_{t-1},r_{t-1},s_{t})$$ made of the system's
+  $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
   history up to an including time $t$.
@@
 Knowledge the agent has about the environment can be encoded in the
-functions $r$, $\mu$, $\rho$, and / or $\pi$, as is done in
+functions $r$, $\mu$, $\rho$, and / or $\mu$, as is done in
 \textit{model-based} RL. For example if a rat in a maze knows, for sure,
@@
 The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
-sampling probabilities over candidate actions to perform from thin state. The  goal of RL is then
-adapt this policy so as to  maximize the cumulative rewards.
-
+sampling probabilities over candidate actions to perform from this state. Starting a time $t=0$,
+a policy $\pi$ generates a trajectory as follows
+\begin{eqnarray*}
+  \begin{split}
+    S_1 &\sim \rho(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
+    A_1 &\sim \pi(.|S_1)\\
+    S_2 &\sim \rho(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
+    \ldots\\
+    A_{t} &\sim \pi(.|S_{t})\\
+    S_{t+1} &\sim \rho(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
+    \ldots
+  \end{split}
+\end{eqnarray*}
 Since an action taken now might have repercussions long into the feature, it turns out that the
@@
   \label{eq:cumr}
-  R^\pi(s_0, a_0) = r(s_0, s_0) + \gamma r(s_1, a_1) + \ldots + 
-  = \sum_{t=0}^{\infty}\gamma^{t}r(s_{t},a_{t}),
-\end{equation}
-which measures the cumulative reward of starting off at state $s_0$ taking action $a_0$, then following a policy $\pi$ for choosing actions henceforth,
-i.e
-
-\begin{eqnarray*}
-  \begin{split}
-    s_0, a_0 &\leftarrow s, a\\
-    s_1 &\sim \rho(.|s_0,a_0) \text{ and collect rewards }r(s_0, a_0)\\
-    a_1 &\sim \pi(.|s_1)\\
-    s_2 &\sim \rho(.|s_1,a_1) \text{ and collect rewards }r(s_1, a_1)\\
-    \ldots\\
-    a_{t-1} &\sim \pi(.|s_{t-1})\\
-    s_{t} &\sim \rho(.|s_{t-1},a_{t-1}) \text{ and collect rewards }r(s_{t-1}, a_{t-1})\\
-    \ldots
-  \end{split}
-\end{eqnarray*}
-
-($\mu$ gives the initially random distributions over the states,
-the initially chaotic model of the agent at birth.)
-%% The goal is then to sample from a policy $a_t \sim \pi$ so that the expected
-%% value of $R(s_t, a_t)$ is maximized.
-Formally, the agent's goal is then to maximize the state-action value for
-taking action $a_t$ and state $s_t$ defined by
-In \eqref{eq:cumr}, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}.
+  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t.
+\end{equation}
+This random variable\footnote{Random as it depends both on the environment's dynamic and the
+  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
+following a
+policy $\pi$. The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
+In the definition of $G^\pi$ above, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}.
 The choice $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their immediate rewards and have no regards for the future.
@@
 
-Note that $R^\pi(s_0, a_0)$ defined above is a stochastic quantity which depends on the trajectory generated by the policy $\pi$ from state $s$ onwards.
-The agent's goal is find a polyciy $\pi$ which maximizes the expected cumulative value of state-stat-pairs $(a_0,s_0)$, namely
-
+The agent's goal can then be formulated as being that of finding policy $\pi$ which maximizes the
+expected cumulative value of state-stat-pairs $(A_0,S_0)$, namely
 \begin{equation}
   \label{eq:qval}
-  Q^{\pi}(s_0,a_0) = \mathbb E [R(s_0, a_0)]
-\end{equation}
-
-where the expectation is over all possible trajectories generate by the
-policy $\pi$, starting from state $s_0$ and action $a_0$ taken at this
-state.
+  Q^{\pi}(s,a) = \mathbb E [G^\pi|S_0=s,A_0=a,A_1 \sim \pi,A_2 \sim \pi,\ldots].
+\end{equation}
+That is, the expectated future reward over all possible trajectories in which we start at $s$ take
+action $a$ and then follow the policy $\pi$ thereafter to select future actions.
 %% Given that each $\pi$ correspondence to an exhaustive description of
-%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
+%% behavioral patterns, $Q^{\pi}(S_t,A_t)$ corresponds to the epistemic value of
 %% the set of actions given states in finite time.
@@
 should be taken at thin state, and therefore define functions from states to actions.
-
 \textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
@@
 \begin{equation}
-  Q^\pi(s, a) = \mathbb E_{s' \sim \rho(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))]
+  Q^\pi(s, a) = \mathbb E_{s' \sim \rho(.|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
+  = \mathbb E_{s' \sim \rho(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))]
   \label{eq:bellman}
@@
 behavior of the agent and is recorded in replay memory buffer:
-$\mathcal D \leftarrow \mathcal D \cup \{e\}$ (possibly discarding the oldest entries to make space).
+$D \leftarrow D \cup \{e\}$ (possibly discarding the oldest entries to make space).
 Now, at iteration $k+1$, replay consists in sampling (uniform or importance-weighted
@@
   \mathcal L(\theta^Q_{k+1})
-  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
-  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
+  % = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
+  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
   \label{eq:oracle}
@@
 representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which
-are by far the current state-of-the-art in RL. In such a case, $\theta^Q$ would correspond to ..
-
-The entire model can then be effectively trained via \textit{backprop} wherein online stochastic
-gradient descent is used to minimize the variational objective \eqref{eq:oracle}.
+are by far the current state-of-the-art in RL. In such a case, $\theta^Q$
+would correspond to ..
+
+The entire model can then be effectively trained via \textit{backprop}
+wherein online stochastic gradient descent is used to minimize the
+variational objective \eqref{eq:oracle}.
 
@@
   % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
-  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
+  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
     \underbrace{\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1})}_{\text{???}}\right],
@@
 \paragraph{Link to classical RL algorithms.}
-Note that most RL algorithms, including \textit{Temporal Difference (TD)} learning [Sutton et al.,
-  Bertsekas] and SARSA can be cast in this general variational framework. For example TD corresponds
-to the above framework using a linear value approximator with feature encoding
-$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid $\mathcal S \times \mathcal A$,
-and so $$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
+Note that most RL algorithms, including \textit{Temporal Difference (TD)}
+learning [Sutton et al., Bertsekas], REINFORCE [William], and SARSA can be
+cast in this general variational framework. For example TD corresponds to
+the above framework using a linear value approximator with feature encoding
+$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
+$\mathcal S \times \mathcal A$, and so
+$$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
 %% Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
@@
 \begin{itemize}
-  \item The policy matrix represents the repertoire of possible actions
-  on the world given a current state. It encodes the probabilites of
+  \item The policy matrix represent the repertoire of possible actions
+  on the world given a current state. It represents the probabilites of
   choosing actions to be executed in a certain situation.
@@
   future scenarios and action outcomes.
-  \item
-  \textit{Inference} in the human brain reduces to generalization of
-  policy and value adaptations from sampled experiences to
-  successful action choices and reward predictions in future states.
-  \item In the RL view, plasticity in the DMN arises naturally.
-  If an agent behaving optimally in a given environment moves
-  to novel, yet unexperienced environments, reward prediction errors will
-  massively increase. This will lead to adaptation of the policy matrix
-  and value function until the system converges to a
-  new steady-state of optimal action decisions given certain states.
 
@@
 \begin{itemize}
-  \item A common point between the Free Energy principle and our framework
-  is placing the minimization of suprise at the core of brain function.
-  In RL, surprise minimization reduces to accurate prediction of 
-  rewarding outcomes in the future.
   \item Ortega et al. 2013 \cite{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
@@
 is a non-trivial problem and usually involves backwards induction
-or some approximation scheme like reinforcement-learning [4–6].
+or some approximation scheme like reinforcement learning [4–6].
 The free-energy formulation circumvents these problems by

### 2016-10-04 paper/paper.tex

--- 2016-10-03 paper/paper.tex
@@
 \textit{state} of the environment
-$s \rightarrow s'$, accompanied with a real-valued \textit{reward}
+$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
 $r = r(s, a)$ (or regret!, since it can be negative) collected by the
@@
 of as an instantaneous satisfaction accompanying the execution of
-an action. Here the subscript $t$ denotes time, and can be discrete or
-continuous.
+an action.
+%% Here the subscript $t$ denotes time, and can be discrete or
+%% continuous.
 The environment $\mathcal E$ is generally taken as stochastic,
-is partially observed and 
-has some kind of memory.
+is partially observed, in the sense that only part of the current state is observed by
+the agent.
+\subsection{Markov Decision Processes}
 A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
-In RL, MDPs are the mathematical instantiation of the environment model,
-which generally can incorporate action probabilities (policy matrix)
-and/or value function given a state.
-(Silver: The environment can be fully defined by state space and
-transition probability matrix.
-Across lifetime, the MDP naturally incorporates always better
-models of the environment due to \points.)
-This is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
+%% In RL, MDPs are the mathematical instantiation of the environment model,
+%% which generally can incorporate action probabilities (policy matrix)
+%% and/or value function given a state.
+%% (Silver: The environment can be fully defined by state space and
+%% transition probability matrix.)
+is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
 \begin{itemize}
-\item $\mathcal S$ is any function of the history.
-The state exhibits the Markov property
-$\mathbb{P}(S_{t+1}|S_{t})=\mathbb{P}(S_{t+1}|S_{1}, \dots, S_{t})$.
-Consequently, the future states are independent of past states and only depend on the present.
-% \item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
+\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
+\item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto \rho(s'|s,a)$,
+  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
+  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present.  
 \item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
@@
 It holds that retrieval of building blocks of world knowledge enables language and other high-level faculties.
-The DMN central involvement in semantic processing results in its implication in many a-priori unrelated aspects of human behavior, such as reasoning, planning, problem solving and decision making (Binder, Desai, Graves, & Conant, 2009; Schilbach et al., 2008; Suddendorf & Corballis, 2007).
+The DMN central involvement in semantic processing results in its implication in many a-priori unrelated aspects of human behavior, such as reasoning, planning, problem solving and decision making (Binder, Desai, Graves, \& Conant, 2009; Schilbach et al., 2008; Suddendorf \& Corballis, 2007).
 This is also illustrated by the substantial relationship between semantic knowledge and words (Binder et al., 2009).
@@
 The proposed role of the DMN in semantic processing is further highlighted by the convergence of these regions in a reciprocal relationship as an interface between the allocentric (bird’s eye view) and egocentric (first person view) perspective; in this way self-reflection, social knowledge, and autobiographical memories are integrated.
-Moreover, the semantic account exquisitely explains the engagement of DMN in both recalling past and envisioning future (Demblon, Bahri, & D'Argembeau, 2016; Hassabis, Kumaran, Vann, & Maguire, 2007; Schacter et al., 2007; Spreng, Mar, & Kim, 2009), in that the internally stored in language world knowledge may lay the contextual foundations for envisioning future events as well as memory processes (Bzdok et al., 2012).
-Hence, the reported retrograde amnesia and the simultaneous impairment of imagining novel experience across patients with hippocampus lesions ties in well with this model (Hassabis & Maguire, 2007).
-In other words, DMN activation might underlie a time dimension independent cognitive process of constructing contextual scenarios out of detached memories (Buckner & Carroll, 2007; Gaesser, Spreng, McLelland, Addis, & Schacter, 2013; Hassabis & Maguire, 2007).
+Moreover, the semantic account exquisitely explains the engagement of DMN in both recalling past and envisioning future (Demblon, Bahri, \& D'Argembeau, 2016; Hassabis, Kumaran, Vann, \& Maguire, 2007; Schacter et al., 2007; Spreng, Mar, \& Kim, 2009), in that the internally stored in language world knowledge may lay the contextual foundations for envisioning future events as well as memory processes (Bzdok et al., 2012).
+Hence, the reported retrograde amnesia and the simultaneous impairment of imagining novel experience across patients with hippocampus lesions ties in well with this model (Hassabis \& Maguire, 2007).
+In other words, DMN activation might underlie a time dimension independent cognitive process of constructing contextual scenarios out of detached memories (Buckner \& Carroll, 2007; Gaesser, Spreng, McLelland, Addis, \& Schacter, 2013; Hassabis \& Maguire, 2007).
 Summing up the 'semantic account', DMN offers the conceptual framework that underlies not only language but also other major high-level functions of human behavior.
@@
  The 'sentinel hypothesis' is a major functional account to explain DMN activity that this canonical network acts as a cohesive unit.
-This coherent neurobiological network would be prominently engaged as a “radar” as well as in self-referential information processing including mind wandering, autobiographical recall, mentalizing, reflection about past and future (Buckner, Andrews-Hanna, & Schacter, 2008; Schilbach, Eickhoff, Rotarska-Jagiela, Fink, & Vogeley, 2008).
+This coherent neurobiological network would be prominently engaged as a “radar” as well as in self-referential information processing including mind wandering, autobiographical recall, mentalizing, reflection about past and future (Buckner, Andrews-Hanna, \& Schacter, 2008; Schilbach, Eickhoff, Rotarska-Jagiela, Fink, \& Vogeley, 2008).
 Acting as a sentinel, DMN functions as a broad focus manager that performs continuous environmental tracking for salient and unexpected events (Buckner et al., 2008).
 This default activity may allow the detection of possible predators and potentially serves an evolutionarily conserved purpose by means of attention allocation (Raichle et al., 2001).
-The “radar” function may be efficiently supported by the activation patterns of DMN in diffuse attention condition tasks with targets appearing in various locations (Hahn, Ross, & Stein, 2007).
-On the other hand, the salient information perceived from the external environment activates self-referential information processing (Krienen, Tu, & Buckner, 2010).
-A conceptualization of these functions suggests two highly adaptive interacting DMN subsystems (Andrews-Hanna, Saxe, & Yarkoni, 2014; Schacter, Addis, & Buckner, 2007).
-However, the sentinel hypothesis may not exhaustively explain the variety of findings on the DMN connectivity with memory-related structures such as hippocampus (Giugni, Vadala, De Vincentiis, Colica, & Bastianello, 2010; McCormick et al., 2014).
+The “radar” function may be efficiently supported by the activation patterns of DMN in diffuse attention condition tasks with targets appearing in various locations (Hahn, Ross, \& Stein, 2007).
+On the other hand, the salient information perceived from the external environment activates self-referential information processing (Krienen, Tu, \& Buckner, 2010).
+A conceptualization of these functions suggests two highly adaptive interacting DMN subsystems (Andrews-Hanna, Saxe, \& Yarkoni, 2014; Schacter, Addis, \& Buckner, 2007).
+However, the sentinel hypothesis may not exhaustively explain the variety of findings on the DMN connectivity with memory-related structures such as hippocampus (Giugni, Vadala, De Vincentiis, Colica, \& Bastianello, 2010; McCormick et al., 2014).
 Summing up this first candidate account for DMN function, the sentinel hypothesis heavily emphasizes on the DMN as a single structure with a continuous sentinel role increasing during self-referential tasks.

### 2016-10-04 paper/paper.tex
more

@@
   \text{laugh}, \text{sympathize}, \text{empathize}\}.$
-\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$, so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
+\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$ is the \textit{reward function},
+  so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
 \item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that

### 2016-10-05 paper/paper.tex
some stuff and some junk

--- 2016-10-04 paper/paper.tex
@@
 (Mason et al., 2007).
-These "offline" processes may however contribute to optimized control of the organism.
+These \"offline\" processes may however contribute to optimized control of the organism.
 We formalize
@@
 based on a brain implementation of Markov Chain Monte Carlo sampling
-(Tenenbaum2011Science). Even (necessarily imperfect) memories
+(Tenenbaum2011Science).
+\textit{Control} refers to the influences an agent exterts when interacting
+with the environment to encourage preferred states.
+Even (necessarily imperfect) memories
 of past experience, random mind-wandering and dreams during sleep
@@
 
-\section{Reinforcement Learning as a computational model for DMN function}
-The main idea of this section is to argue that the DMN has biological substrates (gradient
+\section{Reinforcement-Learning as a computational model for DMN function}
+The main idea is to show that the DMN has biological substrates (gradient
 computation, backprop, experience replay, etc.) for all the
-major components of an efficient \textit{Reinforcement Learning} algorithm, and all these
-components integrate in a plausible way.
-Reinforcement Learning (RL) is learning to achieve a goal by interacting with an external
-environment. In a general RL framework, an \textit{agent}
+major components of an efficient RL algorithm, and all these components
+integrate in a plausible way.
+\subsection{Markov Decision Processes}
+In a general reinforcement-learning (RL) framework, an \textit{agent}
 interacts with an  environment in a trial-by-error fashion by taking
@@
 \textit{state} of the environment
-$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
+$s \rightarrow s'$, accompanied with a real-valued \textit{reward}
 $r = r(s, a)$ (or regret!, since it can be negative) collected by the
@@
 of as an instantaneous satisfaction accompanying the execution of
-an action.
-%% Here the subscript $t$ denotes time, and can be discrete or
-%% continuous.
+an action. Here the subscript $t$ denotes time, and can be discrete or
+continuous.
 The environment $\mathcal E$ is generally taken as stochastic,
-is partially observed, in the sense that only part of the current state is observed by
-the agent.
-\subsection{Markov Decision Processes}
+is partially observed and 
+has some kind of memory.
 A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
-%% In RL, MDPs are the mathematical instantiation of the environment model,
-%% which generally can incorporate action probabilities (policy matrix)
-%% and/or value function given a state.
-%% (Silver: The environment can be fully defined by state space and
-%% transition probability matrix.)
-is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
+In RL, MDPs are the mathematical instantiation of the environment model,
+which generally can incorporate action probabilities (policy matrix)
+and/or value function given a state.
+(Silver: The environment can be fully defined by state space and
+transition probability matrix.
+Across lifetime, the MDP naturally incorporates always better
+models of the environment due to \points.)
+This is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
 \begin{itemize}
-\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
-\item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto \rho(s'|s,a)$,
-  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
-  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present.  
+\item $\mathcal S$ is any function of the history.
+The state exhibits the Markov property
+$\mathbb{P}(S_{t+1}|S_{t})=\mathbb{P}(S_{t+1}|S_{1}, \dots, S_{t})$.
+Consequently, the future states are independent of past states and only depend on the present.
+% \item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
 \item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
@@
   $\mu(s)$ is the probability that the environment starts-off in state $s$.
+  \item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto \rho(s'|s,a)$, the probability of moving to state $s'$ if action $a$ is taken from state $s$.
+In addition, one requires that such transitions be Markovian.
 \end{itemize}
@@
   version of the states
-  $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
+  $$s_t \leftarrow (s_0,a_0,r_0,s_1,a_1,s_2\ldots,s_t,a_{t-1},r_{t-1},s_{t})$$ made of the system's
   history up to an including time $t$.
@@
 Knowledge the agent has about the environment can be encoded in the
-functions $r$, $\mu$, $\rho$, and / or $\mu$, as is done in
+functions $r$, $\mu$, $\rho$, and / or $\pi$, as is done in
 \textit{model-based} RL. For example if a rat in a maze knows, for sure,
@@
 The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
-sampling probabilities over candidate actions to perform from this state. Starting a time $t=0$,
-a policy $\pi$ generates a trajectory as follows
-\begin{eqnarray*}
-  \begin{split}
-    S_1 &\sim \rho(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
-    A_1 &\sim \pi(.|S_1)\\
-    S_2 &\sim \rho(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
-    \ldots\\
-    A_{t} &\sim \pi(.|S_{t})\\
-    S_{t+1} &\sim \rho(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
-    \ldots
-  \end{split}
-\end{eqnarray*}
+sampling probabilities over candidate actions to perform from thin state. The  goal of RL is then
+adapt this policy so as to  maximize the cumulative rewards.
+
 Since an action taken now might have repercussions long into the feature, it turns out that the
@@
   \label{eq:cumr}
-  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t.
-\end{equation}
-This random variable\footnote{Random as it depends both on the environment's dynamic and the
-  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
-following a
-policy $\pi$. The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
-In the definition of $G^\pi$ above, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}.
+  R^\pi(s_0, a_0) = r(s_0, s_0) + \gamma r(s_1, a_1) + \ldots + 
+  = \sum_{t=0}^{\infty}\gamma^{t}r(s_{t},a_{t}),
+\end{equation}
+which measures the cumulative reward of starting off at state $s_0$ taking action $a_0$, then following a policy $\pi$ for choosing actions henceforth,
+i.e
+
+\begin{eqnarray*}
+  \begin{split}
+    s_0, a_0 &\leftarrow s, a\\
+    s_1 &\sim \rho(.|s_0,a_0) \text{ and collect rewards }r(s_0, a_0)\\
+    a_1 &\sim \pi(.|s_1)\\
+    s_2 &\sim \rho(.|s_1,a_1) \text{ and collect rewards }r(s_1, a_1)\\
+    \ldots\\
+    a_{t-1} &\sim \pi(.|s_{t-1})\\
+    s_{t} &\sim \rho(.|s_{t-1},a_{t-1}) \text{ and collect rewards }r(s_{t-1}, a_{t-1})\\
+    \ldots
+  \end{split}
+\end{eqnarray*}
+
+($\mu$ gives the initially random distributions over the states,
+the initially chaotic model of the agent at birth.)
+%% The goal is then to sample from a policy $a_t \sim \pi$ so that the expected
+%% value of $R(s_t, a_t)$ is maximized.
+Formally, the agent's goal is then to maximize the state-action value for
+taking action $a_t$ and state $s_t$ defined by
+In \eqref{eq:cumr}, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}.
 The choice $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their immediate rewards and have no regards for the future.
@@
 Thus the agent is positively biased towards considering rewards which are will
-come soon. The reward importance hence decays with distance in the future,
+come soon. The reward importance hence decays exponentially
+with distance in the future,
 given that there is more uncertainty in the farsighted future.
@@
 
-The agent's goal can then be formulated as being that of finding policy $\pi$ which maximizes the
-expected cumulative value of state-stat-pairs $(A_0,S_0)$, namely
+Note that $R^\pi(s_0, a_0)$ defined above is a stochastic quantity which depends on the trajectory generated by the policy $\pi$ from state $s$ onwards.
+The agent's goal is find a polyciy $\pi$ which maximizes the expected cumulative value of state-stat-pairs $(a_0,s_0)$, namely
+
 \begin{equation}
   \label{eq:qval}
-  Q^{\pi}(s,a) = \mathbb E [G^\pi|S_0=s,A_0=a,A_1 \sim \pi,A_2 \sim \pi,\ldots].
-\end{equation}
-That is, the expectated future reward over all possible trajectories in which we start at $s$ take
-action $a$ and then follow the policy $\pi$ thereafter to select future actions.
+  Q^{\pi}(s_0,a_0) = \mathbb E [R(s_0, a_0)]
+\end{equation}
+
+where the expectation is over all possible trajectories generate by the
+policy $\pi$, starting from state $s_0$ and action $a_0$ taken at this
+state.
 %% Given that each $\pi$ correspondence to an exhaustive description of
-%% behavioral patterns, $Q^{\pi}(S_t,A_t)$ corresponds to the epistemic value of
+%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
 %% the set of actions given states in finite time.
@@
 should be taken at thin state, and therefore define functions from states to actions.
+
 \textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
@@
 \begin{equation}
-  Q^\pi(s, a) = \mathbb E_{s' \sim \rho(.|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
-  = \mathbb E_{s' \sim \rho(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))]
+  Q^\pi(s, a) = \mathbb E_{s' \sim \rho(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))]
   \label{eq:bellman}
 \end{equation}
-which is a fixed-point equation for the deterministic policy $\pi$.
+which provides a recursive decomposition of optimal behavior.
+A complicated dynamic programming optimization can be
+broken into simpler subproblems at different time points.
+It is a fixed-point equation for the deterministic policy $\pi$.
 Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which
@@
 behavior of the agent and is recorded in replay memory buffer:
-$D \leftarrow D \cup \{e\}$ (possibly discarding the oldest entries to make space).
+$\mathcal D \leftarrow \mathcal D \cup \{e\}$ (possibly discarding the oldest entries to make space).
 Now, at iteration $k+1$, replay consists in sampling (uniform or importance-weighted
@@
   \mathcal L(\theta^Q_{k+1})
-  % = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
-  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
+  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
+  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
   \label{eq:oracle}
@@
 representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which
-are by far the current state-of-the-art in RL. In such a case, $\theta^Q$
-would correspond to ..
-
-The entire model can then be effectively trained via \textit{backprop}
-wherein online stochastic gradient descent is used to minimize the
-variational objective \eqref{eq:oracle}.
+are by far the current state-of-the-art in RL. In such a case, $\theta^Q$ would correspond to ..
+
+The entire model can then be effectively trained via \textit{backprop} wherein online stochastic
+gradient descent is used to minimize the variational objective \eqref{eq:oracle}.
 
@@
   % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
-  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
+  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
     \underbrace{\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1})}_{\text{???}}\right],
@@
 \paragraph{Link to classical RL algorithms.}
-Note that most RL algorithms, including \textit{Temporal Difference (TD)}
-learning [Sutton et al., Bertsekas], REINFORCE [William], and SARSA can be
-cast in this general variational framework. For example TD corresponds to
-the above framework using a linear value approximator with feature encoding
-$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
-$\mathcal S \times \mathcal A$, and so
-$$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
+Note that most RL algorithms, including \textit{Temporal Difference (TD)} learning [Sutton et al.,
+  Bertsekas] and SARSA can be cast in this general variational framework. For example TD corresponds
+to the above framework using a linear value approximator with feature encoding
+$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid $\mathcal S \times \mathcal A$,
+and so $$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
 %% Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
@@
   new steady-state of optimal action decisions given certain states.
+  \item The DMN is today known to consistently increase in neural
+  activity when engaging cognitive processes that are detached from
+  current sensory environment. DMN involvement in thinking about the past,
+  the future, and hypothetical possibilities ties in with the
+  MDP's implicit computation of action and state cascades as a function
+  of what happened in the past.
 
@@
 
+as showing that animals could learn a “cognitive map of the environment” in the absence of rewards or penalties, and that they could use the map later when they were motivated to reach a goal (Tolman, 1948).
+
 this perspective can be traced back to Hermann von Helmholtz and the notion of unconscious inference: that is, a pre-rational mechanism by which visual impressions are formed (eg, the seemingly automatic but erroneous belief that the sun rises and sets in the sky, as opposed to the truth that the Earth rotates around it).
@@
 It holds that retrieval of building blocks of world knowledge enables language and other high-level faculties.
-The DMN central involvement in semantic processing results in its implication in many a-priori unrelated aspects of human behavior, such as reasoning, planning, problem solving and decision making (Binder, Desai, Graves, \& Conant, 2009; Schilbach et al., 2008; Suddendorf \& Corballis, 2007).
+The DMN central involvement in semantic processing results in its implication in many a-priori unrelated aspects of human behavior, such as reasoning, planning, problem solving and decision making (Binder, Desai, Graves, & Conant, 2009; Schilbach et al., 2008; Suddendorf & Corballis, 2007).
 This is also illustrated by the substantial relationship between semantic knowledge and words (Binder et al., 2009).
@@
 The proposed role of the DMN in semantic processing is further highlighted by the convergence of these regions in a reciprocal relationship as an interface between the allocentric (bird’s eye view) and egocentric (first person view) perspective; in this way self-reflection, social knowledge, and autobiographical memories are integrated.
-Moreover, the semantic account exquisitely explains the engagement of DMN in both recalling past and envisioning future (Demblon, Bahri, \& D'Argembeau, 2016; Hassabis, Kumaran, Vann, \& Maguire, 2007; Schacter et al., 2007; Spreng, Mar, \& Kim, 2009), in that the internally stored in language world knowledge may lay the contextual foundations for envisioning future events as well as memory processes (Bzdok et al., 2012).
-Hence, the reported retrograde amnesia and the simultaneous impairment of imagining novel experience across patients with hippocampus lesions ties in well with this model (Hassabis \& Maguire, 2007).
-In other words, DMN activation might underlie a time dimension independent cognitive process of constructing contextual scenarios out of detached memories (Buckner \& Carroll, 2007; Gaesser, Spreng, McLelland, Addis, \& Schacter, 2013; Hassabis \& Maguire, 2007).
+Moreover, the semantic account exquisitely explains the engagement of DMN in both recalling past and envisioning future (Demblon, Bahri, & D'Argembeau, 2016; Hassabis, Kumaran, Vann, & Maguire, 2007; Schacter et al., 2007; Spreng, Mar, & Kim, 2009), in that the internally stored in language world knowledge may lay the contextual foundations for envisioning future events as well as memory processes (Bzdok et al., 2012).
+Hence, the reported retrograde amnesia and the simultaneous impairment of imagining novel experience across patients with hippocampus lesions ties in well with this model (Hassabis & Maguire, 2007).
+In other words, DMN activation might underlie a time dimension independent cognitive process of constructing contextual scenarios out of detached memories (Buckner & Carroll, 2007; Gaesser, Spreng, McLelland, Addis, & Schacter, 2013; Hassabis & Maguire, 2007).
 Summing up the 'semantic account', DMN offers the conceptual framework that underlies not only language but also other major high-level functions of human behavior.
@@
 DMN does social cognition -> but it exists in rats and monkeys with much more poorly developped social capacities
+-> our computational framework can be equally applied to humans
+and other primates.
 
 How can a same neurobiological circuit be equally important for baseline house-keeping functions and specific task performance?
+
+Humans have learned to solve ecological problems
+by social means.
 
@@
  The 'sentinel hypothesis' is a major functional account to explain DMN activity that this canonical network acts as a cohesive unit.
-This coherent neurobiological network would be prominently engaged as a “radar” as well as in self-referential information processing including mind wandering, autobiographical recall, mentalizing, reflection about past and future (Buckner, Andrews-Hanna, \& Schacter, 2008; Schilbach, Eickhoff, Rotarska-Jagiela, Fink, \& Vogeley, 2008).
+This coherent neurobiological network would be prominently engaged as a “radar” as well as in self-referential information processing including mind wandering, autobiographical recall, mentalizing, reflection about past and future (Buckner, Andrews-Hanna, & Schacter, 2008; Schilbach, Eickhoff, Rotarska-Jagiela, Fink, & Vogeley, 2008).
 Acting as a sentinel, DMN functions as a broad focus manager that performs continuous environmental tracking for salient and unexpected events (Buckner et al., 2008).
 This default activity may allow the detection of possible predators and potentially serves an evolutionarily conserved purpose by means of attention allocation (Raichle et al., 2001).
-The “radar” function may be efficiently supported by the activation patterns of DMN in diffuse attention condition tasks with targets appearing in various locations (Hahn, Ross, \& Stein, 2007).
-On the other hand, the salient information perceived from the external environment activates self-referential information processing (Krienen, Tu, \& Buckner, 2010).
-A conceptualization of these functions suggests two highly adaptive interacting DMN subsystems (Andrews-Hanna, Saxe, \& Yarkoni, 2014; Schacter, Addis, \& Buckner, 2007).
-However, the sentinel hypothesis may not exhaustively explain the variety of findings on the DMN connectivity with memory-related structures such as hippocampus (Giugni, Vadala, De Vincentiis, Colica, \& Bastianello, 2010; McCormick et al., 2014).
+The “radar” function may be efficiently supported by the activation patterns of DMN in diffuse attention condition tasks with targets appearing in various locations (Hahn, Ross, & Stein, 2007).
+On the other hand, the salient information perceived from the external environment activates self-referential information processing (Krienen, Tu, & Buckner, 2010).
+A conceptualization of these functions suggests two highly adaptive interacting DMN subsystems (Andrews-Hanna, Saxe, & Yarkoni, 2014; Schacter, Addis, & Buckner, 2007).
+However, the sentinel hypothesis may not exhaustively explain the variety of findings on the DMN connectivity with memory-related structures such as hippocampus (Giugni, Vadala, De Vincentiis, Colica, & Bastianello, 2010; McCormick et al., 2014).
 Summing up this first candidate account for DMN function, the sentinel hypothesis heavily emphasizes on the DMN as a single structure with a continuous sentinel role increasing during self-referential tasks.
@@
 is a non-trivial problem and usually involves backwards induction
-or some approximation scheme like reinforcement learning [4–6].
+or some approximation scheme like reinforcement-learning [4–6].
 The free-energy formulation circumvents these problems by
@@
 What single brain function could be most important for existence and survival of the species?
+Estimating quantities that depend on how features of an agent’s environment are expected to unfold over the future.
 From the perspective of the human condition, it remains unclear what the nature and purpose of a domain-overarching computational mechanism implemented in the high association cortex could be to warriant the high energetic costs.
+
+Obtain reward and avoid pain.
+
+The idealized computational framework explains previous observations about the
+DMN in a simple in conception but non-trivially links
+together disparate functional hypotheses in the DMN literature.
+
+biological plausibility of dynamic optimization in
+optimal substructure
 

### 2016-10-05 paper/paper.tex

--- 2016-10-04 paper/paper.tex
@@
 \end{equation}
-which is a fixed-point equation for the deterministic policy $\pi$.
+which provides a recursive decomposition of optimal behavior.
+A complicated dynamic programming optimization can be
+broken into simpler subproblems at different time points.
+It is a fixed-point equation for the deterministic policy $\pi$.
 Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which

### 2016-10-06 paper/paper.tex
completed RL-PC passage

@@
 orchestrating adapted behavior by bottom-up-driven
-processing of “what matters now”.
+processing of “what matters now”,
+probably drawing on model-based value representations
+\cite{doherty2015structure}.
 Quantative lesion findings across 344 human individuals confirmed
@@
 Involvement of this DMN node was repeatedly reported in
+multi-step action execution \cite{hartmann2005takes},
 visuo-proprioceptive conflict (Balslev et al., 2005) and
@@
 supramodal prediction and error signaling.
+It was argued that the RTPJ encodes actions and ensuing outcomes
+without necessarily relating those to outcome value
+\cite{liljeholm2013neural, hamilton2008action,
+jakobs2009effects}.
 Neural activity in the RTPJ has been argued to be responsible 
@@
 In the definition of $G^\pi$ above, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}.
-The choice $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their immediate rewards and have no regards for the future.
-For there to be any learning at all, a technical requirement is that $\gamma < 1$.
+The choice $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their immediate rewards and has no horizon
+for the future. For there to be any learning at all, a technical requirement is that $\gamma < 1$.
 % \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
@@
 broken into simpler subproblems at different time points.
+Using the Bellman equation, each state can be associated with a certain value,
+to guide action (i.e., improve the policy) towards a better state.
 It is a fixed-point equation for the deterministic policy $\pi$.
@@
   future scenarios and action outcomes.
+  Our approach acknowledges the unavoidable stochasticity of
+  computation in neuronal processes \cite{faisal2008noise}.
   \item
@@
 \paragraph{Predictive Coding and Bayesian Brain Hypothesis}
-The predictive coding framework 
-is one of the most frequently evoked hypotheses of default mode function.
-Cortical responses emerge from a functional network of continuously interacting high and low hierarchical levels,
-with a dynamic intercourse that performs inference on the world by reconciling gaps between sensory input and prior information. 
+The predictive coding and Bayesian brain framework 
+is a frequently evoked idea related to the default mode function
+\cite{bar2007}.
+Cortical responses emerge from a functional network of continuously interacting high and low hierarchical levels.
+A dynamic intercourse performs inference on the world by reconciling gaps between sensory input and prior information. 
 Feed-forward sensory experience is constantly modulated by
 modulatory top-down modulation at various hierarchical processing levels.
+It explains the constructive, generative nature of sensory perception
+\cite{friston2010free} and
+why motor action is intimately related to sensory expectations
+\cite{wolpert1995internal}.
+and explains
 At each stage of processing abstraction level,
@@
 back projection outnumber by far the input projection in the brain
-Salin, P.A. and Bullier, J. (1995) Corticocortical connections in the visual system: structure and function. Physiol. Rev. 75, 107–154.
+\cite{salin1995corticocortical}.
 This process permits updates of beliefs about the world
-to best fit the incoming sensory examples (Friston2014).
+to best fit the incoming sensory examples.
 The generative model of how data are generated in the
@@
   neurobiological evidence, that
-  view the brain as a \"statistical organ\" (Friston 2014).
+  view the brain as a \"statistical organ\"
+  that generalizes from the past to new experiences (Friston 2014).
   \item Both have a notion of brain plasticity with increased
@@
 
+Navigation processes require associative processing to link land- marks with spatial locations, episodic memories require associa- tions for binding the constituents of an episode, emotional processing and the concomitant heightened arousal they entail are tightly associated with affective memories, and our percep- tion of ourselves and others require associations formed with experience to help interpret the present and anticipate the future.
+In some respects, predictions may be considered a higher level of thought-element than associations; they also provide the basis to the many processes detailed above as occupying the unconstrained mind (e.g., planning, navigating, anticipating the thoughts and actions of others, stereotyping, imagining our- selves in the future, affective forecasting, etc.).
+The brain seems to be able to extract regularities and familiar patterns from our environment, and associate them with corresponding knowl- edge in memory to afford the generation of predictions.
+This idea of predictions as the basis of many, if not most, cognitive processes in the human brain will have to be developed sepa- rately because the observations we described here cannot speak to it directly
+Bar2007
+
 as showing that animals could learn a “cognitive map of the environment” in the absence of rewards or penalties, and that they could use the map later when they were motivated to reach a goal (Tolman, 1948).
@@
 From the perspective of the human condition, it remains unclear what the nature and purpose of a domain-overarching computational mechanism implemented in the high association cortex could be to warriant the high energetic costs.
-
-Obtain reward and avoid pain.
+MDP conceptualize control of the environment.
+Obtain reward and avoid pain by adaptive behavior.
 
@@
 together disparate functional hypotheses in the DMN literature.
-
+The default brain function may be reducible to a small number
+of well-defined equations.
 biological plausibility of dynamic optimization in
 optimal substructure
-
+mathematical formalism; statistical solution; adaptive fitness
 ; formalize and predict; offer statistical control on the brain's default function/operations
-
+stochastic modelling
 At the least, we propose an alternative vocabulary to

### 2016-10-07 paper/paper.tex
cleanup RL-PC section

@@
 \cite{bar2007}.
-Cortical responses emerge from a functional network of continuously interacting high and low hierarchical levels.
-A dynamic intercourse performs inference on the world by reconciling gaps between sensory input and prior information. 
-Feed-forward sensory experience is constantly modulated by
-modulatory top-down modulation at various hierarchical processing levels.
-It explains the constructive, generative nature of sensory perception
+According to this view,
+cortical responses emerge from continuous functional interaction between
+higher and lower levels of the neural processing hierarchy.
+This dynamic intercourse enables inference on the world by reconciling
+gaps between fresh sensory input and stored prior information. 
+Feed-forward sensory experience is constantly calibrated by
+top-down modulation at various hierarchical processing levels.
+back projections probably outnumber by far the existing input projection in the brain
+\cite{salin1995corticocortical}.
+It can explains the constructive, generative nature of sensory perception
 \cite{friston2010free} and
@@
 \cite{wolpert1995internal}.
-and explains
-At each stage of processing abstraction level,
+At each stage of neural processing,
 an internally generated prediction of the sensory input is
-directly compared against actual external input.
-A prediction error, i.e. the difference between sensory input
-and internal prediction,
-occurring at one of the processing levels
+directly compared against the actual input from the environment.
+A prediction error at one of the processing levels,
+i.e. the difference between sensory input
+and internally predicted sensation,
 incurs plasticity changes of neuronal back-projections (i.e., model parameters)
 for gradually improved future prediction of the environment.
-Indeed,
-back projection outnumber by far the input projection in the brain
-\cite{salin1995corticocortical}.
-This process permits updates of beliefs about the world
-to best fit the incoming sensory examples.
-The generative model of how data are generated in the
-world would hence be incorporated by
-the current neuronal wiring.
 Contextual integration is maintained by top-down modulation of sensorimotor
 processing by context-specific a-priori information.
-formulate perception as a constructive process
-based on internal or generative models.
+The generative model of how data arise in the
+world hence be incorporated into
+the current neuronal wiring.
+Indeed,
+This process permits updates of the internal model of the environment
+to best fit the incoming sensory examples.
 
 \begin{itemize}
-  \item Both accounts are process theories, supported by
-  neurobiological evidence, that
+  \item Both accounts are process theories backed up by plausible
+  neurobiological evidence that
   view the brain as a \"statistical organ\"
-  that generalizes from the past to new experiences (Friston 2014).
-  \item Both have a notion of brain plasticity with increased
+  generalizing from the past to new experiences.
+  \item Both expose a mechanisms of brain plasticity with increased
   adaption of neuronal wiring when faced by unprecedented environmental
   challenges.
-  \item hierarchical aspects from predictive coding in RL lies in the
-  nested prediction of probable upcoming actions and rewards
-  \item Both model consequences of action. In RL, the horizon of that
-  look into the future is manifest in the $\gamma$ parameter, but not
-  explicitly modeled in the PC account.
-  \item mismatch negativity of PC is immanent in reward prediction error
+  \item The hierarchical aspects from predictive coding
+  is re-expressed in MDP in form of
+  nested prediction of probable upcoming actions and rewards.
+  \item Both model the consequences of action. In RL, the horizon of that
+  look into the future is manifest in the $\gamma$ parameter
+  in the Bellman equation, but not
+  explicitly modeled in the PC account that
+  mostly emphasizes prediction about the
+  immediate future.
+  \item Mismatch negativity of PC is immanent in reward prediction error
   by the difference between actually predicted reward of an action given
   a state and the reward predicted by non-linear regression with graded
-  discounting of future rewards
+  discounting of future rewards.
   \item Sensory experience is a generative process from both views.
-  In predictive coding sensory perception of the external world
+  In predictive coding, sensory perception of the external world
   is a generative experience due to modulatory top down experience at
   various hierarchical levels of sensory input processing.
-  In RL, the environmental model is incorporated in the Markov decision
+  In our RL view, the (partially observed)
+  environmental model is incorporated in the Markov decision
   process, \textit{which can be fully recovered based on the last
@@
   novel stimuli are encountered.
-  Similarly, the both propose explanations why
+  Similarly, both propose explanations why
   predicting the future is inherently
-  linked to information from the past; similarly for independence across
-  geographical spaces.
+  linked to information from the past.
 \end{itemize}

### 2016-10-07 paper/paper.tex
RL-semantics part 1

@@
 was related to state encoding in OFC.
-Such a “cognitive map” of action space was argued to encode
+Such a \textit{cognitive map} of action space was argued to encode
 the current task state even when states are unobservable from sensory input,
@@
 A complicated dynamic programming optimization can be
-broken into simpler subproblems at different time points.
+broken into simpler subproblems at different time points,
+which has often been closely related to the medial prefrontal part of the DMN
+\cite{koechlin1999role, braver2002role}.
 Using the Bellman equation, each state can be associated with a certain value,
@@
 
-\section{Relation to other cognitive models (Danilo)}
-\paragraph{Predictive Coding and Bayesian Brain Hypothesis}
+\section{Relation to other cognitive models}
+\subsection{Predictive Coding and Bayesian Brain Hypothesis}
 The predictive coding and Bayesian brain framework 
@@
 
-\paragraph{Semantic Hypothesis}
-
-The proposal promoted here is that the brain is continually engaged in generating predictions, and that these predictions rely on associative activation. 
-what people do when their mental capacity is not completely consumed by a specific task is to generate associations
-Traditionally this process has been considered as recognition, classification, or even a type of memory retrie- val, but in the present context I treat this process as analogy- making instead. 
-in analogy, the emphasis is on ‘what is it like?
-Analogy- based mappings of properties manifest themselves in processes ranging from perception and memory [17] to stereotypic judgments and prejudice
-While our existing memories are used to derive analo- gies and activate predictions, they are constantly being updated.
-The analogical process, in addition to affording the interpretation of our environment, subsequently aug- ments previous representations in a way that fosters increasingly flexible future analogies.
-(Bar)
-
-phantastic organ (from Greek phantastikos, the ability to create mental images)
-Friston2014
-
-Navigation processes require associative processing to link land- marks with spatial locations, episodic memories require associa- tions for binding the constituents of an episode, emotional processing and the concomitant heightened arousal they entail are tightly associated with affective memories, and our percep- tion of ourselves and others require associations formed with experience to help interpret the present and anticipate the future.
-In some respects, predictions may be considered a higher level of thought-element than associations; they also provide the basis to the many processes detailed above as occupying the unconstrained mind (e.g., planning, navigating, anticipating the thoughts and actions of others, stereotyping, imagining our- selves in the future, affective forecasting, etc.).
-The brain seems to be able to extract regularities and familiar patterns from our environment, and associate them with corresponding knowl- edge in memory to afford the generation of predictions.
-This idea of predictions as the basis of many, if not most, cognitive processes in the human brain will have to be developed sepa- rately because the observations we described here cannot speak to it directly
-Bar2007
-
-as showing that animals could learn a “cognitive map of the environment” in the absence of rewards or penalties, and that they could use the map later when they were motivated to reach a goal (Tolman, 1948).
-
-this perspective can be traced back to Hermann von Helmholtz and the notion of unconscious inference: that is, a pre-rational mechanism by which visual impressions are formed (eg, the seemingly automatic but erroneous belief that the sun rises and sets in the sky, as opposed to the truth that the Earth rotates around it).
-
-mem- ory-based predictions, and these predictions are generated continually either based on gist information gleaned from the senses or driven by thought. The emphasis in this.
-The proposed account integrates three primary components.
-The first is associations, which are formed by a lifetime of extracting repeating patterns and statisti- cal regularities from our environment, and storing them in memory.
-The second is the concept of analogies, whereby
-Corresponding author: Bar, M. (bar@nmr.mgh.harvard.edu). Available online 4 June 2007.
-we seek correspondence between a novel input and existing representations in memory (e.g.
-‘what does this look like?’).
-Finally, these analogies activate associated representa- tions that translate into predictions
-(Bar)
-
-A second important functional account of the DMN is the 'semantic hypothesis'.
-It holds that retrieval of building blocks of world knowledge enables language and other high-level faculties.
-The DMN central involvement in semantic processing results in its implication in many a-priori unrelated aspects of human behavior, such as reasoning, planning, problem solving and decision making (Binder, Desai, Graves, \& Conant, 2009; Schilbach et al., 2008; Suddendorf \& Corballis, 2007).
-This is also illustrated by the substantial relationship between semantic knowledge and words (Binder et al., 2009).
-The engagement of DMN in language processing matches the observation that the large majority of the DMN regions such as TPJ receive no direct input from primary sensory areas acting more as a “hermeneutic hub” (Binder et al., 2009; Seghier, 2013).
-The DMN activation rather enables the organization and processing of internally generated information in absence of stimulus presentation (Binder et al., 1999).
-The proposed role of the DMN in semantic processing is further highlighted by the convergence of these regions in a reciprocal relationship as an interface between the allocentric (bird’s eye view) and egocentric (first person view) perspective; in this way self-reflection, social knowledge, and autobiographical memories are integrated.
-Moreover, the semantic account exquisitely explains the engagement of DMN in both recalling past and envisioning future (Demblon, Bahri, \& D'Argembeau, 2016; Hassabis, Kumaran, Vann, \& Maguire, 2007; Schacter et al., 2007; Spreng, Mar, \& Kim, 2009), in that the internally stored in language world knowledge may lay the contextual foundations for envisioning future events as well as memory processes (Bzdok et al., 2012).
-Hence, the reported retrograde amnesia and the simultaneous impairment of imagining novel experience across patients with hippocampus lesions ties in well with this model (Hassabis \& Maguire, 2007).
-In other words, DMN activation might underlie a time dimension independent cognitive process of constructing contextual scenarios out of detached memories (Buckner \& Carroll, 2007; Gaesser, Spreng, McLelland, Addis, \& Schacter, 2013; Hassabis \& Maguire, 2007).
-Summing up the 'semantic account', DMN offers the conceptual framework that underlies not only language but also other major high-level functions of human behavior.
+\subsection{Semantic Hypothesis}
+Another frequently proposed account of DMN function revolves
+around forming meaning associations and abstract analogies between
+the currently experienced environment and
+stored knowledge from the past
+\cite{bar2007proactive, binder1999conceptual}.
+Analogies might naturally tie incoming novel sensory stimuli to
+explicit world knowledge stored in past memories
+\cite{bar2009proactive}.
+Using these semantic analogies to mentally envision scenarios
+would improve optimal behavior in the environment
+integrated as new memories for future analogical associations.
+Such cognitive processes can afford
+the internal generation of necessary information
+that is not presented in the surrounding environment.
+This accounts further explains the reciprocal relationship
+between an egocentric first person perspective and
+an allocentric bird’s eye perspective immersed in
+self-reflection, social knowledge, and autobiographical memories.
+%
+Closely related to episodic memory, language, problem solving,
+planning, others' thoughts, and spatial navigation
+as these necessitate abstract world knowledge and abstract assocations
+for binding the constituent elements in mental scene construction
+\cite{schacter2007remembering}.
+These scene generation processes could contribute to interpreting the
+present and foretelling the future.
+This view is supported by evidence in animals that
+could learn a \textit{cognitive map of the environment},
+even without reward incentives, and exploit it later
+for other means \cite{tolman1948cognitive}.
 
@@
   history.
-  \item Both: there is no evidence to indicate that predictions of various levels of complexity, abstraction, timescale and purpose use mech- anisms that are qualitatively different
+  \item Both: there is no evidence to indicate that predictions of various
+  levels of complexity, abstraction, timescale and purpose
+  use mechanisms that are qualitatively different. This concurs with
+  DMN activity increases across time, space, and content domains
+  demonstrated in many neuroimaging studies. The semantic hypothesis
+  and RL account provide explanations why hippocampus lesion does
+  not only impair retrieving memories, but also hypothetical and future
+  thinking \cite{hassabis2007patients}.
+  \item The semantic hypothesis and MDP expose mechanisms of
+  how alternative decisions could be mentally explored.
+  \item While mental scene construction entails a distinction between
+  environmental perception and internally generated mind-wandering
+  \footnote{``A computational model of how such a process might be structured
+  is far from being defined, but it will probably require a form of
+  regulation by which perception of the current world is suppressed
+  while simulation of possible alternatives are constructed,
+  followed by a return to perception of the present.''\cite{buckner2007self}},
+  the MDP naturally integrates the former egocentric
+  (related to the current action, state and reward) and the later
+  allocentric (related to past and future actions, states, and rewards)
+  angles on the world in a same optimization problem.
+  \item The semantic account of DMN function does not offer 
+  a clear explanation how semantic world knowledge and analogies thereof
+  lead to prediction about future actions and states.
+  \item While both semantic and MDP account propose memory-based internally
+  generated information for probabilistic mental models of action outcomes,
+  only MDPs render explicit the grounds on which the final action is
+  actually chosen (i.e., the estimated cumulative reward).
+  \item In sum, the semantic hypothesis lacks an explicit time and
+  incentive model.
 \end{itemize}
 
-\paragraph{Sentinel Hypothesis}
+\subsection{Sentinel Hypothesis}
 many investigators speculate that this central neural network reflects the brain’s relentless tracking and prediction of environmental events to adaptively optimize the organism's future action (Schacter et al., 2007).
@@
 Cognitive neuroimaging studies linked the DMN to the contemplation of others’ (Mar, 2011) and one’s thoughts (Lombardo et al., 2009) mental states, 
-
+Schilbach et al., 2008
+Clan mentality
 DMN does social cognition -> but it exists in rats and monkeys with much more poorly developped social capacities
@@
 Obtain reward and avoid pain by adaptive behavior.
-
+How representations of past and future are integrated into
+action choice in the present context.
 The idealized computational framework explains previous observations about the

### 2016-10-09 paper/paper.tex
even more stuff

@@
 the more mental activity becomes detached from the actual sensory environment
-and the more stimulus-independent thoughts occur
-(Mason et al., 2007).
+\cite{antrobus1966studies, pope1978regulation}.
+For instance,
+the more the currently executed task is unknown and unpracticed,
+the lower the tendency for stimulus-independent thoughts
+\cite{filler1973daydreaming, teasdale1995stimulus}.
 These \"offline\" processes may however contribute to optimized control of the organism.
@@
 A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
-%% In RL, MDPs are the mathematical instantiation of the environment model,
-%% which generally can incorporate action probabilities (policy matrix)
-%% and/or value function given a state.
-%% (Silver: The environment can be fully defined by state space and
-%% transition probability matrix.)
+Accorind to this model, the environment’s dynamics are fully determined by
+its current state.
+In RL, MDPs are the mathematical instantiation of the environment model,
+which generally can incorporate action probabilities (policy matrix)
+and/or value function given a state.
+(Silver: The environment can be fully defined by state space and
+transition probability matrix.)
 is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
@@
 
+\paragraph{TODO: Monte Carlo optimization.}
+Some text snippets that might be pertinent:
+Monte-Carlo simulation can be used to update the agent’s belief state
+\cite{silver2010monte}.
+Monte-Carlo simulation provides a simple method for evaluating
+a state s.
+In any MDP, there is at least one optimal policy matrix that
+achieves the optimal value function???
+The simulator is used to generate sequences of states, observations and rewards.
+These simulations are used to update the value function, without ever looking inside the black box describing the model’s dynamics.
+In addition, Monte-Carlo methods have a sample complexity that is determined only by the underlying difficulty of the POMDP, rather than the size of the state space or observation space.
+Monte-Carlo simulation can be turned into a simple control algorithm by evaluating all legal actions and selecting the action with highest evaluation [15].
+Monte-Carlo simulation can be extended to partially observable MDPs.
+In these challenging POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
+Link this to experience sampling in the hippocampus.
+
 \paragraph{Link to classical RL algorithms.}
@@
   choosing actions to be executed in a certain situation.
+  \item Interactions between external and internal sources of
+  information are probably commonplace during perception
+  \item The MDP frameworks how the human brain could solve
+  environmental problems over extended time periods.
   \item Given a state, the value function represents the estimated utility of
@@
 \section{Relation to other cognitive models}
-\subsection{Predictive Coding and Bayesian Brain Hypothesis}
+\subsection{Predictive Coding Hypothesis}
 The predictive coding and Bayesian brain framework 
@@
   state alone}.
+  \item Adapting the neuronal connections for improved top-down modulation
+  takes the concrete form of gradient computation
+  \textit{and back-propagation} in MDPs and RL.
   \item Both provide a parsimoneous explanation why
@@
 Another frequently proposed account of DMN function revolves
-around forming meaning associations and abstract analogies between
-the currently experienced environment and
-stored knowledge from the past
+around forming meaning associations and analogies between
+the currently experience and
+abstract knowledge derived from past experiences
 \cite{bar2007proactive, binder1999conceptual}.
 Analogies might naturally tie incoming novel sensory stimuli to
-explicit world knowledge stored in past memories
+explicit world knowledge extractable from past memories
 \cite{bar2009proactive}.
-Using these semantic analogies to mentally envision scenarios
-would improve optimal behavior in the environment
-integrated as new memories for future analogical associations.
+%
+These semantic analogies can be used to mentally envision non-existent scenarios
+that would improve optimal behavior in the environment
+by simulating future analogical associations.
 Such cognitive processes can afford
 the internal generation of necessary information
-that is not presented in the surrounding environment.
-This accounts further explains the reciprocal relationship
+that is not presented in the surrounding environment
+by recombining building blocks of
+concept knowledge and episodic memories.
+Indeed, in aging humans, remembering the past and imaging the future
+equally decreased in the level of detail and are associated with
+concurrent deficits in forming and integrating relationships between
+items \cite{addis2008age, spreng2006temporal}.
+A constructive account further explains the reciprocal relationship
 between an egocentric first person perspective and
@@
 %
-Closely related to episodic memory, language, problem solving,
-planning, others' thoughts, and spatial navigation
+It is also closely related to episodic memory, language, problem solving,
+planning, estimating others' thoughts, and spatial navigation
 as these necessitate abstract world knowledge and abstract assocations
@@
 present and foretelling the future.
-This view is supported by evidence in animals that
+This view is for instance supported by evidence in animals that
 could learn a \textit{cognitive map of the environment},
@@
   \item The notion of semantic or knowledge association is
-  incorporated into the MDP as a function of the state and action
-  history.
-  \item Both: there is no evidence to indicate that predictions of various
+  incorporated into the MDP as the Markov property,
+  that is, the current state directly results from the
+  agent's history of states and actions. The learned
+  value matrix and action transition schedules drive
+  stimulus processing and action choice in the present.
+  \item In both, semantic hypothesis and MDPs,
+  there is no evidence to indicate that predictions of various
   levels of complexity, abstraction, timescale and purpose
@@
   thinking \cite{hassabis2007patients}.
-  \item The semantic hypothesis and MDP expose mechanisms of
-  how alternative decisions could be mentally explored.
-  \item While mental scene construction entails a distinction between
+  \item Both the semantic hypothesis and MDP expose mechanisms of
+  how alternative decision trees could be mentally explored.
+  \item While mental scene construction implies a distinction between
   environmental perception and internally generated mind-wandering
@@
   the MDP naturally integrates the former egocentric
-  (related to the current action, state and reward) and the later
+  (related to current action, state and reward) and the later
   allocentric (related to past and future actions, states, and rewards)
@@
   \item The semantic account of DMN function does not offer 
-  a clear explanation how semantic world knowledge and analogies thereof
-  lead to prediction about future actions and states.
+  a mechanistic explanation how semantic world knowledge and analogies thereof
+  lead to prediction of future actions and states.
+  \item The semantic hypothesis does not explain why memory recall
+  for scene construction in humans is fragmentary and noisy
+  instead of accurate and reliable.
+  Yet, the MDP framework provides an algorithmic explanation in that
+  stochasticity of the parameter space search implemented
+  by Monte Carlo solvers provably yields better models of the world.
+  That is, the purposeful stochasticity of policy and value estimation
+  in MDPs provides a candidate explanation for why humans
+  have evolved imperfect memories as the better adaptation.
   \item While both semantic and MDP account propose memory-based internally
@@
   actually chosen (i.e., the estimated cumulative reward).
-  \item In sum, the semantic hypothesis lacks an explicit time and
-  incentive model.
+  \item In sum, episodic scene construction according to the semantic
+  hypothesis lacks an explicit time and incentive model.
 \end{itemize}
@@
 many investigators speculate that this central neural network reflects the brain’s relentless tracking and prediction of environmental events to adaptively optimize the organism's future action (Schacter et al., 2007).
+
+The contents of memory and prospection may be nec- essarily and adaptively tied to things social in nature, po- tentially the most behaviorally relevant stimuli.
+As such, the evolutionary advancement of autonoetic awareness would correspond to the neural processes of theory of mind.
+The utilization of prospection is arguably of greatest strategic importance, particularly in the secur- ing of mates, protection, and resources (Suddendorf \&
 
@@
 Summing up this first candidate account for DMN function, the sentinel hypothesis heavily emphasizes on the DMN as a single structure with a continuous sentinel role increasing during self-referential tasks.
+
+ \begin{itemize}
+   \item The sentinel account does not provide a formal mechanisms of
+   how attention and memory resources are exactly reallocated when
+    encountering a salient environmental stimulus. In contrast,
+    the Bellman recursion is weighted at each step by
+    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
+    are recursively explorated and which decision trees are neglected,
+    which implies allocation of attention and computational load.
+ \end{itemize}
 
@@
 mathematical formalism; statistical solution; adaptive fitness
+to optimize the organism's fate; contribute to human survival and
+creativity.
 ; formalize and predict; offer statistical control on the brain's default function/operations

### 2016-10-09 paper/paper.tex
a first conclusion

@@
 Link this to experience sampling in the hippocampus.
+Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
 
@@
 \section{Conclusion}
-What single brain function could be most important for existence and survival of the species?
-Estimating quantities that depend on how features of an agent’s environment are expected to unfold over the future.
-From the perspective of the human condition, it remains unclear what the nature and purpose of a domain-overarching computational mechanism implemented in the high association cortex could be to warriant the high energetic costs.
-MDP conceptualize control of the environment.
-Obtain reward and avoid pain by adaptive behavior.
-How representations of past and future are integrated into
-action choice in the present context.
-The idealized computational framework explains previous observations about the
-DMN in a simple in conception but non-trivially links
-together disparate functional hypotheses in the DMN literature.
-The default brain function may be reducible to a small number
-of well-defined equations.
-biological plausibility of dynamic optimization in
-optimal substructure
-mathematical formalism; statistical solution; adaptive fitness
-to optimize the organism's fate; contribute to human survival and
-creativity.
-; formalize and predict; offer statistical control on the brain's default function/operations
-stochastic modelling
-At the least, we propose an alternative vocabulary to
-describe and discuss neuroscientific studies on the DMN.
+What brain function could be important enough
+for the existence and survival of the human species
+to warrant constant, high energy costs?
+MDPs provide an attractive
+computational model how the human association cortex
+might implement high-level control of the environment to
+optimize the organism's fate by obtaining reward and avoiding pain.
+This idealized process model explains
+a number of previous experimental observations in the
+DMN literature by a simple but non-trivial concept.
+%
+Viewed as a MDP, behavior unfolds by integrating across past
+and future experience to guide action choice in the present context.
+This account is more compatible with the DMN's involvement across
+autobiographical memory retrieval,
+abstract reasoning based on internally generated scenes,
+and self-projection into the expected future.
+MDPs provide a mathematical formalism how
+optimal substructure can be recursively exploited
+when contemplating nested decisions.
+Algorithmic gains by injecting stochasticity in past
+behavior and outcomes may plausibilize why
+very accurate memories have been evolutionarily de-selected
+and why human creativity might be adaptive.
 %
@@
 the set of action, value, and state variables that determine
-the behavior of RL systems. This makes proposed machine-learning
+the behavior of RL systems. This makes the proposed machine-learning
 approach to DMN function not only practically computable but
 also neuroscientifically falsifiable.
-%
-The DMN can be viewed as a smoothing kernel extending
+At the least, we propose an alternative vocabulary to
+describe and discuss neuroscientific studies in DMN research.
+%
+Ultimately,
+the DMN can be viewed as a smoothing kernel extending
 from the relatively recent past to the relatively close future
-that constantly estimates complex non-linear gain functions.
--> This is why neuroimaging research revealed activity increases in the DMN
-when 
--> This may be why perfect memory has been evolutionarily de-selected,
-although it is provably feasible by the neurobiological hardware in humans
+to constantly compute complex gain opportunities in the world.
 

### 2016-10-09 paper/paper.tex
finished RL-semantics

@@
 Formal models of the computational processes subserved in the DMN
-could offer statistical control of predictive human behavior.
+could offer access routes to statistical optimality
+of human behavior.
 
@@
 Specific spiking sequences, that were elicited in experimental task conditions,
-have been shown to be reenacted spontaneously during sleep.
+have been shown to be reenacted spontaneously during
+quiet wakefulness and sleep \cite{hartley2014space}.
 Moreover, spike sequences measured in hippocampal place cells of rats
@@
 and neuropsychology,
-the LTPJ probably mediates access to and integrating of world knowlege.
+the LTPJ probably mediates access to and integrating of world knowlege
+to model action concepts
+\cite{binder2011neurobiology, seghier2013angular}.
 For instance, LTPJ lesions also entail problems in recognizing
@@
 Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
+For sampling-based methods it is still almost non-existent beyond the description of receptive fields.
 
@@
 %
-These semantic analogies can be used to mentally envision non-existent scenarios
+Going much beyond language ifself,
+semantic analogies can be used to mentally envision non-existent scenarios
 that would improve optimal behavior in the environment
@@
   lead to prediction of future actions and states.
+  \item In contrast to existing accounts on semantics and
+  mental scene construction, the random and creative aspects of DMN function
+  are here directly related to the advantages of stochastic optimization.
   \item The semantic hypothesis does not explain why memory recall
@@
 computational model how the human association cortex
-might implement high-level control of the environment to
+might implement supra-modal control of the environment to
 optimize the organism's fate by obtaining reward and avoiding pain.
@@
 This account is more compatible with the DMN's involvement across
-autobiographical memory retrieval,
-abstract reasoning based on internally generated scenes,
-and self-projection into the expected future.
+autobiographical memory retrieval, problem solving,
+abstract reasoning based on internally generated scenes, social cognition,
+as well as delay discounting and self-related prospection.
 MDPs provide a mathematical formalism how
 optimal substructure can be recursively exploited
-when contemplating nested decisions.
+when confronted with nested decisions.
 Algorithmic gains by injecting stochasticity in past
@@
 the set of action, value, and state variables that determine
-the behavior of RL systems. This makes the proposed machine-learning
-approach to DMN function not only practically computable but
+the behavior of intelligent RL systems.
+The proposed machine-learning
+approach to DMN function is not only practically computable but
 also neuroscientifically falsifiable.
@@
 the DMN can be viewed as a smoothing kernel extending
-from the relatively recent past to the relatively close future
-to constantly compute complex gain opportunities in the world.
+from the relatively recent events to potentially upcoming events
+to constantly improve sensory input and peformed action in the world.
 

### 2016-10-10 paper/paper.tex
makeup for the paper finish

@@
 \cite{bar2007}.
-According to this view,
+According to this functional account,
 cortical responses emerge from continuous functional interaction between
@@
 top-down modulation at various hierarchical processing levels.
+In fact,
 back projections probably outnumber by far the existing input projection in the brain
 \cite{salin1995corticocortical}.
-It can explains the constructive, generative nature of sensory perception
+This can explains the constructive, generative nature of sensory perception
 \cite{friston2010free} and
@@
 Contextual integration is maintained by top-down modulation of sensorimotor
-processing by context-specific a-priori information.
+processing by a-priori information about the environment.
 The generative model of how data arise in the
-world hence be incorporated into
+world would hence be incorporated into
 the current neuronal wiring.
@@
 \begin{itemize}
-  \item Both accounts are process theories backed up by plausible
+  \item Both the predictive coding account and MPDs
+  are process theories backed up by plausible
   neurobiological evidence that
-  view the brain as a \"statistical organ\"
+  view the brain as a ``statistical organ''
   generalizing from the past to new experiences.
-  \item Both expose a mechanisms of brain plasticity with increased
-  adaption of neuronal wiring when faced by unprecedented environmental
+  \item Both expose a mechanisms of brain plasticity in that 
+  neuronal wiring gets increasingly adapted
+  when faced by unprecedented environmental
   challenges.
   \item The hierarchical aspects from predictive coding
-  is re-expressed in MDP in form of
+  is re-expressed in MDPs in form of
   nested prediction of probable upcoming actions and rewards.
   \item Both model the consequences of action. In RL, the horizon of that
-  look into the future is manifest in the $\gamma$ parameter
+  look into the future is manifested in the $\gamma$ parameter
   in the Bellman equation, but not
@@
   takes the concrete form of gradient computation
-  \textit{and back-propagation} in MDPs and RL.
-  \item Both provide a parsimoneous explanation why
-  humans seem to minimize processing of incoming information
-  when the environment is predictable and allocate attention when
-  novel stimuli are encountered.
+  \textit{and back-propagation} in MDPs and RL,
+  although this may not be neurobiologically plausible.
+  \item Both provide a parsimonious explanation why the
+  human brain decreases processing load devoted to incoming information
+  when the environment is predictable and allocate increased
+  attentional resources when novel stimuli are encountered.
   Similarly, both propose explanations why
@@
 Another frequently proposed account of DMN function revolves
-around forming meaning associations and analogies between
-the currently experience and
-abstract knowledge derived from past experiences
+around forming meaningful associations and analogies between
+the current experience and
+the abstract knowledge derived from past experiences
 \cite{bar2007proactive, binder1999conceptual}.
 Analogies might naturally tie incoming novel sensory stimuli to
-explicit world knowledge extractable from past memories
+explicit world knowledge extractable from the environment
 \cite{bar2009proactive}.
@@
 that would improve optimal behavior in the environment
-by simulating future analogical associations.
+by simulating future events.
 Such cognitive processes can afford
@@
   stimulus processing and action choice in the present.
-  \item In both, semantic hypothesis and MDPs,
+  \item In both semantic hypothesis and MDPs,
   there is no evidence to indicate that predictions of various
@@
   \item While mental scene construction implies a distinction between
-  environmental perception and internally generated mind-wandering
-  \footnote{``A computational model of how such a process might be structured
+  environmental perception and internally generated mind-wandering\footnote{
+  ``A computational model of how such a process might be structured
   is far from being defined, but it will probably require a form of
@@
   mental scene construction, the random and creative aspects of DMN function
-  are here directly related to the advantages of stochastic optimization.
+  are explained in MDPs by the advantages of stochastic optimization.
   \item The semantic hypothesis does not explain why memory recall
-  for scene construction in humans is fragmentary and noisy
+  for scene construction in humans is typically fragmentary and noisy
   instead of accurate and reliable.
@@
   only MDPs render explicit the grounds on which the final action is
-  actually chosen (i.e., the estimated cumulative reward).
+  eventually chosen (i.e., the estimated cumulative reward).
   \item In sum, episodic scene construction according to the semantic
-  hypothesis lacks an explicit time and incentive model.
+  hypothesis is lacking an explicit time and incentive model.
 \end{itemize}
@@
 computational model how the human association cortex
-might implement supra-modal control of the environment to
-optimize the organism's fate by obtaining reward and avoiding pain.
+might implement supra-modal control of the environment by
+obtaining reward and avoiding pain to
+optimize the organism's fate.
 This idealized process model explains
@@
 %
-Viewed as a MDP, behavior unfolds by integrating across past
-and future experience to guide action choice in the present context.
-This account is more compatible with the DMN's involvement across
+Viewed as an MDP, behavior unfolds by integrating happened past events
+and possible future events to guide action choice in the present context.
+This functional account is more compatible with the DMN's
+poorly understood involvement across
 autobiographical memory retrieval, problem solving,
@@
 when confronted with nested decisions.
-Algorithmic gains by injecting stochasticity in past
+Algorithmic gains by injecting stochasticity in the recall of past
 behavior and outcomes may plausibilize why
-very accurate memories have been evolutionarily de-selected
+very accurate memories have been evolutionarily disfavored
 and why human creativity might be adaptive.
@@
 The proposed machine-learning
-approach to DMN function is not only practically computable but
+approach to DMN function is hence not only practically computable but
 also neuroscientifically falsifiable.
 At the least, we propose an alternative vocabulary to
-describe and discuss neuroscientific studies in DMN research.
+describe and interpret neuroscientific studies in DMN research.
 %
@@
 from the relatively recent events to potentially upcoming events
-to constantly improve sensory input and peformed action in the world.
+to constantly improve sensory processing and
+action performance in the world.
 

### 2016-10-12 paper/paper.tex
Starting RL-sentinel

@@
 \subsection{Sentinel Hypothesis}
-many investigators speculate that this central neural network reflects the brain’s relentless tracking and prediction of environmental events to adaptively optimize the organism's future action (Schacter et al., 2007).
-
-The contents of memory and prospection may be nec- essarily and adaptively tied to things social in nature, po- tentially the most behaviorally relevant stimuli.
-As such, the evolutionary advancement of autonoetic awareness would correspond to the neural processes of theory of mind.
-The utilization of prospection is arguably of greatest strategic importance, particularly in the secur- ing of mates, protection, and resources (Suddendorf \&
+Processing self-relevant information was perhaps the first
+functional account that was proposed for the DMN \cite{gusnard2001medial}.
+Since then,
+different investigators have speculated that neural activity in the DMN
+may reflects the brain’s relentless tracking of
+relevance in the environment
+as an advantageous evolutionary adaptation
+\cite{}.
+-> the most behaviorally relevant stimuli
+-> particularly in the secur- ing of mates, protection, and resources
+-> tracking for salient and unexpected events
+
+This coherent neurobiological network would be prominently engaged as a “radar” as well as in self-relevant information processing
+
+Environmental stimuli important for humans are frequently of
+social nature. This is probably unsurprising
+given that a key property of the human species is
+the complexity of their social systems
+\cite{tomasello2009cultural}.
+More specifically,
+according to the ``social brain hypothesis'',
+the human brain did not evolve to solve problems of the
+physical environment, but to form and maintain increasingly complex
+social systems to solve ecological problems in a social way
+\cite{whiten1988machiavellian}.
+Our human ancestors were probably among the few organisms that
+were more likely to be killed by members of their own species
+rather than predators from another species.
+Being able to detect the intention and upcoming actions of conspecific
+turned into a decisive evolutionary advantage.
+
+-> contents of memory and everyday dreaming tied to people and relationships to them
 
@@
 Clan mentality
-DMN does social cognition -> but it exists in rats and monkeys with much more poorly developped social capacities
+
+role increasing during self-referential tasks.
+
+ \begin{itemize}
+    \item DMN does social cognition -> but it exists in rats
+    and monkeys with much more poorly developped social capacities
     -> our computational framework can be equally applied to humans
     and other primates.
-
-How can a same neurobiological circuit be equally important for baseline house-keeping functions and specific task performance?
-
-Humans have learned to solve ecological problems
-by social means.
-
-Our computational account can explain why the DMN is implicated in both a goal-directed task and an idlying rest cognitive set?
--> task is policy/value updates to optmize short-term action / rest updates for mid- and long-term action
-
- The 'sentinel hypothesis' is a major functional account to explain DMN activity that this canonical network acts as a cohesive unit.
-This coherent neurobiological network would be prominently engaged as a “radar” as well as in self-referential information processing including mind wandering, autobiographical recall, mentalizing, reflection about past and future (Buckner, Andrews-Hanna, \& Schacter, 2008; Schilbach, Eickhoff, Rotarska-Jagiela, Fink, \& Vogeley, 2008).
-Acting as a sentinel, DMN functions as a broad focus manager that performs continuous environmental tracking for salient and unexpected events (Buckner et al., 2008).
-This default activity may allow the detection of possible predators and potentially serves an evolutionarily conserved purpose by means of attention allocation (Raichle et al., 2001).
-The “radar” function may be efficiently supported by the activation patterns of DMN in diffuse attention condition tasks with targets appearing in various locations (Hahn, Ross, \& Stein, 2007).
-On the other hand, the salient information perceived from the external environment activates self-referential information processing (Krienen, Tu, \& Buckner, 2010).
-A conceptualization of these functions suggests two highly adaptive interacting DMN subsystems (Andrews-Hanna, Saxe, \& Yarkoni, 2014; Schacter, Addis, \& Buckner, 2007).
-However, the sentinel hypothesis may not exhaustively explain the variety of findings on the DMN connectivity with memory-related structures such as hippocampus (Giugni, Vadala, De Vincentiis, Colica, \& Bastianello, 2010; McCormick et al., 2014).
-Summing up this first candidate account for DMN function, the sentinel hypothesis heavily emphasizes on the DMN as a single structure with a continuous sentinel role increasing during self-referential tasks.
-
- \begin{itemize}
     \item The sentinel account does not provide a formal mechanisms of
@@
     which implies allocation of attention and computational load.
+    \item How can a same neurobiological circuit be equally important for baseline house-keeping functions and specific task performance? Yes, if
+    environmental relevance is processed, manipulated, and retained
+    also as a baseline function of human mental activity.
+    Our computational account can explain why the DMN is implicated in both a goal-directed task and an idlying rest cognitive set?
+-> task is policy/value updates to optmize short-term action / rest updates for mid- and long-term action
  \end{itemize}

### 2016-10-13 paper/paper.tex
Danilo's part quasi-finished

@@
 \usepackage{url}
+\usepackage[round]{natbib}
 \usepackage{amsmath,amsfonts,amsthm}
@@
 
-\author{Elvis Dohmatob, (Bert Kappen), Danilo Bzdok\\
+\author{Elvis Dohmatob, Bert Kappen, Danilo Bzdok\\
   INRIA, Parietal team, Saclay, France\\
@@
 reinforcement learning.
-Neurobiological evidence on the human and animal DMN
+In this way,
+neurobiological evidence on the human and animal DMN
 is linked to recent progress in autonomous systems research
-in the machine-learning domain.
+and machine-learning.
 We argue that this engineering approach to DMN function can parsimoneously
@@
 of human behavior.
-
 \end{abstract}
@@
 % official NIPS keywords
-\textbf{\\keywords}: Systems Biology, mind wandering, cognitive science,
+\textbf{\\keywords}: systems biology, mind wandering, cognitive science,
 autonomous learning systems
-\textbf{\\Get opinion from}: Danielle Bassett, Kai Vogeley, Mohammad Shakir,
-Guillaume Dumas, Smallwood, Tenenbaum
+% \textbf{\\Get opinion from}: Danielle Bassett, Kai Vogeley, Mohammad Shakir,
+% Guillaume Dumas, Smallwood, Tenenbaum
 
@@
 Rather, the brain continues to metabolize large quantities of
-oxygen and glucose energy to maintain inter-neuronal information
-transfer in the absence of a focused behavioral goal (Raichle 2001;
-Shulman, 1999).
+oxygen and glucose energy to maintain
+potentially purposeful neuronal computation
+in the absence of a focused behavioral goal
+\cite{kenet2003spontaneously, fiser2004small}.
 This baseline energy demand is subject to surprisingly little modulations
@@
 across time, splace, and content domains
-interfacing the external world and the self.
-
-Consider an agent faced by choosing the next action
+at the interface of the external world and the self.
+
+Consider an agent faced by the choice of the next action
 and being guided by rich
@@
 \cite{filler1973daydreaming, teasdale1995stimulus}.
-These \"offline\" processes may however contribute to optimized control of the organism.
+These ``offline'' processes may however contribute to optimized control of the organism.
 We formalize
@@
 on the environment given the current state. A value matrix 
-maps maps expected rewards to environmental objects and events.
+maps expected rewards to environmental objects and events.
 Switching between states reduces to a sequential processing model.
@@
 the DMN dynamics are constantly adapted in feed-back loops
-that are calibrated by error occurrence.
+that are calibrated by prediction error.
 A DMN framework incorporating reinforcement learning
@@
 The implication of the DMN in a diversity of human cognitive processes
-can be parsimoneously recast as minimizing prediction error
+can be parsimoneously recast as prediction error minimization
 based on pervasive probabilistic simulations,
@@
 based on a brain implementation of Markov Chain Monte Carlo sampling
-(Tenenbaum2011Science).
+\cite{tenenbaum2011grow}.
 \textit{Control} refers to the influences an agent exterts when interacting
@@
 the predictive DMN machinery to optimize the behavior of the organism.
+%
+In short, it has been proposed earlier that
+the human brain's energy budget is largely dedicated to
+``the development and maintenance of [a]
+probabilistic model of anticipated events''
+\cite{raichle2005intrinsic}.
+This paper proposes a formal model that
+satisfies this contention.
 
@@
 into ongoing action, planning, and problem solving.
-world -> policy matrix driven computation
-self -> value matrix driven computation
+% world -> policy matrix driven computation
+% self -> value matrix driven computation
 
@@
   of what happened in the past.
-
 \end{itemize}
@@
 \cite{bar2007}.
-According to this functional account,
+According to this cognitive account,
 cortical responses emerge from continuous functional interaction between
@@
 why motor action is intimately related to sensory expectations
-\cite{wolpert1995internal}.
+\cite{wolpert1995internal, kording2004bayesian}.
 At each stage of neural processing,
@@
 Processing self-relevant information was perhaps the first
-functional account that was proposed for the DMN \cite{gusnard2001medial}.
+cognitive account that was proposed for the DMN \cite{gusnard2001medial}.
 Since then,
@@
 relevance in the environment
-as an advantageous evolutionary adaptation
-\cite{}.
--> the most behaviorally relevant stimuli
--> particularly in the secur- ing of mates, protection, and resources
--> tracking for salient and unexpected events
-
-This coherent neurobiological network would be prominently engaged as a “radar” as well as in self-relevant information processing
-
-Environmental stimuli important for humans are frequently of
+as an advantageous evolutionary adaptation.
+According to this cognitive account, the brain's baseline realizes
+a ``radar'' function to
+detect subjectively salient and unexpected events that
+are the most important for unfolding behavior,
+including resources,
+the securing of mates, and protection.
+%
+In fact,
+environmental stimuli important for humans are frequently of
 social nature. This is probably unsurprising
@@
 physical environment, but to form and maintain increasingly complex
-social systems to solve ecological problems in a social way
+social systems to solve ecological problems by means of social connections
 \cite{whiten1988machiavellian}.
+Highly efficient neurobiological processing of social cues is exemplified by
+facial judgments being routinely processed in less than 100 ms
+\cite{bar2006very}.
 Our human ancestors were probably among the few organisms that
@@
 rather than predators from another species.
-Being able to detect the intention and upcoming actions of conspecific
-turned into a decisive evolutionary advantage.
-
--> contents of memory and everyday dreaming tied to people and relationships to them
-
-Cognitive neuroimaging studies linked the DMN to the contemplation of others’ (Mar, 2011) and one’s thoughts (Lombardo et al., 2009) mental states, 
-Schilbach et al., 2008
-Clan mentality
-
-role increasing during self-referential tasks.
+Being able to detect the intention and upcoming actions of conspecifics
+turned into a decisive evolutionary advantage
+\cite{frith2010social}.
+%
+This may explain why social topics amount to roughly
+two thirds of human everyday communication \cite{dunbar1997human}
+and
+why mind-wandering and dreams
+are so rich in stories about people and the complex relationships between them.
 
  \begin{itemize}
-    \item DMN does social cognition -> but it exists in rats
-    and monkeys with much more poorly developped social capacities
-    -> our computational framework can be equally applied to humans
-    and other primates.
+    \item Processing social information has been proposed to underlie the
+    physiological baseline of human brain function
+    \cite{schilbach2008minds}. This was later challenged by observing
+    analogues of the DMN in monkeys \cite{mantini2011default}
+    and rats \cite{lu2012rat}, with
+    supposedly less advanced social capacities.
+    The MDPs parsimoneously explain the dominance of social content in
+    human mental scenes by their extremely high relevance to humans
+    reflected by high values for social information in the value matrix.
+    \item How can a same neurobiological circuit be equally important
+    for baseline house-keeping functions and specific task performance?
+    Our computational account can explain why the DMN is implicated
+    in both a goal-directed task and an idlying rest cognitive set,
+    if environmental relevance is processed, manipulated, and retained
+    also as a baseline function of human mental activity.
+    In tasks, the policy and value matrices are updated to optmize short-term action,
+    whereas, at rest, these parameter updates may
+    serve especially mid- and long-term action.
     \item The sentinel account does not provide a formal mechanisms of
@@
     $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
-    are recursively explorated and which decision trees are neglected,
-    which implies allocation of attention and computational load.
-    \item How can a same neurobiological circuit be equally important for baseline house-keeping functions and specific task performance? Yes, if
-    environmental relevance is processed, manipulated, and retained
-    also as a baseline function of human mental activity.
-    Our computational account can explain why the DMN is implicated in both a goal-directed task and an idlying rest cognitive set?
--> task is policy/value updates to optmize short-term action / rest updates for mid- and long-term action
+    are recursively explored by the agend and which decision trees are neglected.
+    This directly implies allocation of attention and computational load.
+    \item In sum,
+    MDPs imply social thinking as a physiological brain baseline
+    as a special case, without excluding other topics.
  \end{itemize}
@@
 \section{Relation to other statistical models (Elvis)}
-\paragraph{Control theory.}
+\subsection{Control theory.}
 \begin{itemize}
@@
 
-\paragraph{Reinforcement Learning.}
+\subsection{Reinforcement Learning.}
 abc
 
-\paragraph{Free Energy Principle.}
+\subsection{Free Energy Principle.}
 \begin{itemize}
@@
 computational model how the human association cortex
-might implement supra-modal control of the environment by
-obtaining reward and avoiding pain to
-optimize the organism's fate.
+might implement supra-modal representation and control of the environment to
+optimize the organism's fate by
+obtaining reward and avoiding pain.
 This idealized process model explains
 a number of previous experimental observations in the
-DMN literature by a simple but non-trivial concept.
+DMN literature by a simple but non-trivial mechanism.
 %
@@
 MDPs provide a mathematical formalism how
-optimal substructure can be recursively exploited
+optimal substructure in the environment can be recursively exploited
 when confronted with nested decisions.
 Algorithmic gains by injecting stochasticity in the recall of past
-behavior and outcomes may plausibilize why
+behavior and outcomes to improve the internal world model may explain why
 very accurate memories have been evolutionarily disfavored
@@
 the DMN can be viewed as a smoothing kernel extending
-from the relatively recent events to potentially upcoming events
+from the relatively recent events to anticipated upcoming events
 to constantly improve sensory processing and
@@
 \small
-\bibliographystyle{splncs03}
+% \bibliographystyle{splncs03}
+% \bibliographystyle{natbib}
+\bibliographystyle{plainnat}
 \bibliography{refs}

### 2016-10-14 paper/paper.tex
fresh stuff

@@
 
-\title{A Computational Account of Default-Mode Function\\
+\title{A Computational Account of Default Mode Function\\
 by Control Theory and Reinforcement Learning}
@@
 Many research streams agree on its likely role in evolutionarily adaptive
-envisioning of past and hypothetical scenarios to predict the environment.
+envisioning of mental scenarios to predict the environment.
 The DMN would hence be dedicated to continuous autobiographical memory retrieval,
@@
 reinforcement learning.
-In this way,
-neurobiological evidence on the human and animal DMN
-is linked to recent progress in autonomous systems research
-and machine-learning.
-We argue that this engineering approach to DMN function can parsimoneously
-explain a variety of existing neuroscientific hypotheses.
+We argue that our formal account of DMN function
+explains existing cognitive accounts as special cases and
+parsimoneously explains a variety of experimental findings
+in animals and humans.
 %
 Formal models of the computational processes subserved in the DMN
-could offer access routes to statistical optimality
-of human behavior.
+could offer access routes to the statistical regularities
+of human predictive behavior.
 \end{abstract}
@@
 in the absence of a focused behavioral goal
-\cite{kenet2003spontaneously, fiser2004small}.
+\citep{kenet2003spontaneously, fiser2004small}.
 This baseline energy demand is subject to surprisingly little modulations
@@
 increased in neural activity during
-unfocused everyday mind wandering (Raichle et al., 2001). That is, in
-the beginning of the 21st century, brain imaging was the first technology
+unfocused everyday mind wandering (Raichle et al., 2001).
+In the beginning of the 21st century, brain imaging was hence the first technology
 to allow for the discovery of a unique brain network that subserves
-baseline mental activity (Bzdok et al., 2012; Bzdok et al., 2015).
+baseline mental activities (Bzdok et al., 2012; Bzdok et al., 2015).
 
@@
 been shown to directly impact human behavior. Goal-directed task performance
-improved with decreased activity in default-mode areas (Weissman et al., 2006)
+improved with decreased activity in default mode areas (Weissman et al., 2006)
 and increased DMN activity was linked to more task-independent,
@@
 the more mental activity becomes detached from the actual sensory environment
-\cite{antrobus1966studies, pope1978regulation}.
-For instance,
+\citep{antrobus1966studies, pope1978regulation}.
+Conversely,
 the more the currently executed task is unknown and unpracticed,
 the lower the tendency for stimulus-independent thoughts
-\cite{filler1973daydreaming, teasdale1995stimulus}.
+\citep{filler1973daydreaming, teasdale1995stimulus}.
 These ``offline'' processes may however contribute to optimized control of the organism.
@@
 based on a brain implementation of Markov Chain Monte Carlo sampling
-\cite{tenenbaum2011grow}.
-\textit{Control} refers to the influences an agent exterts when interacting
+\citep{tenenbaum2011grow}.
+\textit{Control} refers to the influences an agent exerts when interacting
 with the environment to encourage preferred states.
@@
 of past experience, random mind-wandering and dreams during sleep
-may provide meanginful building blocks to iteratively improve
+may provide meaningful building blocks to iteratively improve
 the predictive DMN machinery to optimize the behavior of the organism.
@@
 probabilistic model of anticipated events''
-\cite{raichle2005intrinsic}.
+\citep{raichle2005intrinsic}.
 This paper proposes a formal model that
@@
 
-\section{Known Neurobiological Properties of the Default-Mode Network}
-We will attempt to deconstruct the overall DMN function
+\section{Known Neurobiological Properties of the Default Mode Network}
+We will first attempt to deconstruct the overall DMN function
 by the mechanistic relevance of each individual network node
-based on published evidence from neuroscience experiments.
-%
-The PMC and the dmPFC within the DMN
+based on published findings from neuroscience experiments.
+%
+The posteriormedial cortex (PMC) and
+the dorsomedial prefrontal cortex (dmPFC) within the DMN
 are responsible for the highest metabolic turn-over
@@
 as well as inability to direct action towards
-current unattended environmental objects
+currently unattended environmental objects
 (i.e., optic ataxia).
@@
 Congruently,
-the human PMC was coupled in across two functional connectivity modalities
+the human PMC was coupled in two functional connectivity modalities
 with the amygdala
@@
 %
-The PCC may consequently monitor the subjective outcomes
+The PMC may consequently monitor the subjective outcomes
 of possible decisions and integrates that information
@@
 In doing so, the PMC continuously adapts to changes
-in both the external environment and internal milieu
+in both the external environment and its internal representation
 that modulate strategic behavioral adjustment in volatile environments.
@@
 of one's own and other individuals' action and thoughts.
-Patients with neurological lesions in the prefrontal cortex
+Generally,
+patients with neurological lesions in the prefrontal cortex
 are known to adapt to novel situations and stimuli with difficulty
@@
 dmPFC activity indeed explained the proficiency decline
-of inferring other peoples thoughts in the elderly (Moran et al., 2012).
+of inferring other peoples' thoughts in the elderly (Moran et al., 2012).
 Some dmPFC neurons in macaque monkeys exhibited a preference
@@
 
-The vmPFC subserves less ambiguous subjective-value-related evaluative processes
+The ventromedial prefrontal cortex (vmPFC) subserves
+less ambiguous subjective-value-related evaluative processes
 reward-informed and risk estimation of self-relevant environmental stimuli.
@@
 probably drawing on model-based value representations
-\cite{doherty2015structure}.
+\citep{doherty2015structure}.
 Quantative lesion findings across 344 human individuals confirmed
@@
 have been shown to be reenacted spontaneously during
-quiet wakefulness and sleep \cite{hartley2014space}.
+quiet wakefulness and sleep \citep{hartley2014space}.
 Moreover, spike sequences measured in hippocampal place cells of rats
@@
 have been proposed to be critical in communicating local information
-to the neocortex for long-term storage, potentially also in the DMN.
+to the neocortex for long-term storage, potentially also in the nodes of the DMN.
 These mechanisms of the hippocampus probably make important contributions to the
@@
 ongoing perception in the current environment (Maguire 2016).
--> self-play in deepmind Go
--> Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later
+(TODO: relate to self-play in deepmind Go).
+(TODO: Kai had an interesting idea: the deja-vu effect
+could be when you already accidentally perturbed past
+such that it really occures in reality later.)
 
@@
 Involvement of this DMN node was repeatedly reported in
-multi-step action execution \cite{hartmann2005takes},
+multi-step action execution \citep{hartmann2005takes},
 visuo-proprioceptive conflict (Balslev et al., 2005) and
@@
 without necessarily relating those to outcome value
-\cite{liljeholm2013neural, hamilton2008action,
+\citep{liljeholm2013neural, hamilton2008action,
 jakobs2009effects}.
@@
 to deal instead with immediate environmental objects and contexts.
--> (slow) endogeneous changes in state transitions by the policy and value matrices
-For example, the alarm at the museum is not part of the task set of listening to the guide’s discussion of Hieronymous Bosch, but it is clearly a behaviourally relevant stimulus. <=> MCMC in HC / PCC
+(TODO: slow endogeneous changes in state transitions by the policy and value matrices)
 
@@
 to model action concepts
-\cite{binder2011neurobiology, seghier2013angular}.
+\citep{binder2011neurobiology, seghier2013angular}.
 For instance, LTPJ lesions also entail problems in recognizing
@@
 \paragraph{How many controls do we need at best ?}
-In the thermodynamic limit ($n \rightarrow \infty$), statistical physics \cite{Liu2011} gives the extremely good estimate
+In the thermodynamic limit ($n \rightarrow \infty$), statistical physics \citep{Liu2011} gives the extremely good estimate
 \begin{equation}
@@
 \paragraph{Why linear dynamics ?}
-As explained in \cite{Liu2011}, the choice of linear dynamics over a more general nonlinear dynamics can be justified as follows:
+As explained in \citep{Liu2011}, the choice of linear dynamics over a more general nonlinear dynamics can be justified as follows:
 \begin{itemize}
@@
 \subsection{Demystifying the free-energy principle and active-inference}
-The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
+The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
 
@@
 \subsection{Free-energy minimization via variational auto-encoders}
-Following Kingma et al. 2014 \cite{kingma2013auto}, define the data-dependent auxiliary random function
+Following Kingma et al. 2014 \citep{kingma2013auto}, define the data-dependent auxiliary random function
 \begin{eqnarray}
@@
   \end{tabular}
-  \caption{Reparametrization trick for a variety of models. Kingma et al. 2014 \cite{kingma2013auto}.}
+  \caption{Reparametrization trick for a variety of models. Kingma et al. 2014 \citep{kingma2013auto}.}
 \end{table}
@@
 %%     \end{itemize}
-%% The material presented is a revisit of \cite{braun2011path}. See also \cite{ortega2013thermodynamics}
+%% The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
 %% \paragraph{Utility functions and conjugate pairs.}
@@
 
-%% To conclude this section, let's note that \cite{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
+%% To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
 For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...
@@
 which has often been closely related to the medial prefrontal part of the DMN
-\cite{koechlin1999role, braver2002role}.
+\citep{koechlin1999role, braver2002role}.
 Using the Bellman equation, each state can be associated with a certain value,
@@
 Monte-Carlo simulation can be used to update the agent’s belief state
-\cite{silver2010monte}.
+\citep{silver2010monte}.
 Monte-Carlo simulation provides a simple method for evaluating
@@
   Our approach acknowledges the unavoidable stochasticity of
-  computation in neuronal processes \cite{faisal2008noise}.
+  computation in neuronal processes \citep{faisal2008noise}.
   \item
@@
 is a frequently evoked idea related to the default mode function
-\cite{bar2007}.
+\citep{bar2007}.
 According to this cognitive account,
@@
 back projections probably outnumber by far the existing input projection in the brain
-\cite{salin1995corticocortical}.
-This can explains the constructive, generative nature of sensory perception
-\cite{friston2010free} and
+\citep{salin1995corticocortical}.
+This can explain the constructive, generative nature of sensory perception
+\citep{friston2010free} and
 why motor action is intimately related to sensory expectations
-\cite{wolpert1995internal, kording2004bayesian}.
+\citep{wolpert1995internal, kording2004bayesian}.
 At each stage of neural processing,
@@
 \begin{itemize}
-  \item Both the predictive coding account and MPDs
+  \item Both the predictive coding account and MDPs
   are process theories backed up by plausible
@@
 the abstract knowledge derived from past experiences
-\cite{bar2007proactive, binder1999conceptual}.
+\citep{bar2007proactive, binder1999conceptual}.
 Analogies might naturally tie incoming novel sensory stimuli to
 explicit world knowledge extractable from the environment
-\cite{bar2009proactive}.
+\citep{bar2009proactive}.
 %
@@
 concurrent deficits in forming and integrating relationships between
-items \cite{addis2008age, spreng2006temporal}.
+items \citep{addis2008age, spreng2006temporal}.
 A constructive account further explains the reciprocal relationship
@@
 for binding the constituent elements in mental scene construction
-\cite{schacter2007remembering}.
+\citep{schacter2007remembering}.
 These scene generation processes could contribute to interpreting the
@@
 even without reward incentives, and exploit it later
-for other means \cite{tolman1948cognitive}.
+for other means \citep{tolman1948cognitive}.
 
@@
   not only impair retrieving memories, but also hypothetical and future
-  thinking \cite{hassabis2007patients}.
+  thinking \citep{hassabis2007patients}.
   \item Both the semantic hypothesis and MDP expose mechanisms of
@@
   while simulation of possible alternatives are constructed,
-  followed by a return to perception of the present.''\cite{buckner2007self}},
+  followed by a return to perception of the present.''\citep{buckner2007self}},
   the MDP naturally integrates the former egocentric
@@
 Processing self-relevant information was perhaps the first
-cognitive account that was proposed for the DMN \cite{gusnard2001medial}.
+cognitive account that was proposed for the DMN \citep{gusnard2001medial}.
 Since then,
 different investigators have speculated that neural activity in the DMN
-may reflects the brain’s relentless tracking of
+may reflect the brain’s relentless tracking of
 relevance in the environment
@@
 the complexity of their social systems
-\cite{tomasello2009cultural}.
+\citep{tomasello2009cultural}.
 More specifically,
@@
 social systems to solve ecological problems by means of social connections
-\cite{whiten1988machiavellian}.
+\citep{whiten1988machiavellian}.
 Highly efficient neurobiological processing of social cues is exemplified by
 facial judgments being routinely processed in less than 100 ms
-\cite{bar2006very}.
+\citep{bar2006very}.
 Our human ancestors were probably among the few organisms that
@@
 turned into a decisive evolutionary advantage
-\cite{frith2010social}.
+\citep{frith2010social}.
 %
 This may explain why social topics amount to roughly
-two thirds of human everyday communication \cite{dunbar1997human}
+two thirds of human everyday communication \citep{dunbar1997human}
 and
@@
     physiological baseline of human brain function
-    \cite{schilbach2008minds}. This was later challenged by observing
-    analogues of the DMN in monkeys \cite{mantini2011default}
-    and rats \cite{lu2012rat}, with
+    \citep{schilbach2008minds}. This was later challenged by observing
+    analogues of the DMN in monkeys \citep{mantini2011default}
+    and rats \citep{lu2012rat}, with
     supposedly less advanced social capacities.
@@
     also as a baseline function of human mental activity.
-    In tasks, the policy and value matrices are updated to optmize short-term action,
+    In tasks, the policy and value matrices are updated to optimize short-term action,
     whereas, at rest, these parameter updates may
-    serve especially mid- and long-term action.
+    improve especially mid- and long-term action.
     \item The sentinel account does not provide a formal mechanisms of
@@
     $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
-    are recursively explored by the agend and which decision trees are neglected.
+    are recursively explored by the agent and which decision trees are neglected.
     This directly implies allocation of attention and computational load.
@@
 \begin{itemize}
-  \item The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
+  \item The model proposed in \citep{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
 \end{itemize}
@@
   rewarding outcomes in the future.
-  \item Ortega et al. 2013 \cite{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
-\item Friston's critic of RL \cite{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
+  \item Ortega et al. 2013 \citep{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
+\item Friston's critic of RL \citep{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
 pioneered in the 1950s by Richard Bellman and colleagues [2]. To

### 2016-10-22 paper/paper.tex
fixed some stuff in free energy

@@
     &= -\langle \log  (p_G(., \zeta)) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta))\\
-    &= \langle E_G(., \zeta) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta)) := F_G(\zeta)
+    &= \langle E_G(., \zeta) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta))
   \end{split}
@@
 \begin{equation}
-  E_G(\psi, \zeta) := -\log(p_G(\psi, \zeta)).
+  E_G(\psi, \zeta) := -\log(p_G(\psi, \zeta)),
   \label{eq:gibbs}
@@
 %% In the above, the variable $\psi$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
-The last quantity in \eqref{eq:helm} defined formally as
-\begin{equation}
-  F_G(\zeta) := \langle E_G(., \zeta) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta))
-\end{equation}
-is the \textit{Helmholtz free-energy} (at unit temperature!). %% , modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of macrostates $\psi$
+so that one recognizes The last quantity in \eqref{eq:helm} as the the \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
+\begin{equation}
+  F_G(\zeta) := \langle E_G(., \zeta) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta)).
+\end{equation}
+%% , modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of macrostates $\psi$
 Thus generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
@@
 where $Z_G(\zeta) := \log(p_G(\zeta)) = \sum_{\psi'}\exp(-E_G(\psi',\zeta))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
-Whence $\forall \psi$, $p_G(\zeta) = Z_G(\zeta) = \exp(-E_G(\psi, \zeta)) / p_G(\psi|\zeta)$, and so we have the invariance relation
+Whence, for any macrostate $\psi$, we have $p_G(\zeta) = Z_G(\zeta) = \exp(-E_G(\psi, \zeta)) / p_G(\psi|\zeta)$, and so it holds that
 \begin{equation}
@@
 \end{equation}
-Now, in the equation above, the LHS only depends on the generative model $G$ and the data point $\zeta$: it doesn't depend on the hidden variable $\psi$, etc.
-So, taking expectations w.r.t an arbitrary density\footnote{The conditioning in $P_R(.|\zeta)$ is because this density is selected from a world in which the sensory inputs and internal brain state vector $\zeta$ is assumed already observed.} $p_R(.|\zeta)$ yields
+Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\zeta$:
+it doesn't depend on the hidden variable $\psi$, etc. So, taking expectations w.r.t an arbitrary
+density\footnote{The conditioning in $P_R(.|\zeta)$ is because this density is selected from a world in which the sensory inputs and internal brain
+  state vector $\zeta$ is assumed already observed.} $p_R(.|\zeta)$ yields
 \begin{equation}

### 2016-10-22 paper/paper.tex
simplified math notation

@@
 \newtheorem{cor}[thm]{Corollary}
+\newtheorem{fact}[thm]{Fact}
 
@@
          $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
-             $\zeta$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e $\zeta = (s, b).$ \\ \hline
-             $\psi$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
-             $p_G(.|\zeta)$ & Generative density for ...\\ \hline
-         $p_R(.|\zeta)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
-         $F_G(\zeta)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\zeta$. This measures the surprise incured upon observing $\zeta$ generated by the model $G$.\\ \hline
-         $F^R_G(\zeta)$ & Variational Helmholtz free-energy from model $G$
+             $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e $\x = (s, b).$ \\ \hline
+             $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
+             $p_G(.|\x)$ & Generative density for ...\\ \hline
+         $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
+         $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
+         $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
           to $R$.  Note that $F^G_G = F_G$.\\ \hline
@@
   \begin{split}
-    \text{generative surprise } &= -\log(p_G(\zeta)) = -\log(p_G(\zeta)) \times 1 = -\log(P_G(\zeta))\sum_{\psi}p_G(\psi |\zeta)\\
-    &= -\sum_{\psi}p_G(\psi, \zeta)\log(p_G(\zeta))
-    =-\sum_{\psi}p_G(\psi |\zeta)\log(p_G(\psi, \zeta)/p_G(\psi|\zeta))\\
-    &= \sum_{\psi}p_G(\psi |\zeta)\log(p_G(\psi|\zeta))-\sum_{\psi}p_G(\psi |\zeta)\log  (p_G(\psi, \zeta))\\
-    &= -\langle \log  (p_G(., \zeta)) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta))\\
-    &= \langle E_G(., \zeta) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta))
+    \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(P_G(\x))\sum_{\z}p_G(\z |\x)\\
+    &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
+    =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
+    &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
+    &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
+    &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
   \end{split}
@@
 \end{eqnarray}
-where $E_G(\psi, \zeta)$ is the energy at \textit{macrostate} $\psi$ of a fictive thermodynamic system defined by setting
-\begin{equation}
-  E_G(\psi, \zeta) := -\log(p_G(\psi, \zeta)),
+where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
+\begin{equation}
+  E_G(\z, \x) := -\log(p_G(\z, \x)),
   \label{eq:gibbs}
 \end{equation}
-%% In the above, the variable $\psi$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
+%% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
 so that one recognizes The last quantity in \eqref{eq:helm} as the the \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
 \begin{equation}
-  F_G(\zeta) := \langle E_G(., \zeta) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta)).
-\end{equation}
-%% , modelled as the physical system for which $p_G(\psi|\zeta)$ dictates the occupation probabilities of macrostates $\psi$
-Thus generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
-
-The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|s)$, so as to minimize surprise.
-It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(s)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
-A workaround is then to introduce a second density $p_R(.|\zeta)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\zeta)$, as a trick for doing approximate inference.
+  F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
+\end{equation}
+%% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
+Thus,
+\begin{fact}
+  Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
+\end{fact}
+
+The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
+It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(\x)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
+A workaround is then to introduce a second density $p_R(.|\x)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\x)$, as a trick for doing approximate inference.
 The former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
@@
 \begin{equation}
-  p_G(\psi|\zeta) = \frac{p_G(\psi|\zeta)}{p_G(\zeta)} = \frac{\exp(-E_G(\psi, \zeta))}{Z_G(\zeta)} = \frac{\exp(-E_G(\psi, \zeta))}{Z_G(\zeta)},
-\end{equation}
-where $Z_G(\zeta) := \log(p_G(\zeta)) = \sum_{\psi'}\exp(-E_G(\psi',\zeta))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
-Whence, for any macrostate $\psi$, we have $p_G(\zeta) = Z_G(\zeta) = \exp(-E_G(\psi, \zeta)) / p_G(\psi|\zeta)$, and so it holds that
-\begin{equation}
-F_G(\zeta)\overset{\eqref{eq:helm}}{=} -\log(p_G(\zeta)) = -\log(Z_G(\zeta)) = E_G(\psi, \zeta) + \log(p_G(\psi|\zeta)).
-\end{equation}
-Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\zeta$:
-it doesn't depend on the hidden variable $\psi$, etc. So, taking expectations w.r.t an arbitrary
-density\footnote{The conditioning in $P_R(.|\zeta)$ is because this density is selected from a world in which the sensory inputs and internal brain
-  state vector $\zeta$ is assumed already observed.} $p_R(.|\zeta)$ yields
+  p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
+\end{equation}
+where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
+Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
+\begin{equation}
+F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
+\end{equation}
+Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
+it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
+density\footnote{conditioning in $P_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
+  state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
 \begin{equation}
   \begin{split}
-    F_G(\zeta) &= -\log(Z_G(\zeta)) = \langle E_G(., \zeta)\rangle_{P_R(.|\zeta)} + \sum_{\psi}p_R(\psi|\zeta)\log(p_G(\psi|\zeta))\\
-    &= \langle E_G(., \zeta)\rangle_{P_R(.|\zeta)} - \mathcal H(p_R(.|\zeta)) - \sum_{\psi}p_R(\psi|\zeta)\log(p_R(\psi|\zeta)/p_G(\psi|\zeta))\\
-    &= F^R_G(\zeta) - D_{KL}(P_R(.|\zeta) || P_G(.|\zeta)),
+    F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{P_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
+    &= \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
+    &= F^R_G(\x) - D_{KL}(P_R(.|\x) || P_G(.|\x)),
   \end{split}
@@
 \end{equation}
-where $F^R_G(\zeta)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
-\begin{equation}
-  F^R_G(\zeta) := \langle E_G(., \zeta)\rangle_{P_R(.|\zeta)} - \mathcal H(p_R(.|\zeta))
-\end{equation}
-and $D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))$ is the Kullback-Leibler divergence between the $p_R(.|\zeta)$ and the generative density $p_G(.|\zeta)$. Note that $F^G_G = F_G$.
+where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
+\begin{equation}
+  F^R_G(\x) := \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x))
+\end{equation}
+and $D_{KL}(P_R(.|\x) || P_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
 
@@
   \begin{split}
-    \text{generative surprise } &:= -\log(p_G(\zeta)) = F_G(\zeta) \\
-    &=\underbrace{F^R_G(\zeta)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))}_{\text{complexity}} \\
-    &\le F^R_G(\zeta),
-    \text{ with equality if }p_R(.|\zeta) = p_G(.|\zeta),
+    \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
+    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
+    &\le F^R_G(\x),
+    \text{ with equality if }p_R(.|\x) = p_G(.|\x),
     \end{split}
@@
 %%   \begin{equation}
-%%     \min_{G}F_G(\zeta) = \min_{G,R}F^R_G(\zeta)
+%%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
 %%     \end{equation}
@@
 $$
-p_G(\psi^{(l)}|\zeta) = \Pi_{k=1}^{h_l}p_G(\psi_k^{(l)} | \zeta),\hspace{1cm}
-p_R(\psi^{(l)}|\zeta) = \Pi_{k=1}^{h_l}p_R(\psi_k^{(l)} |\zeta)
+p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
+p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
 $$
@@
 \begin{eqnarray}
-  f_{G,R}(., \zeta) :\psi \mapsto \log(p_G(\psi,\zeta)) - \log(p_R(\psi|\zeta)).
+  f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
 \end{eqnarray}
@@
   \begin{split}
-    F_G^R(\zeta) &:= \langle E_G(., \zeta) \rangle_{p_R(. |\zeta)} - \mathcal H(p_R(. |\zeta)) = \langle E_G(., \zeta) + \log(p_R(.|\zeta))\rangle_{P_R(.|\zeta)}\\
-    &=\langle -\log(p_G(.,\zeta)) + \log(p_R(.|\zeta))\rangle_{P_R(.|\zeta)}\\
-    &= -\langle f_{G,R}\rangle_{P_R(.|\zeta)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\psi^{(m)}), \text{ with }\psi^{(1)},\ldots,\psi^{(M)} \sim p_R(.|\zeta), \text{ and }M \rightarrow \infty.
+    F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
+    &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
+    &= -\langle f_{G,R}\rangle_{P_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
     \end{split}
@@
 \begin{mdframed}
-  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\zeta)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
+  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
 \end{mdframed}
@@
 \item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
-\item Set $\phi = g(\W^{R}, \zeta, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
+\item Set $\phi = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
   \begin{itemize}
-  \item results in a sample $\phi \sim p_R(.|\zeta)$, from the correct
+  \item results in a sample $\phi \sim p_R(.|\x)$, from the correct
     posterior
@@
          \hline
-         Posterior & $p_R(.|\zeta)$ & noise & $g(\W^R,\zeta,\epsilon)$ & Also \\ \hline
+         Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
          Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
@@
 
-%% For fixed $p_G(.|\zeta)$, the optimal choice for $p_R(.|\zeta)$ is given analytically by
+%% For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
 %% \begin{equation}
-%%   p_R(\psi|\zeta) \propto p_G(\psi|\zeta)\exp(-\Delta E_{G \rightarrow R}(\psi, \zeta)).
+%%   p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
 %% \end{equation}

### 2016-10-23 paper/paper.tex
restructured free-energy principles + misc passes on tex

@@
 %% \subsection{Minimal control energy controller}
-
-\subsection{Demystifying the free-energy principle and active-inference}
-The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
-
-In this section, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
-This ideas were first developed by Hinton et al.
-in the early 90s in building their Helmholtz machine.
-Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises.
-For example the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial, e.g) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
-
-\subsubsection{Helmholtz free-energy and the generative model}
-\begin{table}[H]
-  \begin{tabular}{p{2cm}|p{11cm}}
-         \hline
-         \textbf{symbol}    & \textbf{description}  \\ \hline
-         $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
-         random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
-         $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
-          with the usual convention $0 \log(0) := 0$.\\ \hline
-         $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
-             $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e $\x = (s, b).$ \\ \hline
-             $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
-             $p_G(.|\x)$ & Generative density for ...\\ \hline
-         $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
-         $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
-         $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
-          to $R$.  Note that $F^G_G = F_G$.\\ \hline
-  \end{tabular}
-  \caption{Table of notations.}
-\end{table}
-Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
-So,
-\begin{eqnarray}
-  \begin{split}
-    \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(P_G(\x))\sum_{\z}p_G(\z |\x)\\
-    &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
-    =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
-    &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
-    &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
-    &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
-  \end{split}
-  \label{eq:helm}
-\end{eqnarray}
-where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
-\begin{equation}
-  E_G(\z, \x) := -\log(p_G(\z, \x)),
-  \label{eq:gibbs}
-\end{equation}
-%% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
-so that one recognizes The last quantity in \eqref{eq:helm} as the the \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
-\begin{equation}
-  F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
-\end{equation}
-%% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
-Thus,
-\begin{fact}
-  Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
-\end{fact}
-
-The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
-It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(\x)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
-A workaround is then to introduce a second density $p_R(.|\x)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\x)$, as a trick for doing approximate inference.
-The former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
-This primal-dual idea, first proposed in Hinton et al.
-1995, is at the heart of the general free-energy principle that we'll introduce shortly.
-
-\subsubsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
-In this subsection, we'll present an insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
-As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
-
-Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
-distribution
-\begin{equation}
-  p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
-\end{equation}
-where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
-Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
-\begin{equation}
-F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
-\end{equation}
-Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
-it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
-density\footnote{conditioning in $P_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
-  state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
-\begin{equation}
-  \begin{split}
-    F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{P_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
-    &= \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
-    &= F^R_G(\x) - D_{KL}(P_R(.|\x) || P_G(.|\x)),
-  \end{split}
-  \label{eq:fe}
-\end{equation}
-where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
-\begin{equation}
-  F^R_G(\x) := \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x))
-\end{equation}
-and $D_{KL}(P_R(.|\x) || P_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
-
-\subsubsection{A general free-energy principle}
-We can resume the situation as follows:
-\begin{mdframed}
-\begin{equation}
-  \begin{split}
-    \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
-    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
-    &\le F^R_G(\x),
-    \text{ with equality if }p_R(.|\x) = p_G(.|\x),
-    \end{split}
-\end{equation}
-\end{mdframed}
-where we have used the fact that KL divergence is always nonnegative.
-%% In fact, we have the following theorem
-%% \begin{theorem}
-%%   It holds that
-%%   $$
-%%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
-%%   i.e
-%%   \begin{equation}
-%%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
-%%     \end{equation}
-%% \end{theorem}
-
-\subsubsection{Helmholtz machines and the wake-sleep algorithm}
-% \begin{mdframed}
-  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e input from lower more primitive layers).
-Precisely
-$$
-p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
-p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
-$$
-% \end{mdframed}
-
-\subsubsection{Friston's active-inference}
-This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
-
-The following critics can be made:
-\begin{itemize}
-  \item
-    As noted by Dayan et al.
-(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
-  \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
-  \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
-    technical difficulty in getting derivatives of loss function w.r.t
-    recognition weights $\W^R$).
-  \item This difficulty was removed in the 2010s by Kingma et al., an other groups, via a ``reparametrization trick''.
-    \end{itemize}
-
-\subsection{Free-energy minimization via variational auto-encoders}
-Following Kingma et al. 2014 \citep{kingma2013auto}, define the data-dependent auxiliary random function
-\begin{eqnarray}
-  f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
-\end{eqnarray}
-Then we can rewrite the
-variational free-energy as
-\begin{eqnarray*}
-  \begin{split}
-    F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
-    &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
-    &= -\langle f_{G,R}\rangle_{P_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
-    \end{split}
-\end{eqnarray*}
-
-\begin{mdframed}
-  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
-\end{mdframed}
-\paragraph{Solution: The reparametrization trick.}
-\begin{itemize}
-\item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
-\item Set $\phi = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
-  \begin{itemize}
-  \item results in a sample $\phi \sim p_R(.|\x)$, from the correct
-    posterior
-    \end{itemize}
-\end{itemize}
-
-\begin{table}[H]
-  \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
-         \hline
-         Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
-         Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
-         Student’s t, Logistic, Uniform, Triangular, ... \\ \hline
-         Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
-\\ \hline
-         Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
-  \end{tabular}
-  \caption{Reparametrization trick for a variety of models. Kingma et al. 2014 \citep{kingma2013auto}.}
-\end{table}
-
-The result is a scheme for training DBNs via good-old backprop!
-
-%% For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
-%% \begin{equation}
-%%   p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
-%% \end{equation}
-
-%% \subsection{A thermodynamic model for bounded rationality, aka robust
-%%   optimality}
-%% \begin{itemize}
-%%   \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
-%%   \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
-%%   \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
-%%     \end{itemize}
-%% The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
-%% \paragraph{Utility functions and conjugate pairs.}
-%%   Let $(\Omega, \mathcal F, P)$ be a probablity space.
-A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
-%%   \begin{itemize}
-%%   \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.
-%%   \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
-%%     $A,B \in \mathcal F$.
-%%   \item monotonicity: $f_{U}$ is strictly increasing.
-%%     \end{itemize}
-
-%% \begin{theorem}
-%%   The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
-%%   \begin{equation}
-%%     f = \alpha \log(.),
-%%     \label{eq:boltzmann}
-%%   \end{equation}
-%%   where $\alpha > 0$.
-%% \end{theorem}
-
-%% \textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
-
-%% Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
-
-%% \paragraph{Example.}
-%% Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
-%%   \begin{equation}
-%%     P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
-%%   \end{equation}
-%%   where
-
-%%   \begin{equation}
-%%     Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
-%%   \end{equation}
-%%   is a normalization constant called the \textit{partition function} of $U$.
-%%   It's not hard to see that $U$ and the $P$ above form a conjugate pair.
-
-%%   \paragraph{Free-utility functional.}
-%%     Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
-Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
-%%     \begin{equation}
-%%       J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
-%%     \label{eq:free_u}
-%%     \end{equation}
-%%     where
-%%     \begin{equation}
-%%       \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
-%%     \end{equation}
-%%     is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
-%%   \begin{equation}
-%%     J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
-%%   \end{equation}
-%%   In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
-%%   \begin{equation}
-%%     J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
-%%     \end{equation}
-
-%%   \paragraph{The free-energy / utility principle (of Friston ?).}
-%%   We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
-Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
-We want to transform this to a new model by adding constraints represented by the utility function $\Delta U$.
-The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \Delta U$.
-The difference in free-utility is then
-%%   \begin{equation}
-%%     \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
-%%     \label{eq:free}
-%%   \end{equation}
-%%   where
-%%   \begin{equation}
-%%     D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
-%%   \end{equation}
-%%   is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
-%% In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
-%%   The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
-%% a trade-off between an expected utility $\Delta U$ (first term) and
-%% the information cost of transforming Pi
-%% into $P_{\text{fin}}$ (second
-%% term). In this interpretation $P_{\text{ini}}$ represents an initial choice
-%% probability or policy, which includes the special case of the
-%% uniform distribution where the decision-maker has initially no
-%% preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
-%% considers the utility constraint U∗
-%% that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
-%% \begin{equation}
-%% P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
-%%   \end{equation}
-
-%% By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
-
-%% \begin{equation}
-%%   P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
-%% \end{equation}
-
-%% Two limit cases are worth considering.
-%% \begin{itemize}
-%% \item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
-%%   $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
-%%   where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
-%%   \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
-%% \end{itemize}
-
-%% To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
-For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...
 
@@
 \end{itemize}
+
+\section{Relation to other statistical models (Elvis)}
+\subsection{Free-energy principles!}
+The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
+
+In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
+This ideas were first developed by Hinton et al.
+in the early 90s in building their Helmholtz machine.
+Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises.
+For example the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial, e.g) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
+
+\subsubsection{Helmholtz free-energy and the generative model}
+\begin{table}[H]
+  \begin{tabular}{p{2cm}|p{11cm}}
+         \hline
+         \textbf{symbol}    & \textbf{description}  \\ \hline
+         $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
+         random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
+         $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
+          with the usual convention $0 \log(0) := 0$.\\ \hline
+         $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
+             $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e $\x = (s, b).$ \\ \hline
+             $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
+             $p_G(.|\x)$ & Generative density for ...\\ \hline
+         $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
+         $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
+         $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
+          to $R$.  Note that $F^G_G = F_G$.\\ \hline
+  \end{tabular}
+  \caption{Table of notations.}
+\end{table}
+Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
+So,
+\begin{eqnarray}
+  \begin{split}
+    \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(P_G(\x))\sum_{\z}p_G(\z |\x)\\
+    &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
+    =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
+    &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
+    &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
+    &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
+  \end{split}
+  \label{eq:helm}
+\end{eqnarray}
+where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
+\begin{equation}
+  E_G(\z, \x) := -\log(p_G(\z, \x)),
+  \label{eq:gibbs}
+\end{equation}
+%% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
+so that one recognizes The last quantity in \eqref{eq:helm} as the the \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
+\begin{equation}
+  F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
+\end{equation}
+%% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
+Thus,
+\begin{fact}
+  Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
+\end{fact}
+
+The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
+It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(\x)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
+A workaround is then to introduce a second density $p_R(.|\x)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\x)$, as a trick for doing approximate inference.
+The former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
+This primal-dual idea, first proposed in Hinton et al.
+1995, is at the heart of the general free-energy principle that we'll introduce shortly.
+
+\subsubsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
+In this subsection, we'll present an insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
+As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
+
+Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
+distribution
+\begin{equation}
+  p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
+\end{equation}
+where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
+Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
+\begin{equation}
+F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
+\end{equation}
+Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
+it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
+density\footnote{conditioning in $P_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
+  state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
+\begin{equation}
+  \begin{split}
+    F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{P_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
+    &= \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
+    &= F^R_G(\x) - D_{KL}(P_R(.|\x) || P_G(.|\x)),
+  \end{split}
+  \label{eq:fe}
+\end{equation}
+where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
+\begin{equation}
+  F^R_G(\x) := \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x))
+\end{equation}
+and $D_{KL}(P_R(.|\x) || P_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
+
+\subsubsection{A general free-energy principle}
+We can resume the situation as follows:
+\begin{mdframed}
+\begin{equation}
+  \begin{split}
+    \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
+    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
+    &\le F^R_G(\x),
+    \text{ with equality if }p_R(.|\x) = p_G(.|\x),
+    \end{split}
+\end{equation}
+\end{mdframed}
+where we have used the fact that KL divergence is always nonnegative.
+%% In fact, we have the following theorem
+%% \begin{theorem}
+%%   It holds that
+%%   $$
+%%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
+%%   i.e
+%%   \begin{equation}
+%%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
+%%     \end{equation}
+%% \end{theorem}
+
+\subsubsection{Helmholtz machines and the wake-sleep algorithm}
+% \begin{mdframed}
+  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e input from lower more primitive layers).
+Precisely
+$$
+p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
+p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
+$$
+% \end{mdframed}
+
+\subsubsection{Friston's active-inference}
+This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
+
+The following critics can be made:
+\begin{itemize}
+  \item
+    As noted by Dayan et al.
+(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
+  \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
+  \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
+    technical difficulty in getting derivatives of loss function w.r.t
+    recognition weights $\W^R$).
+  \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
+    \end{itemize}
+
+\subsection{Free-energy minimization via variational auto-encoders}
+\label{sec:vae}
+Here, we present a possible way to aleviate some conceptual and computational issues with the
+free-energy framework presented thus far, by using auto-encoder theory.
+Following Kingma et al. 2014 \citep{kingma2013auto}, define the data-dependent auxiliary random
+function
+\begin{eqnarray}
+  f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
+\end{eqnarray}
+Then we can rewrite the
+variational free-energy as
+\begin{eqnarray*}
+  \begin{split}
+    F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
+    &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
+    &= -\langle f_{G,R}\rangle_{P_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
+    \end{split}
+\end{eqnarray*}
+
+\begin{mdframed}
+  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
+\end{mdframed}
+\paragraph{Solution: The reparametrization trick.}
+\begin{itemize}
+\item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
+\item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
+  \begin{itemize}
+  \item results in a sample $\z \sim p_R(.|\x)$, from the correct
+    posterior
+    \end{itemize}
+\end{itemize}
+
+\begin{table}[H]
+  \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
+         \hline
+         Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
+         Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
+         Student’s t, Logistic, Uniform, Triangular, ... \\ \hline
+         Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
+\\ \hline
+         Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
+  \end{tabular}
+  \caption{Reparametrization trick for a variety of models. Kingma et al. 2014 \citep{kingma2013auto}.}
+\end{table}
+
+The result is a scheme for training DBNs via good-old backprop!
+
+%% For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
+%% \begin{equation}
+%%   p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
+%% \end{equation}
+
+%% \subsection{A thermodynamic model for bounded rationality, aka robust
+%%   optimality}
+%% \begin{itemize}
+%%   \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
+%%   \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
+%%   \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
+%%     \end{itemize}
+%% The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
+%% \paragraph{Utility functions and conjugate pairs.}
+%%   Let $(\Omega, \mathcal F, P)$ be a probablity space.
+A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
+%%   \begin{itemize}
+%%   \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.
+%%   \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
+%%     $A,B \in \mathcal F$.
+%%   \item monotonicity: $f_{U}$ is strictly increasing.
+%%     \end{itemize}
+
+%% \begin{theorem}
+%%   The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
+%%   \begin{equation}
+%%     f = \alpha \log(.),
+%%     \label{eq:boltzmann}
+%%   \end{equation}
+%%   where $\alpha > 0$.
+%% \end{theorem}
+
+%% \textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
+
+%% Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
+
+%% \paragraph{Example.}
+%% Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
+%%   \begin{equation}
+%%     P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
+%%   \end{equation}
+%%   where
+
+%%   \begin{equation}
+%%     Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
+%%   \end{equation}
+%%   is a normalization constant called the \textit{partition function} of $U$.
+%%   It's not hard to see that $U$ and the $P$ above form a conjugate pair.
+
+%%   \paragraph{Free-utility functional.}
+%%     Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
+Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
+%%     \begin{equation}
+%%       J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
+%%     \label{eq:free_u}
+%%     \end{equation}
+%%     where
+%%     \begin{equation}
+%%       \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
+%%     \end{equation}
+%%     is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
+%%   \begin{equation}
+%%     J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
+%%   \end{equation}
+%%   In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
+%%   \begin{equation}
+%%     J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
+%%     \end{equation}
+
+%%   \paragraph{The free-energy / utility principle (of Friston ?).}
+%%   We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
+Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
+We want to transform this to a new model by adding constraints represented by the utility function $\Delta U$.
+The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \Delta U$.
+The difference in free-utility is then
+%%   \begin{equation}
+%%     \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
+%%     \label{eq:free}
+%%   \end{equation}
+%%   where
+%%   \begin{equation}
+%%     D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
+%%   \end{equation}
+%%   is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
+%% In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
+%%   The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
+%% a trade-off between an expected utility $\Delta U$ (first term) and
+%% the information cost of transforming Pi
+%% into $P_{\text{fin}}$ (second
+%% term). In this interpretation $P_{\text{ini}}$ represents an initial choice
+%% probability or policy, which includes the special case of the
+%% uniform distribution where the decision-maker has initially no
+%% preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
+%% considers the utility constraint U∗
+%% that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
+%% \begin{equation}
+%% P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
+%%   \end{equation}
+
+%% By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
+
+%% \begin{equation}
+%%   P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
+%% \end{equation}
+
+%% Two limit cases are worth considering.
+%% \begin{itemize}
+%% \item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
+%%   $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
+%%   where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
+%%   \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
+%% \end{itemize}
+
+%% To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
+For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...
+\subsubsection{Concluding remarks}
+In summary:
+\begin{itemize}
+  \item Friston's free-energy principle: brain as inference engine
+  biological systems, including brains, must minimize the long-term average of surprise;
+  \item A common point between the Free Energy principle and our framework
+  is placing the minimization of suprise at the core of brain function.
+  In RL, surprise minimization reduces to accurate prediction of
+  rewarding outcomes in the future.
+  \item Ortega et al. 2013 \citep{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
+\item Friston's critic of RL \citep{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
+pioneered in the 1950s by Richard Bellman and colleagues [2]. To
+optimise control a~pð Þ x~ under this formulation, we have to: (i)
+assume the hidden states are available to the agent and (ii) solve
+Equation 18 for the value-function. Solving for the value-function
+is a non-trivial problem and usually involves backwards induction
+or some approximation scheme like reinforcement learning [4–6].
+The free-energy formulation circumvents these problems by
+prescribing the policy in terms of free-energy, which encodes
+optimal control (Equation 6)"
+\item It is difficult to compute with the free energy principle
+  in practice, which makes it physiologically less plausible. Auto-encoder theory presented
+ \ref{sec:vae} can help to remove these barriers.
+  \end{itemize}
+
+\subsection{Control theory}
+\begin{itemize}
+  \item Cite stuff from Bert K.
+  \item The model proposed in \citep{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
+\end{itemize}
+
+\subsection{Other Leinforcement Learning literature}
+abc
 
@@
 
-\section{Relation to other statistical models (Elvis)}
-\subsection{Control theory.}
-\begin{itemize}
-  \item The model proposed in \citep{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
-\end{itemize}
-
-\subsection{Reinforcement Learning.}
-abc
-
-\subsection{Free Energy Principle.}
-\begin{itemize}
-  \item Closely related to Friston's free-energy principle: brain as inference engine
-  biological systems, including brains, must minimize the long-term average of surprise;
-  \item A common point between the Free Energy principle and our framework
-  is placing the minimization of suprise at the core of brain function.
-  In RL, surprise minimization reduces to accurate prediction of
-  rewarding outcomes in the future.
-  \item Ortega et al. 2013 \citep{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
-\item Friston's critic of RL \citep{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
-pioneered in the 1950s by Richard Bellman and colleagues [2]. To
-optimise control a~pð Þ x~ under this formulation, we have to: (i)
-assume the hidden states are available to the agent and (ii) solve
-Equation 18 for the value-function. Solving for the value-function
-is a non-trivial problem and usually involves backwards induction
-or some approximation scheme like reinforcement learning [4–6].
-The free-energy formulation circumvents these problems by
-prescribing the policy in terms of free-energy, which encodes
-optimal control (Equation 6)"
-\item It is difficult to compute the free energy principle
-in practice, which makes it physiologically less plausible.
-  \end{itemize}
-
 \section{Conclusion}

### 2016-10-23 paper/paper.tex
more
improved and added rl process chart (good shit)

@@
 \tableofcontents
+\listoffigures
+\listoftables
 
@@
 \textbf{XXX Put the diagrams and process charts here.}
+\begin{figure}[!h]
+  \includegraphics[width=.9\linewidth]{rl_process_chart.pdf}
+  \caption{The DMN as an RL agent...}
+  \label{fig:rl_process_chart}
+\end{figure}
+
 \begin{itemize}
@@
 
-\subsection{Free-energy minimization via variational auto-encoders}
+\subsection{Minimizing free-energy via backprop!}
+% \subsection{Free-energy minimization via variational auto-encoders}
 \label{sec:vae}
-Here, we present a possible way to aleviate some conceptual and computational issues with the
+Here, we present a way to aleviate some conceptual and computational issues with the
 free-energy framework presented thus far, by using auto-encoder theory.
@@
   \end{tabular}
-  \caption{Reparametrization trick for a variety of models. Kingma et al. 2014 \citep{kingma2013auto}.}
+  \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
 \end{table}
@@
   in practice, which makes it physiologically less plausible. Auto-encoder theory presented
- \ref{sec:vae} can help to remove these barriers.
+  \ref{sec:vae} can help to remove these barriers, by making the whole scheme trainable via
+  backprop!
   \end{itemize}
@@
 \begin{itemize}
-  \item Cite stuff from Bert K.
+\item An authoritative review of control theory for decision making and its relationship to
+  reinforcement learning can be found in \citep{kappen2007introduction}. 
   \item The model proposed in \citep{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...

### 2016-10-25 paper/paper.tex
add a majority of references

@@
 when mentally processing environmental challenges
-(Bzdok and Eickhoff, 2016a).
+\citep{raichle2001pnas}.
 What has early been described as the "stream of consciousness"
-in psychology (James, 1890), found its neurobiological manifestation
+in psychology \citep{james1890principles},
+found its neurobiological manifestation
 in the so-called "default mode network" (DMN)
-(Shulman et al., 1997). This networked collection of some of the highest regions in
-the human association cortex (Mesulam et al., 1998) consistently
+\citep{shul1997}. This networked collection of some of the highest regions in
+the human association cortex \citep{mesulam1998sensation}
+consistently
 increased in neural activity during
-unfocused everyday mind wandering (Raichle et al., 2001).
+unfocused everyday mind wandering \citep{raichle2001pnas}.
 In the beginning of the 21st century, brain imaging was hence the first technology
 to allow for the discovery of a unique brain network that subserves
-baseline mental activities (Bzdok et al., 2012; Bzdok et al., 2015).
+baseline mental activities
+\citep{bzdok2015resting}.
 
@@
 DMN activity also persists to a substantial degree during
-sleep (Horovitz et al., 2008) and under anesthesia (Greicius et al., 2008).
+sleep and under anesthesia \citep{randy2008}.
 Today, many authors believe that the DMN implements
@@
 been shown to directly impact human behavior. Goal-directed task performance
-improved with decreased activity in default mode areas (Weissman et al., 2006)
+improved with decreased activity in default mode areas \citep{weiss2006}
 and increased DMN activity was linked to more task-independent,
-yet sometimes useful thoughts (Mason et al., 2007; Seli et al., 2016).
+yet sometimes useful thoughts
+\citep{mason2007, seli2016mind}.
 %
@@
 are responsible for the highest metabolic turn-over
-of glucose energy consumption in humans (Raichle).
+of glucose energy consumption in humans \citep{raichle2001pnas}.
 The PMC myelinates relatively late during postnatal development in monkeys
-(Goldman-Rakic, 1987), generally considered to
+\citep{goldman1987development}, generally considered to
 be a sign of phylogenetic sophistication (Flechsig, 1920).
@@
 phenomena of changed conscious awareness,
-including anesthesia (Fiset et al., 1999),
-sleep (Maquet, 2000), and forms of coma (Laureys et al., 1999).
+including anesthesia,
+sleep, and forms of coma \citep{cavanna2006precuneus}.
 %
@@
 the environmental statistics and its internal representation
-as an "inner minds eye" (Cavanna et al., 2006).
+as an "inner minds eye" \citep{cavanna2006precuneus}.
 For instance, B\'alint's syndrome is a neurological disorder of conscious
-awareness that results from damage in the bilateral parietal cortex.
+awareness that results from damage in the bilateral parietal cortex
+\citep{balint1909seelenlahmung}.
 Patients are plagued by an
@@
 (rather than centromedial or superficial) nuclei group
-of the amygdala (Bzdok et al., 2013b).
+of the amygdala \citep{bzdok2015subspecialization}.
 This amygdalar subregion has been proposed to
 continuously scan environmental input
-for biological significance assessment
-(Adolphs, 2010; Aggleton et al., 1980; Bzdok et al., 2011;
-Ghods-Sharifi et al., 2009; LeDoux, 2007).
+for biological significance assessment.
 %
@@
 electrophysiological recordings in animals implicated the PMC in
-strategic selection (Pearson et al., 2009), risk assessment (McCoy and Platt, 2005),
-and outcome-contingent behavioral modulation (Hayden et al., 2008),
+strategic selection \citep{pearson2009neurons},
+risk assessment \citep{mccoy2005risk},
+and outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
 while its retrosplenial portion was
-more specifically implicated in approach-avoidance behavior (Vann et al., 2009).
+more specifically implicated in approach-avoidance behavior
+\citep{vann2009does}.
 Neuron spiking activity in the PMC allowed distinguishing
 whether a monkey would persue an exploratory or exploitative
-behavioral strategy in a complex food foraging task (Pearson2009).
+behavioral strategy in a complex food foraging task \citep{pearson2009neurons}.
 Single-cell recordings in the monkey PMC
 demonstrated this brain region's sensitivity to
-subjective target utility (McCoy and Platt, 2005) and integration
-across individual decision-making instances (Pearson et al., 2009).
+subjective target utility \citep{mccoy2005risk} and integration
+across individual decision-making instances \citep{pearson2009neurons}.
 This DMN node encoded the
@@
 the DMN has been shown to gather and integrate information
-over different paragraphs in an fMRI study on auditory narratives (Simony 2016).
+over different paragraphs in an fMRI study on auditory narratives
+\citep{simony2016dynamic}.
 %
@@
 of various kinds of sensory information
-in rabbits (Gabriel and Talk, 2001) and humans
-(Rudge and Warrington, 1991; Valenstein et al., 1987).
+in rabbits and humans
+\citep{vann2009does}.
 Regarding perspective frames, this PMC subregion has been
@@
 in animals and humans
-(Burgess, 2008; Epstein, 2008; Valiquette and McNamara, 2007).
+\citep{epstein2008parahippocampal, burgess2008spatial, valiquette2007different}.
 %
@@
 sensory-independent top-down pathways.
-This DMN node has been described as a “mental sketchpad” (Goldman-Rakic, 1996),
+This DMN node has been described as a “mental sketchpad”
+\citep{goldman1996prefrontal},
 potentially implicated in de-novo generation and binding
@@
 are known to adapt to novel situations and stimuli with difficulty
-(Stuss, D. T. \& Benson, D. F. The Frontal Lobes, Raven, New York, 1986).
+\citep{stuss1986frontal}.
 Specifically, neural activity in human dmPFC
-reflected expectations about other peoples actions and errors thereof
-(Suzuki et al., 2012; van Schie et al., 2004).
+reflected expectations about other peoples actions and errors thereof.
 dmPFC activity indeed explained the proficiency decline
-of inferring other peoples' thoughts in the elderly (Moran et al., 2012).
+of inferring other peoples' thoughts in the elderly \citep{moran2012social}.
 Some dmPFC neurons in macaque monkeys exhibited a preference
 for processing others', rather than own, behavior
-with fine-grained adjustment of contextual circumstances (Yoshida et al., 2011).
+with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
 Also the topographically neighboring dorsal anterior cingulate cortex
@@
 persisting a behavioral plan versus switching the
-environmental context in several lesion studies (Rushworth NN 2016).
+environmental context in several lesion studies \citep{kolling2016value}.
 %
@@
 whereas the vmPFC activity models value in the environment that leads performed actions
-{Nicolle, 2012 \#3750}.
+\citep{nicolle2012agent}.
 
@@
 a gross impairment in value-based decision making
-(Glaescher et al., 2016).
+\citep{glascher2012lesion}.
 The vmPFC is preferentially connected with limbic and reward-related areas.
@@
 with the nucleus accumbens
-in axonal tracing studies in monkeys (Haber et al., 1995; Ferry et al., 2000).
+in axonal tracing studies in monkeys \citep{haber1995orbital}.
 The NAc is thought to be linked to reward mechanisms that not
 only modulate motivated behavior towards basic survival needs but also
-subserve reinforcement learning in humans more broadly
-(O'Doherty 2015).
+subserve reinforcement learning in humans more broadly.
 This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys
-(Croxson et al., 2005) that
+\citep{croxson2005quantitative} that
 quantified this brain regions to
@@
 as a bioregulatory disposition governing cognition
-and decision making (Damasio, 1996; Nauta, 1971)
+and decision making \citep{damasio1996somatic}
 In neuroeconomic studies of human decision making,
 the vmPFC consistently reflects an individual’s subjective
-valuation (Behrens et al., 2008; Boorman et al., 2009; Knutson et al., 2005; Plassmann et al., 2007)
+valuation
+\citep{behrens2008associative}.
 This may be why performance within and across participants
@@
 correlated with indices of social reward attitudes and
-value-guided behavior (Lebreton et al., 2009).
+value-guided behavior \citep{lebreton2009automatic}.
 Additionally,
@@
 among the most complicated decision that humans and monkeys take
-(Lebreton et al., 2009; Powell et al., 2010; Lewis et al., 2011; Sallet et al., 2011).
+\citep{behrens2009computation}.
 %
@@
 The HC is well known to be involved in memory and
-spatial navigation in animals and humans
-(von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
+spatial navigation in animals and humans.
 Its highly recursive anatomical architecture
@@
 there is now increasing evidence for a role
-in constructing mental models in general (Schachter/Buckner;
-Maguire 2016).
+in constructing mental models in general
+\citep{maguire2016, schacter2007remembering}.
 Indeed,
@@
 but also thinking about one’s own future and
-imagining new fictitious experiences more broadly (Hassabis/PNAS).
+imagining new fictitious experiences more broadly \citep{hassabis2007patients}.
 Mental scenes created by hippocampus-lesioned patients exposed a lack of
@@
 were shown to exhibit increased grey matter volume in the
-posterior hippocampus (Maguire et al. 2006b).
+posterior hippocampus \citep{maguire2000navigation}.
 But encoding and generative reconstruction in the hippocampus extends
@@
 their modification of input-free self-generated patterns
-after environmental events (buszaki2004 theta oscillations).
+after environmental events \citep{buzsaki2004large}.
 Specific spiking sequences, that were elicited in experimental task conditions,
@@
 featured reoccurrence directly after experimental trials
-as well as directly before upcoming experimental trials (Diba + Buzsaki 2007).
+as well as directly before upcoming experimental trials \citep{diba2007forward}.
 Such hippocampal ensemble burts during rest and sleep
@@
 recollection of autobiographical memory episodes and other
-reexperienced or newly generated mental scenarios (Hassabias2007).
+reexperienced or newly generated mental scenarios
+\citep{hassabis2007patients}.
 %
@@
 rich mental scene construction. In this view, the HC my even influence
-ongoing perception in the current environment (Maguire 2016).
-(TODO: relate to self-play in deepmind Go).
-(TODO: Kai had an interesting idea: the deja-vu effect
-could be when you already accidentally perturbed past
-such that it really occures in reality later.)
+ongoing perception in the current environment
+\citep{maguire2016}.
+% (TODO: relate to self-play in deepmind Go).
+% (TODO: Kai had an interesting idea: the deja-vu effect
+% could be when you already accidentally perturbed past
+% such that it really occures in reality later.)
 
@@
 according to their cytoarchitectonic borders and gyrification pattern
-(Caspers et al., 2008; Caspers et al., 2006).
+\citep{seghier2013angular}.
 Neuroscientific investigations on hemispheric functional specialization
 have converged to the right versus left cerebral hemisphere as dominant for
-attention versus language functions (Broca 1865, Wernicke 1874, Lichtheim 1885;
-Gazzaniga, Bogen et al. 1965, Sperry 1982, Stephan, Marshall et al. 2003).
+attention versus language functions.
 %
@@
 sensorimotor control by integrating supramodal stimulus-guided attention
-(Corbetta et al., 2002).
+\citep{corbetta2002control}.
 Involvement of this DMN node was repeatedly reported in
@@
 visual, auditory, or tactile stimulation in a multi-modal fMRI study
-(Downar et al., 2000).
+\citep{downar2000multimodal}.
 In humans, direct electrical stimulation of the
 RTPJ during neurosurgery was associated with altered perception
-and stimulus awareness (Blanke et al., 2002).
+and stimulus awareness \citep{blanke2002neuropsychology}.
 %
@@
 as a ‘‘circuit breaker’’ that recalibrates control and maintenance systems
-(Bzdok et al., 2013; Corbetta et al., 2008).
+\citep{bzdok2013tpj, corbettashul2008}.
+% (Bzdok et al., 2013; Corbetta et al., 2008).
 Indeed, patients with right TPJ damage have particular difficulties
-with multistep actions (Hartmann et al. 2005).
+with multistep actions \citep{hartmann2005takes}.
 In the face of large discrepancies between actual and previously predicted
@@
 Transient RTPJ in humans for instance disrupted the
-impact of predicted intentions of other individuals (Young et al. 2010b),
+impact of predicted intentions of other individuals
+\citep{young2010disruption},
 a capacity believed to be subserved by the DMN.
@@
 to deal instead with immediate environmental objects and contexts.
-(TODO: slow endogeneous changes in state transitions by the policy and value matrices)
+% (TODO: slow endogeneous changes in state transitions by the policy and value matrices)
 
@@
 without obvious relation to processing any language content
-(Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
+\citep{varney1987locus}
+% (Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
 %
@@
 formulated thought ("language of the mind") is closely related to the left TPJ
-(Geva et al., 2011).
+\citep{geva2011neural}.
 A number of neuroimaging studies demonstrated LTPJ activity responses

### 2016-10-26 paper/paper.tex
improve the hell out of it

@@
 \begin{abstract}
-The default mode network (DMN) is the neural representation of the human
+The default mode network (DMN) is the brain manifestations of the human
 baseline mental activity.
@@
 \textit{how} the DMN may actually implement continuous
-environmental assessment and prediction.
+environmental assessment and prediction to guide action choices.
 DMN function is recast in mathematical terms
@@
 What has early been described as the "stream of consciousness"
-in psychology \citep{james1890principles},
+in psychology \citep{james1890principles}
 found its neurobiological manifestation
@@
 \citep{shul1997}. This networked collection of some of the highest regions in
-the human association cortex \citep{mesulam1998sensation}
+the human association cortex \citep{mesulam1998sensation, margulies2016situating}
 consistently
@@
 This is all the more likely because the DMN contains the two
-biggest hotspots of energy consumption in the entire central nervous system.
+biggest hotspots of energy consumption in the entire central nervous system
+\citep{raichle2001pnas}.
 DMN activity also persists to a substantial degree during
@@
 Understanding of the overarching DMN function is complicated by
-simultaneously controlling perception–action cycles and
+simultaneously controlling perception-action cycles and
 maintaining baseline contemplations
@@
 \citep{tenenbaum2011grow}.
-\textit{Control} refers to the influences an agent exerts when interacting
+\textit{Control} refers to the influences that an agent exerts when interacting
 with the environment to encourage preferred states.
@@
 quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
-cumulative version which takes into account the future. A standard for
+cumulative reward estimate which takes into account the future. A standard for
 modeling this accumulation is the so-called
@@
 the delay discounting behavior of the intelligent agent.
+It is also important to appreciate that a stochastic policy estimation
+is more advantageous in many RL settings.
+For instance, a deterministic strategy in playing rock-paper-scissors can
+be easily exploited and a uniformly random choice is optimal.
 
@@
 top-down modulation at various hierarchical processing levels.
-In fact,
-back projections probably outnumber by far the existing input projection in the brain
-\citep{salin1995corticocortical}.
+In fact, axonal
+back projections probably outnumber by far the input projection
+existing in the brain \citep{salin1995corticocortical}.
 This can explain the constructive, generative nature of sensory perception
@@
 and internally predicted sensation,
-incurs plasticity changes of neuronal back-projections (i.e., model parameters)
+incurs plasticity changes of neuronal back projections (i.e., model parameters)
 for gradually improved future prediction of the environment.
-Contextual integration is maintained by top-down modulation of sensorimotor
+Contextual integration is hence maintained by top-down modulation of sensorimotor
 processing by a-priori information about the environment.
-The generative model of how data arise in the
-world would hence be incorporated into
+The generative model of how perceived sensory events arise in the
+world would be incorporated into
 the current neuronal wiring.
@@
 This process permits updates of the internal model of the environment
-to best fit the incoming sensory examples.
+to best fit the constant influx of sensory examples.
 
@@
   generalizing from the past to new experiences.
-  \item Both expose a mechanisms of brain plasticity in that
+  \item Both provide a parsimonious explanation why the
+  human brain decreases processing load devoted to incoming information
+  when the environment is predictable and allocate increased
+  attentional resources when novel stimuli are encountered.
+  Similarly, both propose explanations why
+  predicting the future is inherently
+  linked to information from the past.
+  \item Sensory experience is a generative process from both views.
+  In predictive coding, sensory perception of the external world
+  is a generative experience due to modulatory top down experience at
+  various hierarchical levels of sensory input processing.
+  In our RL view, the (partially observed)
+  environmental model is incorporated in the MDP,
+  \textit{which can be fully recovered based on the last
+  state alone}.
+  \item Both naturally expose a mechanisms of brain plasticity in that
   neuronal wiring gets increasingly adapted
@@
   \item Both model the consequences of action. In RL, the horizon of that
-  look into the future is manifested in the $\gamma$ parameter
+  look into the future is explicitly manifested in the $\gamma$ parameter
   in the Bellman equation, but not
   explicitly modeled in the PC account that
-  mostly emphasizes prediction about the
+  tends to emphasize prediction about the
   immediate future.
@@
   discounting of future rewards.
-  \item Sensory experience is a generative process from both views.
-  In predictive coding, sensory perception of the external world
-  is a generative experience due to modulatory top down experience at
-  various hierarchical levels of sensory input processing.
-  In our RL view, the (partially observed)
-  environmental model is incorporated in the Markov decision
-  process, \textit{which can be fully recovered based on the last
-  state alone}.
   \item Adapting the neuronal connections for improved top-down modulation
@@
   although this may not be neurobiologically plausible.
-  \item Both provide a parsimonious explanation why the
-  human brain decreases processing load devoted to incoming information
-  when the environment is predictable and allocate increased
-  attentional resources when novel stimuli are encountered.
-  Similarly, both propose explanations why
-  predicting the future is inherently
-  linked to information from the past.
+  \item In sum,
+  the PC account in cognitive neuroscience does not expose an explicit
+  mechanism of the horizon of future considerations or
+  how the internal model of the world gets updated,
+  and collapses sensory input processing and action output preparation.
 \end{itemize}
@@
 \subsection{Semantic Hypothesis}
-Another frequently proposed account of DMN function revolves
-around forming meaningful associations and analogies between
+Another frequently proposed cognitive account of DMN function revolves
+around forming logical associations and analogies between
 the current experience and
-the abstract knowledge derived from past experiences
+the conceptual knowledge derived from past experiences
 \citep{bar2007proactive, binder1999conceptual}.
 Analogies might naturally tie incoming novel sensory stimuli to
-explicit world knowledge extractable from the environment
+explicit world knowledge (i.e., semantics) extractable from the environment
 \citep{bar2009proactive}.
 %
-Going much beyond language ifself,
-semantic analogies can be used to mentally envision non-existent scenarios
+Going much beyond human language ifself,
+semantic building blocks provide the basis to mentally envision non-existent scenarios
 that would improve optimal behavior in the environment
@@
 items \citep{addis2008age, spreng2006temporal}.
-A constructive account further explains the reciprocal relationship
+Further,
+a constructive account explains the reciprocal relationship
 between an egocentric first person perspective and
@@
 %
-It is also closely related to episodic memory, language, problem solving,
-planning, estimating others' thoughts, and spatial navigation
+Cognitive aspects of egocentric-allocentric switching
+are also closely related to episodic memory, language, problem solving,
+planning, estimating other people's thoughts, and spatial navigation
 as these necessitate abstract world knowledge and abstract assocations
@@
 \begin{itemize}
-  \item The notion of semantic or knowledge association is
-  incorporated into the MDP as the Markov property,
-  that is, the current state directly results from the
-  agent's history of states and actions. The learned
-  value matrix and action transition schedules drive
-  stimulus processing and action choice in the present.
+  \item Both the semantic hypothesis and MDPs expose mechanisms of
+  how alternative decision trees could be mentally explored.
   \item In both semantic hypothesis and MDPs,
@@
   thinking \citep{hassabis2007patients}.
-  \item Both the semantic hypothesis and MDP expose mechanisms of
-  how alternative decision trees could be mentally explored.
+  \item The notion of semantic or knowledge association is
+  incorporated into the MDP as the Markov property,
+  that is, the current state directly results from the
+  agent's history of states and actions. The learned
+  value matrix and action transition schedules drive
+  stimulus processing and action choice in the present.
   \item While mental scene construction implies a distinction between
@@
   followed by a return to perception of the present.''\citep{buckner2007self}},
-  the MDP naturally integrates the former egocentric
-  (related to current action, state and reward) and the later
-  allocentric (related to past and future actions, states, and rewards)
+  the MDPs naturally integrate the former egocentric
+  (more related to current action, state and reward) and the later
+  allocentric (more related to past and future actions, states, and rewards)
   angles on the world in a same optimization problem.
   \item The semantic account of DMN function does not offer
-  a mechanistic explanation how semantic world knowledge and analogies thereof
+  a mechanistic explanation how explicit world knowledge and semnatic analogies thereof
   lead to prediction of future actions and states.
@@
   in MDPs provides a candidate explanation for why humans
-  have evolved imperfect memories as the better adaptation.
+  have evolved imperfect noisy memories
+  as the more advantageous adaptation.
   \item While both semantic and MDP account propose memory-based internally
@@
   \item In sum, episodic scene construction according to the semantic
-  hypothesis is lacking an explicit time and incentive model.
+  hypothesis is lacking an explicit time and incentive model;
+  neither does it explain how randomness of human mental experience
+  can be beneficial.
 \end{itemize}
@@
 physical environment, but to form and maintain increasingly complex
-social systems to solve ecological problems by means of social connections
+social systems to solve ecological problems by means of social relationships
 \citep{whiten1988machiavellian}.
@@
     \item In sum,
-    MDPs imply social thinking as a physiological brain baseline
-    as a special case, without excluding other topics.
+    MDPs imply a ``radar'' function of montoring the environment for salient information
+    in general and explain social thinking as a physiological brain baseline
+    as an important special case, without excluding other topics
+    contemplated in continuous mental activity.
  \end{itemize}
@@
 might implement supra-modal representation and control of the environment to
-optimize the organism's fate by
-obtaining reward and avoiding pain.
+optimize the organism's fate.
+% by obtaining reward and avoiding pain.
 This idealized process model explains
 a number of previous experimental observations in the
-DMN literature by a simple but non-trivial mechanism.
+DMN by a simple but non-trivial mechanism.
 %
@@
 optimal substructure in the environment can be recursively exploited
-when confronted with nested decisions.
+when confronted with complicated decisions.
 Algorithmic gains by injecting stochasticity in the recall of past
-behavior and outcomes to improve the internal world model may explain why
-very accurate memories have been evolutionarily disfavored
+actions and outcomes to improve the internal world model may explain why
+very accurate memories have been disfavored in human evolution
 and why human creativity might be adaptive.

### 2016-11-09 paper/paper.tex
more
added stuff on inverse reinforcement-learning

@@
 
+\subsection{Inverse reinforcement-learning}
+\begin{mdframed}
+  Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
+  a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
+  (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
+  \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
+  (e.g linear rewards $r(s,a) = \theta^T\phi(s,a)$)?
+  For example, given traces of motor actions from a an adult to grab a cup on a table,
+  can an observing child figure out what ``langragian'' functional is being minimized by the former ?
+  How ? Can they reproduce this optimal behavior ?
+\end{mdframed}
+
+Such questions are of course pertinent to our decision-making theory for the DMN.
+In the general artificial case, the problem has been extensively studied and partially solved by
+\cite{abbeel2004}. They are also been rigourously studied in general optimal control literature under the name ``inverse optimal control'', but in model-based certains (where the physical dynamics
+are known, etc.)...
+
 \section{Relation to other statistical models (Elvis)}
@@
 %% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
-so that one recognizes The last quantity in \eqref{eq:helm} as the the \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
+The last quantity in \eqref{eq:helm} is nothing but \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
 \begin{equation}

### 2016-11-13 paper/paper.tex
changed template to PLOS + some fixups

@@
-\documentclass{article} % For LaTeX2e
-\usepackage{nips14submit_e,times}
-\usepackage{hyperref}
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% Use Unicode characters when possible
+%\usepackage[utf8x]{inputenc}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Use the PLoS provided BiBTeX style
+%\bibliographystyle{plos2015}
+
+%% \documentclass{article} % For LaTeX2e
+%% \usepackage{nips14submit_e,times}
+%% \usepackage{hyperref}
 \usepackage{url}
@@
 \usepackage{multirow}
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
+\date{}
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+\setlength{\headheight}{27.023pt}
+\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\sf PLOS}
+
+%% Include all macros below
+
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
 
@@
 % \nipsfinalcopy % Uncomment for camera-ready version
-\nipsfinaltrue
+%\nipsfinaltrue
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \author{Elvis Dohmatob, Bert Kappen, Danilo Bzdok\\
-  INRIA, Parietal team, Saclay, France\\
-  CEA, Neurospin, Gif-sur-Yvette, France\\
-  firstname.lastname@inria.fr}
+  %% INRIA, Parietal team, Saclay, France\\
+  %% CEA, Neurospin, Gif-sur-Yvette, France\\
+  %% firstname.lastname@inria.fr
+}
 
@@
   \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
-  (e.g linear rewards $r(s,a) = \theta^T\phi(s,a)$)?
+  (e.g linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
   For example, given traces of motor actions from a an adult to grab a cup on a table,

### 2016-11-15 paper/paper.tex
some goodies

@@
 %\bibliographystyle{plos2015}
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+\setlength{\headheight}{27.023pt}
+\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\sf PLOS}
 
@@
 
-\section{Methods}
-
-\subsection{Linear control network model for brain organization}
-Consider the following linear time-invariant (LTI) dynamical system as a (toy) model for high-level brain function
-\begin{equation}
-  \dot{\x}(t) = \A\x(t) + \B\u(t).
-  \label{eq:lti}
-\end{equation}
-Here, the $n$-by-$n$ matrix $\A$ denotes a model of the brain's wiring (for example a resting state connectome computed from an anatomical atlas), while the $n$-by-$k$ ``input matrix'' $\B$ describes which nodes can be controlled by us, an external \textit{controller}, via medical intervention or a careful choice of stimulus presentation, for example.
-At time $t \ge 0$, let  the $n$-vector  $\x(t) := (\x_1(t),\ldots,\x_n(t))$ encodes the state of the network (one value for each node).
-The aim is to supply values of $\u_1(t),\ldots,\u_k(t)$ for $k \le n$ controls as a function of the time $t$,  to take the system from any prescribed initial state $\x(0) = \x^{\text{init}} \in \mathbb R^n$ to any prescribed final state $\x(\tau) = \x^{\text{fin}} \in \mathbb R^n$ in finite time $\tau < \infty$.
-When such a controlling is possible, we say that the system $(\A,\B)$ is \textit{controllable}.
-A precise sufficient and necessary condition for such controllability is the Kalman condition: $(\A,\B)$ is controllable iff the $n \times nk$ \textit{controllability matrix}
-\begin{equation}
-  \C := (\B|\A\B|\ldots|\A^{n-1}\B)
-\end{equation}
-has full rank, i.e
-\begin{equation}
-  rank(\C) = n.
-\end{equation}
-
-\paragraph{How many controls do we need at best ?}
-In the thermodynamic limit ($n \rightarrow \infty$), statistical physics \citep{Liu2011} gives the extremely good estimate
-\begin{equation}
-  n_0(\langle k\rangle, \gamma) \approx \exp\left(-\frac{1}{2}\left(\frac{\gamma-2}{\gamma - 1}\right)\langle k \rangle\right),
-\end{equation}
-where $\gamma \ge 1$ is the \textit{scale-free} parameter for the node-degree distribution of the network described by $\A$, and $\langle k \rangle$ is the mean degree.
-Letting $\gamma \rightarrow \infty$, one recovers the Erdos-Renyi value
-\begin{equation}
-  n_0(\langle k\rangle) = \exp\left(-\frac{1}{2}\langle k\rangle\right),
-\end{equation}
-which agrees perfectly with its known analytic value. One identifies two radically different regimes:
+\section{Free-energy principles!}
+The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
+
+In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
+This ideas were first developed by Hinton et al.
+in the early 90s in building their Helmholtz machine.
+Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises.
+For example the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial, e.g) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
+
+\begin{mdframed}
+  Though our presentation of free-energy principles is thermodynamic in flavor, a completely equivalent
+  information-theoretic formulation is of course possible, simply by exploiting the information-theoretic interpretation of entropy.
+\end{mdframed}
+
+\subsection{Helmholtz free-energy and the generative model}
+\begin{table}[H]
+  \begin{tabular}{p{2cm}|p{11cm}}
+         \hline
+         \textbf{symbol}    & \textbf{description}  \\ \hline
+         $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
+         random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
+         $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
+          with the usual convention $0 \log(0) := 0$.\\ \hline
+         $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
+             $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e $\x = (s, b).$ \\ \hline
+             $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
+             $p_G(.|\x)$ & Generative density for ...\\ \hline
+         $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
+         $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
+         $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
+          to $R$.  Note that $F^G_G = F_G$.\\ \hline
+  \end{tabular}
+  \caption{Table of notations.}
+\end{table}
+Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
+So,
+\begin{eqnarray}
+  \begin{split}
+    \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(P_G(\x))\sum_{\z}p_G(\z |\x)\\
+    &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
+    =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
+    &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
+    &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
+    &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
+  \end{split}
+  \label{eq:helm}
+\end{eqnarray}
+where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
+\begin{equation}
+  E_G(\z, \x) := -\log(p_G(\z, \x)),
+  \label{eq:gibbs}
+\end{equation}
+%% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
+The last quantity in \eqref{eq:helm} is nothing but \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
+\begin{equation}
+  F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
+\end{equation}
+%% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
+Thus,
+\begin{fact}
+  Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
+\end{fact}
+
+The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
+It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(\x)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
+A workaround is then to introduce a second density $p_R(.|\x)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\x)$, as a trick for doing approximate inference.
+The former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
+This primal-dual idea, first proposed in Hinton et al.
+1995, is at the heart of the general free-energy principle that we'll introduce shortly.
+
+\subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
+In this subsection, we'll present an insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
+As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
+
+Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
+distribution
+\begin{equation}
+  p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
+\end{equation}
+where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
+Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
+\begin{equation}
+F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
+\end{equation}
+Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
+it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
+density\footnote{conditioning in $P_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
+  state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
+\begin{equation}
+  \begin{split}
+    F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{P_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
+    &= \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
+    &= F^R_G(\x) - D_{KL}(P_R(.|\x) || P_G(.|\x)),
+  \end{split}
+  \label{eq:fe}
+\end{equation}
+where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
+\begin{equation}
+  F^R_G(\x) := \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x))
+\end{equation}
+and $D_{KL}(P_R(.|\x) || P_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
+
+\subsection{A general free-energy principle}
+We can resume the situation as follows:
+\begin{mdframed}
+\begin{equation}
+  \begin{split}
+    \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
+    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
+    &\le F^R_G(\x),
+    \text{ with equality if }p_R(.|\x) = p_G(.|\x),
+    \end{split}
+\end{equation}
+\end{mdframed}
+where we have used the fact that KL divergence is always nonnegative.
+%% In fact, we have the following theorem
+%% \begin{theorem}
+%%   It holds that
+%%   $$
+%%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
+%%   i.e
+%%   \begin{equation}
+%%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
+%%     \end{equation}
+%% \end{theorem}
+
+\subsection{Helmholtz machines and the wake-sleep algorithm}
+% \begin{mdframed}
+  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e input from lower more primitive layers).
+Precisely
+$$
+p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
+p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
+$$
+% \end{mdframed}
+
+\subsection{Friston's active-inference}
+\label{sec:friston}
+This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
+
+The following critics can be made:
 \begin{itemize}
-  \item the manageable regime $\gamma > \gamma_c := 2$, in which $n_0 < 1$, and so we only need fewer controls than total number of nodes, and
-\item the unmanageable regime ``$\gamma \le \gamma_c$'', in which  $n_0 = 1$, where each node must be controlled explicitly.
+  \item
+    As noted by Dayan et al.
+(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
+  \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
+  \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
+    technical difficulty in getting derivatives of loss function w.r.t
+    recognition weights $\W^R$).
+  \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
     \end{itemize}
 
-\paragraph{Why linear dynamics ?}
-As explained in \citep{Liu2011}, the choice of linear dynamics over a more general nonlinear dynamics can be justified as follows:
+\subsection{Minimizing free-energy via backprop!}
+
+% \subsection{Free-energy minimization via variational auto-encoders}
+\label{sec:vae}
+Here, we present a way to aleviate some conceptual and computational issues with the
+free-energy framework presented thus far, by using auto-encoder theory.
+Following Kingma et al. 2014 \citep{kingma2013auto}, define the data-dependent auxiliary random
+function
+\begin{eqnarray}
+  f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
+\end{eqnarray}
+Then we can rewrite the
+variational free-energy as
+\begin{eqnarray*}
+  \begin{split}
+    F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
+    &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
+    &= -\langle f_{G,R}\rangle_{P_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
+    \end{split}
+\end{eqnarray*}
+
+\begin{mdframed}
+  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
+\end{mdframed}
+\begin{figure}[!tb]
+  \centering
+  \includegraphics[width=.7\linewidth]{vae.pdf}
+  \caption{Variational autoencoders...}
+  \label{fig:vae}
+\end{figure}
+
+\paragraph{Solution: The reparametrization trick.}
 \begin{itemize}
-  \item  Conclusions drawn from linear dynamics can be
-extended to nonlinear systems.
-\item If the controllability matrix of the linearized system
-has full rank at all points, then it is sufficient for most
-systems to say that the actual nonlinear system is
-controllable (i.e. small signal model).
+\item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
+\item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
+  \begin{itemize}
+  \item results in a sample $\z \sim p_R(.|\x)$, from the correct
+    posterior
     \end{itemize}
-
-\paragraph{A note on stability.}
-For stability in the LTI model \eqref{eq:lti} and hence in the constrained path integral \eqref{eq:hj} below, a sufficient (and necessary ?) condition is that all eigenvalues of the $A$ have negative real-part.
-One way to impose this is to add self-loops with a small negative weight.
-\paragraph{Meta brain.}
-
-\subsection{Optimal control: an LQR feedback controller}
-We propose to use a linear quadratic regulator (LQR) for the feedback controller.
-Thus, consider the time-varying \text{value function} $V: [0, \tau] \times \mathbb R^n \rightarrow \mathbb R$, defined by the following Hamilton-Jacobi cost-to-go
-\begin{equation}
-  \begin{split}
-    &V(t, \z) := \min_{\u} \frac{1}{2}\x_{\text{fin}}^T\Q_{\text{fin}}\x_{\text{fin}} + \int_{0}^\tau\left(\frac{1}{2}\x(t)^T\Q\x(t) + \frac{1}{2}\u(t)^T\R\u(t)\right)dt,\\
-    &\text{subject to } \x(t) = \z, \dot{\x}(t) = \A\x(t) + \B\u(t),
-  \end{split}
-  \label{eq:hj}
-\end{equation}
-where the matrices $\Q$ and $\R$, precised by design considerations and subject to meta-optimization, are restricted to be positive semi-definite and positive definite, respectively.
-A classical calculation (reminiscent of the \textit{Pontryagin minimization principle}) reveals that the optimal feedback control in \eqref{eq:hj} is given by
-\begin{equation}
-  \u(t) = -\K(t)\x(t),
-\end{equation}
-where
-\begin{equation}
-  \K(t) := \R^{-1}\B^T\P(t)\text{ (Kalman gain matrix)}
-\end{equation}
-and the time-varying positive semi-definite matrix $\P(t)$ solves following differential Riccati equation (DRE)
-\begin{equation}
-\begin{split}
-&-\dot{\P}(t) = \A^T\P(t) + \P(t)\A - \P(t)\B\R^{-1}\B^T\P(t) + \Q\\
-&\text{subject to }\P(\tau) = \Q_{\text{fin}}.
-\end{split}
-\end{equation}
-%% \subsection{Minimal control energy controller}
+\end{itemize}
+The result is a scheme for training DBNs via good-old backprop! 
+Refer to Fig. \ref{fig:vae}. Some examples of the reparametrization trick for a number of
+choices of the posterior distribution are given in Tab. \ref{tab:rptrick}.
+
+\begin{table}[H]
+  \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
+         \hline
+         Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
+         Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
+         Student’s t, Logistic, Uniform, Triangular, ... \\ \hline
+         Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
+\\ \hline
+         Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
+  \end{tabular}
+  \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
+  \label{tab:rptrick}
+\end{table}
+
+%% For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
+%% \begin{equation}
+%%   p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
+%% \end{equation}
+
+%% \subsection{A thermodynamic model for bounded rationality, aka robust
+%%   optimality}
+%% \begin{itemize}
+%%   \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
+%%   \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
+%%   \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
+%%     \end{itemize}
+%% The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
+%% \paragraph{Utility functions and conjugate pairs.}
+%%   Let $(\Omega, \mathcal F, P)$ be a probablity space.
+A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
+%%   \begin{itemize}
+%%   \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.
+%%   \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
+%%     $A,B \in \mathcal F$.
+%%   \item monotonicity: $f_{U}$ is strictly increasing.
+%%     \end{itemize}
+
+%% \begin{theorem}
+%%   The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
+%%   \begin{equation}
+%%     f = \alpha \log(.),
+%%     \label{eq:boltzmann}
+%%   \end{equation}
+%%   where $\alpha > 0$.
+%% \end{theorem}
+
+%% \textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
+
+%% Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
+
+%% \paragraph{Example.}
+%% Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
+%%   \begin{equation}
+%%     P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
+%%   \end{equation}
+%%   where
+
+%%   \begin{equation}
+%%     Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
+%%   \end{equation}
+%%   is a normalization constant called the \textit{partition function} of $U$.
+%%   It's not hard to see that $U$ and the $P$ above form a conjugate pair.
+
+%%   \paragraph{Free-utility functional.}
+%%     Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
+Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
+%%     \begin{equation}
+%%       J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
+%%     \label{eq:free_u}
+%%     \end{equation}
+%%     where
+%%     \begin{equation}
+%%       \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
+%%     \end{equation}
+%%     is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
+%%   \begin{equation}
+%%     J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
+%%   \end{equation}
+%%   In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
+%%   \begin{equation}
+%%     J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
+%%     \end{equation}
+
+%%   \paragraph{The free-energy / utility principle (of Friston ?).}
+%%   We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
+Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
+We want to transform this to a new model by adding constraints represented by the utility function $\Delta U$.
+The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \Delta U$.
+The difference in free-utility is then
+%%   \begin{equation}
+%%     \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
+%%     \label{eq:free}
+%%   \end{equation}
+%%   where
+%%   \begin{equation}
+%%     D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
+%%   \end{equation}
+%%   is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
+%% In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
+%%   The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
+%% a trade-off between an expected utility $\Delta U$ (first term) and
+%% the information cost of transforming Pi
+%% into $P_{\text{fin}}$ (second
+%% term). In this interpretation $P_{\text{ini}}$ represents an initial choice
+%% probability or policy, which includes the special case of the
+%% uniform distribution where the decision-maker has initially no
+%% preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
+%% considers the utility constraint U∗
+%% that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
+%% \begin{equation}
+%% P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
+%%   \end{equation}
+
+%% By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
+
+%% \begin{equation}
+%%   P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
+%% \end{equation}
+
+%% Two limit cases are worth considering.
+%% \begin{itemize}
+%% \item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
+%%   $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
+%%   where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
+%%   \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
+%% \end{itemize}
+
+%% To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
+For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...
 
@@
 \subsection{Inverse reinforcement-learning}
-\begin{mdframed}
+%\begin{mdframed}
   Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
@@
   How ? Can they reproduce this optimal behavior ?
-\end{mdframed}
+%\end{mdframed}
 
@@
 
+IRL is is suited for problems in which it's hard to define what the reward function could be
+(e.g car-driving, drone maneuvering, etc.
+
+\section{Optimal control}
+\subsection{Linear control network model for brain organization}
+Consider the following linear time-invariant (LTI) dynamical system as a (toy) model for high-level brain function
+\begin{equation}
+  \dot{\x}(t) = \A\x(t) + \B\u(t).
+  \label{eq:lti}
+\end{equation}
+Here, the $n$-by-$n$ matrix $\A$ denotes a model of the brain's wiring (for example a resting state connectome computed from an anatomical atlas), while the $n$-by-$k$ ``input matrix'' $\B$ describes which nodes can be controlled by us, an external \textit{controller}, via medical intervention or a careful choice of stimulus presentation, for example.
+At time $t \ge 0$, let  the $n$-vector  $\x(t) := (\x_1(t),\ldots,\x_n(t))$ encodes the state of the network (one value for each node).
+The aim is to supply values of $\u_1(t),\ldots,\u_k(t)$ for $k \le n$ controls as a function of the time $t$,  to take the system from any prescribed initial state $\x(0) = \x^{\text{init}} \in \mathbb R^n$ to any prescribed final state $\x(\tau) = \x^{\text{fin}} \in \mathbb R^n$ in finite time $\tau < \infty$.
+When such a controlling is possible, we say that the system $(\A,\B)$ is \textit{controllable}.
+A precise sufficient and necessary condition for such controllability is the Kalman condition: $(\A,\B)$ is controllable iff the $n \times nk$ \textit{controllability matrix}
+\begin{equation}
+  \C := (\B|\A\B|\ldots|\A^{n-1}\B)
+\end{equation}
+has full rank, i.e
+\begin{equation}
+  rank(\C) = n.
+\end{equation}
+
+\paragraph{How many controls do we need at best ?}
+In the thermodynamic limit ($n \rightarrow \infty$), statistical physics \citep{Liu2011} gives the extremely good estimate
+\begin{equation}
+  n_0(\langle k\rangle, \gamma) \approx \exp\left(-\frac{1}{2}\left(\frac{\gamma-2}{\gamma - 1}\right)\langle k \rangle\right),
+\end{equation}
+where $\gamma \ge 1$ is the \textit{scale-free} parameter for the node-degree distribution of the network described by $\A$, and $\langle k \rangle$ is the mean degree.
+Letting $\gamma \rightarrow \infty$, one recovers the Erdos-Renyi value
+\begin{equation}
+  n_0(\langle k\rangle) = \exp\left(-\frac{1}{2}\langle k\rangle\right),
+\end{equation}
+which agrees perfectly with its known analytic value. One identifies two radically different regimes:
+\begin{itemize}
+  \item the manageable regime $\gamma > \gamma_c := 2$, in which $n_0 < 1$, and so we only need fewer controls than total number of nodes, and
+\item the unmanageable regime ``$\gamma \le \gamma_c$'', in which  $n_0 = 1$, where each node must be controlled explicitly.
+\end{itemize}
+
+\paragraph{Why linear dynamics ?}
+As explained in \citep{Liu2011}, the choice of linear dynamics over a more general nonlinear dynamics can be justified as follows:
+\begin{itemize}
+  \item  Conclusions drawn from linear dynamics can be
+extended to nonlinear systems.
+\item If the controllability matrix of the linearized system
+has full rank at all points, then it is sufficient for most
+systems to say that the actual nonlinear system is
+controllable (i.e. small signal model).
+\end{itemize}
+
+\paragraph{A note on stability.}
+For stability in the LTI model \eqref{eq:lti} and hence in the constrained path integral \eqref{eq:hj} below, a sufficient (and necessary ?) condition is that all eigenvalues of the $A$ have negative real-part.
+One way to impose this is to add self-loops with a small negative weight.
+\paragraph{Meta brain.}
+
+\subsection{Optimal control: an LQR feedback controller}
+We propose to use a linear quadratic regulator (LQR) for the feedback controller.
+Thus, consider the time-varying \text{value function} $V: [0, \tau] \times \mathbb R^n \rightarrow \mathbb R$, defined by the following Hamilton-Jacobi cost-to-go
+\begin{equation}
+  \begin{split}
+    &V(t, \z) := \min_{\u} \frac{1}{2}\x_{\text{fin}}^T\Q_{\text{fin}}\x_{\text{fin}} + \int_{0}^\tau\left(\frac{1}{2}\x(t)^T\Q\x(t) + \frac{1}{2}\u(t)^T\R\u(t)\right)dt,\\
+    &\text{subject to } \x(t) = \z, \dot{\x}(t) = \A\x(t) + \B\u(t),
+  \end{split}
+  \label{eq:hj}
+\end{equation}
+where the matrices $\Q$ and $\R$, precised by design considerations and subject to meta-optimization, are restricted to be positive semi-definite and positive definite, respectively.
+A classical calculation (reminiscent of the \textit{Pontryagin minimization principle}) reveals that the optimal feedback control in \eqref{eq:hj} is given by
+\begin{equation}
+  \u(t) = -\K(t)\x(t),
+\end{equation}
+where
+\begin{equation}
+  \K(t) := \R^{-1}\B^T\P(t)\text{ (Kalman gain matrix)}
+\end{equation}
+and the time-varying positive semi-definite matrix $\P(t)$ solves following differential Riccati equation (DRE)
+\begin{equation}
+\begin{split}
+&-\dot{\P}(t) = \A^T\P(t) + \P(t)\A - \P(t)\B\R^{-1}\B^T\P(t) + \Q\\
+&\text{subject to }\P(\tau) = \Q_{\text{fin}}.
+\end{split}
+\end{equation}
+%% \subsection{Minimal control energy controller}
+
 \section{Relation to other statistical models (Elvis)}
-\subsection{Free-energy principles!}
-The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
-
-In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
-This ideas were first developed by Hinton et al.
-in the early 90s in building their Helmholtz machine.
-Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises.
-For example the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial, e.g) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
-
-\subsubsection{Helmholtz free-energy and the generative model}
-\begin{table}[H]
-  \begin{tabular}{p{2cm}|p{11cm}}
-         \hline
-         \textbf{symbol}    & \textbf{description}  \\ \hline
-         $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
-         random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
-         $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
-          with the usual convention $0 \log(0) := 0$.\\ \hline
-         $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
-             $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e $\x = (s, b).$ \\ \hline
-             $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
-             $p_G(.|\x)$ & Generative density for ...\\ \hline
-         $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
-         $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
-         $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
-          to $R$.  Note that $F^G_G = F_G$.\\ \hline
-  \end{tabular}
-  \caption{Table of notations.}
-\end{table}
-Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
-So,
-\begin{eqnarray}
-  \begin{split}
-    \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(P_G(\x))\sum_{\z}p_G(\z |\x)\\
-    &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
-    =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
-    &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
-    &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
-    &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
-  \end{split}
-  \label{eq:helm}
-\end{eqnarray}
-where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
-\begin{equation}
-  E_G(\z, \x) := -\log(p_G(\z, \x)),
-  \label{eq:gibbs}
-\end{equation}
-%% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
-The last quantity in \eqref{eq:helm} is nothing but \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
-\begin{equation}
-  F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
-\end{equation}
-%% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
-Thus,
-\begin{fact}
-  Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
-\end{fact}
-
-The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
-It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(\x)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
-A workaround is then to introduce a second density $p_R(.|\x)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\x)$, as a trick for doing approximate inference.
-The former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams!
-This primal-dual idea, first proposed in Hinton et al.
-1995, is at the heart of the general free-energy principle that we'll introduce shortly.
-
-\subsubsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
-In this subsection, we'll present an insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
-As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
-
-Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
-distribution
-\begin{equation}
-  p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
-\end{equation}
-where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
-Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
-\begin{equation}
-F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
-\end{equation}
-Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
-it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
-density\footnote{conditioning in $P_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
-  state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
-\begin{equation}
-  \begin{split}
-    F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{P_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
-    &= \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
-    &= F^R_G(\x) - D_{KL}(P_R(.|\x) || P_G(.|\x)),
-  \end{split}
-  \label{eq:fe}
-\end{equation}
-where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
-\begin{equation}
-  F^R_G(\x) := \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x))
-\end{equation}
-and $D_{KL}(P_R(.|\x) || P_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
-
-\subsubsection{A general free-energy principle}
-We can resume the situation as follows:
-\begin{mdframed}
-\begin{equation}
-  \begin{split}
-    \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
-    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
-    &\le F^R_G(\x),
-    \text{ with equality if }p_R(.|\x) = p_G(.|\x),
-    \end{split}
-\end{equation}
-\end{mdframed}
-where we have used the fact that KL divergence is always nonnegative.
-%% In fact, we have the following theorem
-%% \begin{theorem}
-%%   It holds that
-%%   $$
-%%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
-%%   i.e
-%%   \begin{equation}
-%%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
-%%     \end{equation}
-%% \end{theorem}
-
-\subsubsection{Helmholtz machines and the wake-sleep algorithm}
-% \begin{mdframed}
-  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e input from lower more primitive layers).
-Precisely
-$$
-p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
-p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
-$$
-% \end{mdframed}
-
-\subsubsection{Friston's active-inference}
-This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
-
-The following critics can be made:
-\begin{itemize}
-  \item
-    As noted by Dayan et al.
-(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
-  \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
-  \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
-    technical difficulty in getting derivatives of loss function w.r.t
-    recognition weights $\W^R$).
-  \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
-    \end{itemize}
-
-\subsection{Minimizing free-energy via backprop!}
-% \subsection{Free-energy minimization via variational auto-encoders}
-\label{sec:vae}
-Here, we present a way to aleviate some conceptual and computational issues with the
-free-energy framework presented thus far, by using auto-encoder theory.
-Following Kingma et al. 2014 \citep{kingma2013auto}, define the data-dependent auxiliary random
-function
-\begin{eqnarray}
-  f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
-\end{eqnarray}
-Then we can rewrite the
-variational free-energy as
-\begin{eqnarray*}
-  \begin{split}
-    F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
-    &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
-    &= -\langle f_{G,R}\rangle_{P_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
-    \end{split}
-\end{eqnarray*}
-
-\begin{mdframed}
-  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
-\end{mdframed}
-\paragraph{Solution: The reparametrization trick.}
-\begin{itemize}
-\item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
-\item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
-  \begin{itemize}
-  \item results in a sample $\z \sim p_R(.|\x)$, from the correct
-    posterior
-    \end{itemize}
-\end{itemize}
-
-\begin{table}[H]
-  \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
-         \hline
-         Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
-         Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
-         Student’s t, Logistic, Uniform, Triangular, ... \\ \hline
-         Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
-\\ \hline
-         Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
-  \end{tabular}
-  \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
-\end{table}
-
-The result is a scheme for training DBNs via good-old backprop!
-
-%% For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
-%% \begin{equation}
-%%   p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
-%% \end{equation}
-
-%% \subsection{A thermodynamic model for bounded rationality, aka robust
-%%   optimality}
-%% \begin{itemize}
-%%   \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
-%%   \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
-%%   \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
-%%     \end{itemize}
-%% The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
-%% \paragraph{Utility functions and conjugate pairs.}
-%%   Let $(\Omega, \mathcal F, P)$ be a probablity space.
-A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
-%%   \begin{itemize}
-%%   \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.
-%%   \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
-%%     $A,B \in \mathcal F$.
-%%   \item monotonicity: $f_{U}$ is strictly increasing.
-%%     \end{itemize}
-
-%% \begin{theorem}
-%%   The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
-%%   \begin{equation}
-%%     f = \alpha \log(.),
-%%     \label{eq:boltzmann}
-%%   \end{equation}
-%%   where $\alpha > 0$.
-%% \end{theorem}
-
-%% \textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
-
-%% Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
-
-%% \paragraph{Example.}
-%% Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
-%%   \begin{equation}
-%%     P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
-%%   \end{equation}
-%%   where
-
-%%   \begin{equation}
-%%     Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
-%%   \end{equation}
-%%   is a normalization constant called the \textit{partition function} of $U$.
-%%   It's not hard to see that $U$ and the $P$ above form a conjugate pair.
-
-%%   \paragraph{Free-utility functional.}
-%%     Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
-Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
-%%     \begin{equation}
-%%       J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
-%%     \label{eq:free_u}
-%%     \end{equation}
-%%     where
-%%     \begin{equation}
-%%       \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
-%%     \end{equation}
-%%     is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
-%%   \begin{equation}
-%%     J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
-%%   \end{equation}
-%%   In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
-%%   \begin{equation}
-%%     J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
-%%     \end{equation}
-
-%%   \paragraph{The free-energy / utility principle (of Friston ?).}
-%%   We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
-Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
-We want to transform this to a new model by adding constraints represented by the utility function $\Delta U$.
-The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \Delta U$.
-The difference in free-utility is then
-%%   \begin{equation}
-%%     \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
-%%     \label{eq:free}
-%%   \end{equation}
-%%   where
-%%   \begin{equation}
-%%     D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
-%%   \end{equation}
-%%   is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
-%% In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
-%%   The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
-%% a trade-off between an expected utility $\Delta U$ (first term) and
-%% the information cost of transforming Pi
-%% into $P_{\text{fin}}$ (second
-%% term). In this interpretation $P_{\text{ini}}$ represents an initial choice
-%% probability or policy, which includes the special case of the
-%% uniform distribution where the decision-maker has initially no
-%% preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
-%% considers the utility constraint U∗
-%% that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
-%% \begin{equation}
-%% P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
-%%   \end{equation}
-
-%% By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
-
-%% \begin{equation}
-%%   P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
-%% \end{equation}
-
-%% Two limit cases are worth considering.
-%% \begin{itemize}
-%% \item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
-%%   $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
-%%   where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
-%%   \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
-%% \end{itemize}
-
-%% To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
-For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...
-\subsubsection{Concluding remarks}
+\subsection{Friston's Free-energy principle (refer to \ref{label:friston})}
 In summary:

### 2017-03-14 paper/paper.tex
more refs + writeup

@@
   \centering
-  \includegraphics[width=.7\linewidth]{vae.pdf}
+  \includegraphics[width=.8\linewidth]{vae.pdf}
   \caption{Variational autoencoders...}
@@
 \end{itemize}
+The mapping $g$ should be taught of as a ``blurring'' function which produces noisy versions $\z$,
+called \textit{sensations}, of the true world state $\x$.
 The result is a scheme for training DBNs via good-old backprop! 
@@
 
+\subsection{GANs and other likelihood-free methods}
+...
+
 \section{Reinforcement Learning as a computational model for DMN function}
@@
 A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
-Accorind to this model, the environment’s dynamics are fully determined by
+According to this model, the environment’s dynamics are fully determined by
 its current state.
@@
 transition probability matrix.)
-is simply a quintuplet $(\mathcal S, \mathcal A, r, \rho, \mu)$ where
+is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
 \begin{itemize}
 \item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
-\item $\rho : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto \rho(s'|s,a)$,
+\item $p : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto p(s'|s,a)$,
   the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
@@
 
-\begin{remark}
-  Given a system $(\mathcal S, \mathcal A, r, \rho, \mu)$ satisfying all the axioms for an MDP
-  except the Markov property, we can always transform it into an MDP by considering a compounded
-  version of the states
-  $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
-  history up to an including time $t$.
-\end{remark}
+It is our view that at a sufficiently coarse scale, the brain is a physical system governed by the laws of
+physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
+any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
+step is sufficiently short (defining $s(t)$ --the state at time $t$-- as the position in phase space).
+If the process has $memory$ such that the next state depends not only on the current state but also on some finite
+number of past states, it can reformulated as a stan- dard memoryless Markov process by simply expanding the
+definition of the state $s$ to include elements of the past.
+% \begin{remark}
+%   Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
+%   except the Markov property, we can always transform it into an MDP by considering a compounded
+%   version of the states
+%   $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
+%   history up to an including time $t$.
+% \end{remark}
 
@@
 Knowledge the agent has about the environment can be encoded in the
-functions $r$, $\mu$, $\rho$, and / or $\mu$, as is done in
+functions $r$, $\mu$, $p$, and / or $\mu$, as is done in
 \textit{model-based} RL. For example if a rat in a maze knows, for sure,
@@
 \item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
-\item $\rho(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
+\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
 \item etc.
@@
   \begin{split}
-    S_1 &\sim \rho(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
+    S_1 &\sim p(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
     A_1 &\sim \pi(.|S_1)\\
-    S_2 &\sim \rho(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
+    S_2 &\sim p(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
     \ldots\\
     A_{t} &\sim \pi(.|S_{t})\\
-    S_{t+1} &\sim \rho(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
+    S_{t+1} &\sim p(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
     \ldots
@@
 \begin{equation}
-  Q^\pi(s, a) = \mathbb E_{s' \sim \rho(.|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
-  = \mathbb E_{s' \sim \rho(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))]
+  Q^\pi(s, a) = \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
+  = \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))]
   \label{eq:bellman}

### 2017-03-19 paper/paper.tex
Sunday modifications

@@
   \item In sum,
-  the PC account in cognitive neuroscience does not expose an explicit
+  the MDP account may serve as a concrete conceptualization of the PC account in cognitive neuroscience. MPDs have the advantage of exposing an explicit
   mechanism of the horizon of future considerations or
@@
     \item In sum,
-    MDPs imply a ``radar'' function of montoring the environment for salient information
+    MDPs imply a ``radar'' function of monitoring the environment for salient information
     in general and explain social thinking as a physiological brain baseline
@@
 the behavior of intelligent RL systems.
-The proposed machine-learning
-approach to DMN function is hence not only practically computable but
+The proposed machine-learning perspective
+on DMN function is hence not only practically computable but
 also neuroscientifically falsifiable.

### 2017-03-19 paper/paper.tex
removing unneeded stuff

--- 2017-03-14 paper/paper.tex
@@
 (e.g car-driving, drone maneuvering, etc.
-
-\section{Optimal control}
-\subsection{Linear control network model for brain organization}
-Consider the following linear time-invariant (LTI) dynamical system as a (toy) model for high-level brain function
-\begin{equation}
-  \dot{\x}(t) = \A\x(t) + \B\u(t).
-  \label{eq:lti}
-\end{equation}
-Here, the $n$-by-$n$ matrix $\A$ denotes a model of the brain's wiring (for example a resting state connectome computed from an anatomical atlas), while the $n$-by-$k$ ``input matrix'' $\B$ describes which nodes can be controlled by us, an external \textit{controller}, via medical intervention or a careful choice of stimulus presentation, for example.
-At time $t \ge 0$, let  the $n$-vector  $\x(t) := (\x_1(t),\ldots,\x_n(t))$ encodes the state of the network (one value for each node).
-The aim is to supply values of $\u_1(t),\ldots,\u_k(t)$ for $k \le n$ controls as a function of the time $t$,  to take the system from any prescribed initial state $\x(0) = \x^{\text{init}} \in \mathbb R^n$ to any prescribed final state $\x(\tau) = \x^{\text{fin}} \in \mathbb R^n$ in finite time $\tau < \infty$.
-When such a controlling is possible, we say that the system $(\A,\B)$ is \textit{controllable}.
-A precise sufficient and necessary condition for such controllability is the Kalman condition: $(\A,\B)$ is controllable iff the $n \times nk$ \textit{controllability matrix}
-\begin{equation}
-  \C := (\B|\A\B|\ldots|\A^{n-1}\B)
-\end{equation}
-has full rank, i.e
-\begin{equation}
-  rank(\C) = n.
-\end{equation}
-
-\paragraph{How many controls do we need at best ?}
-In the thermodynamic limit ($n \rightarrow \infty$), statistical physics \citep{Liu2011} gives the extremely good estimate
-\begin{equation}
-  n_0(\langle k\rangle, \gamma) \approx \exp\left(-\frac{1}{2}\left(\frac{\gamma-2}{\gamma - 1}\right)\langle k \rangle\right),
-\end{equation}
-where $\gamma \ge 1$ is the \textit{scale-free} parameter for the node-degree distribution of the network described by $\A$, and $\langle k \rangle$ is the mean degree.
-Letting $\gamma \rightarrow \infty$, one recovers the Erdos-Renyi value
-\begin{equation}
-  n_0(\langle k\rangle) = \exp\left(-\frac{1}{2}\langle k\rangle\right),
-\end{equation}
-which agrees perfectly with its known analytic value. One identifies two radically different regimes:
-\begin{itemize}
-  \item the manageable regime $\gamma > \gamma_c := 2$, in which $n_0 < 1$, and so we only need fewer controls than total number of nodes, and
-\item the unmanageable regime ``$\gamma \le \gamma_c$'', in which  $n_0 = 1$, where each node must be controlled explicitly.
-\end{itemize}
-
-\paragraph{Why linear dynamics ?}
-As explained in \citep{Liu2011}, the choice of linear dynamics over a more general nonlinear dynamics can be justified as follows:
-\begin{itemize}
-  \item  Conclusions drawn from linear dynamics can be
-extended to nonlinear systems.
-\item If the controllability matrix of the linearized system
-has full rank at all points, then it is sufficient for most
-systems to say that the actual nonlinear system is
-controllable (i.e. small signal model).
-\end{itemize}
-
-\paragraph{A note on stability.}
-For stability in the LTI model \eqref{eq:lti} and hence in the constrained path integral \eqref{eq:hj} below, a sufficient (and necessary ?) condition is that all eigenvalues of the $A$ have negative real-part.
-One way to impose this is to add self-loops with a small negative weight.
-\paragraph{Meta brain.}
-
-\subsection{Optimal control: an LQR feedback controller}
-We propose to use a linear quadratic regulator (LQR) for the feedback controller.
-Thus, consider the time-varying \text{value function} $V: [0, \tau] \times \mathbb R^n \rightarrow \mathbb R$, defined by the following Hamilton-Jacobi cost-to-go
-\begin{equation}
-  \begin{split}
-    &V(t, \z) := \min_{\u} \frac{1}{2}\x_{\text{fin}}^T\Q_{\text{fin}}\x_{\text{fin}} + \int_{0}^\tau\left(\frac{1}{2}\x(t)^T\Q\x(t) + \frac{1}{2}\u(t)^T\R\u(t)\right)dt,\\
-    &\text{subject to } \x(t) = \z, \dot{\x}(t) = \A\x(t) + \B\u(t),
-  \end{split}
-  \label{eq:hj}
-\end{equation}
-where the matrices $\Q$ and $\R$, precised by design considerations and subject to meta-optimization, are restricted to be positive semi-definite and positive definite, respectively.
-A classical calculation (reminiscent of the \textit{Pontryagin minimization principle}) reveals that the optimal feedback control in \eqref{eq:hj} is given by
-\begin{equation}
-  \u(t) = -\K(t)\x(t),
-\end{equation}
-where
-\begin{equation}
-  \K(t) := \R^{-1}\B^T\P(t)\text{ (Kalman gain matrix)}
-\end{equation}
-and the time-varying positive semi-definite matrix $\P(t)$ solves following differential Riccati equation (DRE)
-\begin{equation}
-\begin{split}
-&-\dot{\P}(t) = \A^T\P(t) + \P(t)\A - \P(t)\B\R^{-1}\B^T\P(t) + \Q\\
-&\text{subject to }\P(\tau) = \Q_{\text{fin}}.
-\end{split}
-\end{equation}
-%% \subsection{Minimal control energy controller}
 

### 2017-03-26 paper/paper.tex
made a major pass on the paper

@@
 \title{A Computational Account of Default Mode Function\\
-by Control Theory and Reinforcement Learning}
+  by
+  % Control Theory and
+  Reinforcement Learning Theory}
 
@@
 DMN function is recast in mathematical terms
-from the perspective of control theory and
+from the perspective of
+% control theory and
 reinforcement learning.
@@
 \textbf{\\keywords}: systems biology, mind wandering, cognitive science,
-autonomous learning systems
+artificial intelligence (AI),
+reinforcement learning
+% autonomous learning systems
 % \textbf{\\Get opinion from}: Danielle Bassett, Kai Vogeley, Mohammad Shakir,
@@
 
-Consider an agent faced by the choice of the next action
+Consider an agent faced with the choice of the next action
 and being guided by rich
@@
 Switching between states reduces to a sequential processing model.
-In a model-based reinforcement learning approach,
+In a model-based reinforcement learning (RL) approach,
 we try to understand the environment always better
@@
 that are calibrated by prediction error.
-A DMN framework incorporating reinforcement learning
-can naturally embed human behavior
+A DMN framework incorporating RL can naturally embed human behavior
 into the tension between exploitative action with immediate gains and
@@
 \section{Known Neurobiological Properties of the Default Mode Network}
+\textbf{XXX This section is presented as transfinite block of concatenated text.
+  Please reorganize into subsections and paragraphs!}
 We will first attempt to deconstruct the overall DMN function
@@
 %
+\subsection{The PMC and PFC}
 The posteriormedial cortex (PMC) and
@@
 of the amygdala \citep{bzdok2015subspecialization}.
+\textbf{XXX: the previous sentence seems to be bad / incomplete grammar!}
 This amygdalar subregion has been proposed to
@@
 across individual decision-making instances \citep{pearson2009neurons}.
+
 This DMN node encoded the
@@
 
+\subsection{The prefrontal cortex (PFC)}
 The dmPFC subserves predominantly ambiguous amodal processes
@@
 reward-informed and risk estimation of self-relevant environmental stimuli.
+\textbf{XXX: previous sentence needs to be punctuated!}
 This DMN node is more closely associated with
@@
 in axonal tracing studies in monkeys \citep{haber1995orbital}.
-The NAc is thought to be linked to reward mechanisms that not
+The NAc (\textbf{XXX ``NAc'' = ?}) is thought to be linked to reward mechanisms that not
 only modulate motivated behavior towards basic survival needs but also
-subserve reinforcement learning in humans more broadly.
+subserve RL in humans more broadly.
 This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys
@@
 which was shown to be critical for behavior.
-The GMV of the vmPFC and nucleus accumbens
+The GMV (\textbf{``GMV'' = ?}) of the vmPFC and nucleus accumbens
 correlated with indices of social reward attitudes and
 value-guided behavior \citep{lebreton2009automatic}.
+\textbf{XXX Previous sentence is detached and disconnected!}
 Additionally,
@@
 
+\subsection{The TPJ: Situation understanding and context-switching}
 Finally,
@@
 %
-The aRTPJ is a key node network is central for
+The aRTPJ (\textbf{XXX ???})  is a key node network is central for
 action initiation during externally structured tasks and
 sensorimotor control by integrating supramodal stimulus-guided attention
-\citep{corbetta2002control}.
+\citep{corbetta2002control}. \textbf{XXX Previous sentence is weird!}
 Involvement of this DMN node was repeatedly reported in
@@
 agent. In this view, the environmental is partially controlled by
-the action of the agent.
-This reward should be thought
+the action of the agent and the reward should be thought
 of as an instantaneous satisfaction accompanying the execution of
-an action.
+an action, and may be delayed, only becoming substantial when accumulated over an extended period of time.
 %% Here the subscript $t$ denotes time, and can be discrete or
 %% continuous.
-The environment $\mathcal E$ is generally taken as stochastic,
+The environment is generally taken as stochastic,
 is partially observed, in the sense that only part of the current state is observed by
-the agent.
+the agent. This unpredictability
+% (in a first-order sense)
+is realistic in a model which sets out
+to explain DMN function. Indeed, it is known that, the more the external world is predictable,
+the more mental activity becomes detached from the actual sensorium
+\citep{antrobus1966studies, pope1978regulation}.
+Conversely, the more the currently executed task is unknown and unpracticed,
+the lower the DMN activity, i.e the tendency for stimulus-independent thoughts
+\citep{filler1973daydreaming, teasdale1995stimulus}.
+
+These ``offline'' processes may however contribute to optimized control of the organism.
+A DMN framework incorporating reinforcement learning
+can naturally embed human behavior
+into the tension between exploitative action with immediate gains and
+explorative action with longer-term reward schedules.
+The implication of the DMN in a diversity of human cognitive processes
+can be parsimoneously recast as prediction error minimization
+based on pervasive probabilistic simulations,
+thus maximizing reward outcome at various time scales.
+Such a purposeful optimization objective
+may be solved by a stochastic approximation
+based on a brain implementation of Markov Chain Monte Carlo sampling
+\citep{tenenbaum2011grow}.
+
 \subsection{Markov Decision Processes}
-A popular abstract model for the scenario are so-called \textit{Markov Decision Processes (MDPs)}.
-According to this model, the environment’s dynamics are fully determined by
-its current state.
+It is our view that at a sufficiently coarse scale, the brain is a physical system governed by the laws of
+physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
+any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
+step is sufficiently short (defining $s(t)$ --the state at time $t$-- as the position in phase space).
+If the process has $memory$ such that the next state depends not only on the current state but also on some finite
+number of past states, it can reformulated as a standard memoryless Markov process by simply expanding the
+definition of the state $s$ to include elements of the past.
+
+In artificial intelligence (AI), a popular abstract model for such an environment are so-called
+\textit{Markov Decision Processes (MDPs)}. According to this model, the environment's dynamics are
+fully determined by its current state.
 In RL, MDPs are the mathematical instantiation of the environment model,
@@
 
-It is our view that at a sufficiently coarse scale, the brain is a physical system governed by the laws of
-physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
-any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
-step is sufficiently short (defining $s(t)$ --the state at time $t$-- as the position in phase space).
-If the process has $memory$ such that the next state depends not only on the current state but also on some finite
-number of past states, it can reformulated as a stan- dard memoryless Markov process by simply expanding the
-definition of the state $s$ to include elements of the past.
 % \begin{remark}
@@
 following a
-policy $\pi$. The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
+policy $\pi$.
+\begin{mdframed}
+  \textbf{Where is ``value'' buffered in the brain ?}
+  The amygdala is known to be involved in significance evaluation) and
+  the nucleus accumbens is involved in reward evaluation.
+  \end{mdframed}
+The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
 In the definition of $G^\pi$ above, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}.
@@
   backprop!
-  \end{itemize}
-
-\subsection{Control theory}
-\begin{itemize}
-\item An authoritative review of control theory for decision making and its relationship to
-  reinforcement learning can be found in \citep{kappen2007introduction}. 
-  \item The model proposed in \citep{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...
   \end{itemize}

### 2017-03-26 paper/paper.tex
more

@@
 %
-\subsection{The PMC and PFC}
+\subsection{The posteriormedial cortex (PMC)}
 The posteriormedial cortex (PMC) and
@@
 
-The HC is well known to be involved in memory and
+The HC (\textbf{``HC'' = ?}) is well known to be involved in memory and
 spatial navigation in animals and humans.
@@
 
-\subsection{The TPJ: Situation understanding and context-switching}
+\subsection{The TPJ: Situation-understanding and context-switching}
 Finally,
@@
 % self -> value matrix driven computation
+
+\section{Reinforcement Learning (RL) as a computational model for DMN function}
+We now argue that the DMN has biological substrates (gradient
+computation, backprop, experience replay, etc.) for all the
+major components of an efficient \textit{Reinforcement Learning} algorithm, and all these
+components integrate in a plausible way.
+Reinforcement Learning (RL) is learning to achieve a goal by interacting with an external
+environment. In a general RL framework, an \textit{agent}
+interacts with an  environment in a trial-by-error fashion by taking
+actions / decisions $a$, which trigger changes in the
+\textit{state} of the environment
+$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
+$r = r(s, a)$ (or regret!, since it can be negative) collected by the
+agent. In this view, the environmental is partially controlled by
+the action of the agent and the reward should be thought
+of as an instantaneous satisfaction accompanying the execution of
+an action, and may be delayed, only becoming substantial when accumulated over an extended period of time.
+%% Here the subscript $t$ denotes time, and can be discrete or
+%% continuous.
+The environment is generally taken as stochastic,
+is partially observed, in the sense that only part of the current state is observed by
+the agent. This unpredictability
+% (in a first-order sense)
+is realistic in a model which sets out
+to explain DMN function. Indeed, it is known that, the more the external world is predictable,
+the more mental activity becomes detached from the actual sensorium
+\citep{antrobus1966studies, pope1978regulation}.
+Conversely, the more the currently executed task is unknown and unpracticed,
+the lower the DMN activity, i.e the tendency for stimulus-independent thoughts
+\citep{filler1973daydreaming, teasdale1995stimulus}.
+
+These ``offline'' processes may however contribute to optimized control of the organism.
+A DMN framework incorporating reinforcement learning
+can naturally embed human behavior
+into the tension between exploitative action with immediate gains and
+explorative action with longer-term reward schedules.
+The implication of the DMN in a diversity of human cognitive processes
+can be parsimoneously recast as prediction error minimization
+based on pervasive probabilistic simulations,
+thus maximizing reward outcome at various time scales.
+Such a purposeful optimization objective
+may be solved by a stochastic approximation
+based on a brain implementation of Markov Chain Monte Carlo sampling
+\citep{tenenbaum2011grow}.
+
+\subsection{Markov Decision Processes}
+It is our view that at a sufficiently coarse scale, the brain is a physical system governed by the laws of
+physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
+any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
+step is sufficiently short (defining $s(t)$ --the state at time $t$-- as the position in phase space).
+If the process has $memory$ such that the next state depends not only on the current state but also on some finite
+number of past states, it can reformulated as a standard memoryless Markov process by simply expanding the
+definition of the state $s$ to include elements of the past.
+
+In artificial intelligence (AI), a popular abstract model for such an environment are so-called
+\textit{Markov Decision Processes (MDPs)}. According to this model, the environment's dynamics are
+fully determined by its current state.
+In RL, MDPs are the mathematical instantiation of the environment model,
+which generally can incorporate action probabilities (policy matrix)
+and/or value function given a state.
+(Silver: The environment can be fully defined by state space and
+transition probability matrix.)
+is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
+\begin{itemize}
+\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
+\item $p : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto p(s'|s,a)$,
+  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
+  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present.
+\item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
+  \text{laugh}, \text{sympathize}, \text{empathize}\}.$
+\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$ is the \textit{reward function},
+  so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
+\item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
+  $\mu(s)$ is the probability that the environment starts-off in state $s$.
+\end{itemize}
+
+% \begin{remark}
+%   Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
+%   except the Markov property, we can always transform it into an MDP by considering a compounded
+%   version of the states
+%   $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
+%   history up to an including time $t$.
+% \end{remark}
+
+\paragraph{A note on model-based RL.}
+Knowledge the agent has about the environment can be encoded in the
+functions $r$, $\mu$, $p$, and / or $\mu$, as is done in
+\textit{model-based} RL. For example if a rat in a maze knows, for sure,
+that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food, they can assume:
+\begin{itemize}
+\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
+\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
+\item etc.
+\end{itemize}
+
+\subsection{Value functions and policies}
+The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
+sampling probabilities over candidate actions to perform from this state. Starting a time $t=0$,
+a policy $\pi$ generates a trajectory as follows
+\begin{eqnarray*}
+  \begin{split}
+    S_1 &\sim p(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
+    A_1 &\sim \pi(.|S_1)\\
+    S_2 &\sim p(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
+    \ldots\\
+    A_{t} &\sim \pi(.|S_{t})\\
+    S_{t+1} &\sim p(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
+    \ldots
+  \end{split}
+\end{eqnarray*}
+Since an action taken now might have repercussions long into the feature, it turns out that the
+quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
+cumulative reward estimate which takes into account the future. A standard for
+modeling this accumulation is the so-called
+time-discounted \textit{cumulative reward} function
+\begin{equation}
+  \label{eq:cumr}
+  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t.
+\end{equation}
+This random variable\footnote{Random as it depends both on the environment's dynamic and the
+  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
+following a
+policy $\pi$.
+\begin{mdframed}
+  \textbf{Where is ``value'' buffered in the brain ?}
+  The amygdala is known to be involved in significance evaluation) and
+  the nucleus accumbens is involved in reward evaluation.
+  \end{mdframed}
+The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
+In the definition of $G^\pi$ above, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}.
+Taking $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their
+immediate rewards and has no horizon for the future, which is not compatible with long-term goal planning
+as implemented in the DMN. For there to be any learning at all, it is thus necessary that $\gamma > 0$.
+% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
+Thus the agent is positively biased towards considering rewards which are will
+come soon. The reward importance hence decays exponentially
+with distance in the future,
+given that there is more uncertainty in the farsighted future.
+The reward discount factor $\gamma$ can thus be seen as calibrating
+the delay discounting behavior of the intelligent agent.
+It is also important to appreciate that a stochastic policy estimation
+is more advantageous in many RL settings.
+For instance, a deterministic strategy in playing rock-paper-scissors can
+be easily exploited and a uniformly random choice is optimal.
+
+The agent's goal can then be formulated as being that of finding policy $\pi$ which maximizes the
+expected cumulative value of state-stat-pairs $(A_0,S_0)$, namely
+\begin{equation}
+  \label{eq:qval}
+  Q^{\pi}(s,a) = \mathbb E [G^\pi|S_0=s,A_0=a,A_1 \sim \pi,A_2 \sim \pi,\ldots].
+\end{equation}
+That is, the expectated future reward over all possible trajectories in which we start at $s$ take
+action $a$ and then follow the policy $\pi$ thereafter to select future actions.
+%% Given that each $\pi$ correspondence to an exhaustive description of
+%% behavioral patterns, $Q^{\pi}(S_t,A_t)$ corresponds to the epistemic value of
+%% the set of actions given states in finite time.
+
+By construction, in general, maximizing the expectation in  \eqref{eq:qval} is intractable due to
+the enormous number of possible trajectories the agent might need to
+consider.
+(If the Bellman-based computational ideas have a corresponding
+neurobiological implementation, this provides a framework of how
+optimal behavior could be computable in the human brain.)
+One popular simplification that removes this issue is to only consider the sub-family of
+\textit{deterministic policies}, which map each state to the best single best action which
+should be taken at thin state, and therefore define functions from states to actions.
+\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
+\textit{greedy policy}
+\begin{equation}
+  \pi(s) = \argmax_{a \in \mathcal A}{Q}(s, a).
+  \label{eq:qlearning}
+\end{equation}
+% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
+% table.
+The celebrated \textit{Bellman's equation} [Sutton and Barto 1998] then takes the simple form
+\begin{equation}
+  Q^\pi(s, a) = \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
+  = \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))]
+  \label{eq:bellman}
+\end{equation}
+which provides a recursive decomposition of optimal behavior.
+A complicated dynamic programming optimization can be
+broken into simpler subproblems at different time points,
+which has often been closely related to the medial prefrontal part of the DMN
+\citep{koechlin1999role, braver2002role}.
+Using the Bellman equation, each state can be associated with a certain value,
+to guide action (i.e., improve the policy) towards a better state.
+It is a fixed-point equation for the deterministic policy $\pi$.
+Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which
+depend on the environment, and so can be learned off-policy by observing state-transitions
+triggered by another behavioral policy, which can be stochastic.
+
+\subsection{Efficient RL with value approximation and replay}
+%As already mentioned in the previous section, Q-learning optimizes over the class of
+% deterministic policies of the form \eqref{eq:qlearning}.
+We need ways to approximate the state-action value function
+which scale up to large state spaces and / or action spaces.
+In state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
+% (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
+by sampling from its current policy matrix, and collects a reward $r$,
+% (measured on their internal currency of worldly values ?),
+and the environment transitions to a new state $s' \in \mathcal S$. At the end of
+this step, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar
+behavior of the agent and is recorded in replay memory buffer:
+$D \leftarrow D \cup \{e\}$ (possibly discarding the oldest entries to make space).
+Now, at iteration $k+1$, replay consists in sampling (uniform or importance-weighted
+\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
+$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ and trying to
+approximate
+the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
+$r + \gamma \max_{a'} Q(s', a'|\theta^Q_k)$, with the output of a parametrized regression model $(s,a,r,s')
+\mapsto {Q}(s, a|\theta^Q_{k})$, where $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q_{k+1}$.
+Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function
+\begin{equation}
+  \mathcal L(\theta^Q_{k+1})
+  % = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
+  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
+  \label{eq:oracle}
+\end{equation}
+where
+$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
+% While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
+For example, a general linear model with a kernel $\phi$ would be of the
+form
+$${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
+$\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action $(s,a)$ into an appropriate
+feature space. Such are the techniques that have been proposed by [sutton et al. xyzw, Song et al. 2016 (NIPS),
+  and Parr et al. 2003 (JMLR)].
+A breakthrough alternative proposed by Google DeepMind [Minh et al. 2014, Silver et al 2015] is to learn this
+representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which
+are by far the current state-of-the-art in RL. In such a case, $\theta^Q$
+would correspond to ..
+
+The entire model can then be effectively trained via \textit{backprop}
+wherein online stochastic gradient descent is used to minimize the
+variational objective \eqref{eq:oracle}.
+
+\paragraph{Parameter update.}
+\begin{equation}
+  \nabla_{\theta^Q_{k+1}}\mathcal L(\theta^Q_{k+1})
+  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
+  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
+    \underbrace{\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1})}_{\text{???}}\right],
+  \label{eq:oracle}
+\end{equation}
+
+\paragraph{TODO: Monte Carlo optimization.}
+Some text snippets that might be pertinent:
+Monte-Carlo simulation can be used to update the agent’s belief state
+\citep{silver2010monte}.
+Monte-Carlo simulation provides a simple method for evaluating
+a state s.
+In any MDP, there is at least one optimal policy matrix that
+achieves the optimal value function???
+The simulator is used to generate sequences of states, observations and rewards.
+These simulations are used to update the value function, without ever looking inside the black box describing the model’s dynamics.
+In addition, Monte-Carlo methods have a sample complexity that is determined only by the underlying difficulty of the POMDP, rather than the size of the state space or observation space.
+Monte-Carlo simulation can be turned into a simple control algorithm by evaluating all legal actions and selecting the action with highest evaluation [15].
+Monte-Carlo simulation can be extended to partially observable MDPs.
+In these challenging POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
+Link this to experience sampling in the hippocampus.
+Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
+For sampling-based methods it is still almost non-existent beyond the description of receptive fields.
+
+\paragraph{Link to classical RL algorithms.}
+Note that most RL algorithms, including \textit{Temporal Difference (TD)}
+learning [Sutton et al., Bertsekas], REINFORCE [William], and SARSA can be
+cast in this general variational framework. For example TD corresponds to
+the above framework using a linear value approximator with feature encoding
+$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
+$\mathcal S \times \mathcal A$, and so
+$$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
+%% Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
+%% $$\theta^Q(s,a) \leftarrow \theta^Q(s,a) + (Q(s, a|\theta^Q_{k+1}) - y)$$ ??
+
+\subsection{Correspondence between DMN function and RL framework}
+\textbf{XXX Put the diagrams and process charts here.}
+\begin{figure}[!h]
+  \includegraphics[width=.9\linewidth]{rl_process_chart.pdf}
+  \caption{The DMN as an RL agent...}
+  \label{fig:rl_process_chart}
+\end{figure}
+
+\begin{itemize}
+  \item The policy matrix represents the repertoire of possible actions
+  on the world given a current state. It encodes the probabilites of
+  choosing actions to be executed in a certain situation.
+  \item Interactions between external and internal sources of
+  information are probably commonplace during perception
+  \item The MDP frameworks how the human brain could solve
+  environmental problems over extended time periods.
+  \item Given a state, the value function represents the estimated utility of
+  being in that state, which is subjective to the agent.
+  For a state $s$, the value function at $s$ namely $V(s)$ define provides a quantitative answer to the question:
+  "How desirable is it to be in state $s$ ?"
+  For each action $a$ that can be taken from state $s$,
+  the Q-value  $Q(s, a)$ defined in \eqref{eq:qval} provides the subjective
+  utility of executing a specific action; it answers the question
+  "What is the expected utiltiy of taking action $a$ in this situation ?"
+  \item The DMN may subserve
+  constant exploration of possible future actions and their
+  cumulative outcomes. Implicit computation of future choices
+  provides an explanation for purposeful mind-wandering.
+  \item hippocampus -> provides perturbed action-transition-state-reward samples
+  as a block of "imagined", "hypothesized", "recalled" experience;
+  the small variations in these experience blocks allow searching
+  a larger space of parameters and possible experiences.
+  In the absence of environmental input and feedback
+  (e.g., mind-wandering or sleep) pseudo-experiencing possible
+  future scenarios and action outcomes.
+  Our approach acknowledges the unavoidable stochasticity of
+  computation in neuronal processes \citep{faisal2008noise}.
+  \item
+  \textit{Inference} in the human brain reduces to generalization of
+  policy and value adaptations from sampled experiences to
+  successful action choices and reward predictions in future states.
+  \item In the RL view, plasticity in the DMN arises naturally.
+  If an agent behaving optimally in a given environment moves
+  to novel, yet unexperienced environments, reward prediction errors will
+  massively increase. This will lead to adaptation of the policy matrix
+  and value function until the system converges to a
+  new steady-state of optimal action decisions given certain states.
+  \item The DMN is today known to consistently increase in neural
+  activity when engaging cognitive processes that are detached from
+  current sensory environment. DMN involvement in thinking about the past,
+  the future, and hypothetical possibilities ties in with the
+  MDP's implicit computation of action and state cascades as a function
+  of what happened in the past.
+\end{itemize}
+
+\paragraph*{Some notes:}
+\begin{itemize}
+  \item Actions are continuous, whilst states are continuous.
+  \item The brain does backprop (see Hinton's recent talk at Stanford)
+  \item model weights $\theta^Q$ should correspond to connections between neurons
+\item humm, looks like such a knowledge would simply translate into constraints on the DqN weights $\theta^Q$ (... at least, part of such knowledge)
+  ok, gotta think about the model-free / model-based reconsiliation
+\end{itemize}
+
+\subsection{Inverse reinforcement-learning}
+%\begin{mdframed}
+  Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
+  a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
+  (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
+  \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
+  (e.g linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
+  For example, given traces of motor actions from a an adult to grab a cup on a table,
+  can an observing child figure out what ``langragian'' functional is being minimized by the former ?
+  How ? Can they reproduce this optimal behavior ?
+%\end{mdframed}
+
+Such questions are of course pertinent to our decision-making theory for the DMN.
+In the general artificial case, the problem has been extensively studied and partially solved by
+\cite{abbeel2004}. They are also been rigourously studied in general optimal control literature under the name ``inverse optimal control'', but in model-based certains (where the physical dynamics
+are known, etc.)...
+
+IRL is is suited for problems in which it's hard to define what the reward function could be
+(e.g car-driving, drone maneuvering, etc.
 
@@
 ...
-
-\section{Reinforcement Learning as a computational model for DMN function}
-The main idea of this section is to argue that the DMN has biological substrates (gradient
-computation, backprop, experience replay, etc.) for all the
-major components of an efficient \textit{Reinforcement Learning} algorithm, and all these
-components integrate in a plausible way.
-Reinforcement Learning (RL) is learning to achieve a goal by interacting with an external
-environment. In a general RL framework, an \textit{agent}
-interacts with an  environment in a trial-by-error fashion by taking
-actions / decisions $a$, which trigger changes in the
-\textit{state} of the environment
-$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
-$r = r(s, a)$ (or regret!, since it can be negative) collected by the
-agent. In this view, the environmental is partially controlled by
-the action of the agent and the reward should be thought
-of as an instantaneous satisfaction accompanying the execution of
-an action, and may be delayed, only becoming substantial when accumulated over an extended period of time.
-%% Here the subscript $t$ denotes time, and can be discrete or
-%% continuous.
-The environment is generally taken as stochastic,
-is partially observed, in the sense that only part of the current state is observed by
-the agent. This unpredictability
-% (in a first-order sense)
-is realistic in a model which sets out
-to explain DMN function. Indeed, it is known that, the more the external world is predictable,
-the more mental activity becomes detached from the actual sensorium
-\citep{antrobus1966studies, pope1978regulation}.
-Conversely, the more the currently executed task is unknown and unpracticed,
-the lower the DMN activity, i.e the tendency for stimulus-independent thoughts
-\citep{filler1973daydreaming, teasdale1995stimulus}.
-
-These ``offline'' processes may however contribute to optimized control of the organism.
-A DMN framework incorporating reinforcement learning
-can naturally embed human behavior
-into the tension between exploitative action with immediate gains and
-explorative action with longer-term reward schedules.
-The implication of the DMN in a diversity of human cognitive processes
-can be parsimoneously recast as prediction error minimization
-based on pervasive probabilistic simulations,
-thus maximizing reward outcome at various time scales.
-Such a purposeful optimization objective
-may be solved by a stochastic approximation
-based on a brain implementation of Markov Chain Monte Carlo sampling
-\citep{tenenbaum2011grow}.
-
-\subsection{Markov Decision Processes}
-It is our view that at a sufficiently coarse scale, the brain is a physical system governed by the laws of
-physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
-any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
-step is sufficiently short (defining $s(t)$ --the state at time $t$-- as the position in phase space).
-If the process has $memory$ such that the next state depends not only on the current state but also on some finite
-number of past states, it can reformulated as a standard memoryless Markov process by simply expanding the
-definition of the state $s$ to include elements of the past.
-
-In artificial intelligence (AI), a popular abstract model for such an environment are so-called
-\textit{Markov Decision Processes (MDPs)}. According to this model, the environment's dynamics are
-fully determined by its current state.
-In RL, MDPs are the mathematical instantiation of the environment model,
-which generally can incorporate action probabilities (policy matrix)
-and/or value function given a state.
-(Silver: The environment can be fully defined by state space and
-transition probability matrix.)
-is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
-\begin{itemize}
-\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
-\item $p : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto p(s'|s,a)$,
-  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
-  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present.
-\item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
-  \text{laugh}, \text{sympathize}, \text{empathize}\}.$
-\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$ is the \textit{reward function},
-  so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
-\item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
-  $\mu(s)$ is the probability that the environment starts-off in state $s$.
-\end{itemize}
-
-% \begin{remark}
-%   Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
-%   except the Markov property, we can always transform it into an MDP by considering a compounded
-%   version of the states
-%   $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
-%   history up to an including time $t$.
-% \end{remark}
-
-\paragraph{A note on model-based RL.}
-Knowledge the agent has about the environment can be encoded in the
-functions $r$, $\mu$, $p$, and / or $\mu$, as is done in
-\textit{model-based} RL. For example if a rat in a maze knows, for sure,
-that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food, they can assume:
-\begin{itemize}
-\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
-\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
-\item etc.
-\end{itemize}
-
-\subsection{Value functions and policies}
-The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
-sampling probabilities over candidate actions to perform from this state. Starting a time $t=0$,
-a policy $\pi$ generates a trajectory as follows
-\begin{eqnarray*}
-  \begin{split}
-    S_1 &\sim p(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
-    A_1 &\sim \pi(.|S_1)\\
-    S_2 &\sim p(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
-    \ldots\\
-    A_{t} &\sim \pi(.|S_{t})\\
-    S_{t+1} &\sim p(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
-    \ldots
-  \end{split}
-\end{eqnarray*}
-Since an action taken now might have repercussions long into the feature, it turns out that the
-quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
-cumulative reward estimate which takes into account the future. A standard for
-modeling this accumulation is the so-called
-time-discounted \textit{cumulative reward} function
-\begin{equation}
-  \label{eq:cumr}
-  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t.
-\end{equation}
-This random variable\footnote{Random as it depends both on the environment's dynamic and the
-  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
-following a
-policy $\pi$.
-\begin{mdframed}
-  \textbf{Where is ``value'' buffered in the brain ?}
-  The amygdala is known to be involved in significance evaluation) and
-  the nucleus accumbens is involved in reward evaluation.
-  \end{mdframed}
-The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
-In the definition of $G^\pi$ above, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}.
-The choice $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their immediate rewards and has no horizon
-for the future. For there to be any learning at all, a technical requirement is that $\gamma < 1$.
-% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
-Thus the agent is positively biased towards considering rewards which are will
-come soon. The reward importance hence decays exponentially
-with distance in the future,
-given that there is more uncertainty in the farsighted future.
-The reward discount factor $\gamma$ can thus be seen as calibrating
-the delay discounting behavior of the intelligent agent.
-It is also important to appreciate that a stochastic policy estimation
-is more advantageous in many RL settings.
-For instance, a deterministic strategy in playing rock-paper-scissors can
-be easily exploited and a uniformly random choice is optimal.
-
-The agent's goal can then be formulated as being that of finding policy $\pi$ which maximizes the
-expected cumulative value of state-stat-pairs $(A_0,S_0)$, namely
-\begin{equation}
-  \label{eq:qval}
-  Q^{\pi}(s,a) = \mathbb E [G^\pi|S_0=s,A_0=a,A_1 \sim \pi,A_2 \sim \pi,\ldots].
-\end{equation}
-That is, the expectated future reward over all possible trajectories in which we start at $s$ take
-action $a$ and then follow the policy $\pi$ thereafter to select future actions.
-%% Given that each $\pi$ correspondence to an exhaustive description of
-%% behavioral patterns, $Q^{\pi}(S_t,A_t)$ corresponds to the epistemic value of
-%% the set of actions given states in finite time.
-
-By construction, in general, maximizing the expectation in  \eqref{eq:qval} is intractable due to
-the enormous number of possible trajectories the agent might need to
-consider.
-(If the Bellman-based computational ideas have a corresponding
-neurobiological implementation, this provides a framework of how
-optimal behavior could be computable in the human brain.)
-One popular simplification that removes this issue is to only consider the sub-family of
-\textit{deterministic policies}, which map each state to the best single best action which
-should be taken at thin state, and therefore define functions from states to actions.
-\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
-\textit{greedy policy}
-\begin{equation}
-  \pi(s) = \argmax_{a \in \mathcal A}{Q}(s, a).
-  \label{eq:qlearning}
-\end{equation}
-% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
-% table.
-The celebrated \textit{Bellman's equation} [Sutton and Barto 1998] then takes the simple form
-\begin{equation}
-  Q^\pi(s, a) = \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
-  = \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))]
-  \label{eq:bellman}
-\end{equation}
-which provides a recursive decomposition of optimal behavior.
-A complicated dynamic programming optimization can be
-broken into simpler subproblems at different time points,
-which has often been closely related to the medial prefrontal part of the DMN
-\citep{koechlin1999role, braver2002role}.
-Using the Bellman equation, each state can be associated with a certain value,
-to guide action (i.e., improve the policy) towards a better state.
-It is a fixed-point equation for the deterministic policy $\pi$.
-Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which
-depend on the environment, and so can be learned off-policy by observing state-transitions
-triggered by another behavioral policy, which can be stochastic.
-
-\subsection{Efficient RL with value approximation and replay}
-%As already mentioned in the previous section, Q-learning optimizes over the class of
-% deterministic policies of the form \eqref{eq:qlearning}.
-We need ways to approximate the state-action value function
-which scale up to large state spaces and / or action spaces.
-In state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
-% (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
-by sampling from its current policy matrix, and collects a reward $r$,
-% (measured on their internal currency of worldly values ?),
-and the environment transitions to a new state $s' \in \mathcal S$. At the end of
-this step, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar
-behavior of the agent and is recorded in replay memory buffer:
-$D \leftarrow D \cup \{e\}$ (possibly discarding the oldest entries to make space).
-Now, at iteration $k+1$, replay consists in sampling (uniform or importance-weighted
-\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
-$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ and trying to
-approximate
-the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
-$r + \gamma \max_{a'} Q(s', a'|\theta^Q_k)$, with the output of a parametrized regression model $(s,a,r,s')
-\mapsto {Q}(s, a|\theta^Q_{k})$, where $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q_{k+1}$.
-Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function
-\begin{equation}
-  \mathcal L(\theta^Q_{k+1})
-  % = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
-  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
-  \label{eq:oracle}
-\end{equation}
-where
-$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
-% While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
-For example, a general linear model with a kernel $\phi$ would be of the
-form
-$${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
-$\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action $(s,a)$ into an appropriate
-feature space. Such are the techniques that have been proposed by [sutton et al. xyzw, Song et al. 2016 (NIPS),
-  and Parr et al. 2003 (JMLR)].
-A breakthrough alternative proposed by Google DeepMind [Minh et al. 2014, Silver et al 2015] is to learn this
-representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which
-are by far the current state-of-the-art in RL. In such a case, $\theta^Q$
-would correspond to ..
-
-The entire model can then be effectively trained via \textit{backprop}
-wherein online stochastic gradient descent is used to minimize the
-variational objective \eqref{eq:oracle}.
-
-\paragraph{Parameter update.}
-\begin{equation}
-  \nabla_{\theta^Q_{k+1}}\mathcal L(\theta^Q_{k+1})
-  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
-  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
-    \underbrace{\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1})}_{\text{???}}\right],
-  \label{eq:oracle}
-\end{equation}
-
-\paragraph{TODO: Monte Carlo optimization.}
-Some text snippets that might be pertinent:
-Monte-Carlo simulation can be used to update the agent’s belief state
-\citep{silver2010monte}.
-Monte-Carlo simulation provides a simple method for evaluating
-a state s.
-In any MDP, there is at least one optimal policy matrix that
-achieves the optimal value function???
-The simulator is used to generate sequences of states, observations and rewards.
-These simulations are used to update the value function, without ever looking inside the black box describing the model’s dynamics.
-In addition, Monte-Carlo methods have a sample complexity that is determined only by the underlying difficulty of the POMDP, rather than the size of the state space or observation space.
-Monte-Carlo simulation can be turned into a simple control algorithm by evaluating all legal actions and selecting the action with highest evaluation [15].
-Monte-Carlo simulation can be extended to partially observable MDPs.
-In these challenging POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
-Link this to experience sampling in the hippocampus.
-Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
-For sampling-based methods it is still almost non-existent beyond the description of receptive fields.
-
-\paragraph{Link to classical RL algorithms.}
-Note that most RL algorithms, including \textit{Temporal Difference (TD)}
-learning [Sutton et al., Bertsekas], REINFORCE [William], and SARSA can be
-cast in this general variational framework. For example TD corresponds to
-the above framework using a linear value approximator with feature encoding
-$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
-$\mathcal S \times \mathcal A$, and so
-$$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
-%% Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
-%% $$\theta^Q(s,a) \leftarrow \theta^Q(s,a) + (Q(s, a|\theta^Q_{k+1}) - y)$$ ??
-
-\subsection{Correspondence between DMN function and RL framework}
-\textbf{XXX Put the diagrams and process charts here.}
-\begin{figure}[!h]
-  \includegraphics[width=.9\linewidth]{rl_process_chart.pdf}
-  \caption{The DMN as an RL agent...}
-  \label{fig:rl_process_chart}
-\end{figure}
-
-\begin{itemize}
-  \item The policy matrix represents the repertoire of possible actions
-  on the world given a current state. It encodes the probabilites of
-  choosing actions to be executed in a certain situation.
-  \item Interactions between external and internal sources of
-  information are probably commonplace during perception
-  \item The MDP frameworks how the human brain could solve
-  environmental problems over extended time periods.
-  \item Given a state, the value function represents the estimated utility of
-  being in that state, which is subjective to the agent.
-  For a state $s$, the value function at $s$ namely $V(s)$ define provides a quantitative answer to the question:
-  "How desirable is it to be in state $s$ ?"
-  For each action $a$ that can be taken from state $s$,
-  the Q-value  $Q(s, a)$ defined in \eqref{eq:qval} provides the subjective
-  utility of executing a specific action; it answers the question
-  "What is the expected utiltiy of taking action $a$ in this situation ?"
-  \item The DMN may subserve
-  constant exploration of possible future actions and their
-  cumulative outcomes. Implicit computation of future choices
-  provides an explanation for purposeful mind-wandering.
-  \item hippocampus -> provides perturbed action-transition-state-reward samples
-  as a block of "imagined", "hypothesized", "recalled" experience;
-  the small variations in these experience blocks allow searching
-  a larger space of parameters and possible experiences.
-  In the absence of environmental input and feedback
-  (e.g., mind-wandering or sleep) pseudo-experiencing possible
-  future scenarios and action outcomes.
-  Our approach acknowledges the unavoidable stochasticity of
-  computation in neuronal processes \citep{faisal2008noise}.
-  \item
-  \textit{Inference} in the human brain reduces to generalization of
-  policy and value adaptations from sampled experiences to
-  successful action choices and reward predictions in future states.
-  \item In the RL view, plasticity in the DMN arises naturally.
-  If an agent behaving optimally in a given environment moves
-  to novel, yet unexperienced environments, reward prediction errors will
-  massively increase. This will lead to adaptation of the policy matrix
-  and value function until the system converges to a
-  new steady-state of optimal action decisions given certain states.
-  \item The DMN is today known to consistently increase in neural
-  activity when engaging cognitive processes that are detached from
-  current sensory environment. DMN involvement in thinking about the past,
-  the future, and hypothetical possibilities ties in with the
-  MDP's implicit computation of action and state cascades as a function
-  of what happened in the past.
-\end{itemize}
-
-\paragraph*{Some notes:}
-\begin{itemize}
-  \item Actions are continuous, whilst states are continuous.
-  \item The brain does backprop (see Hinton's recent talk at Stanford)
-  \item model weights $\theta^Q$ should correspond to connections between neurons
-\item humm, looks like such a knowledge would simply translate into constraints on the DqN weights $\theta^Q$ (... at least, part of such knowledge)
-  ok, gotta think about the model-free / model-based reconsiliation
-\end{itemize}
-
-\subsection{Inverse reinforcement-learning}
-%\begin{mdframed}
-  Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
-  a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
-  (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
-  \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
-  (e.g linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
-  For example, given traces of motor actions from a an adult to grab a cup on a table,
-  can an observing child figure out what ``langragian'' functional is being minimized by the former ?
-  How ? Can they reproduce this optimal behavior ?
-%\end{mdframed}
-
-Such questions are of course pertinent to our decision-making theory for the DMN.
-In the general artificial case, the problem has been extensively studied and partially solved by
-\cite{abbeel2004}. They are also been rigourously studied in general optimal control literature under the name ``inverse optimal control'', but in model-based certains (where the physical dynamics
-are known, etc.)...
-
-IRL is is suited for problems in which it's hard to define what the reward function could be
-(e.g car-driving, drone maneuvering, etc.
 

### 2017-03-26 paper/paper.tex
more

@@
 The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
-In the definition of $G^\pi$ above, the constant $\gamma \in [0, 1]$ is the \textit{discount factor}.
+In the definition of $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
 Taking $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their
 immediate rewards and has no horizon for the future, which is not compatible with long-term goal planning
-as implemented in the DMN. For there to be any learning at all, it is thus necessary that $\gamma > 0$.
+as implemented in the DMN. For there to be any learning at all, it is thus necessary that $0 < \gamma < 1$.
 % \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
@@
 
-The agent's goal can then be formulated as being that of finding policy $\pi$ which maximizes the
+The agent's goal can then be formulated as finding policy $\pi$ which maximizes the
 expected cumulative value of state-stat-pairs $(A_0,S_0)$, namely
@@
 %% the set of actions given states in finite time.
-
-By construction, in general, maximizing the expectation in  \eqref{eq:qval} is intractable due to
-the enormous number of possible trajectories the agent might need to
+By construction, maximizing the expectation in  \eqref{eq:qval} is intractable in general
+due to the enormous number of possible trajectories the agent might need to
 consider.
@@
 % deterministic policies of the form \eqref{eq:qlearning}.
-We need ways to approximate the state-action value function
+A successful RL theory needs ways to approximate the state-action value function
 which scale up to large state spaces and / or action spaces.
@@
 \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
-$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ and trying to
-approximate
+$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ (hyppocampus ?)
+and trying to approximate
 the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
@@
 are by far the current state-of-the-art in RL. In such a case, $\theta^Q$
-would correspond to ..
-
-The entire model can then be effectively trained via \textit{backprop}
-wherein online stochastic gradient descent is used to minimize the
-variational objective \eqref{eq:oracle}.
+would correspond to the strengths of low-level synaptic or high-level functional connections between different neurons / nodes
+in the network.
+
+Learning of the the entire model parameters can effictively be achieved via \textit{backprop},
+wherein prediction errors are propagated from lower to higher regions to modulate the choice of future
+% online stochastic gradient descent is used to minimize the variational objective
+% \eqref{eq:oracle}.
 

### 2017-03-26 paper/paper.tex
more

@@
 These ``offline'' processes may however contribute to optimized control of the organism.
-We formalize
-a policy matrix to capture the space of possible actions that can be performed
-on the environment given the current state. A value matrix
-maps expected rewards to environmental objects and events.
-Switching between states reduces to a sequential processing model.
+% We formalize
+% a policy matrix to capture the space of possible actions that can be performed
+% on the environment given the current state. A value matrix
+% maps expected rewards to environmental objects and events.
+% Switching between states reduces to a sequential processing model.
 In a model-based reinforcement learning (RL) approach,
@@
 Informed by outcomes of performed action,
-the DMN dynamics are constantly adapted in feed-back loops
+the DMN dynamics are constantly adapted in feedback loops
 that are calibrated by prediction error.
@@
 are by far the current state-of-the-art in RL. In such a case, $\theta^Q$
-would correspond to the strengths of low-level synaptic or high-level functional connections between different neurons / nodes
-in the network.
+would correspond to the strengths of low-level synaptic or high-level functional connections between different
+neurons / nodes / regions in the network.
 
 Learning of the the entire model parameters can effictively be achieved via \textit{backprop},
-wherein prediction errors are propagated from lower to higher regions to modulate the choice of future
+wherein prediction errors (aka \textit{regrets})  are propagated from lower to higher regions to modulate the choice of future actions
 % online stochastic gradient descent is used to minimize the variational objective
@@
 
-\paragraph{Parameter update.}
+% \paragraph{Parameter update.}
 \begin{equation}
-  \nabla_{\theta^Q_{k+1}}\mathcal L(\theta^Q_{k+1})
-  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
-  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
-    \underbrace{\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1})}_{\text{???}}\right],
+  \delta \theta_{k}^Q \propto -\nabla_{\theta^Q_{k}}\mathcal L(\theta^Q_{k})
+  % = -\mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{}))^2]
+  = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k}) - y)}_{\text{regret}}
+    \underbrace{\nabla_{\theta^Q_{k}}Q(s, a|\theta^Q_{k})}_{\text{???}}\right].
   \label{eq:oracle}
 \end{equation}
+
+\paragraph{Link to classical RL algorithms.}
+Note that most RL algorithms, including \textit{Temporal Difference (TD)}
+learning \cite{sutton1998reinforcement}, REINFORCE \cite{williams1992}, and SARSA can be
+cast in this general variational framework. For example TD corresponds to
+the above framework using a linear value approximator with feature encoding
+$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
+$\mathcal S \times \mathcal A$, and so
+$$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
+%% Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
+%% $$\theta^Q(s,a) \leftarrow \theta^Q(s,a) + (Q(s, a|\theta^Q_{k+1}) - y)$$ ??
 
@@
 
-\paragraph{Link to classical RL algorithms.}
-Note that most RL algorithms, including \textit{Temporal Difference (TD)}
-learning [Sutton et al., Bertsekas], REINFORCE [William], and SARSA can be
-cast in this general variational framework. For example TD corresponds to
-the above framework using a linear value approximator with feature encoding
-$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
-$\mathcal S \times \mathcal A$, and so
-$$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
-%% Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
-%% $$\theta^Q(s,a) \leftarrow \theta^Q(s,a) + (Q(s, a|\theta^Q_{k+1}) - y)$$ ??
-
 \subsection{Correspondence between DMN function and RL framework}
-\textbf{XXX Put the diagrams and process charts here.}
 \begin{figure}[!h]

### 2017-03-26 paper/paper.tex
more

@@
 
-\begin{mdframed}
-  Though our presentation of free-energy principles is thermodynamic in flavor, a completely equivalent
-  information-theoretic formulation is of course possible, simply by exploiting the information-theoretic interpretation of entropy.
-\end{mdframed}
+% \begin{mdframed}
+%   Though our presentation of free-energy principles is thermodynamic in flavor, a completely equivalent
+%   information-theoretic formulation is of course possible, simply by exploiting the information-theoretic interpretation of entropy.
+% \end{mdframed}
 
@@
 This primal-dual idea, first proposed in Hinton et al.
-1995, is at the heart of the general free-energy principle that we'll introduce shortly.
+1995, is at the heart of the general free-energy principle that we will introduce shortly.
 
 \subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
-In this subsection, we'll present an insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
+In this subsection, we will present an insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
 As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
@@
 
-\subsection{Friston's active-inference}
+\subsection{Friston's active-inference and agency}
 \label{sec:friston}
@@
 
+\begin{mdframed}
+  \textbf{XXX:} Rather than a heads-on competitor, our proposed formalism should be seen
+  and sold as a solid implementation of the predictive coding hypothesis.
+\end{mdframed}
 \subsection{Semantic Hypothesis}

### 2017-03-26 paper/paper.tex
more passes

@@
 
+\subsection{DMN, a central piece in planning and decision-making}
 The DMN appeared to be exclusive in task-induced
@@
 and being guided by rich
-reenactment of really happened, hypothetically conceived, and
+reenactment of really-happened, hypothetically conceived, and
 predicted future events to perfect ongoing behavioral performance.
@@
 probabilistic model of anticipated events''
-\citep{raichle2005intrinsic}.
-This paper proposes a formal model that
-satisfies this contention.
+\citep{raichle2005intrinsic}. This paper proposes a formal model that satisfies this contention.
+
+\subsection{Sketch of main contributions}
+We show that the DMN is endowed with sufficient neurobiological equipment to
+implement a reinforcement learning (RL) agent, at least in principle. A global
+picture of the current body of knowledge (predictive coding, free-energy principle,
+sentinel hypothesis, etc.) concerning decision-making in the brain
+is drawn, with the DMN at its core. Our work
 
@@
 %\end{mdframed}
-
 Such questions are of course pertinent to our decision-making theory for the DMN.
@@
 The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
-
 In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
@@
 \subsection{A general free-energy principle}
-We can resume the situation as follows:
+We can resume the situation as follows\footnote{Where we have used the fact that KL divergence is always nonnegative.}:
 \begin{mdframed}
@@
     &\le F^R_G(\x),
-    \text{ with equality if }p_R(.|\x) = p_G(.|\x),
+    \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
     \end{split}
@@
 \end{mdframed}
-where we have used the fact that KL divergence is always nonnegative.
 %% In fact, we have the following theorem

### 2017-03-26 paper/paper.tex
more

@@
 %
-\subsection{The posteriormedial cortex (PMC)}
+\subsection{The posteriormedial cortex (PMC): high-order integration}
 The posteriormedial cortex (PMC) and
@@
 the human PMC was coupled in two functional connectivity modalities
-with the amygdala
-(involved in significance evaluation) and
-the nucleus accumbens (involved in reward evaluation).
+with the amygdala involved in significance evaluation and
+the nucleus accumbens (NAc), nvolved in reward evaluation.
 Specifically, the human ventral posterior cingulate cortex was
@@
 in axonal tracing studies in monkeys \citep{haber1995orbital}.
-The NAc (\textbf{XXX ``NAc'' = ?}) is thought to be linked to reward mechanisms that not
+The NAc is thought to be linked to reward mechanisms that not
 only modulate motivated behavior towards basic survival needs but also
@@
 which was shown to be critical for behavior.
-The GMV (\textbf{``GMV'' = ?}) of the vmPFC and nucleus accumbens
+The gray-matter volume (GMV) of the vmPFC and nucleus accumbens
 correlated with indices of social reward attitudes and
@@
 
-The HC (\textbf{``HC'' = ?}) is well known to be involved in memory and
-spatial navigation in animals and humans.
+The HC is well known to be involved in memory and spatial navigation in animals and humans.
 Its highly recursive anatomical architecture
@@
 %
-The aRTPJ (\textbf{XXX ???})  is a key node network is central for
+The anterior RTPJ (\textbf{XXX ???})  is a key node network is central for
 action initiation during externally structured tasks and
@@
   policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
-following a
-policy $\pi$.
+following a policy $\pi$.
 \begin{mdframed}
@@
   The amygdala is known to be involved in significance evaluation) and
-  the nucleus accumbens is involved in reward evaluation.
+  the nucleus accumbens (NAc) is involved in reward evaluation.
 \end{mdframed}
-The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
-In the definition of $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
+  The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards. In
+  the definition of $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
 Taking $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their
@@
 \mapsto {Q}(s, a|\theta^Q_{k})$, where $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q_{k+1}$.
+In the brain, such an experience replay mechanism is probably implemented in the hyppocampus (HC). Indeed, it is well-known that during sleep,
+wake, and rest, the hyppocampal activity is positively correlated with activity in stimulus environment, though at a much faster rate.
 Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function
@@
 \begin{equation}
-  \delta \theta_{k}^Q \propto -\nabla_{\theta^Q_{k}}\mathcal L(\theta^Q_{k})
+  \delta \theta_{k+1}^Q \propto -\nabla_{\theta^Q_{k+1}}\mathcal L(\theta^Q_{k+1})
   % = -\mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{}))^2]
-  = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k}) - y)}_{\text{regret}}
-    \underbrace{\nabla_{\theta^Q_{k}}Q(s, a|\theta^Q_{k})}_{\text{???}}\right].
+  = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
+    \underbrace{\nabla_{\theta^Q_{k}}Q(s, a|\theta^Q_{k+1})}_{\text{???}}\right].
   \label{eq:oracle}

### 2017-03-26 paper/paper.tex
more fixups

@@
 \mapsto {Q}(s, a|\theta^Q_{k})$, where $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q_{k+1}$.
-In the brain, such an experience replay mechanism is probably implemented in the hyppocampus (HC). Indeed, it is well-known that during sleep,
-wake, and rest, the hyppocampal activity is positively correlated with activity in stimulus environment, though at a much faster rate.
+In the brain, such an experience replay mechanism is probably implemented in the hyppocampus (HC). Indeed,
+Hippocampal replay is a phenomenon observed in rats, mice, cats, rabbits, songbirds, and monkeys.
+Indeed, during sleep or awake rest, replay refers to the re-occurrence of a sequence of cell activations that also occurred during activity, but the replay has a much faster time scale.
+% It may be in the same order, or in reverse.
 Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function

### 2017-03-26 paper/paper.tex

@@
 The environment is generally taken as stochastic,
-is partially observed, in the sense that only part of the current state is observed by
+is \textit{partially observed},
+in the sense that only part of the current state is observed by
 the agent. This unpredictability
@@
 \citep{filler1973daydreaming, teasdale1995stimulus}.
-
+%
 These ``offline'' processes may however contribute to optimized control of the organism.

### 2017-03-26 paper/paper.tex
a pass to explain POMDPs

@@
 computation, backprop, experience replay, etc.) for all the
-major components of an efficient \textit{Reinforcement Learning} algorithm, and all these
+major components of an efficient \textit{reinforcement learning (RL)} algorithm, and all these
 components integrate in a plausible way.
@@
 
-\subsection{Markov Decision Processes}
+\subsection{Partially Observable Markov Decision Processes}
 It is our view that at a sufficiently coarse scale, the brain is a physical system governed by the laws of
@@
 In artificial intelligence (AI), a popular abstract model for such an environment are so-called
-\textit{Markov Decision Processes (MDPs)}. According to this model, the environment's dynamics are
-fully determined by its current state.
-In RL, MDPs are the mathematical instantiation of the environment model,
+\textit{Partially Observable Markov Decision Processes (POMDPs)}.
+A POMDP models a decision process in which it is assumed that environment dynamics are determined by a Markov process,
+but the agent cannot directly observe the underlying state.
+Instead, it must maintain a probability distribution over the set of possible states, based on a set of observations and observation probabilities, and the underlying Markov process.
+This is a minimal amount of assumptions that can be made about an environment, and is charastistic of so-called \textit{model-free RL}, the approach adapted by this work.
+In contrast, \textit{model-based RL} starts off with some mechanistic assumptions about the dynamics of world.
+These can be assumptions about the physical laws governing the agent's world or constraints on the state-space and transitions thereupon.
+
+In RL, POMDPs are the mathematical instantiation of the environment model,
 which generally can incorporate action probabilities (policy matrix)

### 2017-03-26 paper/paper.tex
more

@@
 These can be assumptions about the physical laws governing the agent's world or constraints on the state-space and transitions thereupon.
-
-In RL, POMDPs are the mathematical instantiation of the environment model,
-which generally can incorporate action probabilities (policy matrix)
-and/or value function given a state.
-(Silver: The environment can be fully defined by state space and
-transition probability matrix.)
-is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
+For example if a rat in a maze knows, for sure,
+that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food, they can assume:
+\begin{itemize}
+\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
+\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
+\item etc.
+\end{itemize}
+
+Mathematically, a POMDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
 \begin{itemize}
@@
 % \end{remark}
-
-\paragraph{A note on model-based RL.}
-Knowledge the agent has about the environment can be encoded in the
-functions $r$, $\mu$, $p$, and / or $\mu$, as is done in
-\textit{model-based} RL. For example if a rat in a maze knows, for sure,
-that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food, they can assume:
-\begin{itemize}
-\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
-\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
-\item etc.
-\end{itemize}
 

### 2017-03-26 paper/paper.tex
explicity+++

--- 2017-03-26 paper/paper.tex
@@
 
-\subsection{DMN, a central piece in planning and decision-making}
 The DMN appeared to be exclusive in task-induced
@@
 and being guided by rich
-reenactment of really-happened, hypothetically conceived, and
+reenactment of really happened, hypothetically conceived, and
 predicted future events to perfect ongoing behavioral performance.
@@
 probabilistic model of anticipated events''
-\citep{raichle2005intrinsic}. This paper proposes a formal model that satisfies this contention.
-
-\subsection{Sketch of main contributions}
-We show that the DMN is endowed with sufficient neurobiological equipment to
-implement a reinforcement learning (RL) agent, at least in principle. A global
-picture of the current body of knowledge (predictive coding, free-energy principle,
-sentinel hypothesis, etc.) concerning decision-making in the brain
-is drawn, with the DMN at its core. Our work
+\citep{raichle2005intrinsic}.
+This paper proposes a formal model that
+satisfies this contention.
 
@@
 %
-\subsection{The posteriormedial cortex (PMC): high-order integration}
+\subsection{The posteriormedial cortex (PMC)}
 The posteriormedial cortex (PMC) and
@@
 the human PMC was coupled in two functional connectivity modalities
-with the amygdala involved in significance evaluation and
-the nucleus accumbens (NAc), nvolved in reward evaluation.
+with the amygdala
+(involved in significance evaluation) and
+the nucleus accumbens (involved in reward evaluation).
 Specifically, the human ventral posterior cingulate cortex was
@@
 in axonal tracing studies in monkeys \citep{haber1995orbital}.
-The NAc is thought to be linked to reward mechanisms that not
+The NAc (\textbf{XXX ``NAc'' = ?}) is thought to be linked to reward mechanisms that not
 only modulate motivated behavior towards basic survival needs but also
@@
 which was shown to be critical for behavior.
-The gray-matter volume (GMV) of the vmPFC and nucleus accumbens
+The GMV (\textbf{``GMV'' = ?}) of the vmPFC and nucleus accumbens
 correlated with indices of social reward attitudes and
@@
 
-The HC is well known to be involved in memory and spatial navigation in animals and humans.
+The HC (\textbf{``HC'' = ?}) is well known to be involved in memory and
+spatial navigation in animals and humans.
 Its highly recursive anatomical architecture
@@
 %
-The anterior RTPJ (\textbf{XXX ???})  is a key node network is central for
+The aRTPJ (\textbf{XXX ???})  is a key node network is central for
 action initiation during externally structured tasks and
@@
   policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
-following a policy $\pi$.
+following a
+policy $\pi$.
 \begin{mdframed}
@@
   The amygdala is known to be involved in significance evaluation) and
-  the nucleus accumbens (NAc) is involved in reward evaluation.
+  the nucleus accumbens is involved in reward evaluation.
   \end{mdframed}
-  The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards. In
-  the definition of $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
+The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards.
+In the definition of $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
 Taking $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their
@@
 \mapsto {Q}(s, a|\theta^Q_{k})$, where $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q_{k+1}$.
-In the brain, such an experience replay mechanism is probably implemented in the hyppocampus (HC). Indeed,
-Hippocampal replay is a phenomenon observed in rats, mice, cats, rabbits, songbirds, and monkeys.
-Indeed, during sleep or awake rest, replay refers to the re-occurrence of a sequence of cell activations that also occurred during activity, but the replay has a much faster time scale.
-% It may be in the same order, or in reverse.
 Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function
@@
 \begin{equation}
-  \delta \theta_{k+1}^Q \propto -\nabla_{\theta^Q_{k+1}}\mathcal L(\theta^Q_{k+1})
+  \delta \theta_{k}^Q \propto -\nabla_{\theta^Q_{k}}\mathcal L(\theta^Q_{k})
   % = -\mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{}))^2]
-  = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
-    \underbrace{\nabla_{\theta^Q_{k}}Q(s, a|\theta^Q_{k+1})}_{\text{???}}\right].
+  = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k}) - y)}_{\text{regret}}
+    \underbrace{\nabla_{\theta^Q_{k}}Q(s, a|\theta^Q_{k})}_{\text{???}}\right].
   \label{eq:oracle}
@@
 %\end{mdframed}
+
 Such questions are of course pertinent to our decision-making theory for the DMN.
@@
 The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
+
 In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
@@
 \subsection{A general free-energy principle}
-We can resume the situation as follows\footnote{Where we have used the fact that KL divergence is always nonnegative.}:
+We can resume the situation as follows:
 \begin{mdframed}
@@
     &\le F^R_G(\x),
-    \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
+    \text{ with equality if }p_R(.|\x) = p_G(.|\x),
     \end{split}
@@
 \end{mdframed}
+where we have used the fact that KL divergence is always nonnegative.
 %% In fact, we have the following theorem

### 2017-03-26 paper/paper.tex

--- 2017-03-26 paper/paper.tex
@@
 computation, backprop, experience replay, etc.) for all the
-major components of an efficient \textit{reinforcement learning (RL)} algorithm, and all these
+major components of an efficient \textit{Reinforcement Learning} algorithm, and all these
 components integrate in a plausible way.
@@
 
-\subsection{Partially Observable Markov Decision Processes}
+\subsection{Markov Decision Processes}
 It is our view that at a sufficiently coarse scale, the brain is a physical system governed by the laws of
@@
 In artificial intelligence (AI), a popular abstract model for such an environment are so-called
-\textit{Partially Observable Markov Decision Processes (POMDPs)}.
-A POMDP models a decision process in which it is assumed that environment dynamics are determined by a Markov process,
-but the agent cannot directly observe the underlying state.
-Instead, it must maintain a probability distribution over the set of possible states, based on a set of observations and observation probabilities, and the underlying Markov process.
-This is a minimal amount of assumptions that can be made about an environment, and is charastistic of so-called \textit{model-free RL}, the approach adapted by this work.
-In contrast, \textit{model-based RL} starts off with some mechanistic assumptions about the dynamics of world.
-These can be assumptions about the physical laws governing the agent's world or constraints on the state-space and transitions thereupon.
-
-In RL, POMDPs are the mathematical instantiation of the environment model,
+\textit{Markov Decision Processes (MDPs)}. According to this model, the environment's dynamics are
+fully determined by its current state.
+In RL, MDPs are the mathematical instantiation of the environment model,
 which generally can incorporate action probabilities (policy matrix)
@@
 \mapsto {Q}(s, a|\theta^Q_{k})$, where $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q_{k+1}$.
-In the brain, such an experience replay mechanism is probably implemented in the hyppocampus (HC). Indeed,
-Hippocampal replay is a phenomenon observed in rats, mice, cats, rabbits, songbirds, and monkeys.
-Indeed, during sleep or awake rest, replay refers to the re-occurrence of a sequence of cell activations that also occurred during activity, but the replay has a much faster time scale.
-% It may be in the same order, or in reverse.
+In the brain, such an experience replay mechanism is probably implemented in the hyppocampus (HC). Indeed, it is well-known that during sleep,
+wake, and rest, the hyppocampal activity is positively correlated with activity in stimulus environment, though at a much faster rate.
 Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function

### 2017-03-26 paper/paper.tex
polishing + refs

--- 2017-03-26 paper/paper.tex
@@
 $\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action $(s,a)$ into an appropriate
-feature space. Such are the techniques that have been proposed by [sutton et al. xyzw, Song et al. 2016 (NIPS),
-  and Parr et al. 2003 (JMLR)].
-A breakthrough alternative proposed by Google DeepMind [Minh et al. 2014, Silver et al 2015] is to learn this
+feature space. Such are the techniques that have been proposed by \cite{songNIPS2016}, for example.
+%  and Parr et al. 2003 (JMLR)].
+A breakthrough alternative proposed by Google DeepMind \cite{mnih2015,silver2016mastering}, is to learn this
 representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which

### 2017-03-26 paper/paper.tex

@@
 
-\subsection{DMN, a central piece in planning and decision-making}
+\subsection{Default mode network: Higher-order control of the organism}
 The DMN appeared to be exclusive in task-induced
@@
 %
-\subsection{The posteriormedial cortex (PMC): high-order integration}
+\subsection{The posteriormedial cortex: Global monitoring and integration}
 The posteriormedial cortex (PMC) and
@@
 with the amygdala involved in significance evaluation and
-the nucleus accumbens (NAc), nvolved in reward evaluation.
-Specifically, the human ventral posterior cingulate cortex was
-the PMC region most connected to the laterobasal
-(rather than centromedial or superficial) nuclei group
-of the amygdala \citep{bzdok2015subspecialization}.
-\textbf{XXX: the previous sentence seems to be bad / incomplete grammar!}
+the nucleus accumbens (NAc), involved in reward evaluation.
+Specifically, among all parts of the PMC,
+the human ventral posterior cingulate cortex was
+most connected to the laterobasal
+nuclei group of the amygdala
+\citep{bzdok2015subspecialization}.
 This amygdalar subregion has been proposed to
@@
 
-\subsection{The prefrontal cortex (PFC)}
+\subsection{The prefrontal cortex: Stimulus-value assocation and action consideration}
 The dmPFC subserves predominantly ambiguous amodal processes
@@
 The ventromedial prefrontal cortex (vmPFC) subserves
-less ambiguous subjective-value-related evaluative processes
-reward-informed and risk estimation of self-relevant environmental stimuli.
-\textbf{XXX: previous sentence needs to be punctuated!}
+less ambiguous subjective-value-related evaluative processes and
+reward-informed risk estimation of self-relevant environmental stimuli.
 This DMN node is more closely associated with
@@
 in axonal tracing studies in monkeys \citep{haber1995orbital}.
+Additionally, the gray-matter volume of the vmPFC and nucleus accumbens
+correlated with indices of value-guided behavior and social reward attitudes
+\citep{lebreton2009automatic}.
 The NAc is thought to be linked to reward mechanisms that not
@@
 which was shown to be critical for behavior.
-The gray-matter volume (GMV) of the vmPFC and nucleus accumbens
-correlated with indices of social reward attitudes and
-value-guided behavior \citep{lebreton2009automatic}.
-\textbf{XXX Previous sentence is detached and disconnected!}
 Additionally,
@@
 
+\subsection{The hippocampus: Memory, space, and experience replay}
 The HC is well known to be involved in memory and spatial navigation in animals and humans.
@@
 consolidations based on reenactment and for integration into
-rich mental scene construction. In this view, the HC my even influence
+rich mental scene construction. In this way, the HC may even influence
 ongoing perception in the current environment
@@
 
-\subsection{The TPJ: Situation-understanding and context-switching}
+\subsection{The right and left TPJ: Prediction error signaling and world semantics}
 Finally,
@@
 %
-The anterior RTPJ (\textbf{XXX ???})  is a key node network is central for
+The anterior RTPJ is a key node of the DMN that  is central for
 action initiation during externally structured tasks and
 sensorimotor control by integrating supramodal stimulus-guided attention
-\citep{corbetta2002control}. \textbf{XXX Previous sentence is weird!}
+\citep{corbetta2002control}.
 Involvement of this DMN node was repeatedly reported in
@@
 
-\section{Reinforcement Learning (RL) as a computational model for DMN function}
+\section{Reinforcement Learning as a computational model for DMN function}
 We now argue that the DMN has biological substrates (gradient
@@
 \subsection{Friston's Free-energy principle (refer to \ref{label:friston})}
+\textbf{XXX: add tons of more examples and connnections}
 In summary:
@@
   \item In sum,
-  the MDP account may serve as a concrete conceptualization of the PC account in cognitive neuroscience. MPDs have the advantage of exposing an explicit
+  the MDP account may serve as a concrete conceptualization of the PC account
+  in cognitive neuroscience. MPDs have the advantage of exposing an explicit
   mechanism of the horizon of future considerations or
@@
   and collapses sensory input processing and action output preparation.
+  %
+  As such, our account can be viewed as a RL implemnetion of the
+  prediction coding hypothesis.
 \end{itemize}
 
-\begin{mdframed}
-  \textbf{XXX:} Rather than a heads-on competitor, our proposed formalism should be seen
-  and sold as a solid implementation of the predictive coding hypothesis.
-\end{mdframed}
 \subsection{Semantic Hypothesis}
@@
 Going much beyond human language ifself,
-semantic building blocks provide the basis to mentally envision non-existent scenarios
+semantic building blocks provide the basis to
+mentally envision non-existent scenarios
 that would improve optimal behavior in the environment

### 2017-03-26 paper/paper.tex
more fixups
biology section: action titles

--- 2017-03-26 paper/paper.tex
@@
 The ventromedial prefrontal cortex (vmPFC) subserves
-less ambiguous subjective-value-related evaluative processes
-reward-informed and risk estimation of self-relevant environmental stimuli.
+less ambiguous subjective-value-related evaluative processes and
+reward-informed risk estimation of self-relevant environmental stimuli.
 \textbf{XXX: previous sentence needs to be punctuated!}

### 2017-03-26 paper/paper.tex

@@
 In artificial intelligence (AI), a popular abstract model for such an environment are so-called
-\textit{Markov Decision Processes (MDPs)}. According to this model, the environment's dynamics are
-fully determined by its current state.
-In RL, MDPs are the mathematical instantiation of the environment model,
-which generally can incorporate action probabilities (policy matrix)
-and/or value function given a state.
-(Silver: The environment can be fully defined by state space and
-transition probability matrix.)
-is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
+\textit{Partially Observable Markov Decision Processes (POMDPs)}.
+A POMDP models a decision process in which it is assumed that environment dynamics are determined by a Markov process,
+but the agent cannot directly observe the underlying state.
+Instead, it must maintain a probability distribution over the set of possible states, based on a set of observations and observation probabilities, and the underlying Markov process.
+This is a minimal amount of assumptions that can be made about an environment, and is charastistic of so-called \textit{model-free RL}, the approach adapted by this work.
+In contrast, \textit{model-based RL} starts off with some mechanistic assumptions about the dynamics of world.
+These can be assumptions about the physical laws governing the agent's world or constraints on the state-space and transitions thereupon.
+For example if a rat in a maze knows, for sure,
+that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food, they can assume:
+\begin{itemize}
+\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
+\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
+\item etc.
+\end{itemize}
+
+Mathematically, a POMDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
 \begin{itemize}

### 2017-03-26 paper/paper.tex
integrated Elvis' comments

--- 2017-03-26 paper/paper.tex
@@
 $\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action $(s,a)$ into an appropriate
-feature space. Such are the techniques that have been proposed by \cite{songNIPS2016}, for example.
-%  and Parr et al. 2003 (JMLR)].
-A breakthrough alternative proposed by Google DeepMind \cite{mnih2015,silver2016mastering}, is to learn this
+feature space. Such are the techniques that have been proposed by [sutton et al. xyzw, Song et al. 2016 (NIPS),
+  and Parr et al. 2003 (JMLR)].
+A breakthrough alternative proposed by Google DeepMind [Minh et al. 2014, Silver et al 2015] is to learn this
 representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which
@@
 \subsection{Friston's Free-energy principle (refer to \ref{label:friston})}
-\textbf{XXX: add tons of more examples and connnections}
 In summary:
@@
   \item In sum,
-  the MDP account may serve as a concrete conceptualization of the PC account
-  in cognitive neuroscience. MPDs have the advantage of exposing an explicit
+  the MDP account may serve as a concrete conceptualization of the PC account in cognitive neuroscience. MPDs have the advantage of exposing an explicit
   mechanism of the horizon of future considerations or
@@
   and collapses sensory input processing and action output preparation.
-  %
-  As such, our account can be viewed as a RL implemnetion of the
-  prediction coding hypothesis.
 \end{itemize}
 
+\begin{mdframed}
+  \textbf{XXX:} Rather than a heads-on competitor, our proposed formalism should be seen
+  and sold as a solid implementation of the predictive coding hypothesis.
+\end{mdframed}
 \subsection{Semantic Hypothesis}
@@
 Going much beyond human language ifself,
-semantic building blocks provide the basis to
-mentally envision non-existent scenarios
+semantic building blocks provide the basis to mentally envision non-existent scenarios
 that would improve optimal behavior in the environment

### 2017-03-26 paper/paper.tex
blah
added notes for cover letter
new notes

@@
 probabilistic model of anticipated events''
-\citep{raichle2005intrinsic}. This paper proposes a formal model that satisfies this contention.
+\citep{raichle2005intrinsic}.
+The idea has been invigorated by empirical evidence from
+neuroscientific experiments \citep{kording2004bayesian, fiser2004small}.
+This paper proposes a formal model that satisfies this contention.
 
@@
 into the tension between exploitative action with immediate gains and
-explorative action with longer-term reward schedules.
+explorative action with longer-term reward schedules
+\citep{dayan2008decision}.
 The implication of the DMN in a diversity of human cognitive processes
@@
 
-In artificial intelligence (AI), a popular abstract model for such an environment are so-called
+In artificial intelligence (AI), a popular abstract model for
+multi-step decision processes in such an environment are so-called
 \textit{Partially Observable Markov Decision Processes (POMDPs)}.
@@
 but the agent cannot directly observe the underlying state.
+\textbf{XXX say quite early what this control-theoretical object is optimizing
+to inferring solutions to decision problems of what to do next or what action to take;
+sequential decision making}
 Instead, it must maintain a probability distribution over the set of possible states, based on a set of observations and observation probabilities, and the underlying Markov process.
@@
 For example if a rat in a maze knows, for sure,
-that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food, they can assume:
+that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food.
+A core property of human intelligence has been proposed underly to improve
+on an expected utility outcomes as a strategy for action choice in uncertain
+environments \citep{gershman2015computational}.
+For both we can assume:
 \begin{itemize}
@@
 \textbf{XXX: add tons of more examples and connnections}
+\textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
+\textbf{XXX: integrate aspects from WakeNsleep by Hinton}
 In summary:
@@
   %
-  As such, our account can be viewed as a RL implemnetion of the
+  As such, our account can be viewed as a RL implemention of the
   prediction coding hypothesis.

### 2017-03-26 paper/paper.tex

@@
 In the brain, such an experience replay mechanism is probably implemented in the hyppocampus (HC). Indeed,
-Hippocampal replay is a phenomenon observed in rats, mice, cats, rabbits, songbirds, and monkeys.
+hippocampal replay is a phenomenon observed in rats, mice, cats, rabbits, songbirds, and monkeys.
 Indeed, during sleep or awake rest, replay refers to the re-occurrence of a sequence of cell activations that also occurred during activity, but the replay has a much faster time scale.

### 2017-03-26 paper/paper.tex
more notes

@@
 posterior hippocampus \citep{maguire2000navigation}.
+Indeed, place-specific neuronal populations in the hippocampus
+spike one after another when an animal is choosing between alternative
+paths \citep{johnson2007neural}; and
+these neuronal patterns apear to also indicate upcoming behavior
+\citep{pfeiffer2013hippocampal}.
+
 But encoding and generative reconstruction in the hippocampus extends
@@
 \textbf{XXX: integrate aspects from WakeNsleep by Hinton}
+\textbf{XXX: integrate aspects from Sutton/Barto book chapter}
 In summary:
@@
 \subsection{Predictive Coding Hypothesis}
-The predictive coding and Bayesian brain framework
+The predictive coding framework
+\citep{clark2013whatever}
 is a frequently evoked idea related to the default mode function

### 2017-03-26 paper/paper.tex
interim

@@
 \citep{lebreton2009automatic}.
-The NAc is thought to be linked to reward mechanisms that not
-only modulate motivated behavior towards basic survival needs but also
-subserve RL in humans more broadly.
+NAc activity is thought to reflect reward prediction signals from dopaminergic pathways
+\citep{schultz1998predictive}
+that not only modulate motivated behavior towards basic survival needs but also
+subserve RL in humans more broadly \citep{doherty2015structure}.
 This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys
@@
 in constructing mental models in general
-\citep{maguire2016, schacter2007remembering}.
+\citep{maguire2016, schacter2007remembering, gelbard2008internally}.
 Indeed,
@@
 \subsection{Value functions and policies}
-The behavior of the agent is governed by some kind of \textit{policy}, which maps states to
-sampling probabilities over candidate actions to perform from this state. Starting a time $t=0$,
+The behavior of the agent is governed by some kind of \textit{policy},
+which maps states of the world to
+a set of candidate actions to perform given this state. Starting a time $t=0$,
 a policy $\pi$ generates a trajectory as follows
@@
   \item Interactions between external and internal sources of
-  information are probably commonplace during perception
+  information are probably commonplace during perception \citep{gershman2010learning}
   \item The MDP frameworks how the human brain could solve
@@
 \textbf{XXX: integrate aspects from WakeNsleep by Hinton}
+\textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
 \textbf{XXX: integrate aspects from Sutton/Barto book chapter}
@@
 The predictive coding framework
-\citep{clark2013whatever}
+\citep{clark2013whatever, friston2008hierarchical}
 is a frequently evoked idea related to the default mode function
@@
 \citep{bar2009proactive}.
+In this way, the encoding of complex environments could be facilitated
+by relative similarity associations between states.
 %
@@
 by recombining building blocks of
-concept knowledge and episodic memories.
+concept knowledge and episodic memories
+\citep{hassabis2009construction}.
 Indeed, in aging humans, remembering the past and imaging the future
@@
 why mind-wandering and dreams
-are so rich in stories about people and the complex relationships between them.
+are so rich in stories about people and
+the complex relationships between them
+\citep{schilbach2008minds}.
 
@@
 %
-Viewed as an MDP, behavior unfolds by integrating happened past events
+From the computational viewed of a MDP,
+behavior unfolds by integrating happened past events
 and possible future events to guide action choice in the present context.

### 2017-03-26 paper/paper.tex
finish monster-pass

@@
   by
-  % Control Theory and
-  Reinforcement Learning Theory}
+  Control Theory and
+  Reinforcement Learning}
 
@@
 have been shown to be reenacted spontaneously during
-quiet wakefulness and sleep \citep{hartley2014space}.
+quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
 Moreover, spike sequences measured in hippocampal place cells of rats
@@
 If the process has $memory$ such that the next state depends not only on the current state but also on some finite
-number of past states, it can reformulated as a standard memoryless Markov process by simply expanding the
+number of past states,
+rational probabilistic planning can reformulated
+as a standard memoryless Markov process by simply expanding the
 definition of the state $s$ to include elements of the past.
+\textbf{these first sentences are a little too abstract.
+Take inspiration from here: An MDP represents an agent’s model of its interaction with its environment.
+MDPs encode all relevant information about the configuration of the world and the agent with the state variable.
+MDPs also represent the affordances of the environment: what actions the agent can take and a causal model of how these actions change the state of the world.
+Finally, MDPs represent the subjective rewards or costs caused by the agent’s actions in each state.}
 
@@
     human mental scenes by their extremely high relevance to humans
-    reflected by high values for social information in the value matrix.
+    reflected by high values for social information in the value matrix
+    \citep{baker2009action}.
     \item How can a same neurobiological circuit be equally important

### 2017-03-27 paper/paper.tex
addressing some comments + general polishing of the bloody bear

@@
 sentinel hypothesis, etc.) concerning decision-making in the brain
-is drawn, with the DMN at its core. Our work
+is drawn, with the DMN at its core. Our work ...
 
 \section{Known Neurobiological Properties of the Default Mode Network}
-\textbf{XXX This section is presented as transfinite block of concatenated text.
-  Please reorganize into subsections and paragraphs!}
 We will first attempt to deconstruct the overall DMN function
@@
 
-\section{Reinforcement Learning as a computational model for DMN function}
-We now argue that the DMN has biological substrates (gradient
-computation, backprop, experience replay, etc.) for all the
-major components of an efficient \textit{reinforcement learning (RL)} algorithm, and all these
+\section{Reinforcement Learning, a computational model for DMN function}
+We now argue that the DMN is equipped with sufficient neurobiological substrates gradient
+for implementing all the major components (memory, reward signaling, accumulators, random sampling, etc.)
+of an efficient \textit{reinforcement learning (RL)} algorithm, and all these
 components integrate in a plausible way.
 Reinforcement Learning (RL) is learning to achieve a goal by interacting with an external
-environment. In a general RL framework, an \textit{agent}
+\textit{environment / world}. In a general RL framework, an \textit{agent}
 interacts with an  environment in a trial-by-error fashion by taking
@@
 definition of the state $s$ to include elements of the past.
-\textbf{these first sentences are a little too abstract.
-Take inspiration from here: An MDP represents an agent’s model of its interaction with its environment.
-MDPs encode all relevant information about the configuration of the world and the agent with the state variable.
-MDPs also represent the affordances of the environment: what actions the agent can take and a causal model of how these actions change the state of the world.
-Finally, MDPs represent the subjective rewards or costs caused by the agent’s actions in each state.}
 
@@
 A POMDP models a decision process in which it is assumed that environment dynamics are determined by a Markov process,
-but the agent cannot directly observe the underlying state.
-\textbf{XXX say quite early what this control-theoretical object is optimizing
-to inferring solutions to decision problems of what to do next or what action to take;
-sequential decision making}
-Instead, it must maintain a probability distribution over the set of possible states, based on a set of observations and observation probabilities, and the underlying Markov process.
-This is a minimal amount of assumptions that can be made about an environment, and is charastistic of so-called \textit{model-free RL}, the approach adapted by this work.
+but the agent cannot directly observe the underlying state. Instead, it tries to optimize a \textit{subjective} reward
+signal which is tried alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
+states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
+and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.
 In contrast, \textit{model-based RL} starts off with some mechanistic assumptions about the dynamics of world.
@@
 that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food.
-A core property of human intelligence has been proposed underly to improve
-on an expected utility outcomes as a strategy for action choice in uncertain
-environments \citep{gershman2015computational}.
-For both we can assume:
+% A core property of human intelligence has been proposed underly to improve
+% on an expected utility outcomes as a strategy for action choice in uncertain
+% environments \citep{gershman2015computational}.
+Our rat might represent such knowledge about the world as:
 \begin{itemize}
@@
 The agent's goal can then be formulated as finding policy $\pi$ which maximizes the
-expected cumulative value of state-stat-pairs $(A_0,S_0)$, namely
+expected cumulative value of state-stat-pairs $(A_0,S_0)$, also known as the \textit{Q-value}, namely
 \begin{equation}
@@
 \end{equation}
-That is, the expectated future reward over all possible trajectories in which we start at $s$ take
-action $a$ and then follow the policy $\pi$ thereafter to select future actions.
+\begin{mdframed}
+In plain English, the Q-value $Q^{\pi}(s,a)$ is the expected / average / mean future reward
+over all possible trajectories in which the agent starts off starts at with an environment in state
+$s$, takes and action $a$, and then follows the policy $\pi$ thereafter to select future actions.
+\end{mdframed}
 %% Given that each $\pi$ correspondence to an exhaustive description of
@@
 consider.
-(If the Bellman-based computational ideas have a corresponding
-neurobiological implementation, this provides a framework of how
-optimal behavior could be computable in the human brain.)
+% \begin{mdframed}
+%   \textbf{N.B.:}
+%   If the Bellman-based computational ideas have a corresponding
+%   neurobiological implementation, this provides a framework of how
+%   optimal behavior could be computable in the human brain.
+% \end{mdframed}
 One popular simplification that removes this issue is to only consider the sub-family of
@@
 % table.
-The celebrated \textit{Bellman's equation} [Sutton and Barto 1998] then takes the simple form
+The celebrated \textit{Bellman's equation} \cite{sutton1998reinforcement} then takes the simple form
 \begin{equation}
-  Q^\pi(s, a) = \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
-  = \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))]
+  \begin{split}
+    Q^\pi(s, a) &= \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
+    \\
+    &= \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))],
+  \end{split}
   \label{eq:bellman}
 \end{equation}
-which provides a recursive decomposition of optimal behavior.
+a \textit{fixed-point} equation which provides a recursive decomposition of optimal behavior.
 A complicated dynamic programming optimization can be
-broken into simpler subproblems at different time points,
+broken into simpler sub-problems at different time points,
 which has often been closely related to the medial prefrontal part of the DMN
@@
 Using the Bellman equation, each state can be associated with a certain value,
-to guide action (i.e., improve the policy) towards a better state.
-It is a fixed-point equation for the deterministic policy $\pi$.
+to guide action towards a better state, thus improving on the agent's current policy.
+% It is a fixed-point equation for the deterministic policy $\pi$.
 Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which
@@
 \mapsto {Q}(s, a|\theta^Q_{k})$, where $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q_{k+1}$.
+\begin{mdframed}
 In the brain, such an experience replay mechanism is probably implemented in the hyppocampus (HC). Indeed,
@@
 Indeed, during sleep or awake rest, replay refers to the re-occurrence of a sequence of cell activations that also occurred during activity, but the replay has a much faster time scale.
-% It may be in the same order, or in reverse.
+\end{mdframed}
+
 Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function

### 2017-03-27 paper/paper.tex
more

@@
 % Switching between states reduces to a sequential processing model.
-In a model-based reinforcement learning (RL) approach,
-we try to understand the environment always better
-starting, at birth, with an environment that is completely unknown.
+% In a model-based reinforcement learning (RL) approach,
+% we try to understand the environment always better
+% starting, at birth, with an environment that is completely unknown.
 Informed by outcomes of performed action,
@@
 into the tension between exploitative action with immediate gains and
-explorative action with longer-term reward schedules.
-The implication of the DMN in a diversity of human cognitive processes
-can be parsimoneously recast as prediction error minimization
+exploratory action with longer-term reward schedules.
+
+\subsection{Sketch of our main contributions}
+Our theory proposes that the implication of the DMN in a diversity of human cognitive processes
+can be parsimoniously recast as prediction error minimization
 based on pervasive probabilistic simulations,
@@
 
-\subsection{Sketch of main contributions}
 We show that the DMN is endowed with sufficient neurobiological equipment to
@@
 \section{Known Neurobiological Properties of the Default Mode Network}
-We will first attempt to deconstruct the overall DMN function
+We begin by a deconstruction the overall DMN function
 by the mechanistic relevance of each individual network node

### 2017-03-27 paper/paper.tex
on possible connection between iRL and hippocampal replay

@@
 
-\subsection{Inverse reinforcement-learning}
+\subsection{Hippocampal replay = inverse reinforcement-learning ?}
 %\begin{mdframed}
@@
 In the general artificial case, the problem has been extensively studied and partially solved by
-\cite{abbeel2004}. They are also been rigourously studied in general optimal control literature under the name ``inverse optimal control'', but in model-based certains (where the physical dynamics
-are known, etc.)...
+\cite{abbeel2004}. They are also been rigourously studied in general optimal control literature under the name
+``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
 
 IRL is is suited for problems in which it's hard to define what the reward function could be
-(e.g car-driving, drone maneuvering, etc.
+(e.g car-driving, drone maneuvering, etc.) ...
 

### 2017-03-28 paper/paper.tex
start of big pass

@@
 \begin{abstract}
-The default mode network (DMN) is the brain manifestations of the human
-baseline mental activity.
+The default mode network (DMN) reflects the  baseline mental activity
+in humans.
 %
@@
 envisioning of mental scenarios to predict the environment.
-The DMN would hence be dedicated to continuous autobiographical memory retrieval,
+The DMN is believed to be dedicated to
+continuous autobiographical memory retrieval,
 generation of hypothetical outcomes,
 and reward contingency evaluation when letting the mind go.
-It explains its highest energy consumption in the brain and
-its intimate coupling with conscious awareness.
+Indeed, the DMN probably accounts for the
+highest energy consumption in the brain and
+is intimately coupled with conscious awareness.
 %
@@
 in the so-called "default mode network" (DMN)
-\citep{shul1997}. This networked collection of some of the highest regions in
+\citep{shul1997}. This networked collection of some of brain regions at the top of
 the human association cortex \citep{mesulam1998sensation, margulies2016situating}
@@
 unfocused everyday mind wandering \citep{raichle2001pnas}.
-In the beginning of the 21st century, brain imaging was hence the first technology
+In the beginning of the 21st century, brain imaging
+was perhaps the first existing technology
 to allow for the discovery of a unique brain network that subserves
 baseline mental activities
-\citep{bzdok2015resting}.
+\citep{raichle2007default, bzdok2015resting}.
 
@@
 deactivation during various psychological experiments.
-This processing of unknown information categories in the DMN has
+This processing of yet-to-be-discovered information categories in the DMN has
 been argued to mediate an evolutionarily conserved function for the individual.
-This is all the more likely because the DMN contains the two
+This is all the more likely becaus
+the DMN has been shown to contain the two
 biggest hotspots of energy consumption in the entire central nervous system
@@
 A particularly attractive computational framework
-to describe, quantify, and predict autonomously acting systems like the brain
-can be the combination of control theory and reinforcement learning.
-It is known that, the more the external world is predictable,
+to describe, quantify, and experimentally test
+autonomously acting systems like the brain
+can be the combination of control theory and reinforcement learning (RL).
+It is long known that, the more the external world is predictable,
 the more mental activity becomes detached from the actual sensory environment
@@
 These ``offline'' processes may however contribute to optimized control of the organism.
-% We formalize
-% a policy matrix to capture the space of possible actions that can be performed
-% on the environment given the current state. A value matrix
-% maps expected rewards to environmental objects and events.
 % Switching between states reduces to a sequential processing model.
@@
 This paper proposes a formal model that satisfies this contention.
-
 We show that the DMN is endowed with sufficient neurobiological equipment to
-implement a reinforcement learning (RL) agent, at least in principle. A global
-picture of the current body of knowledge (predictive coding, free-energy principle,
-sentinel hypothesis, etc.) concerning decision-making in the brain
-is drawn, with the DMN at its core. Our work ...
+implement a reinforcement learning agent, at least in principle.
 
@@
 that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food.
+We formalize
+a policy matrix to capture the space of possible actions that can be performed
+on the environment given the current state. A value matrix
+maps expected rewards to environmental objects and events.
+
 % A core property of human intelligence has been proposed underly to improve

### 2017-03-28 paper/paper.tex
corrections

@@
 \section{Known Neurobiological Properties of the Default Mode Network}
-We begin by a deconstruction the overall DMN function
-by the mechanistic relevance of each individual network node
-based on published findings from neuroscience experiments.
-%
-\subsection{The posteriormedial cortex: Global monitoring and integration}
-The posteriormedial cortex (PMC) and
+We begin by a deconstruction of the overall DMN function
+by reviewing published findings from neuroscience experiments
+on the mechanistic relevance of each individual network node.
+%
+\subsection{The posteromedial cortex: Global monitoring and integration}
+The posteromedial cortex (PMC) and
 the dorsomedial prefrontal cortex (dmPFC) within the DMN
-are responsible for the highest metabolic turn-over
+are probably responsible for the highest metabolic turn-over
 of glucose energy consumption in humans \citep{raichle2001pnas}.
@@
 including anesthesia,
-sleep, and forms of coma \citep{cavanna2006precuneus}.
+sleep, and coma \citep{cavanna2006precuneus}.
 %
@@
 the environmental statistics and its internal representation
-as an "inner minds eye" \citep{cavanna2006precuneus}.
+to resemble "inner minds eye" \citep{cavanna2006precuneus}.
+
 For instance, B\'alint's syndrome is a neurological disorder of conscious
@@
 environment into an integrated whole (i.e., simultanagnosia)
-as well as inability to direct action towards
+as well as by the inability to direct action towards
 currently unattended environmental objects
@@
 with the amygdala involved in significance evaluation and
-the nucleus accumbens (NAc), involved in reward evaluation.
+the nucleus accumbens (NAc), involved in reward evaluation
+\citep{bzdok2015subspecialization}.
 Specifically, among all parts of the PMC,
@@
 most connected to the laterobasal
-nuclei group of the amygdala
-\citep{bzdok2015subspecialization}.
+nuclei group of the amygdala.
 This amygdalar subregion has been proposed to
@@
 %
-Indeed,
+In support of this contention,
 electrophysiological recordings in animals implicated the PMC in
@@
 consistently associated with anterograde and retrograde memory impairments
-of various kinds of sensory information
-in rabbits and humans
+of various sensory inputs in rabbits and humans
 \citep{vann2009does}.
 Regarding perspective frames, this PMC subregion has been
-proposed to mediate between the organism's egocentric
+proposed to mediate between the organism's egocentric viewpoint
 (i.e., focused on sensory input) and
-allocentric (i.e., focused on world knowledge) viewpoints
+allocentric viewpoint (i.e., focused on world knowledge) 
 in animals and humans
 \citep{epstein2008parahippocampal, burgess2008spatial, valiquette2007different}.
-%
-The PMC may consequently monitor the subjective outcomes
+
+The overall PMC may consequently monitor the subjective outcomes
 of possible decisions and integrates that information
@@
 to predict delayed reward opportunities in the future.
-In doing so, the PMC continuously adapts to changes
+In doing so, the PMC needs to continuously adapts to changes
 in both the external environment and its internal representation
@@
 
-\subsection{The prefrontal cortex: Stimulus-value assocation and action consideration}
-The dmPFC subserves predominantly ambiguous amodal processes
+\subsection{The prefrontal cortex: Stimulus-value association and action consideration}
+The dmPFC subserves predominantly ambiguous processes
 across time, space, and content domains in
@@
 The dmPFC may thus enable inference, representation, and assessment
-of one's own and other individuals' action and thoughts.
+of one's own and other individuals' actions and thoughts.
 Generally,
@@
 According to computational neuroimaging experiments,
-the dmPFC activity preferentially models stimulus-value associations of
+neural activity in the dmPFC
+preferentially models stimulus-value associations of
 possible actions that are not actually executed,
-whereas the vmPFC activity models value in the environment that leads performed actions
+whereas neural activity
+in the ventral portion of the medial PFC models value in the environment that leads performed actions
 \citep{nicolle2012agent}.
@@
 The ventromedial prefrontal cortex (vmPFC) subserves
-less ambiguous subjective-value-related evaluative processes and
-reward-informed risk estimation of self-relevant environmental stimuli.
-This DMN node is more closely associated with
+evaluation of subjective value as well as
+reward and risk estimation of environmental stimuli.
+In contrast to the dmPFC,
+the vmPFC node of the DMN is more closely associated with
 orchestrating adapted behavior by bottom-up-driven
 processing of “what matters now”,
-probably drawing on model-based value representations
+probably drawing on internal value representations
 \citep{doherty2015structure}.
-Quantative lesion findings across 344 human individuals confirmed
+Quantitative lesion findings across 344 human individuals confirmed
 a gross impairment in value-based decision making
 \citep{glascher2012lesion}.
-The vmPFC is preferentially connected with limbic and reward-related areas.
+The vmPFC is preferentially connected with areas of the
+limbic and reward systems.
 The vmPFC has been observed to have monosynaptical connections
-with the nucleus accumbens
+with the NAc
 in axonal tracing studies in monkeys \citep{haber1995orbital}.
-Additionally, the gray-matter volume of the vmPFC and nucleus accumbens
+Additionally, the gray matter volume of the vmPFC and NAc
 correlated with indices of value-guided behavior and social reward attitudes
@@
 subserve RL in humans more broadly \citep{doherty2015structure}.
-This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys
+This is consistent with diffusion MRI tractography in humans and monkeys
 \citep{croxson2005quantitative} that
-quantified this brain regions to
+quantified NAc to
 be substantially more likely connected to the vmPFC than dmPFC in both species.
 Two functional connectivity modalities in humans strongly connected
-the vmPFC with the nucleus accumbens, hippocampus (HC),
-and posteriomedial cortex.
-%
+the vmPFC with the NAc, hippocampus (HC),
+and PMC.
+
+In fact,
 The vmPFC is often proposed to be involved in (external) emotional
@@
 This may be why performance within and across participants
-was related to state encoding in OFC.
+was related to state encoding in this part of the brain.
 Such a \textit{cognitive map} of action space was argued to encode
@@
 independent whole-brain analyses from structural
-neuroimaging studies related the gray-matter volume GMV of the vmPFC,
+neuroimaging studies related the gray matter volume GMV of the vmPFC,
 more consistently than any other
-brain region, to indices of
-social competence and social network complexity,
-among the most complicated decision that humans and monkeys take
-\citep{behrens2009computation}.
+brain region, to complex decision making
+in humans and monkeys \citep{behrens2009computation}.
 %
@@
 London taxi drivers, individuals with high performance in spatial navigation,
-were shown to exhibit increased grey matter volume in the
+were shown to exhibit increased gray matter volume in the
 posterior hippocampus \citep{maguire2000navigation}.
@@
 to the neocortex for long-term storage, potentially also in the nodes of the DMN.
-These mechanisms of the hippocampus probably make important contributions to the
+Such mechanisms of the hippocampus could also make important contributions to the
 recollection of autobiographical memory episodes and other
@@
 %
-The HC thus orchestrates elements of experienced environmental aspects for
-consolidations based on reenactment and for integration into
+The HC thus reenacts elements of experienced environmental aspects for
+consolidations and potentially, for integration into
 rich mental scene construction. In this way, the HC may even influence
@@
   The amygdala is known to be involved in significance evaluation) and
-  the nucleus accumbens (NAc) is involved in reward evaluation.
+  the NAc is involved in reward evaluation.
 \end{mdframed}

### 2017-03-28 paper/paper.tex
monstrous pass some technical parts

--- 2017-03-27 paper/paper.tex
@@
 that are calibrated by prediction error.
+
 A DMN framework incorporating RL can naturally embed human behavior
@@
 
-\subsection{Value functions and policies}
+\subsection{Long-term rewards and policies}
 The behavior of the agent is governed by some kind of \textit{policy},
@@
   \begin{split}
-    S_1 &\sim p(.|S_0,A_0), \text{ collect rewards }R_0 = r(S_0, A_0)\\
-    A_1 &\sim \pi(.|S_1)\\
-    S_2 &\sim p(.|S_1,A_1), \text{ collect rewards }R_1 = r(S_1, A_1)\\
+    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect rewards }R_0 = r(s_0, a_0)\\
+    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
+    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect rewards }R_1 = r(s_1, a_1)\\
     \ldots\\
-    A_{t} &\sim \pi(.|S_{t})\\
-    S_{t+1} &\sim p(.|S_{t},A_{t}), \text{ collect rewards }R_{t} = r(S_{t}, A_{t})\\
+    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
+    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect rewards }R_{t} = r(s_{t}, a_{t})\\
     \ldots
@@
 
+\subsection{The Q-value: a goodness measure for actions}
 The agent's goal can then be formulated as finding policy $\pi$ which maximizes the
-expected cumulative value of state-stat-pairs $(A_0,S_0)$, also known as the \textit{Q-value}, namely
+expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
 \begin{equation}
   \label{eq:qval}
-  Q^{\pi}(s,a) = \mathbb E [G^\pi|S_0=s,A_0=a,A_1 \sim \pi,A_2 \sim \pi,\ldots].
+  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a,a_1 \sim \pi,a_2 \sim \pi,\ldots].
 \end{equation}
-\begin{mdframed}
 In plain English, the Q-value $Q^{\pi}(s,a)$ is the expected / average / mean future reward
@@
 $s$, takes and action $a$, and then follows the policy $\pi$ thereafter to select future actions.
-\end{mdframed}
 %% Given that each $\pi$ correspondence to an exhaustive description of
-%% behavioral patterns, $Q^{\pi}(S_t,A_t)$ corresponds to the epistemic value of
+%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
 %% the set of actions given states in finite time.
+For the brain,
+$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
+utility of executing a specific action; it answers the question
+"What is the expected utiltiy of taking action $a$ in this situation ?".
+When a choice of actions is available, the $Q^{\pi}(s,a)$
+provides a formalization of optimal selection.
+
+\subsection{Computing optimal behavior}
 By construction, maximizing the expectation in  \eqref{eq:qval} is intractable in general
@@
 should be taken at thin state, and therefore define functions from states to actions.
-\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
+
+\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
 \textit{greedy policy}
@@
   \begin{split}
-    Q^\pi(s, a) &= \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
+    Q^\pi(s, a) &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
     \\
-    &= \mathbb E_{s' \sim p(.|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))],
+    &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))],
   \end{split}
@@
 
-\subsection{Efficient RL with value approximation and replay}
+\subsubsection{Efficient learning via value approximation and experience replay}
 %As already mentioned in the previous section, Q-learning optimizes over the class of
@@
 \mapsto {Q}(s, a|\theta^Q_{k})$, where $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q_{k+1}$.
-\begin{mdframed}
+
 In the brain, such an experience replay mechanism is probably implemented in the hyppocampus (HC). Indeed,
@@
 Indeed, during sleep or awake rest, replay refers to the re-occurrence of a sequence of cell activations that also occurred during activity, but the replay has a much faster time scale.
-\end{mdframed}
 
@@
 
-\paragraph{Link to classical RL algorithms.}
+\begin{mdframed}
+\textbf{A note on classical RL algorithms.}
 Note that most RL algorithms, including \textit{Temporal Difference (TD)}
@@
 %% $$\theta^Q(s,a) \leftarrow \theta^Q(s,a) + (Q(s, a|\theta^Q_{k+1}) - y)$$ ??
-
-\paragraph{TODO: Monte Carlo optimization.}
-Some text snippets that might be pertinent:
-Monte-Carlo simulation can be used to update the agent’s belief state
-\citep{silver2010monte}.
-Monte-Carlo simulation provides a simple method for evaluating
-a state s.
-In any MDP, there is at least one optimal policy matrix that
-achieves the optimal value function???
-The simulator is used to generate sequences of states, observations and rewards.
-These simulations are used to update the value function, without ever looking inside the black box describing the model’s dynamics.
-In addition, Monte-Carlo methods have a sample complexity that is determined only by the underlying difficulty of the POMDP, rather than the size of the state space or observation space.
-Monte-Carlo simulation can be turned into a simple control algorithm by evaluating all legal actions and selecting the action with highest evaluation [15].
-Monte-Carlo simulation can be extended to partially observable MDPs.
+\end{mdframed}
+
+\subsubsection{Interlude: zoom on the Hippocampus}
+\paragraph{Does the hippocampus do Monte-Carlo sampling ?}
+In RL, Monte-Carlo simulation can be used to update the agent's belief state
+\citep{silver2010monte}. Such methods have a sample complexity that is determined only by the underlying difficulty of the POMDP, rather than the size of the state space or observation space,
+which can be prohibitively large.
+Monte-Carlo simulation provides a simple method for evaluating a state $s$.
+They provide an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
+
+In the brain, the HC would generate / imagine sequences of world states, observations and rewards.
+These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
+This would be a simple control algorithm by evaluating all legal actions and selecting the action with highest evaluation.
 In these challenging POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
-Link this to experience sampling in the hippocampus.
 Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
@@
 
-\subsection{Correspondence between DMN function and RL framework}
-\begin{figure}[!h]
-  \includegraphics[width=.9\linewidth]{rl_process_chart.pdf}
-  \caption{The DMN as an RL agent...}
-  \label{fig:rl_process_chart}
-\end{figure}
-
-\begin{itemize}
-  \item The policy matrix represents the repertoire of possible actions
-  on the world given a current state. It encodes the probabilites of
-  choosing actions to be executed in a certain situation.
-  \item Interactions between external and internal sources of
-  information are probably commonplace during perception \citep{gershman2010learning}
-  \item The MDP frameworks how the human brain could solve
-  environmental problems over extended time periods.
-  \item Given a state, the value function represents the estimated utility of
-  being in that state, which is subjective to the agent.
-  For a state $s$, the value function at $s$ namely $V(s)$ define provides a quantitative answer to the question:
-  "How desirable is it to be in state $s$ ?"
-  For each action $a$ that can be taken from state $s$,
-  the Q-value  $Q(s, a)$ defined in \eqref{eq:qval} provides the subjective
-  utility of executing a specific action; it answers the question
-  "What is the expected utiltiy of taking action $a$ in this situation ?"
-  \item The DMN may subserve
-  constant exploration of possible future actions and their
-  cumulative outcomes. Implicit computation of future choices
-  provides an explanation for purposeful mind-wandering.
-  \item hippocampus -> provides perturbed action-transition-state-reward samples
-  as a block of "imagined", "hypothesized", "recalled" experience;
-  the small variations in these experience blocks allow searching
-  a larger space of parameters and possible experiences.
-  In the absence of environmental input and feedback
-  (e.g., mind-wandering or sleep) pseudo-experiencing possible
-  future scenarios and action outcomes.
-  Our approach acknowledges the unavoidable stochasticity of
-  computation in neuronal processes \citep{faisal2008noise}.
-  \item
-  \textit{Inference} in the human brain reduces to generalization of
-  policy and value adaptations from sampled experiences to
-  successful action choices and reward predictions in future states.
-  \item In the RL view, plasticity in the DMN arises naturally.
-  If an agent behaving optimally in a given environment moves
-  to novel, yet unexperienced environments, reward prediction errors will
-  massively increase. This will lead to adaptation of the policy matrix
-  and value function until the system converges to a
-  new steady-state of optimal action decisions given certain states.
-  \item The DMN is today known to consistently increase in neural
-  activity when engaging cognitive processes that are detached from
-  current sensory environment. DMN involvement in thinking about the past,
-  the future, and hypothetical possibilities ties in with the
-  MDP's implicit computation of action and state cascades as a function
-  of what happened in the past.
-\end{itemize}
-
-\paragraph*{Some notes:}
-\begin{itemize}
-  \item Actions are continuous, whilst states are continuous.
-  \item The brain does backprop (see Hinton's recent talk at Stanford)
-  \item model weights $\theta^Q$ should correspond to connections between neurons
-\item humm, looks like such a knowledge would simply translate into constraints on the DqN weights $\theta^Q$ (... at least, part of such knowledge)
-  ok, gotta think about the model-free / model-based reconsiliation
-\end{itemize}
-
-\subsection{Hippocampal replay = inverse reinforcement-learning ?}
+\paragraph{Hippocampal replay = inverse reinforcement-learning ?}
 %\begin{mdframed}
@@
 
+\subsection{Conclusion: putting everything together}
+\begin{figure}[!h]
+  \includegraphics[width=.9\linewidth]{rl_process_chart.pdf}
+  \caption{The DMN as an RL agent...}
+  \label{fig:rl_process_chart}
+\end{figure}
+The DMN is today known to consistently increase in neural
+  activity when engaging cognitive processes that are detached from
+  current sensory environment \citep{kenet2003spontaneously, fiser2004small}.
+  Its involvement in thinking about the past,
+  the future, and hypothetical possibilities ties in with the implicit computation of
+  action and state cascades as a function of what happened in the past.
+% A policy represents the repertoire of possible actions
+%   on the world given a current state. It encodes the probabilites of
+%   choosing actions to be executed in a certain situation.
+% Interactions between external and internal sources of
+%   information are probably commonplace during perception \citep{gershman2010learning}
+The DMN may subserve
+  constant exploration of possible future actions and their
+  cumulative outcomes. Implicit computation of future choices
+  provides an explanation for purposeful mind-wandering.
+
+The HC provides perturbed action-transition-state-reward samples
+  as a block of "imagined", "hypothesized", "recalled" experience;
+  the small variations in these experience blocks allow searching
+  a larger space of parameters and possible experiences.
+  In the absence of environmental input and feedback
+  (e.g., mind-wandering or sleep) pseudo-experiencing possible
+  future scenarios and action outcomes.
+  Our approach acknowledges the unavoidable stochasticity of
+  computation in neuronal processes \citep{faisal2008noise}.
+
+In our proposed RL-based view, \textit{inference} in the human brain reduces to generalization of
+  policy and value adaptations from sampled experiences to
+  successful action choices and reward predictions in future states.
+  Plasticity in the DMN also arises naturally:
+  If an agent behaving optimally in a given environment moves
+  to novel, yet unexperienced environments, reward prediction errors will
+  massively increase. This will lead to adaptation of the policy matrix
+  and value function until the system converges to a
+  new steady-state of optimal action decisions given certain states.
+
+% \paragraph*{Some notes:}
+% \begin{itemize}
+%   \item Actions are continuous, whilst states are continuous.
+%   \item The brain does backprop (see Hinton's recent talk at Stanford)
+%   \item model weights $\theta^Q$ should correspond to connections between neurons
+% \item humm, looks like such a knowledge would simply translate into constraints on the DqN weights $\theta^Q$ (... at least, part of such knowledge)
+%   ok, gotta think about the model-free / model-based reconsiliation
+% \end{itemize}
+
+\section{Relation to other models decision-making and optimal control for the brain}
+\subsection{The free-energy principle and active inference}
+% \subsubsection{The theory}
+In Friston's free-energy principle (FEP) \citep{friston2010free,fristonAIorRL}, the brain emerges as biomechanical inference engine which must minimize the long-term average of surprise.
+Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net (DBN)}.
+According to FEP's account, the goal of the brain is then to optimize over the generative model $G$: to iteratively
+modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics, etc., so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles.
+These is called the \textit{generative} model.
+FEP also postulates a dual model called the \textit{recognition} model,
+which works in tandem with the generative model $p_R(\z|\x)$, to accomplish approximate inference.
+The recognition model dreams / fantasizes imaginary worlds $\z$ whilst the latter tries to generate sensations $\x$ which match these dreams!
+
+Because surprisal is intractably difficult to optimize (since we would need to sum over all hidden causes of the sensations), FEP sets out to instead minimize an upper-bound thereupon, namely the \text{free-energy} $F_G^R(\x)$ given by
+
+  \begin{equation}
+    \begin{split}
+      \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
+      &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
+      &\le F^R_G(\x),
+      \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
+    \end{split}
+  \end{equation}
+
+% \begin{figure}[!tb]
+%   \centering
+%   \includegraphics[width=.8\linewidth]{vae.pdf}
+%   \caption{Variational autoencoders...}
+%   \label{fig:vae}
+% \end{figure}
+The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}.
+\citep{dayan1995helmholtz}. As noted already in \citep{dayan1995helmholtz},
+a crucial drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recogntiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
+Thus the brain could not possible be running such a algorithm, not even in principle!
+To the rescue, we note that the recent theory of
+\textit{variational auto-encoders (VAEs)} \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm, as it overcomes the technical limits of the former, by using a reparametrization trick.
+For example, unlike the wake-sleep algorithm for minimizing free-energy, VAEs can be efficiently trained back-propagation of learning errors.
+
+\subsubsection{Comparison to our proposed theory}
+On the surface, a common point between the FEP and our proposed RL-based framework
+  is placing the minimization of a surprise signal at the core of brain function.
+  Indeed in RL, surprise minimization is subsumed  by accurate prediction of
+  rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist internal phase-transitions which lead contradictory with the
+  blurred version of the external world captured as sensations
+  \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}. Such a theory cannot by itself,
+  explain the emergence of strategic behavior inherent in humans (e.g cite dark-room experiment).
+
+% Friston's critic of RL \citep{fristonAIorRL} (he proposes ``active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
+% pioneered in the 1950s by Richard Bellman and colleagues [2]. To
+% optimise control a~pð Þ x~ under this formulation, we have to: (i)
+% assume the hidden states are available to the agent and (ii) solve
+% Equation 18 for the value-function. Solving for the value-function
+% is a non-trivial problem and usually involves backwards induction
+% or some approximation scheme like reinforcement learning [4–6].
+% The free-energy formulation circumvents these problems by
+% prescribing the policy in terms of free-energy, which encodes
+% optimal control (Equation 6)"
+
+% The following critics can be made:
+% \begin{itemize}
+%   \item
+%     As noted by Dayan et al.
+(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
+%   \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
+%   \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
+%     technical difficulty in getting derivatives of loss function w.r.t
+%     recognition weights $\W^R$).
+%   \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
+%   \end{itemize}
+
+\textbf{XXX: add tons of more examples and connnections}
+\textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
+\textbf{XXX: integrate aspects from WakeNsleep by Hinton}
+\textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
+\textbf{XXX: integrate aspects from Sutton/Barto book chapter}
+
+\subsection{Predictive Coding Hypothesis}
+The predictive coding framework
+\citep{clark2013whatever, friston2008hierarchical}
+is a frequently evoked idea related to the default mode function
+\citep{bar2007}.
+According to this cognitive account,
+cortical responses emerge from continuous functional interaction between
+higher and lower levels of the neural processing hierarchy.
+This dynamic intercourse enables inference on the world by reconciling
+gaps between fresh sensory input and stored prior information.
+Feed-forward sensory experience is constantly calibrated by
+top-down modulation at various hierarchical processing levels.
+In fact, axonal
+back projections probably outnumber by far the input projection
+existing in the brain \citep{salin1995corticocortical}.
+This can explain the constructive, generative nature of sensory perception
+\citep{friston2010free} and
+why motor action is intimately related to sensory expectations
+\citep{wolpert1995internal, kording2004bayesian}.
+At each stage of neural processing,
+an internally generated prediction of the sensory input is
+directly compared against the actual input from the environment.
+A prediction error at one of the processing levels,
+i.e. the difference between sensory input
+and internally predicted sensation,
+incurs plasticity changes of neuronal back projections (i.e., model parameters)
+for gradually improved future prediction of the environment.
+Contextual integration is hence maintained by top-down modulation of sensorimotor
+processing by a-priori information about the environment.
+The generative model of how perceived sensory events arise in the
+world would be incorporated into
+the current neuronal wiring.
+Indeed,
+This process permits updates of the internal model of the environment
+to best fit the constant influx of sensory examples.
+
+\begin{itemize}
+  \item Both the predictive coding account and POMDPs
+  are process theories backed up by plausible
+  neurobiological evidence that
+  view the brain as a ``statistical organ''
+  generalizing from the past to new experiences.
+  \item Both provide a parsimonious explanation why the
+  human brain decreases processing load devoted to incoming information
+  when the environment is predictable and allocate increased
+  attentional resources when novel stimuli are encountered.
+  Similarly, both propose explanations why
+  predicting the future is inherently
+  linked to information from the past.
+  \item Sensory experience is a generative process from both views.
+  In predictive coding, sensory perception of the external world
+  is a generative experience due to modulatory top down experience at
+  various hierarchical levels of sensory input processing.
+  In our RL view, the (partially observed)
+  environmental model is incorporated in the POMDP,
+  \textit{which can be fully recovered based on the last
+  state alone}.
+  \item Both naturally expose a mechanisms of brain plasticity in that
+  neuronal wiring gets increasingly adapted
+  when faced by unprecedented environmental
+  challenges.
+  \item The hierarchical aspects from predictive coding
+  is re-expressed in POMDPs in form of
+  nested prediction of probable upcoming actions and rewards.
+  \item Both model the consequences of action. In RL, the horizon of that
+  look into the future is explicitly manifested in the $\gamma$ parameter
+  in the Bellman equation, but not
+  explicitly modeled in the PC account that
+  tends to emphasize prediction about the
+  immediate future.
+  \item Mismatch negativity of PC is immanent in reward prediction error
+  by the difference between actually predicted reward of an action given
+  a state and the reward predicted by non-linear regression with graded
+  discounting of future rewards.
+  \item Adapting the neuronal connections for improved top-down modulation
+  takes the concrete form of gradient computation
+  \textit{and back-propagation} in MDPs and RL,
+  although this may not be neurobiologically plausible.
+  \item In sum,
+  the MDP account may serve as a concrete conceptualization of the PC account
+  in cognitive neuroscience. MPDs have the advantage of exposing an explicit
+  mechanism of the horizon of future considerations or
+  how the internal model of the world gets updated,
+  and collapses sensory input processing and action output preparation.
+  %
+  As such, our account can be viewed as a RL implemention of the
+  prediction coding hypothesis.
+\end{itemize}
+
+\subsection{Semantic Hypothesis}
+Another frequently proposed cognitive account of DMN function revolves
+around forming logical associations and analogies between
+the current experience and
+the conceptual knowledge derived from past experiences
+\citep{bar2007proactive, binder1999conceptual}.
+Analogies might naturally tie incoming novel sensory stimuli to
+explicit world knowledge (i.e., semantics) extractable from the environment
+\citep{bar2009proactive}.
+In this way, the encoding of complex environments could be facilitated
+by relative similarity associations between states.
+%
+Going much beyond human language ifself,
+semantic building blocks provide the basis to
+mentally envision non-existent scenarios
+that would improve optimal behavior in the environment
+by simulating future events.
+Such cognitive processes can afford
+the internal generation of necessary information
+that is not presented in the surrounding environment
+by recombining building blocks of
+concept knowledge and episodic memories
+\citep{hassabis2009construction}.
+Indeed, in aging humans, remembering the past and imaging the future
+equally decreased in the level of detail and are associated with
+concurrent deficits in forming and integrating relationships between
+items \citep{addis2008age, spreng2006temporal}.
+Further,
+a constructive account explains the reciprocal relationship
+between an egocentric first person perspective and
+an allocentric bird’s eye perspective immersed in
+self-reflection, social knowledge, and autobiographical memories.
+%
+Cognitive aspects of egocentric-allocentric switching
+are also closely related to episodic memory, language, problem solving,
+planning, estimating other people's thoughts, and spatial navigation
+as these necessitate abstract world knowledge and abstract assocations
+for binding the constituent elements in mental scene construction
+\citep{schacter2007remembering}.
+These scene generation processes could contribute to interpreting the
+present and foretelling the future.
+This view is for instance supported by evidence in animals that
+could learn a \textit{cognitive map of the environment},
+even without reward incentives, and exploit it later
+for other means \citep{tolman1948cognitive}.
+
+\begin{itemize}
+  \item Both the semantic hypothesis and MDPs expose mechanisms of
+  how alternative decision trees could be mentally explored.
+  \item In both semantic hypothesis and MDPs,
+  there is no evidence to indicate that predictions of various
+  levels of complexity, abstraction, timescale and purpose
+  use mechanisms that are qualitatively different. This concurs with
+  DMN activity increases across time, space, and content domains
+  demonstrated in many neuroimaging studies. The semantic hypothesis
+  and RL account provide explanations why hippocampus lesion does
+  not only impair retrieving memories, but also hypothetical and future
+  thinking \citep{hassabis2007patients}.
+  \item The notion of semantic or knowledge association is
+  incorporated into the MDP as the Markov property,
+  that is, the current state directly results from the
+  agent's history of states and actions. The learned
+  value matrix and action transition schedules drive
+  stimulus processing and action choice in the present.
+  \item While mental scene construction implies a distinction between
+  environmental perception and internally generated mind-wandering\footnote{
+  ``A computational model of how such a process might be structured
+  is far from being defined, but it will probably require a form of
+  regulation by which perception of the current world is suppressed
+  while simulation of possible alternatives are constructed,
+  followed by a return to perception of the present.''\citep{buckner2007self}},
+  the MDPs naturally integrate the former egocentric
+  (more related to current action, state and reward) and the later
+  allocentric (more related to past and future actions, states, and rewards)
+  angles on the world in a same optimization problem.
+  \item The semantic account of DMN function does not offer
+  a mechanistic explanation how explicit world knowledge and semantic analogies thereof
+  lead to prediction of future actions and states.
+  \item In contrast to existing accounts on semantics and
+  mental scene construction, the random and creative aspects of DMN function
+  are explained in MDPs by the advantages of stochastic optimization.
+  \item The semantic hypothesis does not explain why memory recall
+  for scene construction in humans is typically fragmentary and noisy
+  instead of accurate and reliable.
+  Yet, the MDP framework provides an algorithmic explanation in that
+  stochasticity of the parameter space search implemented
+  by Monte Carlo solvers provably yields better models of the world.
+  That is, the purposeful stochasticity of policy and value estimation
+  in MDPs provides a candidate explanation for why humans
+  have evolved imperfect noisy memories
+  as the more advantageous adaptation.
+  \item While both semantic and MDP account propose memory-based internally
+  generated information for probabilistic mental models of action outcomes,
+  only MDPs render explicit the grounds on which the final action is
+  eventually chosen (i.e., the estimated cumulative reward).
+  \item In sum, episodic scene construction according to the semantic
+  hypothesis is lacking an explicit time and incentive model;
+  neither does it explain how randomness of human mental experience
+  can be beneficial.
+\end{itemize}
+
+\subsection{Sentinel Hypothesis}
+Processing self-relevant information was perhaps the first
+cognitive account that was proposed for the DMN \citep{gusnard2001medial}.
+Since then,
+different investigators have speculated that neural activity in the DMN
+may reflect the brain’s relentless tracking of
+relevance in the environment
+as an advantageous evolutionary adaptation.
+According to this cognitive account, the brain's baseline realizes
+a ``radar'' function to
+detect subjectively salient and unexpected events that
+are the most important for unfolding behavior,
+including resources,
+the securing of mates, and protection.
+%
+In fact,
+environmental stimuli important for humans are frequently of
+social nature. This is probably unsurprising
+given that a key property of the human species is
+the complexity of their social systems
+\citep{tomasello2009cultural}.
+More specifically,
+according to the ``social brain hypothesis'',
+the human brain did not evolve to solve problems of the
+physical environment, but to form and maintain increasingly complex
+social systems to solve ecological problems by means of social relationships
+\citep{whiten1988machiavellian}.
+Highly efficient neurobiological processing of social cues is exemplified by
+facial judgments being routinely processed in less than 100 ms
+\citep{bar2006very}.
+Our human ancestors were probably among the few organisms that
+were more likely to be killed by members of their own species
+rather than predators from another species.
+Being able to detect the intention and upcoming actions of conspecifics
+turned into a decisive evolutionary advantage
+\citep{frith2010social}.
+%
+This may explain why social topics amount to roughly
+two thirds of human everyday communication \citep{dunbar1997human}
+and
+why mind-wandering and dreams
+are so rich in stories about people and
+the complex relationships between them
+\citep{schilbach2008minds}.
+
+\textbf{XXX Your focus in this subsection ought to be on the SC itself not POMPDs, etc.}
+ \begin{itemize}
+    \item Processing social information has been proposed to underlie the
+    physiological baseline of human brain function
+    \citep{schilbach2008minds}. This was later challenged by observing
+    analogues of the DMN in monkeys \citep{mantini2011default}
+    and rats \citep{lu2012rat}, with
+    supposedly less advanced social capacities.
+    The MDPs parsimoneously explain the dominance of social content in
+    human mental scenes by their extremely high relevance to humans
+    reflected by high values for social information in the value matrix
+    \citep{baker2009action}.
+    \item How can a same neurobiological circuit be equally important
+    for baseline house-keeping functions and specific task performance?
+    Our computational account can explain why the DMN is implicated
+    in both a goal-directed task and an idlying rest cognitive set,
+    if environmental relevance is processed, manipulated, and retained
+    also as a baseline function of human mental activity.
+    In tasks, the policy and value matrices are updated to optimize short-term action,
+    whereas, at rest, these parameter updates may
+    improve especially mid- and long-term action.
+    \item The sentinel account does not provide a formal mechanisms of
+    how attention and memory resources are exactly reallocated when
+    encountering a salient environmental stimulus. \textbf{XXX remove the next sentence. It's probably false.}
+    In contrast,
+    the Bellman recursion is weighted at each step by
+    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
+    are recursively explored by the agent and which decision trees are neglected.
+    This directly implies allocation of attention and computational load.
+    \item In sum,
+    MDPs imply a ``radar'' function of monitoring the environment for salient information
+    in general and explain social thinking as a physiological brain baseline
+    as an important special case, without excluding other topics
+    contemplated in continuous mental activity.
+ \end{itemize}
+
+\section{Conclusion}
+What brain function could be important enough
+for the existence and survival of the human species
+to warrant constant, high energy costs?
+MDPs provide an attractive
+computational model how the human association cortex
+might implement supra-modal representation and control of the environment to
+optimize the organism's fate.
+% by obtaining reward and avoiding pain.
+This idealized process model explains
+a number of previous experimental observations in the
+DMN by a simple but non-trivial mechanism.
+%
+From the computational viewed of a MDP,
+behavior unfolds by integrating happened past events
+and possible future events to guide action choice in the present context.
+This functional account is more compatible with the DMN's
+poorly understood involvement across
+autobiographical memory retrieval, problem solving,
+abstract reasoning based on internally generated scenes, social cognition,
+as well as delay discounting and self-related prospection.
+MDPs provide a mathematical formalism how
+optimal substructure in the environment can be recursively exploited
+when confronted with complicated decisions.
+Algorithmic gains by injecting stochasticity in the recall of past
+actions and outcomes to improve the internal world model may explain why
+very accurate memories have been disfavored in human evolution
+and why human creativity might be adaptive.
+%
+In principle,
+neuroscientific experiments can be designed that operationalize
+the set of action, value, and state variables that determine
+the behavior of intelligent RL systems.
+The proposed machine-learning perspective
+on DMN function is hence not only practically computable but
+also neuroscientifically falsifiable.
+At the least, we propose an alternative vocabulary to
+describe and interpret neuroscientific studies in DMN research.
+%
+Ultimately,
+the DMN can be viewed as a smoothing kernel extending
+from the relatively recent events to anticipated upcoming events
+to constantly improve sensory processing and
+action performance in the world.
+
+\paragraph{Acknowledgment}
+% {\small The research leading to these results has received funding from the
+% European Union Seventh Framework Programme (FP7/2007-2013)
+% under grant agreement no. 604102 (Human Brain Project).
+% Further support was received from
+% the German National Academic Foundation (D.B.).
+% }
+
+\small
+% \bibliographystyle{splncs03}
+% \bibliographystyle{natbib}
+\bibliographystyle{plainnat}
+\bibliography{refs}
+
+\appendix
 \section{Free-energy principles!}
@@
 
-\subsection{Minimizing free-energy via backprop!}
+\subsection{Minimizing free-energy via backprop: variational auto-encoders}
 
@@
 \label{sec:vae}
-Here, we present a way to aleviate some conceptual and computational issues with the
-free-energy framework presented thus far, by using auto-encoder theory.
-Following Kingma et al. 2014 \citep{kingma2013auto}, define the data-dependent auxiliary random
-function
+Here, we present a way to alleviate some conceptual and computational issues with the
+free-energy framework presented thus far, by using the recent \textit{variational auto-encoder} (VAE) theory \citep{kingma2013auto}. Define the data-dependent
+auxiliary random function
 \begin{eqnarray}
@@
 \end{mdframed}
+
 \begin{figure}[!tb]
@@
 
-\section{Relation to other statistical models (Elvis)}
-\subsection{Friston's Free-energy principle (refer to \ref{label:friston})}
-\textbf{XXX: add tons of more examples and connnections}
-\textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
-\textbf{XXX: integrate aspects from WakeNsleep by Hinton}
-\textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
-\textbf{XXX: integrate aspects from Sutton/Barto book chapter}
-In summary:
-\begin{itemize}
-  \item Friston's free-energy principle: brain as inference engine
-  biological systems, including brains, must minimize the long-term average of surprise;
-  \item A common point between the Free Energy principle and our framework
-  is placing the minimization of suprise at the core of brain function.
-  In RL, surprise minimization reduces to accurate prediction of
-  rewarding outcomes in the future.
-  \item Ortega et al. 2013 \citep{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
-\item Friston's critic of RL \citep{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
-pioneered in the 1950s by Richard Bellman and colleagues [2]. To
-optimise control a~pð Þ x~ under this formulation, we have to: (i)
-assume the hidden states are available to the agent and (ii) solve
-Equation 18 for the value-function. Solving for the value-function
-is a non-trivial problem and usually involves backwards induction
-or some approximation scheme like reinforcement learning [4–6].
-The free-energy formulation circumvents these problems by
-prescribing the policy in terms of free-energy, which encodes
-optimal control (Equation 6)"
-\item It is difficult to compute with the free energy principle
-  in practice, which makes it physiologically less plausible. Auto-encoder theory presented
-  \ref{sec:vae} can help to remove these barriers, by making the whole scheme trainable via
-  backprop!
-  \end{itemize}
-
-\subsection{Other Leinforcement Learning literature}
-abc
-
-\section{Relation to other cognitive models}
-\subsection{Predictive Coding Hypothesis}
-The predictive coding framework
-\citep{clark2013whatever, friston2008hierarchical}
-is a frequently evoked idea related to the default mode function
-\citep{bar2007}.
-According to this cognitive account,
-cortical responses emerge from continuous functional interaction between
-higher and lower levels of the neural processing hierarchy.
-This dynamic intercourse enables inference on the world by reconciling
-gaps between fresh sensory input and stored prior information.
-Feed-forward sensory experience is constantly calibrated by
-top-down modulation at various hierarchical processing levels.
-In fact, axonal
-back projections probably outnumber by far the input projection
-existing in the brain \citep{salin1995corticocortical}.
-This can explain the constructive, generative nature of sensory perception
-\citep{friston2010free} and
-why motor action is intimately related to sensory expectations
-\citep{wolpert1995internal, kording2004bayesian}.
-At each stage of neural processing,
-an internally generated prediction of the sensory input is
-directly compared against the actual input from the environment.
-A prediction error at one of the processing levels,
-i.e. the difference between sensory input
-and internally predicted sensation,
-incurs plasticity changes of neuronal back projections (i.e., model parameters)
-for gradually improved future prediction of the environment.
-Contextual integration is hence maintained by top-down modulation of sensorimotor
-processing by a-priori information about the environment.
-The generative model of how perceived sensory events arise in the
-world would be incorporated into
-the current neuronal wiring.
-Indeed,
-This process permits updates of the internal model of the environment
-to best fit the constant influx of sensory examples.
-
-\begin{itemize}
-  \item Both the predictive coding account and MDPs
-  are process theories backed up by plausible
-  neurobiological evidence that
-  view the brain as a ``statistical organ''
-  generalizing from the past to new experiences.
-  \item Both provide a parsimonious explanation why the
-  human brain decreases processing load devoted to incoming information
-  when the environment is predictable and allocate increased
-  attentional resources when novel stimuli are encountered.
-  Similarly, both propose explanations why
-  predicting the future is inherently
-  linked to information from the past.
-  \item Sensory experience is a generative process from both views.
-  In predictive coding, sensory perception of the external world
-  is a generative experience due to modulatory top down experience at
-  various hierarchical levels of sensory input processing.
-  In our RL view, the (partially observed)
-  environmental model is incorporated in the MDP,
-  \textit{which can be fully recovered based on the last
-  state alone}.
-  \item Both naturally expose a mechanisms of brain plasticity in that
-  neuronal wiring gets increasingly adapted
-  when faced by unprecedented environmental
-  challenges.
-  \item The hierarchical aspects from predictive coding
-  is re-expressed in MDPs in form of
-  nested prediction of probable upcoming actions and rewards.
-  \item Both model the consequences of action. In RL, the horizon of that
-  look into the future is explicitly manifested in the $\gamma$ parameter
-  in the Bellman equation, but not
-  explicitly modeled in the PC account that
-  tends to emphasize prediction about the
-  immediate future.
-  \item Mismatch negativity of PC is immanent in reward prediction error
-  by the difference between actually predicted reward of an action given
-  a state and the reward predicted by non-linear regression with graded
-  discounting of future rewards.
-  \item Adapting the neuronal connections for improved top-down modulation
-  takes the concrete form of gradient computation
-  \textit{and back-propagation} in MDPs and RL,
-  although this may not be neurobiologically plausible.
-  \item In sum,
-  the MDP account may serve as a concrete conceptualization of the PC account
-  in cognitive neuroscience. MPDs have the advantage of exposing an explicit
-  mechanism of the horizon of future considerations or
-  how the internal model of the world gets updated,
-  and collapses sensory input processing and action output preparation.
-  %
-  As such, our account can be viewed as a RL implemention of the
-  prediction coding hypothesis.
-\end{itemize}
-
-\subsection{Semantic Hypothesis}
-Another frequently proposed cognitive account of DMN function revolves
-around forming logical associations and analogies between
-the current experience and
-the conceptual knowledge derived from past experiences
-\citep{bar2007proactive, binder1999conceptual}.
-Analogies might naturally tie incoming novel sensory stimuli to
-explicit world knowledge (i.e., semantics) extractable from the environment
-\citep{bar2009proactive}.
-In this way, the encoding of complex environments could be facilitated
-by relative similarity associations between states.
-%
-Going much beyond human language ifself,
-semantic building blocks provide the basis to
-mentally envision non-existent scenarios
-that would improve optimal behavior in the environment
-by simulating future events.
-Such cognitive processes can afford
-the internal generation of necessary information
-that is not presented in the surrounding environment
-by recombining building blocks of
-concept knowledge and episodic memories
-\citep{hassabis2009construction}.
-Indeed, in aging humans, remembering the past and imaging the future
-equally decreased in the level of detail and are associated with
-concurrent deficits in forming and integrating relationships between
-items \citep{addis2008age, spreng2006temporal}.
-Further,
-a constructive account explains the reciprocal relationship
-between an egocentric first person perspective and
-an allocentric bird’s eye perspective immersed in
-self-reflection, social knowledge, and autobiographical memories.
-%
-Cognitive aspects of egocentric-allocentric switching
-are also closely related to episodic memory, language, problem solving,
-planning, estimating other people's thoughts, and spatial navigation
-as these necessitate abstract world knowledge and abstract assocations
-for binding the constituent elements in mental scene construction
-\citep{schacter2007remembering}.
-These scene generation processes could contribute to interpreting the
-present and foretelling the future.
-This view is for instance supported by evidence in animals that
-could learn a \textit{cognitive map of the environment},
-even without reward incentives, and exploit it later
-for other means \citep{tolman1948cognitive}.
-
-\begin{itemize}
-  \item Both the semantic hypothesis and MDPs expose mechanisms of
-  how alternative decision trees could be mentally explored.
-  \item In both semantic hypothesis and MDPs,
-  there is no evidence to indicate that predictions of various
-  levels of complexity, abstraction, timescale and purpose
-  use mechanisms that are qualitatively different. This concurs with
-  DMN activity increases across time, space, and content domains
-  demonstrated in many neuroimaging studies. The semantic hypothesis
-  and RL account provide explanations why hippocampus lesion does
-  not only impair retrieving memories, but also hypothetical and future
-  thinking \citep{hassabis2007patients}.
-  \item The notion of semantic or knowledge association is
-  incorporated into the MDP as the Markov property,
-  that is, the current state directly results from the
-  agent's history of states and actions. The learned
-  value matrix and action transition schedules drive
-  stimulus processing and action choice in the present.
-  \item While mental scene construction implies a distinction between
-  environmental perception and internally generated mind-wandering\footnote{
-  ``A computational model of how such a process might be structured
-  is far from being defined, but it will probably require a form of
-  regulation by which perception of the current world is suppressed
-  while simulation of possible alternatives are constructed,
-  followed by a return to perception of the present.''\citep{buckner2007self}},
-  the MDPs naturally integrate the former egocentric
-  (more related to current action, state and reward) and the later
-  allocentric (more related to past and future actions, states, and rewards)
-  angles on the world in a same optimization problem.
-  \item The semantic account of DMN function does not offer
-  a mechanistic explanation how explicit world knowledge and semnatic analogies thereof
-  lead to prediction of future actions and states.
-  \item In contrast to existing accounts on semantics and
-  mental scene construction, the random and creative aspects of DMN function
-  are explained in MDPs by the advantages of stochastic optimization.
-  \item The semantic hypothesis does not explain why memory recall
-  for scene construction in humans is typically fragmentary and noisy
-  instead of accurate and reliable.
-  Yet, the MDP framework provides an algorithmic explanation in that
-  stochasticity of the parameter space search implemented
-  by Monte Carlo solvers provably yields better models of the world.
-  That is, the purposeful stochasticity of policy and value estimation
-  in MDPs provides a candidate explanation for why humans
-  have evolved imperfect noisy memories
-  as the more advantageous adaptation.
-  \item While both semantic and MDP account propose memory-based internally
-  generated information for probabilistic mental models of action outcomes,
-  only MDPs render explicit the grounds on which the final action is
-  eventually chosen (i.e., the estimated cumulative reward).
-  \item In sum, episodic scene construction according to the semantic
-  hypothesis is lacking an explicit time and incentive model;
-  neither does it explain how randomness of human mental experience
-  can be beneficial.
-\end{itemize}
-
-\subsection{Sentinel Hypothesis}
-Processing self-relevant information was perhaps the first
-cognitive account that was proposed for the DMN \citep{gusnard2001medial}.
-Since then,
-different investigators have speculated that neural activity in the DMN
-may reflect the brain’s relentless tracking of
-relevance in the environment
-as an advantageous evolutionary adaptation.
-According to this cognitive account, the brain's baseline realizes
-a ``radar'' function to
-detect subjectively salient and unexpected events that
-are the most important for unfolding behavior,
-including resources,
-the securing of mates, and protection.
-%
-In fact,
-environmental stimuli important for humans are frequently of
-social nature. This is probably unsurprising
-given that a key property of the human species is
-the complexity of their social systems
-\citep{tomasello2009cultural}.
-More specifically,
-according to the ``social brain hypothesis'',
-the human brain did not evolve to solve problems of the
-physical environment, but to form and maintain increasingly complex
-social systems to solve ecological problems by means of social relationships
-\citep{whiten1988machiavellian}.
-Highly efficient neurobiological processing of social cues is exemplified by
-facial judgments being routinely processed in less than 100 ms
-\citep{bar2006very}.
-Our human ancestors were probably among the few organisms that
-were more likely to be killed by members of their own species
-rather than predators from another species.
-Being able to detect the intention and upcoming actions of conspecifics
-turned into a decisive evolutionary advantage
-\citep{frith2010social}.
-%
-This may explain why social topics amount to roughly
-two thirds of human everyday communication \citep{dunbar1997human}
-and
-why mind-wandering and dreams
-are so rich in stories about people and
-the complex relationships between them
-\citep{schilbach2008minds}.
-
- \begin{itemize}
-    \item Processing social information has been proposed to underlie the
-    physiological baseline of human brain function
-    \citep{schilbach2008minds}. This was later challenged by observing
-    analogues of the DMN in monkeys \citep{mantini2011default}
-    and rats \citep{lu2012rat}, with
-    supposedly less advanced social capacities.
-    The MDPs parsimoneously explain the dominance of social content in
-    human mental scenes by their extremely high relevance to humans
-    reflected by high values for social information in the value matrix
-    \citep{baker2009action}.
-    \item How can a same neurobiological circuit be equally important
-    for baseline house-keeping functions and specific task performance?
-    Our computational account can explain why the DMN is implicated
-    in both a goal-directed task and an idlying rest cognitive set,
-    if environmental relevance is processed, manipulated, and retained
-    also as a baseline function of human mental activity.
-    In tasks, the policy and value matrices are updated to optimize short-term action,
-    whereas, at rest, these parameter updates may
-    improve especially mid- and long-term action.
-    \item The sentinel account does not provide a formal mechanisms of
-    how attention and memory resources are exactly reallocated when
-    encountering a salient environmental stimulus. In contrast,
-    the Bellman recursion is weighted at each step by
-    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
-    are recursively explored by the agent and which decision trees are neglected.
-    This directly implies allocation of attention and computational load.
-    \item In sum,
-    MDPs imply a ``radar'' function of monitoring the environment for salient information
-    in general and explain social thinking as a physiological brain baseline
-    as an important special case, without excluding other topics
-    contemplated in continuous mental activity.
- \end{itemize}
-
-\section{Conclusion}
-What brain function could be important enough
-for the existence and survival of the human species
-to warrant constant, high energy costs?
-MDPs provide an attractive
-computational model how the human association cortex
-might implement supra-modal representation and control of the environment to
-optimize the organism's fate.
-% by obtaining reward and avoiding pain.
-This idealized process model explains
-a number of previous experimental observations in the
-DMN by a simple but non-trivial mechanism.
-%
-From the computational viewed of a MDP,
-behavior unfolds by integrating happened past events
-and possible future events to guide action choice in the present context.
-This functional account is more compatible with the DMN's
-poorly understood involvement across
-autobiographical memory retrieval, problem solving,
-abstract reasoning based on internally generated scenes, social cognition,
-as well as delay discounting and self-related prospection.
-MDPs provide a mathematical formalism how
-optimal substructure in the environment can be recursively exploited
-when confronted with complicated decisions.
-Algorithmic gains by injecting stochasticity in the recall of past
-actions and outcomes to improve the internal world model may explain why
-very accurate memories have been disfavored in human evolution
-and why human creativity might be adaptive.
-%
-In principle,
-neuroscientific experiments can be designed that operationalize
-the set of action, value, and state variables that determine
-the behavior of intelligent RL systems.
-The proposed machine-learning perspective
-on DMN function is hence not only practically computable but
-also neuroscientifically falsifiable.
-At the least, we propose an alternative vocabulary to
-describe and interpret neuroscientific studies in DMN research.
-%
-Ultimately,
-the DMN can be viewed as a smoothing kernel extending
-from the relatively recent events to anticipated upcoming events
-to constantly improve sensory processing and
-action performance in the world.
-
-\paragraph{Acknowledgment}
-% {\small The research leading to these results has received funding from the
-% European Union Seventh Framework Programme (FP7/2007-2013)
-% under grant agreement no. 604102 (Human Brain Project).
-% Further support was received from
-% the German National Academic Foundation (D.B.).
-% }
-
-\small
-% \bibliographystyle{splncs03}
-% \bibliographystyle{natbib}
-\bibliographystyle{plainnat}
-\bibliography{refs}
-
 \end{document}

### 2017-03-28 paper/paper.tex
comment in Elvis land

--- 2017-03-28 paper/paper.tex
@@
 Neuroscientific investigations on hemispheric functional specialization
-have converged to the right versus left cerebral hemisphere as dominant for
+have highlighted the right versus left cerebral hemisphere as dominant for
 attention versus language functions.
@@
 sensory world and internally-oriented mind sets focussed
-on self-relevant mind-wandering of possible mental scenarios.
+on self-relevant mind wandering through mental scenarios.
 Transient RTPJ in humans for instance disrupted the
@@
 The RTPJ might hence be an important player that shifts away
-from the default active ‘‘internally directed’’ processes
+from the ‘‘internally directed’’ baseline processes
 to deal instead with immediate environmental objects and contexts.
@@
 about the physical and inter-personal world.
-Multivariate volumetric analyses showed that the internal production of
-formulated thought ("language of the mind") is closely related to the left TPJ
+Indeed,
+the internal production of
+formulated thought ("language of the mind") was closely related to the left TPJ
+in a multivariate analysis of brain volume
 \citep{geva2011neural}.
-A number of neuroimaging studies demonstrated LTPJ activity responses
-to sementically aberrant words in sentences, which suggests a role
-in binding a given semantic concept into an integrated whole.
 Further,
@@
 components integrate in a plausible way.
-Reinforcement Learning (RL) is learning to achieve a goal by interacting with an external
-\textit{environment / world}. In a general RL framework, an \textit{agent}
+Reinforcement learning is learning to achieve a goal by interacting with an external
+\textit{environment}. In a general RL framework, an \textit{agent}
 interacts with an  environment in a trial-by-error fashion by taking
-actions / decisions $a$, which trigger changes in the
+actions $a$, which trigger changes in the
 \textit{state} of the environment
@@
 It is our view that at a sufficiently coarse scale, the brain is a physical system governed by the laws of
+\textbf{XXX the paper is about the DMN more specifically, perhaps try to sharpen}
 physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
 any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
-step is sufficiently short (defining $s(t)$ --the state at time $t$-- as the position in phase space).
+step is sufficiently short, defining $s(t)$, the state at time $t$, as the position in phase space.
 If the process has $memory$ such that the next state depends not only on the current state but also on some finite
@@
 These can be assumptions about the physical laws governing the agent's world or constraints on the state-space and transitions thereupon.
-For example if a rat in a maze knows, for sure,
-that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food.
+For example, if a rat in a maze knows
+that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
+\textbf{XXX last sentence not statistically sound}
+\textbf{XXX this paper is mostly about the human DMN, therefore it would be better to rephrase the example into something with a human being}
 We formalize
@@
 
-\subsection{Value functions and policies}
+\subsection{Value mapping and action policies}
 The behavior of the agent is governed by some kind of \textit{policy},
@@
 \begin{mdframed}
-  \textbf{Where is ``value'' buffered in the brain ?}
+  \textbf{Where is ``value'' buffered in the brain ? You mean where in the DMN!
+  then the answer is in the vmPFC which is connected to the NAc}
   The amygdala is known to be involved in significance evaluation) and
@@
 immediate rewards and has no horizon for the future, which is not compatible with long-term goal planning
-as implemented in the DMN. For there to be any learning at all, it is thus necessary that $0 < \gamma < 1$.
+as potentially implemented in the DMN. For there to be any learning at all, it is thus necessary that $0 < \gamma < 1$.
 % \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
-Thus the agent is positively biased towards considering rewards which are will
-come soon. The reward importance hence decays exponentially
+Thus, the agent is positively biased towards considering rewards which with
+a short time delay. The reward importance hence decays exponentially
 with distance in the future,
@@
 The reward discount factor $\gamma$ can thus be seen as calibrating
-the delay discounting behavior of the intelligent agent.
+the delay discounting behavior of the intelligent agent,
+that is the decision related to reward-delay tradeoff schedules.
 It is also important to appreciate that a stochastic policy estimation
@@
 be easily exploited and a uniformly random choice is optimal.
+\textbf{XXX I love this example :-)}
 
@@
 \begin{mdframed}
-In plain English, the Q-value $Q^{\pi}(s,a)$ is the expected / average / mean future reward
-over all possible trajectories in which the agent starts off starts at with an environment in state
+In plain English, the Q-value $Q^{\pi}(s,a)$ is the expected future reward
+\textbf{XXX let's stick to "expected" in these situations because
+it implies that the DMN is doing prediction without needing to 
+emphasize it too much all the time}
+over all possible
+\textbf{XXX action???}
+trajectories in which the agent starts off starts at with an environment in state
 $s$, takes and action $a$, and then follows the policy $\pi$ thereafter to select future actions.
@@
 By construction, maximizing the expectation in  \eqref{eq:qval} is intractable in general
-due to the enormous number of possible trajectories the agent might need to
+\textbf{XXX coputationally or statistically?}
+due to the enormous number of possible
+\textbf{XXX action???}
+trajectories the agent might need to
 consider.
@@
 % \end{mdframed}
-One popular simplification that removes this issue is to only consider the sub-family of
+One popular simplification that removes
+%this issue
+\textbf{excess in complexity on the agent's horizon???}
+is to only consider the sub-family of
 \textit{deterministic policies}, which map each state to the best single best action which
@@
 % table.
-The celebrated \textit{Bellman's equation} \cite{sutton1998reinforcement} then takes the simple form
+The \textit{Bellman's equation} \cite{sutton1998reinforcement} then takes the simple form
 \begin{equation}
@@
 \item Friston's critic of RL \citep{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
-pioneered in the 1950s by Richard Bellman and colleagues [2]. To
+pioneered in the 1950s by Richard Bellman and colleagues. To
 optimise control a~pð Þ x~ under this formulation, we have to: (i)
@@
 is a non-trivial problem and usually involves backwards induction
-or some approximation scheme like reinforcement learning [4–6].
+or some approximation scheme like reinforcement learning.
 The free-energy formulation circumvents these problems by

### 2017-03-28 paper/paper.tex
fixup
fixups
model-free RL is plausible in the brain via dopamine

@@
 and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.
+\begin{mdframed}
+\textbf{Model-free RL in the brain is plausible.}  Indeed, as was noted in \cite{dayan2008decision}, It has long been suggested in the literature that there
+  is a rather direct
+mapping of model-free RL learning algorithms
+onto the brain, with the neuromodulator dopamine serving
+as a teaching signal to train values or policies by controlling
+synaptic plasticity at targets such as the ventral and
+dorsolateral striatum.
+\end{mdframed}
+
 In contrast, \textit{model-based RL} starts off with some mechanistic assumptions about the dynamics of world.
@@
 
-\subsection{Conclusion: putting everything together}
+\subsection{Putting everything together}
 \begin{figure}[!h]
@@
 \textit{variational auto-encoders (VAEs)} \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm, as it overcomes the technical limits of the former, by using a reparametrization trick.
-For example, unlike the wake-sleep algorithm for minimizing free-energy, VAEs can be efficiently trained back-propagation of learning errors.
+For example, unlike the wake-sleep algorithm for minimizing free-energy, VAEs can be efficiently trained via back-propagation of learning errors.
 

### 2017-03-28 paper/paper.tex
more

--- 2017-03-28 paper/paper.tex
@@
 and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.
+\begin{mdframed}
+\textbf{Model-free RL in the brain is plausible.}  Indeed, as was noted in \cite{dayan2008decision}, it has long been suggested in the literature that there
+  is a rather direct mapping of model-free RL learning algorithms
+onto the brain, with the neuromodulator dopamine serving
+as a teaching signal to train values or policies by controlling
+synaptic plasticity at targets such as the ventral and
+dorsolateral striatum.
+\end{mdframed}
+
 In contrast, \textit{model-based RL} starts off with some mechanistic assumptions about the dynamics of world.
@@
 \subsection{Computing optimal behavior}
-By construction, maximizing the expectation in  \eqref{eq:qval} is intractable in general
-\textbf{XXX coputationally or statistically?}
-due to the enormous number of possible
-\textbf{XXX action???}
+By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general
+due to the enormous number of possible action
 trajectories the agent might need to
@@
 One popular simplification that removes
-%this issue
-\textbf{excess in complexity on the agent's horizon???}
-is to only consider the sub-family of
+this issue is to only consider the sub-family of
 \textit{deterministic policies}, which map each state to the best single best action which
@@
 which can be prohibitively large.
-Monte-Carlo simulation provides a simple method for evaluating a state $s$.
+Monte-Carlo simulation provides a simple method for evaluating the value / fitness of a state.
 They provide an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
@@
 
-In our proposed RL-based view, \textit{inference} in the human brain reduces to generalization of
+  In our proposed model-free RL-based view, \textit{inference} in the human brain reduces to
+  generalization of
   policy and value adaptations from sampled experiences to
@@
   to novel, yet unexperienced environments, reward prediction errors will
-  massively increase. This will lead to adaptation of the policy matrix
-  and value function until the system converges to a
+  massively increase. This will lead to adaptation of the policy until the system converges to a 
   new steady-state of optimal action decisions given certain states.

### 2017-03-29 paper/paper.tex

@@
   Indeed in RL, surprise minimization is subsumed  by accurate prediction of
-  rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist internal phase-transitions which lead contradictory with the
-  blurred version of the external world captured as sensations
+  rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist undesired / harmful phase-transitions.
   \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}. Such a theory cannot by itself,

### 2017-03-29 paper/paper.tex
superpass

@@
 
-\author{Elvis Dohmatob, Bert Kappen, Danilo Bzdok\\
+\author{Elvis Dohmatob, Guillaume Dumas, Danilo Bzdok\\
   %% INRIA, Parietal team, Saclay, France\\
@@
 \begin{abstract}
-The default mode network (DMN) is the brain manifestations of the human
+The default mode network (DMN) is the brain manifestation of the human
 baseline mental activity.
@@
 \textbf{\\keywords}: systems biology, mind wandering, cognitive science,
-artificial intelligence (AI),
+artificial intelligence,
 reinforcement learning
@@
 unfocused everyday mind wandering \citep{raichle2001pnas}.
+Indeed, the human brain features "a hierarchy of brain systems with
+the DMN at the top and the salience and dorsal attention systems
+at intermediate levels, above thalamic and unimodal sensory
+cortex" \citep{carhart2010default}.
 In the beginning of the 21st century, brain imaging was hence the first technology
@@
 \subsection{Sketch of our main contributions}
-Our theory proposes that the implication of the DMN in a diversity of human cognitive processes
+Our theory proposes that the implication of the DMN in a
+diversity of humans' most advanced cognitive processes
 can be parsimoniously recast as prediction error minimization
@@
 
-\section{Reinforcement Learning, a computational model for DMN function}
-We now argue that the DMN is equipped with sufficient neurobiological substrates gradient
-for implementing all the major components (memory, reward signaling, accumulators, random sampling, etc.)
-of an efficient \textit{reinforcement learning (RL)} algorithm, and all these
-components integrate in a plausible way.
+\section{Reinforcement learning: A process model for DMN function}
+The reviewed neurobiological properties and likely functions of the DMN
+are now argued to be sufficient
+for implementing all the major components 
+of a full-fledged \textit{reinforcement learning (RL)} algorithm.
+Memory, reward signaling,
+information gathering, random sampling, and other components
+of intelligent RL agents appear to cross-talk in the DMN in a plausible way.
+
 Reinforcement learning is learning to achieve a goal by interacting with an external
-\textit{environment}. In a general RL framework, an \textit{agent}
+\textit{environment}.
+\textbf{XXX Don't define something with the same thing}
+In a general RL framework, an \textit{agent}
 interacts with an  environment in a trial-by-error fashion by taking
-actions $a$, which trigger changes in the
+\textit{actions} $a$, which trigger changes in the
 \textit{state} of the environment
@@
 $r = r(s, a)$ (or regret!, since it can be negative) collected by the
-agent. In this view, the environmental is partially controlled by
+agent.
+In this view, the environment is partially controlled by
 the action of the agent and the reward should be thought
-of as an instantaneous satisfaction accompanying the execution of
+of as an instantaneous
+\textbf{XXX instant is problematic here, because it may change behavior only later in RL}
+satisfaction
+\textbf{XXX or punishment?}
+accompanying the execution of
 an action, and may be delayed, only becoming substantial when accumulated over an extended period of time.
@@
 The environment is generally taken as stochastic,
+\textbf{XXX One sentence, one bullet. Please explain stochast and PO
+in one sentence each.}
 is \textit{partially observed},
 in the sense that only part of the current state is observed by
-the agent. This unpredictability
-% (in a first-order sense)
+the agent.
+This unpredictability
+\textbf{XXX Explain this sentence better because it is important.}
 is realistic in a model which sets out
 to explain DMN function. Indeed, it is known that, the more the external world is predictable,
-the more mental activity becomes detached from the actual sensorium
-\citep{antrobus1966studies, pope1978regulation}.
+the more mental activity becomes detached from the actual sensory inputs
+\citep{antrobus1966studies, pope1978regulation},
+the more stimulus-independent thoughts occur, and
+the higher the neural activity in the DMN.
 Conversely, the more the currently executed task is unknown and unpracticed,
-the lower the DMN activity, i.e the tendency for stimulus-independent thoughts
+the lower the DMN activity
 \citep{filler1973daydreaming, teasdale1995stimulus}.
 %
-These ``offline'' processes may however contribute to optimized control of the organism.
+These ``offline'' processes may however contribute to
+optimizing control of the organism.
 A DMN framework incorporating reinforcement learning
@@
 The implication of the DMN in a diversity of human cognitive processes
-can be parsimoneously recast as prediction error minimization
-based on pervasive probabilistic simulations,
-thus maximizing reward outcome at various time scales.
+can be parsimoneously explained as prediction error minimization
+coupled with probabilistic simulations in form of mental scenes,
+thus calibrating action portfolios and
+maximizing reward outcome at various time scales.
 Such a purposeful optimization objective
 may be solved by a stochastic approximation
-based on a brain implementation of Markov Chain Monte Carlo sampling
+based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
 \citep{tenenbaum2011grow}.
@@
 number of past states,
-rational probabilistic planning can reformulated
+rational probabilistic planning can be reformulated
 as a standard memoryless Markov process by simply expanding the
@@
 but the agent cannot directly observe the underlying state. Instead, it tries to optimize a \textit{subjective} reward
-signal which is tried alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
+signal. This is performed alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
 states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
 and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.
-\begin{mdframed}
-\textbf{Model-free RL in the brain is plausible.}  Indeed, as was noted in \cite{dayan2008decision}, it has long been suggested in the literature that there
+Such model-free RL is plausible in the human brain. Indeed, as was noted in \cite{dayan2008decision}, it has long been suggested in the literature that there
   is a rather direct mapping of model-free RL learning algorithms
-onto the brain, with the neuromodulator dopamine serving
+onto the brain,
+\textbf{XXX New sentence}
+with the neuromodulator dopamine serving
 as a teaching signal to train values or policies by controlling
@@
 dorsolateral striatum.
-\end{mdframed}
-
-In contrast, \textit{model-based RL} starts off with some mechanistic assumptions about the dynamics of world.
-These can be assumptions about the physical laws governing the agent's world or constraints on the state-space and transitions thereupon.
-For example if a rat in a maze knows, for sure,
-that staying still will produce no change in the environment, and in particular will not eventually lead to finding the food.
+\textbf{XXX Loose all boxes by integrating the text please since it is not a usual layouting}
+
+In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of world.
+These can be assumptions about the physical laws governing the agent's world or constraints on the state space and transitions between states.
+For instance, if a rat in a maze knows
+that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
 % A core property of human intelligence has been proposed underly to improve
@@
 % environments \citep{gershman2015computational}.
+\textbf{XXX transition back to model-free RL is not clear. please make explicit}
 Our rat might represent such knowledge about the world as:
@@
 \begin{itemize}
-\item $\mathcal S$ is the set of states, e.g $\mathcal S = \{\text{happy}, \text{sad}\}$.
+\item $\mathcal S$ is the set of states, such as $\mathcal S = \{\text{happy}, \text{sad}\}$.
 \item $p : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto p(s'|s,a)$,
@@
   transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present.
-\item $\mathcal A$ is the set of actions, e.g $\mathcal A = \{\text{read}, \text{run},
+\item $\mathcal A$ is the set of actions, such as $\mathcal A = \{\text{read}, \text{run},
   \text{laugh}, \text{sympathize}, \text{empathize}\}.$
@@
 \item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
-  $\mu(s)$ is the probability that the environment starts-off in state $s$.
+  $\mu(s)$ is the probability that the environment starts off in state $s$.
 \end{itemize}
@@
 
-\subsection{Long-term rewards and policies}
+\subsection{Long-term rewards and action policies}
 The behavior of the agent is governed by some kind of \textit{policy},
@@
 a set of candidate actions to perform given this state. Starting a time $t=0$,
-a policy $\pi$ generates a trajectory as follows
+a policy $\pi$ generates a trajectory of action cascades as follows
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-Since an action taken now might have repercussions long into the feature, it turns out that the
+Since an action taken in the present moment $t$
+might have repercussions long into the future, it turns out that the
 quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
@@
 \end{mdframed}
-  The  goal of RL is then adapt this policy so as to  maximize the cumulative rewards. In
+  The  goal of RL is then to adapt this policy so as to  maximize the cumulative rewards. In
   the definition of $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
@@
 immediate rewards and has no horizon for the future, which is not compatible with long-term goal planning
-as potentially implemented in the DMN. For there to be any learning at all, it is thus necessary that $0 < \gamma < 1$.
+as potentially implemented in the DMN.
+To allow a learning process to arise,
+it is thus necessary that $0 < \gamma < 1$.
 % \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
-Thus, the agent is positively biased towards considering rewards which with
+Thus, the agent is positively biased towards considering rewards with
 a short time delay. The reward importance hence decays exponentially
 with distance in the future,
+\textbf{XXX correct grammar in previous sentences}
 given that there is more uncertainty in the farsighted future.
@@
 
-\subsection{The Q-value: a goodness measure for actions}
-The agent's goal can then be formulated as finding policy $\pi$ which maximizes the
+\subsection{The Q-value: A goodness measure for actions}
+The purpose of the DMN can then be
+formulated as finding a policy $\pi$ for the agent which maximizes the
 expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
@@
 \end{equation}
-In plain English, the Q-value $Q^{\pi}(s,a)$ is the expected / average / mean future reward
-over all possible trajectories in which the agent starts off starts at with an environment in state
-$s$, takes and action $a$, and then follows the policy $\pi$ thereafter to select future actions.
+In other words,
+the Q-value $Q^{\pi}(s,a)$ corresponds to the expected reward
+over all possible action trajectories in which
+the agent starts off starts at with an environment in state
+$s$, chooses action $a$,
+and then follows the policy $\pi$ thereafter to select future actions.
 %% Given that each $\pi$ correspondence to an exhaustive description of
@@
 When a choice of actions is available, the $Q^{\pi}(s,a)$
-provides a formalization of optimal selection.
+provides a formalization of optimal behavior,
+potentially driven by the DMN.
 
@@
 trajectories the agent might need to
-consider.
+consider simultaneously.
 % \begin{mdframed}
@@
 % \end{mdframed}
-One popular simplification that removes
+One popular solution that remedies
 this issue is to only consider the sub-family of
@@
 should be taken at thin state, and therefore define functions from states to actions.
-
-\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
+\textbf{XXX sentence too long, please cut into 2}
+
+\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992), a commonly used off-policy algorithm
+\textbf{XXX WHATTTEEE?}
+, uses the
 \textit{greedy policy}
@@
 % table.
-The \textit{Bellman's equation} \cite{sutton1998reinforcement} then takes the simple form
+The \textit{Bellman equation} \cite{sutton1998reinforcement} then takes the simple form
 \begin{equation}
@@
 % It is a fixed-point equation for the deterministic policy $\pi$.
-Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which
+Notice that in \eqref{eq:bellman}, the sampling is done over only things which
+\textbf{XXX Ahhmm, please make a bigger effort here}
 depend on the environment, and so can be learned off-policy by observing state-transitions
@@
 A successful RL theory needs ways to approximate the state-action value function
-which scale up to large state spaces and / or action spaces.
+which scale up to large state and action spaces.
 In state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
@@
 \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
-$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ (hyppocampus ?)
+$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$
 and trying to approximate
@@
 \mapsto {Q}(s, a|\theta^Q_{k})$, where $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q_{k+1}$.
-
-In the brain, such an experience replay mechanism is probably implemented in the hyppocampus (HC). Indeed,
-hippocampal replay is a phenomenon observed in rats, mice, cats, rabbits, songbirds, and monkeys.
+\textbf{XXX please cut previous sentence into smaller pieces}
+
+In the human brain, such an experience replay mechanism
+is likely to rely on the hyppocampus (HC). Indeed,
+hippocampal replay is a phenomenon observed in rats, mice, cats, rabbits, songbirds, and monkeys \textbf{XXX citation !!!}.
 Indeed, during sleep or awake rest, replay refers to the re-occurrence of a sequence of cell activations that also occurred during activity, but the replay has a much faster time scale.
@@
 where
-$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
+$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target \textbf{XXX unpack what the target is !!!}.
 % While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
-For example, a general linear model with a kernel $\phi$ would be of the
+For instance, a general linear model with a kernel $\phi$ would be of the
 form
@@
 $\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action $(s,a)$ into an appropriate
-feature space. Such are the techniques that have been proposed by \cite{songNIPS2016}, for example.
+feature space. Such are the techniques that have been proposed by \cite{songNIPS2016}.
 %  and Parr et al. 2003 (JMLR)].
@@
 
-\subsubsection{Interlude: zoom on the Hippocampus}
+\subsubsection{Interlude: The role of the hippocampus}
 \paragraph{Does the hippocampus do Monte-Carlo sampling ?}
@@
 
-In the brain, the HC would generate / imagine sequences of world states, observations and rewards.
+In the human brain,
+the HC could contribute to synthesizing imagined sequences of world states,
+observations and rewards.
 These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
@@
   \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
-  (e.g linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
+  (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
   For example, given traces of motor actions from a an adult to grab a cup on a table,
@@
 IRL is is suited for problems in which it's hard to define what the reward function could be
-(e.g car-driving, drone maneuvering, etc.) ...
+(e.g.,  car-driving, drone maneuvering, etc.) ...
 
@@
 The DMN is today known to consistently increase in neural
-  activity when engaging cognitive processes that are detached from
-  current sensory environment \citep{kenet2003spontaneously, fiser2004small}.
+  activity when humans engage in goal-directed behavior that are detached from
+  current sensory environment \citep{kenet2003spontaneously, fiser2004small}
+  and it was proposed to be situated at the top of the functional network hierarchies
+  \citep{carhart2010default, margulies2016situating}.
   Its involvement in thinking about the past,
@@
   In the absence of environmental input and feedback
-  (e.g., mind-wandering or sleep) pseudo-experiencing possible
+  (e.g., mind-wandering or sleep) pseudo-experiencing (i.e., emulating) possible
   future scenarios and action outcomes.
@@
   successful action choices and reward predictions in future states.
-  Plasticity in the DMN also arises naturally:
+  As such,
+  plasticity in the DMN arises naturally:
   If an agent behaving optimally in a given environment moves
@@
   massively increase. This will lead to adaptation of the policy until the system converges to a 
-  new steady-state of optimal action decisions given certain states.
+  new steady-state of optimal action decisions in the volatile environment.
 
@@
 % \subsubsection{The theory}
-In Friston's free-energy principle (FEP) \citep{friston2010free,fristonAIorRL}, the brain emerges as biomechanical inference engine which must minimize the long-term average of surprise.
+In Friston's free-energy principle (FEP) \citep{friston2010free,fristonAIorRL}, the brain is portrayed
+as biomechanical inference engine which must minimize the long-term average of surprise.
 Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net (DBN)}.
@@
   \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}. Such a theory cannot by itself,
-  explain the emergence of strategic behavior inherent in humans (e.g cite dark-room experiment).
+  explain the emergence of strategic behavior inherent in humans (e.g.,  cite dark-room experiment).
 
@@
 MDPs provide an attractive
-computational model how the human association cortex
+process model how the human association cortex
 might implement supra-modal representation and control of the environment to
@@
 %
-From the computational viewed of a MDP,
+From the computational viewed of a Markov decision process,
 behavior unfolds by integrating happened past events
@@
 At the least, we propose an alternative vocabulary to
-describe and interpret neuroscientific studies in DMN research.
+describe and interpret experimental findings in neuroscience studies
+on the default mode network.
 %
@@
 Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises.
-For example the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial, e.g) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
+For example the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
 

### 2017-03-29 paper/paper.tex
latest shit

@@
 This may be why performance within and across participants
-was related to state encoding in OFC.
+was related to state encoding in vmPFC.
 Such a \textit{cognitive map} of action space was argued to encode
@@
 in the sense that only part of the current state is observed by
-the agent.
+the agent \citep{starkweather2017dopamine}.
 This unpredictability
@@
 % table.
-The \textit{Bellman equation} \cite{sutton1998reinforcement} then takes the simple form
+The \textit{Bellman equation} \citep{sutton1998reinforcement} then takes the simple form
 \begin{equation}
@@
 %  and Parr et al. 2003 (JMLR)].
-A breakthrough alternative proposed by Google DeepMind \cite{mnih2015,silver2016mastering}, is to learn this
+A recently proposed breakthrough alternative
+\citep{mnih2015,silver2016mastering}, is to learn this
 representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which
@@
 Note that most RL algorithms, including \textit{Temporal Difference (TD)}
-learning \cite{sutton1998reinforcement}, REINFORCE \cite{williams1992}, and SARSA can be
+learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
 cast in this general variational framework. For example TD corresponds to

### 2017-03-29 paper/paper.tex
blah

@@
 This paper proposes a formal model that satisfies this contention.
-
-We show that the DMN is endowed with sufficient neurobiological equipment to
-implement a reinforcement learning (RL) agent, at least in principle. A global
-picture of the current body of knowledge (predictive coding, free-energy principle,
-sentinel hypothesis, etc.) concerning decision-making in the brain
-is drawn, with the DMN at its core. Our work ...
 

### 2017-03-30 paper/paper.tex
a quick pass

@@
 
-Reinforcement learning is learning to achieve a goal by interacting with an external
-\textit{environment}.
-\textbf{XXX Don't define something with the same thing}
-In a general RL framework, an \textit{agent}
-interacts with an  environment in a trial-by-error fashion by taking
-\textit{actions} $a$, which trigger changes in the
+Reinforcement learning is a problem-solving technique in which, through their interactions with an external environment,
+an agent learns to reach a goal, optimize a reward signal, or minimize some cost,
+in an iterative trial-by-error fashion.
+At any moment, each action $a$ taken triggers a changes in the
 \textit{state} of the environment
@@
 the action of the agent and the reward should be thought
-of as an instantaneous
-\textbf{XXX instant is problematic here, because it may change behavior only later in RL}
-satisfaction
-\textbf{XXX or punishment?}
-accompanying the execution of
+of as satisfaction --or punishment-- accompanying the execution of
 an action, and may be delayed, only becoming substantial when accumulated over an extended period of time.
-%% Here the subscript $t$ denotes time, and can be discrete or
-%% continuous.
-The environment is generally taken as stochastic,
-\textbf{XXX One sentence, one bullet. Please explain stochast and PO
-in one sentence each.}
-is \textit{partially observed},
+The environment is generally taken as stochastic, i.e unpredictable, changing in random ways. In addition,
+it is only \textit{partially observed},
 in the sense that only part of the current state is observed by
 the agent \citep{starkweather2017dopamine}.
-This unpredictability
+This unpredictability of the environment
 \textbf{XXX Explain this sentence better because it is important.}
@@
 \subsection{Partially Observable Markov Decision Processes}
-It is our view that at a sufficiently coarse scale, the brain is a physical system governed by the laws of
-\textbf{XXX the paper is about the DMN more specifically, perhaps try to sharpen}
+It is our view that at a sufficiently coarse scale, the brain --and therefore the DMN in particular--
+is a physical system governed by the laws of
 physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
@@
 and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.
-Such model-free RL is plausible in the human brain. Indeed, as was noted in \cite{dayan2008decision}, it has long been suggested in the literature that there
+
+Model-free RL is plausible in the human brain. Indeed, as was noted in \cite{dayan2008decision}, it has long been suggested in the literature that there
   is a rather direct mapping of model-free RL learning algorithms
 onto the brain,
-\textbf{XXX New sentence}
 with the neuromodulator dopamine serving
@@
 dorsolateral striatum.
-\textbf{XXX Loose all boxes by integrating the text please since it is not a usual layouting}
 
@@
 % environments \citep{gershman2015computational}.
-\textbf{XXX transition back to model-free RL is not clear. please make explicit}
 Our rat might represent such knowledge about the world as:
@@
 
+\begin{definition}
 Mathematically, a POMDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
@@
 % \end{remark}
+\end{definition}
 
@@
 % \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
-Thus, the agent is positively biased towards considering rewards with
-a short time delay. The reward importance hence decays exponentially
-with distance in the future,
-\textbf{XXX correct grammar in previous sentences}
-given that there is more uncertainty in the farsighted future.
-The reward discount factor $\gamma$ can thus be seen as calibrating
+Thus, the agent is more considering of rewards in to come in a short future, paying lesser attention to
+rewards to come in at a later time.
+% given that there is more uncertainty in the farsighted future.
+The reward discount factor $\gamma$ can then be seen as calibrating
 the delay discounting behavior of the intelligent agent,
-that is the decision related to reward-delay tradeoff schedules.
-It is also important to appreciate that a stochastic policy estimation
-is more advantageous in many RL settings.
-For instance, a deterministic strategy in playing rock-paper-scissors can
-be easily exploited and a uniformly random choice is optimal.
-\textbf{XXX I love this example :-)}
+that is the decision related to reward-delay trade-off schedules.
+% It is also important to appreciate that a stochastic policy estimation
+% is more advantageous in many RL settings.
+% For instance, a deterministic strategy in playing rock-paper-scissors can
+% be easily exploited and a uniformly random choice is optimal.
+% \textbf{XXX I love this example :-)}
 
@@
 % \end{mdframed}
-One popular solution that remedies
+One popular solution concept is called \text{Q-learning}. It remedies
 this issue is to only consider the sub-family of
-\textit{deterministic policies}, which map each state to the best single best action which
-should be taken at thin state, and therefore define functions from states to actions.
-\textbf{XXX sentence too long, please cut into 2}
-
-\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992), a commonly used off-policy algorithm
-\textbf{XXX WHATTTEEE?}
-, uses the
+\textit{deterministic policies}, which map each state to the best single best action to
+take at the state. Such policies define deterministic functions from states to actions.
+
+\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992), generates a
 \textit{greedy policy}

### 2017-03-30 paper/paper.tex
second pass

@@
 and trying to approximate
-the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
+the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the Bellman equation \eqref{eq:bellman}, namely
 $r + \gamma \max_{a'} Q(s', a'|\theta^Q_k)$, with the output of a parametrized regression model $(s,a,r,s')
-\mapsto {Q}(s, a|\theta^Q_{k})$, where $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q_{k+1}$.
-\textbf{XXX please cut previous sentence into smaller pieces}
+\mapsto {Q}(s, a|\theta^Q_{k})$. Here, $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by weights $\theta^Q_{k+1}$
+(synaptic strengths, etc.).
 
@@
 where
-$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target \textbf{XXX unpack what the target is !!!}.
+$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
 % While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.

### 2017-03-30 paper/paper.tex
another pass

@@
 cumulative reward estimate which takes into account the future. A standard for
-modeling this accumulation is the so-called
-time-discounted \textit{cumulative reward} function
+modeling this accumulation is the time-discounted \textit{cumulative reward}% function
 \begin{equation}
   \label{eq:cumr}
-  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t.
+  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
 \end{equation}
@@
 \end{mdframed}
-  The  goal of RL is then to adapt this policy so as to  maximize the cumulative rewards. In
-  the definition of $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
+  The  goal of RL is then to adapt this policy so as to  maximize $G^\pi$. In
+  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
 Taking $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their
@@
 
-In the human brain, such an experience replay mechanism
-is likely to rely on the hyppocampus (HC). Indeed,
-hippocampal replay is a phenomenon observed in rats, mice, cats, rabbits, songbirds, and monkeys \textbf{XXX citation !!!}.
-Indeed, during sleep or awake rest, replay refers to the re-occurrence of a sequence of cell activations that also occurred during activity, but the replay has a much faster time scale.
+Biologically, experience replay would be manifested as the re-occurrence of a sequence of cell activations that also occurred during activity.
+%, but the replay has a much faster time scale.
+In the human brain, such a mechanism is likely to be implemented by the
+hyppocampus (HC), as experiments on various animals indicate this is the case.
+Indeed, this phenomenon has been observed in rats, mice, cats, rabbits, songbirds, and
+monkeys \cite{buhry2011,nokia2010,dave2000,skaggs2007}.
 

### 2017-03-30 paper/paper.tex
some comments on biology sections

@@
 \section{Known Neurobiological Properties of the Default Mode Network}
+\textbf{XXX (danilo) this section needs more tight integration of the parts, and no just a separate independent subsection for each portion of the DMN...}
 We begin by a deconstruction the overall DMN function
@@
     environmental perception and internally generated mind-wandering\footnote{
+      \textbf{XXX (danilo) This is more than just a footnote. Make this part of the paper
+      and factorize the lot. Also you probably need a citation here!!!}
   ``A computational model of how such a process might be structured
@@
   instead of accurate and reliable.
-  Yet, the MDP framework provides an algorithmic explanation in that
+\textbf{XXX (danilo) Is this a loose opinion or a fact ? What do you mean by ``probably better models of the world'' ?}  Yet, the MDP framework provides an algorithmic explanation in that
   stochasticity of the parameter space search implemented
@@
 
-\textbf{XXX Your focus in this subsection ought to be on the SC itself not POMPDs, etc.}
+\textbf{XXX (danilo) Your focus in this subsection ought to be on the SC itself not POMPDs, etc.}
  \begin{itemize}
@@
     how attention and memory resources are exactly reallocated when
-    encountering a salient environmental stimulus. \textbf{XXX remove the next sentence. It's probably false.}
+    encountering a salient environmental stimulus. \textbf{XXX (danilo) remove the next sentence. It's probably false.}
     In contrast,

### 2017-03-30 paper/paper.tex
some XXXs

@@
   \textit{and back-propagation} in MDPs and RL,
-  although this may not be neurobiologically plausible.
+  although this may not be neurobiologically plausible \textbf{XXX (danilo) what is not plausible. Be specific.}
   \item In sum,
@@
   As such, our account can be viewed as a RL implemention of the
-  prediction coding hypothesis.
+  prediction coding hypothesis. \textbf{XXX (danilo) Reformulate. Also in general, you should be clear whether
+    you're talking about MDPs, POMDPS, both, or neither. These are being mixed up a lot.}
 \end{itemize}

### 2017-03-30 paper/paper.tex
pass 1/2

--- 2017-03-30 paper/paper.tex
@@
 %
-\subsection{The posteriormedial cortex: Global monitoring and integration}
+\subsection{The posteromedial cortex: Global monitoring and integration}
 The posteriormedial cortex (PMC) and
@@
 
-\subsection{The prefrontal cortex: Stimulus-value assocation and action consideration}
+\subsection{The prefrontal cortex: Stimulus-value association and action consideration}
 The dmPFC subserves predominantly ambiguous amodal processes
@@
 the vmPFC with the nucleus accumbens, hippocampus (HC),
-and posteriomedial cortex.
+and posteromedial cortex.
 %
@@
 
-Reinforcement learning is a problem-solving technique in which, through their interactions with an external environment,
-an agent learns to reach a goal, optimize a reward signal, or minimize some cost,
+RL is a problem-solving technique in which, through their interactions with an external environment,
+an agent learns to reach goals, optimize reward signals, or minimize costs
 in an iterative trial-by-error fashion.
-At any moment, each action $a$ taken triggers a changes in the
+At any moment, each taken \textit{action} $a$ triggers a changes in the
 \textit{state} of the environment
@@
 the action of the agent and the reward should be thought
-of as satisfaction --or punishment-- accompanying the execution of
-an action, and may be delayed, only becoming substantial when accumulated over an extended period of time.
-The environment is generally taken as stochastic, i.e unpredictable, changing in random ways. In addition,
+of as satisfaction -- or punishment -- accompanying the execution of
+an action.
+Action may be delayed to achieve substantial increases in expected reward
+that can increase over time.
+The environment is generally taken as stochastic,
+that is, unpredictable, changing in random ways. In addition,
 it is only \textit{partially observed},
@@
 This unpredictability of the environment
-\textbf{XXX Explain this sentence better because it is important.}
 is realistic in a model which sets out
-to explain DMN function. Indeed, it is known that, the more the external world is predictable,
+to explain DMN function. Indeed, it is known that the more the external world is predictable,
 the more mental activity becomes detached from the actual sensory inputs
@@
 optimizing control of the organism.
-A DMN framework incorporating reinforcement learning
+Such a framework of DMN funtion based on reinforcement learning
 can naturally embed human behavior
-into the tension between exploitative action with immediate gains and
+in the tension between exploitative action with immediate gains and
 explorative action with longer-term reward schedules
 \citep{dayan2008decision}.
-The implication of the DMN in a diversity of human cognitive processes
+The implication of the DMN in a diversity of
+particularly advanced cognitive processes
 can be parsimoneously explained as prediction error minimization
@@
 \subsection{Partially Observable Markov Decision Processes}
-It is our view that at a sufficiently coarse scale, the brain --and therefore the DMN in particular--
+It is our view that at a sufficiently coarse scale, the brain -- and therefore the DMN in particular --
 is a physical system governed by the laws of
@@
 If the process has $memory$ such that the next state depends not only on the current state but also on some finite
-number of past states,
-rational probabilistic planning can be reformulated
+number of past states.
+Rational probabilistic planning can be reformulated
 as a standard memoryless Markov process by simply expanding the
@@
 
-In artificial intelligence (AI), a popular abstract model for
+In artificial intelligence (AI), a popular computational model for
 multi-step decision processes in such an environment are so-called
 \textit{Partially Observable Markov Decision Processes (POMDPs)}.
-A POMDP models a decision process in which it is assumed that environment dynamics are determined by a Markov process,
-but the agent cannot directly observe the underlying state. Instead, it tries to optimize a \textit{subjective} reward
-signal. This is performed alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
+A POMDP formalizes a decision process in which it is assumed that environment dynamics are determined by a Markov process,
+but the agent cannot directly observe the underlying state.
+Instead, the agent tries to optimize a \textit{subjective} reward
+signal (i.e., is likely to be different for another agent in the same state).
+This is performed alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
 states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
@@
 % environments \citep{gershman2015computational}.
-Our rat might represent such knowledge about the world as:
+
+In our present model-free RL framework,
+a rat might represent such knowledge about the world as follows:
 \begin{itemize}
@@
   the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
-Taking $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their
-immediate rewards and has no horizon for the future, which is not compatible with long-term goal planning
+Setting $\gamma = 0$ corresponds to a perfectly myopic agent who is solely concerned about its
+immediate rewards.
+Such an agent would have no horizon for the future,
+which is not compatible with long-term goal planning
 as potentially implemented in the DMN.
 To allow a learning process to arise,
-it is thus necessary that $0 < \gamma < 1$.
+it is necessary that $0 < \gamma < 1$.
 % \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
-Thus, the agent is more considering of rewards in to come in a short future, paying lesser attention to
-rewards to come in at a later time.
+Such an agent is thus better emphasizing consideration of rewards expected in
+a short future and paying less attention to
+rewards expected at a later time.
 % given that there is more uncertainty in the farsighted future.
@@
 the delay discounting behavior of the intelligent agent,
-that is the decision related to reward-delay trade-off schedules.
-% It is also important to appreciate that a stochastic policy estimation
-% is more advantageous in many RL settings.
-% For instance, a deterministic strategy in playing rock-paper-scissors can
-% be easily exploited and a uniformly random choice is optimal.
-% \textbf{XXX I love this example :-)}
-
-\subsection{The Q-value: A goodness measure for actions}
+that is, the decision related to reward-delay trade-off schedules.
+It is also important to appreciate that a stochastic policy estimation
+is more advantageous in many RL settings.
+For instance, a deterministic strategy in playing rock-paper-scissors can
+be easily exploited and a uniformly random choice is optimal.
+
+\subsection{The Q-value: A goodness measure for action choices}
 The purpose of the DMN can then be
@@
 the Q-value $Q^{\pi}(s,a)$ corresponds to the expected reward
-over all possible action trajectories in which
-the agent starts off starts at with an environment in state
+over all possible action trajectories, in which
+the agent sets out in the environment in state
 $s$, chooses action $a$,
@@
 utility of executing a specific action; it answers the question
-"What is the expected utiltiy of taking action $a$ in this situation ?".
+"What is the expected utility of taking action $a$ in this situation ?".
 When a choice of actions is available, the $Q^{\pi}(s,a)$
-provides a formalization of optimal behavior,
+offers a formalization of optimal behavior,
 potentially driven by the DMN.
@@
 \subsection{Computing optimal behavior}
-By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general
-due to the enormous number of possible action
+By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general.
+This is due to the enormous number of possible action
 trajectories the agent might need to
@@
 One popular solution concept is called \text{Q-learning}. It remedies
-this issue is to only consider the sub-family of
-\textit{deterministic policies}, which map each state to the best single best action to
+the issue by only considering the sub-family of
+\textit{deterministic policies}, which map each state to the single best action to
 take at the state. Such policies define deterministic functions from states to actions.
@@
     \\
-    &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))],
+    &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))].
   \end{split}
@@
 \end{equation}
-a \textit{fixed-point} equation which provides a recursive decomposition of optimal behavior.
+This \textit{fixed-point} equation provides a recursive decomposition of optimal behavior.
 A complicated dynamic programming optimization can be
@@
 \citep{koechlin1999role, braver2002role}.
-Using the Bellman equation, each state can be associated with a certain value,
+Using the Bellman equation, each state can be associated with a certain value
 to guide action towards a better state, thus improving on the agent's current policy.
 % It is a fixed-point equation for the deterministic policy $\pi$.
-Notice that in \eqref{eq:bellman}, the sampling is done over only things which
-\textbf{XXX Ahhmm, please make a bigger effort here}
-depend on the environment, and so can be learned off-policy by observing state-transitions
+Notice that in \eqref{eq:bellman}, the sampling is done only over things which
+depend on the environment, and so can be learned off-policy by observing state transitions
 triggered by another behavioral policy, which can be stochastic.
@@
 A successful RL theory needs ways to approximate the state-action value function
-which scale up to large state and action spaces.
+to allow for scaling to large state and action spaces.
 In state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
@@
 % (measured on their internal currency of worldly values ?),
-and the environment transitions to a new state $s' \in \mathcal S$. At the end of
+and the environment transitions to a new state $s' \in \mathcal S$.
+\textbf{XXX Please try to smoothen the previous sentence}
+At the end of
 this step, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar
@@
 $r + \gamma \max_{a'} Q(s', a'|\theta^Q_k)$, with the output of a parametrized regression model $(s,a,r,s')
-\mapsto {Q}(s, a|\theta^Q_{k})$. Here, $Q(.,.|\theta^Q_{k})$ is an approximator for the $Q$-value operator,  parameterized by weights $\theta^Q_{k+1}$
+\mapsto {Q}(s, a|\theta^Q_{k})$.
+Here, $Q(.,.|\theta^Q_{k})$
+\textbf{XXX just double-checking, are the dots int he Q() correct?}
+is an approximator for the $Q$-value operator,  parameterized by weights $\theta^Q_{k+1}$
 (synaptic strengths, etc.).
 
-Biologically, experience replay would be manifested as the re-occurrence of a sequence of cell activations that also occurred during activity.
+From a neurobiological perspective,
+experience replay would be manifested as the re-occurrence of a sequence of cell activations that also occur during specific actions.
 %, but the replay has a much faster time scale.
-In the human brain, such a mechanism is likely to be implemented by the
-hyppocampus (HC), as experiments on various animals indicate this is the case.
-Indeed, this phenomenon has been observed in rats, mice, cats, rabbits, songbirds, and
+In the human brain, the hippocampus is likely to contribute to such a mechanism,
+as neuroscience experiments on various animals have repeatedly indicated
+in rats, mice, cats, rabbits, songbirds, and
 monkeys \cite{buhry2011,nokia2010,dave2000,skaggs2007}.

### 2017-03-30 paper/paper.tex
pass 2/2

@@
 \textit{state} of the environment
-$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
+$s \rightarrow s'$, accompanied by environmental feed signals as \textit{reward}
 $r = r(s, a)$ (or regret!, since it can be negative) collected by the
@@
 $${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
-$\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action $(s,a)$ into an appropriate
-feature space. Such are the techniques that have been proposed by \cite{songNIPS2016}.
+$\phi(s,a)$ would represent a high-level hand-crafted representation in an
+appropriate form of the state-action pairs $(s,a)$,
+as proposed by \cite{songNIPS2016}.
 %  and Parr et al. 2003 (JMLR)].
@@
 are by far the current state-of-the-art in RL. In such a case, $\theta^Q$
-would correspond to the strengths of low-level synaptic or high-level functional connections between different
-neurons / nodes / regions in the network.
-
-Learning of the the entire model parameters can effictively be achieved via \textit{backprop},
-wherein prediction errors (aka \textit{regrets})  are propagated from lower to higher regions to modulate the choice of future actions
+would correspond to the strengths of low-level synaptic or high-level functional connections between different elements
+in brain networks.
+\textbf{XXX perhaps add one sentence to rephrase and detail the idea}
+
+Learning (i.e., fitting)
+of the the entire model parameters can effictively be achieved via \textit{backprop},
+wherein prediction errors (i.e., \textit{regrets})  are propagated from lower to higher regions to modulate the choice of future actions
 % online stochastic gradient descent is used to minimize the variational objective
@@
 learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
-cast in this general variational framework. For example TD corresponds to
+cast in this general variational framework. For instance, TD corresponds to
 the above framework using a linear value approximator with feature encoding
@@
 
-\subsubsection{Interlude: The role of the hippocampus}
-\paragraph{Does the hippocampus do Monte-Carlo sampling ?}
+\subsubsection{Excursion: The role of the hippocampus}
+\paragraph{Does the hippocampus do Monte-Carlo sampling?}
 In RL, Monte-Carlo simulation can be used to update the agent's belief state
@@
 Monte-Carlo simulation provides a simple method for evaluating the value / fitness of a state.
-They provide an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
+They provide an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than a RL agent without stochastic resampling procedures.
 
@@
 These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
-This would be a simple control algorithm by evaluating all legal actions and selecting the action with highest evaluation.
+This would be a simple control algorithm by evaluating all legal actions and selecting the action with
+\textbf{highest expected cumulative rewards [please double-check]}.
 In these challenging POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
@@
 
-\paragraph{Hippocampal replay = inverse reinforcement-learning ?}
+\paragraph{Hippocampal replay corresponds to inverse reinforcement-learning?}
 %\begin{mdframed}
@@
   (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
-  For example, given traces of motor actions from a an adult to grab a cup on a table,
-  can an observing child figure out what ``langragian'' functional is being minimized by the former ?
-  How ? Can they reproduce this optimal behavior ?
+  For instance, given traces of motor actions from a an adult to grab a cup on a table,
+  can an observing child figure out what ``langragian'' functional is being minimized by the former?
+  How? Can they reproduce this optimal behavior?
 %\end{mdframed}
@@
   \includegraphics[width=.9\linewidth]{rl_process_chart.pdf}
-  \caption{The DMN as an RL agent...}
+  \caption{The DMN as an RL agent... \textbf{Please add a figure caption}}
   \label{fig:rl_process_chart}
@@
   cumulative outcomes. Implicit computation of future choices
-  provides an explanation for purposeful mind-wandering.
-
-The HC provides perturbed action-transition-state-reward samples
-  as a block of "imagined", "hypothesized", "recalled" experience;
-  the small variations in these experience blocks allow searching
-  a larger space of parameters and possible experiences.
+  provides an explanation for the
+  evolutionary emergence and practical usefulness of mind-wandering
+  in humans.
+
+The HC may provide perturbed action-transition-state-reward samples
+  as a block of "imagined", "hypothesized", "recalled" experience.
+  The small variations in these experience blocks allow searching
+  a larger space of model parameters and possible experiences.
+  Taken to its extreme, stochastic recombination of experience
+  building blocks can further optimize the behavior of the RL agent
+  by model learning from scenarios in the environment that the agent might
+  only very rarely or never encounter.
   In the absence of environmental input and feedback
@@
   massively increase. This will lead to adaptation of the policy until the system converges to a 
-  new steady-state of optimal action decisions in the volatile environment.
+  new steady-state of optimal action decisions in volatile environments.
 
@@
 \textit{variational auto-encoders (VAEs)} \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm, as it overcomes the technical limits of the former, by using a reparametrization trick.
-For example, unlike the wake-sleep algorithm for minimizing free-energy, VAEs can be efficiently trained via back-propagation of learning errors.
+For instance, unlike the wake-sleep algorithm for minimizing free-energy, VAEs can be efficiently trained via back-propagation of learning errors.
 
@@
     how attention and memory resources are exactly reallocated when
-    encountering a salient environmental stimulus. \textbf{XXX remove the next sentence. It's probably false.}
+    encountering a salient environmental stimulus. \textbf{XXX remove the next sentence. Danilo: can you help me to correct the sentence instead.
+    There appears to be a cool link between cognition and formal theory
+    here.}
     In contrast,
@@
 %
-From the computational viewed of a Markov decision process,
+From the computational view of a Markovian sequential decision process,
 behavior unfolds by integrating happened past events
@@
 when confronted with complicated decisions.
-Algorithmic gains by injecting stochasticity in the recall of past
-actions and outcomes to improve the internal world model may explain why
+Improvement of the internal world model
+by injecting stochasticity in the recall of past
+actions and outcomes may explain why
 very accurate memories have been disfavored in human evolution
@@
 Ultimately,
-the DMN can be viewed as a smoothing kernel extending
+the DMN can be viewed as a functional integrator
 from the relatively recent events to anticipated upcoming events
-to constantly improve sensory processing and
-action performance in the world.
+in order to constantly improve present action in the world.
 
@@
 Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises.
-For example the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
+For instance, the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
 

### 2017-03-30 paper/paper.tex
modify in response to XXX

@@
   \textit{and back-propagation} in MDPs and RL,
-  although this may not be neurobiologically plausible \textbf{XXX (danilo) what is not plausible. Be specific.}
+  although the neurobiological plausibility of
+  the back-propagation procedure is controversal
+  \citep{goodfellow2016deep}.
   \item In sum,
@@
   %
-  As such, our account can be viewed as a RL implemention of the
-  prediction coding hypothesis. \textbf{XXX (danilo) Reformulate. Also in general, you should be clear whether
-    you're talking about MDPs, POMDPS, both, or neither. These are being mixed up a lot.}
+  As such, our MDP account of DMN function can parsimoneously explain
+  prediction coding hypothesis.
 \end{itemize}
@@
   stimulus processing and action choice in the present.
-  \item While mental scene construction implies a distinction between
+  \item Mental scene construction has been proposed by some authors
+  \citep{buckner2007self} to imply a distinction between
     environmental perception and internally generated mind-wandering\footnote{
@@
   while simulation of possible alternatives are constructed,
-  followed by a return to perception of the present.''\citep{buckner2007self}},
-  the MDPs naturally integrate the former egocentric
+  followed by a return to perception of the present.''\citep{buckner2007self}}.
+  The MDPs naturally integrate the former egocentric
   (more related to current action, state and reward) and the later
@@
   instead of accurate and reliable.
-\textbf{XXX (danilo) Is this a loose opinion or a fact ? What do you mean by ``probably better models of the world'' ?}  Yet, the MDP framework provides an algorithmic explanation in that
-  stochasticity of the parameter space search implemented
-  by Monte Carlo solvers provably yields better models of the world.
+  Yet, the MDP framework provides an algorithmic explanation in that
+  stochasticity of the parameter space explored
+  by the Monte Carlo solvers achieves better fine-tuning of the
+  action policies and estimation of expected reward outcomes.
   That is, the purposeful stochasticity of policy and value estimation
@@
 
-\textbf{XXX (danilo) Your focus in this subsection ought to be on the SC itself not POMPDs, etc.}
+\textbf{XXX (danilo) Your focus in this subsection ought to be on the SC itself not POMPDs. -> agree I will fix it later.}
  \begin{itemize}

### 2017-03-30 paper/paper.tex
add cool cit

@@
 \end{equation}
-This \textit{fixed-point} equation provides a recursive decomposition of optimal behavior.
+This \textit{fixed-point} equation provides a recursive decomposition of optimal behavior by dividing the value function into the immediate reward
+and the discounted rewards of the upcoming states.
 A complicated dynamic programming optimization can be
@@
 the HC could contribute to synthesizing imagined sequences of world states,
-observations and rewards.
+observations and rewards \citep{aronov2017}.
 These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
@@
 from the relatively recent events to anticipated upcoming events
-in order to constantly improve present action in the world.
+in order to constantly improve present action in the dynamic world.
 

### 2017-04-01 paper/paper.tex
yet another monster pass

@@
 \section{Known Neurobiological Properties of the Default Mode Network}
-\textbf{XXX (danilo) this section needs more tight integration of the parts, and no just a separate independent subsection for each portion of the DMN...}
+\textbf{XXX Elvis: this section needs more tight integration of the parts, and no just a separate independent subsection for each portion of the DMN...}
 We begin by a deconstruction the overall DMN function
@@
 \begin{mdframed}
-  \textbf{Where is ``value'' buffered in the brain ? You mean where in the DMN!
-  then the answer is in the vmPFC which is connected to the NAc}
-  The amygdala is known to be involved in significance evaluation) and
-  the nucleus accumbens (NAc) is involved in reward evaluation.
+  \textbf{Where is ``value'' buffered in the DMN ?
+    % brain ?
+    }
+  The amygdala is known to be involved in significance evaluation while
+  the nucleus accumbens (NAc), which is connected to the vmPFC, is involved in reward evaluation.
 \end{mdframed}
@@
 % deterministic policies of the form \eqref{eq:qlearning}.
-A successful RL theory needs ways to approximate the state-action value function
+A successful RL theory needs ways to efficiently approximate the state-action value function
 to allow for scaling to large state and action spaces.
-In state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
+In a state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
 % (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
-by sampling from its current policy matrix, and collects a reward $r$,
+by sampling from its current policy, and collects a reward $r$,
 % (measured on their internal currency of worldly values ?),
-and the environment transitions to a new state $s' \in \mathcal S$.
-\textbf{XXX Please try to smoothen the previous sentence}
-At the end of
-this step, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar
+after which the environment transitions to a new state $s' \in \mathcal S$.
+% \textbf{XXX Please try to smoothen the previous sentence}
+At the end of this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced, which represents an exemplar
 behavior of the agent and is recorded in replay memory buffer:
@@
 \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
-$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$
-and trying to approximate
+$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$.
+The agent then tries to approximate
 the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the Bellman equation \eqref{eq:bellman}, namely
-$r + \gamma \max_{a'} Q(s', a'|\theta^Q_k)$, with the output of a parametrized regression model $(s,a,r,s')
-\mapsto {Q}(s, a|\theta^Q_{k})$.
-Here, $Q(.,.|\theta^Q_{k})$
-\textbf{XXX just double-checking, are the dots int he Q() correct?}
+$y := r + \gamma \max_{a'} Q(s', a'|\theta^Q_k)$, with the output of a parametrized regression model $(s,a,r,s')
+\mapsto {Q}(s, a|\theta^Q_{k})$. Here the function $(s,a) \mapsto Q(a,s|\theta^Q_{k})$
+% \textbf{XXX just double-checking, are the dots int he Q() correct?}
 is an approximator for the $Q$-value operator,  parameterized by weights $\theta^Q_{k+1}$
@@
 where
-$ y = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
+$ y: = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
 % While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
@@
 in brain networks.
-\textbf{XXX perhaps add one sentence to rephrase and detail the idea}
-
-Learning (i.e., fitting)
-of the the entire model parameters can effictively be achieved via \textit{backprop},
+\textbf{XXX perhaps add one sentence to rephrase and detail the idea}. \textbf{XXX Elvis: My biology is very poor.
+  The idea of the statement was to say where the parameters $\theta^Q$ would ``live'' in the brain.
+  Would these be synaptic strengths ? effective functional connectivity ?
+If you know this, please could you rephrase ?}
+
+Learning
+of the the entire model parameters can effectively be achieved via \textit{backprop},
 wherein prediction errors (i.e., \textit{regrets})  are propagated from lower to higher regions to modulate the choice of future actions
@@
 
-\begin{mdframed}
-\textbf{A note on classical RL algorithms.}
-Note that most RL algorithms, including \textit{Temporal Difference (TD)}
+\subsubsection{Link to classical reinforcement learning algorithms} One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
 learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
@@
 $$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
-%% Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
-%% $$\theta^Q(s,a) \leftarrow \theta^Q(s,a) + (Q(s, a|\theta^Q_{k+1}) - y)$$ ??
-\end{mdframed}
+Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
+$$\theta^Q(s,a) \leftarrow \theta^Q(s,a) + (Q(s, a|\theta^Q_{k+1}) - y),$$
+the well-known TD update rule \citep{sutton1998reinforcement}.
 
@@
 observations and rewards \citep{aronov2017}.
-These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
+% These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
 This would be a simple control algorithm by evaluating all legal actions and selecting the action with
-\textbf{highest expected cumulative rewards [please double-check]}.
-In these challenging POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
-Much recent work points to Monte Carlo or stochastic sampling– based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
-For sampling-based methods it is still almost non-existent beyond the description of receptive fields.
-
-\paragraph{Hippocampal replay corresponds to inverse reinforcement-learning?}
+% \textbf{highest expected cumulative rewards [please double-check]}.
+In POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
+Much recent work points to Monte Carlo or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
+\textbf{XXX Elvis: refs needed in last sentence.} For sampling-based methods it is still almost non-existent beyond the description of receptive fields.
+\textbf{XXX Elvis: rephrase last sentence}
+
+\paragraph{Hippocampal replay equals inverse reinforcement-learning?}
 %\begin{mdframed}
@@
     environmental perception and internally generated mind-wandering\footnote{
-      \textbf{XXX (danilo) This is more than just a footnote. Make this part of the paper
+      \textbf{XXX Elvis: This is more than just a footnote. Make this part of the paper
       and factorize the lot. Also you probably need a citation here!!!}
@@
 
-\textbf{XXX (danilo) Your focus in this subsection ought to be on the SC itself not POMPDs. -> agree I will fix it later.}
+\textbf{XXX Elvis: Your focus in this subsection ought to be on the SC itself not POMPDs. -> agree I will fix it later.}
  \begin{itemize}
@@
     There appears to be a cool link between cognition and formal theory
-    here.}
+    here. Elvis: I'd've helped you correct it but, i don't quite understand what you wanted to say. For ex, note that $\sum_{a \in \mathcal A}\pi(a|s) = 1$ for any state $s$,
+  and so it doesn't quite make sense to talk about weighting things with ``$\sum_{a \in \mathcal A}\pi(a|s)$''.}
     In contrast,

### 2017-04-01 paper/paper.tex
stuff

@@
 environmental assessment and prediction to guide action choices.
-DMN function is recast in mathematical terms
-from the perspective of
-% control theory and
-reinforcement learning.
+DMN function is recast in the mathematical terms of
+reinforcement learning and control by means of Markov decision processes.
 We argue that our formal account of DMN function
@@
 %  and Parr et al. 2003 (JMLR)].
-A recently proposed breakthrough alternative
-\citep{mnih2015,silver2016mastering}, is to learn this
-representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which
-are by far the current state-of-the-art in RL. In such a case, $\theta^Q$
-would correspond to the strengths of low-level synaptic or high-level functional connections between different elements
-in brain networks.
-\textbf{XXX perhaps add one sentence to rephrase and detail the idea}. \textbf{XXX Elvis: My biology is very poor.
-  The idea of the statement was to say where the parameters $\theta^Q$ would ``live'' in the brain.
-  Would these be synaptic strengths ? effective functional connectivity ?
-If you know this, please could you rephrase ?}
+A recently proposed, practically successful alternative
+\citep{mnih2015,silver2016mastering} is to learn this
+representation using a deep neural network, leading to the so-called \textit{Deep Q-learning} family of methods which
+are by far the current state-of-the-art in RL.
+The set of $\theta^Q$ parameters that instatiate the nonlinear interactions
+between layers of the artifical neural network
+may find a neurobiological correspondence in the adaptive strengths of axonal
+connections between neurons from the different levels
+of the neural processing hierarchy
+\citep{mesulam1998sensation, taylor2015global}.
 

### 2017-04-01 paper/paper.tex
unbelievably awesome links between biological properties

@@
 \section{Known Neurobiological Properties of the Default Mode Network}
-\textbf{XXX Elvis: this section needs more tight integration of the parts, and no just a separate independent subsection for each portion of the DMN...}
 We begin by a deconstruction the overall DMN function
@@
 \subsection{The posteromedial cortex: Global monitoring and integration}
-The posteriormedial cortex (PMC) and
-the dorsomedial prefrontal cortex (dmPFC) within the DMN
+Among all parts of the human brain,
+the midline structures of the DMN,
+including the posteromedial cortex (PMC) and
+the medial prefrontal cortex (mPFC),
 are responsible for the highest metabolic turn-over
-of glucose energy consumption in humans \citep{raichle2001pnas}.
+of glucose energy consumption \citep{raichle2001pnas}.
+These metabolic characteristics go hand-in-hand with
+brain-imaging analyses \citep{andrews2010} in
+suggesting that the PMC and dmPFC
+potentially represent the functional backbone of the DMN.
+
 The PMC myelinates relatively late during postnatal development in monkeys
 \citep{goldman1987development}, generally considered to
-be a sign of phylogenetic sophistication (Flechsig, 1920).
+be a sign of phylogenetic sophistication \citep{flechsig1920}.
 Physiological and disturbed metabolic fluctuations in the
 human PMC have been repeatedly related to
-phenomena of changed conscious awareness,
-including anesthesia,
-sleep, and forms of coma \citep{cavanna2006precuneus}.
-%
-The PMC has long been speculated to reflect constant computation of
+phenomena of changed conscious awareness \citep{cavanna2006precuneus},
+including anesthesia, sleep, and coma.
+%
+Indeed,
+the PMC has long been speculated to reflect constant computation of
 the environmental statistics and its internal representation
-as an "inner minds eye" \citep{cavanna2006precuneus}.
+as an "inner minds eye" \citep{cavanna2006precuneus, leech_pcc2014}.
 For instance, B\'alint's syndrome is a neurological disorder of conscious
@@
 for biological significance assessment.
-%
-Indeed,
-electrophysiological recordings in animals implicated the PMC in
-strategic selection \citep{pearson2009neurons},
+
+This putative role of the PMC in continuous integration of environmental relevance
+and ensuing top-level control of action in the environment is supported
+by a number of neuroscience experiments.
+Electrophysiological recordings in animals implicated the PMC in
+strategic decision making \citep{pearson2009neurons},
 risk assessment \citep{mccoy2005risk},
@@
 behavioral strategy in a complex food foraging task \citep{pearson2009neurons}.
-Single-cell recordings in the monkey PMC
+Further, single-cell recordings in the monkey PMC
 demonstrated this brain region's sensitivity to
@@
 across individual decision-making instances \citep{pearson2009neurons}.
-
 This DMN node encoded the
-preference or aversion to options with uncertain reward outcomes
+preference of or aversion to options with uncertain reward outcomes
 and its spiking activity was more associated with
@@
 distinct from spike firings that indicated choosing either option.
-Also in humans,
+Similarly in humans,
 the DMN has been shown to gather and integrate information
@@
 \citep{epstein2008parahippocampal, burgess2008spatial, valiquette2007different}.
-%
-The PMC may consequently monitor the subjective outcomes
+
+Consequently, the PMC may monitor the subjective outcomes
 of possible decisions and integrates that information
@@
 \subsection{The prefrontal cortex: Stimulus-value association and action consideration}
-The dmPFC subserves predominantly ambiguous amodal processes
+Analogous to the PMC,
+the dorsomedial PFC (dmPFC) is believed to subserve predominantly ambiguous processes
 across time, space, and content domains in
 sensory-independent top-down pathways.
-This DMN node has been described as a “mental sketchpad”
-\citep{goldman1996prefrontal},
-potentially implicated in de-novo generation and binding
+In contrast to the PMC, however,
+the dmPFC may be the DMN node whose function comes closest to a “mental sketchpad”
+\citep{goldman1996prefrontal}, as it 
+potentially subserves the de-novo generation and binding
 of meaning representations instructed by stored semantics and memories.
-%
-The dmPFC may thus enable inference, representation, and assessment
-of one's own and other individuals' action and thoughts.
+
 Generally,
@@
 \citep{stuss1986frontal}.
+The dmPFC may subserve inference, representation, and assessment
+of one's own and other individuals' action considerations.
 Specifically, neural activity in human dmPFC
@@
 with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
-Also the topographically neighboring dorsal anterior cingulate cortex
-has been linked to computing values and efforts of
+In fact, the topographically neighboring dorsal anterior cingulate cortex
+has also been linked to computing values and efforts of
 persisting a behavioral plan versus switching the
@@
 
-The ventromedial prefrontal cortex (vmPFC) subserves
+In contrast to the dorsomedial prefrontal DMN,
+the ventromedial portion of the prefrontal cortex (vmPFC) probably subserves
 less ambiguous subjective-value-related evaluative processes and
@@
 \citep{glascher2012lesion}.
-The vmPFC is preferentially connected with limbic and reward-related areas.
+Indeed,
+the vmPFC is preferentially connected with limbic and reward-related areas.
 The vmPFC has been observed to have monosynaptical connections
-with the nucleus accumbens
+with the NAc
 in axonal tracing studies in monkeys \citep{haber1995orbital}.
-Additionally, the gray-matter volume of the vmPFC and nucleus accumbens
-correlated with indices of value-guided behavior and social reward attitudes
+Additionally, the gray-matter volume of the vmPFC and NAc
+correlated with indices of value-guided behavior and reward attitudes
 \citep{lebreton2009automatic}.
@@
 Two functional connectivity modalities in humans strongly connected
-the vmPFC with the nucleus accumbens, hippocampus (HC),
+the vmPFC with the NAc, hippocampus (HC),
 and posteromedial cortex.
@@
 \subsection{The hippocampus: Memory, space, and experience replay}
+The core structures of the DMN midline are probably in constant communication
+with the HC in the medial temporal lobe \citep{vincet2006}.
 The HC is well known to be involved in memory and spatial navigation in animals and humans.
+
 Its highly recursive anatomical architecture
@@
 %
-While the HC in the medial temporal lobe system
+While the HC
 is traditionally believed to allow remembering the past,
@@
 \citep{hassabis2007patients}.
-%
+
 The HC thus orchestrates elements of experienced environmental aspects for
@@
 \subsection{The right and left TPJ: Prediction error signaling and world semantics}
-Finally,
-the right and left TPJ are known to have hemispheric differences
+While the DMN emerges with its midline structures early in development,
+the right and left TPJ may become fully integrated into this canonical
+network only after birth \citep{doria2010}.
+The TPJs are known to have hemispheric differences
 according to their cytoarchitectonic borders and gyrification pattern
@@
 attention versus language functions.
-%
-The anterior RTPJ is a key node of the DMN that  is central for
+
+The TPJ in the right hemipshere (RTPJ) is central for
 action initiation during externally structured tasks and
@@
 
-The left TPJ in turn exhibits a close topographical relationship to
-Wernicke's area
+The TPJ in the left hemisphere (LTPJ) in turn exhibits a close topographical relationship to Wernicke's area
 involved in the comprehension or understanding of written and spoken language.
@@
   The amygdala is known to be involved in significance evaluation while
-  the nucleus accumbens (NAc), which is connected to the vmPFC, is involved in reward evaluation.
+  the NAc, which is connected to the vmPFC, is involved in reward evaluation.
 \end{mdframed}

### 2017-04-01 paper/paper.tex
addon

@@
 \subsection{The right and left TPJ: Prediction error signaling and world semantics}
-While the DMN emerges with its midline structures early in development,
+While the DMN emerges with its midline structures early in human development,
 the right and left TPJ may become fully integrated into this canonical

### 2017-04-01 paper/paper.tex
improve refs for cognitive maps

@@
 the current task state even when states are unobservable from sensory input,
-which was shown to be critical for behavior.
+which was shown to be critical for behavior \citep{Schuck2016}.
 Additionally,
@@
 in constructing mental models in general
-\citep{maguire2016, schacter2007remembering, gelbard2008internally}.
+\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017}.
 Indeed,
@@
 consolidations based on reenactment and for integration into
-rich mental scene construction. In this way, the HC may even influence
+rich mental scene construction \citep{deuker2016event, bird2010establishing}.
+In this way, the HC may even influence
 ongoing perception in the current environment
@@
 multi-step action execution \citep{hartmann2005takes},
-visuo-proprioceptive conflict (Balslev et al., 2005) and
+visuo-proprioceptive conflict \citep{Balslev2005} and
 multi-modal detection of sensory changes across

### 2017-04-02 paper/paper.tex
pimped abstract

@@
 \begin{abstract}
-The default mode network (DMN) is the brain manifestation of the human
+The default mode network (DMN) is believed to subserve the human
 baseline mental activity.
 %
-Many research streams agree on its likely role in evolutionarily adaptive
-envisioning of mental scenarios to predict the environment.
-The DMN would hence be dedicated to continuous autobiographical memory retrieval,
+Many research streams converge to its likely role in evolutionarily adaptive
+envisioning of mental scenarios to simulate the future.
+The DMN may be dedicated to continuous autobiographical memory retrieval,
 generation of hypothetical outcomes,
 and reward contingency evaluation when letting the mind go.
-It explains its highest energy consumption in the brain and
+Such functional roles offer explanations of
+its highest energy consumption in the brain and
 its intimate coupling with conscious awareness.
@@
 \textit{how} the DMN may actually implement continuous
-environmental assessment and prediction to guide action choices.
+assessment and prediction of the environment to guide action choices.
 DMN function is recast in the mathematical terms of
-reinforcement learning and control by means of Markov decision processes.
+control theory and reinforcement learning based on Markov decision processes.
 We argue that our formal account of DMN function
-explains existing cognitive accounts as special cases and
-parsimoneously explains a variety of experimental findings
-in animals and humans.
-%
-Formal models of the computational processes subserved in the DMN
+naturally accommodates previously proposed cognitive accounts as special cases and
+parsimoneously explains various experimental findings in animals and humans.
+%
+Formal models of the neural computations realized in the DMN
 could offer access routes to the statistical regularities

### 2017-04-02 paper/paper.tex
improve beginning

@@
 in psychology \citep{james1890principles}
-found its neurobiological manifestation
+found a potential neurobiological manifestation
 in the so-called "default mode network" (DMN)
@@
 unfocused everyday mind wandering \citep{raichle2001pnas}.
-Indeed, the human brain features "a hierarchy of brain systems with
+The DMN may exert higher-order control on behavior
+as the human brain was argued to features "a hierarchy of brain systems with
 the DMN at the top and the salience and dorsal attention systems
@@
 cortex" \citep{carhart2010default}.
-In the beginning of the 21st century, brain imaging was hence the first technology
+In the beginning of the 21st century,
+brain imaging may have been first technique
 to allow for the discovery of a unique brain network that subserves
 baseline mental activities
-\citep{bzdok2015resting}.
+\citep{raichle2001pnas, bzdok2015resting}.
 
@@
 been argued to mediate an evolutionarily conserved function for the individual.
-This is all the more likely because the DMN contains the two
-biggest hotspots of energy consumption in the entire central nervous system
+This is all the more likely because the DMN contains hotspots
+of highest energy consumption in the entire central nervous system
 \citep{raichle2001pnas}.
@@
 that are calibrated by prediction error.
-
 A DMN framework incorporating RL can naturally embed human behavior
@@
 \subsection{Sketch of our main contributions}
-Our theory proposes that the implication of the DMN in a
+We argue that the implication of the DMN in a
 diversity of humans' most advanced cognitive processes

### 2017-04-02 paper/paper.tex
pass over biology half

@@
 Even (necessarily imperfect) memories
-of past experience, random mind-wandering and dreams during sleep
+of past experience, random day-time mind-wandering and dreams during sleep
 may provide meaningful building blocks to iteratively improve
@@
 
-\section{Known Neurobiological Properties of the Default Mode Network}
+\section{Known neurobiological properties of the default mode network}
 We begin by a deconstruction the overall DMN function
@@
 %
-\subsection{The posteromedial cortex: Global monitoring and integration}
+\subsection{The posteromedial cortex: Global monitoring and information integration}
 Among all parts of the human brain,
@@
 The PMC myelinates relatively late during postnatal development in monkeys
-\citep{goldman1987development}, generally considered to
+\citep{goldman1987development}, which is generally considered to
 be a sign of phylogenetic sophistication \citep{flechsig1920}.
@@
 the human PMC was coupled in two functional connectivity modalities
-with the amygdala involved in significance evaluation and
+with the amygdala, involved in significance evaluation, and
 the nucleus accumbens (NAc), involved in reward evaluation.
@@
 
-This putative role of the PMC in continuous integration of environmental relevance
+The putative role of the PMC in continuous integration of environmental relevance
 and ensuing top-level control of action in the environment is supported
@@
 and evaluation of reward contingencies by integrating these with
-information from memory and altered perspective frames.
+information from memory and different perspective frames.
 Regarding memory retrieval, retrosplenial lesions have been
@@
 
-Consequently, the PMC may monitor the subjective outcomes
+Consequently, the PMC may contribute to overall DMN function
+by monitoring the subjective outcomes
 of possible decisions and integrates that information
@@
 In contrast to the PMC, however,
-the dmPFC may be the DMN node whose function comes closest to a “mental sketchpad”
-\citep{goldman1996prefrontal}, as it 
+the dmPFC may be the DMN node whose function comes closest to a
+``mental sketchpad'' \citep{goldman1996prefrontal}, as it 
 potentially subserves the de-novo generation and binding
@@
 of one's own and other individuals' action considerations.
-Specifically, neural activity in human dmPFC
+Specifically, neural activity in the human dmPFC
 reflected expectations about other peoples actions and errors thereof.
@@
 This DMN node is more closely associated with
-orchestrating adapted behavior by bottom-up-driven
+orchestrating adaptive behavior by bottom-up-driven
 processing of “what matters now”,
@@
 Indeed,
-the vmPFC is preferentially connected with limbic and reward-related areas.
+the vmPFC is preferentially connected with reward-related and limbic areas.
 The vmPFC has been observed to have monosynaptical connections
@@
 subserve RL in humans more broadly \citep{doherty2015structure}.
-This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys
+This is consistent with diffusion MRI tractography in humans and monkeys
 \citep{croxson2005quantitative} that
@@
 was related to state encoding in vmPFC.
-Such a \textit{cognitive map} of action space was argued to encode
+Such a ``cognitive map'' of action space was argued to encode
 the current task state even when states are unobservable from sensory input,
@@
 independent whole-brain analyses from structural
-neuroimaging studies related the gray-matter volume GMV of the vmPFC,
+neuroimaging studies related the gray-matter volume of the vmPFC,
 more consistently than any other
 brain region, to indices of
-social competence and social network complexity,
+social competence and social network complexity, which are
 among the most complicated decision that humans and monkeys take
@@
 \subsection{The hippocampus: Memory, space, and experience replay}
-The core structures of the DMN midline are probably in constant communication
+The midline core of the DMN is probably in constant communication
 with the HC in the medial temporal lobe \citep{vincet2006}.
-The HC is well known to be involved in memory and spatial navigation in animals and humans.
+The HC is long known to be involved in memory and spatial navigation in animals and humans.
 
@@
 not only associated with an impairment in reexperiencing the past (i.e., amnesia),
-but also thinking about one’s own future and
+but also thinking about one's own future and
 imagining new fictitious experiences more broadly \citep{hassabis2007patients}.
@@
 were shown to exhibit increased grey matter volume in the
-posterior hippocampus \citep{maguire2000navigation}.
+hippocampus \citep{maguire2000navigation}.
 Indeed, place-specific neuronal populations in the hippocampus
@@
 paths \citep{johnson2007neural}; and
-these neuronal patterns apear to also indicate upcoming behavior
+these neuronal patterns appear to also indicate upcoming behavior
 \citep{pfeiffer2013hippocampal}.
@@
 to the neocortex for long-term storage, potentially also in the nodes of the DMN.
-These mechanisms of the hippocampus probably make important contributions to the
+These hippocampus-subserved mechanisms
+probably make important contributions to the
 recollection of autobiographical memory episodes and other
@@
 and neuropsychology,
-the LTPJ probably mediates access to and integrating of world knowlege
+the LTPJ probably mediates access to and integration of world knowlege
 to model action concepts
@@
 others' pantomimed action towards objects
-without obvious relation to processing any language content
+without obvious relation to processing explicit language content
 \citep{varney1987locus}
@@
 %
-Also inner speech hinges on knowledge retrieval of statistical structure
+Inner speech also hinges on knowledge retrieval of statistical structure
 about the physical and inter-personal world.
@@
 in internally generated visual scenarios that
-navigate present action, weigh hypothetical possibilities, and forcast the future.
+navigate present action, weigh hypothetical possibilities, and forecast the future.
 %
@@
 The proposed machine-learning perspective
-on DMN function is hence not only practically computable but
+on DMN biology is hence not only practically computable but
 also neuroscientifically falsifiable.
@@
 describe and interpret experimental findings in neuroscience studies
-on the default mode network.
+on the DMN.
 %
@@
 from the relatively recent events to anticipated upcoming events
-in order to constantly improve present action in the dynamic world.
+in order to constantly improve present action in our dynamic world.
 

### 2017-04-02 paper/paper.tex
monsterissimo pass

@@
 \section{Reinforcement learning: A process model for DMN function}
-The reviewed neurobiological properties and likely functions of the DMN
-are now argued to be sufficient
-for implementing all the major components 
+The outlined neurobiological properties and suspected functions of the DMN
+are argued to be sufficient
+for implementing all components 
 of a full-fledged \textit{reinforcement learning (RL)} algorithm.
-Memory, reward signaling,
-information gathering, random sampling, and other components
-of intelligent RL agents appear to cross-talk in the DMN in a plausible way.
+Integration of past experience, probabilistic mapping of candidate actions,
+random sampling of possible scenarios as well as
+incorporation of instantaneous, delayed, and expected reward outcomes
+are key components of intelligent RL agents
+that are plausible to cross-talk in the DMN.
 
@@
 in an iterative trial-by-error fashion.
-At any moment, each taken \textit{action} $a$ triggers a changes in the
+At a given moment, each taken \textit{action} $a$ triggers a changes in the
 \textit{state} of the environment
 $s \rightarrow s'$, accompanied by environmental feed signals as \textit{reward}
-$r = r(s, a)$ (or regret!, since it can be negative) collected by the
+$r = r(s, a)$ (or regret, since it can be negative) collected by the
 agent.
 In this view, the environment is partially controlled by
-the action of the agent and the reward should be thought
+the action of the agent and the reward can be thought
 of as satisfaction -- or punishment -- accompanying the execution of
@@
 Action may be delayed to achieve substantial increases in expected reward
-that can increase over time.
+that can grow with time.
 The environment is generally taken as stochastic,
-that is, unpredictable, changing in random ways. In addition,
+that is, changing in random ways. In addition,
 it is only \textit{partially observed},
@@
 the agent \citep{starkweather2017dopamine}.
-This unpredictability of the environment
-is realistic in a model which sets out
+The unpredictability of the environment
+is realistic in a computational model which sets out
 to explain DMN function. Indeed, it is known that the more the external world is predictable,
@@
 particularly advanced cognitive processes
-can be parsimoneously explained as prediction error minimization
-coupled with probabilistic simulations in form of mental scenes,
-thus calibrating action portfolios and
-maximizing reward outcome at various time scales.
+can be parsimoneously explained as probabilistic simulations in form of mental scenes coupled with prediction error minimization
+to calibrate action portfolios and
+maximize reward outcome at different time scales.
 Such a purposeful optimization objective
@@
 any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
-step is sufficiently short, defining $s(t)$, the state at time $t$, as the position in phase space.
-If the process has $memory$ such that the next state depends not only on the current state but also on some finite
+step is sufficiently short, defining $s(t)$, the state at time $t$, as the position in phase space \textbf{Replace by "state space"?}.
+The process has $memory$ if the next state depends not only on the current state but also on some finite
 number of past states.
@@
 In artificial intelligence (AI), a popular computational model for
-multi-step decision processes in such an environment are so-called
+multi-step decision processes in such an environment are
 \textit{Partially Observable Markov Decision Processes (POMDPs)}.
-A POMDP formalizes a decision process in which it is assumed that environment dynamics are determined by a Markov process,
+A POMDP formalizes a sequential decision process in which it is assumed that environment dynamics are determined by a Markov process,
 but the agent cannot directly observe the underlying state.
@@
 states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
-and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.
-
-Model-free RL is plausible in the human brain. Indeed, as was noted in \cite{dayan2008decision}, it has long been suggested in the literature that there
+and is characteristic of so-called \textit{model-free RL},
+which we propose as a useful formal framework for DMN function.
+
+Model-free RL is plausible in the human brain. Indeed,
+it has long been proposed \citep{dayan2008decision} that there
 is a rather direct mapping of model-free RL learning algorithms
@@
 dorsolateral striatum.
-
 In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of world.
@@
 quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
-cumulative reward estimate which takes into account the future. A standard for
+cumulative reward estimate which takes into account expected reward in the future.
+A common approach to
 modeling this accumulation is the time-discounted \textit{cumulative reward}% function
@@
 \end{mdframed}
-  The  goal of RL is then to adapt this policy so as to  maximize $G^\pi$. In
+  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$. In
   the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
-Setting $\gamma = 0$ corresponds to a perfectly myopic agent who is solely concerned about its
-immediate rewards.
+Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who is solely concerned about its immediate rewards.
 Such an agent would have no horizon for the future,
@@
 % \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
-Such an agent is thus better emphasizing consideration of rewards expected in
-a short future and paying less attention to
+Such an agent puts relatively more emphasis on rewards expected in
+a short future and pays less attention to
 rewards expected at a later time.
 % given that there is more uncertainty in the farsighted future.
-The reward discount factor $\gamma$ can then be seen as calibrating
+The reward discount factor $\gamma$ can be seen as calibrating
 the delay discounting behavior of the intelligent agent,
@@
 \subsection{The Q-value: A goodness measure for action choices}
-The purpose of the DMN can then be
-formulated as finding a policy $\pi$ for the agent which maximizes the
+We can now formulate the purpose of the DMN
+as finding a policy $\pi$ for the human agent that maximizes the
 expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
@@
 $s$, chooses action $a$,
-and then follows the policy $\pi$ thereafter to select future actions.
+and then follows the policy $\pi$ to select future actions.
 %% Given that each $\pi$ correspondence to an exhaustive description of
@@
 offers a formalization of optimal behavior,
-potentially driven by the DMN.
-
-\subsection{Computing optimal behavior}
+potentially driven by the DMN in human agents.
+
+\subsection{Efficient computation of optimal behavior}
 By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general.
@@
 % \end{mdframed}
-One popular solution concept is called \text{Q-learning}. It remedies
+One popular solution is called \text{Q-learning}. It remedies
 the issue by only considering the sub-family of
@@
 
-\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992), generates a
+\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992) generates a
 \textit{greedy policy}
+\textbf{XXX please add 1-2 sentences why this is interesting in the current
+context. This statement appears a little alone as it stands.}
 \begin{equation}
@@
 \end{equation}
-This \textit{fixed-point} equation provides a recursive decomposition of optimal behavior by dividing the value function into the immediate reward
+This \textit{fixed-point} equation provides a recursive decomposition of optimal behavior by dividing the value function \textbf{XXX please more explicitly introduce
+the notion of value function}
+into the immediate reward
 and the discounted rewards of the upcoming states.
@@
 % It is a fixed-point equation for the deterministic policy $\pi$.
-Notice that in \eqref{eq:bellman}, the sampling is done only over things which
+Notice that in \eqref{eq:bellman} the sampling is done only over things which
 depend on the environment, and so can be learned off-policy by observing state transitions
@@
 appropriate form of the state-action pairs $(s,a)$,
-as proposed by \cite{songNIPS2016}.
+as proposed by \citep{songNIPS2016}.
 %  and Parr et al. 2003 (JMLR)].
-A recently proposed, practically successful alternative
+A recently proposed, practically successful alternative approach
 \citep{mnih2015,silver2016mastering} is to learn this
 representation using a deep neural network, leading to the so-called \textit{Deep Q-learning} family of methods which
-are by far the current state-of-the-art in RL.
-The set of $\theta^Q$ parameters that instatiate the nonlinear interactions
+likely are state-of-the-art in RL.
+The set of $\theta^Q$ parameters that instantiate the nonlinear interactions
 between layers of the artifical neural network
@@
 Learning
-of the the entire model parameters can effectively be achieved via \textit{backprop},
-wherein prediction errors (i.e., \textit{regrets})  are propagated from lower to higher regions to modulate the choice of future actions
+of the the entire model parameters can effectively be achieved
+via \textit{backpropagation},
+where prediction errors (i.e., \textit{regrets}) percolate from lower to higher processing layers to modulate the choice of future actions
 % online stochastic gradient descent is used to minimize the variational objective
@@
 learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
-cast in this general variational framework. For instance, TD corresponds to
+cast in this \textbf{please unpack the "this" here}
+general variational framework. For instance, TD corresponds to
 the above framework using a linear value approximator with feature encoding
@@
 observations and rewards \citep{aronov2017}.
-% These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
+These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
 This would be a simple control algorithm by evaluating all legal actions and selecting the action with
-% \textbf{highest expected cumulative rewards [please double-check]}.
+highest expected cumulative rewards.
 In POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
@@
 
-\paragraph{Hippocampal replay equals inverse reinforcement-learning?}
+\paragraph{Does hippocampal replay equal inverse reinforcement-learning?}
 %\begin{mdframed}
@@
   (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
-  For instance, given traces of motor actions from a an adult to grab a cup on a table,
+  For instance, given traces of motor actions from an adult to grab a cup on a table,
   can an observing child figure out what ``langragian'' functional is being minimized by the former?
@@
 In the general artificial case, the problem has been extensively studied and partially solved by
-\cite{abbeel2004}. They are also been rigourously studied in general optimal control literature under the name
+\cite{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
 ``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
@@
   action and state cascades as a function of what happened in the past.
-% A policy represents the repertoire of possible actions
-%   on the world given a current state. It encodes the probabilites of
-%   choosing actions to be executed in a certain situation.
+  A policy represents the repertoire of possible actions
+    on the world given a current state. It encodes the probabilites of
+    choosing actions to be executed in a certain situation.
 % Interactions between external and internal sources of
@@
   constant exploration of possible future actions and their
-  cumulative outcomes. Implicit computation of future choices
+  cumulative reward outcomes. Implicit computation of future choices
   provides an explanation for the
@@
 The HC may provide perturbed action-transition-state-reward samples
-  as a block of "imagined", "hypothesized", "recalled" experience.
-  The small variations in these experience blocks allow searching
+  as batches of "imagined", "hypothesized", "recalled" experience.
+  The small variations in these experience samplings allow searching
   a larger space of model parameters and possible experiences.
@@
   only very rarely or never encounter.
+  An explanation is thus offered for encountering seemingly familiar situations that
+  a human has however never actually encountered (i.e., d\'{e}j\`{a} vu effect).
+  While such a situation may not have been experienced in the physical world,
+  the DMN may have already stochastically generated, evaluated, and adapted to
+  such a randomly synthesized situation in the past.
   In the absence of environmental input and feedback
@@
   future scenarios and action outcomes.
-  Our approach acknowledges the unavoidable stochasticity of
+  Our approach thus acknowledges the unavoidable stochasticity of
   computation in neuronal processes \citep{faisal2008noise}.
@@
   to novel, yet unexperienced environments, reward prediction errors will
-  massively increase. This will lead to adaptation of the policy until the system converges to a 
+  massively increase \textbf{Please complete sentence}.
+  This will lead to adaptation of the policy until the system converges to a 
   new steady-state of optimal action decisions in volatile environments.

### 2017-04-03 paper/paper.tex
averseness
equivalence table
First test with \suggestremove

@@
 
-\title{A Computational Account of Default Mode Function\\
-  by
-  Control Theory and
-  Reinforcement Learning}
+\title{Dark Control:\\
+       A Unified Account of Default Mode Function\\
+       by Control Theory and Reinforcement Learning}
 
@@
   = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
-    \underbrace{\nabla_{\theta^Q_{k}}Q(s, a|\theta^Q_{k+1})}_{\text{???}}\right].
+    \underbrace{\nabla_{\theta^Q_{k}}Q(s, a|\theta^Q_{k+1})}_{\text{averseness}}\right].
   \label{eq:oracle}
@@
   immediate future.
-  \item Mismatch negativity of PC is immanent in reward prediction error
+  \item \suggestremove{Mismatch negativity of PC is immanent in reward prediction error
   by the difference between actually predicted reward of an action given
   a state and the reward predicted by non-linear regression with graded
-  discounting of future rewards.
+  discounting of future rewards.}
   \item Adapting the neuronal connections for improved top-down modulation
@@
          Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
-         Student’s t, Logistic, Uniform, Triangular, ... \\ \hline
+         Student’s t, Logistic, Uniform, Triangular, \ldots \\ \hline
          Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
@@
 \subsection{GANs and other likelihood-free methods}
-...
+\ldots
 

### 2017-04-03 paper/paper.tex
minor fixups

--- 2017-04-02 paper/paper.tex
@@
   \begin{split}
-    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect rewards }R_0 = r(s_0, a_0)\\
+    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0)\\
     \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
-    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect rewards }R_1 = r(s_1, a_1)\\
+    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1)\\
     \ldots\\
     \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
-    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect rewards }R_{t} = r(s_{t}, a_{t})\\
+    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t})\\
     \ldots
@@
 via \textit{backpropagation},
-where prediction errors (i.e., \textit{regrets}) percolate from lower to higher processing layers to modulate the choice of future actions
+where reward prediction errors (RPES) -- also known as \textit{regret}-- percolate from lower to higher processing layers to modulate the choice of future actions
 % online stochastic gradient descent is used to minimize the variational objective
@@
   % = -\mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{}))^2]
-  = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
-    \underbrace{\nabla_{\theta^Q_{k}}Q(s, a|\theta^Q_{k+1})}_{\text{???}}\right].
+  = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret / RPE}}
+    \underbrace{\nabla_{\theta^Q_{k}}Q(s, a|\theta^Q_{k+1})}_{\text{averseness}}\right].
   \label{eq:oracle}
@@
 \citep{dayan1995helmholtz}. As noted already in \citep{dayan1995helmholtz},
-a crucial drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recogntiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
-Thus the brain could not possible be running such a algorithm, not even in principle!
-To the rescue, we note that the recent theory of
-\textit{variational auto-encoders (VAEs)} \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm, as it overcomes the technical limits of the former, by using a reparametrization trick.
-For instance, unlike the wake-sleep algorithm for minimizing free-energy, VAEs can be efficiently trained via back-propagation of learning errors.
+a crucial drawback of the wake-sleep algorithm is that it postulates a pair forward (generation) and backward (recognition) models,
+which together do not correspond to optimization of (a bound of) the marginal likelihood, because of the incorrect KL used therein, etc.
+Thus the brain could not possibly be running such a algorithm, not even in principle! To the rescue, we note that the recent theory of
+\textit{variational auto-encoders (VAEs)} \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm,
+as it overcomes the technical limits of the former, by using a reparametrization trick. For instance, unlike the wake-sleep algorithm
+for minimizing free-energy, VAEs can be efficiently trained via back-propagation of learning errors.
 

### 2017-04-04 paper/paper.tex
given some psychological meaning to gamma

@@
 any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
-step is sufficiently short, defining $s(t)$, the state at time $t$, as the position in phase space \textbf{Replace by "state space"?}.
+step is sufficiently short, defining $s(t)$, the state at time $t$, as the position in state space.
 The process has $memory$ if the next state depends not only on the current state but also on some finite
@@
 \end{eqnarray*}
+
+\begin{mdframed}
+  \textbf{Time-invariance assumption:} It is assumed that the MDP has reached steady-state conditions so that
+  the dynamics of the process are equivalent over sufficiently long time windows of equal length.
+\end{mdframed}
+
 Since an action taken in the present moment $t$
@@
 % \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
+% The reward discount factor
+$\gamma$ can be seen as calibrating risk-seeking valency of the intelligent agent,
+that is, the decision related to reward-delay trade-off schedules.
 Such an agent puts relatively more emphasis on rewards expected in
 a short future and pays less attention to
-rewards expected at a later time.
+rewards expected at a later time, precisely rewards more than $1/(1 - \gamma)$ time steps into the future are disregarded.
+This reduces the variance of expected rewards to be accumulated across action cascades by limiting the length of this cascades.
 % given that there is more uncertainty in the farsighted future.
-The reward discount factor $\gamma$ can be seen as calibrating
-the delay discounting behavior of the intelligent agent,
-that is, the decision related to reward-delay trade-off schedules.
-It is also important to appreciate that a stochastic policy estimation
-is more advantageous in many RL settings.
-For instance, a deterministic strategy in playing rock-paper-scissors can
-be easily exploited and a uniformly random choice is optimal.
+% It is also important to appreciate that a stochastic policy estimation
+% is more advantageous in many RL settings.
+% For instance, a deterministic strategy in playing rock-paper-scissors can
+% be easily exploited and a uniformly random choice is optimal.
 
@@
 "What is the expected utility of taking action $a$ in this situation ?".
-When a choice of actions is available, the $Q^{\pi}(s,a)$
-offers a formalization of optimal behavior,
+$Q^{\pi}(s,a)$ offers a formalization of optimal behavior,
 potentially driven by the DMN in human agents.
+% The goal of the DMN is then
+% to find then to produce a policy $\pi$ which maximizes the expected cumulative reward $Q^\pi(s_0,\pi(a_0))$
+% given an initial state $s_0$.
 

### 2017-04-05 paper/paper.tex
fix: sentinel hypothesis

@@
 as an advantageous evolutionary adaptation.
-According to this cognitive account, the brain's baseline realizes
+According to this cognitive account, the brain\'s baseline realizes
 a ``radar'' function to
-detect subjectively salient and unexpected events that
-are the most important for unfolding behavior,
-including resources,
-the securing of mates, and protection.
-%
+detect subjectively salient object and unexpected events in the environment.
+The sentinel account does however not provide a formal mechanisms of
+how attention and memory resources are exactly reallocated when
+encountering a salient environmental stimulus.
+In our formal account,
+action cascades
+are recursively explored by human agents. Conversely,
+certain branches of candidate action trajectories
+are detected to be less worthy to get explored.
+This mechanism, mediated by the Bellman equation,
+directly implies allocation of attention and computational load
+to relevant objects and events in the environment.
+%
+Further,
+our computational account provides an explanation for the known implication
+of the DMN
+in both certain goal-directed experimental tasks and
+especially in the idlying rest cognitive set.
+Both cognitive sets may entail DMN recruitment
+if environmental relevance is processed, manipulated, and used for prediction.
+In tasks, the policy and value matrices are updated to optimize short-term action.
+At rest, these parameter updates may
+improve especially mid- and long-term action.
+We thus provide answers to the unsettled question why the involvement
+of the same neurobiological brain circuit, the DMN, has been documented
+for baseline house-keeping functions and specific task performance.
+
 In fact,
-environmental stimuli important for humans are frequently of
+environmental stimuli especially important for humans are frequently of
 social nature. This is probably unsurprising
@@
 \citep{tomasello2009cultural}.
-More specifically,
-according to the ``social brain hypothesis'',
+According to the ``social brain hypothesis'',
 the human brain did not evolve to solve problems of the
@@
 \citep{whiten1988machiavellian}.
-Highly efficient neurobiological processing of social cues is exemplified by
-facial judgments being routinely processed in less than 100 ms
-\citep{bar2006very}.
-Our human ancestors were probably among the few organisms that
-were more likely to be killed by members of their own species
-rather than predators from another species.
-Being able to detect the intention and upcoming actions of conspecifics
-turned into a decisive evolutionary advantage
-\citep{frith2010social}.
-%
 This may explain why social topics amount to roughly
@@
 \citep{schilbach2008minds}.
-
-\textbf{XXX Elvis: Your focus in this subsection ought to be on the SC itself not POMPDs. -> agree I will fix it later.}
- \begin{itemize}
-    \item Processing social information has been proposed to underlie the
+%
+Consequently, the DMN has been proposed to reflect
+continuous processing of social information as a
 physiological baseline of human brain function
-    \citep{schilbach2008minds}. This was later challenged by observing
+\citep{schilbach2008minds}. This view was later challenged by observing
 analogues of the DMN in monkeys \citep{mantini2011default}
-    and rats \citep{lu2012rat}, with
-    supposedly less advanced social capacities.
-    The MDPs parsimoneously explain the dominance of social content in
-    human mental scenes by their extremely high relevance to humans
-    reflected by high values for social information in the value matrix
+and rats \citep{lu2012rat}, species with
+supposedly less advanced social-cognitive capacities.
+Instead,
+our formal account can parsimoneously explain the dominance of social content in
+human mental scenes by their relevance to humans
+reflected in high estimates for social information obtained from the value function
 \citep{baker2009action}.
-    \item How can a same neurobiological circuit be equally important
-    for baseline house-keeping functions and specific task performance?
-    Our computational account can explain why the DMN is implicated
-    in both a goal-directed task and an idlying rest cognitive set,
-    if environmental relevance is processed, manipulated, and retained
-    also as a baseline function of human mental activity.
-    In tasks, the policy and value matrices are updated to optimize short-term action,
-    whereas, at rest, these parameter updates may
-    improve especially mid- and long-term action.
-    \item The sentinel account does not provide a formal mechanisms of
-    how attention and memory resources are exactly reallocated when
-    encountering a salient environmental stimulus. \textbf{XXX remove the next sentence. Danilo: can you help me to correct the sentence instead.
-    There appears to be a cool link between cognition and formal theory
-    here. Elvis: I'd've helped you correct it but, i don't quite understand what you wanted to say. For ex, note that $\sum_{a \in \mathcal A}\pi(a|s) = 1$ for any state $s$,
-  and so it doesn't quite make sense to talk about weighting things with ``$\sum_{a \in \mathcal A}\pi(a|s)$''.}
-    In contrast,
-    the Bellman recursion is weighted at each step by
-    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
-    are recursively explored by the agent and which decision trees are neglected.
-    This directly implies allocation of attention and computational load.
-    \item In sum,
-    MDPs imply a ``radar'' function of monitoring the environment for salient information
-    in general and explain social thinking as a physiological brain baseline
+In sum,
+our account on the DMN directly implies
+its previously proposed ``radar'' function
+of monitoring the environment for salient information
+in general and inherently incorporates the importance of social environmental cues
 as an important special case, without excluding other topics
 contemplated in continuous mental activity.
- \end{itemize}
 

### 2017-04-06 paper/paper.tex
fix: sentinel hypothesis 2.0

@@
 Since then,
-different investigators have speculated that neural activity in the DMN
-may reflect the brain’s relentless tracking of
+many investigators have speculated that neural activity in the DMN
+may reflect the brain's continuous tracking of
 relevance in the environment
 as an advantageous evolutionary adaptation.
-According to this cognitive account, the brain\'s baseline realizes
+According to this cognitive account, the human brain's baseline realizes
 a ``radar'' function to
-detect subjectively salient object and unexpected events in the environment.
-The sentinel account does however not provide a formal mechanisms of
+detect subjectively salient objects and unexpected events in the environment.
+Propositions of such a sentinel function have however seldom detailed
+the mechanisms of
 how attention and memory resources are exactly reallocated when
@@
 In our formal account,
-action cascades
+cascades of candidate actions
 are recursively explored by human agents. Conversely,
 certain branches of candidate action trajectories
-are detected to be less worthy to get explored.
-This mechanism, mediated by the Bellman equation,
-directly implies allocation of attention and computational load
+are detected to be less worthy to get mentally explored.
+This mechanism, expressed by the Bellman equation,
+directly implies stratifications over attention and working memory
 to relevant objects and events in the environment.
@@
 Further,
-our computational account provides an explanation for the known implication
+our account provides an explanation for the known implication
 of the DMN
-in both certain goal-directed experimental tasks and
-especially in the idlying rest cognitive set.
+in certain goal-directed experimental tasks and
+in the idlying task-unconstrained cognitive sets.
 Both cognitive sets may entail DMN recruitment
 if environmental relevance is processed, manipulated, and used for prediction.
-In tasks, the policy and value matrices are updated to optimize short-term action.
+During tasks,
+the policy and value estimates are updated to optimize short-term action.
 At rest, these parameter updates may
@@
 of the same neurobiological brain circuit, the DMN, has been documented
-for baseline house-keeping functions and specific task performance.
+for baseline house-keeping functions and specific task performances.
 
@@
 are so rich in stories about people and
-the complex relationships between them
-\citep{schilbach2008minds}.
-%
-Consequently, the DMN has been proposed to reflect
+the complex relationships between them.
+%
+In line with this, the DMN has been proposed to reflect
 continuous processing of social information as a
@@
 Instead,
-our formal account can parsimoneously explain the dominance of social content in
-human mental scenes by their relevance to humans
-reflected in high estimates for social information obtained from the value function
+our formal account can parsimoneously incorporate
+the dominance of social content in
+human mental activity by reexpressing relevance to humans
+in high value function estimates for social information
 \citep{baker2009action}.
+The DMN may thus exert control on the human agent to prioritize
+appraisal of and action towards social contexts,
+without completely neglecting environmental cues of the physical world.
 In sum,

### 2017-04-06 paper/paper.tex
a small pass

--- 2017-04-04 paper/paper.tex
@@
 \begin{mdframed}
-  \textbf{Time-invariance assumption:} It is assumed that the MDP has reached steady-state conditions so that
-  the dynamics of the process are equivalent over sufficiently long time windows of equal length.
+  \textbf{Time-invariance assumption:} It is assumed that the underlying MDP is stationary --i.e has reached steady-state conditions--
+  so that the dynamics of the process are equivalent over sufficiently long time windows of equal length.
 \end{mdframed}
@@
 
-\subsection{The Q-value: A goodness measure for action choices}
+\subsection{The Q-values, state-values, and the advantage function}
 We can now formulate the purpose of the DMN
@@
   \label{eq:qval}
-  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a,a_1 \sim \pi,a_2 \sim \pi,\ldots].
+  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a].
 \end{equation}
@@
 potentially driven by the DMN in human agents.
+
+One can also define the state-value function
+
+\begin{eqnarray}
+  \V^\pi(s) := \mathbb E_{a \sim \pi}Q^\pi(s,a) = \mathbb E [G^\pi|s_0=s],
+\end{eqnarray}
+which measures the value of each state $s \in \mathcal S$ under the policy $\pi$. This serves
+as a baseline value for any action. Finally, the advantage function is defined by
+\begin{equation}
+A^\pi(s,a) :=  Q^\pi(s,a) - V^\pi(s).
+\end{equation}
+
 % The goal of the DMN is then
@@
 \begin{equation}
-  \pi(s) = \argmax_{a \in \mathcal A}{Q}(s, a).
+  \pi^*(s) = \argmax_{a \in \mathcal A}{Q}^*(s, a),
   \label{eq:qlearning}
 \end{equation}
+which maps each state $s$ to the optimal action which should be taken at that state.
+Here $Q^*(s, a) := \max_{\pi}Q^\pi(s,a)$, the optimal Q-value over all policies.
 % where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
@@
   \begin{split}
-    Q^\pi(s, a) &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
+    Q^*(s_0, a_0) &= \mathbb E_{s_1 \sim p(s_1|s_0,a)} [r(s,a) + \gamma \max_{a_1 \in \mathcal A}Q^*(s_1, a_1)]
     \\
-    &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))].
+    &= \mathbb E_{s_1 \sim p(s_1|s_0,a_0)} [r(s_0,a_0) + \gamma Q^*(s_1, \pi(s_1))].
   \end{split}
@@
 \end{equation}
-This \textit{fixed-point} equation provides a recursive decomposition of optimal behavior by dividing the value function \textbf{XXX please more explicitly introduce
-the notion of value function}
-into the immediate reward
-and the discounted rewards of the upcoming states.
-A complicated dynamic programming optimization can be
+This consistency equation for the optimal (over all policies) Q-value function provides
+a recursive decomposition of optimal behavior by dividing the Q-value function into the immediate
+reward and the discounted rewards of the upcoming states. The optimal Q-value operator $Q^*$
+is a fixed point for this equation. A complicated dynamic programming optimization can be
 broken into simpler sub-problems at different time points,

### 2017-04-06 paper/paper.tex
fixup
more

@@
 $${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
-$\phi(s,a)$ would represent a high-level hand-crafted representation in an
-appropriate form of the state-action pairs $(s,a)$,
-as proposed by \citep{songNIPS2016}.
+$\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action pairs
+$(s,a)$, as proposed by \citep{songNIPS2016}.
 %  and Parr et al. 2003 (JMLR)].
@@
 
-\begin{figure}[!tb]
-  \centering
-  \includegraphics[width=.8\linewidth]{vae.pdf}
-  \caption{Variational autoencoders...}
-  \label{fig:vae}
-\end{figure}
+% \begin{figure}[!tb]
+%   \centering
+%   \includegraphics[width=.8\linewidth]{vae.pdf}
+%   \caption{Variational autoencoders...}
+%   \label{fig:vae}
+% \end{figure}
 
@@
 The result is a scheme for training DBNs via good-old backprop!
-Refer to Fig. \ref{fig:vae}. Some examples of the reparametrization trick for a number of
+Refer to Fig. vae.pdf.
+% \ref{fig:vae}.
+Some examples of the reparametrization trick for a number of
 choices of the posterior distribution are given in Tab. \ref{tab:rptrick}.

### 2017-04-06 paper/paper.tex
Comments abstract.

--- 2017-04-06 paper/paper.tex
@@
 envisioning of mental scenarios to simulate the future.
-The DMN may be dedicated to continuous autobiographical memory retrieval,
+\suggestremove{The DMN may be dedicated to continuous autobiographical memory retrieval,
 generation of hypothetical outcomes,
-and reward contingency evaluation when letting the mind go.
+and reward contingency evaluation when letting the mind go.}
 Such functional roles offer explanations of
 its highest energy consumption in the brain and
-its intimate coupling with conscious awareness.
-%
-This concept paper proposes a process model that describes
+its intimate coupling with conscious awareness
+\textbf{(XXX Guillaume: It escaladed quickly.)}.
+%
+This concept paper proposes a process model \textbf{(XXX Guillaume: It's not clear what a "process model" is for most people?)} that describes
 \textit{how} the DMN may actually implement continuous
@@
 We argue that our formal account of DMN function
-naturally accommodates previously proposed cognitive accounts as special cases and
-parsimoneously explains various experimental findings in animals and humans.
+naturally accommodates \suggestremove{previously} proposed cognitive accounts as special cases and
+parsimoneously explains \suggestremove{various} experimental findings in \suggestadd{both} animals and humans.
 %
@@
 could offer access routes to the statistical regularities
-of human predictive behavior.
+of human predictive behavior \textbf{(XXX Guillaume: Maybe too wordy\ldots)}.
 \end{abstract}
@@
 \begin{mdframed}
-  \textbf{Time-invariance assumption:} It is assumed that the underlying MDP is stationary --i.e has reached steady-state conditions--
-  so that the dynamics of the process are equivalent over sufficiently long time windows of equal length.
+  \textbf{Time-invariance assumption:} It is assumed that the MDP has reached steady-state conditions so that
+  the dynamics of the process are equivalent over sufficiently long time windows of equal length.
 \end{mdframed}
@@
 
-\subsection{The Q-values, state-values, and the advantage function}
+\subsection{The Q-value: A goodness measure for action choices}
 We can now formulate the purpose of the DMN
@@
   \label{eq:qval}
-  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a].
+  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a,a_1 \sim \pi,a_2 \sim \pi,\ldots].
 \end{equation}
@@
 potentially driven by the DMN in human agents.
-
-One can also define the state-value function
-
-\begin{eqnarray}
-  \V^\pi(s) := \mathbb E_{a \sim \pi}Q^\pi(s,a) = \mathbb E [G^\pi|s_0=s],
-\end{eqnarray}
-which measures the value of each state $s \in \mathcal S$ under the policy $\pi$. This serves
-as a baseline value for any action. Finally, the advantage function is defined by
-\begin{equation}
-A^\pi(s,a) :=  Q^\pi(s,a) - V^\pi(s).
-\end{equation}
-
 % The goal of the DMN is then
@@
 \begin{equation}
-  \pi^*(s) = \argmax_{a \in \mathcal A}{Q}^*(s, a),
+  \pi(s) = \argmax_{a \in \mathcal A}{Q}(s, a).
   \label{eq:qlearning}
 \end{equation}
-which maps each state $s$ to the optimal action which should be taken at that state.
-Here $Q^*(s, a) := \max_{\pi}Q^\pi(s,a)$, the optimal Q-value over all policies.
 % where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
@@
   \begin{split}
-    Q^*(s_0, a_0) &= \mathbb E_{s_1 \sim p(s_1|s_0,a)} [r(s,a) + \gamma \max_{a_1 \in \mathcal A}Q^*(s_1, a_1)]
+    Q^\pi(s, a) &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
     \\
-    &= \mathbb E_{s_1 \sim p(s_1|s_0,a_0)} [r(s_0,a_0) + \gamma Q^*(s_1, \pi(s_1))].
+    &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))].
   \end{split}
@@
 \end{equation}
-This consistency equation for the optimal (over all policies) Q-value function provides
-a recursive decomposition of optimal behavior by dividing the Q-value function into the immediate
-reward and the discounted rewards of the upcoming states. The optimal Q-value operator $Q^*$
-is a fixed point for this equation. A complicated dynamic programming optimization can be
+This \textit{fixed-point} equation provides a recursive decomposition of optimal behavior by dividing the value function \textbf{XXX please more explicitly introduce
+the notion of value function}
+into the immediate reward
+and the discounted rewards of the upcoming states.
+A complicated dynamic programming optimization can be
 broken into simpler sub-problems at different time points,

### 2017-04-06 paper/paper.tex
Comments on the introduction

@@
 % \thickhline command for thick horizontal lines that span the table
-\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%@
 \hline
@@
 when mentally processing environmental challenges
-\citep{raichle2001pnas}.
+\citep{raichle2001pnas}. \textbf{(XXX Guillaume: What about the energy consumption compare to sleep?)}
 What has early been described as the "stream of consciousness"
@@
 
-\subsection{Default mode network: Higher-order control of the organism}
+\subsection{Default mode network: higher-order control of the organism}
 The DMN appeared to be exclusive in task-induced
@@
 This processing of unknown information categories in the DMN has
-been argued to mediate an evolutionarily conserved function for the individual.
+been argued to mediate an evolutionarily conserved function for the individual \textbf{(XXX Guillaume: Ref?)}.
 This is all the more likely because the DMN contains hotspots
@@
 Today, many authors believe that the DMN implements
-probabilistic estimation of past, hypothetical, and future events. This
+probabilistic estimation of past, hypothetical, and future events \textbf{(XXX Guillaume: Ref? // many authors)}. This
 brain network
@@
 maintaining baseline contemplations
-across time, splace, and content domains
-at the interface of the external world and the self.
+across time, space, and content domains
+at the interface of the external world and the self \textbf{(XXX Guillaume: This sentence is not clear and too long.)}.
 
@@
 Conversely,
-the more the currently executed task is unknown and unpracticed,
-the lower the tendency for stimulus-independent thoughts
+the more the ongoing executed task is unknown and unpracticed,
+the less stimulus-independent thoughts occur
 \citep{filler1973daydreaming, teasdale1995stimulus}.
-These ``offline'' processes may however contribute to optimized control of the organism.
+These ``offline'' processes may however contribute to optimized control of the organism. \textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment?)}
 % We formalize
@@
 the DMN dynamics are constantly adapted in feedback loops
-that are calibrated by prediction error.
+shaped by prediction error.
 A DMN framework incorporating RL can naturally embed human behavior
@@
 
-\subsection{Sketch of our main contributions}
+\subsection{Sketch of our main contributions\textbf{(XXX Guillaume: What about ``Towards a computational account of DMN function'')}}
 We argue that the implication of the DMN in a
@@
 based on pervasive probabilistic simulations,
-thus maximizing reward outcome at various time scales.
+thus maximizing reward outcome across multiple time scales.
 Such a purposeful optimization objective
@@
 Even (necessarily imperfect) memories
-of past experience, random day-time mind-wandering and dreams during sleep
+\suggestremove{of past experience}, random day-time mind-wandering, and dreams during sleep
 may provide meaningful building blocks to iteratively improve
@@
 \citep{raichle2005intrinsic}.
-The idea has been invigorated by empirical evidence from
-neuroscientific experiments \citep{kording2004bayesian, fiser2004small}.
+Numerous empirical evidence from
+neuroscientific experiments have invigorated this idea \citep{kording2004bayesian, fiser2004small}.
 This paper proposes a formal model that satisfies this contention.

### 2017-04-06 paper/paper.tex
New transition for section 2

@@
 \section{Known neurobiological properties of the default mode network}
-We begin by a deconstruction the overall DMN function
-by the mechanistic relevance of each individual network node
-based on published findings from neuroscience experiments.
+We begin by a deconstruction of the DMN based on experimental findings in neuroscience.
+This walkthrough across each individual network node will describe their mechanistic relevance, paving the way for their functional interpretation within our computational account in the next section.
+
 %

### 2017-04-06 paper/paper.tex
Add the introductory figure about the neurobiology of DMN

@@
 
+\begin{figure}[!h]
+  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
+  \caption{Neurobiological overview of the DMN ... \textbf{Please add a figure caption}}
+  \label{fig:neurobiological_overview_DMN}
+\end{figure}
 %
 \subsection{The posteromedial cortex: Global monitoring and information integration}
-Among all parts of the human brain,
+\suggestremove{Among all parts of the human brain,}
 the midline structures of the DMN,
@@
 brain-imaging analyses \citep{andrews2010} in
-suggesting that the PMC and dmPFC
+suggesting that the PMC and \suggestremove{d}mPFC
 potentially represent the functional backbone of the DMN.

### 2017-04-06 paper/paper.tex
Capitalization of titles Cf. https://en.wikipedia.org/wiki/Colon_(punctuation)#Use_of_capitals

@@
 %
-\subsection{The posteromedial cortex: Global monitoring and information integration}
+\subsection{The posteromedial cortex: global monitoring and information integration}
 \suggestremove{Among all parts of the human brain,}
@@
 
-\subsection{The prefrontal cortex: Stimulus-value association and action consideration}
+\subsection{The prefrontal cortex: stimulus-value association and action consideration}
 Analogous to the PMC,
@@
 
-\subsection{The hippocampus: Memory, space, and experience replay}
+\subsection{The hippocampus: memory, space, and experience replay}
 The midline core of the DMN is probably in constant communication
@@
 
-\subsection{The right and left TPJ: Prediction error signaling and world semantics}
+\subsection{The right and left TPJ: prediction error signaling and world semantics}
 While the DMN emerges with its midline structures early in human development,
@@
 
-\section{Reinforcement learning: A process model for DMN function}
+\section{Reinforcement learning: a process model for DMN function}
 The outlined neurobiological properties and suspected functions of the DMN
@@
 
-\subsubsection{Excursion: The role of the hippocampus}
+\subsubsection{Excursion: the role of the hippocampus}
 \paragraph{Does the hippocampus do Monte-Carlo sampling?}
@@
 
-\subsection{Predictive Coding Hypothesis}
+\subsection{Predictive coding hypothesis}
 The predictive coding framework
@@
 
-\subsection{Semantic Hypothesis}
+\subsection{Semantic hypothesis}
 Another frequently proposed cognitive account of DMN function revolves
@@
 
-\subsection{Sentinel Hypothesis}
+\subsection{Sentinel hypothesis}
 Processing self-relevant information was perhaps the first
@@
 
-\paragraph{Solution: The reparametrization trick.}
+\paragraph{Solution: the reparametrization trick.}
 \begin{itemize}

### 2017-04-06 paper/paper.tex
NEW FIG: neurobiogical implementation of RL in by the DMN
more

@@
 \subsection{Putting everything together}
-\begin{figure}[!h]
-  \includegraphics[width=.9\linewidth]{rl_process_chart.pdf}
-  \caption{The DMN as an RL agent... \textbf{Please add a figure caption}}
-  \label{fig:rl_process_chart}
-\end{figure}
 The DMN is today known to consistently increase in neural
@@
 % information are probably commonplace during perception \citep{gershman2010learning}
+\begin{figure}[!h]
+  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
+  \caption{Neurobiological implementation of RL by the DMN...}
+  \label{fig:rl_process_chart}
+\end{figure}
+
 The DMN may subserve

### 2017-04-06 paper/paper.tex
more fixups
sentinel hypothesis 3.0

@@
 % information are probably commonplace during perception \citep{gershman2010learning}
-\begin{figure}[!h]
-  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
-  \caption{Neurobiological implementation of RL by the DMN...}
-  \label{fig:rl_process_chart}
-\end{figure}
-
 The DMN may subserve
@@
   in humans.    
+\begin{figure}[!h]
+  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
+  \caption{Neurobiological implementation of RL by the DMN...}
+  \label{fig:rl_process_chart}
+\end{figure}
 
@@
 relevance in the environment
-as an advantageous evolutionary adaptation.
+as an advantageous evolutionary adaptation \citep{randy2008, hahn2007cingulate}.
 According to this cognitive account, the human brain's baseline realizes
 a ``radar'' function to
-detect subjectively salient objects and unexpected events in the environment.
+detect subjectively salient cues and unexpected events in the environment.
 Propositions of such a sentinel function have however seldom detailed
@@
 encountering a salient environmental stimulus.
-In our formal account,
-cascades of candidate actions
+In the present formal account,
+promising action trajectories
 are recursively explored by human agents. Conversely,
 certain branches of candidate action trajectories
-are detected to be less worthy to get mentally explored.
+are detected to be less worthy to become mentally explored.
 This mechanism, expressed by the Bellman equation,
-directly implies stratifications over attention and working memory
-to relevant objects and events in the environment.
+directly implies stratified allocation of attention and working memory load
+over relevant cues and events in the environment.
 %
 Further,
-our account provides an explanation for the known implication
-of the DMN
+our account provides a parsimonious explanation for
+the increasingly recognized observation
+of the DMN implication
 in certain goal-directed experimental tasks and
-in the idlying task-unconstrained cognitive sets.
+in task-unconstrained mind-wandering \citep{smith2009, bzdok2016formal}.
 Both cognitive sets may entail DMN recruitment
-if environmental relevance is processed, manipulated, and used for prediction.
+if environmental relevance is processed, manipulated, and
+used for predicting the future.
 During tasks,
-the policy and value estimates are updated to optimize short-term action.
+the policy and value estimates may be
+updated to optimize especially short-term action.
 At rest, these parameter updates may
 improve especially mid- and long-term action.
-We thus provide answers to the unsettled question why the involvement
-of the same neurobiological brain circuit, the DMN, has been documented
+The horizon is expressed in the $\gamma$ parameter in the formal account.
+We thus provide answers for the unsettled question why the involvement
+of the same neurobiological brain circuit (i.e., DMN) has been documented
 for baseline house-keeping functions and specific task performances.
 
-In fact,
+More specifically,
 environmental stimuli especially important for humans are frequently of
 social nature. This is probably unsurprising
-given that a key property of the human species is
-the complexity of their social systems
+given that a likely human-defining property is
+the complexity of the social systems
 \citep{tomasello2009cultural}.
 According to the ``social brain hypothesis'',
-the human brain did not evolve to solve problems of the
+the human brain has not primarily evolved to solve problems of the
 physical environment, but to form and maintain increasingly complex
-social systems to solve ecological problems by means of social relationships
+social systems to solve ecological problems via social relationships
 \citep{whiten1988machiavellian}.
-This may explain why social topics amount to roughly
-two thirds of human everyday communication \citep{dunbar1997human}
-and
-why mind-wandering and dreams
-are so rich in stories about people and
+Indeed, social topics amount to roughly
+two thirds of human everyday communication \citep{dunbar1997human},
+while
+mind-wandering at daytime and dreams during sleep
+are rich in stories about people and
 the complex relationships between them.
@@
 analogues of the DMN in monkeys \citep{mantini2011default}
-and rats \citep{lu2012rat}, species with
+and rats \citep{lu2012rat}, both species with
 supposedly less advanced social-cognitive capacities.
 Instead,
-our formal account can parsimoneously incorporate
+the formal account can parsimoneously incorporate
 the dominance of social content in
-human mental activity by reexpressing relevance to humans
-in high value function estimates for social information
-\citep{baker2009action}.
-The DMN may thus exert control on the human agent to prioritize
+human mental activity by reexpressing relevance of information about humans
+in high value function estimates
+\citep{baker2009action, kampe2001psychology, krienen2010clan}.
+The DMN may thus exert control on the human agent in a way that prioritizes
 appraisal of and action towards social contexts,
-without completely neglecting environmental cues of the physical world.
+without entirely neglecting environmental cues of the physical world.
 In sum,
 our account on the DMN directly implies
-its previously proposed ``radar'' function
+its previously proposed ``sentinel'' function
 of monitoring the environment for salient information
-in general and inherently incorporates the importance of social environmental cues
-as an important special case, without excluding other topics
-contemplated in continuous mental activity.
+in general and
+inherently incorporates the importance of social environmental cues
+as an important special case.
 

### 2017-04-06 paper/paper.tex
fix: PC hypothesis 1.0

@@
 \subsection{Predictive coding hypothesis}
-The predictive coding framework
+The predictive coding account
 \citep{clark2013whatever, friston2008hierarchical}
@@
 
-\begin{itemize}
-  \item Both the predictive coding account and POMDPs
-  are process theories backed up by plausible
-  neurobiological evidence that
-  view the brain as a ``statistical organ''
-  generalizing from the past to new experiences.
-  \item Both provide a parsimonious explanation why the
+There a number of common properties between the predictive coding account
+and the proposed formal account.
+Both frameworks are plausibilized by neurobiological evidence that
+the human brain may be viewed as a ``statistical organ'' aiming to
+generalize from the past to new experiences.
+Both accounts provide a parsimonious explanation why the
 human brain decreases processing load devoted to incoming information
 when the environment is predictable and allocate increased
-  attentional resources when novel stimuli are encountered.
-  Similarly, both propose explanations why
-  predicting the future is inherently
-  linked to information from the past.
-  \item Sensory experience is a generative process from both views.
-  In predictive coding, sensory perception of the external world
-  is a generative experience due to modulatory top down experience at
-  various hierarchical levels of sensory input processing.
-  In our RL view, the (partially observed)
-  environmental model is incorporated in the POMDP,
-  \textit{which can be fully recovered based on the last
-  state alone}.
-  \item Both naturally expose a mechanisms of brain plasticity in that
+computation resources when novel stimuli or unexpected events are encountered.
+The predictive coding and formal account
+naturally evoke a mechanisms of brain plasticity in that
 neuronal wiring gets increasingly adapted
@@
 challenges.
-  \item The hierarchical aspects from predictive coding
-  is re-expressed in POMDPs in form of
-  nested prediction of probable upcoming actions and rewards.
-  \item Both model the consequences of action. In RL, the horizon of that
-  look into the future is explicitly manifested in the $\gamma$ parameter
-  in the Bellman equation, but not
-  explicitly modeled in the PC account that
-  tends to emphasize prediction about the
+
+While sensory experience is a generative process from both views,
+in predictive coding sensory perception of the external world
+is a generative experience due to modulatory top down-experience at
+various hierarchical levels of sensory input processing. Instead,
+the Markov property of our sequential decision making framework
+In our RL view, the environmental model is
+\textbf{effectively expressed in the current state of the human agent
+XXX guys, please double-check}.
+The hierarchical processing aspect from predictive coding
+is re-expressed in our account in form of
+nested prediction of probable upcoming actions, states, and rewards.
+Both model the consequences of action.
+In the present account, the horizon of that
+look into the future is explicitly expressed in the $\gamma$ parameter
+in the Bellman equation.
+The predictive coding framework is typically explained without
+explicitly parameterization of the agent's horizon and
+has a tendency to be presented as emphasizing prediction about the
 immediate future.
-  \item \suggestremove{Mismatch negativity of PC is immanent in reward prediction error
-  by the difference between actually predicted reward of an action given
-  a state and the reward predicted by non-linear regression with graded
-  discounting of future rewards.}
-  \item Adapting the neuronal connections for improved top-down modulation
+% \item \suggestremove{Mismatch negativity of PC is immanent in reward prediction error
+% by the difference between actually predicted reward of an action given
+% a state and the reward predicted by non-linear regression with graded
+% discounting of future rewards.}
+The process of adapting the neuronal connections
+for improved top-down modulation
 takes the concrete form of gradient computation
-  \textit{and back-propagation} in MDPs and RL,
+back-propagation in our RL implementation,
 although the neurobiological plausibility of
@@
 \item In sum,
-  the MDP account may serve as a concrete conceptualization of the PC account
-  in cognitive neuroscience. MPDs have the advantage of exposing an explicit
+the formal account may serve as a concrete implementation of
+the predictive coding account
+from cognitive neuroscience.
+MPDs have the advantage of exposing an explicit
 mechanism of the horizon of future considerations or
@@
 %
-  As such, our MDP account of DMN function can parsimoneously explain
-  prediction coding hypothesis.
-\end{itemize}
+As such, recasting DMN function in terms of MDPs naturally incorporates
+a majority of aspects of the prediction coding hypothesis.
 

### 2017-04-07 paper/paper.tex
fix: PC hypothesis 2.0

@@
 \subsection{Predictive coding hypothesis}
-The predictive coding account
+Predictive coding mechanisms
 \citep{clark2013whatever, friston2008hierarchical}
-is a frequently evoked idea related to the default mode function
+are a frequently evoked idea in the context of default mode function
 \citep{bar2007}.
-According to this cognitive account,
+According to this view,
 cortical responses emerge from continuous functional interaction between
 higher and lower levels of the neural processing hierarchy.
-This dynamic intercourse enables inference on the world by reconciling
-gaps between fresh sensory input and stored prior information.
 Feed-forward sensory experience is constantly calibrated by
-top-down modulation at various hierarchical processing levels.
+top-down modulation realized by more multi-modal and associative brain regions
+further away from the primary sensory cortices.
+The dynamic interplay between cortical processing levels
+may enable inferences about the world by reconciling
+gaps between fresh sensory input and expectations computed
+from stored prior information.
+At each stage of neural processing,
+an internally generated prediction of aspects of sensory input is
+directly compared against the actual environmental input.
+A prediction error at one of the processing levels,
+i.e., the difference between sensory input
+and internally predicted sensation,
+incurs plasticity changes of neuronal back projections
+(i.e., adapting model parameters)
+to allow for gradually improved future prediction of the environment.
 In fact, axonal
-back projections probably outnumber by far the input projection
-existing in the brain \citep{salin1995corticocortical}.
+back projections probably outnumber by far the axonal input projections
+existing in the primate brain \citep{salin1995corticocortical}.
 This can explain the constructive, generative nature of sensory perception
 \citep{friston2010free} and
-why motor action is intimately related to sensory expectations
+why motor action is probably intimately related to sensory expectations
 \citep{wolpert1995internal, kording2004bayesian}.
-At each stage of neural processing,
-an internally generated prediction of the sensory input is
-directly compared against the actual input from the environment.
-A prediction error at one of the processing levels,
-i.e. the difference between sensory input
-and internally predicted sensation,
-incurs plasticity changes of neuronal back projections (i.e., model parameters)
-for gradually improved future prediction of the environment.
-Contextual integration is hence maintained by top-down modulation of sensorimotor
-processing by a-priori information about the environment.
-The generative model of how perceived sensory events arise in the
+Contextual sensorimotor integration of perception-action cycles
+may hence be maintained by top-down modulation
+by a-priori information about the environment.
+
+This process permits updates of the internal representation of the environment
+to best accommodate and prepare for the constant influx of sensory examples.
+There are hence a number of common properties
+between the predictive coding account
+and the proposed formal account.
+Importantly,
+the generative model of how perceived sensory cues arise in the
 world would be incorporated into
 the current neuronal wiring.
-Indeed,
-This process permits updates of the internal model of the environment
-to best fit the constant influx of sensory examples.
-
-There a number of common properties between the predictive coding account
-and the proposed formal account.
-Both frameworks are plausibilized by neurobiological evidence that
-the human brain may be viewed as a ``statistical organ'' aiming to
+Further,
+both frameworks are plausibilized by neurobiological evidence that suggest
+the human brain to be a ``statistical organ'' with the biological purpose to
 generalize from the past to new experiences.
-Both accounts provide a parsimonious explanation why the
+Both accounts also provide a parsimonious explanation why the
 human brain decreases processing load devoted to incoming information
-when the environment is predictable and allocate increased
+when the environment is predictable and allocates increased
 computation resources when novel stimuli or unexpected events are encountered.
@@
 While sensory experience is a generative process from both views,
-in predictive coding sensory perception of the external world
-is a generative experience due to modulatory top down-experience at
+the predictive coding account frames
+sensory perception of the external world
+as a generative experience due to the modulatory top-down influence at
 various hierarchical levels of sensory input processing. Instead,
-the Markov property of our sequential decision making framework
+the Markov property of our sequential decision making framework.
 In our RL view, the environmental model is
@@
 nested prediction of probable upcoming actions, states, and rewards.
-Both model the consequences of action.
-In the present account, the horizon of that
-look into the future is explicitly expressed in the $\gamma$ parameter
-in the Bellman equation.
+Both frameworks model the consequences of action.
 The predictive coding framework is typically explained without
@@
 immediate future.
+In the present account, the horizon of that
+look into the future is explicitly expressed in the $\gamma$ parameter
+of the Bellman equation.
 % \item \suggestremove{Mismatch negativity of PC is immanent in reward prediction error
@@
 for improved top-down modulation
-takes the concrete form of gradient computation
-back-propagation in our RL implementation,
-although the neurobiological plausibility of
+takes the concrete form of gradient computation and
+back-propagation in our RL implementation.
+It is however important to note that
+the neurobiological plausibility of
 the back-propagation procedure is controversal
@@
 how the internal model of the world gets updated,
-and collapses sensory input processing and action output preparation.
+and
+more explicitly reconciles sensory input processing and action output preparation.
 %

### 2017-04-07 paper/paper.tex
a pass on rl sections

@@
 \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
-$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$.
+$(s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$.
 The agent then tries to approximate
 the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the Bellman equation \eqref{eq:bellman}, namely
-$y := r + \gamma \max_{a'} Q(s', a'|\theta^Q_k)$, with the output of a parametrized regression model $(s,a,r,s')
-\mapsto {Q}(s, a|\theta^Q_{k})$. Here the function $(s,a) \mapsto Q(a,s|\theta^Q_{k})$
+\begin{equation}
+  y_k := r_k + \gamma \max_{a'} \tilde{Q}(s', a'|\theta_{k-1}),
+\end{equation}
+with the output of a parametrized regression model $(s,a)
+\mapsto \tilde{Q}(s, a|\theta_{k-1})$. Here the function $(s,a) \mapsto Q(a,s|\theta_{k-1})$
 % \textbf{XXX just double-checking, are the dots int he Q() correct?}
-is an approximator for the $Q$-value operator,  parameterized by weights $\theta^Q_{k+1}$
+is an approximator for the $Q$-value operator,  parameterized by weights $\theta_{k}$
 (synaptic strengths, etc.).
@@
 
-Computing an optimal policy then corresponds to finding the parameters $\theta^Q_{k+1}$ which minimize the following mean-squared loss function
+Computing an optimal policy then corresponds to finding the parameters $\theta_{k}$ which minimize the following mean-squared loss function
 \begin{equation}
-  \mathcal L(\theta^Q_{k+1})
-  % = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{k+1}))^2]
-  = \mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\frac{1}{2}(Q(s, a|\theta^Q_{k+1}) - y)^2\right],
+  \mathcal L(\theta^Q_{k})
+  % = \mathbb E_{(s, a, r, s') \sim \mathcal D}[(\tilde{Q}(s, a|\theta) - \tilde{Q}(s',a'|\theta^Q_{k}))^2]
+  = \mathbb E_{(s, a, r, s') \sim \mathcal D}\left[\frac{1}{2}(\tilde{Q}(s, a|\theta_{k}) - y_k)^2\right].
   \label{eq:oracle}
 \end{equation}
-where
-$ y: = r + \gamma \max_{a' \in \mathcal A} Q(s', a'|\theta^Q_k) = r + \gamma Q(s', \pi(s')|\theta^Q_k)$ is the prediction target.
-% While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.
 For instance, a general linear model with a kernel $\phi$ would be of the
 form
-$${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
+$$\tilde{Q}(s, a|\theta) = \phi(s,a)^T\theta.$$
 $\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action pairs
@@
 \begin{equation}
-  \delta \theta_{k+1}^Q \propto -\nabla_{\theta^Q_{k+1}}\mathcal L(\theta^Q_{k+1})
-  % = -\mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{}))^2]
-  = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret / RPE}}
-    \underbrace{\nabla_{\theta^Q_{k}}Q(s, a|\theta^Q_{k+1})}_{\text{averseness}}\right].
+  \delta \theta_{k} \propto -\nabla_{\theta_{k}}\mathcal L(\theta_{k})
+  = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
+    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{averseness}}].
   \label{eq:oracle}
@@
 $\mathcal S \times \mathcal A$, and so
-$$\nabla_{\theta^Q_{k+1}}Q(s, a|\theta^Q_{k+1}) = \phi(s, a) = \delta_{(s,a)}.$$
+$$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
 Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
-$$\theta^Q(s,a) \leftarrow \theta^Q(s,a) + (Q(s, a|\theta^Q_{k+1}) - y),$$
+$$\theta(s,a) \leftarrow \theta(s,a) + \alpha(\tilde{Q}(s, a|\theta) - y),$$
 the well-known TD update rule \citep{sutton1998reinforcement}.
@@
 \citep{goodfellow2016deep}.
-\item In sum,
+In sum,
 the formal account may serve as a concrete implementation of

### 2017-04-08 paper/paper.tex
cool ref

@@
 the HC could contribute to synthesizing imagined sequences of world states,
-observations and rewards \citep{aronov2017}.
+observations and rewards \citep{aronov2017, chao2017interaction}.
 These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics.

### 2017-04-08 paper/paper.tex
fix: PC hypothesis 3.0

@@
 \citep{bar2007}.
-According to this view,
-cortical responses emerge from continuous functional interaction between
+Cortical responses are explained as
+emerging from continuous functional interaction between
 higher and lower levels of the neural processing hierarchy.
-Feed-forward sensory experience is constantly calibrated by
-top-down modulation realized by more multi-modal and associative brain regions
-further away from the primary sensory cortices.
+Feed-forward sensory processing is constantly calibrated by
+top-down modulation from more multi-modal and associative brain regions
+further away from primary sensory cortical regions.
 The dynamic interplay between cortical processing levels
@@
 At each stage of neural processing,
-an internally generated prediction of aspects of sensory input is
+an internally generated prediction of aspects of sensory sensation is
 directly compared against the actual environmental input.
-A prediction error at one of the processing levels,
-i.e., the difference between sensory input
-and internally predicted sensation,
+A prediction error at one of the processing levels
 incurs plasticity changes of neuronal back projections
@@
 to allow for gradually improved future prediction of the environment.
-In fact, axonal
-back projections probably outnumber by far the axonal input projections
-existing in the primate brain \citep{salin1995corticocortical}.
-This can explain the constructive, generative nature of sensory perception
+In this way,
+the predictive coding hypothesis offers explanations for
+the constructive, generative nature of sensory perception
 \citep{friston2010free} and
-why motor action is probably intimately related to sensory expectations
+the intimate relation of motor action to sensory expectations
 \citep{wolpert1995internal, kording2004bayesian}.
 Contextual sensorimotor integration of perception-action cycles
-may hence be maintained by top-down modulation
+may be maintained by top-down modulation
 by a-priori information about the environment.
 
-This process permits updates of the internal representation of the environment
-to best accommodate and prepare for the constant influx of sensory examples.
+Predictive coding processes
+permit updates of the internal representation of the environment
+to best accommodate and prepare the organism
+for the constant influx of sensory examples and
+for reacting to the environment.
 There are hence a number of common properties
@@
 Importantly,
-the generative model of how perceived sensory cues arise in the
+a generative model of how perceived sensory cues arise in the
 world would be incorporated into
@@
 Further,
-both frameworks are plausibilized by neurobiological evidence that suggest
+both frameworks are plausibilized by neuroscientific evidence that suggest
 the human brain to be a ``statistical organ'' with the biological purpose to
 generalize from the past to new experiences.
-Both accounts also provide a parsimonious explanation why the
+Neuroanatomically, axonal
+back projections probably outnumber by far the axonal input projections
+existing in the primate brain \citep{salin1995corticocortical}.
+Moreover,
+both accounts provide a parsimonious explanation why the
 human brain decreases processing load devoted to incoming information
-when the environment is predictable and allocates increased
-computation resources when novel stimuli or unexpected events are encountered.
+when the environment is predictable because the internal generative
+model only requires updating following descrepancies between
+environmental reality and its internal reinstantiation.
+Increased computation resources are however allocated
+when novel stimuli or
+unexpected events are encountered by the organism.
 The predictive coding and formal account
-naturally evoke a mechanisms of brain plasticity in that
+naturally evoke a mechanism of brain plasticity in that
 neuronal wiring gets increasingly adapted
-when faced by unprecedented environmental
-challenges.
+when faced by unprecedented environmental challenges.
 
@@
 as a generative experience due to the modulatory top-down influence at
-various hierarchical levels of sensory input processing. Instead,
-the Markov property of our sequential decision making framework.
-In our RL view, the environmental model is
+various hierarchical levels of sensory input processing.
+In our RL view, instead,
+the Markov property of our sequential decision making framework entails
+the environmental model to be
 \textbf{effectively expressed in the current state of the human agent
 XXX guys, please double-check}.
-The hierarchical processing aspect from predictive coding
+Further,
+the hierarchical processing aspect from predictive coding
 is re-expressed in our account in form of
@@
 The predictive coding framework is typically explained without
-explicitly parameterization of the agent's horizon and
+explicit parameterization of the agent's horizon and
 has a tendency to be presented as emphasizing prediction about the
@@
 % discounting of future rewards.}
-The process of adapting the neuronal connections
+Finally,
+the process of adapting the neuronal connections
 for improved top-down modulation
@@
 the formal account may serve as a concrete implementation of
-the predictive coding account
-from cognitive neuroscience.
+the long-standing predictive coding account from cognitive neuroscience.
 MPDs have the advantage of exposing an explicit
@@
 and
-more explicitly reconciles sensory input processing and action output preparation.
-%
-As such, recasting DMN function in terms of MDPs naturally incorporates
+more explicitly reconciles sensory input processing and action consideration.
+%
+Recasting DMN function in terms of MDPs therefore naturally incorporate
 a majority of aspects of the prediction coding hypothesis.

### 2017-04-08 paper/paper.tex
fix: semantic hypothesis 1.0

@@
 Analogies might naturally tie incoming novel sensory stimuli to
-explicit world knowledge (i.e., semantics) extractable from the environment
+explicit world knowledge (i.e., semantics) extracted from the environment
 \citep{bar2009proactive}.
-In this way, the encoding of complex environments could be facilitated
+The encoding of complex environments could thus be facilitated
 by relative similarity associations between states.
 %
-Going much beyond human language ifself,
-semantic building blocks provide the basis to
-mentally envision non-existent scenarios
-that would improve optimal behavior in the environment
-by simulating future events.
+Going beyond human language itself,
+semantic building blocks may provide the basis for
+mentally envisioning non-existent scenarios
+that would optimize behavior in the environment
+by constantly simulating possible future events.
 Such cognitive processes can afford
@@
 Indeed, in aging humans, remembering the past and imaging the future
-equally decreased in the level of detail and are associated with
+equally decreased in the level of detail and were associated with
 concurrent deficits in forming and integrating relationships between
@@
 Further,
-a constructive account explains the reciprocal relationship
+a constructive account can explain the reciprocal relationship
 between an egocentric first person perspective and
 an allocentric bird’s eye perspective immersed in
-self-reflection, social knowledge, and autobiographical memories.
+self-reflection, semantic associations, and autobiographical memories.
 %
@@
 are also closely related to episodic memory, language, problem solving,
-planning, estimating other people's thoughts, and spatial navigation
-as these necessitate abstract world knowledge and abstract assocations
+planning, estimating other people's thoughts, and spatial navigation.
+These neural processes are likely to
+build on abstract world knowledge and abstract assocations
 for binding the constituent elements in mental scene construction
 \citep{schacter2007remembering}.
-These scene generation processes could contribute to interpreting the
+Such scene generation processes could contribute to interpreting the
 present and foretelling the future.
-This view is for instance supported by evidence in animals that
+Mental scene construction has been proposed
+to imply a distinction between
+environmental perception and internally generated mind-wandering
+\citep{buckner2007self}.
+These authors stated that
+``A computational model [...] will probably require a form of
+regulation by which perception of the current world is suppressed
+while simulation of possible alternatives are constructed,
+followed by a return to perception of the present.''.
+The semantic hypothesis is for instance supported by evidence in animals that
 could learn a \textit{cognitive map of the environment},
@@
 
-\begin{itemize}
-  \item Both the semantic hypothesis and MDPs expose mechanisms of
-  how alternative decision trees could be mentally explored.
-  \item In both semantic hypothesis and MDPs,
-  there is no evidence to indicate that predictions of various
+In comparison,
+both the semantic hypothesis and MDPs expose mechanisms of
+how action considerations could be mentally explored.
+In both accounts,
+there is also no evidence to indicate that predictions of various
 levels of complexity, abstraction, timescale and purpose
@@
 DMN activity increases across time, space, and content domains
-  demonstrated in many neuroimaging studies. The semantic hypothesis
-  and RL account provide explanations why hippocampus lesion does
+demonstrated in many neuroimaging studies \citep{spreng2009common}.
+The semantic hypothesis
+and MDP account provide explanations why hippocampus lesion does
 not only impair retrieving memories, but also hypothetical and future
 thinking \citep{hassabis2007patients}.
-  \item The notion of semantic or knowledge association is
+While both semantic hypothesis and
+our formal account propose memory-based internally
+generated information for probabilistic mental models of action outcomes,
+MDPs render explicit the grounds on which the final action is
+eventually chosen, namely, the estimated cumulative reward.
+\textbf{The notion of semantic knowledge associations is
 incorporated into the MDP as the Markov property,
@@
 value matrix and action transition schedules drive
-  stimulus processing and action choice in the present.
-  \item Mental scene construction has been proposed by some authors
-  \citep{buckner2007self} to imply a distinction between
-    environmental perception and internally generated mind-wandering\footnote{
-      \textbf{XXX Elvis: This is more than just a footnote. Make this part of the paper
-      and factorize the lot. Also you probably need a citation here!!!}
-  ``A computational model of how such a process might be structured
-  is far from being defined, but it will probably require a form of
-  regulation by which perception of the current world is suppressed
-  while simulation of possible alternatives are constructed,
-  followed by a return to perception of the present.''\citep{buckner2007self}}.
+stimulus processing and action choice in the present.}
 The MDPs naturally integrate the former egocentric
@@
 angles on the world in a same optimization problem.
-  \item The semantic account of DMN function does not offer
+Further,
+the semantic account of DMN function does not offer
 a mechanistic explanation how explicit world knowledge and semantic analogies thereof
 lead to prediction of future actions and states.
-  \item In contrast to existing accounts on semantics and
+The semantic hypothesis does not explain why memory recall
+for scene construction in humans is typically fragmentary and noisy
+instead of accurate and reliable.
+In contrast to existing accounts on semantics and
 mental scene construction, the random and creative aspects of DMN function
 are explained in MDPs by the advantages of stochastic optimization.
-  \item The semantic hypothesis does not explain why memory recall
-  for scene construction in humans is typically fragmentary and noisy
-  instead of accurate and reliable.
 Yet, the MDP framework provides an algorithmic explanation in that
@@
 as the more advantageous adaptation.
-  \item While both semantic and MDP account propose memory-based internally
-  generated information for probabilistic mental models of action outcomes,
-  only MDPs render explicit the grounds on which the final action is
-  eventually chosen (i.e., the estimated cumulative reward).
-  \item In sum, episodic scene construction according to the semantic
+In sum, episodic scene construction according to the semantic
 hypothesis is lacking an explicit time and incentive model;
@@
 can be beneficial.
-\end{itemize}
 

### 2017-04-08 paper/paper.tex
abstract: incorporate Gs comments

@@
 envisioning of mental scenarios to simulate the future.
-\suggestremove{The DMN may be dedicated to continuous autobiographical memory retrieval,
-generation of hypothetical outcomes,
-and reward contingency evaluation when letting the mind go.}
-Such functional roles offer explanations of
-its highest energy consumption in the brain and
+Such functional roles become more likely due to
+this brain network's highest energy consumption in the brain and
 its intimate coupling with conscious awareness
@@
 %
-This concept paper proposes a process model \textbf{(XXX Guillaume: It's not clear what a "process model" is for most people?)} that describes
-\textit{how} the DMN may actually implement continuous
+The present concept paper proposes a process model
+in trying to explain \textit{how} the DMN may actually implement continuous
 assessment and prediction of the environment to guide action choices.
@@
 We argue that our formal account of DMN function
-naturally accommodates \suggestremove{previously} proposed cognitive accounts as special cases and
-parsimoneously explains \suggestremove{various} experimental findings in \suggestadd{both} animals and humans.
-%
-Formal models of the neural computations realized in the DMN
-could offer access routes to the statistical regularities
-of human predictive behavior \textbf{(XXX Guillaume: Maybe too wordy\ldots)}.
+naturally accommodates proposed cognitive accounts as special cases and
+parsimoneously explains
+experimental findings in animals and humans.
+%
+Formal process models of the neural implemention
+of future predictions probably realized in the DMN
+could offer insight into the statistical mechanisms underlying
+human behavior.
 \end{abstract}

### 2017-04-08 paper/paper.tex
Bring darkness into the introduction

@@
 \citep{kenet2003spontaneously, fiser2004small}.
-This baseline energy demand is subject to surprisingly little modulations
+The baseline energy demand is subject to surprisingly little modulations
 when mentally processing environmental challenges
-\citep{raichle2001pnas}. \textbf{(XXX Guillaume: What about the energy consumption compare to sleep?)}
+\citep{raichle2001pnas}. \textbf{(XXX Guillaume: What about the energy consumption compare to sleep? Danilo: I'd love to cite that
+but I am not aware of such studies)}
+This dark matter of brain physiology begs the question
+of underlying biological purpose
+\citep{raichle2006brain}.
 What has early been described as the "stream of consciousness"

### 2017-04-09 paper/paper.tex
some polishing

--- 2017-04-08 paper/paper.tex
@@
 \subsection{Partially Observable Markov Decision Processes}
-It is our view that at a sufficiently coarse scale, the brain -- and therefore the DMN in particular --
+It is our view that at a sufficiently coarse scale, the brain --and therefore the DMN in particular--
 is a physical system governed by the laws of
@@
 any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
-step is sufficiently short, defining $s(t)$, the state at time $t$, as the position in state space.
+step is sufficiently short
+% , defining $s(t)$, the state at time $t$, as the position in state space.
 The process has $memory$ if the next state depends not only on the current state but also on some finite
@@
 a set of candidate actions to perform given this state. Starting a time $t=0$,
-a policy $\pi$ generates a trajectory of action cascades as follows
+a policy $\pi$ generates a trajectory of action cascades / trajectories as follows
 \begin{eqnarray*}
@@
 rewards expected at a later time, precisely rewards more than $1/(1 - \gamma)$ time steps into the future are disregarded.
-This reduces the variance of expected rewards to be accumulated across action cascades by limiting the length of this cascades.
+This reduces the variance of expected rewards accumulated across action cascades by limiting the length of this cascades.
 % given that there is more uncertainty in the farsighted future.
@@
 At the end of this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced, which represents an exemplar
-behavior of the agent and is recorded in replay memory buffer:
-$D \leftarrow D \cup \{e\}$ (possibly discarding the oldest entries to make space).
-Now, at iteration $k+1$, replay consists in sampling (uniform or importance-weighted
+behavior of the agent and is recorded in replay memory buffer
+(possibly discarding the oldest entries to make space):
+$\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
+At iteration $k+1$, replay then consists in sampling (uniform or importance-weighted
 \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
@@
 with the output of a parametrized regression model $(s,a)
-\mapsto \tilde{Q}(s, a|\theta_{k-1})$. Here the function $(s,a) \mapsto Q(a,s|\theta_{k-1})$
+\mapsto \tilde{Q}(s, a|\theta_{k-1})$. Here the function $(s,a) \mapsto \tilde{Q}(a,s|\theta_{k-1})$
 % \textbf{XXX just double-checking, are the dots int he Q() correct?}
@@
 likely are state-of-the-art in RL.
-The set of $\theta^Q$ parameters that instantiate the nonlinear interactions
+The set of $\theta$ parameters that instantiate the nonlinear interactions
 between layers of the artifical neural network
@@
 via \textit{backpropagation},
-where reward prediction errors (RPES) -- also known as \textit{regret}-- percolate from lower to higher processing layers to modulate the choice of future actions
+where reward prediction errors (RPEs) -- also known as \textit{regret}-- percolate
+from lower to higher processing layers to modulate the choice of future actions
 % online stochastic gradient descent is used to minimize the variational objective
@@
 
-\subsubsection{Link to classical reinforcement learning algorithms} One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
+\subsubsection{Link to classical reinforcement learning algorithms}
+One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
 learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
@@
 \citep{bar2007}.
-Cortical responses are explained as
-emerging from continuous functional interaction between
+According to this view,
+cortical responses emerge from continuous functional interaction between
 higher and lower levels of the neural processing hierarchy.
-Feed-forward sensory processing is constantly calibrated by
-top-down modulation from more multi-modal and associative brain regions
-further away from primary sensory cortical regions.
+Feed-forward sensory experience is constantly calibrated by
+top-down modulation realized by more multi-modal and associative brain regions
+further away from the primary sensory cortices.
 The dynamic interplay between cortical processing levels
@@
 At each stage of neural processing,
-an internally generated prediction of aspects of sensory sensation is
+an internally generated prediction of aspects of sensory input is
 directly compared against the actual environmental input.
-A prediction error at one of the processing levels
+A prediction error at one of the processing levels,
+i.e., the difference between sensory input
+and internally predicted sensation,
 incurs plasticity changes of neuronal back projections
@@
 to allow for gradually improved future prediction of the environment.
-In this way,
-the predictive coding hypothesis offers explanations for
-the constructive, generative nature of sensory perception
+In fact, axonal
+back projections probably outnumber by far the axonal input projections
+existing in the primate brain \citep{salin1995corticocortical}.
+This can explain the constructive, generative nature of sensory perception
 \citep{friston2010free} and
-the intimate relation of motor action to sensory expectations
+why motor action is probably intimately related to sensory expectations
 \citep{wolpert1995internal, kording2004bayesian}.
 Contextual sensorimotor integration of perception-action cycles
-may be maintained by top-down modulation
+may hence be maintained by top-down modulation
 by a-priori information about the environment.
 
-Predictive coding processes
-permit updates of the internal representation of the environment
-to best accommodate and prepare the organism
-for the constant influx of sensory examples and
-for reacting to the environment.
+This process permits updates of the internal representation of the environment
+to best accommodate and prepare for the constant influx of sensory examples.
 There are hence a number of common properties
@@
 Importantly,
-a generative model of how perceived sensory cues arise in the
+the generative model of how perceived sensory cues arise in the
 world would be incorporated into
@@
 Further,
-both frameworks are plausibilized by neuroscientific evidence that suggest
+both frameworks are plausibilized by neurobiological evidence that suggest
 the human brain to be a ``statistical organ'' with the biological purpose to
 generalize from the past to new experiences.
-Neuroanatomically, axonal
-back projections probably outnumber by far the axonal input projections
-existing in the primate brain \citep{salin1995corticocortical}.
-Moreover,
-both accounts provide a parsimonious explanation why the
+Both accounts also provide a parsimonious explanation why the
 human brain decreases processing load devoted to incoming information
-when the environment is predictable because the internal generative
-model only requires updating following descrepancies between
-environmental reality and its internal reinstantiation.
-Increased computation resources are however allocated
-when novel stimuli or
-unexpected events are encountered by the organism.
+when the environment is predictable and allocates increased
+computation resources when novel stimuli or unexpected events are encountered.
 The predictive coding and formal account
-naturally evoke a mechanism of brain plasticity in that
+naturally evoke a mechanisms of brain plasticity in that
 neuronal wiring gets increasingly adapted
-when faced by unprecedented environmental challenges.
+when faced by unprecedented environmental
+challenges.
 
@@
 as a generative experience due to the modulatory top-down influence at
-various hierarchical levels of sensory input processing.
-In our RL view, instead,
-the Markov property of our sequential decision making framework entails
-the environmental model to be
+various hierarchical levels of sensory input processing. Instead,
+the Markov property of our sequential decision making framework.
+In our RL view, the environmental model is
 \textbf{effectively expressed in the current state of the human agent
 XXX guys, please double-check}.
-Further,
-the hierarchical processing aspect from predictive coding
+The hierarchical processing aspect from predictive coding
 is re-expressed in our account in form of
@@
 The predictive coding framework is typically explained without
-explicit parameterization of the agent's horizon and
+explicitly parameterization of the agent's horizon and
 has a tendency to be presented as emphasizing prediction about the
@@
 % discounting of future rewards.}
-Finally,
-the process of adapting the neuronal connections
+The process of adapting the neuronal connections
 for improved top-down modulation
@@
 the formal account may serve as a concrete implementation of
-the long-standing predictive coding account from cognitive neuroscience.
+the predictive coding account
+from cognitive neuroscience.
 MPDs have the advantage of exposing an explicit
@@
 and
-more explicitly reconciles sensory input processing and action consideration.
-%
-Recasting DMN function in terms of MDPs therefore naturally incorporate
+more explicitly reconciles sensory input processing and action output preparation.
+%
+As such, recasting DMN function in terms of MDPs naturally incorporates
 a majority of aspects of the prediction coding hypothesis.
@@
 Analogies might naturally tie incoming novel sensory stimuli to
-explicit world knowledge (i.e., semantics) extracted from the environment
+explicit world knowledge (i.e., semantics) extractable from the environment
 \citep{bar2009proactive}.
-The encoding of complex environments could thus be facilitated
+In this way, the encoding of complex environments could be facilitated
 by relative similarity associations between states.
 %
-Going beyond human language itself,
-semantic building blocks may provide the basis for
-mentally envisioning non-existent scenarios
-that would optimize behavior in the environment
-by constantly simulating possible future events.
+Going much beyond human language ifself,
+semantic building blocks provide the basis to
+mentally envision non-existent scenarios
+that would improve optimal behavior in the environment
+by simulating future events.
 Such cognitive processes can afford
@@
 Indeed, in aging humans, remembering the past and imaging the future
-equally decreased in the level of detail and were associated with
+equally decreased in the level of detail and are associated with
 concurrent deficits in forming and integrating relationships between
@@
 Further,
-a constructive account can explain the reciprocal relationship
+a constructive account explains the reciprocal relationship
 between an egocentric first person perspective and
 an allocentric bird’s eye perspective immersed in
-self-reflection, semantic associations, and autobiographical memories.
+self-reflection, social knowledge, and autobiographical memories.
 %
@@
 are also closely related to episodic memory, language, problem solving,
-planning, estimating other people's thoughts, and spatial navigation.
-These neural processes are likely to
-build on abstract world knowledge and abstract assocations
+planning, estimating other people's thoughts, and spatial navigation
+as these necessitate abstract world knowledge and abstract assocations
 for binding the constituent elements in mental scene construction
 \citep{schacter2007remembering}.
-Such scene generation processes could contribute to interpreting the
+These scene generation processes could contribute to interpreting the
 present and foretelling the future.
-Mental scene construction has been proposed
-to imply a distinction between
-environmental perception and internally generated mind-wandering
-\citep{buckner2007self}.
-These authors stated that
-``A computational model [...] will probably require a form of
-regulation by which perception of the current world is suppressed
-while simulation of possible alternatives are constructed,
-followed by a return to perception of the present.''.
-The semantic hypothesis is for instance supported by evidence in animals that
+This view is for instance supported by evidence in animals that
 could learn a \textit{cognitive map of the environment},
@@
 
-In comparison,
-both the semantic hypothesis and MDPs expose mechanisms of
-how action considerations could be mentally explored.
-In both accounts,
-there is also no evidence to indicate that predictions of various
+\begin{itemize}
+  \item Both the semantic hypothesis and MDPs expose mechanisms of
+  how alternative decision trees could be mentally explored.
+  \item In both semantic hypothesis and MDPs,
+  there is no evidence to indicate that predictions of various
   levels of complexity, abstraction, timescale and purpose
@@
   DMN activity increases across time, space, and content domains
-demonstrated in many neuroimaging studies \citep{spreng2009common}.
-The semantic hypothesis
-and MDP account provide explanations why hippocampus lesion does
+  demonstrated in many neuroimaging studies. The semantic hypothesis
+  and RL account provide explanations why hippocampus lesion does
   not only impair retrieving memories, but also hypothetical and future
   thinking \citep{hassabis2007patients}.
-While both semantic hypothesis and
-our formal account propose memory-based internally
-generated information for probabilistic mental models of action outcomes,
-MDPs render explicit the grounds on which the final action is
-eventually chosen, namely, the estimated cumulative reward.
-\textbf{The notion of semantic knowledge associations is
+  \item The notion of semantic or knowledge association is
   incorporated into the MDP as the Markov property,
@@
   value matrix and action transition schedules drive
-stimulus processing and action choice in the present.}
+  stimulus processing and action choice in the present.
+  \item Mental scene construction has been proposed by some authors
+  \citep{buckner2007self} to imply a distinction between
+    environmental perception and internally generated mind-wandering\footnote{
+      \textbf{XXX Elvis: This is more than just a footnote. Make this part of the paper
+      and factorize the lot. Also you probably need a citation here!!!}
+  ``A computational model of how such a process might be structured
+  is far from being defined, but it will probably require a form of
+  regulation by which perception of the current world is suppressed
+  while simulation of possible alternatives are constructed,
+  followed by a return to perception of the present.''\citep{buckner2007self}}.
   The MDPs naturally integrate the former egocentric
@@
   angles on the world in a same optimization problem.
-Further,
-the semantic account of DMN function does not offer
+  \item The semantic account of DMN function does not offer
   a mechanistic explanation how explicit world knowledge and semantic analogies thereof
   lead to prediction of future actions and states.
-The semantic hypothesis does not explain why memory recall
+  \item In contrast to existing accounts on semantics and
+  mental scene construction, the random and creative aspects of DMN function
+  are explained in MDPs by the advantages of stochastic optimization.
+  \item The semantic hypothesis does not explain why memory recall
   for scene construction in humans is typically fragmentary and noisy
   instead of accurate and reliable.
-In contrast to existing accounts on semantics and
-mental scene construction, the random and creative aspects of DMN function
-are explained in MDPs by the advantages of stochastic optimization.
   Yet, the MDP framework provides an algorithmic explanation in that
@@
   as the more advantageous adaptation.
-In sum, episodic scene construction according to the semantic
+  \item While both semantic and MDP account propose memory-based internally
+  generated information for probabilistic mental models of action outcomes,
+  only MDPs render explicit the grounds on which the final action is
+  eventually chosen (i.e., the estimated cumulative reward).
+  \item In sum, episodic scene construction according to the semantic
   hypothesis is lacking an explicit time and incentive model;
@@
   can be beneficial.
+\end{itemize}
 

### 2017-04-09 paper/paper.tex
addressed G's comments

--- 2017-04-08 paper/paper.tex
@@
 \subsection{Default mode network: higher-order control of the organism}
-The DMN appeared to be exclusive in task-induced
-deactivation during various psychological experiments.
-This processing of unknown information categories in the DMN has
-been argued to mediate an evolutionarily conserved function for the individual \textbf{(XXX Guillaume: Ref?)}.
+The human DMN accounts for a large part of the brain's base line activity
+and typically decreases in activity during psychological experiments.
+The standard mode of
+information maintenance and manipulation in the brain has
+been argued to mediate evolutionarily conserved functions
+\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
 This is all the more likely because the DMN contains hotspots
@@
 Today, many authors believe that the DMN implements
-probabilistic estimation of past, hypothetical, and future events \textbf{(XXX Guillaume: Ref? // many authors)}. This
-brain network
+probabilistic estimation of past, hypothetical, and
+future events
+\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
+randy2008}.
+This brain network
 might have emerged to continuously predict environmental events using
@@
 %
-Understanding of the overarching DMN function is complicated by
-simultaneously controlling perception-action cycles and
-maintaining baseline contemplations
-across time, space, and content domains
-at the interface of the external world and the self \textbf{(XXX Guillaume: This sentence is not clear and too long.)}.
+Gaining insight into the overarching DMN function is challenging because
+this network may
+simultaneously control perception-action cycles and
+maintain baseline contemplations
+across time, space, and content domains.
 
@@
 \citep{filler1973daydreaming, teasdale1995stimulus}.
-These ``offline'' processes may however contribute to optimized control of the organism. \textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment?)}
+These ``offline'' processes may however contribute to optimized control of the organism.
+\textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment?
+- danilo: hmm, let's keeping thinking about this)}
 % We formalize
@@
 
-\subsection{Sketch of our main contributions\textbf{(XXX Guillaume: What about ``Towards a computational account of DMN function'')}}
+\subsection{Towards a computational account of DMN function}
 We argue that the implication of the DMN in a
@@
 \citep{tenenbaum2011grow}.
-\textit{Control} refers to the influences that an agent exerts when interacting
+In this context,
+\textit{control} refers to the influences that an agent exerts when interacting
 with the environment to encourage preferred states.
@@
 neuroscientific experiments have invigorated this idea \citep{kording2004bayesian, fiser2004small}.
-This paper proposes a formal model that satisfies this contention.
+This paper proposes a process model that satisfies this contention.
 

### 2017-04-09 paper/paper.tex
addressed G's comments 2

@@
 \author{Elvis Dohmatob, Guillaume Dumas, Danilo Bzdok\\
+  \textbf{Please add your affiliation}
   %% INRIA, Parietal team, Saclay, France\\
@@
 %
-The present concept paper proposes a process model
+The present concept paper proposes a \textit{process model}
 in trying to explain \textit{how} the DMN may actually implement continuous
@@
 with the environment to encourage preferred states.
-Even (necessarily imperfect) memories
-\suggestremove{of past experience}, random day-time mind-wandering, and dreams during sleep
+Even (necessarily imperfect) memories, random day-time mind-wandering, and dreams during sleep
 may provide meaningful building blocks to iteratively improve
@@
 \subsection{The posteromedial cortex: global monitoring and information integration}
-\suggestremove{Among all parts of the human brain,}
-the midline structures of the DMN,
+The midline structures of the DMN,
 including the posteromedial cortex (PMC) and
@@
 brain-imaging analyses \citep{andrews2010} in
-suggesting that the PMC and \suggestremove{d}mPFC
+suggesting that the PMC and mPFC
 potentially represent the functional backbone of the DMN.

### 2017-04-09 paper/paper.tex
monster pass

@@
 We argue that our formal account of DMN function
-naturally accommodates proposed cognitive accounts as special cases and
+naturally accommodates
+previously proposed cognitive accounts as special cases and
 parsimoneously explains
@@
 %
-Formal process models of the neural implemention
-of future predictions probably realized in the DMN
-could offer insight into the statistical mechanisms underlying
-human behavior.
+A process models for the neural implemention
+of continuous future predictions probably realized in the DMN
+could offer insight into the statistical mechanisms underlying human behavior.
 \end{abstract}
@@
 potentially purposeful neuronal computation
-in the absence of a focused behavioral goal
+in the absence of behavioral focus
 \citep{kenet2003spontaneously, fiser2004small}.
@@
 in the so-called "default mode network" (DMN)
-\citep{shul1997}. This networked collection of some of the highest regions in
+\citep{shul1997, raichle2001pnas}.
+This set of some of the highest regions in
 the human association cortex \citep{mesulam1998sensation, margulies2016situating}
@@
 unfocused everyday mind wandering \citep{raichle2001pnas}.
-The DMN may exert higher-order control on behavior
-as the human brain was argued to features "a hierarchy of brain systems with
+The DMN may exert higher-order control on human behavior
+following the notion of "a hierarchy of brain systems with
 the DMN at the top and the salience and dorsal attention systems
@@
 \subsection{Default mode network: higher-order control of the organism}
-The human DMN accounts for a large part of the brain's base line activity
+The human DMN accounts for a large part of the brain's baseline activity
 and typically decreases in activity during psychological experiments.
 The standard mode of
-information maintenance and manipulation in the brain has
+neural information maintenance and manipulation has
 been argued to mediate evolutionarily conserved functions
@@
 Gaining insight into the overarching DMN function is challenging because
-this network may
+this network appears to
 simultaneously control perception-action cycles and
@@
 
-Consider an agent faced with the choice of the next action
+In this work,
+we consider an agent faced with the choice of the next action
 and being guided by rich
@@
 to describe, quantify, and predict autonomously acting systems like the brain
-can be the combination of control theory and reinforcement learning.
+is proposed to be the combination of control theory and
+reinforcement learning (RL).
 It is known that, the more the external world is predictable,
@@
 the less stimulus-independent thoughts occur
-\citep{filler1973daydreaming, teasdale1995stimulus}.
+\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
 These ``offline'' processes may however contribute to optimized control of the organism.
@@
 \section{Known neurobiological properties of the default mode network}
-We begin by a deconstruction of the DMN based on experimental findings in neuroscience.
-This walkthrough across each individual network node will describe their mechanistic relevance, paving the way for their functional interpretation within our computational account in the next section.
+We begin by a neurobiological deconstruction of the DMN
+based on experimental findings in neuroscience.
+This walkthrough across individual network nodes will describe their mechanistic relevance, paving the way for their functional interpretation within our computational account in the next section.
 
@@
 \citep{goldman1987development}, which is generally considered to
-be a sign of phylogenetic sophistication \citep{flechsig1920}.
+be a sign of evolutionary sophistication \citep{flechsig1920}.
 Physiological and disturbed metabolic fluctuations in the
@@
 the PMC has long been speculated to reflect constant computation of
-the environmental statistics and its internal representation
+environmental statistics and its internal representation
 as an "inner minds eye" \citep{cavanna2006precuneus, leech_pcc2014}.
 For instance, B\'alint's syndrome is a neurological disorder of conscious
-awareness that results from damage in the bilateral parietal cortex
+awareness that results from damage in the parietal cortex
 \citep{balint1909seelenlahmung}.
@@
 Specifically, among all parts of the PMC,
-the human ventral posterior cingulate cortex was
+the ventral posterior cingulate cortex was
 most connected to the laterobasal
@@
 The putative role of the PMC in continuous integration of environmental relevance
-and ensuing top-level control of action in the environment is supported
+and ensuing top-level control of action on the environment is supported
 by a number of neuroscience experiments.
@@
 \citep{simony2016dynamic}.
-%
+
 The retrosplenial portion of the PMC can subserve action possibilities
@@
 by monitoring the subjective outcomes
-of possible decisions and integrates that information
+of possible decisions and integrating that information
 with memory, perspective frames, and
 reward schedules into higher-level strategies.
-Perceived value that differs across individuals updates
+Perceived value, that differs across individuals, may update
 statistical assessment of the environment
@@
 in both the external environment and its internal representation
-that modulate strategic behavioral adjustment in volatile environments.
+to enable strategic behavior in volatile environments.
 
@@
 Analogous to the PMC,
-the dorsomedial PFC (dmPFC) is believed to subserve predominantly ambiguous processes
-across time, space, and content domains in
-sensory-independent top-down pathways.
-In contrast to the PMC, however,
-the dmPFC may be the DMN node whose function comes closest to a
+the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
+supra-sensory processes
+across time, space, and content domains to
+exert top-down control.
+In comparison to the PMC, however,
+the dmPFC may be closer to a
 ``mental sketchpad'' \citep{goldman1996prefrontal}, as it
@@
 patients with neurological lesions in the prefrontal cortex
-are known to adapt to novel situations and stimuli with difficulty
+are known to struggle with
+adaption to novel situations and stimuli
 \citep{stuss1986frontal}.
@@
 Specifically, neural activity in the human dmPFC
-reflected expectations about other peoples actions and errors thereof.
-dmPFC activity indeed explained the proficiency decline
+reflected expectations about other peoples' actions and
+errors in these predictions.
+Neural activity in the dmPFC indeed explained the proficiency decline
 of inferring other peoples' thoughts in the elderly \citep{moran2012social}.
@@
 According to computational neuroimaging experiments,
-the dmPFC activity preferentially models stimulus-value associations of
+dmPFC activity preferentially models stimulus-value associations of
 possible actions that are not actually executed,
@@
 
-In contrast to the dorsomedial prefrontal DMN,
+Comparing to the dorsomedial prefrontal DMN,
 the ventromedial portion of the prefrontal cortex (vmPFC) probably subserves
@@
 processing of “what matters now”,
-probably drawing on model-based value representations
+probably drawing on sophisticated value representations
 \citep{doherty2015structure}.
-Quantative lesion findings across 344 human individuals confirmed
-a gross impairment in value-based decision making
+Quantitative lesion findings across 344 human individuals confirmed
+a substantial impairment in value-based decision making
 \citep{glascher2012lesion}.
@@
 the vmPFC is preferentially connected with reward-related and limbic areas.
-The vmPFC has been observed to have monosynaptical connections
+The vmPFC is well known to have monosynaptical connections
 with the NAc
 in axonal tracing studies in monkeys \citep{haber1995orbital}.
-Additionally, the gray-matter volume of the vmPFC and NAc
+Congruently, the gray-matter volume of the vmPFC and NAc
 correlated with indices of value-guided behavior and reward attitudes
+in humans
 \citep{lebreton2009automatic}.
@@
 \citep{croxson2005quantitative} that
-quantified this brain regions to
+quantified the NAc to
 be substantially more likely connected to the vmPFC than dmPFC in both species.
@@
 the vmPFC with the NAc, hippocampus (HC),
-and posteromedial cortex.
+and PMC.
 %
@@
 was related to state encoding in vmPFC.
-Such a ``cognitive map'' of action space was argued to encode
+Such a ``cognitive map'' of the action space was argued to encode
 the current task state even when states are unobservable from sensory input,
@@
 brain region, to indices of
-social competence and social network complexity, which are
+social competence and social network complexity, which are probably
 among the most complicated decision that humans and monkeys take
 \citep{behrens2009computation}.
-%
 
 \subsection{The hippocampus: memory, space, and experience replay}
-The midline core of the DMN is probably in constant communication
+The DMN midline is probably in constant communication
 with the HC in the medial temporal lobe \citep{vincet2006}.
-The HC is long known to be involved in memory and spatial navigation in animals and humans.
-
-Its highly recursive anatomical architecture
-may be specifically designed to allow reconstructing
-entire episodes of experience from memory fragments.
-%
+The HC is long known to be involved in
+memory operations and spatial navigation in animals and humans.
+
 While the HC
@@
 \citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017}.
+Its highly recursive anatomical architecture
+may be specifically designed to allow reconstructing
+entire episodes of experience from memory fragments.
 Indeed,
@@
 some constantly active neuronal ensembles whose firing coincided with
-distinct locations in space while the animal navigated through its surroundings.
+distinct locations in space during environmental navigation.
 London taxi drivers, individuals with high performance in spatial navigation,
-were shown to exhibit increased grey matter volume in the
+were shown to exhibit increased gray-matter volume in the
 hippocampus \citep{maguire2000navigation}.
@@
 spike one after another when an animal is choosing between alternative
-paths \citep{johnson2007neural}; and
-these neuronal patterns appear to also indicate upcoming behavior
+paths \citep{johnson2007neural}.
+These neuronal patterns appear to also indicate upcoming behavior
 \citep{pfeiffer2013hippocampal}.
@@
 Moreover, spike sequences measured in hippocampal place cells of rats
-featured reoccurrence directly after experimental trials
+featured re-occurrence directly after experimental trials
 as well as directly before upcoming experimental trials \citep{diba2007forward}.
-Such hippocampal ensemble burts during rest and sleep
+Such hippocampal ensembles burt during rest and sleep
 have been proposed to be critical in communicating local information
@@
 These hippocampus-subserved mechanisms
-probably make important contributions to the
-recollection of autobiographical memory episodes and other
-reexperienced or newly generated mental scenarios
+probably contribute to
+recollecting autobiographical memory episodes, as well as
+reexperiencing or newly generating mental scenarios
 \citep{hassabis2007patients}.
 
-The HC thus orchestrates elements of experienced environmental aspects for
+The HC thus orchestrates experience of environmental aspects for
 consolidations based on reenactment and for integration into
@@
 network only after birth \citep{doria2010}.
-The TPJs are known to have hemispheric differences
+The TPJs are known to exhibit hemispheric differences
 according to their cytoarchitectonic borders and gyrification pattern
@@
 action initiation during externally structured tasks and
-sensorimotor control by integrating supramodal stimulus-guided attention
+sensorimotor control by integrating supra-modal stimulus-guided attention
 \citep{corbetta2002control}.
@@
 Importantly, the RTPJ has been shown to be intimately related to
-supramodal prediction and error signaling.
+supra-modal prediction and error signaling.
 It was argued that the RTPJ encodes actions and ensuing outcomes
@@
 sensory world and internally-oriented mind sets focussed
-on self-relevant mind wandering through mental scenarios.
+on self-relevant mental scene construction.
 Transient RTPJ in humans for instance disrupted the
@@
 a capacity believed to be subserved by the DMN.
-The RTPJ might hence be an important player that shifts away
+The RTPJ might hence be an important relay that shifts away
 from the ‘‘internally directed’’ baseline processes
-to deal instead with immediate environmental objects and contexts.
+to, instead, deal with environmental objects and events.
 % (TODO: slow endogeneous changes in state transitions by the policy and value matrices)
 
-The TPJ in the left hemisphere (LTPJ) in turn exhibits a close topographical relationship to Wernicke's area
-involved in the comprehension or understanding of written and spoken language.
+The TPJ in the left hemisphere (LTPJ), in turn, exhibits a close topographical relationship to Wernicke's area
+involved in understanding spoken and written language.
 Neurological patients with damage caused to Wernicke's area
@@
 without obvious relation to processing explicit language content
-\citep{varney1987locus}
+\citep{varney1987locus}.
 % (Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
@@
 the internal production of
-formulated thought ("language of the mind") was closely related to the left TPJ
+verbalized thought ("language of the mind") was closely related to the left TPJ
 in a multivariate analysis of brain volume
@@
 The LTPJ may hence facilitate the automated prediction of events
-by incorporating experience-derived models of the world
+by incorporating experience-derived representations of the world
 into ongoing action, planning, and problem solving.
@@
 \section{Reinforcement learning: a process model for DMN function}
-The outlined neurobiological properties and suspected functions of the DMN
-are argued to be sufficient
-for implementing all components
-of a full-fledged \textit{reinforcement learning (RL)} algorithm.
+We now argue the outlined neurobiological properties
+and potential functions of the DMN
+to be sufficient for implementing all components
+of a full-fledged RL algorithm.
 Integration of past experience, probabilistic mapping of candidate actions,
-random sampling of possible scenarios as well as
+random sampling of possible scenarios, as well as
 incorporation of instantaneous, delayed, and expected reward outcomes
@@
 
-RL is a problem-solving technique in which, through their interactions with an external environment,
+RL is a problem-solving technique in which, through interactions with an external environment,
 an agent learns to reach goals, optimize reward signals, or minimize costs
 in an iterative trial-by-error fashion.
-At a given moment, each taken \textit{action} $a$ triggers a changes in the
+At a given moment, each taken \textit{action} $a$ triggers a change in the
 \textit{state} of the environment
 $s \rightarrow s'$, accompanied by environmental feed signals as \textit{reward}
-$r = r(s, a)$ (or regret, since it can be negative) collected by the
+$r = r(s, a)$ (or regret \textbf{XXX exchange by punishment?},
+since it can be negative) collected by the
 agent.
@@
 is realistic in a computational model which sets out
-to explain DMN function. Indeed, it is known that the more the external world is predictable,
-the more mental activity becomes detached from the actual sensory inputs
-\citep{antrobus1966studies, pope1978regulation},
-the more stimulus-independent thoughts occur, and
-the higher the neural activity in the DMN.
-Conversely, the more the currently executed task is unknown and unpracticed,
-the lower the DMN activity
-\citep{filler1973daydreaming, teasdale1995stimulus}.
-%
-These ``offline'' processes may however contribute to
-optimizing control of the organism.
-Such a framework of DMN funtion based on reinforcement learning
+to explain DMN function.
+% XXX: the following passage is already in the introduction
+% Indeed, it is known that the more the external world is predictable,
+% the more mental activity becomes detached from the actual sensory inputs
+% \citep{antrobus1966studies, pope1978regulation},
+% the more stimulus-independent thoughts occur, and
+% the higher the neural activity in the DMN.
+% Conversely, the more the currently executed task is unknown and unpracticed,
+% the lower the DMN activity
+% \citep{filler1973daydreaming, teasdale1995stimulus}.
+% %
+% These ``offline'' processes may however contribute to
+% optimizing control of the organism.
+% Such a framework of DMN funtion based on reinforcement learning
+% can naturally embed human behavior
+% in the tension between exploitative action with immediate gains and
+% explorative action with longer-term reward schedules
+% \citep{dayan2008decision}.
+A framework of DMN funtion based on RL
 can naturally embed human behavior
@@
 \citep{dayan2008decision}.
-The implication of the DMN in a diversity of
+DMN implication in a diversity of
 particularly advanced cognitive processes
@@
 \subsection{Partially Observable Markov Decision Processes}
-It is our view that at a sufficiently coarse scale, the brain -- and therefore the DMN in particular --
-is a physical system governed by the laws of
-physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
+We propose that the brain in general and the DMN in particular 
+are a physical system governed by the laws of
+physics and can be locally described by Markov processes at a sufficiently coarse scale.
+Indeed, as noted in \cite{tegmark2016improved},
 any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
@@
 Since an action taken in the present moment $t$
-might have repercussions long into the future, it turns out that the
+might have repercussions in the far future, it turns out that the
 quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
@@
     }
-  The amygdala is known to be involved in significance evaluation while
-  the NAc, which is connected to the vmPFC, is involved in reward evaluation.
+  The amygdala is known to be involved in evaluation of biological significance,
+  whereas
+  the NAc is involved in reward evaluation. Both are known to
+  be connected to the vmPFC of the DMN.
 \end{mdframed}
@@
 % It is a fixed-point equation for the deterministic policy $\pi$.
-Notice that in \eqref{eq:bellman} the sampling is done only over things which
+Note that in \eqref{eq:bellman} the sampling is done only over things which
 depend on the environment, and so can be learned off-policy by observing state transitions
@@
 From a neurobiological perspective,
-experience replay would be manifested as the re-occurrence of a sequence of cell activations that also occur during specific actions.
+experience replay would be manifested as the re-occurrence of a
+neuron spiking sequences that also occur during specific actions.
 %, but the replay has a much faster time scale.
@@
 highest expected cumulative rewards.
-In POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible.
+In POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
+\textbf{XXX Elvis: refs needed.}.
 Much recent work points to Monte Carlo or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
@@
   Its involvement in thinking about the past,
-  the future, and hypothetical possibilities ties in with the implicit computation of
+  the future, and hypothetical possibilities
+  appears to tie in with the implicit computation of
   action and state cascades as a function of what happened in the past.
@@
 The HC may provide perturbed action-transition-state-reward samples
-  as batches of "imagined", "hypothesized", "recalled" experience.
+  as batches of imagined, hypothesized, recalled experience.
   The small variations in these experience samplings allow searching
@@
   While such a situation may not have been experienced in the physical world,
-  the DMN may have already stochastically generated, evaluated, and adapted to
-  such a randomly synthesized situation in the past.
+  the DMN may have previously stochastically generated, evaluated, and adapted to
+  such a randomly synthesized situation.
   In the absence of environmental input and feedback
-  (e.g., mind-wandering or sleep) pseudo-experiencing (i.e., emulating) possible
+  (e.g., mind-wandering or sleep),
+  mental scence construction allows for pseudo-experiencing possible
   future scenarios and action outcomes.
-  Our approach thus acknowledges the unavoidable stochasticity of
+  Our account of DMN function thus acknowledges the unavoidable stochasticity of
   computation in neuronal processes \citep{faisal2008noise}.
 
-  In our proposed model-free RL-based view, \textit{inference} in the human brain reduces to
+  From the perspective of our model-free RL agent,
+  \textit{inference} in the human brain reduces to
   generalization of
@@
   to novel, yet unexperienced environments, reward prediction errors will
-  massively increase \textbf{Please complete sentence}.
+  massively increase \textbf{Please check/simplify sentence}.
   This will lead to adaptation of the policy until the system converges to a
@@
 \section{Relation to other models decision-making and optimal control for the brain}
-\subsection{The free-energy principle and active inference}
+\subsection{The free-energy principle and active inference \
+textbf{move to Appendix}}
 % \subsubsection{The theory}
@@
 Neuroanatomically, axonal
-back projections probably outnumber by far the axonal input projections
+back projections indeed outnumber by far the axonal input projections
 existing in the primate brain \citep{salin1995corticocortical}.
@@
 as a generative experience due to the modulatory top-down influence at
-various hierarchical levels of sensory input processing.
+various stages of sensory input processing.
 In our RL view, instead,
@@
 In sum,
-the formal account may serve as a concrete implementation of
+the present formal account may serve as a concrete implementation of
 the long-standing predictive coding account from cognitive neuroscience.
 MPDs have the advantage of exposing an explicit
-mechanism of the horizon of future considerations or
+mechanism for the horizon of future considerations or
 how the internal model of the world gets updated,
@@
 %
-Recasting DMN function in terms of MDPs therefore naturally incorporate
+Recasting DMN function in terms of MDPs therefore naturally incorporates
 a majority of aspects of the prediction coding hypothesis.
@@
 The encoding of complex environments could thus be facilitated
-by relative similarity associations between states.
+by association to similar states.
 %
@@
 These neural processes are likely to
-build on abstract world knowledge and abstract assocations
+build on abstract world knowledge and logical assocations
 for binding the constituent elements in mental scene construction
@@
 to imply a distinction between
-environmental perception and internally generated mind-wandering
+engagement in the sensory environment
+and internally generated mind-wandering
 \citep{buckner2007self}.
@@
 agent's history of states and actions. The learned
-value matrix and action transition schedules drive
+value estimations and action transition schedules drive
 stimulus processing and action choice in the present.}
@@
 In sum, episodic scene construction according to the semantic
-hypothesis is lacking an explicit time and incentive model;
-neither does it explain how randomness of human mental experience
-can be beneficial.
+account is lacking an explicit time and incentive model.
 
@@
 social nature. This is probably unsurprising
-given that a likely human-defining property is
+given that 
 the complexity of the social systems
+is a likely human-defining property
 \citep{tomasello2009cultural}.
 According to the ``social brain hypothesis'',
-the human brain has not primarily evolved to solve problems of the
-physical environment, but to form and maintain increasingly complex
-social systems to solve ecological problems via social relationships
+the human brain has especially been shaped for
+forming and maintaining increasingly complex
+social systems,
+which allows solving ecological problems by social relationships
 \citep{whiten1988machiavellian}.
@@
 supposedly less advanced social-cognitive capacities.
-Instead,
-the formal account can parsimoneously incorporate
+Rather than functional specificity for processing social information,
+the present formal account can parsimoneously incorporate
 the dominance of social content in
-human mental activity by reexpressing relevance of information about humans
-in high value function estimates
+human mental activity 
+as high value function estimates for information about humans
 \citep{baker2009action, kampe2001psychology, krienen2010clan}.
@@
 on DMN biology is hence not only practically computable but
-also neuroscientifically falsifiable.
+also yields falsifiable neuroscientific hypotheses.
 At the least, we propose an alternative vocabulary to
@@
 from the relatively recent events to anticipated upcoming events
-in order to constantly improve present action in our dynamic world.
+in order to constantly improve action in our dynamic world.
 

### 2017-04-09 paper/paper.tex
some typos

--- 2017-04-09 paper/paper.tex
@@
 \subsection{Default mode network: higher-order control of the organism}
-The human DMN accounts for a large part of the brain's base line activity
-and typically decreases in activity during psychological experiments.
-The standard mode of
-information maintenance and manipulation in the brain has
-been argued to mediate evolutionarily conserved functions
-\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
+The DMN appeared to be exclusive in task-induced
+deactivation during various psychological experiments.
+This processing of unknown information categories in the DMN has
+been argued to mediate an evolutionarily conserved function for the individual \textbf{(XXX Guillaume: Ref?)}.
 This is all the more likely because the DMN contains hotspots
@@
 Today, many authors believe that the DMN implements
-probabilistic estimation of past, hypothetical, and
-future events
-\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
-randy2008}.
-This brain network
+probabilistic estimation of past, hypothetical, and future events \textbf{(XXX Guillaume: Ref? // many authors)}. This
+brain network
 might have emerged to continuously predict environmental events using
@@
 %
-Gaining insight into the overarching DMN function is challenging because
-this network may
-simultaneously control perception-action cycles and
-maintain baseline contemplations
-across time, space, and content domains.
+Understanding of the overarching DMN function is complicated by
+simultaneously controlling perception-action cycles and
+maintaining baseline contemplations
+across time, space, and content domains
+at the interface of the external world and the self \textbf{(XXX Guillaume: This sentence is not clear and too long.)}.
 
@@
 \citep{filler1973daydreaming, teasdale1995stimulus}.
-These ``offline'' processes may however contribute to optimized control of the organism.
-\textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment?
-- danilo: hmm, let's keeping thinking about this)}
+These ``offline'' processes may however contribute to optimized control of the organism. \textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment?)}
 % We formalize
@@
 
-\subsection{Towards a computational account of DMN function}
+\subsection{Sketch of our main contributions\textbf{(XXX Guillaume: What about ``Towards a computational account of DMN function'')}}
 We argue that the implication of the DMN in a
@@
 \citep{tenenbaum2011grow}.
-In this context,
-\textit{control} refers to the influences that an agent exerts when interacting
+\textit{Control} refers to the influences that an agent exerts when interacting
 with the environment to encourage preferred states.
@@
 neuroscientific experiments have invigorated this idea \citep{kording2004bayesian, fiser2004small}.
-This paper proposes a process model that satisfies this contention.
+This paper proposes a formal model that satisfies this contention.
 
@@
 \begin{equation}
-  y_k := r_k + \gamma \max_{a'} \tilde{Q}(s', a'|\theta_{k-1}),
+  y_k := y_k(s,a,s') =  r + \gamma \max_{a'} \tilde{Q}(s', a'|\theta_{k-1}),
 \end{equation}
-with the output of a parametrized regression model $(s,a)
-\mapsto \tilde{Q}(s, a|\theta_{k-1})$. Here the function $(s,a) \mapsto Q(a,s|\theta_{k-1})$
+with the prediction of a parametrized regression model $(s,a)
+\mapsto \tilde{Q}(s, a|\theta_{k-1})$. Here the function $(s,a) \mapsto \tilde{Q}(a,s|\theta_{k-1})$
 % \textbf{XXX just double-checking, are the dots int he Q() correct?}
@@
 $\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action pairs
-$(s,a)$, as proposed by \citep{songNIPS2016}.
+$(s,a)$, as proposed by \citep{songNIPS2016} for example.
 %  and Parr et al. 2003 (JMLR)].
@@
   \delta \theta_{k} \propto -\nabla_{\theta_{k}}\mathcal L(\theta_{k})
-  = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
+  = -\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
     \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{averseness}}].

### 2017-04-09 paper/paper.tex

@@
 $s \rightarrow s'$, accompanied by environmental feed signals as \textit{reward}
-$r = r(s, a)$ (or regret \textbf{XXX exchange by punishment?},
+$r = r(s, a)$ (or \textit{punishment},
 since it can be negative) collected by the

### 2017-04-10 paper/paper.tex
style++

@@
 previously proposed cognitive accounts as special cases and
-parsimoneously explains
+parsimoniously explains
 experimental findings in animals and humans.
@@
 oxygen and glucose energy to maintain
-potentially purposeful neuronal computation
+neuronal computation
 in the absence of behavioral focus
@@
 In the beginning of the 21st century,
-brain imaging may have been first technique
+brain imaging may have been the first technique
 to allow for the discovery of a unique brain network that subserves
@@
 
-\subsection{Towards a computational account of DMN function}
+\subsection{Towards a formal account of DMN function}
 We argue that the implication of the DMN in a
@@
 \citep{raichle2005intrinsic}.
-Numerous empirical evidence from
-neuroscientific experiments have invigorated this idea \citep{kording2004bayesian, fiser2004small}.
-This paper proposes a process model that satisfies this contention.
+This idea has been invigorated by
+empirical evidence from
+neuroscientific experiments \citep{kording2004bayesian, fiser2004small}.
+The present paper proposes a
+process model that satisfies this contention.
 
@@
 We begin by a neurobiological deconstruction of the DMN
-based on experimental findings in neuroscience.
-This walkthrough across individual network nodes will describe their mechanistic relevance, paving the way for their functional interpretation within our computational account in the next section.
+based on experimental findings in the neuroscience literature.
+This walkthrough across network nodes will
+outline their individual functional profiles,
+paving the way for their functional interpretation within our computational account
+described in the next section.
 
@@
 the medial prefrontal cortex (mPFC),
-are responsible for the highest metabolic turn-over
+are responsible for the highest turn-over
 of glucose energy consumption \citep{raichle2001pnas}.
@@
 be a sign of evolutionary sophistication \citep{flechsig1920}.
-Physiological and disturbed metabolic fluctuations in the
+Normal and disturbed metabolic fluctuations in the
 human PMC have been repeatedly related to
@@
 environmental statistics and its internal representation
-as an "inner minds eye" \citep{cavanna2006precuneus, leech_pcc2014}.
+as an inner "mind's eye" \citep{cavanna2006precuneus, leech_pcc2014}.
 For instance, B\'alint's syndrome is a neurological disorder of conscious
@@
 Patients are plagued by an
-inability to bind various individual features of the visual
+the inability to bind various individual features of the visual
 environment into an integrated whole (i.e., simultanagnosia)
@@
 (i.e., optic ataxia).
-This can be viewed as a high-level impairment in the gathering
-of information about alternative objects (i.e., exploration) as well as
+This can be viewed as a high-level impairment in gathering
+information about alternative objects (i.e., exploration) as well as
 leveraging these environmental opportunities (i.e., exploitation).
 Congruently,
-the human PMC was coupled in two functional connectivity modalities
+the human PMC was coupled in two functional connectivity analyses
 with the amygdala, involved in significance evaluation, and
@@
 
-The putative role of the PMC in continuous integration of environmental relevance
+The putative role of the PMC in continuous integration of
+relevant environmental information
 and ensuing top-level control of action on the environment is supported
@@
 while its retrosplenial portion was
-more specifically implicated in approach-avoidance behavior
+repeatedly implicated in approach-avoidance behavior
 \citep{vann2009does}.
@@
 whether a monkey would persue an exploratory or exploitative
-behavioral strategy in a complex food foraging task \citep{pearson2009neurons}.
+behavioral strategy during food foraging \citep{pearson2009neurons}.
 Further, single-cell recordings in the monkey PMC
@@
 subjectively perceived relevance of a chosen object
-than by its factual value, based on an internal currency of value.
+than by its factual value,
+based on an internal currency of value \citep{mccoy2005risk}.
 In fact, direct stimulation of PMC neurons
-promoted exploratory action towards options
-with unsafe reward outcomes that were previously shunned.
+promoted exploratory action towards actions
+that were previously shunned \citep{hayden2008stim}.
 Graded changes in firing rates of PMC neurons
@@
 to predict delayed reward opportunities in the future.
-In doing so, the PMC continuously adapts to changes
+In doing so, the PMC may continuously adapt to changes
 in both the external environment and its internal representation
@@
 possible actions that are not actually executed,
-whereas the vmPFC activity models value in the environment that leads performed actions
+whereas vmPFC activity models value in the environment that leads performed actions
 \citep{nicolle2012agent}.
@@
 the ventromedial portion of the prefrontal cortex (vmPFC) probably subserves
-less ambiguous subjective-value-related evaluative processes and
-reward-informed risk estimation of self-relevant environmental stimuli.
+subjective-value-related evaluative processes and
+risk estimation of self-relevant environmental stimuli.
 This DMN node is more closely associated with
@@
 be substantially more likely connected to the vmPFC than dmPFC in both species.
-Two functional connectivity modalities in humans strongly connected
+Two functional connectivity analyses in humans strongly connected
 the vmPFC with the NAc, hippocampus (HC),
@@
 reactions and own (visceral) arousal.
-Real or imaginged bodily states could be mapped in the vmPFC
+Real or imagined bodily states could be mapped in the vmPFC
 as a bioregulatory disposition governing cognition
@@
 This may be why performance within and across participants
-was related to state encoding in vmPFC.
+was related to state encoding in the vmPFC.
 Such a ``cognitive map'' of the action space was argued to encode
@@
 which was shown to be critical for behavior \citep{Schuck2016}.
-Additionally,
-independent whole-brain analyses from structural
-neuroimaging studies related the gray-matter volume of the vmPFC,
-more consistently than any other
-brain region, to indices of
-social competence and social network complexity, which are probably
-among the most complicated decision that humans and monkeys take
-\citep{behrens2009computation}.
+% Additionally,
+% independent whole-brain analyses from structural
+% neuroimaging studies related the gray-matter volume of the vmPFC,
+% more consistently than any other
+% brain region, to indices of
+% social competence and social network complexity, which are probably
+% among the most complicated decision that humans and monkeys take
+% \citep{behrens2009computation}.
 
@@
 particularly advanced cognitive processes
-can be parsimoneously explained as probabilistic simulations in form of mental scenes coupled with prediction error minimization
+can be parsimoniously explained as probabilistic simulations in form of mental scenes coupled with prediction error minimization
 to calibrate action portfolios and
@@
 Rather than functional specificity for processing social information,
-the present formal account can parsimoneously incorporate
+the present formal account can parsimoniously incorporate
 the dominance of social content in

### 2017-04-11 paper/paper.tex
more stuff

@@
 \subsection{The hippocampus: memory, space, and experience replay}
-The DMN midline is probably in constant communication
+The DMN midline is probably closely linked
 with the HC in the medial temporal lobe \citep{vincet2006}.
@@
 \citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017}.
-Its highly recursive anatomical architecture
+Its recursive anatomical architecture
 may be specifically designed to allow reconstructing
@@
 imagining new fictitious experiences more broadly \citep{hassabis2007patients}.
-Mental scenes created by hippocampus-lesioned patients exposed a lack of
+Mental scenes created by HC-lesioned patients exposed a lack of
 spatial integrity, richness in detail, and overall coherence.
 %
-Single-cell recordings in the animal hippocampus revealed
+Single-cell recordings in the animal HC revealed
 some constantly active neuronal ensembles whose firing coincided with
@@
 were shown to exhibit increased gray-matter volume in the
-hippocampus \citep{maguire2000navigation}.
-Indeed, place-specific neuronal populations in the hippocampus
+HC \citep{maguire2000navigation}.
+Indeed, place-specific neuronal populations in the HC
 spike one after another when an animal is choosing between alternative
@@
 
-But encoding and generative reconstruction in the hippocampus extends
+But encoding and generative reconstruction in the HC extends
 beyond mere spatial knowledge of the environment.
@@
 to the neocortex for long-term storage, potentially also in the nodes of the DMN.
-These hippocampus-subserved mechanisms
+These HC-subserved mechanisms
 probably contribute to

### 2017-04-12 paper/paper.tex
Add GD affiliations

@@
   %% firstname.lastname@inria.fr
+  %% GD1 Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France
+  %% GD2 CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France
+  %% GD3 University Paris Diderot, Sorbonne Paris Cité, Paris, France
+  %% GD4 Centre de Bioinformatique, Biostatistique et Biologie Intégrative (C3BI, USR 3756 Institut Pasteur et CNRS), Paris, France
 }
@@
 this brain network's highest energy consumption in the brain and
-its intimate coupling with conscious awareness
-\textbf{(XXX Guillaume: It escaladed quickly.)}.
+its intimate coupling with conscious awareness.
 %

### 2017-04-16 paper/paper.tex
stuff

--- 2017-04-11 paper/paper.tex
@@
 The DMN midline is probably closely linked
-with the HC in the medial temporal lobe \citep{vincet2006}.
-The HC is long known to be involved in
+with the HC in the medial temporal lobe \citep{vincet2006} --
+a region long known to be involved in
 memory operations and spatial navigation in animals and humans.
@@
 not only associated with an impairment in reexperiencing the past (i.e., amnesia),
-but also thinking about one's own future and
-imagining new fictitious experiences more broadly \citep{hassabis2007patients}.
+but also imagining about one's own future and
+ new fictitious experiences more broadly \citep{hassabis2007patients}.
 Mental scenes created by HC-lesioned patients exposed a lack of
@@
 HC \citep{maguire2000navigation}.
-Indeed, place-specific neuronal populations in the HC
-spike one after another when an animal is choosing between alternative
-paths \citep{johnson2007neural}.
+Indeed, when an animal is choosing between alternative
+paths, the corresponding neuronal populations in the HC
+spike one after another  \citep{johnson2007neural}.
 These neuronal patterns appear to also indicate upcoming behavior
@@
 rich mental scene construction \citep{deuker2016event, bird2010establishing}.
-In this way, the HC may even influence
+In this way, the HC may influence
 ongoing perception in the current environment
@@
 \subsection{The right and left TPJ: prediction error signaling and world semantics}
-While the DMN emerges with its midline structures early in human development,
+The DMN emerges with its midline structures early in human development.
+However,
 the right and left TPJ may become fully integrated into this canonical
@@
 The TPJs are known to exhibit hemispheric differences
-according to their cytoarchitectonic borders and gyrification pattern
+based on cytoarchitectonic borders and gyrification pattern
 \citep{seghier2013angular}.
@@
 % (Bzdok et al., 2013; Corbetta et al., 2008).
-Indeed, patients with right TPJ damage have particular difficulties
+Indeed, patients with RTPJ damage have particular difficulties
 with multistep actions \citep{hartmann2005takes}.
@@
 on self-relevant mental scene construction.
-Transient RTPJ in humans for instance disrupted the
+Transient RTPJ lesion in humans for instance diminished the
 impact of predicted intentions of other individuals
@@
 from the ‘‘internally directed’’ baseline processes
-to, instead, deal with environmental objects and events.
+to, instead, deal with (unexpected) environmental objects and events.
 % (TODO: slow endogeneous changes in state transitions by the policy and value matrices)
@@
 %
-Inner speech also hinges on knowledge retrieval of statistical structure
+Inner speech also hinges on knowledge retrieval
 about the physical and inter-personal world.
@@
 episodic memory recall and imagination strongly draw on
-complex world knowledge.
+reassembling world knowledge.
 Isolated building blocks of world statistics probably get reassembled
@@
 of a full-fledged RL algorithm.
-Integration of past experience, probabilistic mapping of candidate actions,
+Recalling past experience, considering candidate actions,
 random sampling of possible scenarios, as well as
-incorporation of instantaneous, delayed, and expected reward outcomes
+evaluating instantaneous, delayed, and expected reward outcomes
 are key components of intelligent RL agents
-that are plausible to cross-talk in the DMN.
+that are plausible to intertwine in the DMN.
 
 RL is a problem-solving technique in which, through interactions with an external environment,
-an agent learns to reach goals, optimize reward signals, or minimize costs
+an agent learns to reach goals, optimize reward signals, and minimize costs
 in an iterative trial-by-error fashion.
@@
 \textit{state} of the environment
-$s \rightarrow s'$, accompanied by environmental feed signals as \textit{reward}
+$s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
 $r = r(s, a)$ (or \textit{punishment},
@@
 that is, changing in random ways. In addition,
-it is only \textit{partially observed},
+it is only \textit{partially observed}
 in the sense that only part of the current state is observed by
 the agent \citep{starkweather2017dopamine}.
-The unpredictability of the environment
+The volatility of the environment
 is realistic in a computational model which sets out
@@
 % \citep{dayan2008decision}.
-A framework of DMN funtion based on RL
+We argue that a functional account of the DMN based on RL
 can naturally embed human behavior
@@
 We propose that the brain in general and the DMN in particular 
-are a physical system governed by the laws of
+are a natural system governed by the laws of
 physics and can be locally described by Markov processes at a sufficiently coarse scale.
@@
 any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
-step is sufficiently short
+step is sufficiently short.
 % , defining $s(t)$, the state at time $t$, as the position in state space.
@@
 
-In artificial intelligence (AI), a popular computational model for
+In artificial intelligence and machine learning, a popular computational model for
 multi-step decision processes in such an environment are
@@
 signal (i.e., is likely to be different for another agent in the same state).
-This is performed alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
+This is performed alongside each transition in the environments state space, and maintains a probability distribution over the set of possible
 states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
@@
 is a rather direct mapping of model-free RL learning algorithms
-onto the brain,
-with the neuromodulator dopamine serving
-as a teaching signal to train values or policies by controlling
+onto aspects of the brain.
+The neurotransmitter dopamine could serve
+as a ``teaching signal'' to better estimate value associations
+and action policies by controlling
 synaptic plasticity at targets such as the ventral and
 dorsolateral striatum.
-In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of world.
+In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of the world.
 These can be assumptions about the physical laws governing the agent's world or constraints on the state space and transitions between states.
@@
 \begin{definition}
+\textbf{XXX Danilo: Is there a definition 2 somewhere?}
 Mathematically, a POMDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
@@
 \subsection{Long-term rewards and action policies}
-The behavior of the agent is governed by some kind of \textit{policy},
-which maps states of the world to
+The behavior of the agent is governed by some kind of \textit{policy}.
+It maps states of the world to
 a set of candidate actions to perform given this state. Starting a time $t=0$,
@@
 \begin{mdframed}
-  \textbf{Time-invariance assumption:} It is assumed that the underlying MDP is stationary --i.e has reached steady-state conditions--
+  \textbf{Time-invariance assumption:} It is assumed that the underlying MDP is stationary -- i.e., has reached steady-state conditions --
   so that the dynamics of the process are equivalent over sufficiently long time windows of equal length.
@@
 "What is the expected utility of taking action $a$ in this situation ?".
-$Q^{\pi}(s,a)$ offers a formalization of optimal behavior,
-potentially driven by the DMN in human agents.
+$Q^{\pi}(s,a)$ offers a formalization of optimal behavior that
+may well reflect processing aspects of the DMN in human agents.
 
@@
 This is due to the enormous number of possible action
-trajectories the agent might need to
+trajectories that the agent often needs to
 consider simultaneously.
@@
 \textit{deterministic policies}, which map each state to the single best action to
-take at the state. Such policies define deterministic functions from states to actions.
+take at the current state. Such policies define deterministic functions from states to actions.
 
@@
 \end{equation}
-This consistency equation for the optimal (over all policies) Q-value function provides
+This consistency equation for the optimal (i.e., over all policies) Q-value function provides
 a recursive decomposition of optimal behavior by dividing the Q-value function into the immediate
@@
 is a fixed point for this equation. A complicated dynamic programming optimization can be
-broken into simpler sub-problems at different time points,
-which has often been closely related to the medial prefrontal part of the DMN
+broken into simpler sub-problems at different time points.
+Exploitation of hierarchical structure in actions considerations
+has previously been closely related to the medial prefrontal part of the DMN
 \citep{koechlin1999role, braver2002role}.
@@
 % deterministic policies of the form \eqref{eq:qlearning}.
-A successful RL theory needs ways to efficiently approximate the state-action value function
+A practically successful RL approach
+needs ways to efficiently approximate the state-action value function
 to allow for scaling to large state and action spaces.
@@
 experience replay would be manifested as the re-occurrence of a
-neuron spiking sequences that also occur during specific actions.
+neuron spiking sequence that also occurs during specific actions.
 %, but the replay has a much faster time scale.
@@
 \end{equation}
+\textbf{We still may need to explain y in the formula with one sentence.}
 For instance, a general linear model with a kernel $\phi$ would be of the
@@
 \subsubsection{Excursion: the role of the hippocampus}
-\paragraph{Does the hippocampus do Monte-Carlo sampling?}
+\paragraph{Does the hippocampus subserve Monte-Carlo sampling?}
 In RL, Monte-Carlo simulation can be used to update the agent's belief state
@@
 Much recent work points to Monte Carlo or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
-\textbf{XXX Elvis: refs needed in last sentence.} For sampling-based methods it is still almost non-existent beyond the description of receptive fields.
+\textbf{XXX Elvis: refs needed in last sentence.
+Plus, should be really bring up Bayesian?} For sampling-based methods it is still almost non-existent beyond the description of receptive fields.
 \textbf{XXX Elvis: rephrase last sentence}
@@
   \citep{carhart2010default, margulies2016situating}.
-  Its involvement in thinking about the past,
+  Its putative involvement in thinking about the past,
   the future, and hypothetical possibilities

### 2017-04-16 paper/paper.tex

@@
 We propose that the brain in general and the DMN in particular
-are a natural system governed by the laws of
+are a physical system governed by the laws of
 physics and can be locally described by Markov processes at a sufficiently coarse scale.

### 2017-04-17 paper/paper.tex
clarifying some things + integrating some comments

@@
 \newtheorem{definition}[theorem]{Definition}
+\newtheorem*{definition*}{Definition}
 \newtheorem{remark}{Remark}
@@
 % environments \citep{gershman2015computational}.
-
 In our present model-free RL framework,
@@
 
-\begin{definition}
-\textbf{XXX Danilo: Is there a definition 2 somewhere?}
+\begin{mdframed}
+\begin{definition*}
+% \textbf{XXX Danilo: Is there a definition 2 somewhere?}
 Mathematically, a POMDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
@@
 % \end{remark}
-\end{definition}
+\end{definition*}
+\end{mdframed}
 
@@
   % = \mathbb E_{(s, a, r, s') \sim \mathcal D}[(\tilde{Q}(s, a|\theta) - \tilde{Q}(s',a'|\theta^Q_{k}))^2]
-  = \mathbb E_{(s, a, r, s') \sim \mathcal D}\left[\frac{1}{2}(\tilde{Q}(s, a|\theta_{k}) - y_k)^2\right].
+  = \mathbb E_{(s, a, r, s') \sim \mathcal D}\left[\frac{1}{2}(\tilde{Q}(s, a|\theta_{k}) - y_k)^2\right],
   \label{eq:oracle}
 \end{equation}
-\textbf{We still may need to explain y in the formula with one sentence.}
+where $y_k$ is a defined in \eqref{eq:bellman}
+% \textbf{We still may need to explain y in the formula with one sentence.}
 For instance, a general linear model with a kernel $\phi$ would be of the
@@
 \section{Relation to other models decision-making and optimal control for the brain}
-\subsection{The free-energy principle and active inference \
-textbf{move to Appendix}}
-% \subsubsection{The theory}
-In Friston's free-energy principle (FEP) \citep{friston2010free,fristonAIorRL}, the brain is portrayed
-as biomechanical inference engine which must minimize the long-term average of surprise.
-Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net (DBN)}.
-According to FEP's account, the goal of the brain is then to optimize over the generative model $G$: to iteratively
-modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics, etc., so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles.
-These is called the \textit{generative} model.
-FEP also postulates a dual model called the \textit{recognition} model,
-which works in tandem with the generative model $p_R(\z|\x)$, to accomplish approximate inference.
-The recognition model dreams / fantasizes imaginary worlds $\z$ whilst the latter tries to generate sensations $\x$ which match these dreams!
-
-Because surprisal is intractably difficult to optimize (since we would need to sum over all hidden causes of the sensations), FEP sets out to instead minimize an upper-bound thereupon, namely the \text{free-energy} $F_G^R(\x)$ given by
-
-  \begin{equation}
-    \begin{split}
-      \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
-      &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
-      &\le F^R_G(\x),
-      \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
-    \end{split}
-  \end{equation}
-
-% \begin{figure}[!tb]
-%   \centering
-%   \includegraphics[width=.8\linewidth]{vae.pdf}
-%   \caption{Variational autoencoders...}
-%   \label{fig:vae}
-% \end{figure}
-The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}.
-\citep{dayan1995helmholtz}. As noted already in \citep{dayan1995helmholtz},
-a crucial drawback of the wake-sleep algorithm is that it postulates a pair forward (generation) and backward (recognition) models,
-which together do not correspond to optimization of (a bound of) the marginal likelihood, because of the incorrect KL used therein, etc.
-Thus the brain could not possibly be running such a algorithm, not even in principle! To the rescue, we note that the recent theory of
-\textit{variational auto-encoders (VAEs)} \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm,
-as it overcomes the technical limits of the former, by using a reparametrization trick. For instance, unlike the wake-sleep algorithm
-for minimizing free-energy, VAEs can be efficiently trained via back-propagation of learning errors.
-
-\subsubsection{Comparison to our proposed theory}
-On the surface, a common point between the FEP and our proposed RL-based framework
-  is placing the minimization of a surprise signal at the core of brain function.
-  Indeed in RL, surprise minimization is subsumed  by accurate prediction of
-  rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist undesired / harmful phase-transitions.
-  \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}. Such a theory cannot by itself,
-  explain the emergence of strategic behavior inherent in humans (e.g.,  cite dark-room experiment).
-
-% Friston's critic of RL \citep{fristonAIorRL} (he proposes ``active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
-% pioneered in the 1950s by Richard Bellman and colleagues [2]. To
-% optimise control a~pð Þ x~ under this formulation, we have to: (i)
-% assume the hidden states are available to the agent and (ii) solve
-% Equation 18 for the value-function. Solving for the value-function
-% is a non-trivial problem and usually involves backwards induction
-% or some approximation scheme like reinforcement learning [4–6].
-% The free-energy formulation circumvents these problems by
-% prescribing the policy in terms of free-energy, which encodes
-% optimal control (Equation 6)"
-
-% The following critics can be made:
-% \begin{itemize}
-%   \item
-%     As noted by Dayan et al.
-(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
-%   \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
-%   \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
-%     technical difficulty in getting derivatives of loss function w.r.t
-%     recognition weights $\W^R$).
-%   \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
-%   \end{itemize}
-
-\textbf{XXX: add tons of more examples and connnections}
-\textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
-\textbf{XXX: integrate aspects from WakeNsleep by Hinton}
-\textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
-\textbf{XXX: integrate aspects from Sutton/Barto book chapter}
-
 \subsection{Predictive coding hypothesis}
@@
 
+\subsection{The free-energy principle and active inference \textbf{XXX move to Appendix}}
+% \subsubsection{The theory}
+In Friston's free-energy principle (FEP) \citep{friston2010free,fristonAIorRL}, the brain is portrayed
+as biomechanical inference engine which must minimize the long-term average of surprise.
+Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net (DBN)}.
+According to FEP's account, the goal of the brain is then to optimize over the generative model $G$: to iteratively
+modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics, etc., so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles.
+These is called the \textit{generative} model.
+FEP also postulates a dual model called the \textit{recognition} model,
+which works in tandem with the generative model $p_R(\z|\x)$, to accomplish approximate inference.
+The recognition model dreams / fantasizes imaginary worlds $\z$ whilst the latter tries to generate sensations $\x$ which match these dreams!
+
+Because surprisal is intractably difficult to optimize (since we would need to sum over all hidden causes of the sensations), FEP sets out to instead minimize an upper-bound thereupon, namely the \text{free-energy} $F_G^R(\x)$ given by
+
+  \begin{equation}
+    \begin{split}
+      \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
+      &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
+      &\le F^R_G(\x),
+      \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
+    \end{split}
+  \end{equation}
+
+% \begin{figure}[!tb]
+%   \centering
+%   \includegraphics[width=.8\linewidth]{vae.pdf}
+%   \caption{Variational autoencoders...}
+%   \label{fig:vae}
+% \end{figure}
+The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}.
+\citep{dayan1995helmholtz}. As noted already in \citep{dayan1995helmholtz},
+a crucial drawback of the wake-sleep algorithm is that it postulates a pair forward (generation) and backward (recognition) models,
+which together do not correspond to optimization of (a bound of) the marginal likelihood, because of the incorrect KL used therein, etc.
+Thus the brain could not possibly be running such a algorithm, not even in principle! To the rescue, we note that the recent theory of
+\textit{variational auto-encoders (VAEs)} \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm,
+as it overcomes the technical limits of the former, by using a reparametrization trick. For instance, unlike the wake-sleep algorithm
+for minimizing free-energy, VAEs can be efficiently trained via back-propagation of learning errors.
+
+\subsubsection{Comparison to our proposed theory}
+On the surface, a common point between the FEP and our proposed RL-based framework
+  is placing the minimization of a surprise signal at the core of brain function.
+  Indeed in RL, surprise minimization is subsumed  by accurate prediction of
+  rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist undesired / harmful phase-transitions.
+  \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}. Such a theory cannot by itself,
+  explain the emergence of strategic behavior inherent in humans (e.g.,  cite dark-room experiment).
+
+% Friston's critic of RL \citep{fristonAIorRL} (he proposes ``active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
+% pioneered in the 1950s by Richard Bellman and colleagues [2]. To
+% optimise control a~pð Þ x~ under this formulation, we have to: (i)
+% assume the hidden states are available to the agent and (ii) solve
+% Equation 18 for the value-function. Solving for the value-function
+% is a non-trivial problem and usually involves backwards induction
+% or some approximation scheme like reinforcement learning [4–6].
+% The free-energy formulation circumvents these problems by
+% prescribing the policy in terms of free-energy, which encodes
+% optimal control (Equation 6)"
+
+% The following critics can be made:
+% \begin{itemize}
+%   \item
+%     As noted by Dayan et al.
+(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
+%   \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
+%   \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
+%     technical difficulty in getting derivatives of loss function w.r.t
+%     recognition weights $\W^R$).
+%   \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
+%   \end{itemize}
+
+\textbf{XXX: add tons of more examples and connnections}
+\textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
+\textbf{XXX: integrate aspects from WakeNsleep by Hinton}
+\textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
+\textbf{XXX: integrate aspects from Sutton/Barto book chapter}
+
 \section{Conclusion}

### 2017-04-17 paper/paper.tex
illustrating RL

@@
 in an iterative trial-by-error fashion.
-At a given moment, each taken \textit{action} $a$ triggers a change in the
-\textit{state} of the environment
+At a given moment, each taken \textit{action} $a$ triggers an \textit{partially observed} change
+in the \textit{state} of the environment
 $s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
@@
 since it can be negative) collected by the
-agent.
+agent. This is illustrated in Fig. \ref{fig:rl}.
 In this view, the environment is partially controlled by
@@
 an action.
+
+\begin{wrapfigure}{r}{.4\linewidth}
+  \centering
+  \includegraphics{aeloop-300x183.png}
+  \caption{RL in a nutshell}
+  \label{fig:rl}
+\end{wrapfigure}
+
 Action may be delayed to achieve substantial increases in expected reward

### 2017-04-17 paper/paper.tex
supra-pass part 1

--- 2017-04-16 paper/paper.tex
@@
 \author{Elvis Dohmatob, Guillaume Dumas, Danilo Bzdok\\
-  \textbf{Please add your affiliation}
-  %% INRIA, Parietal team, Saclay, France\\
-  %% CEA, Neurospin, Gif-sur-Yvette, France\\
-  %% firstname.lastname@inria.fr
-  %% GD1 Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France
-  %% GD2 CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France
-  %% GD3 University Paris Diderot, Sorbonne Paris Cité, Paris, France
-  %% GD4 Centre de Bioinformatique, Biostatistique et Biologie Intégrative (C3BI, USR 3756 Institut Pasteur et CNRS), Paris, France
+  INRIA, Parietal team, Saclay, France\\
+  CEA, Neurospin, Gif-sur-Yvette, France\\
+  Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
+  CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
+  University Paris Diderot, Sorbonne Paris Cité, Paris, France\\
+  Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
 }
@@
 %
-A process models for the neural implemention
-of continuous future predictions probably realized in the DMN
+A process model for the neural implemention
+of continuous future predictions potentially realized in the DMN
 could offer insight into the statistical mechanisms underlying human behavior.
@@
 consistently
-increased in neural activity during
+increase in neural activity during
 unfocused everyday mind wandering \citep{raichle2001pnas}.
-The DMN may exert higher-order control on human behavior
+The DMN may subserve important functions in
+higher-order control of human behavior
 following the notion of "a hierarchy of brain systems with
@@
 \subsection{Default mode network: higher-order control of the organism}
-The human DMN accounts for a large part of the brain's baseline activity
-and typically decreases in activity during psychological experiments.
+The human DMN is responsible for a large part of the brain's baseline activity,
+which typically decreases during psychological experiments.
 The standard mode of
@@
 \citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
-This is all the more likely because the DMN contains hotspots
+In line with this, the DMN contains hotspots
 of highest energy consumption in the entire central nervous system
-\citep{raichle2001pnas}.
-DMN activity also persists to a substantial degree during
+\citep{raichle2001pnas},
+which persists to a substantial degree during
 sleep and under anesthesia \citep{randy2008}.
@@
 %
-Gaining insight into the overarching DMN function is challenging because
+Gaining insight into the function of the DMN is
+particularly challenging because
 this network appears to
@@
 
-In this work,
-we consider an agent faced with the choice of the next action
+The present work takes the perspective of
+a human agent faced with the choice of the next action
 and being guided by rich
 reenactment of really-happened, hypothetically conceived, and
-predicted future events to perfect ongoing behavioral performance.
+estimated futures to optmize behavioral performance.
 A particularly attractive computational framework
@@
 \subsection{Towards a formal account of DMN function}
-We argue that the implication of the DMN in a
+We argue that DMN implication in a
 diversity of humans' most advanced cognitive processes
@@
 based on pervasive probabilistic simulations,
-thus maximizing reward outcome across multiple time scales.
+thus maximizing action outcome across multiple time scales.
 Such a purposeful optimization objective
 may be solved by a stochastic approximation
-based on a brain implementation of Markov Chain Monte Carlo sampling
+based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
 \citep{tenenbaum2011grow}.
 In this context,
-\textit{control} refers to the influences that an agent exerts when interacting
+\textit{control} refers to the influence that an agent exerts when interacting
 with the environment to encourage preferred states.
-Even (necessarily imperfect) memories, random day-time mind-wandering, and dreams during sleep
+Even (necessarily imperfect) memory retrieval,
+random day-time mind-wandering, and dreams during sleep
 may provide meaningful building blocks to iteratively improve
@@
 %
-In short, it has been proposed earlier that
+It has indeed been proposed earlier that
 the human brain's energy budget is largely dedicated to
@@
 This walkthrough across network nodes will
-outline their individual functional profiles,
-paving the way for their functional interpretation within our computational account
-described in the next section.
+outline the individual functional profiles,
+paving the way for their functional interpretation
+within our formal account (next section).
 
@@
 the medial prefrontal cortex (mPFC),
-are responsible for the highest turn-over
+are probably responsible for the highest turn-over
 of glucose energy consumption \citep{raichle2001pnas}.
 These metabolic characteristics go hand-in-hand with
-brain-imaging analyses \citep{andrews2010} in
+functional neuroimaging analyses \citep{andrews2010} in
 suggesting that the PMC and mPFC
@@
 human PMC have been repeatedly related to
-phenomena of changed conscious awareness \citep{cavanna2006precuneus},
-including anesthesia, sleep, and coma.
+phenomena of changed conscious awareness \citep{cavanna2006precuneus}.
 %
@@
 Patients are plagued by an
-the inability to bind various individual features of the visual
+inability to combine various individual features of the visual
 environment into an integrated whole (i.e., simultanagnosia)
-as well as inability to direct action towards
+as well as an inability to direct action towards
 currently unattended environmental objects
@@
 ``mental sketchpad'' \citep{goldman1996prefrontal}, as it
-potentially subserves the de-novo generation and binding
+potentially subserves the de-novo generation and manipulation
 of meaning representations instructed by stored semantics and memories.
@@
 may be solved by a stochastic approximation
-based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
+based on a brain implementation of MCMC sampling
 \citep{tenenbaum2011grow}.
@@
 build on abstract world knowledge and logical assocations
-for binding the constituent elements in mental scene construction
+for integrating the constituent elements in mental scene construction
 \citep{schacter2007remembering}.

### 2017-04-17 paper/paper.tex
supra-pass part 2

@@
 risk assessment \citep{mccoy2005risk},
-and outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
-while its retrosplenial portion was
-repeatedly implicated in approach-avoidance behavior
+outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
+as well as approach-avoidance behavior
 \citep{vann2009does}.
@@
 In fact, direct stimulation of PMC neurons
-promoted exploratory action towards actions
-that were previously shunned \citep{hayden2008stim}.
+promoted exploratory actions,
+which would otherwise be shunned \citep{hayden2008stim}.
 Graded changes in firing rates of PMC neurons
@@
 the DMN has been shown to gather and integrate information
-over different paragraphs in an fMRI study on auditory narratives
+over different parts of auditory narratives in an fMRI study
 \citep{simony2016dynamic}.
 
-The retrosplenial portion of the PMC can subserve action possibilities
-and evaluation of reward contingencies by integrating these with
+The retrosplenial portion of the PMC can subserve
+representation of action possibilities
+and evaluation of reward contingencies by integration with
 information from memory and different perspective frames.
@@
 \citep{vann2009does}.
-Regarding perspective frames, this PMC subregion has been
+Regarding perspective frames, the retrosplenial subregion of the PMC has been
 proposed to mediate between the organism's egocentric
@@
 In comparison to the PMC, however,
-the dmPFC may be closer to a
+dmPFC function may be closer to a
 ``mental sketchpad'' \citep{goldman1996prefrontal}, as it
@@
 Neural activity in the dmPFC indeed explained the proficiency decline
-of inferring other peoples' thoughts in the elderly \citep{moran2012social}.
+of inferring other peoples' thoughts in aging humans \citep{moran2012social}.
 Some dmPFC neurons in macaque monkeys exhibited a preference
@@
 %
-According to computational neuroimaging experiments,
-dmPFC activity preferentially models stimulus-value associations of
-possible actions that are not actually executed,
-whereas vmPFC activity models value in the environment that leads performed actions
+In a computational neuroimaging experiment,
+dorsomedial PFC activity preferentially modeled the consequences of
+action choices that were not actually executed,
+whereas ventromedial PFC (vmPFC) activity
+modeled especially value outcomes of performed actions in the environment
 \citep{nicolle2012agent}.
 
-Comparing to the dorsomedial prefrontal DMN,
-the ventromedial portion of the prefrontal cortex (vmPFC) probably subserves
-subjective-value-related evaluative processes and
-risk estimation of self-relevant environmental stimuli.
+Comparing to the dmPFC,
+the vmPFC probably subserves
+subjective value evaluation and
+risk estimation of relevant environmental stimuli.
 This DMN node is more closely associated with
@@
 \citep{schultz1998predictive}
-that not only modulate motivated behavior towards basic survival needs but also
+that not only channel action towards basic survival needs but also
 subserve RL in humans more broadly \citep{doherty2015structure}.
@@
 quantified the NAc to
-be substantially more likely connected to the vmPFC than dmPFC in both species.
+be more connected to the vmPFC than dmPFC in both species.
 Two functional connectivity analyses in humans strongly connected
@@
 %
-The vmPFC is often proposed to be involved in (external) emotional
-reactions and own (visceral) arousal.
+In line with these connectivity findings in animals and humans,
+the vmPFC is often proposed to be involved in (external) emotional
+reactions and own (visceral) arousal \citep{damasio1996somatic}.
 Real or imagined bodily states could be mapped in the vmPFC
 as a bioregulatory disposition governing cognition
-and decision making \citep{damasio1996somatic}
+and decision making.
 In neuroeconomic studies of human decision making,
@@
 which was shown to be critical for behavior \citep{Schuck2016}.
-% Additionally,
-% independent whole-brain analyses from structural
-% neuroimaging studies related the gray-matter volume of the vmPFC,
-% more consistently than any other
-% brain region, to indices of
-% social competence and social network complexity, which are probably
-% among the most complicated decision that humans and monkeys take
-% \citep{behrens2009computation}.
 

### 2017-04-17 paper/paper.tex
supra-pass part 3

@@
 with the environment to encourage preferred states.
-Even (necessarily imperfect) memory retrieval,
+Even (necessarily imperfect) memory recall,
 random day-time mind-wandering, and dreams during sleep
@@
 information from memory and different perspective frames.
-Regarding memory retrieval, retrosplenial lesions have been
+Regarding memory recall, retrosplenial lesions have been
 consistently associated with anterograde and retrograde memory impairments
@@
 generation of probabilistic internal information drawing from
-episodic memory retrieval, generative construction processes,
+episodic memory recall, generative construction processes,
 and explicit knowledge of the external world.
@@
 not only associated with an impairment in reexperiencing the past (i.e., amnesia),
-but also imagining about one's own future and
- new fictitious experiences more broadly \citep{hassabis2007patients}.
+but also imagining one's own future and
+fictitious experiences more broadly \citep{hassabis2007patients}.
 Mental scenes created by HC-lesioned patients exposed a lack of
@@
 some constantly active neuronal ensembles whose firing coincided with
-distinct locations in space during environmental navigation.
+specific locations in space during environmental navigation.
 London taxi drivers, individuals with high performance in spatial navigation,
@@
 
-But encoding and generative reconstruction in the HC extends
-beyond mere spatial knowledge of the environment.
+HC function probably extend beyond simple forms of
+encoding and reconstruction of memory and spatial information.
 Based on large-scale recordings of hippocampal neuronal populations,
@@
 after environmental events \citep{buzsaki2004large}.
-Specific spiking sequences, that were elicited in experimental task conditions,
+Specific spiking sequences, which were elicited in experimental task conditions,
 have been shown to be reenacted spontaneously during
@@
 These HC-subserved mechanisms
-probably contribute to
+probably contribute to advanced cognitive processes that require
 recollecting autobiographical memory episodes, as well as
@@
 rich mental scene construction \citep{deuker2016event, bird2010establishing}.
-In this way, the HC may influence
-ongoing perception in the current environment
+In this way, the HC may contribute to improving
+ongoing perception of and action on the environment
 \citep{maguire2016}.
@@
 \subsection{The right and left TPJ: prediction error signaling and world semantics}
-The DMN emerges with its midline structures early in human development.
+The DMN emerges with its midline structures early in human development
+\citep{doria2010}.
 However,
 the right and left TPJ may become fully integrated into this canonical
-network only after birth \citep{doria2010}.
+network only after birth.
 The TPJs are known to exhibit hemispheric differences
@@
 
-The TPJ in the right hemipshere (RTPJ) is central for
+The TPJ in the right hemipshere (RTPJ) 
+has been shown to be intimately related to
+supra-modal prediction and error signaling.
+It is probably central for
 action initiation during externally structured tasks and
@@
 multi-step action execution \citep{hartmann2005takes},
-visuo-proprioceptive conflict \citep{Balslev2005} and
+visuo-proprioceptive conflict \citep{Balslev2005}, and
 multi-modal detection of sensory changes across
@@
 %
-Importantly, the RTPJ has been shown to be intimately related to
-supra-modal prediction and error signaling.
 It was argued that the RTPJ encodes actions and ensuing outcomes
@@
 In the face of large discrepancies between actual and previously predicted
-environmental events the RTPJ acts a potential switch between
+environmental events the RTPJ acts as a potential switch between
 externally-oriented mind sets focussed on the
@@
 \citep{young2010disruption},
-a capacity believed to be subserved by the DMN.
+a capacity believed to be enabled by the DMN.
 The RTPJ might hence be an important relay that shifts away
@@
 and neuropsychology,
-the LTPJ probably mediates access to and integration of world knowlege
-to model action concepts
+the LTPJ probably mediates access to and integration of world knowlege,
+such as required during action considerations
 \citep{binder2011neurobiology, seghier2013angular}.
@@
 %
-Inner speech also hinges on knowledge retrieval
+Inner speech also hinges on knowledge recall
 about the physical and inter-personal world.
@@
 the internal production of
-verbalized thought ("language of the mind") was closely related to the left TPJ
+verbalized thought ("language of the mind") was closely related to the LTPJ
 in a multivariate analysis of brain volume
@@
 and MDP account provide explanations why hippocampus lesion does
-not only impair retrieving memories, but also hypothetical and future
+not only impair recalling memories, but also hypothetical and future
 thinking \citep{hassabis2007patients}.
@@
 poorly understood involvement across
-autobiographical memory retrieval, problem solving,
+autobiographical memory recall, problem solving,
 abstract reasoning based on internally generated scenes, social cognition,

### 2017-04-17 paper/paper.tex
supra-pass part 4

@@
 into the tension between exploitative action with immediate gains and
-exploratory action with longer-term reward schedules.
+exploratory action with longer-term reward outcomes.
 
@@
 with memory, perspective frames, and
-reward schedules into higher-level strategies.
+reward outcomes into higher-level strategies.
 Perceived value, that differs across individuals, may update
@@
 We now argue the outlined neurobiological properties
-and potential functions of the DMN
+and potential functions of the DMN nodes
 to be sufficient for implementing all components
@@
 are key components of intelligent RL agents
-that are plausible to intertwine in the DMN.
-
-RL is a problem-solving technique in which, through interactions with an external environment,
+that are plausible to intersect in the DMN.
+
+RL is a problem-solving technique in which, through interactions with an environment,
 an agent learns to reach goals, optimize reward signals, and minimize costs
@@
 that is, changing in random ways. In addition,
-it is only \textit{partially observed}
+the environment is only \textit{partially observed}
 in the sense that only part of the current state is observed by
@@
 in the tension between exploitative action with immediate gains and
-explorative action with longer-term reward schedules
+explorative action with longer-term reward outcomes
 \citep{dayan2008decision}.
@@
 particularly advanced cognitive processes
-can be parsimoniously explained as probabilistic simulations in form of mental scenes coupled with prediction error minimization
-to calibrate action portfolios and
-maximize reward outcome at different time scales.
+can be parsimoniously explained as probabilistic mental
+scene simulations coupled with prediction error minimization
+to calibrate action trajectories and
+maximize reward outcomes at different time scales.
 Such a purposeful optimization objective
@@
 \subsection{Partially Observable Markov Decision Processes}
-We propose that the brain in general and the DMN in particular
+We argue that the brain in general and the DMN in particular
 are a physical system governed by the laws of
@@
 
-Model-free RL is plausible in the human brain. Indeed,
+Model-free RL is plausible to be realizable in the human brain. Indeed,
 it has long been proposed \citep{dayan2008decision} that there
@@
 agent's history of states and actions. The learned
-value estimations and action transition schedules drive
+estimations of value and action transitions drive
 stimulus processing and action choice in the present.}

### 2017-04-18 paper/paper.tex
Integrate DM's comments in the abstract

@@
 %
-Many research streams converge to its likely role in evolutionarily adaptive
-envisioning of mental scenarios to simulate the future.
-Such functional roles become more likely due to
-this brain network's highest energy consumption in the brain and
-its intimate coupling with conscious awareness.
-%
-The present concept paper proposes a \textit{process model}
-in trying to explain \textit{how} the DMN may actually implement continuous
-assessment and prediction of the environment to guide action choices.
-DMN function is recast in the mathematical terms of
+Its highest energy consumption in the brain and
+its intimate coupling with conscious awareness are both pointing towards an important functional roles.
+%
+Many research streams support its evolutionarily adaptive role as
+envisioning mental scenarios to simulate the future.
+
+The present concept paper proposes to explain \textit{how} the DMN may actually implement such continuous assessment and prediction of the environment to guide action choices.
+
+DMN function is recast in mathematical terms of
 control theory and reinforcement learning based on Markov decision processes.
-We argue that our formal account of DMN function
-naturally accommodates
-previously proposed cognitive accounts as special cases and
-parsimoniously explains
-experimental findings in animals and humans.
+%
+We argue that our formal account of DMN function naturally accommodates as special cases
+previously proposed cognitive accounts such as \textit{predictive coding}.
 %
@@
 of continuous future predictions potentially realized in the DMN
-could offer insight into the statistical mechanisms underlying human behavior.
+could offer insight into the statistical mechanisms underlying human behavior and provide a parsimonious explanation of experimental findings in animals and humans.
 \end{abstract}

### 2017-04-18 paper/paper.tex
Authors affiliations styling & add ref for DMN and sleep/awareness

@@
 
-\author{Elvis Dohmatob, Guillaume Dumas, Danilo Bzdok\\
-  INRIA, Parietal team, Saclay, France\\
-  CEA, Neurospin, Gif-sur-Yvette, France\\
-  Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
-  CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
-  University Paris Diderot, Sorbonne Paris Cité, Paris, France\\
-  Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
+\author{Elvis Dohmatob\textsuperscript{1,2}, Guillaume Dumas\textsuperscript{3,4,5,6}, Danilo Bzdok\textsuperscript{1,2}
 }
@@
 \maketitle
+\small{
+    \textbf{1} INRIA, Parietal team, Saclay, France\\
+    \textbf{2} CEA, Neurospin, Gif-sur-Yvette, France\\
+    \textbf{3} Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
+    \textbf{4} CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
+    \textbf{5} University Paris Diderot, Sorbonne Paris Cité, Paris, France\\
+    \textbf{6} Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
+}
 
@@
 \citep{kenet2003spontaneously, fiser2004small}.
-The baseline energy demand is subject to surprisingly little modulations
-when mentally processing environmental challenges
-\citep{raichle2001pnas}. \textbf{(XXX Guillaume: What about the energy consumption compare to sleep? Danilo: I'd love to cite that
-but I am not aware of such studies)}
+The baseline energy demand is surprisingly weakly modulated
+with task demand \citep{raichle2001pnas}. At the opposite, during sleep, the
+decoupling of brain structures discarded the idea of the DMN being only a passive network resonance and rather supported its important role in the sustenance of conscious awareness \citep{horovitz2009decoupling}.
 This dark matter of brain physiology begs the question

### 2017-04-18 paper/paper.tex
Finish to integrate DM's comment + some linting

@@
 decoupling of brain structures discarded the idea of the DMN being only a passive network resonance and rather supported its important role in the sustenance of conscious awareness \citep{horovitz2009decoupling}.
-This dark matter of brain physiology begs the question
-of underlying biological purpose
+This dark matter of brain physiology strongly question
+its underlying biological purpose
 \citep{raichle2006brain}.
@@
 \citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
-These ``offline'' processes may however contribute to optimized control of the organism.
+Without requiring explicitly awareness, those ``offline'' processes may however contribute to optimized control of the organism.
 \textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment?
-- danilo: hmm, let's keeping thinking about this)}
+- danilo: hmm, let's keeping thinking about this - Guillaume: where shall we add this, more in the discussion, then?)}
 % We formalize
@@
 \textit{control} refers to the influence that an agent exerts when interacting
-with the environment to encourage preferred states.
+with the environment to encourage preferred states. \textbf{(XXX Guillaume: DM was asking to move this sentence up. Indeed, we may have to define early on the key concepts.}
 Even (necessarily imperfect) memory recall,

### 2017-04-18 paper/paper.tex
Add caption for Fig 1

@@
   \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
-  \caption{Neurobiological overview of the DMN ... \textbf{Please add a figure caption}}
+  \caption{Neurobiological overview of the DMN with its key neural structures and their associated behavioral functional roles within the current proposed model.
+\textbf{XXX Guillaume: Shall we also state explicitely that the functions indicated are only some among many potential others realized by those structures?}}
   \label{fig:neurobiological_overview_DMN}

### 2017-04-19 paper/paper.tex
illustrative rl.png

--- 2017-04-18 paper/paper.tex
@@
 
-\author{Elvis Dohmatob\textsuperscript{1,2}, Guillaume Dumas\textsuperscript{3,4,5,6}, Danilo Bzdok\textsuperscript{1,2}
+\author{Elvis Dohmatob, Guillaume Dumas, Danilo Bzdok\\
+  \textbf{Please add your affiliation}
+  %% INRIA, Parietal team, Saclay, France\\
+  %% CEA, Neurospin, Gif-sur-Yvette, France\\
+  %% firstname.lastname@inria.fr
+  %% GD1 Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France
+  %% GD2 CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France
+  %% GD3 University Paris Diderot, Sorbonne Paris Cité, Paris, France
+  %% GD4 Centre de Bioinformatique, Biostatistique et Biologie Intégrative (C3BI, USR 3756 Institut Pasteur et CNRS), Paris, France
 }
@@
 \maketitle
-\small{
-    \textbf{1} INRIA, Parietal team, Saclay, France\\
-    \textbf{2} CEA, Neurospin, Gif-sur-Yvette, France\\
-    \textbf{3} Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
-    \textbf{4} CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
-    \textbf{5} University Paris Diderot, Sorbonne Paris Cité, Paris, France\\
-    \textbf{6} Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
-}
 
@@
 %
-Its highest energy consumption in the brain and
-its intimate coupling with conscious awareness are both pointing towards an important functional roles.
-%
-Many research streams support its evolutionarily adaptive role as
-envisioning mental scenarios to simulate the future.
-
-The present concept paper proposes to explain \textit{how} the DMN may actually implement such continuous assessment and prediction of the environment to guide action choices.
-
-DMN function is recast in mathematical terms of
+Many research streams converge to its likely role in evolutionarily adaptive
+envisioning of mental scenarios to simulate the future.
+Such functional roles become more likely due to
+this brain network's highest energy consumption in the brain and
+its intimate coupling with conscious awareness.
+%
+The present concept paper proposes a \textit{process model}
+in trying to explain \textit{how} the DMN may actually implement continuous
+assessment and prediction of the environment to guide action choices.
+DMN function is recast in the mathematical terms of
 control theory and reinforcement learning based on Markov decision processes.
-%
-We argue that our formal account of DMN function naturally accommodates as special cases
-previously proposed cognitive accounts such as \textit{predictive coding}.
-%
-A process model for the neural implemention
-of continuous future predictions potentially realized in the DMN
-could offer insight into the statistical mechanisms underlying human behavior and provide a parsimonious explanation of experimental findings in animals and humans.
+We argue that our formal account of DMN function
+naturally accommodates
+previously proposed cognitive accounts as special cases and
+parsimoniously explains
+experimental findings in animals and humans.
+%
+A process models for the neural implemention
+of continuous future predictions probably realized in the DMN
+could offer insight into the statistical mechanisms underlying human behavior.
 \end{abstract}
@@
 \citep{kenet2003spontaneously, fiser2004small}.
-The baseline energy demand is surprisingly weakly modulated
-with task demand \citep{raichle2001pnas}. At the opposite, during sleep, the
-decoupling of brain structures discarded the idea of the DMN being only a passive network resonance and rather supported its important role in the sustenance of conscious awareness \citep{horovitz2009decoupling}.
+The baseline energy demand is subject to surprisingly little modulations
+when mentally processing environmental challenges
+\citep{raichle2001pnas}. \textbf{(XXX Guillaume: What about the energy consumption compare to sleep? Danilo: I'd love to cite that
+but I am not aware of such studies)}
 This dark matter of brain physiology begs the question
@@
 consistently
-increase in neural activity during
+increased in neural activity during
 unfocused everyday mind wandering \citep{raichle2001pnas}.
-The DMN may subserve important functions in
-higher-order control of human behavior
+The DMN may exert higher-order control on human behavior
 following the notion of "a hierarchy of brain systems with
@@
 \subsection{Default mode network: higher-order control of the organism}
-The human DMN is responsible for a large part of the brain's baseline activity,
-which typically decreases during psychological experiments.
+The human DMN accounts for a large part of the brain's baseline activity
+and typically decreases in activity during psychological experiments.
 The standard mode of
@@
 \citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
-In line with this, the DMN contains hotspots
+This is all the more likely because the DMN contains hotspots
 of highest energy consumption in the entire central nervous system
-\citep{raichle2001pnas},
-which persists to a substantial degree during
+\citep{raichle2001pnas}.
+DMN activity also persists to a substantial degree during
 sleep and under anesthesia \citep{randy2008}.
@@
 %
-Gaining insight into the function of the DMN is
-particularly challenging because
+Gaining insight into the overarching DMN function is challenging because
 this network appears to
@@
 
-The present work takes the perspective of
-a human agent faced with the choice of the next action
+In this work,
+we consider an agent faced with the choice of the next action
 and being guided by rich
 reenactment of really-happened, hypothetically conceived, and
-estimated futures to optmize behavioral performance.
+predicted future events to perfect ongoing behavioral performance.
 A particularly attractive computational framework
@@
 into the tension between exploitative action with immediate gains and
-exploratory action with longer-term reward outcomes.
+exploratory action with longer-term reward schedules.
 
 \subsection{Towards a formal account of DMN function}
-We argue that DMN implication in a
+We argue that the implication of the DMN in a
 diversity of humans' most advanced cognitive processes
@@
 based on pervasive probabilistic simulations,
-thus maximizing action outcome across multiple time scales.
+thus maximizing reward outcome across multiple time scales.
 Such a purposeful optimization objective
 may be solved by a stochastic approximation
-based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
+based on a brain implementation of Markov Chain Monte Carlo sampling
 \citep{tenenbaum2011grow}.
 In this context,
-\textit{control} refers to the influence that an agent exerts when interacting
+\textit{control} refers to the influences that an agent exerts when interacting
 with the environment to encourage preferred states.
-Even (necessarily imperfect) memory recall,
-random day-time mind-wandering, and dreams during sleep
+Even (necessarily imperfect) memories, random day-time mind-wandering, and dreams during sleep
 may provide meaningful building blocks to iteratively improve
@@
 %
-It has indeed been proposed earlier that
+In short, it has been proposed earlier that
 the human brain's energy budget is largely dedicated to
@@
 This walkthrough across network nodes will
-outline the individual functional profiles,
-paving the way for their functional interpretation
-within our formal account (next section).
+outline their individual functional profiles,
+paving the way for their functional interpretation within our computational account
+described in the next section.
 
@@
 the medial prefrontal cortex (mPFC),
-are probably responsible for the highest turn-over
+are responsible for the highest turn-over
 of glucose energy consumption \citep{raichle2001pnas}.
 These metabolic characteristics go hand-in-hand with
-functional neuroimaging analyses \citep{andrews2010} in
+brain-imaging analyses \citep{andrews2010} in
 suggesting that the PMC and mPFC
@@
 human PMC have been repeatedly related to
-phenomena of changed conscious awareness \citep{cavanna2006precuneus}.
+phenomena of changed conscious awareness \citep{cavanna2006precuneus},
+including anesthesia, sleep, and coma.
 %
@@
 Patients are plagued by an
-inability to combine various individual features of the visual
+the inability to bind various individual features of the visual
 environment into an integrated whole (i.e., simultanagnosia)
-as well as an inability to direct action towards
+as well as inability to direct action towards
 currently unattended environmental objects
@@
 risk assessment \citep{mccoy2005risk},
-outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
-as well as approach-avoidance behavior
+and outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
+while its retrosplenial portion was
+repeatedly implicated in approach-avoidance behavior
 \citep{vann2009does}.
@@
 In fact, direct stimulation of PMC neurons
-promoted exploratory actions,
-which would otherwise be shunned \citep{hayden2008stim}.
+promoted exploratory action towards actions
+that were previously shunned \citep{hayden2008stim}.
 Graded changes in firing rates of PMC neurons
@@
 the DMN has been shown to gather and integrate information
-over different parts of auditory narratives in an fMRI study
+over different paragraphs in an fMRI study on auditory narratives
 \citep{simony2016dynamic}.
 
-The retrosplenial portion of the PMC can subserve
-representation of action possibilities
-and evaluation of reward contingencies by integration with
+The retrosplenial portion of the PMC can subserve action possibilities
+and evaluation of reward contingencies by integrating these with
 information from memory and different perspective frames.
-Regarding memory recall, retrosplenial lesions have been
+Regarding memory retrieval, retrosplenial lesions have been
 consistently associated with anterograde and retrograde memory impairments
@@
 \citep{vann2009does}.
-Regarding perspective frames, the retrosplenial subregion of the PMC has been
+Regarding perspective frames, this PMC subregion has been
 proposed to mediate between the organism's egocentric
@@
 with memory, perspective frames, and
-reward outcomes into higher-level strategies.
+reward schedules into higher-level strategies.
 Perceived value, that differs across individuals, may update
@@
 In comparison to the PMC, however,
-dmPFC function may be closer to a
+the dmPFC may be closer to a
 ``mental sketchpad'' \citep{goldman1996prefrontal}, as it
-potentially subserves the de-novo generation and manipulation
+potentially subserves the de-novo generation and binding
 of meaning representations instructed by stored semantics and memories.
@@
 Neural activity in the dmPFC indeed explained the proficiency decline
-of inferring other peoples' thoughts in aging humans \citep{moran2012social}.
+of inferring other peoples' thoughts in the elderly \citep{moran2012social}.
 Some dmPFC neurons in macaque monkeys exhibited a preference
@@
 generation of probabilistic internal information drawing from
-episodic memory recall, generative construction processes,
+episodic memory retrieval, generative construction processes,
 and explicit knowledge of the external world.
 %
-In a computational neuroimaging experiment,
-dorsomedial PFC activity preferentially modeled the consequences of
-action choices that were not actually executed,
-whereas ventromedial PFC (vmPFC) activity
-modeled especially value outcomes of performed actions in the environment
+According to computational neuroimaging experiments,
+dmPFC activity preferentially models stimulus-value associations of
+possible actions that are not actually executed,
+whereas vmPFC activity models value in the environment that leads performed actions
 \citep{nicolle2012agent}.
 
-Comparing to the dmPFC,
-the vmPFC probably subserves
-subjective value evaluation and
-risk estimation of relevant environmental stimuli.
+Comparing to the dorsomedial prefrontal DMN,
+the ventromedial portion of the prefrontal cortex (vmPFC) probably subserves
+subjective-value-related evaluative processes and
+risk estimation of self-relevant environmental stimuli.
 This DMN node is more closely associated with
@@
 \citep{schultz1998predictive}
-that not only channel action towards basic survival needs but also
+that not only modulate motivated behavior towards basic survival needs but also
 subserve RL in humans more broadly \citep{doherty2015structure}.
@@
 quantified the NAc to
-be more connected to the vmPFC than dmPFC in both species.
+be substantially more likely connected to the vmPFC than dmPFC in both species.
 Two functional connectivity analyses in humans strongly connected
@@
 %
-In line with these connectivity findings in animals and humans,
-the vmPFC is often proposed to be involved in (external) emotional
-reactions and own (visceral) arousal \citep{damasio1996somatic}.
+The vmPFC is often proposed to be involved in (external) emotional
+reactions and own (visceral) arousal.
 Real or imagined bodily states could be mapped in the vmPFC
 as a bioregulatory disposition governing cognition
-and decision making.
+and decision making \citep{damasio1996somatic}
 In neuroeconomic studies of human decision making,
@@
 which was shown to be critical for behavior \citep{Schuck2016}.
+% Additionally,
+% independent whole-brain analyses from structural
+% neuroimaging studies related the gray-matter volume of the vmPFC,
+% more consistently than any other
+% brain region, to indices of
+% social competence and social network complexity, which are probably
+% among the most complicated decision that humans and monkeys take
+% \citep{behrens2009computation}.
 
@@
 not only associated with an impairment in reexperiencing the past (i.e., amnesia),
-but also imagining one's own future and
-fictitious experiences more broadly \citep{hassabis2007patients}.
+but also imagining about one's own future and
+ new fictitious experiences more broadly \citep{hassabis2007patients}.
 Mental scenes created by HC-lesioned patients exposed a lack of
@@
 some constantly active neuronal ensembles whose firing coincided with
-specific locations in space during environmental navigation.
+distinct locations in space during environmental navigation.
 London taxi drivers, individuals with high performance in spatial navigation,
@@
 
-HC function probably extend beyond simple forms of
-encoding and reconstruction of memory and spatial information.
+But encoding and generative reconstruction in the HC extends
+beyond mere spatial knowledge of the environment.
 Based on large-scale recordings of hippocampal neuronal populations,
@@
 after environmental events \citep{buzsaki2004large}.
-Specific spiking sequences, which were elicited in experimental task conditions,
+Specific spiking sequences, that were elicited in experimental task conditions,
 have been shown to be reenacted spontaneously during
@@
 These HC-subserved mechanisms
-probably contribute to advanced cognitive processes that require
+probably contribute to
 recollecting autobiographical memory episodes, as well as
@@
 rich mental scene construction \citep{deuker2016event, bird2010establishing}.
-In this way, the HC may contribute to improving
-ongoing perception of and action on the environment
+In this way, the HC may influence
+ongoing perception in the current environment
 \citep{maguire2016}.
@@
 \subsection{The right and left TPJ: prediction error signaling and world semantics}
-The DMN emerges with its midline structures early in human development
-\citep{doria2010}.
+The DMN emerges with its midline structures early in human development.
 However,
 the right and left TPJ may become fully integrated into this canonical
-network only after birth.
+network only after birth \citep{doria2010}.
 The TPJs are known to exhibit hemispheric differences
@@
 
-The TPJ in the right hemipshere (RTPJ)
-has been shown to be intimately related to
-supra-modal prediction and error signaling.
-It is probably central for
+The TPJ in the right hemipshere (RTPJ) is central for
 action initiation during externally structured tasks and
@@
 multi-step action execution \citep{hartmann2005takes},
-visuo-proprioceptive conflict \citep{Balslev2005}, and
+visuo-proprioceptive conflict \citep{Balslev2005} and
 multi-modal detection of sensory changes across
@@
 %
+Importantly, the RTPJ has been shown to be intimately related to
+supra-modal prediction and error signaling.
 It was argued that the RTPJ encodes actions and ensuing outcomes
@@
 In the face of large discrepancies between actual and previously predicted
-environmental events the RTPJ acts as a potential switch between
+environmental events the RTPJ acts a potential switch between
 externally-oriented mind sets focussed on the
@@
 \citep{young2010disruption},
-a capacity believed to be enabled by the DMN.
+a capacity believed to be subserved by the DMN.
 The RTPJ might hence be an important relay that shifts away
@@
 and neuropsychology,
-the LTPJ probably mediates access to and integration of world knowlege,
-such as required during action considerations
+the LTPJ probably mediates access to and integration of world knowlege
+to model action concepts
 \citep{binder2011neurobiology, seghier2013angular}.
@@
 %
-Inner speech also hinges on knowledge recall
+Inner speech also hinges on knowledge retrieval
 about the physical and inter-personal world.
@@
 the internal production of
-verbalized thought ("language of the mind") was closely related to the LTPJ
+verbalized thought ("language of the mind") was closely related to the left TPJ
 in a multivariate analysis of brain volume
@@
 We now argue the outlined neurobiological properties
-and potential functions of the DMN nodes
+and potential functions of the DMN
 to be sufficient for implementing all components
@@
 are key components of intelligent RL agents
-that are plausible to intersect in the DMN.
-
-RL is a problem-solving technique in which, through interactions with an environment,
+that are plausible to intertwine in the DMN.
+
+RL is a problem-solving technique in which, through interactions with an external environment,
 an agent learns to reach goals, optimize reward signals, and minimize costs
@@
   \centering
-  \includegraphics{aeloop-300x183.png}
+  \includegraphics{rl.png}
   \caption{RL in a nutshell}
@@
 that is, changing in random ways. In addition,
-the environment is only \textit{partially observed}
+it is only \textit{partially observed}
 in the sense that only part of the current state is observed by
@@
 in the tension between exploitative action with immediate gains and
-explorative action with longer-term reward outcomes
+explorative action with longer-term reward schedules
 \citep{dayan2008decision}.
@@
 particularly advanced cognitive processes
-can be parsimoniously explained as probabilistic mental
-scene simulations coupled with prediction error minimization
-to calibrate action trajectories and
-maximize reward outcomes at different time scales.
+can be parsimoniously explained as probabilistic simulations in form of mental scenes coupled with prediction error minimization
+to calibrate action portfolios and
+maximize reward outcome at different time scales.
 Such a purposeful optimization objective
 may be solved by a stochastic approximation
-based on a brain implementation of MCMC sampling
+based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
 \citep{tenenbaum2011grow}.
@@
 \subsection{Partially Observable Markov Decision Processes}
-We argue that the brain in general and the DMN in particular
+We propose that the brain in general and the DMN in particular
 are a physical system governed by the laws of
@@
 
-Model-free RL is plausible to be realizable in the human brain. Indeed,
+Model-free RL is plausible in the human brain. Indeed,
 it has long been proposed \citep{dayan2008decision} that there
@@
 build on abstract world knowledge and logical assocations
-for integrating the constituent elements in mental scene construction
+for binding the constituent elements in mental scene construction
 \citep{schacter2007remembering}.
@@
 and MDP account provide explanations why hippocampus lesion does
-not only impair recalling memories, but also hypothetical and future
+not only impair retrieving memories, but also hypothetical and future
 thinking \citep{hassabis2007patients}.
@@
 agent's history of states and actions. The learned
-estimations of value and action transitions drive
+value estimations and action transition schedules drive
 stimulus processing and action choice in the present.}
@@
 poorly understood involvement across
-autobiographical memory recall, problem solving,
+autobiographical memory retrieval, problem solving,
 abstract reasoning based on internally generated scenes, social cognition,

### 2017-04-19 paper/paper.tex
hot verses

--- 2017-04-18 paper/paper.tex
@@
 
-\author{Elvis Dohmatob\textsuperscript{1,2}, Guillaume Dumas\textsuperscript{3,4,5,6}, Danilo Bzdok\textsuperscript{1,2}
+\author{Elvis Dohmatob\textsuperscript{1,2}, Guillaume Dumas\textsuperscript{4,5,6,7}, Danilo Bzdok\textsuperscript{1,2,3}
 }
@@
     \textbf{2} CEA, Neurospin, Gif-sur-Yvette, France\\
-    \textbf{3} Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
-    \textbf{4} CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
-    \textbf{5} University Paris Diderot, Sorbonne Paris Cité, Paris, France\\
-    \textbf{6} Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
+    \textbf{3} Department of Psychiatry, Psychotherapy and Psychosomatics, RWTH Aachen University, Aachen, Germany\\
+    \textbf{4} Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
+    \textbf{5} CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
+    \textbf{6} University Paris Diderot, Sorbonne Paris Cité, Paris, France\\
+    \textbf{7} Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
 }
@@
 Its highest energy consumption in the brain and
-its intimate coupling with conscious awareness are both pointing towards an important functional roles.
-%
-Many research streams support its evolutionarily adaptive role as
+its intimate coupling with conscious awareness are both pointing towards
+an important functional role.
+%
+Many research streams support an evolutionarily adaptive role in
 envisioning mental scenarios to simulate the future.
 
-The present concept paper proposes to explain \textit{how} the DMN may actually implement such continuous assessment and prediction of the environment to guide action choices.
-
+The present paper proposes a \textit{process model}
+that tries to explain \textit{how}
+the DMN may implement
+continuous assessment and prediction of the environment to guide action choices.
+%
 DMN function is recast in mathematical terms of
@@
 We argue that our formal account of DMN function naturally accommodates as special cases
-previously proposed cognitive accounts such as \textit{predictive coding}.
-%
-A process model for the neural implemention
-of continuous future predictions potentially realized in the DMN
-could offer insight into the statistical mechanisms underlying human behavior and provide a parsimonious explanation of experimental findings in animals and humans.
+previously proposed cognitive accounts on
+\textit{i}) predictive coding,
+\textit{ii}) semantic association, and
+\textit{iii}) a ``radar'' function.
+%
+A process model for the neural optimization of complex behavior
+can offer parsimonious explanations for
+various experimental findings in animals and humans.
 \end{abstract}
@@
 
-\tableofcontents
-\listoffigures
-\listoftables
+% \tableofcontents
+% \listoffigures
+% \listoftables
 
@@
 When left unperturbed, the human brain is not at rest.
-Rather, the brain continues to metabolize large quantities of
+Especially the so-called "default mode network" (DMN) continues
+to metabolize large quantities of
 oxygen and glucose energy to maintain
@@
 with task demand \citep{raichle2001pnas}. At the opposite, during sleep, the
-decoupling of brain structures discarded the idea of the DMN being only a passive network resonance and rather supported its important role in the sustenance of conscious awareness \citep{horovitz2009decoupling}.
-This dark matter of brain physiology strongly question
-its underlying biological purpose
+decoupling of brain structures discarded the idea of the DMN being
+only a passive network resonance and rather supported its
+important role in sustaining conscious awareness \citep{horovitz2009decoupling}.
+This dark matter of brain physiology begs the question
+of the biological purpose underlying the DMN
 \citep{raichle2006brain}.
@@
 found a potential neurobiological manifestation
-in the so-called "default mode network" (DMN)
+in the DMN
 \citep{shul1997, raichle2001pnas}.
-This set of some of the highest regions in
+This set of some of the most advanced regions in
 the human association cortex \citep{mesulam1998sensation, margulies2016situating}
@@
 
-\subsection{Default mode network: higher-order control of the organism}
-The human DMN is responsible for a large part of the brain's baseline activity,
+\subsection{Towards a formal account of default mode function: higher-order control of the organism}
+The human DMN is responsible for a large part of the baseline neural activity,
 which typically decreases during psychological experiments.
@@
 \citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
-In line with this, the DMN contains hotspots
-of highest energy consumption in the entire central nervous system
-\citep{raichle2001pnas},
-which persists to a substantial degree during
-sleep and under anesthesia \citep{randy2008}.
-Today, many authors believe that the DMN implements
+Today, many investigators believe that the DMN implements some form of
 probabilistic estimation of past, hypothetical, and
@@
 The present work takes the perspective of
-a human agent faced with the choice of the next action
-and being guided by rich
-reenactment of really-happened, hypothetically conceived, and
-estimated futures to optmize behavioral performance.
+a human agent faced with the choice of the next actions
+and being guided by outcome estimations
+of really-happened, hypothetically conceived, and
+expected futures to optimize behavioral performance.
 A particularly attractive computational framework
@@
 \citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
-Without requiring explicitly awareness, those ``offline'' processes may however contribute to optimized control of the organism.
-\textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment?
-- danilo: hmm, let's keeping thinking about this - Guillaume: where shall we add this, more in the discussion, then?)}
-% We formalize
-% a policy matrix to capture the space of possible actions that can be performed
-% on the environment given the current state. A value matrix
-% maps expected rewards to environmental objects and events.
-% Switching between states reduces to a sequential processing model.
-% In a model-based reinforcement learning (RL) approach,
-% we try to understand the environment always better
-% starting, at birth, with an environment that is completely unknown.
+Without requiring explicit awareness,
+these ``offline'' processes may however contribute
+to optimizing control of the organism.
+We formalize
+a policy matrix to capture the space of possible actions that can be performed
+on the environment given the current state. A value function
+maps expected rewards to environmental objects and events.
+Switching between states reduces to a sequential processing model.
 Informed by outcomes of performed action,
@@
 shaped by prediction error.
-A DMN framework incorporating RL can naturally embed human behavior
+A DMN account incorporating RL can naturally embed human behavior
 into the tension between exploitative action with immediate gains and
@@
 
-\subsection{Towards a formal account of DMN function}
 We argue that DMN implication in a
@@
 
-\section{Relation to other models decision-making and optimal control for the brain}
+\section{Relation to existing cognitive accounts on default mode function}
 \subsection{Predictive coding hypothesis}
@@
 \citep{buckner2007self}.
-These authors stated that
+These investigators stated that
 ``A computational model [...] will probably require a form of

### 2017-04-19 paper/paper.tex
major pass on RL sections

--- 2017-04-18 paper/paper.tex
@@
 
-RL is a problem-solving technique in which, through interactions with an environment,
+RL is a problem-solving technique in which, through interactions with an unknown \textit{environment / world},
 an agent learns to reach goals, optimize reward signals, and minimize costs
-in an iterative trial-by-error fashion.
+in an iterative trial-by-error fashion. See Fig. \ref{fig:rl}.
 At a given moment, each taken \textit{action} $a$ triggers an \textit{partially observed} change
@@
 
-Model-free RL is plausible to be realizable in the human brain. Indeed,
+Model-free RL is plausibly realizable in the human brain. Indeed,
 it has long been proposed \citep{dayan2008decision} that there
@@
 
-\subsection{The Q-values, state-values, and the advantage function}
-We can now formulate the purpose of the DMN
-as finding a policy $\pi$ for the human agent that maximizes the
-expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
+\subsection{Expected values, Bellman equation, and optimal control}
+Given only the limited information available from a POMDP, at a state $s$ the average quality or usefulness of taking an action $a$ under a policy $\pi$ is completely captured by the single number
+% We can now formulate the purpose of the DMN
+% as finding a policy $\pi$ for the human agent that maximizes the
+% expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
 \begin{equation}
   \label{eq:qval}
-  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a].
+  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a],
 \end{equation}
-In other words,
-the Q-value $Q^{\pi}(s,a)$ corresponds to the expected reward
+called the $Q$-value for the state-action pair $(s,a)$.
+In other words, $Q^{\pi}(s,a)$ corresponds to the expected reward
 over all possible action trajectories, in which
 the agent sets out in the environment in state
-$s$, chooses action $a$,
-and then follows the policy $\pi$ to select future actions.
+$s$, chooses action $a$, and then follows the policy $\pi$ to select future actions.
 %% Given that each $\pi$ correspondence to an exhaustive description of
@@
 
-One can also define the state-value function
-
-\begin{eqnarray}
-  \V^\pi(s) := \mathbb E_{a \sim \pi}Q^\pi(s,a) = \mathbb E [G^\pi|s_0=s],
-\end{eqnarray}
-which measures the value of each state $s \in \mathcal S$ under the policy $\pi$. This serves
-as a baseline value for any action. Finally, the advantage function is defined by
-\begin{equation}
-A^\pi(s,a) :=  Q^\pi(s,a) - V^\pi(s).
-\end{equation}
+% One can also define the state-value function
+
+% \begin{eqnarray}
+%   \V^\pi(s) := \mathbb E_{a \sim \pi}Q^\pi(s,a) = \mathbb E [G^\pi|s_0=s],
+% \end{eqnarray}
+% which measures the value of each state $s \in \mathcal S$ under the policy $\pi$. This serves
+% as a baseline value for any action. Finally, the advantage function is defined by
+% \begin{equation}
+% A^\pi(s,a) :=  Q^\pi(s,a) - V^\pi(s).
+% \end{equation}
 
@@
 
-\subsection{Efficient computation of optimal behavior}
-By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general.
-This is due to the enormous number of possible action
-trajectories that the agent often needs to
-consider simultaneously.
-% \begin{mdframed}
-%   \textbf{N.B.:}
-%   If the Bellman-based computational ideas have a corresponding
-%   neurobiological implementation, this provides a framework of how
-%   optimal behavior could be computable in the human brain.
-% \end{mdframed}
-One popular solution is called \text{Q-learning}. It remedies
-the issue by only considering the sub-family of
-\textit{deterministic policies}, which map each state to the single best action to
-take at the current state. Such policies define deterministic functions from states to actions.
-
-\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992) generates a
-\textit{greedy policy}
-\textbf{XXX please add 1-2 sentences why this is interesting in the current
-context. This statement appears a little alone as it stands.}
+% \subsection{Efficient computation of optimal behavior}
+% By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general.
+% This is due to the enormous number of possible action
+% trajectories that the agent often needs to
+% consider simultaneously.
+% % \begin{mdframed}
+% %   \textbf{N.B.:}
+% %   If the Bellman-based computational ideas have a corresponding
+% %   neurobiological implementation, this provides a framework of how
+% %   optimal behavior could be computable in the human brain.
+% % \end{mdframed}
+% One popular solution is called \text{Q-learning}. It remedies
+% the issue by only considering the sub-family of
+% \textit{deterministic policies}, which map each state to the single best action to
+% take at the current state. Such policies define deterministic functions from states to actions.
+
+% \subsubsection{The Bellman equation:  characterizing optimal behavior}
+% Q-learning (Watkins \& Dayan, 1992) generates a
+% \textit{greedy policy}
+Optimal behavior corresponds to a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. This can be formalized by
+% \textbf{XXX please add 1-2 sentences why this is interesting in the current
+% context. This statement appears a little alone as it stands.}
 \begin{equation}
@@
 \end{equation}
-which maps each state $s$ to the optimal action which should be taken at that state.
-Here $Q^*(s, a) := \max_{\pi}Q^\pi(s,a)$, the optimal Q-value over all policies.
+where
+\begin{equation}
+  Q^*(s, a) := \max_{\pi}Q^\pi(s,a),
+\end{equation}
+the optimal Q-value over all policies.
+The goal of learning is to approach the policy $\pi^*$ as ``close as possible'', that is to \text{solve} the POMDP.
+Of course \eqref{eq:qlearning} is just a definition and
+represents by no means, a candidate schema for solving POMDPs with even moderately-sized action and state spaces: it's intractable!
 % where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
 % table.
-The \textit{Bellman equation} \citep{sutton1998reinforcement} then takes the simple form
+Fortunately, the \textit{Bellman equation} \citep{sutton1998reinforcement} provides a fixed-point relation which defines $Q^*$ implicitly via a sampling procedure, without the intractable problem
+over all every policy, viz
+\begin{equation}
+  Q^* = \text{Bel}(Q^*),
+  \label{eq:bellman}
+\end{equation}
+where the ``Bellman transform'' $\text{Bel}(Q)$ of an arbitrary $Q$-value function $Q: \mathcal S \times \mathcal A \rightarrow \mathbb R$  is defined by setting
 \begin{equation}
   \begin{split}
-    Q^*(s_0, a_0) &= \mathbb E_{s_1 \sim p(s_1|s_0,a)} [r(s,a) + \gamma \max_{a_1 \in \mathcal A}Q^*(s_1, a_1)]
-    \\
-    &= \mathbb E_{s_1 \sim p(s_1|s_0,a_0)} [r(s_0,a_0) + \gamma Q^*(s_1, \pi(s_1))].
+   \text{Bel}(Q)(s,a) &:=
+   \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q(s', a')]\\
+   &= r(s,a) + \mathbb E_{s' \sim p(s'|s,a)} [\gamma \max_{a' \in \mathcal A}Q(s', a')]\\
+   &= \text{instantaneous reward} + \text{expected reward for acting greedily thereafter}
+%    \; \forall (s,a) \in \mathcal S \times \mathcal A
+    % \\&=
+    % \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^*(s', \pi^*(s'))].
     \end{split}
-  \label{eq:bellman}
   \end{equation}
-This consistency equation for the optimal (i.e., over all policies) Q-value function provides
-a recursive decomposition of optimal behavior by dividing the Q-value function into the immediate
-reward and the discounted rewards of the upcoming states. The optimal Q-value operator $Q^*$
-is a fixed point for this equation. A complicated dynamic programming optimization can be
-broken into simpler sub-problems at different time points.
-Exploitation of hierarchical structure in actions considerations
+The Bellman equation \eqref{eq:bellman} is a ``self-consistency'' equation which provides
+a recursive decomposition of optimal behavior by dividing the $Q$-value function into the immediate
+reward and the discounted rewards of the upcoming states.
+% The optimal Q-value operator $Q^*$
+% is a fixed point for this equation.
+As a consequence, the complicated dynamic programming optimization problem \eqref{eq:qlearning}
+is broken into simpler sub-problems at different time points.
+Exploitation of hierarchical structure in actions consideration
 has previously been closely related to the medial prefrontal part of the DMN
@@
 
-\subsubsection{Efficient learning via value approximation and experience replay}
+\begin{mdframed}
+  \textbf{ NCs for the Bellman equation:}
+  ...
+\end{mdframed}
+
+\subsubsection{Value approximation and greedy action consideration, and experience replay}
 %As already mentioned in the previous section, Q-learning optimizes over the class of
 % deterministic policies of the form \eqref{eq:qlearning}.
-A practically successful RL approach
-needs ways to efficiently approximate the state-action value function
-to allow for scaling to large state and action spaces.
-In a state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
-% (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
-by sampling from its current policy, and collects a reward $r$,
+At a time step $k-1$, the world is in a state $s \in \mathcal S$, and the agent greedily takes an
+action which it expects to be the must valuable on average, namely
+
+\begin{equation}
+  \pi^{\text{greedy}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a|\theta),
+  \label{eq:policy}
+\end{equation}
+by querying an approximate version $a \mapsto \tilde{Q}(a,s|\theta)$ of the $Q$-value function ``frozen'' at state $s$.
+In the DMN, The dmPFC (BA9) would implement such a greedy lookup over the action space.
+The tensor $\theta$ encodes the model
+  parameters and would correspond to synaptic weights and / or connection strengths between
+  DMN regions.
+It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the greedy policy \eqref{eq:policy}.
+Learning consists in rewriting this program on the fly, so that the approximating functional $\tilde{Q}(.,.|\theta)$ becomes closer and closer to the optimum $Q^*$.
+
+The action $a$ taken in \eqref{eq:policy} triggers a reward $r$ collected by the agent,
 % (measured on their internal currency of worldly values ?),
@@
 $\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
-At iteration $k+1$, replay then consists in sampling (uniform or importance-weighted
+At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1} + \delta \theta_{k}$ of the parameters for its approximate model of $Q$-value functional. This warrants learning,
+and a loss function is in demand. The Bellman equation \eqref{eq:bellman}
+provides a way to produce such a loss function \eqref{eq:oracle} as we now outline below.
+
+% A practically successful RL approach
+% needs ways to efficiently approximate the state-action value function
+% to allow for scaling to large state and action spaces.
+Replay consists in sampling (uniform or importance-weighted
 \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
@@
 with the prediction of a parametrized regression model $(s,a)
-\mapsto \tilde{Q}(s, a|\theta_{k-1})$. Here the function $(s,a) \mapsto \tilde{Q}(a,s|\theta_{k-1})$
-% \textbf{XXX just double-checking, are the dots int he Q() correct?}
-is an approximator for the $Q$-value operator,  parameterized by weights $\theta_{k}$
-(synaptic strengths, etc.).
-
+\mapsto \tilde{Q}(s, a|\theta_{k-1})$.
 From a neurobiological perspective,
@@
 
-Computing an optimal policy then corresponds to finding the parameters $\theta_{k}$ which minimize the following mean-squared loss function
+\subsubsection{Stochastic gradient descent in the DMN}
+At the current iteration $k$, computing an optimal parameter update then corresponds to finding the parameters $\theta_{k}$ which minimize the following mean-squared loss function
 \begin{equation}
@@
 where reward prediction errors (RPEs) -- also known as \textit{regret}-- percolate
-from lower to higher processing layers to modulate the choice of future actions
+from lower to higher processing layers to modulate the choice of future actions.
+Over time (i.e in the limit $k \rightarrow \infty$), the approximate
+$Q$-values $\tilde{Q}(a,s|\theta_{k-1})$ provably\footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
+converge uniformly to the unique fixed-point $Q^*(a,s)$ defined implicitly by the Bellman
+equation \eqref{eq:bellman}
+
 % online stochastic gradient descent is used to minimize the variational objective

### 2017-04-19 paper/paper.tex
more

@@
 % \textit{greedy policy}
-Optimal behavior corresponds to a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. This can be formalized by
+Optimal behavior corresponds to a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. Formally,
 % \textbf{XXX please add 1-2 sentences why this is interesting in the current
@@
 \begin{equation}
-  \pi^*(s) = \argmax_{a \in \mathcal A}{Q}^*(s, a),
+  \pi^*(s) := \argmax_{a \in \mathcal A}{Q}^*(s, a),
   \label{eq:qlearning}
@@
 % table.
-Fortunately, the \textit{Bellman equation} \citep{sutton1998reinforcement} provides a fixed-point relation which defines $Q^*$ implicitly via a sampling procedure, without the intractable problem
-over all every policy, viz
+Fortunately, the \textit{Bellman equation} \citep{sutton1998reinforcement} provides a fixed-point relation which defines $Q^*$ implicitly via a sampling procedure, without querying the entire space of policies, viz
 \begin{equation}

### 2017-04-19 paper/paper.tex
interm

--- 2017-04-19 paper/paper.tex
@@
 When left unperturbed, the human brain is not at rest.
+In the beginning of the 21st century,
+brain imaging may have been the first technique
+to allow for the discovery of a unique brain network that subserves
+baseline mental activities
+\citep{raichle2001pnas, bzdok2015resting}.
 Especially the so-called "default mode network" (DMN) continues
@@
 cortex" \citep{carhart2010default}.
-In the beginning of the 21st century,
-brain imaging may have been the first technique
-to allow for the discovery of a unique brain network that subserves
-baseline mental activities
-\citep{raichle2001pnas, bzdok2015resting}.
 
@@
 been shown to directly impact human behavior. Goal-directed task performance
-improved with decreased activity in default mode areas \citep{weiss2006}
+improved with decreased activity in default mode regions \citep{weiss2006}
 and increased DMN activity was linked to more task-independent,
@@
 A particularly attractive computational framework
-to describe, quantify, and predict autonomously acting systems like the brain
+to describe, quantify, and predict intelligent systems, such as the brain,
 is proposed to be the combination of control theory and
@@
 Without requiring explicit awareness,
-these ``offline'' processes may however contribute
+these ``offline'' processes may contribute
 to optimizing control of the organism.
@@
 empirical evidence from
-neuroscientific experiments \citep{kording2004bayesian, fiser2004small}.
+neuroscience experiments \citep{kording2004bayesian, fiser2004small}.
 The present paper proposes a
@@
 Indeed,
-the vmPFC is preferentially connected with reward-related and limbic areas.
+the vmPFC is preferentially connected with reward-related and limbic regions.
 The vmPFC is well known to have monosynaptical connections
@@
 
-RL is a problem-solving technique in which, through interactions with an unknown \textit{environment / world},
+RL is a problem-solving technique in which, through interactions with an environment,
 an agent learns to reach goals, optimize reward signals, and minimize costs
-in an iterative trial-by-error fashion. See Fig. \ref{fig:rl}.
+in an iterative trial-by-error fashion.
 At a given moment, each taken \textit{action} $a$ triggers an \textit{partially observed} change
@@
   \centering
-  \includegraphics{rl.png}
+  \includegraphics{aeloop-300x183.png}
   \caption{RL in a nutshell}
@@
 
-Model-free RL is plausibly realizable in the human brain. Indeed,
+Model-free RL is plausible to be realizable in the human brain. Indeed,
 it has long been proposed \citep{dayan2008decision} that there
@@
 
-\subsection{Expected values, Bellman equation, and optimal control}
-Given only the limited information available from a POMDP, at a state $s$ the average quality or usefulness of taking an action $a$ under a policy $\pi$ is completely captured by the single number
-% We can now formulate the purpose of the DMN
-% as finding a policy $\pi$ for the human agent that maximizes the
-% expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
+\subsection{The Q-values, state-values, and the advantage function}
+We can now formulate the purpose of the DMN
+as finding a policy $\pi$ for the human agent that maximizes the
+expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
 \begin{equation}
   \label{eq:qval}
-  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a],
+  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a].
 \end{equation}
-called the $Q$-value for the state-action pair $(s,a)$.
-In other words, $Q^{\pi}(s,a)$ corresponds to the expected reward
+In other words,
+the Q-value $Q^{\pi}(s,a)$ corresponds to the expected reward
 over all possible action trajectories, in which
 the agent sets out in the environment in state
-$s$, chooses action $a$, and then follows the policy $\pi$ to select future actions.
+$s$, chooses action $a$,
+and then follows the policy $\pi$ to select future actions.
 %% Given that each $\pi$ correspondence to an exhaustive description of
@@
 
-% One can also define the state-value function
-
-% \begin{eqnarray}
-%   \V^\pi(s) := \mathbb E_{a \sim \pi}Q^\pi(s,a) = \mathbb E [G^\pi|s_0=s],
-% \end{eqnarray}
-% which measures the value of each state $s \in \mathcal S$ under the policy $\pi$. This serves
-% as a baseline value for any action. Finally, the advantage function is defined by
-% \begin{equation}
-% A^\pi(s,a) :=  Q^\pi(s,a) - V^\pi(s).
-% \end{equation}
+One can also define the state-value function
+
+\begin{eqnarray}
+  \V^\pi(s) := \mathbb E_{a \sim \pi}Q^\pi(s,a) = \mathbb E [G^\pi|s_0=s],
+\end{eqnarray}
+which measures the value of each state $s \in \mathcal S$ under the policy $\pi$. This serves
+as a baseline value for any action. Finally, the advantage function is defined by
+\begin{equation}
+A^\pi(s,a) :=  Q^\pi(s,a) - V^\pi(s).
+\end{equation}
 
@@
 
-% \subsection{Efficient computation of optimal behavior}
-% By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general.
-% This is due to the enormous number of possible action
-% trajectories that the agent often needs to
-% consider simultaneously.
-% % \begin{mdframed}
-% %   \textbf{N.B.:}
-% %   If the Bellman-based computational ideas have a corresponding
-% %   neurobiological implementation, this provides a framework of how
-% %   optimal behavior could be computable in the human brain.
-% % \end{mdframed}
-% One popular solution is called \text{Q-learning}. It remedies
-% the issue by only considering the sub-family of
-% \textit{deterministic policies}, which map each state to the single best action to
-% take at the current state. Such policies define deterministic functions from states to actions.
-
-% \subsubsection{The Bellman equation:  characterizing optimal behavior}
-% Q-learning (Watkins \& Dayan, 1992) generates a
-% \textit{greedy policy}
-Optimal behavior corresponds to a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. This can be formalized by
-% \textbf{XXX please add 1-2 sentences why this is interesting in the current
-% context. This statement appears a little alone as it stands.}
+\subsection{Efficient computation of optimal behavior}
+By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general.
+This is due to the enormous number of possible action
+trajectories that the agent often needs to
+consider simultaneously.
+% \begin{mdframed}
+%   \textbf{N.B.:}
+%   If the Bellman-based computational ideas have a corresponding
+%   neurobiological implementation, this provides a framework of how
+%   optimal behavior could be computable in the human brain.
+% \end{mdframed}
+One popular solution is called \text{Q-learning}. It remedies
+the issue by only considering the sub-family of
+\textit{deterministic policies}, which map each state to the single best action to
+take at the current state. Such policies define deterministic functions from states to actions.
+
+\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992) generates a
+\textit{greedy policy}
+\textbf{XXX please add 1-2 sentences why this is interesting in the current
+context. This statement appears a little alone as it stands.}
 \begin{equation}
@@
 \end{equation}
-where
-\begin{equation}
-  Q^*(s, a) := \max_{\pi}Q^\pi(s,a),
-\end{equation}
-the optimal Q-value over all policies.
-The goal of learning is to approach the policy $\pi^*$ as ``close as possible'', that is to \text{solve} the POMDP.
-Of course \eqref{eq:qlearning} is just a definition and
-represents by no means, a candidate schema for solving POMDPs with even moderately-sized action and state spaces: it's intractable!
+which maps each state $s$ to the optimal action which should be taken at that state.
+Here $Q^*(s, a) := \max_{\pi}Q^\pi(s,a)$, the optimal Q-value over all policies.
 % where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
 % table.
-Fortunately, the \textit{Bellman equation} \citep{sutton1998reinforcement} provides a fixed-point relation which defines $Q^*$ implicitly via a sampling procedure, without the intractable problem
-over all every policy, viz
+The \textit{Bellman equation} \citep{sutton1998reinforcement} then takes the simple form
 \begin{equation}
-  Q^* = \text{Bel}(Q^*),
+  \begin{split}
+    Q^*(s_0, a_0) &= \mathbb E_{s_1 \sim p(s_1|s_0,a)} [r(s,a) + \gamma \max_{a_1 \in \mathcal A}Q^*(s_1, a_1)]
+    \\
+    &= \mathbb E_{s_1 \sim p(s_1|s_0,a_0)} [r(s_0,a_0) + \gamma Q^*(s_1, \pi(s_1))].
+  \end{split}
   \label{eq:bellman}
 \end{equation}
-where the ``Bellman transform'' $\text{Bel}(Q)$ of an arbitrary $Q$-value function $Q: \mathcal S \times \mathcal A \rightarrow \mathbb R$  is defined by setting
-\begin{equation}
-  \begin{split}
-   \text{Bel}(Q)(s,a) &:=
-   \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q(s', a')]\\
-   &= r(s,a) + \mathbb E_{s' \sim p(s'|s,a)} [\gamma \max_{a' \in \mathcal A}Q(s', a')]\\
-   &= \text{instantaneous reward} + \text{expected reward for acting greedily thereafter}
-%    \; \forall (s,a) \in \mathcal S \times \mathcal A
-    % \\&=
-    % \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^*(s', \pi^*(s'))].
-    \end{split}
-  \end{equation}
-The Bellman equation \eqref{eq:bellman} is a ``self-consistency'' equation which provides
-a recursive decomposition of optimal behavior by dividing the $Q$-value function into the immediate
-reward and the discounted rewards of the upcoming states.
-% The optimal Q-value operator $Q^*$
-% is a fixed point for this equation.
-As a consequence, the complicated dynamic programming optimization problem \eqref{eq:qlearning}
-is broken into simpler sub-problems at different time points.
-Exploitation of hierarchical structure in actions consideration
+This consistency equation for the optimal (i.e., over all policies) Q-value function provides
+a recursive decomposition of optimal behavior by dividing the Q-value function into the immediate
+reward and the discounted rewards of the upcoming states. The optimal Q-value operator $Q^*$
+is a fixed point for this equation. A complicated dynamic programming optimization can be
+broken into simpler sub-problems at different time points.
+Exploitation of hierarchical structure in actions considerations
 has previously been closely related to the medial prefrontal part of the DMN
@@
 
-\begin{mdframed}
-  \textbf{ NCs for the Bellman equation:}
-  ...
-\end{mdframed}
-
-\subsubsection{Value approximation and greedy action consideration, and experience replay}
+\subsubsection{Efficient learning via value approximation and experience replay}
 %As already mentioned in the previous section, Q-learning optimizes over the class of
 % deterministic policies of the form \eqref{eq:qlearning}.
-At a time step $k-1$, the world is in a state $s \in \mathcal S$, and the agent greedily takes an
-action which it expects to be the must valuable on average, namely
-
-\begin{equation}
-  \pi^{\text{greedy}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a|\theta),
-  \label{eq:policy}
-\end{equation}
-by querying an approximate version $a \mapsto \tilde{Q}(a,s|\theta)$ of the $Q$-value function ``frozen'' at state $s$.
-In the DMN, The dmPFC (BA9) would implement such a greedy lookup over the action space.
-The tensor $\theta$ encodes the model
-  parameters and would correspond to synaptic weights and / or connection strengths between
-  DMN regions.
-It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the greedy policy \eqref{eq:policy}.
-Learning consists in rewriting this program on the fly, so that the approximating functional $\tilde{Q}(.,.|\theta)$ becomes closer and closer to the optimum $Q^*$.
-
-The action $a$ taken in \eqref{eq:policy} triggers a reward $r$ collected by the agent,
+A practically successful RL approach
+needs ways to efficiently approximate the state-action value function
+to allow for scaling to large state and action spaces.
+In a state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
+% (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
+by sampling from its current policy, and collects a reward $r$,
 % (measured on their internal currency of worldly values ?),
@@
 $\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
-At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1} + \delta \theta_{k}$ of the parameters for its approximate model of $Q$-value functional. This warrants learning,
-and a loss function is in demand. The Bellman equation \eqref{eq:bellman}
-provides a way to produce such a loss function \eqref{eq:oracle} as we now outline below.
-
-% A practically successful RL approach
-% needs ways to efficiently approximate the state-action value function
-% to allow for scaling to large state and action spaces.
-Replay consists in sampling (uniform or importance-weighted
+At iteration $k+1$, replay then consists in sampling (uniform or importance-weighted
 \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
@@
 with the prediction of a parametrized regression model $(s,a)
-\mapsto \tilde{Q}(s, a|\theta_{k-1})$.
+\mapsto \tilde{Q}(s, a|\theta_{k-1})$. Here the function $(s,a) \mapsto \tilde{Q}(a,s|\theta_{k-1})$
+% \textbf{XXX just double-checking, are the dots int he Q() correct?}
+is an approximator for the $Q$-value operator,  parameterized by weights $\theta_{k}$
+(synaptic strengths, etc.).
+
 From a neurobiological perspective,
@@
 
-\subsubsection{Stochastic gradient descent in the DMN}
-At the current iteration $k$, computing an optimal parameter update then corresponds to finding the parameters $\theta_{k}$ which minimize the following mean-squared loss function
+Computing an optimal policy then corresponds to finding the parameters $\theta_{k}$ which minimize the following mean-squared loss function
 \begin{equation}
@@
 where reward prediction errors (RPEs) -- also known as \textit{regret}-- percolate
-from lower to higher processing layers to modulate the choice of future actions.
-Over time (i.e in the limit $k \rightarrow \infty$), the approximate
-$Q$-values $\tilde{Q}(a,s|\theta_{k-1})$ provably\footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
-converge uniformly to the unique fixed-point $Q^*(a,s)$ defined implicitly by the Bellman
-equation \eqref{eq:bellman}
-
+from lower to higher processing layers to modulate the choice of future actions
 % online stochastic gradient descent is used to minimize the variational objective
@@
 gaps between fresh sensory input and expectations computed
-from stored prior information.
+based on stored prior information.
 At each stage of neural processing,
-an internally generated prediction of aspects of sensory sensation is
+an internally generated prediction of aspects of environmental sensations is
 directly compared against the actual environmental input.
@@
 may be maintained by top-down modulation
-by a-priori information about the environment.
-
-Predictive coding processes
-permit updates of the internal representation of the environment
+using a-priori information about the environment.
+
+In short, predictive coding processes
+conceptualize updates of the internal representation of the environment
 to best accommodate and prepare the organism
 for the constant influx of sensory examples and
-for reacting to the environment.
+performing action on the environment.
 There are hence a number of common properties
 between the predictive coding account
-and the proposed formal account.
+and the proposed formal account based on MDPs.
 Importantly,
@@
 Further,
-both frameworks are plausibilized by neuroscientific evidence that suggest
+both accounts are plausibilized by neuroscientific evidence that suggest
 the human brain to be a ``statistical organ'' with the biological purpose to
@@
 nested prediction of probable upcoming actions, states, and rewards.
-Both frameworks model the consequences of action.
-The predictive coding framework is typically explained without
+Both accounts model the consequences of action.
+The predictive coding account is typically explained without
 explicit parameterization of the agent's horizon and
@@
 are explained in MDPs by the advantages of stochastic optimization.
-Yet, the MDP framework provides an algorithmic explanation in that
+Yet, the MDP-based account provides an algorithmic explanation in that
 stochasticity of the parameter space explored
@@
 for the existence and survival of the human species
-to warrant constant, high energy costs?
+to warrant constantly high energy costs?
 MDPs provide an attractive
-process model how the human association cortex
+formal account how the human association cortex
 might implement supra-modal representation and control of the environment to
@@
 and possible future events to guide action choice in the present context.
-This functional account is more compatible with the DMN's
+MDPs provide a mathematical formalism how
+optimal substructure in the environment can be recursively exploited
+when confronted with complicated decisions.
+%
+This functional interpretation is more compatible with the DMN's
 poorly understood involvement across
@@
 as well as delay discounting and self-related prospection.
-MDPs provide a mathematical formalism how
-optimal substructure in the environment can be recursively exploited
-when confronted with complicated decisions.
 Improvement of the internal world model
@@
 In principle,
-neuroscientific experiments can be designed that operationalize
+neuroscience experiments can be designed that operationalize
 the set of action, value, and state variables that determine
 the behavior of intelligent RL systems.
+
 The proposed machine-learning perspective
@@
 also yields falsifiable neuroscientific hypotheses.
+%
 At the least, we propose an alternative vocabulary to
 describe and interpret experimental findings in neuroscience studies
-on the DMN.
+on this macroscopic brain network.
 %
@@
 from the relatively recent events to anticipated upcoming events
-in order to constantly improve action in our dynamic world.
+in to constantly refine action in our dynamic world.
 

### 2017-04-19 paper/paper.tex
stuffy stuff

@@
 \textit{ii}) semantic association, and
-\textit{iii}) a ``radar'' function.
+\textit{iii}) a ``sentinel'' function.
 %
@@
 various stages of sensory input processing.
-In our RL view, instead,
+In our MDP view, instead,
 the Markov property of our sequential decision making framework entails
@@
 nested prediction of probable upcoming actions, states, and rewards.
-Both accounts model the consequences of action.
+Both accounts capture the consequences of action.
 The predictive coding account is typically explained without
@@
 takes the concrete form of gradient computation and
-back-propagation in our RL implementation.
+back-propagation in our MDP implementation.
 It is however important to note that
@@
 the long-standing predictive coding account from cognitive neuroscience.
-MPDs have the advantage of exposing an explicit
+MDPs have the advantage of exposing an explicit
 mechanism for the horizon of future considerations or
@@
 Recasting DMN function in terms of MDPs therefore naturally incorporates
-a majority of aspects of the prediction coding hypothesis.
+a majority of aspects from the prediction coding hypothesis.
 
@@
 Another frequently proposed cognitive account of DMN function revolves
-around forming logical associations and analogies between
+around forming logical associations and abstract analogies between
 the current experience and
 the conceptual knowledge derived from past experiences
-\citep{bar2007proactive, binder1999conceptual}.
+\citep{bar2007proactive, binder1999conceptual, constantinescu2016organizing}.
 Analogies might naturally tie incoming novel sensory stimuli to
@@
 semantic building blocks may provide the basis for
-mentally envisioning non-existent scenarios
+mentally envisioning non-existing scenarios
 that would optimize behavior in the environment
@@
 Further,
-a constructive account can explain the reciprocal relationship
-between an egocentric first person perspective and
-an allocentric bird’s eye perspective immersed in
-self-reflection, semantic associations, and autobiographical memories.
-%
-Cognitive aspects of egocentric-allocentric switching
-are also closely related to episodic memory, language, problem solving,
-planning, estimating other people's thoughts, and spatial navigation.
-These neural processes are likely to
+% a constructive account can explain the reciprocal relationship
+% between an egocentric first person perspective and
+% an allocentric bird’s eye perspective immersed in
+% self-reflection, semantic associations, and autobiographical memories.
+% %
+% Cognitive aspects of egocentric-allocentric switching
+episodic memory, language, problem solving,
+planning, estimating other people's thoughts, and spatial navigation
+represent neural processes that are likely to
 build on abstract world knowledge and logical assocations
@@
 The semantic hypothesis is for instance supported by evidence in animals that
-could learn a \textit{cognitive map of the environment},
-even without reward incentives, and exploit it later
-for other means \citep{tolman1948cognitive}.
+could learn a \textit{cognitive map of the environment}
+and exploit it later for other means \citep{tolman1948cognitive}.
 
 In comparison,
-both the semantic hypothesis and MDPs expose mechanisms of
-how action considerations could be mentally explored.
+both the semantic hypothesis and the present account based on MDPs
+expose mechanisms of how action considerations could be mentally explored.
 In both accounts,
-there is also no evidence to indicate that predictions of various
+there is also no reason to assume that predictions of various
 levels of complexity, abstraction, timescale and purpose
@@
 DMN activity increases across time, space, and content domains
-demonstrated in many neuroimaging studies \citep{spreng2009common}.
+demonstrated in many neuroimaging studies
+\citep{spreng2009common, laird2009, bzdok2012morality, binder2009}.
 The semantic hypothesis
@@
 stimulus processing and action choice in the present.}
-The MDPs naturally integrate the former egocentric
+The MDPs naturally integrate the former egocentric view
 (more related to current action, state and reward) and the later
-allocentric (more related to past and future actions, states, and rewards)
-angles on the world in a same optimization problem.
+world view (more related to past and future actions, states, and rewards)
+on the world in a same optimization problem.
 Further,
@@
 in task-unconstrained mind-wandering \citep{smith2009, bzdok2016formal}.
-Both cognitive sets may entail DMN recruitment
+Both environment-detached and environment-engaged
+cognitive sets may entail DMN recruitment
 if environmental relevance is processed, manipulated, and
@@
 The horizon is expressed in the $\gamma$ parameter in the formal account.
-We thus provide answers for the unsettled question why the involvement
+We thus provide answers for the currently unsettled question why the involvement
 of the same neurobiological brain circuit (i.e., DMN) has been documented
@@
 a number of previous experimental observations in the
-DMN by a simple but non-trivial mechanism.
+DMN by simple but non-trivial mechanisms.
 %
@@
 autobiographical memory recall, problem solving,
-abstract reasoning based on internally generated scenes, social cognition,
+abstract reasoning, social cognition,
 as well as delay discounting and self-related prospection.
@@
 by injecting stochasticity in the recall of past
-actions and outcomes may explain why
+actions and the estimation of action outcomes may explain why
 very accurate memories have been disfavored in human evolution
@@
 describe and interpret experimental findings in neuroscience studies
-on this macroscopic brain network.
+on this major brain network.
 %
@@
 from the relatively recent events to anticipated upcoming events
-in to constantly refine action in our dynamic world.
+in to constantly refine the organism's interaction with the world.
 

### 2017-04-19 paper/paper.tex
doing ambidextrous passes over the manuscript

--- 2017-04-19 paper/paper.tex
@@
 When left unperturbed, the human brain is not at rest.
-In the beginning of the 21st century,
-brain imaging may have been the first technique
-to allow for the discovery of a unique brain network that subserves
-baseline mental activities
-\citep{raichle2001pnas, bzdok2015resting}.
 Especially the so-called "default mode network" (DMN) continues
@@
 cortex" \citep{carhart2010default}.
+In the beginning of the 21st century,
+brain imaging may have been the first technique
+to allow for the discovery of a unique brain network that subserves
+baseline mental activities
+\citep{raichle2001pnas, bzdok2015resting}.
 
@@
 been shown to directly impact human behavior. Goal-directed task performance
-improved with decreased activity in default mode regions \citep{weiss2006}
+improved with decreased activity in default mode areas \citep{weiss2006}
 and increased DMN activity was linked to more task-independent,
@@
 A particularly attractive computational framework
-to describe, quantify, and predict intelligent systems, such as the brain,
+to describe, quantify, and predict autonomously acting systems like the brain
 is proposed to be the combination of control theory and
@@
 Without requiring explicit awareness,
-these ``offline'' processes may contribute
+these ``offline'' processes may however contribute
 to optimizing control of the organism.
@@
 empirical evidence from
-neuroscience experiments \citep{kording2004bayesian, fiser2004small}.
+neuroscientific experiments \citep{kording2004bayesian, fiser2004small}.
 The present paper proposes a
@@
 Indeed,
-the vmPFC is preferentially connected with reward-related and limbic regions.
+the vmPFC is preferentially connected with reward-related and limbic areas.
 The vmPFC is well known to have monosynaptical connections
@@
 
-\subsection{Long-term rewards and action policies}
+\subsubsection{Long-term rewards and action policies}
 The behavior of the agent is governed by some kind of \textit{policy}.
@@
 
-\subsection{Expected values, Bellman equation, and optimal control}
+\subsection{The knots and bolts for RL in the DMN}
+\subsubsection{The objective value of an action given a state}
 Given only the limited information available from a POMDP, at a state $s$ the average quality or usefulness of taking an action $a$ under a policy $\pi$ is completely captured by the single number
@@
 % \textit{greedy policy}
-Optimal behavior corresponds to a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. Formally,
+\subsubsection{Optimal behavior and the Bellman equation} Optimal behavior corresponds to a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average.
+Formally,
 % \textbf{XXX please add 1-2 sentences why this is interesting in the current
@@
 \end{equation}
-the optimal Q-value over all policies.
-The goal of learning is to approach the policy $\pi^*$ as ``close as possible'', that is to \text{solve} the POMDP.
+the optimal Q-value over all policies. The goal of learning is to approach the policy $\pi^*$ as ``close as possible'', that is to \text{solve} the POMDP.
 Of course \eqref{eq:qlearning} is just a definition and
@@
    \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q(s', a')]\\
-   &= r(s,a) + \mathbb E_{s' \sim p(s'|s,a)} [\gamma \max_{a' \in \mathcal A}Q(s', a')]\\
+   &= r(s,a) + \gamma\mathbb E_{s' \sim p(s'|s,a)} [\max_{a' \in \mathcal A}Q(s', a')]\\
    &= \text{instantaneous reward} + \text{expected reward for acting greedily thereafter}
@@
 
-\subsubsection{Value approximation and greedy action consideration, and experience replay}
+% \subsection{Learning in the DMN}
+\subsubsection{Value approximation and the policy matrix}
 %As already mentioned in the previous section, Q-learning optimizes over the class of
@@
 \end{equation}
-by querying an approximate version $a \mapsto \tilde{Q}(a,s|\theta)$ of the $Q$-value function ``frozen'' at state $s$.
+by querying an approximate version $a \mapsto \tilde{Q}(a,s|\theta)$ of the $Q$-value function ``frozen'' at state $s$. For instance, a simplistic linear model with a kernel $\phi$ would be of the
+form $\tilde{Q}(s, a|\theta) = \phi(s,a)^T\theta$, where
+$\phi(s,a)$ would represent a high-level representation of the state-action pairs
+$(s,a)$, as proposed by \citep{songNIPS2016} for example.
+%  and Parr et al. 2003 (JMLR)].
 In the DMN, The dmPFC (BA9) would implement such a greedy lookup over the action space.
@@
   parameters and would correspond to synaptic weights and / or connection strengths between
-  DMN regions.
-It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the greedy policy \eqref{eq:policy}.
-Learning consists in rewriting this program on the fly, so that the approximating functional $\tilde{Q}(.,.|\theta)$ becomes closer and closer to the optimum $Q^*$.
-
-The action $a$ taken in \eqref{eq:policy} triggers a reward $r$ collected by the agent,
+  DMN regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the greedy policy \eqref{eq:policy}.
+The oracle Q-values $Q(s, a|\theta)$ which tell the DMN the (expected) usefulness of taking an action a in state / situation  $s$.
+The brain (BA9 ?) would then choose, a given $s$, an action $a$ which maximizes this tabulated values.
+This mapping from states to actions defines a matrix, which can be conveniently called a "policy matrix".
+Learning consists in starting from a given instantiation of such a table and tweaking it around while using it to take actions which take the agent to different areas of the table, and so on, and so forth.
+% like solving a problem which is changing...), etc..
+
+\subsubsection{The loss function}
+There can be no learning without a loss function in sight.
+The action $a$ taken in state $s$ according to the policy matrix defined in
+\eqref{eq:policy} triggers a reward $r$ collected by the agent,
 % (measured on their internal currency of worldly values ?),
@@
 At the end of this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced, which represents an exemplar
-behavior of the agent and is recorded in replay memory buffer
-(possibly discarding the oldest entries to make space):
+behavior of the agent and is recorded in replay memory buffer --which we hypothesize to be the
+hippocampus--, possibly discarding the oldest entries to make space:
 $\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
 At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1} + \delta \theta_{k}$ of the parameters for its approximate model of $Q$-value functional. This warrants learning,
-and a loss function is in demand. The Bellman equation \eqref{eq:bellman}
-provides a way to produce such a loss function \eqref{eq:oracle} as we now outline below.
-
+and a loss function is in demand.
+% \subsubsection{Stochastic gradient in the DMN via experience replay}
+The Bellman equation \eqref{eq:bellman} provides a way to produce such a loss function \eqref{eq:oracle} as we now outline.
 % A practically successful RL approach
@@
 
-\subsubsection{Stochastic gradient descent in the DMN}
 At the current iteration $k$, computing an optimal parameter update then corresponds to finding the parameters $\theta_{k}$ which minimize the following mean-squared loss function
@@
 % \textbf{We still may need to explain y in the formula with one sentence.}
-For instance, a general linear model with a kernel $\phi$ would be of the
-form
-$$\tilde{Q}(s, a|\theta) = \phi(s,a)^T\theta.$$
-$\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action pairs
-$(s,a)$, as proposed by \citep{songNIPS2016} for example.
-%  and Parr et al. 2003 (JMLR)].
 A recently proposed, practically successful alternative approach
@@
 
-Learning
-of the the entire model parameters can effectively be achieved
-via \textit{backpropagation},
-where reward prediction errors (RPEs) -- also known as \textit{regret}-- percolate
-from lower to higher processing layers to modulate the choice of future actions.
-Over time (i.e in the limit $k \rightarrow \infty$), the approximate
-$Q$-values $\tilde{Q}(a,s|\theta_{k-1})$ provably\footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
-converge uniformly to the unique fixed-point $Q^*(a,s)$ defined implicitly by the Bellman
-equation \eqref{eq:bellman}
-
-% online stochastic gradient descent is used to minimize the variational objective
-% \eqref{eq:oracle}.
-
-% \paragraph{Parameter update.}
-\begin{equation}
-  \delta \theta_{k} \propto -\nabla_{\theta_{k}}\mathcal L(\theta_{k})
-  = -\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
-    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{averseness}}].
-  \label{eq:oracle}
-\end{equation}
-
-\subsubsection{Link to classical reinforcement learning algorithms}
-One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
-learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
-cast in this \textbf{please unpack the "this" here}
-general variational framework. For instance, TD corresponds to
-the above framework using a linear value approximator with feature encoding
-$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
-$\mathcal S \times \mathcal A$, and so
-$$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
-Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
-$$\theta(s,a) \leftarrow \theta(s,a) + \alpha(\tilde{Q}(s, a|\theta) - y),$$
-the well-known TD update rule \citep{sutton1998reinforcement}.
-
-\subsubsection{Excursion: the role of the hippocampus}
-\paragraph{Does the hippocampus subserve Monte-Carlo sampling?}
+\subsubsection{Does the hippocampus subserve Monte-Carlo sampling?}
 In RL, Monte-Carlo simulation can be used to update the agent's belief state
@@
 \textbf{XXX Elvis: rephrase last sentence}
+
+\subsubsection{Optimal control via stochastic gradient descent in the DMN}
+Learning of the the entire model parameters can effectively be achieved
+via \textit{gradient descent},
+where reward prediction errors (RPEs) -- also known as \textit{regret}-- percolate
+from lower to higher processing layers to modulate the choice of future actions.
+Over time (i.e in the limit $k \rightarrow \infty$), the approximate
+$Q$-values $\tilde{Q}(a,s|\theta_{k-1})$ provably\footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
+converge uniformly to the unique fixed-point $Q^*(a,s)$ defined implicitly by the Bellman
+equation \eqref{eq:bellman}
+
+% online stochastic gradient descent is used to minimize the variational objective
+% \eqref{eq:oracle}.
+
+% \paragraph{Parameter update.}
+\begin{equation}
+  \delta \theta_{k} \propto -\nabla_{\theta_{k}}\mathcal L(\theta_{k})
+  = -\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
+    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{averseness}}].
+  \label{eq:oracle}
+\end{equation}
+
+% \subsubsection{Link to classical reinforcement learning algorithms}
+% One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
+% learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
+% cast in this \textbf{please unpack the "this" here}
+% general variational framework. For instance, TD corresponds to
+% the above framework using a linear value approximator with feature encoding
+% $\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
+% $\mathcal S \times \mathcal A$, and so
+% $$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
+% Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
+% $$\theta(s,a) \leftarrow \theta(s,a) + \alpha(\tilde{Q}(s, a|\theta) - y),$$
+% the well-known TD update rule \citep{sutton1998reinforcement}.
 
@@
 gaps between fresh sensory input and expectations computed
-based on stored prior information.
+from stored prior information.
 At each stage of neural processing,
-an internally generated prediction of aspects of environmental sensations is
+an internally generated prediction of aspects of sensory sensation is
 directly compared against the actual environmental input.
@@
 may be maintained by top-down modulation
-using a-priori information about the environment.
-
-In short, predictive coding processes
-conceptualize updates of the internal representation of the environment
+by a-priori information about the environment.
+
+Predictive coding processes
+permit updates of the internal representation of the environment
 to best accommodate and prepare the organism
 for the constant influx of sensory examples and
-performing action on the environment.
+for reacting to the environment.
 There are hence a number of common properties
 between the predictive coding account
-and the proposed formal account based on MDPs.
+and the proposed formal account.
 Importantly,
@@
 Further,
-both accounts are plausibilized by neuroscientific evidence that suggest
+both frameworks are plausibilized by neuroscientific evidence that suggest
 the human brain to be a ``statistical organ'' with the biological purpose to
@@
 nested prediction of probable upcoming actions, states, and rewards.
-Both accounts model the consequences of action.
-The predictive coding account is typically explained without
+Both frameworks model the consequences of action.
+The predictive coding framework is typically explained without
 explicit parameterization of the agent's horizon and
@@
 are explained in MDPs by the advantages of stochastic optimization.
-Yet, the MDP-based account provides an algorithmic explanation in that
+Yet, the MDP framework provides an algorithmic explanation in that
 stochasticity of the parameter space explored
@@
 for the existence and survival of the human species
-to warrant constantly high energy costs?
+to warrant constant, high energy costs?
 MDPs provide an attractive
-formal account how the human association cortex
+process model how the human association cortex
 might implement supra-modal representation and control of the environment to
@@
 and possible future events to guide action choice in the present context.
-MDPs provide a mathematical formalism how
-optimal substructure in the environment can be recursively exploited
-when confronted with complicated decisions.
-%
-This functional interpretation is more compatible with the DMN's
+This functional account is more compatible with the DMN's
 poorly understood involvement across
@@
 as well as delay discounting and self-related prospection.
+MDPs provide a mathematical formalism how
+optimal substructure in the environment can be recursively exploited
+when confronted with complicated decisions.
 Improvement of the internal world model
@@
 In principle,
-neuroscience experiments can be designed that operationalize
+neuroscientific experiments can be designed that operationalize
 the set of action, value, and state variables that determine
 the behavior of intelligent RL systems.
-
 The proposed machine-learning perspective
@@
 also yields falsifiable neuroscientific hypotheses.
-%
 At the least, we propose an alternative vocabulary to
 describe and interpret experimental findings in neuroscience studies
-on this macroscopic brain network.
+on the DMN.
 %
@@
 from the relatively recent events to anticipated upcoming events
-in to constantly refine action in our dynamic world.
+in order to constantly improve action in our dynamic world.
 

### 2017-04-19 paper/paper.tex
some affiliation politics

@@
 \small{
-    \textbf{1} INRIA, Parietal team, Saclay, France\\
-    \textbf{2} CEA, Neurospin, Gif-sur-Yvette, France\\
+    \textbf{1} Universit\'e Paris-Saclay, INRIA, Parietal team, Saclay, France\\
+    \textbf{2} Universit\'e Paris-Saclay, CEA, Neurospin, Gif-sur-Yvette, France\\
     \textbf{3} Department of Psychiatry, Psychotherapy and Psychosomatics, RWTH Aachen University, Aachen, Germany\\
@@
     \textbf{5} CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
-    \textbf{6} University Paris Diderot, Sorbonne Paris Cité, Paris, France\\
+    \textbf{6} University Paris Diderot, Sorbonne Paris Cit\'e, Paris, France\\
     \textbf{7} Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
@@
 
-\subsubsection{Long-term rewards and action policies}
+\subsubsection{Accumulated rewards and policies}
 The behavior of the agent is governed by some kind of \textit{policy}.
@@
 \end{mdframed}
-  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$. In
+  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$, in an average sense to be made precise latter. In
   the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.

### 2017-04-19 paper/paper.tex
more

@@
 
-\subsection{The free-energy principle and active inference \textbf{XXX move to Appendix}}
+\subsection{The free-energy principle and active inference}
 % \subsubsection{The theory}
@@
 Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net (DBN)}.
-According to FEP's account, the goal of the brain is then to optimize over the generative model $G$: to iteratively
+According to FEP's account, the goal of the brain is then to optimize over a generative model $G$ of sensations: to iteratively
 modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics, etc., so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles.
@@
 as it overcomes the technical limits of the former, by using a reparametrization trick. For instance, unlike the wake-sleep algorithm
-for minimizing free-energy, VAEs can be efficiently trained via back-propagation of learning errors.
+for minimizing free-energy, VAEs can be efficiently trained via back-propagation of prediction errors.
 

### 2017-04-19 paper/paper.tex
fixing some glitches

@@
 a short future and pays less attention to
-rewards expected at a later time, precisely rewards more than $1/(1 - \gamma)$ time steps into the future are disregarded.
+rewards expected at a later time. Precisely, rewards that are not expected to come within $\tau := 1/(1 - \gamma)$ time steps from the present point are disregarded.
 This reduces the variance of expected rewards accumulated across action cascades by limiting the length of this cascades.
@@
   DMN regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the greedy policy \eqref{eq:policy}.
-The oracle Q-values $Q(s, a|\theta)$ which tell the DMN the (expected) usefulness of taking an action a in state / situation  $s$.
-The brain (BA9 ?) would then choose, a given $s$, an action $a$ which maximizes this tabulated values.
+  The oracle Q-values $Q(s, a|\theta)$ tell the DMN the (expected) usefulness of taking an action a in state / situation  $s$.
+The DMN (via BA9 ?) would then choose, at a given $s$, an action $a$ which maximizes this tabulated values.
 This mapping from states to actions defines a matrix, which can be conveniently called a "policy matrix".
 Learning consists in starting from a given instantiation of such a table and tweaking it around while using it to take actions which take the agent to different areas of the table, and so on, and so forth.
-% like solving a problem which is changing...), etc..
 
@@
 
+\subsubsection{Optimal control via stochastic gradient descent in the DMN}
+Learning of the the entire model parameters can effectively be achieved
+via \textit{gradient descent},
+where reward prediction errors (RPEs) -- also known as \textit{regret}-- percolate
+from lower to higher processing layers to modulate the choice of future actions
+
+% online stochastic gradient descent is used to minimize the variational objective
+% \eqref{eq:oracle}.
+
+% \paragraph{Parameter update.}
+\begin{equation}
+  \delta \theta_{k} \propto -\nabla_{\theta_{k}}\mathcal L(\theta_{k})
+  = -\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
+    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{averseness}}].
+  \label{eq:oracle}
+\end{equation}
+Over time (i.e in the limit $k \rightarrow \infty$), the so-generated approximating sequence of
+$Q$-value functions provably\footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.} $Q$-value functions $\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$
+are attracted and absorbed by the optimal $Q$-value function $Q^*$ defined implicitly by the Bellman equation \eqref{eq:bellman}.
+
+% \subsubsection{Link to classical reinforcement learning algorithms}
+% One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
+% learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
+% cast in this \textbf{please unpack the "this" here}
+% general variational framework. For instance, TD corresponds to
+% the above framework using a linear value approximator with feature encoding
+% $\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
+% $\mathcal S \times \mathcal A$, and so
+% $$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
+% Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
+% $$\theta(s,a) \leftarrow \theta(s,a) + \alpha(\tilde{Q}(s, a|\theta) - y),$$
+% the well-known TD update rule \citep{sutton1998reinforcement}.
+
 \subsubsection{Does the hippocampus subserve Monte-Carlo sampling?}
@@
 \textbf{XXX Elvis: rephrase last sentence}
-
-\subsubsection{Optimal control via stochastic gradient descent in the DMN}
-Learning of the the entire model parameters can effectively be achieved
-via \textit{gradient descent},
-where reward prediction errors (RPEs) -- also known as \textit{regret}-- percolate
-from lower to higher processing layers to modulate the choice of future actions.
-Over time (i.e in the limit $k \rightarrow \infty$), the approximate
-$Q$-values $\tilde{Q}(a,s|\theta_{k-1})$ provably\footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
-converge uniformly to the unique fixed-point $Q^*(a,s)$ defined implicitly by the Bellman
-equation \eqref{eq:bellman}
-
-% online stochastic gradient descent is used to minimize the variational objective
-% \eqref{eq:oracle}.
-
-% \paragraph{Parameter update.}
-\begin{equation}
-  \delta \theta_{k} \propto -\nabla_{\theta_{k}}\mathcal L(\theta_{k})
-  = -\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
-    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{averseness}}].
-  \label{eq:oracle}
-\end{equation}
-
-% \subsubsection{Link to classical reinforcement learning algorithms}
-% One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
-% learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
-% cast in this \textbf{please unpack the "this" here}
-% general variational framework. For instance, TD corresponds to
-% the above framework using a linear value approximator with feature encoding
-% $\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
-% $\mathcal S \times \mathcal A$, and so
-% $$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
-% Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
-% $$\theta(s,a) \leftarrow \theta(s,a) + \alpha(\tilde{Q}(s, a|\theta) - y),$$
-% the well-known TD update rule \citep{sutton1998reinforcement}.
 

### 2017-04-20 paper/paper.tex
gentle morning pass

@@
 \begin{abstract}
-The default mode network (DMN) is believed to subserve the human
-baseline mental activity.
-%
-Its highest energy consumption in the brain and
+The default mode network (DMN) is believed to subserve the
+baseline mental activity in humans.
+%
+Its highest energy consumption compared to other brain networks and
 its intimate coupling with conscious awareness are both pointing towards
@@
 Many research streams support an evolutionarily adaptive role in
-envisioning mental scenarios to simulate the future.
+envisioning mental scenarios to anticipate the future.
 
@@
 We argue that our formal account of DMN function naturally accommodates as special cases
-previously proposed cognitive accounts on
+the previously proposed cognitive accounts on
 \textit{i}) predictive coding,
-\textit{ii}) semantic association, and
+\textit{ii}) semantic associations, and
 \textit{iii}) a ``sentinel'' function.
 %
-A process model for the neural optimization of complex behavior
-can offer parsimonious explanations for
+A process model for the neural optimization of complex behavior in the DMN
+offers parsimonious explanations for
 various experimental findings in animals and humans.
@@
 \citep{kenet2003spontaneously, fiser2004small}.
-The baseline energy demand is surprisingly weakly modulated
-with task demand \citep{raichle2001pnas}. At the opposite, during sleep, the
+The baseline energy demand is only weakly modulated
+at the onset of psychological tasks \citep{raichle2001pnas}.
+At the opposite, during sleep, the
 decoupling of brain structures discarded the idea of the DMN being
@@
 increase in neural activity during
-unfocused everyday mind wandering \citep{raichle2001pnas}.
-The DMN may subserve important functions in
-higher-order control of human behavior
+unfocused mind wandering \citep{raichle2001pnas}.
+Indeed, the DMN may exert
+higher-order control on human behavior
 following the notion of "a hierarchy of brain systems with
@@
 diversity of humans' most advanced cognitive processes
-can be parsimoniously recast as prediction error minimization
+can be recast as prediction error minimization
 based on pervasive probabilistic simulations,
@@
 %
-It has indeed been proposed earlier that
+It has indeed been proposed that
 the human brain's energy budget is largely dedicated to
@@
 \citep{raichle2005intrinsic}.
-This idea has been invigorated by
+This idea is invigorated by
 empirical evidence from
@@
 based on experimental findings in the neuroscience literature.
-This walkthrough across network nodes will
+This walkthrough across each node of the DMN will
 outline the individual functional profiles,
@@
   \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
-  \caption{Neurobiological overview of the DMN with its key neural structures and their associated behavioral functional roles within the current proposed model.
-\textbf{XXX Guillaume: Shall we also state explicitely that the functions indicated are only some among many potential others realized by those structures?}}
+  \caption{Neurobiological overview of the DMN with its key constituent nodes and their associated functional roles relevant in our functional account.
   \label{fig:neurobiological_overview_DMN}

### 2017-04-20 paper/paper.tex
style up figure captions

@@
   \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
-  \caption{Neurobiological overview of the DMN with its key constituent nodes and their associated functional roles relevant in our functional account.
+  \caption{\textbf{Default mode network: key functions.}
+  Neurobiological overview of the DMN with its major constituent nodes and
+  their associated functional roles relevant in our functional account.}
   \label{fig:neurobiological_overview_DMN}
@@
   \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
-  \caption{Neurobiological implementation of RL by the DMN...}
+  \caption{\textbf{Default mode network: reinforcement learning implementation.}
+  Overview of how the constituent nodes of the DMN
+  map onto computational components necessary for a Markov decision process.}
   \label{fig:rl_process_chart}
@@
 from the relatively recent events to anticipated upcoming events
-in to constantly refine the organism's interaction with the world.
+to constantly refine the organism's interaction with the world.
 

### 2017-04-20 paper/paper.tex
saying POMDPs are cool

--- 2017-04-20 paper/paper.tex
@@
   \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
-  \caption{Neurobiological overview of the DMN with its key constituent nodes and their associated functional roles relevant in our functional account.
+  \caption{Neurobiological overview of the DMN with its key constituent nodes and their associated functional roles relevant in our functional account.}
   \label{fig:neurobiological_overview_DMN}
@@
 which we propose as a useful formal framework for DMN function.
-
-Model-free RL is plausibly realizable in the human brain. Indeed,
+%
+
+Model-free RL has had great success in real-world problems including robotics\cite{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering},
+etc., just to name a few. Such a learning scheme is plausibly realizable in the human brain. Indeed,
+a core property of human intelligence has been proposed underly to improve
+on an expected utility outcomes as a strategy for action choice in uncertain
+environments \citep{gershman2015computational}, a situation perfectly captured by the formalism of POMDPs. Also,
 it has long been proposed \citep{dayan2008decision} that there
@@
 that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
-% A core property of human intelligence has been proposed underly to improve
-% on an expected utility outcomes as a strategy for action choice in uncertain
-% environments \citep{gershman2015computational}.
-In our present model-free RL framework,
+%
+
+In our adopted model-free RL framework,
 a rat might represent such knowledge about the world as follows:
@@
 \begin{equation}
-  \pi^*(s) := \argmax_{a \in \mathcal A}{Q}^*(s, a),
+  \pi^*(s) := \argmax_{a \in \mathcal A}{Q}^*(s, a),\text{ where }Q^*(s, a) := \max_{\pi}Q^\pi(s,a).
   \label{eq:qlearning}
 \end{equation}
-where
-\begin{equation}
-  Q^*(s, a) := \max_{\pi}Q^\pi(s,a),
-\end{equation}
-the optimal Q-value over all policies. The goal of learning is to approach the policy $\pi^*$ as ``close as possible'', that is to \text{solve} the POMDP.
+The goal of learning is to approach the policy $\pi^*$ as ``close as possible'', that is to \text{solve} the POMDP.
 Of course \eqref{eq:qlearning} is just a definition and
@@
 
-\paragraph{Does hippocampal replay equal inverse reinforcement-learning?}
-%\begin{mdframed}
-  Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
-  a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
-  (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
-  \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
-  (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
-  For instance, given traces of motor actions from an adult to grab a cup on a table,
-  can an observing child figure out what ``langragian'' functional is being minimized by the former?
-  How? Can they reproduce this optimal behavior?
-%\end{mdframed}
-Such questions are of course pertinent to our decision-making theory for the DMN.
-In the general artificial case, the problem has been extensively studied and partially solved by
-\cite{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
-``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
-
-IRL is is suited for problems in which it's hard to define what the reward function could be
-(e.g.,  car-driving, drone maneuvering, etc.) ...
+% \paragraph{Does hippocampal replay equal inverse reinforcement-learning?}
+% %\begin{mdframed}
+%   Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
+%   a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
+%   (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
+%   \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
+%   (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
+%   For instance, given traces of motor actions from an adult to grab a cup on a table,
+%   can an observing child figure out what ``langragian'' functional is being minimized by the former?
+%   How? Can they reproduce this optimal behavior?
+% %\end{mdframed}
+% Such questions are of course pertinent to our decision-making theory for the DMN.
+% In the general artificial case, the problem has been extensively studied and partially solved by
+% \cite{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
+% ``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
+
+% IRL is is suited for problems in which it's hard to define what the reward function could be
+% (e.g.,  car-driving, drone maneuvering, etc.) ...
 

### 2017-04-20 paper/paper.tex

@@
   \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
-  \caption{Neurobiological overview of the DMN with its key constituent nodes and their associated functional roles relevant in our functional account.}
+  \caption{\textbf{Default mode network: key functions.}
+  Neurobiological overview of the DMN with its major constituent nodes and
+  their associated functional roles relevant in our functional account.}
   \label{fig:neurobiological_overview_DMN}

### 2017-04-20 paper/paper.tex
more

@@
 physics and can be locally described by Markov processes at a sufficiently coarse scale.
-Indeed, as noted in \cite{tegmark2016improved},
+Indeed, as noted in \citep{tegmark2016improved},
 any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
@@
 multi-step decision processes in such an environment are
-\textit{Partially Observable Markov Decision Processes (POMDPs)}.
+\textit{Partially Observable Markov Decision Processes (POMDPs)}\citep{sutton1998reinforcement}.
 A POMDP formalizes a sequential decision process in which it is assumed that environment dynamics are determined by a Markov process,
@@
 
-Model-free RL has had great success in real-world problems including robotics\cite{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering},
+\paragraph{Model-free Reinforcement Learning.}
+Model-free RL has had great success in real-world problems including robotics\citep{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering},
 etc., just to name a few. Such a learning scheme is plausibly realizable in the human brain. Indeed,
@@
 in rats, mice, cats, rabbits, songbirds, and
-monkeys \cite{buhry2011,nokia2010,dave2000,skaggs2007}.
+monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}.
 
@@
 % In the general artificial case, the problem has been extensively studied and partially solved by
-% \cite{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
+% \citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
 % ``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...

### 2017-04-20 paper/paper.tex
some fixups + improved fig captions

@@
   \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
-  \caption{\textbf{Default mode network: reinforcement learning implementation.}
-  Overview of how the constituent nodes of the DMN
-  map onto computational components necessary for a Markov decision process.}
+  \caption{\textbf{DMN: neurobiological implementation of Reinforcement Learning.}
+  Overview of how the constituent nodes of the Default Mode Network (DMN)
+  map onto computational components necessary for a RL agent.}
   \label{fig:rl_process_chart}
@@
 
-\subsubsection{Comparison to our proposed theory}
+% \paragraph{Comparison to our proposed theory.}
 On the surface, a common point between the FEP and our proposed RL-based framework

### 2017-04-20 paper/paper.tex
more
some smoothing

@@
 
-\subsection{The knots and bolts for RL in the DMN}
+\subsection{The knots and bolts of Reinforcement Learning in the DMN}
+\textbf{XXX Elvis: a short preface is needed here, explaining the roadmap of this subsection from a high-level perspective!}
 \subsubsection{The objective value of an action given a state}
@@
 \begin{mdframed}
-  \textbf{ NCs for the Bellman equation:}
+  \textbf{ Neural correlates of the Bellman equation in the DMN:}
   ...
@@
 
-\subsubsection{The loss function}
+  \subsubsection{Self-training and the loss function}
 There can be no learning without a loss function in sight.
@@
   Overview of how the constituent nodes of the Default Mode Network (DMN)
-  map onto computational components necessary for a RL agent.}
+  map onto computational components necessary for an RL agent.}
   \label{fig:rl_process_chart}
@@
 
-\paragraph{Acknowledgment}
+\paragraph{Acknowledgment.}
+Starwars!
 % {\small The research leading to these results has received funding from the

### 2017-04-20 paper/paper.tex
relocating iRL stuff to the appendix

@@
 
-% \paragraph{Does hippocampal replay equal inverse reinforcement-learning?}
-% %\begin{mdframed}
-%   Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
-%   a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
-%   (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
-%   \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
-%   (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
-%   For instance, given traces of motor actions from an adult to grab a cup on a table,
-%   can an observing child figure out what ``langragian'' functional is being minimized by the former?
-%   How? Can they reproduce this optimal behavior?
-% %\end{mdframed}
-% Such questions are of course pertinent to our decision-making theory for the DMN.
-% In the general artificial case, the problem has been extensively studied and partially solved by
-% \citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
-% ``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
-
-% IRL is is suited for problems in which it's hard to define what the reward function could be
-% (e.g.,  car-driving, drone maneuvering, etc.) ...
-
 \subsection{Putting everything together}
@@
 \appendix
+
+\section{More on RL in the brain}
+\subsection{Does hippocampal replay equal inverse reinforcement-learning?}
+%\begin{mdframed}
+  Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
+  a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
+  (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
+  \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
+  (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
+  For instance, given traces of motor actions from an adult to grab a cup on a table,
+  can an observing child figure out what ``langragian'' functional is being minimized by the former?
+  How? Can they reproduce this optimal behavior?
+%\end{mdframed}
+Such questions are of course pertinent to our decision-making theory for the DMN.
+In the general artificial case, the problem has been extensively studied and partially solved by
+\citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
+``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
+
+IRL is is suited for problems in which it's hard to define what the reward function could be
+(e.g.,  car-driving, drone maneuvering, etc.) ...
+
 \section{Free-energy principles!}

### 2017-04-20 paper/paper.tex
better RL nutshell figure

@@
 \small{
-    \textbf{1} Universit\'e Paris-Saclay, INRIA, Parietal team, Saclay, France\\
+    \textbf{1} Universit\'e Paris-Saclay, INRIA, Parietal Team, Saclay, France\\
     \textbf{2} Universit\'e Paris-Saclay, CEA, Neurospin, Gif-sur-Yvette, France\\
@@
 \begin{wrapfigure}{r}{.4\linewidth}
+  \caption{Reinforcement Learning in a nutshell}  
   \centering
-  \includegraphics{rl.png}
-  \caption{RL in a nutshell}
+  \includegraphics[width=.9\linewidth]{rl.png}
   \label{fig:rl}
@@
 a set of candidate actions to perform given this state. Starting a time $t=0$,
-a policy $\pi$ generates a trajectory of action cascades / trajectories as follows
+a policy $\pi$ generates a trajectory of action cascades / trajectories as follows:
 \begin{eqnarray*}
   \begin{split}
+    \textbf{choose action: }a_0 &\sim \pi(a|s_0)\\
     \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0)\\
@@
     \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1)\\
-    \ldots\\
+    \vdots\\
     \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
     \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t})\\
-    \ldots
+    \vdots
   \end{split}
@@
 
-Since an action taken in the present moment $t$
-might have repercussions in the far future, it turns out that the
+Since an action taken in the present moment might have repercussions in the far future, it turns out that the
 quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
@@
 \subsection{The knots and bolts of Reinforcement Learning in the DMN}
-\textbf{XXX Elvis: a short preface is needed here, explaining the roadmap of this subsection from a high-level perspective!}
+\textbf{XXX Elvis: ==> Elvis a short preface is needed here, explaining the roadmap of this subsection from a high-level perspective!}
 \subsubsection{The objective value of an action given a state}

### 2017-04-20 paper/paper.tex
improved RL nutshell fig

@@
 
-\begin{wrapfigure}{r}{.4\linewidth}
-  \caption{Reinforcement Learning in a nutshell}  
+\begin{wrapfigure}{r}{.45\linewidth}
+  \caption{\textbf{Reinforcement Learning in a nutshell.} At each step, given the current state of the environment,
+    agent takes and action. This generates a reward and an observed state transition of the environment.}  
   \centering
@@
 % \textit{greedy policy}
-\subsubsection{Optimal behavior and the Bellman equation} Optimal behavior corresponds to a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average.
-Formally,
+\subsubsection{Optimal behavior and the Bellman equation}
+
+Optimal behavior corresponds to a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. Formally,
 % \textbf{XXX please add 1-2 sentences why this is interesting in the current
@@
 \end{equation}
+
 The goal of learning is to approach the policy $\pi^*$ as ``close as possible'', that is to \text{solve} the POMDP.
@@
 % is a fixed point for this equation.
+
+% \begin{wrapfigure}{r}{.4\linewidth}
+%     \caption{Graphical illustration of the Bellman transform}   
+%   \includegraphics[width=.9\linewidth]{bellman.png}
+%   \label{fig:rl}
+% \end{wrapfigure}
+
 As a consequence, the complicated dynamic programming optimization problem \eqref{eq:qlearning}

### 2017-04-20 paper/paper.tex
more

@@
 
+\begin{wrapfigure}{r}{.4\linewidth}
+  \caption{\textbf{Reinforcement Learning in a nutshell.} Given the current state of the environment,
+    the agent takes an action (by following some policy). This generates a reward and an observed
+    state transition of the environment. This process goes on until interrupted or a goal state is reached.}
+  \centering
+  \includegraphics[width=1.\linewidth]{rl.png}
+  \label{fig:rl}
+\end{wrapfigure}
+
 RL is a problem-solving technique in which, through interactions with an unknown \textit{environment / world},
@@
 an action.
-
-\begin{wrapfigure}{r}{.45\linewidth}
-  \caption{\textbf{Reinforcement Learning in a nutshell.} At each step, given the current state of the environment,
-    agent takes and action. This generates a reward and an observed state transition of the environment.}  
-  \centering
-  \includegraphics[width=.9\linewidth]{rl.png}
-  \label{fig:rl}
-\end{wrapfigure}
 

### 2017-04-20 paper/paper.tex
integrated more papers

--- 2017-04-20 paper/paper.tex
@@
 \small{
-    \textbf{1} Universit\'e Paris-Saclay, INRIA, Parietal Team, Saclay, France\\
+    \textbf{1} Universit\'e Paris-Saclay, INRIA, Parietal team, Saclay, France\\
     \textbf{2} Universit\'e Paris-Saclay, CEA, Neurospin, Gif-sur-Yvette, France\\
@@
 These metabolic characteristics go hand-in-hand with
-functional neuroimaging analyses \citep{andrews2010} in
+functional and structural neuroimaging analyses
 suggesting that the PMC and mPFC
-potentially represent the functional backbone of the DMN.
+potentially represent the functional backbone of the DMN
+\citep{andrews2010, hagmann2008mapping}.
 
@@
 The DMN midline is probably closely linked
-with the HC in the medial temporal lobe \citep{vincet2006} --
+with the HC in the medial temporal lobe \citep{vincet2006, shannon2013morning} --
 a region long known to be involved in
@@
 but also imagining one's own future and
-fictitious experiences more broadly \citep{hassabis2007patients}.
+imagined experiences more broadly \citep{hassabis2007patients}.
 Mental scenes created by HC-lesioned patients exposed a lack of
@@
 to the neocortex for long-term storage, potentially also in the nodes of the DMN.
+Moreover, invasively triggering experience recall
+in the HC of mice during sleep
+has been shown to subsequently alter action choice during wakefulness
+\citep{lavilleon2015}.
 These HC-subserved mechanisms
@@
 ongoing perception of and action on the environment
-\citep{maguire2016}.
+\citep{maguire2016, lavilleon2015}.
 % (TODO: relate to self-play in deepmind Go).
@@
 \begin{wrapfigure}{r}{.4\linewidth}
-  \caption{Reinforcement Learning in a nutshell}  
   \centering
-  \includegraphics[width=.9\linewidth]{rl.png}
+  \includegraphics{rl.png}
+  \caption{RL in a nutshell}
   \label{fig:rl}
@@
 a set of candidate actions to perform given this state. Starting a time $t=0$,
-a policy $\pi$ generates a trajectory of action cascades / trajectories as follows:
+a policy $\pi$ generates a trajectory of action cascades / trajectories as follows
 \begin{eqnarray*}
   \begin{split}
-    \textbf{choose action: }a_0 &\sim \pi(a|s_0)\\
     \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0)\\
@@
     \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1)\\
-    \vdots\\
+    \ldots\\
     \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
     \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t})\\
-    \vdots
+    \ldots
   \end{split}
@@
 
-Since an action taken in the present moment might have repercussions in the far future, it turns out that the
+Since an action taken in the present moment $t$
+might have repercussions in the far future, it turns out that the
 quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
@@
 
-\subsection{The knots and bolts of Reinforcement Learning in the DMN}
-\textbf{XXX Elvis: ==> Elvis a short preface is needed here, explaining the roadmap of this subsection from a high-level perspective!}
+\subsection{The knots and bolts for RL in the DMN}
 \subsubsection{The objective value of an action given a state}
@@
 \begin{mdframed}
-  \textbf{ Neural correlates of the Bellman equation in the DMN:}
+  \textbf{ NCs for the Bellman equation:}
   ...
@@
 
-  \subsubsection{Self-training and the loss function}
+\subsubsection{The loss function}
 There can be no learning without a loss function in sight.
@@
 observations and rewards \citep{aronov2017, chao2017interaction}.
-These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
+These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics \citep{lavilleon2015}.
 This would be a simple control algorithm by evaluating all legal actions and selecting the action with
@@
 \textbf{XXX Elvis: rephrase last sentence}
+
+% \paragraph{Does hippocampal replay equal inverse reinforcement-learning?}
+% %\begin{mdframed}
+%   Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
+%   a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
+%   (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
+%   \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
+%   (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
+%   For instance, given traces of motor actions from an adult to grab a cup on a table,
+%   can an observing child figure out what ``langragian'' functional is being minimized by the former?
+%   How? Can they reproduce this optimal behavior?
+% %\end{mdframed}
+% Such questions are of course pertinent to our decision-making theory for the DMN.
+% In the general artificial case, the problem has been extensively studied and partially solved by
+% \citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
+% ``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
+
+% IRL is is suited for problems in which it's hard to define what the reward function could be
+% (e.g.,  car-driving, drone maneuvering, etc.) ...
 
@@
   \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
-  \caption{\textbf{DMN: neurobiological implementation of Reinforcement Learning.}
-  Overview of how the constituent nodes of the Default Mode Network (DMN)
-  map onto computational components necessary for an RL agent.}
+  \caption{\textbf{Default mode network: reinforcement learning implementation.}
+  Overview of how the constituent nodes of the DMN
+  map onto computational components necessary for a Markov decision process.}
   \label{fig:rl_process_chart}
@@
 \citep{schilbach2008minds}. This view was later challenged by observing
-analogues of the DMN in monkeys \citep{mantini2011default}
-and rats \citep{lu2012rat}, both species with
+analogues of the DMN in monkeys \citep{mantini2011default},
+cats \citep{popa2009contrasting},
+and rats \citep{lu2012rat}, three species with
 supposedly less advanced social-cognitive capacities.
@@
 
-% \paragraph{Comparison to our proposed theory.}
+\subsubsection{Comparison to our proposed theory}
 On the surface, a common point between the FEP and our proposed RL-based framework
@@
 
-\paragraph{Acknowledgment.}
-Starwars!
+\paragraph{Acknowledgment}
 % {\small The research leading to these results has received funding from the
@@
 \appendix
-
-\section{More on RL in the brain}
-\subsection{Does hippocampal replay equal inverse reinforcement-learning?}
-%\begin{mdframed}
-  Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
-  a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
-  (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
-  \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
-  (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
-  For instance, given traces of motor actions from an adult to grab a cup on a table,
-  can an observing child figure out what ``langragian'' functional is being minimized by the former?
-  How? Can they reproduce this optimal behavior?
-%\end{mdframed}
-Such questions are of course pertinent to our decision-making theory for the DMN.
-In the general artificial case, the problem has been extensively studied and partially solved by
-\citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
-``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
-
-IRL is is suited for problems in which it's hard to define what the reward function could be
-(e.g.,  car-driving, drone maneuvering, etc.) ...
-
 \section{Free-energy principles!}

### 2017-04-21 paper/paper.tex
some more smoothing

--- 2017-04-20 paper/paper.tex
@@
 \subsection{The knots and bolts of Reinforcement Learning in the DMN}
-\textbf{XXX Elvis: ==> Elvis a short preface is needed here, explaining the roadmap of this subsection from a high-level perspective!}
+\textbf{XXX Elvis to Elvis:a short preface is needed here, explaining the roadmap of this subsection from a high-level perspective!}
 \subsubsection{The objective value of an action given a state}
@@
 \end{equation}
-
 The goal of learning is to approach the policy $\pi^*$ as ``close as possible'', that is to \text{solve} the POMDP.
@@
 \end{equation}
-where the ``Bellman transform'' $\text{Bel}(Q)$ of an arbitrary $Q$-value function $Q: \mathcal S \times \mathcal A \rightarrow \mathbb R$  is defined by setting
+where the ``Bellman transform'' $\text{Bel}(Q)$ of an arbitrary $Q$-value function $Q: \mathcal S \times \mathcal A \rightarrow \mathbb R$  is another $Q$-value function defined by
 \begin{equation}
@@
 reward and the discounted rewards of the upcoming states.
-% The optimal Q-value operator $Q^*$
-% is a fixed point for this equation.
-
-% \begin{wrapfigure}{r}{.4\linewidth}
-%     \caption{Graphical illustration of the Bellman transform}   
-%   \includegraphics[width=.9\linewidth]{bellman.png}
-%   \label{fig:rl}
-% \end{wrapfigure}
-
-As a consequence, the complicated dynamic programming optimization problem \eqref{eq:qlearning}
+The optimal Q-value operator $Q^*$ is a fixed point for this equation.
+As a consequence of this decomposition, the complicated dynamic programming optimization
+problem \eqref{eq:qlearning}
 is broken into simpler sub-problems at different time points.

### 2017-04-21 paper/paper.tex
1st half of super pass

--- 2017-04-20 paper/paper.tex
@@
 Switching between states reduces to a sequential processing model.
-Informed by outcomes of performed action,
+Informed by outcomes of really performed action,
 the DMN dynamics are constantly adapted in feedback loops
@@
 outline the individual functional profiles,
-paving the way for their functional interpretation
-within our formal account (next section).
+paving the way for their algorithmic interpretation
+in our formal account (next section).
 
@@
 
-The PMC myelinates relatively late during postnatal development in monkeys
+The PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
 \citep{goldman1987development}, which is generally considered to
@@
 Normal and disturbed metabolic fluctuations in the
-human PMC have been repeatedly related to
-phenomena of changed conscious awareness \citep{cavanna2006precuneus}.
+human PMC have been closely related to
+changes of conscious awareness \citep{cavanna2006precuneus}.
 %
@@
 and ensuing top-level control of action on the environment is supported
-by a number of neuroscience experiments.
-Electrophysiological recordings in animals implicated the PMC in
+by many neuroscience experiments.
+Electrophysiological recordings in animals implicated PMC neurons in
 strategic decision making \citep{pearson2009neurons},
 risk assessment \citep{mccoy2005risk},
-outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
+outcome-dependent behavioral modulation \citep{hayden2009electrophysiological},
 as well as approach-avoidance behavior
@@
 indicated changing choices in upcoming trials and their neural patterns were
-distinct from spike firings that indicated choosing either option.
+distinct from neuronal spike firings that indicated choosing either option.
 Similarly in humans,
@@
 
-The retrosplenial portion of the PMC can subserve
+Moreover, the retrosplenial portion of the PMC can subserve
 representation of action possibilities
-and evaluation of reward contingencies by integration with
+and evaluation of reward outcomes by integration with
 information from memory and different perspective frames.
@@
 proposed to mediate between the organism's egocentric
-(i.e., focused on sensory input) and
+(i.e., focused on sensory environment) and
 allocentric (i.e., focused on world knowledge) viewpoints
@@
 by monitoring the subjective outcomes
-of possible decisions and integrating that information
+of possible actions and integrating that information
 with memory, perspective frames, and
 reward outcomes into higher-level strategies.
-Perceived value, that differs across individuals, may update
+Estimated value, that differs across individuals, may update
 statistical assessment of the environment
 to predict delayed reward opportunities in the future.
-In doing so, the PMC may continuously adapt to changes
+In doing so, the PMC may continuously adapt the organism to changes
 in both the external environment and its internal representation
-to enable strategic behavior in volatile environments.
+to enable strategic behavior.
 
@@
 across time, space, and content domains to
-exert top-down control.
-In comparison to the PMC, however,
+exert top-level control on behavior.
+Comparing to the PMC, however,
 dmPFC function may be closer to a
@@
 potentially subserves the de-novo generation and manipulation
-of meaning representations instructed by stored semantics and memories.
+of meaning representations instructed by stored semantics and memories
+\citep{bzdok2013segregation}.
 
@@
 are known to struggle with
-adaption to novel situations and stimuli
+adaption to new situations and stimuli
 \citep{stuss1986frontal}.
@@
 episodic memory recall, generative construction processes,
-and explicit knowledge of the external world.
+and stored knowledge of the world.
 %
@@
 dorsomedial PFC activity preferentially modeled the consequences of
-action choices that were not actually executed,
+action choices that were considered but not actually executed,
 whereas ventromedial PFC (vmPFC) activity
@@
 Quantitative lesion findings across 344 human individuals confirmed
-a substantial impairment in value-based decision making
+a substantial impairment in value-based action choice
 \citep{glascher2012lesion}.
 Indeed,
-the vmPFC is preferentially connected with reward-related and limbic regions.
-The vmPFC is well known to have monosynaptical connections
+this DMN node is preferentially connected with reward-related and limbic regions.
+The vmPFC is well known to have direct connections
 with the NAc
@@
 \citep{lebreton2009automatic}.
-NAc activity is thought to reflect reward prediction signals from dopaminergic pathways
+NAc activity is thought to reflect reward prediction signals
+from dopaminergic neurotransmitter pathways
 \citep{schultz1998predictive}
@@
 In line with these connectivity findings in animals and humans,
-the vmPFC is often proposed to be involved in (external) emotional
-reactions and own (visceral) arousal \citep{damasio1996somatic}.
-Real or imagined bodily states could be mapped in the vmPFC
-as a bioregulatory disposition governing cognition
+the vmPFC is often proposed to represent externally or internally triggered
+emotional and motivational states \citep{damasio1996somatic}.
+Such real or imagined arousal states could be mapped in the vmPFC
+as a bioregulatory disposition influencing cognition
 and decision making.
@@
 the vmPFC consistently reflects an individual’s subjective
-valuation
+value estimates
 \citep{behrens2008associative}.
@@
 Such a ``cognitive map'' of the action space was argued to encode
-the current task state even when states are unobservable from sensory input,
-which was shown to be critical for behavior \citep{Schuck2016}.
+the current task state even when states are unobservable from the sensory environment
+\citep{Schuck2016}.
 
@@
 Single-cell recordings in the animal HC revealed
-some constantly active neuronal ensembles whose firing coincided with
+some constantly active neuronal populations whose firing coincided with
 specific locations in space during environmental navigation.
@@
 spike one after another  \citep{johnson2007neural}.
-These neuronal patterns appear to also indicate upcoming behavior
+These neuronal patterns appear to directly indicate upcoming behavior
 \citep{pfeiffer2013hippocampal}.
@@
 quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
-Moreover, spike sequences measured in hippocampal place cells of rats
+Moreover, neuronal spike sequences measured in hippocampal place cells of rats
 featured re-occurrence directly after experimental trials
 as well as directly before upcoming experimental trials \citep{diba2007forward}.
-Such hippocampal ensembles burt during rest and sleep
+Such hippocampal neurons burst during rest and sleep
 have been proposed to be critical in communicating local information
@@
 
-The HC thus orchestrates experience of environmental aspects for
+The HC thus orchestrates re-experience of environmental aspects for
 consolidations based on reenactment and for integration into
@@
 However,
-the right and left TPJ may become fully integrated into this canonical
+the right and left TPJ may become fully integrated into this major brain
 network only after birth.
@@
 \citep{seghier2013angular}.
-Neuroscientific investigations on hemispheric functional specialization
+Globally, neuroscientific investigations on hemispheric functional specialization
 have highlighted the right versus left cerebral hemisphere as dominant for
-attention versus language functions.
-
-The TPJ in the right hemipshere (RTPJ)
+attention versus language functions
+\citep{seghier2013angular, bzdok2013tpj, bzdok2016left}.
+
+The TPJ in the right-hemispheric DMN (RTPJ)
 has been shown to be intimately related to
@@
 
-The TPJ in the left hemisphere (LTPJ), in turn, exhibits a close topographical relationship to Wernicke's area
+The TPJ in the left-hemispheric hemisphere (LTPJ),
+in turn, exhibits a close topographical relationship to Wernicke's area
 involved in understanding spoken and written language.
@@
   If an agent behaving optimally in a given environment moves
-  to novel, yet unexperienced environments, reward prediction errors will
+  to new, yet unexperienced environments, reward prediction errors will
   massively increase \textbf{Please check/simplify sentence}.
@@
 Increased computation resources are however allocated
-when novel stimuli or
+when unknown stimuli or
 unexpected events are encountered by the organism.
@@
 \citep{bar2007proactive, binder1999conceptual, constantinescu2016organizing}.
-Analogies might naturally tie incoming novel sensory stimuli to
+Analogies might naturally tie incoming new sensory stimuli to
 explicit world knowledge (i.e., semantics) extracted from the environment

### 2017-04-21 paper/paper.tex
2nd half of super pass

@@
-\documentclass[10pt,letterpaper]{article}
+sensory\documentclass[10pt,letterpaper]{article}
 \usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
@@
 % official NIPS keywords
-\textbf{\\keywords}: systems biology, mind wandering, cognitive science,
+\textbf{\\keywords}: systems biology, mind-wandering, cognitive science,
 artificial intelligence,
@@
 the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
-supra-sensory processes
+multi-sensory processes
 across time, space, and content domains to
@@
 have highlighted the right versus left cerebral hemisphere as dominant for
-attention versus language functions
+attentional versus semantic functions
 \citep{seghier2013angular, bzdok2013tpj, bzdok2016left}.
@@
 has been shown to be intimately related to
-supra-modal prediction and error signaling.
+multi-sensory prediction and error signaling.
 It is probably central for
-action initiation during externally structured tasks and
-sensorimotor control by integrating supra-modal stimulus-guided attention
+action initiation during goal-dircted psychological tasks and for
+sensorimotor control by integrating multi-sensory attention
 \citep{corbetta2002control}.
@@
 visuo-proprioceptive conflict \citep{Balslev2005}, and
-multi-modal detection of sensory changes across
-visual, auditory, or tactile stimulation in a multi-modal fMRI study
+multi-sensory detection of environmental changes across
+visual, auditory, or tactile stimulation in an fMRI experiment
 \citep{downar2000multimodal}.
@@
 It was argued that the RTPJ encodes actions and ensuing outcomes
-without necessarily relating those to outcome value
+without necessarily relating those to value estimation
 \citep{liljeholm2013neural, hamilton2008action,
 jakobs2009effects}.
-Neural activity in the RTPJ has been argued to be responsible
-for stimulus-driven attentional reallocation to
-salient and surprising sources of information
+Neural activity in the RTPJ has been argued to reflect
+stimulus-driven attentional reallocation to
+self-relevant and unexpected sources of information
 as a ‘‘circuit breaker’’ that recalibrates control and maintenance systems
@@
 externally-oriented mind sets focussed on the
-sensory world and internally-oriented mind sets focussed
+sensory environment and internally-oriented mind sets focussed
 on self-relevant mental scene construction.
-Transient RTPJ lesion in humans for instance diminished the
+For instance, temporally induced RTPJ lesion in humans diminished the
 impact of predicted intentions of other individuals
@@
 The TPJ in the left-hemispheric hemisphere (LTPJ),
-in turn, exhibits a close topographical relationship to Wernicke's area
-involved in understanding spoken and written language.
-Neurological patients with damage caused to Wernicke's area
+in turn, exhibits a close relationship to Wernicke's area
+involved in understanding semantics, such as in spoken and written language.
+Neurological patients with damage in Wernicke's area
 have a major impairment of language comprehension
 when listening to others or reading a book.
-Their speech
-preserves natural rhythm and about normal syntax, yet the
-voiced sentences are devoid of meaning (i.e., aphasia).
-Abstracting from the typical semantic interpretations in linguistics
+Patient speech
+preserves natural rhythm and normal syntax, yet the
+voiced sentences lack meaning (i.e., aphasia).
+Abstracting from the typical speech interpretations in linguistics
 and neuropsychology,
@@
 \citep{binder2011neurobiology, seghier2013angular}.
-For instance, LTPJ lesions also entail problems in recognizing
+For instance, LTPJ lesions in humans also entail problems in recognizing
 others' pantomimed action towards objects
@@
 Inner speech also hinges on knowledge recall
-about the physical and inter-personal world.
+about the physical and social world.
 Indeed,
-the internal production of
+the internal generation of
 verbalized thought ("language of the mind") was closely related to the LTPJ
-in a multivariate analysis of brain volume
+in a pattern analysis of brain volume
 \citep{geva2011neural}.
@@
 Isolated building blocks of world statistics probably get reassembled
-in internally generated visual scenarios that
-navigate present action, weigh hypothetical possibilities, and forecast the future.
+in internally generated mental scenarios that
+guide present action choice,
+weigh hypothetical possibilities, and forecast the future.
 %
 The LTPJ may hence facilitate the automated prediction of events
-by incorporating experience-derived representations of the world
+by incorporating experience-derived building blocks of world statistics
 into ongoing action, planning, and problem solving.
@@
 We now argue the outlined neurobiological properties
-and potential functions of the DMN nodes
+and functional involvements of the DMN nodes
 to be sufficient for implementing all components
-of a full-fledged RL algorithm.
+of a full-fledged RL system.
 Recalling past experience, considering candidate actions,
-random sampling of possible scenarios, as well as
+random sampling of possible experiences, as well as
 evaluating instantaneous, delayed, and expected reward outcomes
@@
 
-RL is a problem-solving technique in which, through interactions with an unknown \textit{environment / world},
+RL is a problem-solving technique in which, through interactions with an unknown \textit{environment},
 an agent learns to reach goals, optimize reward signals, and minimize costs
@@
 $s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
-$r = r(s, a)$ (or \textit{punishment},
-since it can be negative) collected by the
-agent. This is illustrated in Fig. \ref{fig:rl}.
+$r = r(s, a)$ collected by the agent (Fig. \ref{fig:rl}).
+If the collected reward outcome yields a negative value it can be
+more naturally interpreted as \textit{punishment}.
 In this view, the environment is partially controlled by
 the action of the agent and the reward can be thought
-of as satisfaction -- or punishment -- accompanying the execution of
+of as satisfaction -- or aversion -- accompanying the execution of
 an action.
@@
 Action may be delayed to achieve substantial increases in expected reward
-that can grow with time.
-The environment is generally taken as stochastic,
+that is expected to grow with time.
+The environment is generally taken as \textit{stochastic},
 that is, changing in random ways. In addition,
@@
 the agent \citep{starkweather2017dopamine}.
-The volatility of the environment
+Assuming volatility of the environment
 is realistic in a computational model which sets out
@@
 \citep{dayan2008decision}.
-DMN implication in a diversity of
+
+In short, DMN implication in a diversity of
 particularly advanced cognitive processes
@@
 scene simulations coupled with prediction error minimization
-to calibrate action trajectories and
+to calibrate action trajectories to
 maximize reward outcomes at different time scales.
@@
 are a physical system governed by the laws of
-physics and can be locally described by Markov processes at a sufficiently coarse scale.
-Indeed, as noted in \citep{tegmark2016improved},
+physics and can be formally described
+by Markov processes at a sufficiently coarse scale.
+It has indeed been previously proposed \citep{tegmark2016improved} that
 any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
@@
 as a standard memoryless Markov process by simply expanding the
-definition of the state $s$ to include elements of the past.
+definition of the state $s$ to include experience episodes of the past.
 
@@
 \textit{Partially Observable Markov Decision Processes (POMDPs)}\citep{sutton1998reinforcement}.
-A POMDP formalizes a sequential decision process in which it is assumed that environment dynamics are determined by a Markov process,
+A POMDP operationalizes a sequential decision process
+in which it is assumed that environment dynamics are determined by a Markov process,
 but the agent cannot directly observe the underlying state.
@@
 Model-free RL has had great success in real-world problems including robotics\citep{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering},
-etc., just to name a few. Such a learning scheme is plausibly realizable in the human brain. Indeed,
+etc., just to name a few. Such a learning scheme can be
+plausibly realized in the human brain. Indeed,
 a core property of human intelligence has been proposed underly to improve
 on an expected utility outcomes as a strategy for action choice in uncertain
-environments \citep{gershman2015computational}, a situation perfectly captured by the formalism of POMDPs. Also,
-it has long been proposed \citep{dayan2008decision} that there
+environments \citep{gershman2015computational}, a situation perfectly captured by the formalism of POMDPs.
+It has also long been proposed \citep{dayan2008decision} that there
 is a rather direct mapping of model-free RL learning algorithms
@@
 and action policies by controlling
-synaptic plasticity at targets such as the ventral and
-dorsolateral striatum.
+synaptic plasticity at targets such as the ventral striatum.
 In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of the world.
@@
 
-\subsection{The knots and bolts of Reinforcement Learning in the DMN}
+\subsection{The nuts and bolts of Reinforcement Learning in the DMN}
 \textbf{XXX Elvis: ==> Elvis a short preface is needed here, explaining the roadmap of this subsection from a high-level perspective!}
@@
 In other words, $Q^{\pi}(s,a)$ corresponds to the expected reward
-over all possible action trajectories, in which
+over all considered action trajectories, in which
 the agent sets out in the environment in state
@@
 
-The goal of learning is to approach the policy $\pi^*$ as ``close as possible'', that is to \text{solve} the POMDP.
+The learning goal is to approach the
+policy $\pi^*$ as ``close as possible'', that is to \text{solve} the POMDP.
 Of course \eqref{eq:qlearning} is just a definition and
@@
 is broken into simpler sub-problems at different time points.
-Exploitation of hierarchical structure in actions consideration
+Exploitation of hierarchical structure in action considerations
 has previously been closely related to the medial prefrontal part of the DMN
@@
   The oracle Q-values $Q(s, a|\theta)$ tell the DMN the (expected) usefulness of taking an action a in state / situation  $s$.
-The DMN (via BA9 ?) would then choose, at a given $s$, an action $a$ which maximizes this tabulated values.
+  The DMN, and in particular its dmPFC node, would then contribute to the choice, at a given $s$, of an action $a$ which maximizes this tabulated values.
 This mapping from states to actions defines a matrix, which can be conveniently called a "policy matrix".
-Learning consists in starting from a given instantiation of such a table and tweaking it around while using it to take actions which take the agent to different areas of the table, and so on, and so forth.
+Learning consists in starting from a given instantiation of such a table and tweaking it around while using it to take actions which take the agent to different areas of the table, and so forth.
 
   \subsubsection{Self-training and the loss function}
-There can be no learning without a loss function in sight.
-The action $a$ taken in state $s$ according to the policy matrix defined in
-\eqref{eq:policy} triggers a reward $r$ collected by the agent,
+Successful learning is challenging without a clear learning goal --
+the \textit{loss function}.
+The action $a$ chosen in state $s$ according to the policy matrix defined in
+\eqref{eq:policy} yields a reward $r$ collected by the agent,
 % (measured on their internal currency of worldly values ?),
@@
 \citep{mnih2015,silver2016mastering} is to learn this
-representation using a deep neural network, leading to the so-called \textit{Deep Q-learning} family of methods which
+representation using a deep neural network, leading to the
+so-called \textit{deep Q-learning} family of methods which
 likely are state-of-the-art in RL.
@@
   = -\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
-    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{averseness}}].
+    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{aversion}}].
   \label{eq:oracle}
 \end{equation}
-Over time (i.e in the limit $k \rightarrow \infty$), the so-generated approximating sequence of
+Over time (i.e., in the limit $k \rightarrow \infty$),
+the thus generated approximating sequence of
 $Q$-value functions provably\footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.} $Q$-value functions $\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$
@@
 which can be prohibitively large.
-Monte-Carlo simulation provides a simple method for evaluating the value / fitness of a state.
-They provide an effective mechanism both for tree search and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than a RL agent without stochastic resampling procedures.
+Monte-Carlo simulation provides a simple method for evaluating the value of a state.
+They provide an effective mechanism both for tree search (of the considered
+action trajectories)
+and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than a RL agent without stochastic resampling procedures.
 
@@
 The DMN is today known to consistently increase in neural
-  activity when humans engage in goal-directed behavior that are detached from
+  activity when humans engage in cognitive processes that are detached from
   current sensory environment \citep{kenet2003spontaneously, fiser2004small}
-  and it was proposed to be situated at the top of the functional network hierarchies
+  and it was proposed to be situated at the top of the brain network hierarchy
   \citep{carhart2010default, margulies2016situating}.
   Its putative involvement in thinking about the past,
-  the future, and hypothetical possibilities
+  the future, and hypothetical experiences
   appears to tie in with the implicit computation of
   action and state cascades as a function of what happened in the past.
-  A policy represents the repertoire of possible actions
+  A policy encapsulates the repertoire of possible actions
     on the world given a current state. It encodes the probabilites of
@@
 The DMN may subserve
-  constant exploration of possible future actions and their
+  constant exploration of possible future actions and estimaton of their
   cumulative reward outcomes. Implicit computation of future choices
   provides an explanation for the
-  evolutionary emergence and practical usefulness of mind-wandering
+  evolutionary emergence and practical usefulness of
+  mind-wandering at day-time and dreams during sleep
   in humans.
@@
   only very rarely or never encounter.
-  An explanation is thus offered for encountering seemingly familiar situations that
+  An explanation is thus offered for experiencing seemingly familiar situations that
   a human has however never actually encountered (i.e., d\'{e}j\`{a} vu effect).
@@
   future scenarios and action outcomes.
-  Our account of DMN function thus acknowledges the unavoidable stochasticity of
-  computation in neuronal processes \citep{faisal2008noise}.
+  Our formal account of DMN function
+  thus acknowledges the unavoidable stochasticity of
+  computation in neural systems \citep{faisal2008noise}.
 
@@
 Feed-forward sensory processing is constantly calibrated by
-top-down modulation from more multi-modal and associative brain regions
+top-down modulation from more multi-sensory and associative brain regions
 further away from primary sensory cortical regions.
@@
 a ``radar'' function to
-detect subjectively salient cues and unexpected events in the environment.
+detect subjectively relevant cues and unexpected events in the environment.
 Propositions of such a sentinel function have however seldom detailed
@@
 how attention and memory resources are exactly reallocated when
-encountering a salient environmental stimulus.
+encountering a self-relevant environmental stimulus.
 In the present formal account,
@@
 its previously proposed ``sentinel'' function
-of monitoring the environment for salient information
+of monitoring the environment for self-relevant information
 in general and
@@
 formal account how the human association cortex
-might implement supra-modal representation and control of the environment to
+might implement multi-sensory representation and control of the environment to
 optimize the organism's fate.
@@
          $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
-             $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e $\x = (s, b).$ \\ \hline
+             $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e., $\x = (s, b).$ \\ \hline
              $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
@@
 \subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
-In this subsection, we will present an insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
+In this subsection, we will present an insightful upper bound for the generative surprise (i.e., generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
 As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
@@
 % \begin{mdframed}
-  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e input from lower more primitive layers).
+  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e., input from lower more primitive layers).
 Precisely

### 2017-04-21 paper/paper.tex
fix abnobxious typo

@@
-sensory\documentclass[10pt,letterpaper]{article}
+\documentclass[10pt,letterpaper]{article}
 \usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
@@
 \subsection{The nuts and bolts of Reinforcement Learning in the DMN}
-\textbf{XXX Elvis: ==> Elvis a short preface is needed here, explaining the roadmap of this subsection from a high-level perspective!}
+\textbf{XXX Elvis to Elvis a short preface is needed here, explaining the roadmap of this subsection from a high-level perspective!}
+
 \subsubsection{The objective value of an action given a state}

### 2017-04-21 paper/paper.tex
i ==> 1, ii ==> 2, etc. in abstract

@@
 the previously proposed cognitive accounts on
-\textit{i}) predictive coding,
-\textit{ii}) semantic associations, and
-\textit{iii}) a ``sentinel'' function.
+(1) predictive coding,
+(2) semantic associations, and
+(2) a ``sentinel'' function.
 %

### 2017-04-21 paper/paper.tex
a small pass on manuscript

@@
 
-\begin{mdframed}
 \begin{definition*}
-% \textbf{XXX Danilo: Is there a definition 2 somewhere?}
 Mathematically, a POMDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
@@
 
-% \begin{remark}
-%   Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
-%   except the Markov property, we can always transform it into an MDP by considering a compounded
-%   version of the states
-%   $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
-%   history up to an including time $t$.
-% \end{remark}
 \end{definition*}
-\end{mdframed}
+
+\begin{remark}
+  Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
+  except the Markov property, we can always transform it into an MDP by considering a compounded
+  version of the states
+  $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
+  history up to an including time $t$.
+\end{remark}
 

### 2017-04-22 paper/paper.tex
more stuff

@@
 Its highest energy consumption compared to other brain networks and
-its intimate coupling with conscious awareness are both pointing towards
+its intimate coupling with conscious awareness are both pointing to
 an important functional role.
@@
 the DMN may implement
-continuous assessment and prediction of the environment to guide action choices.
+continuous evaluation and prediction of the environment to guide action choices.
 %
@@
 (2) semantic associations, and
-(2) a ``sentinel'' function.
+(2) a ``sentinel'' role.
 %
@@
 offers parsimonious explanations for
-various experimental findings in animals and humans.
+recent experimental findings in animals and humans.
 \end{abstract}
@@
 brain imaging may have been the first technique
-to allow for the discovery of a unique brain network that subserves
+to allow for the discovery of a unique brain network that probably maintains
 baseline mental activities
-\citep{raichle2001pnas, bzdok2015resting}.
-Especially the so-called "default mode network" (DMN) continues
+\citep{raichle2001pnas, randy2008, bzdok2015resting}.
+The "default mode network" (DMN) continues
 to metabolize large quantities of
@@
 The baseline energy demand is only weakly modulated
-at the onset of psychological tasks \citep{raichle2001pnas}.
+at the onset of defined psychological tasks \citep{raichle_baseline}.
 At the opposite, during sleep, the
 decoupling of brain structures discarded the idea of the DMN being
-only a passive network resonance and rather supported its
+only a passive network resonance and rather supported an
 important role in sustaining conscious awareness \citep{horovitz2009decoupling}.
 This dark matter of brain physiology begs the question
-of the biological purpose underlying the DMN
+of the biological purpose underlying DMN activity
 \citep{raichle2006brain}.
@@
 the human association cortex \citep{mesulam1998sensation, margulies2016situating}
-consistently
-increase in neural activity during
-unfocused mind wandering \citep{raichle2001pnas}.
-Indeed, the DMN may exert
-higher-order control on human behavior
-following the notion of "a hierarchy of brain systems with
+may exert higher-order control on human behavior.
+It follows the notion of "a hierarchy of brain systems with
 the DMN at the top and the salience and dorsal attention systems
@@
 The human DMN is responsible for a large part of the baseline neural activity,
-which typically decreases during psychological experiments.
+which typically decreases when engaged in psychological experiments.
 The standard mode of
@@
 this network appears to
-simultaneously control perception-action cycles and
+simultaneously influence perception-action cycles and
 maintain baseline contemplations
@@
 and being guided by outcome estimations
-of really-happened, hypothetically conceived, and
+of really happened, hypothetically imagined, and
 expected futures to optimize behavioral performance.
-A particularly attractive computational framework
+Formally, a particularly attractive framework
 to describe, quantify, and predict intelligent systems, such as the brain,
@@
 reinforcement learning (RL).
-It is known that, the more the external world is predictable,
+An intelligent agent improves its interaction with the environment
+by continuously updating its computation of value estimates and action predispositions
+through integration of feedback outcomes.
+Biologically,
+it is known that, the more the external world is predictable,
 the more mental activity becomes detached from the actual sensory environment
@@
 \citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
-Without requiring explicit awareness,
+Without requiring explicit awareness by the agent,
 these ``offline'' processes may contribute
 to optimizing control of the organism.
+In this context,
+\textit{control} refers to the influence that an agent exerts when interacting
+with the environment to encourage preferred states. \textbf{(XXX Guillaume: DM was asking to move this sentence up. Indeed, we may have to define early on the key concepts.}
 We formalize
@@
 into the tension between exploitative action with immediate gains and
-exploratory action with longer-term reward outcomes.
+exploratory action with longer-term gratification.
 
@@
 can be recast as prediction error minimization
-based on pervasive probabilistic simulations,
+based on probabilistic mental simulations,
 thus maximizing action outcome across multiple time scales.
@@
 \citep{tenenbaum2011grow}.
-In this context,
-\textit{control} refers to the influence that an agent exerts when interacting
-with the environment to encourage preferred states. \textbf{(XXX Guillaume: DM was asking to move this sentence up. Indeed, we may have to define early on the key concepts.}
 Even (necessarily imperfect) memory recall,
-random day-time mind-wandering, and dreams during sleep
+random day-time mind-wandering, and
+seemingly arbitrary dreams during sleep
 may provide meaningful building blocks to iteratively improve
@@
 These metabolic characteristics go hand-in-hand with
-functional and structural neuroimaging analyses
-suggesting that the PMC and mPFC
+neuroimaging analyses that suggested the PMC and mPFC to
 potentially represent the functional backbone of the DMN
@@
 subjectively perceived relevance of a chosen object
-than by its factual value,
-based on an internal currency of value \citep{mccoy2005risk}.
+than by its actual value,
+based on an ``internal currency of value'' \citep{mccoy2005risk}.
 In fact, direct stimulation of PMC neurons
@@
 information from memory and different perspective frames.
-Regarding memory recall, retrosplenial lesions have been
+Regarding memory recall, retrosplenial damage have been
 consistently associated with anterograde and retrograde memory impairments
@@
 In doing so, the PMC may continuously adapt the organism to changes
-in both the external environment and its internal representation
+in both the external environment and its internally modeled representation
 to enable strategic behavior.
@@
 Generally,
-patients with neurological lesions in the prefrontal cortex
+neurological patients with tissue damage in the prefrontal cortex
 are known to struggle with
@@
 be more connected to the vmPFC than dmPFC in both species.
-Two functional connectivity analyses in humans strongly connected
+Two different functional connectivity analyses in humans strongly connected
 the vmPFC with the NAc, hippocampus (HC),
-and PMC.
+and PMC \citep{bzdok2015subspecialization}.
 %
@@
 hippocampal damage is
-not only associated with an impairment in reexperiencing the past (i.e., amnesia),
+not only associated with an impairment in re-experiencing the past (i.e., amnesia),
 but also imagining one's own future and
 imagined experiences more broadly \citep{hassabis2007patients}.
-Mental scenes created by HC-lesioned patients exposed a lack of
+Mental scenes created by patients with HC lesion exposed a lack of
 spatial integrity, richness in detail, and overall coherence.
@@
 spike one after another  \citep{johnson2007neural}.
-These neuronal patterns appear to directly indicate upcoming behavior
+Such neuronal patterns in the HC appear to directly indicate upcoming behavior
 \citep{pfeiffer2013hippocampal}.
@@
 as well as directly before upcoming experimental trials \citep{diba2007forward}.
-Such hippocampal neurons burst during rest and sleep
+Such hippocampal neurons spking patterns during rest and sleep
 have been proposed to be critical in communicating local information
@@
 recollecting autobiographical memory episodes, as well as
-reexperiencing or newly generating mental scenarios
+re-experiencing or newly generating mental scenarios
 \citep{hassabis2007patients}.
@@
 \citep{downar2000multimodal}.
-In humans, direct electrical stimulation of the
+Direct electrical stimulation of the human
 RTPJ during neurosurgery was associated with altered perception
@@
 self-relevant and unexpected sources of information
-as a ‘‘circuit breaker’’ that recalibrates control and maintenance systems
+as a ‘‘circuit breaker’’ that recalibrates control and maintenance brain networks
 \citep{bzdok2013tpj, corbettashul2008}.
@@
 on self-relevant mental scene construction.
-For instance, temporally induced RTPJ lesion in humans diminished the
+For instance, temporally induced RTPJ damage in humans diminished the
 impact of predicted intentions of other individuals
@@
 \citep{binder2011neurobiology, seghier2013angular}.
-For instance, LTPJ lesions in humans also entail problems in recognizing
+For instance, LTPJ damage in humans also entail problems in recognizing
 others' pantomimed action towards objects
@@
 \begin{wrapfigure}{r}{.4\linewidth}
-  \caption{\textbf{Reinforcement Learning in a nutshell.} Given the current state of the environment,
+  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
     the agent takes an action (by following some policy). This generates a reward and an observed
@@
 
-\paragraph{Model-free Reinforcement Learning.}
+\paragraph{Model-free reinforcement learning.}
 Model-free RL has had great success in real-world problems including robotics\citep{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering},
@@
 
-\subsection{The nuts and bolts of Reinforcement Learning in the DMN}
+\subsection{The nuts and bolts of reinforcement learning in the DMN}
 \textbf{XXX Elvis to Elvis a short preface is needed here, explaining the roadmap of this subsection from a high-level perspective!}
@@
 The semantic hypothesis
-and MDP account provide explanations why hippocampus lesion does
+and MDP account provide explanations why hippocampus damage does
 not only impair recalling memories, but also hypothetical and future
@@
 for the existence and survival of the human species
-to warrant constantly high energy costs?
+to justify constantly high energy costs?
 MDPs provide an attractive
@@
 might implement multi-sensory representation and control of the environment to
-optimize the organism's fate.
+optimize the organism's interaction with the world.
 % by obtaining reward and avoiding pain.
@@
 From the computational view of a Markovian sequential decision process,
-behavior unfolds by integrating happened past events
-and possible future events to guide action choice in the present context.
+behavior unfolds by integrating happened past experience
+and expected future events to guide action choice in the present context.
 MDPs provide a mathematical formalism how
@@
 neuroscience experiments can be designed that operationalize
-the set of action, value, and state variables that determine
+the set of action, value, and state variables that operationalize
 the behavior of intelligent RL systems.
@@
 The proposed machine-learning perspective
-on DMN biology is hence not only practically computable but
-also yields falsifiable neuroscientific hypotheses.
+on DMN biology is hence practically computable and
+yields falsifiable neuroscientific hypotheses.
 %
@@
 describe and interpret experimental findings in neuroscience studies
-on this major brain network.
+on this top-level brain network.
 %
 Ultimately,
-the DMN can be viewed as a functional integrator
-from the relatively recent events to anticipated upcoming events
-to constantly refine the organism's interaction with the world.
+DMN activity may instantiate a holistic integration
+ranging from real experience over purposeful dreams to anticipated futures
+for constant refinement of the organism's fate.
 

### 2017-04-22 paper/paper.tex
morer stuffer

@@
 an action.
-
 Action may be delayed to achieve substantial increases in expected reward
 that is expected to grow with time.
+
 The environment is generally taken as \textit{stochastic},
@@
 plausibly realized in the human brain. Indeed,
-a core property of human intelligence has been proposed underly to improve
-on an expected utility outcomes as a strategy for action choice in uncertain
+a core property of human intelligence has been proposed to underly improvement
+of expected utility outcomes as a strategy for action choice in uncertain
 environments \citep{gershman2015computational}, a situation perfectly captured by the formalism of POMDPs.
@@
 The tensor $\theta$ encodes the model
-  parameters and would correspond to synaptic weights and / or connection strengths between
+  parameters and would correspond to synaptic weights and connection strengths between
   DMN regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the greedy policy \eqref{eq:policy}.
@@
 the thus generated approximating sequence of
-$Q$-value functions provably\footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.} $Q$-value functions $\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$
+$Q$-value functions provably\footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
+$\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$
 are attracted and absorbed by the optimal $Q$-value function $Q^*$ defined implicitly by the Bellman equation \eqref{eq:bellman}.

### 2017-04-23 paper/paper.tex
Sunday election pass

@@
 its intimate coupling with conscious awareness are both pointing to
-an important functional role.
+an overarching function.
 %
 Many research streams support an evolutionarily adaptive role in
-envisioning mental scenarios to anticipate the future.
+envisioning experience to anticipate the future.
 
@@
 the DMN may implement
-continuous evaluation and prediction of the environment to guide action choices.
+continuous evaluation and prediction of the environment to guide behavior.
 %
@@
 %
-When left unperturbed, the human brain is not at rest.
+In the absence of external stimulation, the human brain is not at rest.
 In the beginning of the 21st century,
@@
 \citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
-Today, many investigators believe that the DMN implements some form of
+Today, many psychologists and neuroscientists believe that
+the DMN implements some form of
 probabilistic estimation of past, hypothetical, and
@@
 \citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
-randy2008}.
+randy2008, spreng2009common}.
 This brain network
@@
 this network appears to
-simultaneously influence perception-action cycles and
-maintain baseline contemplations
+simultaneously impact perception-action cycles and
+maintain probabilistic contemplations
 across time, space, and content domains.
 
-The present work takes the perspective of
+The present work adopts the perspective of
 a human agent faced with the choice of the next actions
-and being guided by outcome estimations
+and guided by outcome estimations
 of really happened, hypothetically imagined, and
@@
 through integration of feedback outcomes.
-Biologically,
+In this context,
+\textit{control} refers to the influence that an agent exerts when interacting
+with the environment to encourage preferred states.
+Psychologically,
 it is known that, the more the external world is predictable,
@@
 to optimizing control of the organism.
-In this context,
-\textit{control} refers to the influence that an agent exerts when interacting
-with the environment to encourage preferred states. \textbf{(XXX Guillaume: DM was asking to move this sentence up. Indeed, we may have to define early on the key concepts.}
 We formalize
@@
 We argue that DMN implication in a
-diversity of humans' most advanced cognitive processes
+diversity of most advanced cognitive processes in humans
 can be recast as prediction error minimization
@@
 RL is a problem-solving technique in which, through interactions with an unknown \textit{environment},
-an agent learns to reach goals, optimize reward signals, and minimize costs
-in an iterative trial-by-error fashion. See Fig. \ref{fig:rl}.
+an agent learns to reach goals and optimize reward signals
+in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
 At a given moment, each taken \textit{action} $a$ triggers an \textit{partially observed} change
@@
 scene simulations coupled with prediction error minimization
-to calibrate action trajectories to
-maximize reward outcomes at different time scales.
+to calibrate action trajectories for
+reward outcome maximization at different time scales.
 Such a purposeful optimization objective
@@
 
-\subsection{Partially Observable Markov Decision Processes}
+\subsection{Markov Decision Processes}
 We argue that the brain in general and the DMN in particular
@@
 Model-free RL has had great success in real-world problems including robotics\citep{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering},
-etc., just to name a few. Such a learning scheme can be
-plausibly realized in the human brain. Indeed,
+etc., just to name a few \textbf{XXX Please expand with 1-2 additional sentences}.
+Such a learning scheme can be
+plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
 a core property of human intelligence has been proposed to underly improvement
@@
 and action policies by controlling
-synaptic plasticity at targets such as the ventral striatum.
+synaptic plasticity at targets such as the NAc.
 In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of the world.
-These can be assumptions about the physical laws governing the agent's world or constraints on the state space and transitions between states.
-For instance, if a rat in a maze knows
-that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
+These can be assumptions about the physical laws governing the agent's environment or constraints on the state space and transitions between states.
+% For instance, if a rat in a maze knows
+% that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
 %
@@
 which is not compatible with long-term goal planning
-as potentially implemented in the DMN.
+as potentially realized in the DMN.
 To allow a learning process to arise,
@@
 % The reward discount factor
-$\gamma$ can be seen as calibrating risk-seeking valency of the intelligent agent,
-that is, the decision related to reward-delay trade-off schedules.
+$\gamma$ can be seen as calibrating risk-seeking trait of the intelligent agent,
+that is, the behavioral predispositions related to reward-delay trade-off schedules.
 Such an agent puts relatively more emphasis on rewards expected in
@@
 %  and Parr et al. 2003 (JMLR)].
-In the DMN, The dmPFC (BA9) would implement such a greedy lookup over the action space.
-The tensor $\theta$ encodes the model
-  parameters and would correspond to synaptic weights and connection strengths between
-  DMN regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the greedy policy \eqref{eq:policy}.
-  The oracle Q-values $Q(s, a|\theta)$ tell the DMN the (expected) usefulness of taking an action a in state / situation  $s$.
-  The DMN, and in particular its dmPFC node, would then contribute to the choice, at a given $s$, of an action $a$ which maximizes this tabulated values.
-This mapping from states to actions defines a matrix, which can be conveniently called a "policy matrix".
+In the DMN, the dmPFC (BA9) would implement such a greedy (i.e., heuristic but efficient) lookup
+over the action space.
+The model
+  parameters $\theta$ would correspond to synaptic weights and connection strengths between
+  brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the greedy policy \eqref{eq:policy}.
+  The oracle Q-values $Q(s, a|\theta)$ tell the DMN the (expected) usefulness of taking an action a in state $s$.
+  The DMN, and in particular its dmPFC node, would then contribute to the choice, at a given state $s$, of an action $a$ which maximizes this tabulated values.
+This mapping from states to actions is conventionally called "policy matrix".
 Learning consists in starting from a given instantiation of such a table and tweaking it around while using it to take actions which take the agent to different areas of the table, and so forth.
@@
 \subsubsection{Self-training and the loss function}
-Successful learning is challenging without a clear learning goal --
+Successful learning in brains and computer algorithms is challenging without a clear learning goal --
 the \textit{loss function}.
@@
 % \textbf{XXX Please try to smoothen the previous sentence}
-At the end of this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced, which represents an exemplar
-behavior of the agent and is recorded in replay memory buffer --which we hypothesize to be the
+At the end of this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced, which represents an 
+behavior unit of the agent and is recorded in replay memory buffer --which we hypothesize to be the
 hippocampus--, possibly discarding the oldest entries to make space:
 $\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
-At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1} + \delta \theta_{k}$ of the parameters for its approximate model of $Q$-value functional. This warrants learning,
-and a loss function is in demand.
+At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1} + \delta \theta_{k}$ of the parameters for its approximate model of $Q$-value function. This warrants a learning process
+and definition of a loss function.
 % \subsubsection{Stochastic gradient in the DMN via experience replay}
@@
 % to allow for scaling to large state and action spaces.
-Replay consists in sampling (uniform or importance-weighted
-\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
+Replay consists in sampling
+% (uniform or importance-weighted
+% \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.})
+mini-batches of experiences $e$
 $(s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$.
@@
 From a neurobiological perspective,
-experience replay would be manifested as the re-occurrence of a
+experience replay can be manifested in the re-occurrence of a
 neuron spiking sequence that also occurs during specific actions.
@@
 \end{equation}
-where $y_k$ is a defined in \eqref{eq:bellman}
+where $y_k$ is a defined in \eqref{eq:bellman}.
 % \textbf{We still may need to explain y in the formula with one sentence.}
@@
 likely are state-of-the-art in RL.
-The set of $\theta$ parameters that instantiate the nonlinear interactions
+The set of model parameters $\theta$ that instantiate the non-linear interactions
 between layers of the artifical neural network
@@
 the thus generated approximating sequence of
-$Q$-value functions provably\footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
+$Q$-value functions provably
+% \footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
 $\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$
@@
 \textbf{XXX Elvis: refs needed in last sentence.
-Plus, should be really bring up Bayesian?} For sampling-based methods it is still almost non-existent beyond the description of receptive fields.
-\textbf{XXX Elvis: rephrase last sentence}
+Plus, should be really bring up Bayesian?} For sampling-based methods it is still scarce
+beyond the description of receptive fields. \textbf{XXX Elvis: rephrase last sentence}
 
@@
 
-\textbf{XXX: add tons of more examples and connnections}
 \textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
@@
 ranging from real experience over purposeful dreams to anticipated futures
-for constant refinement of the organism's fate.
+for continued refinement of the organism's fate.
 

### 2017-04-23 paper/paper.tex
rare perls

@@
 (2) semantic associations, and
-(2) a ``sentinel'' role.
+(3) a ``sentinel'' role.
 %
@@
 maintain probabilistic contemplations
-across time, space, and content domains.
+across time, space, and content domains
+\citep{boyer2008evolutionary}.
 
@@
 in constructing mental models in general
-\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017}.
+\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017,
+boyer2008evolutionary}.
 Its recursive anatomical architecture
@@
 via \textit{gradient descent},
-where reward prediction errors (RPEs) -- also known as \textit{regret}-- percolate
+where reward prediction errors -- also known as \textit{regret}-- percolate
 from lower to higher processing layers to modulate the choice of future actions
@@
 the HC could contribute to synthesizing imagined sequences of world states,
-observations and rewards \citep{aronov2017, chao2017interaction}.
+observations and rewards \citep{aronov2017, chao2017interaction, boyer2008evolutionary}.
 These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics \citep{lavilleon2015}.
@@
   activity when humans engage in cognitive processes that are detached from
-  current sensory environment \citep{kenet2003spontaneously, fiser2004small}
+  current sensory environment
+  \citep{kenet2003spontaneously, fiser2004small, bzdok2013segregation}
   and it was proposed to be situated at the top of the brain network hierarchy
@@
   \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
-  \caption{\textbf{DMN: neurobiological implementation of Reinforcement Learning.}
+  \caption{\textbf{DMN: neurobiological implementation of reinforcement learning.}
   Overview of how the constituent nodes of the Default Mode Network (DMN)
@@
 %
-Going beyond human language itself,
+Going beyond meaning and concepts extracted from the world,
 semantic building blocks may provide the basis for
@@
 that would optimize behavior in the environment
-by constantly simulating possible future events.
+by constantly simulating possible future events
+\citep{boyer2008evolutionary, binder2011neurobiology}.
 Such cognitive processes can afford
@@
 \subsection{Sentinel hypothesis}
+DMN regions have been associated with processing the
+experienced or expected relevance of a diversity of environment stimuli
+\citep{montague2006imaging}.
 Processing self-relevant information was perhaps the first

### 2017-04-23 paper/paper.tex
XXX epsilon-greedy

@@
 \end{equation}
-
 The learning goal is to approach the
@@
 % deterministic policies of the form \eqref{eq:qlearning}.
+\textbf{XXX Elvis to Elvis: clarify the notion of greed. Add a paragraph on the exploitation vs exploration dilemma, and end with a conclusion saying the DMN probably does epsilon-greedy.}
+
 At a time step $k-1$, the world is in a state $s \in \mathcal S$, and the agent greedily takes an
@@
 %  and Parr et al. 2003 (JMLR)].
-In the DMN, the dmPFC (BA9) would implement such a greedy (i.e., heuristic but efficient) lookup
+In the DMN, the dmPFC (BA9) would implement such a greedy lookup
 over the action space.

### 2017-04-24 paper/paper.tex
some glitches

@@
 % deterministic policies of the form \eqref{eq:qlearning}.
-\textbf{XXX Elvis to Elvis: clarify the notion of greed. Add a paragraph on the exploitation vs exploration dilemma, and end with a conclusion saying the DMN probably does epsilon-greedy.}
-
 At a time step $k-1$, the world is in a state $s \in \mathcal S$, and the agent greedily takes an
@@
 \end{equation}
-by querying an approximate version $a \mapsto \tilde{Q}(a,s|\theta)$ of the $Q$-value function ``frozen'' at state $s$. For instance, a simplistic linear model with a kernel $\phi$ would be of the
+by querying an oracle which maps from states directly to actions. Here $\tilde{Q}(s,a|\theta)$ is an
+approximate version of the $Q$-value function parameterized by the tensor $\theta$.
+% $a \mapsto \tilde{Q}(a,s|\theta)$ of the $Q$-value function ``frozen'' at state $s$.
+For instance, a simplistic linear model with a kernel $\phi$ would be of the
 form $\tilde{Q}(s, a|\theta) = \phi(s,a)^T\theta$, where
@@
   brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the greedy policy \eqref{eq:policy}.
-  The oracle Q-values $Q(s, a|\theta)$ tell the DMN the (expected) usefulness of taking an action a in state $s$.
+  The oracle Q-values $\tilde{Q}(s, a|\theta)$ tell the DMN the (expected) usefulness of taking an action a in state $s$.
   The DMN, and in particular its dmPFC node, would then contribute to the choice, at a given state $s$, of an action $a$ which maximizes this tabulated values.
@@
 Learning consists in starting from a given instantiation of such a table and tweaking it around while using it to take actions which take the agent to different areas of the table, and so forth.
+
+\textbf{XXX Elvis to Elvis: clarify the notion of greed. Add a paragraph on the exploitation vs exploration dilemma, and end with a conclusion saying the DMN probably does epsilon-greedy.}
 

### 2017-04-24 paper/paper.tex
standardize quotes

@@
 \citep{raichle2001pnas, randy2008, bzdok2015resting}.
-The "default mode network" (DMN) continues
+The ``default mode network'' (DMN) continues
 to metabolize large quantities of
@@
 neuronal computation
-in the absence of behavioral focus
+during task-unconstrained behavior
 \citep{kenet2003spontaneously, fiser2004small}.
@@
 important role in sustaining conscious awareness \citep{horovitz2009decoupling}.
+
 This dark matter of brain physiology begs the question
@@
 \citep{raichle2006brain}.
-What has early been described as the "stream of consciousness"
+What has early been described as the ``stream of consciousness''
 in psychology \citep{james1890principles}
@@
 \citep{shul1997, raichle2001pnas}.
-This set of some of the most advanced regions in
+We propose that this set of some of the most advanced regions in
 the human association cortex \citep{mesulam1998sensation, margulies2016situating}
-may exert higher-order control on human behavior.
-It follows the notion of "a hierarchy of brain systems with
+are responsible for higher-order control on human behavior.
+Our functional account follows the notion of
+``a hierarchy of brain systems with
 the DMN at the top and the salience and dorsal attention systems
 at intermediate levels, above thalamic and unimodal sensory
-cortex" \citep{carhart2010default}.
+cortex'' \citep{carhart2010default}.
 
@@
 environmental statistics and its internal representation
-as an inner "mind's eye" \citep{cavanna2006precuneus, leech_pcc2014}.
+as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
 For instance, B\'alint's syndrome is a neurological disorder of conscious
@@
 the internal generation of
-verbalized thought ("language of the mind") was closely related to the LTPJ
+verbalized thought (``language of the mind'') was closely related to the LTPJ
 in a pattern analysis of brain volume
@@
 
-\subsection{Markov Decision Processes}
+\subsection{Markov decision processes}
 We argue that the brain in general and the DMN in particular
@@
 utility of executing a specific action; it answers the question
-"What is the expected utility of taking action $a$ in this situation ?".
+``What is the expected utility of taking action $a$ in this situation ?''.
 $Q^{\pi}(s,a)$ offers a formalization of optimal behavior that
@@
   The DMN, and in particular its dmPFC node, would then contribute to the choice, at a given state $s$, of an action $a$ which maximizes this tabulated values.
-This mapping from states to actions is conventionally called "policy matrix".
+This mapping from states to actions is conventionally called ``policy matrix''.
 Learning consists in starting from a given instantiation of such a table and tweaking it around while using it to take actions which take the agent to different areas of the table, and so forth.

### 2017-04-24 paper/paper.tex
styler collection upgrade

@@
 the human association cortex \citep{mesulam1998sensation, margulies2016situating}
-are responsible for higher-order control on human behavior.
+are responsible for higher-order control of human behavior.
 Our functional account follows the notion of
@@
 %
-Gaining insight into the function of the DMN is
+Gaining insight into DMN function is
 particularly challenging because
@@
 through integration of feedback outcomes.
-In this context,
+Here,
 \textit{control} refers to the influence that an agent exerts when interacting
 with the environment to encourage preferred states.
+
 Psychologically,
-it is known that, the more the external world is predictable,
-the more mental activity becomes detached from the actual sensory environment
+it is known that, the more the world is easy to predict,
+the more human mental activity
+becomes detached from the actual sensory environment
 \citep{antrobus1966studies, pope1978regulation}.
@@
 shaped by prediction error.
-A DMN account incorporating RL can naturally embed human behavior
+A RL account of DMN function can naturally embed human behavior
 into the tension between exploitative action with immediate gains and
@@
 \subsection{The posteromedial cortex: global monitoring and information integration}
-The midline structures of the DMN,
+The midline structures of the human DMN,
 including the posteromedial cortex (PMC) and
@@
 are probably responsible for the highest turn-over
-of glucose energy consumption \citep{raichle2001pnas}.
+of energy consumption \citep{raichle2001pnas, raichle_baseline}.
 These metabolic characteristics go hand-in-hand with
@@
 the human PMC was coupled in two functional connectivity analyses
+\citep{bzdok2015subspecialization}
 with the amygdala, involved in significance evaluation, and
@@
 
-The putative role of the PMC in continuous integration of
-relevant environmental information
+The putative role of the PMC in continuous information integration of
+environmental relevance
 and ensuing top-level control of action on the environment is supported
@@
 with memory, perspective frames, and
-reward outcomes into higher-level strategies.
+reward outcomes into short- and longer-term behavioral agendas.
 Estimated value, that differs across individuals, may update
@@
 are known to struggle with
-adaption to new situations and stimuli
+adaption to new  stimuli and events
 \citep{stuss1986frontal}.
@@
 errors in these predictions.
-Neural activity in the dmPFC indeed explained the proficiency decline
+Neural activity in the dmPFC indeed explained the performance decline
 of inferring other peoples' thoughts in aging humans \citep{moran2012social}.
-Some dmPFC neurons in macaque monkeys exhibited a preference
+Certain dmPFC neurons in macaque monkeys exhibited a preference
 for processing others', rather than own, behavior
 with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
-In fact, the topographically neighboring dorsal anterior cingulate cortex
+In fact, the spatially neighboring dorsal anterior cingulate cortex
 has also been linked to computing values and efforts of
@@
 From the computational view of a Markovian sequential decision process,
-behavior unfolds by integrating happened past experience
-and expected future events to guide action choice in the present context.
+behavior unfolds by integrating reward outcomes from stored past events
+and expected upcoming events to guide action choice in the present context.
 MDPs provide a mathematical formalism how
-optimal substructure in the environment can be recursively exploited
-when confronted with complicated decisions.
-%
-This functional interpretation is more compatible with the DMN's
+opportunity structure in the environment can be recursively exploited
+when confronted with challenging decisions.
+%
+This functional interpretation may be more compatible with the DMN's
 poorly understood involvement across
@@
 and why human creativity might be adaptive.
-%
-In principle,
-neuroscience experiments can be designed that operationalize
-the set of action, value, and state variables that operationalize
-the behavior of intelligent RL systems.
 
@@
 %
+Neuroscience experiments could be designed that operationalize
+the set of action, value, and state variables that govern
+the behavior of intelligent RL systems.
+%
 At the least, we propose an alternative vocabulary to
 describe and interpret experimental findings in neuroscience studies
-on this top-level brain network.
+on the DMN situated at top of the brain network hiearchy.
 %

### 2017-04-24 paper/paper.tex
wip

--- 2017-04-24 paper/paper.tex
@@
 
-% \subsubsection{Link to classical reinforcement learning algorithms}
-% One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
-% learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
-% cast in this \textbf{please unpack the "this" here}
-% general variational framework. For instance, TD corresponds to
-% the above framework using a linear value approximator with feature encoding
-% $\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
-% $\mathcal S \times \mathcal A$, and so
-% $$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
-% Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
-% $$\theta(s,a) \leftarrow \theta(s,a) + \alpha(\tilde{Q}(s, a|\theta) - y),$$
-% the well-known TD update rule \citep{sutton1998reinforcement}.
+\subsubsection{Link to classical reinforcement learning algorithms}
+One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
+learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
+cast in this \textbf{please unpack the "this" here}
+general variational framework. For instance, TD corresponds to
+the above framework using a linear value approximator with feature encoding
+$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
+$\mathcal S \times \mathcal A$, and so
+$$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
+Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
+$$\tilde{Q}(s, a) \leftarrow \tilde{Q}(s, a) + \alpha(y-\tilde{Q}(s, a)),$$
+the well-known TD update rule \citep{sutton1998reinforcement}.
 

### 2017-04-25 paper/paper.tex
awesomify

--- 2017-04-24 paper/paper.tex
@@
 
+The dmPFC may subserve inference, representation, and assessment
+of one's own and other individuals' action considerations.
 Generally,
@@
 are known to struggle with
-adaption to new  stimuli and events
+adaptation to new stimuli and events
 \citep{stuss1986frontal}.
-The dmPFC may subserve inference, representation, and assessment
-of one's own and other individuals' action considerations.
 Specifically, neural activity in the human dmPFC
@@
 with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
-In fact, the spatially neighboring dorsal anterior cingulate cortex
-has also been linked to computing values and efforts of
-persisting a behavioral plan versus switching the
-environmental context in several lesion studies \citep{kolling2016value}.
+% In fact, the spatially neighboring dorsal anterior cingulate cortex
+% has also been linked to computing values and efforts of
+% persisting a behavioral plan versus switching the
+% environmental context in several lesion studies \citep{kolling2016value}.
 %
@@
 generation of probabilistic internal information drawing from
-episodic memory recall, generative construction processes,
+episodic memory recall, mental scene elaboration processes,
 and stored knowledge of the world.
 %
-In a computational neuroimaging experiment,
-dorsomedial PFC activity preferentially modeled the consequences of
+Moreover,
+in a computational neuroimaging experiment,
+dorsomedial PFC activity preferentially processed the consequences of
 action choices that were considered but not actually executed,
 whereas ventromedial PFC (vmPFC) activity
-modeled especially value outcomes of performed actions in the environment
+processed especially value outcomes of performed actions in the environment
 \citep{nicolle2012agent}.
@@
 risk estimation of relevant environmental stimuli.
-This DMN node is more closely associated with
-orchestrating adaptive behavior by bottom-up-driven
+This DMN node may
+orchestrate adaptive behavior by bottom-up-driven
 processing of “what matters now”,
@@
 \textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
-\textbf{XXX: integrate aspects from WakeNsleep by Hinton}
 \textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
@@
 From the computational view of a Markovian sequential decision process,
-behavior unfolds by integrating reward outcomes from stored past events
+human behavior unfolds by integrating reward outcomes from stored past events
 and expected upcoming events to guide action choice in the present context.

### 2017-04-25 paper/paper.tex
awesomify 2.0

@@
 ``mental sketchpad'' \citep{goldman1996prefrontal}, as it
-potentially subserves the de-novo generation and manipulation
+potentially subserves the de-novo construction and manipulation
 of meaning representations instructed by stored semantics and memories
@@
 Such highly abstract neural computations necessarily rely on the
-generation of probabilistic internal information drawing from
+construction of probabilistic internal information drawing from
 episodic memory recall, mental scene elaboration processes,
@@
 this DMN node is preferentially connected with reward-related and limbic regions.
-The vmPFC is well known to have direct connections
+The vmPFC is well known to have strong axonal connections
 with the NAc
@@
 subserve RL in humans more broadly \citep{doherty2015structure}.
-This is consistent with diffusion MRI tractography in humans and monkeys
-\citep{croxson2005quantitative} that
+Consistently, diffusion MRI tractography in humans and monkeys
+\citep{croxson2005quantitative}
 quantified the NAc to
 be more connected to the vmPFC than dmPFC in both species.
-Two different functional connectivity analyses in humans strongly connected
+Two different functional connectivity analyses in humans also strongly connected
 the vmPFC with the NAc, hippocampus (HC),
@@
 This may be why performance within and across participants
-was related to state encoding in the vmPFC.
+was related to state encoding in the vmPFC \citep{Schuck2016}.
 Such a ``cognitive map'' of the action space was argued to encode
-the current task state even when states are unobservable from the sensory environment
-\citep{Schuck2016}.
+the current task state even when states are unobservable from the sensory environment.
 
@@
 While the HC
-is traditionally believed to allow remembering the past,
+is traditionally believed to allow recalling past experience,
 there is now increasing evidence for a role
@@
 spike one after another  \citep{johnson2007neural}.
-Such neuronal patterns in the HC appear to directly indicate upcoming behavior
+Such neuronal patterns in the HC appear to directly indicate upcoming behavior,
+such as in \textbf{XXX COMPLETE-ME}
 \citep{pfeiffer2013hippocampal}.
@@
 as well as directly before upcoming experimental trials \citep{diba2007forward}.
-Such hippocampal neurons spking patterns during rest and sleep
+Such spking patterns in hippocampal neurons during rest and sleep
 have been proposed to be critical in communicating local information
 to the neocortex for long-term storage, potentially also in the nodes of the DMN.
-Moreover, invasively triggering experience recall
-in the HC of mice during sleep
+Moreover, in mice, invasively triggering spatial experience recall
+in the HC during sleep
 has been shown to subsequently alter action choice during wakefulness
@@
 recollecting autobiographical memory episodes, as well as
-re-experiencing or newly generating mental scenarios
+re-experiencing or newly constructed mental scenarios
 \citep{hassabis2007patients}.
 
-The HC thus orchestrates re-experience of environmental aspects for
+Thus, the HC probably orchestrates re-experience of environmental aspects for
 consolidations based on reenactment and for integration into
 rich mental scene construction \citep{deuker2016event, bird2010establishing}.
-In this way, the HC may contribute to improving
+In this way, the HC may impact
 ongoing perception of and action on the environment
@@
 The DMN emerges with its midline structures early in human development
-\citep{doria2010}.
-However,
+\citep{doria2010}, while
 the right and left TPJ may become fully integrated into this major brain
@@
 The TPJs are known to exhibit hemispheric differences
-based on cytoarchitectonic borders and gyrification pattern
+based on microanatomical properties and gyrification pattern
 \citep{seghier2013angular}.
@@
 The TPJ in the right-hemispheric DMN (RTPJ)
-has been shown to be intimately related to
-multi-sensory prediction and error signaling.
+has been shown to be closely related to
+multi-sensory prediction and prediction error signaling.
 It is probably central for
@@
 visuo-proprioceptive conflict \citep{Balslev2005}, and
-multi-sensory detection of environmental changes across
+detection of environmental changes across
 visual, auditory, or tactile stimulation in an fMRI experiment
@@
 %
-It was argued that the RTPJ encodes actions and ensuing outcomes
+It was argued that the RTPJ encodes actions and ensuing outcomes,
 without necessarily relating those to value estimation
@@
 jakobs2009effects}.
-Neural activity in the RTPJ has been argued to reflect
+Additionally, neural activity in the RTPJ has been proposed to reflect
 stimulus-driven attentional reallocation to
@@
 Indeed, patients with RTPJ damage have particular difficulties
-with multistep actions \citep{hartmann2005takes}.
+with multi-step actions \citep{hartmann2005takes}.
 In the face of large discrepancies between actual and previously predicted
@@
 
-The TPJ in the left-hemispheric hemisphere (LTPJ),
-in turn, exhibits a close relationship to Wernicke's area
+The TPJ in the left-hemispheric DMN (LTPJ),
+in turn, has a close relationship to Wernicke's area
 involved in understanding semantics, such as in spoken and written language.
@@
 voiced sentences lack meaning (i.e., aphasia).
-Abstracting from the typical speech interpretations in linguistics
+Abstracting from speech interpretations in linguistics
 and neuropsychology,
@@
 Indeed,
-the internal generation of
+the internal production of
 verbalized thought (``language of the mind'') was closely related to the LTPJ
@@
 reassembling world knowledge.
-Isolated building blocks of world statistics probably get reassembled
-in internally generated mental scenarios that
+Isolated building blocks of world structure probably get reassembled
+in internally constructed mental scenarios that
 guide present action choice,
@@
 The LTPJ may hence facilitate the automated prediction of events
-by incorporating experience-derived building blocks of world statistics
+by incorporating experience-derived building blocks of world regularities
 into ongoing action, planning, and problem solving.
@@
 random sampling of possible experiences, as well as
-evaluating instantaneous, delayed, and expected reward outcomes
+estimation of instantaneous and expected delayed reward outcomes
 are key components of intelligent RL agents
@@
   \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
-    the agent takes an action (by following some policy). This generates a reward and an observed
-    state transition of the environment. This process goes on until interrupted or a goal state is reached.}
+    the agent takes an action (by following some policy). This yields a reward and an observed
+    state transition of the environment. This process goes on until interrupted or a goal state is reached.
+    \textbf{XXX Please increase size of the font in the figure}}
   \centering
@@
 
-RL is a problem-solving technique in which, through interactions with an unknown \textit{environment},
+RL is a problem-solving technique in which,
+through repeated interactions with an unknown \textit{environment},
 an agent learns to reach goals and optimize reward signals
 in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
-At a given moment, each taken \textit{action} $a$ triggers an \textit{partially observed} change
+At a given moment, each taken \textit{action} $a$ triggers a change
 in the \textit{state} of the environment
@@
 the environment is only \textit{partially observed}
-in the sense that only part of the current state is observed by
+in the sense that \textbf{XXX [Please Define without using the same words]}
+only part of the current state is observed by
 the agent \citep{starkweather2017dopamine}.
-Assuming volatility of the environment
+We assume that volatility of the environment
 is realistic in a computational model which sets out
-to explain DMN function.
+to explain DMN functions of the human brain.
 % XXX: the following passage is already in the introduction
@@
 \subsection{Markov decision processes}
-We argue that the brain in general and the DMN in particular
-are a physical system governed by the laws of
+We argue that the brain in general, and the DMN in particular,
+is a physical system governed by the laws of
 physics and can be formally described
@@
 % , defining $s(t)$, the state at time $t$, as the position in state space.
-The process has $memory$ if the next state depends not only on the current state but also on some finite
-number of past states.
+The process has $memory$ if the next state depends not only on the current state 
+but also on a finite number of past states.
 Rational probabilistic planning can be reformulated
@@
 multi-step decision processes in such an environment are
-\textit{Partially Observable Markov Decision Processes (POMDPs)}\citep{sutton1998reinforcement}.
-A POMDP operationalizes a sequential decision process
+\textit{Markov decision processes (MDPs)}\citep{sutton1998reinforcement}.
+For the remainder of this paper, for the sake of simplicity
+we use the abbreviation ``MDP'' to refer to the special case
+of \textit{Partially Observable Markov Decision Processes}.
+A MDP operationalizes a sequential decision process
 in which it is assumed that environment dynamics are determined by a Markov process,
@@
 \paragraph{Model-free reinforcement learning.}
-Model-free RL has had great success in real-world problems including robotics\citep{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering},
+Model-free RL has had great success in real-world problems including robotics \citep{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering},
 etc., just to name a few \textbf{XXX Please expand with 1-2 additional sentences}.
@@
 of expected utility outcomes as a strategy for action choice in uncertain
-environments \citep{gershman2015computational}, a situation perfectly captured by the formalism of POMDPs.
+environments \citep{gershman2015computational}, a situation perfectly captured by the formalism of MDPs.
 It has also long been proposed \citep{dayan2008decision} that there
@@
 In our adopted model-free RL framework,
-a rat might represent such knowledge about the world as follows:
+an agent might represent such knowledge about the world as follows:
 \begin{itemize}
@@
 \begin{definition*}
-Mathematically, a POMDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
+Mathematically, a MDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
 \begin{itemize}
@@
 
-\begin{remark}
-  Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
-  except the Markov property, we can always transform it into an MDP by considering a compounded
-  version of the states
-  $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
-  history up to an including time $t$.
-\end{remark}
-
 \subsubsection{Accumulated rewards and policies}
@@
 It maps states of the world to
-a set of candidate actions to perform given this state. Starting a time $t=0$,
-a policy $\pi$ generates a trajectory of action cascades / trajectories as follows:
+a set of candidate actions that can be performed given this state. Starting at time $t=0$,
+a policy $\pi$ generates a trajectory of action choices as follows:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-
-\begin{mdframed}
-\textbf{Time-invariance assumption:} It is assumed that the underlying MDP is stationary -- i.e., has reached steady-state conditions --
-  so that the dynamics of the process are equivalent over sufficiently long time windows of equal length.
-\end{mdframed}
-
-Since an action taken in the present moment might have repercussions in the far future, it turns out that the
+We assume time-invariance
+in that we expect the dynamics of the process
+to be equivalent over sufficiently long time windows of equal length (i.e., stationarity).
+
+Since an action executed in the present moment might have repercussions in the far future, it turns out that the
 quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
@@
 following a policy $\pi$.
-\begin{mdframed}
-  \textbf{Where is ``value'' buffered in the DMN ?
-    % brain ?
-    }
-  The amygdala is known to be involved in evaluation of biological significance,
-  whereas
-  the NAc is involved in reward evaluation. Both are known to
-  be connected to the vmPFC of the DMN.
-\end{mdframed}
-  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$, in an average sense to be made precise latter. In
-  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
-Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who is solely concerned about its immediate rewards.
+%
+Note that value buffering may be realized in the vmPFC.
+This DMN node has direct connections to
+to the NAc, known to be involved in reward evaluation,
+and
+the amygdala, known to be involved in evaluation of biological significance.
+
+The goal of the RL agent is then to update this policy in order
+to maximize $G^\pi$, in an average sense to be made precise latter. In
+the definition of cumulative reward $G^\pi$ \eqref{eq:cumr},
+the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}, viewed to be specific an agent.
+Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who
+is hedonistic in being exclusively concerned with immediate rewards.
 Such an agent would have no horizon for the future,
@@
 To allow a learning process to arise,
-it is necessary that $0 < \gamma < 1$.
+it is necessary to set $0 < \gamma < 1$.
 % \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
@@
 rewards expected at a later time. Precisely, rewards that are not expected to come within $\tau := 1/(1 - \gamma)$ time steps from the present point are disregarded.
-This reduces the variance of expected rewards accumulated across action cascades by limiting the length of this cascades.
+This reduces the variance of expected rewards accumulated across
+considered action cascades by limiting the depth of the search tree.
 % given that there is more uncertainty in the farsighted future.
@@
 \subsubsection{The objective value of an action given a state}
-Given only the limited information available from a POMDP, at a state $s$ the average quality or usefulness of taking an action $a$ under a policy $\pi$ is completely captured by the single number
+Given only the limited information available from a MDP, at a state $s$ the average quality or usefulness of taking an action $a$ under a policy $\pi$ is completely captured by the single number
 % We can now formulate the purpose of the DMN
@@
 The learning goal is to approach the
-policy $\pi^*$ as ``close as possible'', that is to \text{solve} the POMDP.
+policy $\pi^*$ as ``close as possible'', that is to \text{solve} the MDP.
 Of course \eqref{eq:qlearning} is just a definition and
-represents by no means, a candidate schema for solving POMDPs with even moderately-sized action and state spaces: it's intractable!
+represents by no means, a candidate schema for solving MDPs with even moderately-sized action and state spaces: it's intractable!
 % where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
@@
 At the end of this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced, which represents an 
-behavior unit of the agent and is recorded in replay memory buffer --which we hypothesize to be the
-hippocampus--, possibly discarding the oldest entries to make space:
+behavior unit of the agent and is recorded in replay memory buffer -- which we hypothesize to be the
+HC --, possibly discarding the oldest entries to make space:
 $\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
@@
 %, but the replay has a much faster time scale.
-In the human brain, the hippocampus is likely to contribute to such a mechanism,
+In the human brain, the HC is likely to contribute to such a mechanism,
 as neuroscience experiments on various animals have repeatedly indicated
@@
 In RL, Monte-Carlo simulation can be used to update the agent's belief state
-\citep{silver2010monte}. Such methods have a sample complexity that is determined only by the underlying difficulty of the POMDP, rather than the size of the state space or observation space,
+\citep{silver2010monte}. Such methods have a sample complexity that is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
 which can be prohibitively large.
@@
 highest expected cumulative rewards.
-In POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
+In MDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
 \textbf{XXX Elvis: refs needed.}.
@@
 Such cognitive processes can afford
-the internal generation of necessary information
+the internal construction and elaboration of necessary information
 that is not presented in the surrounding environment
@@
 \citep{schacter2007remembering}.
-Such scene generation processes could contribute to interpreting the
+Such scene construction processes could contribute to interpreting the
 present and foretelling the future.
@@
 The semantic hypothesis
-and MDP account provide explanations why hippocampus damage does
+and MDP account provide explanations why HC damage does
 not only impair recalling memories, but also hypothetical and future
@@
 \paragraph{Acknowledgment.}
-Starwars!
 % {\small The research leading to these results has received funding from the
@@
 
+\begin{remark}
+  Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
+  except the Markov property, we can always transform it into an MDP by considering a compounded
+  version of the states
+  $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
+  history up to an including time $t$.
+\end{remark}
+
 \section{More on RL in the brain}
@@
   Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
-  a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
+  a MDP (called a \textit{teacher's demonstration}), can we figure out what is the
   (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A

### 2017-04-25 paper/paper.tex
awesomify 3.0

@@
 
-The goal of the RL agent is then to update this policy in order
-to maximize $G^\pi$, in an average sense to be made precise latter. In
-the definition of cumulative reward $G^\pi$ \eqref{eq:cumr},
-the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}, viewed to be specific an agent.
-Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who
-is hedonistic in being exclusively concerned with immediate rewards.
-Such an agent would have no horizon for the future,
-which is not compatible with long-term goal planning
-as potentially realized in the DMN.
-To allow a learning process to arise,
-it is necessary to set $0 < \gamma < 1$.
+The goal of the RL agent is then to update this action policy in order
+to maximize $G^\pi$ on average (cf. below). In \eqref{eq:cumr},
+the definition of cumulative reward $G^\pi$,
+the constant $\gamma$ $(0 \le \gamma < 1$) is the \textit{reward discount factor},
+viewed to be specific an agent.
+On the one hand,
+setting $\gamma = 0$ yields perfectly hedonistic behavior.
+An agent with such a shortsighted horizon is exclusively
+concerned with immediate rewards.
+This is however not compatible with coordinated planning if long-term goal that is
+potentially subserved by neural activity in the DMN.
 % \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
 % The reward discount factor
-$\gamma$ can be seen as calibrating risk-seeking trait of the intelligent agent,
-that is, the behavioral predispositions related to reward-delay trade-off schedules.
+On the other hand,
+setting $0 < \gamma < 1$ allows a learning process to arise.
+A positive $\gamma$ can be seen as calibrating risk-seeking trait of the intelligent agent,
+that is, the behavioral predispositions related to trading longer delays for higher rewards.
 Such an agent puts relatively more emphasis on rewards expected in
-a short future and pays less attention to
+a longer-term future and pays less attention to
 rewards expected at a later time. Precisely, rewards that are not expected to come within $\tau := 1/(1 - \gamma)$ time steps from the present point are disregarded.
@@
 considered action cascades by limiting the depth of the search tree.
-% given that there is more uncertainty in the farsighted future.
-% It is also important to appreciate that a stochastic policy estimation
-% is more advantageous in many RL settings.
+Given that there is more uncertainty in the farsighted future,
+it is important to appreciate that a stochastic policy estimation
+is more advantageous in many RL settings.
 % For instance, a deterministic strategy in playing rock-paper-scissors can
@@
 \subsubsection{The objective value of an action given a state}
-Given only the limited information available from a MDP, at a state $s$ the average quality or usefulness of taking an action $a$ under a policy $\pi$ is completely captured by the single number
+Given only the limited information available from a MDP, at a state $s$ the average 
+utility of choosing an action $a$ under a policy $\pi$ can be captured by the single number
 % We can now formulate the purpose of the DMN
@@
 $Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
-utility of executing a specific action; it answers the question
+utility of executing a specific action.
+Thus, it answers the question
 ``What is the expected utility of taking action $a$ in this situation ?''.
@@
 
-Optimal behavior corresponds to a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. Formally,
+Optimal behavior of the agent corresponds to
+a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. Formally,
 % \textbf{XXX please add 1-2 sentences why this is interesting in the current
@@
 % table.
-Fortunately, the \textit{Bellman equation} \citep{sutton1998reinforcement} provides a fixed-point relation which defines $Q^*$ implicitly via a sampling procedure, without querying the entire space of policies, viz
+Fortunately, the \textit{Bellman equation} \citep{sutton1998reinforcement} provides a fixed-point relation which defines $Q^*$ implicitly via a sampling procedure, without querying the entire space of policies, with the form
 \begin{equation}
@@
 a recursive decomposition of optimal behavior by dividing the $Q$-value function into the immediate
-reward and the discounted rewards of the upcoming states.
+reward and the discounted rewards of the upcoming states in the environment.
 The optimal Q-value operator $Q^*$ is a fixed point for this equation.
@@
 Exploitation of hierarchical structure in action considerations
-has previously been closely related to the medial prefrontal part of the DMN
+has previously been related to the medial prefrontal part of the DMN
 \citep{koechlin1999role, braver2002role}.
 Using the Bellman equation, each state can be associated with a certain value
-to guide action towards a better state, thus improving on the agent's current policy.
+to guide action towards a preferred state, thus improving on the agent's current policy.
 % It is a fixed-point equation for the deterministic policy $\pi$.
-Note that in \eqref{eq:bellman} the sampling is done only over things which
-depend on the environment, and so can be learned off-policy by observing state transitions
+Note that in \eqref{eq:bellman} the random sampling
+is performed only over quantities which
+depend on the environment.
+This aspect of the learning process
+can unroll off-policy by observing state transitions
 triggered by another behavioral policy, which can be stochastic.
@@
 At a time step $k-1$, the world is in a state $s \in \mathcal S$, and the agent greedily takes an
-action which it expects to be the must valuable on average, namely
+action which it expects to be the most valuable on average, namely
 
@@
 \end{equation}
-by querying an oracle which maps from states directly to actions. Here $\tilde{Q}(s,a|\theta)$ is an
+by querying an oracle \textbf{XXX Please let's not use "oracle" anywhere in
+the paper, if possible, because it will confuse a lot of peoople.}
+which maps from states directly to actions. Here $\tilde{Q}(s,a|\theta)$ is an
 approximate version of the $Q$-value function parameterized by the tensor $\theta$.
@@
 $\phi(s,a)$ would represent a high-level representation of the state-action pairs
-$(s,a)$, as proposed by \citep{songNIPS2016} for example.
+$(s,a)$, as proposed by \citep{songNIPS2016}.
 %  and Parr et al. 2003 (JMLR)].
-In the DMN, the dmPFC (BA9) would implement such a greedy lookup
+In the DMN, the dmPFC would implement such a greedy lookup
 over the action space.
@@
   The oracle Q-values $\tilde{Q}(s, a|\theta)$ tell the DMN the (expected) usefulness of taking an action a in state $s$.
-  The DMN, and in particular its dmPFC node, would then contribute to the choice, at a given state $s$, of an action $a$ which maximizes this tabulated values.
+  The DMN, and in particular its dmPFC node, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes this tabulated values.
 This mapping from states to actions is conventionally called ``policy matrix''.
-Learning consists in starting from a given instantiation of such a table and tweaking it around while using it to take actions which take the agent to different areas of the table, and so forth.
+  Learning consists in starting from a given table and
+  updating it during action choices,
+  which take the agent to different table entries.
 
@@
 \subsubsection{Self-training and the loss function}
-Successful learning in brains and computer algorithms is challenging without a clear learning goal --
-the \textit{loss function}.
+Successful learning in brains and computer algorithms is challenging without a
+defined learning goal -- the \textit{loss function}.
 The action $a$ chosen in state $s$ according to the policy matrix defined in
@@
 % \textbf{XXX Please try to smoothen the previous sentence}
-At the end of this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced, which represents an 
-behavior unit of the agent and is recorded in replay memory buffer -- which we hypothesize to be the
+One such cycle yields a new \textit{experience} $e = (s,a,r,s')$.
+Each cycle represents a behavior unit of the agent
+and is recorded in replay memory buffer -- which we hypothesize to be the
 HC --, possibly discarding the oldest entries to make space:
 $\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
-At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1} + \delta \theta_{k}$ of the parameters for its approximate model of $Q$-value function. This warrants a learning process
+
+At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1} + \delta \theta_{k}$ of the parameters for its approximate model of the $Q$-value function.
+This warrants a learning process
 and definition of a loss function.
 % \subsubsection{Stochastic gradient in the DMN via experience replay}
-The Bellman equation \eqref{eq:bellman} provides a way to produce such a loss function \eqref{eq:oracle} as we now outline.
+The Bellman equation \eqref{eq:bellman} provides a way to obtain such a loss function \eqref{eq:oracle} as we outline in the following.
 % A practically successful RL approach
@@
 % \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.})
-mini-batches of experiences $e$
+batches of experiences $e$
 $(s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$.
@@
 experience replay can be manifested in the re-occurrence of a
-neuron spiking sequence that also occurs during specific actions.
+neuron spiking sequence that also occurred during specific prior actions.
 %, but the replay has a much faster time scale.
-In the human brain, the HC is likely to contribute to such a mechanism,
+The HC is likely to contribute to such a mechanism,
 as neuroscience experiments on various animals have repeatedly indicated
@@
 
-At the current iteration $k$, computing an optimal parameter update then corresponds to finding the parameters $\theta_{k}$ which minimize the following mean-squared loss function
+At the current step $k$, computing an optimal parameter update then corresponds to
+finding the model parameters $\theta_{k}$ which minimize the following mean-squared loss function
 \begin{equation}
@@
 \end{equation}
-where $y_k$ is a defined in \eqref{eq:bellman}.
+where $y_k$ is defined in \eqref{eq:bellman}.
 % \textbf{We still may need to explain y in the formula with one sentence.}
@@
 \subsubsection{Optimal control via stochastic gradient descent in the DMN}
-Learning of the the entire model parameters can effectively be achieved
-via \textit{gradient descent},
-where reward prediction errors -- also known as \textit{regret}-- percolate
+Efficient learning of the entire set of model parameters can effectively be achieved
+via \textit{gradient descent}.
+\textbf{XXX Please add one explanatory sentence for GD.}
+Reward prediction errors -- also known as \textit{regret}-- percolate
 from lower to higher processing layers to modulate the choice of future actions
@@
 are attracted and absorbed by the optimal $Q$-value function $Q^*$ defined implicitly by the Bellman equation \eqref{eq:bellman}.
+\textbf{XXX Please say something about *stochasti*, which you can then
+also use as transition into the next passage.}
 
@@
 In RL, Monte-Carlo simulation can be used to update the agent's belief state
-\citep{silver2010monte}. Such methods have a sample complexity that is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
+\citep{silver2010monte}.
+Such methods have scaling as a function of available data (i.e., sample complexity) that
+is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
 which can be prohibitively large.
@@
 observations and rewards \citep{aronov2017, chao2017interaction, boyer2008evolutionary}.
-These simulations would be used to update the value function, without ever looking inside the black box describing the model's dynamics \citep{lavilleon2015}.
+These simulations of experience batches
+would be used to update the value function, without ever looking inside the black box describing the model's dynamics \citep{lavilleon2015}.
 This would be a simple control algorithm by evaluating all legal actions and selecting the action with
@@
 The HC may provide perturbed action-transition-state-reward samples
-  as batches of imagined, hypothesized, recalled experience.
+  as batches of imagined, hypothesized, and recalled experience.
   The small variations in these experience samplings allow searching
@@
 to best accommodate and prepare the organism
-for the constant influx of sensory examples and
+for processing the constant influx of sensory stimulation and
 performing action on the environment.
@@
 both accounts are plausibilized by neuroscientific evidence that suggest
-the human brain to be a ``statistical organ'' with the biological purpose to
+the human brain to be a ``statistical organ'' \citep{friston2014phantastic}
+with the biological purpose to
 generalize from the past to new experiences.
@@
 back projections indeed outnumber by far the axonal input projections
-existing in the primate brain \citep{salin1995corticocortical}.
+existing in the monkey and probably also human brain 
+\citep{salin1995corticocortical}.
 Moreover,
@@
 In our MDP view, instead,
-the Markov property of our sequential decision making framework entails
+the Markov property of the sequential decision making framework entails
 the environmental model to be
@@
 In the present account, the horizon of that
-look into the future is explicitly expressed in the $\gamma$ parameter
+look into the future is made explicit in the $\gamma$ parameter
 of the Bellman equation.
@@
 \citep{goodfellow2016deep}.
+
 In sum,
@@
 represent neural processes that are likely to
-build on abstract world knowledge and logical assocations
+build on abstract world knowledge and logical associations
 for integrating the constituent elements in mental scene construction
@@
 In comparison,
-both the semantic hypothesis and the present account based on MDPs
+both the semantic hypothesis and the present formal account based on MDPs
 expose mechanisms of how action considerations could be mentally explored.
@@
 the semantic account of DMN function does not offer
-a mechanistic explanation how explicit world knowledge and semantic analogies thereof
+a mechanistic explanation \textit{how}
+explicit world knowledge and semantic analogies thereof
 lead to prediction of future actions and states.
@@
 are explained in MDPs by the advantages of stochastic optimization.
-Yet, the MDP-based account provides an algorithmic explanation in that
+Our MDP account provides an algorithmic explanation in that
 stochasticity of the parameter space explored
@@
 In sum, episodic scene construction according to the semantic
-account is lacking an explicit time and incentive model.
+account is lacking an explicit time and incentive model,
+both of which are incorporated into MDPs.
 
@@
 DMN regions have been associated with processing the
-experienced or expected relevance of a diversity of environment stimuli
+experienced or expected relevance of environment stimuli
 \citep{montague2006imaging}.
 Processing self-relevant information was perhaps the first
-cognitive account that was proposed for the DMN \citep{gusnard2001medial}.
+cognitive account that was proposed for DMN function \citep{gusnard2001medial}.
 Since then,
@@
 detect subjectively relevant cues and unexpected events in the environment.
-Propositions of such a sentinel function have however seldom detailed
+Propositions of a sentinel function to underlie DMN activity
+have however seldom detailed
 the mechanisms of
@@
 our account provides a parsimonious explanation for
-the increasingly recognized observation
-of the DMN implication
+the consistently observed DMN implication
 in certain goal-directed experimental tasks and
@@
 Both environment-detached and environment-engaged
-cognitive sets may entail DMN recruitment
-if environmental relevance is processed, manipulated, and
+cognitive processes may entail DMN recruitment
+if real or imagined experience is processed, manipulated, and
 used for predicting the future.
@@
 the complexity of the social systems
-is a likely human-defining property
+is a likely to be a human-defining property
 \citep{tomasello2009cultural}.
@@
 FEP also postulates a dual model called the \textit{recognition} model,
-which works in tandem with the generative model $p_R(\z|\x)$, to accomplish approximate inference.
-The recognition model dreams / fantasizes imaginary worlds $\z$ whilst the latter tries to generate sensations $\x$ which match these dreams!
-
-Because surprisal is intractably difficult to optimize (since we would need to sum over all hidden causes of the sensations), FEP sets out to instead minimize an upper-bound thereupon, namely the \text{free-energy} $F_G^R(\x)$ given by
+which works in tandem with the generative model $p_R(\z|\x)$, to accomplish approximate inference. The recognition model dreams imaginary worlds $\z$,
+while the latter tries to generate sensations $\x$ which match these dreams!
+
+Because surprise is intractably difficult to optimize (since we would need to sum over all hidden causes of the sensations), FEP sets out to instead minimize an upper-bound thereof,
+namely the \text{free-energy} $F_G^R(\x)$ given by
 
@@
 % \end{figure}
-The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}.
+The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
 \citep{dayan1995helmholtz}. As noted already in \citep{dayan1995helmholtz},

### 2017-04-27 paper/paper.tex
stuff

@@
 For instance, B\'alint's syndrome is a neurological disorder of conscious
-awareness that results from damage in the parietal cortex
+awareness that results from medial damage in the parietal cortex
 \citep{balint1909seelenlahmung}.
@@
 Processing self-relevant information was perhaps the first
-cognitive account that was proposed for DMN function \citep{gusnard2001medial}.
+cognitive account that was proposed for DMN function
+\citep{gusnard2001medial, raichle2001pnas}.
 Since then,
@@
 may reflect the brain's continuous tracking of
-relevance in the environment
+relevance in the environment, such as spotting predators,
 as an advantageous evolutionary adaptation \citep{randy2008, hahn2007cingulate}.

### 2017-04-28 paper/paper.tex
some fixups

--- 2017-04-25 paper/paper.tex
@@
 ``mental sketchpad'' \citep{goldman1996prefrontal}, as it
-potentially subserves the de-novo construction and manipulation
+potentially subserves the de-novo generation and manipulation
 of meaning representations instructed by stored semantics and memories
@@
 
-The dmPFC may subserve inference, representation, and assessment
-of one's own and other individuals' action considerations.
 Generally,
@@
 are known to struggle with
-adaptation to new stimuli and events
+adaption to new  stimuli and events
 \citep{stuss1986frontal}.
+The dmPFC may subserve inference, representation, and assessment
+of one's own and other individuals' action considerations.
 Specifically, neural activity in the human dmPFC
@@
 with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
-% In fact, the spatially neighboring dorsal anterior cingulate cortex
-% has also been linked to computing values and efforts of
-% persisting a behavioral plan versus switching the
-% environmental context in several lesion studies \citep{kolling2016value}.
+In fact, the spatially neighboring dorsal anterior cingulate cortex
+has also been linked to computing values and efforts of
+persisting a behavioral plan versus switching the
+environmental context in several lesion studies \citep{kolling2016value}.
 %
 Such highly abstract neural computations necessarily rely on the
-construction of probabilistic internal information drawing from
-episodic memory recall, mental scene elaboration processes,
+generation of probabilistic internal information drawing from
+episodic memory recall, generative construction processes,
 and stored knowledge of the world.
 %
-Moreover,
-in a computational neuroimaging experiment,
-dorsomedial PFC activity preferentially processed the consequences of
+In a computational neuroimaging experiment,
+dorsomedial PFC activity preferentially modeled the consequences of
 action choices that were considered but not actually executed,
 whereas ventromedial PFC (vmPFC) activity
-processed especially value outcomes of performed actions in the environment
+modeled especially value outcomes of performed actions in the environment
 \citep{nicolle2012agent}.
@@
 risk estimation of relevant environmental stimuli.
-This DMN node may
-orchestrate adaptive behavior by bottom-up-driven
+This DMN node is more closely associated with
+orchestrating adaptive behavior by bottom-up-driven
 processing of “what matters now”,
@@
 this DMN node is preferentially connected with reward-related and limbic regions.
-The vmPFC is well known to have strong axonal connections
+The vmPFC is well known to have direct connections
 with the NAc
@@
 subserve RL in humans more broadly \citep{doherty2015structure}.
-Consistently, diffusion MRI tractography in humans and monkeys
-\citep{croxson2005quantitative}
+This is consistent with diffusion MRI tractography in humans and monkeys
+\citep{croxson2005quantitative} that
 quantified the NAc to
 be more connected to the vmPFC than dmPFC in both species.
-Two different functional connectivity analyses in humans also strongly connected
+Two different functional connectivity analyses in humans strongly connected
 the vmPFC with the NAc, hippocampus (HC),
@@
 This may be why performance within and across participants
-was related to state encoding in the vmPFC \citep{Schuck2016}.
+was related to state encoding in the vmPFC.
 Such a ``cognitive map'' of the action space was argued to encode
-the current task state even when states are unobservable from the sensory environment.
+the current task state even when states are unobservable from the sensory environment
+\citep{Schuck2016}.
 
@@
 While the HC
-is traditionally believed to allow recalling past experience,
+is traditionally believed to allow remembering the past,
 there is now increasing evidence for a role
@@
 spike one after another  \citep{johnson2007neural}.
-Such neuronal patterns in the HC appear to directly indicate upcoming behavior,
-such as in \textbf{XXX COMPLETE-ME}
+Such neuronal patterns in the HC appear to directly indicate upcoming behavior
 \citep{pfeiffer2013hippocampal}.
@@
 as well as directly before upcoming experimental trials \citep{diba2007forward}.
-Such spking patterns in hippocampal neurons during rest and sleep
+Such hippocampal neurons spking patterns during rest and sleep
 have been proposed to be critical in communicating local information
 to the neocortex for long-term storage, potentially also in the nodes of the DMN.
-Moreover, in mice, invasively triggering spatial experience recall
-in the HC during sleep
+Moreover, invasively triggering experience recall
+in the HC of mice during sleep
 has been shown to subsequently alter action choice during wakefulness
@@
 recollecting autobiographical memory episodes, as well as
-re-experiencing or newly constructed mental scenarios
+re-experiencing or newly generating mental scenarios
 \citep{hassabis2007patients}.
 
-Thus, the HC probably orchestrates re-experience of environmental aspects for
+The HC thus orchestrates re-experience of environmental aspects for
 consolidations based on reenactment and for integration into
 rich mental scene construction \citep{deuker2016event, bird2010establishing}.
-In this way, the HC may impact
+In this way, the HC may contribute to improving
 ongoing perception of and action on the environment
@@
 The DMN emerges with its midline structures early in human development
-\citep{doria2010}, while
+\citep{doria2010}.
+However,
 the right and left TPJ may become fully integrated into this major brain
@@
 The TPJs are known to exhibit hemispheric differences
-based on microanatomical properties and gyrification pattern
+based on cytoarchitectonic borders and gyrification pattern
 \citep{seghier2013angular}.
@@
 The TPJ in the right-hemispheric DMN (RTPJ)
-has been shown to be closely related to
-multi-sensory prediction and prediction error signaling.
+has been shown to be intimately related to
+multi-sensory prediction and error signaling.
 It is probably central for
@@
 visuo-proprioceptive conflict \citep{Balslev2005}, and
-detection of environmental changes across
+multi-sensory detection of environmental changes across
 visual, auditory, or tactile stimulation in an fMRI experiment
@@
 %
-It was argued that the RTPJ encodes actions and ensuing outcomes,
+It was argued that the RTPJ encodes actions and ensuing outcomes
 without necessarily relating those to value estimation
@@
 jakobs2009effects}.
-Additionally, neural activity in the RTPJ has been proposed to reflect
+Neural activity in the RTPJ has been argued to reflect
 stimulus-driven attentional reallocation to
@@
 Indeed, patients with RTPJ damage have particular difficulties
-with multi-step actions \citep{hartmann2005takes}.
+with multistep actions \citep{hartmann2005takes}.
 In the face of large discrepancies between actual and previously predicted
@@
 
-The TPJ in the left-hemispheric DMN (LTPJ),
-in turn, has a close relationship to Wernicke's area
+The TPJ in the left-hemispheric hemisphere (LTPJ),
+in turn, exhibits a close relationship to Wernicke's area
 involved in understanding semantics, such as in spoken and written language.
@@
 voiced sentences lack meaning (i.e., aphasia).
-Abstracting from speech interpretations in linguistics
+Abstracting from the typical speech interpretations in linguistics
 and neuropsychology,
@@
 Indeed,
-the internal production of
+the internal generation of
 verbalized thought (``language of the mind'') was closely related to the LTPJ
@@
 reassembling world knowledge.
-Isolated building blocks of world structure probably get reassembled
-in internally constructed mental scenarios that
+Isolated building blocks of world statistics probably get reassembled
+in internally generated mental scenarios that
 guide present action choice,
@@
 The LTPJ may hence facilitate the automated prediction of events
-by incorporating experience-derived building blocks of world regularities
+by incorporating experience-derived building blocks of world statistics
 into ongoing action, planning, and problem solving.
@@
 random sampling of possible experiences, as well as
-estimation of instantaneous and expected delayed reward outcomes
+evaluating instantaneous, delayed, and expected reward outcomes
 are key components of intelligent RL agents
@@
   \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
-    the agent takes an action (by following some policy). This yields a reward and an observed
-    state transition of the environment. This process goes on until interrupted or a goal state is reached.
-    \textbf{XXX Please increase size of the font in the figure}}
+    the agent takes an action (by following some policy). This generates a reward and an observed
+    state transition of the environment. This process goes on until interrupted or a goal state is reached.}
   \centering
@@
 
-RL is a problem-solving technique in which,
-through repeated interactions with an unknown \textit{environment},
+RL is a problem-solving technique in which, through interactions with an unknown \textit{environment},
 an agent learns to reach goals and optimize reward signals
 in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
-At a given moment, each taken \textit{action} $a$ triggers a change
+At a given moment, each taken \textit{action} $a$ triggers an \textit{partially observed} change
 in the \textit{state} of the environment
@@
 the environment is only \textit{partially observed}
-in the sense that \textbf{XXX [Please Define without using the same words]}
-only part of the current state is observed by
+in the sense that only part of the current state is observed by
 the agent \citep{starkweather2017dopamine}.
-We assume that volatility of the environment
+Assuming volatility of the environment
 is realistic in a computational model which sets out
-to explain DMN functions of the human brain.
+to explain DMN function.
 % XXX: the following passage is already in the introduction
@@
 \subsection{Markov decision processes}
-We argue that the brain in general, and the DMN in particular,
-is a physical system governed by the laws of
+We argue that the brain in general and the DMN in particular
+are a physical system governed by the laws of
 physics and can be formally described
@@
 % , defining $s(t)$, the state at time $t$, as the position in state space.
-The process has $memory$ if the next state depends not only on the current state 
-but also on a finite number of past states.
+The process has $memory$ if the next state depends not only on the current state but also on some finite
+number of past states.
 Rational probabilistic planning can be reformulated
@@
 multi-step decision processes in such an environment are
-\textit{Markov decision processes (MDPs)}\citep{sutton1998reinforcement}.
-For the remainder of this paper, for the sake of simplicity
-we use the abbreviation ``MDP'' to refer to the special case
-of \textit{Partially Observable Markov Decision Processes}.
-A MDP operationalizes a sequential decision process
+\textit{Partially Observable Markov Decision Processes (POMDPs)}\citep{sutton1998reinforcement}.
+A POMDP operationalizes a sequential decision process
 in which it is assumed that environment dynamics are determined by a Markov process,
@@
 \paragraph{Model-free reinforcement learning.}
-Model-free RL has had great success in real-world problems including robotics \citep{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering},
+Model-free RL has had great success in real-world problems including robotics\citep{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering},
 etc., just to name a few \textbf{XXX Please expand with 1-2 additional sentences}.
@@
 of expected utility outcomes as a strategy for action choice in uncertain
-environments \citep{gershman2015computational}, a situation perfectly captured by the formalism of MDPs.
+environments \citep{gershman2015computational}, a situation perfectly captured by the formalism of POMDPs.
 It has also long been proposed \citep{dayan2008decision} that there
@@
 In our adopted model-free RL framework,
-an agent might represent such knowledge about the world as follows:
+a rat might represent such knowledge about the world as follows:
 \begin{itemize}
@@
 \begin{definition*}
-Mathematically, a MDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
+Mathematically, a POMDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
 \begin{itemize}
@@
 
+\begin{remark}
+  Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
+  except the Markov property, we can always transform it into an MDP by considering a compounded
+  version of the states
+  $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
+  history up to an including time $t$.
+\end{remark}
+
 \subsubsection{Accumulated rewards and policies}
@@
 It maps states of the world to
-a set of candidate actions that can be performed given this state. Starting at time $t=0$,
-a policy $\pi$ generates a trajectory of action choices as follows:
+a set of candidate actions to perform given this state. Starting a time $t=0$,
+a policy $\pi$ generates a trajectory of action cascades / trajectories as follows:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-We assume time-invariance
-in that we expect the dynamics of the process
-to be equivalent over sufficiently long time windows of equal length (i.e., stationarity).
-
-Since an action executed in the present moment might have repercussions in the far future, it turns out that the
+
+\begin{mdframed}
+\textbf{Time-invariance assumption:} It is assumed that the underlying MDP is stationary -- i.e., has reached steady-state conditions --
+  so that the dynamics of the process are equivalent over sufficiently long time windows of equal length.
+\end{mdframed}
+
+Since an action taken in the present moment might have repercussions in the far future, it turns out that the
 quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
@@
 following a policy $\pi$.
-%
-Note that value buffering may be realized in the vmPFC.
-This DMN node has direct connections to
-to the NAc, known to be involved in reward evaluation,
-and
-the amygdala, known to be involved in evaluation of biological significance.
-
-The goal of the RL agent is then to update this policy in order
-to maximize $G^\pi$, in an average sense to be made precise latter. In
-the definition of cumulative reward $G^\pi$ \eqref{eq:cumr},
-the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}, viewed to be specific an agent.
-Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who
-is hedonistic in being exclusively concerned with immediate rewards.
+\begin{mdframed}
+  \textbf{Where is ``value'' buffered in the DMN ?
+    % brain ?
+    }
+  The amygdala is known to be involved in evaluation of biological significance,
+  whereas
+  the NAc is involved in reward evaluation. Both are known to
+  be connected to the vmPFC of the DMN.
+\end{mdframed}
+  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$, in an average sense to be made precise latter. In
+  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
+Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who is solely concerned about its immediate rewards.
 Such an agent would have no horizon for the future,
@@
 To allow a learning process to arise,
-it is necessary to set $0 < \gamma < 1$.
+it is necessary that $0 < \gamma < 1$.
 % \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
@@
 rewards expected at a later time. Precisely, rewards that are not expected to come within $\tau := 1/(1 - \gamma)$ time steps from the present point are disregarded.
-This reduces the variance of expected rewards accumulated across
-considered action cascades by limiting the depth of the search tree.
+This reduces the variance of expected rewards accumulated across action cascades by limiting the length of this cascades.
 % given that there is more uncertainty in the farsighted future.
@@
 \subsubsection{The objective value of an action given a state}
-Given only the limited information available from a MDP, at a state $s$ the average quality or usefulness of taking an action $a$ under a policy $\pi$ is completely captured by the single number
+Given only the limited information available from a POMDP, at a state $s$ the average quality or usefulness of taking an action $a$ under a policy $\pi$ is completely captured by the single number
 % We can now formulate the purpose of the DMN
@@
 % \textit{greedy policy}
-\subsubsection{Optimal behavior and the Bellman equation}
+\subsubsection{Optimal behavior and the Bellman temporal consistency}
 
@@
 The learning goal is to approach the
-policy $\pi^*$ as ``close as possible'', that is to \text{solve} the MDP.
+policy $\pi^*$ as ``close as possible'', that is to \text{solve} the POMDP.
 Of course \eqref{eq:qlearning} is just a definition and
-represents by no means, a candidate schema for solving MDPs with even moderately-sized action and state spaces: it's intractable!
+represents by no means, a candidate schema for solving POMDPs with even moderately-sized action and state spaces: it's intractable!
 % where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
@@
   \end{equation}
-The Bellman equation \eqref{eq:bellman} is a ``self-consistency'' equation which provides
-a recursive decomposition of optimal behavior by dividing the $Q$-value function into the immediate
+The Bellman equation \eqref{eq:bellman} is a temporal consistency equation which provides
+a dynamic decomposition of optimal behavior by dividing the $Q$-value function into the immediate
 reward and the discounted rewards of the upcoming states.
@@
 
+\subsubsection{Bounded rationality}
+$$\max_{a \in \mathcal A}\tilde{Q}(s, a|\theta)  \propto \lim_{T \rightarrow 0^+}\exp(\tilde{Q}(s,a|\theta)/T)$$
+
 \subsubsection{Self-training and the loss function}
@@
 At the end of this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced, which represents an 
-behavior unit of the agent and is recorded in replay memory buffer -- which we hypothesize to be the
-HC --, possibly discarding the oldest entries to make space:
+behavior unit of the agent and is recorded in replay memory buffer --which we hypothesize to be the
+hippocampus--, possibly discarding the oldest entries to make space:
 $\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
@@
 %, but the replay has a much faster time scale.
-In the human brain, the HC is likely to contribute to such a mechanism,
+In the human brain, the hippocampus is likely to contribute to such a mechanism,
 as neuroscience experiments on various animals have repeatedly indicated
@@
 In RL, Monte-Carlo simulation can be used to update the agent's belief state
-\citep{silver2010monte}. Such methods have a sample complexity that is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
+\citep{silver2010monte}. Such methods have a sample complexity that is determined only by the underlying difficulty of the POMDP, rather than the size of the state space or observation space,
 which can be prohibitively large.
@@
 highest expected cumulative rewards.
-In MDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
+In POMDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
 \textbf{XXX Elvis: refs needed.}.
@@
 Such cognitive processes can afford
-the internal construction and elaboration of necessary information
+the internal generation of necessary information
 that is not presented in the surrounding environment
@@
 \citep{schacter2007remembering}.
-Such scene construction processes could contribute to interpreting the
+Such scene generation processes could contribute to interpreting the
 present and foretelling the future.
@@
 The semantic hypothesis
-and MDP account provide explanations why HC damage does
+and MDP account provide explanations why hippocampus damage does
 not only impair recalling memories, but also hypothetical and future
@@
 \textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
+\textbf{XXX: integrate aspects from WakeNsleep by Hinton}
 \textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
@@
 From the computational view of a Markovian sequential decision process,
-human behavior unfolds by integrating reward outcomes from stored past events
+behavior unfolds by integrating reward outcomes from stored past events
 and expected upcoming events to guide action choice in the present context.
@@
 \paragraph{Acknowledgment.}
+Starwars!
 % {\small The research leading to these results has received funding from the
@@
 
-\begin{remark}
-  Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
-  except the Markov property, we can always transform it into an MDP by considering a compounded
-  version of the states
-  $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
-  history up to an including time $t$.
-\end{remark}
-
 \section{More on RL in the brain}
@@
   Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
-  a MDP (called a \textit{teacher's demonstration}), can we figure out what is the
+  a POMDP (called a \textit{teacher's demonstration}), can we figure out what is the
   (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A

### 2017-04-28 paper/paper.tex

--- 2017-04-27 paper/paper.tex
@@
   \end{equation}
-The Bellman equation \eqref{eq:bellman} is a ``self-consistency'' equation which provides
-a recursive decomposition of optimal behavior by dividing the $Q$-value function into the immediate
-reward and the discounted rewards of the upcoming states in the environment.
+The Bellman equation \eqref{eq:bellman} is a temporal consistency equation which provides
+a dynamic decomposition of optimal behavior by dividing the $Q$-value function into the immediate
+reward and the discounted rewards of the upcoming states.
 The optimal Q-value operator $Q^*$ is a fixed point for this equation.
@@
 \textbf{XXX Elvis to Elvis: clarify the notion of greed. Add a paragraph on the exploitation vs exploration dilemma, and end with a conclusion saying the DMN probably does epsilon-greedy.}
+
+% \subsubsection{Bounded rationality}
+% $$\max_{a \in \mathcal A}\tilde{Q}(s, a|\theta)  \propto \lim_{T \rightarrow 0^+}\exp(\tilde{Q}(s,a|\theta)/T)$$
 

### 2017-04-30 paper/paper.tex
Figure RL => ODP/PDF + Bigger font size
Solve emdash "—" => \textemdash

@@
 \def\trans{\mathsf{T}}
-
-%\renewcommand{\labelitemi}{--}
 
@@
 The DMN midline is probably closely linked
-with the HC in the medial temporal lobe \citep{vincet2006, shannon2013morning} --
+with the HC in the medial temporal lobe \citep{vincet2006, shannon2013morning} \textemdash
 a region long known to be involved in
@@
     the agent takes an action (by following some policy). This yields a reward and an observed
-    state transition of the environment. This process goes on until interrupted or a goal state is reached.
-    \textbf{XXX Please increase size of the font in the figure}}
+    state transition of the environment. This process goes on until interrupted or a goal state is reached.}
   \centering
-  \includegraphics[width=1.\linewidth]{rl.png}
+  \includegraphics[width=1.\linewidth]{rl.pdf}
   \label{fig:rl}
@@
 the action of the agent and the reward can be thought
-of as satisfaction -- or aversion -- accompanying the execution of
+of as satisfaction \textemdash or aversion \textemdash accompanying the execution of
 an action.
@@
 Successful learning in brains and computer algorithms is challenging without a
-defined learning goal -- the \textit{loss function}.
+defined learning goal \textemdash the \textit{loss function}.
 The action $a$ chosen in state $s$ according to the policy matrix defined in
@@
 Each cycle represents a behavior unit of the agent
-and is recorded in replay memory buffer -- which we hypothesize to be the
-HC --, possibly discarding the oldest entries to make space:
+and is recorded in replay memory buffer \textemdash which we hypothesize to be the
+HC \textemdash, possibly discarding the oldest entries to make space:
 $\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
@@
 \textbf{XXX Please add one explanatory sentence for GD.}
-Reward prediction errors -- also known as \textit{regret}-- percolate
+Reward prediction errors \textemdash also known as \textit{regret}\textemdash percolate
 from lower to higher processing layers to modulate the choice of future actions

### 2017-04-30 paper/paper.tex
philadelphia notes

--- 2017-04-28 paper/paper.tex
@@
 
-This dark matter of brain physiology begs the question
-of the biological purpose underlying DMN activity
-\citep{raichle2006brain}.
+This dark matter of brain physiology \citep{raichle2006brain} begs the question
+of the biological purpose underlying DMN activity.
 What has early been described as the ``stream of consciousness''
@@
 a human agent faced with the choice of the next actions
-and guided by outcome estimations
+and guided by outcomes
 of really happened, hypothetically imagined, and
@@
 
-We argue that DMN implication in a
-diversity of most advanced cognitive processes in humans
+We argue that DMN implication in many
+advanced cognitive processes in humans
 can be recast as prediction error minimization
@@
   Neurobiological overview of the DMN with its major constituent nodes and
-  their associated functional roles relevant in our functional account.}
+  the associated functional roles relevant in our functional account.}
   \label{fig:neurobiological_overview_DMN}
@@
 
-The PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
-\citep{goldman1987development}, which is generally considered to
-be a sign of evolutionary sophistication \citep{flechsig1920}.
 Normal and disturbed metabolic fluctuations in the
@@
 changes of conscious awareness \citep{cavanna2006precuneus}.
-%
 Indeed,
-the PMC has long been speculated to reflect constant computation of
+the PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
+\citep{goldman1987development}, which is generally considered to
+be a sign of evolutionary sophistication \citep{flechsig1920}.
+%
+This DMN node has long been speculated to
+reflect constant computation of
 environmental statistics and its internal representation
@@
 environmental relevance
-and ensuing top-level control of action on the environment is supported
+and ensuing top-level guidance of action on the environment is supported
 by many neuroscience experiments.
@@
 of possible actions and integrating that information
-with memory, perspective frames, and
-reward outcomes into short- and longer-term behavioral agendas.
-Estimated value, that differs across individuals, may update
+with memory and perspective frames
+into short- and longer-term behavioral agendas.
+Estimated value, that differs across individuals, probably enriches
 statistical assessment of the environment
@@
 In doing so, the PMC may continuously adapt the organism to changes
-in both the external environment and its internally modeled representation
+in both the external environment and its internal representation
 to enable strategic behavior.
@@
 In line with these connectivity findings in animals and humans,
-the vmPFC is often proposed to represent externally or internally triggered
+the vmPFC is often proposed to represent triggered
 emotional and motivational states \citep{damasio1996somatic}.
@@
 specific locations in space during environmental navigation.
-London taxi drivers, individuals with high performance in spatial navigation,
-were shown to exhibit increased gray-matter volume in the
-HC \citep{maguire2000navigation}.
 Indeed, when an animal is choosing between alternative
@@
 Such neuronal patterns in the HC appear to directly indicate upcoming behavior,
-such as in \textbf{XXX COMPLETE-ME}
-\citep{pfeiffer2013hippocampal}.
-
-HC function probably extend beyond simple forms of
+such as in planning navigational trajectories
+\citep{pfeiffer2013hippocampal} and
+memory consolidation of choice relevance \citep{lavilleon2015}.
+Congruently,
+London taxi drivers, humans with high performance in spatial navigation,
+were shown to exhibit increased gray-matter volume in the
+HC \citep{maguire2000navigation}.
+
+There is increasing evidence that
+HC function extends beyond simple forms of
 encoding and reconstruction of memory and spatial information.
-Based on large-scale recordings of hippocampal neuronal populations,
+Based on neuronal spike recordings of hippocampal neuronal populations,
 complex spiking patterns can be followed across extended periods including
@@
 after environmental events \citep{buzsaki2004large}.
-Specific spiking sequences, which were elicited in experimental task conditions,
+Specific spiking sequences, which were elicited by experimental task design,
 have been shown to be reenacted spontaneously during
@@
 probably contribute to advanced cognitive processes that require
-recollecting autobiographical memory episodes, as well as
-re-experiencing or newly constructed mental scenarios
+re-experiencing or newly constructed mental scenarios,
+such as in recalling autobiographical memory episodes
 \citep{hassabis2007patients}.
@@
 in turn, has a close relationship to Wernicke's area
-involved in understanding semantics, such as in spoken and written language.
+involved in semantic processes, such as in spoken and written language.
 Neurological patients with damage in Wernicke's area
@@
 \citep{binder2011neurobiology, seghier2013angular}.
-For instance, LTPJ damage in humans also entail problems in recognizing
+For instance, LTPJ damage in humans also entails problems in recognizing
 others' pantomimed action towards objects
@@
 reassembling world knowledge.
-Isolated building blocks of world structure probably get reassembled
+Isolated building blocks of world structure probably get rebuilt
 in internally constructed mental scenarios that
@@
 We now argue the outlined neurobiological properties
-and functional involvements of the DMN nodes
+of the DMN nodes
 to be sufficient for implementing all components
@@
 are key components of intelligent RL agents
-that are plausible to intersect in the DMN.
+that are plausible to functionally intersect in the DMN.
 
@@
 and action policies by controlling
-synaptic plasticity at targets such as the NAc.
+synaptic plasticity in the reward system, including the NAc.
 In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of the world.
@@
 $Q^{\pi}(s,a)$ offers a formalization of optimal behavior that
-may well reflect processing aspects of the DMN in human agents.
+may well capture processing aspects of the DMN in human agents.
 
@@
 policy $\pi^*$ as ``close as possible'', that is to \text{solve} the MDP.
-Of course \eqref{eq:qlearning} is just a definition and
-represents by no means, a candidate schema for solving MDPs with even moderately-sized action and state spaces: it's intractable!
+Note that \eqref{eq:qlearning} presents merely a definition and
+does not lend itself as a candidate schema
+for solving MDPs with even moderately-sized action and state spaces
+(i.e., intractability).
 % where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
@@
 % to allow for scaling to large state and action spaces.
-Replay consists in sampling
+Experience replay consists in sampling
 % (uniform or importance-weighted
@@
   A policy encapsulates the repertoire of possible actions
-    on the world given a current state. It encodes the probabilites of
+    on the world given a current state.
+    The policy matrix encodes the probabilites of
     choosing actions to be executed in a certain situation.
@@
   (e.g., mind-wandering or sleep),
-  mental scence construction allows for pseudo-experiencing possible
+  mental scene construction allows for pseudo-experiencing possible
   future scenarios and action outcomes.
@@
 \subsection{Semantic hypothesis}
-Another frequently proposed cognitive account of DMN function revolves
+Another frequently proposed cognitive account to explain DMN function revolves
 around forming logical associations and abstract analogies between
-the current experience and
-the conceptual knowledge derived from past experiences
+experiences and conceptual knowledge derived from past behavior
 \citep{bar2007proactive, binder1999conceptual, constantinescu2016organizing}.
 Analogies might naturally tie incoming new sensory stimuli to
-explicit world knowledge (i.e., semantics) extracted from the environment
+explicit world knowledge (i.e., semantics) about the environment
 \citep{bar2009proactive}.
@@
 mentally envisioning non-existing scenarios
-that would optimize behavior in the environment
+that would optimize behavior
 by constantly simulating possible future events
@@
 levels of complexity, abstraction, timescale and purpose
-use mechanisms that are qualitatively different. This concurs with
+rely on mechanisms that are qualitatively different. This concurs with
 DMN activity increases across time, space, and content domains
@@
 \citep{spreng2009common, laird2009, bzdok2012morality, binder2009}.
-The semantic hypothesis
+Further, the semantic hypothesis
 and MDP account provide explanations why HC damage does
@@
 eventually chosen, namely, the estimated cumulative reward.
-\textbf{The notion of semantic knowledge associations is
-incorporated into the MDP as the Markov property,
-that is, the current state directly results from the
-agent's history of states and actions. The learned
+\textbf{The learned
 estimations of value and action transitions drive
 stimulus processing and action choice in the present.}
-The MDPs naturally integrate the former egocentric view
-(more related to current action, state and reward) and the later
+Additionally,
+the MDPs naturally integrate the egocentric view
+(more related to current action, state and reward) and the
 world view (more related to past and future actions, states, and rewards)
 on the world in a same optimization problem.
-Further,
+Finally,
 the semantic account of DMN function does not offer
@@
 account is lacking an explicit time and incentive model,
-both of which are incorporated into MDPs.
+both of which are integral parts of MDP frameworks.
 
 \subsection{Sentinel hypothesis}
-DMN regions have been associated with processing the
+The DMN regions have been associated with processing the
 experienced or expected relevance of environment stimuli
@@
 as an advantageous evolutionary adaptation \citep{randy2008, hahn2007cingulate}.
-According to this cognitive account, the human brain's baseline realizes
+According to this cognitive account, the human brain's baseline maintains
 a ``radar'' function to
@@
 encountering a self-relevant environmental stimulus.
-In the present formal account,
+In the present MDP account,
 promising action trajectories
-are recursively explored by human agents. Conversely,
+are recursively explored by the human DMN. Conversely,
 certain branches of candidate action trajectories
@@
 if real or imagined experience is processed, manipulated, and
-used for predicting the future.
+used for predictions.
 During tasks,
@@
 improve especially mid- and long-term action.
-The horizon is expressed in the $\gamma$ parameter in the formal account.
+This horizon of the agent
+is expressed in the $\gamma$ parameter in the MDP account.
 We thus provide answers for the currently unsettled question why the involvement
 of the same neurobiological brain circuit (i.e., DMN) has been documented
-for baseline house-keeping functions and specific task performances.
+for specific task performances and baseline house-keeping functions.
 
@@
 \citep{whiten1988machiavellian}.
-Indeed, social topics amount to roughly
+Indeed, social topics amounted to roughly
 two thirds of human everyday communication \citep{dunbar1997human},
@@
 %
-In line with this, the DMN has been proposed to reflect
+In line with this, the DMN has been proposed to be specialzed in
 continuous processing of social information as a
@@
 \citep{baker2009action, kampe2001psychology, krienen2010clan}.
-The DMN may thus exert control on the human agent in a way that prioritizes
+The DMN may thus modulate reward processing
+in the human agent in a way that prioritizes
 appraisal of and action towards social contexts,
-without entirely neglecting environmental cues of the physical world.
+without excluding relevance of environmental cues of the physical world.
 In sum,
@@
 in general and
-inherently incorporates the importance of social environmental cues
-as an important special case.
+inherently accommodates the importance of social environmental cues
+as a special case.
 
@@
 to justify constantly high energy costs?
-MDPs provide an attractive
+MDPs motivate an attractive
 formal account how the human association cortex
@@
 %
-From the computational view of a Markovian sequential decision process,
+From the view of a Markovian sequential decision process,
 human behavior unfolds by integrating reward outcomes from stored past events
-and expected upcoming events to guide action choice in the present context.
-MDPs provide a mathematical formalism how
+and extrapolation to
+upcoming events for guiding action choice in the present context.
+MDPs also provide a mathematical formalism how
 opportunity structure in the environment can be recursively exploited
@@
 as well as delay discounting and self-related prospection.
-Improvement of the internal world model
+Improvement of the internal world representation
 by injecting stochasticity in the recall of past
@@
 
-The proposed machine-learning perspective
+The proposed artificiale intelligence perspective
 on DMN biology is hence practically computable and

### 2017-04-30 paper/paper.tex
started dominik notes

@@
 
-%\renewcommand{\labelitemi}{--}
-
 \newtheorem{theorem}{Theorem} \newtheorem{lemma}[theorem]{Lemma}
@@
 on the environment given the current state. A value function
-maps expected rewards to environmental objects and events.
+maps environmental objects and events (i.e., states) to expected rewards.
 Switching between states reduces to a sequential processing model.

### 2017-05-01 paper/paper.tex
a major pass on RL sections

@@
 \textit{Markov decision processes (MDPs)}\citep{sutton1998reinforcement}.
-For the remainder of this paper, for the sake of simplicity
-we use the abbreviation ``MDP'' to refer to the special case
-of \textit{Partially Observable Markov Decision Processes}.
+% For the remainder of this paper, for the sake of simplicity
+% we use the abbreviation ``MDP'' to refer to the special case
+% of \textit{Partially Observable Markov Decision Processes}.
 A MDP operationalizes a sequential decision process
@@
 states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
-and is characteristic of so-called \textit{model-free RL},
-which we propose as a useful formal framework for DMN function.
-%
-
-\paragraph{Model-free reinforcement learning.}
-Model-free RL has had great success in real-world problems including robotics \citep{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering},
-etc., just to name a few \textbf{XXX Please expand with 1-2 additional sentences}.
-Such a learning scheme can be
-plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
+and is characteristic of so-called \textit{model-free RL}. Model-free RL has had great success in real-world problems including robotics
+\citep{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering}, etc., just to name a few.
+We propose to use such a model as formal basis to model DMN function.
+
+Such a learning scheme can be plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
 a core property of human intelligence has been proposed to underly improvement
@@
 \begin{definition*}
-Mathematically, a MDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
+Mathematically, an MDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
 \begin{itemize}
@@
 \subsection{The nuts and bolts of reinforcement learning in the DMN}
-\textbf{XXX Elvis to Elvis a short preface is needed here, explaining the roadmap of this subsection from a high-level perspective!}
-
-\subsubsection{The objective value of an action given a state}
-Given only the limited information available from a MDP, at a state $s$ the average
+Given only the limited information available from an MDP, at a state $s$ the average
 utility of choosing an action $a$ under a policy $\pi$ can be captured by the single number
@@
 \end{equation}
-by querying an oracle \textbf{XXX Please let's not use "oracle" anywhere in
-the paper, if possible, because it will confuse a lot of peoople.}
-which maps from states directly to actions. Here $\tilde{Q}(s,a|\theta)$ is an
-approximate version of the $Q$-value function parameterized by the tensor $\theta$.
+by querying an an approximate version $\tilde{Q}(s,a|\theta)$ of the $Q$-value function, parameterized
+by the tensor $\theta$. This defines a mapping from states directly to actions.
 % $a \mapsto \tilde{Q}(a,s|\theta)$ of the $Q$-value function ``frozen'' at state $s$.
@@
 Efficient learning of the entire set of model parameters can effectively be achieved
-via \textit{gradient descent}.
-\textbf{XXX Please add one explanatory sentence for GD.}
-Reward prediction errors \textemdash also known as \textit{regret}\textemdash percolate
-from lower to higher processing layers to modulate the choice of future actions
+via stochastic \textit{gradient descent}, a first-order\footnote{ i.e using only differential information up to order 1.} algorithm for finding
+local minimal of differentiable functions. Stochastic here that the true gradient is estimated from batches of training examples samples from the replay memory, the HC
 
@@
 \end{equation}
-Over time (i.e., in the limit $k \rightarrow \infty$),
+Thus actions are taken to drive reward prediction errors \textemdash also known as \textit{regret}\textemdash percolate
+from lower to higher processing layers to modulate the choice of future actions .Over time (i.e., in the limit $k \rightarrow \infty$),
 the thus generated approximating sequence of
@@
 are attracted and absorbed by the optimal $Q$-value function $Q^*$ defined implicitly by the Bellman equation \eqref{eq:bellman}.
-\textbf{XXX Please say something about *stochasti*, which you can then
-also use as transition into the next passage.}
-
-\subsubsection{Link to classical reinforcement learning algorithms}
+
 One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
 learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
-cast in this \textbf{please unpack the "this" here}
-general variational framework. For instance, TD corresponds to
+cast in the general variational framework \eqref{eq:oracle}. For instance, TD corresponds to
 the above framework using a linear value approximator with feature encoding
@@
   Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
-  a MDP (called a \textit{teacher's demonstration}), can we figure out what is the
+  an MDP (called a \textit{teacher's demonstration}), can we figure out what is the
   (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A

### 2017-05-01 paper/paper.tex
more

@@
 
-\textbf{XXX Elvis to Elvis: clarify the notion of greed. Add a paragraph on the exploitation vs exploration dilemma, and end with a conclusion saying the DMN probably does epsilon-greedy.}
+  \textbf{XXX Elvis to Elvis: clarify the notion of greed. Add a paragraph on the exploitation vs exploration dilemma, and end with a conclusion saying the DMN probably does epsilon-greedy
+    or softmax.}
 
@@
 \end{equation}
-Thus actions are taken to drive reward prediction errors \textemdash also known as \textit{regret}\textemdash percolate
+Thus next action is taken to drive reward prediction errors \textemdash also known as \textit{regret}\textemdash percolate
 from lower to higher processing layers to modulate the choice of future actions .Over time (i.e., in the limit $k \rightarrow \infty$),
@@
 various stages of sensory input processing.
-In our MDP view, instead,
-the Markov property of the sequential decision making framework entails
-the environmental model to be
-\textbf{effectively expressed in the current state of the human agent
-XXX guys, please double-check}.
+The sequential decision-making framework inherent in our MDP view of the DMN,
+replaces such a top-down generative design.
 Further,

### 2017-05-01 paper/paper.tex
restoring intended meaning

@@
 \subsubsection{Self-training and the loss function}
-Successful learning in brains and computer algorithms is challenging without a
+Successful learning in brains and computer algorithms is not possible without a
 defined learning goal \textemdash the \textit{loss function}.

### 2017-05-03 paper/paper.tex
itemr1

@@
 We propose that this set of some of the most advanced regions in
-the human association cortex \citep{mesulam1998sensation, margulies2016situating}
+the association cortex \citep{mesulam1998sensation, margulies2016situating}
 are responsible for higher-order control of human behavior.
@@
 The human DMN is responsible for a large part of the baseline neural activity,
-which typically decreases when engaged in psychological experiments.
+which typically decreases when engaged in psychological experiments
+\citep{raichle_baseline}.
 The standard mode of
@@
 This brain network
-might have emerged to continuously predict environmental events using
-mental imagery as an evolutionary advantage.
+might have emerged to continuously predict the environment using
+mental imagery as an evolutionary advantage \citep{suddendorf2007evolution}.
 %
@@
 through integration of feedback outcomes.
-Here,
+Henceforth,
 \textit{control} refers to the influence that an agent exerts when interacting
-with the environment to encourage preferred states.
+with the environment to reach preferred states.
 
 Psychologically,
+the more the ongoing executed task is unknown and unpracticed,
+the less stimulus-independent thoughts occur
+\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
+Conversely,
 it is known that, the more the world is easy to predict,
@@
 \citep{antrobus1966studies, pope1978regulation}.
-Conversely,
-the more the ongoing executed task is unknown and unpracticed,
-the less stimulus-independent thoughts occur
-\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
 Without requiring explicit awareness by the agent,
@@
 We formalize
-a policy matrix to capture the space of possible actions that can be performed
-on the environment given the current state. A value function
+a \textit{policy matrix} to capture the space of possible actions that can be performed
+on the environment given the current state. A \textit{value function}
 maps environmental objects and events (i.e., states) to expected rewards.
@@
 Informed by outcomes of really performed action,
-the DMN dynamics are constantly adapted in feedback loops
+the DMN dynamics could be constantly adapted in feedback loops
 shaped by prediction error.
-A RL account of DMN function can naturally embed human behavior
+Such an RL account of DMN function can naturally embed human behavior
 into the tension between exploitative action with immediate gains and
@@
 \citep{tenenbaum2011grow}.
-Even (necessarily imperfect) memory recall,
+Even necessarily imperfect memory recall,
 random day-time mind-wandering, and
 seemingly arbitrary dreams during sleep
-may provide meaningful building blocks to iteratively improve
-the predictive DMN machinery to optimize the behavior of the organism.
+may provide blocks of pseudo-experience to iteratively
+optimize the behavior of the organism.
 %
@@
 continuously scan environmental input
-for biological significance assessment.
+for biological significance assessment
+\citep{trust_attract}.
 
@@
 This DMN node encoded the
-preference of or aversion to options with uncertain reward outcomes
+preference for or aversion to options with uncertain reward outcomes
 and its spiking activity was more associated with
@@
 probably drawing on sophisticated value representations
-\citep{doherty2015structure}.
+\citep{rolls_OFC, doherty2015structure}.
 Quantitative lesion findings across 344 human individuals confirmed
@@
 that not only channel action towards basic survival needs but also
-subserve RL in humans more broadly \citep{doherty2015structure}.
+subserve reward processing, and thus perhaps RL,
+in humans more broadly \citep{doherty2015structure}.
 Consistently, diffusion MRI tractography in humans and monkeys
@@
 \subsection{The hippocampus: memory, space, and experience replay}
-The DMN midline is probably closely linked
+The DMN midline has close functional links
 with the HC in the medial temporal lobe \citep{vincet2006, shannon2013morning} \textemdash
@@
 action initiation during goal-dircted psychological tasks and for
-sensorimotor control by integrating multi-sensory attention
+sensorimotor behavior by integrating multi-sensory attention
 \citep{corbetta2002control}.
@@
 from the ‘‘internally directed’’ baseline processes
-to, instead, deal with (unexpected) environmental objects and events.
+to, instead, deal with (unexpected) environmental stimuli and events.
 % (TODO: slow endogeneous changes in state transitions by the policy and value matrices)
@@
 Further,
-episodic memory recall and imagination strongly draw on
+episodic memory recall and mental imagery strongly draw on
 reassembling world knowledge.
@@
 %
-The LTPJ may hence facilitate the automated prediction of events
+The LTPJ may hence facilitate the automated environmental predictions
 by incorporating experience-derived building blocks of world regularities
@@
     the agent takes an action (by following some policy). This yields a reward and an observed
-    state transition of the environment. This process goes on until interrupted or a goal state is reached.}
+    state transition of the environment. The process goes on until interrupted or a goal state is reached.}
   \centering
@@
 multi-step decision processes in such an environment are
-\textit{Markov decision processes (MDPs)}\citep{sutton1998reinforcement}.
+\textit{Markov decision processes} (MDPs) \citep{sutton1998reinforcement}.
 % For the remainder of this paper, for the sake of simplicity
@@
 % of \textit{Partially Observable Markov Decision Processes}.
-A MDP operationalizes a sequential decision process
+An MDP operationalizes a sequential decision process
 in which it is assumed that environment dynamics are determined by a Markov process,
@@
 signal (i.e., is likely to be different for another agent in the same state).
-This is performed alongside each transition in the environments state space, and maintains a probability distribution over the set of possible
-states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
+This is performed alongside each transition in the state space of the environment
+ and maintains a probability distribution over the set of possible
+states, based on a set of observations and observation probabilities.
+This is a minimal set of assumptions that can be made about an environment,
 and is characteristic of so-called \textit{model-free RL}. Model-free RL has had great success in real-world problems including robotics
@@
 quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
-cumulative reward estimate which takes into account expected reward in the future.
+cumulative reward estimate which takes into account expected reward
+from action choices in the future.
 A common approach to
@@
 the constant $\gamma$ $(0 \le \gamma < 1$) is the \textit{reward discount factor},
-viewed to be specific an agent.
+viewed to be characteristic for a certain agent.
 On the one hand,
@@
 Such an agent puts relatively more emphasis on rewards expected in
-a longer-term future and pays less attention to
-rewards expected at a later time. Precisely, rewards that are not expected to come within $\tau := 1/(1 - \gamma)$ time steps from the present point are disregarded.
+a longer-term future. 
+Precisely, rewards that are not expected to come within $\tau := 1/(1 - \gamma)$ time steps from the present point are disregarded.
 This reduces the variance of expected rewards accumulated across
@@
 % \textit{greedy policy}
-\subsubsection{Optimal behavior and the Bellman temporal consistency}
+\subsubsection{Optimal behavior and the Bellman equation}
 
@@
 \end{equation}
-where the ``Bellman transform'' $\text{Bel}(Q)$ of an arbitrary $Q$-value function $Q: \mathcal S \times \mathcal A \rightarrow \mathbb R$  is another $Q$-value function defined by
+where the so-called Bellman transform $\text{Bel}(Q)$
+of an arbitrary $Q$-value function $Q: \mathcal S \times \mathcal A \rightarrow \mathbb R$  is another $Q$-value function defined by
 \begin{equation}
@@
 The optimal Q-value operator $Q^*$ is a fixed point for this equation.
-As a consequence of this decomposition, the complicated dynamic programming optimization
+As a consequence of this decomposition, the complicated dynamic programming
 problem \eqref{eq:qlearning}
-is broken into simpler sub-problems at different time points.
+is broken down into simpler sub-problems at different time points.
 Exploitation of hierarchical structure in action considerations
@@
 can unroll off-policy by observing state transitions
-triggered by another behavioral policy, which can be stochastic.
+triggered by another (possibly stochastic) behavioral policy.
 
@@
 % $a \mapsto \tilde{Q}(a,s|\theta)$ of the $Q$-value function ``frozen'' at state $s$.
-For instance, a simplistic linear model with a kernel $\phi$ would be of the
+For instance, a simple linear model with a kernel $\phi$ would be of the
 form $\tilde{Q}(s, a|\theta) = \phi(s,a)^T\theta$, where
 $\phi(s,a)$ would represent a high-level representation of the state-action pairs
-$(s,a)$, as proposed by \citep{songNIPS2016}.
+$(s,a)$, as recently proposed \citep{songNIPS2016}.
 %  and Parr et al. 2003 (JMLR)].
@@
   The DMN, and in particular its dmPFC node, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes this tabulated values.
-This mapping from states to actions is conventionally called ``policy matrix''.
+This mapping from states to actions is conventionally called \textit{policy matrix}.
   Learning consists in starting from a given table and
@@
 From a neurobiological perspective,
-experience replay can be manifested in the re-occurrence of a
-neuron spiking sequence that also occurred during specific prior actions.
+experience replay can be manifested as the re-occurrence of
+neuron spiking sequences that have also occurred during specific prior actions.
 %, but the replay has a much faster time scale.
-The HC is likely to contribute to such a mechanism,
-as neuroscience experiments on various animals have repeatedly indicated
+The HC is a strong candidate to contribute to such a mechanism,
+as neuroscience experiments have repeatedly indicated
 in rats, mice, cats, rabbits, songbirds, and
@@
 via stochastic \textit{gradient descent}, a first-order\footnote{ i.e using only differential information up to order 1.} algorithm for finding
-local minimal of differentiable functions. Stochastic here that the true gradient is estimated from batches of training examples samples from the replay memory, the HC
+local minima of differentiable functions. Stochastic here that the true gradient is estimated from batches of training examples samples from the replay memory, the HC
 
@@
 action trajectories)
-and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than a RL agent without stochastic resampling procedures.
+and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than an RL agent without stochastic resampling procedures.
 
@@
   \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
-  \caption{\textbf{DMN: neurobiological implementation of reinforcement learning.}
-  Overview of how the constituent nodes of the Default Mode Network (DMN)
+  \caption{\textbf{Default mode network:
+  neurobiological implementation of reinforcement learning.}
+  Overview of how the constituent nodes of the DMN
   map onto computational components necessary for an RL agent.}

### 2017-05-03 paper/paper.tex
interm2

@@
 via stochastic \textit{gradient descent}, a first-order\footnote{ i.e using only differential information up to order 1.} algorithm for finding
-local minima of differentiable functions. Stochastic here that the true gradient is estimated from batches of training examples samples from the replay memory, the HC
+local minima of differentiable functions.
+Stochastic here means that the true gradient is estimated from batches of training examples samples from the replay memory, the HC
 
@@
 \end{equation}
-Thus next action is taken to drive reward prediction errors \textemdash also known as \textit{regret}\textemdash percolate
+Thus, the next action is taken to drive reward prediction errors
+(or \textit{regret}) to
+percolate
 from lower to higher processing layers to modulate the choice of future actions .Over time (i.e., in the limit $k \rightarrow \infty$),
@@
 $$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
-Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
+Hence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
 $$\tilde{Q}(s, a) \leftarrow \tilde{Q}(s, a) + \alpha(y-\tilde{Q}(s, a)),$$
@@
 \subsubsection{Does the hippocampus subserve Monte-Carlo sampling?}
-In RL, Monte-Carlo simulation can be used to update the agent's belief state
+In RL, Monte-Carlo simulation is a common means to update the agent's belief state
 \citep{silver2010monte}.
-Such methods have scaling as a function of available data (i.e., sample complexity) that
-is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
-which can be prohibitively large.
 Monte-Carlo simulation provides a simple method for evaluating the value of a state.
@@
 and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than an RL agent without stochastic resampling procedures.
+Such methods have scaling as a function of available data (i.e., sample complexity) that
+is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
+which can be prohibitively large.
 
@@
   activity when humans engage in cognitive processes that are detached from
-  current sensory environment
+  the current sensory environment
   \citep{kenet2003spontaneously, fiser2004small, bzdok2013segregation}
-  and it was proposed to be situated at the top of the brain network hierarchy
+  and this network was proposed to be situated
+  at the top of the brain network hierarchy
   \citep{carhart2010default, margulies2016situating}.
@@
 The HC may provide perturbed action-transition-state-reward samples
-  as batches of imagined, hypothesized, and recalled experience.
+  as batches of pseudo-experience
+  (i.e., imagined, hypothesized, and recalled mental scenarios).
   The small variations in these experience samplings allow searching

### 2017-05-03 paper/paper.tex
this and that

@@
 for biological significance assessment
-\citep{trust_attract}.
+\citep{amygdala_bzdok}.
 
@@
 A prediction error at one of the processing levels
-incurs plasticity changes of neuronal back projections
+incurs plasticity changes of neuronal projections
 (i.e., adapting model parameters)
@@
 the predictive coding hypothesis offers explanations for
-the constructive, generative nature of sensory perception
-\citep{friston2010free} and
+the constructive, non-deterministic nature of sensory perception
+\citep{friston2010free, buzsaki2006rhythms} and
 the intimate relation of motor action to sensory expectations
 \citep{wolpert1995internal, kording2004bayesian}.
-Contextual sensorimotor integration of perception-action cycles
+Contextual integration of sensorimotor perception-action cycles
 may be maintained by top-down modulation
@@
 Further,
-both accounts are plausibilized by neuroscientific evidence that suggest
+both functional accounts are plausibilized by
+neuroscientific evidence that suggest
 the human brain to be a ``statistical organ'' \citep{friston2014phantastic}
@@
 \citep{salin1995corticocortical}.
+These many and diverse modulatory influences from higher cortical areas
+can inject prior knowledge
+at every stage of processing environmental information.
 Moreover,
@@
 human brain decreases processing load devoted to incoming information
-when the environment is predictable because the internal generative
-model only requires updating following descrepancies between
-environmental reality and its internal reinstantiation.
+when the environment is predictable.
+This is because the internal generative
+model only requires updates after descrepancies have occurred between
+environmental reality and its internally reinstantiated representation.
 Increased computation resources are however allocated
@@
 unexpected events are encountered by the organism.
-The predictive coding and formal account
+The predictive coding and MDP account
 naturally evoke a mechanism of brain plasticity in that
 neuronal wiring gets increasingly adapted
-when faced by unprecedented environmental challenges.
-
-While sensory experience is a generative process from both views,
+when faced by unanticipated environmental challenges.
+
+While sensory experience is a constructive process from both views,
 the predictive coding account frames
@@
 various stages of sensory input processing.
-The sequential decision-making framework inherent in our MDP view of the DMN,
-replaces such a top-down generative design.
+This top-down generative design is replaced in our MDP view of the DMN
+by a sequential decision-making framework.
 Further,
@@
 nested prediction of probable upcoming actions, states, and rewards.
-Both accounts capture the consequences of action.
-The predictive coding account is typically explained without
+While both accounts capture the consequences of action,
+the predictive coding account is typically explained without
 explicit parameterization of the agent's horizon and
@@
 for improved top-down modulation
-takes the concrete form of gradient computation and
+takes the concrete form of stochastic gradient computation and
 back-propagation in our MDP implementation.
@@
 In sum,
-the present formal account may serve as a concrete implementation of
+the present MDP account of DMN function may serve as a concrete implementation of
 the long-standing predictive coding account from cognitive neuroscience.
 MDPs have the advantage of exposing an explicit
-mechanism for the horizon of future considerations or
-how the internal model of the world gets updated,
-and
-more explicitly reconciles sensory input processing and action consideration.
+mechanism for the horizon of future considerations and for
+how the internal representation of the world is updated,
+as well as
+more explicitly reconciles sensory input processing and action considerations.
 %
@@
 Analogies might naturally tie incoming new sensory stimuli to
-explicit world knowledge (i.e., semantics) about the environment
+explicit world knowledge (i.e., semantics)
 \citep{bar2009proactive}.
-The encoding of complex environments could thus be facilitated
+The encoding of complex environmental features could thus be facilitated
 by association to similar states.
 %
-Going beyond meaning and concepts extracted from the world,
-semantic building blocks may provide the basis for
-mentally envisioning non-existing scenarios
-that would optimize behavior
-by constantly simulating possible future events
+Going beyond isolated meaning and concepts extracted from the world,
+semantic building blocks may need to get recombined to enable
+mentally envisioning non-existing scenarios.
+As such, semantic knowledge would be a prerequisite for optimizing behavior
+by constantly simulating possible future scenarios
 \citep{boyer2008evolutionary, binder2011neurobiology}.
@@
 the internal construction and elaboration of necessary information
-that is not presented in the surrounding environment
+that is not presented in the surrounding sensory environment
 by recombining building blocks of
@@
 build on abstract world knowledge and logical associations
-for integrating the constituent elements in mental scene construction
+for integrating the constituent elements in rich and coherent mental scenes
 \citep{schacter2007remembering}.
@@
 present and foretelling the future.
-Mental scene construction has been proposed
+Further,
+mental scene construction has been proposed
 to imply a distinction between
@@
 there is also no reason to assume that predictions of various
-levels of complexity, abstraction, timescale and purpose
+levels of complexity, abstraction, timescale, and purpose
 rely on mechanisms that are qualitatively different. This concurs with
@@
 generated information for probabilistic mental models of action outcomes,
-MDPs render explicit the grounds on which the final action is
+MDPs render explicit the grounds on which an action is
 eventually chosen, namely, the estimated cumulative reward.
-\textbf{The learned
-estimations of value and action transitions drive
-stimulus processing and action choice in the present.}
-Additionally,
+In contrast to many versions of the semantic hypothesis,
 the MDPs naturally integrate the egocentric view
-(more related to current action, state and reward) and the
+(more related to current action, state, and reward) and the
 world view (more related to past and future actions, states, and rewards)
@@
 stochasticity of the parameter space explored
-by the Monte Carlo solvers achieves better fine-tuning of the
+by Monte Carlo approximation achieves better fine-tuning of the
 action policies and estimation of expected reward outcomes.
@@
 The DMN regions have been associated with processing the
-experienced or expected relevance of environment stimuli
+experienced or expected relevance of environment cues
 \citep{montague2006imaging}.
@@
 %
-In line with this, the DMN has been proposed to be specialzed in
+In line with this, the DMN has been proposed to be specialized in
 continuous processing of social information as a
@@
 and rats \citep{lu2012rat}, three species with
-supposedly less advanced social-cognitive capacities.
+social-cognitive capacities that are supposedly less advanced than in humans.
 Rather than functional specificity for processing social information,
-the present formal account can parsimoniously incorporate
+the present MDP account can parsimoniously incorporate
 the dominance of social content in
@@
 In Friston's free-energy principle (FEP) \citep{friston2010free,fristonAIorRL}, the brain is portrayed
-as biomechanical inference engine which must minimize the long-term average of surprise.
+as biomechanical inference engine
+dedicated to minimizing the long-term average of surprise.
 Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net (DBN)}.
-According to FEP's account, the goal of the brain is then to optimize over a generative model $G$ of sensations: to iteratively
-modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics, etc., so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles.
+According to the FEP account,
+the goal of the brain is then to optimize over a generative model $G$ of sensations: to iteratively
+modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics,
+so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles.
 These is called the \textit{generative} model.
@@
   Indeed in RL, surprise minimization is subsumed  by accurate prediction of
-  rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist undesired / harmful phase-transitions.
+  rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist undesired phase-transitions
   \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}. Such a theory cannot by itself,

### 2017-05-06 paper/paper.tex
saturday pass

@@
 neuronal computation
-during task-unconstrained behavior
+during free-ranging thought
 \citep{kenet2003spontaneously, fiser2004small}.
@@
 this network appears to
-simultaneously impact perception-action cycles and
+simultaneously modulate perception-action cycles and
 maintain probabilistic contemplations
@@
 The present work adopts the perspective of
-a human agent faced with the choice of the next actions
+a human \textit{agent} faced with the choice of the next actions
 and guided by outcomes
@@
 reinforcement learning (RL).
-An intelligent agent improves its interaction with the environment
+An intelligent agent improves the interaction with the environment
 by continuously updating its computation of value estimates and action predispositions
@@
 \citep{antrobus1966studies, pope1978regulation}.
-Without requiring explicit awareness by the agent,
+Without requiring explicit awareness,
 these ``offline'' processes may contribute
@@
 We formalize
-a \textit{policy matrix} to capture the space of possible actions that can be performed
+a \textit{policy matrix} to capture the space of possible actions that
+the agent can perform
 on the environment given the current state. A \textit{value function}
@@
 Switching between states reduces to a sequential processing model.
-Informed by outcomes of really performed action,
-the DMN dynamics could be constantly adapted in feedback loops
+Informed by outcomes of performed actions, neural computation reflected in
+DMN dynamics could be constantly adapted in feedback loops
 shaped by prediction error.
@@
 \citep{raichle2005intrinsic}.
-This idea is invigorated by
+This idea has been invigorated by
 empirical evidence from
@@
 \citep{goldman1987development}, which is generally considered to
-be a sign of evolutionary sophistication \citep{flechsig1920}.
+be a sign of evolutionary sophistication.
 %
@@
 continuously scan environmental input
-for biological significance assessment
+for biological relevance assessment
 \citep{amygdala_bzdok}.
 
-The putative role of the PMC in continuous information integration of
+The putative role of the PMC in continuous abstract integration of
 environmental relevance
@@
 Neuron spiking activity in the PMC allowed distinguishing
-whether a monkey would persue an exploratory or exploitative
+whether a monkey would pursue an exploratory or exploitative
 behavioral strategy during food foraging \citep{pearson2009neurons}.
@@
 Graded changes in firing rates of PMC neurons
-indicated changing choices in upcoming trials and their neural patterns were
+indicated changes in upcoming choice trials, while their neural patterns were
 distinct from neuronal spike firings that indicated choosing either option.
@@
 
-Moreover, the retrosplenial portion of the PMC can subserve
+Moreover, the retrosplenial portion of the PMC could support
 representation of action possibilities
-and evaluation of reward outcomes by integration with
-information from memory and different perspective frames.
+and evaluation of reward outcomes by integrating
+information from memory recall and different perspective frames.
 Regarding memory recall, retrosplenial damage have been
@@
 of various kinds of sensory information
-in rabbits and humans
+in animals and humans
 \citep{vann2009does}.
@@
 statistical assessment of the environment
-to predict delayed reward opportunities in the future.
+to map and predict delayed reward opportunities in the future.
 In doing so, the PMC may continuously adapt the organism to changes
@@
 reflected expectations about other peoples' actions and
-errors in these predictions.
+outcomes of these predictions.
 Neural activity in the dmPFC indeed explained the performance decline
@@
 construction of probabilistic internal information drawing from
-episodic memory recall, mental scene elaboration processes,
-and stored knowledge of the world.
+memory recall, mental scene elaboration processes,
+and stored knowledge of world regularities.
 %
@@
 whereas ventromedial PFC (vmPFC) activity
-processed especially value outcomes of performed actions in the environment
+processed especially value outcomes of actions performed on the environment
 \citep{nicolle2012agent}.
@@
 risk estimation of relevant environmental stimuli.
-This DMN node may
+The ventromedial prefrontal DMN may
 orchestrate adaptive behavior by bottom-up-driven
@@
 is traditionally believed to allow recalling past experience,
-there is now increasing evidence for a role
+there is now increasing evidence for an important role
 in constructing mental models in general
@@
 not only associated with an impairment in re-experiencing the past (i.e., amnesia),
-but also imagining one's own future and
-imagined experiences more broadly \citep{hassabis2007patients}.
+but also imagination dedicated to one's own future and
+imagination of experiences more broadly \citep{hassabis2007patients}.
 Mental scenes created by patients with HC lesion exposed a lack of
@@
 HC function extends beyond simple forms of
-encoding and reconstruction of memory and spatial information.
-Based on neuronal spike recordings of hippocampal neuronal populations,
+encoding and reconstruction of memory and space information.
+Based on spike recordings of hippocampal neuronal populations,
 complex spiking patterns can be followed across extended periods including
@@
 Specific spiking sequences, which were elicited by experimental task design,
-have been shown to be reenacted spontaneously during
+have been shown to be re-enacted spontaneously during
 quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
@@
 as well as directly before upcoming experimental trials \citep{diba2007forward}.
-Such spking patterns in hippocampal neurons during rest and sleep
+Similar spiking patterns in hippocampal neurons during rest and sleep
 have been proposed to be critical in communicating local information
@@
 Thus, the HC probably orchestrates re-experience of environmental aspects for
-consolidations based on reenactment and for integration into
+consolidations based on re-enactment and for integration into
 rich mental scene construction \citep{deuker2016event, bird2010establishing}.
-In this way, the HC may impact
+As such, the HC may impact
 ongoing perception of and action on the environment
@@
 \citep{doria2010}, while
-the right and left TPJ may become fully integrated into this major brain
+the right and left TPJs may become fully integrated into this major brain
 network only after birth.
@@
 detection of environmental changes across
-visual, auditory, or tactile stimulation in an fMRI experiment
+visual, auditory, or tactile stimulation
 \citep{downar2000multimodal}.
@@
 sensory environment and internally-oriented mind sets focussed
-on self-relevant mental scene construction.
+on mental scene construction.
 For instance, temporally induced RTPJ damage in humans diminished the
@@
 from the ‘‘internally directed’’ baseline processes
-to, instead, deal with (unexpected) environmental stimuli and events.
+to, instead, deal with unexpected environmental stimuli and events.
 % (TODO: slow endogeneous changes in state transitions by the policy and value matrices)
@@
 guide present action choice,
-weigh hypothetical possibilities, and forecast the future.
+weigh hypothetical possibilities, and forecast future events.
 %

### 2017-05-06 paper/paper.tex
integrated dominik's comments

--- 2017-05-03 paper/paper.tex
@@
   \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
-    the agent takes an action (by following some policy). This yields a reward and an observed
-    state transition of the environment. The process goes on until interrupted or a goal state is reached.}
+    the agent takes an action (by following some policy), receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
   \centering
@@
 RL is a problem-solving technique in which,
-through repeated interactions with an unknown \textit{environment},
+through repeated interactions with an\textit{environment},
 an agent learns to reach goals and optimize reward signals
@@
 an action.
-Action may be delayed to achieve substantial increases in expected reward
-that is expected to grow with time.
+% Action may be delayed to achieve substantial increases in expected reward
+% that is expected to grow with time.
 
@@
 \begin{definition*}
-Mathematically, an MDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
+Mathematically, an MDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p)$ where
 \begin{itemize}
 \item $\mathcal S$ is the set of states, such as $\mathcal S = \{\text{happy}, \text{sad}\}$.
-\item $p : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto p(s'|s,a)$,
-  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
-  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present.
 \item $\mathcal A$ is the set of actions, such as $\mathcal A = \{\text{read}, \text{run},
@@
    so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
-\item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
-  $\mu(s)$ is the probability that the environment starts off in state $s$.
+% \item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
+%   $\mu(s)$ is the probability that the environment starts off in state $s$.
+\item $p : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto p(s'|s,a)$,
+  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
+  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present state and action taken.
+
 \end{itemize}
@@
 \subsubsection{Accumulated rewards and policies}
-The behavior of the agent is governed by some kind of \textit{policy}.
-It maps states of the world to
-a set of candidate actions that can be performed given this state. Starting at time $t=0$,
-a policy $\pi$ generates a trajectory of action choices as follows:
+The behavior of the agent is governed by a \textit{policy}, which maps states of the world
+to probability distributions over actions. Starting at time $t=0$,
+followung a policy $\pi$ generates a trajectory of action choices as follows:
 \begin{eqnarray*}
@@
 \subsubsection{Value approximation and the policy matrix}
-%As already mentioned in the previous section, Q-learning optimizes over the class of
-% deterministic policies of the form \eqref{eq:qlearning}.
-At a time step $k-1$, the world is in a state $s \in \mathcal S$, and the agent greedily takes an
+As already mentioned in the previous section, Q-learning optimizes over the class of
+deterministic policies of the form \eqref{eq:qlearning}. State spaces may be extremely large,
+and tracking all possible states and actions may require excessive space.
+The need of maintaining an explicit table of
+states can be eliminated by instead using of an approximate $Q$-value function $\tilde{Q}(s,a|\theta)$
+by keeping track of an approximating parameter $\theta$ of much lower dimension than the number of states.
+At a time step, the world is in a state $s \in \mathcal S$, and the agent takes an
 action which it expects to be the most valuable on average, namely
@@
 \end{equation}
-by querying an an approximate version $\tilde{Q}(s,a|\theta)$ of the $Q$-value function, parameterized
-by the tensor $\theta$. This defines a mapping from states directly to actions.
+This defines a mapping from states directly to actions.
 % $a \mapsto \tilde{Q}(a,s|\theta)$ of the $Q$-value function ``frozen'' at state $s$.
@@
 $\phi(s,a)$ would represent a high-level representation of the state-action pairs
-$(s,a)$, as recently proposed \citep{songNIPS2016}.
+$(s,a)$, as was done proposed \citep{songNIPS2016}, or a eural network as recently proposed by in the
+breakthrough works
+\citep{mnih2015,silver2016mastering} for playing complex games (atari, Go, etc.) at superhuman levels.
 %  and Parr et al. 2003 (JMLR)].
-In the DMN, the dmPFC would implement such a greedy lookup
+In the DMN, the dmPFC would implement such a hard-max lookup
 over the action space.
@@
   parameters $\theta$ would correspond to synaptic weights and connection strengths between
-  brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the greedy policy \eqref{eq:policy}.
+  brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the hard-max policy \eqref{eq:policy}.
   The oracle Q-values $\tilde{Q}(s, a|\theta)$ tell the DMN the (expected) usefulness of taking an action a in state $s$.
@@
 
-  \textbf{XXX Elvis to Elvis: clarify the notion of greed. Add a paragraph on the exploitation vs exploration dilemma, and end with a conclusion saying the DMN probably does epsilon-greedy
-    or softmax.}
+  % \textbf{XXX Elvis to Elvis: clarify the notion of greed. Add a paragraph on the exploitation vs exploration dilemma, and end with a conclusion saying the DMN probably does epsilon-greedy
+  %   or softmax.}
 
@@
 \begin{equation}
-  \delta \theta_{k} \propto -\nabla_{\theta_{k}}\mathcal L(\theta_{k})
-  = -\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
-    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{aversion}}].
+  \delta= -\alpha_k \nabla_{\theta_{k}}\mathcal L(\theta_{k})
+  = -\alpha_k\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
+    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{aversion}}],
   \label{eq:oracle}
 \end{equation}
+where the positive constants $\alpha_1, \alpha_2,\ldots$ are learning rates.
 Thus, the next action is taken to drive reward prediction errors
-(or \textit{regret}) to
-percolate
-from lower to higher processing layers to modulate the choice of future actions .Over time (i.e., in the limit $k \rightarrow \infty$),
-the thus generated approximating sequence of
-$Q$-value functions provably
+(or \textit{regret}) to percolate from lower to higher processing layers to modulate the
+choice of future actions. It is a standard result that under special conditions on the learning rates $\alpha_k$
+--namely that the sum $\sum_{k=0}^\infty\alpha_k$
+diverges while $\sum_{k=0}^\infty\alpha_k^2$ converges, over time (i.e., in the limit $k \rightarrow \infty$)
+the thus generated approximating sequence of $Q$-value functions
 % \footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
-$\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$
+$$\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$$
 are attracted and absorbed by the optimal $Q$-value function $Q^*$ defined implicitly by the Bellman equation \eqref{eq:bellman}.
 
-One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
-learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
-cast in the general variational framework \eqref{eq:oracle}. For instance, TD corresponds to
-the above framework using a linear value approximator with feature encoding
-$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
-$\mathcal S \times \mathcal A$, and so
-$$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
-Hence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
-$$\tilde{Q}(s, a) \leftarrow \tilde{Q}(s, a) + \alpha(y-\tilde{Q}(s, a)),$$
-the well-known TD update rule \citep{sutton1998reinforcement}.
+% \paragraph{Link to classical reinforcement learning.}
+% One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
+% learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
+% cast in the general variational framework \eqref{eq:oracle}. For instance, TD corresponds to
+% the above framework using a linear value approximator with feature encoding
+% $\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
+% $\mathcal S \times \mathcal A$, and so
+% $$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
+% Hence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
+% $$\tilde{Q}(s, a) \leftarrow \tilde{Q}(s, a) + \alpha(y-\tilde{Q}(s, a)),$$
+% the well-known TD update rule \citep{sutton1998reinforcement}.
 

### 2017-05-06 paper/paper.tex
Major pass

@@
 and is characteristic of so-called \textit{model-free RL}. Model-free RL has had great success in real-world problems including robotics
-\citep{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering}, etc., just to name a few.
+\citep{ng2004,abbeel2004}, super-human play in complex games like Atari\citep{mnih2015} and Go\citep{silver2016mastering}, etc., just to name a few.
 We propose to use such a model as formal basis to model DMN function.
 
-Such a learning scheme can be plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
+Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
 a core property of human intelligence has been proposed to underly improvement
@@
   \text{laugh}, \text{sympathize}, \text{empathize}\}.$
-\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$ is the \textit{reward function},
-   so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
+\item $r : \mathcal S \times \mathcal A \times \mathcal S \rightarrow \mathbb R$ is the \textit{reward function},
+   so that $r(s, a, s')$ is the instant reward for taking action $a$ in state $s$ followed by a state-transition $s \rightarrow s'$.
 % \item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
@@
     \textbf{choose action: }a_0 &\sim \pi(a|s_0)\\
-    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0)\\
+    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0, s_1)\\
     \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
-    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1)\\
+    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1,s_2)\\
     \vdots\\
     \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
-    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t})\\
+    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t},s_{t+1})\\
     \vdots
@@
 a longer-term future. 
-Precisely, rewards that are not expected to come within $\tau := 1/(1 - \gamma)$ time steps from the present point are disregarded.
-This reduces the variance of expected rewards accumulated across
+Precisely, rewards that are not expected to come within\footnote{\textbf{N.B.:} $\frac{1}{1-\gamma} = 1 + \gamma + \gamma^2 + \gamma^3 + \ldots$}
+$\tau := 1/(1 - \gamma)$
+time steps from the present point are disregarded. This reduces the variance of expected rewards accumulated across
 considered action cascades by limiting the depth of the search tree.
@@
   brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the hard-max policy \eqref{eq:policy}.
-  The oracle Q-values $\tilde{Q}(s, a|\theta)$ tell the DMN the (expected) usefulness of taking an action a in state $s$.
-  The DMN, and in particular its dmPFC node, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes this tabulated values.
-This mapping from states to actions is conventionally called \textit{policy matrix}.
+  The approximating $Q$-value function $\tilde{Q}(s, a|\theta)$ would tell the DMN the (expected) usefulness of taking an action $a$ in state $s$.
+  The DMN, and in particular its dmPFC node, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes these approximate
+  $Q$-values. Such a mapping from states to actions is conventionally called \textit{policy matrix}\citep{mnih2015,silver2016mastering}.
   Learning consists in starting from a given table and
@@
 choice of future actions. It is a standard result that under special conditions on the learning rates $\alpha_k$
---namely that the sum $\sum_{k=0}^\infty\alpha_k$
-diverges while $\sum_{k=0}^\infty\alpha_k^2$ converges, over time (i.e., in the limit $k \rightarrow \infty$)
+(namely that the sum $\sum_{k=0}^\infty\alpha_k$ diverges while $\sum_{k=0}^\infty\alpha_k^2$ converges) over time (i.e., in the limit $k \rightarrow \infty$)
 the thus generated approximating sequence of $Q$-value functions
@@
   plasticity in the DMN arises naturally:
-  If an agent behaving optimally in a given environment moves
+  if an agent behaving optimally in a given environment moves
   to new, yet unexperienced environments, reward prediction errors will

### 2017-05-06 paper/paper.tex

--- 2017-05-06 paper/paper.tex
@@
 \begin{definition*}
-Mathematically, an MDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
+Mathematically, an MDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p)$ where
 \begin{itemize}
 \item $\mathcal S$ is the set of states, such as $\mathcal S = \{\text{happy}, \text{sad}\}$.
+\item $\mathcal A$ is the set of actions, such as $\mathcal A = \{\text{read}, \text{run},
+  \text{laugh}, \text{sympathize}, \text{empathize}\}.$
+\item $r : \mathcal S \times \mathcal A \times \mathcal S \rightarrow \mathbb R$ is the \textit{reward function},
+   so that $r(s, a, s')$ is the instant reward for taking action $a$ in state $s$ followed by a state-transition $s \rightarrow s'$.
+% \item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
+%   $\mu(s)$ is the probability that the environment starts off in state $s$.
 \item $p : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto p(s'|s,a)$,
   the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
-  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present.
-\item $\mathcal A$ is the set of actions, such as $\mathcal A = \{\text{read}, \text{run},
-  \text{laugh}, \text{sympathize}, \text{empathize}\}.$
-\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$ is the \textit{reward function},
-  so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
-\item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
-  $\mu(s)$ is the probability that the environment starts off in state $s$.
+  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present state and action taken.
+
 \end{itemize}
@@
   parameters $\theta$ would correspond to synaptic weights and connection strengths between
-  brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the greedy policy \eqref{eq:policy}.
-  The oracle Q-values $\tilde{Q}(s, a|\theta)$ tell the DMN the (expected) usefulness of taking an action a in state $s$.
-  The DMN, and in particular its dmPFC node, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes this tabulated values.
-This mapping from states to actions is conventionally called \textit{policy matrix}.
+  brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the hard-max policy \eqref{eq:policy}.
+  The approximating $Q$-value function $\tilde{Q}(s, a|\theta)$ would tell the DMN the (expected) usefulness of taking an action $a$ in state $s$.
+  The DMN, and in particular its dmPFC node, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes these approximate
+  $Q$-values. Such a mapping from states to actions is conventionally called \textit{policy matrix}\citep{mnih2015,silver2016mastering}.
   Learning consists in starting from a given table and
@@
 \begin{equation}
-  \delta \theta_{k} \propto -\nabla_{\theta_{k}}\mathcal L(\theta_{k})
-  = -\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
-    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{aversion}}].
+  \delta= -\alpha_k \nabla_{\theta_{k}}\mathcal L(\theta_{k})
+  = -\alpha_k\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
+    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{aversion}}],
   \label{eq:oracle}
 \end{equation}
+where the positive constants $\alpha_1, \alpha_2,\ldots$ are learning rates.
 Thus, the next action is taken to drive reward prediction errors
-(or \textit{regret}) to
-percolate
-from lower to higher processing layers to modulate the choice of future actions .Over time (i.e., in the limit $k \rightarrow \infty$),
-the thus generated approximating sequence of
-$Q$-value functions provably
+(or \textit{regret}) to percolate from lower to higher processing layers to modulate the
+choice of future actions. It is a standard result that under special conditions on the learning rates $\alpha_k$
+(namely that the sum $\sum_{k=0}^\infty\alpha_k$ diverges while $\sum_{k=0}^\infty\alpha_k^2$ converges) over time (i.e., in the limit $k \rightarrow \infty$)
+the thus generated approximating sequence of $Q$-value functions
 % \footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
-$\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$
+$$\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$$
 are attracted and absorbed by the optimal $Q$-value function $Q^*$ defined implicitly by the Bellman equation \eqref{eq:bellman}.
 
-One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
-learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
-cast in the general variational framework \eqref{eq:oracle}. For instance, TD corresponds to
-the above framework using a linear value approximator with feature encoding
-$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
-$\mathcal S \times \mathcal A$, and so
-$$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
-Hence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
-$$\tilde{Q}(s, a) \leftarrow \tilde{Q}(s, a) + \alpha(y-\tilde{Q}(s, a)),$$
-the well-known TD update rule \citep{sutton1998reinforcement}.
+% \paragraph{Link to classical reinforcement learning.}
+% One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
+% learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
+% cast in the general variational framework \eqref{eq:oracle}. For instance, TD corresponds to
+% the above framework using a linear value approximator with feature encoding
+% $\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
+% $\mathcal S \times \mathcal A$, and so
+% $$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
+% Hence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
+% $$\tilde{Q}(s, a) \leftarrow \tilde{Q}(s, a) + \alpha(y-\tilde{Q}(s, a)),$$
+% the well-known TD update rule \citep{sutton1998reinforcement}.
 

### 2017-05-06 paper/paper.tex
stuff

@@
   \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
-    the agent takes an action (by following some policy), receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
+    the agent takes an action (by following some policy). This yields a reward and an observed
+    state transition of the environment. The process goes on until interrupted or a goal state is reached.}
   \centering
@@
 RL is a problem-solving technique in which,
-through repeated interactions with an\textit{environment},
+through repeated interactions with an unknown \textit{environment},
 an agent learns to reach goals and optimize reward signals
@@
 an action.
-% Action may be delayed to achieve substantial increases in expected reward
-% that is expected to grow with time.
+Action may be delayed to achieve substantial increases in expected reward
+that is expected to grow with time.
 
@@
 and is characteristic of so-called \textit{model-free RL}. Model-free RL has had great success in real-world problems including robotics
-\citep{ng2004,abbeel2004}, super-human play in complex games like Atari\citep{mnih2015} and Go\citep{silver2016mastering}, etc., just to name a few.
+\citep{abbeel2004,andrewng}, computer game play, \citep{mnih2015,silver2016mastering}, etc., just to name a few.
 We propose to use such a model as formal basis to model DMN function.
 
-Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
+Such a learning scheme can be plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
 a core property of human intelligence has been proposed to underly improvement
@@
 \begin{definition*}
-Mathematically, an MDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p)$ where
+Mathematically, an MDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p, \mu)$ where
 \begin{itemize}
 \item $\mathcal S$ is the set of states, such as $\mathcal S = \{\text{happy}, \text{sad}\}$.
+\item $p : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto p(s'|s,a)$,
+  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
+  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present.
 \item $\mathcal A$ is the set of actions, such as $\mathcal A = \{\text{read}, \text{run},
   \text{laugh}, \text{sympathize}, \text{empathize}\}.$
-\item $r : \mathcal S \times \mathcal A \times \mathcal S \rightarrow \mathbb R$ is the \textit{reward function},
-   so that $r(s, a, s')$ is the instant reward for taking action $a$ in state $s$ followed by a state-transition $s \rightarrow s'$.
-% \item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
-%   $\mu(s)$ is the probability that the environment starts off in state $s$.
-\item $p : \mathcal S \times \mathcal A \times \mathcal S \rightarrow [0, 1],\; (s,a,s') \mapsto p(s'|s,a)$,
-  the probability of moving to state $s'$ if action $a$ is taken from state $s$. In addition, one requires that such
-  transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present state and action taken.
-
+\item $r : \mathcal S \times \mathcal A \rightarrow \mathbb R$ is the \textit{reward function},
+  so that $r(s, a)$ is the instant reward for taking action $a$ in state $s$.
+\item $\mu: \mathcal S \rightarrow [0, 1]$ is the prior probability on the states so that
+  $\mu(s)$ is the probability that the environment starts off in state $s$.
 \end{itemize}
@@
 \subsubsection{Accumulated rewards and policies}
-The behavior of the agent is governed by a \textit{policy}, which maps states of the world
-to probability distributions over actions. Starting at time $t=0$,
-followung a policy $\pi$ generates a trajectory of action choices as follows:
+The behavior of the agent is governed by some kind of \textit{policy}.
+It maps states of the world to
+a set of candidate actions that can be performed given this state. Starting at time $t=0$,
+a policy $\pi$ generates a trajectory of action choices as follows:
 \begin{eqnarray*}
@@
     \textbf{choose action: }a_0 &\sim \pi(a|s_0)\\
-    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0, s_1)\\
+    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0)\\
     \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
-    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1,s_2)\\
+    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1)\\
     \vdots\\
     \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
-    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t},s_{t+1})\\
+    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t})\\
     \vdots
@@
 a longer-term future. 
-Precisely, rewards that are not expected to come within\footnote{\textbf{N.B.:} $\frac{1}{1-\gamma} = 1 + \gamma + \gamma^2 + \gamma^3 + \ldots$}
-$\tau := 1/(1 - \gamma)$
-time steps from the present point are disregarded. This reduces the variance of expected rewards accumulated across
+Precisely, rewards that are not expected to come within $\tau := 1/(1 - \gamma)$ time steps from the present point are disregarded.
+This reduces the variance of expected rewards accumulated across
 considered action cascades by limiting the depth of the search tree.
@@
 \subsubsection{Value approximation and the policy matrix}
-As already mentioned in the previous section, Q-learning optimizes over the class of
-deterministic policies of the form \eqref{eq:qlearning}. State spaces may be extremely large,
-and tracking all possible states and actions may require excessive space.
-The need of maintaining an explicit table of
-states can be eliminated by instead using of an approximate $Q$-value function $\tilde{Q}(s,a|\theta)$
-by keeping track of an approximating parameter $\theta$ of much lower dimension than the number of states.
-At a time step, the world is in a state $s \in \mathcal S$, and the agent takes an
+%As already mentioned in the previous section, Q-learning optimizes over the class of
+% deterministic policies of the form \eqref{eq:qlearning}.
+At a time step $k-1$, the world is in a state $s \in \mathcal S$, and the agent greedily takes an
 action which it expects to be the most valuable on average, namely
@@
 \end{equation}
-This defines a mapping from states directly to actions.
+by querying an an approximate version $\tilde{Q}(s,a|\theta)$ of the $Q$-value function, parameterized
+by the tensor $\theta$. This defines a mapping from states directly to actions.
 % $a \mapsto \tilde{Q}(a,s|\theta)$ of the $Q$-value function ``frozen'' at state $s$.
@@
 $\phi(s,a)$ would represent a high-level representation of the state-action pairs
-$(s,a)$, as was done proposed \citep{songNIPS2016}, or a eural network as recently proposed by in the
-breakthrough works
-\citep{mnih2015,silver2016mastering} for playing complex games (atari, Go, etc.) at superhuman levels.
+$(s,a)$, as recently proposed \citep{songNIPS2016}.
 %  and Parr et al. 2003 (JMLR)].
-In the DMN, the dmPFC would implement such a hard-max lookup
+In the DMN, the dmPFC would implement such a greedy lookup
 over the action space.
@@
   parameters $\theta$ would correspond to synaptic weights and connection strengths between
-  brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the hard-max policy \eqref{eq:policy}.
-  The approximating $Q$-value function $\tilde{Q}(s, a|\theta)$ would tell the DMN the (expected) usefulness of taking an action $a$ in state $s$.
-  The DMN, and in particular its dmPFC node, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes these approximate
-  $Q$-values. Such a mapping from states to actions is conventionally called \textit{policy matrix}\citep{mnih2015,silver2016mastering}.
+  brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the greedy policy \eqref{eq:policy}.
+  The oracle Q-values $\tilde{Q}(s, a|\theta)$ tell the DMN the (expected) usefulness of taking an action a in state $s$.
+  The DMN, and in particular its dmPFC node, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes this tabulated values.
+This mapping from states to actions is conventionally called \textit{policy matrix}.
   Learning consists in starting from a given table and
@@
 
-  % \textbf{XXX Elvis to Elvis: clarify the notion of greed. Add a paragraph on the exploitation vs exploration dilemma, and end with a conclusion saying the DMN probably does epsilon-greedy
-  %   or softmax.}
+  \textbf{XXX Elvis to Elvis: clarify the notion of greed. Add a paragraph on the exploitation vs exploration dilemma, and end with a conclusion saying the DMN probably does epsilon-greedy
+    or softmax.}
 
@@
 \begin{equation}
-  \delta= -\alpha_k \nabla_{\theta_{k}}\mathcal L(\theta_{k})
-  = -\alpha_k\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
-    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{aversion}}],
+  \delta \theta_{k} \propto -\nabla_{\theta_{k}}\mathcal L(\theta_{k})
+  = -\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
+    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{aversion}}].
   \label{eq:oracle}
 \end{equation}
-where the positive constants $\alpha_1, \alpha_2,\ldots$ are learning rates.
 Thus, the next action is taken to drive reward prediction errors
-(or \textit{regret}) to percolate from lower to higher processing layers to modulate the
-choice of future actions. It is a standard result that under special conditions on the learning rates $\alpha_k$
-(namely that the sum $\sum_{k=0}^\infty\alpha_k$ diverges while $\sum_{k=0}^\infty\alpha_k^2$ converges) over time (i.e., in the limit $k \rightarrow \infty$)
-the thus generated approximating sequence of $Q$-value functions
+(or \textit{regret}) to
+percolate
+from lower to higher processing layers to modulate the choice of future actions .Over time (i.e., in the limit $k \rightarrow \infty$),
+the thus generated approximating sequence of
+$Q$-value functions provably
 % \footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
-$$\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$$
+$\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$
 are attracted and absorbed by the optimal $Q$-value function $Q^*$ defined implicitly by the Bellman equation \eqref{eq:bellman}.
 
-% \paragraph{Link to classical reinforcement learning.}
-% One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
-% learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
-% cast in the general variational framework \eqref{eq:oracle}. For instance, TD corresponds to
-% the above framework using a linear value approximator with feature encoding
-% $\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
-% $\mathcal S \times \mathcal A$, and so
-% $$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
-% Hence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
-% $$\tilde{Q}(s, a) \leftarrow \tilde{Q}(s, a) + \alpha(y-\tilde{Q}(s, a)),$$
-% the well-known TD update rule \citep{sutton1998reinforcement}.
+One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
+learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
+cast in the general variational framework \eqref{eq:oracle}. For instance, TD corresponds to
+the above framework using a linear value approximator with feature encoding
+$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
+$\mathcal S \times \mathcal A$, and so
+$$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
+Hence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
+$$\tilde{Q}(s, a) \leftarrow \tilde{Q}(s, a) + \alpha(y-\tilde{Q}(s, a)),$$
+the well-known TD update rule \citep{sutton1998reinforcement}.
 
@@
   plasticity in the DMN arises naturally:
-  if an agent behaving optimally in a given environment moves
+  If an agent behaving optimally in a given environment moves
   to new, yet unexperienced environments, reward prediction errors will
@@
 \subsection{The free-energy principle and active inference}
-% \subsubsection{The theory}
+\textbf{XXX Passage: Explain FEP and active inference}
 In Friston's free-energy principle (FEP) \citep{friston2010free,fristonAIorRL}, the brain is portrayed
@@
 dedicated to minimizing the long-term average of surprise.
-Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net (DBN)}.
+Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net}.
 According to the FEP account,
@@
 modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics,
-so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles.
-These is called the \textit{generative} model.
-FEP also postulates a dual model called the \textit{recognition} model,
+so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles (i.e., the \textit{generative} model).
+The FEP also postulates a dual model called the \textit{recognition} model,
 which works in tandem with the generative model $p_R(\z|\x)$, to accomplish approximate inference. The recognition model dreams imaginary worlds $\z$,
@@
 % \end{figure}
+\textbf{XXX Passage: Explain the relation of FEP and active inference to our MDPs}
 The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
@@
 From the view of a Markovian sequential decision process,
-human behavior unfolds by integrating reward outcomes from stored past events
+human behavior unfolds by integrating action outcomes from stored past events
 and extrapolation to
@@
 actions and the estimation of action outcomes may explain why
-very accurate memories have been disfavored in human evolution
+highly accurate memories have been disfavored in human evolution
 and why human creativity might be adaptive.
@@
 on DMN biology is hence practically computable and
-yields falsifiable neuroscientific hypotheses.
+yields falsifiable neurobiological hypotheses.
 %
@@
 describe and interpret experimental findings in neuroscience studies
-on the DMN situated at top of the brain network hiearchy.
+on the DMN situated at top of the brain network hierarchy.
 %

### 2017-05-06 paper/paper.tex
rephrasing stuff

--- 2017-05-06 paper/paper.tex
@@
 risk estimation of relevant environmental stimuli.
-The ventromedial prefrontal DMN may
+The ventromedial prefrontal DMN (\textbf{XXX Elvis: you mean ``... ventromedial prefrontal part of the DMN'' ?}) may
 orchestrate adaptive behavior by bottom-up-driven
@@
    &= r(s,a) + \gamma\mathbb E_{s' \sim p(s'|s,a)} [\max_{a' \in \mathcal A}Q(s', a')]\\
-   &= \text{instantaneous reward} + \text{expected reward for acting greedily thereafter}
+   &= \text{instantaneous reward} + \text{expected maximum reward thereafter}
 %    \; \forall (s,a) \in \mathcal S \times \mathcal A

### 2017-05-06 paper/paper.tex
more stuff

--- 2017-05-06 paper/paper.tex
@@
 may be solved by a stochastic approximation
-based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
+based on a brain implementation of Markov Chain MCMC (MCMC) sampling
 \citep{tenenbaum2011grow}.
@@
   \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
-    the agent takes an action (by following some policy), receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
+    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
   \centering
@@
 $s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
-$r = r(s, a)$ collected by the agent (Fig. \ref{fig:rl}).
+$r = r(s, a)$ collected by the agent.
 If the collected reward outcome yields a negative value it can be
@@
 of as satisfaction \textemdash or aversion \textemdash accompanying the execution of
-an action.
+a particular action.
 % Action may be delayed to achieve substantial increases in expected reward
@@
 can be parsimoniously explained as probabilistic mental
-scene simulations coupled with prediction error minimization
+scene of experience simulations coupled with prediction error minimization
 to calibrate action trajectories for
@@
 \subsection{Markov decision processes}
-We argue that the brain in general, and the DMN in particular,
+Model-free RL has had great success in real-world problems including robotics
+\citep{ng2004,abbeel2004},
+super-human performance in complex games like Atari\citep{mnih2015} and Go\citep{silver2016mastering}, etc., just to name a few.
+We emphasize that the brain in general, and the DMN in particular,
 is a physical system governed by the laws of
@@
  and maintains a probability distribution over the set of possible
-states, based on a set of observations and observation probabilities.
+states.
+% , based on a set of observations and observation probabilities.
 This is a minimal set of assumptions that can be made about an environment,
-and is characteristic of so-called \textit{model-free RL}. Model-free RL has had great success in real-world problems including robotics
-\citep{ng2004,abbeel2004}, super-human play in complex games like Atari\citep{mnih2015} and Go\citep{silver2016mastering}, etc., just to name a few.
-We propose to use such a model as formal basis to model DMN function.
+and is characteristic of so-called \textit{model-free RL}.
 
 Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
-a core property of human intelligence has been proposed to underly improvement
+a core property of human intelligence has been proposed to underlie improvement
 of expected utility outcomes as a strategy for action choice in uncertain
@@
 and action policies by controlling
-synaptic plasticity in the reward system, including the NAc.
+synaptic plasticity in the reward-processing circuitry, including the NAc.
 In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of the world.
@@
 \begin{itemize}
-\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
+\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a location
+offering relevant resources.
 \item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
@@
 The behavior of the agent is governed by a \textit{policy}, which maps states of the world
-to probability distributions over actions. Starting at time $t=0$,
-followung a policy $\pi$ generates a trajectory of action choices as follows:
+to probability distributions over actions.
+Starting at time $t=0$,
+following a policy $\pi$ generates a trajectory of action choices as follows:
 \begin{eqnarray*}
@@
 concerned with immediate rewards.
-This is however not compatible with coordinated planning if long-term goal that is
+This is however not compatible with coordinated planning of long-term goal that is
 potentially subserved by neural activity in the DMN.
@@
 
-\subsection{The nuts and bolts of reinforcement learning in the DMN}
+\subsection{The components of reinforcement learning in the DMN}
 Given only the limited information available from an MDP, at a state $s$ the average
@@
 is broken down into simpler sub-problems at different time points.
-Exploitation of hierarchical structure in action considerations
+Indeed,
+exploitation of hierarchical structure in action considerations
 has previously been related to the medial prefrontal part of the DMN
@@
 Efficient learning of the entire set of model parameters can effectively be achieved
-via stochastic \textit{gradient descent}, a first-order\footnote{ i.e using only differential information up to order 1.} algorithm for finding
-local minima of differentiable functions.
+via stochastic \textit{gradient descent}, a common solver algorithm for finding
+local minima based on the first derivate of the optimization objective.
 Stochastic here means that the true gradient is estimated from batches of training examples samples from the replay memory, the HC
@@
 Thus, the next action is taken to drive reward prediction errors
-(or \textit{regret}) to percolate from lower to higher processing layers to modulate the
+to percolate from lower to higher processing layers to modulate the
 choice of future actions. It is a standard result that under special conditions on the learning rates $\alpha_k$
@@
 
-\subsubsection{Does the hippocampus subserve Monte-Carlo sampling?}
-In RL, Monte-Carlo simulation is a common means to update the agent's belief state
+\subsubsection{Does the hippocampus subserve MCMC sampling?}
+In RL, MCMC simulation is a common means to update the agent's belief state
 \citep{silver2010monte}.
-Monte-Carlo simulation provides a simple method for evaluating the value of a state.
+MCMC simulation provides a simple method for evaluating the value of a state.
 They provide an effective mechanism both for tree search (of the considered
@@
 highest expected cumulative rewards.
-In MDPs, Monte-Carlo simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
+In MDPs, MCMC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
 \textbf{XXX Elvis: refs needed.}.
-Much recent work points to Monte Carlo or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
+Much recent work points to MCMC or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
 \textbf{XXX Elvis: refs needed in last sentence.
@@
   the current sensory environment
-  \citep{kenet2003spontaneously, fiser2004small, bzdok2013segregation}
   and this network was proposed to be situated
@@
   Its putative involvement in thinking about the past,
-  the future, and hypothetical experiences
+  hypothetical experiences, and the future
   appears to tie in with the implicit computation of
@@
   This will lead to adaptation of the policy until the system converges to a
-  new steady-state of optimal action decisions in volatile environments.
+  new steady-state of optimal action decisions in the volatile environment.
 
@@
 This is because the internal generative
-model only requires updates after descrepancies have occurred between
+model only requires updates after discrepancies have occurred between
 environmental reality and its internally reinstantiated representation.
@@
 various stages of sensory input processing.
-This top-down generative design is replaced in our MDP view of the DMN
+This generative top-down design is replaced in our MDP view of the DMN
 by a sequential decision-making framework.
@@
 is re-expressed in our account in form of
-nested prediction of probable upcoming actions, states, and rewards.
+nested prediction of probable upcoming actions, states, and outcomes.
 While both accounts capture the consequences of action,
@@
 the present MDP account of DMN function may serve as a concrete implementation of
-the long-standing predictive coding account from cognitive neuroscience.
+the predictive coding account from cognitive neuroscience.
 MDPs have the advantage of exposing an explicit
@@
 episodic memory, language, problem solving,
-planning, estimating other people's thoughts, and spatial navigation
+planning, estimating other others' thoughts, and spatial navigation
 represent neural processes that are likely to
@@
 followed by a return to perception of the present.''.
-The semantic hypothesis is for instance supported by evidence in animals that
-could learn a \textit{cognitive map of the environment}
-and exploit it later for other means \citep{tolman1948cognitive}.
+% The semantic hypothesis is for instance supported by evidence in animals that
+% could learn a \textit{cognitive map of the environment}
+% and exploit it later for other means \citep{tolman1948cognitive}.
 
@@
 stochasticity of the parameter space explored
-by Monte Carlo approximation achieves better fine-tuning of the
+by MCMC approximation achieves better fine-tuning of the
 action policies and estimation of expected reward outcomes.
@@
 as the more advantageous adaptation.
-In sum, episodic scene construction according to the semantic
+In sum, mental scene construction according to the semantic
 account is lacking an explicit time and incentive model,
-both of which are integral parts of MDP frameworks.
+both of which are integral parts of the MDP frameworks.
 
@@
 encountering a self-relevant environmental stimulus.
-In the present MDP account,
+However,
+in the present MDP account,
 promising action trajectories
@@
 social systems,
-which allows solving ecological problems by social relationships
+which allows solving ecological problems by means of social relationships
 \citep{whiten1988machiavellian}.
@@
 %
-In line with this, the DMN has been proposed to be specialized in
+In line with this, the DMN was argued to be specialized in
 continuous processing of social information as a
@@
 
-\textbf{XXX: integrate aspects from section "ThE MARkOv DECISION PROBLEM" in Dayan and Daw}
 \textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
-\textbf{XXX: integrate aspects from Sutton/Barto book chapter}
 

### 2017-05-06 paper/paper.tex
more more stuff
major pass: link between fep and rl

--- 2017-05-06 paper/paper.tex
@@
 \begin{equation}
-  \pi^{\text{greedy}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a|\theta),
+  \pi^{\text{hard-max}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a|\theta),
   \label{eq:policy}
@@
 \textbf{XXX Passage: Explain FEP and active inference}
-In Friston's free-energy principle (FEP) \citep{friston2010free,fristonAIorRL}, the brain is portrayed
+In Friston's free-energy principle (FEP) \citep{friston2010free} and active inference \citep{fristonAIorRL} theories, the brain is portrayed
 as biomechanical inference engine
-dedicated to minimizing the long-term average of surprise.
+dedicated to minimizing the long-term average of surprise: the log-likelihood of the observe sensory input.
+It does this by actively generating hypothetical explanations or predictions of its sensory input.
 Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net}.
@@
 % \end{figure}
-\textbf{XXX Passage: Explain the relation of FEP and active inference to our MDPs}
+
 The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
@@
 
+\subsubsection{RL reinterpretation of FEP and active inference}  FEP and active inference can be reinterpreted in our model-free RL framework by postulating that the reward signal maximized by the DMN is the negative of surprisal, i.e the log-likehood of sensory the current priors the agent has about the world.
+This is worth writing down explicitly,
+
+\begin{equation}
+  r^{\text{FEP}} = \underbrace{\log (p_G)}_{-\text{FEP generative surprisal}}.
+    \end{equation}
 % \paragraph{Comparison to our proposed theory.}
-On the surface, a common point between the FEP and our proposed RL-based framework
-  is placing the minimization of a surprise signal at the core of brain function.
-  Indeed in RL, surprise minimization is subsumed  by accurate prediction of
-  rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist undesired phase-transitions
-  \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}. Such a theory cannot by itself,
-  explain the emergence of strategic behavior inherent in humans (e.g.,  cite dark-room experiment).
+% On the surface, a common point between the FEP and our proposed RL-based framework
+%   is placing the minimization of a surprise signal at the core of brain function.
+%   Indeed in RL, surprise minimization is subsumed  by accurate prediction of
+%   rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist undesired phase-transitions
+%   \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}.
+    Such a surprisal-based reinforcement learning scheme has been advocated for in \citep{schmidhuber2010,mohamed2015} under the equivalent framework of information compression.
+    However, we note that though relevant, minimization of surprisal alone is insufficient to explain intelligent behavior observed in humans and other ``intelligent'' animals.
+The dark-room experiment, for example, provides a formidable argument against in favor of the insufficiency of the FEP and active inference theories to explain intelligent behavior in humans: when a human is put in a dark room, they don't just sit down in a corner forever, what would be the FEP-optimal thing to do; they actively search for the light switch, and start exploring the depths of the room, possibly looking for food or the way out!
 

### 2017-05-06 paper/paper.tex

--- 2017-05-06 paper/paper.tex
@@
 local minima based on the first derivate of the optimization objective.
-Stochastic here means that the true gradient is estimated from batches of training examples samples from the replay memory, the HC
+Stochastic here means that the true gradient is estimated from batches of training examples samples (in our case blocks of experience from the replay memory):
 

### 2017-05-06 paper/paper.tex
defining 'partially observable'

@@
 $s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
-$r = r(s, a)$ collected by the agent.
+$r = r(s, a,s')$ collected by the agent.
 If the collected reward outcome yields a negative value it can be
@@
 The environment is generally taken as \textit{stochastic},
-that is, changing in random ways. In addition,
-the environment is only \textit{partially observed}
-in the sense that \textbf{XXX [Please Define without using the same words]}
-only part of the current state is observed by
-the agent \citep{starkweather2017dopamine}.
+that is, changing in random ways. In addition, the environment is only
+\textit{partially observable} in the sense only limited aspects of the environment's
+state are accessible to --or can be sensed by-- the agent's sensors.
+\citep{starkweather2017dopamine}.
 We assume that volatility of the environment
@@
 
-  % \textbf{XXX Elvis to Elvis: clarify the notion of greed. Add a paragraph on the exploitation vs exploration dilemma, and end with a conclusion saying the DMN probably does epsilon-greedy
-  %   or softmax.}
-
 % \subsubsection{Bounded rationality}

### 2017-05-07 paper/paper.tex
wuhu
rmed circularity

@@
 that is, changing in random ways. In addition, the environment is only
-\textit{partially observable} in the sense only limited aspects of the environment's
-state are accessible to --or can be sensed by-- the agent's sensors.
+\textit{partially observable} in the sense that only limited aspects of the environment's
+state are accessible to the agent's sensors.
 \citep{starkweather2017dopamine}.
@@
 In MDPs, MCMC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
-\textbf{XXX Elvis: refs needed.}.
+\textbf{XXX: refs needed.}.
 Much recent work points to MCMC or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
-\textbf{XXX Elvis: refs needed in last sentence.
+\textbf{XXX: refs needed in last sentence.
 Plus, should be really bring up Bayesian?} For sampling-based methods it is still scarce
-beyond the description of receptive fields. \textbf{XXX Elvis: rephrase last sentence}
+beyond the description of receptive fields. \textbf{XXX: rephrase last sentence}
 

### 2017-05-09 paper/paper.tex
Multiple edits, especially mention of the synaptic epigenesis with link to policy matrix, and, of the phylogeny of connectivity with link to social interaction in human

@@
     \textbf{6} University Paris Diderot, Sorbonne Paris Cit\'e, Paris, France\\
-    \textbf{7} Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
+    \textbf{7} Centre de Bioinformatique, Biostatistique et Biologie Int\'egrative, Paris, France
 }
@@
 %
-A process model for the neural optimization of complex behavior in the DMN
+Moreover, this process model for the neural optimization of complex behavior in the DMN
 offers parsimonious explanations for
-recent experimental findings in animals and humans.
+recent experimental findings\suggestremove{ in animals and humans}.
 \end{abstract}
@@
 reinforcement learning
+\textbf{XXX: Maybe the keywords "systems biology" and "cognitive science" are too broad.
+I would also remove "reinforcement learning", which is already in the title.
+The keywords are the occasion to enhance reachability through web and bibliography search.
+What about "process model", "Markov decision processes", or "Bellman equation"?}
 % autonomous learning systems
@@
 DMN dynamics could be constantly adapted in feedback loops
-shaped by prediction error.
+shaped by prediction error. \textbf{XXX: "could be constantly adapted in feedback loops shaped by prediction error" => "are constantly shaped by prediction error through feedback loops"}
 Such an RL account of DMN function can naturally embed human behavior
@@
 \citep{ng2004,abbeel2004},
-super-human performance in complex games like Atari\citep{mnih2015} and Go\citep{silver2016mastering}, etc., just to name a few.
+super-human performance in complex video games \citep{mnih2015} and Go\citep{silver2016mastering}, etc., just to name a few.
 We emphasize that the brain in general, and the DMN in particular,
@@
   ...
+  \textbf{XXX Guillaume: I agree that this would be BIG!}
 \end{mdframed}
@@
   which take the agent to different table entries.
+\suggestadd{An additional layer of learning concerns the addition of new entries in the table (i.e.
+state and action).
+The framing of the policy matrix is not described here but could be supported by the formation of new attractor in the neural dynamics through synaptic epigenesis \citep{gisiger_acquisition_2005}.}
 
@@
 social-cognitive capacities that are supposedly less advanced than in humans.
+
+\suggestadd{The principal connectivity gradient in the cortex appears to be greatly expanded in humans compare to macaques, suggesting a phylogenetically conserved axis of cortical expansion with the DMN emerging at the extreme-end in humans \citep{margulies_situating_2016}.
+Neurocomputational models of dyadic whole brain demonstrated how the human connectivity topology, on top of facilitating processing at the intra-individual level, can explain our propensity to coordinate through sensorimotor loops with others at the inter-individual level \citep{dumas_anatomical_2012}.}
+
 Rather than functional specificity for processing social information,
@@
 \section{Conclusion}
-What brain function could be important enough
+Which brain function could be important enough
 for the existence and survival of the human species

### 2017-05-14 paper/paper.tex
NC of Bellman equation

@@
 
-\begin{mdframed}
-  \textbf{ Neural correlates of the Bellman equation in the DMN:}
-  ...
-  \textbf{XXX Guillaume: I agree that this would be BIG!}
-\end{mdframed}
+% neuroscience of Bellman equation
+%\textbf{Neural correlates of the Bellman equation in the DMN:}
+In neuroscience,
+processing of ``goal-tree sequences''
+has been linked to specific neural activity in the
+dorsal prefrontal cortex (BA9) based on human brain-imaging experiments
+\citep{koechlin1999role, koechlin2000dissociating}.
+Sub-goal exploration may require multi-task switching between cognitive processes
+as later parts of a solution
+frequently depend on respective earlier steps of a given solution path, which
+probably necessitates storage of expected intermediate outcomes.
+``Cognitive branching'' operations for nested solution processing
+are thus likely to
+entail secondary reallocation of attention and working-memory resources.
+The various advanced human mental abilities subserved by the DMN, such as
+planning and abstract reasoning, can be viewed to involve some form of branching
+to enable higher-order executive control.
+Further neuroimaging experiments \citep{braver2002role} corroborated
+the prefrontal DMN to subserve
+``processes related to the management and monitoring of sub-goals while maintaining information in working memory''.
+%
+Moreover,
+neurological patients with lesions in this DMN node were reported
+to be impaired in aspects of realizing ``multiple sub-goal scheduling''
+\citep{burgess2000cognitive}.
 

### 2017-05-14 paper/paper.tex
add awesome DMN prediction figure

@@
 %
+\begin{wrapfigure}{r}{.50\linewidth}
+  \includegraphics[width=1.\linewidth]{fig_dmn.png}
+  \centering
+  \caption{\textbf{Morphological coupling between reward system and
+  default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11+/-0.02$) and right NAc volume ($R^2=0.14+/-0.02$) could be predicted from volume in the DMN nodes.
+These out-of-sample generalization performances were obtained from support vector regression applied to normalized node volumes in the DMN in a 10-fold cross-validation framework.
+Consistent for the left and right reward system, NAc volume in a given subject has an especially close relation to vmPFC and HC.
+  Code and data for reproduction and visualization:
+  % \url{www.github.com/banilo/darkcontrol_paper}.
+  }
+  \label{fig:dmn}
+\end{wrapfigure}
 Such highly abstract neural computations necessarily rely on the
@@
 subjective value evaluation and
-risk estimation of relevant environmental stimuli.
+risk estimation of relevant environmental stimuli
+(Fig. \ref{fig:dmn}).
 The ventromedial prefrontal DMN may

### 2017-05-15 paper/paper.tex
update Bellman/Bio

@@
 %
-\begin{wrapfigure}{r}{.50\linewidth}
+\begin{wrapfigure}{r}{.5 0\linewidth}
   \includegraphics[width=1.\linewidth]{fig_dmn.png}
@@
 It is probably central for
-action initiation during goal-dircted psychological tasks and for
+action initiation during goal-directed psychological tasks and for
 sensorimotor behavior by integrating multi-sensory attention
@@
 %\textbf{Neural correlates of the Bellman equation in the DMN:}
-In neuroscience,
-processing of ``goal-tree sequences''
-has been linked to specific neural activity in the
-dorsal prefrontal cortex (BA9) based on human brain-imaging experiments
+Relating decomposition of consecutive action choices by the Bellman equation
+to neuroscience,
+specific neural activity in the dorsal prefrontal cortex (BA9)
+was linked to processing ``goal-tree sequences''
+in human neuroimaging experiments
 \citep{koechlin1999role, koechlin2000dissociating}.
-Sub-goal exploration may require multi-task switching between cognitive processes
+Sub-goal exploration may require multi-task switching
+between cognitive processes
 as later parts of a solution
-frequently depend on respective earlier steps of a given solution path, which
-probably necessitates storage of expected intermediate outcomes.
-``Cognitive branching'' operations for nested solution processing
-are thus likely to
+frequently depend on respective earlier steps in a given solution path, which
+necessitates storage of expected intermediate outcomes.
+As such,
+``cognitive branching'' operations for nested processing of behavioral strategies
+are likely to
 entail secondary reallocation of attention and working-memory resources.
-The various advanced human mental abilities subserved by the DMN, such as
-planning and abstract reasoning, can be viewed to involve some form of branching
-to enable higher-order executive control.
 Further neuroimaging experiments \citep{braver2002role} corroborated
@@
 \citep{burgess2000cognitive}.
+Hence,
+the various advanced human mental abilities subserved by the DMN, such as
+planning and abstract reasoning, can be viewed to involve some form of
+action-decision branching to enable higher-order executive control.
 

### 2017-05-18 paper/paper.tex
big pokemon

@@
 \textbf{\\keywords}: systems biology, mind-wandering, cognitive science,
-artificial intelligence,
-reinforcement learning
+artificial intelligence, Markov decision process
 \textbf{XXX: Maybe the keywords "systems biology" and "cognitive science" are too broad.
@@
 %
-\begin{wrapfigure}{r}{.5 0\linewidth}
+\begin{wrapfigure}{r}{.4\linewidth}
   \includegraphics[width=1.\linewidth]{fig_dmn.png}
@@
 
-\begin{wrapfigure}{r}{.4\linewidth}
-  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
-    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
-  \centering
-  \includegraphics[width=1.\linewidth]{rl.pdf}
-  \label{fig:rl}
-\end{wrapfigure}
-
 RL is a problem-solving technique in which,
@@
 
+\begin{wrapfigure}{r}{.4\linewidth}
+  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
+    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
+  \centering
+  \includegraphics[width=1.\linewidth]{rl.pdf}
+  \label{fig:rl}
+\end{wrapfigure}
 The environment is generally taken as \textit{stochastic},
@@
 
+\begin{mdframed}
 % neuroscience of Bellman equation
-%\textbf{Neural correlates of the Bellman equation in the DMN:}
+\textbf{Neural correlates of the Bellman equation in the DMN:}
 Relating decomposition of consecutive action choices by the Bellman equation
@@
 action-decision branching to enable higher-order executive control.
+\end{mdframed}
 
@@
   \subsubsection{Self-training and the loss function}
+  \label{sec:self}
 Successful learning in brains and computer algorithms is not possible without a
@@
 In MDPs, MCMC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
-\textbf{XXX: refs needed.}.
+\textbf{XXX elvis: refs needed.}.
 Much recent work points to MCMC or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
-\textbf{XXX: refs needed in last sentence.
+\textbf{XXX elvis: refs needed in last sentence.
 Plus, should be really bring up Bayesian?} For sampling-based methods it is still scarce
-beyond the description of receptive fields. \textbf{XXX: rephrase last sentence}
+beyond the description of receptive fields. \textbf{XXX elvis: rephrase last sentence}
 
@@
 \subsection{The free-energy principle and active inference}
-\textbf{XXX Passage: Explain FEP and active inference}
-In Friston's free-energy principle (FEP) \citep{friston2010free} and active inference \citep{fristonAIorRL} theories, the brain is portrayed
+In Friston's free-energy principle (FEP) \citep{friston2010free} and active inference  (i.e action under the FEP) \citep{fristonAIorRL} theories, the brain is portrayed
 as biomechanical inference engine
-dedicated to minimizing the long-term average of surprise: the log-likelihood of the observe sensory input.
+dedicated to minimizing the long-term average of surprise: the log-likelihood of the observed sensory input, or more precisely, an upper bound thereof.
 It does this by actively generating hypothetical explanations or predictions of its sensory input.
@@
 
+\subsubsection{Criticisms of the FEP}
+  Despite it's popularity, a number of critics about the FEP have surfaced over the years. We outline the main ones here.
+
+\paragraph{The wake-sleep algorithm is not neurobiologically plausible.}
 The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
@@
 for minimizing free-energy, VAEs can be efficiently trained via back-propagation of prediction errors.
+
+\paragraph{The ``dark room'' experiment.}
+Since the FEP postulates that all ongoing intelligent behavior in an organism is geared towards reaching the grand goal of minimizing surprisal, an organism which enters a \textit{dark room}\footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.} would strive to remain there as its sensory input would be perfectly predictable in such a state \citep{darkroom2012}.
+\begin{wrapfigure}{r}{.45\linewidth}
+  \centering
+  \includegraphics[width=1.\linewidth]{darkroom.jpg}
+  \caption{\textbf{The \textit{dark room} experiment.} The adjective \textit{dark} is used as an allegory to indicate the complete absence of any \textit{external} sensory input.}  
+  \label{fig:darkroom}
+\end{wrapfigure}
+However, this is definitely not what is observed, neither in humans nor in animals: if you get into a dark room, you try to figure out where the light switch is, so as to \textit{explore} this mysterious room, or simply get out of it!
+Defenders of the FEP have retaliated by advancing the `full package''\citep{darkroom2012}: FEP is multi-scale, and there would be a meta-scale at which the organism would
+be \textit{surprised} by such a \textit{lack of surprise}.
+According to this argument, a dark room is thus paradoxically, a state of extremely high surprise.
+Driven by the zeal to always diminish surprise, the FEP agent would eventually bootstrap itself out of such saddles, to explore more interesting corners of the environment.
+Recourse to this argument is essential equivalent to invoking a \textit{halting oracle}, which is problematic as it sucks up the problem the theory is trying to address in the first place.
+
+In contrast, operating under our RL-based theory, an organism would quickly notice that this dark room is a local minimum: hippocampal experience replay (see \ref{sec:self}) would set it and the organism would sample memories or fantasies of alternative situations in which there was reward to mine.
+Such considerations would artificially generate \textit{internal} sensory input and eventually let the organism to explore the room (for example look for and find the light switch), or simply leave the room.
 
@@
 
-\textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
+% \textbf{XXX: integrate aspects from Friston2014(Dropbox) on active inference}
 

### 2017-05-18 paper/paper.tex
more on DRE

@@
 as biomechanical inference engine
-dedicated to minimizing the long-term average of surprise: the log-likelihood of the observed sensory input, or more precisely, an upper bound thereof.
+dedicated to minimizing the long-term average of surprise: the log-likelihood of the observed sensory input --or more precisely, an upper bound thereof-- \textit{relative} to the expectations of, and the internal representations the brain has about the external world.
 It does this by actively generating hypothetical explanations or predictions of its sensory input.
@@
 \subsubsection{Criticisms of the FEP}
-  Despite it's popularity, a number of critics about the FEP have surfaced over the years. We outline the main ones here.
+  Despite its popularity, a number of criticisms against the FEP have surfaced over the years. We outline the most notable ones here.
 
@@
 \paragraph{The ``dark room'' experiment.}
-Since the FEP postulates that all ongoing intelligent behavior in an organism is geared towards reaching the grand goal of minimizing surprisal, an organism which enters a \textit{dark room}\footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.} would strive to remain there as its sensory input would be perfectly predictable in such a state \citep{darkroom2012}.
+Since the FEP postulates that all ongoing intelligent behavior in an organism is geared towards reaching the grand goal of minimizing surprisal, an organism which enters a \textit{dark room}\footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.} (Fig.
+\ref{fig:darkroom}) would strive to remain there as its sensory input would be perfectly predictable in such a state \citep{darkroom2012}.
+However, such a tendency is definitely not what is observed, neither in humans nor in animals: if you get into a dark room,
 \begin{wrapfigure}{r}{.45\linewidth}
@@
 \end{wrapfigure}
-However, this is definitely not what is observed, neither in humans nor in animals: if you get into a dark room, you try to figure out where the light switch is, so as to \textit{explore} this mysterious room, or simply get out of it!
+you try to figure out where the light switch is, so as to \textit{explore} this mysterious room, or simply get out of it!
 Defenders of the FEP have retaliated by advancing the `full package''\citep{darkroom2012}: FEP is multi-scale, and there would be a meta-scale at which the organism would
@@
 Driven by the zeal to always diminish surprise, the FEP agent would eventually bootstrap itself out of such saddles, to explore more interesting corners of the environment.
-Recourse to this argument is essential equivalent to invoking a \textit{halting oracle}, which is problematic as it sucks up the problem the theory is trying to address in the first place.
-
-In contrast, operating under our RL-based theory, an organism would quickly notice that this dark room is a local minimum: hippocampal experience replay (see \ref{sec:self}) would set it and the organism would sample memories or fantasies of alternative situations in which there was reward to mine.
-Such considerations would artificially generate \textit{internal} sensory input and eventually let the organism to explore the room (for example look for and find the light switch), or simply leave the room.
+Recourse to this argument is equivalent to invoking a \textit{halting oracle}\citep{haltingoracle}, an arbitrarily powerful \textit{Turing machine} which can solve the \textit{halting problem} (an arbitrarily difficult computational problem which cannot be solved under any reasonable \textit{model of computation}, not even in principle!).
+But this is problematic as it sucks up the problem (that of decision-making and active agency) the theory is intends to elude in the first place.
+
+In contrast, operating under our RL-based theory, an organism would quickly notice that this dark room is a local minimum.
+Indeed, hippocampal experience replay (see \ref{sec:self}) would set it and the organism would sample memories or fantasies of alternative situations in which there was reward to mine.
+Such considerations would artificially generate \textit{internal} sensory input for the DMN and eventually let the organism to explore the room, for example look for and find the light switch, or simply leave the room.
 

### 2017-05-18 paper/paper.tex
typo

@@
 \paragraph{The ``dark room'' experiment.}
-Since the FEP postulates that all ongoing intelligent behavior in an organism is geared towards reaching the grand goal of minimizing surprisal, an organism which enters a \textit{dark room}\footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.} (Fig.
+Since FEP-based theories \citep{friston2010free,fristonAIorRL} postulate that all ongoing intelligent behavior in an organism is geared towards reaching the grand goal of minimizing surprisal, an organism which enters a \textit{dark room}\footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.} (Fig.
 \ref{fig:darkroom}) would strive to remain there as its sensory input would be perfectly predictable in such a state \citep{darkroom2012}.
@@
 Recourse to this argument is equivalent to invoking a \textit{halting oracle}\citep{haltingoracle}, an arbitrarily powerful \textit{Turing machine} which can solve the \textit{halting problem} (an arbitrarily difficult computational problem which cannot be solved under any reasonable \textit{model of computation}, not even in principle!).
-But this is problematic as it sucks up the problem (that of decision-making and active agency) the theory is intends to elude in the first place.
+But this is problematic as it sucks up the problem (that of decision-making and active agency) the theory intends to elude in the first place.
 

### 2017-05-18 paper/paper.tex
Additional infos for RTPJ and synaptic epigenesis

--- 2017-05-18 paper/paper.tex
@@
 state and action).
-The framing of the policy matrix is not described here but could be supported by the formation of new attractor in the neural dynamics through synaptic epigenesis \citep{gisiger_acquisition_2005}.}
+The framing of the policy matrix is not described here but could be supported by synaptic epigenesis \citep{gisiger_acquisition_2005}.
+Indeed, the tuning of synaptic weights through learning can stabilize additional patterns of activity by
+creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}. Those attractors will then constrain both
+the number of factors taken into account by decision processes and the number of behaviors at reach by the agent \citep{wang_decision_2008}.}
 
@@
 \suggestadd{The principal connectivity gradient in the cortex appears to be greatly expanded in humans compare to macaques, suggesting a phylogenetically conserved axis of cortical expansion with the DMN emerging at the extreme-end in humans \citep{margulies_situating_2016}.
-Neurocomputational models of dyadic whole brain demonstrated how the human connectivity topology, on top of facilitating processing at the intra-individual level, can explain our propensity to coordinate through sensorimotor loops with others at the inter-individual level \citep{dumas_anatomical_2012}.}
+Neurocomputational models of dyadic whole brain demonstrated how the human connectivity topology, on top of facilitating processing at the intra-individual level, can explain our propensity to coordinate through sensorimotor loops with others at the inter-individual level \citep{dumas_anatomical_2012}.
+DMN is moreover largely overlapping with neural networks associated with higher level social cognition \citep{schilbach_introspective_2012}.
+For instance, the ventromedial prefrontal cortex, the posterior cingulate/precuneus, and the temporo-parietal junction (rTPJ) play together a key role in bridging the gap between self and other by
+integrating low-level embodied processes within higher level inference-based mentalizing \citep{lombardo_shared_2009}.}
 

### 2017-05-18 paper/paper.tex
removed XXXs on MCMC part

@@
 highest expected cumulative rewards.
-In MDPs, MCMC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible
-\textbf{XXX elvis: refs needed.}.
-Much recent work points to MCMC or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
+In MDPs, MCMC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible \citep{silver2016mastering}.
+% Much recent work points to MCMC or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
 \textbf{XXX elvis: refs needed in last sentence.
-Plus, should be really bring up Bayesian?} For sampling-based methods it is still scarce
-beyond the description of receptive fields. \textbf{XXX elvis: rephrase last sentence}
+Plus, should be really bring up Bayesian?}
 

### 2017-05-21 paper/paper.tex
more fixups

@@
 Driven by the zeal to always diminish surprise, the FEP agent would eventually bootstrap itself out of such saddles, to explore more interesting corners of the environment.
-Recourse to this argument is equivalent to invoking a \textit{halting oracle}\citep{haltingoracle}, an arbitrarily powerful \textit{Turing machine} which can solve the \textit{halting problem} (an arbitrarily difficult computational problem which cannot be solved under any reasonable \textit{model of computation}, not even in principle!).
+Recourse to this argument is equivalent to invoking a \textit{halting oracle}\citep{haltingoracle}, an arbitrarily powerful machine which can solve the \textit{halting problem} (an arbitrarily difficult computational problem which cannot be solved under any reasonable \textit{model of computation}, not even in principle!).
 But this is problematic as it sucks up the problem (that of decision-making and active agency) the theory intends to elude in the first place.

### 2017-06-10 paper/paper.tex
fixing Fig 3 caption layout
pass over FEP

@@
 
-\begin{wrapfigure}{r}{.4\linewidth}
+\begin{wrapfigure}{r}{.3\linewidth}
   \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
@@
 \subsection{The free-energy principle and active inference}
-In Friston's free-energy principle (FEP) \citep{friston2010free} and active inference  (i.e action under the FEP) \citep{fristonAIorRL} theories, the brain is portrayed
-as biomechanical inference engine
-dedicated to minimizing the long-term average of surprise: the log-likelihood of the observed sensory input --or more precisely, an upper bound thereof-- \textit{relative} to the expectations of, and the internal representations the brain has about the external world.
-It does this by actively generating hypothetical explanations or predictions of its sensory input.
-Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net}.
-According to the FEP account,
-the goal of the brain is then to optimize over a generative model $G$ of sensations: to iteratively
-modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics,
-so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles (i.e., the \textit{generative} model).
+% INTRO
+According to the free-energy principle (FEP) and
+active inference theories \citep{friston2010free, fristonAIorRL},
+the brain corresponds to
+a biomechanical inference engine. It is
+dedicated to minimizing the long-term average of surprise: the log-likelihood of the observed sensory input --or more precisely, an upper bound thereof-- \textit{relative} to the expectations about the external world derived from the brain's internal representations.
+The brain
+would continuously generate hypothetical explanations of the world
+and predict its sensory input.
+Precursors of this theory can be traced back to Dayan and colleagues \citep{dayan1995helmholtz}
+who introduced the \textit{Helmholtz machine}, a hierarchical factorial directional deep belief network. According to the FEP account,
+the goal of the brain is to optimize over a generative model $G$ of sensations: to iteratively
+modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics. The model then allows
+to minimize surprise when these representations are confronted with sensory input $\x$ during action-perception cycles (i.e., the \textit{generative} model).
 The FEP also postulates a dual model called the \textit{recognition} model,
-which works in tandem with the generative model $p_R(\z|\x)$, to accomplish approximate inference. The recognition model dreams imaginary worlds $\z$,
-while the latter tries to generate sensations $\x$ which match these dreams!
-
-Because surprise is intractably difficult to optimize (since we would need to sum over all hidden causes of the sensations), FEP sets out to instead minimize an upper-bound thereof,
+which works in tandem with the generative model $p_R(\z|\x)$ to accomplish approximate inference. Put differently, the recognition model dreams imaginary worlds $\z$,
+while the generative model tries to produce sensations $\x$ which match these
+imagined mental scenarios.
+Howewer, surprise is challenging to optimize numerically
+because we need to sum over all hidden causes of the sensations.
+Instead, FEP therefore minimizes an upper-bound of surprise,
 namely the \text{free-energy} $F_G^R(\x)$ given by
@@
 
-\subsubsection{Criticisms of the FEP}
-  Despite its popularity, a number of criticisms against the FEP have surfaced over the years. We outline the most notable ones here.
-
-\paragraph{The wake-sleep algorithm is not neurobiologically plausible.}
+% CRITICISM
+Despite its popularity, criticism against the FEP has arisen over the years,
+some of which we outline in the following.
 The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
-\citep{dayan1995helmholtz}. As noted already in \citep{dayan1995helmholtz},
-a crucial drawback of the wake-sleep algorithm is that it postulates a pair forward (generation) and backward (recognition) models,
-which together do not correspond to optimization of (a bound of) the marginal likelihood, because of the incorrect KL used therein, etc.
-Thus the brain could not possibly be running such a algorithm, not even in principle! To the rescue, we note that the recent theory of
-\textit{variational auto-encoders (VAEs)} \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm,
-as it overcomes the technical limits of the former, by using a reparametrization trick. For instance, unlike the wake-sleep algorithm
+\citep{dayan1995helmholtz}. As previously noted \citep{dayan1995helmholtz},
+a crucial drawback of the wake-sleep algorithm is that it postulates a pair
+of forward (generation) and backward (recognition) models that
+together do not correspond to optimization of (a bound of) the marginal likelihood.
+Thus, the brain is unlikely to implement a variant of the wake-sleep algorithm.
+We note that the recent theory of
+\textit{variational auto-encoders} (VAEs) \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm. This is because VAEs
+overcome a number of the technical limits of the wake-sleep algorithm
+by using a reparametrization trick. For instance, unlike the wake-sleep algorithm
 for minimizing free-energy, VAEs can be efficiently trained via back-propagation of prediction errors.
 
-\paragraph{The ``dark room'' experiment.}
-Since FEP-based theories \citep{friston2010free,fristonAIorRL} postulate that all ongoing intelligent behavior in an organism is geared towards reaching the grand goal of minimizing surprisal, an organism which enters a \textit{dark room}\footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.} (Fig.
-\ref{fig:darkroom}) would strive to remain there as its sensory input would be perfectly predictable in such a state \citep{darkroom2012}.
-However, such a tendency is definitely not what is observed, neither in humans nor in animals: if you get into a dark room,
+Since FEP-based theories \citep{friston2010free,fristonAIorRL} conceptualize ongoing behavior
+in an organism to be geared towards the surprise-minimizing goal,
+an organism entering a dark room
+% \footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.}
+(Fig. \ref{fig:darkroom}) would strive to remain in this location
+because its sensory inputs are perfectly predictable
+in this environmental state \citep{darkroom2012}.
+However, such a behavioral tendency is
+seldom observed in animals or humans in the real world: In a dark room,
 \begin{wrapfigure}{r}{.45\linewidth}
@@
   \includegraphics[width=1.\linewidth]{darkroom.jpg}
-  \caption{\textbf{The \textit{dark room} experiment.} The adjective \textit{dark} is used as an allegory to indicate the complete absence of any \textit{external} sensory input.}
+  \caption{\textbf{The dark room experiment.} An intelligent agent situated
+  in a closed space is used as a thought experiment for the complete absence
+  of external sensory input.}
   \label{fig:darkroom}
 \end{wrapfigure}
-you try to figure out where the light switch is, so as to \textit{explore} this mysterious room, or simply get out of it!
-Defenders of the FEP have retaliated by advancing the `full package''\citep{darkroom2012}: FEP is multi-scale, and there would be a meta-scale at which the organism would
-be \textit{surprised} by such a \textit{lack of surprise}.
-According to this argument, a dark room is thus paradoxically, a state of extremely high surprise.
-Driven by the zeal to always diminish surprise, the FEP agent would eventually bootstrap itself out of such saddles, to explore more interesting corners of the environment.
-Recourse to this argument is equivalent to invoking a \textit{halting oracle}\citep{haltingoracle}, an arbitrarily powerful machine which can solve the \textit{halting problem} (an arbitrarily difficult computational problem which cannot be solved under any reasonable \textit{model of computation}, not even in principle!).
+intelligent agents would search for light sources to \textit{explore}
+its surroundings or leave it.
+Defenders of the FEP have retorted by advancing the ``full package'' \citep{darkroom2012}:
+FEP is proposed to be multi-scale and there would be a meta-scale at which the organism would
+be \textit{surprised} by such a \textit{lack of surprise}. According to this argument, a dark room would paradoxically correspond to a state of particularly high surprise.
+Driven by surprise-minimization objective, the FEP agent would eventually bootstrap itself out of such saddles to explore more interesting parts of the environment.
+% Recourse to this argument is equivalent to invoking a \textit{halting oracle}\citep{haltingoracle}, an arbitrarily powerful machine which can solve the \textit{halting problem} (an arbitrarily difficult computational problem which cannot be solved under any reasonable \textit{model of computation}, not even in principle!).
 But this is problematic as it sucks up the problem (that of decision-making and active agency) the theory intends to elude in the first place.
-
-In contrast, operating under our RL-based theory, an organism would quickly notice that this dark room is a local minimum.
-Indeed, hippocampal experience replay (see \ref{sec:self}) would set it and the organism would sample memories or fantasies of alternative situations in which there was reward to mine.
-Such considerations would artificially generate \textit{internal} sensory input for the DMN and eventually let the organism to explore the room, for example look for and find the light switch, or simply leave the room.
-
-\subsubsection{RL reinterpretation of FEP and active inference}  FEP and active inference can be reinterpreted in our model-free RL framework by postulating that the reward signal maximized by the DMN is the negative of surprisal, i.e the log-likehood of sensory the current priors the agent has about the world.
-This is worth writing down explicitly,
+In contrast, an organism operating under our RL-based theory would readily
+identify the visual-input-free room as a local minimum. Indeed, hippocampal experience replay (see \ref{sec:self}) would serve to sample memories or fantasies of alternative situations
+with reward structure. Such artificially generated \textit{internal} sensory input
+subserved by the DMN can entice the organism to explore the room,
+for instance by looking for and using the light switch or simply leaving the room.
+
+Finally, we note that FEP and active inference can be recast in our model-free RL framework.
+This becomes possible by postulating that the reward signal maximized by the DMN is
+the negative surprise,
+that is, the log-likehood of sensory the current priors the agent has about the world. More explicitly, this corresponds to
 
@@
 %   \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}.
-    Such a surprisal-based reinforcement learning scheme has been advocated for in \citep{schmidhuber2010,mohamed2015} under the equivalent framework of information compression.
-    However, we note that though relevant, minimization of surprisal alone is insufficient to explain intelligent behavior observed in humans and other ``intelligent'' animals.
-The dark-room experiment, for example, provides a formidable argument against in favor of the insufficiency of the FEP and active inference theories to explain intelligent behavior in humans: when a human is put in a dark room, they don't just sit down in a corner forever, what would be the FEP-optimal thing to do; they actively search for the light switch, and start exploring the depths of the room, possibly looking for food or the way out!
+Such a surprisal-based reinforcement learning scheme has previously been advocated \citep{schmidhuber2010,mohamed2015} under the equivalent framework of information compression.
+Nevertheless, minimization of surprise alone
+may be insufficient to explain behavior observed in humans and other intelligent animals.
+% The dark-room experiment, for example, provides a formidable argument against in favor of the insufficiency of the FEP and active inference theories to explain intelligent behavior in humans: when a human is put in a dark room, they don't just sit down in a corner forever, what would be the FEP-optimal thing to do; they actively search for the light switch, and start exploring the depths of the room, possibly looking for food or the way out!
 

### 2017-06-10 paper/paper.tex
more fixups

@@
 offers parsimonious explanations for
-recent experimental findings\suggestremove{ in animals and humans}.
+recent experimental findings in animals and humans.
 \end{abstract}
@@
 a longer-term future.
-Precisely, rewards that are not expected to come within\footnote{\textbf{N.B.:} $\frac{1}{1-\gamma} = 1 + \gamma + \gamma^2 + \gamma^3 + \ldots$}
+Precisely, rewards that are not expected to come within
+% \footnote{\textbf{N.B.:} $\frac{1}{1-\gamma} = 1 + \gamma + \gamma^2 + \gamma^3 + \ldots$}
 $\tau := 1/(1 - \gamma)$
@@
 which take the agent to different table entries.
-\suggestadd{An additional layer of learning concerns the addition of new entries in the table (i.e.
-state and action).
+An additional layer of learning concerns the addition of new entries in the table (i.e., state and action).
 The framing of the policy matrix is not described here but could be supported by synaptic epigenesis \citep{gisiger_acquisition_2005}.
@@
 
-\suggestadd{The principal connectivity gradient in the cortex appears to be greatly expanded in humans compare to macaques, suggesting a phylogenetically conserved axis of cortical expansion with the DMN emerging at the extreme-end in humans \citep{margulies_situating_2016}.
+The principal connectivity gradient in the cortex appears to be greatly expanded in humans compare to macaques, suggesting a phylogenetically conserved axis of cortical expansion with the DMN emerging at the extreme-end in humans \citep{margulies_situating_2016}.
 Neurocomputational models of dyadic whole brain demonstrated how the human connectivity topology, on top of facilitating processing at the intra-individual level, can explain our propensity to coordinate through sensorimotor loops with others at the inter-individual level \citep{dumas_anatomical_2012}.
@@
 For instance, the ventromedial prefrontal cortex, the posterior cingulate/precuneus, and the temporo-parietal junction (rTPJ) play together a key role in bridging the gap between self and other by
-integrating low-level embodied processes within higher level inference-based mentalizing \citep{lombardo_shared_2009}.}
+integrating low-level embodied processes within higher level inference-based mentalizing \citep{lombardo_shared_2009}.
 
@@
 the goal of the brain is to optimize over a generative model $G$ of sensations: to iteratively
-modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics. The model then allows
+modify its internal representation $p_G(\z|\x)$ of objects in the world, their interactions and dynamics. The model then allows
 to minimize surprise when these representations are confronted with sensory input $\x$ during action-perception cycles (i.e., the \textit{generative} model).
-The FEP also postulates a dual model called the \textit{recognition} model,
+The FEP also postulates a companion model called the \textit{recognition} model,
 which works in tandem with the generative model $p_R(\z|\x)$ to accomplish approximate inference. Put differently, the recognition model dreams imaginary worlds $\z$,
@@
 together do not correspond to optimization of (a bound of) the marginal likelihood.
-Thus, the brain is unlikely to implement a variant of the wake-sleep algorithm.
+Thus, the brain may therefore be unlikely to implement a variant of the wake-sleep algorithm.
 We note that the recent theory of
@@
 subserved by the DMN can entice the organism to explore the room,
-for instance by looking for and using the light switch or simply leaving the room.
+for instance by looking for and using the light switch or simply finding the room exit.
 

### 2017-06-10 paper/paper.tex
more more fixes

@@
 % official NIPS keywords
-\textbf{\\keywords}: systems biology, mind-wandering, cognitive science,
-artificial intelligence, Markov decision process
-\textbf{XXX: Maybe the keywords "systems biology" and "cognitive science" are too broad.
-I would also remove "reinforcement learning", which is already in the title.
-The keywords are the occasion to enhance reachability through web and bibliography search.
-What about "process model", "Markov decision processes", or "Bellman equation"?}
+\textbf{\\keywords}: mind-wandering, artificial intelligence, Markov decision process
+
 % autonomous learning systems

### 2017-06-10 paper/paper.tex
fixes

@@
 creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}. Those attractors will then constrain both
-the number of factors taken into account by decision processes and the number of behaviors at reach by the agent \citep{wang_decision_2008}.}
+the number of factors taken into account by decision processes and the number of behaviors at reach by the agent \citep{wang_decision_2008}.
 

### 2017-06-11 paper/paper.tex
fixed broken figure layout
typpass 2.0 over FEP

@@
 
-\begin{wrapfigure}{r}{.3\linewidth}
-  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
-    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
-  \centering
-  \includegraphics[width=1.\linewidth]{rl.pdf}
-  \label{fig:rl}
-\end{wrapfigure}
 The environment is generally taken as \textit{stochastic},
@@
 \citep{dayan2008decision}.
+
+\begin{wrapfigure}{r}{.3\linewidth}
+  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
+    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
+  \centering
+  \includegraphics[width=1.\linewidth]{rl.pdf}
+  \label{fig:rl}
+\end{wrapfigure}
 
@@
 
-\section{Relation to existing cognitive accounts on default mode function}
+\section{Relation to existing functional accounts}
 \subsection{Predictive coding hypothesis}
@@
 % INTRO
-According to the free-energy principle (FEP) and
-active inference theories \citep{friston2010free, fristonAIorRL},
+According to theories of the free-energy principle (FEP) and
+active inference \citep{friston2010free, fristonAIorRL},
 the brain corresponds to
-a biomechanical inference engine. It is
-dedicated to minimizing the long-term average of surprise: the log-likelihood of the observed sensory input --or more precisely, an upper bound thereof-- \textit{relative} to the expectations about the external world derived from the brain's internal representations.
+a biomechanical reasoning engine. It is
+dedicated to minimizing the long-term average of surprise: the log-likelihood of the observed sensory input --more precisely, an upper bound thereof-- \textit{relative} to the expectations about the external world derived from internal representations.
 The brain
@@
 and predict its sensory input.
-Precursors of this theory can be traced back to Dayan and colleagues \citep{dayan1995helmholtz}
+Precursors of this interpretation can be traced back to Dayan and colleagues \citep{dayan1995helmholtz}
 who introduced the \textit{Helmholtz machine}, a hierarchical factorial directional deep belief network. According to the FEP account,
@@
 to minimize surprise when these representations are confronted with sensory input $\x$ during action-perception cycles (i.e., the \textit{generative} model).
-The FEP also postulates a companion model called the \textit{recognition} model,
+The FEP also proposes a companion model called the \textit{recognition} model,
 which works in tandem with the generative model $p_R(\z|\x)$ to accomplish approximate inference. Put differently, the recognition model dreams imaginary worlds $\z$,
-while the generative model tries to produce sensations $\x$ which match these
+while the generative model tries to produce sensory sensations $\x$ which match these
 imagined mental scenarios.
@@
 Instead, FEP therefore minimizes an upper-bound of surprise,
-namely the \text{free-energy} $F_G^R(\x)$ given by
+namely the \text{free energy} $F_G^R(\x)$ given by
 
@@
 Despite its popularity, criticism against the FEP has arisen over the years,
-some of which we outline in the following.
-The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
-\citep{dayan1995helmholtz}. As previously noted \citep{dayan1995helmholtz},
-a crucial drawback of the wake-sleep algorithm is that it postulates a pair
+some of which outlined in the following.
+The main algorithm for minimizing free energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
+\citep{dayan1995helmholtz}. As these authors noted,
+a crucial drawback of the wake-sleep algorithm is that it involves a pair
 of forward (generation) and backward (recognition) models that
 together do not correspond to optimization of (a bound of) the marginal likelihood.
-Thus, the brain may therefore be unlikely to implement a variant of the wake-sleep algorithm.
-We note that the recent theory of
-\textit{variational auto-encoders} (VAEs) \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm. This is because VAEs
+The brain may therefore be unlikely to implement a variant of the wake-sleep algorithm.
+The recent theory of
+\textit{variational auto-encoders} (VAEs) \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm. VAEs
 overcome a number of the technical limits of the wake-sleep algorithm
 by using a reparametrization trick. For instance, unlike the wake-sleep algorithm
-for minimizing free-energy, VAEs can be efficiently trained via back-propagation of prediction errors.
-
-Since FEP-based theories \citep{friston2010free,fristonAIorRL} conceptualize ongoing behavior
+for minimizing free energy, VAEs can be efficiently trained via back-propagation of prediction errors.
+
+Since theories based on the FEP \citep{friston2010free,fristonAIorRL}
+conceptualize ongoing behavior
 in an organism to be geared towards the surprise-minimizing goal,
@@
 because its sensory inputs are perfectly predictable
-in this environmental state \citep{darkroom2012}.
+given the environmental state \citep{darkroom2012}.
 However, such a behavioral tendency is
@@
   \caption{\textbf{The dark room experiment.} An intelligent agent situated
-  in a closed space is used as a thought experiment for the complete absence
+  in a light-deprived closed space can be used
+  as a thought experiment for the complete absence
   of external sensory input.}
@@
 be \textit{surprised} by such a \textit{lack of surprise}. According to this argument, a dark room would paradoxically correspond to a state of particularly high surprise.
-Driven by surprise-minimization objective, the FEP agent would eventually bootstrap itself out of such saddles to explore more interesting parts of the environment.
+Driven by surprise-minimization objective, the FEP agent would eventually bootstrap itself out of such saddle points to explore more interesting parts of the environment.
 % Recourse to this argument is equivalent to invoking a \textit{halting oracle}\citep{haltingoracle}, an arbitrarily powerful machine which can solve the \textit{halting problem} (an arbitrarily difficult computational problem which cannot be solved under any reasonable \textit{model of computation}, not even in principle!).
@@
 In contrast, an organism operating under our RL-based theory would readily
-identify the visual-input-free room as a local minimum. Indeed, hippocampal experience replay (see \ref{sec:self}) would serve to sample memories or fantasies of alternative situations
+identify the visual-input-deprived room as a local minimum. Indeed, hippocampal experience replay (see \ref{sec:self}) would serve to sample memories or fantasies of alternative situations
 with reward structure. Such artificially generated \textit{internal} sensory input
@@
 
-Finally, we note that FEP and active inference can be recast in our model-free RL framework.
-This becomes possible by postulating that the reward signal maximized by the DMN is
-the negative surprise,
-that is, the log-likehood of sensory the current priors the agent has about the world. More explicitly, this corresponds to
+Finally, we note that FEP and active inference can be reframed in terms of
+our model-free RL framework.
+This becomes possible by recasting the reward signal maximized by the DMN to 
+correspond to negative surprise,
+that is, the log-likehood of current sensory priors
+the agent has about the world. More explicitly, this corresponds to
 
 \begin{equation}
-  r^{\text{FEP}} = \underbrace{\log (p_G)}_{-\text{FEP generative surprisal}}.
+  r^{\text{FEP}} = \underbrace{\log (p_G)}_{-\text{FEP generative surprise}}.
     \end{equation}
@@
 %   \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}.
-Such a surprisal-based reinforcement learning scheme has previously been advocated \citep{schmidhuber2010,mohamed2015} under the equivalent framework of information compression.
-Nevertheless, minimization of surprise alone
-may be insufficient to explain behavior observed in humans and other intelligent animals.
+Such a surprise-guided reinforcement learning scheme has previously been advocated
+under the equivalent framework of information compression \citep{schmidhuber2010,mohamed2015}.
+Nevertheless, minimization of surprise quantities alone
+may be insufficient to explain
+the diversity of behaviors observed in humans and other intelligent animals.
 % The dark-room experiment, for example, provides a formidable argument against in favor of the insufficiency of the FEP and active inference theories to explain intelligent behavior in humans: when a human is put in a dark room, they don't just sit down in a corner forever, what would be the FEP-optimal thing to do; they actively search for the light switch, and start exploring the depths of the room, possibly looking for food or the way out!

### 2017-06-11 paper/paper.tex
new logo

@@
 % official NIPS keywords
-\textbf{\\keywords}: mind-wandering, artificial intelligence, Markov decision process
+\textbf{\\keywords}: mind-wandering, artificial intelligence, Markov decision processes
 
@@
 brain imaging may have been the first technique
-to allow for the discovery of a unique brain network that probably maintains
+to allow for the discovery of a unique brain network that probably subserves
 baseline mental activities
@@
 at the onset of defined psychological tasks \citep{raichle_baseline}.
-At the opposite, during sleep, the
+At its opposite, during sleep, the
 decoupling of brain structures discarded the idea of the DMN being
@@
 
-This dark matter of brain physiology \citep{raichle2006brain} begs the question
+This \textit{dark matter of brain physiology}
+\citep{raichle2006brain} begs the question
 of the biological purpose underlying DMN activity.
@@
 But this is problematic as it sucks up the problem (that of decision-making and active agency) the theory intends to elude in the first place.
-In contrast, an organism operating under our RL-based theory would readily
-identify the visual-input-deprived room as a local minimum. Indeed, hippocampal experience replay (see \ref{sec:self}) would serve to sample memories or fantasies of alternative situations
+In contrast, an organism operating under our RL-based theory would inevitably
+identify the sensory-stimulus-deprived room as a local minimum. Indeed, hippocampal experience replay (see \ref{sec:self}) would serve to sample memories or fantasies of alternative situations
 with reward structure. Such artificially generated \textit{internal} sensory input
@@
 MDPs also provide a mathematical formalism how
-opportunity structure in the environment can be recursively exploited
+opportunity in the environment can be recursively exploited
 when confronted with challenging decisions.
@@
 Improvement of the internal world representation
-by injecting stochasticity in the recall of past
+by injecting stochasticity into the recall of past
 actions and the estimation of action outcomes may explain why
@@
 
-The proposed artificiale intelligence perspective
+The proposed artificial intelligence perspective
 on DMN biology is hence practically computable and
-yields falsifiable neurobiological hypotheses.
+yields falsifiable neuroscientific hypotheses.
 %
@@
 the set of action, value, and state variables that govern
-the behavior of intelligent RL systems.
+the behavior of intelligent RL agents.
 %
@@
 
-\appendix
-
-\begin{remark}
-  Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
-  except the Markov property, we can always transform it into an MDP by considering a compounded
-  version of the states
-  $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
-  history up to an including time $t$.
-\end{remark}
-
-\section{More on RL in the brain}
-\subsection{Does hippocampal replay equal inverse reinforcement-learning?}
-%\begin{mdframed}
-  Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
-  an MDP (called a \textit{teacher's demonstration}), can we figure out what is the
-  (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
-  \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
-  (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
-  For instance, given traces of motor actions from an adult to grab a cup on a table,
-  can an observing child figure out what ``langragian'' functional is being minimized by the former?
-  How? Can they reproduce this optimal behavior?
-%\end{mdframed}
-Such questions are of course pertinent to our decision-making theory for the DMN.
-In the general artificial case, the problem has been extensively studied and partially solved by
-\citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
-``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
-
-IRL is is suited for problems in which it's hard to define what the reward function could be
-(e.g.,  car-driving, drone maneuvering, etc.) ...
-
-\section{Free-energy principles!}
-The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
-In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
+% \appendix
+% 
+% \begin{remark}
+%   Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
+%   except the Markov property, we can always transform it into an MDP by considering a compounded
+%   version of the states
+%   $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
+%   history up to an including time $t$.
+% \end{remark}
+% 
+% \section{More on RL in the brain}
+% \subsection{Does hippocampal replay equal inverse reinforcement-learning?}
+% %\begin{mdframed}
+%   Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
+%   an MDP (called a \textit{teacher's demonstration}), can we figure out what is the
+%   (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
+%   \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
+%   (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
+%   For instance, given traces of motor actions from an adult to grab a cup on a table,
+%   can an observing child figure out what ``langragian'' functional is being minimized by the former?
+%   How? Can they reproduce this optimal behavior?
+% %\end{mdframed}
+% Such questions are of course pertinent to our decision-making theory for the DMN.
+% In the general artificial case, the problem has been extensively studied and partially solved by
+% \citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
+% ``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
+% 
+% IRL is is suited for problems in which it's hard to define what the reward function could be
+% (e.g.,  car-driving, drone maneuvering, etc.) ...
+% 
+% \section{Free-energy principles!}
+% The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
+% In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
 This ideas were first developed by Hinton et al.
@@
 For instance, the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
-
-% \begin{mdframed}
-%   Though our presentation of free-energy principles is thermodynamic in flavor, a completely equivalent
-%   information-theoretic formulation is of course possible, simply by exploiting the information-theoretic interpretation of entropy.
-% \end{mdframed}
-
-\subsection{Helmholtz free-energy and the generative model}
-\begin{table}[H]
-  \begin{tabular}{p{2cm}|p{11cm}}
-         \hline
-         \textbf{symbol}    & \textbf{description}  \\ \hline
-         $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
-         random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
-         $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
-          with the usual convention $0 \log(0) := 0$.\\ \hline
-         $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
-             $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e., $\x = (s, b).$ \\ \hline
-             $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
-             $p_G(.|\x)$ & Generative density for ...\\ \hline
-         $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
-         $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
-         $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
-          to $R$.  Note that $F^G_G = F_G$.\\ \hline
-  \end{tabular}
-  \caption{Table of notations.}
-\end{table}
-Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
+% 
+% % \begin{mdframed}
+% %   Though our presentation of free-energy principles is thermodynamic in flavor, a completely equivalent
+% %   information-theoretic formulation is of course possible, simply by exploiting the information-theoretic interpretation of entropy.
+% % \end{mdframed}
+% 
+% \subsection{Helmholtz free-energy and the generative model}
+% \begin{table}[H]
+%   \begin{tabular}{p{2cm}|p{11cm}}
+%          \hline
+%          \textbf{symbol}    & \textbf{description}  \\ \hline
+%          $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
+%          random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
+%          $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
+%           with the usual convention $0 \log(0) := 0$.\\ \hline
+%          $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
+%              $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e., $\x = (s, b).$ \\ \hline
+%              $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
+%              $p_G(.|\x)$ & Generative density for ...\\ \hline
+%          $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
+%          $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
+%          $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
+%           to $R$.  Note that $F^G_G = F_G$.\\ \hline
+%   \end{tabular}
+%   \caption{Table of notations.}
+% \end{table}
+% Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
 So,
-\begin{eqnarray}
-  \begin{split}
-    \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(P_G(\x))\sum_{\z}p_G(\z |\x)\\
-    &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
-    =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
-    &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
-    &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
-    &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
-  \end{split}
-  \label{eq:helm}
-\end{eqnarray}
-where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
-\begin{equation}
-  E_G(\z, \x) := -\log(p_G(\z, \x)),
-  \label{eq:gibbs}
-\end{equation}
-%% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
-The last quantity in \eqref{eq:helm} is nothing but \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
-\begin{equation}
-  F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
-\end{equation}
-%% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
-Thus,
-\begin{fact}
-  Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
-\end{fact}
-
-The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
+% \begin{eqnarray}
+%   \begin{split}
+%     \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(P_G(\x))\sum_{\z}p_G(\z |\x)\\
+%     &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
+%     =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
+%     &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
+%     &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
+%     &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
+%   \end{split}
+%   \label{eq:helm}
+% \end{eqnarray}
+% where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
+% \begin{equation}
+%   E_G(\z, \x) := -\log(p_G(\z, \x)),
+%   \label{eq:gibbs}
+% \end{equation}
+% %% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
+% The last quantity in \eqref{eq:helm} is nothing but \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
+% \begin{equation}
+%   F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
+% \end{equation}
+% %% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
+% Thus,
+% \begin{fact}
+%   Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
+% \end{fact}
+% 
+% The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
 It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(\x)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
@@
 1995, is at the heart of the general free-energy principle that we will introduce shortly.
-
-\subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
-In this subsection, we will present an insightful upper bound for the generative surprise (i.e., generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
+% 
+% \subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
+% In this subsection, we will present an insightful upper bound for the generative surprise (i.e., generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
 As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
-
-Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
-distribution
-\begin{equation}
-  p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
-\end{equation}
-where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
-Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
-\begin{equation}
-F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
-\end{equation}
-Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
-it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
-density\footnote{conditioning in $P_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
-  state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
-\begin{equation}
-  \begin{split}
-    F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{P_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
-    &= \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
-    &= F^R_G(\x) - D_{KL}(P_R(.|\x) || P_G(.|\x)),
-  \end{split}
-  \label{eq:fe}
-\end{equation}
-where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
-\begin{equation}
-  F^R_G(\x) := \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x))
-\end{equation}
-and $D_{KL}(P_R(.|\x) || P_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
-
-\subsection{A general free-energy principle}
-We can resume the situation as follows\footnote{Where we have used the fact that KL divergence is always nonnegative.}:
-\begin{mdframed}
-\begin{equation}
-  \begin{split}
-    \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
-    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
-    &\le F^R_G(\x),
-    \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
-    \end{split}
-\end{equation}
-\end{mdframed}
-%% In fact, we have the following theorem
-%% \begin{theorem}
-%%   It holds that
-%%   $$
-%%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
-%%   i.e
-%%   \begin{equation}
-%%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
-%%     \end{equation}
-%% \end{theorem}
-
-\subsection{Helmholtz machines and the wake-sleep algorithm}
+% 
+% Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
+% distribution
+% \begin{equation}
+%   p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
+% \end{equation}
+% where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
+% Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
+% \begin{equation}
+% F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
+% \end{equation}
+% Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
+% it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
+% density\footnote{conditioning in $P_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
+%   state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
+% \begin{equation}
+%   \begin{split}
+%     F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{P_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
+%     &= \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
+%     &= F^R_G(\x) - D_{KL}(P_R(.|\x) || P_G(.|\x)),
+%   \end{split}
+%   \label{eq:fe}
+% \end{equation}
+% where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
+% \begin{equation}
+%   F^R_G(\x) := \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x))
+% \end{equation}
+% and $D_{KL}(P_R(.|\x) || P_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
+% 
+% \subsection{A general free-energy principle}
+% We can resume the situation as follows\footnote{Where we have used the fact that KL divergence is always nonnegative.}:
 % \begin{mdframed}
-  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e., input from lower more primitive layers).
+% \begin{equation}
+%   \begin{split}
+%     \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
+%     &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
+%     &\le F^R_G(\x),
+%     \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
+%     \end{split}
+% \end{equation}
+% \end{mdframed}
+% %% In fact, we have the following theorem
+% %% \begin{theorem}
+% %%   It holds that
+% %%   $$
+% %%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
+% %%   i.e
+% %%   \begin{equation}
+% %%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
+% %%     \end{equation}
+% %% \end{theorem}
+% 
+% \subsection{Helmholtz machines and the wake-sleep algorithm}
+% % \begin{mdframed}
+%   \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e., input from lower more primitive layers).
 Precisely
-$$
-p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
-p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
-$$
+% $$
+% p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
+% p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
+% $$
+% % \end{mdframed}
+% 
+% \subsection{Friston's active-inference and agency}
+% \label{sec:friston}
+% This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
+% 
+% The following critics can be made:
+% \begin{itemize}
+%   \item
+%     As noted by Dayan et al.
+(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
+%   \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
+%   \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
+%     technical difficulty in getting derivatives of loss function w.r.t
+%     recognition weights $\W^R$).
+%   \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
+%     \end{itemize}
+% 
+% 
+% \subsection{Minimizing free-energy via backprop: variational auto-encoders}
+% 
+% % \subsection{Free-energy minimization via variational auto-encoders}
+% \label{sec:vae}
+% Here, we present a way to alleviate some conceptual and computational issues with the
+% free-energy framework presented thus far, by using the recent \textit{variational auto-encoder} (VAE) theory \citep{kingma2013auto}. Define the data-dependent
+% auxiliary random function
+% \begin{eqnarray}
+%   f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
+% \end{eqnarray}
+% Then we can rewrite the
+% variational free-energy as
+% \begin{eqnarray*}
+%   \begin{split}
+%     F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
+%     &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
+%     &= -\langle f_{G,R}\rangle_{P_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
+%     \end{split}
+% \end{eqnarray*}
+% 
+% \begin{mdframed}
+%   \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
 % \end{mdframed}
-
-\subsection{Friston's active-inference and agency}
-\label{sec:friston}
-This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
-
-The following critics can be made:
-\begin{itemize}
-  \item
-    As noted by Dayan et al.
-(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
-  \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
-  \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
-    technical difficulty in getting derivatives of loss function w.r.t
-    recognition weights $\W^R$).
-  \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
-    \end{itemize}
-
-\subsection{Minimizing free-energy via backprop: variational auto-encoders}
-
-% \subsection{Free-energy minimization via variational auto-encoders}
-\label{sec:vae}
-Here, we present a way to alleviate some conceptual and computational issues with the
-free-energy framework presented thus far, by using the recent \textit{variational auto-encoder} (VAE) theory \citep{kingma2013auto}. Define the data-dependent
-auxiliary random function
-\begin{eqnarray}
-  f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
-\end{eqnarray}
-Then we can rewrite the
-variational free-energy as
-\begin{eqnarray*}
-  \begin{split}
-    F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
-    &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
-    &= -\langle f_{G,R}\rangle_{P_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
-    \end{split}
-\end{eqnarray*}
-
-\begin{mdframed}
-  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
-\end{mdframed}
-
-% \begin{figure}[!tb]
-%   \centering
-%   \includegraphics[width=.8\linewidth]{vae.pdf}
-%   \caption{Variational autoencoders...}
-%   \label{fig:vae}
-% \end{figure}
-
-\paragraph{Solution: the reparametrization trick.}
-\begin{itemize}
-\item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
-\item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
-  \begin{itemize}
-  \item results in a sample $\z \sim p_R(.|\x)$, from the correct
-    posterior
-    \end{itemize}
-\end{itemize}
-The mapping $g$ should be taught of as a ``blurring'' function which produces noisy versions $\z$,
-called \textit{sensations}, of the true world state $\x$.
-The result is a scheme for training DBNs via good-old backprop!
-Refer to Fig. vae.pdf.
-% \ref{fig:vae}.
-Some examples of the reparametrization trick for a number of
-choices of the posterior distribution are given in Tab. \ref{tab:rptrick}.
-
-\begin{table}[H]
-  \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
-         \hline
-         Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
-         Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
-         Student’s t, Logistic, Uniform, Triangular, \ldots \\ \hline
-         Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
+% 
+% % \begin{figure}[!tb]
+% %   \centering
+% %   \includegraphics[width=.8\linewidth]{vae.pdf}
+% %   \caption{Variational autoencoders...}
+% %   \label{fig:vae}
+% % \end{figure}
+% 
+% \paragraph{Solution: the reparametrization trick.}
+% \begin{itemize}
+% \item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
+% \item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
+%   \begin{itemize}
+%   \item results in a sample $\z \sim p_R(.|\x)$, from the correct
+%     posterior
+%     \end{itemize}
+% \end{itemize}
+% The mapping $g$ should be taught of as a ``blurring'' function which produces noisy versions $\z$,
+% called \textit{sensations}, of the true world state $\x$.
+% The result is a scheme for training DBNs via good-old backprop!
+% Refer to Fig. vae.pdf.
+% % \ref{fig:vae}.
+% Some examples of the reparametrization trick for a number of
+% choices of the posterior distribution are given in Tab. \ref{tab:rptrick}.
+% 
+% \begin{table}[H]
+%   \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
+%          \hline
+%          Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
+%          Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
+%          Student’s t, Logistic, Uniform, Triangular, \ldots \\ \hline
+%          Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
 \\ \hline
-         Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
-  \end{tabular}
-  \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
-  \label{tab:rptrick}
-\end{table}
+%          Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
+%   \end{tabular}
+%   \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
+%   \label{tab:rptrick}
+% \end{table}
 
@@
 
-\subsection{GANs and other likelihood-free methods}
-\ldots
+% \subsection{GANs and other likelihood-free methods}
+% \ldots
 

### 2017-06-11 paper/paper.tex
big overall pass

@@
 \subsection{Towards a formal account of default mode function: higher-order control of the organism}
-The human DMN is responsible for a large part of the baseline neural activity,
+The human DMN is responsible for extended parts of the baseline neural activity,
 which typically decreases when engaged in psychological experiments
@@
 simultaneously modulate perception-action cycles and
-maintain probabilistic contemplations
+entertain probabilistic contemplations
 across time, space, and content domains
@@
 Informed by outcomes of performed actions, neural computation reflected in
-DMN dynamics could be constantly adapted in feedback loops
-shaped by prediction error. \textbf{XXX: "could be constantly adapted in feedback loops shaped by prediction error" => "are constantly shaped by prediction error through feedback loops"}
+DMN dynamics could be constantly shaped by prediction error through feedback loops.
 Such an RL account of DMN function can naturally embed human behavior
@@
 may be solved by a stochastic approximation
-based on a brain implementation of Markov Chain MCMC (MCMC) sampling
+based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
 \citep{tenenbaum2011grow}.
@@
 \citep{raichle2005intrinsic}.
-This idea has been invigorated by
-empirical evidence from
+This idea is invigorated by empirical evidence from
 neuroscience experiments \citep{kording2004bayesian, fiser2004small}.
@@
 paving the way for their algorithmic interpretation
-in our formal account (next section).
+in our formal account (section 3).
 
@@
   Neurobiological overview of the DMN with its major constituent nodes and
-  the associated functional roles relevant in our functional account.}
+  the associated functional roles relevant in our functional interpretation.}
   \label{fig:neurobiological_overview_DMN}
@@
 \citep{balint1909seelenlahmung}.
-Patients are plagued by an
+Neurological patients are plagued by an
 inability to combine various individual features of the visual
@@
 information from memory recall and different perspective frames.
-Regarding memory recall, retrosplenial damage have been
+Regarding memory recall, retrosplenial damage has been
 consistently associated with anterograde and retrograde memory impairments
@@
 
-\subsection{The prefrontal cortex: stimulus-value association and action consideration}
+\subsection{The prefrontal cortex: action consideration and stimulus-value association}
 Analogous to the PMC,
@@
   default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11+/-0.02$) and right NAc volume ($R^2=0.14+/-0.02$) could be predicted from volume in the DMN nodes.
-These out-of-sample generalization performances were obtained from support vector regression applied to normalized node volumes in the DMN in a 10-fold cross-validation framework.
-Consistent for the left and right reward system, NAc volume in a given subject has an especially close relation to vmPFC and HC.
+These out-of-sample generalization performances were obtained from support vector regression applied to normalized node volumes in the DMN in a 10-fold cross-validation procedure.
+  Consistent for the left and right reward system, NAc volume in a given subject is
+  positively coupled with the vmPFC and HC.
   Code and data for reproduction and visualization:
-  % \url{www.github.com/banilo/darkcontrol_paper}.
+  \url{www.github.com/banilo/to_be_added_later}.
   }
@@
 whereas ventromedial PFC (vmPFC) activity
-processed especially value outcomes of actions performed on the environment
+processed especially value outcomes of actions actually performed on the environment
 \citep{nicolle2012agent}.
@@
 The ventromedial prefrontal DMN may
-orchestrate adaptive behavior by bottom-up-driven
+subserve adaptive behavior by bottom-up-driven
 processing of “what matters now”,
@@
 that not only channel action towards basic survival needs but also
-subserve reward processing, and thus perhaps RL,
-in humans more broadly \citep{doherty2015structure}.
+enable more abstract reward processings, and thus perhaps RL,
+in humans \citep{doherty2015structure}.
 Consistently, diffusion MRI tractography in humans and monkeys
@@
 Such a ``cognitive map'' of the action space was argued to encode
-the current task state even when states are unobservable from the sensory environment.
+the current task state even when states are unobservable from
+the sensory environment.
 
@@
 imagination of experiences more broadly \citep{hassabis2007patients}.
-Mental scenes created by patients with HC lesion exposed a lack of
+Mental scenes created by neurological patients with HC lesion exposed a lack of
 spatial integrity, richness in detail, and overall coherence.
@@
 Single-cell recordings in the animal HC revealed
-some constantly active neuronal populations whose firing coincided with
+constantly active neuronal populations whose firing coincided with
 specific locations in space during environmental navigation.
@@
 
-There is increasing evidence that
+There is hence increasing evidence that
 HC function extends beyond simple forms of
@@
 in the HC during sleep
-has been shown to subsequently alter action choice during wakefulness
+has been demonstrated to subsequently alter action choice during wakefulness
 \citep{lavilleon2015}.
@@
 The TPJs are known to exhibit hemispheric differences
-based on microanatomical properties and gyrification pattern
+based on microanatomical properties and gyrification patterns
 \citep{seghier2013angular}.
@@
 attentional versus semantic functions
-\citep{seghier2013angular, bzdok2013tpj, bzdok2016left}.
+\citep{seghier2013angular, bzdok2013tpj, bzdok2016left,
+stephan2007mechanisms}.
 
@@
 % (Bzdok et al., 2013; Corbetta et al., 2008).
-Indeed, patients with RTPJ damage have particular difficulties
+Indeed, neurological patients with RTPJ damage have particular difficulties
 with multi-step actions \citep{hartmann2005takes}.
@@
 \citep{binder2011neurobiology, seghier2013angular}.
-For instance, LTPJ damage in humans also entails problems in recognizing
+Consistent with this view,
+LTPJ damage in humans also entails problems in recognizing
 others' pantomimed action towards objects
@@
 episodic memory recall and mental imagery strongly draw on
-reassembling world knowledge.
+re-assembling world knowledge.
 Isolated building blocks of world structure probably get rebuilt
@@
 RL is a problem-solving technique in which,
-through repeated interactions with an\textit{environment},
+through repeated interactions with an \textit{environment},
 an agent learns to reach goals and optimize reward signals
@@
 \textit{partially observable} in the sense that only limited aspects of the environment's
-state are accessible to the agent's sensors.
+state are accessible to the agent's sensory perception.
 \citep{starkweather2017dopamine}.
@@
 \subsection{Markov decision processes}
-Model-free RL has had great success in real-world problems including robotics
-\citep{ng2004,abbeel2004},
-super-human performance in complex video games \citep{mnih2015} and Go\citep{silver2016mastering}, etc., just to name a few.
+Model-free RL has had great success in many real-world problems,
+including robotics \citep{ng2004,abbeel2004},
+super-human performance in complex video games \citep{mnih2015},
+and strategic board games \citep{silver2016mastering}.
 We emphasize that the brain in general, and the DMN in particular,
@@
 % , based on a set of observations and observation probabilities.
-This is a minimal set of assumptions that can be made about an environment,
-and is characteristic of so-called \textit{model-free RL}.
+This is a minimal set of assumptions that can be made about an environment.
 
@@
 quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
-cumulative reward estimate which takes into account expected reward
+\textit{cumulative reward} estimate which takes into account expected reward
 from action choices in the future.
 A common approach to
-modeling this accumulation is the time-discounted \textit{cumulative reward}% function
+modeling this accumulation is the time-discounted cumulative reward% function
 \begin{equation}
@@
   policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
-following a policy $\pi$.
+following an action policy $\pi$.
 %
@@
 This DMN node has direct connections to
-to the NAc, known to be involved in reward evaluation,
-and
-the amygdala, known to be involved in evaluation of biological significance.
+to the NAc, known to be involved in reward evaluation.
 
@@
 setting $\gamma = 0$ yields perfectly hedonistic behavior.
-An agent with such a shortsighted horizon is exclusively
+An agent with such a shortsighted time horizon is exclusively
 concerned with immediate rewards.
@@
 A positive $\gamma$ can be seen as calibrating risk-seeking trait of the intelligent agent,
-that is, the behavioral predispositions related to trading longer delays for higher rewards.
+that is, the behavioral predispositions related to trading longer delays
+for higher reward outcomes.
 Such an agent puts relatively more emphasis on rewards expected in
 a longer-term future.
-Precisely, rewards that are not expected to come within
+More specifically,
+rewards that are not expected to come within
 % \footnote{\textbf{N.B.:} $\frac{1}{1-\gamma} = 1 + \gamma + \gamma^2 + \gamma^3 + \ldots$}
@@
 utility of executing a specific action.
-Thus, it answers the question
+It thus answers the question
 ``What is the expected utility of taking action $a$ in this situation ?''.
@@
 The learning goal is to approach the
-policy $\pi^*$ as ``close as possible'', that is to \text{solve} the MDP.
+policy $\pi^*$ as close as possible, that is to \text{solve} the MDP.
 Note that \eqref{eq:qlearning} presents merely a definition and
@@
 Using the Bellman equation, each state can be associated with a certain value
-to guide action towards a preferred state, thus improving on the agent's current policy.
+to guide action towards a preferred state, thus improving on the 
+current action policy of the agent.
 % It is a fixed-point equation for the deterministic policy $\pi$.
@@
 deterministic policies of the form \eqref{eq:qlearning}. State spaces may be extremely large,
-and tracking all possible states and actions may require excessive space.
+and tracking all possible states and actions may require excessive
+computation and memory resources.
 The need of maintaining an explicit table of
@@
 by keeping track of an approximating parameter $\theta$ of much lower dimension than the number of states.
-At a time step, the world is in a state $s \in \mathcal S$, and the agent takes an
+At a given time step, the world is in a state $s \in \mathcal S$, and the agent takes an
 action which it expects to be the most valuable on average, namely
@@
 $\phi(s,a)$ would represent a high-level representation of the state-action pairs
-$(s,a)$, as was done proposed \citep{songNIPS2016}, or a eural network as recently proposed by in the
-breakthrough works
-\citep{mnih2015,silver2016mastering} for playing complex games (atari, Go, etc.) at superhuman levels.
+$(s,a)$, as was previously proposed \citep{songNIPS2016},
+or artificial neural-network models as demonstrated in recent
+seminal investigations
+\citep{mnih2015,silver2016mastering} for playing complex games (atari, Go, etc.) at super-human levels.
 %  and Parr et al. 2003 (JMLR)].
@@
 The DMN, and in particular its dmPFC node, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes these approximate
-$Q$-values. Such a mapping from states to actions is conventionally called \textit{policy matrix}\citep{mnih2015,silver2016mastering}.
+$Q$-values.
+It is this mapping from states to actions that is conventionally called
+policy matrix \citep{mnih2015,silver2016mastering}.
 Learning consists in starting from a given table and
@@
 which take the agent to different table entries.
+
 An additional layer of learning concerns the addition of new entries in the table (i.e., state and action).
@@
 Each cycle represents a behavior unit of the agent
-and is recorded in replay memory buffer \textemdash which we hypothesize to be the
+and is recorded in replay memory buffer \textemdash
+which we hypothesize to be subserved by the
 HC \textemdash, possibly discarding the oldest entries to make space:
@@
 experience replay can be manifested as the re-occurrence of
-neuron spiking sequences that have also occurred during specific prior actions.
+neuron spiking sequences that have also occurred during specific prior actions
+and environmental states.
 %, but the replay has a much faster time scale.
-The HC is a strong candidate to contribute to such a mechanism,
-as neuroscience experiments have repeatedly indicated
+The HC is a strong candidate to contribute to such a mechanism
+because neuroscience experiments have repeatedly indicated
 in rats, mice, cats, rabbits, songbirds, and
@@
 \citep{mnih2015,silver2016mastering} is to learn this
-representation using a deep neural network, leading to the
+representation using a deep neural-network model, leading to the
 so-called \textit{deep Q-learning} family of methods which
@@
 local minima based on the first derivate of the optimization objective.
-Stochastic here means that the true gradient is estimated from batches of training examples samples (in our case blocks of experience from the replay memory):
+Stochastic here means that the true gradient is estimated from batches of training examples samples,
+which, in our case, corresponds to blocks of experience from the replay memory:
 
@@
 the HC could contribute to synthesizing imagined sequences of world states,
-observations and rewards \citep{aronov2017, chao2017interaction, boyer2008evolutionary}.
+actions and rewards \citep{aronov2017, chao2017interaction, boyer2008evolutionary}.
 These simulations of experience batches
@@
   action and state cascades as a function of what happened in the past.
-  A policy encapsulates the repertoire of possible actions
+  A policy matrix encapsulates the repertoire of possible actions
     on the world given a current state.
@@
 The DMN may subserve
-  constant exploration of possible future actions and estimaton of their
+  constant exploration of possible action trajectories and
+  nested estimaton of their
   cumulative reward outcomes. Implicit computation of future choices
@@
   neurobiological implementation of reinforcement learning.}
-  Overview of how the constituent nodes of the DMN
+  Overview of how the constituent nodes of the DMN may
   map onto computational components necessary for an RL agent.}
@@
 
-The HC may provide perturbed action-transition-state-reward samples
+The HC may contribute to generation of
+perturbed action-transition-state-reward samples
 as batches of pseudo-experience
@@
 
-  From the perspective of our model-free RL agent,
+From the perspective of a model-free RL agent,
 \textit{inference} in the human brain reduces to
 generalization of
-  policy and value adaptations from sampled experiences to
+policy and value computations from sampled experiences to
 successful action choices and reward predictions in future states.
 As such,
-  plasticity in the DMN arises naturally:
-  if an agent behaving optimally in a given environment moves
-  to new, yet unexperienced environments, reward prediction errors will
-  massively increase \textbf{Please check/simplify sentence}.
-  This will lead to adaptation of the policy until the system converges to a
-  new steady-state of optimal action decisions in the volatile environment.
+plasticity in the DMN arises naturally.
+If an agent behaving optimally in a certain environment moves
+to new, yet unexperienced environment, reward prediction errors will
+largely increase.
+This feedback will lead to adaptation of policy considerations
+and value estimations until the intelligent system converges to a
+new steady state of optimal action decisions in a volatile world.
 
@@
 
-\section{Relation to existing functional accounts}
+\section{Relation to existing accounts}
 \subsection{Predictive coding hypothesis}
@@
 The dynamic interplay between cortical processing levels
-may enable inferences about the world by reconciling
+may enable learning about world aspects by reconciling
 gaps between fresh sensory input and expectations computed
@@
 between the predictive coding account
-and the proposed formal account based on MDPs.
+and the proposed formal account of DMN function based on MDPs.
 Importantly,
@@
 \citep{salin1995corticocortical}.
-These many and diverse modulatory influences from higher cortical areas
+These many and diverse modulatory influences
+from higher onto downstream cortical areas
 can inject prior knowledge
 at every stage of processing environmental information.
+%
 Moreover,
 both accounts provide a parsimonious explanation why the
-human brain decreases processing load devoted to incoming information
-when the environment is predictable.
+human brain's processing load devoted to incoming information decreases
+when the environment becomes predictable.
 This is because the internal generative
@@
 unexpected events are encountered by the organism.
-The predictive coding and MDP account
+The predictive coding and MDP account hence
 naturally evoke a mechanism of brain plasticity in that
@@
 the predictive coding account is typically explained without
-explicit parameterization of the agent's horizon and
+explicit parameterization of the agent's time horizon and
 has a tendency to be presented as emphasizing prediction about the
@@
 In sum,
-the present MDP account of DMN function may serve as a concrete implementation of
-the predictive coding account from cognitive neuroscience.
+recasting DMN function in terms of MDPs therefore naturally incorporates
+a majority of aspects from the prediction coding hypothesis.
+The present MDP account of DMN function may therefore
+serve as a concrete implementation of
+predictive coding ideas from cognitive neuroscience.
 MDPs have the advantage of exposing an explicit
@@
 as well as
-more explicitly reconciles sensory input processing and action considerations.
-%
-Recasting DMN function in terms of MDPs therefore naturally incorporates
-a majority of aspects from the prediction coding hypothesis.
+why certain predictions may be more relevant to the agent than others.
 
@@
 The encoding of complex environmental features could thus be facilitated
-by association to similar states.
+by association to known similar states.
 %
@@
 semantic building blocks may need to get recombined to enable
-mentally envisioning non-existing scenarios.
+mental imagery of non-existing scenarios.
 As such, semantic knowledge would be a prerequisite for optimizing behavior
@@
 lead to prediction of future actions and states.
-The semantic hypothesis does not explain why memory recall
+The semantic hypothesis does also not explain why memory recall
 for scene construction in humans is typically fragmentary and noisy
@@
 Our MDP account provides an algorithmic explanation in that
-stochasticity of the parameter space explored
+stochasticity of the parameter space exploration
 by MCMC approximation achieves better fine-tuning of the
@@
 account is lacking an explicit time and incentive model,
-both of which are integral parts of the MDP frameworks.
+both of which are integral parts of the MDP interpretation of DMN function.
 
@@
 
-More specifically,
+In particular,
 environmental stimuli especially important for humans are frequently of
@@
 
-The principal connectivity gradient in the cortex appears to be greatly expanded in humans compare to macaques, suggesting a phylogenetically conserved axis of cortical expansion with the DMN emerging at the extreme-end in humans \citep{margulies_situating_2016}.
-Neurocomputational models of dyadic whole brain demonstrated how the human connectivity topology, on top of facilitating processing at the intra-individual level, can explain our propensity to coordinate through sensorimotor loops with others at the inter-individual level \citep{dumas_anatomical_2012}.
-DMN is moreover largely overlapping with neural networks associated with higher level social cognition \citep{schilbach_introspective_2012}.
+% Further,
+% the principal connectivity gradient in the cortex appears to be greatly expanded in humans compared to monkeys, suggesting a phylogenetically conserved axis of cortical expansion with the DMN emerging at the extreme end in humans \citep{margulies_situating_2016}.
+Neurocomputational models of dyadic whole-brain dynamics
+% demonstrated how the human connectivity topology, on top of facilitating processing at the intra-individual level, can explain our propensity to coordinate through sensorimotor loops with others at the inter-individual level \citep{dumas_anatomical_2012}.
+% The DMN is moreover largely overlapping with neural networks associated with higher level social cognition \citep{schilbach_introspective_2012}.
 For instance, the ventromedial prefrontal cortex, the posterior cingulate/precuneus, and the temporo-parietal junction (rTPJ) play together a key role in bridging the gap between self and other by
-integrating low-level embodied processes within higher level inference-based mentalizing \citep{lombardo_shared_2009}.
+% integrating low-level embodied processes within higher level inference-based mentalizing \citep{lombardo_shared_2009}.
 
@@
 %
-This functional interpretation may be more compatible with the DMN's
+This functional interpretation may be well compatible with the DMN's
 poorly understood involvement across
@@
 
-The proposed artificial intelligence perspective
-on DMN biology is hence practically computable and
+It is an important feature of the proposed artificial intelligence perspective
+on DMN biology that it is practically computable and
 yields falsifiable neuroscientific hypotheses.
@@
 At the least, we propose an alternative vocabulary to
-describe and interpret experimental findings in neuroscience studies
-on the DMN situated at top of the brain network hierarchy.
+describe, contextualize, and interpret experimental findings in neuroscience studies
+on the DMN.
 %
@@
 
-\paragraph{Acknowledgment.}
+% \paragraph{Acknowledgment.}
 % {\small The research leading to these results has received funding from the

### 2017-06-11 paper/paper.tex
finished my pass

@@
 super-human performance in complex video games \citep{mnih2015},
-and strategic board games \citep{silver2016mastering}.
+and strategic board games like the breakthrough results upon recently on the game of Go \citep{silver2016mastering}, considered to be a golden benchmark problem in artificial intelligence.
 We emphasize that the brain in general, and the DMN in particular,
@@
 Instead, the agent tries to optimize a \textit{subjective} reward
-signal (i.e., is likely to be different for another agent in the same state).
-This is performed alongside each transition in the state space of the environment
- and maintains a probability distribution over the set of possible
-states.
-% , based on a set of observations and observation probabilities.
-This is a minimal set of assumptions that can be made about an environment.
+signal (i.e., is likely to be different for another agent in the same state),
+by maintaining probability distributions over actions according to their expected utility.
+% This is performed alongside each transition in the state space of the environment
+%  and maintains a probability distribution over the set of possible
+% states.
+% % , based on a set of observations and observation probabilities.
+This is a minimal set of assumptions that can be made about an environment faced by an agent engaged in active learning.
 
 Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
-a core property of human intelligence has been proposed to underlie improvement
+it has been proposed \citep{gershman2015computational} that a core property of human intelligence  underlie improvement
 of expected utility outcomes as a strategy for action choice in uncertain
-environments \citep{gershman2015computational}, a situation perfectly captured by the formalism of MDPs.
+environments, a situation perfectly captured by the formalism of MDPs.
 It has also long been proposed \citep{dayan2008decision} that there
@@
 to be equivalent over sufficiently long time windows of equal length (i.e., stationarity).
-
 Since an action executed in the present moment might have repercussions in the far future, it turns out that the
@@
 deterministic policies of the form \eqref{eq:qlearning}. State spaces may be extremely large,
-and tracking all possible states and actions may require excessive
+and tracking all possible states and actions may require prohibitively excessive
 computation and memory resources.
@@
 The model
-parameters $\theta$ would correspond to synaptic weights and connection strengths between
+parameters $\theta$ would correspond to synaptic weights and connection strengths within and between
 brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the hard-max policy \eqref{eq:policy}.
@@
 $Q$-values.
-It is this mapping from states to actions that is conventionally called
-policy matrix \citep{mnih2015,silver2016mastering}.
+This mapping from states to actions that is conventionally called
+\textit{policy matrix} \citep{mnih2015,silver2016mastering}.
 Learning consists in starting from a given table and
@@
 $\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
-
 At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1} + \delta \theta_{k}$ of the parameters for its approximate model of the $Q$-value function.
@@
 Efficient learning of the entire set of model parameters can effectively be achieved
-via stochastic \textit{gradient descent}, a common solver algorithm for finding
-local minima based on the first derivate of the optimization objective.
+via stochastic \textit{gradient descent}, a universal algorithm for finding
+local minima based on the first derivative of the optimization objective.
 Stochastic here means that the true gradient is estimated from batches of training examples samples,
@@
 to percolate from lower to higher processing layers to modulate the
-choice of future actions. It is a standard result that under special conditions on the learning rates $\alpha_k$
-(namely that the sum $\sum_{k=0}^\infty\alpha_k$ diverges while $\sum_{k=0}^\infty\alpha_k^2$ converges) over time (i.e., in the limit $k \rightarrow \infty$)
+choice of future actions.
+It is a standard result that under special conditions on the learning rates $\alpha_k$ --namely that the learning rates are neither too large nor too small, or more precisely that the sum $\sum_{k=0}^\infty\alpha_k$ diverges while $\sum_{k=0}^\infty\alpha_k^2$--
 the thus generated approximating sequence of $Q$-value functions
@@
 
-Since theories based on the FEP \citep{friston2010free,fristonAIorRL}
+On another front, since theories based on the FEP \citep{friston2010free,fristonAIorRL}
 conceptualize ongoing behavior
@@
 seldom observed in animals or humans in the real world: In a dark room,
-\begin{wrapfigure}{r}{.45\linewidth}
-  \centering
-  \includegraphics[width=1.\linewidth]{darkroom.jpg}
+intelligent agents would search for light sources to \textit{explore}
+its surroundings or leave it.
+Defenders of the FEP have retorted by advancing the ``full package'' \citep{darkroom2012}:
+FEP is proposed to be multi-scale and there would be a meta-scale at which the organism would
+be \textit{surprised} by such a \textit{lack of surprise}. According to this argument, a dark room would paradoxically correspond to a state of particularly high surprise.
+\begin{wrapfigure}{r}{.4\linewidth}
   \caption{\textbf{The dark room experiment.} An intelligent agent situated
@@
   \label{fig:darkroom}  
+  \centering
+  \includegraphics[width=1.\linewidth]{darkroom_resized.png}
 \end{wrapfigure}
-intelligent agents would search for light sources to \textit{explore}
-its surroundings or leave it.
-Defenders of the FEP have retorted by advancing the ``full package'' \citep{darkroom2012}:
-FEP is proposed to be multi-scale and there would be a meta-scale at which the organism would
-be \textit{surprised} by such a \textit{lack of surprise}. According to this argument, a dark room would paradoxically correspond to a state of particularly high surprise.
 Driven by surprise-minimization objective, the FEP agent would eventually bootstrap itself out of such saddle points to explore more interesting parts of the environment.
@@
 our model-free RL framework.
-This becomes possible by recasting the reward signal maximized by the DMN to 
+This becomes possible by recasting the Q-value function (i.e expected long-term reward) maximized by the DMN to 
 correspond to negative surprise,
@@
 
-\begin{equation}
-  r^{\text{FEP}} = \underbrace{\log (p_G)}_{-\text{FEP generative surprise}}.
-    \end{equation}
+\begin{eqnarray*}
+  -Q^{\text{FEP}} \sim \underbrace{\log (p_G)}_{\text{FEP generative surprise}}.
+    \end{eqnarray*}
 % \paragraph{Comparison to our proposed theory.}
@@
 Nevertheless, minimization of surprise quantities alone
-may be insufficient to explain
-the diversity of behaviors observed in humans and other intelligent animals.
+may be insufficient to explain the diversity of behaviors observed in humans and other intelligent animals.
 % The dark-room experiment, for example, provides a formidable argument against in favor of the insufficiency of the FEP and active inference theories to explain intelligent behavior in humans: when a human is put in a dark room, they don't just sit down in a corner forever, what would be the FEP-optimal thing to do; they actively search for the light switch, and start exploring the depths of the room, possibly looking for food or the way out!

### 2017-06-11 paper/paper.tex
final touches

@@
 
+\pagebreak[2]
 \begin{mdframed}
+\hspace{2mm}
+\vspace{2mm}
 % neuroscience of Bellman equation
@@
 
-% Further,
-% the principal connectivity gradient in the cortex appears to be greatly expanded in humans compared to monkeys, suggesting a phylogenetically conserved axis of cortical expansion with the DMN emerging at the extreme end in humans \citep{margulies_situating_2016}.
+Further,
+the principal connectivity gradient in the cortex appears to be greatly expanded in humans compared to monkeys, suggesting a phylogenetically conserved axis of cortical expansion with the DMN emerging at the extreme end in humans \citep{margulies_situating_2016}.
 Neurocomputational models of dyadic whole-brain dynamics
-% demonstrated how the human connectivity topology, on top of facilitating processing at the intra-individual level, can explain our propensity to coordinate through sensorimotor loops with others at the inter-individual level \citep{dumas_anatomical_2012}.
-% The DMN is moreover largely overlapping with neural networks associated with higher level social cognition \citep{schilbach_introspective_2012}.
-For instance, the ventromedial prefrontal cortex, the posterior cingulate/precuneus, and the temporo-parietal junction (rTPJ) play together a key role in bridging the gap between self and other by
-% integrating low-level embodied processes within higher level inference-based mentalizing \citep{lombardo_shared_2009}.
+demonstrated how the human connectivity topology, on top of facilitating processing at the intra-individual level, can explain our propensity to coordinate through sensorimotor loops with others at the inter-individual level \citep{dumas_anatomical_2012}.
+The DMN is moreover largely overlapping with neural networks associated with higher level social cognition \citep{schilbach_introspective_2012}. For instance, the vmPFC,
+PMC, and RTPJ
+together play a key role in bridging the gap between self and other by
+integrating low-level embodied processes within higher level inference-based mentalizing \citep{lombardo_shared_2009}.
 
@@
   \centering
-  \includegraphics[width=1.\linewidth]{darkroom_resized.png}
+  \includegraphics[width=1.\linewidth]{darkroom.jpg}
 \end{wrapfigure}

### 2017-06-11 paper/paper.tex
minor style tweaks

@@
 %
+Such highly abstract neural computations necessarily rely on the
+construction of probabilistic internal information drawing from
+memory recall, mental scene elaboration processes,
+and stored knowledge of world regularities.
 \begin{wrapfigure}{r}{.4\linewidth}
@@
 \end{wrapfigure}
-Such highly abstract neural computations necessarily rely on the
-construction of probabilistic internal information drawing from
-memory recall, mental scene elaboration processes,
-and stored knowledge of world regularities.
 %
@@
 
-\pagebreak[2]
+\pagebreak[4]
 \begin{mdframed}
@@
 because its sensory inputs are perfectly predictable
+\begin{wrapfigure}{r}{.4\linewidth}
+  \caption{\textbf{The dark room experiment.} An intelligent agent situated
+  in a light-deprived closed space can be used
+  as a thought experiment for the complete absence
+  of external sensory input.}
+  \label{fig:darkroom}  
+  \centering
+  \includegraphics[width=1.\linewidth]{darkroom_resized.png}
+\end{wrapfigure}
 given the environmental state \citep{darkroom2012}.
@@
 be \textit{surprised} by such a \textit{lack of surprise}. According to this argument, a dark room would paradoxically correspond to a state of particularly high surprise.
-\begin{wrapfigure}{r}{.4\linewidth}
-  \caption{\textbf{The dark room experiment.} An intelligent agent situated
-  in a light-deprived closed space can be used
-  as a thought experiment for the complete absence
-  of external sensory input.}
-  \label{fig:darkroom}  
-  \centering
-  \includegraphics[width=1.\linewidth]{darkroom.jpg}
-\end{wrapfigure}
 Driven by surprise-minimization objective, the FEP agent would eventually bootstrap itself out of such saddle points to explore more interesting parts of the environment.

### 2017-06-20 paper/paper.tex
correct latex typesetting form plus-or-minus

--- 2017-06-11 paper/paper.tex
@@
   \caption{\textbf{Morphological coupling between reward system and
-  default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11+/-0.02$) and right NAc volume ($R^2=0.14+/-0.02$) could be predicted from volume in the DMN nodes.
+  default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from volume in the DMN nodes.
 These out-of-sample generalization performances were obtained from support vector regression applied to normalized node volumes in the DMN in a 10-fold cross-validation procedure.

### 2017-06-20 paper/paper.tex
a small pass on the abstract

@@
 
-The present paper proposes a \textit{process model}
-that tries to explain \textit{how}
-the DMN may implement
+In the present manuscript, we propose a \textit{process model}
+that attempts to explain \textit{how} the DMN may implement
 continuous evaluation and prediction of the environment to guide behavior.
 %
-DMN function is recast in mathematical terms of
+We first review the rich literature on the DMN and recognize that the DMN contains the sufficient neuro-biological substrates to implement a full-blown reinforcement-learning agent, at least in principle.
+DMN function is then recast in mathematical terms of
 control theory and reinforcement learning based on Markov decision processes.

### 2017-06-27 paper/paper.tex
more passes

@@
 through repeated interactions with an \textit{environment},
-an agent learns to reach goals and optimize reward signals
-in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
+an agent learns to reach goals and optimize reward signals occasionally emitted by the environment, in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
 At a given moment, each taken \textit{action} $a$ triggers a change
@@
 of as satisfaction \textemdash or aversion \textemdash accompanying the execution of
-a particular action.
+a particular action.  Thus the agent learns behaviors.
 % Action may be delayed to achieve substantial increases in expected reward
@@
 The optimal Q-value operator $Q^*$ is a fixed point for this equation.
-As a consequence of this decomposition, the complicated dynamic programming
-problem \eqref{eq:qlearning}
+As a consequence of this decomposition, the complicated problem \eqref{eq:qlearning}
 is broken down into simpler sub-problems at different time points.
-Indeed,
-exploitation of hierarchical structure in action considerations
+Indeed, exploitation of hierarchical structure in action considerations
 has previously been related to the medial prefrontal part of the DMN
@@
 for minimizing free energy, VAEs can be efficiently trained via back-propagation of prediction errors.
+Most importantly, the reparametrization trick allows for a drastic reduction in the variance of the in estimates of the model's gradients, leading to effective learning, something which was not possible in wake-sleep\citep{dayan1995helmholtz}.
+The details of this trick can be found in \citep{kingma2013auto}.
 

### 2017-07-01 paper/paper.tex
integrate conversation with David Smith

--- 2017-06-20 paper/paper.tex
@@
 
-Consequently, the PMC may contribute to overall DMN function
-by monitoring the subjective outcomes
-of possible actions and integrating that information
+Consequently, the PMC may contribute to overall DMN function by
+``altering behavior in response to unexpected change''
+\citep{pearson2011posterior}.
+This DMN node probably
+monitors sudden and gradual environmental changes in subjective outcomes
+of action choices for integration
 with memory and perspective frames
-into short- and longer-term behavioral agendas.
+to improve short- and longer-term behavioral agendas.
 Estimated value, that differs across individuals, probably enriches
@@
   \caption{\textbf{Morphological coupling between reward system and
-  default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from volume in the DMN nodes.
+  default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11+/-0.02$) and right NAc volume ($R^2=0.14+/-0.02$) could be predicted from volume in the DMN nodes.
 These out-of-sample generalization performances were obtained from support vector regression applied to normalized node volumes in the DMN in a 10-fold cross-validation procedure.

### 2017-07-05 paper/paper.tex
some fixups + uncommented supp. materials

--- 2017-06-27 paper/paper.tex
@@
 Instead, FEP therefore minimizes an upper-bound of surprise,
-namely the \text{free energy} $F_G^R(\x)$ given by
+namely the \text{free energy} $F_G^R(\x) := \langle \log p_G(., \x) \rangle_{p_R(. |\x)} - \langle \log p_R(. |\x))\rangle_{p_R(. |\x)}$ given by
 
@@
       &\le F^R_G(\x),
-      \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
+      \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z.
     \end{split}
@@
 
-% \appendix
-% 
-% \begin{remark}
-%   Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
-%   except the Markov property, we can always transform it into an MDP by considering a compounded
-%   version of the states
-%   $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
-%   history up to an including time $t$.
-% \end{remark}
-% 
-% \section{More on RL in the brain}
-% \subsection{Does hippocampal replay equal inverse reinforcement-learning?}
-% %\begin{mdframed}
-%   Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
-%   an MDP (called a \textit{teacher's demonstration}), can we figure out what is the
-%   (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
-%   \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
-%   (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
-%   For instance, given traces of motor actions from an adult to grab a cup on a table,
-%   can an observing child figure out what ``langragian'' functional is being minimized by the former?
-%   How? Can they reproduce this optimal behavior?
-% %\end{mdframed}
-% Such questions are of course pertinent to our decision-making theory for the DMN.
-% In the general artificial case, the problem has been extensively studied and partially solved by
-% \citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
-% ``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
-% 
-% IRL is is suited for problems in which it's hard to define what the reward function could be
-% (e.g.,  car-driving, drone maneuvering, etc.) ...
-% 
-% \section{Free-energy principles!}
-% The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
-% In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
+\appendix
+
+\begin{remark}
+  Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
+  except the Markov property, we can always transform it into an MDP by considering a compounded
+  version of the states
+  $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
+  history up to an including time $t$.
+\end{remark}
+
+\section{Free-energy principles!}
+The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
+In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
 This ideas were first developed by Hinton et al.
@@
 For instance, the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
-% 
-% % \begin{mdframed}
-% %   Though our presentation of free-energy principles is thermodynamic in flavor, a completely equivalent
-% %   information-theoretic formulation is of course possible, simply by exploiting the information-theoretic interpretation of entropy.
-% % \end{mdframed}
-% 
-% \subsection{Helmholtz free-energy and the generative model}
-% \begin{table}[H]
-%   \begin{tabular}{p{2cm}|p{11cm}}
-%          \hline
-%          \textbf{symbol}    & \textbf{description}  \\ \hline
-%          $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
-%          random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
-%          $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
-%           with the usual convention $0 \log(0) := 0$.\\ \hline
-%          $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
-%              $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e., $\x = (s, b).$ \\ \hline
-%              $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
-%              $p_G(.|\x)$ & Generative density for ...\\ \hline
-%          $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
-%          $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
-%          $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
-%           to $R$.  Note that $F^G_G = F_G$.\\ \hline
-%   \end{tabular}
-%   \caption{Table of notations.}
-% \end{table}
-% Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
+
+% \begin{mdframed}
+%   Though our presentation of free-energy principles is thermodynamic in flavor, a completely equivalent
+%   information-theoretic formulation is of course possible, simply by exploiting the information-theoretic interpretation of entropy.
+% \end{mdframed}
+
+\subsection{Helmholtz free-energy and the generative model}
+\begin{table}[H]
+  \begin{tabular}{p{2cm}|p{11cm}}
+         \hline
+         \textbf{symbol}    & \textbf{description}  \\ \hline
+         $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
+         random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
+         $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
+          with the usual convention $0 \log(0) := 0$.\\ \hline
+         $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
+             $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e., $\x = (s, b).$ \\ \hline
+             $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
+             $p_G(.|\x)$ & Generative density for ...\\ \hline
+         $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
+         $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
+         $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
+          to $R$.  Note that $F^G_G = F_G$.\\ \hline
+  \end{tabular}
+  \caption{Table of notations.}
+\end{table}
+Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
 So,
-% \begin{eqnarray}
-%   \begin{split}
-%     \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(P_G(\x))\sum_{\z}p_G(\z |\x)\\
-%     &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
-%     =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
-%     &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
-%     &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
-%     &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
-%   \end{split}
-%   \label{eq:helm}
-% \end{eqnarray}
-% where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
-% \begin{equation}
-%   E_G(\z, \x) := -\log(p_G(\z, \x)),
-%   \label{eq:gibbs}
-% \end{equation}
-% %% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
-% The last quantity in \eqref{eq:helm} is nothing but \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
-% \begin{equation}
-%   F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
-% \end{equation}
-% %% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
-% Thus,
-% \begin{fact}
-%   Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
-% \end{fact}
-% 
-% The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
+\begin{eqnarray}
+  \begin{split}
+    \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(P_G(\x))\sum_{\z}p_G(\z |\x)\\
+    &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
+    =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
+    &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
+    &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
+    &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
+  \end{split}
+  \label{eq:helm}
+\end{eqnarray}
+where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
+\begin{equation}
+  E_G(\z, \x) := -\log(p_G(\z, \x)),
+  \label{eq:gibbs}
+\end{equation}
+%% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
+The last quantity in \eqref{eq:helm} is nothing but \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
+\begin{equation}
+  F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
+\end{equation}
+%% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
+Thus,
+\begin{fact}
+  Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
+\end{fact}
+
+The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
 It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(\x)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
@@
 1995, is at the heart of the general free-energy principle that we will introduce shortly.
-% 
-% \subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
-% In this subsection, we will present an insightful upper bound for the generative surprise (i.e., generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
+
+\subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
+In this subsection, we will present an insightful upper bound for the generative surprise (i.e., generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
 As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
-% 
-% Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
-% distribution
-% \begin{equation}
-%   p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
-% \end{equation}
-% where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
-% Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
-% \begin{equation}
-% F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
-% \end{equation}
-% Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
-% it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
-% density\footnote{conditioning in $P_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
-%   state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
-% \begin{equation}
-%   \begin{split}
-%     F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{P_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
-%     &= \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
-%     &= F^R_G(\x) - D_{KL}(P_R(.|\x) || P_G(.|\x)),
-%   \end{split}
-%   \label{eq:fe}
-% \end{equation}
-% where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
-% \begin{equation}
-%   F^R_G(\x) := \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x))
-% \end{equation}
-% and $D_{KL}(P_R(.|\x) || P_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
-% 
-% \subsection{A general free-energy principle}
-% We can resume the situation as follows\footnote{Where we have used the fact that KL divergence is always nonnegative.}:
+
+Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
+distribution
+\begin{equation}
+  p_G(\z|\x) = \frac{p_G(\z,\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
+\end{equation}
+where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
+Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
+\begin{equation}
+F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
+\end{equation}
+Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
+it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
+density\footnote{conditioning in $P_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
+  state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
+\begin{equation}
+  \begin{split}
+    F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{P_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
+    &= \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
+    &= F^R_G(\x) - D_{KL}(P_R(.|\x) || P_G(.|\x)),
+  \end{split}
+  \label{eq:fe}
+\end{equation}
+where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
+\begin{equation}
+  F^R_G(\x) := \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x))
+\end{equation}
+and $D_{KL}(P_R(.|\x) || P_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
+
+\subsection{A general free-energy principle}
+We can resume the situation as follows\footnote{Where we have used the fact that KL divergence is always nonnegative.}:
+\begin{mdframed}
+\begin{equation}
+  \begin{split}
+    \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
+    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
+    &\le F^R_G(\x),
+    \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
+    \end{split}
+\end{equation}
+\end{mdframed}
+%% In fact, we have the following theorem
+%% \begin{theorem}
+%%   It holds that
+%%   $$
+%%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
+%%   i.e
+%%   \begin{equation}
+%%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
+%%     \end{equation}
+%% \end{theorem}
+
+\subsection{Helmholtz machines and the wake-sleep algorithm}
 % \begin{mdframed}
-% \begin{equation}
-%   \begin{split}
-%     \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
-%     &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
-%     &\le F^R_G(\x),
-%     \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
-%     \end{split}
-% \end{equation}
+  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e., input from lower more primitive layers).
+Precisely
+$$
+p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
+p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
+$$
 % \end{mdframed}
-% %% In fact, we have the following theorem
-% %% \begin{theorem}
-% %%   It holds that
-% %%   $$
-% %%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
-% %%   i.e
-% %%   \begin{equation}
-% %%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
-% %%     \end{equation}
-% %% \end{theorem}
-% 
-% \subsection{Helmholtz machines and the wake-sleep algorithm}
-% % \begin{mdframed}
-%   \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e., input from lower more primitive layers).
-Precisely
-% $$
-% p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
-% p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
-% $$
-% % \end{mdframed}
-% 
-% \subsection{Friston's active-inference and agency}
-% \label{sec:friston}
-% This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
-% 
-% The following critics can be made:
-% \begin{itemize}
-%   \item
-%     As noted by Dayan et al.
+
+\subsection{Friston's active-inference and agency}
+\label{sec:friston}
+This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
+
+The following critics can be made:
+\begin{itemize}
+  \item
+    As noted by Dayan et al.
 (\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
-%   \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
-%   \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
-%     technical difficulty in getting derivatives of loss function w.r.t
-%     recognition weights $\W^R$).
-%   \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
-%     \end{itemize}
-% 
-% 
-% \subsection{Minimizing free-energy via backprop: variational auto-encoders}
-% 
-% % \subsection{Free-energy minimization via variational auto-encoders}
-% \label{sec:vae}
-% Here, we present a way to alleviate some conceptual and computational issues with the
-% free-energy framework presented thus far, by using the recent \textit{variational auto-encoder} (VAE) theory \citep{kingma2013auto}. Define the data-dependent
-% auxiliary random function
-% \begin{eqnarray}
-%   f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
-% \end{eqnarray}
-% Then we can rewrite the
-% variational free-energy as
-% \begin{eqnarray*}
-%   \begin{split}
-%     F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
-%     &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
-%     &= -\langle f_{G,R}\rangle_{P_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
-%     \end{split}
-% \end{eqnarray*}
-% 
-% \begin{mdframed}
-%   \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
-% \end{mdframed}
-% 
-% % \begin{figure}[!tb]
-% %   \centering
-% %   \includegraphics[width=.8\linewidth]{vae.pdf}
-% %   \caption{Variational autoencoders...}
-% %   \label{fig:vae}
-% % \end{figure}
-% 
-% \paragraph{Solution: the reparametrization trick.}
-% \begin{itemize}
-% \item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
-% \item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
-%   \begin{itemize}
-%   \item results in a sample $\z \sim p_R(.|\x)$, from the correct
-%     posterior
-%     \end{itemize}
-% \end{itemize}
-% The mapping $g$ should be taught of as a ``blurring'' function which produces noisy versions $\z$,
-% called \textit{sensations}, of the true world state $\x$.
-% The result is a scheme for training DBNs via good-old backprop!
-% Refer to Fig. vae.pdf.
-% % \ref{fig:vae}.
-% Some examples of the reparametrization trick for a number of
-% choices of the posterior distribution are given in Tab. \ref{tab:rptrick}.
-% 
-% \begin{table}[H]
-%   \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
-%          \hline
-%          Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
-%          Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
-%          Student’s t, Logistic, Uniform, Triangular, \ldots \\ \hline
-%          Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
+  \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
+  \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
+    technical difficulty in getting derivatives of loss function w.r.t
+    recognition weights $\W^R$).
+  \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
+    \end{itemize}
+
+\subsection{Minimizing free-energy via backprop: variational auto-encoders}
+
+% \subsection{Free-energy minimization via variational auto-encoders}
+\label{sec:vae}
+Here, we present a way to alleviate some conceptual and computational issues with the
+free-energy framework presented thus far, by using the recent \textit{variational auto-encoder} (VAE) theory \citep{kingma2013auto}. Define the data-dependent
+auxiliary random function
+\begin{eqnarray}
+  f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
+\end{eqnarray}
+Then we can rewrite the
+variational free-energy as
+\begin{eqnarray*}
+  \begin{split}
+    F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
+    &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
+    &= -\langle f_{G,R}\rangle_{P_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
+    \end{split}
+\end{eqnarray*}
+
+\begin{mdframed}
+  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
+\end{mdframed}
+
+% \begin{figure}[!tb]
+%   \centering
+%   \includegraphics[width=.8\linewidth]{vae.pdf}
+%   \caption{Variational autoencoders...}
+%   \label{fig:vae}
+% \end{figure}
+
+\paragraph{Solution: the reparametrization trick.}
+\begin{itemize}
+\item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
+\item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
+  \begin{itemize}
+  \item results in a sample $\z \sim p_R(.|\x)$, from the correct
+    posterior
+    \end{itemize}
+\end{itemize}
+The mapping $g$ should be taught of as a ``blurring'' function which produces noisy versions $\z$,
+called \textit{sensations}, of the true world state $\x$.
+The result is a scheme for training DBNs via good-old backprop!
+Refer to Fig. vae.pdf.
+% \ref{fig:vae}.
+Some examples of the reparametrization trick for a number of
+choices of the posterior distribution are given in Tab. \ref{tab:rptrick}.
+
+\begin{table}[H]
+  \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
+         \hline
+         Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
+         Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
+         Student’s t, Logistic, Uniform, Triangular, \ldots \\ \hline
+         Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
 \\ \hline
-%          Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
-%   \end{tabular}
-%   \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
-%   \label{tab:rptrick}
-% \end{table}
-
-%% For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
-%% \begin{equation}
-%%   p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
-%% \end{equation}
-
-%% \subsection{A thermodynamic model for bounded rationality, aka robust
-%%   optimality}
-%% \begin{itemize}
-%%   \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
-%%   \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
-%%   \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
-%%     \end{itemize}
-%% The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
-%% \paragraph{Utility functions and conjugate pairs.}
-%%   Let $(\Omega, \mathcal F, P)$ be a probablity space.
+         Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
+  \end{tabular}
+  \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
+  \label{tab:rptrick}
+\end{table}
+
+For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
+\begin{equation}
+  p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
+\end{equation}
+
+\subsection{A thermodynamic model for bounded rationality, aka robust
+  optimality}
+\begin{itemize}
+  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
+  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
+  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
+    \end{itemize}
+The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
+\paragraph{Utility functions and conjugate pairs.}
+  Let $(\Omega, \mathcal F, P)$ be a probablity space.
 A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
-%%   \begin{itemize}
-%%   \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.
-%%   \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
-%%     $A,B \in \mathcal F$.
-%%   \item monotonicity: $f_{U}$ is strictly increasing.
-%%     \end{itemize}
-
-%% \begin{theorem}
-%%   The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
-%%   \begin{equation}
-%%     f = \alpha \log(.),
-%%     \label{eq:boltzmann}
-%%   \end{equation}
-%%   where $\alpha > 0$.
-%% \end{theorem}
-
-%% \textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
-
-%% Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
-
-%% \paragraph{Example.}
-%% Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
-%%   \begin{equation}
-%%     P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
-%%   \end{equation}
-%%   where
-
-%%   \begin{equation}
-%%     Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
-%%   \end{equation}
-%%   is a normalization constant called the \textit{partition function} of $U$.
-%%   It's not hard to see that $U$ and the $P$ above form a conjugate pair.
-
-%%   \paragraph{Free-utility functional.}
-%%     Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
+  \begin{itemize}
+  \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.
+  \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
+    $A,B \in \mathcal F$.
+  \item monotonicity: $f_{U}$ is strictly increasing.
+    \end{itemize}
+
+\begin{theorem}
+  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
+  \begin{equation}
+    f = \alpha \log(.),
+    \label{eq:boltzmann}
+  \end{equation}
+  where $\alpha > 0$.
+\end{theorem}
+
+\textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
+
+Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
+
+\paragraph{Example.}
+Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
+  \begin{equation}
+    P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
+  \end{equation}
+  where
+
+  \begin{equation}
+    Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
+  \end{equation}
+  is a normalization constant called the \textit{partition function} of $U$.
+  It's not hard to see that $U$ and the $P$ above form a conjugate pair.
+
+  \paragraph{Free-utility functional.}
+    Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
 Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
-%%     \begin{equation}
-%%       J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
-%%     \label{eq:free_u}
-%%     \end{equation}
-%%     where
-%%     \begin{equation}
-%%       \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
-%%     \end{equation}
-%%     is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
-%%   \begin{equation}
-%%     J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
-%%   \end{equation}
-%%   In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
-%%   \begin{equation}
-%%     J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
-%%     \end{equation}
-
-%%   \paragraph{The free-energy / utility principle (of Friston ?).}
-%%   We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
+    \begin{equation}
+      J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
+    \label{eq:free_u}
+    \end{equation}
+    where
+    \begin{equation}
+      \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
+    \end{equation}
+    is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
+  \begin{equation}
+    J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
+  \end{equation}
+  In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
+  \begin{equation}
+    J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
+    \end{equation}
+
+  \paragraph{The free-energy / utility principle (of Friston ?).}
+  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
 Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
@@
 The difference in free-utility is then
-%%   \begin{equation}
-%%     \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
-%%     \label{eq:free}
-%%   \end{equation}
-%%   where
-%%   \begin{equation}
-%%     D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
-%%   \end{equation}
-%%   is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
-%% In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
-%%   The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
-%% a trade-off between an expected utility $\Delta U$ (first term) and
-%% the information cost of transforming Pi
-%% into $P_{\text{fin}}$ (second
-%% term). In this interpretation $P_{\text{ini}}$ represents an initial choice
-%% probability or policy, which includes the special case of the
-%% uniform distribution where the decision-maker has initially no
-%% preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
-%% considers the utility constraint U∗
-%% that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
-%% \begin{equation}
-%% P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
-%%   \end{equation}
-
-%% By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
-
-%% \begin{equation}
-%%   P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
-%% \end{equation}
-
-%% Two limit cases are worth considering.
-%% \begin{itemize}
-%% \item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
-%%   $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
-%%   where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
-%%   \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
-%% \end{itemize}
-
-%% To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
+  \begin{equation}
+    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
+    \label{eq:free}
+  \end{equation}
+  where
+  \begin{equation}
+    D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
+  \end{equation}
+  is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
+In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
+  The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
+a trade-off between an expected utility $\Delta U$ (first term) and
+the information cost of transforming Pi
+into $P_{\text{fin}}$ (second
+term). In this interpretation $P_{\text{ini}}$ represents an initial choice
+probability or policy, which includes the special case of the
+uniform distribution where the decision-maker has initially no
+preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
+considers the utility constraint U∗
+that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
+\begin{equation}
+P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
+  \end{equation}
+
+By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
+
+\begin{equation}
+  P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
+\end{equation}
+
+Two limit cases are worth considering.
+\begin{itemize}
+\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
+  $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
+  where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
+  \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
+\end{itemize}
+
+To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
 For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...
 
-% \subsection{GANs and other likelihood-free methods}
-% \ldots
+\subsection{GANs and other likelihood-free methods}
+\ldots
+
+\section{More on RL in the brain}
+\subsection{Does hippocampal replay equal inverse reinforcement-learning?}
+%\begin{mdframed}
+  Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
+  an MDP (called a \textit{teacher's demonstration}), can we figure out what is the
+  (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
+  \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
+  (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
+  For instance, given traces of motor actions from an adult to grab a cup on a table,
+  can an observing child figure out what ``langragian'' functional is being minimized by the former?
+  How? Can they reproduce this optimal behavior?
+%\end{mdframed}
+Such questions are of course pertinent to our decision-making theory for the DMN.
+In the general artificial case, the problem has been extensively studied and partially solved by
+\citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
+``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
+
+IRL is is suited for problems in which it's hard to define what the reward function could be
+(e.g.,  car-driving, drone maneuvering, etc.) ...
 

### 2017-08-11 paper/paper.tex
new refs (pomdp) + new ideas (pomdp, hippocampal stuff, etc.)

@@
 for biological relevance assessment
-\citep{amygdala_bzdok}.
+\citep{Averbeck2017,amygdala_bzdok}.
 
@@
 
+\begin{wrapfigure}{r}{.35\linewidth}
+  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
+    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
+  \centering
+  \includegraphics[width=1.\linewidth]{rl.pdf}
+  \label{fig:rl}
+\end{wrapfigure}
+
 The environment is generally taken as \textit{stochastic},
@@
 state are accessible to the agent's sensory perception.
-\citep{starkweather2017dopamine}.
+\citep{starkweather2017dopamine,o2006making,o2006making}.
 We assume that volatility of the environment
@@
 
-\begin{wrapfigure}{r}{.3\linewidth}
-  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
-    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
-  \centering
-  \includegraphics[width=1.\linewidth]{rl.pdf}
-  \label{fig:rl}
-\end{wrapfigure}
-
 In short, DMN implication in a diversity of
@@
 
+\subsection{Interlude: Rethinking the Action-State-Reward model of RL}
+``Remarkably, Barto proposes to decouple extrinsic rewards (i.e generated by the external environment) from intrinsic reward (generated from withing the "organism")
+biologically, extrinsic rewards would correspond to so-called "sensations"
+while intrinsic rewards would correspond to "reward signals".
+
+Thus there are two enviroments: the external environment, and the internal environment built-into the "organism".
+The organism takes highly level DECISIONS (the DMN can confortably sit here!) which act on the internal environment (this environment probably corresponds to low-level subsystems like motor, etc.) which then transform theses  into ACTIONS executed in the external environment.
+The external environment generates sensations which get mapped into reward signals (via the internal environment), etc.
+Il y a 6 min
+Guillaume Dumas (deep.introspection@gmail.com)
+Yeah, good for creating Skynet :D!
+(Half joking 🙃)
+Guillaume • Il y a 6 min
+haha
+The interesting thing is that with this rethinking, the organism can have a perception-action loop even in the absence of external stimulus.
+This is called curiosity.''
+
 \subsection{Markov decision processes}
@@
 and strategic board games like the breakthrough results upon recently on the game of Go \citep{silver2016mastering}, considered to be a golden benchmark problem in artificial intelligence.
-We emphasize that the brain in general, and the DMN in particular,
-is a physical system governed by the laws of
-physics and can be formally described
-by Markov processes at a sufficiently coarse scale.
-It has indeed been previously proposed \citep{tegmark2016improved} that
-any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
-step is sufficiently short.
-% , defining $s(t)$, the state at time $t$, as the position in state space.
-The process has $memory$ if the next state depends not only on the current state
-but also on a finite number of past states.
-Rational probabilistic planning can be reformulated
-as a standard memoryless Markov process by simply expanding the
-definition of the state $s$ to include experience episodes of the past.
-
 In artificial intelligence and machine learning, a popular computational model for
@@
 \end{definition*}
+
+\begin{mdframed}
+\paragraph*{Why MDPS ?} At this point, it is righteous to question the applicability of MPDs as a model for something as complex as human behavior ?
+Indeed a critique could argue that Markov chains are state-machines without memory.
+These may describe insects, or other behaviorally primitive beings like ATM machines, insects, etc., but won't be sufficient to describe human beings, as for example, human behavior is heavily influenced not only by the current state but by past states as well (memory).
+
+\paragraph*{Answer:} We emphasize that the brain in general, and the DMN in particular,
+is a physical system governed by the laws of
+physics and can be formally described
+by Markov processes at a sufficiently coarse scale.
+It has indeed been previously proposed \citep{tegmark2016improved} that
+any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
+step is sufficiently short.
+
+Rational probabilistic planning can be reformulated
+as a standard memory Markov process by simply expanding the
+definition of the state $s$ to include experience episodes of the past.
+This adds the capacity for memory, since then, the next state depends not only on the current state but also on a finite number of past states.
+
+Also, it should be noted that MDPs have been successfully used as neural model for animal behavior. One can mention ~\citep{Rao2010}, just to name a few.
+\end{mdframed}
 
@@
 This would be a simple control algorithm by evaluating all legal actions and selecting the action with
-highest expected cumulative rewards.
+highest expected cumulative rewards.As a result, quick-to-learn instance-based control policies serve as a rough approximation while a slower more generalisable decision-making system system is trained up~\citep{pritzel2017neural,lengyel2008hippocampal}.
 In MDPs, MCMC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible \citep{silver2016mastering}.

### 2017-08-11 paper/paper.tex
more refs + barto ideas

@@
 \textit{partially observable} in the sense that only limited aspects of the environment's
-state are accessible to the agent's sensory perception.
-\citep{starkweather2017dopamine,o2006making,o2006making}.
+state are accessible to the agent's sensory perception~\citep{starkweather2017dopamine,o2006making,o2006making}.
 We assume that volatility of the environment
@@
 \subsection{Interlude: Rethinking the Action-State-Reward model of RL}
-``Remarkably, Barto proposes to decouple extrinsic rewards (i.e generated by the external environment) from intrinsic reward (generated from withing the "organism")
+Barto proposes to decouple extrinsic rewards (i.e generated by the external environment) from intrinsic reward (generated from withing the "organism")
 biologically, extrinsic rewards would correspond to so-called "sensations"
@@
 
-Thus there are two enviroments: the external environment, and the internal environment built-into the "organism".
+Thus there are two environments: the external environment, and the internal environment built-into the "organism".
 The organism takes highly level DECISIONS (the DMN can confortably sit here!) which act on the internal environment (this environment probably corresponds to low-level subsystems like motor, etc.) which then transform theses  into ACTIONS executed in the external environment.
 The external environment generates sensations which get mapped into reward signals (via the internal environment), etc.
-Il y a 6 min
-Guillaume Dumas (deep.introspection@gmail.com)
-Yeah, good for creating Skynet :D!
-(Half joking 🙃)
-Guillaume • Il y a 6 min
-haha
+
 The interesting thing is that with this rethinking, the organism can have a perception-action loop even in the absence of external stimulus.
-This is called curiosity.''
+This is called curiosity...
 
@@
 This adds the capacity for memory, since then, the next state depends not only on the current state but also on a finite number of past states.
+This is precisely what partially observable MDPs do ~\citep{starkweather2017dopamine,o2006making,o2006making}.
 

### 2017-08-16 paper/paper.tex
more stuff

@@
 
-\subsection{Interlude: Rethinking the Action-State-Reward model of RL}
-Barto proposes to decouple extrinsic rewards (i.e generated by the external environment) from intrinsic reward (generated from withing the "organism")
-biologically, extrinsic rewards would correspond to so-called "sensations"
-while intrinsic rewards would correspond to "reward signals".
-
-Thus there are two environments: the external environment, and the internal environment built-into the "organism".
-The organism takes highly level DECISIONS (the DMN can confortably sit here!) which act on the internal environment (this environment probably corresponds to low-level subsystems like motor, etc.) which then transform theses  into ACTIONS executed in the external environment.
-The external environment generates sensations which get mapped into reward signals (via the internal environment), etc.
-
-The interesting thing is that with this rethinking, the organism can have a perception-action loop even in the absence of external stimulus.
-This is called curiosity...
-
 \subsection{Markov decision processes}
@@
 
+\paragraph*{Interlude: Rethinking the Action-State-Reward model of RL.}
+Barto proposes to decouple extrinsic rewards (i.e generated by the external environment) from intrinsic reward (generated from withing the "organism")
+biologically, extrinsic rewards would correspond to so-called "sensations"
+while intrinsic rewards would correspond to "reward signals".
+
+Thus there are two environments: the external environment, and the internal environment built-into the "organism".
+The organism takes high-level DECISIONS (the DMN can confortably sit here!) which act on the internal environment (this environment probably corresponds to low-level subsystems like motor, etc.) which then transform theses  into ACTIONS executed in the external environment.
+The external environment generates sensations which get mapped into reward signals (via the internal environment), etc.
+
+The interesting thing is that with this rethinking, the organism can have a perception-action loop even in the absence of external stimulus.
+This is called curiosity...
+
 \begin{mdframed}
-\paragraph*{Why MDPS ?} At this point, it is righteous to question the applicability of MPDs as a model for something as complex as human behavior ?
+\paragraph*{Why MDPS ?} At this point, it is righteous to question the applicability of MPDs as a model for something as complex as human behavior.
 Indeed a critique could argue that Markov chains are state-machines without memory.
@@
 
-\paragraph*{Answer:} We emphasize that the brain in general, and the DMN in particular,
+\paragraph*{Answer:}
+\begin{itemize}
+  \item First, let us emphasize that the brain in general, and the DMN in particular,
 is a physical system governed by the laws of
@@
 
-Rational probabilistic planning can be reformulated
+\item Rational probabilistic planning can be reformulated
 as a standard memory Markov process by simply expanding the
@@
 This adds the capacity for memory, since then, the next state depends not only on the current state but also on a finite number of past states.
-This is precisely what partially observable MDPs do ~\citep{starkweather2017dopamine,o2006making,o2006making}.
-
-Also, it should be noted that MDPs have been successfully used as neural model for animal behavior. One can mention ~\citep{Rao2010}, just to name a few.
+This is precisely what partially observable MDPs do ~\citep{starkweather2017dopamine,o2006making}.
+
+\item Also, it should be noted that MDPs have been successfully used as neural model for animal behavior.
+One can mention ~\citep{Rao2010} (random dots), ~\citep{Valton2011} (mal-adaptive decision-making in a rat version of a gambling task) just to name a few.
+
+\item Successes in AI/ML: \citep{ng2004,abbeel2004,mnih2015,silver2016mastering,pritzel2017neural}
+
+\end{itemize}
 \end{mdframed}

### 2017-08-18 paper/paper.tex
more refs on MDPs in algorithmic trading

@@
 
-\begin{mdframed}
-\paragraph*{Why MDPS ?} At this point, it is righteous to question the applicability of MPDs as a model for something as complex as human behavior.
+% \begin{mdframed}
+\paragraph*{Why MDPs ?} At this point, it is righteous to question the applicability of MPDs as a model for something as complex as human behavior.
 Indeed a critique could argue that Markov chains are state-machines without memory.
@@
 
-\item Successes in AI/ML: \citep{ng2004,abbeel2004,mnih2015,silver2016mastering,pritzel2017neural}
-
+\item Successes in AI/ML:
+  \begin{itemize}
+    \item Robotics: \citep{ng2004,abbeel2004}
+    \item Games and other: ~\citep{mnih2015,silver2016mastering,pritzel2017neural}
 \end{itemize}
-\end{mdframed}
+
+  \item Algorithmic trading.
+In financial trading, the main actions which the agent (human or computer acting on behalf of a human operator) must execute strategically are: sell (sell a given quantity of a commodity), buy (by a given quantity of a commodity), do nothing.
+According to how the market responds, a definite gain / loss is recorded as a consequence of the executed action, perhaps with some delay.
+Over the years, research in algorithmic trading has identified MDPs as a powerful framework for modelling this behavioral process  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010}, for which dynamic-programming based algorithms can be employed to produce numerical solutions.
+~\citep{abergel2017} is an excellent reference with the mathematical details on how MDPs work in finance.
+
+\end{itemize}
+% \end{mdframed}
 

### 2017-09-08 paper/paper.tex
Danilos revision 1 for NHB

--- 2017-08-11 paper/paper.tex
@@
 
-In the present manuscript, we propose a \textit{process model}
-that attempts to explain \textit{how} the DMN may implement
+The present paper proposes a \textit{process model}
+that tries to explain \textit{how}
+the DMN may implement
 continuous evaluation and prediction of the environment to guide behavior.
 %
-We first review the rich literature on the DMN and recognize that the DMN contains the sufficient neuro-biological substrates to implement a full-blown reinforcement-learning agent, at least in principle.
-DMN function is then recast in mathematical terms of
+DMN function is recast in mathematical terms of
 control theory and reinforcement learning based on Markov decision processes.
@@
 process model that satisfies this contention.
+We also contribute to the discussion of DMN function
+by providing some of the first empirical evidence that
+morphological variability in DMN nodes
+is linked to the reward circuitry (Fig. 2).
+Finally, we detail how our process model relates
+to previous accounts of DMN function and
+we derive explicit hypotheses to be tested in future neuroscience
+experiments (Box 1).
+
+\begin{mdframed}
+\hspace{2mm}
+\vspace{2mm}
+\textbf{BOX 1: Hypotheses for testing the MDP account of DMN function}
+
+\begin{enumerate}
+\item Experiment (Humans): We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
+During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigms.
+We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, and the more the fMRI signals in the reward circuits should be independent of the reward context
+in the sensory environment.
+
+\item Experiment (Humans): We hypothesize a functional dissociation between computations for behavioral policy versus adapting stimulus-value associations as these are probably implemented in different subsystems of the DMN.
+We first expect that fMRI signals in the right temporo-parietal junction can predict behavioral changes caused by adaptation in the action choice tendencies (policy matrix) related to non-value-related prediction error.
+Second, fMRI signals in the ventromedial prefrontal cortex should predict behavioral changes caused by adaptation in value estimation (value matrix) due to reward-related stimulus-value association.
+We finally predict that fMRI signals in the posteromedial cortex, as a potential global information integrator, is able to predict shifts in overt behavior based on previous adaptations in policy or value estimation.
+
+\item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is necessary for choice behavior in mice.
+Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before make analogous choices in the future.
+Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal dysruption of DMN nodes, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
+
+\end{enumerate}
+\end{mdframed}
 
@@
 for biological relevance assessment
-\citep{Averbeck2017,amygdala_bzdok}.
+\citep{amygdala_bzdok}.
 
@@
 
-Consequently, the PMC may contribute to overall DMN function by
-``altering behavior in response to unexpected change''
-\citep{pearson2011posterior}.
-This DMN node probably
-monitors sudden and gradual environmental changes in subjective outcomes
-of action choices for integration
+Consequently, the PMC may contribute to overall DMN function
+by monitoring the subjective outcomes
+of possible actions and integrating that information
 with memory and perspective frames
-to improve short- and longer-term behavioral agendas.
+into short- and longer-term behavioral agendas.
 Estimated value, that differs across individuals, probably enriches
@@
 through repeated interactions with an \textit{environment},
-an agent learns to reach goals and optimize reward signals occasionally emitted by the environment, in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
+an agent learns to reach goals and optimize reward signals
+in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
 At a given moment, each taken \textit{action} $a$ triggers a change
@@
 of as satisfaction \textemdash or aversion \textemdash accompanying the execution of
-a particular action.  Thus the agent learns behaviors.
+a particular action.
 % Action may be delayed to achieve substantial increases in expected reward
 % that is expected to grow with time.
-
-\begin{wrapfigure}{r}{.35\linewidth}
-  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
-    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
-  \centering
-  \includegraphics[width=1.\linewidth]{rl.pdf}
-  \label{fig:rl}
-\end{wrapfigure}
 
@@
 \textit{partially observable} in the sense that only limited aspects of the environment's
-state are accessible to the agent's sensory perception~\citep{starkweather2017dopamine,o2006making,o2006making}.
+state are accessible to the agent's sensory perception.
+\citep{starkweather2017dopamine}.
 We assume that volatility of the environment
@@
 
+\begin{wrapfigure}{r}{.3\linewidth}
+  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
+    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
+  \centering
+  \includegraphics[width=1.\linewidth]{rl.pdf}
+  \label{fig:rl}
+\end{wrapfigure}
+
 In short, DMN implication in a diversity of
@@
 and strategic board games like the breakthrough results upon recently on the game of Go \citep{silver2016mastering}, considered to be a golden benchmark problem in artificial intelligence.
+We emphasize that the brain in general, and the DMN in particular,
+is a physical system governed by the laws of
+physics and can be formally described
+by Markov processes at a sufficiently coarse scale.
+It has indeed been previously proposed \citep{tegmark2016improved} that
+any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
+step is sufficiently short.
+% , defining $s(t)$, the state at time $t$, as the position in state space.
+The process has $memory$ if the next state depends not only on the current state
+but also on a finite number of past states.
+Rational probabilistic planning can be reformulated
+as a standard memoryless Markov process by simply expanding the
+definition of the state $s$ to include experience episodes of the past.
+
 In artificial intelligence and machine learning, a popular computational model for
@@
 \end{definition*}
-
-\begin{mdframed}
-\paragraph*{Why MDPS ?} At this point, it is righteous to question the applicability of MPDs as a model for something as complex as human behavior ?
-Indeed a critique could argue that Markov chains are state-machines without memory.
-These may describe insects, or other behaviorally primitive beings like ATM machines, insects, etc., but won't be sufficient to describe human beings, as for example, human behavior is heavily influenced not only by the current state but by past states as well (memory).
-
-\paragraph*{Answer:} We emphasize that the brain in general, and the DMN in particular,
-is a physical system governed by the laws of
-physics and can be formally described
-by Markov processes at a sufficiently coarse scale.
-It has indeed been previously proposed \citep{tegmark2016improved} that
-any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
-step is sufficiently short.
-
-Rational probabilistic planning can be reformulated
-as a standard memory Markov process by simply expanding the
-definition of the state $s$ to include experience episodes of the past.
-This adds the capacity for memory, since then, the next state depends not only on the current state but also on a finite number of past states.
-This is precisely what partially observable MDPs do ~\citep{starkweather2017dopamine,o2006making,o2006making}.
-
-Also, it should be noted that MDPs have been successfully used as neural model for animal behavior. One can mention ~\citep{Rao2010}, just to name a few.
-\end{mdframed}
 
@@
 The optimal Q-value operator $Q^*$ is a fixed point for this equation.
-As a consequence of this decomposition, the complicated problem \eqref{eq:qlearning}
+As a consequence of this decomposition, the complicated dynamic programming
+problem \eqref{eq:qlearning}
 is broken down into simpler sub-problems at different time points.
-Indeed, exploitation of hierarchical structure in action considerations
+Indeed,
+exploitation of hierarchical structure in action considerations
 has previously been related to the medial prefrontal part of the DMN
@@
 % neuroscience of Bellman equation
-\textbf{Neural correlates of the Bellman equation in the DMN:}
+\textbf{Box 2: Neural correlates of the Bellman equation in the DMN}
 Relating decomposition of consecutive action choices by the Bellman equation
@@
 This would be a simple control algorithm by evaluating all legal actions and selecting the action with
-highest expected cumulative rewards.As a result, quick-to-learn instance-based control policies serve as a rough approximation while a slower more generalisable decision-making system system is trained up~\citep{pritzel2017neural,lengyel2008hippocampal}.
+highest expected cumulative rewards.
 In MDPs, MCMC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible \citep{silver2016mastering}.
@@
 Instead, FEP therefore minimizes an upper-bound of surprise,
-namely the \text{free energy} $F_G^R(\x) := \langle \log p_G(., \x) \rangle_{p_R(. |\x)} - \langle \log p_R(. |\x))\rangle_{p_R(. |\x)}$ given by
+namely the \text{free energy} $F_G^R(\x)$ given by
 
@@
       &\le F^R_G(\x),
-      \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z.
+      \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
     \end{split}
@@
 for minimizing free energy, VAEs can be efficiently trained via back-propagation of prediction errors.
-Most importantly, the reparametrization trick allows for a drastic reduction in the variance of the in estimates of the model's gradients, leading to effective learning, something which was not possible in wake-sleep\citep{dayan1995helmholtz}.
-The details of this trick can be found in \citep{kingma2013auto}.
 
@@
 
-\appendix
-
-\begin{remark}
-  Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
-  except the Markov property, we can always transform it into an MDP by considering a compounded
-  version of the states
-  $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
-  history up to an including time $t$.
-\end{remark}
-
-\section{Free-energy principles!}
-The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
-In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
+% \appendix
+% 
+% \begin{remark}
+%   Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
+%   except the Markov property, we can always transform it into an MDP by considering a compounded
+%   version of the states
+%   $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
+%   history up to an including time $t$.
+% \end{remark}
+% 
+% \section{More on RL in the brain}
+% \subsection{Does hippocampal replay equal inverse reinforcement-learning?}
+% %\begin{mdframed}
+%   Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
+%   an MDP (called a \textit{teacher's demonstration}), can we figure out what is the
+%   (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
+%   \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
+%   (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
+%   For instance, given traces of motor actions from an adult to grab a cup on a table,
+%   can an observing child figure out what ``langragian'' functional is being minimized by the former?
+%   How? Can they reproduce this optimal behavior?
+% %\end{mdframed}
+% Such questions are of course pertinent to our decision-making theory for the DMN.
+% In the general artificial case, the problem has been extensively studied and partially solved by
+% \citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
+% ``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
+% 
+% IRL is is suited for problems in which it's hard to define what the reward function could be
+% (e.g.,  car-driving, drone maneuvering, etc.) ...
+% 
+% \section{Free-energy principles!}
+% The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
+% In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
 This ideas were first developed by Hinton et al.
@@
 1995, is at the heart of the general free-energy principle that we will introduce shortly.
-
-\subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
-In this subsection, we will present an insightful upper bound for the generative surprise (i.e., generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
+% 
+% \subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
+% In this subsection, we will present an insightful upper bound for the generative surprise (i.e., generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
 As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
-
-Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
-distribution
-\begin{equation}
-  p_G(\z|\x) = \frac{p_G(\z,\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
-\end{equation}
-where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
-Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
-\begin{equation}
-F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
-\end{equation}
-Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
-it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
-density\footnote{conditioning in $P_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
-  state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
-\begin{equation}
-  \begin{split}
-    F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{P_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
-    &= \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
-    &= F^R_G(\x) - D_{KL}(P_R(.|\x) || P_G(.|\x)),
-  \end{split}
-  \label{eq:fe}
-\end{equation}
-where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
-\begin{equation}
-  F^R_G(\x) := \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x))
-\end{equation}
-and $D_{KL}(P_R(.|\x) || P_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
-
-\subsection{A general free-energy principle}
-We can resume the situation as follows\footnote{Where we have used the fact that KL divergence is always nonnegative.}:
-\begin{mdframed}
-\begin{equation}
-  \begin{split}
-    \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
-    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
-    &\le F^R_G(\x),
-    \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
-    \end{split}
-\end{equation}
-\end{mdframed}
-%% In fact, we have the following theorem
+% 
+% Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
+% distribution
+% \begin{equation}
+%   p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
+% \end{equation}
+% where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
+% Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
+% \begin{equation}
+% F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
+% \end{equation}
+% Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
+% it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
+% density\footnote{conditioning in $P_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
+%   state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
+% \begin{equation}
+%   \begin{split}
+%     F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{P_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
+%     &= \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
+%     &= F^R_G(\x) - D_{KL}(P_R(.|\x) || P_G(.|\x)),
+%   \end{split}
+%   \label{eq:fe}
+% \end{equation}
+% where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
+% \begin{equation}
+%   F^R_G(\x) := \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x))
+% \end{equation}
+% and $D_{KL}(P_R(.|\x) || P_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
+% 
+% \subsection{A general free-energy principle}
+% We can resume the situation as follows\footnote{Where we have used the fact that KL divergence is always nonnegative.}:
+% \begin{mdframed}
+% \begin{equation}
+%   \begin{split}
+%     \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
+%     &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
+%     &\le F^R_G(\x),
+%     \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
+%     \end{split}
+% \end{equation}
+% \end{mdframed}
+% %% In fact, we have the following theorem
+% %% \begin{theorem}
+% %%   It holds that
+% %%   $$
+% %%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
+% %%   i.e
+% %%   \begin{equation}
+% %%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
+% %%     \end{equation}
+% %% \end{theorem}
+% 
+% \subsection{Helmholtz machines and the wake-sleep algorithm}
+% % \begin{mdframed}
+%   \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e., input from lower more primitive layers).
+Precisely
+% $$
+% p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
+% p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
+% $$
+% % \end{mdframed}
+% 
+% \subsection{Friston's active-inference and agency}
+% \label{sec:friston}
+% This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
+% 
+% The following critics can be made:
+% \begin{itemize}
+%   \item
+%     As noted by Dayan et al.
+(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
+%   \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
+%   \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
+%     technical difficulty in getting derivatives of loss function w.r.t
+%     recognition weights $\W^R$).
+%   \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
+%     \end{itemize}
+% 
+% 
+% \subsection{Minimizing free-energy via backprop: variational auto-encoders}
+% 
+% % \subsection{Free-energy minimization via variational auto-encoders}
+% \label{sec:vae}
+% Here, we present a way to alleviate some conceptual and computational issues with the
+% free-energy framework presented thus far, by using the recent \textit{variational auto-encoder} (VAE) theory \citep{kingma2013auto}. Define the data-dependent
+% auxiliary random function
+% \begin{eqnarray}
+%   f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
+% \end{eqnarray}
+% Then we can rewrite the
+% variational free-energy as
+% \begin{eqnarray*}
+%   \begin{split}
+%     F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
+%     &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
+%     &= -\langle f_{G,R}\rangle_{P_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
+%     \end{split}
+% \end{eqnarray*}
+% 
+% \begin{mdframed}
+%   \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
+% \end{mdframed}
+% 
+% % \begin{figure}[!tb]
+% %   \centering
+% %   \includegraphics[width=.8\linewidth]{vae.pdf}
+% %   \caption{Variational autoencoders...}
+% %   \label{fig:vae}
+% % \end{figure}
+% 
+% \paragraph{Solution: the reparametrization trick.}
+% \begin{itemize}
+% \item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
+% \item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
+%   \begin{itemize}
+%   \item results in a sample $\z \sim p_R(.|\x)$, from the correct
+%     posterior
+%     \end{itemize}
+% \end{itemize}
+% The mapping $g$ should be taught of as a ``blurring'' function which produces noisy versions $\z$,
+% called \textit{sensations}, of the true world state $\x$.
+% The result is a scheme for training DBNs via good-old backprop!
+% Refer to Fig. vae.pdf.
+% % \ref{fig:vae}.
+% Some examples of the reparametrization trick for a number of
+% choices of the posterior distribution are given in Tab. \ref{tab:rptrick}.
+% 
+% \begin{table}[H]
+%   \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
+%          \hline
+%          Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
+%          Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
+%          Student’s t, Logistic, Uniform, Triangular, \ldots \\ \hline
+%          Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
+\\ \hline
+%          Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
+%   \end{tabular}
+%   \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
+%   \label{tab:rptrick}
+% \end{table}
+
+%% For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
+%% \begin{equation}
+%%   p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
+%% \end{equation}
+
+%% \subsection{A thermodynamic model for bounded rationality, aka robust
+%%   optimality}
+%% \begin{itemize}
+%%   \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
+%%   \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
+%%   \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
+%%     \end{itemize}
+%% The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
+%% \paragraph{Utility functions and conjugate pairs.}
+%%   Let $(\Omega, \mathcal F, P)$ be a probablity space.
+A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
+%%   \begin{itemize}
+%%   \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.
+%%   \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
+%%     $A,B \in \mathcal F$.
+%%   \item monotonicity: $f_{U}$ is strictly increasing.
+%%     \end{itemize}
+
 %% \begin{theorem}
-%%   It holds that
-%%   $$
-%%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
-%%   i.e
+%%   The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
 %%   \begin{equation}
-%%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
+%%     f = \alpha \log(.),
+%%     \label{eq:boltzmann}
 %%   \end{equation}
+%%   where $\alpha > 0$.
 %% \end{theorem}
 
-\subsection{Helmholtz machines and the wake-sleep algorithm}
-% \begin{mdframed}
-  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e., input from lower more primitive layers).
-Precisely
-$$
-p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
-p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
-$$
-% \end{mdframed}
-
-\subsection{Friston's active-inference and agency}
-\label{sec:friston}
-This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
-
-The following critics can be made:
-\begin{itemize}
-  \item
-    As noted by Dayan et al.
-(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
-  \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
-  \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
-    technical difficulty in getting derivatives of loss function w.r.t
-    recognition weights $\W^R$).
-  \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
-    \end{itemize}
-
-\subsection{Minimizing free-energy via backprop: variational auto-encoders}
-
-% \subsection{Free-energy minimization via variational auto-encoders}
-\label{sec:vae}
-Here, we present a way to alleviate some conceptual and computational issues with the
-free-energy framework presented thus far, by using the recent \textit{variational auto-encoder} (VAE) theory \citep{kingma2013auto}. Define the data-dependent
-auxiliary random function
-\begin{eqnarray}
-  f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
-\end{eqnarray}
-Then we can rewrite the
-variational free-energy as
-\begin{eqnarray*}
-  \begin{split}
-    F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
-    &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
-    &= -\langle f_{G,R}\rangle_{P_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
-    \end{split}
-\end{eqnarray*}
-
-\begin{mdframed}
-  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
-\end{mdframed}
-
-% \begin{figure}[!tb]
-%   \centering
-%   \includegraphics[width=.8\linewidth]{vae.pdf}
-%   \caption{Variational autoencoders...}
-%   \label{fig:vae}
-% \end{figure}
-
-\paragraph{Solution: the reparametrization trick.}
-\begin{itemize}
-\item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
-\item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
-  \begin{itemize}
-  \item results in a sample $\z \sim p_R(.|\x)$, from the correct
-    posterior
-    \end{itemize}
-\end{itemize}
-The mapping $g$ should be taught of as a ``blurring'' function which produces noisy versions $\z$,
-called \textit{sensations}, of the true world state $\x$.
-The result is a scheme for training DBNs via good-old backprop!
-Refer to Fig. vae.pdf.
-% \ref{fig:vae}.
-Some examples of the reparametrization trick for a number of
-choices of the posterior distribution are given in Tab. \ref{tab:rptrick}.
-
-\begin{table}[H]
-  \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
-         \hline
-         Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
-         Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
-         Student’s t, Logistic, Uniform, Triangular, \ldots \\ \hline
-         Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
-\\ \hline
-         Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
-  \end{tabular}
-  \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
-  \label{tab:rptrick}
-\end{table}
-
-For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
-\begin{equation}
-  p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
-\end{equation}
-
-\subsection{A thermodynamic model for bounded rationality, aka robust
-  optimality}
-\begin{itemize}
-  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
-  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
-  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
-    \end{itemize}
-The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
-\paragraph{Utility functions and conjugate pairs.}
-  Let $(\Omega, \mathcal F, P)$ be a probablity space.
-A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
-  \begin{itemize}
-  \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.
-  \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
-    $A,B \in \mathcal F$.
-  \item monotonicity: $f_{U}$ is strictly increasing.
-    \end{itemize}
-
-\begin{theorem}
-  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
-  \begin{equation}
-    f = \alpha \log(.),
-    \label{eq:boltzmann}
-  \end{equation}
-  where $\alpha > 0$.
-\end{theorem}
-
-\textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
-
-Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
-
-\paragraph{Example.}
-Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
-  \begin{equation}
-    P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
-  \end{equation}
-  where
-
-  \begin{equation}
-    Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
-  \end{equation}
-  is a normalization constant called the \textit{partition function} of $U$.
-  It's not hard to see that $U$ and the $P$ above form a conjugate pair.
-
-  \paragraph{Free-utility functional.}
-    Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
+%% \textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
+
+%% Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
+
+%% \paragraph{Example.}
+%% Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
+%%   \begin{equation}
+%%     P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
+%%   \end{equation}
+%%   where
+
+%%   \begin{equation}
+%%     Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
+%%   \end{equation}
+%%   is a normalization constant called the \textit{partition function} of $U$.
+%%   It's not hard to see that $U$ and the $P$ above form a conjugate pair.
+
+%%   \paragraph{Free-utility functional.}
+%%     Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
 Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
-    \begin{equation}
-      J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
-    \label{eq:free_u}
-    \end{equation}
-    where
-    \begin{equation}
-      \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
-    \end{equation}
-    is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
-  \begin{equation}
-    J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
-  \end{equation}
-  In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
-  \begin{equation}
-    J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
-    \end{equation}
-
-  \paragraph{The free-energy / utility principle (of Friston ?).}
-  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
+%%     \begin{equation}
+%%       J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
+%%     \label{eq:free_u}
+%%     \end{equation}
+%%     where
+%%     \begin{equation}
+%%       \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
+%%     \end{equation}
+%%     is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
+%%   \begin{equation}
+%%     J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
+%%   \end{equation}
+%%   In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
+%%   \begin{equation}
+%%     J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
+%%     \end{equation}
+
+%%   \paragraph{The free-energy / utility principle (of Friston ?).}
+%%   We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
 Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
@@
 The difference in free-utility is then
-  \begin{equation}
-    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
-    \label{eq:free}
-  \end{equation}
-  where
-  \begin{equation}
-    D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
-  \end{equation}
-  is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
-In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
-  The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
-a trade-off between an expected utility $\Delta U$ (first term) and
-the information cost of transforming Pi
-into $P_{\text{fin}}$ (second
-term). In this interpretation $P_{\text{ini}}$ represents an initial choice
-probability or policy, which includes the special case of the
-uniform distribution where the decision-maker has initially no
-preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
-considers the utility constraint U∗
-that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
-\begin{equation}
-P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
-  \end{equation}
-
-By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
-
-\begin{equation}
-  P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
-\end{equation}
-
-Two limit cases are worth considering.
-\begin{itemize}
-\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
-  $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
-  where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
-  \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
-\end{itemize}
-
-To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
+%%   \begin{equation}
+%%     \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
+%%     \label{eq:free}
+%%   \end{equation}
+%%   where
+%%   \begin{equation}
+%%     D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
+%%   \end{equation}
+%%   is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
+%% In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
+%%   The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
+%% a trade-off between an expected utility $\Delta U$ (first term) and
+%% the information cost of transforming Pi
+%% into $P_{\text{fin}}$ (second
+%% term). In this interpretation $P_{\text{ini}}$ represents an initial choice
+%% probability or policy, which includes the special case of the
+%% uniform distribution where the decision-maker has initially no
+%% preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
+%% considers the utility constraint U∗
+%% that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
+%% \begin{equation}
+%% P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
+%%   \end{equation}
+
+%% By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
+
+%% \begin{equation}
+%%   P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
+%% \end{equation}
+
+%% Two limit cases are worth considering.
+%% \begin{itemize}
+%% \item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
+%%   $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
+%%   where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
+%%   \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
+%% \end{itemize}
+
+%% To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
 For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...
 
-\subsection{GANs and other likelihood-free methods}
-\ldots
-
-\section{More on RL in the brain}
-\subsection{Does hippocampal replay equal inverse reinforcement-learning?}
-%\begin{mdframed}
-  Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
-  an MDP (called a \textit{teacher's demonstration}), can we figure out what is the
-  (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
-  \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
-  (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
-  For instance, given traces of motor actions from an adult to grab a cup on a table,
-  can an observing child figure out what ``langragian'' functional is being minimized by the former?
-  How? Can they reproduce this optimal behavior?
-%\end{mdframed}
-Such questions are of course pertinent to our decision-making theory for the DMN.
-In the general artificial case, the problem has been extensively studied and partially solved by
-\citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
-``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
-
-IRL is is suited for problems in which it's hard to define what the reward function could be
-(e.g.,  car-driving, drone maneuvering, etc.) ...
+% \subsection{GANs and other likelihood-free methods}
+% \ldots
 

### 2017-09-08 paper/paper.tex
Add Venn Diagram figure + Typos corrections

@@
 local minima based on the first derivative of the optimization objective.
-Stochastic here means that the true gradient is estimated from batches of training examples samples,
+Stochastic here means that the true gradient is estimated from batches of training samples,
 which, in our case, corresponds to blocks of experience from the replay memory:
@@
 
+\begin{figure}[!h]
+  \includegraphics[width=.9\linewidth]{VennDiagram-2017-09-08.pdf}
+  \caption{\textbf{Where does Markov Decision Process fit in the theoretical landscape?}}
+  This Venn diagram summarizes the relationship between four existing accounts of DMN: in blue colors, the Free Energy Principle \& Predictive Coding, more related to the ”How” DMN works, and in yellow, the Sentinel \& the Semantic hypotheses, more related to "What" DMN does.
+Our account of DMN by Markov Decision Process is at the intersection of those accounts and thus corresponds to a "Process model" meaning a model of both "How" and "What" a process occur; here the mental processes and how it relate to observed behavior.
+For each account, 4 icons describe the core concepts of Time horizon, Neural plasticity, Hierarchy, and Behavioral goal.
+The "Time horizon" should be understood as a weighting of the expected future outcomes shorter to longer terms into present decision making.
+Then, "online monitoring" of the environment means detecting and processing salient cues in the sensory environment.
+The "Predictive Coding" is understood here as the joint minimization of feedforward prediction error and feedback modulation to optimize neural processing; there is thus no notion of an agent with behavioral goal.}
+  \label{fig:VennDiagram}
+\end{figure}
+
 \section{Relation to existing accounts}
+
 \subsection{Predictive coding hypothesis}
@@
 The dynamic interplay between cortical processing levels
-may enable learning about world aspects by reconciling
+may enable learning about aspects of the world by reconciling
 gaps between fresh sensory input and expectations computed
@@
 Moreover,
-both accounts provide a parsimonious explanation why the
+both accounts provide a parsimonious explanation for why the
 human brain's processing load devoted to incoming information decreases
@@
 the hierarchical processing aspect from predictive coding
-is re-expressed in our account in form of
+is re-expressed in our account in the form of
 nested prediction of probable upcoming actions, states, and outcomes.
@@
 the neurobiological plausibility of
-the back-propagation procedure is controversal
+the back-propagation procedure is controversial
 \citep{goodfellow2016deep}.
@@
 \citep{hassabis2009construction}.
-Indeed, in aging humans, remembering the past and imaging the future
+Indeed, in aging humans, remembering the past and imagining the future
 equally decreased in the level of detail and were associated with
@@
 episodic memory, language, problem solving,
-planning, estimating other others' thoughts, and spatial navigation
+planning, estimating others' thoughts, and spatial navigation
 represent neural processes that are likely to
@@
 the complexity of the social systems
-is a likely to be a human-defining property
+is likely to be a human-defining property
 \citep{tomasello2009cultural}.
@@
 Despite its popularity, criticism against the FEP has arisen over the years,
-some of which outlined in the following.
+some of which is outlined in the following.
 The main algorithm for minimizing free energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
@@
 overcome a number of the technical limits of the wake-sleep algorithm
-by using a reparametrization trick. For instance, unlike the wake-sleep algorithm
+by using a reparametrization maneuver. For instance, unlike the wake-sleep algorithm
 for minimizing free energy, VAEs can be efficiently trained via back-propagation of prediction errors.
@@
 MDPs motivate an attractive
-formal account how the human association cortex
+formal account of how the human association cortex
 might implement multi-sensory representation and control of the environment to
@@
 %
-This functional interpretation may be well compatible with the DMN's
+This functional interpretation may well be compatible with the DMN's
 poorly understood involvement across

### 2017-09-09 paper/paper.tex
integrating other comments for R1

@@
        A Unified Account of Default Mode Function\\
-       by Control Theory and Reinforcement Learning}
+       by Markov Decision Processes}
 
@@
 %
-DMN function is recast in mathematical terms of
-control theory and reinforcement learning based on Markov decision processes.
+DMN function is recast in terms of
+control theory and reinforcement learning by basing it on Markov decision processes.
 %
@@
 \subsection{Towards a formal account of default mode function: higher-order control of the organism}
-The human DMN is responsible for extended parts of the baseline neural activity,
+The network nodes that compose the human DMN
+are responsible for extended parts of the baseline neural activity,
 which typically decreases when engaged in psychological experiments
@@
 by providing some of the first empirical evidence that
-morphological variability in DMN nodes
+morphological variability in DMN regions
 is linked to the reward circuitry (Fig. 2).
@@
 Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before make analogous choices in the future.
-Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal dysruption of DMN nodes, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
+Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal dysruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
 
@@
 based on experimental findings in the neuroscience literature.
-This walkthrough across each node of the DMN will
+This walkthrough across each region of the DMN will
 outline the individual functional profiles,
@@
 %
-This DMN node has long been speculated to
+This DMN region has long been speculated to
 reflect constant computation of
@@
 across individual decision-making instances \citep{pearson2009neurons}.
-This DMN node encoded the
+This DMN region encoded the
 preference for or aversion to options with uncertain reward outcomes
@@
   \caption{\textbf{Morphological coupling between reward system and
-  default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11+/-0.02$) and right NAc volume ($R^2=0.14+/-0.02$) could be predicted from volume in the DMN nodes.
-These out-of-sample generalization performances were obtained from support vector regression applied to normalized node volumes in the DMN in a 10-fold cross-validation procedure.
+  default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11+/-0.02$) and right NAc volume ($R^2=0.14+/-0.02$) could be predicted from volume in the DMN regions.
+These out-of-sample generalization performances were obtained from support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure.
   Consistent for the left and right reward system, NAc volume in a given subject is
@@
 Indeed,
-this DMN node is preferentially connected with reward-related and limbic regions.
+this DMN region is preferentially connected with reward-related and limbic regions.
 The vmPFC is well known to have direct connections
@@
 have been proposed to be critical in communicating local information
-to the neocortex for long-term storage, potentially also in the nodes of the DMN.
+to the neocortex for long-term storage, potentially also in the regions of the DMN.
 Moreover, in mice, invasively triggering spatial experience recall
@@
 \citep{corbetta2002control}.
-Involvement of this DMN node was repeatedly reported in
+Involvement of this DMN region was repeatedly reported in
 multi-step action execution \citep{hartmann2005takes},
@@
 self-relevant and unexpected sources of information
-as a ‘‘circuit breaker’’ that recalibrates control and maintenance brain networks
+as a ‘‘circuit breaker’’ that recalibrates functional control of brain networks
 \citep{bzdok2013tpj, corbettashul2008}.
@@
 
-\section{Reinforcement learning: a process model for DMN function}
+\section{Reinforcement learning control: a process model for DMN function}
 We now argue the outlined neurobiological properties
-of the DMN nodes
+of the DMN regions
 to be sufficient for implementing all components
@@
 Note that value buffering may be realized in the vmPFC.
-This DMN node has direct connections to
+This DMN region has direct connections to
 to the NAc, known to be involved in reward evaluation.
@@
 Moreover,
-neurological patients with lesions in this DMN node were reported
+neurological patients with lesions in this DMN region were reported
 to be impaired in aspects of realizing ``multiple sub-goal scheduling''
@@
 
-An additional layer of learning concerns the addition of new entries in the table (i.e., state and action).
+% An additional layer of learning concerns the addition of new entries in the table (i.e., state and action).
 The framing of the policy matrix is not described here but could be supported by synaptic epigenesis \citep{gisiger_acquisition_2005}.
 Indeed, the tuning of synaptic weights through learning can stabilize additional patterns of activity by
-creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}. Those attractors will then constrain both
-the number of factors taken into account by decision processes and the number of behaviors at reach by the agent \citep{wang_decision_2008}.
+% creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}. Those attractors will then constrain both
+% the number of factors taken into account by decision processes and the number of behaviors at reach by the agent \citep{wang_decision_2008}.
 
@@
   activity when humans engage in cognitive processes that are detached from
-  the current sensory environment
-  and this network was proposed to be situated
-  at the top of the brain network hierarchy
+  the current sensory environment. Additionally,
+  this network was proposed to be situated
+  at the top of the brain network hierarchy, with 
+  the subordinate salience and dorsal attention network in the middle and
+  the unimodal sensory cortices at the bottom
   \citep{carhart2010default, margulies2016situating}.

### 2017-09-09 paper/paper.tex
- BF: fixed latex errors due broken Vienn diagram caption (unmatched curly parantheses) - BF: broken layout of figures (floating beyond page border) - BF: multiple problems introduced lately

@@
 \citep{bzdok2013segregation}.
-
 The dmPFC may subserve inference, representation, and assessment
@@
 and stored knowledge of world regularities.
-\begin{wrapfigure}{r}{.4\linewidth}
-  \includegraphics[width=1.\linewidth]{fig_dmn.png}
-  \centering
-  \caption{\textbf{Morphological coupling between reward system and
-  default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11+/-0.02$) and right NAc volume ($R^2=0.14+/-0.02$) could be predicted from volume in the DMN regions.
-These out-of-sample generalization performances were obtained from support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure.
-  Consistent for the left and right reward system, NAc volume in a given subject is
-  positively coupled with the vmPFC and HC.
-  Code and data for reproduction and visualization:
-  \url{www.github.com/banilo/to_be_added_later}.
-  }
-  \label{fig:dmn}
-\end{wrapfigure}
 %
@@
 in humans \citep{doherty2015structure}.
+
+% \begin{wrapfigure}{r}{.4\linewidth}
+\begin{figure}[!htbp]
+  \begin{minipage}[c]{0.63\textwidth}  
+    \includegraphics[width=.8\linewidth]{fig_dmn.png}
+  \end{minipage}
+  \hspace{-4em}
+  \begin{minipage}[c]{0.4\textwidth}  
+    \caption{\textbf{Morphological coupling between reward system and
+        default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from volume in the DMN regions.
+These out-of-sample generalization performances were obtained from support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure.
+      Consistent for the left and right reward system, NAc volume in a given subject is
+  positively coupled with the vmPFC and HC.
+  Code and data for reproduction and visualization:
+  \url{www.github.com/banilo/to_be_added_later}.
+  }
+  \label{fig:dmn}
+\end{minipage}
+% \end{wrapfigure}
+\end{figure}
+
 Consistently, diffusion MRI tractography in humans and monkeys
@@
 memory operations and spatial navigation in animals and humans.
-
 While the HC
@@
 imagination of experiences more broadly \citep{hassabis2007patients}.
+
 Mental scenes created by neurological patients with HC lesion exposed a lack of
@@
 \citep{hassabis2007patients}.
-
 Thus, the HC probably orchestrates re-experience of environmental aspects for
@@
 definition of the state $s$ to include experience episodes of the past.
+
+\begin{wrapfigure}{r}{.35\linewidth}
+  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
+    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
+  \centering
+  \includegraphics[width=1.\linewidth]{rl.pdf}
+  \label{fig:rl}
+\end{wrapfigure}
 
@@
   \includegraphics[width=.9\linewidth]{VennDiagram-2017-09-08.pdf}
-  \caption{\textbf{Where does Markov Decision Process fit in the theoretical landscape?}}
+  \caption{\textbf{Where does Markov Decision Process fit in the theoretical landscape?}
   This Venn diagram summarizes the relationship between four existing accounts of DMN: in blue colors, the Free Energy Principle \& Predictive Coding, more related to the ”How” DMN works, and in yellow, the Sentinel \& the Semantic hypotheses, more related to "What" DMN does.
@@
 Then, "online monitoring" of the environment means detecting and processing salient cues in the sensory environment.
-The "Predictive Coding" is understood here as the joint minimization of feedforward prediction error and feedback modulation to optimize neural processing; there is thus no notion of an agent with behavioral goal.}
+The "Predictive Coding" is understood here as the joint minimization of feedforward prediction error and feedback modulation to optimize neural processing; there is thus no notion of an agent with behavioral goal.
+  }
   \label{fig:VennDiagram}
@@
 and predict its sensory input.
+\begin{wrapfigure}{r}{.45\linewidth}
+  \caption{\textbf{The dark room experiment.} An intelligent agent situated
+  in a light-deprived closed space can be used
+  as a thought experiment for the complete absence
+  of external sensory input.}
+  \label{fig:darkroom}
+  \centering
+  \includegraphics[width=1.\linewidth]{darkroom_resized.png}
+\end{wrapfigure}
 Precursors of this interpretation can be traced back to Dayan and colleagues \citep{dayan1995helmholtz}
@@
 imagined mental scenarios.
-Howewer, surprise is challenging to optimize numerically
+
+However, surprise is challenging to optimize numerically
 because we need to sum over all hidden causes of the sensations.
@@
 because its sensory inputs are perfectly predictable
-\begin{wrapfigure}{r}{.4\linewidth}
-  \caption{\textbf{The dark room experiment.} An intelligent agent situated
-  in a light-deprived closed space can be used
-  as a thought experiment for the complete absence
-  of external sensory input.}
-  \label{fig:darkroom}
-  \centering
-  \includegraphics[width=1.\linewidth]{darkroom_resized.png}
-\end{wrapfigure}
 given the environmental state \citep{darkroom2012}.
@@
 its surroundings or leave it.
+
 Defenders of the FEP have retorted by advancing the ``full package'' \citep{darkroom2012}:

### 2017-09-09 paper/paper.tex
making the abstract a single paragraph (I've never seen a multi-paragraph abstract...)

@@
 envisioning experience to anticipate the future.
-
-The present paper proposes a \textit{process model}
+In the present work, we propose a \textit{process model}
 that tries to explain \textit{how}
@@
 \citep{dayan2008decision}.
-
 In short, DMN implication in a diversity of

### 2017-09-09 paper/paper.tex
a general pass

@@
 \begin{figure}[!h]
-  \includegraphics[width=.9\linewidth]{VennDiagram-2017-09-08.pdf}
+  \includegraphics[width=1\linewidth]{VennDiagram-2017-09-08.pdf}
   \caption{\textbf{Where does Markov Decision Process fit in the theoretical landscape?}
@@
 According to theories of the free-energy principle (FEP) and
-active inference \citep{friston2010free, fristonAIorRL},
+active inference \citep{friston2010free, fristonAIorRL} (see also \citep{dayan1995helmholtz}),
 the brain corresponds to
@@
 and predict its sensory input.
-\begin{wrapfigure}{r}{.45\linewidth}
-  \caption{\textbf{The dark room experiment.} An intelligent agent situated
-  in a light-deprived closed space can be used
-  as a thought experiment for the complete absence
-  of external sensory input.}
-  \label{fig:darkroom}
-  \centering
-  \includegraphics[width=1.\linewidth]{darkroom_resized.png}
-\end{wrapfigure}
-Precursors of this interpretation can be traced back to Dayan and colleagues \citep{dayan1995helmholtz}
-who introduced the \textit{Helmholtz machine}, a hierarchical factorial directional deep belief network. According to the FEP account,
-the goal of the brain is to optimize over a generative model $G$ of sensations: to iteratively
-modify its internal representation $p_G(\z|\x)$ of objects in the world, their interactions and dynamics. The model then allows
-to minimize surprise when these representations are confronted with sensory input $\x$ during action-perception cycles (i.e., the \textit{generative} model).
-The FEP also proposes a companion model called the \textit{recognition} model,
-which works in tandem with the generative model $p_R(\z|\x)$ to accomplish approximate inference. Put differently, the recognition model dreams imaginary worlds $\z$,
-while the generative model tries to produce sensory sensations $\x$ which match these
-imagined mental scenarios.
-
+% Precursors of this interpretation can be traced back to Dayan and colleagues \citep{dayan1995helmholtz}
+% who introduced the \textit{Helmholtz machine}, a hierarchical factorial directional deep belief network. According to the FEP account,
+% the goal of the brain is to optimize over a generative model $G$ of sensations: to iteratively
+% modify its internal representation $p_G(\z|\x)$ of objects in the world, their interactions and dynamics. The model then allows
+% to minimize surprise when these representations are confronted with sensory input $\x$ during action-perception cycles (i.e., the \textit{generative} model).
+% The FEP also proposes a companion model called the \textit{recognition} model $p_R(\z|\x)$,
+% which works in tandem with the generative model to accomplish approximate inference. Put differently, the recognition model dreams imaginary worlds $\z$,
+% while the generative model tries to produce sensory sensations $\x \sim p_G(\x|\z)$ which match these
+% imagined mental scenarios.
 However, surprise is challenging to optimize numerically
-because we need to sum over all hidden causes of the sensations.
-Instead, FEP therefore minimizes an upper-bound of surprise,
-namely the \text{free energy} $F_G^R(\x)$ given by
+because we need to sum over all hidden causes $\z$ of the sensations (an intractable problem).
+Instead, FEP therefore minimizes an upper-bound on surprise given by
 
@@
     \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
-      &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
+    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{\operatorname{{KL}}(p_R(\z|\x) || p_G(\z|\x))}_{\text{complexity}} \\
     &\le F^R_G(\x),
-      \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
+    \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z.
   \end{split}
 \end{equation}
-
+where
+\begin{equation}
+  F_G^R(\x) := \langle -\log(p_G(\z, \x))\rangle_{p_R(\z|\x)} - \mathcal H(p_R(\z|\x))
+\end{equation}
+is the \textit{free energy}.
+Here, $\langle . \rangle$ is the \textit{expectation} operator, $\mathcal H$ is the \textit{entropy} functional while $\operatorname{KL}(.|.)$ is the usual \textit{Kullback-Leibler (KL) divergence},
+which is a measure of how different two probability distributions are.
+In this framework, the goal of the agent is then to iterative refine the generative model $p_G$ and the recognition model $p_R$ so as to minimize the free energy $F_G^R(\x)$ over sensory input $\x$.
+% The optimization procedure proposed by the authors derives from the \textit{wake-sleep algorithm} \citep{dayan1995helmholtz}.
 % \begin{figure}[!tb]
@@
 % CRITICISM
+\paragraph{Main criticisms.}
 Despite its popularity, criticism against the FEP has arisen over the years,
@@
 of forward (generation) and backward (recognition) models that
-together do not correspond to optimization of (a bound of) the marginal likelihood.
+together do not correspond to optimization of (a bound of) the marginal likelihood,
+because KL divergence is not symmetric in its arguments.
 The brain may therefore be unlikely to implement a variant of the wake-sleep algorithm.
@@
 overcome a number of the technical limits of the wake-sleep algorithm
-by using a reparametrization maneuver. For instance, unlike the wake-sleep algorithm
+by using a reparametrization maneuver, which makes it possible to do differential calculus on random sampling procedures without exploding variance. As a result, unlike the wake-sleep algorithm
 for minimizing free energy, VAEs can be efficiently trained via back-propagation of prediction errors.
@@
 its surroundings or leave it.
-
 Defenders of the FEP have retorted by advancing the ``full package'' \citep{darkroom2012}:
@@
 be \textit{surprised} by such a \textit{lack of surprise}. According to this argument, a dark room would paradoxically correspond to a state of particularly high surprise.
+\begin{wrapfigure}{r}{.37\linewidth}
+  \centering
+  \includegraphics[width=1.\linewidth]{darkroom_resized.png}
+  \caption{\textbf{The dark room experiment.} An intelligent agent situated
+    in a light-deprived closed space can be used
+    as a thought experiment for the complete absence
+    of external sensory input.
+    % The free-energy principle theory paradoxally predicts that such an agent would remain quietly in this room, such it perfectly predict all sensory input from such an environment.
+    % However such behavior is not observed in humans or any animals: the agent would strive to leave such an environment in search of more ``interesting'' worlds.
+  }
+  \label{fig:darkroom}
+\end{wrapfigure}
 Driven by surprise-minimization objective, the FEP agent would eventually bootstrap itself out of such saddle points to explore more interesting parts of the environment.
@@
 %   \begin{split}
-%     \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(P_G(\x))\sum_{\z}p_G(\z |\x)\\
+%     \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(p_G(\x))\sum_{\z}p_G(\z |\x)\\
 %     &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
@@
 % it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
-% density\footnote{conditioning in $P_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
+% density\footnote{conditioning in $p_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
 %   state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
@@
 %   \begin{split}
-%     F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{P_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
-%     &= \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
-%     &= F^R_G(\x) - D_{KL}(P_R(.|\x) || P_G(.|\x)),
+%     F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{p_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
+%     &= \langle E_G(., \x)\rangle_{p_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
+%     &= F^R_G(\x) - D_{KL}(p_R(.|\x) || p_G(.|\x)),
 %   \end{split}
@@
 % \begin{equation}
-%   F^R_G(\x) := \langle E_G(., \x)\rangle_{P_R(.|\x)} - \mathcal H(p_R(.|\x))
+%   F^R_G(\x) := \langle E_G(., \x)\rangle_{p_R(.|\x)} - \mathcal H(p_R(.|\x))
 % \end{equation}
-% and $D_{KL}(P_R(.|\x) || P_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
+% and $D_{KL}(p_R(.|\x) || p_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
 %
@@
 %     \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
-%     &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(P_R(.|\x) || P_G(.|\x))}_{\text{complexity}} \\
+%     &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(p_R(.|\x) || p_G(.|\x))}_{\text{complexity}} \\
 %     &\le F^R_G(\x),
@@
 %   \begin{split}
-%     F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
-%     &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{P_R(.|\x)}\\
-%     &= -\langle f_{G,R}\rangle_{P_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
+%     F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{p_R(.|\x)}\\
+%     &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{p_R(.|\x)}\\
+%     &= -\langle f_{G,R}\rangle_{p_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
 %     \end{split}

### 2017-09-09 paper/paper.tex
another big pass

@@
 \begin{mdframed}
-\hspace{2mm}
-\vspace{2mm}
-\textbf{BOX 1: Hypotheses for testing the MDP account of DMN function}
-
+  \textbf{Box 1: Hypotheses for testing the MDP account of DMN function}
+  \vspace{1em}
 \begin{enumerate}
@@
 step is sufficiently short.
+\begin{wrapfigure}{l}{.4\linewidth}
+  \centering  
+  \includegraphics[width=1.\linewidth]{rl.pdf}
+  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
+    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
+  \label{fig:rl}
+\end{wrapfigure}
 % , defining $s(t)$, the state at time $t$, as the position in state space.
@@
 definition of the state $s$ to include experience episodes of the past.
-
-\begin{wrapfigure}{r}{.35\linewidth}
-  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
-    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
-  \centering
-  \includegraphics[width=1.\linewidth]{rl.pdf}
-  \label{fig:rl}
-\end{wrapfigure}
 
@@
   \end{equation}
+
 The Bellman equation \eqref{eq:bellman} is a temporal consistency equation which provides
@@
 
-\pagebreak[4]
 \begin{mdframed}
-\hspace{2mm}
-\vspace{2mm}
 % neuroscience of Bellman equation
   \textbf{Box 2: Neural correlates of the Bellman equation in the DMN}
-Relating decomposition of consecutive action choices by the Bellman equation
+  \vspace{1em}
+
+\hspace{-1.5em}Relating decomposition of consecutive action choices by the Bellman equation
 to neuroscience,

### 2017-09-09 paper/paper.tex
recovering lost MDP motivation section

@@
 \end{definition*}
+
+% \paragraph*{Interlude: Rethinking the Action-State-Reward model of RL.}
+% Barto proposes to decouple extrinsic rewards (i.e generated by the external environment) from intrinsic reward (generated from withing the "organism")
+% biologically, extrinsic rewards would correspond to so-called "sensations"
+% while intrinsic rewards would correspond to "reward signals".
+
+% Thus there are two environments: the external environment, and the internal environment built-into the "organism".
+The organism takes high-level DECISIONS (the DMN can confortably sit here!) which act on the internal environment (this environment probably corresponds to low-level subsystems like motor, etc.) which then transform theses  into ACTIONS executed in the external environment.
+The external environment generates sensations which get mapped into reward signals (via the internal environment), etc.
+
+% The interesting thing is that with this rethinking, the organism can have a perception-action loop even in the absence of external stimulus.
+% This is called curiosity...
+
+% \begin{mdframed}
+\paragraph*{Why MDPs ?} At this point, it is righteous to question the applicability of MPDs as a model for something as complex as human behavior.
+Indeed a critique could argue that Markov chains are state-machines without memory.
+These may describe insects, or other behaviorally primitive beings like ATM machines, insects, etc., but won't be sufficient to describe human beings, as for example, human behavior is heavily influenced not only by the current state but by past states as well (memory).
+
+\paragraph*{Answer:}
+\begin{itemize}
+  \item First, let us emphasize that the brain in general, and the DMN in particular,
+is a physical system governed by the laws of
+physics and can be formally described
+by Markov processes at a sufficiently coarse scale.
+It has indeed been previously proposed \citep{tegmark2016improved} that
+any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
+step is sufficiently short.
+
+\item Rational probabilistic planning can be reformulated
+as a standard memory Markov process by simply expanding the
+definition of the state $s$ to include experience episodes of the past.
+This adds the capacity for memory, since then, the next state depends not only on the current state but also on a finite number of past states.
+This is precisely what partially observable MDPs do ~\citep{starkweather2017dopamine,o2006making}.
+
+\item Also, it should be noted that MDPs have been successfully used as neural model for animal behavior.
+One can mention ~\citep{Rao2010} (random dots), ~\citep{Valton2011} (mal-adaptive decision-making in a rat version of a gambling task) just to name a few.
+
+\item Successes in AI/ML:
+  \begin{itemize}
+    \item Robotics: \citep{ng2004,abbeel2004}
+    \item Games and other: ~\citep{mnih2015,silver2016mastering,pritzel2017neural}
+\end{itemize}
+
+  \item Algorithmic trading.
+In financial trading, the main actions which the agent (human or computer acting on behalf of a human operator) must execute strategically are: sell (sell a given quantity of a commodity), buy (by a given quantity of a commodity), do nothing.
+According to how the market responds, a definite gain / loss is recorded as a consequence of the executed action, perhaps with some delay.
+Over the years, research in algorithmic trading has identified MDPs as a powerful framework for modelling this behavioral process  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010}, for which dynamic-programming based algorithms can be employed to produce numerical solutions.
+~\citep{abergel2017} is an excellent reference with the mathematical details on how MDPs work in finance.
+
+\end{itemize}
 

### 2017-09-10 paper/paper.tex
Typo + Shorten Venn diagram caption

@@
   This Venn diagram summarizes the relationship between four existing accounts of DMN: in blue colors, the Free Energy Principle \& Predictive Coding, more related to the ”How” DMN works, and in yellow, the Sentinel \& the Semantic hypotheses, more related to "What" DMN does.
-Our account of DMN by Markov Decision Process is at the intersection of those accounts and thus corresponds to a "Process model" meaning a model of both "How" and "What" a process occur; here the mental processes and how it relate to observed behavior.
+Our account of DMN by Markov Decision Process is at the intersection of those accounts and thus corresponds to a "Process model" meaning a model of both "How" and "What" a process occur.
+  % ; here the mental processes and how it relate to observed behavior.
 For each account, 4 icons describe the core concepts of Time horizon, Neural plasticity, Hierarchy, and Behavioral goal.
@@
 A prediction error at one of the processing levels
-incurs plasticity changes of neuronal projections
+induces plasticity changes of neuronal projections
 (i.e., adapting model parameters)
@@
 Further,
-both functional accounts are plausibilized by
+both functional accounts are supported by
 neuroscientific evidence that suggest

### 2017-09-10 paper/paper.tex
some more refs to FEP and FEP-RL links

@@
 would continuously generate hypothetical explanations of the world
-and predict its sensory input.
+and predict its sensory input $\bf{x}$ (analogous to the state-action $(s, a)$ pair in an MDP
+framework).
 % Precursors of this interpretation can be traced back to Dayan and colleagues \citep{dayan1995helmholtz}
@@
     \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
-    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{\operatorname{{KL}}(p_R(\z|\x) || p_G(\z|\x))}_{\text{complexity}} \\
+    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{\operatorname{{KL}}(p_R(\z|\x) \| p_G(\z|\x))}_{\text{complexity}} \\
     &\le F^R_G(\x),
@@
 is the \textit{free energy}.
-Here, $\langle . \rangle$ is the \textit{expectation} operator, $\mathcal H$ is the \textit{entropy} functional while $\operatorname{KL}(.|.)$ is the usual \textit{Kullback-Leibler (KL) divergence},
+Here, the angular brackets denote the expectation of the joint negative log-likelihood $-\log(p_G(\z, \x))$ w.r.t the recognition density $p_R(\z|\x)$, $\mathcal H$ is the \textit{entropy} functional defined by $\mathcal H(p) := -\sum_{\z}p(\z)\log(p(\z)$, while $\operatorname{KL}(.\|.)$ is the usual \textit{Kullback-Leibler (KL) divergence} defined by $\operatorname{KL}(p\|q) := \sum_{z}p(\z)\log(p(\z)/q(\z))$,
 which is a measure of how different two probability distributions are.
-In this framework, the goal of the agent is then to iterative refine the generative model $p_G$ and the recognition model $p_R$ so as to minimize the free energy $F_G^R(\x)$ over sensory input $\x$.
+In this framework, the goal of the agent is then to iteratively refine the generative model $p_G$ and the recognition model $p_R$ (i.e two Bayesian belief nets) so as to minimize the free energy $F_G^R(\x)$ over sensory input $\x$.
+$F^R_G(x)$ is low just in case
+\begin{itemize}
+\item  $p_R(\z|\x)$ puts a lot of mass on configurations $(\z,\x)$ which are $p_G$-likely, % (i.e the sensation $\x$ is likely caused by $\z$)
+  and
+\item $p_R(\z|\x)$ as uniform as possible (i.e have high entropy), so as not to concentrate its mass only on a small subset of possible causes for the sensation $\x$.
+  \end{itemize}
 % The optimization procedure proposed by the authors derives from the \textit{wake-sleep algorithm} \citep{dayan1995helmholtz}.
@@
 \begin{eqnarray*}
-  -Q^{\text{FEP}} \sim \underbrace{\log (p_G)}_{\text{FEP generative surprise}}.
+  -Q^{\text{FEP}} \approx \underbrace{F^R_G(\x)}_{\text{negative free energy}} \approx \underbrace{-\log (p_G)}_{\text{FEP generative surprise}}.
     \end{eqnarray*}
@@
 Such a surprise-guided reinforcement learning scheme has previously been advocated
-under the equivalent framework of information compression \citep{schmidhuber2010,mohamed2015}.
+under the equivalent framework of energy-based reinforcement-learning ~\citep{sallans2004,elfwing2016free} and information compression \citep{schmidhuber2010,mohamed2015}.
 Nevertheless, minimization of surprise quantities alone

### 2017-09-10 paper/paper.tex
more passes on FEP

@@
 % INTRO
-According to theories of the free-energy principle (FEP) and
+According to theories of the \textit{free-energy principle} (FEP) and
 active inference \citep{friston2010free, fristonAIorRL} (see also \citep{dayan1995helmholtz}),
@@
 The main algorithm for minimizing free energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
-\citep{dayan1995helmholtz}. As these authors noted,
-a crucial drawback of the wake-sleep algorithm is that it involves a pair
-of forward (generation) and backward (recognition) models that
+\citep{dayan1995helmholtz}. As these P. Dayan and G. Hinton  noted,
+a crucial drawback of the wake-sleep algorithm (and therefore of theories like the FEP \citep{friston2010free} based on it) is that it involves a pair
+of forward (generation) and backward (recognition) models $p_G$ and $p_R$ respectively that
 together do not correspond to optimization of (a bound of) the marginal likelihood,
@@
 Such a surprise-guided reinforcement learning scheme has previously been advocated
-under the equivalent framework of energy-based reinforcement-learning ~\citep{sallans2004,elfwing2016free} and information compression \citep{schmidhuber2010,mohamed2015}.
+under the equivalent framework of energy-based reinforcement-learning\footnote{These methods assume that the posterior distribution  of the hidden states $\z$ given the visible states $\x$ factorizes, leading to so-called RBMs --\textit{Restricted Boltzmann machines}.
+One notes, that even with these simplifications, such a model can be very hard to train.} ~\citep{sallans2004,elfwing2016free} and information compression \citep{schmidhuber2010,mohamed2015}.
 Nevertheless, minimization of surprise quantities alone

### 2017-09-10 paper/paper.tex
more

@@
 According to theories of the \textit{free-energy principle} (FEP) and
-active inference \citep{friston2010free, fristonAIorRL} (see also \citep{dayan1995helmholtz}),
+\textit{active inference} \citep{friston2010free, fristonAIorRL} (see also \citep{dayan1995helmholtz}),
 the brain corresponds to
@@
 is the \textit{free energy}.
-Here, the angular brackets denote the expectation of the joint negative log-likelihood $-\log(p_G(\z, \x))$ w.r.t the recognition density $p_R(\z|\x)$, $\mathcal H$ is the \textit{entropy} functional defined by $\mathcal H(p) := -\sum_{\z}p(\z)\log(p(\z)$, while $\operatorname{KL}(.\|.)$ is the usual \textit{Kullback-Leibler (KL) divergence} defined by $\operatorname{KL}(p\|q) := \sum_{z}p(\z)\log(p(\z)/q(\z))$,
+Here, the angular brackets denote the \textit{expectation} of the joint negative log-likelihood $-\log(p_G(\z, \x))$ w.r.t the recognition density $p_R(\z|\x)$, $\mathcal H$ is the \textit{entropy} functional defined by $\mathcal H(p) := -\sum_{\z}p(\z)\log(p(\z)$, while $\operatorname{KL}(.\|.)$ is the usual \textit{Kullback-Leibler (KL) divergence} (also known as \textit{relative entropy}) defined by $\operatorname{KL}(p\|q) := \sum_{z}p(\z)\log(p(\z)/q(\z)) \ge 0$,
 which is a measure of how different two probability distributions are.
 In this framework, the goal of the agent is then to iteratively refine the generative model $p_G$ and the recognition model $p_R$ (i.e two Bayesian belief nets) so as to minimize the free energy $F_G^R(\x)$ over sensory input $\x$.
-$F^R_G(x)$ is low just in case
+
+Importantly, one notes that $F^R_G(x)$ is low just in case:
 \begin{itemize}
 \item  $p_R(\z|\x)$ puts a lot of mass on configurations $(\z,\x)$ which are $p_G$-likely, % (i.e the sensation $\x$ is likely caused by $\z$)
-  and
-\item $p_R(\z|\x)$ as uniform as possible (i.e have high entropy), so as not to concentrate its mass only on a small subset of possible causes for the sensation $\x$.
+  \textbf{and}
+\item $p_R(\z|\x)$ is as uniform as possible (i.e have high entropy), so as not to concentrate
+  all its mass on a small subset of possible causes for the sensation $\x$.
   \end{itemize}

### 2017-09-10 paper/paper.tex
some more facts and justifications

@@
 
-Finally, we note that FEP and active inference can be reframed in terms of
+\paragraph{Links between FEP and RL.}
+We note that FEP and active inference can be reframed in terms of
 our model-free RL framework.
@@
 that is, the log-likehood of current sensory priors
-the agent has about the world. More explicitly, this corresponds to
+the agent has about the world. More explicitly, this corresponds to using free-energy as a Q-value approximator for the MDP, i.e
 
 \begin{eqnarray*}
-  -Q^{\text{FEP}} \approx \underbrace{F^R_G(\x)}_{\text{negative free energy}} \approx \underbrace{-\log (p_G)}_{\text{FEP generative surprise}}.
+  -Q \approx \underbrace{F^R_G(\x)}_{\text{negative free energy}} \approx \underbrace{-\log (p_G)}_{\text{FEP generative surprise}}.
     \end{eqnarray*}
@@
 Such a surprise-guided reinforcement learning scheme has previously been advocated
-under the equivalent framework of energy-based reinforcement-learning\footnote{These methods assume that the posterior distribution  of the hidden states $\z$ given the visible states $\x$ factorizes, leading to so-called RBMs --\textit{Restricted Boltzmann machines}.
+under the equivalent framework of energy-based reinforcement-learning\footnote{These methods assume that the posterior distribution  of the hidden states $\z$ given the visible states $\x$ factorizes, leading to so-called RBMs --\textit{Restricted Boltzmann Machines}.
 One notes, that even with these simplifications, such a model can be very hard to train.} ~\citep{sallans2004,elfwing2016free} and information compression \citep{schmidhuber2010,mohamed2015}.

### 2017-09-17 paper/paper.tex
addressing some comments

@@
 brain imaging may have been the first technique
-to allow for the discovery of a unique brain network that probably subserves
+to allow for the discovery of a unique brain network that would subserve
 baseline mental activities
@@
 
-\item Experiment (Humans): We hypothesize a functional dissociation between computations for behavioral policy versus adapting stimulus-value associations as these are probably implemented in different subsystems of the DMN.
+\item Experiment (Humans): We hypothesize a functional dissociation between computations for behavioral policy versus adapting stimulus-value associations as these are conceivably implemented in different subsystems of the DMN.
 We first expect that fMRI signals in the right temporo-parietal junction can predict behavioral changes caused by adaptation in the action choice tendencies (policy matrix) related to non-value-related prediction error.
@@
 \section{Known neurobiological properties of the default mode network}
+\label{sec:nodes}
 We begin by a neurobiological deconstruction of the DMN
@@
 into short- and longer-term behavioral agendas.
-Estimated value, that differs across individuals, probably enriches
+Estimated value, that differs across individuals, might enrich
 statistical assessment of the environment
@@
 processing of “what matters now”,
-probably drawing on sophisticated value representations
+drawing on sophisticated value representations
 \citep{rolls_OFC, doherty2015structure}.
@@
   \hspace{-4em}
-  \begin{minipage}[c]{0.4\textwidth}
+  \begin{minipage}[c]{0.45\textwidth}
     \caption{\textbf{Morphological coupling between reward system and
@@
       Consistent for the left and right reward system, NAc volume in a given subject is
-  positively coupled with the vmPFC and HC.
+      positively coupled with the vmPFC and HC. The colors are indicative of the (\textcolor{red}{red} = positive,
+      \textcolor{blue}{blue} = negative) and size (the lighter the higher) of the regression coefficients.
   Code and data for reproduction and visualization:
@@
 These HC-subserved mechanisms
-probably contribute to advanced cognitive processes that require
+conceivably contribute to advanced cognitive processes that require
 re-experiencing or newly constructed mental scenarios,
@@
 \citep{hassabis2007patients}.
-Thus, the HC probably orchestrates re-experience of environmental aspects for
+Thus, the HC would orchestrate re-experience of environmental aspects for
 consolidations based on re-enactment and for integration into
@@
 multi-sensory prediction and prediction error signaling.
-It is probably central for
-action initiation during goal-directed psychological tasks and for
+It is central for action initiation during goal-directed psychological tasks and for
 sensorimotor behavior by integrating multi-sensory attention
@@
 and neuropsychology,
-the LTPJ probably mediates access to and integration of world knowlege,
+the LTPJ mediates access to and integration of world knowledge,
 such as required during action considerations
@@
 re-assembling world knowledge.
-Isolated building blocks of world structure probably get rebuilt
+Isolated building blocks of world structure get rebuilt
 in internally constructed mental scenarios that
@@
 
-% \paragraph*{Interlude: Rethinking the Action-State-Reward model of RL.}
-% Barto proposes to decouple extrinsic rewards (i.e generated by the external environment) from intrinsic reward (generated from withing the "organism")
-% biologically, extrinsic rewards would correspond to so-called "sensations"
-% while intrinsic rewards would correspond to "reward signals".
-
-% Thus there are two environments: the external environment, and the internal environment built-into the "organism".
-The organism takes high-level DECISIONS (the DMN can confortably sit here!) which act on the internal environment (this environment probably corresponds to low-level subsystems like motor, etc.) which then transform theses  into ACTIONS executed in the external environment.
-The external environment generates sensations which get mapped into reward signals (via the internal environment), etc.
-
-% The interesting thing is that with this rethinking, the organism can have a perception-action loop even in the absence of external stimulus.
-% This is called curiosity...
-
-% \begin{mdframed}
 \paragraph*{Why MDPs ?} At this point, it is righteous to question the applicability of MPDs as a model for something as complex as human behavior.
@@
     choosing actions to be executed in a certain situation.
-% Interactions between external and internal sources of
-% information are probably commonplace during perception \citep{gershman2010learning}
 The DMN may subserve
@@
   neurobiological implementation of reinforcement learning.}
-  Overview of how the constituent nodes of the DMN may
+  Overview of how the constituent nodes of the DMN (refer to section \ref{sec:nodes}) may
   map onto computational components necessary for an RL agent.}
@@
 environmental stimuli especially important for humans are frequently of
-social nature. This is probably unsurprising
+social nature. This is not surprising
 given that
@@
 
-% \appendix
-%
-% \begin{remark}
-%   Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
-%   except the Markov property, we can always transform it into an MDP by considering a compounded
-%   version of the states
-%   $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
-%   history up to an including time $t$.
-% \end{remark}
-%
-% \section{More on RL in the brain}
-% \subsection{Does hippocampal replay equal inverse reinforcement-learning?}
-% %\begin{mdframed}
-%   Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
-%   an MDP (called a \textit{teacher's demonstration}), can we figure out what is the
-%   (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
-%   \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
-%   (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
-%   For instance, given traces of motor actions from an adult to grab a cup on a table,
-%   can an observing child figure out what ``langragian'' functional is being minimized by the former?
-%   How? Can they reproduce this optimal behavior?
-% %\end{mdframed}
-% Such questions are of course pertinent to our decision-making theory for the DMN.
-% In the general artificial case, the problem has been extensively studied and partially solved by
-% \citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
-% ``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
-%
-% IRL is is suited for problems in which it's hard to define what the reward function could be
-% (e.g.,  car-driving, drone maneuvering, etc.) ...
-%
-% \section{Free-energy principles!}
-% The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
-% In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
+\appendix
+
+\begin{remark}
+  Given a system $(\mathcal S, \mathcal A, r, p, \mu)$ satisfying all the axioms for an MDP
+  except the Markov property, we can always transform it into an MDP by considering a compounded
+  version of the states
+  $$S_t \leftarrow (S_0,A_0,R_0,S_1,A_1,S_2\ldots,S_t,A_{t-1},R_{t-1},S_{t})$$ made of the system's
+  history up to an including time $t$.
+\end{remark}
+
+\section{More on RL in the brain}
+\subsection{Does hippocampal replay equal inverse reinforcement-learning?}
+%\begin{mdframed}
+  Given the trace $s_0,a_0,s_1,\ldots,$ of an optimal agent's strategy $\pi^*$ in
+  an MDP (called a \textit{teacher's demonstration}), can we figure out what is the
+  (instantaneous) reward function $r: \mathcal S \times \mathcal \mathcal A
+  \rightarrow \mathbb R$ that the agent is optimizing over a prescribed class of reward functions
+  (e.g., linear rewards $r(s,a) \equiv \theta^T\phi(s,a)$)?
+  For instance, given traces of motor actions from an adult to grab a cup on a table,
+  can an observing child figure out what ``langragian'' functional is being minimized by the former?
+  How? Can they reproduce this optimal behavior?
+%\end{mdframed}
+Such questions are of course pertinent to our decision-making theory for the DMN.
+In the general artificial case, the problem has been extensively studied and partially solved by
+\citep{abbeel2004}. They are also been rigorously studied in general optimal control literature under the name
+``inverse optimal control'', but in model-based certain (where the physical dynamics are known, etc.)...
+
+IRL is is suited for problems in which it's hard to define what the reward function could be
+(e.g.,  car-driving, drone maneuvering, etc.) ...
+
+\section{Free-energy principles!}
+The so-called free-energy principle in its present form (including notions like ``generative density'', ``recognition density'', etc.) can be traced back to works of Dayan \& Hinton \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial directional deep belief-net (DBN).
+In this subsection, we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making.
 This ideas were first developed by Hinton et al.
@@
 For instance, the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
-%
-% % \begin{mdframed}
-% %   Though our presentation of free-energy principles is thermodynamic in flavor, a completely equivalent
-% %   information-theoretic formulation is of course possible, simply by exploiting the information-theoretic interpretation of entropy.
-% % \end{mdframed}
-%
-% \subsection{Helmholtz free-energy and the generative model}
-% \begin{table}[H]
-%   \begin{tabular}{p{2cm}|p{11cm}}
-%          \hline
-%          \textbf{symbol}    & \textbf{description}  \\ \hline
-%          $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
-%          random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
-%          $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
-%           with the usual convention $0 \log(0) := 0$.\\ \hline
-%          $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
-%              $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e., $\x = (s, b).$ \\ \hline
-%              $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
-%              $p_G(.|\x)$ & Generative density for ...\\ \hline
-%          $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
-%          $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
-%          $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
-%           to $R$.  Note that $F^G_G = F_G$.\\ \hline
-%   \end{tabular}
-%   \caption{Table of notations.}
-% \end{table}
-% Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
+
+% \begin{mdframed}
+%   Though our presentation of free-energy principles is thermodynamic in flavor, a completely equivalent
+%   information-theoretic formulation is of course possible, simply by exploiting the information-theoretic interpretation of entropy.
+% \end{mdframed}
+
+\subsection{Helmholtz free-energy and the generative model}
+\begin{table}[H]
+  \begin{tabular}{p{2cm}|p{11cm}}
+         \hline
+         \textbf{symbol}    & \textbf{description}  \\ \hline
+         $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
+         random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
+         $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
+          with the usual convention $0 \log(0) := 0$.\\ \hline
+         $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
+             $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e., $\x = (s, b).$ \\ \hline
+             $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
+             $p_G(.|\x)$ & Generative density for ...\\ \hline
+         $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
+         $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
+         $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
+          to $R$.  Note that $F^G_G = F_G$.\\ \hline
+  \end{tabular}
+  \caption{Table of notations.}
+\end{table}
+Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
 So,
-% \begin{eqnarray}
-%   \begin{split}
-%     \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(p_G(\x))\sum_{\z}p_G(\z |\x)\\
-%     &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
-%     =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
-%     &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
-%     &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
-%     &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
-%   \end{split}
-%   \label{eq:helm}
-% \end{eqnarray}
-% where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
-% \begin{equation}
-%   E_G(\z, \x) := -\log(p_G(\z, \x)),
-%   \label{eq:gibbs}
-% \end{equation}
-% %% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
-% The last quantity in \eqref{eq:helm} is nothing but \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
-% \begin{equation}
-%   F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
-% \end{equation}
-% %% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
-% Thus,
-% \begin{fact}
-%   Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
-% \end{fact}
-%
-% The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
+\begin{eqnarray}
+  \begin{split}
+    \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(p_G(\x))\sum_{\z}p_G(\z |\x)\\
+    &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
+    =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
+    &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
+    &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
+    &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
+  \end{split}
+  \label{eq:helm}
+\end{eqnarray}
+where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
+\begin{equation}
+  E_G(\z, \x) := -\log(p_G(\z, \x)),
+  \label{eq:gibbs}
+\end{equation}
+%% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
+The last quantity in \eqref{eq:helm} is nothing but \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
+\begin{equation}
+  F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
+\end{equation}
+%% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
+Thus,
+\begin{fact}
+  Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
+\end{fact}
+
+The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
 It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(\x)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
@@
 1995, is at the heart of the general free-energy principle that we will introduce shortly.
-%
-% \subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
-% In this subsection, we will present an insightful upper bound for the generative surprise (i.e., generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
+
+\subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
+In this subsection, we will present an insightful upper bound for the generative surprise (i.e., generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
 As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
-%
-% Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
-% distribution
-% \begin{equation}
-%   p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
-% \end{equation}
-% where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
-% Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
-% \begin{equation}
-% F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
-% \end{equation}
-% Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
-% it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
-% density\footnote{conditioning in $p_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
-%   state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
-% \begin{equation}
-%   \begin{split}
-%     F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{p_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
-%     &= \langle E_G(., \x)\rangle_{p_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
-%     &= F^R_G(\x) - D_{KL}(p_R(.|\x) || p_G(.|\x)),
-%   \end{split}
-%   \label{eq:fe}
-% \end{equation}
-% where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
-% \begin{equation}
-%   F^R_G(\x) := \langle E_G(., \x)\rangle_{p_R(.|\x)} - \mathcal H(p_R(.|\x))
-% \end{equation}
-% and $D_{KL}(p_R(.|\x) || p_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
-%
-% \subsection{A general free-energy principle}
-% We can resume the situation as follows\footnote{Where we have used the fact that KL divergence is always nonnegative.}:
+
+Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
+distribution
+\begin{equation}
+  p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
+\end{equation}
+where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
+Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
+\begin{equation}
+F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
+\end{equation}
+Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
+it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
+density\footnote{conditioning in $p_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
+  state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
+\begin{equation}
+  \begin{split}
+    F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{p_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
+    &= \langle E_G(., \x)\rangle_{p_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
+    &= F^R_G(\x) - D_{KL}(p_R(.|\x) || p_G(.|\x)),
+  \end{split}
+  \label{eq:fe}
+\end{equation}
+where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
+\begin{equation}
+  F^R_G(\x) := \langle E_G(., \x)\rangle_{p_R(.|\x)} - \mathcal H(p_R(.|\x))
+\end{equation}
+and $D_{KL}(p_R(.|\x) || p_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
+
+\subsection{A general free-energy principle}
+We can resume the situation as follows\footnote{Where we have used the fact that KL divergence is always nonnegative.}:
+\begin{mdframed}
+\begin{equation}
+  \begin{split}
+    \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
+    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(p_R(.|\x) || p_G(.|\x))}_{\text{complexity}} \\
+    &\le F^R_G(\x),
+    \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
+    \end{split}
+\end{equation}
+\end{mdframed}
+%% In fact, we have the following theorem
+%% \begin{theorem}
+%%   It holds that
+%%   $$
+%%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
+%%   i.e
+%%   \begin{equation}
+%%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
+%%     \end{equation}
+%% \end{theorem}
+
+\subsection{Helmholtz machines and the wake-sleep algorithm}
 % \begin{mdframed}
-% \begin{equation}
-%   \begin{split}
-%     \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
-%     &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(p_R(.|\x) || p_G(.|\x))}_{\text{complexity}} \\
-%     &\le F^R_G(\x),
-%     \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
-%     \end{split}
-% \end{equation}
+  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e., input from lower more primitive layers).
+Precisely
+$$
+p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
+p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
+$$
 % \end{mdframed}
-% %% In fact, we have the following theorem
-% %% \begin{theorem}
-% %%   It holds that
-% %%   $$
-% %%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
-% %%   i.e
-% %%   \begin{equation}
-% %%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
-% %%     \end{equation}
-% %% \end{theorem}
-%
-% \subsection{Helmholtz machines and the wake-sleep algorithm}
-% % \begin{mdframed}
-%   \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e., input from lower more primitive layers).
-Precisely
-% $$
-% p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
-% p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
-% $$
-% % \end{mdframed}
-%
-% \subsection{Friston's active-inference and agency}
-% \label{sec:friston}
-% This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
-%
-% The following critics can be made:
-% \begin{itemize}
-%   \item
-%     As noted by Dayan et al.
+
+\subsection{Friston's active-inference and agency}
+\label{sec:friston}
+This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
+
+The following critics can be made:
+\begin{itemize}
+  \item
+    As noted by Dayan et al.
 (\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
-%   \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
-%   \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
-%     technical difficulty in getting derivatives of loss function w.r.t
-%     recognition weights $\W^R$).
-%   \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
-%     \end{itemize}
-%
-%
-% \subsection{Minimizing free-energy via backprop: variational auto-encoders}
-%
-% % \subsection{Free-energy minimization via variational auto-encoders}
-% \label{sec:vae}
-% Here, we present a way to alleviate some conceptual and computational issues with the
-% free-energy framework presented thus far, by using the recent \textit{variational auto-encoder} (VAE) theory \citep{kingma2013auto}. Define the data-dependent
-% auxiliary random function
-% \begin{eqnarray}
-%   f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
-% \end{eqnarray}
-% Then we can rewrite the
-% variational free-energy as
-% \begin{eqnarray*}
-%   \begin{split}
-%     F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{p_R(.|\x)}\\
-%     &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{p_R(.|\x)}\\
-%     &= -\langle f_{G,R}\rangle_{p_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
-%     \end{split}
-% \end{eqnarray*}
-%
-% \begin{mdframed}
-%   \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
-% \end{mdframed}
-%
-% % \begin{figure}[!tb]
-% %   \centering
-% %   \includegraphics[width=.8\linewidth]{vae.pdf}
-% %   \caption{Variational autoencoders...}
-% %   \label{fig:vae}
-% % \end{figure}
-%
-% \paragraph{Solution: the reparametrization trick.}
-% \begin{itemize}
-% \item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
-% \item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
-%   \begin{itemize}
-%   \item results in a sample $\z \sim p_R(.|\x)$, from the correct
-%     posterior
-%     \end{itemize}
-% \end{itemize}
-% The mapping $g$ should be taught of as a ``blurring'' function which produces noisy versions $\z$,
-% called \textit{sensations}, of the true world state $\x$.
-% The result is a scheme for training DBNs via good-old backprop!
-% Refer to Fig. vae.pdf.
-% % \ref{fig:vae}.
-% Some examples of the reparametrization trick for a number of
-% choices of the posterior distribution are given in Tab. \ref{tab:rptrick}.
-%
-% \begin{table}[H]
-%   \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
-%          \hline
-%          Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
-%          Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
-%          Student’s t, Logistic, Uniform, Triangular, \ldots \\ \hline
-%          Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
+  \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
+  \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
+    technical difficulty in getting derivatives of loss function w.r.t
+    recognition weights $\W^R$).
+  \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
+    \end{itemize}
+
+\subsection{Minimizing free-energy via backprop: variational auto-encoders}
+
+% \subsection{Free-energy minimization via variational auto-encoders}
+\label{sec:vae}
+Here, we present a way to alleviate some conceptual and computational issues with the
+free-energy framework presented thus far, by using the recent \textit{variational auto-encoder} (VAE) theory \citep{kingma2013auto}. Define the data-dependent
+auxiliary random function
+\begin{eqnarray}
+  f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
+\end{eqnarray}
+Then we can rewrite the
+variational free-energy as
+\begin{eqnarray*}
+  \begin{split}
+    F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{p_R(.|\x)}\\
+    &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{p_R(.|\x)}\\
+    &= -\langle f_{G,R}\rangle_{p_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
+    \end{split}
+\end{eqnarray*}
+
+\begin{mdframed}
+  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
+\end{mdframed}
+
+% \begin{figure}[!tb]
+%   \centering
+%   \includegraphics[width=.8\linewidth]{vae.pdf}
+%   \caption{Variational autoencoders...}
+%   \label{fig:vae}
+% \end{figure}
+
+\paragraph{Solution: the reparametrization trick.}
+\begin{itemize}
+\item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
+\item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
+  \begin{itemize}
+  \item results in a sample $\z \sim p_R(.|\x)$, from the correct
+    posterior
+    \end{itemize}
+\end{itemize}
+The mapping $g$ should be taught of as a ``blurring'' function which produces noisy versions $\z$,
+called \textit{sensations}, of the true world state $\x$.
+The result is a scheme for training DBNs via good-old backprop!
+Refer to Fig. vae.pdf.
+% \ref{fig:vae}.
+Some examples of the reparametrization trick for a number of
+choices of the posterior distribution are given in Tab. \ref{tab:rptrick}.
+
+\begin{table}[H]
+  \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
+         \hline
+         Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
+         Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
+         Student’s t, Logistic, Uniform, Triangular, \ldots \\ \hline
+         Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
 \\ \hline
-%          Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
-%   \end{tabular}
-%   \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
-%   \label{tab:rptrick}
-% \end{table}
-
-%% For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
-%% \begin{equation}
-%%   p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
-%% \end{equation}
-
-%% \subsection{A thermodynamic model for bounded rationality, aka robust
-%%   optimality}
-%% \begin{itemize}
-%%   \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
-%%   \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
-%%   \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
-%%     \end{itemize}
-%% The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
-%% \paragraph{Utility functions and conjugate pairs.}
-%%   Let $(\Omega, \mathcal F, P)$ be a probablity space.
+         Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
+  \end{tabular}
+  \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
+  \label{tab:rptrick}
+\end{table}
+
+For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
+\begin{equation}
+  p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
+\end{equation}
+
+\subsection{A thermodynamic model for bounded rationality, aka robust
+  optimality}
+\begin{itemize}
+  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
+  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
+  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
+    \end{itemize}
+The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
+\paragraph{Utility functions and conjugate pairs.}
+  Let $(\Omega, \mathcal F, P)$ be a probablity space.
 A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
-%%   \begin{itemize}
-%%   \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.
-%%   \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
-%%     $A,B \in \mathcal F$.
-%%   \item monotonicity: $f_{U}$ is strictly increasing.
-%%     \end{itemize}
-
-%% \begin{theorem}
-%%   The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
-%%   \begin{equation}
-%%     f = \alpha \log(.),
-%%     \label{eq:boltzmann}
-%%   \end{equation}
-%%   where $\alpha > 0$.
-%% \end{theorem}
-
-%% \textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
-
-%% Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
-
-%% \paragraph{Example.}
-%% Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
-%%   \begin{equation}
-%%     P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
-%%   \end{equation}
-%%   where
-
-%%   \begin{equation}
-%%     Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
-%%   \end{equation}
-%%   is a normalization constant called the \textit{partition function} of $U$.
-%%   It's not hard to see that $U$ and the $P$ above form a conjugate pair.
-
-%%   \paragraph{Free-utility functional.}
-%%     Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
+  \begin{itemize}
+  \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.
+  \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
+    $A,B \in \mathcal F$.
+  \item monotonicity: $f_{U}$ is strictly increasing.
+    \end{itemize}
+
+\begin{theorem}
+  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
+  \begin{equation}
+    f = \alpha \log(.),
+    \label{eq:boltzmann}
+  \end{equation}
+  where $\alpha > 0$.
+\end{theorem}
+
+\textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
+
+Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
+
+\paragraph{Example.}
+Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
+  \begin{equation}
+    P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
+  \end{equation}
+  where
+
+  \begin{equation}
+    Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
+  \end{equation}
+  is a normalization constant called the \textit{partition function} of $U$.
+  It's not hard to see that $U$ and the $P$ above form a conjugate pair.
+
+  \paragraph{Free-utility functional.}
+    Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
 Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
-%%     \begin{equation}
-%%       J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
-%%     \label{eq:free_u}
-%%     \end{equation}
-%%     where
-%%     \begin{equation}
-%%       \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
-%%     \end{equation}
-%%     is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
-%%   \begin{equation}
-%%     J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
-%%   \end{equation}
-%%   In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
-%%   \begin{equation}
-%%     J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
-%%     \end{equation}
-
-%%   \paragraph{The free-energy / utility principle (of Friston ?).}
-%%   We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
+    \begin{equation}
+      J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
+    \label{eq:free_u}
+    \end{equation}
+    where
+    \begin{equation}
+      \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
+    \end{equation}
+    is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
+  \begin{equation}
+    J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
+  \end{equation}
+  In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
+  \begin{equation}
+    J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
+    \end{equation}
+
+  \paragraph{The free-energy / utility principle (of Friston ?).}
+  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
 Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
@@
 The difference in free-utility is then
-%%   \begin{equation}
-%%     \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
-%%     \label{eq:free}
-%%   \end{equation}
-%%   where
-%%   \begin{equation}
-%%     D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
-%%   \end{equation}
-%%   is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
-%% In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
-%%   The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
-%% a trade-off between an expected utility $\Delta U$ (first term) and
-%% the information cost of transforming Pi
-%% into $P_{\text{fin}}$ (second
-%% term). In this interpretation $P_{\text{ini}}$ represents an initial choice
-%% probability or policy, which includes the special case of the
-%% uniform distribution where the decision-maker has initially no
-%% preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
-%% considers the utility constraint U∗
-%% that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
-%% \begin{equation}
-%% P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
-%%   \end{equation}
-
-%% By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
-
-%% \begin{equation}
-%%   P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
-%% \end{equation}
-
-%% Two limit cases are worth considering.
-%% \begin{itemize}
-%% \item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
-%%   $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
-%%   where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
-%%   \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
-%% \end{itemize}
-
-%% To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
+  \begin{equation}
+    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
+    \label{eq:free}
+  \end{equation}
+  where
+  \begin{equation}
+    D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
+  \end{equation}
+  is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
+In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
+  The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
+a trade-off between an expected utility $\Delta U$ (first term) and
+the information cost of transforming Pi
+into $P_{\text{fin}}$ (second
+term). In this interpretation $P_{\text{ini}}$ represents an initial choice
+probability or policy, which includes the special case of the
+uniform distribution where the decision-maker has initially no
+preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
+considers the utility constraint U∗
+that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
+\begin{equation}
+P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
+  \end{equation}
+
+By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
+
+\begin{equation}
+  P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
+\end{equation}
+
+Two limit cases are worth considering.
+\begin{itemize}
+\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
+  $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
+  where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
+  \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
+\end{itemize}
+
+To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
 For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...
 
-% \subsection{GANs and other likelihood-free methods}
-% \ldots
+\subsection{GANs and other likelihood-free methods}
+\ldots
 

### 2017-09-17 paper/paper.tex
more

@@
 \citep{mnih2015,silver2016mastering} is to learn this
-representation using a deep neural-network model, leading to the
+representation using an artificial deep neural-network model, leading to the
 so-called \textit{deep Q-learning} family of methods which
-likely are state-of-the-art in RL.
+This is the current state-of-the-art in RL research.
 The set of model parameters $\theta$ that instantiate the non-linear interactions

### 2017-09-17 paper/paper.tex
finished addressing comments

@@
 Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before make analogous choices in the future.
-Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal dysruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
+Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal disruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
 
@@
 %
-Such highly abstract neural computations necessarily rely on the
-construction of probabilistic internal information drawing from
-memory recall, mental scene elaboration processes,
-and stored knowledge of world regularities.
-%
-Moreover,
-in a computational neuroimaging experiment,
-dorsomedial PFC activity preferentially processed the consequences of
-action choices that were considered but not actually executed,
-whereas ventromedial PFC (vmPFC) activity
-processed especially value outcomes of actions actually performed on the environment
-\citep{nicolle2012agent}.
+% Such highly abstract neural computations necessarily rely on the
+% construction of probabilistic internal information drawing from
+% memory recall, mental scene elaboration processes,
+% and stored knowledge of world regularities.
+% %
+% Moreover,
+% in a computational neuroimaging experiment,
+% dorsomedial PFC activity preferentially processed the consequences of
+% action choices that were considered but not actually executed,
+% whereas ventromedial PFC (vmPFC) activity
+% processed especially value outcomes of actions actually performed on the environment
+% \citep{nicolle2012agent}.
 
@@
 to be sufficient for implementing all components
-of a full-fledged RL system.
+of a full-fledged reinforcement-learning (RL) system.
 Recalling past experience, considering candidate actions,
@@
 
-RL is a problem-solving technique in which,
-through repeated interactions with an \textit{environment},
-an agent learns to reach goals and optimize reward signals
+RL is an area of machine learning concerned with learning optimal behavioral through interactions
+with an \textit{environment}, the goal being to maximize some notion of \textit{cumulative reward}.
+The optimal behavior typical
+takes the future into account, as some rewards could be \textit{delayed}.
+Through repeated interactions with the environment,
+the agent learns to reach goals and optimize reward signals
 in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
@@
 \subsubsection{Value approximation and the policy matrix}
+\label{sec:policymat}
 As already mentioned in the previous section, Q-learning optimizes over the class of
@@
 From the view of a Markovian sequential decision process,
-human behavior unfolds by integrating action outcomes from stored past events
+human behavior unfolds by integrating action outcomes from policy matrix (mapping from states to actions which maximize expected cumulative reward; see subsection \ref{sec:policymat})
+% stored past events
 and extrapolation to
@@
 for continued refinement of the organism's fate.
+Indeed, a major hurdle in explaining DMN activity from brain-imaging studies has been its very similar engaged across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., dreams during sleep or daytime mind-wandering), and thinking about future scenarios (e.g., prospection or delay discounting).
+The MDP account of DMN function naturally integrates this a-priori diverging mental processes into a common framework.
+% It is this feature of our functional view that we underline in the referred sentence.
 

### 2017-09-17 paper/paper.tex
Update of Fig 3 w/ Bellman equation

@@
   \centering
-  \includegraphics[width=1.\linewidth]{rl.pdf}
+  \includegraphics[width=1.\linewidth]{rl2.pdf}
   \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
-    the agent takes an action by following some policy, receives a consequential reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
+    the agent takes an action by following the policy matrix as updated by the Bellman equation.
+The agent receives a consequential reward and observes the next state.
+The process goes on until interrupted or a goal state is reached.}
   \label{fig:rl}

### 2017-09-18 paper/paper.tex
Update of Fig 3 width
Rephrasing for Reviewer #2

@@
 this network appears to
-simultaneously modulate perception-action cycles and
-entertain probabilistic contemplations
+simultaneously modulate perception-action cycles in the present
+and to support mental travel
 across time, space, and content domains
@@
 optimize the behavior of the organism.
-%
-It has indeed been proposed that
-the human brain's energy budget is largely dedicated to
+
+Evidence form computional modeling in human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
+the human brain is largely dedicated to
 ``the development and maintenance of [a]
@@
 \citep{raichle2005intrinsic}.
-This idea is invigorated by empirical evidence from
-neuroscience experiments \citep{kording2004bayesian, fiser2004small}.
 The present paper proposes a
@@
 step is sufficiently short.
-\begin{wrapfigure}{l}{.4\linewidth}
+\begin{figure}[!h]
   \centering
@@
   \label{fig:rl}
-\end{wrapfigure}
+\end{figure}
 % , defining $s(t)$, the state at time $t$, as the position in state space.

### 2017-09-18 paper/paper.tex
Add chunk about MEG and DMN

@@
 \begin{abstract}
-The default mode network (DMN) is believed to subserve the
+The Default Mode Network (DMN) is believed to subserve the
 baseline mental activity in humans.
@@
 \citep{raichle2001pnas, randy2008, bzdok2015resting}.
-The ``default mode network'' (DMN) continues
+The ``Default Mode Network'' (DMN) continues
 to metabolize large quantities of
@@
 of the biological purpose underlying DMN activity.
+Despite observation in MEG of similar large-scale networks of co-varying spontaneous activity \citep{de2010temporal,brookes2011investigating,baker2014fast}, the link between the fMRI BOLD signal and population-level neural activity is still unclear.
+If those frequency-specific electrophysiological correlations are proposed as complementary to those observed with BOLD \citep{hipp2015bold}, their link to DMN remains challenging \citep{maldjian2014graph}.
+
 What has early been described as the ``stream of consciousness''

### 2017-09-18 paper/paper.tex
update Venn figure caption

--- 2017-09-18 paper/paper.tex
@@
 
-Evidence form computional modeling in human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
+Evidence from computional modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
 the human brain is largely dedicated to
@@
   \includegraphics[width=1\linewidth]{VennDiagram-2017-09-08.pdf}
-  \caption{\textbf{Where does Markov Decision Process fit in the theoretical landscape?}
-  This Venn diagram summarizes the relationship between four existing accounts of DMN: in blue colors, the Free Energy Principle \& Predictive Coding, more related to the ”How” DMN works, and in yellow, the Sentinel \& the Semantic hypotheses, more related to "What" DMN does.
+  \caption{\textbf{Situating Markov Decision Processes among other accounts of default mode function}
+  This Venn diagram summarizes the relationship between four existing explanations of the functional role of the DMN and our present account.
+Viewing empirical findings in the DMN from the MDP viewpoint incorporates important aspects of the free energy principle, predictive coding, sentinal hypothesis, and semantic hypothesis.
+  % more related to "What" DMN does.
 Our account of DMN by Markov Decision Process is at the intersection of those accounts and thus corresponds to a "Process model" meaning a model of both "How" and "What" a process occur.
+  The MDP account may reconcile several strengths of these functional accounts in a process model that simultaneously acknowledges environmental input and behavioral choices as well as the computational and algorithmic properties (How?
+and What?) underlying higher-order control of the organism.
+
   % ; here the mental processes and how it relate to observed behavior.

### 2017-09-18 paper/paper.tex
roundup final sentence
hide appendix

@@
 Indeed, a major hurdle in explaining DMN activity from brain-imaging studies has been its very similar engaged across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., dreams during sleep or daytime mind-wandering), and thinking about future scenarios (e.g., prospection or delay discounting).
-The MDP account of DMN function naturally integrates this a-priori diverging mental processes into a common framework.
+The MDP account of DMN function offers a natural integration of these a-priori diverging mental processes into a common framework.
 % It is this feature of our functional view that we underline in the referred sentence.
@@
 For instance, the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
-
-% \begin{mdframed}
-%   Though our presentation of free-energy principles is thermodynamic in flavor, a completely equivalent
-%   information-theoretic formulation is of course possible, simply by exploiting the information-theoretic interpretation of entropy.
-% \end{mdframed}
-
-\subsection{Helmholtz free-energy and the generative model}
-\begin{table}[H]
-  \begin{tabular}{p{2cm}|p{11cm}}
-         \hline
-         \textbf{symbol}    & \textbf{description}  \\ \hline
-         $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
-         random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
-         $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
-          with the usual convention $0 \log(0) := 0$.\\ \hline
-         $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
-             $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e., $\x = (s, b).$ \\ \hline
-             $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
-             $p_G(.|\x)$ & Generative density for ...\\ \hline
-         $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
-         $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
-         $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
-          to $R$.  Note that $F^G_G = F_G$.\\ \hline
-  \end{tabular}
-  \caption{Table of notations.}
-\end{table}
-Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
+% 
+% % \begin{mdframed}
+% %   Though our presentation of free-energy principles is thermodynamic in flavor, a completely equivalent
+% %   information-theoretic formulation is of course possible, simply by exploiting the information-theoretic interpretation of entropy.
+% % \end{mdframed}
+% 
+% \subsection{Helmholtz free-energy and the generative model}
+% \begin{table}[H]
+%   \begin{tabular}{p{2cm}|p{11cm}}
+%          \hline
+%          \textbf{symbol}    & \textbf{description}  \\ \hline
+%          $\langle X\rangle_p$ & Expectation (a.k.a average, a.k.a mean value) of the
+%          random quantity $X$ w.r.t to the probability density $p$, formally defined by $\langle E\rangle_p := \sum_{z}p(z)X(z)$.\\ \hline
+%          $\mathcal H(p)$ & Information-theoretic entropy of a probability density $p$, formally defined by $\mathcal H(p) := -\sum_{z}p(z)\log(p(z))$,
+%           with the usual convention $0 \log(0) := 0$.\\ \hline
+%          $D_{KL}(q||p)$ & The Kullback-Leibler divergence between the probability densities $q$ and $p$ respectively, formally defined by $D_{KL}(q||p) := \sum_{z}q(z)\log(q(z)/p(z))$.\\ \hline
+%              $\x$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e., $\x = (s, b).$ \\ \hline
+%              $\z$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
+%              $p_G(.|\x)$ & Generative density for ...\\ \hline
+%          $p_R(.|\x)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
+%          $F_G(\x)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\x$. This measures the surprise incured upon observing $\x$ generated by the model $G$.\\ \hline
+%          $F^R_G(\x)$ & Variational Helmholtz free-energy from model $G$
+%           to $R$.  Note that $F^G_G = F_G$.\\ \hline
+%   \end{tabular}
+%   \caption{Table of notations.}
+% \end{table}
+% Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system.
 So,
-\begin{eqnarray}
-  \begin{split}
-    \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(p_G(\x))\sum_{\z}p_G(\z |\x)\\
-    &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
-    =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
-    &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
-    &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
-    &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
-  \end{split}
-  \label{eq:helm}
-\end{eqnarray}
-where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
-\begin{equation}
-  E_G(\z, \x) := -\log(p_G(\z, \x)),
-  \label{eq:gibbs}
-\end{equation}
-%% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
-The last quantity in \eqref{eq:helm} is nothing but \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
-\begin{equation}
-  F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
-\end{equation}
-%% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
-Thus,
-\begin{fact}
-  Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
-\end{fact}
-
-The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
+% \begin{eqnarray}
+%   \begin{split}
+%     \text{generative surprise } &= -\log(p_G(\x)) = -\log(p_G(\x)) \times 1 = -\log(p_G(\x))\sum_{\z}p_G(\z |\x)\\
+%     &= -\sum_{\z}p_G(\z, \x)\log(p_G(\x))
+%     =-\sum_{\z}p_G(\z |\x)\log(p_G(\z, \x)/p_G(\z|\x))\\
+%     &= \sum_{\z}p_G(\z |\x)\log(p_G(\z|\x))-\sum_{\z}p_G(\z |\x)\log  (p_G(\z, \x))\\
+%     &= -\langle \log  (p_G(., \x)) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))\\
+%     &= \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x))
+%   \end{split}
+%   \label{eq:helm}
+% \end{eqnarray}
+% where $E_G(\z, \x)$ is the energy at \textit{macrostate} $\z$ of a fictive thermodynamic system defined by setting
+% \begin{equation}
+%   E_G(\z, \x) := -\log(p_G(\z, \x)),
+%   \label{eq:gibbs}
+% \end{equation}
+% %% In the above, the variable $\z$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
+% The last quantity in \eqref{eq:helm} is nothing but \textit{Helmholtz free-energy} (at unit temperature!), defined formally by
+% \begin{equation}
+%   F_G(\x) := \langle E_G(., \x) \rangle_{p_G(. |\x)} - \mathcal H(p_G(. |\x)).
+% \end{equation}
+% %% , modelled as the physical system for which $p_G(\z|\x)$ dictates the occupation probabilities of macrostates $\z$
+% Thus,
+% \begin{fact}
+%   Generative surprise and generative Helmholtz free-energy are different views on exactly the same object.
+% \end{fact}
+% 
+% The goal of the brain is then to optimize over the generative model $G$: to iteratively or analytically modify the generative density $p_G(.|\x)$, so as to minimize surprise.
 It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(\x)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations.
@@
 1995, is at the heart of the general free-energy principle that we will introduce shortly.
-
-\subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
-In this subsection, we will present an insightful upper bound for the generative surprise (i.e., generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
+% 
+% \subsection{Variational Helmholtz free-energy and the bottom-up recognition sub-model}
+% In this subsection, we will present an insightful upper bound for the generative surprise (i.e., generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy.
 As an avant-gout of what is to come shorty, let's just note that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.
-
-Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
-distribution
-\begin{equation}
-  p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
-\end{equation}
-where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
-Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
-\begin{equation}
-F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
-\end{equation}
-Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
-it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
-density\footnote{conditioning in $p_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
-  state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
-\begin{equation}
-  \begin{split}
-    F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{p_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
-    &= \langle E_G(., \x)\rangle_{p_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
-    &= F^R_G(\x) - D_{KL}(p_R(.|\x) || p_G(.|\x)),
-  \end{split}
-  \label{eq:fe}
-\end{equation}
-where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
-\begin{equation}
-  F^R_G(\x) := \langle E_G(., \x)\rangle_{p_R(.|\x)} - \mathcal H(p_R(.|\x))
-\end{equation}
-and $D_{KL}(p_R(.|\x) || p_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
-
-\subsection{A general free-energy principle}
-We can resume the situation as follows\footnote{Where we have used the fact that KL divergence is always nonnegative.}:
-\begin{mdframed}
-\begin{equation}
-  \begin{split}
-    \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
-    &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(p_R(.|\x) || p_G(.|\x))}_{\text{complexity}} \\
-    &\le F^R_G(\x),
-    \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
-    \end{split}
-\end{equation}
-\end{mdframed}
-%% In fact, we have the following theorem
-%% \begin{theorem}
-%%   It holds that
-%%   $$
-%%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
-%%   i.e
-%%   \begin{equation}
-%%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
-%%     \end{equation}
-%% \end{theorem}
-
-\subsection{Helmholtz machines and the wake-sleep algorithm}
+% 
+% Invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
+% distribution
+% \begin{equation}
+%   p_G(\z|\x) = \frac{p_G(\z|\x)}{p_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)} = \frac{\exp(-E_G(\z, \x))}{Z_G(\x)},
+% \end{equation}
+% where $Z_G(\x) := \log(p_G(\x)) = \sum_{\z'}\exp(-E_G(\z',\x))$, the normalizing \textit{partition function} for the model \ref{eq:gibbs}.
+% Whence, for any macrostate $\z$, we have $p_G(\x) = Z_G(\x) = \exp(-E_G(\z, \x)) / p_G(\z|\x)$, and so it holds that
+% \begin{equation}
+% F_G(\x)\overset{\eqref{eq:helm}}{=} -\log(p_G(\x)) = -\log(Z_G(\x)) = E_G(\z, \x) + \log(p_G(\z|\x)).
+% \end{equation}
+% Now, in the above equation, the LHS only depends on the generative model $G$ and the data point $\x$:
+% it doesn't depend on the hidden variable $\z$, etc. So, taking expectations w.r.t an arbitrary
+% density\footnote{conditioning in $p_R(.|\x)$ is because this density is selected from a world in which the sensory inputs and internal brain
+%   state vector $\x$ is assumed already observed.} $p_R(.|\x)$ yields
+% \begin{equation}
+%   \begin{split}
+%     F_G(\x) &= -\log(Z_G(\x)) = \langle E_G(., \x)\rangle_{p_R(.|\x)} + \sum_{\z}p_R(\z|\x)\log(p_G(\z|\x))\\
+%     &= \langle E_G(., \x)\rangle_{p_R(.|\x)} - \mathcal H(p_R(.|\x)) - \sum_{\z}p_R(\z|\x)\log(p_R(\z|\x)/p_G(\z|\x))\\
+%     &= F^R_G(\x) - D_{KL}(p_R(.|\x) || p_G(.|\x)),
+%   \end{split}
+%   \label{eq:fe}
+% \end{equation}
+% where $F^R_G(\x)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
+% \begin{equation}
+%   F^R_G(\x) := \langle E_G(., \x)\rangle_{p_R(.|\x)} - \mathcal H(p_R(.|\x))
+% \end{equation}
+% and $D_{KL}(p_R(.|\x) || p_G(.|\x))$ is the Kullback-Leibler divergence between the $p_R(.|\x)$ and the generative density $p_G(.|\x)$. Note that $F^G_G = F_G$.
+% 
+% \subsection{A general free-energy principle}
+% We can resume the situation as follows\footnote{Where we have used the fact that KL divergence is always nonnegative.}:
 % \begin{mdframed}
-  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e., input from lower more primitive layers).
+% \begin{equation}
+%   \begin{split}
+%     \text{generative surprise } &:= -\log(p_G(\x)) = F_G(\x) \\
+%     &=\underbrace{F^R_G(\x)}_{\text{accuracy}} - \underbrace{D_{KL}(p_R(.|\x) || p_G(.|\x))}_{\text{complexity}} \\
+%     &\le F^R_G(\x),
+%     \text{ with equality if }p_R(\z|\x) = p_G(\z|\x)\text{ for all } \z
+%     \end{split}
+% \end{equation}
+% \end{mdframed}
+% %% In fact, we have the following theorem
+% %% \begin{theorem}
+% %%   It holds that
+% %%   $$
+% %%   \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
+% %%   i.e
+% %%   \begin{equation}
+% %%     \min_{G}F_G(\x) = \min_{G,R}F^R_G(\x)
+% %%     \end{equation}
+% %% \end{theorem}
+% 
+% \subsection{Helmholtz machines and the wake-sleep algorithm}
+% % \begin{mdframed}
+%   \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data (i.e., input from lower more primitive layers).
 Precisely
-$$
-p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
-p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
-$$
+% $$
+% p_G(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_G(\z_k^{(l)} | \x),\hspace{1cm}
+% p_R(\z^{(l)}|\x) = \Pi_{k=1}^{h_l}p_R(\z_k^{(l)} |\x)
+% $$
+% % \end{mdframed}
+% 
+% \subsection{Friston's active-inference and agency}
+% \label{sec:friston}
+% This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
+% 
+% The following critics can be made:
+% \begin{itemize}
+%   \item
+%     As noted by Dayan et al.
+(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
+%   \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
+%   \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
+%     technical difficulty in getting derivatives of loss function w.r.t
+%     recognition weights $\W^R$).
+%   \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
+%     \end{itemize}
+% 
+% 
+% \subsection{Minimizing free-energy via backprop: variational auto-encoders}
+% 
+% % \subsection{Free-energy minimization via variational auto-encoders}
+% \label{sec:vae}
+% Here, we present a way to alleviate some conceptual and computational issues with the
+% free-energy framework presented thus far, by using the recent \textit{variational auto-encoder} (VAE) theory \citep{kingma2013auto}. Define the data-dependent
+% auxiliary random function
+% \begin{eqnarray}
+%   f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
+% \end{eqnarray}
+% Then we can rewrite the
+% variational free-energy as
+% \begin{eqnarray*}
+%   \begin{split}
+%     F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{p_R(.|\x)}\\
+%     &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{p_R(.|\x)}\\
+%     &= -\langle f_{G,R}\rangle_{p_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
+%     \end{split}
+% \end{eqnarray*}
+% 
+% \begin{mdframed}
+%   \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
 % \end{mdframed}
-
-\subsection{Friston's active-inference and agency}
-\label{sec:friston}
-This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...
-
-The following critics can be made:
-\begin{itemize}
-  \item
-    As noted by Dayan et al.
-(\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence assumption which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
-  \item A drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recognitiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.).
-  \item Also, note that the wake-sleep algorithm doesn't do backprop! This is due to
-    technical difficulty in getting derivatives of loss function w.r.t
-    recognition weights $\W^R$).
-  \item This difficulty was removed in the 2010s by \citep{kingma2013auto}, an other groups, via a ``reparametrization trick''.
-    \end{itemize}
-
-\subsection{Minimizing free-energy via backprop: variational auto-encoders}
-
-% \subsection{Free-energy minimization via variational auto-encoders}
-\label{sec:vae}
-Here, we present a way to alleviate some conceptual and computational issues with the
-free-energy framework presented thus far, by using the recent \textit{variational auto-encoder} (VAE) theory \citep{kingma2013auto}. Define the data-dependent
-auxiliary random function
-\begin{eqnarray}
-  f_{G,R}(., \x) :\z \mapsto \log(p_G(\z,\x)) - \log(p_R(\z|\x)).
-\end{eqnarray}
-Then we can rewrite the
-variational free-energy as
-\begin{eqnarray*}
-  \begin{split}
-    F_G^R(\x) &:= \langle E_G(., \x) \rangle_{p_R(. |\x)} - \mathcal H(p_R(. |\x)) = \langle E_G(., \x) + \log(p_R(.|\x))\rangle_{p_R(.|\x)}\\
-    &=\langle -\log(p_G(.,\x)) + \log(p_R(.|\x))\rangle_{p_R(.|\x)}\\
-    &= -\langle f_{G,R}\rangle_{p_R(.|\x)} \approx -\frac{1}{M}\sum_{m=1}^Mf_{G,R}(\z^{(m)}), \text{ with }\z^{(1)},\ldots,\z^{(M)} \sim p_R(.|\x), \text{ and }M \rightarrow \infty.
-    \end{split}
-\end{eqnarray*}
-
-\begin{mdframed}
-  \textbf{Problem:} How do we sample from the recognition density $p_R(.|\x)$ in such a way that the sampling process is differentiable w.r.t the weights of the recognition network $\W^R$ ?
-\end{mdframed}
-
-% \begin{figure}[!tb]
-%   \centering
-%   \includegraphics[width=.8\linewidth]{vae.pdf}
-%   \caption{Variational autoencoders...}
-%   \label{fig:vae}
-% \end{figure}
-
-\paragraph{Solution: the reparametrization trick.}
-\begin{itemize}
-\item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
-\item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
-  \begin{itemize}
-  \item results in a sample $\z \sim p_R(.|\x)$, from the correct
-    posterior
-    \end{itemize}
-\end{itemize}
-The mapping $g$ should be taught of as a ``blurring'' function which produces noisy versions $\z$,
-called \textit{sensations}, of the true world state $\x$.
-The result is a scheme for training DBNs via good-old backprop!
-Refer to Fig. vae.pdf.
-% \ref{fig:vae}.
-Some examples of the reparametrization trick for a number of
-choices of the posterior distribution are given in Tab. \ref{tab:rptrick}.
-
-\begin{table}[H]
-  \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
-         \hline
-         Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
-         Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
-         Student’s t, Logistic, Uniform, Triangular, \ldots \\ \hline
-         Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
+% 
+% % \begin{figure}[!tb]
+% %   \centering
+% %   \includegraphics[width=.8\linewidth]{vae.pdf}
+% %   \caption{Variational autoencoders...}
+% %   \label{fig:vae}
+% % \end{figure}
+% 
+% \paragraph{Solution: the reparametrization trick.}
+% \begin{itemize}
+% \item Choose $\epsilon \sim p_{\text{noise}}$ (noise distribution, independent of $\W^R$!)
+% \item Set $\z = g(\W^{R}, \x, \epsilon)$, where $g$ is an appropriate class $\mathcal C^1$ function
+%   \begin{itemize}
+%   \item results in a sample $\z \sim p_R(.|\x)$, from the correct
+%     posterior
+%     \end{itemize}
+% \end{itemize}
+% The mapping $g$ should be taught of as a ``blurring'' function which produces noisy versions $\z$,
+% called \textit{sensations}, of the true world state $\x$.
+% The result is a scheme for training DBNs via good-old backprop!
+% Refer to Fig. vae.pdf.
+% % \ref{fig:vae}.
+% Some examples of the reparametrization trick for a number of
+% choices of the posterior distribution are given in Tab. \ref{tab:rptrick}.
+% 
+% \begin{table}[H]
+%   \begin{tabular}{p{1.7cm}|p{1.5cm}|p{1.9cm}|p{2.1cm}|p{5cm}}
+%          \hline
+%          Posterior & $p_R(.|\x)$ & noise & $g(\W^R,\x,\epsilon)$ & Also \\ \hline
+%          Normal & $\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\mu + \sigma \odot \epsilon$ & Location-scale family: Laplace, Elliptical,
+%          Student’s t, Logistic, Uniform, Triangular, \ldots \\ \hline
+%          Exponential & $\mathcal \exp(\lambda)$ & $\epsilon \sim \mathcal U([0, 1))$ & $-\log(1-\epsilon)/\lambda$ & Invertible CDF: Cauchy, Logistic, Rayleigh, Pareta, Weibull, Reciprocal, Gompert, Gumbel, Erlan, ...
 \\ \hline
-         Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
-  \end{tabular}
-  \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
-  \label{tab:rptrick}
-\end{table}
-
-For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
-\begin{equation}
-  p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
-\end{equation}
-
-\subsection{A thermodynamic model for bounded rationality, aka robust
-  optimality}
-\begin{itemize}
-  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
-  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
-  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
-    \end{itemize}
-The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
-\paragraph{Utility functions and conjugate pairs.}
-  Let $(\Omega, \mathcal F, P)$ be a probablity space.
+%          Other & $\log\mathcal N(\mu,\sigma)$ & $\epsilon \sim \mathcal N(0, 1)$ & $\exp(\mu + \sigma \odot \epsilon)$ & Gamma, Dirichlet, Beta, Chi-squared, F, ... \\ \hline
+%   \end{tabular}
+%   \caption{Reparametrization trick \citep{kingma2013auto} for a variety of models.}
+%   \label{tab:rptrick}
+% \end{table}
+% 
+% For fixed $p_G(.|\x)$, the optimal choice for $p_R(.|\x)$ is given analytically by
+% \begin{equation}
+%   p_R(\z|\x) \propto p_G(\z|\x)\exp(-\Delta E_{G \rightarrow R}(\z, \x)).
+% \end{equation}
+% 
+% \subsection{A thermodynamic model for bounded rationality, aka robust
+%   optimality}
+% \begin{itemize}
+%   \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
+%   \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
+%   \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
+%     \end{itemize}
+% The material presented is a revisit of \citep{braun2011path}. See also \citep{ortega2013thermodynamics}
+% \paragraph{Utility functions and conjugate pairs.}
+%   Let $(\Omega, \mathcal F, P)$ be a probablity space.
 A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
-  \begin{itemize}
-  \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.
-  \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
-    $A,B \in \mathcal F$.
-  \item monotonicity: $f_{U}$ is strictly increasing.
-    \end{itemize}
-
-\begin{theorem}
-  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
-  \begin{equation}
-    f = \alpha \log(.),
-    \label{eq:boltzmann}
-  \end{equation}
-  where $\alpha > 0$.
-\end{theorem}
-
-\textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
-
-Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
-
-\paragraph{Example.}
-Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
-  \begin{equation}
-    P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
-  \end{equation}
-  where
-
-  \begin{equation}
-    Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
-  \end{equation}
-  is a normalization constant called the \textit{partition function} of $U$.
-  It's not hard to see that $U$ and the $P$ above form a conjugate pair.
-
-  \paragraph{Free-utility functional.}
-    Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
+%   \begin{itemize}
+%   \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.
+%   \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
+%     $A,B \in \mathcal F$.
+%   \item monotonicity: $f_{U}$ is strictly increasing.
+%     \end{itemize}
+% 
+% \begin{theorem}
+%   The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
+%   \begin{equation}
+%     f = \alpha \log(.),
+%     \label{eq:boltzmann}
+%   \end{equation}
+%   where $\alpha > 0$.
+% \end{theorem}
+% 
+% \textbf{XXX: equation \eqref{eq:boltzmann} above looks like Boltzmann's formula (on his gravestone...)!}
+% 
+% Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.
+% 
+% 
+% \paragraph{Example.}
+% Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
+%   \begin{equation}
+%     P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
+%   \end{equation}
+%   where
+% 
+%   \begin{equation}
+%     Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)
+%   \end{equation}
+%   is a normalization constant called the \textit{partition function} of $U$.
+%   It's not hard to see that $U$ and the $P$ above form a conjugate pair.
+% 
+%   \paragraph{Free-utility functional.}
+%     Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$.
 Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
-    \begin{equation}
-      J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
-    \label{eq:free_u}
-    \end{equation}
-    where
-    \begin{equation}
-      \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
-    \end{equation}
-    is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
-  \begin{equation}
-    J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
-  \end{equation}
-  In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
-  \begin{equation}
-    J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
-    \end{equation}
-
-  \paragraph{The free-energy / utility principle (of Friston ?).}
-  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
+%     \begin{equation}
+%       J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
+%     \label{eq:free_u}
+%     \end{equation}
+%     where
+%     \begin{equation}
+%       \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
+%     \end{equation}
+%     is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
+%   \begin{equation}
+%     J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
+%   \end{equation}
+%   In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
+%   \begin{equation}
+%     J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
+%     \end{equation}
+% 
+%   \paragraph{The free-energy / utility principle (of Friston ?).}
+%   We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it.
 Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$.
@@
 The difference in free-utility is then
-  \begin{equation}
-    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
-    \label{eq:free}
-  \end{equation}
-  where
-  \begin{equation}
-    D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
-  \end{equation}
-  is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
-In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
-  The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
-a trade-off between an expected utility $\Delta U$ (first term) and
-the information cost of transforming Pi
-into $P_{\text{fin}}$ (second
-term). In this interpretation $P_{\text{ini}}$ represents an initial choice
-probability or policy, which includes the special case of the
-uniform distribution where the decision-maker has initially no
-preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
-considers the utility constraint U∗
-that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
-\begin{equation}
-P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
-  \end{equation}
-
-By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
-
-\begin{equation}
-  P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
-\end{equation}
-
-Two limit cases are worth considering.
-\begin{itemize}
-\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
-  $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
-  where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
-  \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
-\end{itemize}
-
-To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
+%   \begin{equation}
+%     \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
+%     \label{eq:free}
+%   \end{equation}
+%   where
+%   \begin{equation}
+%     D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
+%   \end{equation}
+%   is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
+% In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
+%   The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
+% a trade-off between an expected utility $\Delta U$ (first term) and
+% the information cost of transforming Pi
+% into $P_{\text{fin}}$ (second
+% term). In this interpretation $P_{\text{ini}}$ represents an initial choice
+% probability or policy, which includes the special case of the
+% uniform distribution where the decision-maker has initially no
+% preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
+% considers the utility constraint U∗
+% that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
+% \begin{equation}
+% P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
+%   \end{equation}
+% 
+% By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution
+% 
+% \begin{equation}
+%   P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
+% \end{equation}
+% 
+% Two limit cases are worth considering.
+% \begin{itemize}
+% \item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
+%   $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
+%   where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
+%   \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
+% \end{itemize}
+% 
+% To conclude this section, let's note that \citep{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework.
 For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...
-
-\subsection{GANs and other likelihood-free methods}
-\ldots
+% 
+% 
+% \subsection{GANs and other likelihood-free methods}
+% \ldots
 

### 2017-09-18 paper/paper.tex
fix Latex compile error
homogenize the 4 func acc

@@
 Viewing empirical findings in the DMN from the MDP viewpoint incorporates important aspects of the free energy principle, predictive coding, sentinal hypothesis, and semantic hypothesis.
-  % more related to "What" DMN does.
-Our account of DMN by Markov Decision Process is at the intersection of those accounts and thus corresponds to a "Process model" meaning a model of both "How" and "What" a process occur.
   The MDP account may reconcile several strengths of these functional accounts in a process model that simultaneously acknowledges environmental input and behavioral choices as well as the computational and algorithmic properties (How?
-and What?) underlying higher-order control of the organism.
-
-  % ; here the mental processes and how it relate to observed behavior.
-For each account, 4 icons describe the core concepts of Time horizon, Neural plasticity, Hierarchy, and Behavioral goal.
-The "Time horizon" should be understood as a weighting of the expected future outcomes shorter to longer terms into present decision making.
-Then, "online monitoring" of the environment means detecting and processing salient cues in the sensory environment.
-The "Predictive Coding" is understood here as the joint minimization of feedforward prediction error and feedback modulation to optimize neural processing; there is thus no notion of an agent with behavioral goal.
-  }
+and What?) underlying higher-order control of the organism.}
   \label{fig:VennDiagram}
@@
 % CRITICISM
-\paragraph{Main criticisms.}
+% \paragraph{Main criticisms.}
 Despite its popularity, criticism against the FEP has arisen over the years,
@@
 
-\paragraph{Links between FEP and RL.}
+% \paragraph{Links between FEP and RL.}
 We note that FEP and active inference can be reframed in terms of

### 2017-09-18 paper/paper.tex
R1 material

@@
 of the biological purpose underlying DMN activity.
-Despite observation in MEG of similar large-scale networks of co-varying spontaneous activity \citep{de2010temporal,brookes2011investigating,baker2014fast}, the link between the fMRI BOLD signal and population-level neural activity is still unclear.
-If those frequency-specific electrophysiological correlations are proposed as complementary to those observed with BOLD \citep{hipp2015bold}, their link to DMN remains challenging \citep{maldjian2014graph}.
+Despite observation of similar large-scale networks of co-varying spontaneous activity in MEG investigations \citep{de2010temporal,brookes2011investigating,baker2014fast}, the link between the fMRI BOLD signal and population-level neural activity is still unclear.
+If those frequency-specific electrophysiological correlations are proposed as complementary to those observed with BOLD \citep{hipp2015bold}, their role in DMN function remains challenging \citep{maldjian2014graph}.
 
@@
 particularly challenging because
-this network appears to
+this brain network appears to
 simultaneously modulate perception-action cycles in the present
@@
 Conversely,
-it is known that, the more the world is easy to predict,
+it has been empirically shown that, the more the world is easy to predict,
 the more human mental activity
 becomes detached from the actual sensory environment
-\citep{antrobus1966studies, pope1978regulation}.
+\citep{antrobus1966studies, pope1978regulation, mason2007, weiss2006}.
 Without requiring explicit awareness,
@@
 
-Evidence from computional modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
+Evidence from computational modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
 the human brain is largely dedicated to
@@
 \begin{mdframed}
+  \vspace{1em}
   \textbf{Box 1: Hypotheses for testing the MDP account of DMN function}
-  \vspace{1em}
 \begin{enumerate}
@@
       positively coupled with the vmPFC and HC. The colors are indicative of the (\textcolor{red}{red} = positive,
-      \textcolor{blue}{blue} = negative) and size (the lighter the higher) of the regression coefficients.
+      \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
   Code and data for reproduction and visualization:
@@
 % neuroscience of Bellman equation
+  \vspace{2em}
   \textbf{Box 2: Neural correlates of the Bellman equation in the DMN}
-  \vspace{1em}
-
-\hspace{-1.5em}Relating decomposition of consecutive action choices by the Bellman equation
+% \hspace{-1.5em}
+Relating decomposition of consecutive action choices by the Bellman equation
 to neuroscience,
@@
 
+\begin{wrapfigure}{r}{.37\linewidth}
+  \centering
+  \includegraphics[width=1.\linewidth]{darkroom_resized.png}
+  \caption{\textbf{The dark room experiment.} An intelligent agent situated
+  in a light-deprived closed space can be used
+  as a thought experiment for the complete absence
+  of external sensory input.
+  % The free-energy principle theory paradoxally predicts that such an agent would remain quietly in this room, such it perfectly predict all sensory input from such an environment.
+  % However such behavior is not observed in humans or any animals: the agent would strive to leave such an environment in search of more ``interesting'' worlds.
+  }
+  \label{fig:darkroom}
+\end{wrapfigure}
 On another front, since theories based on the FEP \citep{friston2010free,fristonAIorRL}
@@
 be \textit{surprised} by such a \textit{lack of surprise}. According to this argument, a dark room would paradoxically correspond to a state of particularly high surprise.
-\begin{wrapfigure}{r}{.37\linewidth}
-  \centering
-  \includegraphics[width=1.\linewidth]{darkroom_resized.png}
-  \caption{\textbf{The dark room experiment.} An intelligent agent situated
-    in a light-deprived closed space can be used
-    as a thought experiment for the complete absence
-    of external sensory input.
-    % The free-energy principle theory paradoxally predicts that such an agent would remain quietly in this room, such it perfectly predict all sensory input from such an environment.
-    % However such behavior is not observed in humans or any animals: the agent would strive to leave such an environment in search of more ``interesting'' worlds.
-  }
-  \label{fig:darkroom}
-\end{wrapfigure}
 Driven by surprise-minimization objective, the FEP agent would eventually bootstrap itself out of such saddle points to explore more interesting parts of the environment.

### 2017-10-12 paper/paper.tex
some corrections on tex + improved rl2.odp figure

@@
 The main algorithm for minimizing free energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
-\citep{dayan1995helmholtz}. As these P. Dayan and G. Hinton  noted,
+\citep{dayan1995helmholtz}. As P. Dayan and G. Hinton  noted,
 a crucial drawback of the wake-sleep algorithm (and therefore of theories like the FEP \citep{friston2010free} based on it) is that it involves a pair
@@
 because KL divergence is not symmetric in its arguments.
-The brain may therefore be unlikely to implement a variant of the wake-sleep algorithm.
-The recent theory of
-\textit{variational auto-encoders} (VAEs) \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm. VAEs
-overcome a number of the technical limits of the wake-sleep algorithm
-by using a reparametrization maneuver, which makes it possible to do differential calculus on random sampling procedures without exploding variance. As a result, unlike the wake-sleep algorithm
-for minimizing free energy, VAEs can be efficiently trained via back-propagation of prediction errors.
-
+The brain is therefore unlikely to implement the wake-sleep algorithm or a variant thereof.
 \begin{wrapfigure}{r}{.37\linewidth}
@@
 \end{wrapfigure}
+
+The recent theory of
+\textit{variational auto-encoders} (VAEs) \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm. VAEs
+overcome a number of the technical limits of the wake-sleep algorithm
+by using a reparametrization maneuver, which makes it possible to do differential calculus on random sampling procedures without exploding variance. As a result, unlike the wake-sleep algorithm
+for minimizing free energy, VAEs can be efficiently trained via back-propagation of prediction errors.
+
 On another front, since theories based on the FEP \citep{friston2010free,fristonAIorRL}
@@
 intelligent agents would search for light sources to \textit{explore}
-its surroundings or leave it.
+the surroundings or leave the exit the room.
 Defenders of the FEP have retorted by advancing the ``full package'' \citep{darkroom2012}:

### 2017-11-28 paper/paper.tex
Reshape the why

@@
 \end{definition*}
-
-\paragraph*{Why MDPs ?} At this point, it is righteous to question the applicability of MPDs as a model for something as complex as human behavior.
-Indeed a critique could argue that Markov chains are state-machines without memory.
-These may describe insects, or other behaviorally primitive beings like ATM machines, insects, etc., but won't be sufficient to describe human beings, as for example, human behavior is heavily influenced not only by the current state but by past states as well (memory).
-
-\paragraph*{Answer:}
-\begin{itemize}
-  \item First, let us emphasize that the brain in general, and the DMN in particular,
-is a physical system governed by the laws of
-physics and can be formally described
-by Markov processes at a sufficiently coarse scale.
-It has indeed been previously proposed \citep{tegmark2016improved} that
-any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
-step is sufficiently short.
-
-\item Rational probabilistic planning can be reformulated
+Rational probabilistic planning can be reformulated
 as a standard memory Markov process by simply expanding the
@@
 
-\item Also, it should be noted that MDPs have been successfully used as neural model for animal behavior.
-One can mention ~\citep{Rao2010} (random dots), ~\citep{Valton2011} (mal-adaptive decision-making in a rat version of a gambling task) just to name a few.
-
-\item Successes in AI/ML:
-  \begin{itemize}
-    \item Robotics: \citep{ng2004,abbeel2004}
-    \item Games and other: ~\citep{mnih2015,silver2016mastering,pritzel2017neural}
-\end{itemize}
-
-  \item Algorithmic trading.
-In financial trading, the main actions which the agent (human or computer acting on behalf of a human operator) must execute strategically are: sell (sell a given quantity of a commodity), buy (by a given quantity of a commodity), do nothing.
-According to how the market responds, a definite gain / loss is recorded as a consequence of the executed action, perhaps with some delay.
-Over the years, research in algorithmic trading has identified MDPs as a powerful framework for modelling this behavioral process  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010}, for which dynamic-programming based algorithms can be employed to produce numerical solutions.
-~\citep{abergel2017} is an excellent reference with the mathematical details on how MDPs work in finance.
-
-\end{itemize}
+\paragraph*{Why MDPs ?} One may wonder whether MDP models are applicable to something as complex as human behavior.
+For instance, financial trading is largely a manifestation of strategic decision-making of interacting human agents.
+According to how the market responds, the agent incurs definite gain or loss as environmental feedback of the executed financial actions.
+Recent research on automatizing market exchanges by algorithmic trading has successfully used MDPs as a framework for modeling these sophisticated behavioral process  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010}.
+~\citep{abergel2017} is an excellent reference with the mathematical details on how MDPs can be readily exploited in the financial world.
+%
+MDPs have also been effective as a behavioral model in robotics \citep{ng2004,abbeel2004} and in challenging multistep strategy games ~\citep{mnih2015,silver2016mastering,pritzel2017neural}.
+%
+As such, we aim to expand MDPs as a useful model from "online" decision-making to the realms of "offline" mental operations most associated with the DMN.
+
+% Indeed a critique could argue that Markov chains are state-machines without memory.
+These may describe insects, or other behaviorally primitive beings like ATM machines, insects, etc., but won't be sufficient to describe human beings, as for example, human behavior is heavily influenced not only by the current state but by past states as well (memory).
+
+% \paragraph*{Answer:}
+% \begin{itemize}
+%   \item First, let us emphasize that the brain in general, and the DMN in particular,
+% is a physical system governed by the laws of
+% physics and can be formally described
+% by Markov processes at a sufficiently coarse scale.
+% It has indeed been previously proposed \citep{tegmark2016improved} that
+% any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
+% step is sufficiently short.
+
+% \end{itemize}
 
@@
 Such a surprise-guided reinforcement learning scheme has previously been advocated
-under the equivalent framework of energy-based reinforcement-learning\footnote{These methods assume that the posterior distribution  of the hidden states $\z$ given the visible states $\x$ factorizes, leading to so-called RBMs --\textit{Restricted Boltzmann Machines}.
-One notes, that even with these simplifications, such a model can be very hard to train.} ~\citep{sallans2004,elfwing2016free} and information compression \citep{schmidhuber2010,mohamed2015}.
+under the equivalent framework of energy-based reinforcement-learning
+% \footnote{These methods assume that the posterior distribution  of the hidden states $\z$ given the visible states $\x$ factorizes, leading to so-called RBMs --\textit{Restricted Boltzmann Machines}.
+One notes, that even with these simplifications, such a model can be very hard to train.}
+~\citep{sallans2004,elfwing2016free} and information compression \citep{schmidhuber2010,mohamed2015}.
 Nevertheless, minimization of surprise quantities alone

### 2017-11-28 paper/paper.tex
modestify title
fixed some bugoos

--- 2017-10-12 paper/paper.tex
@@
 The main algorithm for minimizing free energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
-\citep{dayan1995helmholtz}. As P. Dayan and G. Hinton  noted,
+\citep{dayan1995helmholtz}. As P. Dayan and G. Hinton themselves noted,
 a crucial drawback of the wake-sleep algorithm (and therefore of theories like the FEP \citep{friston2010free} based on it) is that it involves a pair
@@
   \caption{\textbf{The dark room experiment.} An intelligent agent situated
-  in a light-deprived closed space can be used
+  in a light-deprived closed space is used
   as a thought experiment for the complete absence

### 2017-11-28 paper/paper.tex

--- 2017-11-28 paper/paper.tex
@@
 \title{Dark Control:\\
-       A Unified Account of Default Mode Function\\
+       Towards a Unified Account of Default Mode Function\\
        by Markov Decision Processes}

### 2017-11-28 paper/paper.tex
Tweak refs + Move Box1 + Change Box numbers

--- 2017-11-28 paper/paper.tex
@@
 experiments (Box 1).
-
-\begin{mdframed}
-  \vspace{1em}
-  \textbf{Box 1: Hypotheses for testing the MDP account of DMN function}
-\begin{enumerate}
-\item Experiment (Humans): We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
-During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigms.
-We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, and the more the fMRI signals in the reward circuits should be independent of the reward context
-in the sensory environment.
-
-\item Experiment (Humans): We hypothesize a functional dissociation between computations for behavioral policy versus adapting stimulus-value associations as these are conceivably implemented in different subsystems of the DMN.
-We first expect that fMRI signals in the right temporo-parietal junction can predict behavioral changes caused by adaptation in the action choice tendencies (policy matrix) related to non-value-related prediction error.
-Second, fMRI signals in the ventromedial prefrontal cortex should predict behavioral changes caused by adaptation in value estimation (value matrix) due to reward-related stimulus-value association.
-We finally predict that fMRI signals in the posteromedial cortex, as a potential global information integrator, is able to predict shifts in overt behavior based on previous adaptations in policy or value estimation.
-
-\item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is necessary for choice behavior in mice.
-Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before make analogous choices in the future.
-Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal disruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
-
-\end{enumerate}
-\end{mdframed}
 
@@
   \vspace{2em}
-  \textbf{Box 2: Neural correlates of the Bellman equation in the DMN}
+  \textbf{Box 1: Neural correlates of the Bellman equation in the DMN}
 % \hspace{-1.5em}
@@
 % \end{itemize}
+
+\begin{mdframed}
+  \vspace{1em}
+  \textbf{Box 2: Hypotheses for testing the MDP account of DMN function}
+\begin{enumerate}
+\item Experiment (Humans): We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
+During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigms.
+We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, and the more the fMRI signals in the reward circuits should be independent of the reward context
+in the sensory environment.
+
+\item Experiment (Humans): We hypothesize a functional dissociation between computations for behavioral policy versus adapting stimulus-value associations as these are conceivably implemented in different subsystems of the DMN.
+We first expect that fMRI signals in the right temporo-parietal junction can predict behavioral changes caused by adaptation in the action choice tendencies (policy matrix) related to non-value-related prediction error.
+Second, fMRI signals in the ventromedial prefrontal cortex should predict behavioral changes caused by adaptation in value estimation (value matrix) due to reward-related stimulus-value association.
+We finally predict that fMRI signals in the posteromedial cortex, as a potential global information integrator, is able to predict shifts in overt behavior based on previous adaptations in policy or value estimation.
+
+\item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is necessary for choice behavior in mice.
+Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before make analogous choices in the future.
+Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal disruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
+
+\end{enumerate}
+\end{mdframed}
 

### 2017-11-28 paper/paper.tex
Add embryo of two new experiments in Box 2

@@
 Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal disruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
+
+\item Experiment (Humans): We hypotherize that time horizon should be modulated by different factors such as age, stress, and impulsivity.
+Using a temporal discounting task, we can measure how the time horizon could be affected at the behavioral level, and then trace back its potential representation at the neural level.
+Experimental design can be between groups (by age or with a impulsive population like chronic gamblers or drug addicts), or within group using a stress.
+
+% To exam-ine this question, Fellows and Farah (2005) investigated the effects of dorsolateral frontal (DLF) and ventromedial frontal (VMF) lobe damage on two aspects of temporal foresight—namely, temporal discounting and the futuretime perspective (measuredby an adaptation of Wallace’s [1956] list of future life events).
+
+\item Experiment (Animals): how synaptic epigenesis between DMN nodes can form a neural correlate of the policy matrix update? % Simulations:
 

### 2017-12-02 paper/paper.tex
added some flesh to MF / BF bs

--- 2017-11-28 paper/paper.tex
@@
 \subsection{Markov decision processes}
-Model-free RL has had great success in many real-world problems,
-including robotics \citep{ng2004,abbeel2004},
+Reinforcement learning (RL) has had great success in many real-world problems, including robotics \citep{ng2004,abbeel2004},
 super-human performance in complex video games \citep{mnih2015},
@@
 
-Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
-it has been proposed \citep{gershman2015computational} that a core property of human intelligence  underlie improvement
-of expected utility outcomes as a strategy for action choice in uncertain
-environments, a situation perfectly captured by the formalism of MDPs.
-It has also long been proposed \citep{dayan2008decision} that there
-is a rather direct mapping of model-free RL learning algorithms
-onto aspects of the brain.
-The neurotransmitter dopamine could serve
-as a ``teaching signal'' to better estimate value associations
-and action policies by controlling
-synaptic plasticity in the reward-processing circuitry, including the NAc.
-In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of the world.
-These can be assumptions about the physical laws governing the agent's environment or constraints on the state space and transitions between states.
-% For instance, if a rat in a maze knows
-% that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
-%
-
-In our adopted model-free RL framework,
-an agent might represent such knowledge about the world as follows:
-\begin{itemize}
-\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a location
-offering relevant resources.
-\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
-\item etc.
-\end{itemize}
-
 \begin{definition*}
@@
   transitions be Markovian. Consequently, the future states are independent of past states and only depend on the present state and action taken.
-
 \end{itemize}
-
 \end{definition*}
+
 Rational probabilistic planning can be reformulated
@@
 % \end{itemize}
+
+\paragraph*{Towards Model-free reinforcement learning for the DMN.}
+Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
+it has been proposed \citep{gershman2015computational} that a core property of human intelligence  underlie improvement
+of expected utility outcomes as a strategy for action choice in uncertain
+environments, a situation perfectly captured by the formalism of MDPs.
+It has also long been proposed \citep{dayan2008decision} that there
+is a rather direct mapping of model-free RL learning algorithms
+onto aspects of the brain.
+The neurotransmitter dopamine could serve
+as a ``teaching signal'' to better estimate value associations
+and action policies by controlling
+synaptic plasticity in the reward-processing circuitry, including the NAc.
+In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of the world.
+These can be assumptions about the physical laws governing the agent's environment or constraints on the state space, transition probabilities between states,
+reward contingencies, etc.
+% For instance, if a rat in a maze knows
+% that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
+An agent might represent such knowledge about the world as follows:
+\begin{itemize}
+\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a location
+offering relevant resources.
+\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
+\item etc.
+\end{itemize}
+% Model-based assumes knowledge of the environment in terms of the transition probabilities between states, the reward contingencies for state-action pairs, or any form of representation of the environment.
+Such knowledge can of course be partly learned from the environment: the agent learns a model for the world alongside learning to make optimal decisions using this model.
+These methods learn what the effect is going to be of taking an particular action in a particular state.
+The result is an estimate of the underlying MDP which can then be either solved exactly or approximately, depending on the setting and what is feasible.
+The main problem then is to "connect the dots", by using dynamic programming --for example.
+
+In contrast, model-free methods require no knowledge of the environment (transition probabilities, identities of sensory input, etc.) or representation thereof, and the agent learns which state-action pairs lead to reward through sampling the world in trial-and-error manner, and predicting long-run reward aggregates, using prediction error as an incentive.
+In so doing, such methods ultimately learn both a policy and  implicit representations of the world / environment.
+This is what algorithms TD learning~\citep{sutton1998reinforcement}.
+In terms of number of states, model-based RL scales better than model-free RL; in terms of the time horizon, model-free RL scales better~\citep{strehl2006}.
+Drawing from a vast literature, we will show that the DMN contains sufficient neurobiological hardware to implement model-free RL, at least in principle.
 
@@
 \label{sec:policymat}
-As already mentioned in the previous section, Q-learning optimizes over the class of
+As already mentioned in the previous section, Q-learning~\citep{watkins92} optimizes over the class of
 deterministic policies of the form \eqref{eq:qlearning}. State spaces may be extremely large,
@@
 % \textbf{We still may need to explain y in the formula with one sentence.}
-A recently proposed, practically successful alternative approach
-\citep{mnih2015,silver2016mastering} is to learn this
+A recently proposed, practically successful alternative approach is to learn this
 representation using an artificial deep neural-network model, leading to the
-so-called \textit{deep Q-learning} family of methods which
-This is the current state-of-the-art in RL research.
+so-called \textit{deep Q-learning}~\citep{mnih2015,silver2016mastering} family of methods which
+is the current state-of-the-art in RL research.
 The set of model parameters $\theta$ that instantiate the non-linear interactions
-between layers of the artifical neural network
+between layers of the artificial neural network
 may find a neurobiological correspondence in the adaptive strengths of axonal
@@
 \subsubsection{Does the hippocampus subserve MCMC sampling?}
-In RL, MCMC simulation is a common means to update the agent's belief state
-\citep{silver2010monte}.
+In RL, MCMC simulation is a common means to update the agent's belief state.
+% \citep{silver2010monte}.
 MCMC simulation provides a simple method for evaluating the value of a state.

### 2017-12-02 paper/paper.tex

@@
 % \end{itemize}
+
+\begin{mdframed}
+  \vspace{1em}
+  \textbf{Box 2: Hypotheses for testing the MDP account of DMN function}
+\begin{enumerate}
+\item Experiment (Humans): We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
+During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigms.
+We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, and the more the fMRI signals in the reward circuits should be independent of the reward context
+in the sensory environment.
+
+\item Experiment (Humans): We hypothesize a functional dissociation between computations for behavioral policy versus adapting stimulus-value associations as these are conceivably implemented in different subsystems of the DMN.
+We first expect that fMRI signals in the right temporo-parietal junction can predict behavioral changes caused by adaptation in the action choice tendencies (policy matrix) related to non-value-related prediction error.
+Second, fMRI signals in the ventromedial prefrontal cortex should predict behavioral changes caused by adaptation in value estimation (value matrix) due to reward-related stimulus-value association.
+We finally predict that fMRI signals in the posteromedial cortex, as a potential global information integrator, is able to predict shifts in overt behavior based on previous adaptations in policy or value estimation.
+
+\item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is necessary for choice behavior in mice.
+Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before make analogous choices in the future.
+Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal disruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
+
+\item Experiment (Humans): We hypotherize that time horizon should be modulated by different factors such as age, stress, and impulsivity.
+Using a temporal discounting task, we can measure how the time horizon could be affected at the behavioral level, and then trace back its potential representation at the neural level.
+Experimental design can be between groups (by age or with a impulsive population like chronic gamblers or drug addicts), or within group using a stress.
+
+% To exam-ine this question, Fellows and Farah (2005) investigated the effects of dorsolateral frontal (DLF) and ventromedial frontal (VMF) lobe damage on two aspects of temporal foresight—namely, temporal discounting and the futuretime perspective (measuredby an adaptation of Wallace’s [1956] list of future life events).
+
+\item Experiment (Animals): how synaptic epigenesis between DMN nodes can form a neural correlate of the policy matrix update? % Simulations:
+
+\end{enumerate}
+\end{mdframed}
 

### 2017-12-04 paper/paper.tex
rainbow stuff (response to reviewer)

@@
 \citep{mesulam1998sensation, taylor2015global}.
+
+\paragraph*{Self-training may introduce bias. How to alleviate it ?}
+As pointed out by one of the reviewers, self-training may introduce bias due to information shortage caused by the absence of external stimulus (task).
+ way to solve this issue is to use importance sampling as was done in ~\citep{schaul2015} and ~\citep{hessel2017}, to replay more often state-transitions from which there is more to learn.
+New transitions are inserted into the replay buffer with maximum priority, providing a bias towards recent transitions.
+Such an insertion bias would help counter-balance the bias introduced by the information shortage caused by the absence of external stimulus (task).
+Also, ~\citep{hessel2017} noticed that such prioritized replay reduces the data-complexity and the agent can start learning much sooner than conventionally.
 

### 2017-12-04 paper/paper.tex
typos pointed-out by rev #2

@@
 with an \textit{environment}, the goal being to maximize some notion of \textit{cumulative reward}.
-The optimal behavior typical
+The optimal behavior typically
 takes the future into account, as some rewards could be \textit{delayed}.
@@
 % \textbf{We still may need to explain y in the formula with one sentence.}
-A recently proposed, practically successful alternative approach is to learn this
+A recently proposed, practically successful alternative approach is, to learn this
 representation using an artificial deep neural-network model, leading to the
@@
 \item Experiment (Humans): We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
-During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigms.
+During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigm.
 We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, and the more the fMRI signals in the reward circuits should be independent of the reward context
@@
 In particular,
-environmental stimuli especially important for humans are frequently of
+environmental stimuli are especially important for humans are frequently of
 social nature. This is not surprising
@@
 for continued refinement of the organism's fate.
-Indeed, a major hurdle in explaining DMN activity from brain-imaging studies has been its very similar engaged across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., dreams during sleep or daytime mind-wandering), and thinking about future scenarios (e.g., prospection or delay discounting).
+Indeed, a major hurdle in explaining DMN activity from brain-imaging studies has been its very similar engagement across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., dreams during sleep or daytime mind-wandering), and thinking about future scenarios (e.g., prospection or delay discounting).
 The MDP account of DMN function offers a natural integration of these a-priori diverging mental processes into a common framework.

### 2017-12-05 paper/paper.tex
more fixes
Add a pararaph on the levels of observation + correction of typos + adjustement of figures position

@@
 Conversely,
-it has been empirically shown that, the more the world is easy to predict,
+it has been empirically shown that, the more the world is easy to foresee,
 the more human mental activity
@@
 we derive explicit hypotheses to be tested in future neuroscience
-experiments (Box 1).
+experiments (Box 2).
+At this stage, it is important to clarify how different levels of observation
+are at play in the current account. A process model is not solely intended to 
+capture behavior of the agent but also the neurocomputational specifics of the agent.
+In the following, we will use the word ``inference`` when describing the formal aspects
+of the statistical model, the word ``prediction`` when discussing the neural implementation,
+and words like ``forecast`` or ``forsee`` when refering to the behavior of the agent.
 
@@
 may be specifically designed to allow reconstructing
-entire episodes of experience from memory fragments.
+entire sequences of experience from memory fragments.
 Indeed,
@@
 
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=1.\linewidth]{rl2.pdf}
+  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
+    the agent takes an action by following the policy matrix as updated by the Bellman equation.
+The agent receives a consequential reward and observes the next state.
+The process goes on until interrupted or a goal state is reached.}
+  \label{fig:rl}
+\end{figure}
+
 \subsection{Markov decision processes}
@@
 step is sufficiently short.
-\begin{figure}[!h]
-  \centering
-  \includegraphics[width=1.\linewidth]{rl2.pdf}
-  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
-    the agent takes an action by following the policy matrix as updated by the Bellman equation.
-The agent receives a consequential reward and observes the next state.
-The process goes on until interrupted or a goal state is reached.}
-  \label{fig:rl}
-\end{figure}
+
 % , defining $s(t)$, the state at time $t$, as the position in state space.
@@
 The brain is therefore unlikely to implement the wake-sleep algorithm or a variant thereof.
+
+The recent theory of
+\textit{variational auto-encoders} (VAEs) \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm. VAEs
+overcome a number of the technical limits of the wake-sleep algorithm
+by using a reparametrization maneuver, which makes it possible to do differential calculus on random sampling procedures without exploding variance. As a result, unlike the wake-sleep algorithm
+for minimizing free energy, VAEs can be efficiently trained via back-propagation of prediction errors.
+
 \begin{wrapfigure}{r}{.37\linewidth}
@@
 \end{wrapfigure}
-
-The recent theory of
-\textit{variational auto-encoders} (VAEs) \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm. VAEs
-overcome a number of the technical limits of the wake-sleep algorithm
-by using a reparametrization maneuver, which makes it possible to do differential calculus on random sampling procedures without exploding variance. As a result, unlike the wake-sleep algorithm
-for minimizing free energy, VAEs can be efficiently trained via back-propagation of prediction errors.
 

### 2017-12-05 paper/paper.tex
Remove a redundant mention of memory in Markov process

@@
 
-% , defining $s(t)$, the state at time $t$, as the position in state space.
-The process has $memory$ if the next state depends not only on the current state
-but also on a finite number of past states.
-Rational probabilistic planning can be reformulated
-as a standard memoryless Markov process by simply expanding the
-definition of the state $s$ to include experience episodes of the past.
-
 In artificial intelligence and machine learning, a popular computational model for
@@
 
-Rational probabilistic planning can be reformulated
-as a standard memory Markov process by simply expanding the
+% , defining $s(t)$, the state at time $t$, as the position in state space.
+The process has $memory$ if the next state depends not only on the current state
+but also on a finite number of past states.
+Rational probabilistic planning can thus be reformulated
+as a standard memoryless Markov process by simply expanding the
 definition of the state $s$ to include experience episodes of the past.

### 2017-12-05 paper/paper.tex
Clarification of memory and POMDPs

@@
 This adds the capacity for memory, since then, the next state depends not only on the current state but also on a finite number of past states.
-This is precisely what partially observable MDPs do ~\citep{starkweather2017dopamine,o2006making}.
+This is precisely what Partially Observable MDPs (POMDPs) do ~\citep{starkweather2017dopamine,o2006making}.
+Nevertheless, this mathematical procedure in POMDPs only accounts for implicit $memory$.
+Since the current paper is aiming at both plausibility at the behavioral and neural level, we will address later in the paper how our model captures the neurophysiological constraints of the DMN and can also account for the explicit part of $memory$ from a human agent perspective.
 

### 2017-12-05 paper/paper.tex
Add second part of the memory aspects of MDP in DMN with addition of Pezzulo 2011 ref.

@@
 
+\paragraph*{On implicit and explicit memory in the current model.}
+Markov Processes are memoryless but this is possible mathematically to artificially insert the previous states of such model into the current state.
+While this may partially account for implicit memory at the behavioral level, it does not explain the underlying implementation at the neurophysiological level nor accounts for the explicit part of memory.
+First, it is possible to incorporate explicit memory capacities into the MDP.
+
+ We address this question in the paper by clarifying what memory retrieval would mean in the current account.
+In fact, implicit memory-based processing is already present in our MDP account of DMN function.
+It occurs in 3 three different forms: a) the action policy, b) the value function $\textemdash$as both of these are a result/product of the past$\textemdash$ and c) the deep architecture with many non-linear links in the hierarchical connections as realized by biological neural networks.
+Indeed, we reason about the brain has deep neural networks that present an implicit form of information compression obtained from life experience.
+This memory traces are implicitly stored in the neural machinery and can be retrieve as a form of Knowledge by simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}.
+The neural processes in the hippocampus can be seen as MCMC sampling from a formal statistical level but those memory recall are also the basis for scene construction processes across time-scales such as in the semantic hypothesis \citep{schacter2007remembering}.
+In a complementary way, neural encoding of reward information is not limited to value but incorporates specific information about the identity of the reward outcome, the predictive stimulus, or the action necessary to obtain the reward \citep{kahnt2017decade}.
+
 \subsection{Putting everything together}
@@
 From the perspective of a model-free RL agent,
-\textit{inference} in the human brain reduces to
+\textit{prediction} in the human brain reduces to
 generalization of

### 2017-12-05 paper/paper.tex
Correct ref to Fig 1 & 2

@@
 risk estimation of relevant environmental stimuli
-(Fig. \ref{fig:dmn}).
+(see Fig. \ref{fig:neurobiological_overview_DMN} and \ref{fig:dmn}).
 The ventromedial prefrontal DMN may

### 2017-12-05 paper/paper.tex
Polish the 2nd part of the discussion on memory

@@
 While this may partially account for implicit memory at the behavioral level, it does not explain the underlying implementation at the neurophysiological level nor accounts for the explicit part of memory.
-First, it is possible to incorporate explicit memory capacities into the MDP.
-
- We address this question in the paper by clarifying what memory retrieval would mean in the current account.
-In fact, implicit memory-based processing is already present in our MDP account of DMN function.
-It occurs in 3 three different forms: a) the action policy, b) the value function $\textemdash$as both of these are a result/product of the past$\textemdash$ and c) the deep architecture with many non-linear links in the hierarchical connections as realized by biological neural networks.
+The implicit memory-based processing is already present in our MDP account of DMN function; it occurs in 3 three different forms: a) the action policy, b) the value function $\textemdash$as both of these are a result/product of the past$\textemdash$ and c) the deep architecture with many non-linear links in the hierarchical connections as realized by biological neural networks.
 Indeed, we reason about the brain has deep neural networks that present an implicit form of information compression obtained from life experience.
 This memory traces are implicitly stored in the neural machinery and can be retrieve as a form of Knowledge by simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}.
-The neural processes in the hippocampus can be seen as MCMC sampling from a formal statistical level but those memory recall are also the basis for scene construction processes across time-scales such as in the semantic hypothesis \citep{schacter2007remembering}.
+The neural processes in the hippocampus can be seen as MCMC sampling from a formal statistical level but those memory recall are also the basis for scene construction processes across time-scales such as in the semantic hypothesis \citep{schacter2007remembering,Axelrod2017}.
 In a complementary way, neural encoding of reward information is not limited to value but incorporates specific information about the identity of the reward outcome, the predictive stimulus, or the action necessary to obtain the reward \citep{kahnt2017decade}.

### 2017-12-06 paper/paper.tex
fix affiliations
Danilo's changes for R1

@@
 \small{
-    \textbf{1} Universit\'e Paris-Saclay, INRIA, Parietal Team, Saclay, France\\
-    \textbf{2} Universit\'e Paris-Saclay, CEA, Neurospin, Gif-sur-Yvette, France\\
+    \textbf{1} INRIA, Parietal Team, Saclay, France\\
+    \textbf{2} Neurospin, CEA, Gif-sur-Yvette, France\\
     \textbf{3} Department of Psychiatry, Psychotherapy and Psychosomatics, RWTH Aachen University, Aachen, Germany\\
@@
 %
-DMN function is recast in terms of
-control theory and reinforcement learning by basing it on Markov decision processes.
+% DMN function is recast in terms of
+% control theory and reinforcement learning by basing it on Markov decision processes.
+%
+Specifically, we answer the question whether the neurobiological properties of the default mode network collectively provide the computational building blocks necessary for a Markov Decision Process.
 %
@@
 through integration of feedback outcomes.
+That is, ``[agents], with their actions, modify the environment and in doing so partially determine their next stimuli, in particular stimuli that are necessary for triggering the next action''
+\citep{pezzulo2011grounding}.
+Agents with other behavioral policies therefore sample different
+distributions of action-perception trajectories \citep{ghavamzadeh2015bayesian}.
 Henceforth,
@@
 can be recast as prediction error minimization
-based on probabilistic mental simulations,
+based on probabilistic mental simulations
+- ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
 thus maximizing action outcome across multiple time scales.
@@
 \paragraph*{Towards Model-free reinforcement learning for the DMN.}
-Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure}. Indeed,
+Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure, daw2014algorithmic}.
+Indeed,
 it has been proposed \citep{gershman2015computational} that a core property of human intelligence  underlie improvement
@@
 entail secondary reallocation of attention and working-memory resources.
-Further neuroimaging experiments \citep{braver2002role} corroborated
+Further neuroimaging experiments corroborated
 the prefrontal DMN to subserve
-``processes related to the management and monitoring of sub-goals while maintaining information in working memory''.
+``processes related to the management and monitoring of sub-goals while
+maintaining information in working memory''
+\citep{braver2002role}
+and to functionally couple with the hippocampus conditioned by ``deep versus shallow planning''
+\citep{kaplan2017neural}.
 %
@@
 \subsubsection{Does the hippocampus subserve MCMC sampling?}
-In RL, MCMC simulation is a common means to update the agent's belief state.
+In RL, MCMC simulation is a common means to update the agent's belief state
+by stochastic sampling around states and possible transitions \citep{daw2014algorithmic}.
 % \citep{silver2010monte}.
@@
 would be used to update the value function, without ever looking inside the black box describing the model's dynamics \citep{lavilleon2015}.
+A neuroimaging experiment in humans for instance identified hippocampal signals that specifically preceded upcoming choice performance in prospective planning in new environments
+\citep{kaplan2017neural}.
 This would be a simple control algorithm by evaluating all legal actions and selecting the action with
@@
 In MDPs, MCMC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible \citep{silver2016mastering}.
+This is because expected consequences of action choices can be well evaluated although only a subset of the states are actually considered \citep{daw2014algorithmic}.
 % Much recent work points to MCMC or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines.
@@
 The DMN is today known to consistently increase in neural
-  activity when humans engage in cognitive processes that are detached from
-  the current sensory environment. Additionally,
+activity when humans engage in cognitive processes that are relatively detached from
+the current sensory environment.
+
+The more familiar and predictable the current environment, the more resources remain for allocating DMN activity to MDP processes extending beyond the present time and sensory context.
+In line with this perspective, DMN engagement was shown to heighten and relate to effective behavioral responses in the practiced phase of a demanding cognitive flexibility task, as compared to acquisition phase when participants learned context-specific rules.
+This involvement in automated decision-making lead the authors to propose an ``autopilot'' role for the DMN \citep{vatansever2017pnas}, which may contribute to optimizing control of the organism in general.
+Among all DMN nodes, the RTPJ has probably a most evident role as network-switching relay calibrating between processing of environment-engaged versus internally generated information \citep{downar2000multimodal, golland2006extrinsic, bzdok2013tpj}.
+
+Additionally,
 this network was proposed to be situated
@@
 such a randomly synthesized situation.
+Generated representations arguably are ``internally manipulable,
+and can be used for attempting actions internally,
+before or instead of acting in the external reality,
+and in diverse goal and sensory contexts, i.e. even outside the context in which they were learned''
+\citep{pezzulo2011grounding}.
 In the absence of environmental input and feedback
@@
 \citep{schacter2007remembering}.
+``[P]redictions and simulations are not only automatically elicited by external events but can be endogenously generated when needed.
+[...] The mechanism of access via simulation could be a widespread method for accessing and producing knowledge, and represents a valid alternative to the traditional idea of storage and retrieval''
+\citep{pezzulo2011grounding}.
 Such scene construction processes could contribute to interpreting the

### 2017-12-06 paper/paper.tex
fixing broken wrapfigure (guys check yo stuff)

@@
 
+On another front, since theories based on the FEP \citep{friston2010free,fristonAIorRL}
+conceptualize ongoing behavior
+in an organism to be geared towards the surprise-minimizing goal,
+an organism entering a dark room
+% \footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.}
+(Fig. \ref{fig:darkroom}) would strive to remain in this location
+because its sensory inputs are perfectly predictable
+given the environmental state \citep{darkroom2012}.
 \begin{wrapfigure}{r}{.37\linewidth}
@@
 \end{wrapfigure}
-
-On another front, since theories based on the FEP \citep{friston2010free,fristonAIorRL}
-conceptualize ongoing behavior
-in an organism to be geared towards the surprise-minimizing goal,
-an organism entering a dark room
-% \footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.}
-(Fig. \ref{fig:darkroom}) would strive to remain in this location
-because its sensory inputs are perfectly predictable
-given the environmental state \citep{darkroom2012}.
-However, such a behavioral tendency is
-seldom observed in animals or humans in the real world: In a dark room,
-intelligent agents would search for light sources to \textit{explore}
+However, such a behavior is seldom observed in animals or humans in the
+real world: In a dark room, intelligent agents would search for light sources to \textit{explore}
 the surroundings or leave the exit the room.

### 2017-12-06 paper/paper.tex
Corrections of the memory/levels edits for reviewer 2

@@
 may be solved by a stochastic approximation
-based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
-\citep{tenenbaum2011grow}.
+based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling.
 Even necessarily imperfect memory recall,
@@
 experiments (Box 2).
-At this stage, it is important to clarify how different levels of observation
-are at play in the current account. A process model is not solely intended to 
-capture behavior of the agent but also the neurocomputational specifics of the agent.
-In the following, we will use the word ``inference`` when describing the formal aspects
-of the statistical model, the word ``prediction`` when discussing the neural implementation,
-and words like ``forecast`` or ``forsee`` when refering to the behavior of the agent.
+At this stage, we emphasize the importance of differentiating which levels of observation are at play in the present account.
+A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent.
+Henceforth, we will use ``inference`` when describing aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
 
@@
 may be solved by a stochastic approximation
-based on a brain implementation of MCMC sampling
-\citep{tenenbaum2011grow}.
+based on a brain implementation of MCMC sampling.
 
@@
 % , defining $s(t)$, the state at time $t$, as the position in state space.
-The process has $memory$ if the next state depends not only on the current state
-but also on a finite number of past states.
-Rational probabilistic planning can thus be reformulated
-as a standard memoryless Markov process by simply expanding the
-definition of the state $s$ to include experience episodes of the past.
+The process has $memory$ if the next state depends not only on the current state but also on a finite number of past states.
+Rational probabilistic planning can thus be reformulated as a standard memoryless Markov process by simply expanding the definition of the state $s$ to include experience episodes of the past.
 This adds the capacity for memory, since then, the next state depends not only on the current state but also on a finite number of past states.
-This is precisely what Partially Observable MDPs (POMDPs) do ~\citep{starkweather2017dopamine,o2006making}.
-Nevertheless, this mathematical procedure in POMDPs only accounts for implicit $memory$.
-Since the current paper is aiming at both plausibility at the behavioral and neural level, we will address later in the paper how our model captures the neurophysiological constraints of the DMN and can also account for the explicit part of $memory$ from a human agent perspective.
+This is precisely what Partially Observable MDPs (POMDPs) aim to realize ~\citep{starkweather2017dopamine,o2006making}.
+Nevertheless, this mathematical property in POMDPs mostly accounts for implicit $memory$.
+Since the current paper is focused on both plausibility at the behavioral and neurobiological level, we will address below how our model captures the neurophysiological constraints of the DMN and can also account for the explicit part of $memory$ characteristic for human agents.
 
@@
 \paragraph*{On implicit and explicit memory in the current model.}
-Markov Processes are memoryless but this is possible mathematically to artificially insert the previous states of such model into the current state.
-While this may partially account for implicit memory at the behavioral level, it does not explain the underlying implementation at the neurophysiological level nor accounts for the explicit part of memory.
-The implicit memory-based processing is already present in our MDP account of DMN function; it occurs in 3 three different forms: a) the action policy, b) the value function $\textemdash$as both of these are a result/product of the past$\textemdash$ and c) the deep architecture with many non-linear links in the hierarchical connections as realized by biological neural networks.
+Markov Processes are memoryless but this it is mathematically feasible to insert the previous states of such model into the current state.
+While this may partially account for implicit memory at the behavioral level, it may not explain the underlying implementation at the neurophysiological level nor accounts for the explicit part of memory.
+Implicit memory-based processing arises in our MDP account of DMN function in three different forms: successive updates of a) the action policy and b) the value function, both being products of the past and c) the deep non-linear links in the hierarchical connections realized by biological neural networks.
 Indeed, we reason about the brain has deep neural networks that present an implicit form of information compression obtained from life experience.
-This memory traces are implicitly stored in the neural machinery and can be retrieve as a form of Knowledge by simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}.
+Such memory traces are implicitly stored in the neural machinery and can be retrieve as a form of knowledge during simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}.
 The neural processes in the hippocampus can be seen as MCMC sampling from a formal statistical level but those memory recall are also the basis for scene construction processes across time-scales such as in the semantic hypothesis \citep{schacter2007remembering,Axelrod2017}.
-In a complementary way, neural encoding of reward information is not limited to value but incorporates specific information about the identity of the reward outcome, the predictive stimulus, or the action necessary to obtain the reward \citep{kahnt2017decade}.
 

### 2017-12-06 paper/paper.tex
Corrections of Box 2 experiments

@@
 
-\item Experiment (Humans): We hypotherize that time horizon should be modulated by different factors such as age, stress, and impulsivity.
-Using a temporal discounting task, we can measure how the time horizon could be affected at the behavioral level, and then trace back its potential representation at the neural level.
-Experimental design can be between groups (by age or with a impulsive population like chronic gamblers or drug addicts), or within group using a stress.
+\item Experiment (Humans): We hypothesize that the relevant time horizon is modulated by various factors such as organism age, acute stress, and time-enduring impulsivity traits.
+Using a temporal discounting experiment, it can be measured how the time horizon could be affected at the behavioral level, and then trace back its potential representation at the neural level.
+The experiment can be designed to examine between-group (by age or with a impulsive population like chronic gamblers or drug addicts) and within group effects using a stress; and brought in context with the participant’s age and impulsivity tendencies.
 
@@
 
-\item Experiment (Animals): how synaptic epigenesis between DMN nodes can form a neural correlate of the policy matrix update? % Simulations:
+\item Experiment (Humans \& Animals): An additional layer of learning concerns the addition of new entries in the table (i.e., state and action).
+The framing of the policy matrix is not described here but could be supported by synaptic epigenesis \citep{gisiger_acquisition_2005}.
+Indeed, the tuning of synaptic weights through learning can stabilize additional patterns of activity by creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}.
+Those attractors will then constrain both the number of factors taken into account by decision processes and the number of behaviors at reach by the agent \citep{wang_decision_2008}.
+To probe this aspects, we propose two experiments probing how synaptic epigenesis forms a neural correlate of the policy matrix update: one in human, investigating the changes of functional connectivity between DMN nodes following a temporal discounting experiment; the second in rodents, directly testing with anterograde tracing how homolog regions of the DMN present increased synaptic changes compare to other brain regions.
 

### 2017-12-06 paper/paper.tex
Typos

@@
 \item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is necessary for choice behavior in mice.
-Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before make analogous choices in the future.
+Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future.
 Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal disruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.

### 2017-12-06 paper/paper.tex
Add MDP def at beginning

@@
 DMN dynamics could be constantly shaped by prediction error through feedback loops.
+This computational account of DMN will be described in the mathematical framework of Markov Decision Processes (MDPs).
+MDPs indeed address specifically the case of decision making in stochastic context with rewards.
+
 Such an RL account of DMN function can naturally embed human behavior
@@
   cumulative reward outcomes. Implicit computation of future choices
-  provides an explanation for the
+  provides a potential explanation for the
   evolutionary emergence and practical usefulness of

### 2017-12-06 paper/paper.tex
Add MDP def at beginning
Correct unimodal ctx => primary sensory ctx

@@
 the DMN at the top and the salience and dorsal attention systems
-at intermediate levels, above thalamic and unimodal sensory
+at intermediate levels, above thalamic and primary sensory
 cortex'' \citep{carhart2010default}.
@@
 This computational account of DMN will be described in the mathematical framework of Markov Decision Processes (MDPs).
-MDPs indeed address specifically the case of decision making in stochastic context with rewards.
+MDPs indeed address specifically the case of decision making in stochastic context with reward feedback.
 
@@
 the subordinate salience and dorsal attention network in the middle and
-the unimodal sensory cortices at the bottom
+the primary sensory cortices at the bottom
 \citep{carhart2010default, margulies2016situating}.

### 2017-12-06 paper/paper.tex
exponential delay discounting addition in footnote

@@
 This random variable\footnote{Random as it depends both on the environment's dynamic and the
-  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
+  policy $\pi$ being played (which can be stochastic).
+The exponential delay discounting function used here refere to the usual formulation in the field of reinforcement learning but psychological experiments may also reveal other discounting regimes \citep{green2004discounting}}  measures the cumulative reward of
 following an action policy $\pi$.

### 2017-12-06 paper/paper.tex
reread part 1

--- 2017-12-06 paper/paper.tex
@@
 %
-Many research streams support an evolutionarily adaptive role in
+Many research streams speak in favor of an evolutionarily adaptive role in
 envisioning experience to anticipate the future.
@@
 of the biological purpose underlying DMN activity.
-Despite observation of similar large-scale networks of co-varying spontaneous activity in MEG investigations \citep{de2010temporal,brookes2011investigating,baker2014fast}, the link between the fMRI BOLD signal and population-level neural activity is still unclear.
-If those frequency-specific electrophysiological correlations are proposed as complementary to those observed with BOLD \citep{hipp2015bold}, their role in DMN function remains challenging \citep{maldjian2014graph}.
+Despite observation of similar large-scale networks of co-varying spontaneous activity
+in electrophysiological investigations \citep{de2010temporal,brookes2011investigating,baker2014fast}, the link between the fMRI BOLD signal and population-level neural activity is still unclear.
+If those frequency-specific electrophysiological correlations are proposed as complementary to those observed with BOLD \citep{hipp2015bold}, their role in DMN function remains elusive \citep{maldjian2014graph}.
 
@@
 are responsible for extended parts of the baseline neural activity,
-which typically decreases when engaged in psychological experiments
+which typically decreases when engaged in controlled psychological experiments
 \citep{raichle_baseline}.
@@
 DMN dynamics could be constantly shaped by prediction error through feedback loops.
-This computational account of DMN will be described in the mathematical framework of Markov Decision Processes (MDPs).
-MDPs indeed address specifically the case of decision making in stochastic context with rewards.
-
 Such an RL account of DMN function can naturally embed human behavior
@@
 - ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
-thus maximizing action outcome across multiple time scales.
+thus maximizing action outcomes across multiple time scales.
 Such a purposeful optimization objective
@@
 based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling.
-Even necessarily imperfect memory recall,
+Even (unavoidably) imperfect memory recall,
 random day-time mind-wandering, and
 seemingly arbitrary dreams during sleep
-may provide blocks of pseudo-experience to iteratively
+may provide randomly sampled blocks of pseudo-experience instrumental to iteratively
 optimize the behavior of the organism.
@@
 Evidence from computational modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
-the human brain is largely dedicated to
+the brain is largely dedicated to
 ``the development and maintenance of [a]
@@
 The present paper proposes a
-process model that satisfies this contention.
+process model that satisfies this previously proposed contention.
 We also contribute to the discussion of DMN function
@@
 morphological variability in DMN regions
-is linked to the reward circuitry (Fig. 2).
+is linked to the reward circuitry (Fig. 2),
+thus linking two literatures with currently scarce cross-references.
 Finally, we detail how our process model relates
@@
 experiments (Box 2).
-At this stage, we emphasize the importance of differentiating which levels of observation are at play in the present account.
-A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent.
-Henceforth, we will use ``inference`` when describing aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
+At this stage, it is important to clarify how different levels of observation
+are at play in the current account. A process model is not solely intended to 
+capture behavior of the agent but also the neurocomputational specifics of the agent.
+In the following, we will use the word ``inference'' when describing the formal aspects
+of the statistical model, the word ``prediction'' when discussing the neural implementation,
+and words like ``forecast'' or ``forsee'' when refering to the behavior of the agent.
 
@@
 based on experimental findings in the neuroscience literature.
-This walkthrough across each region of the DMN will
+This walkthrough across main regions of the DMN will
 outline the individual functional profiles,
@@
 the medial prefrontal cortex (mPFC),
-are probably responsible for the highest turn-over
+are probably responsible for highest turn-overs
 of energy consumption \citep{raichle2001pnas, raichle_baseline}.
 These metabolic characteristics go hand-in-hand with
-neuroimaging analyses that suggested the PMC and mPFC to
-potentially represent the functional backbone of the DMN
+neuroimaging findings that suggested the PMC and mPFC to
+potentially represent the functional core of the DMN
 \citep{andrews2010, hagmann2008mapping}.
@@
 \citep{balint1909seelenlahmung}.
-Neurological patients are plagued by an
-inability to combine various individual features of the visual
+Such neurological patients are plagued by an
+inability to bind various individual features of the visual
 environment into an integrated whole (i.e., simultanagnosia)
@@
 (i.e., optic ataxia).
-This can be viewed as a high-level impairment in gathering
+This dysfunction can be viewed as a high-level impairment in gathering
 information about alternative objects (i.e., exploration) as well as
@@
 Congruently,
-the human PMC was coupled in two functional connectivity analyses
+the human PMC was coupled in two different functional connectivity analyses
 \citep{bzdok2015subspecialization}
@@
 for biological relevance assessment
-\citep{amygdala_bzdok}.
+\citep{amygdala_bzdok, ghods2009fundamental, baxter2002amygdala}.
 
@@
 preference for or aversion to options with uncertain reward outcomes
-and its spiking activity was more associated with
+and its neural spiking activity was more associated with
 subjectively perceived relevance of a chosen object
@@
 based on an ``internal currency of value'' \citep{mccoy2005risk}.
-In fact, direct stimulation of PMC neurons
+In fact, direct stimulation of PMC neurons in monkeys
 promoted exploratory actions,
@@
 proposed to mediate between the organism's egocentric
-(i.e., focused on sensory environment) and
-allocentric (i.e., focused on world knowledge) viewpoints
+(i.e., focused on external sensory environment) and
+allocentric (i.e., focused on internal world knowledge) viewpoints
 in animals and humans
@@
 into short- and longer-term behavioral agendas.
-Estimated value, that differs across individuals, might enrich
+Estimated value, found to differs across individuals, might enrich
 statistical assessment of the environment
@@
 dmPFC function may be closer to a
-``mental sketchpad'' \citep{goldman1996prefrontal}, as it
+``mental sketchpad'' \citep{goldman1996prefrontal}, as this DMN part
 potentially subserves the de-novo construction and manipulation
@@
 \citep{bzdok2013segregation}.
-The dmPFC may subserve inference, representation, and assessment
+The dmPFC may subserve representation and assessment
 of one's own and other individuals' action considerations.
@@
 Certain dmPFC neurons in macaque monkeys exhibited a preference
-for processing others', rather than own, behavior
-with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
+for processing others', rather than own, action
+with fine-grained adjustment of contextual aspects \citep{yoshida2010neural}.
 % In fact, the spatially neighboring dorsal anterior cingulate cortex
@@
 Comparing to the dmPFC,
-the vmPFC probably subserves
+the vmPFC is probably more specifically devoted to
 subjective value evaluation and
 risk estimation of relevant environmental stimuli
-(see Fig. \ref{fig:neurobiological_overview_DMN} and \ref{fig:dmn}).
+(Fig. \ref{fig:neurobiological_overview_DMN} and \ref{fig:dmn}).
 The ventromedial prefrontal DMN may
@@
 \citep{lebreton2009automatic}.
-NAc activity is thought to reflect reward prediction signals
+NAc activity is further thought to reflect reward prediction signals
 from dopaminergic neurotransmitter pathways
@@
       \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
-  Code and data for reproduction and visualization:
-  \url{www.github.com/banilo/to_be_added_later}.
+  Code for reproduction and visualization:
+  \url{www.github.com/banilo/darkcontrol_PCB2018}.
   }
@@
 be more connected to the vmPFC than dmPFC in both species.
-Two different functional connectivity analyses in humans also strongly connected
-the vmPFC with the NAc, hippocampus (HC),
+Two different functional connectivity analyses in humans also revealed
+strong vmPFC connections
+with the NAc, hippocampus (HC),
 and PMC \citep{bzdok2015subspecialization}.
@@
 the vmPFC consistently reflects an individual’s subjective
-value estimates
-\citep{behrens2008associative}.
-This may be why performance within and across participants
-was related to state encoding in the vmPFC \citep{Schuck2016}.
+value predictions
+\citep{behrens2008associative},
+which may also explain why performance within and across participants
+was reported to relate to state encoding in the vmPFC \citep{Schuck2016}.
 Such a ``cognitive map'' of the action space was argued to encode
@@
 may be solved by a stochastic approximation
-based on a brain implementation of MCMC sampling.
+based on a brain implementation of MCMC sampling
+\citep{tenenbaum2011grow}.
 
@@
 % , defining $s(t)$, the state at time $t$, as the position in state space.
-The process has $memory$ if the next state depends not only on the current state but also on a finite number of past states.
-Rational probabilistic planning can thus be reformulated as a standard memoryless Markov process by simply expanding the definition of the state $s$ to include experience episodes of the past.
+The process has $memory$ if the next state depends not only on the current state
+but also on a finite number of past states.
+Rational probabilistic planning can thus be reformulated
+as a standard memoryless Markov process by simply expanding the
+definition of the state $s$ to include experience episodes of the past.
 This adds the capacity for memory, since then, the next state depends not only on the current state but also on a finite number of past states.
-This is precisely what Partially Observable MDPs (POMDPs) aim to realize ~\citep{starkweather2017dopamine,o2006making}.
-Nevertheless, this mathematical property in POMDPs mostly accounts for implicit $memory$.
-Since the current paper is focused on both plausibility at the behavioral and neurobiological level, we will address below how our model captures the neurophysiological constraints of the DMN and can also account for the explicit part of $memory$ characteristic for human agents.
+This is precisely what Partially Observable MDPs (POMDPs) do ~\citep{starkweather2017dopamine,o2006making}.
+Nevertheless, this mathematical procedure in POMDPs only accounts for implicit $memory$.
+Since the current paper is aiming at both plausibility at the behavioral and neural level, we will address later in the paper how our model captures the neurophysiological constraints of the DMN and can also account for the explicit part of $memory$ from a human agent perspective.
 
@@
 \mapsto \tilde{Q}(s, a|\theta_{k-1})$.
-From a neurobiological perspective,
-experience replay can be manifested as the re-occurrence of
-neuron spiking sequences that have also occurred during specific prior actions
-and environmental states.
-%, but the replay has a much faster time scale.
-The HC is a strong candidate to contribute to such a mechanism
-because neuroscience experiments have repeatedly indicated
-in rats, mice, cats, rabbits, songbirds, and
-monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}.
+From a neurobiological perspective, experience replay can be manifested as the re-occurrence of neuron spiking sequences that have also occurred during specific prior actions and environmental states.
+The HC is a strong candidate for contributing to such neural reinstantiation of behavioral episodes as neuroscience experiments have repeatedly indicated in rats, mice, cats, rabbits, songbirds, and monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}.
 
@@
 \paragraph*{On implicit and explicit memory in the current model.}
-Markov Processes are memoryless but this it is mathematically feasible to insert the previous states of such model into the current state.
-While this may partially account for implicit memory at the behavioral level, it may not explain the underlying implementation at the neurophysiological level nor accounts for the explicit part of memory.
-Implicit memory-based processing arises in our MDP account of DMN function in three different forms: successive updates of a) the action policy and b) the value function, both being products of the past and c) the deep non-linear links in the hierarchical connections realized by biological neural networks.
+Markov Processes are memoryless but this is possible mathematically to artificially insert the previous states of such model into the current state.
+While this may partially account for implicit memory at the behavioral level, it does not explain the underlying implementation at the neurophysiological level nor accounts for the explicit part of memory.
+The implicit memory-based processing is already present in our MDP account of DMN function; it occurs in 3 three different forms: a) the action policy, b) the value function $\textemdash$as both of these are a result/product of the past$\textemdash$ and c) the deep architecture with many non-linear links in the hierarchical connections as realized by biological neural networks.
 Indeed, we reason about the brain has deep neural networks that present an implicit form of information compression obtained from life experience.
-Such memory traces are implicitly stored in the neural machinery and can be retrieve as a form of knowledge during simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}.
+This memory traces are implicitly stored in the neural machinery and can be retrieve as a form of Knowledge by simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}.
 The neural processes in the hippocampus can be seen as MCMC sampling from a formal statistical level but those memory recall are also the basis for scene construction processes across time-scales such as in the semantic hypothesis \citep{schacter2007remembering,Axelrod2017}.
+In a complementary way, neural encoding of reward information is not limited to value but incorporates specific information about the identity of the reward outcome, the predictive stimulus, or the action necessary to obtain the reward \citep{kahnt2017decade}.
 
@@
 appears to tie in with the implicit computation of
-action and state cascades as a function of what happened in the past.
+action and state cascades as a function of
+experienced events and collected feedback from the past.
+
 A policy matrix encapsulates the repertoire of possible actions
@@
   cumulative reward outcomes. Implicit computation of future choices
-  provides a potential explanation for the
+  provides an explanation for the
   evolutionary emergence and practical usefulness of
@@
 \item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is necessary for choice behavior in mice.
-Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future.
+Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before make analogous choices in the future.
 Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal disruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
 
-\item Experiment (Humans): We hypothesize that the relevant time horizon is modulated by various factors such as organism age, acute stress, and time-enduring impulsivity traits.
-Using a temporal discounting experiment, it can be measured how the time horizon could be affected at the behavioral level, and then trace back its potential representation at the neural level.
-The experiment can be designed to examine between-group (by age or with a impulsive population like chronic gamblers or drug addicts) and within group effects using a stress; and brought in context with the participant’s age and impulsivity tendencies.
+\item Experiment (Humans): We hypotherize that time horizon should be modulated by different factors such as age, stress, and impulsivity.
+Using a temporal discounting task, we can measure how the time horizon could be affected at the behavioral level, and then trace back its potential representation at the neural level.
+Experimental design can be between groups (by age or with a impulsive population like chronic gamblers or drug addicts), or within group using a stress.
 
@@
 
-\item Experiment (Humans \& Animals): An additional layer of learning concerns the addition of new entries in the table (i.e., state and action).
-The framing of the policy matrix is not described here but could be supported by synaptic epigenesis \citep{gisiger_acquisition_2005}.
-Indeed, the tuning of synaptic weights through learning can stabilize additional patterns of activity by creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}.
-Those attractors will then constrain both the number of factors taken into account by decision processes and the number of behaviors at reach by the agent \citep{wang_decision_2008}.
-To probe this aspects, we propose two experiments probing how synaptic epigenesis forms a neural correlate of the policy matrix update: one in human, investigating the changes of functional connectivity between DMN nodes following a temporal discounting experiment; the second in rodents, directly testing with anterograde tracing how homolog regions of the DMN present increased synaptic changes compare to other brain regions.
+\item Experiment (Animals): how synaptic epigenesis between DMN nodes can form a neural correlate of the policy matrix update? % Simulations:
 
@@
 
-On another front, since theories based on the FEP \citep{friston2010free,fristonAIorRL}
-conceptualize ongoing behavior
-in an organism to be geared towards the surprise-minimizing goal,
-an organism entering a dark room
-% \footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.}
-(Fig. \ref{fig:darkroom}) would strive to remain in this location
-because its sensory inputs are perfectly predictable
-given the environmental state \citep{darkroom2012}.
 \begin{wrapfigure}{r}{.37\linewidth}
@@
 \end{wrapfigure}
-However, such a behavior is seldom observed in animals or humans in the
-real world: In a dark room, intelligent agents would search for light sources to \textit{explore}
+
+On another front, since theories based on the FEP \citep{friston2010free,fristonAIorRL}
+conceptualize ongoing behavior
+in an organism to be geared towards the surprise-minimizing goal,
+an organism entering a dark room
+% \footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.}
+(Fig. \ref{fig:darkroom}) would strive to remain in this location
+because its sensory inputs are perfectly predictable
+given the environmental state \citep{darkroom2012}.
+However, such a behavioral tendency is
+seldom observed in animals or humans in the real world: In a dark room,
+intelligent agents would search for light sources to \textit{explore}
 the surroundings or leave the exit the room.

### 2017-12-06 paper/paper.tex

@@
 DMN dynamics could be constantly shaped by prediction error through feedback loops.
+This computational account of DMN will be described in the mathematical framework of Markov Decision Processes (MDPs).
+MDPs indeed address specifically the case of decision making in stochastic context with reward feedback.
+
 Such an RL account of DMN function can naturally embed human behavior
@@
 Even (unavoidably) imperfect memory recall,
+Even necessarily imperfect memory recall,
 random day-time mind-wandering, and
@@
 experiments (Box 2).
-At this stage, it is important to clarify how different levels of observation
-are at play in the current account. A process model is not solely intended to 
-capture behavior of the agent but also the neurocomputational specifics of the agent.
-In the following, we will use the word ``inference'' when describing the formal aspects
-of the statistical model, the word ``prediction'' when discussing the neural implementation,
-and words like ``forecast'' or ``forsee'' when refering to the behavior of the agent.
+At this stage, we emphasize the importance of differentiating which levels of observation are at play in the present account.
+A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent.
+Henceforth, we will use ``inference`` when describing aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
 
@@
 \item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is necessary for choice behavior in mice.
-Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before make analogous choices in the future.
+Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future.
 Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal disruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
 
-\item Experiment (Humans): We hypotherize that time horizon should be modulated by different factors such as age, stress, and impulsivity.
-Using a temporal discounting task, we can measure how the time horizon could be affected at the behavioral level, and then trace back its potential representation at the neural level.
-Experimental design can be between groups (by age or with a impulsive population like chronic gamblers or drug addicts), or within group using a stress.
+\item Experiment (Humans): We hypothesize that the relevant time horizon is modulated by various factors such as organism age, acute stress, and time-enduring impulsivity traits.
+Using a temporal discounting experiment, it can be measured how the time horizon could be affected at the behavioral level, and then trace back its potential representation at the neural level.
+The experiment can be designed to examine between-group (by age or with a impulsive population like chronic gamblers or drug addicts) and within group effects using a stress; and brought in context with the participant’s age and impulsivity tendencies.
 

### 2017-12-07 paper/paper.tex
reread part 2

@@
       \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
-  Code for reproduction and visualization:
+  The code for reproduction and visualization:
   \url{www.github.com/banilo/darkcontrol_PCB2018}.
@@
 Indeed,
-hippocampal damage is
+hippocampal damage was
 not only associated with an impairment in re-experiencing the past (i.e., amnesia),
-but also imagination dedicated to one's own future and
+but also forecasting of one's own future and
 imagination of experiences more broadly \citep{hassabis2007patients}.
@@
 Congruently,
-London taxi drivers, humans with high performance in spatial navigation,
+London taxi drivers, humans with high performance in forecasting spatial navigation,
 were shown to exhibit increased gray-matter volume in the
@@
 featured re-occurrence directly after experimental trials
-as well as directly before upcoming experimental trials \citep{diba2007forward}.
+as well as directly before (prediction of) upcoming experimental trials \citep{diba2007forward}.
 Similar spiking patterns in hippocampal neurons during rest and sleep
 have been proposed to be critical in communicating local information
-to the neocortex for long-term storage, potentially also in the regions of the DMN.
+to the neocortex for long-term storage, potentially including DMN regions.
 Moreover, in mice, invasively triggering spatial experience recall
@@
 The TPJs are known to exhibit hemispheric differences
-based on microanatomical properties and gyrification patterns
+based on microanatomical properties and cortical gyrification patterns
 \citep{seghier2013angular}.
 Globally, neuroscientific investigations on hemispheric functional specialization
-have highlighted the right versus left cerebral hemisphere as dominant for
-attentional versus semantic functions
+have highlighted the right cerebral hemisphere as dominant for attentional functions
+and the left side for semantic functions
 \citep{seghier2013angular, bzdok2013tpj, bzdok2016left,
@@
 
-The TPJ in the right-hemispheric DMN (RTPJ)
+The TPJ in the right hemisphere (RTPJ)
 has been shown to be closely related to
 multi-sensory prediction and prediction error signaling.
-It is central for action initiation during goal-directed psychological tasks and for
+This DMN region is probably central for
+action initiation during goal-directed psychological tasks and for
 sensorimotor behavior by integrating multi-sensory attention
 \citep{corbetta2002control}.
-Involvement of this DMN region was repeatedly reported in
+Its involvement was repeatedly reported in
 multi-step action execution \citep{hartmann2005takes},
@@
 %
-It was argued that the RTPJ encodes actions and ensuing outcomes,
-without necessarily relating those to value estimation
+It was argued that the RTPJ encodes actions and predicted outcomes,
+without necessarily relating these neural processes to value estimation
 \citep{liljeholm2013neural, hamilton2008action,
@@
 In the face of large discrepancies between actual and previously predicted
-environmental events the RTPJ acts as a potential switch between
+environmental events, the RTPJ acts as a potential switch between
 externally-oriented mind sets focussed on the
@@
 
-The TPJ in the left-hemispheric DMN (LTPJ),
+The left TPJ of the DMN (LTPJ),
 in turn, has a close relationship to Wernicke's area
@@
 and neuropsychology,
-the LTPJ mediates access to and integration of world knowledge,
+the LTPJ appears to mediate access to and binding of world knowledge,
 such as required during action considerations
@@
 Consistent with this view,
-LTPJ damage in humans also entails problems in recognizing
+LTPJ damage in humans also entailed problems in recognizing
 others' pantomimed action towards objects
@@
 Further,
-episodic memory recall and mental imagery strongly draw on
+episodic memory recall and mental imagery to forecast future events strongly draw on
 re-assembling world knowledge.
@@
 guide present action choice,
-weigh hypothetical possibilities, and forecast future events.
-%
-The LTPJ may hence facilitate the automated environmental predictions
+weigh hypothetical possibilities, and forecast event outcomes.
+%
+Neural processes in the LTPJ may hence
+contribute to the automated predictions of the environment
 by incorporating experience-derived building blocks of world regularities
@@
 \mapsto \tilde{Q}(s, a|\theta_{k-1})$.
-From a neurobiological perspective, experience replay can be manifested as the re-occurrence of neuron spiking sequences that have also occurred during specific prior actions and environmental states.
+From a neurobiological perspective, experience replay can be manifested as the re-occurrence of neuron spiking sequences that have also been measured during specific prior actions and environmental states.
 The HC is a strong candidate for contributing to such neural reinstantiation of behavioral episodes as neuroscience experiments have repeatedly indicated in rats, mice, cats, rabbits, songbirds, and monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}.

### 2017-12-07 paper/paper.tex
reread part 3

@@
 takes the future into account, as some rewards could be \textit{delayed}.
-Through repeated interactions with the environment,
+Through repeated action on and feedback from the environment,
 the agent learns to reach goals and optimize reward signals
-in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
+in a trial-and-error fashion (Fig. \ref{fig:rl}).
 At a given moment, each taken \textit{action} $a$ triggers a change
@@
 \textit{partially observable} in the sense that only limited aspects of the environment's
-state are accessible to the agent's sensory perception.
+state are accessible to the agent's sensory perception
 \citep{starkweather2017dopamine}.
@@
 particularly advanced cognitive processes
-can be parsimoniously explained as probabilistic mental
-scene of experience simulations coupled with prediction error minimization
+can be parsimoniously explained as instantiating probabilistic mental
+simulations of experience coupled with prediction error minimization
 to calibrate action trajectories for
@@
 Such a purposeful optimization objective
-may be solved by a stochastic approximation
+may be subserved by a stochastic approximation
 based on a brain implementation of MCMC sampling.
@@
     the agent takes an action by following the policy matrix as updated by the Bellman equation.
-The agent receives a consequential reward and observes the next state.
+The agent receives a triggered reward and observes the next state.
 The process goes on until interrupted or a goal state is reached.}
@@
 \subsection{Markov decision processes}
-Reinforcement learning (RL) has had great success in many real-world problems, including robotics \citep{ng2004,abbeel2004},
-super-human performance in complex video games \citep{mnih2015},
-and strategic board games like the breakthrough results upon recently on the game of Go \citep{silver2016mastering}, considered to be a golden benchmark problem in artificial intelligence.
+Reinforcement learning (RL) has had considerable success in modeling many real-world problems, including super-human performance in complex video games \citep{mnih2015},
+robotics \citep{ng2004,abbeel2004},
+and strategic board games like the breakthrough results upon recently on the game of Go \citep{silver2016mastering} considered to be a milestone benchmark in artificial intelligence.
 We emphasize that the brain in general, and the DMN in particular,
@@
 Instead, the agent tries to optimize a \textit{subjective} reward
-signal (i.e., is likely to be different for another agent in the same state),
+signal (i.e., likely to be different for another agent in the same state),
 by maintaining probability distributions over actions according to their expected utility.
@@
 % % , based on a set of observations and observation probabilities.
-This is a minimal set of assumptions that can be made about an environment faced by an agent engaged in active learning.
+This is a minimal set of assumptions that can be made about an environment faced by an agent engaged in interactive learning.
 
@@
 According to how the market responds, the agent incurs definite gain or loss as environmental feedback of the executed financial actions.
-Recent research on automatizing market exchanges by algorithmic trading has successfully used MDPs as a framework for modeling these sophisticated behavioral process  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010}.
-~\citep{abergel2017} is an excellent reference with the mathematical details on how MDPs can be readily exploited in the financial world.
+Recent research on automatizing market exchanges by algorithmic trading has successfully used MDPs as a framework for modeling these sophisticated behavioral process  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
 %
@@
 Indeed,
-it has been proposed \citep{gershman2015computational} that a core property of human intelligence  underlie improvement
+it has been proposed \citep{gershman2015computational} that a core property of human intelligence is the improvement
 of expected utility outcomes as a strategy for action choice in uncertain
@@
 It has also long been proposed \citep{dayan2008decision} that there
-is a rather direct mapping of model-free RL learning algorithms
-onto aspects of the brain.
+can be a direct mapping between model-free RL learning algorithms
+and aspects of the brain.
 The neurotransmitter dopamine could serve
@@
 % Model-based assumes knowledge of the environment in terms of the transition probabilities between states, the reward contingencies for state-action pairs, or any form of representation of the environment.
-Such knowledge can of course be partly learned from the environment: the agent learns a model for the world alongside learning to make optimal decisions using this model.
-These methods learn what the effect is going to be of taking an particular action in a particular state.
+Such knowledge can be partly extracted from the environment: the agent learns a model for the world alongside learning to make optimal decisions based on this model.
+These methods learn what the effect is going to be of taking a particular action in a particular state.
 The result is an estimate of the underlying MDP which can then be either solved exactly or approximately, depending on the setting and what is feasible.
-The main problem then is to "connect the dots", by using dynamic programming --for example.
-
-In contrast, model-free methods require no knowledge of the environment (transition probabilities, identities of sensory input, etc.) or representation thereof, and the agent learns which state-action pairs lead to reward through sampling the world in trial-and-error manner, and predicting long-run reward aggregates, using prediction error as an incentive.
-In so doing, such methods ultimately learn both a policy and  implicit representations of the world / environment.
-This is what algorithms TD learning~\citep{sutton1998reinforcement}.
-In terms of number of states, model-based RL scales better than model-free RL; in terms of the time horizon, model-free RL scales better~\citep{strehl2006}.
-Drawing from a vast literature, we will show that the DMN contains sufficient neurobiological hardware to implement model-free RL, at least in principle.
+
+In contrast, model-free methods require no prespecified knowledge of the environment (transition probabilities, identities of sensory input, etc.) or representation thereof, and the agent learns which state-action pairs lead to reward by sampling the world in a trial-and-error manner, and inferring long-run reward aggregates, using prediction error as an incentive.
+In so doing, such an intelligent agent ultimately
+learns both a policy how to act on the environment and an implicit representations of the world 
+\citep{sutton1998reinforcement}.
+Model-free RL may struggle with high numbers of states, but scales better with the time horizon than model-based RL \citep{strehl2006}.
+Drawing from a vast literature, we hope to show that the DMN can be expected to harbor the neurobiological hardware necessary to implement the principles of a model-free RL agent.
 
@@
 The behavior of the agent is governed by a \textit{policy}, which maps states of the world
-to probability distributions over actions.
+to probability distributions over candidate actions.
 Starting at time $t=0$,
-following a policy $\pi$ generates a trajectory of action choices as follows:
+following a policy $\pi$ generates a trajectory of action choices:
 \begin{eqnarray*}
@@
 formal account of how the human association cortex
-might implement multi-sensory representation and control of the environment to
-optimize the organism's interaction with the world.
+can be thought to implement multi-sensory representation and decision making to
+optimize the organism's intervention on the world.
 % by obtaining reward and avoiding pain.
-This idealized process model explains
-a number of previous experimental observations in the
+This idealized process model accommodates
+a number of previous observations from neuroscience experiments on the
 DMN by simple but non-trivial mechanisms.
@@
 From the view of a Markovian sequential decision process,
-human behavior unfolds by integrating action outcomes from policy matrix (mapping from states to actions which maximize expected cumulative reward; see subsection \ref{sec:policymat})
+human behavior unfolds by inferring cumulative reward outcomes from hypothetical action cascades
 % stored past events
-and extrapolation to
-upcoming events for guiding action choice in the present context.
+and extrapolation from past experience to
+upcoming events for guiding behavior in the present.
 MDPs also provide a mathematical formalism how
@@
 abstract reasoning, social cognition,
-as well as delay discounting and self-related prospection.
+as well as delay discounting and self prospection into the future.
+%
 Improvement of the internal world representation
@@
 highly accurate memories have been disfavored in human evolution
-and why human creativity might be adaptive.
-
+and why human creativity may be adaptive.
+
+A major hurdle in explaining DMN activity from brain-imaging studies has been its very similar engagement across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., daytime mind-wandering), and thinking about future scenarios (e.g., delay discounting).
+The MDP account of DMN function offers a natural integration of these a-priori diverging mental processes into a common framework.
+%
 It is an important feature of the proposed artificial intelligence perspective
 on DMN biology that it is practically computable and
-yields falsifiable neuroscientific hypotheses.
-%
-Neuroscience experiments could be designed that operationalize
-the set of action, value, and state variables that govern
+proposes neuroscientific hypotheses that can be put to the test in future research.
+%
+Neuroscience experiments on the DMN should be designed that operationalize
+the set of action, value, and state variables governing
 the behavior of intelligent RL agents.
@@
 describe, contextualize, and interpret experimental findings in neuroscience studies
-on the DMN.
+on higher-level cognition.
 %
@@
 for continued refinement of the organism's fate.
-Indeed, a major hurdle in explaining DMN activity from brain-imaging studies has been its very similar engagement across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., dreams during sleep or daytime mind-wandering), and thinking about future scenarios (e.g., prospection or delay discounting).
-The MDP account of DMN function offers a natural integration of these a-priori diverging mental processes into a common framework.
-% It is this feature of our functional view that we underline in the referred sentence.
 

### 2017-12-07 paper/paper.tex
dayan2014 citation

@@
 
-In contrast, model-free methods require no prespecified knowledge of the environment (transition probabilities, identities of sensory input, etc.) or representation thereof, and the agent learns which state-action pairs lead to reward by sampling the world in a trial-and-error manner, and inferring long-run reward aggregates, using prediction error as an incentive.
-In so doing, such an intelligent agent ultimately
-learns both a policy how to act on the environment and an implicit representations of the world 
-\citep{sutton1998reinforcement}.
-Model-free RL may struggle with high numbers of states, but scales better with the time horizon than model-based RL \citep{strehl2006}.
-Drawing from a vast literature, we hope to show that the DMN can be expected to harbor the neurobiological hardware necessary to implement the principles of a model-free RL agent.
+In contrast, model-free methods require no knowledge of the environment (transition probabilities, identities of sensory input, etc.) or representation thereof, and the agent learns which state-action pairs lead to reward through sampling the world in trial-and-error manner, and predicting long-run reward aggregates, using prediction error as an incentive.
+In so doing, such methods ultimately learn both a policy and  implicit representations of the world / environment.
+This is what algorithms TD learning~\citep{sutton1998reinforcement}.
+The distinction outlined above between model-free and model-based RL is similar to ~\citep{dayan2014}.
+
+In terms of number of states, model-based RL scales better than model-free RL; in terms of the time horizon, model-free RL scales better~\citep{strehl2006}.
+Drawing from a vast literature, we will show that the DMN contains sufficient neurobiological hardware to implement model-free RL, at least in principle.
 

### 2017-12-07 paper/paper.tex
minor edits

--- 2017-12-07 paper/paper.tex
@@
 % official NIPS keywords
-\textbf{\\keywords}: mind-wandering, artificial intelligence, Markov decision processes
+\textbf{\\keywords}: systems neuroscience, artificial intelligence, reinforcement learning, mind-wandering
 
@@
 This idealized process model accommodates
-a number of previous observations from neuroscience experiments on the
+a number of previous observations from neuroscience studies on the
 DMN by simple but non-trivial mechanisms.
@@
 opportunity in the environment can be recursively exploited
-when confronted with challenging decisions.
+when confronted with challenging interdependent decisions.
 %
@@
 
-A major hurdle in explaining DMN activity from brain-imaging studies has been its very similar engagement across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., daytime mind-wandering), and thinking about future scenarios (e.g., delay discounting).
-The MDP account of DMN function offers a natural integration of these a-priori diverging mental processes into a common framework.
-%
-It is an important feature of the proposed artificial intelligence perspective
+A major hurdle in understanding DMN activity from cognitive neuroscienceå studies has been its similar engagement across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., daytime mind-wandering), and thinking about predicted scenarios (e.g., delay discounting).
+The MDP account of DMN function offers a natural integration of a-priori diverging neural processes into a common framework.
+%
+It is an important advantage of the proposed artificial intelligence perspective
 on DMN biology that it is practically computable and
-proposes neuroscientific hypotheses that can be put to the test in future research.
+readily motivates neuroscientific hypotheses that can be put to the test in future research.
 %
@@
 Ultimately,
-DMN activity may instantiate a holistic integration
-ranging from real experience over purposeful dreams to anticipated futures
-for continued refinement of the organism's fate.
+neural processes in the DMN may instantiate a holistic integration
+ranging from real experience over purposeful dreams to predicted futures
+to continuously refine the organism's fate.
 

### 2017-12-07 paper/paper.tex
a

@@
 %
-Specifically, we answer the question whether the neurobiological properties of the default mode network collectively provide the computational building blocks necessary for a Markov Decision Process.
+Specifically, we answer the question whether the neurobiological properties of the DMN
+collectively provide the computational building blocks necessary for a Markov Decision Process.
 %
@@
 exploratory action with longer-term gratification.
-
 We argue that DMN implication in many
@@
 the current sensory environment.
-
-The more familiar and predictable the current environment, the more resources remain for allocating DMN activity to MDP processes extending beyond the present time and sensory context.
+The more familiar and predictable the current environment, the more brain resources may remain for allocating DMN activity to MDP processes extending beyond the present time and sensory context.
 In line with this perspective, DMN engagement was shown to heighten and relate to effective behavioral responses in the practiced phase of a demanding cognitive flexibility task, as compared to acquisition phase when participants learned context-specific rules.
-This involvement in automated decision-making lead the authors to propose an ``autopilot'' role for the DMN \citep{vatansever2017pnas}, which may contribute to optimizing control of the organism in general.
-Among all DMN nodes, the RTPJ has probably a most evident role as network-switching relay calibrating between processing of environment-engaged versus internally generated information \citep{downar2000multimodal, golland2006extrinsic, bzdok2013tpj}.
+This involvement in automated decision-making has led the authors to propose an ``autopilot'' role for the DMN \citep{vatansever2017pnas}, which may contribute to optimizing control of the organism in general.
+Among all DMN nodes, the RTPJ is perhaps the most evident candidate for a network-switching relay that calibrates between processing of environment-engaged versus internally generated information \citep{downar2000multimodal, golland2006extrinsic, bzdok2013tpj}.
 
 Additionally,
-this network was proposed to be situated
+the DMN was proposed to be situated
 at the top of the brain network hierarchy, with
@@
 \citep{carhart2010default, margulies2016situating}.
-Its putative involvement in thinking about the past,
-hypothetical experiences, and the future
+Its putative involvement in thinking about 
+hypothetical experiences and future outcomes
 appears to tie in with the implicit computation of
@@
 experienced events and collected feedback from the past.
-
-A policy matrix encapsulates the repertoire of possible actions
-on the world given a current state.
-The policy matrix encodes the probabilites of
-choosing actions to be executed in a certain situation.
+A policy matrix encapsulates the choice probabilities of action repertoire
+on the world given a current situation (i.e., state).
 The DMN may subserve
@@
 as batches of pseudo-experience
-(i.e., imagined, hypothesized, and recalled mental scenarios).
+(i.e., recalled, hypothesized, and forecasted mental scenarios).
 The small variations in these experience samplings allow searching
@@
 the DMN may have previously stochastically generated, evaluated, and adapted to
-such a randomly synthesized situation.
+such a randomly synthesized event.
 Generated representations arguably are ``internally manipulable,
@@
 \citep{pezzulo2011grounding}.
-In the absence of environmental input and feedback
+In the context of scarce environmental input and feedback
 (e.g., mind-wandering or sleep),
-mental scene construction allows for pseudo-experiencing possible
+mental scene construction allows pseudo-experiencing possible
 future scenarios and action outcomes.
-Our formal account of DMN function
-thus acknowledges the unavoidable stochasticity of
-computation in neural systems \citep{faisal2008noise}.
 
 From the perspective of a model-free RL agent,
-\textit{prediction} in the human brain reduces to
+prediction in the DMN reduces to
 generalization of
@@
   \vspace{1em}
-  \textbf{Box 2: Hypotheses for testing the MDP account of DMN function}
+  \textbf{Box 2: Future studies for testing the MDP account of DMN function}
 \begin{enumerate}

### 2017-12-08 paper/paper.tex
reread part 4

@@
 the DMN at the top and the salience and dorsal attention systems
-at intermediate levels, above thalamic and primary sensory
+at intermediate levels, above thalamic and unimodal sensory
 cortex'' \citep{carhart2010default}.
@@
 DMN dynamics could be constantly shaped by prediction error through feedback loops.
-This computational account of DMN will be described in the mathematical framework of Markov Decision Processes (MDPs).
-MDPs indeed address specifically the case of decision making in stochastic context with reward feedback.
+The present computational account of DMN function will be described in the mathematical framework of Markov Decision Processes (MDP).
+MDPs specifically formalize decision making in stochastic contexts with reward feedback.
 
@@
 multi-step decision processes in such an environment are
-\textit{Markov decision processes} (MDPs) \citep{sutton1998reinforcement}.
+MDPs \citep{sutton1998reinforcement}.
 % For the remainder of this paper, for the sake of simplicity
@@
 \end{equation}
-This random variable\footnote{Random as it depends both on the environment's dynamic and the
-  policy $\pi$ being played (which can be stochastic).
-The exponential delay discounting function used here refere to the usual formulation in the field of reinforcement learning but psychological experiments may also reveal other discounting regimes \citep{green2004discounting}}  measures the cumulative reward of
+This random variable
+
+  \footnote{Random as it depends both on the environment's dynamics and the
+  policy $\pi$ being executed.
+The exponential delay discounting function used here refers to the usual formulation in the field of reinforcement learning, although psychological experiments may also reveal other discounting regimes \citep{green2004discounting}}
+
+measures the cumulative reward of
 following an action policy $\pi$.

### 2017-12-08 paper/paper.tex
finalizing rebuttal

@@
 and strategic board games like the breakthrough results upon recently on the game of Go \citep{silver2016mastering} considered to be a milestone benchmark in artificial intelligence.
-We emphasize that the brain in general, and the DMN in particular,
-is a physical system governed by the laws of
-physics and can be formally described
-by Markov processes at a sufficiently coarse scale.
-It has indeed been previously proposed \citep{tegmark2016improved} that
-any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
-step is sufficiently short.
-
+% We emphasize that the brain in general, and the DMN in particular,
+% is a physical system governed by the laws of
+% physics and can be formally described
+% by Markov processes at a sufficiently coarse scale.
+% It has indeed been previously proposed \citep{tegmark2016improved} that
+% any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
+% step is sufficiently short.
 In artificial intelligence and machine learning, a popular computational model for

### 2017-12-08 paper/paper.tex
reread part 5

--- 2017-12-08 paper/paper.tex
@@
 % , defining $s(t)$, the state at time $t$, as the position in state space.
-The process has $memory$ if the next state depends not only on the current state but also on a finite number of past states.
+The process has $memory$ if the next state depends not only on the current state but also on a number of past states.
 Rational probabilistic planning can thus be reformulated as a standard memoryless Markov process by simply expanding the definition of the state $s$ to include experience episodes of the past.
-This adds the capacity for memory, since then, the next state depends not only on the current state but also on a finite number of past states.
-This is precisely what Partially Observable MDPs (POMDPs) aim to realize ~\citep{starkweather2017dopamine,o2006making}.
-Nevertheless, this mathematical property in POMDPs mostly accounts for implicit $memory$.
-Since the current paper is focused on both plausibility at the behavioral and neurobiological level, we will address below how our model captures the neurophysiological constraints of the DMN and can also account for the explicit part of $memory$ characteristic for human agents.
+This adds the capacity for memory to the model because the next state then depends not only on the current situation but also on previously experienced events, which is the motivation behind Partially Observable MDPs (POMDPs)
+\citep{starkweather2017dopamine,o2006making}.
+Nevertheless, this mathematical property of POMDPs mostly accounts for implicit $memory$.
+Since the current paper is concerned with plausibility at the behavioral and neurobiological level, we will address below how our account can accommodate the neurophysiological constraints of the DMN and the explicit $memory$ characteristics of human agents.
 
@@
 
-\paragraph*{On implicit and explicit memory in the current model.}
-Markov Processes are memoryless but this it is mathematically feasible to insert the previous states of such model into the current state.
-While this may partially account for implicit memory at the behavioral level, it may not explain the underlying implementation at the neurophysiological level nor accounts for the explicit part of memory.
-Implicit memory-based processing arises in our MDP account of DMN function in three different forms: successive updates of a) the action policy and b) the value function, both being products of the past and c) the deep non-linear links in the hierarchical connections realized by biological neural networks.
-Indeed, we reason about the brain has deep neural networks that present an implicit form of information compression obtained from life experience.
+\paragraph*{A note on implicit and explicit memory.}
+While Markov Processes are usually memoryless, it is mathematically feasible to incorporate the previous states of such model into the current state.
+This extension may partially account for implicit memory at the behavioral level, but may not explain the underlying neurobiological implementation or accommodate explicit memory.
+Implicit memory-based processing arises in our MDP account of DMN function in several different forms: successive updates of a) the action policy and the value function, both being products of the past, as well as b) the deep non-linear relationships within the hierarchical connections of biological neural networks.
+The brain's adaptive synaptic connections can be thought as deep neural networks affording an implicit form of information compression of life experience.
 Such memory traces are implicitly stored in the neural machinery and can be retrieve as a form of knowledge during simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}.
-The neural processes in the hippocampus can be seen as MCMC sampling from a formal statistical level but those memory recall are also the basis for scene construction processes across time-scales such as in the semantic hypothesis \citep{schacter2007remembering,Axelrod2017}.
+c) The neural processes in the hippocampus can be seen as some type of MCMC sampling for memory recall, which can also be a basis for scene construction processes across time-scales \citep{schacter2007remembering,Axelrod2017}.
 

### 2017-12-08 paper/paper.tex
interim

@@
 The brain's adaptive synaptic connections can be thought as deep neural networks affording an implicit form of information compression of life experience.
-Such memory traces are implicitly stored in the neural machinery and can be retrieve as a form of knowledge during simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}.
-c) The neural processes in the hippocampus can be seen as some type of MCMC sampling for memory recall, which can also be a basis for scene construction processes across time-scales \citep{schacter2007remembering,Axelrod2017}.
+Such memory traces are stored in the neural machinery and can be implicitly retrieve as a form of knowledge during simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}.
+c) Certain neural processes in the hippocampus can be seen as some type of MCMC sampling for memory recall, which can also be a basis for mental imagery across time-scales \citep{schacter2007remembering,Axelrod2017}.
 

### 2017-12-09 paper/paper.tex
a

@@
 We argue that our formal account of DMN function naturally accommodates as special cases
-the previously proposed cognitive accounts on
+the previously proposed functional accounts on
 (1) predictive coding,
@@
 we derive explicit hypotheses to be tested in future neuroscience
-experiments (Box 2).
+experiments.
 At this stage, we emphasize the importance of differentiating which levels of observation are at play in the present account.
@@
 % Model-based assumes knowledge of the environment in terms of the transition probabilities between states, the reward contingencies for state-action pairs, or any form of representation of the environment.
-Such knowledge can be partly extracted from the environment: the agent learns a model for the world alongside learning to make optimal decisions based on this model.
+Such knowledge can be partly extracted from the environment: the agent learns a model of the world while learning to take optimal decisions based on this model.
 These methods learn what the effect is going to be of taking a particular action in a particular state.
@@
 
-In contrast, model-free methods require no knowledge of the environment (transition probabilities, identities of sensory input, etc.) or representation thereof, and the agent learns which state-action pairs lead to reward through sampling the world in trial-and-error manner, and predicting long-run reward aggregates, using prediction error as an incentive.
-In so doing, such methods ultimately learn both a policy and  implicit representations of the world / environment.
-This is what algorithms TD learning~\citep{sutton1998reinforcement}.
-The distinction outlined above between model-free and model-based RL is similar to ~\citep{dayan2014}.
-
-In terms of number of states, model-based RL scales better than model-free RL; in terms of the time horizon, model-free RL scales better~\citep{strehl2006}.
+In contrast, model-free methods require no particular knowledge of the environment (transition probabilities, types of sensory input, etc.) or representation thereof.
+The agent learns which state-action pairs lead to reward through sampling the world in a trial-and-error manner and infers longer-term reward aggregates using environmental feedback information as an incentive.
+In so doing, model-free agents ultimately learn both an action policy and an implicit representations of the world.
+Our distinction between model-free and model-based RL is hence similar to previous views \citep{dayan2014}.
+
+% In terms of number of states, model-based RL scales better than model-free RL; in terms of the time horizon, model-free RL scales better~\citep{strehl2006}.
 Drawing from a vast literature, we will show that the DMN contains sufficient neurobiological hardware to implement model-free RL, at least in principle.
@@
 This random variable
-
   \footnote{Random as it depends both on the environment's dynamics and the
@@
 The exponential delay discounting function used here refers to the usual formulation in the field of reinforcement learning, although psychological experiments may also reveal other discounting regimes \citep{green2004discounting}}
-
 measures the cumulative reward of
@@
 
-\paragraph*{Self-training may introduce bias. How to alleviate it ?}
-As pointed out by one of the reviewers, self-training may introduce bias due to information shortage caused by the absence of external stimulus (task).
- way to solve this issue is to use importance sampling as was done in ~\citep{schaul2015} and ~\citep{hessel2017}, to replay more often state-transitions from which there is more to learn.
-New transitions are inserted into the replay buffer with maximum priority, providing a bias towards recent transitions.
-Such an insertion bias would help counter-balance the bias introduced by the information shortage caused by the absence of external stimulus (task).
-Also, ~\citep{hessel2017} noticed that such prioritized replay reduces the data-complexity and the agent can start learning much sooner than conventionally.
+\paragraph*{A note on bias in self-training.}
+Some bias may be introduced by self-training due to information shortage caused by the absence of external stimulation.
+ One way to address this issue is using importance sampling to replay especially those state-transitions from which there is more to learn \citep{schaul2015, hessel2017}.
+New transitions are inserted into the replay buffer with maximum priority, thus shifting emphasis to more recent transitions.
+Such insertion strategy would help counter-balance the bias introduced by the information shortage incurred by absent external input.
+Other authors noticed \citep{hessel2017} that such prioritized replay reduces the data complexity and the agent shows faster increases in learning performance.
 
@@
   \vspace{1em}
-  \textbf{Box 2: Future studies for testing the MDP account of DMN function}
+  \textbf{Box 2: Proposed studies for testing the MDP account of DMN function}
 \begin{enumerate}
@@
 During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigm.
-We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, and the more the fMRI signals in the reward circuits should be independent of the reward context
-in the sensory environment.
-
-\item Experiment (Humans): We hypothesize a functional dissociation between computations for behavioral policy versus adapting stimulus-value associations as these are conceivably implemented in different subsystems of the DMN.
-We first expect that fMRI signals in the right temporo-parietal junction can predict behavioral changes caused by adaptation in the action choice tendencies (policy matrix) related to non-value-related prediction error.
-Second, fMRI signals in the ventromedial prefrontal cortex should predict behavioral changes caused by adaptation in value estimation (value matrix) due to reward-related stimulus-value association.
-We finally predict that fMRI signals in the posteromedial cortex, as a potential global information integrator, is able to predict shifts in overt behavior based on previous adaptations in policy or value estimation.
-
-\item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is necessary for choice behavior in mice.
+We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, the more the fMRI signals in the reward circuits should be independent of the reward context in the current sensory environment.
+
+\item Experiment (Humans): We hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
+First, we expect that fMRI signals in the right temporo-parietal junction relate to behavioral changes subsequent to adaptation in the action choice tendencies (policy matrix) involved in non-value-related prediction error.
+Second, fMRI signals in the ventromedial prefrontal cortex should relate to behavioral changes following adaptation in value estimation (value matrix) due to reward-related stimulus-value association.
+We finally expect that fMRI signals in the posteromedial cortex, as a potential global information integrator, are related to shifts in overt behavior based on previous adaptations in both policy or value estimation.
+
+\item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is contributing to choice behavior in mice.
 Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future.
-Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal disruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
-
-\item Experiment (Humans): We hypothesize that the relevant time horizon is modulated by various factors such as organism age, acute stress, and time-enduring impulsivity traits.
-Using a temporal discounting experiment, it can be measured how the time horizon could be affected at the behavioral level, and then trace back its potential representation at the neural level.
-The experiment can be designed to examine between-group (by age or with a impulsive population like chronic gamblers or drug addicts) and within group effects using a stress; and brought in context with the participant’s age and impulsivity tendencies.
+Necessity of cortical DMN regions, in addition to the hippocampus, for “mind-searching” candidate actions during choice behavior can be experimentally corroborated by causal disruption of DMN regions, such as by circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
+
+\item Experiment (Humans): We hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits.
+Using a temporal discounting experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced-back to its corresponding neural representation.
+Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age and personality traits.
 
@@
 
-\item Experiment (Humans \& Animals): An additional layer of learning concerns the addition of new entries in the table (i.e., state and action).
-The framing of the policy matrix is not described here but could be supported by synaptic epigenesis \citep{gisiger_acquisition_2005}.
+\item Experiment (Humans \& Animals): An additional layer of learning concerns the addition of new entries in the state and action spaces.
+Extension of the action repertoire could be biologically realized by synaptic epigenesis \citep{gisiger_acquisition_2005}.
 Indeed, the tuning of synaptic weights through learning can stabilize additional patterns of activity by creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}.
-Those attractors will then constrain both the number of factors taken into account by decision processes and the number of behaviors at reach by the agent \citep{wang_decision_2008}.
-To probe this aspects, we propose two experiments probing how synaptic epigenesis forms a neural correlate of the policy matrix update: one in human, investigating the changes of functional connectivity between DMN nodes following a temporal discounting experiment; the second in rodents, directly testing with anterograde tracing how homolog regions of the DMN present increased synaptic changes compare to other brain regions.
+Those attractors can then constrain both the number of factors taken into account by decision processes and the possible behaviors of the agent \citep{wang_decision_2008}.
+To examine this potential higher-level mechanism, we propose to probe how synaptic epigenesis is related to neural correlates underlying policy matrix updates: in humans the changes of functional connectivity between DMN regions can be investigated following a temporal discounting experiment and in monkeys or rodents anterograde tracing can be used to study how homolog regions of the DMN present increased synaptic changes compare to other parts of the brain.
 
@@
 may enable learning about aspects of the world by reconciling
-gaps between fresh sensory input and expectations computed
+gaps between fresh sensory input and predictions computed
 based on stored prior information.
 At each stage of neural processing,
-an internally generated prediction of aspects of environmental sensations is
+an internally generated expectation of aspects of environmental sensations is
 directly compared against the actual environmental input.
@@
 induces plasticity changes of neuronal projections
-(i.e., adapting model parameters)
+% (i.e., adapting model parameters)
 to allow for gradually improved future prediction of the environment.
@@
 \citep{friston2010free, buzsaki2006rhythms} and
-the intimate relation of motor action to sensory expectations
+the intimate relation of motor movement to sensory expectations
 \citep{wolpert1995internal, kording2004bayesian}.
@@
 may be maintained by top-down modulation
-using a-priori information about the environment.
+using internally generated information about the environment.
 
@@
 to best accommodate and prepare the organism
-for processing the constant influx of sensory stimulation and
+for processing the constant influx of sensory stimuli and
 performing action on the environment.
@@
 Neuroanatomically, axonal
-back projections indeed outnumber by far the axonal input projections
-existing in the monkey and probably also human brain
+back projections indeed outnumber by far the axonal connections mediating feedforward input processing
+in the monkey brain and probably also in humans
 \citep{salin1995corticocortical}.
-These many and diverse modulatory influences
+These many and diverse top-down modulations
 from higher onto downstream cortical areas
@@
 serve as a concrete implementation of
-predictive coding ideas from cognitive neuroscience.
+predictive coding ideas.
 MDPs have the advantage of exposing an explicit
-mechanism for the horizon of future considerations and for
+mechanisms for controlling the horizon of future considerations and for
 how the internal representation of the world is updated,
-as well as
-why certain predictions may be more relevant to the agent than others.
+as well as why certain predictions may be more relevant to the agent than others.
 
 \subsection{Semantic hypothesis}
-Another frequently proposed cognitive account to explain DMN function revolves
+This frequently proposed cognitive account to explain DMN function revolves
 around forming logical associations and abstract analogies between
@@
 semantic building blocks may need to get recombined to enable
-mental imagery of non-existing scenarios.
-As such, semantic knowledge would be a prerequisite for optimizing behavior
+mental imagery to (fore)see never-experienced scenarios.
+As such, semantic knowledge would be an important ingredient for optimizing behavior
 by constantly simulating possible future scenarios
@@
 the internal construction and elaboration of necessary information
-that is not presented in the surrounding sensory environment
+that is not presented in the immediate sensory environment
 by recombining building blocks of
@@
 \citep{schacter2007remembering}.
-``[P]redictions and simulations are not only automatically elicited by external events but can be endogenously generated when needed.
+``[Foresight] and simulations are not only automatically elicited by external events but can be endogenously generated when needed.
 [...] The mechanism of access via simulation could be a widespread method for accessing and producing knowledge, and represents a valid alternative to the traditional idea of storage and retrieval''
 \citep{pezzulo2011grounding}.
-Such scene construction processes could contribute to interpreting the
-present and foretelling the future.
-Further,
-mental scene construction has been proposed
+Such mental scene-construction processes could contribute to interpreting the
+present and foreseeing the future.
+Further, mental scene imagery has been proposed
 to imply a distinction between
@@
 while simulation of possible alternatives are constructed,
-followed by a return to perception of the present.''.
+followed by a return to perception of the present''.
 % The semantic hypothesis is for instance supported by evidence in animals that
@@
 both the semantic hypothesis and the present formal account based on MDPs
-expose mechanisms of how action considerations could be mentally explored.
+expose mechanisms of how action considerations could be explored.
 In both accounts,
-there is also no reason to assume that predictions of various
+there is also little reason to assume that contemplating alternative realities of various
 levels of complexity, abstraction, timescale, and purpose
-rely on mechanisms that are qualitatively different. This concurs with
+rely on mechanisms that are qualitatively different. This interpretation concurs with
 DMN activity increases across time, space, and content domains
@@
 Further, the semantic hypothesis
-and MDP account provide explanations why HC damage does
-not only impair recalling memories, but also hypothetical and future
-thinking \citep{hassabis2007patients}.
+and MDP account offer explanations why HC damage does
+not only impair recalling past events, but also imagining hypothetical and future
+scenarios \citep{hassabis2007patients}.
 While both semantic hypothesis and
-our formal account propose memory-based internally
-generated information for probabilistic mental models of action outcomes,
+our formal account propose memory-enabled, internally
+generated information for probabilistic representation of action outcomes,
 MDPs render explicit the grounds on which an action is
@@
 Finally,
-the semantic account of DMN function does not offer
-a mechanistic explanation \textit{how}
-explicit world knowledge and semantic analogies thereof
-lead to prediction of future actions and states.
-The semantic hypothesis does also not explain why memory recall
+the semantic account of DMN function does not provide
+suffcient explanation of \textit{how}
+explicit world knowledge and logical analogies thereof
+lead to foresight of future actions and states.
+The semantic hypothesis does also not fully explain why memory recall
 for scene construction in humans is typically fragmentary and noisy
@@
 by MCMC approximation achieves better fine-tuning of the
-action policies and estimation of expected reward outcomes.
-That is, the purposeful stochasticity of policy and value estimation
+action policies and inference of expected reward outcomes.
+That is, the purposeful stochasticity of policy and value updates
 in MDPs provides a candidate explanation for why humans
@@
 \subsection{Sentinel hypothesis}
-The DMN regions have been associated with processing the
+Regions of the DMN have been proposed to process the
 experienced or expected relevance of environment cues
@@
 Processing self-relevant information was perhaps the first
-cognitive account that was proposed for DMN function
+functional account that was proposed for the DMN
 \citep{gusnard2001medial, raichle2001pnas}.
@@
 encountering a self-relevant environmental stimulus.
-However,
+Instead,
 in the present MDP account,
@@
 certain branches of candidate action trajectories
-are detected to be less worthy to become mentally explored.
+are detected to be less worthy to get explored.
 This mechanism, expressed by the Bellman equation,
@@
 if real or imagined experience is processed, manipulated, and
-used for predictions.
-During tasks,
+used in service of organism control.
+During active engagement in tasks,
 the policy and value estimates may be
 updated to optimize especially short-term action.
-At rest, these parameter updates may
+At passive rest, these parameter updates may
 improve especially mid- and long-term action.
@@
 of the same neurobiological brain circuit (i.e., DMN) has been documented
-for specific task performances and baseline house-keeping functions.
+for specific task performances and baseline "house-keeping" functions.
 
 In particular,
-environmental stimuli are especially important for humans are frequently of
-social nature. This is not surprising
+environmental cues that are especially important for humans are frequently of
+social nature. This may not be surprising
 given that
@@
 \citep{whiten1988machiavellian}.
-Indeed, social topics amounted to roughly
-two thirds of human everyday communication \citep{dunbar1997human},
-while
-mind-wandering at daytime and dreams during sleep
-are rich in stories about people and
-the complex relationships between them.
-%
-In line with this, the DMN was argued to be specialized in
+In fact, social topics probably amount to roughly
+two thirds of human everyday communication \citep{dunbar1997human}.
+Mind-wandering at daytime and dreams during sleep
+are also rich in stories about people and
+the complex interactions between them.
+%
+In line with this, DMN activity was advocated to be specialized in
 continuous processing of social information as a
@@
 and rats \citep{lu2012rat}, three species with
-social-cognitive capacities that are supposedly less advanced than in humans.
+social capacities that are supposedly less advanced than in humans.
 
@@
 the principal connectivity gradient in the cortex appears to be greatly expanded in humans compared to monkeys, suggesting a phylogenetically conserved axis of cortical expansion with the DMN emerging at the extreme end in humans \citep{margulies_situating_2016}.
-Neurocomputational models of dyadic whole-brain dynamics
+Computational models of dyadic whole-brain dynamics
 demonstrated how the human connectivity topology, on top of facilitating processing at the intra-individual level, can explain our propensity to coordinate through sensorimotor loops with others at the inter-individual level \citep{dumas_anatomical_2012}.
-The DMN is moreover largely overlapping with neural networks associated with higher level social cognition \citep{schilbach_introspective_2012}. For instance, the vmPFC,
+The DMN is moreover largely overlapping with neural networks associated with higher-level social processes \citep{schilbach_introspective_2012}. For instance, the vmPFC,
 PMC, and RTPJ
-together play a key role in bridging the gap between self and other by
+together may play a key role in bridging the gap between self and other by
 integrating low-level embodied processes within higher level inference-based mentalizing \citep{lombardo_shared_2009}.
@@
 According to theories of the \textit{free-energy principle} (FEP) and
-\textit{active inference} \citep{friston2010free, fristonAIorRL} (see also \citep{dayan1995helmholtz}),
+\textit{active inference} \citep{friston2010free, fristonAIorRL,dayan1995helmholtz},
 the brain corresponds to
 a biomechanical reasoning engine. It is
-dedicated to minimizing the long-term average of surprise: the log-likelihood of the observed sensory input --more precisely, an upper bound thereof-- \textit{relative} to the expectations about the external world derived from internal representations.
+dedicated to minimizing the long-term average of surprise: the log-likelihood of the observed sensory input --more precisely, an upper bound thereof-- relative to the expectations about the external world derived from internal representations.
 The brain
@@
 which is a measure of how different two probability distributions are.
-In this framework, the goal of the agent is then to iteratively refine the generative model $p_G$ and the recognition model $p_R$ (i.e two Bayesian belief nets) so as to minimize the free energy $F_G^R(\x)$ over sensory input $\x$.
-
-Importantly, one notes that $F^R_G(x)$ is low just in case:
+In this framework, the goal of the agent is then to iteratively refine the generative model $p_G$ and the recognition model $p_R$ so as to minimize the free energy $F_G^R(\x)$ over sensory input $\x$.
+
+Importantly, $F^R_G(x)$ gets low in the following cases:
 \begin{itemize}
@@
 % \paragraph{Main criticisms.}
-Despite its popularity, criticism against the FEP has arisen over the years,
+Despite its popularity, criticism against the FEP has been voiced over the years,
 some of which is outlined in the following.
 The main algorithm for minimizing free energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
-\citep{dayan1995helmholtz}. As P. Dayan and G. Hinton themselves noted,
-a crucial drawback of the wake-sleep algorithm (and therefore of theories like the FEP \citep{friston2010free} based on it) is that it involves a pair
-of forward (generation) and backward (recognition) models $p_G$ and $p_R$ respectively that
-together do not correspond to optimization of (a bound of) the marginal likelihood,
+\citep{dayan1995helmholtz}. As these authors noted,
+a crucial drawback of the wake-sleep algorithm (and therefore of theories like the FEP \citep{friston2010free}) is that it involves a pair
+of forward (generation) and backward (recognition) models $p_G$ and $p_R$ that
+together does not correspond to optimization of (a bound of) the marginal likelihood
 because KL divergence is not symmetric in its arguments.
-The brain is therefore unlikely to implement the wake-sleep algorithm or a variant thereof.
-
-The recent theory of
-\textit{variational auto-encoders} (VAEs) \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm. VAEs
+
+These considerations render the brain less likely to implement the wake-sleep algorithm or a variant thereof. More recently,
+\textit{variational auto-encoders} (VAEs) \citep{kingma2013auto} emerged that may provide an efficient alternative to the wake-sleep algorithm. VAEs
 overcome a number of the technical limits of the wake-sleep algorithm
@@
 
-On another front, since theories based on the FEP \citep{friston2010free,fristonAIorRL}
+The difference between the FEP and the MDP account may be further clarified by a thought experiment.
+Since theories based on the FEP \citep{friston2010free,fristonAIorRL}
 conceptualize ongoing behavior
@@
 % \footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.}
-(Fig. \ref{fig:darkroom}) would strive to remain in this location
+(Fig. \ref{fig:darkroom}) would remain trapped in this location
 because its sensory inputs are perfectly predictable
@@
 \end{wrapfigure}
-However, such a behavior is seldom observed in animals or humans in the
-real world: In a dark room, intelligent agents would search for light sources to \textit{explore}
-the surroundings or leave the exit the room.
+However, such a behavior is seldom observed in humans in the
+real world. In a dark room, the intelligent agents would search for light sources by explore
+the surroundings or aim to exit the room.
 Defenders of the FEP have retorted by advancing the ``full package'' \citep{darkroom2012}:
 FEP is proposed to be multi-scale and there would be a meta-scale at which the organism would
-be \textit{surprised} by such a \textit{lack of surprise}. According to this argument, a dark room would paradoxically correspond to a state of particularly high surprise.
-Driven by surprise-minimization objective, the FEP agent would eventually bootstrap itself out of such saddle points to explore more interesting parts of the environment.
+be \textit{surprised} by such a \textit{lack of surprise}. According to this argument, a dark room would paradoxically correspond to a state of particularly high relevance.
+Driven by the surprise-minimization objective, the FEP agent would eventually bootstrap itself out of such saddle points to explore more interesting parts of the environment.
 % Recourse to this argument is equivalent to invoking a \textit{halting oracle}\citep{haltingoracle}, an arbitrarily powerful machine which can solve the \textit{halting problem} (an arbitrarily difficult computational problem which cannot be solved under any reasonable \textit{model of computation}, not even in principle!).
@@
 In contrast, an organism operating under our RL-based theory would inevitably
-identify the sensory-stimulus-deprived room as a local minimum. Indeed, hippocampal experience replay (see \ref{sec:self}) would serve to sample memories or fantasies of alternative situations
+identify the sensory-stimulus-deprived room as a local minimum. Indeed, hippocampal experience replay (see \ref{sec:self}) could serve to sample memories or fantasies of alternative situations
 with reward structure. Such artificially generated \textit{internal} sensory input
 subserved by the DMN can entice the organism to explore the room,
-for instance by looking for and using the light switch or simply finding the room exit.
+for instance by looking for and using the light switch or finding the room exit.
 
 % \paragraph{Links between FEP and RL.}
-We note that FEP and active inference can be reframed in terms of
+We finally note that FEP and active inference can be reframed in terms of
 our model-free RL framework.
-This becomes possible by recasting the Q-value function (i.e expected long-term reward) maximized by the DMN to
+This becomes possible by recasting the Q-value function (i.e., expected long-term reward) maximized by the DMN to
 correspond to negative surprise,
 that is, the log-likehood of current sensory priors
-the agent has about the world. More explicitly, this corresponds to using free-energy as a Q-value approximator for the MDP, i.e
+the agent has about the world. More explicitly, this corresponds to using free-energy as a Q-value approximator for the MDP such that
 
@@
 %   \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}.
-Such a surprise-guided reinforcement learning scheme has previously been advocated
-under the equivalent framework of energy-based reinforcement-learning
+Such a surprise-guided RL scheme has previously been advocated
+under the equivalent framework of energy-based RL
 % \footnote{These methods assume that the posterior distribution  of the hidden states $\z$ given the visible states $\x$ factorizes, leading to so-called RBMs --\textit{Restricted Boltzmann Machines}.
@@
 Nevertheless, minimization of surprise quantities alone
-may be insufficient to explain the diversity of behaviors observed in humans and other intelligent animals.
+may be insufficient to explain the diversity of behaviors that humans and other intelligent animals can perform.
 % The dark-room experiment, for example, provides a formidable argument against in favor of the insufficiency of the FEP and active inference theories to explain intelligent behavior in humans: when a human is put in a dark room, they don't just sit down in a corner forever, what would be the FEP-optimal thing to do; they actively search for the light switch, and start exploring the depths of the room, possibly looking for food or the way out!
@@
 %
-From the view of a Markovian sequential decision process,
+Viewed as a Markovian sequential decision process,
 human behavior unfolds by inferring cumulative reward outcomes from hypothetical action cascades
@@
 by injecting stochasticity into the recall of past
-actions and the estimation of action outcomes may explain why
+actions and inference of action outcomes may explain why
 highly accurate memories have been disfavored in human evolution
@@
 
-A major hurdle in understanding DMN activity from cognitive neuroscienceå studies has been its similar engagement across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., daytime mind-wandering), and thinking about predicted scenarios (e.g., delay discounting).
+A major hurdle in understanding DMN activity from cognitive brain-imaging studies has been its similar neural engagement in different time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., daytime mind-wandering), and thinking about predicted scenarios (e.g., delay discounting).
 The MDP account of DMN function offers a natural integration of a-priori diverging neural processes into a common framework.

### 2017-12-09 paper/paper.tex
reread part 6

@@
 
-\author{Elvis Dohmatob\textsuperscript{1,2}, Guillaume Dumas\textsuperscript{4,5,6,7}, Danilo Bzdok\textsuperscript{1,2,3}
+\author{Elvis Dohmatob\textsuperscript{1,2}, Guillaume Dumas\textsuperscript{5,6,7,8}, Danilo Bzdok\textsuperscript{1,2,3,4}
 }
@@
     \textbf{3} Department of Psychiatry, Psychotherapy and Psychosomatics, RWTH Aachen University, Aachen, Germany\\
-    \textbf{4} Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
-    \textbf{5} CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
-    \textbf{6} University Paris Diderot, Sorbonne Paris Cit\'e, Paris, France\\
-    \textbf{7} Centre de Bioinformatique, Biostatistique et Biologie Int\'egrative, Paris, France
+    \textbf{4} JARA-BRAIN, J\"ulich-Aachen Research Alliance, Germany\\
+    \textbf{5} Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
+    \textbf{6} CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
+    \textbf{7} University Paris Diderot, Sorbonne Paris Cit\'e, Paris, France\\
+    \textbf{8} Centre de Bioinformatique, Biostatistique et Biologie Int\'egrative, Paris, France
 }
@@
 We argue that DMN implication in many
-advanced cognitive processes in humans
+of the most advanced human capacities
 can be recast as prediction error minimization
-based on probabilistic mental simulations
+informed by internally generated probabilistic simulations
 - ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
-thus maximizing action outcomes across multiple time scales.
+allowing maximization of action outcomes across multiple time-scales.
 Such a purposeful optimization objective
@@
   \caption{\textbf{Default mode network: key functions.}
-  Neurobiological overview of the DMN with its major constituent nodes and
+  Neurobiological overview of the DMN with its major constituent parts and
   the associated functional roles relevant in our functional interpretation.}
@@
 These metabolic characteristics go hand-in-hand with
-neuroimaging findings that suggested the PMC and mPFC to
+brain-imaging findings that suggested the PMC and mPFC to
 potentially represent the functional core of the DMN
@@
 % Moreover,
-% in a computational neuroimaging experiment,
+% in a computational brain-imaging experiment,
 % dorsomedial PFC activity preferentially processed the consequences of
@@
 \section{Reinforcement learning control: a process model for DMN function}
-We now argue the outlined neurobiological properties
+We argue the outlined neurobiological properties
 of the DMN regions
@@
 
-RL is an area of machine learning concerned with learning optimal behavioral through interactions
-with an \textit{environment}, the goal being to maximize some notion of \textit{cumulative reward}.
+RL is an area of machine learning concerned with searching optimal behavioral strategies through interactions
+with an \textit{environment} with the goal to maximize some \textit{cumulative reward}.
 The optimal behavior typically
-takes the future into account, as some rewards could be \textit{delayed}.
+takes the future into account as some rewards could be \textit{delayed}.
 Through repeated action on and feedback from the environment,
-the agent learns to reach goals and optimize reward signals
+the agent learns how to reach goals and continuously improve the collection of reward signals
 in a trial-and-error fashion (Fig. \ref{fig:rl}).
@@
 $s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
-$r = r(s, a,s')$ collected by the agent.
+$r = r(s, a,s')$ obtained by the agent.
 If the collected reward outcome yields a negative value it can be
 more naturally interpreted as \textit{punishment}.
-In this view, the environment is partially controlled by
+In this setting, the environment is partially controlled by
 the action of the agent and the reward can be thought
-of as satisfaction \textemdash or aversion \textemdash accompanying the execution of
+of as satisfaction \textemdash or aversion \textemdash that accompany the execution of
 a particular action.
@@
 \textit{partially observable} in the sense that only limited aspects of the environment's
-state are accessible to the agent's sensory perception
+state are accessible to the agent's sensory input
 \citep{starkweather2017dopamine}.
@@
 In short, DMN implication in a diversity of
-particularly advanced cognitive processes
-can be parsimoniously explained as instantiating probabilistic mental
+particularly sophisticated human behaviors
+can be parsimoniously explained as instantiating probabilistic
 simulations of experience coupled with prediction error minimization
 to calibrate action trajectories for
-reward outcome maximization at different time scales.
+reward outcome maximization at different time-scales.
 Such a purposeful optimization objective
@@
 \subsection{Markov decision processes}
-Reinforcement learning (RL) has had considerable success in modeling many real-world problems, including super-human performance in complex video games \citep{mnih2015},
+RL has had considerable success in modeling many real-world problems, including super-human performance in complex video games \citep{mnih2015},
 robotics \citep{ng2004,abbeel2004},
@@
 \begin{itemize}
-\item $\mathcal S$ is the set of states, such as $\mathcal S = \{\text{happy}, \text{sad}\}$.
+\item $\mathcal S$ is the set of states, such as $\mathcal S = \{\text{happy}, \text{sad}, \text{puzzeled}\}$.
 \item $\mathcal A$ is the set of actions, such as $\mathcal A = \{\text{read}, \text{run},
@@
 % , defining $s(t)$, the state at time $t$, as the position in state space.
-The process has $memory$ if the next state depends not only on the current state but also on a number of past states.
+The process has $memory$ if the subsequent state depends not only on the current state but also on a number of past states.
 Rational probabilistic planning can thus be reformulated as a standard memoryless Markov process by simply expanding the definition of the state $s$ to include experience episodes of the past.
-This adds the capacity for memory to the model because the next state then depends not only on the current situation but also on previously experienced events, which is the motivation behind Partially Observable MDPs (POMDPs)
+This extension adds the capacity for memory to the model because the next state then depends not only on the current situation but also on previously experienced events, which is the motivation behind Partially Observable MDPs (POMDPs)
 \citep{starkweather2017dopamine,o2006making}.
@@
 
-\paragraph*{Why MDPs ?} One may wonder whether MDP models are applicable to something as complex as human behavior.
+\paragraph*{Why Markov Decision Processes?} One may wonder whether MDP models are applicable to something as complex as human behavior.
 For instance, financial trading is largely a manifestation of strategic decision-making of interacting human agents.
-According to how the market responds, the agent incurs definite gain or loss as environmental feedback of the executed financial actions.
-Recent research on automatizing market exchanges by algorithmic trading has successfully used MDPs as a framework for modeling these sophisticated behavioral process  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
+According to how the market responds, the agent incurs gain or loss as environmental feedback of the executed financial actions.
+Recent research on automatizing market exchanges by algorithmic trading has successfully used MDPs as a framework for modeling these elaborate behavioral dynamics  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
 %
@@
 %
-As such, we aim to expand MDPs as a useful model from "online" decision-making to the realms of "offline" mental operations most associated with the DMN.
+As such, we aim to expand MDP applications as a useful model from "online" decision-making to the realms of "offline" behaviors most associated with the DMN.
 
@@
 
-\paragraph*{Towards Model-free reinforcement learning for the DMN.}
+\paragraph*{Towards model-free reinforcement learning for the DMN.}
 Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure, daw2014algorithmic}.
@@
 The neurotransmitter dopamine could serve
-as a ``teaching signal'' to better estimate value associations
+as a 'teaching signal' to better estimate value associations
 and action policies by controlling
@@
 % Model-based assumes knowledge of the environment in terms of the transition probabilities between states, the reward contingencies for state-action pairs, or any form of representation of the environment.
-Such knowledge can be partly extracted from the environment: the agent learns a model of the world while learning to take optimal decisions based on this model.
+Such knowledge can be partly extracted from the environment: the agent infers a model of the world while learning to take optimal decisions based on the current representation of the environment.
 These methods learn what the effect is going to be of taking a particular action in a particular state.
@@
 
-In contrast, model-free methods require no particular knowledge of the environment (transition probabilities, types of sensory input, etc.) or representation thereof.
-The agent learns which state-action pairs lead to reward through sampling the world in a trial-and-error manner and infers longer-term reward aggregates using environmental feedback information as an incentive.
+In contrast, model-free methods require no prespecified knowledge of the environment (transition probabilities, types of sensory input, etc.) or representation thereof.
+The agent infers which state-action pairs lead to reward through sampling the world in a trial-and-error manner and derives longer-term reward aggregates using environmental feedback information as an incentive.
 In so doing, model-free agents ultimately learn both an action policy and an implicit representations of the world.
-Our distinction between model-free and model-based RL is hence similar to previous views \citep{dayan2014}.
+This distinction between model-free and model-based RL is similar to previous views \citep{dayan2014}.
 
@@
 A common approach to
-modeling this accumulation is the time-discounted cumulative reward% function
+modeling this gathered outcome is the time-discounted cumulative reward% function
 \begin{equation}
@@
   policy $\pi$ being executed.
-The exponential delay discounting function used here refers to the usual formulation in the field of reinforcement learning, although psychological experiments may also reveal other discounting regimes \citep{green2004discounting}}
+The exponential delay discounting function used here refers to the usual formulation in the field of reinforcement learning, although psychological experiments may also reveal other discounting regimes \citep{green2004discounting}.}
 measures the cumulative reward of
@@
 
-The goal of the RL agent is then to update this action policy in order
+The goal of the RL agent is then to successively update this action policy in order
 to maximize $G^\pi$ on average (cf. below). In \eqref{eq:cumr},
@@
 concerned with immediate rewards.
-This is however not compatible with coordinated planning of long-term goal that is
+This is however not compatible with coordinated planning of longer-term agendas that is
 potentially subserved by neural activity in the DMN.
@@
 Such an agent puts relatively more emphasis on rewards expected in
-a longer-term future.
-More specifically,
+a more distant future.
+Concretely,
 rewards that are not expected to come within
@@
 $\tau := 1/(1 - \gamma)$
-time steps from the present point are disregarded. This reduces the variance of expected rewards accumulated across
+time steps from the present point are ignored.
+The complexity reduction by time discounting
+alleviates the variance of expected rewards accumulated across
 considered action cascades by limiting the depth of the search tree.
-Given that there is more uncertainty in the farsighted future,
+Given that there is more uncertainty in the far future,
 it is important to appreciate that a stochastic policy estimation
@@
 It thus answers the question
-``What is the expected utility of taking action $a$ in this situation ?''.
+``What is the expected utility of choosing action $a$, and its ramifications, in this situation?''.
 $Q^{\pi}(s,a)$ offers a formalization of optimal behavior that
-may well capture processing aspects of the DMN in human agents.
+may well capture such processing aspects subserved by the DMN in human agents.
 
@@
 does not lend itself as a candidate schema
-for solving MDPs with even moderately-sized action and state spaces
+for solving MDPs with even moderately sized action and state spaces
 (i.e., intractability).
@@
 The optimal Q-value operator $Q^*$ is a fixed point for this equation.
-As a consequence of this decomposition, the complicated dynamic programming
+As a consequence of this outcome stratification, the complicated dynamic programming
 problem \eqref{eq:qlearning}
@@
 Relating decomposition of consecutive action choices by the Bellman equation
-to neuroscience,
+to neuroscientific insights,
 specific neural activity in the dorsal prefrontal cortex (BA9)
 was linked to processing ``goal-tree sequences''
-in human neuroimaging experiments
+in human brain-imaging experiments
 \citep{koechlin1999role, koechlin2000dissociating}.
@@
 entail secondary reallocation of attention and working-memory resources.
-Further neuroimaging experiments corroborated
+Further brain-imaging experiments corroborated
 the prefrontal DMN to subserve
@@
 Hence,
-the various advanced human mental abilities subserved by the DMN, such as
+the various advanced human abilities subserved by the DMN, such as
 planning and abstract reasoning, can be viewed to involve some form of
@@
 As already mentioned in the previous section, Q-learning~\citep{watkins92} optimizes over the class of
-deterministic policies of the form \eqref{eq:qlearning}. State spaces may be extremely large,
+deterministic policies of the form \eqref{eq:qlearning}. State spaces may be extremely large
 and tracking all possible states and actions may require prohibitively excessive
@@
 \begin{equation}
-  \pi^{\text{hard-max}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a|\theta),
+  \pi^{\text{hard-max}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a|\theta).
   \label{eq:policy}
@@
 brain regions. It is a time-varying neuronal program which dictates how to move from world states $s$ to actions $a$ via the hard-max policy \eqref{eq:policy}.
-The approximating $Q$-value function $\tilde{Q}(s, a|\theta)$ would tell the DMN the (expected) usefulness of taking an action $a$ in state $s$.
-The DMN, and in particular its dmPFC node, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes these approximate
+The approximating $Q$-value function $\tilde{Q}(s, a|\theta)$ would inform the DMN with the (expected) usefulness of choosing an action $a$ in state $s$.
+The DMN, and in particular its dmPFC part, could then contribute to the choice, at a given state $s$, of an action $a$ which maximizes the approximate
 $Q$-values.
@@
   \label{sec:self}
-Successful learning in brains and computer algorithms is not possible without a
+Successful learning in brains and computer algorithms may not be possible without a
 defined learning goal \textemdash the \textit{loss function}.
@@
 \end{equation}
-with the prediction of a parametrized regression model $(s,a)
+with the estimation of a parametrized regression model $(s,a)
 \mapsto \tilde{Q}(s, a|\theta_{k-1})$.
@@
 \end{equation}
-where $y_k$ is defined in \eqref{eq:bellman}.
+where $y_k$ is obtained from \eqref{eq:bellman}.
 % \textbf{We still may need to explain y in the formula with one sentence.}
-A recently proposed, practically successful alternative approach is, to learn this
-representation using an artificial deep neural-network model, leading to the
+A recently proposed, practically successful alternative approach is to learn this
+representation using an artificial deep neural-network model. This approach leads to the
 so-called \textit{deep Q-learning}~\citep{mnih2015,silver2016mastering} family of methods which
@@
 Some bias may be introduced by self-training due to information shortage caused by the absence of external stimulation.
- One way to address this issue is using importance sampling to replay especially those state-transitions from which there is more to learn \citep{schaul2015, hessel2017}.
+ One way to address this issue is using importance sampling to replay especially those state-transitions from which there is more to learn for the agent \citep{schaul2015, hessel2017}.
 New transitions are inserted into the replay buffer with maximum priority, thus shifting emphasis to more recent transitions.
-Such insertion strategy would help counter-balance the bias introduced by the information shortage incurred by absent external input.
+Such insertion strategy would help counterbalance the bias introduced by the information shortage incurred by absent external input.
 Other authors noticed \citep{hessel2017} that such prioritized replay reduces the data complexity and the agent shows faster increases in learning performance.
@@
 local minima based on the first derivative of the optimization objective.
-Stochastic here means that the true gradient is estimated from batches of training samples,
-which, in our case, corresponds to blocks of experience from the replay memory:
+Stochastic here means that the gradient is estimated from batches of training samples,
+which here corresponds to blocks of experience from the replay memory:
 
@@
 where the positive constants $\alpha_1, \alpha_2,\ldots$ are learning rates.
-Thus, the next action is taken to drive reward prediction errors
+Thus, the subsequent action is taken to drive reward prediction errors
 to percolate from lower to higher processing layers to modulate the
 choice of future actions.
-It is a standard result that under special conditions on the learning rates $\alpha_k$ --namely that the learning rates are neither too large nor too small, or more precisely that the sum $\sum_{k=0}^\infty\alpha_k$ diverges while $\sum_{k=0}^\infty\alpha_k^2$--
+It is known that under special conditions on the learning rates $\alpha_k$ --namely that the learning rates are neither too large nor too small, or more precisely that the sum $\sum_{k=0}^\infty\alpha_k$ diverges while $\sum_{k=0}^\infty\alpha_k^2$--
 the thus generated approximating sequence of $Q$-value functions
@@
 In RL, MCMC simulation is a common means to update the agent's belief state
-by stochastic sampling around states and possible transitions \citep{daw2014algorithmic}.
+based on stochastic sampling around states and possible transitions \citep{daw2014algorithmic}.
 % \citep{silver2010monte}.
 MCMC simulation provides a simple method for evaluating the value of a state.
-They provide an effective mechanism both for tree search (of the considered
+This inference procedure provides an effective mechanism both for tree search (of the considered
 action trajectories)
@@
 These simulations of experience batches
-would be used to update the value function, without ever looking inside the black box describing the model's dynamics \citep{lavilleon2015}.
-A neuroimaging experiment in humans for instance identified hippocampal signals that specifically preceded upcoming choice performance in prospective planning in new environments
+would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
+A brain-imaging experiment in humans for instance identified hippocampal signals that specifically preceded upcoming choice performance in prospective planning in new environments
 \citep{kaplan2017neural}.
@@
 \paragraph*{A note on implicit and explicit memory.}
-While Markov Processes are usually memoryless, it is mathematically feasible to incorporate the previous states of such model into the current state.
+While Markov processes are usually memoryless, it is mathematically feasible to incorporate the previous states of such model into the current state.
 This extension may partially account for implicit memory at the behavioral level, but may not explain the underlying neurobiological implementation or accommodate explicit memory.
 Implicit memory-based processing arises in our MDP account of DMN function in several different forms: successive updates of a) the action policy and the value function, both being products of the past, as well as b) the deep non-linear relationships within the hierarchical connections of biological neural networks.
-The brain's adaptive synaptic connections can be thought as deep neural networks affording an implicit form of information compression of life experience.
-Such memory traces are stored in the neural machinery and can be implicitly retrieve as a form of knowledge during simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}.
-c) Certain neural processes in the hippocampus can be seen as some type of MCMC sampling for memory recall, which can also be a basis for mental imagery across time-scales \citep{schacter2007remembering,Axelrod2017}.
+The brain's adaptive synaptic connections can be thought as a deep neural-network architecture affording an implicit form of information compression of life experience.
+Such memory traces are stored in the neural machinery and can be implicitly retrieved as a form of knowledge during simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}.
+c) Certain neural processes in the hippocampus can be seen as some type of MCMC sampling for memory recall, which can also be a basis for probabilistic simulations across time-scales \citep{schacter2007remembering,Axelrod2017}.
 
@@
 This involvement in automated decision-making has led the authors to propose an ``autopilot'' role for the DMN \citep{vatansever2017pnas}, which may contribute to optimizing control of the organism in general.
-Among all DMN nodes, the RTPJ is perhaps the most evident candidate for a network-switching relay that calibrates between processing of environment-engaged versus internally generated information \citep{downar2000multimodal, golland2006extrinsic, bzdok2013tpj}.
+Among all parts of the DMN, the RTPJ is perhaps the most evident candidate for a network-switching relay that calibrates between processing of environment-engaged versus internally generated information \citep{downar2000multimodal, golland2006extrinsic, bzdok2013tpj}.
 
@@
 experienced events and collected feedback from the past.
-A policy matrix encapsulates the choice probabilities of action repertoire
+A policy matrix encapsulates the choice probabilities of possible actions
 on the world given a current situation (i.e., state).
 The DMN may subserve
-  constant exploration of possible action trajectories and
+  constant exploration of candidate action trajectories and
   nested estimaton of their
@@
   \caption{\textbf{Default mode network:
-  neurobiological implementation of reinforcement learning.}
-  Overview of how the constituent nodes of the DMN (refer to section \ref{sec:nodes}) may
+  possible neurobiological implementation of reinforcement learning.}
+  Overview of how the constituent regions of the DMN (refer to section \ref{sec:nodes}) may
   map onto computational components necessary for an RL agent.}
@@
 
-The HC may contribute to generation of
+The HC may contribute to generating
 perturbed action-transition-state-reward samples
 as batches of pseudo-experience
-(i.e., recalled, hypothesized, and forecasted mental scenarios).
+(i.e., recalled, hypothesized, and forecasted scenarios).
 The small variations in these experience samplings allow searching
@@
   \includegraphics[width=1\linewidth]{VennDiagram-2017-09-08.pdf}
-  \caption{\textbf{Situating Markov Decision Processes among other accounts of default mode function}
-  This Venn diagram summarizes the relationship between four existing explanations of the functional role of the DMN and our present account.
+  \caption{\textbf{Situating Markov Decision Processes among other accounts of default mode function.}
+  The Venn diagram summarizes the relationship between four previously proposed explanations for the functional role of the DMN and our present account.
 Viewing empirical findings in the DMN from the MDP viewpoint incorporates important aspects of the free energy principle, predictive coding, sentinal hypothesis, and semantic hypothesis.
@@
 there is also little reason to assume that contemplating alternative realities of various
-levels of complexity, abstraction, timescale, and purpose
+levels of complexity, abstraction, time-scale, and purpose
 rely on mechanisms that are qualitatively different. This interpretation concurs with
 DMN activity increases across time, space, and content domains
-demonstrated in many neuroimaging studies
+demonstrated in many brain-imaging studies
 \citep{spreng2009common, laird2009, bzdok2012morality, binder2009}.
@@
 of the same neurobiological brain circuit (i.e., DMN) has been documented
-for specific task performances and baseline "house-keeping" functions.
+for specific task performances and baseline 'house-keeping' functions.
 
@@
 formal account of how the human association cortex
-can be thought to implement multi-sensory representation and decision making to
+can be thought to implement multi-sensory representation and high-level decision-making to
 optimize the organism's intervention on the world.
@@
 upcoming events for guiding behavior in the present.
-MDPs also provide a mathematical formalism how
-opportunity in the environment can be recursively exploited
+MDPs also provide a formalism how
+opportunity in the environment can be deconstructed, evaluated, and exploited
 when confronted with challenging interdependent decisions.
@@
 
-A major hurdle in understanding DMN activity from cognitive brain-imaging studies has been its similar neural engagement in different time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., daytime mind-wandering), and thinking about predicted scenarios (e.g., delay discounting).
+A major hurdle in understanding DMN activity from cognitive brain-imaging studies has been its similar neural engagement in different time-scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., daytime mind-wandering), and thinking about anticipated scenarios (e.g., delay discounting).
 The MDP account of DMN function offers a natural integration of a-priori diverging neural processes into a common framework.
@@
 Ultimately,
-neural processes in the DMN may instantiate a holistic integration
+neural processes in the DMN may realize a holistic integration
 ranging from real experience over purposeful dreams to predicted futures

### 2018-01-06 paper/paper.tex
Correct typo p.3/34, 2nd paragraph, "Even (unavoidably) imperfect memory recall, Even necessarily imperfect memory recall, ..."

@@
 \begin{abstract}
-The Default Mode Network (DMN) is believed to subserve the
+The default mode network (DMN) is believed to subserve the
 baseline mental activity in humans.
@@
 We argue that our formal account of DMN function naturally accommodates as special cases
-the previously proposed functional accounts on
+previous interpretations based on
 (1) predictive coding,
 (2) semantic associations, and
-(3) a ``sentinel'' role.
+(3) a sentinel role.
 %
@@
 In the absence of external stimulation, the human brain is not at rest.
-In the beginning of the 21st century,
-brain imaging may have been the first technique
+At the turn to the 21st century,
+brain-imaging may have been the first technique
 to allow for the discovery of a unique brain network that would subserve
@@
 \citep{raichle2001pnas, randy2008, bzdok2015resting}.
-The ``Default Mode Network'' (DMN) continues
+The ``default mode network'' (DMN) continues
 to metabolize large quantities of
@@
 based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling.
-Even (unavoidably) imperfect memory recall,
 Even necessarily imperfect memory recall,

### 2018-03-10 paper/paper.tex
did a huge pass to integrate 2nd round of reviews

@@
 
-\title{Dark Control:\\
-       Towards a Unified Account of Default Mode Function\\
-       by Markov Decision Processes}
+% \title{Dark Control:\\
+%        Towards a Unified Account of Default Mode Function\\
+%        by Markov Decision Processes}
+\title{The Default Mode Network is a Reinforcement Learning Agent!}
 
@@
 expected futures to optimize behavioral performance.
-Formally, a particularly attractive framework
-to describe, quantify, and predict intelligent systems, such as the brain,
-is proposed to be the combination of control theory and
-reinforcement learning (RL). 
+Formally, we propose reinforcement learning to be a particularly attractive framework
+to describe, quantify, DMN function and the brain.
 An intelligent agent improves the interaction with the environment
@@
 may be subserved by a stochastic approximation
-based on a brain implementation of MCMC sampling.
+based on a brain implementation of Monte Carlo sampling.
 
@@
 
-\paragraph*{Why Markov Decision Processes?} One may wonder whether MDP models are applicable to something as complex as human behavior.
+\paragraph*{Why Markov Decision Processes?} One may wonder whether MDP models
+are applicable to something as complex as human behavior. The success of such
+models in diverse domains are numerous.
 For instance, financial trading is largely a manifestation of strategic decision-making of interacting human agents.
@@
 %
-MDPs have also been effective as a behavioral model in robotics \citep{ng2004,abbeel2004} and in challenging multistep strategy games ~\citep{mnih2015,silver2016mastering,pritzel2017neural}.
-%
-As such, we aim to expand MDP applications as a useful model from "online" decision-making to the realms of "offline" behaviors most associated with the DMN.
+MDPs have also been effective as a behavioral model in robotics
+\citep{ng2004,abbeel2004} and in challenging multistep strategy games
+~\citep{mnih2015,silver2016mastering,pritzel2017neural}. An even recent work~\citep{robinowitz2018}
+develops an MDP-based Theory of Mind for learning agents (the \textit{ToM-Net}).
+The idea is to use-meta learning (learning-to-learn) to build strong priors
+about the behavior of of a population of other agents.
+
+% As such, we aim to expand MDP applications as a useful model from "online" decision-making to the realms of "offline" behaviors most associated with the DMN.
 
@@
 
-\paragraph*{Towards model-free reinforcement learning for the DMN.}
+\paragraph*{Reinforcement learning in the brain.}
 Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure, daw2014algorithmic}.
@@
 
-In contrast, model-free methods require no prespecified knowledge of the environment (transition probabilities, types of sensory input, etc.) or representation thereof.
+% In contrast, model-free methods require no pre-specified knowledge of the environment (transition probabilities, types of sensory input, etc.) or representation thereof.
 The agent infers which state-action pairs lead to reward through sampling the world in a trial-and-error manner and derives longer-term reward aggregates using environmental feedback information as an incentive.
@@
 
-\subsubsection{Does the hippocampus subserve MCMC sampling?}
-In RL, MCMC simulation is a common means to update the agent's belief state
-based on stochastic sampling around states and possible transitions \citep{daw2014algorithmic}.
-% \citep{silver2010monte}.
-MCMC simulation provides a simple method for evaluating the value of a state.
+\subsubsection{Does the hippocampus subserve Monte Carlo sampling?}
+In RL, Monte Carlo simulation is a common means to update the agent's belief state
+based on stochastic sampling around states and possible transitions \citep{daw2014algorithmic,silver2010monte}.
+Monte Carlo simulation provides a simple method for evaluating the value of a state.
 This inference procedure provides an effective mechanism both for tree search (of the considered
@@
 highest expected cumulative rewards.
-In MDPs, MCMC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible \citep{silver2016mastering}.
+In MDPs, MC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible \citep{silver2016mastering}.
 This is because expected consequences of action choices can be well evaluated although only a subset of the states are actually considered \citep{daw2014algorithmic}.
@@
 \paragraph*{A note on implicit and explicit memory.}
-While Markov processes are usually memoryless, it is mathematically feasible to incorporate the previous states of such model into the current state.
-This extension may partially account for implicit memory at the behavioral level, but may not explain the underlying neurobiological implementation or accommodate explicit memory.
-Implicit memory-based processing arises in our MDP account of DMN function in several different forms: successive updates of a) the action policy and the value function, both being products of the past, as well as b) the deep non-linear relationships within the hierarchical connections of biological neural networks.
-The brain's adaptive synaptic connections can be thought as a deep neural-network architecture affording an implicit form of information compression of life experience.
-Such memory traces are stored in the neural machinery and can be implicitly retrieved as a form of knowledge during simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}.
-c) Certain neural processes in the hippocampus can be seen as some type of MCMC sampling for memory recall, which can also be a basis for probabilistic simulations across time-scales \citep{schacter2007remembering,Axelrod2017}.
+While Markov processes are usually memoryless, it is mathematically feasible to
+incorporate the previous states of such model into the current state. This
+extension may partially account for implicit memory at the behavioral level, but
+may not explain the underlying neurobiological implementation or accommodate
+explicit memory. Implicit memory-based processing arises in our MDP account of
+DMN function in several different forms: successive updates of a) the action
+policy and the value function, both being products of the past, as well as b)
+the deep non-linear relationships within the hierarchical connections of
+biological neural networks. The brain's adaptive synaptic connections can be
+thought as a deep neural-network architecture affording an implicit form of
+information compression of life experience. Such memory traces are stored in the
+neural machinery and can be implicitly retrieved as a form of knowledge during
+simulation of action rather than accessed as a stored explicit representation
+\citep{pezzulo2011grounding}. c) Certain neural processes in the hippocampus can
+be seen as some type of Monte Carlo sampling for memory recall, which can also
+be a basis for probabilistic simulations across time-scales
+\citep{schacter2007remembering,Axelrod2017}.
 
@@
 stochasticity of the parameter space exploration
-by MCMC approximation achieves better fine-tuning of the
+by Monte Carlo approximation achieves better fine-tuning of the
 action policies and inference of expected reward outcomes.
@@
 % \footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.}
-(Fig. \ref{fig:darkroom}) would remain trapped in this location
+
+% (Fig. \ref{fig:darkroom})
+
+would remain trapped in this location
 because its sensory inputs are perfectly predictable
 given the environmental state \citep{darkroom2012}.
-\begin{wrapfigure}{r}{.37\linewidth}
-  \centering
-  \includegraphics[width=1.\linewidth]{darkroom_resized.png}
-  \caption{\textbf{The dark room experiment.} An intelligent agent situated
-  in a light-deprived closed space is used
-  as a thought experiment for the complete absence
-  of external sensory input.
-  % The free-energy principle theory paradoxally predicts that such an agent would remain quietly in this room, such it perfectly predict all sensory input from such an environment.
-  % However such behavior is not observed in humans or any animals: the agent would strive to leave such an environment in search of more ``interesting'' worlds.
-  }
-  \label{fig:darkroom}
-\end{wrapfigure}
+% \begin{wrapfigure}{r}{.37\linewidth}
+%   \centering
+%   \includegraphics[width=1.\linewidth]{darkroom_resized.png}
+%   \caption{\textbf{The dark room experiment.} An intelligent agent situated
+%   in a light-deprived closed space is used
+%   as a thought experiment for the complete absence
+%   of external sensory input.
+%   % The free-energy principle theory paradoxally predicts that such an agent would remain quietly in this room, such it perfectly predict all sensory input from such an environment.
+%   % However such behavior is not observed in humans or any animals: the agent would strive to leave such an environment in search of more ``interesting'' worlds.
+%   }
+%   \label{fig:darkroom}
+% \end{wrapfigure}
 However, such a behavior is seldom observed in humans in the

### 2018-03-13 paper/paper.tex
last changes for sub2

@@
 %        by Markov Decision Processes}
-\title{The Default Mode Network is a Reinforcement Learning Agent!}
+\title{Dark Control: The Default Mode Network as a Reinforcement Learning Agent}
 
@@
 %
-Specifically, we answer the question whether the neurobiological properties of the DMN
-collectively provide the computational building blocks necessary for a Markov Decision Process.
-%
-We argue that our formal account of DMN function naturally accommodates as special cases
+The main purpose of the DMN, we argue,
+may be to perform optimization based on Markov Decision Processes
+through vicarious trial and error.
+%
+Our formal account of DMN function naturally accommodates as special cases
 previous interpretations based on
@@
 may be solved by a stochastic approximation
-based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling.
+based on a brain implementation of Monte Carlo sampling.
 Even necessarily imperfect memory recall,
@@
 Henceforth, we will use ``inference`` when describing aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
+%
+Notably, our account does not claim that the brain or neural activity in the DMN are completely identical with reinforcement learning algorithms.
+Rather, we advocate feedback-based learning strategies as an attractive alternative perspective
+to understand DMN function.
 
@@
 to be sufficient for implementing all components
-of a full-fledged reinforcement-learning (RL) system.
+of a full-fledged reinforcement learning (RL) system.
 Recalling past experience, considering candidate actions,
@@
 \paragraph*{Why Markov Decision Processes?} One may wonder whether MDP models
-are applicable to something as complex as human behavior. The success of such
-models in diverse domains are numerous.
+are applicable to something as complex as human behavior. This class of
+reinforcement-learning models has had numerous successes in diverse domains.
 For instance, financial trading is largely a manifestation of strategic decision-making of interacting human agents.
 According to how the market responds, the agent incurs gain or loss as environmental feedback of the executed financial actions.
-Recent research on automatizing market exchanges by algorithmic trading has successfully used MDPs as a framework for modeling these elaborate behavioral dynamics  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
+Recent research on automatizing market exchanges by algorithmic trading has effectively deployed MDPs as a framework for modeling these elaborate behavioral dynamics  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
 %
@@
 ~\citep{mnih2015,silver2016mastering,pritzel2017neural}. An even recent work~\citep{robinowitz2018}
-develops an MDP-based Theory of Mind for learning agents (the \textit{ToM-Net}).
-The idea is to use-meta learning (learning-to-learn) to build strong priors
+develops an MDP-based reasoning about future behavior
+of other agents (the \textit{ToM-Net}).
+The idea is to use meta-learning (learning to learn) to build strong priors
 about the behavior of of a population of other agents.
@@
 \paragraph*{Reinforcement learning in the brain.}
-Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure, daw2014algorithmic}.
+RL can be plausibly realized in the human brain \citep{doherty2015structure, daw2014algorithmic}.
 Indeed,
@@
 It has also long been proposed \citep{dayan2008decision} that there
-can be a direct mapping between model-free RL learning algorithms
-and aspects of the brain.
+can be a mapping between algorithmic aspects
+underlying model-free and model-based RL learning
+and neurobiological aspects underlying brain function.
 The neurotransmitter dopamine could serve
 as a 'teaching signal' to better estimate value associations
-and action policies by controlling
+and action policies by modulating
 synaptic plasticity in the reward-processing circuitry, including the NAc.
-In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of the world.
+In contrast, model-based RL would start off with some mechanistic assumptions about the dynamics of the world.
 These can be assumptions about the physical laws governing the agent's environment or constraints on the state space, transition probabilities between states,
@@
 Concretely,
-rewards that are not expected to come within
+rewards that are not expected to occur only within
 % \footnote{\textbf{N.B.:} $\frac{1}{1-\gamma} = 1 + \gamma + \gamma^2 + \gamma^3 + \ldots$}
-$\tau := 1/(1 - \gamma)$
+a very large number of
 time steps from the present point are ignored.
@@
 \citep{kaplan2017neural}.
-This would be a simple control algorithm by evaluating all legal actions and selecting the action with
+It would be a simple strategy to evaluate
+all legal actions and selecting the action with
 highest expected cumulative rewards.
@@
 In line with this perspective, DMN engagement was shown to heighten and relate to effective behavioral responses in the practiced phase of a demanding cognitive flexibility task, as compared to acquisition phase when participants learned context-specific rules.
-This involvement in automated decision-making has led the authors to propose an ``autopilot'' role for the DMN \citep{vatansever2017pnas}, which may contribute to optimizing control of the organism in general.
+This involvement in automated decision-making has led the authors to propose an ``autopilot'' role for the DMN \citep{vatansever2017pnas}, which may contribute to optimizing intervention of the organism on the world in general.
 Among all parts of the DMN, the RTPJ is perhaps the most evident candidate for a network-switching relay that calibrates between processing of environment-engaged versus internally generated information \citep{downar2000multimodal, golland2006extrinsic, bzdok2013tpj}.
@@
 
-From the perspective of a model-free RL agent,
+From the perspective of a RL agent,
 prediction in the DMN reduces to
@@
 
-\item Experiment (Humans): We hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits.
+\item Experiment (Humans): We hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits
+\citep{luksys2009stress, haushofer2014psychology}.
 Using a temporal discounting experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced-back to its corresponding neural representation.
@@
 MDPs have the advantage of exposing an explicit
-mechanisms for controlling the horizon of future considerations and for
+mechanisms for modulating the horizon of future considerations and for
 how the internal representation of the world is updated,
@@
 We finally note that FEP and active inference can be reframed in terms of
-our model-free RL framework.
+our RL framework.
 This becomes possible by recasting the Q-value function (i.e., expected long-term reward) maximized by the DMN to

### 2018-07-30 paper/Cortex_submission/darkcontrol_paper.tex
new Cortex

--- 2018-01-06 paper/paper.tex
@@
 based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling.
+Even (unavoidably) imperfect memory recall,
 Even necessarily imperfect memory recall,

### 2018-08-01 paper/paper.tex
new Cortex
Change RL/POMDP figure
Toned down UK Biobank claims

@@
 We also contribute to the discussion of DMN function
-by providing some of the first empirical evidence that
-morphological variability in DMN regions
+by providing tentative evidence that
+variation of the grey-matter volume in DMN regions
 is linked to the reward circuitry (Fig. 2),
@@
   The code for reproduction and visualization:
-  \url{www.github.com/banilo/darkcontrol_PCB2018}.
+  \url{www.github.com/banilo/darkcontrol_2018}.
   }
@@
   \centering
-  \includegraphics[width=1.\linewidth]{rl2.pdf}
+  \includegraphics[width=1.\linewidth]{POMDP_Fig.pdf}
   \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,

### 2018-08-01 paper/paper.tex
Update title of Figure 3

@@
   \includegraphics[width=1.\linewidth]{POMDP_Fig.pdf}
-  \caption{\textbf{Reinforcement learning in a nutshell.} Given the current state of the environment,
+  \caption{\textbf{Illustration of Partially Observable Markov Decision Process (POMDP).} Given the current state of the environment,
     the agent takes an action by following the policy matrix as updated by the Bellman equation.

### 2018-08-07 paper/paper.tex
Danilo:additions for cortexrevision

@@
 \citep{raichle2006brain} begs the question
-of the biological purpose underlying DMN activity.
-Despite observation of similar large-scale networks of co-varying spontaneous activity
-in electrophysiological investigations \citep{de2010temporal,brookes2011investigating,baker2014fast}, the link between the fMRI BOLD signal and population-level neural activity is still unclear.
+of the biological purpose underlying DMN activity,
+which however still remains elusive at the electrophysiological level
+\citep{de2010temporal,brookes2011investigating,baker2014fast}.
+% the link between the fMRI BOLD signal and population-level neural activity is still unclear.
 If those frequency-specific electrophysiological correlations are proposed as complementary to those observed with BOLD \citep{hipp2015bold}, their role in DMN function remains elusive \citep{maldjian2014graph}.
-
 What has early been described as the ``stream of consciousness''
@@
 \citep{shul1997, raichle2001pnas}.
+%
+Axonal tracing injection in such parts of the association cortex in monkeys
+were shown to resemble connectivity links between nodes of the human
+DMN (see here for details on anatomical connections: \citep{randy2008}).
+Additionally, myelination patterns of axon connections were found to finish
+particularly late in these cortical areas \citep{flechsig1920}, often believed to suggest
+sophistication of subserved neural processes
+\citep{Sowell2003, Yakovlev1967}.
+%
 We propose that this set of some of the most advanced regions in
@@
 We begin by a neurobiological deconstruction of the DMN
-based on experimental findings in the neuroscience literature.
-This walkthrough across main regions of the DMN will
-outline the individual functional profiles,
+based on integrating experimental findings in the neuroscience literature
+from different species.
+This walkthrough across main functional zones of the DMN
+(i.e., de-emphasizing their precise anatomical properties)
+will outline the individual functional profiles with the goal of
 paving the way for their algorithmic interpretation
 in our formal account (section 3).
+As our focus will be on major \textit{functional} zones of the DMN,
+please see elsewhere for excellent surveys on their \textit{anatomical}
+boundaries and connectivity patterns \citep{randy2008, binder2009, seghier2013angular}.
 
@@
   Neurobiological overview of the DMN with its major constituent parts and
-  the associated functional roles relevant in our functional interpretation.}
+  the associated functional roles relevant in our functional interpretation.
+  The blue horizontal line indicates the cytoarchitectonic border between the
+  more dorsal BA9 and the more ventral BA10 (Brodmann, 1909).
+  %
+  Axonal tracing in monkeys and diffusion tractography in humans suggested
+  that the NAc of the reward circuitry has monosynaptic fiber connections
+  to the vmPFC \citep{haber1995, croxson2005quantitative}. Evaluation of
+  propagated value information and triggered affective states encoded
+  in the vmPFC may then feed into the functionally connected partner nodes of
+  the DMN, such as the dmPFC and PMC \citep{andrews2010, bzdok2013segregation}.
+  }
   \label{fig:neurobiological_overview_DMN}
@@
 For instance, B\'alint's syndrome is a neurological disorder of conscious
-awareness that results from medial damage in the parietal cortex
-\citep{balint1909seelenlahmung}.
+awareness that can result from tissue damage in the posterior medial cortex
+\citep{balint1909seelenlahmung, randy2008}.
 Such neurological patients are plagued by an
@@
 (i.e., optic ataxia).
+%
+Scanning complex scenes is impaired in that statistic or moving objects in the
+environment may be invisible or disappear in the subject perception of the
+patient \citep{mesulam_book, blumenfeld_book}.
+%
 This dysfunction can be viewed as a high-level impairment in gathering
 information about alternative objects (i.e., exploration) as well as
-leveraging these environmental opportunities (i.e., exploitation).
+using these environmental opportunities towards a behavioral goal (i.e., exploitation).
 Congruently,
@@
 to map and predict delayed reward opportunities in the future.
-In doing so, the PMC may continuously adapt the organism to changes
+Viewed from a RL perspective, the PMC may continuously adapt the organism to changes
 in both the external environment and its internal representation
@@
 The dmPFC may subserve representation and assessment
-of one's own and other individuals' action considerations.
+of one's own and other individuals' action considerations - a necessary
+component of a full-blown RL agent.
 Generally,
@@
   \begin{minipage}[c]{0.45\textwidth}
-    \caption{\textbf{Morphological coupling between reward system and
-        default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from volume in the DMN regions.
-These out-of-sample generalization performances were obtained from support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure.
+    \caption{\textbf{Predictive structural association between reward system and
+        default mode network.} Reward tasks \citep{doherty2015structure} and neural processign in the DMN \citep{randy2008}, often call task-negative, have been studied so far in largely separate niches of the neuroscience literature.
+A currently underappreciated link is however suggested here based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$ [standard deviation across cross-validation folds]) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from (z-scored) volume in the DMN regions.
+These out-of-sample generalizations reflect the expected performance in yet-to-be observed individuals obtained from linear support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure \citep{hastie}.
       Consistent for the left and right reward system, NAc volume in a given subject is
-      positively coupled with the vmPFC and HC. The colors are indicative of the (\textcolor{red}{red} = positive,
+      positively coupled with the vmPFC and HC.
+The congruence of our structural association results for both NAc targets speaks to the robustness of our pattern-prediction findings.
+The opposite relation of the left and right TPJ to the NAc appears to reflect a repeatedly recognized hemispheric asymmetry with respect to functional implications \citep{seghier2013angular}, impairments in neurological patients \citep{corbetta2000}, different types of connectivity \citep{uddin_tpj, caspers2011} as well as micro- and macroanatomy \citep{caspers2006, caspers2008}.
+The colors are indicative of the (\textcolor{red}{red} = positive,
       \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
@@
 The DMN midline has close functional links
-with the HC in the medial temporal lobe \citep{vincet2006, shannon2013morning} \textemdash
+with the HC \citep{henceforth implying to include also parahippocampal regions}
+in the medial temporal lobe \citep{vincet2006, shannon2013morning} \textemdash
 a region long known to be involved in
@@
 \citep{hassabis2007patients}.
-Thus, the HC would orchestrate re-experience of environmental aspects for
+Within a RL framework, the HC could thus orchestrate re-experience of environmental aspects for
 consolidations based on re-enactment and for integration into
@@
 
-The TPJ in the right hemisphere (RTPJ)
-has been shown to be closely related to
-multi-sensory prediction and prediction error signaling.
+The TPJ in the right hemisphere (RTPJ) denotes a broad functional zone with
+varying anatomical nomenclature \citep{marstpj2011, seghier2010, seghier2013angular}
+that has been shown to be closely related to
+multi-sensory prediction and prediction error signaling
+\citep{downar2000multimodal, vetter2011, shulman2010, shulman2007}.
 This DMN region is probably central for
@@
 a capacity believed to be enabled by the DMN.
-The RTPJ might hence be an important relay that shifts away
+Viewed from a RL account,
+the RTPJ might reflect an important relay that shifts away
 from the ‘‘internally directed’’ baseline processes
@@
 The left TPJ of the DMN (LTPJ),
-in turn, has a close relationship to Wernicke's area
-involved in semantic processes, such as in spoken and written language.
+in turn, may have a functional relationship to Wernicke's area
+involved in semantic processes \citep{blumenfeld_book} and has
+been described as ``a temporoparietal transmodal gateway for language''
+by some investigators \citep{mesulam_book}.
 Neurological patients with damage in Wernicke's area
@@
 %
-Neural processes in the LTPJ may hence
+As a candidate component of a RL agent,
+neural processes in the LTPJ may
 contribute to the automated predictions of the environment
@@
 The more familiar and predictable the current environment, the more brain resources may remain for allocating DMN activity to MDP processes extending beyond the present time and sensory context.
+%
+This speculation receives quantitative support in that connectional links between
+nodes of the DMN have been reported to be more consistent and reliable than
+functional couplings within any other macroscopical networks \citep{shehz2009}.
+As such, random-sampling-related baseline evaluation of possibilities and
+events may be subserved by the DMN and get partly suspended when novelty in
+the external environment is encountered or immediate action is required.
+%
 In line with this perspective, DMN engagement was shown to heighten and relate to effective behavioral responses in the practiced phase of a demanding cognitive flexibility task, as compared to acquisition phase when participants learned context-specific rules.
@@
   Overview of how the constituent regions of the DMN (refer to section \ref{sec:nodes}) may
-  map onto computational components necessary for an RL agent.}
+  map onto computational components necessary for an RL agent.
+  %
+  Axonal tracing in monkeys and diffusion tractography in humans suggested
+  that the NAc of the reward circuitry has monosynaptic fiber connections
+  to the vmPFC \citep{haber1995, croxson2005quantitative}. Evaluation of
+  propagated value information and triggered affective states encoded
+  in the vmPFC may then feed into the functionally connected partner nodes of
+  the DMN, such as the dmPFC and PMC \citep{andrews2010, bzdok2013segregation}.
+}
   \label{fig:rl_process_chart}
@@
 to justify constantly high energy costs?
+While previous experiments on the DMN frequently set out to investigate
+\textit{what} its subserved function may be,
+we have proposed a way of reasoning
+\textit{how} this major network may do what it is doing.
+%
 MDPs motivate an attractive
@@
 can be thought to implement multi-sensory representation and high-level decision-making to
-optimize the organism's intervention on the world.
+optimize the organism's behavioral strategies.
 % by obtaining reward and avoiding pain.
@@
 Viewed as a Markovian sequential decision process,
-human behavior unfolds by inferring cumulative reward outcomes from hypothetical action cascades
+human behavior unfolds by inferring expected reward outcomes
+from hypothetical action cascades
 % stored past events
@@
 %
-This functional interpretation may well be compatible with the DMN's
+This abstract process interpretation may well be compatible with the DMN's
 poorly understood involvement across
@@
 abstract reasoning, social cognition,
-as well as delay discounting and self prospection into the future.
-%
-Improvement of the internal world representation
+as well as delay discounting and self-prospection into the future.
+%
+For instance, improvement of the internal world representation
 by injecting stochasticity into the recall of past
@@
 
-A major hurdle in understanding DMN activity from cognitive brain-imaging studies has been its similar neural engagement in different time-scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., daytime mind-wandering), and thinking about anticipated scenarios (e.g., delay discounting).
-The MDP account of DMN function offers a natural integration of a-priori diverging neural processes into a common framework.
+A major hurdle in guessing DMN function from cognitive brain-imaging studies has been its similar neural engagement in different time-scales: thinking about the past (e.g., autobiographical memory retrieval), thinking about hypothetical presents (e.g., daytime mind-wandering), and
+anticipating scenarios yet to come (e.g., delay discounting).
+The MDP account of DMN activity offers a natural integration of a-priori diverging neural processes into a common framework.
 %
@@
 %
-Neuroscience experiments on the DMN should be designed that operationalize
+Future neuroscience experiments on the DMN should operationalize
 the set of action, value, and state variables governing
@@
 Ultimately,
-neural processes in the DMN may realize a holistic integration
+neural processes in the DMN may realize a brain-wide information integration
 ranging from real experience over purposeful dreams to predicted futures
-to continuously refine the organism's fate.
+to continuously refine the organism's intervention on the world.
 

### 2018-08-07 paper/paper.tex
some style edits

@@
 serve as a concrete implementation of
-predictive coding ideas.
+many predictive coding ideas.
 MDPs have the advantage of exposing an explicit
@@
 items \citep{addis2008age, spreng2006temporal}.
+
 Further,
@@
 That is, the purposeful stochasticity of policy and value updates
-in MDPs provides a candidate explanation for why humans
+in MDPs provides a candidate explanation for why humans may
 have evolved imperfect noisy memories
@@
 In sum, mental scene construction according to the semantic
-account is lacking an explicit time and incentive model,
+account is lacking an explicit time and incentive structure,
 both of which are integral parts of the MDP interpretation of DMN function.
@@
 over relevant cues and events in the environment.
-%
+
 Further,
@@
 is likely to be a human-defining property
-\citep{tomasello2009cultural}.
+\citep{tomasello2009cultural, dunbar2007science}.
 According to the ``social brain hypothesis'',
@@
 and rats \citep{lu2012rat}, three species with
-social capacities that are supposedly less advanced than in humans.
-
-Further,
+social capacities that can be expected to be less advanced than in humans.
+
+Moreover,
 the principal connectivity gradient in the cortex appears to be greatly expanded in humans compared to monkeys, suggesting a phylogenetically conserved axis of cortical expansion with the DMN emerging at the extreme end in humans \citep{margulies_situating_2016}.
@@
 together may play a key role in bridging the gap between self and other by
-integrating low-level embodied processes within higher level inference-based mentalizing \citep{lombardo_shared_2009}.
-
-Rather than functional specificity for processing social information,
+integrating low-level embodied processes within higher level inference-based mentalizing
+\citep{lombardo_shared_2009, social_brain_atlas}.
+
+Rather than functional specificity for processing social information in particular,
 the present MDP account can parsimoniously incorporate

### 2018-08-07 paper/paper.tex
Update Fg 1 & 4 + correct BibTex compilation error

@@
   the associated functional roles relevant in our functional interpretation.
-  The blue horizontal line indicates the cytoarchitectonic border between the
-  more dorsal BA9 and the more ventral BA10 (Brodmann, 1909).
+  The blue horizontal dashed line indicates the cytoarchitectonic border between the
+  more dorsal BA9 and the more ventral BA10 \citep{brodmann1909vergleichende}.
   %
@@
 The DMN midline has close functional links
-with the HC \citep{henceforth implying to include also parahippocampal regions}
+with the HC (henceforth implying to include also parahippocampal regions)
 in the medial temporal lobe \citep{vincet2006, shannon2013morning} \textemdash
@@
   possible neurobiological implementation of reinforcement learning.}
-  Overview of how the constituent regions of the DMN (refer to section \ref{sec:nodes}) may
+  Overview of how the constituent regions of the DMN (refer to section \ref{sec:nodes}; blue horizontal dashed line indicates the border between BA9 and BA10) may
   map onto computational components necessary for an RL agent.

### 2018-08-07 paper/paper.tex
addressed comments of rev. #2

@@
 a human \textit{agent} faced with the choice of the next actions
-and guided by outcomes
-of really happened, hypothetically imagined, and
-expected futures to optimize behavioral performance.
-Formally, we propose reinforcement learning to be a particularly attractive framework
-to describe, quantify, DMN function and the brain.
+and guided by outcomes to optimize behavioral performance. These outcomes
+can be experienced, hypothetically imagined, or simply future expectations.
+Formally, we propose reinforcement learning as a particularly attractive framework
+for describing and quantifying the DMN function.
 An intelligent agent improves the interaction with the environment
@@
 through integration of feedback outcomes.
-That is, ``[agents], with their actions, modify the environment and in doing so partially determine their next stimuli, in particular stimuli that are necessary for triggering the next action''
+That is, ``[agents], with their actions, modify the environment and in doing so
+partially determine their next stimuli, in particular stimuli that are necessary
+for triggering the next action''
 \citep{pezzulo2011grounding}.
@@
     \textbf{choose action: }a_0 &\sim \pi(a|s_0)\\
-    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0, s_1)\\
+    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect
+      reward }R_0 = r(s_0, a_0, s_1)\\
     \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
-    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1,s_2)\\
+    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect
+      reward }R_1 = r(s_1, a_1,s_2)\\
     \vdots\\
     \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
-    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t},s_{t+1})\\
+    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and
+      collect reward }R_{t} = r(s_{t}, a_{t},s_{t+1})\\
     \vdots
@@
 to be equivalent over sufficiently long time windows of equal length (i.e., stationarity).
-Since an action executed in the present moment might have repercussions in the far future, it turns out that the
-quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
-\textit{cumulative reward} estimate which takes into account expected reward
-from action choices in the future.
-A common approach to
+Since an action executed in the present moment might have repercussions in the
+far future, it turns out that the quantity to optimize is not the instantaneous
+rewards $r(s, a)$, but a \textit{cumulative reward} estimate which takes into
+account expected reward from action choices in the future. A common approach to
 modeling this gathered outcome is the time-discounted cumulative reward% function
@@
   \footnote{Random as it depends both on the environment's dynamics and the
-  policy $\pi$ being executed.
-The exponential delay discounting function used here refers to the usual formulation in the field of reinforcement learning, although psychological experiments may also reveal other discounting regimes \citep{green2004discounting}.}
+  policy $\pi$ being executed. The exponential delay discounting function used
+  here refers to the usual formulation in the field of reinforcement learning,
+  although psychological experiments may also reveal other discounting regimes
+  \citep{green2004discounting}.}
 measures the cumulative reward of
@@
 Note that value buffering may be realized in the vmPFC.
-This DMN region has direct connections to
-to the NAc, known to be involved in reward evaluation.
+This DMN region has direct connections to the NAc, known to be involved in
+reward evaluation~\citep{carlezon2009}.
 
@@
 $\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
-At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1} + \delta \theta_{k}$ of the parameters for its approximate model of the $Q$-value function.
-This warrants a learning process
-and definition of a loss function.
-% \subsubsection{Stochastic gradient in the DMN via experience replay}
-The Bellman equation \eqref{eq:bellman} provides a way to obtain such a loss function \eqref{eq:oracle} as we outline in the following.
+At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1}
++ \delta \theta_{k}$ of the parameters for its approximate model of the
+$Q$-value function. This warrants a learning process and definition of a loss
+function. % \subsubsection{Stochastic gradient in the DMN via experience replay}
+The Bellman equation \eqref{eq:bellman} provides a way to obtain such a loss
+function \eqref{eq:oracle} as we outline in the following.
 % A practically successful RL approach
@@
 The agent then tries to approximate
-the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the Bellman equation \eqref{eq:bellman}, namely
+the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the
+Bellman equation \eqref{eq:bellman}, namely
 \begin{equation}
@@
 \mapsto \tilde{Q}(s, a|\theta_{k-1})$.
-From a neurobiological perspective, experience replay can be manifested as the re-occurrence of neuron spiking sequences that have also been measured during specific prior actions and environmental states.
-The HC is a strong candidate for contributing to such neural reinstantiation of behavioral episodes as neuroscience experiments have repeatedly indicated in rats, mice, cats, rabbits, songbirds, and monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}.
+From a neurobiological perspective, experience replay can be manifested as the
+re-occurrence of neuron spiking sequences that have also been measured during
+specific prior actions and environmental states. The HC is a strong candidate
+for contributing to such neural reinstantiation of behavioral episodes as
+neuroscience experiments have repeatedly indicated in rats, mice, cats, rabbits,
+songbirds, and monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}. Refer to
+Fig. \ref{fig:rl_process_chart}.
 
@@
 % \bibliographystyle{natbib}
-\bibliographystyle{plainnat}
+% \bibliographystyle{plainnat}
+\bibliographystyle{abbrvnat}
 \bibliography{refs}

### 2018-08-08 paper/paper.tex
more refs required rev2

@@
 Mental scenes created by neurological patients with HC lesion exposed a lack of
-spatial integrity, richness in detail, and overall coherence.
+spatial integrity, richness in detail, and overall coherence (c.f ~\citep{hassabis2007patients}).
 %

### 2018-08-08 paper/paper.tex
some edits of edits

--- 2018-08-07 paper/paper.tex
@@
 and guided by outcomes to optimize behavioral performance. These outcomes
-can be experienced, hypothetically imagined, or simply future expectations.
+can be really experienced, hypothetically imagined, or simply expected in the future.
 Formally, we propose reinforcement learning as a particularly attractive framework
-for describing and quantifying the DMN function.
+for describing, containing, and quantifying the DMN function.
 An intelligent agent improves the interaction with the environment
@@
 + \delta \theta_{k}$ of the parameters for its approximate model of the
-$Q$-value function. This warrants a learning process and definition of a loss
+$Q$-value function. Step-by-step model parameter updates warrant a learning process and definition of a loss
 function. % \subsubsection{Stochastic gradient in the DMN via experience replay}
@@
 re-occurrence of neuron spiking sequences that have also been measured during
-specific prior actions and environmental states. The HC is a strong candidate
+specific prior actions or environmental states. The HC is a strong candidate
 for contributing to such neural reinstantiation of behavioral episodes as

### 2018-08-08 paper/paper.tex
edits to the stars

@@
 %
-The main purpose of the DMN, we argue,
-may be to perform optimization based on Markov Decision Processes
-through vicarious trial and error.
-%
-Our formal account of DMN function naturally accommodates as special cases
+The main purpose of DMN activity, we argue,
+may be to optimize action policies via value estimates
+based on Markov Decision Processes through vicarious trial and error.
+%
+Our formal perspective on DMN function naturally accommodates as special cases
 previous interpretations based on
@@
 \citep{raichle2006brain} begs the question
-of the biological purpose underlying DMN activity,
+of the biological purpose underlying neural activity in the DMN,
 which however still remains elusive at the electrophysiological level
@@
 The network nodes that compose the human DMN
-are responsible for extended parts of the baseline neural activity,
+are hubs of high baseline neural activity,
 which typically decreases when engaged in controlled psychological experiments
@@
 
-The present work adopts the perspective of
+We aim at proposing an alternative to reasoning about the DMN
+based on longstanding cognitive theory.
+The present work adopts the control-theoretical perspective of
 a human \textit{agent} faced with the choice of the next actions
@@
 Formally, we propose reinforcement learning as a particularly attractive framework
-for describing, containing, and quantifying the DMN function.
+for describing, containing, and quantifying the unknown function underlying DMN activity.
 An intelligent agent improves the interaction with the environment
@@
 - ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
-allowing maximization of action outcomes across multiple time-scales.
+allowing maximization of action outcomes across different time scales.
 Such a purposeful optimization objective
@@
 seemingly arbitrary dreams during sleep
-may provide randomly sampled blocks of pseudo-experience instrumental to iteratively
-optimize the behavior of the organism.
+may provide randomly sampled blocks of pseudo-experience  that are
+instrumental to iteratively
+optimize the behavioral agenda of the organism.
 
 Evidence from computational modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
-the brain is largely dedicated to
+much of brain activity is dedicated to
 ``the development and maintenance of [a]
@@
 thus linking two literatures with currently scarce cross-references.
-Finally, we detail how our process model relates
-to previous accounts of DMN function and
-we derive explicit hypotheses to be tested in future neuroscience
-experiments.
-At this stage, we emphasize the importance of differentiating which levels of observation are at play in the present account.
+Finally,
+we derive explicit hypotheses that could be tested in targeted neuroscience
+experiments in the future
+and
+we detail how our process model relates
+to previous cognitive and theoretical accounts of DMN function.
+
+Please appreciate the importance of differentiating which levels of observation are at play in the present account.
 A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent.
-Henceforth, we will use ``inference`` when describing aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
-%
-Notably, our account does not claim that the brain or neural activity in the DMN are completely identical with reinforcement learning algorithms.
+Henceforth, we will use ``inference`` when referring to aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
+%
+It is moreover important to note that
+our account does not claim that neural activity in the DMN in particular
+or the brain in general are identical with reinforcement learning algorithms.
 Rather, we advocate feedback-based learning strategies as an attractive alternative perspective
-to understand DMN function.
+to describe, quantify, and interpret DMN function in neuroscience research.
 
@@
 is the \textit{free energy}.
-Here, the angular brackets denote the \textit{expectation} of the joint negative log-likelihood $-\log(p_G(\z, \x))$ w.r.t the recognition density $p_R(\z|\x)$, $\mathcal H$ is the \textit{entropy} functional defined by $\mathcal H(p) := -\sum_{\z}p(\z)\log(p(\z)$, while $\operatorname{KL}(.\|.)$ is the usual \textit{Kullback-Leibler (KL) divergence} (also known as \textit{relative entropy}) defined by $\operatorname{KL}(p\|q) := \sum_{z}p(\z)\log(p(\z)/q(\z)) \ge 0$,
-which is a measure of how different two probability distributions are.
-In this framework, the goal of the agent is then to iteratively refine the generative model $p_G$ and the recognition model $p_R$ so as to minimize the free energy $F_G^R(\x)$ over sensory input $\x$.
+Here, the angular brackets denote the \textit{expectation} of the joint negative log-likelihood $-\log(p_G(\z, \x))$ w.r.t the recognition density $p_R(\z|\x)$, $\mathcal H$ is the \textit{entropy} function defined by $\mathcal H(p) := -\sum_{\z}p(\z)\log(p(\z)$, while $\operatorname{KL}(.\|.)$ is the usual \textit{Kullback-Leibler (KL) divergence} (also known as \textit{relative entropy}) defined by $\operatorname{KL}(p\|q) := \sum_{z}p(\z)\log(p(\z)/q(\z)) \ge 0$,
+which is a measure of difference between two probability distributions.
+In this framework, the goal of the agent is to iteratively refine the generative model $p_G$ and the recognition model $p_R$ so as to minimize the free energy $F_G^R(\x)$ over sensory input $\x$.
 
@@
 \begin{itemize}
-\item  $p_R(\z|\x)$ puts a lot of mass on configurations $(\z,\x)$ which are $p_G$-likely, % (i.e the sensation $\x$ is likely caused by $\z$)
-  \textbf{and}
+\item  $p_R(\z|\x)$ puts a lot of mass on configurations $(\z,\x)$ which are $p_G$-likely
+% (i.e the sensation $\x$ is likely caused by $\z$)
+%  \textbf{and}
 \item $p_R(\z|\x)$ is as uniform as possible (i.e have high entropy), so as not to concentrate
-  all its mass on a small subset of possible causes for the sensation $\x$.
+  all its mass on a small subset of possible causes for the sensation $\x$
   \end{itemize}
@@
 Despite its popularity, criticism against the FEP has been voiced over the years,
-some of which is outlined in the following.
+which we allude to in the following.
 The main algorithm for minimizing free energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
@@
 Defenders of the FEP have retorted by advancing the ``full package'' \citep{darkroom2012}:
-FEP is proposed to be multi-scale and there would be a meta-scale at which the organism would
-be \textit{surprised} by such a \textit{lack of surprise}. According to this argument, a dark room would paradoxically correspond to a state of particularly high relevance.
+FEP is framed as being multi-scale, and there would be one scale at which
+the organism would
+be \textit{surprised} by a \textit{lack of surprise}.
+One may object that a dark room would paradoxically correspond to a state of particularly high relevance.
 Driven by the surprise-minimization objective, the FEP agent would eventually bootstrap itself out of such saddle points to explore more interesting parts of the environment.
@@
 identify the sensory-stimulus-deprived room as a local minimum. Indeed, hippocampal experience replay (see \ref{sec:self}) could serve to sample memories or fantasies of alternative situations
-with reward structure. Such artificially generated \textit{internal} sensory input
-subserved by the DMN can entice the organism to explore the room,
+with reward structure. Such artificially generated \textit{internal} sensory input,
+potentially subserved by the DMN, could then entice the organism to explore the room,
 for instance by looking for and using the light switch or finding the room exit.
@@
 that is, the log-likehood of current sensory priors
-the agent has about the world. More explicitly, this corresponds to using free-energy as a Q-value approximator for the MDP such that
+the agent has about the world. More explicitly, this corresponds to using free-energy as a Q-value approximator for the MDP in the following way:
 

### 2018-08-08 paper/paper.tex
sentence with 'bootstrap' in section 4 negated

@@
 One may object that a dark room would paradoxically correspond to a state of particularly high relevance.
-Driven by the surprise-minimization objective, the FEP agent would eventually bootstrap itself out of such saddle points to explore more interesting parts of the environment.
+Driven by the surprise-minimization objective, the FEP agent would eventually not bootstrap itself out of such saddle points to explore more interesting parts of the environment.
 % Recourse to this argument is equivalent to invoking a \textit{halting oracle}\citep{haltingoracle}, an arbitrarily powerful machine which can solve the \textit{halting problem} (an arbitrarily difficult computational problem which cannot be solved under any reasonable \textit{model of computation}, not even in principle!).

### 2018-08-08 paper/paper.tex
Unboxing...

@@
 triggered by another (possibly stochastic) behavioral policy.
-
-\begin{mdframed}
-% neuroscience of Bellman equation
-  \vspace{2em}
-  \textbf{Box 1: Neural correlates of the Bellman equation in the DMN}
-% \hspace{-1.5em}
-Relating decomposition of consecutive action choices by the Bellman equation
-to neuroscientific insights,
-specific neural activity in the dorsal prefrontal cortex (BA9)
-was linked to processing ``goal-tree sequences''
-in human brain-imaging experiments
-\citep{koechlin1999role, koechlin2000dissociating}.
-Sub-goal exploration may require multi-task switching
-between cognitive processes
-as later parts of a solution
-frequently depend on respective earlier steps in a given solution path, which
-necessitates storage of expected intermediate outcomes.
-As such,
-``cognitive branching'' operations for nested processing of behavioral strategies
-are likely to
-entail secondary reallocation of attention and working-memory resources.
-Further brain-imaging experiments corroborated
-the prefrontal DMN to subserve
-``processes related to the management and monitoring of sub-goals while
-maintaining information in working memory''
-\citep{braver2002role}
-and to functionally couple with the hippocampus conditioned by ``deep versus shallow planning''
-\citep{kaplan2017neural}.
-%
-Moreover,
-neurological patients with lesions in this DMN region were reported
-to be impaired in aspects of realizing ``multiple sub-goal scheduling''
-\citep{burgess2000cognitive}.
-Hence,
-the various advanced human abilities subserved by the DMN, such as
-planning and abstract reasoning, can be viewed to involve some form of
-action-decision branching to enable higher-order executive control.
-\end{mdframed}
 
@@
 \subsection{Putting everything together}
+\subsubsection{Consequences for the DMN function}
 The DMN is today known to consistently increase in neural
@@
 
-\begin{mdframed}
-  \vspace{1em}
-  \textbf{Box 2: Proposed studies for testing the MDP account of DMN function}
-\begin{enumerate}
-\item Experiment (Humans): We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
-During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigm.
-We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, the more the fMRI signals in the reward circuits should be independent of the reward context in the current sensory environment.
-
-\item Experiment (Humans): We hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
+% \begin{mdframed}
+%   \vspace{1em}
+%   \textbf{Box 2: Proposed studies for testing the MDP account of DMN function}
+% \begin{enumerate}
+
+\subsubsection{Experimental predictions}
+Following our process model of DMN, several empirical predictions can be drawn.
+
+\textbf{Neural correlates of the Bellman equation in the DMN (Humans)}
+Relating decomposition of consecutive action choices by the Bellman equation
+to neuroscientific insights,
+specific neural activity in the dorsal prefrontal cortex (BA9)
+was linked to processing ``goal-tree sequences''
+in human brain-imaging experiments
+\citep{koechlin1999role, koechlin2000dissociating}.
+Sub-goal exploration may require multi-task switching
+between cognitive processes
+as later parts of a solution
+frequently depend on respective earlier steps in a given solution path, which
+necessitates storage of expected intermediate outcomes.
+As such,
+``cognitive branching'' operations for nested processing of behavioral strategies
+are likely to
+entail secondary reallocation of attention and working-memory resources.
+Further brain-imaging experiments corroborated
+the prefrontal DMN to subserve
+``processes related to the management and monitoring of sub-goals while
+maintaining information in working memory''
+\citep{braver2002role}
+and to functionally couple with the hippocampus conditioned by ``deep versus shallow planning''
+\citep{kaplan2017neural}.
+%
+Moreover,
+neurological patients with lesions in this DMN region were reported
+to be impaired in aspects of realizing ``multiple sub-goal scheduling''
+\citep{burgess2000cognitive}.
+Hence,
+the various advanced human abilities subserved by the DMN, such as
+planning and abstract reasoning, can be viewed to involve some form of
+action-decision branching to enable higher-order executive control.
+
+We hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
 First, we expect that fMRI signals in the right temporo-parietal junction relate to behavioral changes subsequent to adaptation in the action choice tendencies (policy matrix) involved in non-value-related prediction error.
@@
 
-\item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is contributing to choice behavior in mice.
-Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future.
-Necessity of cortical DMN regions, in addition to the hippocampus, for “mind-searching” candidate actions during choice behavior can be experimentally corroborated by causal disruption of DMN regions, such as by circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
-
-\item Experiment (Humans): We hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits
+\textbf{Continuous learning paradigm (Humans)} We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
+During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigm.
+We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, the more the fMRI signals in the reward circuits should be independent of the reward context in the current sensory environment.
+
+\textbf{Temporal discounting experiment (Humans)} We hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits
 \citep{luksys2009stress, haushofer2014psychology}.
@@
 Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age and personality traits.
-
 % To exam-ine this question, Fellows and Farah (2005) investigated the effects of dorsolateral frontal (DLF) and ventromedial frontal (VMF) lobe damage on two aspects of temporal foresight—namely, temporal discounting and the futuretime perspective (measuredby an adaptation of Wallace’s [1956] list of future life events).
 
-\item Experiment (Humans \& Animals): An additional layer of learning concerns the addition of new entries in the state and action spaces.
+\textbf{Synaptic epigenesis and creation of new entry in policy matrix (Humans \& Animals)} An additional layer of learning concerns the addition of new entries in the state and action spaces.
 Extension of the action repertoire could be biologically realized by synaptic epigenesis \citep{gisiger_acquisition_2005}.
@@
 Those attractors can then constrain both the number of factors taken into account by decision processes and the possible behaviors of the agent \citep{wang_decision_2008}.
-To examine this potential higher-level mechanism, we propose to probe how synaptic epigenesis is related to neural correlates underlying policy matrix updates: in humans the changes of functional connectivity between DMN regions can be investigated following a temporal discounting experiment and in monkeys or rodents anterograde tracing can be used to study how homolog regions of the DMN present increased synaptic changes compare to other parts of the brain.
-
-\end{enumerate}
-\end{mdframed}
+To examine this potential higher-level mechanism, we propose to probe how synaptic epigenesis is related to neural correlates underlying policy matrix updates: in humans the changes of functional connectivity between DMN regions can be investigated following a temporal discounting experiment and in monkeys or rodents anterograde tracing can be used to study how homolog regions of the DMN present increased synapti\c changes compare to other parts of the brain.
+
+\textbf{Interplay with memory and “mind-searching” (Animals)} We hypothesize that experience replay for browsing problem solutions subserved by the DMN is contributing to choice behavior in mice.
+Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future.
+Necessity of cortical DMN regions, in addition to the hippocampus, for “mind-searching” candidate actions during choice behavior can be experimentally corroborated by causal disruption of DMN regions, such as by circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
+
+% \end{enumerate}
+% \end{mdframed}
 

### 2018-08-08 paper/paper.tex
Update Figure 2 size & style

@@
 \begin{figure}[!htbp]
-  \begin{minipage}[c]{0.63\textwidth}
-    \includegraphics[width=.8\linewidth]{fig_dmn.png}
-  \end{minipage}
-  \hspace{-4em}
-  \begin{minipage}[c]{0.45\textwidth}
+
+  \includegraphics[width=.75\linewidth]{fig_dmn.png}
   \caption{\textbf{Predictive structural association between reward system and
-        default mode network.} Reward tasks \citep{doherty2015structure} and neural processign in the DMN \citep{randy2008}, often call task-negative, have been studied so far in largely separate niches of the neuroscience literature.
+      DMN.} Reward tasks \citep{doherty2015structure} and neural processign in the DMN \citep{randy2008}, often call task-negative, have been studied so far in largely separate niches of the neuroscience literature.
 A currently underappreciated link is however suggested here based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$ [standard deviation across cross-validation folds]) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from (z-scored) volume in the DMN regions.
@@
     \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
-  The code for reproduction and visualization:
-  \url{www.github.com/banilo/darkcontrol_2018}.
-  }
+  The code for reproduction and visualization: \url{www.github.com/banilo/darkcontrol_2018}.}
   \label{fig:dmn}
-\end{minipage}
 % \end{wrapfigure}

### 2018-08-08 paper/paper.tex
Unboxing with update of Section 3.3

@@
 \subsection{Putting everything together}
-\subsubsection{Consequences for the DMN function}
+
 The DMN is today known to consistently increase in neural
@@
   in humans.
+
+\paragraph*{Neural correlates of the Bellman equation in the DMN (Humans)}
+Relating decomposition of consecutive action choices by the Bellman equation
+to neuroscientific insights,
+specific neural activity in the dorsal prefrontal cortex (BA9)
+was linked to processing ``goal-tree sequences''
+in human brain-imaging experiments
+\citep{koechlin1999role, koechlin2000dissociating}.
+Sub-goal exploration may require multi-task switching
+between cognitive processes
+as later parts of a solution
+frequently depend on respective earlier steps in a given solution path, which
+necessitates storage of expected intermediate outcomes.
+As such,
+``cognitive branching'' operations for nested processing of behavioral strategies
+are likely to
+entail secondary reallocation of attention and working-memory resources.
+Further brain-imaging experiments corroborated
+the prefrontal DMN to subserve
+``processes related to the management and monitoring of sub-goals while
+maintaining information in working memory''
+\citep{braver2002role}
+and to functionally couple with the hippocampus conditioned by ``deep versus shallow planning''
+\citep{kaplan2017neural}.
+%
+Moreover,
+neurological patients with lesions in this DMN region were reported
+to be impaired in aspects of realizing ``multiple sub-goal scheduling''
+\citep{burgess2000cognitive}.
+Hence,
+the various advanced human abilities subserved by the DMN, such as
+planning and abstract reasoning, can be viewed to involve some form of
+action-decision branching to enable higher-order executive control.
+
+We hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
+First, we expect that fMRI signals in the right temporo-parietal junction relate to behavioral changes subsequent to adaptation in the action choice tendencies (policy matrix) involved in non-value-related prediction error.
+Second, fMRI signals in the ventromedial prefrontal cortex should relate to behavioral changes following adaptation in value estimation (value matrix) due to reward-related stimulus-value association.
+We finally expect that fMRI signals in the posteromedial cortex, as a potential global information integrator, are related to shifts in overt behavior based on previous adaptations in both policy or value estimation.
+
+\paragraph*{Continuous learning paradigm (Humans)} We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
+During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigm.
+We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, the more the fMRI signals in the reward circuits should be independent of the reward context in the current sensory environment.
+
+\paragraph*{Temporal discounting experiment (Humans)} We hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits
+\citep{luksys2009stress, haushofer2014psychology}.
+Using a temporal discounting experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced-back to its corresponding neural representation.
+Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age and personality traits.
+% To exam-ine this question, Fellows and Farah (2005) investigated the effects of dorsolateral frontal (DLF) and ventromedial frontal (VMF) lobe damage on two aspects of temporal foresight—namely, temporal discounting and the futuretime perspective (measuredby an adaptation of Wallace’s [1956] list of future life events).
+
 \begin{figure}[!h]
@@
 
+\paragraph*{Interplay with memory and “mind-searching” (Animals)} We hypothesize that experience replay for browsing problem solutions subserved by the DMN is contributing to choice behavior in mice.
+Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future.
+Necessity of cortical DMN regions, in addition to the hippocampus, for “mind-searching” candidate actions during choice behavior can be experimentally corroborated by causal disruption of DMN regions, such as by circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
+
 From the perspective of a RL agent,
@@
 
+\paragraph*{Synaptic epigenesis and creation of new entry in policy matrix (Humans \& Animals)} An additional layer of learning concerns the addition of new entries in the state and action spaces.
+Extension of the action repertoire could be biologically realized by synaptic epigenesis \citep{gisiger_acquisition_2005}.
+Indeed, the tuning of synaptic weights through learning can stabilize additional patterns of activity by creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}.
+Those attractors can then constrain both the number of factors taken into account by decision processes and the possible behaviors of the agent \citep{wang_decision_2008}.
+To examine this potential higher-level mechanism, we propose to probe how synaptic epigenesis is related to neural correlates underlying policy matrix updates: in humans the changes of functional connectivity between DMN regions can be investigated following a temporal discounting experiment and in monkeys or rodents anterograde tracing can be used to study how homolog regions of the DMN present increased synapti\c changes compare to other parts of the brain.
+
 % \paragraph*{Some notes:}
@@
 % \begin{enumerate}
-
-\subsubsection{Experimental predictions}
-Following our process model of DMN, several empirical predictions can be drawn.
-
-\textbf{Neural correlates of the Bellman equation in the DMN (Humans)}
-Relating decomposition of consecutive action choices by the Bellman equation
-to neuroscientific insights,
-specific neural activity in the dorsal prefrontal cortex (BA9)
-was linked to processing ``goal-tree sequences''
-in human brain-imaging experiments
-\citep{koechlin1999role, koechlin2000dissociating}.
-Sub-goal exploration may require multi-task switching
-between cognitive processes
-as later parts of a solution
-frequently depend on respective earlier steps in a given solution path, which
-necessitates storage of expected intermediate outcomes.
-As such,
-``cognitive branching'' operations for nested processing of behavioral strategies
-are likely to
-entail secondary reallocation of attention and working-memory resources.
-Further brain-imaging experiments corroborated
-the prefrontal DMN to subserve
-``processes related to the management and monitoring of sub-goals while
-maintaining information in working memory''
-\citep{braver2002role}
-and to functionally couple with the hippocampus conditioned by ``deep versus shallow planning''
-\citep{kaplan2017neural}.
-%
-Moreover,
-neurological patients with lesions in this DMN region were reported
-to be impaired in aspects of realizing ``multiple sub-goal scheduling''
-\citep{burgess2000cognitive}.
-Hence,
-the various advanced human abilities subserved by the DMN, such as
-planning and abstract reasoning, can be viewed to involve some form of
-action-decision branching to enable higher-order executive control.
-
-We hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
-First, we expect that fMRI signals in the right temporo-parietal junction relate to behavioral changes subsequent to adaptation in the action choice tendencies (policy matrix) involved in non-value-related prediction error.
-Second, fMRI signals in the ventromedial prefrontal cortex should relate to behavioral changes following adaptation in value estimation (value matrix) due to reward-related stimulus-value association.
-We finally expect that fMRI signals in the posteromedial cortex, as a potential global information integrator, are related to shifts in overt behavior based on previous adaptations in both policy or value estimation.
-
-\textbf{Continuous learning paradigm (Humans)} We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
-During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigm.
-We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, the more the fMRI signals in the reward circuits should be independent of the reward context in the current sensory environment.
-
-\textbf{Temporal discounting experiment (Humans)} We hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits
-\citep{luksys2009stress, haushofer2014psychology}.
-Using a temporal discounting experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced-back to its corresponding neural representation.
-Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age and personality traits.
-% To exam-ine this question, Fellows and Farah (2005) investigated the effects of dorsolateral frontal (DLF) and ventromedial frontal (VMF) lobe damage on two aspects of temporal foresight—namely, temporal discounting and the futuretime perspective (measuredby an adaptation of Wallace’s [1956] list of future life events).
-
-\textbf{Synaptic epigenesis and creation of new entry in policy matrix (Humans \& Animals)} An additional layer of learning concerns the addition of new entries in the state and action spaces.
-Extension of the action repertoire could be biologically realized by synaptic epigenesis \citep{gisiger_acquisition_2005}.
-Indeed, the tuning of synaptic weights through learning can stabilize additional patterns of activity by creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}.
-Those attractors can then constrain both the number of factors taken into account by decision processes and the possible behaviors of the agent \citep{wang_decision_2008}.
-To examine this potential higher-level mechanism, we propose to probe how synaptic epigenesis is related to neural correlates underlying policy matrix updates: in humans the changes of functional connectivity between DMN regions can be investigated following a temporal discounting experiment and in monkeys or rodents anterograde tracing can be used to study how homolog regions of the DMN present increased synapti\c changes compare to other parts of the brain.
-
-\textbf{Interplay with memory and “mind-searching” (Animals)} We hypothesize that experience replay for browsing problem solutions subserved by the DMN is contributing to choice behavior in mice.
-Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future.
-Necessity of cortical DMN regions, in addition to the hippocampus, for “mind-searching” candidate actions during choice behavior can be experimentally corroborated by causal disruption of DMN regions, such as by circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
-
-% \end{enumerate}
-% \end{mdframed}
 

### 2018-08-09 paper/paper.tex
interm

@@
 %
-Its highest energy consumption compared to other brain networks and
+Its higher energy consumption compared to other brain networks and
 its intimate coupling with conscious awareness are both pointing to
-an overarching function.
+an unknown overarching function.
 %
@@
 The main purpose of DMN activity, we argue,
-may be to optimize action policies via value estimates
-based on Markov Decision Processes through vicarious trial and error.
+may be described by Markov Decision Processes
+that optimize action policies via value estimates
+based through vicarious trial and error.
 %
@@
 
-Psychologically,
+At the psychological level,
 the more the ongoing executed task is unknown and unpracticed,
@@
 these ``offline'' processes may contribute
-to optimizing control of the organism.
+to optimizing control of the organism in general.
 We formalize
@@
 on the environment given the current state. A \textit{value function}
-maps environmental objects and events (i.e., states) to expected rewards.
+maps environmental objects and events (i.e., states) to expected reward outcomes.
 Switching between states reduces to a sequential processing model.
@@
 
-Such an RL account of DMN function can naturally embed human behavior
+Such a RL perspective on DMN function can naturally embed human behavior
 into the tension between exploitative action with immediate gains and
@@
 Rather, we advocate feedback-based learning strategies as an attractive alternative perspective
-to describe, quantify, and interpret DMN function in neuroscience research.
+to describe, quantify, and interpret research findings on the DMN.
 
@@
 human PMC have been closely related to
-changes of conscious awareness \citep{cavanna2006precuneus}.
+changes of conscious awareness \citep{cavanna2006precuneus, leech_pcc2014}.
 Indeed,
@@
 Analogous to the PMC,
-the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
+the dorsomedial PFC (dmPFC, related to BA9) of the DMN is believed to subserve
 multi-sensory processes
@@
 dmPFC function may be closer to a
-``mental sketchpad'' \citep{goldman1996prefrontal}, as this DMN part
+``mental sketchpad'' \citep{goldman1996prefrontal}. This DMN part
 potentially subserves the de-novo construction and manipulation
@@
 Comparing to the dmPFC,
-the vmPFC is probably more specifically devoted to
+the ventromedial PFC (vmPFC, related to BA10) is probably more specifically devoted to
 subjective value evaluation and
@@
   \caption{\textbf{Predictive structural association between reward system and
-      DMN.} Reward tasks \citep{doherty2015structure} and neural processign in the DMN \citep{randy2008}, often call task-negative, have been studied so far in largely separate niches of the neuroscience literature.
+      DMN nodes.} Reward tasks \citep{doherty2015structure} and neural processing in the DMN \citep{randy2008}, often considered task-negative, have been studied so far in largely separate niches of the neuroscience literature.
 A currently underappreciated link is however suggested here based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$ [standard deviation across cross-validation folds]) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from (z-scored) volume in the DMN regions.
@@
 value predictions
-\citep{behrens2008associative},
-which may also explain why performance within and across participants
+\citep{behrens2008associative}. This finding
+may also explain why performance within and across participants
 was reported to relate to state encoding in the vmPFC \citep{Schuck2016}.
-Such a ``cognitive map'' of the action space was argued to encode
+Such a ``cognitive map'' of the action space - an integral part of a RL
+agent -
+was argued to encode
 the current task state even when states are unobservable from
@@
 rich mental scene construction \citep{deuker2016event, bird2010establishing}.
-As such, the HC may impact
+In this way, the HC may impact
 ongoing perception of and action on the environment
@@
 may be subserved by a stochastic approximation
-based on a brain implementation of Monte Carlo sampling.
+based on a brain implementation for Monte Carlo sampling of events and outcomes.
 
@@
 action trajectories)
-and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than an RL agent without stochastic resampling procedures.
+and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than a RL agent without stochastic resampling procedures.
 Such methods have scaling as a function of available data (i.e., sample complexity) that
@@
   Overview of how the constituent regions of the DMN (refer to section \ref{sec:nodes}; blue horizontal dashed line indicates the border between BA9 and BA10) may
-  map onto computational components necessary for an RL agent.
+  map onto computational components necessary for a RL agent.
   %

### 2018-08-09 paper/paper.tex
interm2

@@
 \citep{doria2010}, while
-the right and left TPJs may become fully integrated into this major brain
+the right and left TPJs may become fully functionally integrated into this major brain
 network only after birth.
@@
 \citep{seghier2013angular}.
-Globally, neuroscientific investigations on hemispheric functional specialization
-have highlighted the right cerebral hemisphere as dominant for attentional functions
-and the left side for semantic functions
+In general, neuroscientific investigations on hemispheric functional specialization
+have highlighted the right cerebral hemisphere as more dominant for attentional functions
+and the left side more for semantic functions
 \citep{seghier2013angular, bzdok2013tpj, bzdok2016left,
@@
 multi-step action execution \citep{hartmann2005takes},
-visuo-proprioceptive conflict \citep{Balslev2005}, and
+visuo-proprioceptive conflict \citep{Balslev2005},
+spatial reorienting \citep{corbetta2000},
+and
 detection of environmental changes across
@@
 jakobs2009effects}.
-Additionally, neural activity in the RTPJ has been proposed to reflect
+More specifically, neural activity in the RTPJ has been proposed to reflect
 stimulus-driven attentional reallocation to
@@
 % (Bzdok et al., 2013; Corbetta et al., 2008).
-Indeed, neurological patients with RTPJ damage have particular difficulties
-with multi-step actions \citep{hartmann2005takes}.
+% Indeed, neurological patients with RTPJ damage have particular difficulties
+% with multi-step actions \citep{hartmann2005takes}.
 In the face of large discrepancies between actual and previously predicted
@@
 by some investigators \citep{mesulam_book}.
-Neurological patients with damage in Wernicke's area
+Neurological patients with damage in this region
 have a major impairment of language comprehension
@@
 to calibrate action trajectories for
-reward outcome maximization at different time-scales.
+reward outcome maximization at different time scales.
 Such a purposeful optimization objective
@@
   \includegraphics[width=1.\linewidth]{POMDP_Fig.pdf}
-  \caption{\textbf{Illustration of Partially Observable Markov Decision Process (POMDP).} Given the current state of the environment,
-    the agent takes an action by following the policy matrix as updated by the Bellman equation.
+  \caption{\textbf{Illustration of a Partially Observable Markov Decision Process (POMDP).} Given the current state of the environment,
+    the agent takes an action by following the policy matrix, which is iteratively updated by the Bellman equation.
 The agent receives a triggered reward and observes the next state.
@@
 \citep{ng2004,abbeel2004} and in challenging multistep strategy games
-~\citep{mnih2015,silver2016mastering,pritzel2017neural}. An even recent work~\citep{robinowitz2018}
-develops an MDP-based reasoning about future behavior
-of other agents (the \textit{ToM-Net}).
+~\citep{mnih2015,silver2016mastering,pritzel2017neural}. 
+More recent work has
+developped an MDP-based way of reasoning about future behavior
+of other agents \citep{robinowitz2018}.
 The idea is to use meta-learning (learning to learn) to build strong priors
@@
 Indeed,
-it has been proposed \citep{gershman2015computational} that a core property of human intelligence is the improvement
+previous authors have proposed
+\citep{gershman2015computational} that a core property of human intelligence is the improvement
 of expected utility outcomes as a strategy for action choice in uncertain
@@
 \end{eqnarray*}
-We assume time-invariance
+We assume time invariance
 in that we expect the dynamics of the process
@@
 %
-Note that value buffering may be realized in the vmPFC.
-This DMN region has direct connections to the NAc, known to be involved in
-reward evaluation~\citep{carlezon2009}.
+Note that value buffering may be realized in the vmPFC by virtue of
+this region's connections to the NAc of the reward sytem
+\citep{carlezon2009}.
 
@@
 be seen as some type of Monte Carlo sampling for memory recall, which can also
-be a basis for probabilistic simulations across time-scales
+be a basis for probabilistic simulations across time scales
 \citep{schacter2007remembering,Axelrod2017}.
@@
 there is also little reason to assume that contemplating alternative realities of various
-levels of complexity, abstraction, time-scale, and purpose
+levels of complexity, abstraction, time scale, and purpose
 rely on mechanisms that are qualitatively different. This interpretation concurs with
@@
 
-A major hurdle in guessing DMN function from cognitive brain-imaging studies has been its similar neural engagement in different time-scales: thinking about the past (e.g., autobiographical memory retrieval), thinking about hypothetical presents (e.g., daytime mind-wandering), and
+A major hurdle in guessing DMN function from cognitive brain-imaging studies has been its similar neural engagement in different time scales: thinking about the past (e.g., autobiographical memory retrieval), thinking about hypothetical presents (e.g., daytime mind-wandering), and
 anticipating scenarios yet to come (e.g., delay discounting).

### 2018-08-10 paper/paper.tex
interm3

@@
 RL is an area of machine learning concerned with searching optimal behavioral strategies through interactions
-with an \textit{environment} with the goal to maximize some \textit{cumulative reward}.
+with an \textit{environment} with the goal to maximize the \textit{cumulative reward} over time.
 The optimal behavior typically
-takes the future into account as some rewards could be \textit{delayed}.
+takes the future into account as certain rewards could be \textit{delayed}.
 Through repeated action on and feedback from the environment,
@@
 more naturally interpreted as \textit{punishment}.
-In this setting, the environment is partially controlled by
+The environment be partly controlled by
 the action of the agent and the reward can be thought
@@
 % \citep{dayan2008decision}.
-We argue that a functional account of the DMN based on RL
+We argue that an abstract description of DMN activity based on RL
 can naturally embed human behavior
@@
 \subsection{Markov decision processes}
-RL has had considerable success in modeling many real-world problems, including super-human performance in complex video games \citep{mnih2015},
-robotics \citep{ng2004,abbeel2004},
-and strategic board games like the breakthrough results upon recently on the game of Go \citep{silver2016mastering} considered to be a milestone benchmark in artificial intelligence.
+% RL has had considerable success in modeling complex decision-making problems,
 % We emphasize that the brain in general, and the DMN in particular,
@@
 In artificial intelligence and machine learning, a popular computational model for
-multi-step decision processes in such an environment are
+multi-step decision processes are
 MDPs \citep{sutton1998reinforcement}.
@@
 Instead, the agent tries to optimize a \textit{subjective} reward
-signal (i.e., likely to be different for another agent in the same state),
+signal (i.e., likely to be different for another agent in the same state
+and possibly driven by neural processing in the vmPFC)
 by maintaining probability distributions over actions according to their expected utility.
@@
 of other agents \citep{robinowitz2018}.
-The idea is to use meta-learning (learning to learn) to build strong priors
-about the behavior of of a population of other agents.
+The idea is to use meta-learning (i.e., learning to learn) to build strong priors
+about the behavior of a population of other agents.
 
@@
 \paragraph*{Reinforcement learning in the brain.}
-RL can be plausibly realized in the human brain \citep{doherty2015structure, daw2014algorithmic}.
+RL has been argued to be a biologically plausible mechanism in the human brain
+\citep{doherty2015structure, daw2014algorithmic}.
 Indeed,
@@
 of expected utility outcomes as a strategy for action choice in uncertain
-environments, a situation perfectly captured by the formalism of MDPs.
+environments, a situation captured by the formalism of MDPs.
 It has also long been proposed \citep{dayan2008decision} that there
@@
 The neurotransmitter dopamine could serve
-as a 'teaching signal' to better estimate value associations
+as a 'teaching signal' to guide estimation of value associations
 and action policies by modulating
@@
 In contrast, model-based RL would start off with some mechanistic assumptions about the dynamics of the world.
-These can be assumptions about the physical laws governing the agent's environment or constraints on the state space, transition probabilities between states,
-reward contingencies, etc.
+These assumptions could relate to the
+physical laws governing the agent's environment, constraints on the state space, transition probabilities between states, or reward contingencies.
 % For instance, if a rat in a maze knows
@@
 This random variable
-  \footnote{Random as it depends both on the environment's dynamics and the
+  \footnote{The reward outcome is random as it depends both on the environment's dynamics and the
   policy $\pi$ being executed. The exponential delay discounting function used
@@
 utility of executing a specific action.
-It thus answers the question
+In this way, we can answer the question
 ``What is the expected utility of choosing action $a$, and its ramifications, in this situation?''.
 $Q^{\pi}(s,a)$ offers a formalization of optimal behavior that
-may well capture such processing aspects subserved by the DMN in human agents.
+may well capture processing aspects such as subserved by the DMN in human agents.
 
@@
 for solving MDPs with even moderately sized action and state spaces
-(i.e., intractability).
+(i.e., computational intractability).
 % where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
@@
 and tracking all possible states and actions may require prohibitively excessive
-computation and memory resources.
+computation and memory resources, perhaps reflect in the especially high metabolic
+turn-over of the posterior medial DMN (PMC).
 The need of maintaining an explicit table of
@@
 $(s,a)$, as was previously proposed \citep{songNIPS2016},
-or artificial neural-network models as demonstrated in recent
-seminal investigations
+or artificial neural-network models as demonstrated in 
+seminal machine-learning models
 \citep{mnih2015,silver2016mastering} for playing complex games (atari, Go, etc.) at super-human levels.
@@
 Learning consists in starting from a given table and
-updating it during action choices,
+updating it during action choices, potentially reflected in neural processing in the PMC,
 which take the agent to different table entries.
@@
 and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than a RL agent without stochastic resampling procedures.
-Such methods have scaling as a function of available data (i.e., sample complexity) that
+Such methods scale as a function of available data (i.e., sample complexity) that
 is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
@@
 While Markov processes are usually memoryless, it is mathematically feasible to
-incorporate the previous states of such model into the current state. This
-extension may partially account for implicit memory at the behavioral level, but
+incorporate a set of previous states of such model into the current state. This
+extension may partly account for implicit memory at the behavioral level, but
 may not explain the underlying neurobiological implementation or accommodate
@@
 the deep non-linear relationships within the hierarchical connections of
-biological neural networks. The brain's adaptive synaptic connections can be
-thought as a deep neural-network architecture affording an implicit form of
+biological neural networks (especially in the association cortex).
+The brain's adaptive synaptic connections can be
+viewed as a deep neural-network architecture affording an implicit form of
 information compression of life experience. Such memory traces are stored in the
@@
 
-\subsection{Putting everything together}
+\subsection{Putting everything together: hypotheses for future studies}
 
@@
 functional couplings within any other macroscopical networks \citep{shehz2009}.
-As such, random-sampling-related baseline evaluation of possibilities and
-events may be subserved by the DMN and get partly suspended when novelty in
+As such, random-sampling-related baseline evaluation of action possibilities and
+their consequences may be subserved by the DMN and get partly suspended when novelty in
 the external environment is encountered or immediate action is required.
@@
 \citep{luksys2009stress, haushofer2014psychology}.
-Using a temporal discounting experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced-back to its corresponding neural representation.
+Using a temporal discounting experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced back to its corresponding neural representation.
 Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age and personality traits.
@@
 building blocks can further optimize the behavior of the RL agent
-by model learning from scenarios in the environment that the agent might
-only very rarely or never encounter.
+by learning from scenarios in the environment that the agent might
+encounter only very rarely or never.
 An explanation is thus offered for experiencing seemingly familiar situations that
@@
 Those attractors can then constrain both the number of factors taken into account by decision processes and the possible behaviors of the agent \citep{wang_decision_2008}.
-To examine this potential higher-level mechanism, we propose to probe how synaptic epigenesis is related to neural correlates underlying policy matrix updates: in humans the changes of functional connectivity between DMN regions can be investigated following a temporal discounting experiment and in monkeys or rodents anterograde tracing can be used to study how homolog regions of the DMN present increased synapti\c changes compare to other parts of the brain.
+To examine this potential higher-level mechanism, we propose to probe how synaptic epigenesis is related to neural correlates underlying policy matrix updates: in humans the changes of functional connectivity between DMN regions can be investigated following a temporal discounting experiment and in monkeys or rodents anterograde tracing can be used to study how homolog regions of the DMN present increased synaptic changes compare to other parts of the brain.
 

### 2018-08-11 paper/paper.tex
added haber and croxson refs for rev#2

--- 2018-08-09 paper/paper.tex
@@
 %
-Its higher energy consumption compared to other brain networks and
+Its highest energy consumption compared to other brain networks and
 its intimate coupling with conscious awareness are both pointing to
-an unknown overarching function.
+an overarching function.
 %
@@
 %
-The main purpose of DMN activity, we argue,
-may be described by Markov Decision Processes
-that optimize action policies via value estimates
-based through vicarious trial and error.
-%
-Our formal perspective on DMN function naturally accommodates as special cases
+The main purpose of the DMN, we argue,
+may be to perform optimization based on Markov Decision Processes
+through vicarious trial and error.
+%
+Our formal account of DMN function naturally accommodates as special cases
 previous interpretations based on
@@
 \citep{raichle2006brain} begs the question
-of the biological purpose underlying neural activity in the DMN,
+of the biological purpose underlying DMN activity,
 which however still remains elusive at the electrophysiological level
@@
 The network nodes that compose the human DMN
-are hubs of high baseline neural activity,
+are responsible for extended parts of the baseline neural activity,
 which typically decreases when engaged in controlled psychological experiments
@@
 
-We aim at proposing an alternative to reasoning about the DMN
-based on longstanding cognitive theory.
-The present work adopts the control-theoretical perspective of
+The present work adopts the perspective of
 a human \textit{agent} faced with the choice of the next actions
 and guided by outcomes to optimize behavioral performance. These outcomes
-can be really experienced, hypothetically imagined, or simply expected in the future.
+can be experienced, hypothetically imagined, or simply future expectations.
 Formally, we propose reinforcement learning as a particularly attractive framework
-for describing, containing, and quantifying the unknown function underlying DMN activity.
+for describing and quantifying the DMN function.
 An intelligent agent improves the interaction with the environment
@@
 
-At the psychological level,
+Psychologically,
 the more the ongoing executed task is unknown and unpracticed,
@@
 these ``offline'' processes may contribute
-to optimizing control of the organism in general.
+to optimizing control of the organism.
 We formalize
@@
 on the environment given the current state. A \textit{value function}
-maps environmental objects and events (i.e., states) to expected reward outcomes.
+maps environmental objects and events (i.e., states) to expected rewards.
 Switching between states reduces to a sequential processing model.
@@
 
-Such a RL perspective on DMN function can naturally embed human behavior
+Such an RL account of DMN function can naturally embed human behavior
 into the tension between exploitative action with immediate gains and
@@
 - ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
-allowing maximization of action outcomes across different time scales.
+allowing maximization of action outcomes across multiple time-scales.
 Such a purposeful optimization objective
@@
 seemingly arbitrary dreams during sleep
-may provide randomly sampled blocks of pseudo-experience  that are
-instrumental to iteratively
-optimize the behavioral agenda of the organism.
+may provide randomly sampled blocks of pseudo-experience instrumental to iteratively
+optimize the behavior of the organism.
 
 Evidence from computational modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
-much of brain activity is dedicated to
+the brain is largely dedicated to
 ``the development and maintenance of [a]
@@
 thus linking two literatures with currently scarce cross-references.
-Finally,
-we derive explicit hypotheses that could be tested in targeted neuroscience
-experiments in the future
-and
-we detail how our process model relates
-to previous cognitive and theoretical accounts of DMN function.
-
-Please appreciate the importance of differentiating which levels of observation are at play in the present account.
+Finally, we detail how our process model relates
+to previous accounts of DMN function and
+we derive explicit hypotheses to be tested in future neuroscience
+experiments.
+At this stage, we emphasize the importance of differentiating which levels of observation are at play in the present account.
 A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent.
-Henceforth, we will use ``inference`` when referring to aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
-%
-It is moreover important to note that
-our account does not claim that neural activity in the DMN in particular
-or the brain in general are identical with reinforcement learning algorithms.
+Henceforth, we will use ``inference`` when describing aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
+%
+Notably, our account does not claim that the brain or neural activity in the DMN are completely identical with reinforcement learning algorithms.
 Rather, we advocate feedback-based learning strategies as an attractive alternative perspective
-to describe, quantify, and interpret research findings on the DMN.
+to understand DMN function.
 
@@
 human PMC have been closely related to
-changes of conscious awareness \citep{cavanna2006precuneus, leech_pcc2014}.
+changes of conscious awareness \citep{cavanna2006precuneus}.
 Indeed,
@@
 Analogous to the PMC,
-the dorsomedial PFC (dmPFC, related to BA9) of the DMN is believed to subserve
+the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
 multi-sensory processes
@@
 dmPFC function may be closer to a
-``mental sketchpad'' \citep{goldman1996prefrontal}. This DMN part
+``mental sketchpad'' \citep{goldman1996prefrontal}, as this DMN part
 potentially subserves the de-novo construction and manipulation
@@
 Comparing to the dmPFC,
-the ventromedial PFC (vmPFC, related to BA10) is probably more specifically devoted to
+the vmPFC is probably more specifically devoted to
 subjective value evaluation and
@@
 \begin{figure}[!htbp]
-
-  \includegraphics[width=.75\linewidth]{fig_dmn.png}
+  \begin{minipage}[c]{0.63\textwidth}
+    \includegraphics[width=.8\linewidth]{fig_dmn.png}
+  \end{minipage}
+  \hspace{-4em}
+  \begin{minipage}[c]{0.45\textwidth}
     \caption{\textbf{Predictive structural association between reward system and
-      DMN nodes.} Reward tasks \citep{doherty2015structure} and neural processing in the DMN \citep{randy2008}, often considered task-negative, have been studied so far in largely separate niches of the neuroscience literature.
+        default mode network.} Reward tasks \citep{doherty2015structure} and neural processign in the DMN \citep{randy2008}, often call task-negative, have been studied so far in largely separate niches of the neuroscience literature.
 A currently underappreciated link is however suggested here based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$ [standard deviation across cross-validation folds]) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from (z-scored) volume in the DMN regions.
@@
       \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
-  The code for reproduction and visualization: \url{www.github.com/banilo/darkcontrol_2018}.}
+  The code for reproduction and visualization:
+  \url{www.github.com/banilo/darkcontrol_2018}.
+  }
   \label{fig:dmn}
+\end{minipage}
 % \end{wrapfigure}
@@
 value predictions
-\citep{behrens2008associative}. This finding
-may also explain why performance within and across participants
+\citep{behrens2008associative},
+which may also explain why performance within and across participants
 was reported to relate to state encoding in the vmPFC \citep{Schuck2016}.
-Such a ``cognitive map'' of the action space - an integral part of a RL
-agent -
-was argued to encode
+Such a ``cognitive map'' of the action space was argued to encode
 the current task state even when states are unobservable from
@@
 rich mental scene construction \citep{deuker2016event, bird2010establishing}.
-In this way, the HC may impact
+As such, the HC may impact
 ongoing perception of and action on the environment
@@
 may be subserved by a stochastic approximation
-based on a brain implementation for Monte Carlo sampling of events and outcomes.
+based on a brain implementation of Monte Carlo sampling.
 
@@
   \label{eq:cumr}
-  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
+  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t =
+  R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
 \end{equation}
@@
 This DMN region has direct connections to the NAc, known to be involved in
-reward evaluation~\citep{carlezon2009}.
+reward evaluation~\citep{carlezon2009,haber1995,croxson2005quantitative}.
 
@@
 
+\begin{mdframed}
+% neuroscience of Bellman equation
+  \vspace{2em}
+  \textbf{Box 1: Neural correlates of the Bellman equation in the DMN}
+% \hspace{-1.5em}
+Relating decomposition of consecutive action choices by the Bellman equation
+to neuroscientific insights,
+specific neural activity in the dorsal prefrontal cortex (BA9)
+was linked to processing ``goal-tree sequences''
+in human brain-imaging experiments
+\citep{koechlin1999role, koechlin2000dissociating}.
+Sub-goal exploration may require multi-task switching
+between cognitive processes
+as later parts of a solution
+frequently depend on respective earlier steps in a given solution path, which
+necessitates storage of expected intermediate outcomes.
+As such,
+``cognitive branching'' operations for nested processing of behavioral strategies
+are likely to
+entail secondary reallocation of attention and working-memory resources.
+Further brain-imaging experiments corroborated
+the prefrontal DMN to subserve
+``processes related to the management and monitoring of sub-goals while
+maintaining information in working memory''
+\citep{braver2002role}
+and to functionally couple with the hippocampus conditioned by ``deep versus shallow planning''
+\citep{kaplan2017neural}.
+%
+Moreover,
+neurological patients with lesions in this DMN region were reported
+to be impaired in aspects of realizing ``multiple sub-goal scheduling''
+\citep{burgess2000cognitive}.
+Hence,
+the various advanced human abilities subserved by the DMN, such as
+planning and abstract reasoning, can be viewed to involve some form of
+action-decision branching to enable higher-order executive control.
+\end{mdframed}
+
 % \subsection{Learning in the DMN}
@@
 + \delta \theta_{k}$ of the parameters for its approximate model of the
-$Q$-value function. Step-by-step model parameter updates warrant a learning process and definition of a loss
+$Q$-value function. This warrants a learning process and definition of a loss
 function. % \subsubsection{Stochastic gradient in the DMN via experience replay}
@@
 re-occurrence of neuron spiking sequences that have also been measured during
-specific prior actions or environmental states. The HC is a strong candidate
+specific prior actions and environmental states. The HC is a strong candidate
 for contributing to such neural reinstantiation of behavioral episodes as
@@
 action trajectories)
-and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than a RL agent without stochastic resampling procedures.
+and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than an RL agent without stochastic resampling procedures.
 Such methods have scaling as a function of available data (i.e., sample complexity) that
@@
 \subsection{Putting everything together}
-
 The DMN is today known to consistently increase in neural
@@
   in humans.
-
-\paragraph*{Neural correlates of the Bellman equation in the DMN (Humans)}
-Relating decomposition of consecutive action choices by the Bellman equation
-to neuroscientific insights,
-specific neural activity in the dorsal prefrontal cortex (BA9)
-was linked to processing ``goal-tree sequences''
-in human brain-imaging experiments
-\citep{koechlin1999role, koechlin2000dissociating}.
-Sub-goal exploration may require multi-task switching
-between cognitive processes
-as later parts of a solution
-frequently depend on respective earlier steps in a given solution path, which
-necessitates storage of expected intermediate outcomes.
-As such,
-``cognitive branching'' operations for nested processing of behavioral strategies
-are likely to
-entail secondary reallocation of attention and working-memory resources.
-Further brain-imaging experiments corroborated
-the prefrontal DMN to subserve
-``processes related to the management and monitoring of sub-goals while
-maintaining information in working memory''
-\citep{braver2002role}
-and to functionally couple with the hippocampus conditioned by ``deep versus shallow planning''
-\citep{kaplan2017neural}.
-%
-Moreover,
-neurological patients with lesions in this DMN region were reported
-to be impaired in aspects of realizing ``multiple sub-goal scheduling''
-\citep{burgess2000cognitive}.
-Hence,
-the various advanced human abilities subserved by the DMN, such as
-planning and abstract reasoning, can be viewed to involve some form of
-action-decision branching to enable higher-order executive control.
-
-We hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
-First, we expect that fMRI signals in the right temporo-parietal junction relate to behavioral changes subsequent to adaptation in the action choice tendencies (policy matrix) involved in non-value-related prediction error.
-Second, fMRI signals in the ventromedial prefrontal cortex should relate to behavioral changes following adaptation in value estimation (value matrix) due to reward-related stimulus-value association.
-We finally expect that fMRI signals in the posteromedial cortex, as a potential global information integrator, are related to shifts in overt behavior based on previous adaptations in both policy or value estimation.
-
-\paragraph*{Continuous learning paradigm (Humans)} We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
-During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigm.
-We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, the more the fMRI signals in the reward circuits should be independent of the reward context in the current sensory environment.
-
-\paragraph*{Temporal discounting experiment (Humans)} We hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits
-\citep{luksys2009stress, haushofer2014psychology}.
-Using a temporal discounting experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced-back to its corresponding neural representation.
-Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age and personality traits.
-% To exam-ine this question, Fellows and Farah (2005) investigated the effects of dorsolateral frontal (DLF) and ventromedial frontal (VMF) lobe damage on two aspects of temporal foresight—namely, temporal discounting and the futuretime perspective (measuredby an adaptation of Wallace’s [1956] list of future life events).
-
 \begin{figure}[!h]
@@
   Overview of how the constituent regions of the DMN (refer to section \ref{sec:nodes}; blue horizontal dashed line indicates the border between BA9 and BA10) may
-  map onto computational components necessary for a RL agent.
+  map onto computational components necessary for an RL agent.
   %
@@
 
-\paragraph*{Interplay with memory and “mind-searching” (Animals)} We hypothesize that experience replay for browsing problem solutions subserved by the DMN is contributing to choice behavior in mice.
-Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future.
-Necessity of cortical DMN regions, in addition to the hippocampus, for “mind-searching” candidate actions during choice behavior can be experimentally corroborated by causal disruption of DMN regions, such as by circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
-
 From the perspective of a RL agent,
@@
 
-\paragraph*{Synaptic epigenesis and creation of new entry in policy matrix (Humans \& Animals)} An additional layer of learning concerns the addition of new entries in the state and action spaces.
-Extension of the action repertoire could be biologically realized by synaptic epigenesis \citep{gisiger_acquisition_2005}.
-Indeed, the tuning of synaptic weights through learning can stabilize additional patterns of activity by creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}.
-Those attractors can then constrain both the number of factors taken into account by decision processes and the possible behaviors of the agent \citep{wang_decision_2008}.
-To examine this potential higher-level mechanism, we propose to probe how synaptic epigenesis is related to neural correlates underlying policy matrix updates: in humans the changes of functional connectivity between DMN regions can be investigated following a temporal discounting experiment and in monkeys or rodents anterograde tracing can be used to study how homolog regions of the DMN present increased synapti\c changes compare to other parts of the brain.
-
 % \paragraph*{Some notes:}
@@
 
-% \begin{mdframed}
-%   \vspace{1em}
-%   \textbf{Box 2: Proposed studies for testing the MDP account of DMN function}
-% \begin{enumerate}
+\begin{mdframed}
+  \vspace{1em}
+  \textbf{Box 2: Proposed studies for testing the MDP account of DMN function}
+\begin{enumerate}
+\item Experiment (Humans): We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
+During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigm.
+We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, the more the fMRI signals in the reward circuits should be independent of the reward context in the current sensory environment.
+
+\item Experiment (Humans): We hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
+First, we expect that fMRI signals in the right temporo-parietal junction relate to behavioral changes subsequent to adaptation in the action choice tendencies (policy matrix) involved in non-value-related prediction error.
+Second, fMRI signals in the ventromedial prefrontal cortex should relate to behavioral changes following adaptation in value estimation (value matrix) due to reward-related stimulus-value association.
+We finally expect that fMRI signals in the posteromedial cortex, as a potential global information integrator, are related to shifts in overt behavior based on previous adaptations in both policy or value estimation.
+
+\item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is contributing to choice behavior in mice.
+Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future.
+Necessity of cortical DMN regions, in addition to the hippocampus, for “mind-searching” candidate actions during choice behavior can be experimentally corroborated by causal disruption of DMN regions, such as by circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
+
+\item Experiment (Humans): We hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits
+\citep{luksys2009stress, haushofer2014psychology}.
+Using a temporal discounting experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced-back to its corresponding neural representation.
+Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age and personality traits.
+
+% To exam-ine this question, Fellows and Farah (2005) investigated the effects of dorsolateral frontal (DLF) and ventromedial frontal (VMF) lobe damage on two aspects of temporal foresight—namely, temporal discounting and the futuretime perspective (measuredby an adaptation of Wallace’s [1956] list of future life events).
+
+\item Experiment (Humans \& Animals): An additional layer of learning concerns the addition of new entries in the state and action spaces.
+Extension of the action repertoire could be biologically realized by synaptic epigenesis \citep{gisiger_acquisition_2005}.
+Indeed, the tuning of synaptic weights through learning can stabilize additional patterns of activity by creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}.
+Those attractors can then constrain both the number of factors taken into account by decision processes and the possible behaviors of the agent \citep{wang_decision_2008}.
+To examine this potential higher-level mechanism, we propose to probe how synaptic epigenesis is related to neural correlates underlying policy matrix updates: in humans the changes of functional connectivity between DMN regions can be investigated following a temporal discounting experiment and in monkeys or rodents anterograde tracing can be used to study how homolog regions of the DMN present increased synaptic changes compare to other parts of the brain.
+
+\end{enumerate}
+\end{mdframed}
 
@@
 is the \textit{free energy}.
-Here, the angular brackets denote the \textit{expectation} of the joint negative log-likelihood $-\log(p_G(\z, \x))$ w.r.t the recognition density $p_R(\z|\x)$, $\mathcal H$ is the \textit{entropy} function defined by $\mathcal H(p) := -\sum_{\z}p(\z)\log(p(\z)$, while $\operatorname{KL}(.\|.)$ is the usual \textit{Kullback-Leibler (KL) divergence} (also known as \textit{relative entropy}) defined by $\operatorname{KL}(p\|q) := \sum_{z}p(\z)\log(p(\z)/q(\z)) \ge 0$,
-which is a measure of difference between two probability distributions.
-In this framework, the goal of the agent is to iteratively refine the generative model $p_G$ and the recognition model $p_R$ so as to minimize the free energy $F_G^R(\x)$ over sensory input $\x$.
+Here, the angular brackets denote the \textit{expectation} of the joint negative log-likelihood $-\log(p_G(\z, \x))$ w.r.t the recognition density $p_R(\z|\x)$, $\mathcal H$ is the \textit{entropy} functional defined by $\mathcal H(p) := -\sum_{\z}p(\z)\log(p(\z)$, while $\operatorname{KL}(.\|.)$ is the usual \textit{Kullback-Leibler (KL) divergence} (also known as \textit{relative entropy}) defined by $\operatorname{KL}(p\|q) := \sum_{z}p(\z)\log(p(\z)/q(\z)) \ge 0$,
+which is a measure of how different two probability distributions are.
+In this framework, the goal of the agent is then to iteratively refine the generative model $p_G$ and the recognition model $p_R$ so as to minimize the free energy $F_G^R(\x)$ over sensory input $\x$.
 
@@
 \begin{itemize}
-\item  $p_R(\z|\x)$ puts a lot of mass on configurations $(\z,\x)$ which are $p_G$-likely
-% (i.e the sensation $\x$ is likely caused by $\z$)
-%  \textbf{and}
+\item  $p_R(\z|\x)$ puts a lot of mass on configurations $(\z,\x)$ which are $p_G$-likely, % (i.e the sensation $\x$ is likely caused by $\z$)
+  \textbf{and}
 \item $p_R(\z|\x)$ is as uniform as possible (i.e have high entropy), so as not to concentrate
-  all its mass on a small subset of possible causes for the sensation $\x$
+  all its mass on a small subset of possible causes for the sensation $\x$.
   \end{itemize}
@@
 Despite its popularity, criticism against the FEP has been voiced over the years,
-which we allude to in the following.
+some of which is outlined in the following.
 The main algorithm for minimizing free energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}
@@
 Defenders of the FEP have retorted by advancing the ``full package'' \citep{darkroom2012}:
-FEP is framed as being multi-scale, and there would be one scale at which
-the organism would
-be \textit{surprised} by a \textit{lack of surprise}.
-One may object that a dark room would paradoxically correspond to a state of particularly high relevance.
-Driven by the surprise-minimization objective, the FEP agent would eventually not bootstrap itself out of such saddle points to explore more interesting parts of the environment.
+FEP is proposed to be multi-scale and there would be a meta-scale at which the organism would
+be \textit{surprised} by such a \textit{lack of surprise}. According to this argument, a dark room would paradoxically correspond to a state of particularly high relevance.
+Driven by the surprise-minimization objective, the FEP agent would eventually bootstrap itself out of such saddle points to explore more interesting parts of the environment.
 % Recourse to this argument is equivalent to invoking a \textit{halting oracle}\citep{haltingoracle}, an arbitrarily powerful machine which can solve the \textit{halting problem} (an arbitrarily difficult computational problem which cannot be solved under any reasonable \textit{model of computation}, not even in principle!).
@@
 identify the sensory-stimulus-deprived room as a local minimum. Indeed, hippocampal experience replay (see \ref{sec:self}) could serve to sample memories or fantasies of alternative situations
-with reward structure. Such artificially generated \textit{internal} sensory input,
-potentially subserved by the DMN, could then entice the organism to explore the room,
+with reward structure. Such artificially generated \textit{internal} sensory input
+subserved by the DMN can entice the organism to explore the room,
 for instance by looking for and using the light switch or finding the room exit.
@@
 that is, the log-likehood of current sensory priors
-the agent has about the world. More explicitly, this corresponds to using free-energy as a Q-value approximator for the MDP in the following way:
+the agent has about the world. More explicitly, this corresponds to using free-energy as a Q-value approximator for the MDP such that
 

### 2018-08-11 paper/paper.tex

--- 2018-08-10 paper/paper.tex
@@
 Note that value buffering may be realized in the vmPFC by virtue of
-this region's connections to the NAc of the reward sytem
-\citep{carlezon2009}.
+this region's connections to the NAc of the reward system
+~\citep{carlezon2009,haber1995,croxson2005quantitative}.
 

### 2018-08-11 paper/paper.tex
Update Section 3.3 & VennDiagram

@@
 
-\section{Introduction}
+\section{Towards a formal account of default mode function: higher-order control of the organism}
 %
@@
 
-\subsection{Towards a formal account of default mode function: higher-order control of the organism}
+% \subsection{Towards a formal account of default mode function: higher-order control of the organism}
 The network nodes that compose the human DMN
@@
 
-\subsubsection{Accumulated rewards and policies}
+\paragraph*{Accumulated rewards and policies}
 The behavior of the agent is governed by a \textit{policy}, which maps states of the world
@@
 
-\subsubsection{Optimal control via stochastic gradient descent in the DMN}
+\subsubsection{Optimal control via stochastic gradient descent}
 Efficient learning of the entire set of model parameters can effectively be achieved
@@
 
-\subsection{Putting everything together: hypotheses for future studies}
+\subsection{Summary and empirical hypotheses}
 
@@
 
-\paragraph*{Neural correlates of the Bellman equation in the DMN (Humans)}
-Relating decomposition of consecutive action choices by the Bellman equation
-to neuroscientific insights,
+The current account of DMN has multiple empirical consequences and predictions.
+Perhaps one of the first experimental venue concerns the neural correlates of the Bellman equation in the DMN.
+There are already relationship between the decomposition of consecutive action choices by the Bellman equation and neuroscientific insights:
 specific neural activity in the dorsal prefrontal cortex (BA9)
-was linked to processing ``goal-tree sequences''
+was for instance linked to processing ``goal-tree sequences''
 in human brain-imaging experiments
@@
 action-decision branching to enable higher-order executive control.
-
-We hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
+Here, we hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
 First, we expect that fMRI signals in the right temporo-parietal junction relate to behavioral changes subsequent to adaptation in the action choice tendencies (policy matrix) involved in non-value-related prediction error.
@@
 
-\paragraph*{Continuous learning paradigm (Humans)} We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
+Beyond the Bellman equation and its potential underlying neural mechanisms, our process model of the DMN has also implication for experiments in neuroeconomy, especially for temporal discounting and continuous learning paradigms.
+Regarding the later, we for instance hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
 During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigm.
 We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, the more the fMRI signals in the reward circuits should be independent of the reward context in the current sensory environment.
-
-\paragraph*{Temporal discounting experiment (Humans)} We hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits
+In the case of temporal discounting, we hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits
 \citep{luksys2009stress, haushofer2014psychology}.
@@
 
-The HC may contribute to generating
+Another prediction from the current model is that the HC may contribute to generating
 perturbed action-transition-state-reward samples
@@
 
-\paragraph*{Interplay with memory and “mind-searching” (Animals)} We hypothesize that experience replay for browsing problem solutions subserved by the DMN is contributing to choice behavior in mice.
+This interplay between memory and “mind-searching” also supports that experience replay for browsing problem solutions subserved by the DMN is contributing to choice behavior in mice.
 Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future.
-Necessity of cortical DMN regions, in addition to the hippocampus, for “mind-searching” candidate actions during choice behavior can be experimentally corroborated by causal disruption of DMN regions, such as by circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
-
+We hypothesize that, in addition to the hippocampus, there is a necessity of cortical DMN regions for “mind-searching” candidate actions during choice behavior.
+It can be experimentally corroborated by causal disruption of DMN regions, such as by circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
 From the perspective of a RL agent,
@@
 
-\paragraph*{Synaptic epigenesis and creation of new entry in policy matrix (Humans \& Animals)} An additional layer of learning concerns the addition of new entries in the state and action spaces.
+A last empirical perspective concerns the role of synaptic epigenesis in the creation of new entry in policy matrix.
+Indeed, we did not adress here the additional layer of learning which concerns the addition of new entries in the state and action spaces.
 Extension of the action repertoire could be biologically realized by synaptic epigenesis \citep{gisiger_acquisition_2005}.
@@
 \begin{figure}[!h]
-  \includegraphics[width=1\linewidth]{VennDiagram-2017-09-08.pdf}
+  \includegraphics[width=1\linewidth]{VennDiagram-2018-08-11.pdf}
   \caption{\textbf{Situating Markov Decision Processes among other accounts of default mode function.}

### 2018-08-11 paper/paper.tex
some edits to summary

@@
 
-\subsection{Summary and empirical hypotheses}
+\subsection{Summary and hypotheses for future studies}
 
@@
 
-The current account of DMN has multiple empirical consequences and predictions.
+Our formal account on the DMN readily motivates several empirical predictions for future neuroscience research.
 Perhaps one of the first experimental venue concerns the neural correlates of the Bellman equation in the DMN.
@@
 action-decision branching to enable higher-order executive control.
-Here, we hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
+
+We therefore hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
 First, we expect that fMRI signals in the right temporo-parietal junction relate to behavioral changes subsequent to adaptation in the action choice tendencies (policy matrix) involved in non-value-related prediction error.
 Second, fMRI signals in the ventromedial prefrontal cortex should relate to behavioral changes following adaptation in value estimation (value matrix) due to reward-related stimulus-value association.
-We finally expect that fMRI signals in the posteromedial cortex, as a potential global information integrator, are related to shifts in overt behavior based on previous adaptations in both policy or value estimation.
-
-Beyond the Bellman equation and its potential underlying neural mechanisms, our process model of the DMN has also implication for experiments in neuroeconomy, especially for temporal discounting and continuous learning paradigms.
-Regarding the later, we for instance hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
-During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigm.
+We further expect that fMRI signals in the posteromedial cortex, as a potential global information integrator, are related to shifts in overt behavior based on previous adaptations in both policy or value estimation.
+
+Our process model of the DMN has also implications for experiments in neuroeconomy; especially for temporal discounting and continuous learning paradigms.
+More specifically, we hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
+During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a multi-step learning paradigm.
 We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, the more the fMRI signals in the reward circuits should be independent of the reward context in the current sensory environment.
@@
 \citep{luksys2009stress, haushofer2014psychology}.
-Using a temporal discounting experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced back to its corresponding neural representation.
-Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age and personality traits.
+Using such a delayed-reward experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced back to its corresponding neural representation.
+Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age, education, IQ, and personality traits.
 % To exam-ine this question, Fellows and Farah (2005) investigated the effects of dorsolateral frontal (DLF) and ventromedial frontal (VMF) lobe damage on two aspects of temporal foresight—namely, temporal discounting and the futuretime perspective (measuredby an adaptation of Wallace’s [1956] list of future life events).
@@
 
-Another prediction from the current model is that the HC may contribute to generating
+As another experimental prediction derived from our MDP approach to the DMN,
+the HC may contribute to generating
 perturbed action-transition-state-reward samples
@@
 (i.e., recalled, hypothesized, and forecasted scenarios).
-The small variations in these experience samplings allow searching
-a larger space of model parameters and possible experiences.
+The small variations in these experience samplings allow searching through
+a larger space of model parameters and candidate experiences.
 Taken to its extreme, stochastic recombination of experience
@@
 
-This interplay between memory and “mind-searching” also supports that experience replay for browsing problem solutions subserved by the DMN is contributing to choice behavior in mice.
+A possible interplay between memory retrieval and ``mind-searching'' moreover suggests that experience replay for browsing problem solutions subserved by the DMN contributes to choice behavior in mice.
 Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future.
-We hypothesize that, in addition to the hippocampus, there is a necessity of cortical DMN regions for “mind-searching” candidate actions during choice behavior.
+We hypothesize that, in addition to the hippocampus, there is a necessity of cortical DMN regions for ``mind-searching'' candidate actions during choice behavior.
 It can be experimentally corroborated by causal disruption of DMN regions, such as by circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
@@
 
-A last empirical perspective concerns the role of synaptic epigenesis in the creation of new entry in policy matrix.
+A last experimental prediction for future studies concerns the role of synaptic epigenesis in the creation of new entries in policy matrix.
 Indeed, we did not adress here the additional layer of learning which concerns the addition of new entries in the state and action spaces.
 Extension of the action repertoire could be biologically realized by synaptic epigenesis \citep{gisiger_acquisition_2005}.
-Indeed, the tuning of synaptic weights through learning can stabilize additional patterns of activity by creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}.
+The tuning of synaptic weights through learning can stabilize additional patterns of activity by creating new attractors in the neural dynamics landscape \citep{takeuchi_synaptic_2014}.
 Those attractors can then constrain both the number of factors taken into account by decision processes and the possible behaviors of the agent \citep{wang_decision_2008}.

### 2018-08-12 paper/paper.tex
gros commit

@@
 
-\section{Towards a formal account of default mode function: higher-order control of the organism}
+\section{Introduction}
 %
@@
 DMN (see here for details on anatomical connections: \citep{randy2008}).
-Additionally, myelination patterns of axon connections were found to finish
-particularly late in these cortical areas \citep{flechsig1920}, often believed to suggest
+Additionally, myelination patterns of axon connections were found to complete
+particularly late in these cortical areas \citep{flechsig1920}, often believed to reflect
 sophistication of subserved neural processes
@@
 are responsible for higher-order control of human behavior.
-Our functional account follows the notion of
+Our perspective therefore follows the notion of
 ``a hierarchy of brain systems with
@@
 
-% \subsection{Towards a formal account of default mode function: higher-order control of the organism}
+\section{Towards a formal account of default mode function: higher-order control of the organism}
 The network nodes that compose the human DMN
 are hubs of high baseline neural activity,
-which typically decreases when engaged in controlled psychological experiments
+which typically decreases when engaged in well-defined psychological experiments
 \citep{raichle_baseline}.
@@
 this brain network appears to
-simultaneously modulate perception-action cycles in the present
+simultaneously influence perception-action cycles in the present
 and to support mental travel
@@
 a human \textit{agent} faced with the choice of the next actions
-and guided by outcomes to optimize behavioral performance. These outcomes
-can be really experienced, hypothetically imagined, or simply expected in the future.
+guided by outcomes to optimize behavioral performance. These outcomes
+can be really experienced, hypothetically imagined, or expected in the future.
 Formally, we propose reinforcement learning as a particularly attractive framework
@@
 Henceforth,
-\textit{control} refers to the influence that an agent exerts when interacting
+\textit{control} refers to the influence that an agent exerts by interacting
 with the environment to reach preferred states.
@@
 
-Such a RL perspective on DMN function can naturally embed human behavior
+Such a RL perspective on DMN activity can naturally embed human behavior
 into the tension between exploitative action with immediate gains and
@@
 A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent.
-Henceforth, we will use ``inference`` when referring to aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
+Henceforth, we will use ``inference`` when referring to aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the
+cognitive behavior of the agent.
 %
@@
   \caption{\textbf{Predictive structural association between reward system and
-      DMN nodes.} Reward tasks \citep{doherty2015structure} and neural processing in the DMN \citep{randy2008}, often considered task-negative, have been studied so far in largely separate niches of the neuroscience literature.
+      DMN nodes.} Reward tasks \citep{doherty2015structure} and neural processing in the DMN \citep{randy2008}, often considered ``task-negative'', have been studied so far in largely separate niches of the neuroscience literature.
 A currently underappreciated link is however suggested here based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$ [standard deviation across cross-validation folds]) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from (z-scored) volume in the DMN regions.
@@
 \citep{doria2010}, while
-the right and left TPJs may become fully functionally integrated into this major brain
+the right and left TPJs may become fully functionally integrated into this macroscopical
 network only after birth.
@@
 that has been shown to be closely related to
-multi-sensory prediction and prediction error signaling
+multi-sensory event representation and prediction error signaling
 \citep{downar2000multimodal, vetter2011, shulman2010, shulman2007}.
@@
 \citep{corbetta2002control}.
-Its involvement was repeatedly reported in
+Its involvement was repeatedly reported in monitoring
 multi-step action execution \citep{hartmann2005takes},
 visuo-proprioceptive conflict \citep{Balslev2005},
-spatial reorienting \citep{corbetta2000},
+spatial re-orientation \citep{corbetta2000},
 and
@@
 a capacity believed to be enabled by the DMN.
-Viewed from a RL account,
+Viewed from a RL perspective,
 the RTPJ might reflect an important relay that shifts away
 from the ‘‘internally directed’’ baseline processes
-to, instead, deal with unexpected environmental stimuli and events.
+to, instead, deal with unexpected environmental cues and events.
 % (TODO: slow endogeneous changes in state transitions by the policy and value matrices)
@@
 random sampling of possible experiences, as well as
-estimation of instantaneous and expected delayed reward outcomes
+estimation of instantaneous and delayed reward outcomes
 are key components of intelligent RL agents
@@
 RL is an area of machine learning concerned with searching optimal behavioral strategies through interactions
-with an \textit{environment} with the goal to maximize the \textit{cumulative reward} over time.
-The optimal behavior typically
+with an \textit{environment} with the goal to maximize the \textit{cumulative reward} over time
+\citep{sutton1998reinforcement}.
+Optimal behavior typically
 takes the future into account as certain rewards could be \textit{delayed}.
@@
 more naturally interpreted as \textit{punishment}.
-The environment be partly controlled by
+The environment can be partly controlled by
 the action of the agent and the reward can be thought
@@
 
-The environment is generally taken as \textit{stochastic},
+The environment is assumed to be \textit{stochastic},
 that is, changing in random ways. In addition, the environment is only
 \textit{partially observable} in the sense that only limited aspects of the environment's
-state are accessible to the agent's sensory input
+state are accessible to the agent's sensory perception
 \citep{starkweather2017dopamine}.
@@
 
-\subsection{Markov decision processes}
+\subsection{Markov Decision Processes}
 % RL has had considerable success in modeling complex decision-making problems,
@@
 and possibly driven by neural processing in the vmPFC)
-by maintaining probability distributions over actions according to their expected utility.
+by maintaining probability distributions over actions
+(possibly represented in the dmPFC)
+according to their expected utility.
 % This is performed alongside each transition in the state space of the environment
@@
 \begin{definition*}
-Mathematically, an MDP is simply a quintuplet $(\mathcal S, \mathcal A, r, p)$ where
+Mathematically, an MDP involves a quadruple $(\mathcal S, \mathcal A, r, p)$ where
 \begin{itemize}
@@
 \citep{starkweather2017dopamine,o2006making}.
-Nevertheless, this mathematical property of POMDPs mostly accounts for implicit $memory$.
-Since the current paper is concerned with plausibility at the behavioral and neurobiological level, we will address below how our account can accommodate the neurophysiological constraints of the DMN and the explicit $memory$ characteristics of human agents.
+Nevertheless, this mathematical property of POMDPs mostly accounts for implicit memory.
+Since the current paper is concerned with plausibility at the behavioral and neurobiological level, we will address below how our account can accommodate the neurophysiological constraints of the DMN and the explicit memory characteristics of human agents.
 
@@
 are applicable to something as complex as human behavior. This class of
-reinforcement-learning models has had numerous successes in diverse domains.
-For instance, financial trading is largely a manifestation of strategic decision-making of interacting human agents.
+reinforcement-learning models has had numerous successes in diverse applied domains.
+For instance, MDPs have been successfully used in financial trading is largely a manifestation of strategic decision-making of interacting human agents.
 According to how the market responds, the agent incurs gain or loss as environmental feedback of the executed financial actions.
@@
 More recent work has
-developped an MDP-based way of reasoning about future behavior
+developped an MDP-related way of reasoning about future behavior
 of other agents \citep{robinowitz2018}.
@@
 
-\paragraph*{Reinforcement learning in the brain.}
+\paragraph*{Reinforcement learning in the brain?}
 RL has been argued to be a biologically plausible mechanism in the human brain
@@
 of expected utility outcomes as a strategy for action choice in uncertain
-environments, a situation captured by the formalism of MDPs.
+environments, a view captured by the formalism of MDPs.
 It has also long been proposed \citep{dayan2008decision} that there
 can be a mapping between algorithmic aspects
-underlying model-free and model-based RL learning
-and neurobiological aspects underlying brain function.
+underlying model-free and model-based RL
+and neurobiological aspects underlying decision-making, which involves parts of the DMN.
 The neurotransmitter dopamine could serve
-as a 'teaching signal' to guide estimation of value associations
+as a ``teaching signal'' to guide estimation of value associations
 and action policies by modulating
@@
 The behavior of the agent is governed by a \textit{policy}, which maps states of the world
-to probability distributions over candidate actions.
+to probability distributions over candidate actions
+(potentially represented in the dmPFC).
 Starting at time $t=0$,
@@
 Since an action executed in the present moment might have repercussions in the
-far future, it turns out that the quantity to optimize is not the instantaneous
+far future. It turns out that the quantity to optimize is not the instantaneous
 rewards $r(s, a)$, but a \textit{cumulative reward} estimate which takes into
 account expected reward from action choices in the future. A common approach to
-modeling this gathered outcome is the time-discounted cumulative reward% function
+modeling this gathered outcome,
+which is likely to involve extended parts of the DMN,
+is the time-discounted cumulative reward% function
 \begin{equation}
@@
 This random variable
-  \footnote{The reward outcome is random as it depends both on the environment's dynamics and the
+measures the cumulative reward of
+following an action policy $\pi$.
+The reward outcome is random because it depends both on the environment's dynamics and the
 policy $\pi$ being executed. The exponential delay discounting function used
@@
 although psychological experiments may also reveal other discounting regimes
-  \citep{green2004discounting}.}
-measures the cumulative reward of
-following an action policy $\pi$.
+\citep{green2004discounting}.
 %
@@
 
-The goal of the RL agent is then to successively update this action policy in order
+The goal of the RL agent is then to successively update this action policy
+(perhaps most closely related to the PMC) in order
 to maximize $G^\pi$ on average (cf. below). In \eqref{eq:cumr},
@@
 the constant $\gamma$ $(0 \le \gamma < 1$) is the \textit{reward discount factor},
-viewed to be characteristic for a certain agent.
+viewed to be characteristic trait for a certain agent.
 On the one hand,
@@
 setting $0 < \gamma < 1$ allows a learning process to arise.
-A positive $\gamma$ can be seen as calibrating risk-seeking trait of the intelligent agent,
+A positive $\gamma$ can be seen as calibrating the risk-seeking trait of the intelligent agent,
 that is, the behavioral predispositions related to trading longer delays
@@
 Given only the limited information available from an MDP, at a state $s$ the average
-utility of choosing an action $a$ under a policy $\pi$ can be captured by the single number
+utility of choosing an action $a$ under a policy $\pi$ can be captured by the single quantity
 % We can now formulate the purpose of the DMN
@@
 \end{equation}
-The learning goal is to approach the
-policy $\pi^*$ as close as possible, that is to \text{solve} the MDP.
+The learning goal is to approach the ideal
+policy $\pi^*$ as close as possible, that is, to \text{solve} the MDP.
 Note that \eqref{eq:qlearning} presents merely a definition and
 does not lend itself as a candidate schema
-for solving MDPs with even moderately sized action and state spaces
+for fully computing MDPs with even moderately sized action and state spaces
 (i.e., computational intractability).
@@
 a dynamic decomposition of optimal behavior by dividing the $Q$-value function into the immediate
-reward and the discounted rewards of the upcoming states.
+reward component and the discounted reward component of the upcoming states.
 The optimal Q-value operator $Q^*$ is a fixed point for this equation.
@@
 computation and memory resources, perhaps reflect in the especially high metabolic
-turn-over of the posterior medial DMN (PMC).
+turn-over of the posterior medial DMN (i.e., PMC).
 The need of maintaining an explicit table of
@@
 %  and Parr et al. 2003 (JMLR)].
-In the DMN, the dmPFC would implement such a hard-max lookup
+In the DMN, the dmPFC is conceivable to implement such a hard-max lookup
 over the action space.
@@
 Successful learning in brains and computer algorithms may not be possible without a
-defined learning goal \textemdash the \textit{loss function}.
+defined optimization goal \textemdash the \textit{loss function}.
 The action $a$ chosen in state $s$ according to the policy matrix defined in
@@
 and is recorded in replay memory buffer \textemdash
-which we hypothesize to be subserved by the
+which we hypothesize to involve especially the
 HC \textemdash, possibly discarding the oldest entries to make space:
@@
 neuroscience experiments have repeatedly indicated in rats, mice, cats, rabbits,
-songbirds, and monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}. Refer to
+songbirds, and monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}.
+Importantly, neural encoding of abstract representations of space and meaning 
+may extent to several parts of the DMN \citep{behrens_grid}.
+Refer to
 Fig. \ref{fig:rl_process_chart}.
@@
 At the current step $k$, computing an optimal parameter update then corresponds to
-finding the model parameters $\theta_{k}$ which minimize the following mean-squared loss function
+finding the model parameters $\theta_{k}$ which minimize the following mean-squared
+optimization loss
 \begin{equation}
@@
 % \textbf{We still may need to explain y in the formula with one sentence.}
-A recently proposed, practically successful alternative approach is to learn this
+A recently proposed, practically successful alternative approach is to estimate the
 representation using an artificial deep neural-network model. This approach leads to the
-so-called \textit{deep Q-learning}~\citep{mnih2015,silver2016mastering} family of methods which
+so-called \textit{deep Q-learning}~\citep{mnih2015,silver2016mastering} - a family of methods which
 is the current state-of-the-art in RL research.
@@
 In RL, Monte Carlo simulation is a common means to update the agent's belief state
-based on stochastic sampling around states and possible transitions \citep{daw2014algorithmic,silver2010monte}.
+based on stochastic sampling of environmental states and
+possible transitions \citep{daw2014algorithmic,silver2010monte}.
 Monte Carlo simulation provides a simple method for evaluating the value of a state.
-This inference procedure provides an effective mechanism both for tree search (of the considered
-action trajectories)
+This inference procedure provides an effective mechanism both for tree search of the considered
+action trajectories
 and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than a RL agent without stochastic resampling procedures.
@@
 actions and rewards \citep{aronov2017, chao2017interaction, boyer2008evolutionary}.
-These simulations of experience batches
+These stochastic simulations of experience batches re-assembled from memory
 would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
@@
 The brain's adaptive synaptic connections can be
-viewed as a deep neural-network architecture affording an implicit form of
+viewed as a deep artificial neural-network architecture affording an implicit form of
 information compression of life experience. Such memory traces are stored in the
@@
 
-\subsection{Predictive coding hypothesis}
+\subsection{Predictive coding}
 Predictive coding mechanisms
@@
 
-\subsection{Semantic hypothesis}
+\subsection{The semantic account}
 This frequently proposed cognitive account to explain DMN function revolves
@@
 
-\subsection{Sentinel hypothesis}
+\subsection{The sentinel account}
 Regions of the DMN have been proposed to process the
@@
 
-\subsection{The free-energy principle and active inference}
+\subsection{A note on the free-energy principle and active inference}
 % INTRO
-According to theories of the \textit{free-energy principle} (FEP) and
+According the \textit{free-energy principle} (FEP) and theories of
 \textit{active inference} \citep{friston2010free, fristonAIorRL,dayan1995helmholtz},
 the brain corresponds to
-a biomechanical reasoning engine. It is
+a biomechanical reasoning engine. Much of neural computation is
 dedicated to minimizing the long-term average of surprise: the log-likelihood of the observed sensory input --more precisely, an upper bound thereof-- relative to the expectations about the external world derived from internal representations.
@@
 However, surprise is challenging to optimize numerically
-because we need to sum over all hidden causes $\z$ of the sensations (an intractable problem).
+because we need to solve the intractable problems of
+summing over all hidden causes $\z$ of the sensations (an intractable problem).
 Instead, FEP therefore minimizes an upper-bound on surprise given by
@@
 %  \textbf{and}
-\item $p_R(\z|\x)$ is as uniform as possible (i.e have high entropy), so as not to concentrate
+\item $p_R(\z|\x)$ is as uniform as possible (i.e., have high entropy), so as not to concentrate
   all its mass on a small subset of possible causes for the sensation $\x$
@@
 % \paragraph{Main criticisms.}
-Despite its popularity, criticism against the FEP has been voiced over the years,
+Despite its popularity, criticism against the FEP has been voiced repeatedly,
 which we allude to in the following.
@@
 of forward (generation) and backward (recognition) models $p_G$ and $p_R$ that
-together does not correspond to optimization of (a bound of) the marginal likelihood
+together does not correspond to optimization of a bound of the marginal likelihood
 because KL divergence is not symmetric in its arguments.
 
-These considerations render the brain less likely to implement the wake-sleep algorithm or a variant thereof. More recently,
-\textit{variational auto-encoders} (VAEs) \citep{kingma2013auto} emerged that may provide an efficient alternative to the wake-sleep algorithm. VAEs
+These considerations render the brain less likely to implement
+a variant of the wake-sleep algorithm. More recently,
+\textit{variational auto-encoders} \citep{kingma2013auto} emerged that may provide an efficient alternative to the wake-sleep algorithm. Such compression-and-reconstruction models
 overcome a number of the technical limits of the wake-sleep algorithm
 by using a reparametrization maneuver, which makes it possible to do differential calculus on random sampling procedures without exploding variance. As a result, unlike the wake-sleep algorithm
-for minimizing free energy, VAEs can be efficiently trained via back-propagation of prediction errors.
+for minimizing free energy,
+variational auto-encoders can be efficiently trained via back-propagation of prediction errors.
 
@@
 conceptualize ongoing behavior
-in an organism to be geared towards the surprise-minimizing goal,
-an organism entering a dark room
-% \footnote{\textit{Dark} is used here as an allegory to indicate the complete absence of any \textit{external} sensory input.}
-
-% (Fig. \ref{fig:darkroom})
-
+in an organism to be geared towards the surprise-minimizing goal.
+Hence, an organism entering a dark room
 would remain trapped in this location
@@
 However, such a behavior is seldom observed in humans in the
-real world. In a dark room, the intelligent agents would search for light sources by explore
+real world. In a dark room, the intelligent agents would search for light sources to explore
 the surroundings or aim to exit the room.
-Defenders of the FEP have retorted by advancing the ``full package'' \citep{darkroom2012}:
-FEP is framed as being multi-scale, and there would be one scale at which
-the organism would
-be \textit{surprised} by a \textit{lack of surprise}.
-One may object that a dark room would paradoxically correspond to a state of particularly high relevance.
+% Defenders of the FEP have retorted by advancing the ``full package'' \citep{darkroom2012}:
+% FEP is framed as being multi-scale, and there would be one scale at which
+% the organism would
+% be \textit{surprised} by a \textit{lack of surprise}.
+One may object that, for the FEP agent,
+a dark room would paradoxically correspond to a state of particularly high relevance.
 Driven by the surprise-minimization objective, the FEP agent would eventually not bootstrap itself out of such saddle points to explore more interesting parts of the environment.
@@
 our RL framework.
-This becomes possible by recasting the Q-value function (i.e., expected long-term reward) maximized by the DMN to
+This is possible by recasting the Q-value function (i.e., expected long-term reward) maximized by the DMN to
 correspond to negative surprise,
 that is, the log-likehood of current sensory priors
-the agent has about the world. More explicitly, this corresponds to using free-energy as a Q-value approximator for the MDP in the following way:
+the agent has about the world. More explicitly,
+this formulation corresponds to using free-energy as a Q-value approximator for the MDP in the following way:
 
@@
 we have proposed a way of reasoning
-\textit{how} this major network may do what it is doing.
+\textit{how} this major brain network may do what it is doing.
 %
@@
 opportunity in the environment can be deconstructed, evaluated, and exploited
-when confronted with challenging interdependent decisions.
+when an agent is confronted with challenging interdependent decisions.
 %
@@
 
-A major hurdle in guessing DMN function from cognitive brain-imaging studies has been its similar neural engagement in different time scales: thinking about the past (e.g., autobiographical memory retrieval), thinking about hypothetical presents (e.g., daytime mind-wandering), and
+A major hurdle in guessing DMN function from cognitive brain-imaging studies has been its similar neural engagement in different time scales: thinking about the past (e.g., autobiographical memory retrieval),
+imagining hypothetical presents (e.g., daytime mind-wandering), and
 anticipating scenarios yet to come (e.g., delay discounting).
-The MDP account of DMN activity offers a natural integration of a-priori diverging neural processes into a common framework.
+The MDP account of DMN activity offers a natural integration of a-priori diverging
+cognitive processes into a common framework.
 %
@@
 %
-Future neuroscience experiments on the DMN should operationalize
-the set of action, value, and state variables governing
+We encourage neuroscience experiments on the DMN to operationalize
+the set of action, value, and state variables that govern
 the behavior of intelligent RL agents.

### 2018-08-12 paper/paper.tex
Typos

@@
   constant exploration of candidate action trajectories and
-  nested estimaton of their
+  nested estimation of their
   cumulative reward outcomes. Implicit computation of future choices
@@
 
-We therefore hypothesize a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
+We therefore hypothesize in humans a functional dissociation between computations pertaining to action policy versus adapting stimulus-value associations as we expect implementation in different subsystems of the DMN.
 First, we expect that fMRI signals in the right temporo-parietal junction relate to behavioral changes subsequent to adaptation in the action choice tendencies (policy matrix) involved in non-value-related prediction error.
@@
 Our process model of the DMN has also implications for experiments in neuroeconomy; especially for temporal discounting and continuous learning paradigms.
-More specifically, we hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
+More specifically, we hypothesize in humans a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry.
 During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a multi-step learning paradigm.
 We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, the more the fMRI signals in the reward circuits should be independent of the reward context in the current sensory environment.
-In the case of temporal discounting, we hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits
+In the case of temporal discounting, we hypothesize in humans that the relevant time horizon is modulated by various factors such as age, acute stress, and time-enduring impulsivity traits
 \citep{luksys2009stress, haushofer2014psychology}.
@@
 
-A last experimental prediction for future studies concerns the role of synaptic epigenesis in the creation of new entries in policy matrix.
+A last experimental prediction for future studies concerns how synaptic epigenesis may shape the policy matrix.
 Indeed, we did not adress here the additional layer of learning which concerns the addition of new entries in the state and action spaces.

### 2018-08-12 paper/paper.tex
More typos

@@
 variation of the grey-matter volume in DMN regions
-is linked to the reward circuitry (Fig. 2),
+is linked to the reward circuitry (Fig. \ref{fig:dmn}),
 thus linking two literatures with currently scarce cross-references.
@@
 Importantly, neural encoding of abstract representations of space and meaning 
-may extent to several parts of the DMN \citep{behrens_grid}.
-Refer to
-Fig. \ref{fig:rl_process_chart}.
+may extent to several parts of the DMN \citep{behrens_grid} (see
+Fig. \ref{fig:rl_process_chart}).
 

### 2018-08-12 paper/paper.tex
adapted kw
added submitted rev

@@
 % official NIPS keywords
-\textbf{\\keywords}: systems neuroscience, artificial intelligence, reinforcement learning, mind-wandering
+\textbf{\\keywords}: systems neuroscience, artificial intelligence, mind-wandering
 
@@
 for processing the constant influx of sensory stimuli and
-performing action on the environment.
+performing action on the environment (Fig. \ref{fig:VennDiagram}).
 There are hence a number of common properties
@@
 Analogies might naturally tie incoming new sensory stimuli to
-explicit world knowledge (i.e., semantics)
+explicit world knowledge (i.e., semantics, Fig. \ref{fig:VennDiagram})
 \citep{bar2009proactive}.
@@
 a ``radar'' function to
-detect subjectively relevant cues and unexpected events in the environment.
+detect subjectively relevant cues and unexpected events in the environment
+(Fig. \ref{fig:VennDiagram}).
 Propositions of a sentinel function to underlie DMN activity
@@
 and rats \citep{lu2012rat}, three species with
-social capacities that can be expected to be less advanced than in humans.
+social capacities that can be expected to be less advanced than in humans
+\citep{mars2012_social_brain}.
 
